### 2017-09-11 simulation_studies/fdr_simulation_studies.Rmd
started R Markdown for evaluating methods using simulated data

---
title: "Benchmarking FDR Methods"
author: "Rafalab Journal Club Members"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The goal of this working group is to benchmark recent covariate-adjusted FDR methods. For 
complete details, check out our [benchmark-fdr GitHub repo](https://github.com/pkimes/benchmark-fdr).

In this R Markdown, we use a Monte Carlo simulation study  to evaluate different simulation scenarios:

1. The `All nulls` simulation: All tests are simulated from the null distribution (no difference). 
2. add here... 
3. add here...
4. etc

In all cases, we define

* $m$ = number of tests
* $M$ = Monte Carlo replications
*

```{r}
m = 20000 # number of tests
M = 1000 # number of MC reps
```

**Things to do**: add a summary of all the simulation studies we plan to do.

## Methods to evaluate

- IHW
- ASH
- BL
- Scott
- locfdr
- mixfdr

**Things to do**: add details, links, etc of methods here

## Metrics to evaluate

- Number of discoveries
- Control FDR at $\alpha$
- Power
- Relationship between $pi_0$ and covariate
- Assumptions: unimodal, etc

**Things to do**: Better describe metrics to evaluate our simulation studies

# The `All nulls` simulation

Simulate two standard normals with mean 0, sd = 1. Calculate t-statistic. 
Independent covariate: Uniform between 0, 1. Apply different methods.

## Control for FDR at $\alpha$

Evaluate methods control for FDR at nominal $\alpha$ level ranging from [0.01, 0.1].

```{r}
# notes: convert this into R code
# - IHW: p-values from t-test, independent covariate
# - ASH: difference in means, independent covariate
# - BL: p-values from t-test, multiple covariates (with and without cov structure)
#   - lambda: 0.75, 0.85, 0.95
# - Scott: t-test statistic, independent covariate
#   - assumption of test statistic follow a standard normal. this is good here. 
# - BH: p-values only. 
# - locfdr: zz = t-test statistic, nulltype (try both). 
# - mixfdr: ? do we want to include?

```

## add here...

### 2017-09-12 datasets/GWAS/GWAS.Rmd
start adding GWAS datasets

---
title: "GWAS datasets for FDR benchmarking"
author: "Keegan Korthauer"
date: "9/12/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Summary

Here are the GWAS datasets we will use to benchmark the FDR methods. For each one
we provide the script to download the data and read into R. We'd like to verify 
that each dataset contains the desired inputs:

*p-value
*effect size
*standard error
*additional covariates (optional)

Note that while knitting this document will download the data, I'm not committing 
the data itself to this repo. See the knitted `.html` document if you do not 
wish to download the datasets but still wish to see the output.

### GWAS meta-analysis for BMI (included in Boca-Leek)

Here we download the GWAS BMI dataset analyzed in the Boca-Leek paper.

```{bash, GWAS1-download}
curl -O "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
```

Next we unzip the file, and delete the files we won't use (they provide results
subsetted by ancestry and sex), keeping only the file with all ancestries and
both sexes `BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt`. Note that it seems that
the Boca-Leek paper actually used the 'EuropeanOnly' subset, but it's not clear
why (to me), and there's no reason we need to do the same.

```{bash, GWAS1-unzip}
unzip BMI.SNPadjSMK.zip
rm BMI.SNPadjSMK.zip
rm BMI.SNPadjSMK.Women.EuropeanOnly.txt
rm BMI.SNPadjSMK.Women.AllAncestry.txt
rm BMI.SNPadjSMK.Men.EuropeanOnly.txt
rm BMI.SNPadjSMK.Men.AllAncestry.txt
rm BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt
```

Next, we'll read in the unzipped `.txt` file into R and verify that it contains
the necessary inputs to run the FDR benchmark comparisons.

```{r, GWAS1-verify-contents}
library(data.table)
bmi <- fread("BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt", header=TRUE)
head(bmi)
```

It looks like we have:

- [x] p-value
- [x] effect size
- [x] standard error
- [x] additional covariates: 
  *`N`: Number of samples with this SNP - BMI association measured
  *`Freq_Allele1_HapMapCEU`: Allele Frequency of Allele1, as measured in HapMapCEU

### 2017-09-13 simulation_studies/fdr_simulation_studies.Rmd
started null simulations using bonf, bh, qvalue, ihw and ash

@@
 ```{r}
-m = 20000 # number of tests
-M = 1000 # number of MC reps
+m = 2000 # number of tests
+M = 10 # number of MC reps
 ```
@@
 
+- Bonferroni 
+- BH
 - IHW
@@
 
-## Control for FDR at $\alpha$
+## Control for FDR and FWER at $\alpha$
 
-Evaluate methods control for FDR at nominal $\alpha$ level ranging from [0.01, 0.1].
+Evaluate methods control for false discovery rate (FDR) and 
+family-wise error rate (FWER) at nominal $\alpha$ level ranging from [0.01, 0.1].
 
-```{r}
+```{r, load-libraries, message=FALSE}
+library(dplyr)
+library(ggplot2)
+library(doParallel)
+
+library(IHW)
+library(ashr)
+library(qvalue)
+library(swfdr)
+
 # notes: convert this into R code
-# - IHW: p-values from t-test, independent covariate
-# - ASH: difference in means, independent covariate
 # - BL: p-values from t-test, multiple covariates (with and without cov structure)
@@
 #   - assumption of test statistic follow a standard normal. this is good here. 
-# - BH: p-values only. 
 # - locfdr: zz = t-test statistic, nulltype (try both). 
 # - mixfdr: ? do we want to include? 
+```
+
+```{r simulate-null-data-helper-functions}
+du_ttest_sim <- function(m, pi0, effect_size, n_samples=20, n_groups = 2, seed=NULL){
+  if (!is.null(seed)) set.seed(seed)
+    m0 <- ceiling(m*pi0)
+    false_nulls <- sample(1:m, m-m0)
+    z_table <- matrix(rnorm(n_samples*m), ncol=n_samples)
+    z_table[false_nulls,(n_samples/2+1):n_samples] <- matrix(rnorm(n_samples/2*(m-m0), effect_size), ncol=n_samples/2)
+    H <- rep(0,m)
+    H[false_nulls] <- 1
+    gF <- factor(rep(1:n_groups, each=n_samples/n_groups))
+    t_test <- genefilter::rowttests(z_table, gF)
+    sds <- genefilter::rowSds(z_table) # pooled var reduces to unpooled var b/c same sample sizes across groups 
+    SE <- sds / sqrt(n_samples)
+    simDf <- data.frame(H = H, t_stat = t_test$statistic, t_dm = t_test$dm, 
+                      t_pval = t_test$p.value, sd = sds, SE = SE)
+    return(simDf)
+}
+
+calculate_test_stats <- function(sim, adj_p, alpha){
+  rejected <- adj_p <= alpha
+  rjs <- sum(rejected) # number of rejected nulls
+  false_rjs <- sum(sim$H == 0 & rejected) # number of false rejected nulls
+  rj_ratio <- rjs/nrow(sim)
+  FDP <- ifelse(rjs == 0, 0, false_rjs / rjs) # False Discovery Proportion
+  power <- ifelse(sum(sim$H) == 0, NA, 
+                  sum(sim$H == 1 & rejected) / sum(sim$H == 1)) # power
+  FPR <- sum(sim$H == 0 & rejected) / sum(sim$H == 0)
+ 	FWER <- as.numeric(false_rjs > 0)
+	df <- data.frame(n_rejects = rjs, FWER=FWER, FPR=FPR, 
+	                 FDP=FDP, alpha = alpha) #  power=power, 
+	return(df)
+ }
 
@@
 
+Simulate an example "null" dataset. `t_stat` is the t-test statistic, `t_dm` is the difference in means, 
+`t_pval` is the p-value, and `sd` is the pooled standard deviation. 
+```{r example-simulate-null-data}
+# set these to a smaller number temporarily 
+m = 5000 # number of tests
+M = 500 # number of MC reps
+
+n_samples = 20
+n_groups = 2
+
+sim_df <- du_ttest_sim(m=m, pi0=1, effect_size=0, n_samples=n_samples, 
+                       n_groups = 2)
+head(sim_df)
+```
+
+```{r evaluate-methods-null-data}
+alphas = seq(0.01, 0.1, length.out = 10)
+
+nCores <- 4
+registerDoParallel(cores = nCores)
+workers <- getDoParWorkers()
+backend <- getDoParName()
+version <- getDoParVersion()
+
+df.out <- foreach(i = 1:M) %dopar% 
+    {        
+        sim_df <- du_ttest_sim(m=m, pi0=1, effect_size=0, n_samples=n_samples, n_groups=2)
+
+        # Bonferroni 
+        adj_p <- p.adjust(sim_df$t_pval, method="bonferroni")
+        df.bonf <- plyr::ldply(alphas, function(a){
+                      calculate_test_stats(sim=sim_df, adj_p = adj_p, alpha = a) })
+        # BH
+        adj_p <- p.adjust(sim_df$t_pval, method="BH")
+        df.bh <- plyr::ldply(alphas, function(a){
+                      calculate_test_stats(sim=sim_df, adj_p = adj_p, alpha = a) })
+
+        # qvalue (Storey)
+        adj_p <- qvalue::qvalue(p=sim_df$t_pval)$qvalues
+        df.qvalue <- plyr::ldply(alphas, function(a){
+                      calculate_test_stats(sim=sim_df, adj_p = adj_p, alpha = a) })
+
+        # IHW (Wolfgang)
+        ihw_fdr <- ihw(pvalues = sim_df$t_pval, covariates = sim_df$sd, alpha =  0.1)
+        adj_p <- adj_pvalues(ihw_fdr)
+        df.ihw <- plyr::ldply(alphas, function(a){
+                      calculate_test_stats(sim=sim_df, adj_p = adj_p, alpha = a) })
+
+        # ash (Stephens)
+        beta.ash = ash(betahat = sim_df$t_dm, sebetahat = sim_df$SE)
+        adj_p <- get_svalue(beta.ash)
+        df.ash <- plyr::ldply(alphas, function(a){
+                      calculate_test_stats(sim=sim_df, adj_p = adj_p, alpha = a) })
+
+        # swfdr (Boca-Leek)
+
+        # locfdr (Cai)
+
+        # Scott
+
+        # Summary table of FDP for each method at each alpha
+        rbind(data.frame(df.bonf, "method" = "bonferroni"), 
+              data.frame(df.bh, "method" = "bh"), 
+              data.frame(df.qvalue, "method" = "qvalue"), 
+              data.frame(df.ihw, "method" = "ihw"), 
+              data.frame(df.ash, "method" = "ash"))
+  }
+
+df.out <- dplyr::bind_rows(df.out)
+```
+
+Number of discoveries as a function of $\alpha$
+```{r}
+df.out %>% dplyr::group_by(method, alpha) %>% 
+      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
+                       FWER = mean(FWER), FPR = mean(FPR)) %>%  
+    ggplot(aes(x = alpha, y = n_rejects, color = method)) + 
+         geom_line() + geom_abline(linetype="dashed") + 
+         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("Num of Discoveries") +
+         theme(axis.title = element_text(face="bold") )
+
+```
+
+False discovery rate as a function of $\alpha$
+```{r}
+df.out %>% dplyr::group_by(method, alpha) %>% 
+      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
+                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
+      ggplot(aes(x = alpha, y = FDR, color = method)) + 
+         geom_line() + geom_abline(linetype="dashed") + 
+         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FDR") +
+         theme(axis.title = element_text(face="bold") )
+
+```
+
+Family wise error rate as a function of $\alpha$
+```{r}
+df.out %>% dplyr::group_by(method, alpha) %>% 
+      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
+                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
+      ggplot(aes(x = alpha, y = FWER, color = method)) + 
+         geom_line() + geom_abline(linetype="dashed") + 
+         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FWER") +
+         theme(axis.title = element_text(face="bold") )
+```
+
 ## add here...

### 2017-09-13 datasets/GWAS/GWAS.Rmd
add chronotype dataset

@@
 
-*p-value
-*effect size
-*standard error
-*additional covariates (optional)
+- p-value
+- effect size
+- standard error
+- additional covariates (optional)
 
-Note that while knitting this document will download the data, I'm not committing 
-the data itself to this repo. See the knitted `.html` document if you do not 
+Note that while running all the bash code in this document will download the data, 
+some of it doesn't seem to work properly with knitting (e.g. the ftp download 
+in the second dataset), and I'm not committing 
+the data itself to this repo. Run the bash code in the terminal separately to 
+download the data yourself, and see the knitted `.html` document if you do not 
 wish to download the datasets but still wish to see the output.
@@
 
-```{bash, GWAS1-download}
+```{bash, GWAS1-download, eval=FALSE}
 curl -O "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
@@
 
-```{bash, GWAS1-unzip}
+```{bash, GWAS1-unzip, eval=FALSE}
 unzip BMI.SNPadjSMK.zip
@@
 bmi <- fread("BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt", header=TRUE)
+dim(bmi)
 head(bmi)
@@
 
-- [x] p-value
-- [x] effect size
-- [x] standard error
-- [x] additional covariates: 
-  *`N`: Number of samples with this SNP - BMI association measured
-  *`Freq_Allele1_HapMapCEU`: Allele Frequency of Allele1, as measured in HapMapCEU
+- `p_value`:p-value
+- `effect`:effect size
+- `stderr`:standard error
+- additional covariates: 
+  - `N`: Number of samples with this SNP - BMI association measured
+  - `Freq_Allele1_HapMapCEU`: Allele Frequency of Allele1, as measured in HapMapCEU
+
+for 2,458,133 SNPs.
+
+```{r, GWAS1-cleanup}
+rm(bmi)
+```
+
+### GWAS for Self-reported Chronotype
+
+A [study by Hu et al.
+(Nature Comm, 2016)](https://www.ncbi.nlm.nih.gov/pubmed/?term=26835600) presented a GWAS of 89,283 individuals that identified genetic variants associated with self-reporting of being a morning person.
+
+Let's download the data from the `ebi.ac.uk` ftp site:
+
+```{bash, GWAS2-download, eval=FALSE}
+ftp ftp://ftp.ebi.ac.uk/pub/databases/gwas/summary_statistics/
+cd KhorCC_27064256_GCST003467
+get "Khor-27064256.txt"
+exit
+```
+Next, we'll read the `.txt` file into R and verify that it contains
+the necessary inputs to run the FDR benchmark comparisons.
+
+```{r, GWAS2-verify-contents}
+glauc <- fread("Khor-27064256.txt", header=TRUE)
+dim(glauc)
+head(glauc)
+```
+
+It looks like we have:
+
+- `pvalue`:p-value
+- `effect`:effect size
+- `stderr`:standard error
+- additional covariates: 
+  - `gt.rate`: ?
+  - `freq.b`: ?
+  - `dose.b`: ?
+
+for 10,000 SNPs - unfortunately, looks like they only give us the *top* 
+10K SNPs, so this is not useful for our purposes.
+
+```{r, GWAS2-cleanup}
+rm(chrono)
+```
+
+### GWAS for Asthma
+
+Let's download the data from the Gabriel site
+
+```{bash, GWAS2-download, eval=FALSE}
+curl -O "https://beaune.cng.fr/gabriel/gabriel_results.zip"
+```

### 2017-09-13 simulation_studies/fdr_simulation_studies.Rmd

--- 2017-09-11 simulation_studies/fdr_simulation_studies.Rmd
@@
 
-## Control for FDR at $\alpha$
+## Control for FDR and FWER at $\alpha$
 
-Evaluate methods control for FDR at nominal $\alpha$ level ranging from [0.01, 0.1].
+In this section, we will plot the number of discoveries 
+as a function of $\alpha$ (should be 0). We will also evaluate
+methods control for false discovery rate (FDR) and family-wise 
+error rate (FWER) at nominal $\alpha$ level ranging from [0.01, 0.1].
 
-```{r}
+```{r, load-libraries, message=FALSE}
+# library(genefilter)
+library(dplyr)
+library(ggplot2)
+library(doParallel)
+
+# library(IHW)
+# library(ashr)
+# library(qvalue)
+# library(swfdr)
+  ## to use IHWpaper::scott_fdrreg wrapper, we need to install FDRreg v2.0 (only available github):   
+  ## devtools::install_github(repo= 'jgscott/FDRreg', subdir='R_pkg/') 
+  ## however, I couldn't get this to install. So I did the v1.0 via install.packages("FDRreg") on CRAN
+# library(FDRreg)
+
 # notes: convert this into R code
-# - IHW: p-values from t-test, independent covariate
-# - ASH: difference in means, independent covariate
-# - BL: p-values from t-test, multiple covariates (with and without cov structure)
-#   - lambda: 0.75, 0.85, 0.95
 # - Scott: t-test statistic, independent covariate
 #   - assumption of test statistic follow a standard normal. this is good here. 
-# - BH: p-values only. 
 # - locfdr: zz = t-test statistic, nulltype (try both). 
 # - mixfdr: ? do we want to include? 
+```
+
+Additional packages that you should install are: 
+```{r, eval=FALSE}
+library(IHW)
+library(ashr)
+library(qvalue)
+library(swfdr)
+  ## to use IHWpaper::scott_fdrreg wrapper, we need to install FDRreg v2.0 (only available github):   
+  ## devtools::install_github(repo= 'jgscott/FDRreg', subdir='R_pkg/') 
+  ## however, I couldn't get this to install. So I did the v1.0 via install.packages("FDRreg") on CRAN
+library(FDRreg)
+```
+
+Load some helper functions: 
+```{r simulate-null-data-helper-functions}
+du_ttest_sim <- function(m, pi0, effect_size, n_samples=20, n_groups = 2, seed=NULL){
+  if (!is.null(seed)) set.seed(seed)
+    m0 <- ceiling(m*pi0)
+    false_nulls <- sample(1:m, m-m0)
+    z_table <- matrix(rnorm(n_samples*m), ncol=n_samples)
+    z_table[false_nulls,(n_samples/2+1):n_samples] <- matrix(rnorm(n_samples/2*(m-m0), effect_size), ncol=n_samples/2)
+    H <- rep(0,m)
+    H[false_nulls] <- 1
+    gF <- factor(rep(1:n_groups, each=n_samples/n_groups))
+    t_test <- genefilter::rowttests(z_table, gF)
+    sds <- genefilter::rowSds(z_table) # pooled var reduces to unpooled var b/c same sample sizes across groups 
+    SE <- sds / sqrt(n_samples)
+    simDf <- data.frame(H = H, t_stat = t_test$statistic, t_dm = t_test$dm, 
+                      t_pval = t_test$p.value, sd = sds, SE = SE)
+    return(simDf)
+}
+
+calculate_test_stats <- function(sim, adj_p, alpha){
+  rejected <- adj_p <= alpha
+  rjs <- sum(rejected) # number of rejected nulls
+  false_rjs <- sum(sim$H == 0 & rejected) # number of false rejected nulls
+  rj_ratio <- rjs/nrow(sim)
+  FDP <- ifelse(rjs == 0, 0, false_rjs / rjs) # False Discovery Proportion
+  power <- ifelse(sum(sim$H) == 0, NA, 
+                  sum(sim$H == 1 & rejected) / sum(sim$H == 1)) # power
+  FPR <- sum(sim$H == 0 & rejected) / sum(sim$H == 0)
+ 	FWER <- as.numeric(false_rjs > 0)
+	df <- data.frame(n_rejects = rjs, FWER=FWER, FPR=FPR, 
+	                 FDP=FDP, alpha = alpha) #  power=power, 
+	return(df)
+ }
+
+# adapted from IHWpaper::scott_fdrreg()
+scott_fdrreg_hickswrapper <- function(unadj_p, filterstat, df=3)
+  {
+	   # no automated way to choose function space over which we optimize
+	   # so we just use bs(df=3) as done in their analysis
+     b <- splines::bs(filterstat, df=df)
+
+     Xs <- model.matrix( ~  b - 1)
+     fdrreg_res <- FDRreg::FDRreg(z=qnorm(unadj_p), covars=Xs) # assumption of test statistic follow a standard normal
+     adj_p <- fdrreg_res$FDR
+     return(adj_p)
+  }
 
@@
 
+Simulate an example "null" dataset. `H` is an indicator representing the true null (0) or alternative status (1). `t_stat` is the t-test statistic, `t_dm` is the difference in means, 
+`t_pval` is the $p$-value, `sd` is the pooled standard deviation ($\sigma$), which reduces to the unpooled sd in this case, 
+and `SE` is the standard error ($\frac{\sigma}{\sqrt{n}}$)
+```{r example-simulate-null-data}
+n_samples = 20
+n_groups = 2
+
+sim_df <- du_ttest_sim(m=m, pi0=1, effect_size=0, 
+                       n_samples=n_samples, n_groups = 2)
+head(sim_df)
+```
+
+Run following on odyssey and save output as RDS: 
+```{r evaluate-methods-null-data, eval=FALSE}
+alphas = seq(0.01, 0.1, length.out = 10)
+
+nCores <- 4
+registerDoParallel(cores = nCores)
+workers <- getDoParWorkers()
+backend <- getDoParName()
+version <- getDoParVersion()
+
+df.out <- foreach(i = 1:M) %dopar% 
+    {        
+        sim_df <- du_ttest_sim(m=m, pi0=1, effect_size=0, n_samples=n_samples, n_groups=2)
+
+        # Bonferroni 
+        adj_p <- p.adjust(sim_df$t_pval, method="bonferroni")
+        df.bonf <- plyr::ldply(alphas, function(a){
+                      calculate_test_stats(sim=sim_df, adj_p = adj_p, alpha = a) })
+        # BH
+        adj_p <- p.adjust(sim_df$t_pval, method="BH")
+        df.bh <- plyr::ldply(alphas, function(a){
+                      calculate_test_stats(sim=sim_df, adj_p = adj_p, alpha = a) })
+
+        # qvalue (Storey)
+        adj_p <- qvalue::qvalue(p=sim_df$t_pval)$qvalues
+        df.qvalue <- plyr::ldply(alphas, function(a){
+                      calculate_test_stats(sim=sim_df, adj_p = adj_p, alpha = a) })
+
+        # IHW (Wolfgang)
+        ihw_fdr <- IHW::ihw(pvalues = sim_df$t_pval, covariates = sim_df$sd, alpha =  0.1)
+        adj_p <- IHW::adj_pvalues(ihw_fdr)
+        df.ihw <- plyr::ldply(alphas, function(a){
+                      calculate_test_stats(sim=sim_df, adj_p = adj_p, alpha = a) })
+
+        # ash (Stephens)
+        beta.ash = ashr::ash(betahat = sim_df$t_dm, sebetahat = sim_df$SE)
+        adj_p <- ashr::get_svalue(beta.ash)
+        df.ash <- plyr::ldply(alphas, function(a){
+                      calculate_test_stats(sim=sim_df, adj_p = adj_p, alpha = a) })
+
+        # swfdr::lm_pi0 (Boca-Leek) **think about multiple covariates** or **correlation structure**
+        pi0x <- swfdr::lm_pi0(pValues = sim_df$t_pval, X = sim_df$sd, smooth.df =3)
+        adj_p <- p.adjust(sim_df$t_pval, method="BH")*pi0x$pi0
+        df.BL <- plyr::ldply(alphas, function(a){
+                      calculate_test_stats(sim=sim_df, adj_p = adj_p, alpha = a) })
+
+        # Scott et al. (2015) (FDRreg)
+        adj_p <- scott_fdrreg_hickswrapper(unadj_p = sim_df$t_pval, filterstat = sim_df$sd, df=3)
+        df.scott <- plyr::ldply(alphas, function(a){
+                      calculate_test_stats(sim=sim_df, adj_p = adj_p, alpha = a) })
+
+        # locfdr (Cai)
+
+        # Summary table of FDP for each method at each alpha
+        rbind(data.frame(df.bonf, "method" = "bonferroni"), 
+              data.frame(df.bh, "method" = "bh"), 
+              data.frame(df.qvalue, "method" = "qvalue"), 
+              data.frame(df.ihw, "method" = "ihw"), 
+              data.frame(df.ash, "method" = "ash"), 
+              data.frame(df.BL, "method" = "boca-leek"), 
+              data.frame(df.scott, "method" = "scott"))
+  }
+
+df.out <- dplyr::bind_rows(df.out)
+saveRDS(df.out, file=file.path(workingPath, "nullSims-Results.RDS"))
+```
+
+Load and plot results from RDS file: 
+```{r, eval=FALSE}
+workingPath <- "/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/shicks/projects/benchmark-fdr"
+readRDS(file=file.path(workingPath, "nullSims-Results.RDS"))
+
+# Number of discoveries as a function of $\alpha$
+df.out %>% dplyr::group_by(method, alpha) %>% 
+      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
+                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
+    ggplot(aes(x = alpha, y = n_rejects, color = method)) + 
+         geom_line() + geom_abline(linetype="dashed") + 
+         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("Number of Discoveries") +
+         theme(axis.title = element_text(face="bold") )
+
+# False discovery rate as a function of $\alpha$
+df.out %>% dplyr::group_by(method, alpha) %>% 
+      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
+                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
+      ggplot(aes(x = alpha, y = FDR, color = method)) + 
+         geom_line() + geom_abline(linetype="dashed") + 
+         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FDR") +
+         theme(axis.title = element_text(face="bold") )
+
+# Family wise error rate as a function of $\alpha$
+df.out %>% dplyr::group_by(method, alpha) %>% 
+      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
+                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
+      ggplot(aes(x = alpha, y = FWER, color = method)) + 
+         geom_line() + geom_abline(linetype="dashed") + 
+         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FWER") +
+         theme(axis.title = element_text(face="bold") )
+```
+
 ## add here...

### 2017-09-13 datasets/GWAS/GWAS.Rmd
remove chrono study; add asthma study

@@
 
-### GWAS for Self-reported Chronotype
+### GWAS for Asthma
 
-A [study by Hu et al.
-(Nature Comm, 2016)](https://www.ncbi.nlm.nih.gov/pubmed/?term=26835600) presented a GWAS of 89,283 individuals that identified genetic variants associated with self-reporting of being a morning person.
+[This study by Moffatt et al. (2010, NEJM)](http://www.nejm.org/doi/full/10.1056/NEJMoa0906312?query=TOC) performed a large
+GWAS on  10,365 people with asthma and 16,110 controls 23 individual studies.
 
-Let's download the data from the `ebi.ac.uk` ftp site:
+Let's download the data from the Gabriel site. The second file contains metadata
+for the columns in the main data file.
 
 ```{bash, GWAS2-download, eval=FALSE}
-ftp ftp://ftp.ebi.ac.uk/pub/databases/gwas/summary_statistics/
-cd KhorCC_27064256_GCST003467
-get "Khor-27064256.txt"
-exit
+curl -O "https://beaune.cng.fr/gabriel/gabriel_results.zip"
+curl -O "https://beaune.cng.fr/gabriel/gabriel_results_description.xls"
 ```
-Next, we'll read the `.txt` file into R and verify that it contains
+
+Next we unzip the file and remove the zipped copy.
+
+```{bash, GWAS2-unzip, eval=FALSE}
+unzip gabriel_results.zip
+rm gabriel_results.zip
+```
+
+Next, we'll read in the unzipped `.txt` file and the results desription `.xls` 
+file into R and verify that it contains
 the necessary inputs to run the FDR benchmark comparisons.
@@
 ```{r, GWAS2-verify-contents}
-glauc <- fread("Khor-27064256.txt", header=TRUE)
-dim(glauc)
-head(glauc)
+# read in meta-data
+library(readxl)
+meta <- read_xls("gabriel_results_description.xls")
+head(meta)
+dim(meta)
+
+# read in data
+asthma <- fread("gabriel_asthma_meta-analysis_36studies_format_repository_NEJM.txt", header=TRUE)
+dim(asthma)
+
+# verify meta-data rows = data cols
+nrow(meta) == ncol(asthma)
+
+# remove study-specific columns
+library(dplyr)
+empty <- which(!is.na(meta$description))
+asthma <- asthma %>%
+            select(empty)
+meta <- meta %>% 
+            slice(empty)
+head(asthma)
 ```
@@
 
-- `pvalue`:p-value
-- `effect`:effect size
-- `stderr`:standard error
+- `P_fix` and `P_ran`:p-values for fixed and random effect models
+- `theta_fix` and `theta_fix`:effect size for fixed and random effect models
+- `se_theta_fix` and `se_theta_fix`:standard error for fixed and random effect models
 - additional covariates: 
-  - `gt.rate`: ?
-  - `freq.b`: ?
-  - `dose.b`: ?
+    - `no_studies_meta`: Number of studies included in the meta-analysis with this SNP
+    - `freq_all_1_max` and `freq_all_1_min`: Maximum and Minimum Allele Frequency of Allele 1 across samples
 
-for 10,000 SNPs - unfortunately, looks like they only give us the *top* 
-10K SNPs, so this is not useful for our purposes.
+for 567,589 SNPs.
 
 ```{r, GWAS2-cleanup}
-rm(chrono)
+rm(asthma)
 ```
-
-### GWAS for Asthma
-
-Let's download the data from the Gabriel site
-
-```{bash, GWAS2-download, eval=FALSE}
-curl -O "https://beaune.cng.fr/gabriel/gabriel_results.zip"
-```

### 2017-09-13 datasets/GWAS/GWAS.Rmd
add triglyceride gwas

@@
 ```
+
+### GWAS for Lipid Levels
+
+[This study by Willer et al. (2013, Nature Genetics)](http://www.nature.com/ng/journal/v45/n11/full/ng.2797.html) performed a large
+GWAS on 188,577 and measured blood lipid levels (as well as several cardiovascular traits).
+
+Let's download the data from the University of Michigan site. The download page 
+has a handy [metadata table](http://csg.sph.umich.edu//abecasis/public/lipids2013/)
+that describes the contents of the columns in the results files.
+
+There are several types of results files to download (in terms of which phenotype to analyze):
+
+- LDL Cholesterol
+- HDL Cholesterol
+- Triglycerides
+- Total Cholesterol
+
+As well as two types of analyses:
+
+- Joint analyses of GWAS and Metabochip
+- analysis of Metabochip
+
+We would like to look at the joint analysis of GWAS + Metabochip since we'd like 
+to look at all available SNPs (not just those that have previously shown a possible
+association with CVD). We could choose any or all of the traits, but I'll go ahead
+and proceed with Triglycerides since I'm curious about the role of genetic variation 
+on Triglyceride levels.
+
+```{bash, GWAS3-download, eval=FALSE}
+curl -O "http://csg.sph.umich.edu/abecasis/public/lipids2013/jointGwasMc_TG.txt.gz"
+```
+
+Next we uncompress the file.
+
+```{bash, GWAS3-unzip, eval=FALSE}
+gunzip jointGwasMc_TG.txt.gz 
+```
+
+Next, we'll read in the uncompressed `.txt` file into R and verify that it contains
+the necessary inputs to run the FDR benchmark comparisons.
+
+```{r, GWAS3-verify-contents}
+# read in data
+trig <- fread("jointGwasMc_TG.txt", header=TRUE)
+dim(trig)
+head(trig)
+```
+
+It looks like we have:
+
+- `P_value`:p-value
+- `beta`:effect size
+- `SE`:standard error
+- additional covariates: 
+    - `N`: The number of individuals analyzed for this marker.
+    - `Freq.A1.1000G.EUR`: Frequency of allele A1 from 1000G EUR sample.
+
+for 2,439,432 SNPs.
+
+```{r, GWAS3-cleanup}
+rm(trig)
+```

### 2017-09-14 simulation_studies/fdr_simulation_studies.Rmd
overhauled null simulations using /R/simulation-helpers.R; made du_ttest_sim() output generic for any test statistic, effect size, p value, and independent covariate

--- 2017-09-11 simulation_studies/fdr_simulation_studies.Rmd
@@
 ```{r}
-m = 20000 # number of tests
-M = 1000 # number of MC reps
+m = 5000 # number of tests
+M = 4 # number of MC reps
+
+ # load helper functions
+RPROJ <- list(PROJHOME = normalizePath(getwd()))
+source(file.path(RPROJ$PROJHOME, "R/simulation-helpers.R"))
 ```
@@
 
-## Control for FDR at $\alpha$
+## Control for FDR and FWER at $\alpha$
 
-Evaluate methods control for FDR at nominal $\alpha$ level ranging from [0.01, 0.1].
+In this section, we will plot the number of discoveries 
+as a function of $\alpha$ (should be 0). We will also evaluate
+methods control for false discovery rate (FDR) and family-wise 
+error rate (FWER) at nominal $\alpha$ level ranging from [0.01, 0.1].
 
-```{r}
-# notes: convert this into R code
-# - IHW: p-values from t-test, independent covariate
-# - ASH: difference in means, independent covariate
-# - BL: p-values from t-test, multiple covariates (with and without cov structure)
-#   - lambda: 0.75, 0.85, 0.95
-# - Scott: t-test statistic, independent covariate
-#   - assumption of test statistic follow a standard normal. this is good here. 
-# - BH: p-values only. 
-# - locfdr: zz = t-test statistic, nulltype (try both). 
-# - mixfdr: ? do we want to include?
+```{r, load-libraries, message=FALSE}
+library(dplyr)
+library(ggplot2)
+library(doParallel)
+```
 
+Additional packages that you should install are: 
+```{r, eval=FALSE}
+library(genefilter)
+
+library(IHW)
+library(ashr)
+library(qvalue)
+library(swfdr)
+library(fdrtool)
+  ## to use IHWpaper::scott_fdrreg wrapper, need to install FDRreg v2.0 (only available github):   
+  ## devtools::install_github(repo= 'jgscott/FDRreg', subdir='R_pkg/') 
+library(FDRreg)
+```
+
+Simulate an example "null" dataset. `H` is an indicator representing the true null (0) or alternative status (1). `test_statistic` is the t-test statistic, `effect_size` is the difference in means, 
+`pval` is the $p$-value from the t-test, `ind_covariate` (in this case) is the pooled standard deviation ($\sigma$), which reduces to the unpooled sd in this case, 
+and `SE` is the standard error ($\frac{\sigma}{\sqrt{n}}$)
+```{r example-simulate-null-data}
+n_samples = 20
+n_groups = 2
+
+sim_df <- du_ttest_sim(m=m, pi0=1, effect_size=0, 
+                       n_samples=n_samples, n_groups = 2)
+head(sim_df)
+```
+
+Run following on odyssey and save output as RDS: 
+```{r evaluate-methods-null-data, eval=FALSE}
+library(dplyr)
+library(ggplot2)
+library(doParallel)
+
+workingPath <- "/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/shicks/projects/benchmark-fdr"
+source(file.path(workingPath, "simulation-helpers.R"))
+
+# set these to a smaller number temporarily 
+m = 5000 # number of tests
+M = 500 # number of MC reps
+
+n_samples = 20
+n_groups = 2
+
+alphas = seq(0.01, 0.1, length.out = 10)
+
+nCores <- 10
+registerDoParallel(cores = nCores)
+workers <- getDoParWorkers()
+backend <- getDoParName()
+version <- getDoParVersion()
+
+df.out <- foreach(i = 1:M, .verbose=T) %dopar% {
+  sim_df <- du_ttest_sim(m=m, pi0=1, effect_size=0, n_samples=n_samples, n_groups=2)
+  sim_runner(sim = sim_df, alphas = alphas)
+}
+
+df.out <- dplyr::bind_rows(df.out)
+
+saveRDS(df.out, file=file.path(workingPath, "nullSims-Results.RDS"))
+```
+
+Load and plot results from RDS file: 
+```{r, eval=FALSE}
+# workingPath <- "/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/shicks/projects/benchmark-fdr"
+# df.out <- readRDS(file=file.path(workingPath, "nullSims-Results.RDS"))
+df.out <- readRDS(file=file.path(RPROJ$PROJHOME, "R/nullSims-Results.RDS"))
+
+# Number of discoveries as a function of $\alpha$
+df.out %>% dplyr::group_by(method, alpha) %>% 
+      dplyr::filter(!(method %in% c("ash", "scott", "locfdr"))) %>%
+      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
+                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
+    ggplot(aes(x = alpha, y = n_rejects, color = method)) + 
+         geom_line() + geom_abline(linetype="dashed") + 
+         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("Number of Discoveries") +
+         theme(axis.title = element_text(face="bold") )
+
+# False discovery rate as a function of $\alpha$
+df.out %>% dplyr::group_by(method, alpha) %>%
+        # dplyr::filter(!(method %in% c("ash", "scott", "locfdr"))) %>%
+      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
+                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
+      ggplot(aes(x = alpha, y = FDR, color = method)) + 
+         geom_line() + geom_abline(linetype="dashed") + 
+         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FDR") +
+         theme(axis.title = element_text(face="bold") )
+
+# Family wise error rate as a function of $\alpha$
+df.out %>% dplyr::group_by(method, alpha) %>% 
+         #  dplyr::filter(!(method %in% c("ash", "scott"))) %>%
+      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
+                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
+      ggplot(aes(x = alpha, y = FWER, color = method)) + 
+         geom_line() + geom_abline(linetype="dashed") + 
+         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FWER") +
+         theme(axis.title = element_text(face="bold") )
 ```

### 2017-09-14 simulation_studies/R/simulation-helpers.R

@@ -14,11 +14,11 @@
 #' @return
 #' data.frame of test results for `m` simulated data sets, with columns:
 #' * `H`: 0/1 indicator whether data simulated under null (0) or alternative (1)
-#' * `t_stat`: t-test statistic
-#' * `t_dm`: group mean difference
-#' * `t_pval`: t-test p-value
-#' * `sd`: standard deviation across all samples
-#' * `SE`: standard error across all samples
+#' * `test_statistic`: t-test statistic
+#' * `effect_size`: group mean difference
+#' * `pval`: t-test p-value
+#' * `ind_covariate`: in this case, it's standard deviation (sd) across all samples 
+#' * `SE`: standard error across all samples (sd / sqrt(n))
 #'
 #' @md
 #' @importFrom genefilter rowtests
@@ -41,8 +41,8 @@ du_ttest_sim <- function(m, pi0, effect_size, n_samples = 20, n_groups = 2, seed
     t_test <- genefilter::rowttests(z_table, gF)
     sds <- genefilter::rowSds(z_table) # pooled var reduces to unpooled var b/c same sample sizes across groups 
     SE <- sds / sqrt(n_samples)
-    simDf <- data.frame(H = H, t_stat = t_test$statistic, t_dm = t_test$dm, 
-                        t_pval = t_test$p.value, sd = sds, SE = SE)
+    simDf <- data.frame(H = H, test_statistic = t_test$statistic, effect_size = t_test$dm, 
+                        pval = t_test$p.value, ind_covariate = sds, SE = SE)
     return(simDf)
 }
 
@@ -67,6 +67,7 @@ du_ttest_sim <- function(m, pi0, effect_size, n_samples = 20, n_groups = 2, seed
 #' * `FWER`: family-wise error rate
 #' * `FPR`: false positive rate
 #' * `FDP`: false discovery proportion
+#' * `power`: power (only returned if at least one H = 1)
 #' * `alpha`: alpha input parameter
 #'
 #' @md
@@ -84,8 +85,13 @@ calculate_test_stats <- function(sim, adj_p, alpha) {
                     sum(sim$H == 1 & rejected) / sum(sim$H == 1)) # power
     FPR <- sum(sim$H == 0 & rejected) / sum(sim$H == 0)
     FWER <- as.numeric(false_rjs > 0)
[…]

### 2017-09-15 simulation_studies/fdr_sims_alt.Rmd
Draft setup for analysis of single sim

---
title: "Benchmarking FDR Methods - Alternative Settings"
author: "Rafalab Journal Club Members"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    highlight: tango
    number_sections: true
    code_folding: hide
---

<!-- ######################################################################## -->
<!-- preliminaries                                                            --> 
<!-- ######################################################################## -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, install-benchmarked-methods, eval=FALSE}
## from CRAN
install.packages("ashr")

## from Bioconductor
BiocInstaller::biocLite("IHW")
BiocInstaller::biocLite("qvalue")
BiocInstaller::biocLite("swfdr")

## from GitHub
devtools::install_github('jgscott/FDRreg', subdir="R_pkg/")
devtools::install_github("nignatiadis/IHWpaper")
```

<!-- ######################################################################## -->
# Workspace Setup
<!-- ######################################################################## -->

```{r, load-libraries, message=FALSE}
library("genefilter")
library("dplyr")
library("ggplot2")
library("doParallel")

library("IHW")
library("ashr")
library("qvalue")
library("swfdr")
library("FDRreg")
library("IHWpaper")
```

```{r, load-helper-functions}
source("R/simulation-helpers.R")
```

```{r, set-location}
projdir <- "/n/irizarryfs01_backed_up/pkimes/project/project-02-benchmark-fdr"
```

<!-- ######################################################################## -->
# Overview
<!-- ######################################################################## -->

- Alternative Setttings
- Single Monte Carlo Simulations

In the alternative simulations, we consider:

- different magnitudes of effect sizes / difference in means
- different shapes of effect sizes / difference in means 
  - independent tests: (Boca-Leek): beta, normal, t, chi2; (ASH): spiky, near-normal, flat-top, skew, bimodal
  - correlated tests: (Boca-Leek): mutivar-normal, multivar-t (blocked covariance design)
- different sample sizes (e.g. impact on SE estimates)
  - same sample size for all tests
  - different sample sizes for some tests (i.e. for tests w/ different precision levels)
- different in relationship between pi_0 and independent covariate
- different correlation structure of multiple covariates
- different global $\pi_0$ values (between 0, 1) - setting 1 in Boca-Leek

```{r, define-base}
## setting 0: (base) null simulation setting
setting_base <- list(m = 2000, pi0 = 1, effect_size = 0, n_samples = 20, n_groups = 2)
```

<!-- ######################################################################## -->
# Alternative Setting 1: Varying Null Proportion ($\pi_0$)
<!-- ######################################################################## -->

First we define the set of simulation settings to be tested.

```{r, define-setting-1}
## create settings w/ varying pi0 (null proportions)
setting_alts <- lapply(seq(0, 1, by=.1),
                       function(p) { replace(setting_base,
                                             c("pi0", "effect_size"),
                                             c(p, 1.0))
                       })
names(setting_alts) <- paste0("altp0_", 0:10)

## verify settings are as expected
do.call(rbind, setting_alts)
```

```{r, run-setting-1}
## simulate single datasets for each setting
sim_dfl <- lapply(setting_alts, do.call, what=du_ttest_sim)

## calculate adjusted p-values and return pvals along w/ mets
adj_dfl <- lapply(sim_dfl, sim_runner, alphas=seq(0.01, 0.10, by=.01), pvals=TRUE)

## create 
sim_res <- cbind(sim_dfl$alteff_20, bb$alteff_20)
```

<!-- ------------------------------------------------------------------------ -->
## Visual Inspection of Single Monte Carlo Simulation
<!-- ------------------------------------------------------------------------ -->

```{r, ASH-svals-vs-qvals}
## plot ASH s-values against ASH q-values
gp <- ggplot(sim_res) +
    geom_point(aes(ashs, ashq), alpha=1/4) +
    expand_limits(x=c(0, 1), y=c(0, 1)) +
    theme_bw()
ggsave("plots/sims-01-null-ash_s_vs_q.png", gp, width=5, height=5)
```

```{r, effect-vs-adjp}
## base of plot for: 'effect size' vs. 'adjusted pval (method X)'
gp_base <- ggplot(sim_res, aes(x=effect_size, 
                               color=cut_width(SE, width=.05, boundary=0))) +
    expand_limits(y=c(0, 1)) + 
    scale_color_brewer("Binned StdErr", palette="Set1") + 
    theme_bw()

## 'effect size' vs. t-test p-values
gp0 <- gp_base + 
    geom_point(aes(y=pval), alpha=1/2) +
    ggtitle("dist. of unadjusted p-values (20k null tests)")

## 'effect size' vs. ASH adjusted p-values
gp1 <- gp_base + 
    geom_point(aes(y=ashs), alpha=1/2) +
    ggtitle("dist. of ASH s-values (20k null tests)")

## 'effect size' vs. IHW adjusted p-values
gp3 <- gp_base +
    geom_point(aes(y=ihw), alpha=1/2) + 
    ggtitle("dist. of IHW adjusted p-values (20k null tests)")

## 'effect size' vs. Boca-Leek adjusted p-values
gp4 <- gp_base +
    geom_point(aes(y=bl), alpha=1/2) +
    ggtitle("dist. of Boca-Leek adjusted p-values (20% alt tests; e.size=2)")

## plot together
png("plots/sims-01-null-ash.png", units="in", width=6, height=7, res=200)
pkimes::ggmulti(gp1, gp2, gp3, gp4, cols=2)
dev.off()
```

```{r}
gp <- ggplot(sim_res) +
    geom_point(aes(x=pval, y=ashs), alpha=1/10) +
    theme_bw() +
    xlab("unadjusted p-value") +
    ylab("ASH s-value")
ggsave("plots/sims-01-null-pval_vs_ash.png", gp, width=5, height=5)
```

```{r}
## distribution of rejected/non-rejected; sort reverse 
ggplot(sim_res) +
    geom_histogram(aes(x=qnorm(pval)*sign(test_statistic),
                       fill=cut_width(ashs, width=.05, boundary=0) %>%
                           factor(., levels=rev(levels(.)))),
                   color='black', binwidth=.1, boundary=0,
             position="stack") +
    scale_fill_brewer("ASH (s-values)", palette="Blues", direction=-1) +
    theme_bw()
```

<!-- ------------------------------------------------------------------------ -->
## Numerically Evaluate of Single Monte Carlo Simulation
<!-- ------------------------------------------------------------------------ -->

Analysis of metrics calculated for various FDR cutoffs.

<!-- ######################################################################## -->
# Alternative Setting 2: Varying Effect Size
<!-- ######################################################################## -->

First we define the set of simulation settings to be tested.

```{r}
## create settings w/ varying effect_size (mean difference)
setting_alts <- lapply(seq(0, 2, by=.2),
                       function(x) { replace(setting_base,
                                             c("pi0", "effect_size"),
                                             c(0.8, x))
                       })
names(setting_alts) <- paste0("alteff_", seq(0, 20, by=2))

## verify settings are as expected
do.call(rbind, setting_alts)
```

<!-- ------------------------------------------------------------------------ -->
## Visual Inspection of Single Monte Carlo Simulation
<!-- ------------------------------------------------------------------------ -->

<!-- ------------------------------------------------------------------------ -->
## Numerically Evaluate of Single Monte Carlo Simulation
<!-- ------------------------------------------------------------------------ -->

<!-- ######################################################################## -->
# Alternative Setting 3:
<!-- ######################################################################## -->

### 2017-09-15 datasets/ChIP-seq/ChIP-seq.Rmd
ChIP-seq

---
title: "ChIP-seq datasets for FDR benchmarking"
author: "Mingxiang Teng"
date: "9/14/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Summary

For FDR benchmark using ChIP-seq datasets, we may focus on peak calling and
differential peaks. Peak calling is the most common way that people analyze 
ChIP-seq since this technique was invented, though some groups started to care
about differential peaks in some complex experiment designs. For peak calling,
peak callers always report effect size (they call "enrichment signal") 
and p-value, but standard error. I think it is too complicate if we dig into
their code for generating standard error. For differential peaks, classical
pipeline would call peaks first, then recruit existing algorithms (*e.g.* DESeq2)
to test the covariant of interest, in a way that all effect size, p-value and 
standard error can be generated. However, this pipeline probably gives 
very similar patterns (distributions) of effect size / p-value to those
based on differential gene expression analysis. Thus, differential peak 
analysis might not be useful in our FDR benchmark.

### Oestrogen receptor binding case

This dataset have plenty of replicates so that we can focus on both peak
calling and differential peaks. And it is the first dataset people used to 
illustrate differential peak detection, where the tools DiffBind was 
introduced.

We can download the processed data from GEO.

```{bash, chipseq-download, eval=FALSE}
curl -O "https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE32222&format=file"
tar -xvf "?acc=GSE32222&format=file"
rm "?acc=GSE32222&format=file"
```

For one sample, extract the peak calling results.

```{bash, chipseq-unzip1, eval=FALSE}
gunzip "GSM798383_SLX-1201.250.s_4_SLX-1202.250.s_1_peaks.txt.gz"
rm *peaks.txt.gz
```

And load in to R.
```{r, chipseq-read1}
peaks <- read.table("~/GSM798383_SLX-1201.250.s_4_SLX-1202.250.s_1_peaks.txt", header=TRUE, sep='\t')
dim(peaks)
head(peaks)
```

Here, "X.10.log10.pvalue." is the -log10P, and fold_enrichment is the effect size.

### 2017-09-15 simulation_studies/fdr_sims_alt.Rmd
Add better plots

@@
 <!-- ######################################################################## -->
-# Workspace Setup
+## Workspace Setup
 <!-- ######################################################################## -->
@@
 library("dplyr")
+library("data.table")
 library("ggplot2")
-library("doParallel")
 
@@
 <!-- ######################################################################## -->
-# Overview
+## Overview
 <!-- ######################################################################## -->
 
-- Alternative Setttings
-- Single Monte Carlo Simulations
+We first investigate a broad range of settings with only a **single Monte Carlo replication** under each condition.
+In each simulation, a large set of univariate two-sample $t$-tests are carried out based on raw data sampled from two mean-shifted Gaussian distributions.
+Each method for controlling FDR is applied to the results of the $t$-tests, and performance is compared using metrics include: FPR and power.
 
-In the alternative simulations, we consider:
-
-- different magnitudes of effect sizes / difference in means
-- different shapes of effect sizes / difference in means 
-  - independent tests: (Boca-Leek): beta, normal, t, chi2; (ASH): spiky, near-normal, flat-top, skew, bimodal
-  - correlated tests: (Boca-Leek): mutivar-normal, multivar-t (blocked covariance design)
-- different sample sizes (e.g. impact on SE estimates)
-  - same sample size for all tests
-  - different sample sizes for some tests (i.e. for tests w/ different precision levels)
-- different in relationship between pi_0 and independent covariate
-- different correlation structure of multiple covariates
-- different global $\pi_0$ values (between 0, 1) - setting 1 in Boca-Leek
+A range of simulations are considered by varying the following baseline parameters.
 
@@
 ## setting 0: (base) null simulation setting
-setting_base <- list(m = 2000, pi0 = 1, effect_size = 0, n_samples = 20, n_groups = 2)
+setting_base <- list(m = 20000,         # number of hypothesis tests
+                     pi0 = 1,           # proportion of null hypotheses
+                     effect_size = 0,   # expected mean diff of non-null tests
+                     n_samples = 20,    # total number of samples
+                     n_groups = 2)      # number of groups in contrast
+```
+
+At a high level, we consider the following set of simulation settings:
+
+- **no informative covariate**
+  - *varying null proportion*
+  - *varying sample size*
+  - *varying effect size* (constant 'effect size')
+  - *varying effect size* (random 'effect size')
+    - independent tests: Beta, Normal, t, chi2, spiky, near-normal, flat-top, bimodal, skewed 
+    - dependent tests: multivariate-Normal, multivariate-t
+- **informative covariate**
+  - *varying null proportion*
+  - *varying sample size*
+  - *varying effect size* (constant 'effect size')
+  - *varying effect size* (random 'effect size')
+    - independent tests: Beta, Normal, t, chi2, spiky, near-normal, flat-top, bimodal, skewed 
+    - dependent tests: multivariate-Normal, multivariate-t
+  - *varying covariate, pi0 relation*
+    - single informative covariate
+    - multiple informative covariates
+
+In this analysis, we compare the following methods.
+For methods which accept user-specified covariates, under the "no informative covariate" settings, the pooled standard deviation was used specified to the methods as a dummy covariate.
+
+```{r, define-plot-variables}
+## string abb for each method (should match output of `sim_runner`)
+methods_str <- c("pval", "bh", "qvalue", "ihw",
+                 "ashs", "bl", "lfdr", "scott")
+
+## description of each method
+methods_desc <- list(pval = "unadjusted p-values",
+                     bh = "BH adjusted p-values",
+                     qvalue = "Storey's q-value",
+                     ihw = "IHW adjusted p-values",
+                     ashs = "ASH s-values",
+                     bl = "Boca-Leek adjusted p-values",
+                     lfdr = "local FDR",
+                     scott = "FDRreg adjusted p-values")
 ```
@@
 <!-- ######################################################################## -->
-# Alternative Setting 1: Varying Null Proportion ($\pi_0$)
+## Setting 1.1: No Informative Covariate, Varying Null Proportion ($\pi_0$)
 <!-- ######################################################################## -->
@@
 First we define the set of simulation settings to be tested.
+A range of $\pi_0$ values are considered between 0 and 1.
+For each setting, a fixed effect size (group mean differences) of `1.5` was used in the simulation of all non-null (alternative) tests.
 
@@
                                              c("pi0", "effect_size"),
-                                             c(p, 1.0))
+                                             c(p, 1.5))
                        })
@@
 ## verify settings are as expected
-do.call(rbind, setting_alts)
+ptab <- do.call(rbind.data.frame, setting_alts)
+ptab$setting <- rownames(ptab)
+setnames(ptab, "effect_size", "alt_effsize") 
+ptab[] <- lapply(ptab, as.character)
+ptab
 ```
 
+Next, we perform one iteration of simulation for each setting and keep both the adjusted p-values (`pvals`) and the evaluation metrics (`stats`) computed for each method.
+
 ```{r, run-setting-1}
-## simulate single datasets for each setting
+## simulate single datasets, calculate adj p-values
 sim_dfl <- lapply(setting_alts, do.call, what=du_ttest_sim)
-
-## calculate adjusted p-values and return pvals along w/ mets
 adj_dfl <- lapply(sim_dfl, sim_runner, alphas=seq(0.01, 0.10, by=.01), pvals=TRUE)
 
-## create 
-sim_res <- cbind(sim_dfl$alteff_20, bb$alteff_20)
+## combine results, settings for analysis
+sim_pv <- mapply(function(x, y) { cbind(x, y$pvals) }, sim_dfl, adj_dfl, SIMPLIFY=FALSE)
+sim_pv <- data.table::rbindlist(sim_pv, idcol="setting")
+sim_pv <- merge(sim_pv, ptab, by="setting", all.x=TRUE)
+
+sim_st <- lapply(adj_dfl, `[[`, "stats")
+sim_st <- data.table::rbindlist(sim_st, fill=TRUE, idcol="setting")
+sim_st <- merge(sim_st, ptab, by="setting", all.x=TRUE)
 ```
@@
 <!-- ------------------------------------------------------------------------ -->
-## Visual Inspection of Single Monte Carlo Simulation
+### Visual Inspection of Single Monte Carlo Simulation
 <!-- ------------------------------------------------------------------------ -->
 
-```{r, ASH-svals-vs-qvals}
-## plot ASH s-values against ASH q-values
-gp <- ggplot(sim_res) +
-    geom_point(aes(ashs, ashq), alpha=1/4) +
-    expand_limits(x=c(0, 1), y=c(0, 1)) +
-    theme_bw()
-ggsave("plots/sims-01-null-ash_s_vs_q.png", gp, width=5, height=5)
+All figures generated in this section will be saved with the following prefix.
+
+```{r, define-plot-prefix-1}
+plt_pfix <- "plots/sims-01_01-"
 ```
 
-```{r, effect-vs-adjp}
-## base of plot for: 'effect size' vs. 'adjusted pval (method X)'
-gp_base <- ggplot(sim_res, aes(x=effect_size, 
-                               color=cut_width(SE, width=.05, boundary=0))) +
+First, we take a look at the distribution of adjusted p-values as a function of the observed effect sizes over the `r setting_base$m` hypothesis tests.
+Each test is also colored by the corresponding standard error across the `r setting_base$n_samples` samples in the simulation.
+We plot this distribution across the values of pi0.
+
+```{r, plot-effect-vs-adjp, fig.width=14, fig.height=10}
+## set ID columns to keep when melting table for plot 
+plotvars <- c("effect_size", "SE", "pi0")
+pi0set <- c("0", "0.1", "0.3", "0.7", "0.9", "1")
+
+## generate plot for subset of pi0 values
+gp_all <- ggplot(sim_pv %>%
+                 melt(id.vars=plotvars, measure.vars=methods_str,
+                      variable.name='method', value.name='adjp') %>%
+                 filter(pi0 %in% pi0set) %>%
+                 mutate(method = plyr::revalue(method, unlist(methods_desc)),
+                        pi0 = paste("pi0 =", pi0))) + 
+    geom_point(aes(x=effect_size, y=adjp,
+                   color=cut_width(SE, width=.05, boundary=0)),
+               alpha=1/2) +
+    scale_color_brewer("Binned StdErr", palette="Set1") + 
     expand_limits(y=c(0, 1)) +
-    scale_color_brewer("Binned StdErr", palette="Set1") + 
+    ylab("adjusted p-value") + 
+    facet_grid(pi0 ~ method) + 
     theme_bw()
+gp_all
 
-## 'effect size' vs. t-test p-values
-gp0 <- gp_base + 
-    geom_point(aes(y=pval), alpha=1/2) +
-    ggtitle("dist. of unadjusted p-values (20k null tests)")
-
-## 'effect size' vs. ASH adjusted p-values
-gp1 <- gp_base + 
-    geom_point(aes(y=ashs), alpha=1/2) +
-    ggtitle("dist. of ASH s-values (20k null tests)")
-
-## 'effect size' vs. IHW adjusted p-values
-gp3 <- gp_base +
-    geom_point(aes(y=ihw), alpha=1/2) + 
-    ggtitle("dist. of IHW adjusted p-values (20k null tests)")
-
-## 'effect size' vs. Boca-Leek adjusted p-values
-gp4 <- gp_base +
-    geom_point(aes(y=bl), alpha=1/2) +
-    ggtitle("dist. of Boca-Leek adjusted p-values (20% alt tests; e.size=2)")
-
-## plot together
-png("plots/sims-01-null-ash.png", units="in", width=6, height=7, res=200)
-pkimes::ggmulti(gp1, gp2, gp3, gp4, cols=2)
-dev.off()
+## save to file
+ggsave(paste0(plt_pfix, "effectsize_vs_adjustedp.png"), gp_all, width=14, height=10)
 ```
 
-```{r}
-gp <- ggplot(sim_res) +
-    geom_point(aes(x=pval, y=ashs), alpha=1/10) +
-    theme_bw() +
-    xlab("unadjusted p-value") +
-    ylab("ASH s-value")
-ggsave("plots/sims-01-null-pval_vs_ash.png", gp, width=5, height=5)
+Next, we generate a similar plot, but look at the distribution of significant p-values as a function of the z-score, as was done in the original ASH paper.
+This was originally motivated by an interest in illustrating the effect of the unimodal assumption proposed with the ASH method.
+
+```{r, plot-zscale-vs-adjp, fig.width=14, fig.height=10}
+## set ID columns to keep when melting table for plot 
+plotvars <- c("effect_size", "SE", "pi0", "pval", "test_statistic")
+pi0set <- c("0", "0.1", "0.3", "0.7", "0.9", "1")
+
+## helper; bin adjusted p-vals, flip order of bins
+.p2disc <- function(x) {
+    cut(x, breaks=c(0, .01, .05, .10, .20, 1), include.lowest=TRUE) %>%
+        factor(., levels=rev(levels(.)))
+}
+
+## similar plot to Figure 1 of ASH paper
+gp_all <- ggplot(sim_pv %>%
+                 melt(id.vars=plotvars, measure.vars=methods_str,
+                      variable.name='method', value.name='adjp') %>%
+                 filter(pi0 %in% pi0set) %>%
+                 mutate(method = plyr::revalue(method, unlist(methods_desc)),
+                        pi0 = paste("pi0 =", pi0))) + 
+    geom_histogram(aes(fill=.p2disc(adjp),
+                       x=qnorm(pval)*sign(test_statistic)),
+                   color='black', size=1/4, binwidth=.2, boundary=0,
+                   position="stack") +
+    xlab("z-score") + 
+    viridis::scale_fill_viridis("Binned\nadj p-val", discrete=TRUE,
+                                option="plasma", direction=-1,
+                                begin=.1, end=.9, drop=FALSE) +
+    facet_grid(pi0 ~ method) + 
+    theme_bw()
+gp_all
+
+## save to file
+ggsave(paste0(plt_pfix, "zscore_vs_adjustedp.png"), gp_all, width=14, height=10)
 ```
 
-```{r}
-## distribution of rejected/non-rejected; sort reverse 
-ggplot(sim_res) +
-    geom_histogram(aes(x=qnorm(pval)*sign(test_statistic),
-                       fill=cut_width(ashs, width=.05, boundary=0) %>%
-                           factor(., levels=rev(levels(.)))),
-                   color='black', binwidth=.1, boundary=0,
-             position="stack") +
-    scale_fill_brewer("ASH (s-values)", palette="Blues", direction=-1) +
+A similar plot, is simply to look at the pairwise plots of original p-values against the adjusted p-values generated by each method
+
+```{r, plot-p-vs-adjp, fig.width=14, fig.height=10}
+## set ID columns to keep when melting table for plot 
+plotvars <- c("SE", "H", "pi0", "pval")
+pi0set <- c("0", "0.1", "0.3", "0.7", "0.9", "1")
+
+## color by SE
+gp_all <- ggplot(bb <- sim_pv %>%
+                 melt(id.vars=plotvars, measure.vars=methods_str,
+                      variable.name='method', value.name='adjp') %>%
+                 filter(pi0 %in% pi0set) %>%
+                 mutate(method = plyr::revalue(method, unlist(methods_desc)),
+                        pi0 = paste("pi0 =", pi0))) + 
+    geom_point(aes(x=pval, y=adjp,
+                   color=cut_width(SE, width=.05, boundary=0)),
+               alpha=1/10) +
+    scale_color_brewer("Binned StdErr", palette="Set1") + 
+    expand_limits(x=c(0, 1), y=c(0, 1)) +
+    xlab("unadjusted p-value") + 
+    ylab("adjusted p-value") + 
+    facet_grid(pi0 ~ method) + 
     theme_bw()
+gp_all
+
+## save to file
+ggsave(paste0(plt_pfix, "pvalue_vs_adjustedp_bySE.png"), gp_all, width=14, height=10)
+
+## color by H: 0/1
+gp_all <- ggplot(bb <- sim_pv %>%
+                 melt(id.vars=plotvars, measure.vars=methods_str,
+                      variable.name='method', value.name='adjp') %>%
+                 filter(pi0 %in% pi0set) %>%
+                 mutate(method = plyr::revalue(method, unlist(methods_desc)),
+                        pi0 = paste("pi0 =", pi0))) + 
+    geom_point(aes(x=pval, y=adjp, color=factor(H)), alpha=1/10) +
+    scale_color_brewer("null|alt", palette="Set1") + 
+    expand_limits(x=c(0, 1), y=c(0, 1)) +
+    xlab("unadjusted p-value") + 
+    ylab("adjusted p-value") + 
+    facet_grid(pi0 ~ method) + 
+    theme_bw()
+gp_all
+
+## save to file
+ggsave(paste0(plt_pfix, "pvalue_vs_adjustedp_byH.png"), gp_all, width=14, height=10)
+ggsave(paste0(plt_pfix, "pvalue_vs_adjustedp_byH_zoom.png"),
+       gp_all + coord_cartesian(x=c(0, .1), y=c(0, .1)),
+       width=14, height=10)
 ```
@@
 <!-- ------------------------------------------------------------------------ -->
-## Numerically Evaluate of Single Monte Carlo Simulation
+### Numeric Evaluate of Single Monte Carlo Simulation
 <!-- ------------------------------------------------------------------------ -->
 
-Analysis of metrics calculated for various FDR cutoffs.
+To be completed.
 
 <!-- ######################################################################## -->
-# Alternative Setting 2: Varying Effect Size
+## Setting 1.2: No Informative Covariate, Varying Effect Size (constant)
 <!-- ######################################################################## -->
@@
 First we define the set of simulation settings to be tested.
+A range of (constant) effect sizes are considered between 0 and 2.
+For each setting, a fixed null proportion of `pi0 = 0.2` was used.
+More complex settings with non-null effect sizes sampled from various distirbutions are considered in a later section.
 
-```{r}
+```{r, define-setting-2}
 ## create settings w/ varying effect_size (mean difference)
@@
 ## verify settings are as expected
-do.call(rbind, setting_alts)
+ptab <- do.call(rbind.data.frame, setting_alts)
+ptab$setting <- rownames(ptab)
+setnames(ptab, "effect_size", "alt_effsize") 
+ptab[] <- lapply(ptab, as.character)
+ptab
+```
+
+Next, we perform one iteration of simulation for each setting and keep both the adjusted p-values (`pvals`) and the evaluation metrics (`stats`) computed for each method.
+
+```{r, run-setting-2}
+## simulate single datasets, calculate adj p-values
+sim_dfl <- lapply(setting_alts, do.call, what=du_ttest_sim)
+adj_dfl <- lapply(sim_dfl, sim_runner, alphas=seq(0.01, 0.10, by=.01), pvals=TRUE)
+
+## combine results, settings for analysis
+sim_pv <- mapply(function(x, y) { cbind(x, y$pvals) }, sim_dfl, adj_dfl, SIMPLIFY=FALSE)
+sim_pv <- data.table::rbindlist(sim_pv, idcol="setting")
+sim_pv <- merge(sim_pv, ptab, by="setting", all.x=TRUE)
+
+sim_st <- lapply(adj_dfl, `[[`, "stats")
+sim_st <- data.table::rbindlist(sim_st, fill=TRUE, idcol="setting")
+sim_st <- merge(sim_st, ptab, by="setting", all.x=TRUE)
 ```
@@
 <!-- ------------------------------------------------------------------------ -->
-## Visual Inspection of Single Monte Carlo Simulation
+### Visual Inspection of Single Monte Carlo Simulation
 <!-- ------------------------------------------------------------------------ -->
 
+All figures generated in this section will be saved with the following prefix.
+
+```{r, define-plot-prefix-2}
+plt_pfix <- "plots/sims-01_02-"
+```
+
+First, we take a look at the distribution of adjusted p-values as a function of the observed effect sizes over the `r setting_base$m` hypothesis tests.
+Each test is also colored by the corresponding standard error across the `r setting_base$n_samples` samples in the simulation.
+We plot this distribution across the values of `effect_size`.
+
+```{r, plot-effect-vs-adjp-2, fig.width=14, fig.height=10}
+## set ID columns to keep when melting table for plot 
+plotvars <- c("effect_size", "H", "alt_effsize")
+es0set <- c("0.2", "0.4", "0.8", "1.2", "1.6", "2")
+
+## generate plot for subset of pi0 values
+gp_all <- ggplot(sim_pv %>%
+                 melt(id.vars=plotvars, measure.vars=methods_str,
+                      variable.name='method', value.name='adjp') %>%
+                 filter(alt_effsize %in% es0set) %>%
+                 mutate(method = plyr::revalue(method, unlist(methods_desc)),
+                        alt_effsize = paste("eff size =", alt_effsize))) + 
+    geom_point(aes(x=effect_size, y=adjp, color=factor(H)), alpha=1/10) +
+    scale_color_brewer("null|alt", palette="Set1") + 
+    expand_limits(y=c(0, 1)) +
+    ylab("p/q/s-value") + 
+    facet_grid(alt_effsize ~ method) + 
+    theme_bw()
+gp_all
+
+## save to file
+ggsave(paste0(plt_pfix, "effectsize_vs_adjustedp.png"), gp_all, width=14, height=10)
+```
+
+Next, we generate a similar plot, but look at the distribution of significant p-values as a function of the z-score, as was done in the original ASH paper.
+This was originally motivated by an interest in illustrating the effect of the unimodal assumption proposed with the ASH method.
+
+```{r, plot-zscale-vs-adjp-2, fig.width=14, fig.height=10}
+## set ID columns to keep when melting table for plot 
+plotvars <- c("H", "alt_effsize", "pval", "test_statistic")
+es0set <- c("0.2", "0.4", "0.8", "1.2", "1.6", "2")
+
+## helper; bin adjusted p-vals, flip order of bins
+.p2disc <- function(x) {
+    cut(x, breaks=c(0, .01, .05, .10, .20, 1), include.lowest=TRUE) %>%
+        factor(., levels=rev(levels(.)))
+}
+
+## similar plot to Figure 1 of ASH paper
+gp_all <- ggplot(sim_pv %>%
+                 melt(id.vars=plotvars, measure.vars=methods_str,
+                      variable.name='method', value.name='adjp') %>%
+                 filter(alt_effsize %in% es0set) %>%
+                 mutate(method = plyr::revalue(method, unlist(methods_desc)),
+                        alt_effsize = paste("eff size =", alt_effsize))) + 
+    geom_histogram(aes(fill=.p2disc(adjp),
+                       x=qnorm(pval)*sign(test_statistic)),
+                   color='black', size=1/4, binwidth=.2, boundary=0,
+                   position="stack") +
+    xlab("z-score") + 
+    viridis::scale_fill_viridis("Binned\nadj p-val", discrete=TRUE,
+                                option="plasma", direction=-1,
+                                begin=.1, end=.9, drop=FALSE) +
+    facet_grid(alt_effsize ~ method) + 
+    theme_bw()
+gp_all
+
+## save to file
+ggsave(paste0(plt_pfix, "zscore_vs_adjustedp.png"), gp_all, width=14, height=10)
+```
+
+A similar plot, is simply to look at the pairwise plots of original p-values against the adjusted p-values generated by each method
+
+```{r, plot-p-vs-adjp-1, fig.width=14, fig.height=10}
+## set ID columns to keep when melting table for plot 
+plotvars <- c("SE", "H", "alt_effsize", "pval")
+es0set <- c("0.2", "0.4", "0.8", "1.2", "1.6", "2")
+
+## color by SE
+gp_all <- ggplot(bb <- sim_pv %>%
+                 melt(id.vars=plotvars, measure.vars=methods_str,
+                      variable.name='method', value.name='adjp') %>%
+                 filter(alt_effsize %in% es0set) %>%
+                 mutate(method = plyr::revalue(method, unlist(methods_desc)),
+                        alt_effsize = paste("eff size =", alt_effsize))) + 
+    geom_point(aes(x=pval, y=adjp,
+                   color=cut_width(SE, width=.05, boundary=0)),
+               alpha=1/10) +
+    scale_color_brewer("Binned StdErr", palette="Set1") + 
+    expand_limits(x=c(0, 1), y=c(0, 1)) +
+    xlab("unadjusted p-value") + 
+    ylab("adjusted p-value") + 
+    facet_grid(alt_effsize ~ method) + 
+    theme_bw()
+gp_all
+
+## save to file
+ggsave(paste0(plt_pfix, "pvalue_vs_adjustedp_bySE.png"), gp_all, width=14, height=10)
+ggsave(paste0(plt_pfix, "pvalue_vs_adjustedp_bySE_zoom.png"),
+       gp_all + coord_cartesian(x=c(0, .1), y=c(0, .1)),
+       width=14, height=10)
+
+## color by H: 0/1
+gp_all <- ggplot(bb <- sim_pv %>%
+                 melt(id.vars=plotvars, measure.vars=methods_str,
+                      variable.name='method', value.name='adjp') %>%
+                 filter(alt_effsize %in% es0set) %>%
+                 mutate(method = plyr::revalue(method, unlist(methods_desc)),
+                        alt_effsize = paste("eff size =", alt_effsize))) + 
+    geom_point(aes(x=pval, y=adjp, color=factor(H)), alpha=1/10) +
+    scale_color_brewer("null|alt", palette="Set1") + 
+    expand_limits(x=c(0, 1), y=c(0, 1)) +
+    xlab("unadjusted p-value") + 
+    ylab("adjusted p-value") + 
+    facet_grid(alt_effsize ~ method) + 
+    theme_bw()
+gp_all
+
+## save to file
+ggsave(paste0(plt_pfix, "pvalue_vs_adjustedp_byH.png"), gp_all, width=14, height=10)
+ggsave(paste0(plt_pfix, "pvalue_vs_adjustedp_byH_zoom.png"),
+       gp_all + coord_cartesian(x=c(0, .1), y=c(0, .1)),
+       width=14, height=10)
+```
+
 <!-- ------------------------------------------------------------------------ -->
-## Numerically Evaluate of Single Monte Carlo Simulation
+### Numeric Evaluation of Single Monte Carlo Simulation
 <!-- ------------------------------------------------------------------------ -->
 
-<!-- ######################################################################## -->
-# Alternative Setting 3:
-<!-- ######################################################################## -->
+To be completed.

### 2017-09-20 datasets/GWAS/GWAS.Rmd
Add benchmarking evaluation for GWAS data

@@
 
+We'll also run the benchmark methods on each dataset and report the number of 
+discoveries for each one.
+
+### Workspace setup
+
+To run the benchmark methods on the GWAS datasets, we'll use some of the helper 
+and wrapper functions from the simulated data analysis. We also create and source an analgous version of the `sim_runner` function to apply to the real data.
+
+```{r, wkspace-setup}
+# Note that swfdr package depends on R version 3.4 or higher
+setwd("/n/irizarryfs01_backed_up/kkorthauer/benchmark-fdr/")
+source("simulation_studies/R/simulation-helpers.R")
+source("datasets/R/evaluation-helpers.R")
+setwd("./datasets/GWAS")
+
+# wrangling & plotting tools
+library(data.table)
+library(readxl)
+library(dplyr)
+library(ggplot2)
+
+# benchmark methods
+library(IHW)
+library(ashr)
+library(qvalue)
+library(swfdr)
+library(fdrtool)
+```
+
 ### GWAS meta-analysis for BMI (included in Boca-Leek)
 
-Here we download the GWAS BMI dataset analyzed in the Boca-Leek paper.
+Here we download and analyze the GWAS BMI dataset analyzed in the Boca-Leek paper.
+
+#### Data download
 
@@
 ```{r, GWAS1-verify-contents}
-library(data.table)
 bmi <- fread("BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt", header=TRUE)
@@
 
+#### Running benchmark methods
+
+First, we'll standardize the data frame to use the wrapper function. For now, we'll
+assume that there is only a single 'independent covariate' for those methods that 
+utilize this, and we'll use the sample size (number of subjects with that SNP
+measured). **Still to-do: consider multiple covariates for methods that allow this.**
+
+```{r, GWAS1-dat}
+bmi <- bmi %>% 
+        rename(pval = p_value,
+               SE = stderr,
+               effect_size = effect,
+               ind_covariate = N)
+```
+
+Now, we'll run the `run_benchmarks` helper function on the `bmi` data frame for
+several alpha values and plot the results. This is fairly computationally intensive 
+for some of the benchmark methods (since there are more than 2.5 million tests)
+and best run on a machine with lots of memory.
+
+```{r, GWAS1-run}
+cutoffs <- c(0.01, 0.025, 0.05, 0.10, 0.20)
+if (!file.exists("bmi_results.RData")){
+  bmi_results <- run_benchmarks(dat=bmi, alphas=cutoffs)
+  save(bmi_results, file="bmi_results.RData")
+}else{
+  load("bmi_results.RData")
+}
+
+# plot
+ggplot(bmi, aes(alpha, n_rejects, color=method)) +
+  geom_point() +
+  geom_line() +
+  xlab("FDR cutoff") + ylab("Number of rejections") +
+  ggtitle("BMI GWAS Metanalysis with Sample Size Covariate")
+```
+
+And we'll clean up the workspace before moving on to the next dataset.
+
 ```{r, GWAS1-cleanup}
 rm(bmi)
+rm(asthma_results)
 ```
@@
 GWAS on  10,365 people with asthma and 16,110 controls 23 individual studies.
+
+#### Data Download
 
@@
 # read in meta-data
-library(readxl)
 meta <- read_xls("gabriel_results_description.xls")
@@
 # remove study-specific columns
-library(dplyr)
 empty <- which(!is.na(meta$description))
@@
             slice(empty)
+
+# remove SNPs with missing effect size/pvalues
+asthma <- asthma %>% 
+            filter(!is.na(P_fix) & !is.na(theta_fix))
+
+dim(asthma)
 head(asthma)
@@
 - `P_fix` and `P_ran`:p-values for fixed and random effect models
-- `theta_fix` and `theta_fix`:effect size for fixed and random effect models
-- `se_theta_fix` and `se_theta_fix`:standard error for fixed and random effect models
+- `theta_fix` and `theta_ran`:effect size for fixed and random effect models
+- `se_theta_fix` and `se_theta_ran`:standard error for fixed and random effect models
 - additional covariates: 
@@
 
-for 567,589 SNPs.
+for 546,182 SNPs.
+
+#### Running benchmark methods
+
+First, we'll standardize the data frame to use the wrapper function. For now, we'll
+assume that there is only a single 'independent covariate' for those methods that 
+utilize this, and we'll use the sample size (number of studies with that SNP
+measured). We'll also focus on the fixed effects tests first.
+**Still to-do: consider multiple covariates for methods that allow this.**
+
+```{r, GWAS2-dat}
+asthma <- asthma %>% 
+        rename(pval = P_fix,
+               SE = se_theta_fix,
+               effect_size = theta_fix,
+               ind_covariate = no_studies_meta)
+```
+
+Now, we'll run the `run_benchmarks` helper function on the `asthma` data frame for
+several alpha values and plot the results. This is fairly computationally intensive 
+for some of the benchmark methods
+(since there are more than 500 thousand tests)
+and best run on a machine with lots of memory.
+
+```{r, GWAS2-run}
+# use same cutoffs as for the bmi study (defined above)
+if (!file.exists("asthma_results.RData")){
+  asthma_results <- run_benchmarks(dat=asthma, alphas=cutoffs)
+  save(asthma_results, file="asthma_results.RData")
+}else{
+  load("asthma_results.RData")
+}
+
+# plot
+ggplot(asthma_results, aes(alpha, n_rejects, color=method)) +
+  geom_point() +
+  geom_line() +
+  xlab("FDR cutoff") + ylab("Number of rejections") +
+  ggtitle("Asthma GWAS Metanalysis with Sample Size Covariate")
+```
+
+And we'll clean up the workspace before moving on to the next dataset.
 
@@
 rm(asthma)
+rm(asthma_results)
 ```
@@
 GWAS on 188,577 and measured blood lipid levels (as well as several cardiovascular traits).
+
+#### Data Download
 
@@
 
+#### Running benchmark methods
+
+First, we'll standardize the data frame to use the wrapper function. For now, we'll
+assume that there is only a single 'independent covariate' for those methods that 
+utilize this, and we'll use the sample size (number of individuals with that SNP
+measured). 
+**Still to-do: consider multiple covariates for methods that allow this.**
+
+```{r, GWAS3-dat}
+trig <- trig %>% 
+        rename(pval = P_value,
+               SE = SE,
+               effect_size = beta,
+               ind_covariate = N)
+```
+
+Now, we'll run the `run_benchmarks` helper function on the `trig` data frame for
+several alpha values and plot the results. This is fairly computationally intensive 
+for some of the benchmark methods
+(since there are more than 500 thousand tests)
+and best run on a machine with lots of memory.
+
+```{r, GWAS2-run}
+# use same cutoffs as for the bmi study (defined above)
+if (!file.exists("triglyceride_results.RData")){
+  trig_results <- run_benchmarks(dat=trig, alphas=cutoffs)
+  save(trig_results, file="triglyceride_results.RData")
+}else{
+  load("triglyceride_results.RData")
+}
+
+# plot
+ggplot(trig_results, aes(alpha, n_rejects, color=method)) +
+  geom_point() +
+  geom_line() +
+  xlab("FDR cutoff") + ylab("Number of rejections") +
+  ggtitle("Asthma GWAS Metanalysis with Sample Size Covariate")
+```
+
+And we'll clean up the workspace before moving on to the next dataset.
+
 ```{r, GWAS3-cleanup}
 rm(trig)
+rm(trig_results)
 ```

### 2017-09-20 datasets/R/evaluation-helpers.R

@@ -1,0 +1,136 @@
+#' Dataset Evaluation
+#'
+#' Wrapper function for computing adjusted p-values (or comparable values) and
+#' number of rejections at various alphas for
+#' all methods in benchmkaring study.
+#' 
+#' @param dat data.frame containing the following columns: `pval`= p-value for each
+#'   test, `ind_covariate` = an independent covariate under the null, 
+#'   `effect_size` =  the effect size for each test, `SE` = the standard error 
+#'   for each effect. (STILL TO DO - ALLOW MULTIPLE COVARIATES)
+#' @param alphas vector of FDR cutoffs between 0, 1
+#' @param pvals logical whether to return table of 'adjusted p-values' along with
+#'        summary metrics (default = FALSE)
+#' @param verbose logical whether to print out progress messages to stdout 
+#' 
+#' @return
+#' If `pvals = FALSE`, a data.frame with evaluation metrics for each method at the specified alpha
+#' thresholds. Should have number of rows equal to `length(alphas) * (n_methods)`,
+#' where `n_methods` is the number of different methods (or methods and parameters)
+#' in the benchmark study.
+#' If `pvals = TRUE`, a list with two elements:
+#' * `stats`: data.frame containing evaluation metrics
+#' * `pvals`: data.frame of the adjusted p-values calculated by
+#'    each method, with number of rows equal to `nrows(dat)`.
+#'
+#' @md
+#' @author Stephanie Hicks
+run_benchmarks <- function(dat, alphas, pvals = FALSE, verbose = TRUE) { 
+  
+  ## keep adjusted p-values
+  adj_pset <- list()
+  
+  ## Bonferroni 
+  if(verbose)
+    message("Running Bonferonni.")
+  adj_p <- p.adjust(dat$pval, method="bonferroni")
+  df.bonf <- plyr::ldply(alphas, function(a){
+    rjs <- sum(adj_p <= a)
+    data.frame(n_rejects = rjs, alpha = a) 
+    })
+  adj_pset$bonf <- adj_p
+  
+  ## BH
+  if(verbose)
+    message("Running BH.")
+  adj_p <- p.adjust(dat$pval, method="BH")
[…]

### 2017-09-21 datasets/GWAS/GWAS.Rmd
Add demo evaluation on 10K SNPs

@@
 author: "Keegan Korthauer"
-date: "9/12/2017"
+date: "9/21/2017"
 output: html_document
@@
 We'll also run the benchmark methods on each dataset and report the number of 
-discoveries for each one.
+discoveries for each one. At first, we'll demo on just a __**small subset of 
+SNPs**__, since the computation is time-intensive. After that is working 
+smoothly, we'll scale it up to the full set of SNPs.
 
@@
 To run the benchmark methods on the GWAS datasets, we'll use some of the helper 
-and wrapper functions from the simulated data analysis. We also create and source an analgous version of the `sim_runner` function to apply to the real data.
+and wrapper functions from the simulated data analysis. We also create and source
+an analgous version of the `sim_runner` function to apply to the real data.
 
@@
 # Note that swfdr package depends on R version 3.4 or higher
-setwd("/n/irizarryfs01_backed_up/kkorthauer/benchmark-fdr/")
-source("simulation_studies/R/simulation-helpers.R")
-source("datasets/R/evaluation-helpers.R")
-setwd("./datasets/GWAS")
+setwd("/n/irizarryfs01_backed_up/kkorthauer/FDR")
+source("benchmark-fdr/simulation_studies/R/simulation-helpers.R")
+source("benchmark-fdr/datasets/R/evaluation-helpers.R")
+setwd("DATA/GWAS")
 
@@
                effect_size = effect,
-               ind_covariate = N)
+               ind_covariate = N) %>%
+        mutate(zscore = effect_size/SE)
+
+# Demo on a small subset of SNPs first (remove this later)
+bmi <- bmi[sample(1:nrow(bmi), 10000),]
 ```
@@
 
-```{r, GWAS1-run}
+```{r, GWAS1-run, results="hide"}
 cutoffs <- c(0.01, 0.025, 0.05, 0.10, 0.20)
-if (!file.exists("bmi_results.RData")){
+resfile <- paste0("bmi_results_", nrow(bmi), ".RData")
+if (!file.exists(resfile)){
   bmi_results <- run_benchmarks(dat=bmi, alphas=cutoffs)
-  save(bmi_results, file="bmi_results.RData")
+  save(bmi_results, file=resfile)
 }else{
-  load("bmi_results.RData")
+  load(resfile)
 }
 
-# plot
-ggplot(bmi, aes(alpha, n_rejects, color=method)) +
+# plot N rejects vs FDR cutoff
+p <- ggplot(bmi_results, aes(alpha, n_rejects, color=method)) +
   geom_point() +
@@
   xlab("FDR cutoff") + ylab("Number of rejections") +
-  ggtitle("BMI GWAS Metanalysis with Sample Size Covariate")
+  ggtitle("BMI GWAS Metanalysis with Sample Size Covariate") +
+  theme_classic()
+print(p)
+
+# log-scale y axis
+print(p + scale_y_continuous(trans="log2") + 
+        ylab("Number of rejections (log2 scale)"))
 ```
@@
                effect_size = theta_fix,
-               ind_covariate = no_studies_meta)
+               ind_covariate = no_studies_meta) %>%
+        mutate(zscore = effect_size/SE)
+
+# Demo on a small subset of SNPs first (remove this later)
+asthma <- asthma[sample(1:nrow(asthma), 10000),]
 ```
@@
 
-```{r, GWAS2-run}
+```{r, GWAS2-run, results="hide"}
 # use same cutoffs as for the bmi study (defined above)
-if (!file.exists("asthma_results.RData")){
+resfile <- paste0("asthma_results_", nrow(asthma), ".RData")
+if (!file.exists(resfile)){
   asthma_results <- run_benchmarks(dat=asthma, alphas=cutoffs)
-  save(asthma_results, file="asthma_results.RData")
+  save(asthma_results, file=resfile)
 }else{
-  load("asthma_results.RData")
+  load(resfile)
 }
 
-# plot
-ggplot(asthma_results, aes(alpha, n_rejects, color=method)) +
+# plot N rejects vs FDR cutoff
+p <- ggplot(asthma_results, aes(alpha, n_rejects, color=method)) +
   geom_point() +
@@
   xlab("FDR cutoff") + ylab("Number of rejections") +
-  ggtitle("Asthma GWAS Metanalysis with Sample Size Covariate")
+  ggtitle("Asthma GWAS Metanalysis with Sample Size Covariate") +
+  theme_classic()
+print(p)
+
+# log-scale y axis
+print(p + scale_y_continuous(trans="log2") + 
+        ylab("Number of rejections (log2 scale)"))
 ```
@@
 
+Note that for this dataset, all reported effect sizes are positive. This is because
+the alleles (A1 and A2) have been ordered such that the direction of effect are 
+consistent. Since Scott assumes that the Z-scores are standard normal, we can't 
+apply this method to a set of all positive Z-scores. To recover the direction of 
+effect, we use the `Freq.A1.1000G.EUR` (allele frequency of A1) variable,
+and use it to determine whether
+`A1` (allele 1) is the reference allele or not. If the allele frequence of A1 is
+greater than 0.5, we consider it the reference allele. Then we'll use the 
+comparison of reference to minor as the positive direction. That is,
+if A2 was reference,
+then effect direction is negative. Otherwise it is positive. Since some of the
+SNPs are missing the reference allele information, we cannot determine their 
+direction of effect and they will be removed.
+
 ```{r, GWAS3-dat}
 trig <- trig %>% 
-        rename(pval = P_value,
-               SE = SE,
+        rename(pval = "P-value",
+               SE = se,
                effect_size = beta,
-               ind_covariate = N)
+               ind_covariate = N)%>%
+        mutate(zscore = effect_size/SE)
+
+# add effect size
+trig <- trig %>% 
+       mutate(A2.reference = Freq.A1.1000G.EUR < 0.5) %>%
+       mutate(zscore = zscore * (1-2*A2.reference))
+
+# remove SNPs with missing values for zscore (couldn't determine direction of 
+# effect)
+trig <- trig %>% 
+  filter(!is.na(zscore))
+
+# Demo on a small subset of SNPs first (remove this later)
+trig <- trig[sample(1:nrow(trig), 10000),]
 ```
@@
 
-```{r, GWAS2-run}
+```{r, GWAS3-run, results="hide"}
 # use same cutoffs as for the bmi study (defined above)
-if (!file.exists("triglyceride_results.RData")){
+resfile <- paste0("trig_results_", nrow(trig), ".RData")
+if (!file.exists(resfile)){
   trig_results <- run_benchmarks(dat=trig, alphas=cutoffs)
-  save(trig_results, file="triglyceride_results.RData")
+  save(trig_results, file=resfile)
 }else{
-  load("triglyceride_results.RData")
+  load(resfile)
 }
 
-# plot
-ggplot(trig_results, aes(alpha, n_rejects, color=method)) +
+# plot N rejects vs FDR cutoff
+p <- ggplot(trig_results, aes(alpha, n_rejects, color=method)) +
   geom_point() +
@@
   xlab("FDR cutoff") + ylab("Number of rejections") +
-  ggtitle("Asthma GWAS Metanalysis with Sample Size Covariate")
+  ggtitle("Triglyceride GWAS Metanalysis with Sample Size Covariate") +
+  theme_classic()
+print(p)
+
+# log-scale y axis
+print(p + scale_y_continuous(trans="log2") + 
+        ylab("Number of rejections (log2 scale)"))
+
 ```

### 2017-09-21 datasets/R/evaluation-helpers.R

@@ -25,6 +25,7 @@
 #'
 #' @md
 #' @author Stephanie Hicks
+#' @author Keegan Korthauer
 run_benchmarks <- function(dat, alphas, pvals = FALSE, verbose = TRUE) { 
   
   ## keep adjusted p-values
@@ -37,7 +38,7 @@ run_benchmarks <- function(dat, alphas, pvals = FALSE, verbose = TRUE) {
   df.bonf <- plyr::ldply(alphas, function(a){
     rjs <- sum(adj_p <= a)
     data.frame(n_rejects = rjs, alpha = a) 
-    })
+  })
   adj_pset$bonf <- adj_p
   
   ## BH
@@ -107,16 +108,22 @@ run_benchmarks <- function(dat, alphas, pvals = FALSE, verbose = TRUE) {
   adj_pset$lfdr <- adj_p
   
   ## Scott et al. (2015) (FDR regression (FDRreg) available via GitHub for version 2.0)
-  if(verbose)
-    message("Running Scott.")
-  adj_p <- scott_fdrreg_hickswrapper(unadj_p = dat$pval, effect_size = dat$effect_size,
-                                     filterstat = dat$ind_covariate, 
-                                     df=3, lambda=0.1, nulltype='theoretical')
-  df.scott <- plyr::ldply(alphas, function(a){
-    rjs <- sum(adj_p <= a)
-    data.frame(n_rejects = rjs, alpha = a) 
-  })
-  adj_pset$scott <- adj_p
+  # can only run if z-scores are present.
+  if("zscore" %in% colnames(dat)){
+    if(verbose)
+      message("Running Scott.")
+    adj_p <- scott_fdrreg_hickswrapper(zscores = dat$zscore, 
+                                       filterstat = dat$ind_covariate, 
+                                       df=3, lambda=0.1, nulltype='theoretical')
+    df.scott <- plyr::ldply(alphas, function(a){
+      rjs <- sum(adj_p <= a)
+      data.frame(n_rejects = rjs, alpha = a) 
+    })
+    adj_pset$scott <- adj_p
+  }else{
+    if(verbose)
+      message("Skipping Scott; no Z-scores present")
+  }
   
   ## Summary table of FDP for each method at each alpha
   stat_df <- rbind(data.frame(df.bonf, "method" = "bonferroni"), 
@@ -135,3 +142,32 @@ run_benchmarks <- function(dat, alphas, pvals = FALSE, verbose = TRUE) {
     return(stat_df)
   }
 }
+
[…]

### 2017-09-21 datasets/GWAS/GWAS.Rmd
Increase to 100K SNPs and add directory pointers

@@
 source("benchmark-fdr/datasets/R/evaluation-helpers.R")
-setwd("DATA/GWAS")
+datdir <- "/n/irizarryfs01_backed_up/kkorthauer/FDR/DATA/GWAS/"
+resdir <- "/n/irizarryfs01_backed_up/kkorthauer/FDR/RESULTS/GWAS/"
 
@@
 ```{r, GWAS1-verify-contents}
-bmi <- fread("BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt", header=TRUE)
+bmi <- fread(paste0(datdir, "BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt"),
+             header=TRUE)
 dim(bmi)
@@
 # Demo on a small subset of SNPs first (remove this later)
-bmi <- bmi[sample(1:nrow(bmi), 10000),]
+bmi <- bmi[sample(1:nrow(bmi), 100000),]
 ```
@@
 cutoffs <- c(0.01, 0.025, 0.05, 0.10, 0.20)
-resfile <- paste0("bmi_results_", nrow(bmi), ".RData")
+resfile <- paste0(resdir, "bmi_results_",
+                  nrow(bmi), ".RData")
 if (!file.exists(resfile)){
+  t1 <- proc.time()
   bmi_results <- run_benchmarks(dat=bmi, alphas=cutoffs)
   save(bmi_results, file=resfile)
+  message(paste0("Took ", round((proc.time()-t1)[3],1),
+                 " seconds for ", nrow(bmi), " SNPs."))
 }else{
@@
 rm(bmi)
-rm(asthma_results)
+rm(bmi_results)
 ```
@@
 # read in meta-data
-meta <- read_xls("gabriel_results_description.xls")
+meta <- read_xls(paste0(datdir, "gabriel_results_description.xls"))
 head(meta)
@@
 # read in data
-asthma <- fread("gabriel_asthma_meta-analysis_36studies_format_repository_NEJM.txt", header=TRUE)
+asthma <- fread(paste0(datdir, 
+                "gabriel_asthma_meta-analysis_36studies_format_repository_NEJM.txt"),
+                header=TRUE)
 dim(asthma)
@@
 # Demo on a small subset of SNPs first (remove this later)
-asthma <- asthma[sample(1:nrow(asthma), 10000),]
+asthma <- asthma[sample(1:nrow(asthma), 100000),]
 ```
@@
 # use same cutoffs as for the bmi study (defined above)
-resfile <- paste0("asthma_results_", nrow(asthma), ".RData")
+resfile <- paste0(resdir, "asthma_results_",
+                  nrow(asthma), ".RData")
 if (!file.exists(resfile)){
+  t1 <- proc.time()
   asthma_results <- run_benchmarks(dat=asthma, alphas=cutoffs)
   save(asthma_results, file=resfile)
+  message(paste0("Took ", round((proc.time()-t1)[3],1),
+                 " seconds for ", nrow(asthma), " SNPs."))
 }else{
@@
 # read in data
-trig <- fread("jointGwasMc_TG.txt", header=TRUE)
+trig <- fread(paste0(datdir, "jointGwasMc_TG.txt"), header=TRUE)
 dim(trig)
@@
 # Demo on a small subset of SNPs first (remove this later)
-trig <- trig[sample(1:nrow(trig), 10000),]
+trig <- trig[sample(1:nrow(trig), 100000),]
 ```
@@
 # use same cutoffs as for the bmi study (defined above)
-resfile <- paste0("trig_results_", nrow(trig), ".RData")
+resfile <- paste0(resdir, "trig_results_",
+                  nrow(trig), ".RData")
 if (!file.exists(resfile)){
+  t1 <- proc.time()
   trig_results <- run_benchmarks(dat=trig, alphas=cutoffs)
   save(trig_results, file=resfile)
+  message(paste0("Took ", round((proc.time()-t1)[3],1),
+                 " seconds for ", nrow(trig), " SNPs."))
 }else{

### 2017-09-21 datasets/R/evaluation-helpers.R

@@ -143,31 +143,3 @@ run_benchmarks <- function(dat, alphas, pvals = FALSE, verbose = TRUE) {
   }
 }
 
-## adapted from IHWpaper::scott_fdrreg()
-scott_fdrreg_hickswrapper <- function(zscores=NULL, filterstat, df=3, lambda=0.01, nulltype = 'theoretical') {
-  if (! as.character(packageVersion("FDRreg")) %in% c('0.2.1', '0.2')){
-    stop(paste("Benchmarks were run against version 0.2 of FDRreg",
-               "available on github via:",
-               "devtools::install_github(repo= 'jgscott/FDRreg', subdir='R_pkg/')"
-    ))
-  }
-  
-  if (is.null(zscores) | 
-      sum(zscores < 1 & zscores > 0, na.rm=TRUE)==sum(!is.na(zscores))){
-    stop(paste0("The function scott_fdrreg_hickswrapper now takes as input ",
-                "Z-scores instead of p-values. Please calculate Z-scores ",
-                "first, either from p-values (considering whether the test ",
-                "is 1- or 2-sided), or from effect sizes and SEs."))
-  }
-  
-  ## no automated way to choose function space over which we optimize
-  ## so we just use bs(df=3) as done in their analysis
-  b <- splines::bs(filterstat, df=df)
-  
-  Xs <- model.matrix( ~  b - 1)
-  fdrreg_res <- FDRreg::FDRreg(z=zscores, features=Xs, nulltype = nulltype,
-                               control=list(lambda = lambda)) # assumption of test statistic follow a standard normal
-  adj_p <- fdrreg_res$FDR
-  return(adj_p)
-}
-

### 2017-09-22 datasets/microbiome/cdi_schubert.Rmd
run CDI Schubert through all tests

---
title: "CDI Schubert"
author: "Claire Duvallet"
date: "9/21/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary

This script will download one processed OTU table from Zenodo, calculate p-values for
OTU abundances in cases vs. controls, and also calculate some covariates like ubiquity,
abundance, and average reads.

## Data

Note: later, we'll be using the following datasets:

- cdi_schubert: 154 healthy, 93 CDI, 89 nonCDI (we can either do CDI vs. healthy or any diarrhea vs. healthy)
- ob_goodrich: 428 healthy, 185 obese
- crc_baxter: 172 healthy, 120 CRC
- ibd_gevers: 16 nonIBD, 146 Crohn's
      - note: the two largest IBD datasets are very uneven (ibd_morgan has 18 healthy, 108 IBD). We could 
  aso use ibd_papa, which has 24 nonIBD and 66 IBD

### CDI Schubert

Let's start with the cdi_schubert dataset first...

```{r}
library(dplyr)
library(ggplot2)

setwd("/Users/claire/github/benchmark-fdr/datasets/microbiome")
source("../../simulation_studies/R/simulation-helpers.R")
source("../R/evaluation-helpers.R")
resdir <- "/Users/claire/github/benchmark-fdr/datasets/microbiome/"
```

```{bash, schubert-download, eval=FALSE}
curl -O "https://zenodo.org/record/840333/files/cdi_schubert_results.tar.gz"
tar -xzvf cdi_schubert_results.tar.gz
```

```{r, results='hide'}
# load in dbOTU table.
otu <- read.table('cdi_schubert_results/RDP/cdi_schubert.otu_table.dbOTU.rdp_assigned')
# and metadata
meta <- read.csv('cdi_schubert_results/cdi_schubert.metadata.txt', sep='\t')
head(otu)
```

```{r, results='hide'}
## Clean up the data

# Keep only samples with DiseaseState metadata
labels <- c("H", "nonCDI", "CDI")

print(dim(meta))
meta <- meta %>% filter(DiseaseState %in% labels)
print(dim(meta))

# Metadata and 16S data
keep_samples <- intersect(colnames(otu), meta$sample_id)
otu <- otu[, keep_samples]
meta <- meta %>% filter(sample_id %in% keep_samples)
print(dim(otu))
print(dim(meta))

# Remove OTUs and samples without enough reads
n_reads_samples <- 1000
n_reads_otus <- 10

remove_shallow_smpls <- function(df, n_reads) {
  # Removes samples with fewer than n_reads from dataframe df.
  # df has OTUs in rows and samples in columns
  return(df[, colSums(df) > n_reads])
  }

remove_shallow_otus <- function(df, n_reads){
  return(df[rowSums(df) > n_reads, ])
  }

print(dim(otu))
otu <- remove_shallow_smpls(otu, n_reads_samples)
otu <- remove_shallow_otus(otu, n_reads_otus)
print(dim(otu))

# Update metadata with new samples
meta <- meta %>% filter(sample_id %in% colnames(otu))

```

```{r, results='hide'}
## Calculate some covariates

# Mean read depth
reads <- rowMeans(otu)

# Ubiquity
ubi <- rowSums(otu > 0) / length(rownames(otu))

# Mean relative abundance (only considering people who have the OTU present)
abun_otu <- t(t(otu) / rowSums(t(otu)))
mean_abun <- rowSums(abun_otu) / rowSums(abun_otu > 0)

# Make new dataframe to store covariates and eventual p-values
schubert <- data.frame(otu = rownames(otu), mean_abun = mean_abun,
                       ubiquity = ubi, mean_reads = reads)
head(schubert)
```

```{r}
## Calculate p-values between case (CDI or nonCDI) and controls (H)
## Use the relative abundances!

# just do a for loop and ask how to vectorize l8er
case_idx <- meta$DiseaseState %in% c("CDI", "nonCDI")
ctrl_idx <- meta$DiseaseState %in% c("H")
pvals <- c()
ses <- c()
effs <- c()
for (o in rownames(abun_otu)) {
  p <- wilcox.test(abun_otu[o, ctrl_idx], 
                   abun_otu[o, case_idx])$p.value
  pvals <- c(pvals, p)

  # standard error
  ses <- c(ses, sd(abun_otu[o,])/sqrt(length(abun_otu[o,])))

  # effect (difference of means)
  effs <- c(effs, mean(abun_otu[o, ctrl_idx]) - mean(abun_otu[o, case_idx]))
}
```

```{r}
schubert$pval <- pvals
schubert$SE <- ses
schubert$effect_size <- effs

# Set up ubiquity as the first covariate to test.
# Ranges from 0.001 to 0.4
schubert <- schubert %>% 
              mutate(zscore = effect_size/SE) %>%
              mutate(ind_covariate = log10(ubiquity) )
```

```{r, results='hide', eval = FALSE}
head(schubert)
```

```{r}
# Start with ubiquity as the covariate
# TODO: check independence under the null

# Copied from Keegan's code :)
cutoffs <- c(0.01, 0.025, 0.05, 0.10, 0.20)
resfile <- paste0(resdir, "cdi_schubert_results_",
                  nrow(schubert), ".RData")
t1 <- proc.time()
schu_results <- run_benchmarks(dat=schubert, alphas=cutoffs)
save(schu_results, file=resfile)
message(paste0("Took ", round((proc.time()-t1)[3],1),
               " seconds for ", nrow(schubert), " OTUs"))
```

```{r}
# plot % rejects vs FDR cutoff
schu_results <- schu_results %>% mutate(perc_reject = n_rejects / length(rownames(abun_otu)))
p <- ggplot(schu_results, aes(alpha, perc_reject, color=method)) +
  geom_point() +
  geom_line() +
  xlab("FDR cutoff") + ylab("Fraction of hypotheses rejected") +
  ggtitle("CDI Schubert with log10(Ubiquity) Covariate") +
  theme_classic()
print(p)

```

### 2017-09-28 datasets/GWAS/GWAS.Rmd
Add diagnostic plots for covariates (Issue #13)

@@
 author: "Keegan Korthauer"
-date: "9/21/2017"
+date: "9/28/2017"
 output: html_document
@@
 library(ggplot2)
+library(cowplot)
 
@@
 
+#### Covariate Diagnostics
+
+Here we examine whether the available covariates appear to be suitable for the  
+assumptions of the benchmark methods. The main assumption used by many methods 
+is that the covariate is independent of the p-value conditional on the null
+hypothesis. In other words, the covariate itself is not informative for whether
+the test should be rejected or not. If this is satisfied, the covariate still
+may or may not be useful. To assess whether the covariate is useful or not, we 
+look to see if there is an association between covariate values and p-values. 
+This is because under the alternative, we expect to see an enrichment of 
+rejections for certain values of the covariate (if it is indeed informative).
+In this section we examine two diagnostic plots for each of the covariates to 
+see if these conditions hold in this dataset.
+
+First, we will look at histograms of p-values stratified by covariate values. To
+stratify by continuous covariate values, we'll use quantiles.
+
+```{r, GWAS1-diag1, fig.width=9, fig.height=2.5}
+# Demo on a small subset of SNPs first (remove this later)
+bmi <- bmi[sample(1:nrow(bmi), 100000),]
+
+# First for the sample size covariate (N)
+strat_hist(bmi, pval="p_value", covariate="N", maxy=3)
+
+# Next for the MAF covariate (Freq_Allele1_HapMapCEU)
+strat_hist(bmi %>% filter(!is.na(Freq_Allele1_HapMapCEU)),
+           pval="p_value", covariate="Freq_Allele1_HapMapCEU", maxy=3)
+
+```
+
+Next we'll look at a scatter plot of p-values by the ranks of covariate values.
+
+```{r, GWAS1-diag2, fig.width=7, fig.height=4}
+# First for the sample size covariate (N)
+rank_scatter(bmi, pval="p_value", covariate="N")
+
+# Next for the MAF covariate (Freq_Allele1_HapMapCEU)
+rank_scatter(bmi, pval="p_value", covariate="Freq_Allele1_HapMapCEU")
+
+# Take proximity to 0.5 as the actual covariate (according to previous plot)
+bmi <- bmi %>%
+        mutate(MAF.5=-abs(Freq_Allele1_HapMapCEU-0.5))
+rank_scatter(bmi, pval="p_value", covariate="MAF.5") 
+```
+
 #### Running benchmark methods
@@
         mutate(zscore = effect_size/SE)
-
-# Demo on a small subset of SNPs first (remove this later)
-bmi <- bmi[sample(1:nrow(bmi), 100000),]
 ```
@@
 
+#### Covariate Diagnostics
+
+First, we will look at histograms of p-values stratified by covariate values. To
+stratify by continuous covariate values, we'll use quantiles.
+
+```{r, GWAS2-diag1, fig.width=9, fig.height=2.5}
+# Demo on a small subset of SNPs first (remove this later)
+asthma <- asthma[sample(1:nrow(asthma), 100000),]
+
+# First for the sample size covariate (no_studies_meta)
+strat_hist(asthma, pval="P_fix", covariate="no_studies_meta", maxy=1.3)
+
+# Next for the MAF covariate (freq_all_1_max)
+strat_hist(asthma %>% filter(!is.na(freq_all_1_max)),
+           pval="P_fix", covariate="freq_all_1_max", maxy=1.3)
+```
+
+Next we'll look at a scatter plot of p-values by the ranks of covariate values.
+
+```{r, GWAS2-diag2, fig.width=7, fig.height=4}
+# First for the sample size covariate (N)
+rank_scatter(asthma, pval="P_fix", covariate="no_studies_meta")
+
+# Next for the MAF covariate (freq_all_1_max)
+rank_scatter(asthma, pval="P_fix", covariate="freq_all_1_max")
+
+# Take distance to 0.5 as the actual covariate 
+asthma <- asthma %>%
+        mutate(MAF.5=abs(freq_all_1_max-0.5))
+rank_scatter(asthma, pval="P_fix", covariate="MAF.5") 
+```
+
 #### Running benchmark methods
@@
         mutate(zscore = effect_size/SE)
-
-# Demo on a small subset of SNPs first (remove this later)
-asthma <- asthma[sample(1:nrow(asthma), 100000),]
 ```
@@
 
+#### Covariate Diagnostics
+
+First, we will look at histograms of p-values stratified by covariate values. To
+stratify by continuous covariate values, we'll use quantiles.
+
+```{r, GWAS3-diag1, fig.width=9, fig.height=2.5}
+# Demo on a small subset of SNPs first (remove this later)
+trig <- trig[sample(1:nrow(trig), 100000),]
+
+# First for the sample size covariate (no_studies_meta)
+strat_hist(trig, pval="P-value", covariate="N", maxy=1.3)
+
+# Next for the MAF covariate (Freq.A1.1000G.EUR)
+strat_hist(trig %>% filter(!is.na(Freq.A1.1000G.EUR)),
+           pval="P-value", covariate="Freq.A1.1000G.EUR", maxy=1.3)
+```
+
+Next we'll look at a scatter plot of p-values by the ranks of covariate values.
+
+```{r, GWAS3-diag2, fig.width=7, fig.height=4}
+# First for the sample size covariate (N)
+rank_scatter(trig, pval="P-value", covariate="N")
+
+# Next for the MAF covariate (Freq.A1.1000G.EUR)
+rank_scatter(trig, pval="P-value", covariate="Freq.A1.1000G.EUR") 
+```
+
 #### Running benchmark methods
@@
   filter(!is.na(zscore))
-
-# Demo on a small subset of SNPs first (remove this later)
-trig <- trig[sample(1:nrow(trig), 100000),]
 ```

### 2017-09-28 datasets/R/evaluation-helpers.R

@@ -143,3 +143,101 @@ run_benchmarks <- function(dat, alphas, pvals = FALSE, verbose = TRUE) {
   }
 }
 
+
+# strat hist function to input data frame and the covariate of interest, 
+# and output a cowplot object that displays 4 histograms: one overall, 
+# and three for the tertiles of the covariate
+# also takes as input the bin width and the max y value for the density
+
+# dat = dataframe with each row as a test
+# pval = name of p-value covariate
+# covariate = name of covariate to stratify by
+# binwidth = binwidth parameter for geom_hist
+# maximum y-value for density (so that all histograms are on the same scale
+strat_hist <- function(dat, pval, covariate, binwidth=0.025, maxy=3){
+  
+  # First for the sample size covariate (N)
+  gg_all <- ggplot(data=dat, aes(x=get(pval))) + 
+    geom_histogram(binwidth = binwidth, boundary = 0, 
+                   colour="grey", fill="lightgrey") +
+    aes(y=..density..)+
+    theme_classic() +
+    theme(axis.title = element_text(face="bold"),
+          plot.title = element_text(face="bold")) +
+    scale_x_continuous(expand = c(0.02, 0)) + 
+    scale_y_continuous(expand = c(0.02, 0), limits=c(0,maxy)) + 
+    xlab("p-value") +
+    ylab("Density") +
+    ggtitle("All SNPs")
+  
+  dat.strat <- dat %>% 
+    filter(-rank(get(covariate), ties="first") > quantile(-rank(get(covariate), ties="first"), 2/3))
+  gg_top <- ggplot(data=dat.strat, aes(x=get(pval))) + 
+    geom_histogram(binwidth = binwidth, boundary = 0, 
+                   colour="grey", fill="lightgrey") +
+    aes(y=..density..)+
+    theme_classic() +
+    theme(axis.title = element_text(face="bold"),
+          plot.title = element_text(face="bold")) +
+    scale_x_continuous(expand = c(0.02, 0)) + 
+    scale_y_continuous(expand = c(0.02, 0), limits=c(0,maxy)) + 
+    xlab("p-value") +
+    ylab("Density") +
+    ggtitle(paste0("Top third by ", covariate))
+  
+  dat.strat <- dat %>% 
+    filter(-rank(get(covariate), ties="first") <= quantile(-rank(get(covariate), ties="first"), 2/3) &
[…]

### 2017-09-28 datasets/GWAS/GWAS.Rmd
Add UpSet plots for GWAS results (Issue #13)

@@
 library(cowplot)
+library(UpSetR)
 
@@
 library(fdrtool)
+
+# constant parameters (for UpSet Plots)
+alpha.cutoff <- 0.05
 ```
@@
   t1 <- proc.time()
-  bmi_results <- run_benchmarks(dat=bmi, alphas=cutoffs)
+  bmi_results <- run_benchmarks(dat=bmi, alphas=cutoffs, pvals=TRUE)
   save(bmi_results, file=resfile)
@@
 # plot N rejects vs FDR cutoff
-p <- ggplot(bmi_results, aes(alpha, n_rejects, color=method)) +
+p <- ggplot(bmi_results$stats, aes(alpha, n_rejects, color=method)) +
   geom_point() +
@@
         ylab("Number of rejections (log2 scale)"))
+```
+
+Next, we look beyond the number of rejections for each method and examine
+the overlap between them. We'll use the Venn diagram alternative provided in the
+UpsetR package.
+
+```{r, GWAS1-upset, fig.width=9, fig.height=9}
+sig_df <- bmi_results$pvals %>%
+            mutate_all(function(x) 1*(x < alpha.cutoff))
+upset(sig_df, nsets = 9, nintersects = 30, mb.ratio = c(0.5, 0.5),
+      order.by = c("freq"), decreasing = c(TRUE), 
+      sets.bar.color = "#56B4E9")
 ```
@@
   t1 <- proc.time()
-  asthma_results <- run_benchmarks(dat=asthma, alphas=cutoffs)
+  asthma_results <- run_benchmarks(dat=asthma, alphas=cutoffs, pvals=TRUE)
   save(asthma_results, file=resfile)
@@
 # plot N rejects vs FDR cutoff
-p <- ggplot(asthma_results, aes(alpha, n_rejects, color=method)) +
+p <- ggplot(asthma_results$stats, aes(alpha, n_rejects, color=method)) +
   geom_point() +
@@
         ylab("Number of rejections (log2 scale)"))
+```
+
+Next, we look beyond the number of rejections for each method and examine
+the overlap between them. We'll use the Venn diagram alternative provided in the
+UpsetR package.
+
+```{r, GWAS2-upset, fig.width=9, fig.height=9}
+sig_df <- asthma_results$pvals %>%
+            mutate_all(function(x) 1*(x < alpha.cutoff))
+upset(sig_df, nsets = 9, nintersects = 30, mb.ratio = c(0.5, 0.5),
+      order.by = c("freq"), decreasing = c(TRUE), 
+      sets.bar.color = "#56B4E9")
 ```
@@
   t1 <- proc.time()
-  trig_results <- run_benchmarks(dat=trig, alphas=cutoffs)
+  trig_results <- run_benchmarks(dat=trig, alphas=cutoffs, pvals=TRUE)
   save(trig_results, file=resfile)
@@
 # plot N rejects vs FDR cutoff
-p <- ggplot(trig_results, aes(alpha, n_rejects, color=method)) +
+p <- ggplot(trig_results$stats, aes(alpha, n_rejects, color=method)) +
   geom_point() +
@@
 
+Next, we look beyond the number of rejections for each method and examine
+the overlap between them. We'll use the Venn diagram alternative provided in the
+UpsetR package.
+
+```{r, GWAS3-upset, fig.width=9, fig.height=9}
+sig_df <- trig_results$pvals %>%
+            mutate_all(function(x) 1*(x < alpha.cutoff))
+upset(sig_df, nsets = 9, nintersects = 30, mb.ratio = c(0.5, 0.5),
+      order.by = c("freq"), decreasing = c(TRUE), 
+      sets.bar.color = "#56B4E9")
+```
+
 And we'll clean up the workspace before moving on to the next dataset.

### 2017-09-29 datasets/microbiome/cdi_schubert.Rmd
Add Goodrich and diagnostic plots

@@
 library(ggplot2)
+library(cowplot)
 
-setwd("/Users/claire/github/benchmark-fdr/datasets/microbiome")
+wd <- "/Users/claire/github/benchmark-fdr/datasets/microbiome/"
+setwd(wd)
 source("../../simulation_studies/R/simulation-helpers.R")
@@
 
-```{r, results='hide'}
-# load in dbOTU table.
-otu <- read.table('cdi_schubert_results/RDP/cdi_schubert.otu_table.dbOTU.rdp_assigned')
-# and metadata
-meta <- read.csv('cdi_schubert_results/cdi_schubert.metadata.txt', sep='\t')
-head(otu)
-```
-
-```{r, results='hide'}
-## Clean up the data
-
-# Keep only samples with DiseaseState metadata
-labels <- c("H", "nonCDI", "CDI")
-
-print(dim(meta))
-meta <- meta %>% filter(DiseaseState %in% labels)
-print(dim(meta))
-
-# Metadata and 16S data
-keep_samples <- intersect(colnames(otu), meta$sample_id)
-otu <- otu[, keep_samples]
-meta <- meta %>% filter(sample_id %in% keep_samples)
-print(dim(otu))
-print(dim(meta))
-
-# Remove OTUs and samples without enough reads
-n_reads_samples <- 1000
-n_reads_otus <- 10
+```{r, data_functions, echo = FALSE, results='hide'}
+# Define functions to read and clean up data
 
@@
 
-print(dim(otu))
-otu <- remove_shallow_smpls(otu, n_reads_samples)
-otu <- remove_shallow_otus(otu, n_reads_otus)
-print(dim(otu))
+read_and_clean_data <- function(otu_path, meta_path, labels, col_label,
+                                sample_reads = 1000, otu_reads = 10) {
+  # Reads the OTU table and metadata from otu_path and meta_path,
+  # keeps samples which have both metadata and 16S data and whose
+  # DiseaseState is in the list labels. Removes samples with fewer
+  # than sample_reads reads, and and OTUs with fewer than otu_reads
+  # reads. Metadata should have sample IDs in first column and a column
+  # labeled DiseaseState
+
+  # load OTU table and metadata
+  otu <- read.table(otu_path)
+  meta <- read.csv(meta_path, sep='\t')
+
+  # Keep only samples with the right DiseaseState metadata
+  meta <- meta %>% filter(DiseaseState %in% labels)
+
+  # Keep only samples with both metadata and 16S data
+  keep_samples <- intersect(colnames(otu), meta[, 1])
+  otu <- otu[, keep_samples]
+  meta <- meta %>% filter(get(col_label) %in% keep_samples)
+  # Other ways to do this:
+  #meta[meta[, col_label] %in% keep_samples, ]
+  #meta[col_label]
+  #`[[`(meta, col_label)
+
+  # Remove OTUs and samples without enough reads
+  otu <- remove_shallow_smpls(otu, sample_reads)
+  otu <- remove_shallow_otus(otu, otu_reads)
 
   # Update metadata with new samples
-meta <- meta %>% filter(sample_id %in% colnames(otu))
+  meta <- meta %>% filter(get(col_label) %in% colnames(otu))
 
+  return(list(otu, meta))
+}
 ```
 
-```{r, results='hide'}
-## Calculate some covariates
+```{r, covariate_calcs, echo = FALSE, results='hide'}
+calculate_covariates <- function(otu, abun_otu){
+  # Calculates covariates for the OTU table.
+  # OTU table should have OTUs in rows and samples in columns.
+  # This function calculates (per OTU): mean read depth, 
+  # ubiquity, mean relative abundance in non-zero samples.
+  # It returns a dataframe with each of these as columns, in addition
+  # to "otu".
+  # zeroabun is the value in abun_otu corresponding to zero abundance.
 
@@
   # Mean relative abundance (only considering people who have the OTU present)
-abun_otu <- t(t(otu) / rowSums(t(otu)))
   mean_abun <- rowSums(abun_otu) / rowSums(abun_otu > 0)
 
-# Make new dataframe to store covariates and eventual p-values
-schubert <- data.frame(otu = rownames(otu), mean_abun = mean_abun,
+  covdf <- data.frame(otu = rownames(otu), mean_abun = mean_abun,
                       ubiquity = ubi, mean_reads = reads)
-head(schubert)
+  return(covdf)
+}
 ```
 
-```{r}
-## Calculate p-values between case (CDI or nonCDI) and controls (H)
-## Use the relative abundances!
+```{r, calculate_pvals, echo = FALSE}
+calculate_pvals <- function(abun_otu, ctrl_idx, case_idx){
+  # Does wilcoxon rank sum test for all OTUs between cases and
+  # controls. abun_otu is a dataframe with OTUs in rows and 
+  # samples in columns. ctrl_idx and case_idx are boolean
+  # vectors indicating the columns correspnding to control
+  # and case samples, respectively.
+  # This function also calculate the effect size 
+  # (difference of mean abundances) and standard error
 
-# just do a for loop and ask how to vectorize l8er
-case_idx <- meta$DiseaseState %in% c("CDI", "nonCDI")
-ctrl_idx <- meta$DiseaseState %in% c("H")
   pvals <- c()
@@
   effs <- c()
+
+  casedf <- abun_otu[, case_idx]
+  ctrldf <- abun_otu[, ctrl_idx]
   for (o in rownames(abun_otu)) {
-  p <- wilcox.test(abun_otu[o, ctrl_idx], 
-                   abun_otu[o, case_idx])$p.value
+    p <- wilcox.test(casedf[o, ], 
+                     ctrldf[o, ])$p.value
     pvals <- c(pvals, p)
@@
     effs <- c(effs, mean(abun_otu[o, ctrl_idx]) - mean(abun_otu[o, case_idx]))
+
+  }
+  res <- data.frame(otu = rownames(abun_otu), pval = pvals, SE = ses, effect_size = effs)
+  return(res)
+}
+
+```
+
+```{r, def_strat_hist}
+strat_hist <- function(dat, pval, covariate, binwidth=0.025, maxy=3){
+  gg_all <- ggplot(data=schubertdat, aes(x=get(pcol))) + 
+    geom_histogram(binwidth = binwidth, boundary = 0, 
+                   colour="grey", fill="lightgrey") +
+    aes(y=..density..)+
+    theme_classic() +
+    theme(axis.title = element_text(face="bold"),
+          plot.title = element_text(face="bold")) +
+    scale_x_continuous(expand = c(0.02, 0)) + 
+    scale_y_continuous(expand = c(0.02, 0), limits=c(0,maxy)) + 
+    xlab("p-value") +
+    ylab("Density") +
+    ggtitle("All OTUs")
+
+  dat.strat <- schubertdat %>% 
+      filter(-rank(get(covcol), ties="first") > quantile(-rank(get(covcol), ties="first"), 2/3))
+  gg_top <- ggplot(data=dat.strat, aes(x=get(pcol))) + 
+    geom_histogram(binwidth = binwidth, boundary = 0, 
+                   colour="grey", fill="lightgrey") +
+    aes(y=..density..) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold"),
+          plot.title = element_text(face="bold")) +
+    scale_x_continuous(expand = c(0.02, 0)) + 
+    scale_y_continuous(expand = c(0.02, 0), limits=c(0,maxy)) + 
+    xlab("p-value") +
+    ylab("Density") +
+    ggtitle(paste0("Top third by ", covcol))
+
+  dat.strat <- dat %>% 
+    filter(-rank(get(covcol), ties="first") <= quantile(-rank(get(covcol), ties="first"), 2/3) &
+           -rank(get(covcol), ties="first") > quantile(-rank(get(covcol), ties="first"), 1/3) )
+  gg_mid <- ggplot(data=dat.strat, aes(x=get(pcol))) + 
+    geom_histogram(binwidth = binwidth, boundary = 0, 
+                   colour="grey", fill="lightgrey") +
+    aes(y=..density..)+
+    theme_classic() +
+    theme(axis.title = element_text(face="bold"),
+          plot.title = element_text(face="bold")) +
+    scale_x_continuous(expand = c(0.02, 0)) + 
+    scale_y_continuous(expand = c(0.02, 0), limits=c(0,maxy)) + 
+    xlab("p-value") +
+    ylab("Density") +
+    ggtitle(paste0("Middle third by ", covcol))
+
+  dat.strat <- dat %>% 
+    filter(-rank(get(covcol), ties="first") <= quantile(-rank(get(covcol), ties="first"), 1/3))
+  gg_bot <- ggplot(data=dat.strat, aes(x=get(pcol))) + 
+    geom_histogram(binwidth = binwidth, boundary = 0, 
+                   colour="grey", fill="lightgrey") +
+    aes(y=..density..) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold"),
+          plot.title = element_text(face="bold")) +
+    scale_x_continuous(expand = c(0.02, 0)) + 
+    scale_y_continuous(expand = c(0.02, 0), limits=c(0,maxy)) + 
+    xlab("p-value") +
+    ylab("Density") +
+    ggtitle(paste0("Lower third by ", covcol))
+
+  gg_stratified <- plot_grid(gg_all, gg_top, gg_mid, gg_bot,
+                             nrow=1,
+                             labels=c("(a)", "(b)", "(c)","(d)"),
+                             hjust = 0.1)
+  return(gg_stratified)
 }
@@
 ```{r}
-schubert$pval <- pvals
-schubert$SE <- ses
-schubert$effect_size <- effs
+# Read and clean cdi_schubert
+otu_path = paste0(wd, "cdi_schubert_results/RDP/cdi_schubert.otu_table.100.denovo.rdp_assigned")
+meta_path = paste0(wd, "cdi_schubert_results/cdi_schubert.metadata.txt")
+labels = c("H", "nonCDI", "CDI")
 
-# Set up ubiquity as the first covariate to test.
-# Ranges from 0.001 to 0.4
-schubert <- schubert %>% 
-              mutate(zscore = effect_size/SE) %>%
-              mutate(ind_covariate = log10(ubiquity) )
+dat <- read_and_clean_data(otu_path, meta_path, labels, 'sample_id')
+otu <- dat[[1]]
+meta <- dat[[2]]
+
+abun_otu <- t(t(otu) / rowSums(t(otu)))
+minabun <- min(abun_otu[abun_otu > 0])
+zeroabun = 1e-6
+abun_otu <- abun_otu + zeroabun
 ```
 
-```{r, results='hide', eval = FALSE}
-head(schubert)
+```{r, schubert-pvals}
+case_idx <- meta$DiseaseState %in% c("CDI", "nonCDI")
+ctrl_idx <- meta$DiseaseState %in% c("H")
+
+# Calculate some basic covariates for each OTU
+schubertdat <- calculate_covariates(otu, abun_otu)
+# Get the pvalues, effects, and SEs
+schu_pvals <- calculate_pvals(abun_otu, ctrl_idx, case_idx)
+schubertdat <- merge(schubertdat, schu_pvals, by="otu")
+schubertdat <- schubertdat %>% 
+                 mutate(logpval = log10(pval)) %>%
+                 mutate(neglogpval = -log10(pval)) %>%
+                 mutate(logubiquity = log10(ubiquity)) %>%
+                 mutate(logmean_reads = log10(mean_reads))
+head(schubertdat)
 ```
@@
 ```{r}
-# Start with ubiquity as the covariate
-# TODO: check independence under the null
-
-# Copied from Keegan's code :)
-cutoffs <- c(0.01, 0.025, 0.05, 0.10, 0.20)
-resfile <- paste0(resdir, "cdi_schubert_results_",
-                  nrow(schubert), ".RData")
-t1 <- proc.time()
-schu_results <- run_benchmarks(dat=schubert, alphas=cutoffs)
-save(schu_results, file=resfile)
-message(paste0("Took ", round((proc.time()-t1)[3],1),
-               " seconds for ", nrow(schubert), " OTUs"))
+pcol <- "pval"
+covcol <- "logubiquity"
+binwidth <- 0.025
+maxy <- 3
+strat_hist(schubertdat, pcol, covcol)
+```
+```{r}
+g <- ggplot(schubertdat, aes(x = get(covcol), y = pval)) + geom_point()
+g 
 ```
@@
 ```{r}
+covcol <- "logmean_reads"
+strat_hist(schubertdat, pcol, covcol)
+```
+
+```{r}
+g <- ggplot(schubertdat, aes(x = get(covcol), y = pval)) + geom_point()
+g 
+```
+
+```{r, run_cdi_schubert}
+# Set up ubiquity as the first covariate to test.
+# Ranges from 0.001 to 0.4
+schubertdat <- schubertdat %>% 
+              mutate(zscore = effect_size/SE) %>%
+              mutate(ind_covariate = log10(ubiquity) )
+
+# Run all the tests and plot
+cutoffs <- c(0.01, 0.025, 0.05, 0.10, 0.20)
+resfile <- paste0(resdir, "cdi_schubert_results_",
+                  nrow(schubertdat), ".RData")
+t1 <- proc.time()
+schu_results <- run_benchmarks(dat=schubertdat, alphas=cutoffs)
+save(schu_results, file=resfile)
+message(paste0("Took ", round((proc.time()-t1)[3],1),
+               " seconds for ", nrow(schubertdat), " OTUs"))
+
 # plot % rejects vs FDR cutoff
@@
 print(p)
+```
 
+```{bash, goodrich-download, eval=FALSE}
+curl -O "https://zenodo.org/record/840333/files/ob_goodrich_results.tar.gz"
+tar -xzvf ob_goodrich_results.tar.gz
 ```
+
+```{r, read-ob-goodrich}
+# Read and clean ob_goodrich
+otu_path = paste0(wd, "ob_goodrich_results/RDP/ob_goodrich.otu_table.100.denovo.rdp_assigned")
+meta_path = paste0(wd, "ob_goodrich_results/ob_goodrich.metadata.txt")
+labels = c("H", "OB")
+
+dat <- read_and_clean_data(otu_path, meta_path, labels, "X")
+otu <- dat[[1]]
+meta <- dat[[2]]
+
+abun_otu <- t(t(otu) / rowSums(t(otu)))
+minabun <- min(abun_otu[abun_otu > 0])
+zeroabun = 1e-6
+abun_otu <- abun_otu + zeroabun
+```
+
+```{r, goodrich-pvals}
+case_idx <- meta$DiseaseState %in% c("OB")
+ctrl_idx <- meta$DiseaseState %in% c("H")
+
+# Calculate some basic covariates for each OTU
+goodrichdat <- calculate_covariates(otu, abun_otu)
+# Get the pvalues, effects, and SEs
+goodrich_pvals <- calculate_pvals(abun_otu, ctrl_idx, case_idx)
+goodrichdat <- merge(goodrichdat, goodrich_pvals, by="otu")
+```
+
+```{r}
+goodrichdat <- goodrichdat %>% 
+                 mutate(logpval = log10(pval)) %>%
+                 mutate(neglogpval = -log10(pval)) %>%
+                 mutate(logubiquity = log10(ubiquity)) %>%
+                 mutate(logmean_reads = log10(mean_reads))
+head(goodrichdat)
+```
+
+```{r}
+covcol <- "logubiquity"
+strat_hist(goodrichdat, pcol, covcol)
+```
+
+```{r, explore-goodrich-covariates}
+var <- "logubiquity"
+g <- ggplot(goodrichdat, aes(x = get(var), y = pval)) + geom_point()
+g
+```
+
+```{r}
+covcol <- "logmean_reads"
+strat_hist(goodrichdat, pcol, covcol)
+```
+
+```{r}
+g <- ggplot(goodrichdat, aes(x = get(covcol), y = pval)) + geom_point()
+g
+```
+
+```{r, test-fdrs-ubiquity}
+# Set up ubiquity as the first covariate to test.
+# Ranges from 0.001 to 0.4
+goodrichdat <- goodrichdat %>% 
+              mutate(zscore = effect_size/SE) %>%
+              mutate(ind_covariate = log10(ubiquity) )
+
+# Run all the tests and plot
+cutoffs <- c(0.01, 0.025, 0.05, 0.10, 0.20)
+resfile <- paste0(resdir, "ob_goodrich_results_",
+                  nrow(goodrichdat), ".RData")
+t1 <- proc.time()
+goodrich_results <- run_benchmarks(dat=goodrichdat, alphas=cutoffs)
+save(goodrich_results, file=resfile)
+message(paste0("Took ", round((proc.time()-t1)[3],1),
+               " seconds for ", nrow(goodrichdat), " OTUs"))
+
+# plot % rejects vs FDR cutoff
+goodrich_results <- goodrich_results %>% 
+  mutate(perc_reject = n_rejects / length(rownames(abun_otu)))
+
+p <- ggplot(goodrich_results, aes(alpha, perc_reject, color=method)) +
+  geom_point() +
+  geom_line() +
+  xlab("FDR cutoff") + ylab("Fraction of hypotheses rejected") +
+  ggtitle("OB Goodrich with log10(Ubiquity) Covariate") +
+  theme_classic()
+print(p)
+```

### 2017-10-06 simulation_studies/fdr_simulation_steph.Rmd
Add diagnostic and FDR plots for All Null sims

---
title: "Benchmarking FDR Methods - Null settings"
author: "Rafalab Journal Club Members"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load helper functions
RPROJ <- list(PROJHOME = normalizePath(getwd()))
source(file.path(RPROJ$PROJHOME, "R/simulation-helpers.R"))
source(file.path(RPROJ$PROJHOME, "../datasets/R/evaluation-helpers.R"))

m = 20000 # number of tests
```

# Introduction

The goal of this working group is to benchmark recent covariate-adjusted FDR methods. For 
complete details, check out our [benchmark-fdr GitHub repo](https://github.com/pkimes/benchmark-fdr).

In this R Markdown, we use a Monte Carlo simulation study to evaluate 
the `All nulls` simulation (all tests are simulated from the null distribution
i.e. no difference). We define

* $m$ = number of tests
* $M$ = Monte Carlo replications

Methods to evaluate:

- Bonferroni 
- BH
- IHW
- ASH
- BL
- Scott
- locfdr
- mixfdr

Metrics to evaluate:

- Number of discoveries
- Control FDR at $\alpha$
- Power
- Relationship between $\pi_0$ and covariate
- Assumptions: unimodal, etc

Load libraries: 
```{r, load-libraries, message=FALSE}
library(dplyr)
library(ggplot2)
library(cowplot)
library(doParallel)
```

Additional packages that should be installed are: 
```{r, eval=FALSE}
library(genefilter)

library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)
  ## to use IHWpaper::scott_fdrreg wrapper, need to install FDRreg v2.0 (only available github):   
  ## devtools::install_github(repo= 'jgscott/FDRreg', subdir='R_pkg/') 
library(FDRreg)
```

# The "All Nulls" simulation scenerio

Descriptive summary of simulation study:

- Simulate a "null" dataset using `du_ttest_sim()`:
    * Two standard normals with mean = 0, sd = 1. 
    * $\pi_0$ = 1
- Calculate t-statistic. 
- Independent covariate (independent of the p-value under the null hypothesis): 
    * Uninformative covariate: uniform between 0, 1. 
    * Informative covariate: pooled standard deviation
- Apply different methods to control for false discoveries

Generate example data set with an **uninformative covariate** and 
**informative covariate**: 
```{r example-simulate-null-data}
n_samples = 20
n_groups = 2

# uninformative covariate = uniform(0,1)
sim_df_uninform <- du_ttest_sim(m=m, pi0=1, effect_size=0, 
                       n_samples=n_samples, n_groups = 2, 
                       informative_ind_covariate = FALSE)

# informative covariate = pooled variance
sim_df_inform <- du_ttest_sim(m=m, pi0=1, effect_size=0, 
                       n_samples=n_samples, n_groups = 2, 
                       informative_ind_covariate = TRUE)

head(sim_df_inform)
```

Inside the simulated "null" dataset object:

- `H` is an indicator representing the true null (0) or alternative status (1)
- `test_statistic` is the t-test statistic
- `effect_size` is the difference in means
- `pval` is the $p$-value from the t-test 
- `ind_covariate` the independent covariate (can be informative or uninformative). Here it's informative
- `SE` is the standard error ($\frac{\sigma}{\sqrt{n}}$)

## Diagnostic plots

Need to check two things:

### Conditional independence property

1. Conditional independence property. Need to check if the covariate is independent of the $p$-value under the null hypothesis. In other words, the covariate itself is not informative for whether the test should be rejected or not. Can be checked by splitting histogram of p-values into groups based on covariate. If independent, the right tails of the distribution will be the same. If not independent, the tails may no longer look uniform.

#### Using the uninformative covariate

```{r example-diagnostic-plot-null-indepedence-uninform}
colnames(sim_df_uninform)[which(colnames(sim_df_uninform) == "pval")] <- "pvals"
strat_hist(sim_df_uninform, pval="pvals", covariate = "ind_covariate", maxy=2)
```

#### Using the informative covariate

```{r example-diagnostic-plot-null-indepedence-inform}
colnames(sim_df_inform)[which(colnames(sim_df_inform) == "pval")] <- "pvals"
strat_hist(sim_df_inform, pval="pvals", covariate = "ind_covariate", maxy=2)
```

**This does not mean the covariate is informative**. We use the second diagnostic plot to check this.

### Informative property

2. Informative property. Need to check if the covariate is informative or not. In other words, the covariate should be associated with the p-values. This is because under the alternative hypothesis, we expect to see an enrichment of rejections for certain values of the covariate (if it is indeed informative). If all histograms look the same, the covariate is uninformative, and its use will not lead to an increase in power.

#### Using the uninformative covariate

```{r example-diagnostic-plot-null-uninform}
rank_scatter(sim_df_uninform, pval="pvals", covariate = "ind_covariate", bins=100)
```

#### Using the informative covariate

```{r example-diagnostic-plot-null-inform}
rank_scatter(sim_df_inform, pval="pvals", covariate = "ind_covariate", bins=100)
```

Run following on odyssey and save output as RDS: 
```{r evaluate-methods-null-data, eval=FALSE}
library(dplyr)
library(ggplot2)
library(doParallel)

workingPath <- "/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/shicks/projects/benchmark-fdr"
source(file.path(workingPath, "simulation-helpers.R"))
source(file.path(workingPath, "evaluation-helpers.R"))

# set these to a smaller number temporarily 
m = 20000 # number of tests
M = 500 # number of MC reps

n_samples = 20
n_groups = 2

alphas = seq(0.01, 0.1, length.out = 10)

nCores <- 10
registerDoParallel(cores = nCores)
workers <- getDoParWorkers()
backend <- getDoParName()
version <- getDoParVersion()

# uninformative covariate (uniform[0,1])
df.out.uninform <- foreach(i = 1:M, .verbose=T) %dopar% {
  sim_df_uninform <- du_ttest_sim(m=m, pi0=1, effect_size=0, n_samples=n_samples, 
                                  n_groups=n_groups, informative_ind_covariate = FALSE)
  sim_runner(sim = sim_df_uninform, alphas = alphas, pvals = FALSE)
}

df.out.uninform <- dplyr::bind_rows(df.out.uninform)
saveRDS(df.out.uninform, file=file.path(workingPath, "Results-nullSims-uninform.RDS"))

# informative covariate (pooled variance)
df.out.inform <- foreach(i = 1:M, .verbose=T) %dopar% {
  sim_df_inform <- du_ttest_sim(m=m, pi0=1, effect_size=0, n_samples=n_samples, 
                                  n_groups=n_groups, informative_ind_covariate = TRUE)
  sim_runner(sim = sim_df_inform, alphas = alphas, pvals = FALSE)
}

df.out.inform <- dplyr::bind_rows(df.out.inform)
saveRDS(df.out.inform, file=file.path(workingPath, "Results-nullSims-inform.RDS"))
```

## Control for FDR and FWER at $\alpha$

Here we will plot the number of discoveries as a function 
of $\alpha$ (should be 0 false discoveries). We will also evaluate
methods control for false discovery rate (FDR) and family-wise 
error rate (FWER) at nominal $\alpha$ level ranging from [0.01, 0.1].

### Using the uninformative covariate
Load and plot results from RDS file: 
```{r, eval=TRUE}
# workingPath <- "/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/shicks/projects/benchmark-fdr"
# df.out.uninform <- readRDS(file=file.path(workingPath, "Results-nullSims-uninform.RDS"))
df.out.uninform <- readRDS(file=file.path(RPROJ$PROJHOME, "R/Results-nullSims-uninform.RDS"))

# Number of discoveries as a function of $\alpha$
df.out.uninform %>% dplyr::group_by(method, alpha) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
    ggplot(aes(x = alpha, y = n_rejects, color = method)) + 
         geom_line() + geom_point() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("Number of Discoveries") +
         theme(axis.title = element_text(face="bold") )

# False discovery rate as a function of $\alpha$
df.out.uninform %>% dplyr::group_by(method, alpha) %>%
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FDR, color = method)) + 
         geom_line() + geom_point() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FDR") +
         theme(axis.title = element_text(face="bold") )

# Family wise error rate as a function of $\alpha$
df.out.uninform %>% dplyr::group_by(method, alpha) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FWER, color = method)) + 
         geom_line() + geom_point() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FWER") +
         theme(axis.title = element_text(face="bold") )
```

### Using the informative covariate
Load and plot results from RDS file: 
```{r, eval=TRUE}
# workingPath <- "/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/shicks/projects/benchmark-fdr"
# df.out.inform <- readRDS(file=file.path(workingPath, "Results-nullSims-inform.RDS"))
df.out.inform <- readRDS(file=file.path(RPROJ$PROJHOME, "R/Results-nullSims-inform.RDS"))

# Number of discoveries as a function of $\alpha$
df.out.inform %>% dplyr::group_by(method, alpha) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
    ggplot(aes(x = alpha, y = n_rejects, color = method)) + 
         geom_line() + geom_point() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("Number of Discoveries") +
         theme(axis.title = element_text(face="bold") )

# False discovery rate as a function of $\alpha$
df.out.inform %>% dplyr::group_by(method, alpha) %>%
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FDR, color = method)) + 
         geom_line() + geom_point() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FDR") +
         theme(axis.title = element_text(face="bold") )

# Family wise error rate as a function of $\alpha$
df.out.inform %>% dplyr::group_by(method, alpha) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FWER, color = method)) + 
         geom_line() + geom_point() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FWER") +
         theme(axis.title = element_text(face="bold") )
```

# The "Varying $\pi_0$" simulation scenerio

Descriptive summary of simulation study:

- Simulate a dataset using `du_ttest_sim()`:
    * Two standard normals with difference in means of 1.5, sd = 1. 
    * $\pi_0$ varying between 0 and 1 (e.g. $\pi_0$= 0.7)
- Calculate t-statistic. 
- Independent covariate (independent of the p-value under the null hypothesis): 
    * Uninformative covariate: uniform between 0, 1. 
    * Informative covariate: pooled standard deviation
- Apply different methods to control for false discoveries

Generate example data set with an **uninformative covariate** and 
**informative covariate**: 
```{r example-simulate-varyingpi0-data}
n_samples = 20
n_groups = 2

# uninformative covariate = uniform(0,1)
sim_df_uninform <- du_ttest_sim(m=m, pi0=0.7, effect_size=1.5, 
                       n_samples=n_samples, n_groups = 2, 
                       informative_ind_covariate = FALSE)

# informative covariate = pooled variance
sim_df_inform <- du_ttest_sim(m=m, pi0=0.7, effect_size=1.5, 
                       n_samples=n_samples, n_groups = 2, 
                       informative_ind_covariate = TRUE)

head(sim_df_inform)
```

## Diagnostic plots

Need to check two things:

### Conditional independence property

1. Conditional independence property. Need to check if the covariate is independent of the $p$-value under the null hypothesis. In other words, the covariate itself is not informative for whether the test should be rejected or not. Can be checked by splitting histogram of p-values into groups based on covariate. If independent, the right tails of the distribution will be the same. If not independent, the tails may no longer look uniform.

#### Using the uninformative covariate

```{r example-diagnostic-plot-varyingpi0-indepedence-uninform, warning=FALSE}
colnames(sim_df_uninform)[which(colnames(sim_df_uninform) == "pval")] <- "pvals"
strat_hist(sim_df_uninform, pval="pvals", covariate = "ind_covariate", maxy=2)
```

#### Using the informative covariate

```{r example-diagnostic-plot-varyingpi0-indepedence-inform, warning=FALSE}
colnames(sim_df_inform)[which(colnames(sim_df_inform) == "pval")] <- "pvals"
strat_hist(sim_df_inform, pval="pvals", covariate = "ind_covariate", maxy=2)
```

**This does not mean the covariate is informative**. We use the second diagnostic plot to check this.

### Informative property

2. Informative property. Need to check if the covariate is informative or not. In other words, the covariate should be associated with the p-values. This is because under the alternative hypothesis, we expect to see an enrichment of rejections for certain values of the covariate (if it is indeed informative). If all histograms look the same, the covariate is uninformative, and its use will not lead to an increase in power.

#### Using the uninformative covariate

```{r example-diagnostic-plot-varyingpi0-uninform}
rank_scatter(sim_df_uninform, pval="pvals", covariate = "ind_covariate", bins=100)
```

#### Using the informative covariate

```{r example-diagnostic-plot-varyingpi0-inform}
rank_scatter(sim_df_inform, pval="pvals", covariate = "ind_covariate", bins=100)
```

Run following on odyssey and save output as RDS: 
```{r evaluate-methods-varyingpi0-data, eval=FALSE}
library(dplyr)
library(ggplot2)
library(doParallel)

workingPath <- "/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/shicks/projects/benchmark-fdr"
source(file.path(workingPath, "simulation-helpers.R"))
source(file.path(workingPath, "evaluation-helpers.R"))

# set these to a smaller number temporarily 
m = 20000 # number of tests
M = 500 # number of MC reps

n_samples = 20
n_groups = 2

alphas = seq(0.01, 0.1, length.out = 10)
pi0set <- c(0, 0.1, 0.3, 0.7, 0.9, 1)

nCores <- 10
registerDoParallel(cores = nCores)
workers <- getDoParWorkers()
backend <- getDoParName()
version <- getDoParVersion()

# uninformative covariate (uniform[0,1])
df.out.uninform <- foreach(i = 1:M, .verbose=T) %dopar% {
  dat <- NULL
  for(i in 1:length(pi0set)){
    sim_df_uninform <- du_ttest_sim(m=m, pi0=pi0set[i], effect_size=1.5, n_samples=n_samples, 
                                    n_groups=n_groups, informative_ind_covariate = FALSE)
    dat <- rbind(dat, data.frame(sim_runner(sim = sim_df_uninform, alphas = alphas, pvals = FALSE), 
                                 "pi0" = pi0set[i]))
  }
  dat
}

df.out.uninform <- dplyr::bind_rows(df.out.uninform)
saveRDS(df.out.uninform, file=file.path(workingPath, "Results-varyingpi0Sims-uninform.RDS"))

# informative covariate (pooled variance)
df.out.inform <- foreach(i = 1:M, .verbose=T) %dopar% {
    dat <- NULL
  for(i in 1:length(pi0set)){
    sim_df_inform <- du_ttest_sim(m=m, pi0=pi0set[i], effect_size=1.5, n_samples=n_samples, 
                                    n_groups=n_groups, informative_ind_covariate = TRUE)
    dat <- rbind(dat, data.frame(sim_runner(sim = sim_df_inform, alphas = alphas, pvals = FALSE), 
                                 "pi0" = pi0set[i]))
  }
  dat
}

df.out.inform <- dplyr::bind_rows(df.out.inform)
saveRDS(df.out.inform, file=file.path(workingPath, "Results-varyingpi0Sims-inform.RDS"))
```

## Control for FDR, FWER, Power at $\alpha$

Here we will plot the number of discoveries as a function 
of $\alpha$. We will also evaluate
methods control for false discovery rate (FDR), family-wise 
error rate (FWER) and Power at nominal $\alpha$ level 
ranging from [0.01, 0.1].

### Using the uninformative covariate
Load and plot results from RDS file: 
```{r, eval=FALSE}
# workingPath <- "/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/shicks/projects/benchmark-fdr"
# df.out.uninform <- readRDS(file=file.path(workingPath, "Results-varyingpi0Sims-uninform"))
df.out.uninform <- readRDS(file=file.path(RPROJ$PROJHOME, "R/Results-varyingpi0Sims-uninform"))

# Number of discoveries as a function of $\alpha$
df.out.uninform %>% dplyr::group_by(method, alpha, pi0) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
    ggplot(aes(x = alpha, y = n_rejects, color = method)) + 
         geom_line() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("Number of Discoveries") +
         theme(axis.title = element_text(face="bold") ) + 
  facet_grid(~ pi0)

# False discovery rate as a function of $\alpha$
df.out.uninform %>% dplyr::group_by(method, alpha, pi0) %>%
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FDR, color = method)) + 
         geom_line() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FDR") +
         theme(axis.title = element_text(face="bold") ) + 
  facet_grid(~ pi0)

# Family wise error rate as a function of $\alpha$
df.out.uninform %>% dplyr::group_by(method, alpha, pi0) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FWER, color = method)) + 
         geom_line() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FWER") +
         theme(axis.title = element_text(face="bold") ) + 
  facet_grid(~ pi0)

# Power as a function of $\alpha$
df.out.uninform %>% dplyr::group_by(method, alpha, pi0) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), Power = mean(power)) %>% 
      ggplot(aes(x = alpha, y = Power, color = method)) + 
         geom_line() + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("Power") +
         theme(axis.title = element_text(face="bold") ) + 
  facet_grid(~ pi0)

```

### Using the informative covariate
Load and plot results from RDS file: 
```{r, eval=FALSE}
# workingPath <- "/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/shicks/projects/benchmark-fdr"
# df.out.inform <- readRDS(file=file.path(workingPath, "Results-varyingpi0Sims-inform"))
df.out.inform <- readRDS(file=file.path(RPROJ$PROJHOME, "R/Results-varyingpi0Sims-inform"))

# Number of discoveries as a function of $\alpha$
df.out.inform %>% dplyr::group_by(method, alpha, pi0) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
    ggplot(aes(x = alpha, y = n_rejects, color = method)) + 
         geom_line() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("Number of Discoveries") +
         theme(axis.title = element_text(face="bold") ) + 
  facet_grid(~ pi0)

# False discovery rate as a function of $\alpha$
df.out.inform %>% dplyr::group_by(method, alpha, pi0) %>%
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FDR, color = method)) + 
         geom_line() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FDR") +
         theme(axis.title = element_text(face="bold") ) + 
  facet_grid(~ pi0)

# Family wise error rate as a function of $\alpha$
df.out.inform %>% dplyr::group_by(method, alpha, pi0) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FWER, color = method)) + 
         geom_line() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FWER") +
         theme(axis.title = element_text(face="bold") ) + 
  facet_grid(~ pi0)

# Power as a function of $\alpha$
df.out.inform %>% dplyr::group_by(method, alpha, pi0) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), Power = mean(power)) %>% 
      ggplot(aes(x = alpha, y = Power, color = method)) + 
         geom_line() + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("Power") +
         theme(axis.title = element_text(face="bold") ) + 
  facet_grid(~ pi0)

```

### 2017-10-06 simulation_studies/R/simulation-helpers.R

@@ -9,6 +9,9 @@
 #' @param effect_size expected mean difference
 #' @param n_samples number of total samples, should be multiple of `n_groups` (default = 20)
 #' @param n_groups number of groups in comparison [NOT SUPPORTED] (default = 2)
+#' @param informative_ind_covariate default is TRUE. If informative covariate (TRUE), 
+#' then it's the pooled variance (in this case, it's the sd across all samples). 
+#' If uninformative covariate (FALSE), then we randomly sample from Uniform[0,1].
 #' @param seed integer seed for random number generator, ignored if NULL (default = NULL) 
 #'
 #' @return
@@ -17,8 +20,8 @@
 #' * `test_statistic`: t-test statistic
 #' * `effect_size`: group mean difference
 #' * `pval`: t-test p-value
-#' * `ind_covariate`: in this case, it's standard deviation (sd) across all samples 
-#' * `SE`: standard error across all samples (sd / sqrt(n))
+#' * `ind_covariate`: the independent covariate 
+#' * `SE`: standard error across all samples (estimated using the ratio of difference in means / t-test statistic)
 #'
 #' @details
 #' This function builds on `IHWpaper::du_ttest_sim` originally written by Nikos Ignatiadis
@@ -29,7 +32,8 @@
 #' @importFrom genefilter rowtests
 #' @importFrom genefilter rowSds
 #' @author Stephanie Hicks
-du_ttest_sim <- function(m, pi0, effect_size, n_samples = 20, n_groups = 2, seed = NULL) {
+du_ttest_sim <- function(m, pi0, effect_size, n_samples = 20, n_groups = 2, 
+                         informative_ind_covariate = TRUE, seed = NULL) {
     if (!is.null(seed)) {
         set.seed(seed)
     }
@@ -44,7 +48,11 @@ du_ttest_sim <- function(m, pi0, effect_size, n_samples = 20, n_groups = 2, seed
     H[false_nulls] <- 1
     gF <- factor(rep(1:n_groups, each=n_samples/n_groups))
     t_test <- genefilter::rowttests(z_table, gF)
-    sds <- genefilter::rowSds(z_table) # pooled var reduces to unpooled var b/c same sample sizes across groups 
+    if(informative_ind_covariate){
+      sds <- genefilter::rowSds(z_table) # pooled var reduces to unpooled var b/c same sample sizes across groups 
+    } else {
+      sds <- runif(m, 0, 1)
+    }
[…]

### 2017-10-19 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
Add skeleton for SummarizedBenchmark usage on GWAS

---
title: "Comparison of FDR methods on GWAS with SummarizedBenchmark"
author: "Keegan Korthauer"
date: "10/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Summary

In the previous document `GWAS.Rmd`, we identified three GWAS datasets for FDR
benchmarking. Here we will make use of the new `SummarizedBenchmark` package to
perform the benchmarking in a standardized way.

### Workspace setup

```{r, wkspace-setup}
# project directory & data/results folders
setwd("/n/irizarryfs01_backed_up/kkorthauer/FDR")
datdir <- "/n/irizarryfs01_backed_up/kkorthauer/FDR/DATA/GWAS/"
resdir <- "/n/irizarryfs01_backed_up/kkorthauer/FDR/RESULTS/GWAS/"

# wrangling & plotting tools
library(data.table)
library(readxl)
library(dplyr)
library(ggplot2)
library(magrittr)
library(R.utils)

# benchmark methods
library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)
library(FDRreg)

# comparison tools/functions
library(SummarizedBenchmark)
source("benchmark-fdr/simulation_studies/R/simulation-helpers.R")
sourceDirectory("benchmark-fdr/datasets/R/")
```

### GWAS meta-analysis for BMI (included in Boca-Leek)

Here we download and analyze the GWAS BMI dataset analyzed in the Boca-Leek paper.

#### Data download

```{bash, GWAS1-download, eval=FALSE}
curl -O "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
```

Next we unzip the file, and delete the files we won't use (they provide results
subsetted by ancestry and sex), keeping only the file with all ancestries and
both sexes `BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt`. Note that it seems that
the Boca-Leek paper actually used the 'EuropeanOnly' subset, but it's not clear
why (to me), and there's no reason we need to do the same.

```{bash, GWAS1-unzip, eval=FALSE}
unzip BMI.SNPadjSMK.zip
rm BMI.SNPadjSMK.zip
rm BMI.SNPadjSMK.Women.EuropeanOnly.txt
rm BMI.SNPadjSMK.Women.AllAncestry.txt
rm BMI.SNPadjSMK.Men.EuropeanOnly.txt
rm BMI.SNPadjSMK.Men.AllAncestry.txt
rm BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt
```

Next, we'll read in the unzipped `.txt` file into R and verify that it contains
the necessary inputs to run the FDR benchmark comparisons.

```{r, GWAS1-verify-contents}
bmi <- fread(paste0(datdir, "BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt"),
             header=TRUE)
dim(bmi)
head(bmi)
```

It looks like we have:

- `p_value`:p-value
- `effect`:effect size
- `stderr`:standard error
- additional covariates: 
    - `N`: Number of samples with this SNP - BMI association measured
    - `Freq_Allele1_HapMapCEU`: Allele Frequency of Allele1, as measured in HapMapCEU

for 2,458,133 SNPs.

Before moving on, we'll rename the relevant columns so that they will use common terms
across the different datasets. We'll also add a 'zscore' column, and subset on a 
toy example before moving on to the full dataset.

```{r, GWAS1-reformat}
bmi <- bmi %>% 
        dplyr::rename(pval = p_value,
               SE = stderr,
               effect_size = effect,
               ind_covariate = N) %>%
        mutate(zscore = effect_size/SE)

### experimenting on a subset of the full dataset for now
bmi <- bmi[1:10000,]
```

#### Check Covariate Diagnostics

```{r, GWAS1-diag, fig.width=10, fig.height=3.5}
strat_hist(bmi, pval="pval", covariate="ind_covariate", maxy=3)
rank_scatter(bmi, pval="pval", covariate="ind_covariate")
```

#### Set up BenchDesign object

First, we'll create an object of `BenchDesign` class to hold the data.

**Still to-do: consider multiple covariates for methods that allow this.**

```{r, GWAS1-benchdesign}
bd <- BenchDesign(bmi)
```

Next, we'll add the benchmark methods to the `BenchDesign` object.

```{r, GWAS1-methods}
# This should maybe be sourced from a common script that everyone uses...
bd %<>% addBMethod(blabel = "BH", bfunc = p.adjust,
                 p = pval, method = "BH")

bd %<>% addBMethod("qv", qvalue::qvalue,
                  p = pval,
                  bpost = function(x) { x$qvalues })

bd %<>% addBMethod("ihw", IHW::ihw,
                  pvalues = pval,
                  covariates = IHW::groups_by_filter(ind_covariate, 5),
                  alpha = 0.1,
                  bpost = IHW::adj_pvalues)
```

Now, we're ready to construct the `SummarizedBenchmark` object, which will run
the functions specified in each method (these are actually sourced in from the
helper scripts).

```{r, GWAS1-sb}
resfile <- paste0(resdir, "bmi_summarizedBenchmark_",
                  nrow(bmi), ".RData")
if (!file.exists(resfile)){
  t1 <- proc.time()
  sb <- bd %>% buildBench
  metadata(sb)$data_download_link <- "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
  save(sb, file=resfile)
  message(paste0("Took ", round((proc.time()-t1)[3],1),
                 " seconds for ", nrow(bmi), " SNPs."))
}else{
  load(resfile)
}

```

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, GWAS1-metrics}
# rename assay to qvalue
names(assays(sb)) <- names(rowData(sb)) <- names(sb@performanceMetrics) <- "qvalue"
sb <- addDefaultMetrics(sb)
sb

head(assay(sb))
colData(sb)
sb@performanceMetrics[[1]]

# estimate perf metrics
sb <- estimatePerformanceMetrics(sb, alpha=seq(0.01, 0.10, 0.01), addColData=TRUE)
colData( sb )
```

Now, we'll plot the results.

```{r, GWAS1-plot, results="hide"}
## to go here - standardized plotting functions from SummarizedBenchmarks -
## to discuss at meeting 10/19

# alpha versus nrejects

# upset plot

```

And we'll clean up the workspace before moving on to the next dataset.

```{r, GWAS1-cleanup}
rm(bmi)
rm(bmi_results)
```

### 2017-10-19 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
Add skeleton for SummarizedBenchmark usage on GWAS

@@
 
-Before moving on, we'll rename the relevant columns so that they will use common terms
-across the different datasets. We'll also add a 'zscore' column, and subset on a 
-toy example before moving on to the full dataset.
+#### Check Covariate Diagnostics
 
-```{r, GWAS1-reformat}
+```{r, GWAS1-diag, fig.width=10, fig.height=3.5}
+strat_hist(bmi, pval="pval", covariate="ind_covariate", maxy=3)
+rank_scatter(bmi, pval="pval", covariate="ind_covariate")
+```
+
+#### Set up BenchDesign object
+
+First, we'll create an object of `BenchDesign` class to hold the data. Before 
+doing so, we'll rename the relevant columns so that they will use common terms
+across the different datasets. We'll also add a 'zscore' column.
+
+**Still to-do: consider multiple covariates for methods that allow this.**
+
+```{r, GWAS1-benchdesign}
 bmi <- bmi %>% 
@@
 bmi <- bmi[1:10000,]
-```
 
-#### Check Covariate Diagnostics
-
-```{r, GWAS1-diag, fig.width=10, fig.height=3.5}
-strat_hist(bmi, pval="pval", covariate="ind_covariate", maxy=3)
-rank_scatter(bmi, pval="pval", covariate="ind_covariate")
-```
-
-#### Set up BenchDesign object
-
-First, we'll create an object of `BenchDesign` class to hold the data.
-
-**Still to-do: consider multiple covariates for methods that allow this.**
-
-```{r, GWAS1-benchdesign}
 bd <- BenchDesign(bmi)

### 2017-10-19 datasets/GWAS/GWAS-knit.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+#SBATCH -J knitGWAS                 # A single job name for the array
+#SBATCH -n 1                        # Number of cores
+#SBATCH -N 1                        # All cores on one machine
+#SBATCH -p irizarry,serial_requeue  # Partition
+#SBATCH --mem 20000                 # Memory request 
+#SBATCH -t 0-1:00                   # Maximum execution time (D-HH:MM)
+#SBATCH -o SLURM/render-%j.out  # Standard output
+#SBATCH -e SLURM/render-%j.err  # Standard error
+ 
+module load rstudio
+export SLURM_NTASKS
+export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/rstudio/0.98.1103-fasrc01/bin/pandoc/"
+
+# change filename to Rmd to be knitted. ncores is passed in through environment
+# variable SLURM_NTASKS
+R -e "rmarkdown::render('/n/irizarryfs01_backed_up/kkorthauer/FDR/benchmark-fdr/datasets/GWAS/GWAS-SummarizedBenchmark.Rmd')"

### 2017-10-20 datasets/simulations/analysis-01-plot.Rmd
Fix sim core to point to renamed settings functions

---
title: "Benchmarking FDR Methods - Alternative Settings"
author: "Rafalab Journal Club Members"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    highlight: tango
    number_sections: true
    code_folding: hide
---

<!-- ######################################################################## -->
<!-- preliminaries                                                            --> 
<!-- ######################################################################## -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, install-benchmarked-methods, eval=FALSE}
## from CRAN
install.packages("ashr")

## from Bioconductor
BiocInstaller::biocLite("IHW")
BiocInstaller::biocLite("qvalue")
BiocInstaller::biocLite("swfdr")

## from GitHub
devtools::install_github('jgscott/FDRreg', subdir="R_pkg/")
devtools::install_github("nignatiadis/IHWpaper")
```

<!-- ######################################################################## -->
## Workspace Setup
<!-- ######################################################################## -->

```{r, load-libraries, message=FALSE}
library("dplyr")
library("ggplot2")
library("data.table")
library("SummarizedBenchmark")
```

```{r, load-helper-functions}
source("analysis-01-plotters.R")
source("../datasets/R/evaluation-helpers.R")
```

<!-- ######################################################################## -->
## Plot Performance Metrics (100 replications)
<!-- ######################################################################## -->

```{r}
## list of results from 100 replication experiments
fl <- list.files("data", pattern="results-.*-M100.RDS", full.names=TRUE)

## generate same plots for all parameter settings
for (f in fl) {
    ## read process data
    res <- readRDS(f)
    tsb <- standardize_results(res, alpha = seq(0.01, 0.10, 0.01))
    tsbs <- tsb %>%
        filter(!(blabel %in% c("bl-df02", "bl-df04", "bl-df05", "bonf")))

    ## look at FPR for Boca-Leek variations
    gp <- tsb %>%
        filter(grepl("^bl-", blabel)) %>%
        average_plots("FPR")
    ggsave(paste0("plots/FPR-bocaleek-",
                  gsub("(.*)\\.RDS", "\\1", basename(f)), ".png"), gp,
           width=7, height=6, units="in", dpi=200)

    ## look at FPR for IHW variations
    gp <- tsb %>%
        filter(grepl("^ihw-", blabel)) %>%
        average_plots("FPR", merge_ihw = FALSE)
    ggsave(paste0("plots/FPR-ihw-",
                  gsub("(.*)\\.RDS", "\\1", basename(f)), ".png"), gp,
           width=7, height=6, units="in", dpi=200)

    ## make plots for all metrics
    for (im in unique(tsbs$performanceMetric)) {
        gp <- average_plots(tsbs, im, clean_names = TRUE)
        ggsave(paste0("plots/", im, "-", gsub("(.*)\\.RDS", "\\1", basename(f)), ".png"), gp,
               width=7, height=6, units="in", dpi=200)
    }
}
```

<!-- ######################################################################## -->
## Plot Adjusted P-Values (1 replication)
<!-- ######################################################################## -->

```{r}
## list of results from 1 replication experiments
fl <- list.files("data", pattern = "results-.*-M1.rdata", full.names = TRUE)

## generate same plots for all parameter settings
for (f in fl) {
    ## read process data
    load(f)
    sba <- data.frame(assay(sb))

    ## for plots, only interested in Boca-Leek w/ 3 dof, IHw w/ alpha=0.05
    sbas <- select(sba, -c(`bl.df02`, `bl.df04`, `bl.df05`, `bonf`,
                           `ihw.a01`:`ihw.a04`, `ihw.a06`:`ihw.a10`))

    ## adjusted p-values vs. effect size
    gp <- single_plots(sbas, sim_df, "effect_size", TRUE)
    ggsave(paste0("plots/effectsize-adjp-", gsub("(.*)\\.rdata", "\\1", basename(f)), ".png"),
           gp, width=10, height=9, units="in", dpi=200)

    ## adjusted p-values vs. independent covariate (uniformative)
    gp <- single_plots(sbas, sim_df, "ind_covarate", TRUE)
    ggsave(paste0("plots/indcov-adjp-", gsub("(.*)\\.rdata", "\\1", basename(f)), ".png"),
           gp, width=10, height=9, units="in", dpi=200)

    ## adjusted p-values vs. standard error
    gp <- single_plots(sbas, sim_df, "SE", TRUE)
    ggsave(paste0("plots/SE-adjp-", gsub("(.*)\\.rdata", "\\1", basename(f)), ".png"),
           gp, width=10, height=9, units="in", dpi=200)
}
```

### 2017-10-20 datasets/simulations/psim-core.R

@@ -163,7 +163,7 @@ bd %<>% addBMethod("scott-empirical",
 ## generate simulation settings
 ## ##############################################################################
 
-settings <- define_settings(setting_base, setting_vparam, setting_icparam)
+settings <- psim_settings(setting_base, setting_vparam, setting_icparam)
 
 
 ## ##############################################################################

### 2017-10-20 datasets/simulations/tsim-core.R

@@ -10,13 +10,13 @@
 #'
 #' For a specified simulation settings ('setting_vparam' + 'setting_icparam'),
 #' this script applied FDR correction to results from a collection of independent
-#' t-tests. For more deetails on the simulation settings, see the `define_settings`
+#' t-tests. For more details on the simulation settings, see the `tsim_settings`
 #' function defined in `sim-core-settings.R`. 
 #'
 #' This script works with the following other scripts:
 #' - `R/du_tsim.R`: `du_tsim`
 #'     This function is used for actually simulating t-tests.
-#' - `sim-core-settings.R`: `define_settings`
+#' - `sim-core-settings.R`: `tsim_settings`
 #'     This function is used for defining the settings to be passed to
 #'     `du_tsim` for the simulations.
 #' - `tsim-submit.R`: 
@@ -176,7 +176,7 @@ bd %<>% addBMethod("scott-empirical",
 ## generate simulation settings
 ## ##############################################################################
 
-settings <- define_settings(setting_base, setting_vparam, setting_icparam)
+settings <- tsim_settings(setting_base, setting_vparam, setting_icparam)
 
 
 ## ##############################################################################

### 2017-10-20 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
Update to use common bench design function

--- 2017-10-19 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
@@
 library(R.utils)
+library(cowplot)
 
@@
 library(SummarizedBenchmark)
-source("benchmark-fdr/simulation_studies/R/simulation-helpers.R")
 sourceDirectory("benchmark-fdr/datasets/R/")
@@
 
-First, we'll create an object of `BenchDesign` class to hold the data.
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object.
 
@@
 ```{r, GWAS1-benchdesign}
-bd <- BenchDesign(bmi)
-```
-
-Next, we'll add the benchmark methods to the `BenchDesign` object.
-
-```{r, GWAS1-methods}
-# This should maybe be sourced from a common script that everyone uses...
-bd %<>% addBMethod(blabel = "BH", bfunc = p.adjust,
-                 p = pval, method = "BH")
-
-bd %<>% addBMethod("qv", qvalue::qvalue,
-                  p = pval,
-                  bpost = function(x) { x$qvalues })
-
-bd %<>% addBMethod("ihw", IHW::ihw,
-                  pvalues = pval,
-                  covariates = IHW::groups_by_filter(ind_covariate, 5),
-                  alpha = 0.1,
-                  bpost = IHW::adj_pvalues)
+bd <- initializeBenchDesign()
 ```
@@
 
-```{r, GWAS1-sb}
+```{r, GWAS1-sb, results="hide"}
 resfile <- paste0(resdir, "bmi_summarizedBenchmark_",
@@
   t1 <- proc.time()
-  sb <- bd %>% buildBench
+  sb <- bd %>% buildBench(data=bmi)
   metadata(sb)$data_download_link <- "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
@@
 ```{r, GWAS1-plot, results="hide"}
-## to go here - standardized plotting functions from SummarizedBenchmarks -
+#
+
+## to add here - standardized plotting functions from SummarizedBenchmarks -
 ## to discuss at meeting 10/19
@@
 rm(bmi)
-rm(bmi_results)
+rm(bd)
+rm(sb)
 ```

### 2017-10-20 datasets/simulations/analysis-01-plot.Rmd
Update analysis to point to new dir structure

@@
 <!-- ######################################################################## -->
-<!-- preliminaries                                                            --> 
-<!-- ######################################################################## -->
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
-
-```{r, install-benchmarked-methods, eval=FALSE}
-## from CRAN
-install.packages("ashr")
-
-## from Bioconductor
-BiocInstaller::biocLite("IHW")
-BiocInstaller::biocLite("qvalue")
-BiocInstaller::biocLite("swfdr")
-
-## from GitHub
-devtools::install_github('jgscott/FDRreg', subdir="R_pkg/")
-devtools::install_github("nignatiadis/IHWpaper")
-```
-
-<!-- ######################################################################## -->
 ## Workspace Setup
@@
 ```{r, load-helper-functions}
-source("analysis-01-plotters.R")
-source("../datasets/R/evaluation-helpers.R")
-```
-
-<!-- ######################################################################## -->
-## Plot Performance Metrics (100 replications)
-<!-- ######################################################################## -->
-
-```{r}
-## list of results from 100 replication experiments
-fl <- list.files("data", pattern="results-.*-M100.RDS", full.names=TRUE)
-
-## generate same plots for all parameter settings
-for (f in fl) {
-    ## read process data
-    res <- readRDS(f)
-    tsb <- standardize_results(res, alpha = seq(0.01, 0.10, 0.01))
-    tsbs <- tsb %>%
-        filter(!(blabel %in% c("bl-df02", "bl-df04", "bl-df05", "bonf")))
-
-    ## look at FPR for Boca-Leek variations
-    gp <- tsb %>%
-        filter(grepl("^bl-", blabel)) %>%
-        average_plots("FPR")
-    ggsave(paste0("plots/FPR-bocaleek-",
-                  gsub("(.*)\\.RDS", "\\1", basename(f)), ".png"), gp,
-           width=7, height=6, units="in", dpi=200)
-
-    ## look at FPR for IHW variations
-    gp <- tsb %>%
-        filter(grepl("^ihw-", blabel)) %>%
-        average_plots("FPR", merge_ihw = FALSE)
-    ggsave(paste0("plots/FPR-ihw-",
-                  gsub("(.*)\\.RDS", "\\1", basename(f)), ".png"), gp,
-           width=7, height=6, units="in", dpi=200)
-
-    ## make plots for all metrics
-    for (im in unique(tsbs$performanceMetric)) {
-        gp <- average_plots(tsbs, im, clean_names = TRUE)
-        ggsave(paste0("plots/", im, "-", gsub("(.*)\\.RDS", "\\1", basename(f)), ".png"), gp,
-               width=7, height=6, units="in", dpi=200)
-    }
-}
+source("../R/plotsim_standardize.R")
+source("../R/plotsim_average.R")
+source("../R/plotsim_single.R")
 ```
@@
 ```{r}
-## list of results from 1 replication experiments
-fl <- list.files("data", pattern = "results-.*-M1.rdata", full.names = TRUE)
+## list of results from 100 replication experiments
+dl <- list.dirs("data-tsim/M1", recursive=FALSE, full.names=TRUE)
+
+for (dd in dl) {
+    fl <- list.files(dd, pattern=".rdata", full.names=TRUE)
+
+    outp <- gsub("data", "plots", dd)
+    dir.create(outp, showWarnings=FALSE, recursive=TRUE)
 
@@
         ## adjusted p-values vs. effect size
-    gp <- single_plots(sbas, sim_df, "effect_size", TRUE)
-    ggsave(paste0("plots/effectsize-adjp-", gsub("(.*)\\.rdata", "\\1", basename(f)), ".png"),
+        gp <- plotsim_single(sbas, sim_df, "effect_size", TRUE)
+        ggsave(paste0(outp, "/effectsize-adjp-", gsub("(.*)\\.rdata", "\\1", basename(f)), ".png"),
                gp, width=10, height=9, units="in", dpi=200)
@@
         ## adjusted p-values vs. independent covariate (uniformative)
-    gp <- single_plots(sbas, sim_df, "ind_covarate", TRUE)
-    ggsave(paste0("plots/indcov-adjp-", gsub("(.*)\\.rdata", "\\1", basename(f)), ".png"),
+        gp <- plotsim_single(sbas, sim_df, "ind_covariate", TRUE)
+        ggsave(paste0(outp, "/indcov-adjp-", gsub("(.*)\\.rdata", "\\1", basename(f)), ".png"),
                gp, width=10, height=9, units="in", dpi=200)
@@
         ## adjusted p-values vs. standard error
-    gp <- single_plots(sbas, sim_df, "SE", TRUE)
-    ggsave(paste0("plots/SE-adjp-", gsub("(.*)\\.rdata", "\\1", basename(f)), ".png"),
+        gp <- plotsim_single(sbas, sim_df, "SE", TRUE)
+        ggsave(paste0(outp, "/SE-adjp-", gsub("(.*)\\.rdata", "\\1", basename(f)), ".png"),
                gp, width=10, height=9, units="in", dpi=200)
     }
+}
 ```
+
+<!-- ######################################################################## -->
+## Plot Performance Metrics (100 replications)
+<!-- ######################################################################## -->
+
+```{r}
+## list of results from 100 replication experiments
+dl <- list.dirs("data-tsim/M100", recursive=FALSE, full.names=TRUE)
+
+for (dd in dl) {
+    fl <- list.files(dd, pattern=".rds", full.names=TRUE)
+
+    outp <- gsub("data", "plots", dd)
+    dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+    ## generate same plots for all parameter settings
+    for (f in fl) {
+        ## read process data
+        res <- readRDS(f)
+        tsb <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+        tsbs <- tsb %>%
+            filter(!(blabel %in% c("bl-df02", "bl-df04", "bl-df05", "bonf")))
+
+        ## look at FPR for Boca-Leek variations
+        gp <- tsb %>%
+            filter(grepl("^bl-", blabel)) %>%
+            plotsim_average("FPR")
+        ggsave(paste0(outp, "/FPR-bocaleek-",
+                      gsub("(.*)\\.rds", "\\1", basename(f)), ".png"), gp,
+               width=7, height=6, units="in", dpi=200)
+
+        ## look at FPR for IHW variations
+        gp <- tsb %>%
+            filter(grepl("^ihw-", blabel)) %>%
+            plotsim_average("FPR", merge_ihw = FALSE)
+        ggsave(paste0(outp, "/FPR-ihw-",
+                      gsub("(.*)\\.rds", "\\1", basename(f)), ".png"), gp,
+               width=7, height=6, units="in", dpi=200)
+
+        ## make plots for all metrics
+        for (im in unique(tsbs$performanceMetric)) {
+            gp <- plotsim_average(tsbs, im, clean_names = TRUE)
+            ggsave(paste0(outp, "/", im, "-", gsub("(.*)\\.rds", "\\1", basename(f)), ".png"), gp,
+                   width=7, height=6, units="in", dpi=200)
+        }
+    }
+}
+```

### 2017-10-20 datasets/R/plotsim_single.R
Handle NAs when filtering negative adjusted p-values

@@ -52,7 +52,7 @@ plotsim_single <- function(sba, df, strat = "effect_size", clean_names = FALSE) 
     }
 
     ## check for negative adjusted p-values and throw warning; still plot with 0s
-    if (any(sba$adjp < 0)) {
+    if (any(sba$adjp < 0, na.rm=TRUE)) {
         negp <- which(sba$adjp < 0)
         nneg <- length(negp)
         mneg <- paste(unique(sba$method[negp]), collapse=",")

### 2017-10-20 datasets/simulations/analysis-01-plot.Rmd
Add p-value vs. covariate plots

@@
 library("ggplot2")
+library("viridis")
 library("data.table")
@@
 source("../R/plotsim_single.R")
+source("../R/rank_scatter.R")
+source("../R/strat_hist.R")
 ```
@@
                gp, width=10, height=9, units="in", dpi=200)
+
+        ## covariate rank vs. nlp (unadjusted)
+        gp <- rank_scatter(sim_df, "pval", "ind_covariate") +
+            scale_fill_viridis(option="viridis", direction=-1) +
+            ggtitle("Informative Covariate Check")
+        ggsave(paste0(outp, "/truep-heatmap-", gsub("(.*)\\.rdata", "\\1", basename(f)), ".png"),
+               gp, width=10, height=6, units="in", dpi=200)
+
+        ## stratified covariate vs. nlp (unadjusted)
+        gp <- strat_hist(sim_df, "pval", "ind_covariate", maxy=5) +
+            ggtitle("Informative Covariate Check")
+        ggsave(paste0(outp, "/truep-barchart-", gsub("(.*)\\.rdata", "\\1", basename(f)), ".png"),
+               gp, width=10, height=4, units="in", dpi=200)
     }

### 2017-10-20 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
Fix naming of zscore to test_statistic

@@
 
-```{r, wkspace-setup}
+```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
 # project directory & data/results folders
@@
 Before moving on, we'll rename the relevant columns so that they will use common terms
-across the different datasets. We'll also add a 'zscore' column, and subset on a 
+across the different datasets. We'll also add a 'test_statistic' column, and subset on a 
 toy example before moving on to the full dataset.
@@
                ind_covariate = N) %>%
-        mutate(zscore = effect_size/SE)
+        mutate(test_statistic = effect_size/SE)
 
 ### experimenting on a subset of the full dataset for now
-bmi <- bmi[1:10000,]
+bmi <- bmi[1:50000,]
 ```
@@
 
-```{r, GWAS1-plot, results="hide"}
-#
+```{r, GWAS1-plot, results="hide", width=15, height=12}
+# plot adjusted p-values by effect_size (diagnostic from ASH paper)
+plotsim_single(data.frame(assay(sb)), bmi, strat = "effect_size")
 

### 2017-10-20 datasets/GSEA/gsea.Rmd
Initial GSEA analysis

---
title: "GSEA datasets for FDR benchmarking"
author: "Alejandro Reyes (on behalf of the Rafalab Benchmarking Consortium)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Differential expression analysis with DESeq2

Downloading *DESeqDataSet* if not present locally,

```{r, message=FALSE}

file <- "dsd.RData"
if(!file.exists( file ) ){
    download.file("https://www.dropbox.com/s/rdmv3tyqt4g37b0/dsd.RData?dl=1",destfile="dsd.RData")
}
load(file)

library(DESeq2)
levels(colData(dsdObject)$individual) <- gsub("GTEX-", "",levels(colData(dsdObject)$individual))

```

Specify model matrix design and run DESeq2 to get differentially expressed genes.

```{r deseq2}

library(EnsDb.Hsapiens.v86)
gnType <-  genes(EnsDb.Hsapiens.v86, column="gene_biotype")
dsdObject <- dsdObject[rownames(dsdObject) %in% gnType[gnType$gene_biotype == "protein_coding"]$gene_id,]
dsdObject <- DESeq(dsdObject)
res <- results( dsdObject, lfcThreshold=1 )

```

Performing a gene set enrichment analysis with goseq.

```{r goseq, message=FALSE}

library(goseq)

file="goSets.RData"
if(!file.exists(file)){
  stopifnot( require( biomaRt ) )
  mart <- useMart( "ensembl", "hsapiens_gene_ensembl" )
  goSets <- getBM( c("ensembl_gene_id", "go_id"), mart=mart, filters="ensembl_gene_id", values=rownames(res))
  goSets <- goSets[!nchar( goSets$go_id ) == 0,]
  goSets <- with( goSets, split( go_id, ensembl_gene_id ) )
  save(goSets, file=file)
}else{
  load(file)
}

genes <- as.numeric( res$padj < 0.1 )
names( genes ) <- rownames( res )

### getting median transcript length
txByGene <- transcriptsBy(EnsDb.Hsapiens.v86, "gene")
geneLength <- sapply( width( txByGene ), median )
geneLength <- geneLength[names(genes)]
genes[is.na( genes )] <- 0

any( is.na(genes) )
stopifnot(!any(is.na(genes)))
pwf <- nullp(genes, bias.data=geneLength )
goRes <- goseq( pwf, gene2cat=goSets )

```

```{r, fig.height=3.5, fig.width=4, message=FALSE}
library(ggplot2)
library(magrittr)
library(cowplot)

ggplot( goRes,
       aes( log10(numInCat), -log10(over_represented_pvalue) ) ) +
    geom_hex(aes( alpha=log2(..count..)), fill="black", color="black", bins=100, show.legend=FALSE) +
    ylim(0, 12) + ggtitle("Over-represented gene sets") +
    xlab(expression(log[10]~"(# of genes)")) +
    ylab(expression(-log[10]~"(p-value)"))

```

```{r, message=FALSE}
library(IHW)
goRes$groups <- groups_by_filter( goRes$numInCat, 10 )

## filter numInCat below only used for plotting
## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
#dplyr:::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)

dplyr:::filter( goRes, numInCat > 10, numDEInCat > 0 ) %>%
  dplyr:::mutate( groups = groups_by_filter( numInCat, 4 ) ) %>%
  ggplot(aes(x=over_represented_pvalue)) + 
  geom_histogram(binwidth = 0.025, boundary = 0) +
  facet_wrap( ~ groups, nrow = 2) +
  ggtitle("Distribution of p-values") +
  ylab("Frequency") + xlab("p-value")

```

 Generating the *SummarizedBenchmark* object:

```{r}

library(R.utils)
sourceDirectory("../R/")

designBen <- initializeBenchDesign()
res <- dplyr:::select( goRes, c("over_represented_pvalue", "numInCat") ) %>%
  dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
res$pval <- pmin( res$pval, 1 )

sGSEA <- designBen %>% buildBench( res )
assayNames(sGSEA) <- names( rowData(sGSEA) ) <-  names(sGSEA@performanceMetrics) <- "qvalue"

sGSEA <- addDefaultMetrics( sGSEA )

estimatePerformanceMetrics( sGSEA, alpha=seq(0, 0.6, 0.1), tidy=TRUE ) %>%
  dplyr:::filter(pkg_name == "IHW") %>%
  head

  dplyr:::select(blabel, key, value, assay, performanceMetric, alpha ) %>%
  dplyr:::filter(blabel!="unadjusted") %>%
  ggplot( aes(alpha, value, col=blabel) ) +
  geom_line() +
  geom_point()

```

### 2017-10-25 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
Add results plots (overlap and nrejects)

@@
 author: "Keegan Korthauer"
-date: "10/19/2017"
+date: '`r format(Sys.Date(), "%B %d, %Y")`'
 output: html_document
@@
 
-for 2,458,133 SNPs.
+for `r nrow(bmi)` SNPs.
 
@@
 ### experimenting on a subset of the full dataset for now
-bmi <- bmi[1:50000,]
+bmi <- bmi[1:500000,]
 ```
@@
 # rename assay to qvalue
-names(assays(sb)) <- names(rowData(sb)) <- names(sb@performanceMetrics) <- "qvalue"
+assayNames(sb) <- "qvalue"
+names(rowData(sb)) <- names(sb@performanceMetrics) <- "qvalue"
 sb <- addDefaultMetrics(sb)
-sb
 
 head(assay(sb))
-colData(sb)
+head(colData(sb))
 sb@performanceMetrics[[1]]
 
-# estimate perf metrics
-sb <- estimatePerformanceMetrics(sb, alpha=seq(0.01, 0.10, 0.01), addColData=TRUE)
-colData( sb )
+# estimate perf metrics and store in tidy format for plotting
+pf <- estimatePerformanceMetrics(sb, alpha=seq(0.01, 0.10, 0.01), tidy=TRUE)
 ```
@@
 
-## to add here - standardized plotting functions from SummarizedBenchmarks -
-## to discuss at meeting 10/19
+# plot adjusted p-values by independent covariate
+plotsim_single(data.frame(assay(sb)), bmi, strat = "ind_covariate")
 
 # alpha versus nrejects
+plotsim_average(pf, met="rejections", merge_ihw = TRUE) +
+  ggtitle("Number of rejections for BMI dataset")
+
+# alpha versus nrejects (excluding "unadjusted")
+plotsim_average(pf, met="rejections", filter_set = "unadjusted", merge_ihw = TRUE) +
+  ggtitle("Number of rejections for BMI dataset")
 
 # upset plot
+plotMethodsOverlap(sb, alpha=0.1, nsets=ncol(sb))
 
+# upset plot excluding "unadjusted"
+plotMethodsOverlap(sb[,-1], alpha=0.1, nsets=ncol(sb)-1)
 ```
@@
 rm(sb)
+rm(pf)
 ```

### 2017-10-25 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
Add results plots (overlap and nrejects)

@@
 
-```{r, GWAS1-plot, results="hide", width=15, height=12}
+```{r, GWAS1-plot, results="hide", width=15, height=15}
 # plot adjusted p-values by effect_size (diagnostic from ASH paper)

### 2017-10-25 datasets/GWAS/GWAS-knit.sh

@@ -4,7 +4,7 @@
 #SBATCH -N 1                        # All cores on one machine
 #SBATCH -p irizarry,serial_requeue  # Partition
 #SBATCH --mem 20000                 # Memory request 
-#SBATCH -t 0-1:00                   # Maximum execution time (D-HH:MM)
+#SBATCH -t 0-5:00                   # Maximum execution time (D-HH:MM)
 #SBATCH -o SLURM/render-%j.out  # Standard output
 #SBATCH -e SLURM/render-%j.err  # Standard error
  

### 2017-10-26 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
Use AssayNames Method
Select only necessary data columns for buildBench

@@
 knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_knit$set(root.dir = "/n/irizarryfs01_backed_up/kkorthauer/FDR")
 ```
@@
 # project directory & data/results folders
-setwd("/n/irizarryfs01_backed_up/kkorthauer/FDR")
+#setwd("/n/irizarryfs01_backed_up/kkorthauer/FDR")
 datdir <- "/n/irizarryfs01_backed_up/kkorthauer/FDR/DATA/GWAS/"
@@
                ind_covariate = N) %>%
-        mutate(test_statistic = effect_size/SE)
+        mutate(test_statistic = effect_size/SE) %>%
+        select(pval, SE, effect_size, ind_covariate, test_statistic)
 
@@
 assayNames(sb) <- "qvalue"
-names(rowData(sb)) <- names(sb@performanceMetrics) <- "qvalue"
 sb <- addDefaultMetrics(sb)

### 2017-10-26 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
Calculate Zscore from pvals instead of effect/SE

@@
                ind_covariate = N) %>%
-        mutate(test_statistic = effect_size/SE) %>%
+        mutate(test_statistic = qnorm(exp(log(pval)-log(2)), lower.tail=FALSE) * sign(effect_size)) %>%
         select(pval, SE, effect_size, ind_covariate, test_statistic)

### 2017-10-27 datasets/GSEA/gsea.Rmd
Added functions to plot rejections vs alpha. Updated GSEA analysis.

@@
 
-```{r deseq2}
+```{r deseq2, message=FALSE}
 
@@
 
-```{r, fig.height=3.5, fig.width=4, message=FALSE}
+```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
 library(ggplot2)
@@
 library(cowplot)
+library(R.utils)
+library(scales)
+sourceDirectory("../R/")
 
-ggplot( goRes,
-       aes( log10(numInCat), -log10(over_represented_pvalue) ) ) +
-    geom_hex(aes( alpha=log2(..count..)), fill="black", color="black", bins=100, show.legend=FALSE) +
+rank_scatter( dat=goRes, pval="over_represented_pvalue", 
+              covariate="numInCat", bins=50, funx=log2, 
+              funfill=log10_trans()) +
     ylim(0, 12) + ggtitle("Over-represented gene sets") +
-    xlab(expression(log[10]~"(# of genes)")) +
-    ylab(expression(-log[10]~"(p-value)"))
+    xlab( expression(log[10]~"(# of genes)")) +
+    ylab( expression(-log[10]~"(p-value)") )
 
@@
 
-```{r, message=FALSE}
-library(IHW)
-goRes$groups <- groups_by_filter( goRes$numInCat, 10 )
+```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
+
+library(dplyr)
 
@@
 
-dplyr:::filter( goRes, numInCat > 10, numDEInCat > 0 ) %>%
-  dplyr:::mutate( groups = groups_by_filter( numInCat, 4 ) ) %>%
-  ggplot(aes(x=over_represented_pvalue)) + 
-  geom_histogram(binwidth = 0.025, boundary = 0) +
-  facet_wrap( ~ groups, nrow = 2) +
-  ggtitle("Distribution of p-values") +
-  ylab("Frequency") + xlab("p-value")
+dplyr:::filter( goRes, numDEInCat > 0, numInCat > 15 ) %>%
+  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=7)
 
@@
 
-```{r}
-
-library(R.utils)
-sourceDirectory("../R/")
-
+```{r runningBenckmark, message=FALSE}
 designBen <- initializeBenchDesign()
@@
 res$pval <- pmin( res$pval, 1 )
+sGSEA <- designBen %>% buildBench( res )
+assayNames(sGSEA) <- "qvalue"
+sGSEA <- addDefaultMetrics( sGSEA )
+```
 
-sGSEA <- designBen %>% buildBench( res )
-assayNames(sGSEA) <- names( rowData(sGSEA) ) <-  names(sGSEA@performanceMetrics) <- "qvalue"
+```{r plottingResults}
+rejections_scatter( sGSEA, as_fraction=FALSE )
 
-sGSEA <- addDefaultMetrics( sGSEA )
+stopifnot( all(assays(sGSEA)[["qvalue"]][,"unadjusted"] == pmin(goRes$over_represented_pvalue, 1) ) )
+rowData(sGSEA)$numInCat <- goRes$numInCat
 
-estimatePerformanceMetrics( sGSEA, alpha=seq(0, 0.6, 0.1), tidy=TRUE ) %>%
-  dplyr:::filter(pkg_name == "IHW") %>%
-  head
-
-  dplyr:::select(blabel, key, value, assay, performanceMetric, alpha ) %>%
-  dplyr:::filter(blabel!="unadjusted") %>%
-  ggplot( aes(alpha, value, col=blabel) ) +
-  geom_line() +
-  geom_point()
-
+rejection_scatter_bins( sGSEA, "numInCat", threshold=15, bins=4, as_fraction=TRUE )
 ```

### 2017-10-27 datasets/R/rejections_scatter.R

@@ -12,11 +12,11 @@
 #' 
 #' 
 #' @author Alejandro Reyes
-rejections_scatter <- function( sb, asFraction=FALSE ){
+rejections_scatter <- function( sb, as_fraction=FALSE ){
   stopifnot( is(sb, "SummarizedBenchmark" ) )
   alphas <- unique( as.numeric( as.character( colData( sb )$param.alpha ) ) )
   alphas <- alphas[!is.na(alphas)]
-  if( asFraction ){
+  if( as_fraction ){
     deno <- nrow(sb)
     yl <- "Fraction of hypotheses rejected"
   }else{
@@ -27,9 +27,59 @@ rejections_scatter <- function( sb, asFraction=FALSE ){
     dplyr:::filter( !(pkg_name == "IHW" & param.alpha != alpha ) ) %>%
     dplyr:::mutate( blabel=gsub("(ihw)-a\\d+", "\\1", blabel ) ) %>%
     dplyr:::select( blabel, key, value, assay, performanceMetric, alpha ) %>%
-    dplyr:::filter( blabel!="unadjusted") %>%
+    dplyr:::filter( blabel!="unadjusted", performanceMetric == "rejections" ) %>%
     ggplot( aes(alpha, value/deno, col=blabel) ) +
     geom_line() + geom_point() +
+    xlab(expression(paste("Nominal"~alpha))) +
+    ylab(yl)
+}
+
+#' Scatterplot of nominal alpha vs rejections stratified by covariate bins
+#' 
+#' Similar to the function rejections_scatter, but it stratifies the scatter 
+#' plot by covariate bins.  
+#'   
+#' @param sb A SummarizedBenchmark object
+#' @param as_fraction Logical. Whether to plot the fraction of hypotheses
+#' that were rejected (TRUE) or the absolute number of rejections (FALSE).
+#' @param threshold A numeric threshold. If specified, only test where the covariate
+#' is larger than this number are considered. 
+#' @param bins Number of bins.
+#' @param ncol_facet Number of columns for facet_grid
+#'        
+#' @return a ggplot2 object
+#' 
+#' 
+#' @author Alejandro Reyes
+rejection_scatter_bins <- function( sb, covariate, threshold=NULL, bins= 4, ncol_facet=2, as_fraction=FALSE ){
+  stopifnot(is(sb, "SummarizedExperiment"))
[…]

### 2017-10-27 datasets/scrnaseq/scRNAseq.Rmd

---
title: "scRNAseq"
author: "Ayshwarya Subramanian"
date: "9/15/2017"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## scRNAseq datasets

For the application of differential gene expression, we look at scRNAseq and bulk RNAseq data. Here we discuss the analyses of scRNAseq.

We selected two datasets, one each from mouse and human spp, from the `Conquer` (Soneson et al., 2017) scRNAseq data collection. `Conquer` provides uniformly processed gene expression data summarized to both count, TPM, length-scaled TPM and x units at the gene level. The authors use `salmon` to quantify raw reads and `tximport` to summarize to gene level.

## Data Download

We first download the data in our working directory. First, let's download the Human data:GSE63818-GPL16791

```{bash, humandata-download, eval=FALSE}
# all data and results are on the wd: /net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/fdrbenchmark-scrnaseq

#these look well separated on the tsne plots on the conquer manuscript supplement
wget http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE63818-GPL16791.rds
# qc report: 
wget http://imlspenticton.uzh.ch/robinson_lab/conquer/report-multiqc/GSE63818-GPL16791_multiqc_report.html
# scater report
wget http://imlspenticton.uzh.ch/robinson_lab/conquer/report-scater/GSE63818-GPL16791_scater.html
```

Next, we download the Mouse data:GSE45719

```{bash, mousedata-download, eval=FALSE}
#another well-separated dataset on tsne: GSE60749−GPL13112
wget http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE45719.rds
```

## R workspace setup

The downloaded data are stored in R objects called `MultiAssayExperiment`; this package requires BioConductor release 3.5 and R version 3.4 or greater for installation.

```{r, workspace-setup, eval=FALSE}
#installation
# source("https://bioconductor.org/biocLite.R")
# biocLite("swfdr")
# biocLite("IHW")
# biocLite("qvalue")
# 
# install.packages("ashr")
# install.packages("fdrtool")
# biocLite("edgeR")
# source("https://bioconductor.org/biocLite.R")
# biocLite("scran")
#  Samseq installation
# source("https://bioconductor.org/biocLite.R")
# biocLite("impute") 
# install.packages("samr")
# source("https://bioconductor.org/biocLite.R")
#biocLite("MultiAssayExperiment")

# packages for data analyses
library(MultiAssayExperiment)
library(dplyr)
library(ggplot2)
library(cowplot)
library(UpSetR)
library(hexbin)

# packages for differential expression testing
library(impute)
library(samr)
library(scDD)
library(EBSeq)
library(edgeR)

# benchmark methods
library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)

```

## Data processing and filtering

For data processing, we follow practices and recommendations from Soneson et al., 2017. Briefly, we limit ourselves to 2-group comparisons, and use the Wilcoxon test at the TPM level for differential expresssion testing and SAMseq for additional comparison. We select the maximal size defined as the smallest number of cells in either group as the group-level sample size for the DE analyses.

First, we get the human data for the two groups of interest.

We look at results with and without prefiltering genes (keeping genes with TPM > 1 in atleast 25% of cells)

```{r, dataload, eval=FALSE}
setwd("/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/fdrbenchmark-scrnaseq")
#load compressed mae
humandata=readRDS("GSE63818-GPL16791.rds")
#get gene-level data
genes=experiments(humandata)[1]

#TPMS
#get tpms
tpms=assays(genes$gene)[1]
#get tpm dataframe
human_tpms=tpms$TPM
# get column data/metadata for sample
cdata=colData(humandata)
cdata=as.data.frame(cdata)
# select groups of interest
g1 = cdata%>%subset(source_name_ch1=="Somatic Cells" & characteristics_ch1=="developmental stage: 7 week gestation")
g2 = cdata%>%subset(source_name_ch1=="Primordial Germ Cells" & characteristics_ch1=="developmental stage: 7 week gestation")
# select maximal sample size
sample_size = c(dim(g1)[1],dim(g2)[1])
ssize = which.min(sample_size)
g1_ids=sample.int(dim(g1)[1], size = sample_size[ssize], replace = FALSE)
g2_ids=sample.int(dim(g2)[1], size = sample_size[ssize], replace = FALSE)
# get group data of maximal size
g1_data=human_tpms[,colnames(human_tpms)%in%g1$geo_accession[g1_ids]]
g2_data=human_tpms[,colnames(human_tpms)%in%g2$geo_accession[g2_ids]]
#save data
saveRDS(g1_data,"h_g1_data.rds")
saveRDS(g2_data,"h_g2_data.rds")

#COUNTS
#get counts
count=assays(genes$gene)[2]
#get count dataframe
human_count=count$count
# get column data/metadata for sample
cdata=colData(humandata)
cdata=as.data.frame(cdata)
# get group data of maximal size
g1_cdata=human_count[,colnames(human_count)%in%g1$geo_accession[g1_ids]]
g2_cdata=human_count[,colnames(human_count)%in%g2$geo_accession[g2_ids]]
#save data
saveRDS(g1_cdata,"h_g1_cdata.rds")
saveRDS(g2_cdata,"h_g2_cdata.rds")
```

We see that each group has 26 cells and 65218 genes. We merge the data and remove all genes which have 0s in all samples

```{r, nofilter, eval=FALSE}
g_data=cbind(g1_data,g2_data)
g_data_sub=g_data[rowSums(g_data==0)!=52,] # 0.25*52=13
saveRDS(g_data_sub,"h_g_nofilter_subdata.rds") # removing rows with all zeros reduces the number of genes to 33169

g_cdata=cbind(g1_cdata,g2_cdata)
g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=52,]
saveRDS(g_cdata_sub,"h_g_nofilter_subcdata.rds") # removing rows with all zeros reduces the number of genes to 33169

saveRDS(g1_ids,"g1_ids.rds")
saveRDS(g2_ids,"g2_ids.rds")
```

### Filtering

Now we merge the data and filter the data to retain only genes with have atleast a 25% detection rate (defined as TPM > 1).

```{r, filter, eval=FALSE}
g_data=cbind(g1_data,g2_data)
g_data_sub=g_data[rowSums(g_data>1)>=13,] # 0.25*52=13
saveRDS(g_data_sub,"h_g_subdata.rds")
```

Filtering leaves us with 11415 genes (retaining about 17.5% of the genes).

## Differential expression testing

We perform differential expression testing using 3 methods on the unfiltered data:

### Wilcoxon rank-sum test

Per conquer, we run TMM normalization to the TPM units before running Wilcoxon.
```{r, wilcoxon, eval=FALSE}
groups=c(rep(1,26),rep(2,26))
## code borrowed from https://github.com/csoneson/conquer_comparison/blob/master/scripts/apply_Wilcoxon.R
tmm <- edgeR::calcNormFactors(g_data_sub)
    tpmtmm <- edgeR::cpm(g_data_sub, lib.size = tmm * colSums(g_data_sub))
    idx <- 1:nrow(tpmtmm)
    names(idx) <- rownames(tpmtmm)
    timing <- system.time(wilcox_p <- sapply(idx, function(i) {
      wilcox.test(tpmtmm[i, ] ~ groups)$p.value
    }))
saveRDS(wilcox_p,"h_data_wilcox_p.rds")

## warnings:
## In wilcox.test.default(x = structure(c(0, 0, 14.2107316559397,  ... :
## cannot compute exact p-value with ties

```

### Differences in distributions of gene expression in the 2 conditions

We apply scDD to log-transformed counts as below:

```{r, scdd, eval=FALSE}

Sizes=MedianNorm(g_cdata_sub)

ng_cdata_sub = Sizes*g_cdata_sub
saveRDS(Sizes,"sizefactors.rds")
library(SummarizedExperiment)
condition=c(rep(1,26),rep(2,26))
rownames(ng_cdata_sub) <- paste0(rownames(ng_cdata_sub), 1:nrow(ng_cdata_sub), sep="")
colnames(ng_cdata_sub) <- names(condition) <- paste0("Sample",
1:ncol(ng_cdata_sub), sep="")
SDSumExp <- SummarizedExperiment(assays=list("NormCounts"=ng_cdata_sub),
colData=data.frame(condition))
saveRDS(SDSumExp,"data_sumexp.rds")
#

prior_param=list(alpha=0.01, mu0=0, s0=0.01, a0=0.01, b0=0.01)
param=BiocParallel::MulticoreParam(workers=12)
scDatEx <- scDD(SDSumExp, prior_param=prior_param, param=param, testZeroes=FALSE)
## Notice: 16073 genes have less than 3 nonzero cells per condition.  Skipping these genes.
saveRDS(scDatEx,"scDatEx.rds")

perm=1000
scDatExperm <- scDD(SDSumExp, prior_param=prior_param, param=param, permutations=1000,testZeroes=FALSE)
saveRDS(scDatExperm,"/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/fdrbenchmark-scrnaseq/scDatExperm.rds")

```

### SAMseq

We next run SAMseq

```{r, samseq, eval=FALSE}
#below code is borrowed from https://github.com/csoneson/conquer_comparison/blob/master/scripts/apply_SAMseq.R
#
  message("SAMseq")
  session_info <- sessionInfo()
  tryCatch({
    timing <- system.time({
      SAMseq.test <- SAMseq(round(g_cdata_sub), as.numeric(as.factor(groups)),
                            resp.type = "Two class unpaired",
                            geneid = rownames(g_cdata_sub), genenames = rownames(g_cdata_sub),
                            nperms = 100, nresamp = 20, fdr.output = 1)
      SAMseq.result.table <- rbind(SAMseq.test$siggenes.table$genes.up,
                                   SAMseq.test$siggenes.table$genes.lo)
      SAMseq.FDR <- rep(NA, nrow(L$count))
      SAMseq.FDR[match(SAMseq.result.table[, 1], rownames(L$count))] <- 
        as.numeric(SAMseq.result.table[, 5])/100
    })

    hist(SAMseq.FDR, 50)

    list(session_info = session_info,
         timing = timing,
         res = SAMseq.result.table, 
         df = data.frame(padj = SAMseq.FDR, 
                         row.names = rownames(L$count)))
  }, error = function(e) {
    "SAMseq results could not be calculated"
    list(session_info = session_info)
  })

  saveRDS(SAMseq.test,"SAMseq.test.RDS")
  saveRDS(SAMseq.result.table,"SAMseq.result.table.RDS")

```

## Covariate Diagnostics

We could use a few different covariates: strength of the signal defined as the median gene expression unit for the gene, detection rate or a combination of the two.

We split the genes into 4 bins based on values of covariates (1 quartile, 2nd quartile, etc)
```{r, cov, eval=FALSE}
scdd_m=metadata(scDatEx)
scdd_p_df=data.frame(genes=scdd_m$Genes$gene,scdd_pval=matrix(scdd_m$Genes$nonzero.pvalue))
wilcox_p_df=data.frame(genes=names(wilcox_p),wilcox_pval=matrix(wilcox_p))
samseq_p_df=data.frame(genes=as.character(SAMseq.result.table[,2]),samseq_qval=matrix(SAMseq.result.table[,5]))
g_cdata_sub_med=apply(g_cdata_sub,1,median)
g_cdata_sub_mean=apply(g_cdata_sub,1,mean)
g_data_sub_dr=apply(g_cdata_sub,1,function(x) length(which(x==1))/length(x))
cov_df=data.frame(genes=names(g_cdata_sub_med),median_exp=g_cdata_sub_med,mean_exp=g_cdata_sub_mean,detection_rate=g_data_sub_dr)
dat=cov_df%>%left_join(wilcox_p_df,by="genes")%>%left_join(scdd_p_df,by="genes")%>%left_join(samseq_p_df,by="genes")

dat=data.frame(cov_df%>%left_join(wilcox_p_df,by="genes")%>%left_join(samseq_p_df,by="genes"),scdd_pval=scdd_p_df$scdd_pval)
source("benchmark-fdr/datasets/R/simulation-helpers.R")
source("benchmark-fdr/datasets/R/evaluation-helpers.R")
# First for the median exp level and wilcoxon p-val
q=quantile(rank(dat$median_exp, ties="first"))
r=rank(dat$median_exp, ties="first")

pdf("diagnosticplot_wilcox_medexp.pdf",width=20,height=10)
q=quantile(rank(dat$median_exp, ties="first"))
r=rank(dat$median_exp, ties="first")
bin2 = which((dat$median_exp >=dat$median_exp[which(r==q[1])])&(dat$median_exp <=  dat$median_exp[which(r==q[2])]))
bin3 = which((dat$median_exp >dat$median_exp[which(r==q[2])])&(dat$median_exp <=  dat$median_exp[which(r==q[3])]))
bin4 = which((dat$median_exp >dat$median_exp[which(r==q[3])])&(dat$median_exp <=  dat$median_exp[which(r==q[4])]))
bin5 = which(dat$median_exp >  dat$median_exp[which(r==q[4])]) 
bin=dat$median_exp
bin[bin2]="bin2"
bin[bin3]="bin3"
bin[bin4]="bin4"
bin[bin5]="bin5"
df=data.frame(pval=dat$wilcox_pval,bins=bin)
df%>%ggplot(aes(x=pval,facet=bin))+geom_histogram() +aes(y=..density..)+facet_wrap( ~ bins, nrow = 4)
dev.off()

pdf("diagnosticplot_wilcox_meanexp.pdf",width=20,height=10)
q=quantile(rank(dat$mean_exp, ties="first"))
r=rank(dat$mean_exp, ties="first")
bin2 = which((dat$mean_exp >=dat$mean_exp[which(r==q[1])])&(dat$mean_exp <=  dat$mean_exp[which(r==q[2])]))
bin3 = which((dat$mean_exp >dat$mean_exp[which(r==q[2])])&(dat$mean_exp <=  dat$mean_exp[which(r==q[3])]))
bin4 = which((dat$mean_exp >dat$mean_exp[which(r==q[3])])&(dat$mean_exp <=  dat$mean_exp[which(r==q[4])]))
bin5 = which(dat$mean_exp >  dat$mean_exp[which(r==q[4])]) 
bin=dat$mean_exp
bin[bin2]="bin2"
bin[bin3]="bin3"
bin[bin4]="bin4"
bin[bin5]="bin5"
df=data.frame(pval=dat$wilcox_pval,bins=bin)
df%>%ggplot(aes(x=pval,facet=bin))+geom_histogram() +aes(y=..density..)+facet_wrap( ~ bins, nrow = 4)
dev.off()

pdf("diagnosticplot_wilcox_detrate.pdf",width=20,height=10)
q=quantile(rank(dat$detection_rate, ties="first"))
r=rank(dat$detection_rate, ties="first")
bin2 = which((dat$detection_rate >=dat$detection_rate[which(r==q[1])])&(dat$detection_rate <=  dat$detection_rate[which(r==q[2])]))
bin3 = which((dat$detection_rate >dat$detection_rate[which(r==q[2])])&(dat$detection_rate <=  dat$detection_rate[which(r==q[3])]))
bin4 = which((dat$detection_rate >dat$detection_rate[which(r==q[3])])&(dat$detection_rate <=  dat$detection_rate[which(r==q[4])]))
bin5 = which(dat$detection_rate >  dat$detection_rate[which(r==q[4])]) 
bin=dat$detection_rate
bin[bin2]="bin2"
bin[bin3]="bin3"
bin[bin4]="bin4"
bin[bin5]="bin5"
df=data.frame(pval=dat$wilcox_pval,bins=bin)
df%>%ggplot(aes(x=pval,facet=bin))+geom_histogram() +aes(y=..density..)+facet_wrap( ~ bins, nrow = 4)
dev.off()

## samseq

pdf("diagnosticplot_samseq_medexp.pdf",width=20,height=10)
q=quantile(rank(dat$median_exp, ties="first"))
r=rank(dat$median_exp, ties="first")
bin2 = which((dat$median_exp >=dat$median_exp[which(r==q[1])])&(dat$median_exp <=  dat$median_exp[which(r==q[2])]))
bin3 = which((dat$median_exp >dat$median_exp[which(r==q[2])])&(dat$median_exp <=  dat$median_exp[which(r==q[3])]))
bin4 = which((dat$median_exp >dat$median_exp[which(r==q[3])])&(dat$median_exp <=  dat$median_exp[which(r==q[4])]))
bin5 = which(dat$median_exp >  dat$median_exp[which(r==q[4])]) 
bin=dat$median_exp
bin[bin2]="bin2"
bin[bin3]="bin3"
bin[bin4]="bin4"
bin[bin5]="bin5"
df=data.frame(pval=as.numeric(paste(dat$samseq_qval))/100,bins=bin)
df%>%ggplot(aes(x=pval,facet=bin))+geom_histogram() +aes(y=..density..)+facet_wrap( ~ bins, nrow = 4)
dev.off()

pdf("diagnosticplot_samseq_meanexp.pdf",width=20,height=10)
q=quantile(rank(dat$mean_exp, ties="first"))
r=rank(dat$mean_exp, ties="first")
bin2 = which((dat$mean_exp >=dat$mean_exp[which(r==q[1])])&(dat$mean_exp <=  dat$mean_exp[which(r==q[2])]))
bin3 = which((dat$mean_exp >dat$mean_exp[which(r==q[2])])&(dat$mean_exp <=  dat$mean_exp[which(r==q[3])]))
bin4 = which((dat$mean_exp >dat$mean_exp[which(r==q[3])])&(dat$mean_exp <=  dat$mean_exp[which(r==q[4])]))
bin5 = which(dat$mean_exp >  dat$mean_exp[which(r==q[4])]) 
bin=dat$mean_exp
bin[bin2]="bin2"
bin[bin3]="bin3"
bin[bin4]="bin4"
bin[bin5]="bin5"
df=data.frame(pval=as.numeric(paste(dat$samseq_qval))/100,bins=bin)
df%>%ggplot(aes(x=pval,facet=bin))+geom_histogram() +aes(y=..density..)+facet_wrap( ~ bins, nrow = 4)
dev.off()

pdf("diagnosticplot_samseq_detrate.pdf",width=20,height=10)
q=quantile(rank(dat$detection_rate, ties="first"))
r=rank(dat$detection_rate, ties="first")
bin2 = which((dat$detection_rate >=dat$detection_rate[which(r==q[1])])&(dat$detection_rate <=  dat$detection_rate[which(r==q[2])]))
bin3 = which((dat$detection_rate >dat$detection_rate[which(r==q[2])])&(dat$detection_rate <=  dat$detection_rate[which(r==q[3])]))
bin4 = which((dat$detection_rate >dat$detection_rate[which(r==q[3])])&(dat$detection_rate <=  dat$detection_rate[which(r==q[4])]))
bin5 = which(dat$detection_rate >  dat$detection_rate[which(r==q[4])]) 
bin=dat$detection_rate
bin[bin2]="bin2"
bin[bin3]="bin3"
bin[bin4]="bin4"
bin[bin5]="bin5"
df=data.frame(pval=as.numeric(paste(dat$samseq_qval))/100,bins=bin)
df%>%ggplot(aes(x=pval,facet=bin))+geom_histogram() +aes(y=..density..)+facet_wrap( ~ bins, nrow = 4)
dev.off()

## scdd
pdf("diagnosticplot_scdd_medexp.pdf",width=20,height=10)
q=quantile(rank(dat$median_exp, ties="first"))
r=rank(dat$median_exp, ties="first")
bin2 = which((dat$median_exp >=dat$median_exp[which(r==q[1])])&(dat$median_exp <=  dat$median_exp[which(r==q[2])]))
bin3 = which((dat$median_exp >dat$median_exp[which(r==q[2])])&(dat$median_exp <=  dat$median_exp[which(r==q[3])]))
bin4 = which((dat$median_exp >dat$median_exp[which(r==q[3])])&(dat$median_exp <=  dat$median_exp[which(r==q[4])]))
bin5 = which(dat$median_exp >  dat$median_exp[which(r==q[4])]) 
bin=dat$median_exp
bin[bin2]="bin2"
bin[bin3]="bin3"
bin[bin4]="bin4"
bin[bin5]="bin5"
df=data.frame(pval=dat$scdd_pval,bins=bin)
df%>%ggplot(aes(x=pval,facet=bin))+geom_histogram() +aes(y=..density..)+facet_wrap( ~ bins, nrow = 4)
dev.off()

pdf("diagnosticplot_scdd_meanexp.pdf",width=20,height=10)
q=quantile(rank(dat$mean_exp, ties="first"))
r=rank(dat$mean_exp, ties="first")
bin2 = which((dat$mean_exp >=dat$mean_exp[which(r==q[1])])&(dat$mean_exp <=  dat$mean_exp[which(r==q[2])]))
bin3 = which((dat$mean_exp >dat$mean_exp[which(r==q[2])])&(dat$mean_exp <=  dat$mean_exp[which(r==q[3])]))
bin4 = which((dat$mean_exp >dat$mean_exp[which(r==q[3])])&(dat$mean_exp <=  dat$mean_exp[which(r==q[4])]))
bin5 = which(dat$mean_exp >  dat$mean_exp[which(r==q[4])]) 
bin=dat$mean_exp
bin[bin2]="bin2"
bin[bin3]="bin3"
bin[bin4]="bin4"
bin[bin5]="bin5"
df=data.frame(pval=dat$scdd_pval,bins=bin)
df%>%ggplot(aes(x=pval,facet=bin))+geom_histogram() +aes(y=..density..)+facet_wrap( ~ bins, nrow = 4)
dev.off()

pdf("diagnosticplot_scdd_detrate.pdf",width=20,height=10)
q=quantile(rank(dat$detection_rate, ties="first"))
r=rank(dat$detection_rate, ties="first")
bin2 = which((dat$detection_rate >=dat$detection_rate[which(r==q[1])])&(dat$detection_rate <=  dat$detection_rate[which(r==q[2])]))
bin3 = which((dat$detection_rate >dat$detection_rate[which(r==q[2])])&(dat$detection_rate <=  dat$detection_rate[which(r==q[3])]))
bin4 = which((dat$detection_rate >dat$detection_rate[which(r==q[3])])&(dat$detection_rate <=  dat$detection_rate[which(r==q[4])]))
bin5 = which(dat$detection_rate >  dat$detection_rate[which(r==q[4])]) 
bin=dat$detection_rate
bin[bin2]="bin2"
bin[bin3]="bin3"
bin[bin4]="bin4"
bin[bin5]="bin5"
df=data.frame(pval=dat$scdd_pval,bins=bin)
df%>%ggplot(aes(x=pval,facet=bin))+geom_histogram() +aes(y=..density..)+facet_wrap( ~ bins, nrow = 4)
dev.off()

# diagnostic scatter plots

pdf("diagnosticscatterplots.pdf",width=20,height=10)
gg_scat <- ggplot(dat, aes(y=-log10(wilcox_pval), x=rank(detection_rate, ties="first")/nrow(dat))) + geom_hex(bins = 100) + ylab(expression(-log[10]~p)) + xlab("Covariate Rank") + theme_classic() + ggtitle("wilcoxon, detection rate")
gg_scat
gg_scat <- ggplot(dat, aes(y=-log10(as.numeric(paste(dat$samseq_qval))/100), x=rank(detection_rate, ties="first")/nrow(dat))) + geom_hex(bins = 100) + ylab(expression(-log[10]~p)) + xlab("Covariate Rank") + theme_classic() + ggtitle("samseq q, detection rate")
gg_scat
gg_scat <- ggplot(dat, aes(y=-log10(scdd_pval), x=rank(detection_rate, ties="first")/nrow(dat))) + geom_hex(bins = 100) + ylab(expression(-log[10]~p)) + xlab("Covariate Rank") + theme_classic() + ggtitle("scdd, detection rate")
gg_scat

gg_scat <- ggplot(dat, aes(y=-log10(wilcox_pval), x=rank(median_exp, ties="first")/nrow(dat))) + geom_hex(bins = 100) + ylab(expression(-log[10]~p)) + xlab("Covariate Rank") + theme_classic() + ggtitle("wilcoxon, median exp")
gg_scat
gg_scat <- ggplot(dat, aes(y=-log10(as.numeric(paste(dat$samseq_qval))/100), x=rank(median_exp, ties="first")/nrow(dat))) + geom_hex(bins = 100) + ylab(expression(-log[10]~p)) + xlab("Covariate Rank") + theme_classic() + ggtitle("samseq q, median exp")
gg_scat
gg_scat <- ggplot(dat, aes(y=-log10(scdd_pval), x=rank(median_exp, ties="first")/nrow(dat))) + geom_hex(bins = 100) + ylab(expression(-log[10]~p)) + xlab("Covariate Rank") + theme_classic() + ggtitle("scdd, median exp")
gg_scat

gg_scat <- ggplot(dat, aes(y=-log10(wilcox_pval), x=rank(mean_exp, ties="first")/nrow(dat))) + geom_hex(bins = 100) + ylab(expression(-log[10]~p)) + xlab("Covariate Rank") + theme_classic() + ggtitle("wilcoxon, mean exp")
gg_scat
gg_scat <- ggplot(dat, aes(y=-log10(as.numeric(paste(dat$samseq_qval))/100), x=rank(mean_exp, ties="first")/nrow(dat))) + geom_hex(bins = 100) + ylab(expression(-log[10]~p)) + xlab("Covariate Rank") + theme_classic() + ggtitle("samseq q, mean exp")
gg_scat
gg_scat <- ggplot(dat, aes(y=-log10(scdd_pval), x=rank(mean_exp, ties="first")/nrow(dat))) + geom_hex(bins = 100) + ylab(expression(-log[10]~p)) + xlab("Covariate Rank") + theme_classic() + ggtitle("scdd, mean exp")
gg_scat
dev.off()

```

## Benchmarking covariate adjusted FDR tools for application

```{r, benchmarking, eval=FALSE}
g_cdata_sub_se=apply(g_cdata_sub,1,function(x) sd(x)/sqrt(length(x))) #standard error
g_cdata_sub_es=apply(g_cdata_sub,1,function(x) mean(x[1:26])-mean(x[27:52])) #difference of means

b_dat=data.frame(pval=dat$wilcox_pval, ind_covariate=dat$mean_exp,effect_size=matrix(g_cdata_sub_es),SE=matrix(g_cdata_sub_se))
b_dat=b_dat%>%mutate(zscore = effect_size/SE)
saveRDS(b_dat,"wilcox_meanexp.RDS")
cutoffs <- c(0.01, 0.025, 0.05, 0.10, 0.20)
resfile <- paste0(getwd(), "/scrnaseqhuman_wilcox_meanexp_results_",
                  nrow(b_dat), ".RData")
library(devtools)
source("https://install-github.me/mangothecat/callr")
source("https://install-github.me/r-lib/remotes")
remotes::install_github('jgscott/FDRreg', subdir="R_pkg/")

source("benchmark-fdr/datasets/R/evaluation-helpers_mod.R")
if (!file.exists(resfile)){
  t1 <- proc.time()
  b_dat_results <- run_benchmarks_mod(dat=b_dat, alphas=cutoffs, pvals=TRUE)
  save(b_dat_results, file=resfile)
  message(paste0("Took ", round((proc.time()-t1)[3],1),
                 " seconds for ", nrow(b_dat), " genes"))
}else{
  load(resfile)
}

# plot N rejects vs FDR cutoff

pdf("scrnaseq_wilcox_results.pdf",width=20,height=10)
p <- ggplot(b_dat_results$stats, aes(alpha, n_rejects, color=method)) +
  geom_point() +
  geom_line() +
  xlab("FDR cutoff") + ylab("Number of rejections") +
  ggtitle("scRNAseq with Mean gene expression Covariate, wilcox") +
  theme_classic()
print(p)

# log-scale y axis
print(p + scale_y_continuous(trans="log2") + 
        ylab("Number of rejections (log2 scale)"))
dev.off()

b_dat=data.frame(pval=dat$scdd_pval, ind_covariate=dat$mean_exp,effect_size=matrix(g_cdata_sub_es),SE=matrix(g_cdata_sub_se))
b_dat=b_dat%>%mutate(zscore = effect_size/SE)
saveRDS(b_dat,"wilcox_meanexp.RDS")

resfile <- paste0(getwd(), "/scrnaseqhuman_scdd_meanexp_results_",
                  nrow(b_dat), ".RData")
if (!file.exists(resfile)){
  t1 <- proc.time()
  b_dat_results <- run_benchmarks_mod(dat=b_dat%>%filter(!is.na(pval)), alphas=cutoffs, pvals=TRUE)
  save(b_dat_results, file=resfile)
  message(paste0("Took ", round((proc.time()-t1)[3],1),
                 " seconds for ", nrow(b_dat), " genes"))
}else{
  load(resfile)
}

# plot N rejects vs FDR cutoff

pdf("scrnaseq_scdd_results.pdf",width=20,height=10)
p <- ggplot(b_dat_results$stats, aes(alpha, n_rejects, color=method)) +
  geom_point() +
  geom_line() +
  xlab("FDR cutoff") + ylab("Number of rejections") +
  ggtitle("scRNAseq with Mean gene expression Covariate, scdd") +
  theme_classic()
print(p)

# log-scale y axis
print(p + scale_y_continuous(trans="log2") + 
        ylab("Number of rejections (log2 scale)"))
dev.off()

```

### 2017-10-29 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
Add parallel buildBench, new plotting functions

@@
 sourceDirectory("benchmark-fdr/datasets/R/")
+
+# set up parallel backend
+library(BiocParallel)
+cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
+multicoreParam <- MulticoreParam(workers = cores)
 ```
@@
 ### experimenting on a subset of the full dataset for now
-bmi <- bmi[1:500000,]
+bmi <- bmi[1:750000,]
 ```
@@
 ```{r, GWAS1-diag, fig.width=10, fig.height=3.5}
-strat_hist(bmi, pval="pval", covariate="ind_covariate", maxy=3)
-rank_scatter(bmi, pval="pval", covariate="ind_covariate")
+strat_hist(bmi, pvalue="pval", covariate="ind_covariate", maxy=3)
+rank_scatter(bmi, pvalue="pval", covariate="ind_covariate")
 ```
@@
   t1 <- proc.time()
-  sb <- bd %>% buildBench(data=bmi)
+  sb <- bd %>% buildBench(data=bmi, parallel=TRUE, BPPARAM=multicoreParam)
   metadata(sb)$data_download_link <- "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
@@
 
-# estimate perf metrics and store in tidy format for plotting
-pf <- estimatePerformanceMetrics(sb, alpha=seq(0.01, 0.10, 0.01), tidy=TRUE)
+# estimate perf metrics 
+sb <- estimatePerformanceMetrics(sb, alpha=seq(0.01, 0.10, 0.01), addColData=TRUE)
 ```
@@
 ```{r, GWAS1-plot, results="hide", width=15, height=15}
-# plot adjusted p-values by effect_size (diagnostic from ASH paper)
-plotsim_single(data.frame(assay(sb)), bmi, strat = "effect_size")
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter(sb)
 
-# plot adjusted p-values by independent covariate
-plotsim_single(data.frame(assay(sb)), bmi, strat = "ind_covariate")
+# Add ind covariate to rowData
+rowData(sb)$ind_covariate <- bmi$ind_covariate
 
-# alpha versus nrejects
-plotsim_average(pf, met="rejections", merge_ihw = TRUE) +
-  ggtitle("Number of rejections for BMI dataset")
+rejection_scatter_bins(sb, covariate="ind_covariate", bins=4)
 
-# alpha versus nrejects (excluding "unadjusted")
-plotsim_average(pf, met="rejections", filter_set = "unadjusted", merge_ihw = TRUE) +
-  ggtitle("Number of rejections for BMI dataset")
+# upset plot (only plot IHW for the plotted alpha)
+rmv.ihw <- which(grepl("ihw", colnames(sb)) &
+                              colData(sb)$param.alpha != 0.05)
 
-# upset plot
-plotMethodsOverlap(sb, alpha=0.1, nsets=ncol(sb))
+plotMethodsOverlap(sb[,-rmv.ihw], 
+                   alpha=0.05, nsets=ncol(sb)-length(rmv.ihw),
+                   order.by="freq", decreasing=TRUE)
 
 # upset plot excluding "unadjusted"
-plotMethodsOverlap(sb[,-1], alpha=0.1, nsets=ncol(sb)-1)
+plotMethodsOverlap(sb[,-c(1,rmv.ihw)], 
+                   alpha=0.05, nsets=ncol(sb)-length(rmv.ihw)-1,
+                   order.by="freq", decreasing=TRUE)
 ```

### 2017-10-29 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
Don't add performance metrics to colData before plotting nrejects by bins

@@
 
-**Still to-do: consider multiple covariates for methods that allow this.**
+**Still to-do: repeat for additional covariate (MAF).**
 
@@
 sb@performanceMetrics[[1]]
-
-# estimate perf metrics 
-sb <- estimatePerformanceMetrics(sb, alpha=seq(0.01, 0.10, 0.01), addColData=TRUE)
 ```

### 2017-10-29 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
Add ind_covariate to buildBench; run all SNPs

@@
 Before moving on, we'll rename the relevant columns so that they will use common terms
-across the different datasets. We'll also add a 'test_statistic' column, and subset on a 
-toy example before moving on to the full dataset.
+across the different datasets. We'll also add a 'test_statistic' column.
 
@@
         select(pval, SE, effect_size, ind_covariate, test_statistic) 
-
-### experimenting on a subset of the full dataset for now
-bmi <- bmi[1:750000,]
 ```
@@
 
-```{r, GWAS1-sb, results="hide"}
+```{r, GWAS1-sb, results="hide", message=FALSE}
 resfile <- paste0(resdir, "bmi_summarizedBenchmark_",
@@
   t1 <- proc.time()
-  sb <- bd %>% buildBench(data=bmi, parallel=TRUE, BPPARAM=multicoreParam)
+  sb <- bd %>% buildBench(data=bmi, ftCols = "ind_covariate",
+                          parallel=TRUE, BPPARAM=multicoreParam)
   metadata(sb)$data_download_link <- "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
   save(sb, file=resfile)
-  message(paste0("Took ", round((proc.time()-t1)[3],1),
-                 " seconds for ", nrow(bmi), " SNPs."))
+  duration <- round((proc.time()-t1)[3]/60,1)
 }else{
@@
 }
+```
 
-```
+This step took `r duration` minutes for `r nrow(bmi)` SNPs using `r cores` cores.
 
@@
 rejections_scatter(sb)
-
-# Add ind covariate to rowData
-rowData(sb)$ind_covariate <- bmi$ind_covariate
 

### 2017-10-30 datasets/GSEA/gsea.Rmd
Added new buildBench ftCols code to GSEA

@@
 ```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
+
 library(ggplot2)
@@
 res$pval <- pmin( res$pval, 1 )
-sGSEA <- designBen %>% buildBench( res )
+head(res)
+sGSEA <- designBen %>% buildBench( res, ftCols="ind_covariate")
+elementMetadata( rowData(sGSEA) )
 assayNames(sGSEA) <- "qvalue"
@@
 stopifnot( all(assays(sGSEA)[["qvalue"]][,"unadjusted"] == pmin(goRes$over_represented_pvalue, 1) ) )
-rowData(sGSEA)$numInCat <- goRes$numInCat
 
-rejection_scatter_bins( sGSEA, "numInCat", threshold=15, bins=4, as_fraction=TRUE )
+sGSEA <- estimatePerformanceMetrics(sGSEA, addColData=TRUE)
+rejection_scatter_bins( sGSEA, "ind_covariate", threshold=15, bins=4, as_fraction=TRUE )
+
 ```

### 2017-10-31 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
Move timestamp message to separate code chunk

@@
 
+Here we look to see if the covariate does indeed look informative.
+
 ```{r, GWAS1-diag, fig.width=10, fig.height=3.5}
@@
 
+**Still to-do: repeat for additional covariate (MAF).**
+
 #### Set up BenchDesign object
@@
 add the benchmark methods to the `BenchDesign` object.
-
-**Still to-do: repeat for additional covariate (MAF).**
 
@@
                   nrow(bmi), ".RData")
+duration <- NA
 if (!file.exists(resfile)){
@@
 
-This step took `r duration` minutes for `r nrow(bmi)` SNPs using `r cores` cores.
+```{r, echo=FALSE}
+if(!is.na(duration)){
+  message("This step took ", duration, " minutes for ", nrow(bmi), " SNPs using ",
+        cores, " cores.")
+}
+```
 
@@
 sb <- addDefaultMetrics(sb)
-
-head(assay(sb))
-head(colData(sb))
-sb@performanceMetrics[[1]]
 ```

### 2017-10-31 datasets/GWAS/GWAS-knit.sh

@@ -1,9 +1,9 @@
 #!/bin/bash
 #SBATCH -J knitGWAS                 # A single job name for the array
-#SBATCH -n 1                        # Number of cores
+#SBATCH -n 8                        # Number of cores
 #SBATCH -N 1                        # All cores on one machine
 #SBATCH -p irizarry,serial_requeue  # Partition
-#SBATCH --mem 20000                 # Memory request 
+#SBATCH --mem 60000                 # Memory request 
 #SBATCH -t 0-5:00                   # Maximum execution time (D-HH:MM)
 #SBATCH -o SLURM/render-%j.out  # Standard output
 #SBATCH -e SLURM/render-%j.err  # Standard error

### 2017-11-02 datasets/simulations/analysis-02-try-adapt.Rmd
Add quick analysis of AdaPT

---
title: "Testing out AdaPT"
author: "Rafalab Journal Club Members"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
---

```{r}
library("tidyverse")
```

The AdaPT method was proposed in Lei and Fithian (2016) with accompanying code (`AdaPT.R`). While based on the paper, the method appears to be relevant to the present benchmark study, because the paper does not include substantial empirical studies (no simulations, one data application), it is unclear whether [1] the method performs reasonably well in practice, and [2] the corresponding implementation is sufficient for data practicioners.

Here, we first consider a simple simulation setting to develop an understanding of the output of the method, since documentation is sparse. We use the `du_tsim` function to simulate a simple t-test based dataset with the rank of the sample variance used as the informative covariate.

```{r}
source("../R/du_tsim.R")
```

We use the AdaPT implementation fron the author's GitHub repo.

```{r}
library(RCurl)
script <- getURL("https://raw.githubusercontent.com/lihualei71/AdaPT/master/AdaPT.R")
eval(parse(text = script))
```

We simulate the simple data set and apply AdaPT.

```{r}

sim_df <- du_tsim(m = 2e4, pi0 = .9, effect_size = 2,
                  n_sample = 10, n_group = 2, icovariate = TRUE)
res <- AdaPT(x = rank(sim_df$ind_covariate),
             pvals = sim_df$pval,
             quiet = TRUE)

```

The returned object is a list with `r length(res)` entries.

```{r}
lapply(res, class)
lapply(res, function(x) if (is.matrix(x)) dim(x) else length(x))
```

Entries are either vectors of the same length or matrices with the same number of columns, and number of rows equal to the number of tested hypotheses. Presumably, the columns of each matrix element correspond to the entries of each vector element in the output.

First, we can take a quick look at the vector elements.

```{r}
vec_tib <- tibble(step = res$step,
                  fdp = res$fdp,
                  num_rej = res$num_rej)
```

The `step` vector is likely the iteration index for the updates used in the AdaPT algorithm. We can plot the two other vectors, `fdp` and `num_rej` as a function of `step`.

```{r, fig.width = }
vec_tib %>%
    gather(var, value, -step) %>%
    ggplot(aes(x = step, y = value, label = step)) +
    geom_line(alpha = 1/2) + geom_point() +
    geom_text(vjust = 1/2, hjust = -0.5) + 
    facet_grid(var ~ ., scales = "free_y") +
    theme_bw()
```

Both the false discovery proportion and number of rejects follow a strictly decreasing relationship with the step index. Only a subset of steps are included in the final output.

For each of the matrix elements in the output, we only examine the first column since this corresponds to the largest (and presumably final) entry in the `res$step` vector.

```{r}
mat_tib <- tibble(s = res$s[, 1],
                  pi = res$pi[, 1],
                  mu = res$mu[, 1])
```

We can also merge this table with the original `sim_df` data.frame of p-values.

```{r}
mat_tib <- mat_tib %>%
    bind_cols(sim_df)
```

Based on the notation used in Section 4 of Lei and Fithian (2016), `s` is the significance thresholds for the p-values calculated as a function of the informative covariate, here `rank(ind_covariate)`. The other two variables, `pi` and `mu` are the parameters of the two-group Gamma GLM mixture model (fit to the negative log p-values) used to determine the AdaPT threshold, `s`.

We can start by plotting the `s` as a function of the informative covariate and faceted by the truth (`H = 0` or `H = 1`).

```{r, fig.width = 7, fig.height = 4}
mat_tib %>%
    ggplot(aes(x = rank(ind_covariate), y = s)) +
    geom_point(alpha = 1/20) + 
    theme_bw() + facet_grid(factor(H) ~ .)
```

That's a very interesting looking curve, and the trend follows what we expect. We can also check if our understanding of `s` is correct by thresholding the p-values against the `s` cutoff and comparing against the `num_rej` values of the output.

```{r}
all(colSums(sim_df$pval < res$s) == res$num_rej)
```

Wonderful!

### 2017-11-03 datasets/simulations/analysis-01-plot.Rmd
Parse t-test simulations at alpha = 0.05

@@
 ```{r, load-libraries, message=FALSE}
-library("dplyr")
-library("ggplot2")
-library("viridis")
 library("data.table")
 library("SummarizedBenchmark")
+library("tidyverse")
+library("magrittr")
 ```
@@
 ```
+
+<!-- ######################################################################## -->
+## Plot Cross-Setting Metrics (100 replications) just "alpha = 0.05"
+<!-- ######################################################################## -->
+
+In the previous section, we generated a single plot for each metric under each simulation setting as a function of the nominal FDR threshold over a range of 0.01 to 0.10.
+In this section, we instead just focus on the FDR threshold of 0.05 and generate plots as a function of the simulation settings.
+Thus, each family of simulation settings is reduced to a single plot for each metric, making it easier to summarize and compare the performance of each method across a large set of settings.
+
+Here, we simply parse and save the merged performance metrics. Results are plotted in `analysis-03-plot.Rmd`.
+
+```{r}
+library("tidyverse")
+library("SummarizedBenchmark")
+
+## list of results from 100 replication experiments
+dl <- list.dirs("data-tsim/M100", recursive=FALSE, full.names=TRUE)
+
+## running in batch
+for (dd in dl) {
+    fl <- list.files(dd, pattern=".rds", full.names=TRUE)
+
+    outp <- gsub("data", "plots", dd)
+    dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+    ## generate same plots for all parameter settings
+    tsbl <- list()
+    for (f in fl) {
+        t1 <- proc.time()
+        ## read process data
+        res <- readRDS(f)
+        tsb <- plotsim_standardize(res, alpha = 0.05)
+        tsbs <- tsb %>%
+            filter(!(blabel %in% c("bl-df02", "bl-df04", "bl-df05", 
+                                   paste0("ihw-a0", c(1:4, 6:9)), "ihw-a10")),
+                   alpha == 0.05) %>%
+            select(blabel, performanceMetric, value, rep)
+        tsbl[[gsub(".*/(.*)\\.rds", "\\1", f)]] <- tsbs
+        print(paste0("done with ", f, "; time: ", proc.time()[3] - t1[3]))
+    }
+    tsbc <- bind_rows(tsbl, .id = "param") %>% as.tibble
+
+    ## save the parsed tables for easier future access
+    saveRDS(tsbc, file.path(dd, "results-performanceMetrics-alpha05.rds"))
+}
+```

### 2017-11-03 datasets/simulations/analysis-03-plot-summaries.Rmd
Parse t-test simulations at alpha = 0.05

---
title: "Benchmarking FDR Methods - Summarizing Simulation Results"
author: "Rafalab Journal Club Members"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    highlight: tango
    number_sections: true
    code_folding: hide
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, fig.width = 12, fig.height = 10)
```

<!-- ######################################################################## -->
## Workspace Setup
<!-- ######################################################################## -->

```{r, load-libraries, message=FALSE}
library("data.table")
library("SummarizedBenchmark")
library("tidyverse")
library("magrittr")
library("forcats")
```

```{r, load-helper-functions}
source("../R/plotsim_standardize.R")
```

<!-- ######################################################################## -->
## Overview
<!-- ######################################################################## -->

In this analysis, we look at summarizing the results for the various simulation settings considered under the two-sample t-test setting. Each simulation included 20,000 two-sample t-tests performed over a variable number of null (group difference 0) and alternative (group difference > 0) datasets. Different simulation settings were considered to evaluate the impact of varying group sample sizes, effect sizes, effect size distributions, proportions of null hypotheses, and presence or absence of an informative covariate. More details on the simulation settings considered are included in the header of the `tsim-core-settings.R` script. Each simulation setting was replicated 100 times.

In a previous analysis (`analysis-01-plot.Rmd`), each simulation setting was considered and plotted separately with performance metrics reported over a range of FDR ("alpha") thresholds. However, since the total number of settings considered is quite large, processing and interpreting these results was both cumbersome and difficult. To get a higher-level picture of the simulation results, here we consider a single FDR threshold, `alpha = 0.05` and plot performance metrics for each method within sets of related simulation settings. Plots for FPR and "rejectprop" (rejection proportion) include blue horizontal lines at the expected FDR level (`0.05`) and proportion of alternative hypotheses (`1 - pi0`), respectively.

For simulation settings where the effect size (expected group differece) is not varying, a fixed difference of `1.5` was used for all alternative hypotheses. For simulation settings with the uniform (un-informative) or SE informative covariate, if the proportion of null hypotheses is not varying, `80%` of hypotheses were simulated under the null. For simulation settings with the "Boca-Leek" style informative covariate, a functional relationship was first defined between the probability of a test being null and the "informative covariate". Then, the informative covariates were sampled from Uniform(0, 1) and used to randomly assign each test as null or alternative according to the predefined functional relationship. The form of the relationship used here was taken from Setting II of the Boca-Leek simulations.

For easier comparison across settings, we fix the order of the methods first. The ordering is roughly based on the anti-conservative/conservative-ness of each method with covariate-aware methods included as a block.

```{r}
lab_order <- c("Bonferroni", "Benjamini-Hochberg", "q-value", "local FDR",
               "IHW", "Boca-Leek", "FDRreg (theoretical)", "FDRreg (empirical)",
               "ASH", "unadjusted")
```

Similarly, we define a consistent re-naming for each method for easier printing, and also add a column for differentiating covariate-aware methods from the other methods.

```{r}
re_blabel <- function(tab) {
    tsb %>%
        mutate(blabel = fct_recode(f = blabel, "IHW" = "ihw-a05", "ASH" = "ashs",
                                   "Boca-Leek" = "bl-df03", "local FDR" = "lfdr",
                                   "FDRreg (theoretical)" = "scott-theoretical",
                                   "FDRreg (empirical)" = "scott-empirical",
                                   "Benjamini-Hochberg" = "bh",
                                   "Bonferroni" = "bonf", "q-value" = "qvalue"),
               takes_cov = blabel %in%
                   c("IHW", "ASH", "Boca-Leek", "FDRreg (theoretical)", "FDRreg (empirical)")) %>%
        mutate(blabel = factor(blabel, levels = lab_order))
}
```

Results are reported using the following common plotting style.

```{r}
## summarize results in plot
gen_plot <- function(ptitle, tab, co) {
    tab %>%
        ggplot(aes(x = blabel, y = value, color = takes_cov)) +
        geom_boxplot() +
        geom_jitter(alpha = 1/10) +
        scale_y_continuous("", labels = scales::percent) +
        scale_color_brewer(palette = "Paired", guide = FALSE) + 
        xlab("") + 
        theme_bw() +
        theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
        facet_grid(performanceMetric ~ param) +
        geom_hline(aes(yintercept = expected), data = co,
                   color = "blue", alpha = 1/2) +
        ggtitle(ptitle)
}
```

<!-- ######################################################################## -->
## Varying Constant Effect Size
<!-- ######################################################################## -->

<!-- ######################################################################## -->
### Non-Informative Covariate

```{r}
f <- "data-tsim/M100/esize_fixed-uniform/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

outp <- gsub("data", "plots", dirname(f))
dir.create(outp, showWarnings=FALSE, recursive=TRUE)

## remove non-percentage metric (rejections),
## subset on settings (effect < 1.0 not really informative)
tsb %<>%
    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
           performanceMetric =
               factor(performanceMetric,
                      levels = c("rejections", "rejectprop", "TPR",
                                 "TNR", "FPR", "FNR", "FWER"))) %>%
    filter(!(performanceMetric %in% c("rejections", "FWER")),
           param >= 10) %>%
    mutate(param = paste("effect size =", sprintf("%.1f", param / 10)))

## relabel methods for easier reading
tsb %<>% re_blabel

## expected thresholds based on simulation setting
cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
                                                 levels=levels(tsb$performanceMetric)),
                      expected = c(0.05, 0.20))

## summarize results in plot
gen_plot("Two-Sample T-Tests w/ Varying Effect Size, Non-Informative Covariate (100 replications)",
         tsb, cutoffs)
```

<!-- ######################################################################## -->
### SE as Informative Covariate

```{r}
f <- "data-tsim/M100/esize_fixed-se/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

outp <- gsub("data", "plots", dirname(f))
dir.create(outp, showWarnings=FALSE, recursive=TRUE)

## remove non-percentage metric (rejections),
## subset on settings (effect < 1.0 not really informative)
tsb %<>%
    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
           performanceMetric =
               factor(performanceMetric,
                      levels = c("rejections", "rejectprop", "TPR",
                                 "TNR", "FPR", "FNR", "FWER"))) %>%
    filter(!(performanceMetric %in% c("rejections", "FWER")),
           param >= 10) %>%
    mutate(param = paste("effect size =", sprintf("%.1f", param / 10)))

## relabel methods for easier reading
tsb %<>% re_blabel

## expected thresholds based on simulation setting
cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
                                                 levels=levels(tsb$performanceMetric)),
                      expected = c(0.05, 0.20))

## summarize results in plot
gen_plot("Two-Sample T-Tests w/ Varying Effect Size, SE as Covariate (100 replications)",
         tsb, cutoffs)
```

<!-- ######################################################################## -->
### Boca-Leek Style Informative Covariate

```{r}
f <- "data-tsim/M100/esize_fixed-bl/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

outp <- gsub("data", "plots", dirname(f))
dir.create(outp, showWarnings=FALSE, recursive=TRUE)

## remove non-percentage metric (rejections),
## subset on settings (effect < 1.0 not really informative)
tsb %<>%
    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
           performanceMetric =
               factor(performanceMetric,
                      levels = c("rejections", "rejectprop", "TPR",
                                 "TNR", "FPR", "FNR", "FWER"))) %>%
    filter(!(performanceMetric %in% c("rejections", "FWER")),
           param >= 10) %>%
    mutate(param = paste("effect size =", sprintf("%.1f", param / 10)))

## relabel methods for easier reading
tsb %<>% re_blabel

## expected thresholds based on simulation setting
cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
                                                 levels=levels(tsb$performanceMetric)),
                      expected = c(0.05, 0.074))

## summarize results in plot
gen_plot("Two-Sample T-Tests w/ Varying Effect Size, Boca-Leek Style Covariate (100 replications)",
         tsb, cutoffs)
```

<!-- ######################################################################## -->
## Varying Null Proportion
<!-- ######################################################################## -->

<!-- ######################################################################## -->
### Non-Informative Covariate

```{r}
f <- "data-tsim/M100/pi0-uniform/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

outp <- gsub("data", "plots", dirname(f))
dir.create(outp, showWarnings=FALSE, recursive=TRUE)

## remove rejections: scaling is off from rest (also dup'd w/ rejprop)
## remove FWER: not informative as a boxplot
tsb %<>%
    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
           performanceMetric =
               factor(performanceMetric,
                      levels = c("rejections", "rejectprop", "TPR",
                                 "TNR", "FPR", "FNR", "FWER"))) %>%
    filter(!(performanceMetric %in% c("rejections", "FWER"))) %>%
    mutate(param = factor(paste0("null = ", param * 10, "%"),
                          levels = paste0("null = ", seq(10, 100, 10), "%")))

## relabel methods for easier reading
tsb %<>% re_blabel

## expected thresholds based on simulation setting
cutoffs <- data.frame(performanceMetric =
                          factor(c(rep("FPR", 10), rep("rejectprop", 10)),
                                 levels=levels(tsb$performanceMetric)),
                      param = rep(paste0("null = ", seq(10, 100, 10), "%"), 2),
                      expected = c(rep(0.05, 10), seq(.9, 0, -.1)))

## summarize results in plot
gen_plot("Two-Sample T-Tests w/ Varying Null Proportion, Non-Informative Covariate (100 replications)",
         tsb, cutoffs)
```

<!-- ######################################################################## -->
### SE as Informative Covariate

```{r}
f <- "data-tsim/M100/pi0-se/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

outp <- gsub("data", "plots", dirname(f))
dir.create(outp, showWarnings=FALSE, recursive=TRUE)

## remove rejections: scaling is off from rest (also dup'd w/ rejprop)
## remove FWER: not informative as a boxplot
tsb %<>%
    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
           performanceMetric =
               factor(performanceMetric,
                      levels = c("rejections", "rejectprop", "TPR",
                                 "TNR", "FPR", "FNR", "FWER"))) %>%
    filter(!(performanceMetric %in% c("rejections", "FWER"))) %>%
    mutate(param = factor(paste0("null = ", param * 10, "%"),
                          levels = paste0("null = ", seq(10, 100, 10), "%")))

## relabel methods for easier reading
tsb %<>% re_blabel

## expected thresholds based on simulation setting
cutoffs <- data.frame(performanceMetric =
                          factor(c(rep("FPR", 10), rep("rejectprop", 10)),
                                 levels=levels(tsb$performanceMetric)),
                      param = rep(paste0("null = ", seq(10, 100, 10), "%"), 2),
                      expected = c(rep(0.05, 10), seq(.9, 0, -.1)))

## summarize results in plot
gen_plot("Two-Sample T-Tests w/ Varying Null Proportion, SE as Covariate (100 replications)",
         tsb, cutoffs)
```

<!-- ######################################################################## -->
## Varying Sample Sizes
<!-- ######################################################################## -->

<!-- ######################################################################## -->
### Non-Informative Covariate

```{r}
f <- "data-tsim/M100/n-uniform/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

## can also add default of n = 10 from alt pi0 simulation set
## (same effect_size = 1.5, pi0 = 0.8)
f2 <- "data-tsim/M100/pi0-uniform/results-performanceMetrics-alpha05.rds"
tsb2 <- readRDS(f2)
tsb2 %<>%
    filter(param == "results-altp0_8-M100") %>%
    mutate(param = "results-altn_10-M100")

tsb <- bind_rows(tsb, tsb2)

outp <- gsub("data", "plots", dirname(f))
dir.create(outp, showWarnings=FALSE, recursive=TRUE)

## remove rejections: scaling is off from rest (also dup'd w/ rejprop)
## remove FWER: not informative as a boxplot
tsb %<>%
    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
           performanceMetric =
               factor(performanceMetric,
                      levels = c("rejections", "rejectprop", "TPR",
                                 "TNR", "FPR", "FNR", "FWER"))) %>%
    filter(!(performanceMetric %in% c("rejections", "FWER"))) %>%
    mutate(param = factor(paste0("n (per group) = ", param),
                          levels = paste0("n (per group) = ", c(3, 6, 10, 20))))

## relabel methods for easier reading
tsb %<>% re_blabel

## expected thresholds based on simulation setting
cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
                                                 levels=levels(tsb$performanceMetric)),
                      expected = c(0.05, 0.20))

## summarize results in plot
gen_plot("Two-Sample T-Tests w/ Varying Sample Sizes, Non-Informative Covariate (100 replications)",
         tsb, cutoffs)
```

<!-- ######################################################################## -->
### SE as Informative Covariate

```{r}
f <- "data-tsim/M100/n-se/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

## can also add default of n = 10 from alt pi0 simulation set
## (same effect_size = 1.5, pi0 = 0.8)
f2 <- "data-tsim/M100/pi0-se/results-performanceMetrics-alpha05.rds"
tsb2 <- readRDS(f2)
tsb2 %<>%
    filter(param == "results-altp0_8-M100") %>%
    mutate(param = "results-altn_10-M100")

tsb <- bind_rows(tsb, tsb2)

outp <- gsub("data", "plots", dirname(f))
dir.create(outp, showWarnings=FALSE, recursive=TRUE)

## remove rejections: scaling is off from rest (also dup'd w/ rejprop)
## remove FWER: not informative as a boxplot
tsb %<>%
    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
           performanceMetric =
               factor(performanceMetric,
                      levels = c("rejections", "rejectprop", "TPR",
                                 "TNR", "FPR", "FNR", "FWER"))) %>%
    filter(!(performanceMetric %in% c("rejections", "FWER"))) %>%
    mutate(param = factor(paste0("n (per group) = ", param),
                          levels = paste0("n (per group) = ", c(3, 6, 10, 20))))

## relabel methods for easier reading
tsb %<>% re_blabel

## expected thresholds based on simulation setting
cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
                                                 levels=levels(tsb$performanceMetric)),
                      expected = c(0.05, 0.20))

## summarize results in plot
gen_plot("Two-Sample T-Tests w/ Varying Sample Sizes, SE as Covariate (100 replications)",
         tsb, cutoffs)
```

<!-- ######################################################################## -->
### Boca-Leek Style Informative Covariate

```{r}
f <- "data-tsim/M100/n-bl/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

outp <- gsub("data", "plots", dirname(f))
dir.create(outp, showWarnings=FALSE, recursive=TRUE)

## remove rejections: scaling is off from rest (also dup'd w/ rejprop)
## remove FWER: not informative as a boxplot
tsb %<>%
    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
           performanceMetric =
               factor(performanceMetric,
                      levels = c("rejections", "rejectprop", "TPR",
                                 "TNR", "FPR", "FNR", "FWER"))) %>%
    filter(!(performanceMetric %in% c("rejections", "FWER"))) %>%
    mutate(param = factor(paste0("n (per group) = ", param),
                          levels = paste0("n (per group) = ", c(3, 6, 20))))

## relabel methods for easier reading
tsb %<>% re_blabel

## expected thresholds based on simulation setting
cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
                                                 levels=levels(tsb$performanceMetric)),
                      expected = c(0.05, 0.074))

## summarize results in plot
gen_plot("Two-Sample T-Tests w/ Varying Sample Sizes, Boca-Leek Style Covariate (100 replications)",
         tsb, cutoffs)
```

<!-- ######################################################################## -->
## Randomly-Sampled Effect Size
<!-- ######################################################################## -->

<!-- ######################################################################## -->
### Non-Informative Covariate

```{r}
f <- "data-tsim/M100/esize_random-uniform/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

outp <- gsub("data", "plots", dirname(f))
dir.create(outp, showWarnings=FALSE, recursive=TRUE)

## remove non-percentage metric (rejections),
## subset on settings (effect < 1.0 not really informative)
tsb %<>%
    mutate(param = gsub(".*?-.*?_(.*)-.+", "\\1", param),
           performanceMetric =
               factor(performanceMetric,
                      levels = c("rejections", "rejectprop", "TPR",
                                 "TNR", "FPR", "FNR", "FWER"))) %>%
    filter(!(performanceMetric %in% c("rejections", "FWER")))

## relabel methods for easier reading
tsb %<>% re_blabel

## expected thresholds based on simulation setting
cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
                                                 levels=levels(tsb$performanceMetric)),
                      expected = c(0.05, 0.20))

## summarize results in plot
gen_plot("Two-Sample T-Tests w/ Varying Randomly Sampled Effect, Non-Informative Covariate (100 replications)",
         tsb, cutoffs)
```

<!-- ######################################################################## -->
### SE as Informative Covariate

```{r}
f <- "data-tsim/M100/esize_random-se/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

outp <- gsub("data", "plots", dirname(f))
dir.create(outp, showWarnings=FALSE, recursive=TRUE)

## remove non-percentage metric (rejections),
## subset on settings (effect < 1.0 not really informative)
tsb %<>%
    mutate(param = gsub(".*?-.*?_(.*)-.+", "\\1", param),
           performanceMetric =
               factor(performanceMetric,
                      levels = c("rejections", "rejectprop", "TPR",
                                 "TNR", "FPR", "FNR", "FWER"))) %>%
    filter(!(performanceMetric %in% c("rejections", "FWER")))

## relabel methods for easier reading
tsb %<>% re_blabel

## expected thresholds based on simulation setting
cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
                                                 levels=levels(tsb$performanceMetric)),
                      expected = c(0.05, 0.20))

## summarize results in plot
gen_plot("Two-Sample T-Tests w/ Varying Randomly Sampled Effect, SE as Covariate (100 replications)",
         tsb, cutoffs)
```

<!-- ######################################################################## -->
### Boca-Leek Style Informative Covariate

```{r}
f <- "data-tsim/M100/esize_random-bl/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

outp <- gsub("data", "plots", dirname(f))
dir.create(outp, showWarnings=FALSE, recursive=TRUE)

## remove non-percentage metric (rejections),
## subset on settings (effect < 1.0 not really informative)
tsb %<>%
    mutate(param = gsub(".*?-.*?_(.*)-.+", "\\1", param),
           performanceMetric =
               factor(performanceMetric,
                      levels = c("rejections", "rejectprop", "TPR",
                                 "TNR", "FPR", "FNR", "FWER"))) %>%
    filter(!(performanceMetric %in% c("rejections", "FWER")))

## relabel methods for easier reading
tsb %<>% re_blabel

## expected thresholds based on simulation setting
cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
                                                 levels=levels(tsb$performanceMetric)),
                      expected = c(0.05, 0.074))

## summarize results in plot
gen_plot("Two-Sample T-Tests w/ Varying Randomly Sampled Effect, Boca-Leek Style Covariate (100 replications)",
         tsb, cutoffs)
```

### 2017-11-03 datasets/simulations/psim-submit.R
Fix psim simulation parameter names

@@ -13,8 +13,8 @@ params <- list(c("pi0", "uniform"),
                c("esize_random_ua", "bl"),
                c("esize_random_shift", "uniform"),
                c("esize_random_shift", "bl"),
-               c("esize_random_altnoise", "uniform"),
-               c("esize_random_altnoise", "bl"))
+               c("altnoise", "uniform"),
+               c("altnoise", "bl"))
 
 for (ip in params) {
     cmd <- paste("sbatch",

### 2017-11-03 datasets/simulations/analysis-03-plot-tsim-summaries.Rmd
Fix typos in t-test sim summary plots

--- 2017-11-03 datasets/simulations/analysis-03-plot-summaries.Rmd
@@
 
-In this analysis, we look at summarizing the results for the various simulation settings considered under the two-sample t-test setting.
+In this analysis, we look at summarizing the results for the various simulation settings considered under the two-sample t-test setup.
 Each simulation included 20,000 two-sample t-tests performed over a variable number of null (group difference 0) and alternative (group difference > 0) datasets.
-Different simulation settings were considered to evaluate the impact of varying group sample sizes, effect sizes, effect size distributions, proportions of null hypotheses, and presence or absence of an informative covariate.
+Different simulation settings were considered to evaluate the impact of varying sample sizes, effect sizes, effect size distributions, proportions of null hypotheses, and presence or absence of an informative covariate.
 More details on the simulation settings considered are included in the header of the `tsim-core-settings.R` script.
-Each simulation setting was replicated 100 times.
+Each setting was replicated 100 times.
 
@@
 However, since the total number of settings considered is quite large, processing and interpreting these results was both cumbersome and difficult.
-To get a higher-level picture of the simulation results, here we consider a single FDR threshold, `alpha = 0.05` and plot performance metrics for each method within sets of related simulation settings.
+To get a higher-level picture of the simulation results, here we consider a single FDR threshold, `alpha = 0.05`, and plot performance metrics across settings for each method.
 Plots for FPR and "rejectprop" (rejection proportion) include blue horizontal lines at the expected FDR level (`0.05`) and proportion of alternative hypotheses (`1 - pi0`), respectively.
@@
 <!-- ######################################################################## -->
-### SE as Informative Covariate
+### SE as Covariate
 
@@
 <!-- ######################################################################## -->
-### Boca-Leek Style Informative Covariate
+### Boca-Leek Style Covariate
 

### 2017-11-03 datasets/simulations/analysis-03-plot-tsim-summaries.Rmd
Move t-test sim summary parsing from 01 to 03

@@
 source("../R/plotsim_standardize.R")
+```
+
+```{r, process-data, message=FALSE}
+## list of results from 100 replication experiments
+dl <- list.dirs("data-tsim/M100", recursive=FALSE, full.names=TRUE)
+
+## parse each setting-block separately and save metrics in single file
+for (dd in dl) {
+    if (!file.exists(file.path(dd, "results-performanceMetrics-alpha05.rds"))) {
+        fl <- list.files(dd, pattern=".rds", full.names=TRUE)
+        tsbl <- list()
+        for (f in fl) {
+            ## read process data
+            res <- readRDS(f)
+            tsb <- plotsim_standardize(res, alpha = 0.05)
+            tsbs <- tsb %>%
+                filter(!(blabel %in% c("bl-df02", "bl-df04", "bl-df05", 
+                                       paste0("ihw-a0", c(1:4, 6:9)), "ihw-a10")),
+                       alpha == 0.05) %>%
+                select(blabel, performanceMetric, value, rep)
+            tsbl[[gsub(".*/(.*)\\.rds", "\\1", f)]] <- tsbs
+        }
+        tsbc <- as.tibble(bind_rows(tsbl, .id = "param"))
+        saveRDS(tsbc, file.path(dd, "results-performanceMetrics-alpha05.rds"))
+    }
+}
 ```

### 2017-11-06 datasets/simulations/analysis-04-plot-psim-summaries.Rmd
Add initial analysis of psim results

--- 2017-11-03 datasets/simulations/analysis-03-plot-tsim-summaries.Rmd
@@
 ---
-title: "Benchmarking FDR Methods - Summarizing Simulation Results"
+title: "Benchmarking FDR Methods - Summarizing Simulation Results (Part II)"
 author: "Rafalab Journal Club Members"
@@
 ## list of results from 100 replication experiments
-dl <- list.dirs("data-tsim/M100", recursive=FALSE, full.names=TRUE)
+dl <- list.dirs("data-psim/M100", recursive=FALSE, full.names=TRUE)
 
@@
 
-In this analysis, we look at summarizing the results for the various simulation settings considered under the two-sample t-test setup.
-Each simulation included 20,000 two-sample t-tests performed over a variable number of null (group difference 0) and alternative (group difference > 0) datasets.
-Different simulation settings were considered to evaluate the impact of varying sample sizes, effect sizes, effect size distributions, proportions of null hypotheses, and presence or absence of an informative covariate.
-More details on the simulation settings considered are included in the header of the `tsim-core-settings.R` script.
+This analysis is similar to the one presented in `analysis-03-plot-tsim-summaries.[Rmd|html]`.
+However, rather than looking at the results of two-sample t-tests, here, we consider simulations performed directly at the level of effect sizes and p-values without any underlying data.
+This approach to simulating data follows similar simulations carried out in the Boca-Leek, IHW ("size investing strategy" simulations), and ASH papers.
+Different simulation settings were considered to evaluate the impact of varying effect sizes, effect size distributions, effect size sampling distributions, proportions of null hypotheses, and presence or absence of an informative covariate.
+More details on the simulation settings considered are included in the header of the `psim-core-settings.R` script.
 Each setting was replicated 100 times.
 
-In a previous analysis (`analysis-01-plot.Rmd`), each simulation setting was considered and plotted separately with performance metrics reported over a range of FDR ("alpha") thresholds.
-However, since the total number of settings considered is quite large, processing and interpreting these results was both cumbersome and difficult.
-To get a higher-level picture of the simulation results, here we consider a single FDR threshold, `alpha = 0.05`, and plot performance metrics across settings for each method.
-Plots for FPR and "rejectprop" (rejection proportion) include blue horizontal lines at the expected FDR level (`0.05`) and proportion of alternative hypotheses (`1 - pi0`), respectively.
+Since these simulations do not rely on underlying simulated data, but rather simulate summary statistics directly, there is no way of calculating a corresponding standard error (SE) for each hypothesis test.
+*As a result, the ASH method, which requires the SE for each test, is not included in these simulations.*
 
-For simulation settings where the effect size (expected group differece) is not varying, a fixed difference of `1.5` was used for all alternative hypotheses.
-For simulation settings with the uniform (un-informative) or SE informative covariate, if the proportion of null hypotheses is not varying, `80%` of hypotheses were simulated under the null.
-For simulation settings with the "Boca-Leek" style informative covariate, a functional relationship was first defined between the probability of a test being null and the "informative covariate".
-Then, the informative covariates were sampled from Uniform(0, 1) and used to randomly assign each test as null or alternative according to the predefined functional relationship.
-The form of the relationship used here was taken from Setting II of the Boca-Leek simulations.
+As in `analysis-03-plot-tsim-summaries.[Rmd|html]`, simulation results are reported across settings for each method at the `alpha = 0.05` FDR threshold.
 
@@
                "IHW", "Boca-Leek", "FDRreg (theoretical)", "FDRreg (empirical)",
-               "ASH", "unadjusted")
+               "unadjusted")
 ```
@@
     tsb %>%
-        mutate(blabel = fct_recode(f = blabel, "IHW" = "ihw-a05", "ASH" = "ashs",
+        mutate(blabel = fct_recode(f = blabel, "IHW" = "ihw-a05",
                                    "Boca-Leek" = "bl-df03", "local FDR" = "lfdr",
@@
                takes_cov = blabel %in%
-                   c("IHW", "ASH", "Boca-Leek", "FDRreg (theoretical)", "FDRreg (empirical)")) %>%
+                   c("IHW", "Boca-Leek", "FDRreg (theoretical)", "FDRreg (empirical)")) %>%
         mutate(blabel = factor(blabel, levels = lab_order))
@@
 ```{r}
-f <- "data-tsim/M100/esize_fixed-uniform/results-performanceMetrics-alpha05.rds"
+f <- "data-psim/M100/esize_fixed-uniform/results-performanceMetrics-alpha05.rds"
 tsb <- readRDS(f)
@@
 ## summarize results in plot
-gen_plot("Two-Sample T-Tests w/ Varying Effect Size, Non-Informative Covariate (100 replications)",
-         tsb, cutoffs)
-```
-
-<!-- ######################################################################## -->
-### SE as Covariate
-
-```{r}
-f <- "data-tsim/M100/esize_fixed-se/results-performanceMetrics-alpha05.rds"
-tsb <- readRDS(f)
-
-outp <- gsub("data", "plots", dirname(f))
-dir.create(outp, showWarnings=FALSE, recursive=TRUE)
-
-## remove non-percentage metric (rejections),
-## subset on settings (effect < 1.0 not really informative)
-tsb %<>%
-    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
-           performanceMetric =
-               factor(performanceMetric,
-                      levels = c("rejections", "rejectprop", "TPR",
-                                 "TNR", "FPR", "FNR", "FWER"))) %>%
-    filter(!(performanceMetric %in% c("rejections", "FWER")),
-           param >= 10) %>%
-    mutate(param = paste("effect size =", sprintf("%.1f", param / 10)))
-
-## relabel methods for easier reading
-tsb %<>% re_blabel
-
-## expected thresholds based on simulation setting
-cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
-                                                 levels=levels(tsb$performanceMetric)),
-                      expected = c(0.05, 0.20))
-
-## summarize results in plot
-gen_plot("Two-Sample T-Tests w/ Varying Effect Size, SE as Covariate (100 replications)",
+gen_plot("Direct Parameter Simulation w/ Varying Effect Size, Non-Informative Covariate (100 replications)",
          tsb, cutoffs) 
@@
 ```{r}
-f <- "data-tsim/M100/esize_fixed-bl/results-performanceMetrics-alpha05.rds"
+f <- "data-psim/M100/esize_fixed-bl/results-performanceMetrics-alpha05.rds"
 tsb <- readRDS(f)
@@
 ## summarize results in plot
-gen_plot("Two-Sample T-Tests w/ Varying Effect Size, Boca-Leek Style Covariate (100 replications)",
+gen_plot("Direct Parameter Simulation w/ Varying Effect Size, Boca-Leek Style Covariate (100 replications)",
          tsb, cutoffs)
@@
 
+Need to re-run these simulations with larger test statistic value under the alternative - TPR is too low to be particularly interesting/informative for comparing across methods.
+
 <!-- ######################################################################## -->
@@
 ```{r}
-f <- "data-tsim/M100/pi0-uniform/results-performanceMetrics-alpha05.rds"
+## older run with alternative tests at 1.5.
+f <- "data-psim/M100/_pi0-uniform-delta15/results-performanceMetrics-alpha05.rds"
 tsb <- readRDS(f)
@@
 ## summarize results in plot
-gen_plot("Two-Sample T-Tests w/ Varying Null Proportion, Non-Informative Covariate (100 replications)",
+gen_plot("Direct Parameter Simulation w/ Varying Null Proportion, Non-Informative Covariate (100 replications)",
          tsb, cutoffs)
@@
 <!-- ######################################################################## -->
-### SE as Covariate
+## Randomly-Sampled UA Effect Size
+<!-- ######################################################################## -->
 
-```{r}
-f <- "data-tsim/M100/pi0-se/results-performanceMetrics-alpha05.rds"
-tsb <- readRDS(f)
-
-outp <- gsub("data", "plots", dirname(f))
-dir.create(outp, showWarnings=FALSE, recursive=TRUE)
-
-## remove rejections: scaling is off from rest (also dup'd w/ rejprop)
-## remove FWER: not informative as a boxplot
-tsb %<>%
-    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
-           performanceMetric =
-               factor(performanceMetric,
-                      levels = c("rejections", "rejectprop", "TPR",
-                                 "TNR", "FPR", "FNR", "FWER"))) %>%
-    filter(!(performanceMetric %in% c("rejections", "FWER"))) %>%
-    mutate(param = factor(paste0("null = ", param * 10, "%"),
-                          levels = paste0("null = ", seq(10, 100, 10), "%")))
-
-## relabel methods for easier reading
-tsb %<>% re_blabel
-
-## expected thresholds based on simulation setting
-cutoffs <- data.frame(performanceMetric =
-                          factor(c(rep("FPR", 10), rep("rejectprop", 10)),
-                                 levels=levels(tsb$performanceMetric)),
-                      param = rep(paste0("null = ", seq(10, 100, 10), "%"), 2),
-                      expected = c(rep(0.05, 10), seq(.9, 0, -.1)))
-
-## summarize results in plot
-gen_plot("Two-Sample T-Tests w/ Varying Null Proportion, SE as Covariate (100 replications)",
-         tsb, cutoffs)
-```
-
-<!-- ######################################################################## -->
-## Varying Sample Sizes
-<!-- ######################################################################## -->
+Need to re-run these simulations with lower sampling variance - TPR is too low to be particularly interesting/informative for comparing across methods.
 
@@
 ```{r}
-f <- "data-tsim/M100/n-uniform/results-performanceMetrics-alpha05.rds"
-tsb <- readRDS(f)
-
-## can also add default of n = 10 from alt pi0 simulation set
-## (same effect_size = 1.5, pi0 = 0.8)
-f2 <- "data-tsim/M100/pi0-uniform/results-performanceMetrics-alpha05.rds"
-tsb2 <- readRDS(f2)
-tsb2 %<>%
-    filter(param == "results-altp0_8-M100") %>%
-    mutate(param = "results-altn_10-M100")
-
-tsb <- bind_rows(tsb, tsb2)
-
-outp <- gsub("data", "plots", dirname(f))
-dir.create(outp, showWarnings=FALSE, recursive=TRUE)
-
-## remove rejections: scaling is off from rest (also dup'd w/ rejprop)
-## remove FWER: not informative as a boxplot
-tsb %<>%
-    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
-           performanceMetric =
-               factor(performanceMetric,
-                      levels = c("rejections", "rejectprop", "TPR",
-                                 "TNR", "FPR", "FNR", "FWER"))) %>%
-    filter(!(performanceMetric %in% c("rejections", "FWER"))) %>%
-    mutate(param = factor(paste0("n (per group) = ", param),
-                          levels = paste0("n (per group) = ", c(3, 6, 10, 20))))
-
-## relabel methods for easier reading
-tsb %<>% re_blabel
-
-## expected thresholds based on simulation setting
-cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
-                                                 levels=levels(tsb$performanceMetric)),
-                      expected = c(0.05, 0.20))
-
-## summarize results in plot
-gen_plot("Two-Sample T-Tests w/ Varying Sample Sizes, Non-Informative Covariate (100 replications)",
-         tsb, cutoffs)
-```
-
-<!-- ######################################################################## -->
-### SE as Covariate
-
-```{r}
-f <- "data-tsim/M100/n-se/results-performanceMetrics-alpha05.rds"
-tsb <- readRDS(f)
-
-## can also add default of n = 10 from alt pi0 simulation set
-## (same effect_size = 1.5, pi0 = 0.8)
-f2 <- "data-tsim/M100/pi0-se/results-performanceMetrics-alpha05.rds"
-tsb2 <- readRDS(f2)
-tsb2 %<>%
-    filter(param == "results-altp0_8-M100") %>%
-    mutate(param = "results-altn_10-M100")
-
-tsb <- bind_rows(tsb, tsb2)
-
-outp <- gsub("data", "plots", dirname(f))
-dir.create(outp, showWarnings=FALSE, recursive=TRUE)
-
-## remove rejections: scaling is off from rest (also dup'd w/ rejprop)
-## remove FWER: not informative as a boxplot
-tsb %<>%
-    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
-           performanceMetric =
-               factor(performanceMetric,
-                      levels = c("rejections", "rejectprop", "TPR",
-                                 "TNR", "FPR", "FNR", "FWER"))) %>%
-    filter(!(performanceMetric %in% c("rejections", "FWER"))) %>%
-    mutate(param = factor(paste0("n (per group) = ", param),
-                          levels = paste0("n (per group) = ", c(3, 6, 10, 20))))
-
-## relabel methods for easier reading
-tsb %<>% re_blabel
-
-## expected thresholds based on simulation setting
-cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
-                                                 levels=levels(tsb$performanceMetric)),
-                      expected = c(0.05, 0.20))
-
-## summarize results in plot
-gen_plot("Two-Sample T-Tests w/ Varying Sample Sizes, SE as Covariate (100 replications)",
-         tsb, cutoffs)
-```
-
-<!-- ######################################################################## -->
-### Boca-Leek Style Covariate
-
-```{r}
-f <- "data-tsim/M100/n-bl/results-performanceMetrics-alpha05.rds"
-tsb <- readRDS(f)
-
-outp <- gsub("data", "plots", dirname(f))
-dir.create(outp, showWarnings=FALSE, recursive=TRUE)
-
-## remove rejections: scaling is off from rest (also dup'd w/ rejprop)
-## remove FWER: not informative as a boxplot
-tsb %<>%
-    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
-           performanceMetric =
-               factor(performanceMetric,
-                      levels = c("rejections", "rejectprop", "TPR",
-                                 "TNR", "FPR", "FNR", "FWER"))) %>%
-    filter(!(performanceMetric %in% c("rejections", "FWER"))) %>%
-    mutate(param = factor(paste0("n (per group) = ", param),
-                          levels = paste0("n (per group) = ", c(3, 6, 20))))
-
-## relabel methods for easier reading
-tsb %<>% re_blabel
-
-## expected thresholds based on simulation setting
-cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
-                                                 levels=levels(tsb$performanceMetric)),
-                      expected = c(0.05, 0.074))
-
-## summarize results in plot
-gen_plot("Two-Sample T-Tests w/ Varying Sample Sizes, Boca-Leek Style Covariate (100 replications)",
-         tsb, cutoffs)
-```
-
-<!-- ######################################################################## -->
-## Randomly-Sampled Effect Size
-<!-- ######################################################################## -->
-
-<!-- ######################################################################## -->
-### Non-Informative Covariate
-
-```{r}
-f <- "data-tsim/M100/esize_random-uniform/results-performanceMetrics-alpha05.rds"
+f <- "data-psim/M100/esize_random_ua-uniform/results-performanceMetrics-alpha05.rds"
 tsb <- readRDS(f)
@@
 ## summarize results in plot
-gen_plot("Two-Sample T-Tests w/ Varying Randomly Sampled Effect, Non-Informative Covariate (100 replications)",
+gen_plot("Direct Parameter Simulation w/ Varying Randomly Sampled UA Effect, Non-Informative Covariate (100 replications)",
          tsb, cutoffs)
@@
 <!-- ######################################################################## -->
-### SE as Covariate
+### Boca-Leek Style Covariate
 
 ```{r}
-f <- "data-tsim/M100/esize_random-se/results-performanceMetrics-alpha05.rds"
+f <- "data-psim/M100/esize_random_ua-bl/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*)-.+", "\\1", param),
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.074))
+
+## summarize results in plot
+gen_plot("Direct Parameter Simulation w/ Varying Randomly Sampled UA Effect, Boca-Leek Style Covariate (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
+## Randomly-Sampled Mean-Shift Effect Size
+<!-- ######################################################################## -->
+
+Chi-sq simulations look reasonable, but probably need to re-run the normal and t-test simulations with larger test statistic value under the alternative (possibly 3) - TPR is too low with shift of 1 and 2 to be particularly interesting/informative for comparing across methods.
+
+<!-- ######################################################################## -->
+### Non-Informative Covariate
+
+```{r}
+f <- "data-psim/M100/esize_random_shift-uniform/results-performanceMetrics-alpha05.rds"
 tsb <- readRDS(f)
@@
 ## summarize results in plot
-gen_plot("Two-Sample T-Tests w/ Varying Randomly Sampled Effect, SE as Covariate (100 replications)",
+gen_plot("Direct Parameter Simulation w/ Varying Randomly Sampled Mean Shifted Effect, Non-Informative Covariate (100 replications)",
          tsb, cutoffs)
@@
 ```{r}
-f <- "data-tsim/M100/esize_random-bl/results-performanceMetrics-alpha05.rds"
+f <- "data-psim/M100/esize_random_shift-bl/results-performanceMetrics-alpha05.rds"
 tsb <- readRDS(f)
@@
 ## summarize results in plot
-gen_plot("Two-Sample T-Tests w/ Varying Randomly Sampled Effect, Boca-Leek Style Covariate (100 replications)",
+gen_plot("Direct Parameter Simulation w/ Varying Randomly Sampled Mean Shifted Effect, Boca-Leek Style Covariate (100 replications)",
          tsb, cutoffs)
 ```
+
+<!-- ######################################################################## -->
+## Varying Parameter Sampling Distribution
+<!-- ######################################################################## -->
+
+Similar as the simulation settings considered above: chi-sq simulations look reasonable, but probably need to re-run the normal and t-test simulations with larger test statistic value under the alternative (possibly 3) - TPR is too low with shift of 1 and 2 to be particularly interesting/informative for comparing across methods.
+
+<!-- ######################################################################## -->
+### Non-Informative Covariate
+
+```{r}
+f <- "data-psim/M100/altnoise-uniform/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*)-.+", "\\1", param),
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Direct Parameter Simulation w/ Varying Sampling Distribution, Non-Informative Covariate (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
+### Boca-Leek Style Covariate
+
+```{r}
+f <- "data-psim/M100/altnoise-bl/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*)-.+", "\\1", param),
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.074))
+
+## summarize results in plot
+gen_plot("Direct Parameter Simulation w/ Varying Sampling Distribution, Boca-Leek Style Covariate (100 replications)",
+         tsb, cutoffs)
+```

### 2017-11-08 datasets/GSEA/gsea.Rmd
Added supplementary option to plots. Added plotFDRMethodsOverlap.

@@
 res$pval <- pmin( res$pval, 1 )
-head(res)
 sGSEA <- designBen %>% buildBench( res, ftCols="ind_covariate")
-elementMetadata( rowData(sGSEA) )
 assayNames(sGSEA) <- "qvalue"
@@
 ```{r plottingResults}
-rejections_scatter( sGSEA, as_fraction=FALSE )
-
-stopifnot( all(assays(sGSEA)[["qvalue"]][,"unadjusted"] == pmin(goRes$over_represented_pvalue, 1) ) )
+rejections_scatter( sGSEA, as_fraction=TRUE, supplementary=FALSE)
 
 sGSEA <- estimatePerformanceMetrics(sGSEA, addColData=TRUE)
-rejection_scatter_bins( sGSEA, "ind_covariate", threshold=15, bins=4, as_fraction=TRUE )
+rejection_scatter_bins( sGSEA, "ind_covariate", threshold=15, bins=4, as_fraction=TRUE, supplementary=FALSE)
+```
+
+```{r}
+
+plotFDRMethodsOverlap( sGSEA, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
 

### 2017-11-08 datasets/R/rejections_scatter.R

@@ -7,12 +7,13 @@
 #' @param sb A SummarizedBenchmark object
 #' @param asFraction Logical. Whether to plot the fraction of hypotheses
 #' that were rejected (TRUE) or the absolute number of rejections (FALSE).
+#' @param supplementary Logical. Is the figure a supplementary figure?
 #'        
 #' @return a ggplot2 object
 #' 
 #' 
 #' @author Alejandro Reyes
-rejections_scatter <- function( sb, as_fraction=FALSE ){
+rejections_scatter <- function( sb, as_fraction=FALSE, supplementary=TRUE ){
   stopifnot( is(sb, "SummarizedBenchmark" ) )
   alphas <- unique( as.numeric( as.character( colData( sb )$param.alpha ) ) )
   alphas <- alphas[!is.na(alphas)]
@@ -23,11 +24,19 @@ rejections_scatter <- function( sb, as_fraction=FALSE ){
     deno <- 1
     yl <- "Number of rejections"
   }
-  estimatePerformanceMetrics( sb, alpha=alphas, tidy=TRUE ) %>%
+  plotDF <- estimatePerformanceMetrics( sb, alpha=alphas, tidy=TRUE ) %>%
     dplyr:::filter( !(pkg_name == "IHW" & param.alpha != alpha ) ) %>%
     dplyr:::mutate( blabel=gsub("(ihw)-a\\d+", "\\1", blabel ) ) %>%
-    dplyr:::select( blabel, key, value, assay, performanceMetric, alpha ) %>%
-    dplyr:::filter( blabel!="unadjusted", performanceMetric == "rejections" ) %>%
+    #dplyr:::select( blabel, key, value, assay, performanceMetric, alpha ) %>%
+    dplyr:::filter( blabel!="unadjusted", performanceMetric == "rejections" ) 
+  if( !supplementary ){
+    plotDF <- plotDF %>%
+      dplyr:::mutate( param.smooth.df=gsub("L", "", param.smooth.df ) ) %>%
+      dplyr:::filter( !( pkg_name == "swfdr" & 
+                      as.numeric(as.character(param.smooth.df) != 3 ) ) ) %>%
+      dplyr:::mutate( blabel=gsub("-df03", "", blabel))
+  }
+  plotDF %>%
     ggplot( aes(alpha, value/deno, col=blabel) ) +
     geom_line() + geom_point() +
     xlab(expression(paste("Nominal"~alpha))) +
[…]

### 2017-11-10 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd

@@
 ```{bash, GWAS1-download, eval=FALSE}
+cd /n/irizarryfs01_backed_up/kkorthauer/FDR/DATA/GWAS/
 curl -O "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
@@
 Next we unzip the file, and delete the files we won't use (they provide results
-subsetted by ancestry and sex), keeping only the file with all ancestries and
-both sexes `BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt`. Note that it seems that
-the Boca-Leek paper actually used the 'EuropeanOnly' subset, but it's not clear
-why (to me), and there's no reason we need to do the same.
+subsetted by ancestry and sex), keeping only the file with European ancestry and
+both sexes `BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt`. Note that
+the Boca-Leek paper also used the 'EuropeanOnly' subset, likely because using 
+all ancestries would require that we assess the impact of population stratification
+and adjust for it if present. For simplicity, we follow suit and use the homogeneous 
+subset to avoid the impact of population stratification on our results.
 
@@
 rm BMI.SNPadjSMK.Men.AllAncestry.txt
-rm BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt
+rm BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt
 ```
@@
 ```{r, GWAS1-verify-contents}
-bmi <- fread(paste0(datdir, "BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt"),
+bmi <- fread(paste0(datdir, "BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt"),
              header=TRUE)
@@
                effect_size = effect,
-               ind_covariate = N) %>%
+               ind_covar_N = N,
+               ind_covar_AF = Freq_Allele1_HapMapCEU) %>%
         mutate(test_statistic = qnorm(exp(log(pval)-log(2)), lower.tail=FALSE) * sign(effect_size)) %>%
-        select(pval, SE, effect_size, ind_covariate, test_statistic) 
+        select(pval, SE, effect_size, ind_covar_N, ind_covar_AF, test_statistic) 
 ```
@@
 
-Here we look to see if the covariate does indeed look informative.
+Here we look to see if the covariates do indeed look informative. First we look at 
+the sample size covariate.
 
 ```{r, GWAS1-diag, fig.width=10, fig.height=3.5}
-strat_hist(bmi, pvalue="pval", covariate="ind_covariate", maxy=3)
-rank_scatter(bmi, pvalue="pval", covariate="ind_covariate")
+strat_hist(bmi, pvalue="pval", covariate="ind_covar_N", maxy=3)
+rank_scatter(bmi, pvalue="pval", covariate="ind_covar_N")
 ```
 
-**Still to-do: repeat for additional covariate (MAF).**
+Next we look at the allele frequency covariate.
+
+```{r, GWAS1-diag2, fig.width=10, fig.height=3.5}
+strat_hist(bmi, pvalue="pval", covariate="ind_covar_AF", maxy=3)
+rank_scatter(bmi, pvalue="pval", covariate="ind_covar_AF")
+```
 
@@
 
+#### Run buildBench for the sample size covariate and examine results
+
 Now, we're ready to construct the `SummarizedBenchmark` object, which will run
@@
 ```{r, GWAS1-sb, results="hide", message=FALSE}
-resfile <- paste0(resdir, "bmi_summarizedBenchmark_",
+resN <- paste0(resdir, "bmi_N_summarizedBenchmark_",
                   nrow(bmi), ".RData")
+
 duration <- NA
-if (!file.exists(resfile)){
+if (!file.exists(resN)){
   t1 <- proc.time()
-  sb <- bd %>% buildBench(data=bmi, ftCols = "ind_covariate",
+  sb <- bd %>% buildBench(data=(bmi %>% mutate(ind_covariate=ind_covar_N)), 
+                          ftCols = c("ind_covar_N"),
                           parallel=TRUE, BPPARAM=multicoreParam)
   metadata(sb)$data_download_link <- "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
-  save(sb, file=resfile)
+  save(sb, file=resN)
   duration <- round((proc.time()-t1)[3]/60,1)
 }else{
-  load(resfile)
+  load(resN)
 }
@@
 # plot nrejects by method overall and stratified by covariate
-rejections_scatter(sb)
+rejections_scatter(sb,
+                   supplementary=FALSE)
 
-rejection_scatter_bins(sb, covariate="ind_covariate", bins=4)
+rejection_scatter_bins(sb, covariate="ind_covar_N", bins=4,
+                   supplementary=FALSE)
 
-# upset plot (only plot IHW for the plotted alpha)
-rmv.ihw <- which(grepl("ihw", colnames(sb)) &
-                              colData(sb)$param.alpha != 0.05)
+# upset plot 
+plotFDRMethodsOverlap(sb, 
+                   alpha=0.05, nsets=ncol(sb),
+                   order.by="freq", decreasing=TRUE,
+                   supplementary=FALSE)
+```
 
-plotMethodsOverlap(sb[,-rmv.ihw], 
-                   alpha=0.05, nsets=ncol(sb)-length(rmv.ihw),
-                   order.by="freq", decreasing=TRUE)
+#### Run buildBench for the allele frequency covariate and examine results
 
-# upset plot excluding "unadjusted"
-plotMethodsOverlap(sb[,-c(1,rmv.ihw)], 
-                   alpha=0.05, nsets=ncol(sb)-length(rmv.ihw)-1,
-                   order.by="freq", decreasing=TRUE)
+Now, we'll repeat the previous steps for the other covariate (AF):
+
+```{r, GWAS1-sb2, results="hide", message=FALSE}
+bd <- initializeBenchDesign()
+resAF <- paste0(resdir, "bmi_AF_summarizedBenchmark_",
+                  nrow(bmi), ".RData")
+
+duration <- NA
+if (!file.exists(resAF)){
+  t1 <- proc.time()
+  sb <- bd %>% buildBench(data=(bmi %>% mutate(ind_covariate=ind_covar_AF) %>%
+                                        filter(!is.na(ind_covariate))), 
+                          ftCols = c("ind_covar_AF"),
+                          parallel=TRUE, BPPARAM=multicoreParam)
+  metadata(sb)$data_download_link <- "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
+  save(sb, file=resAF)
+  duration <- round((proc.time()-t1)[3]/60,1)
+}else{
+  load(resAF)
+}
+```
+
+```{r, echo=FALSE}
+if(!is.na(duration)){
+  message("This step took ", duration, " minutes for ", nrow(bmi), " SNPs using ",
+        cores, " cores.")
+}
+```
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, GWAS1-metrics2}
+# rename assay to qvalue
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, GWAS1-plot2, results="hide", width=15, height=15}
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter(sb,
+                   supplementary=FALSE)
+
+rejection_scatter_bins(sb, covariate="ind_covar_AF", bins=4,
+                   supplementary=FALSE)
+
+# upset plot 
+plotFDRMethodsOverlap(sb, 
+                   alpha=0.05, nsets=ncol(sb),
+                   order.by="freq", decreasing=TRUE,
+                   supplementary=FALSE)
 ```

### 2017-11-10 datasets/R/du_psim.R
Add more null sim settings to psim

@@ -71,8 +71,10 @@ du_psim <- function(m, pi0, tstat, tstat_dist, null_dist, icovariate, seed = NUL
     
     ## generate set of null (0) and alternative test statistics
     ts <- rep(0, m)
-    ts[alts] <- tstat(length(alts))
-
+    if (length(alts) > 0) {
+        ts[alts] <- tstat(length(alts))
+    }
+    
     ## determine (true) SD of sampling dist used to perturb each stat
     SE <- tstat_dist(ts, se = TRUE)
 

### 2017-11-10 datasets/R/funcs_tstat.R

@@ -109,18 +109,21 @@ rchisq_perturber <- function(df) {
 ## ##############################################################################
 
 ## function factory: two-sided normal p-value calculator
+## @param s standard deviation of null zero-centered normal distribution
 rnorm_2pvaluer <- function(s) {
     function(x) { 2 * (1 - pnorm(abs(x), 0, s)) }
 }
 
 
 ## function factory: two-sided t p-value calculator
+## @param df degrees of freedom of null zero-centered t-distribution
 rt_2pvaluer <- function(df) {
     function(x) { 2 * (1 - pt(abs(x), df)) }
 }
 
 
 ## function factory: chi-sq p-value calculator
+## @param df degrees of freedom of null zero-centered t-distribution
 rchisq_pvaluer <- function(df) {
     function(x) { 1 - pchisq(x, df) }
 }

### 2017-11-10 datasets/simulations/psim-core-settings.R

@@ -7,14 +7,14 @@
 #' @param sbase default simulation settings to be modified.
 #' @param vparam variable parameter; should be one of: "pi0",
 #'        "esize_fixed", "esize_random_ua", "esize_random_shift",
-#'        "altnoise".
+#'        "altnoise", "allnull".
 #' @param icparam independent/informative covariate parameter; should be
 #'        one of: "uniform", "bl".
 #'
 #' @description
 #' The following simulation groups are defined:
 #' - `vparam`: variable parameter
-#'     - `pi0`: for fixed effect size, varying null prop
+#'     - `pi0`: for fixed effect size, varying null proportion.
 #'     - `esize_fixed`: varying effect size taken to be the same fixed
 #'                      value across all alternative tests.
 #'     - `esize_random_ua`: varying effect size sampled from a (mostly)
@@ -26,6 +26,8 @@
 #'     - `altnoise`: varying sampling distributions for perturbing the
 #'                   test statistics (both null and alternative)
 #'                   (shifted normal, non-central t, non-central chi-sq).
+#'     - `allnull`: similar to `altnoise` setting, but with pi0 = 1, i.e. all test
+#'                  statistics are simulated under the null distribution.
 #' - `icparam`: independent covariate structure
 #'     - `uniform`: covariate sampled from uniform (0, 1) interval AND
 #'                  no association between covariate and distribution of
@@ -39,7 +41,8 @@
 psim_settings <- function(sbase, vparam, icparam) {
 
     ## check that variable parameter is one of specified set
-    vp <- c("pi0", "esize_fixed", "esize_random_ua", "esize_random_shift", "altnoise")
+    vp <- c("pi0", "esize_fixed", "esize_random_ua", "esize_random_shift",
+            "altnoise", "allnull")
     stopifnot(vparam %in% vp)
     
     ## check that indep/inform covariate parameter is one of specified set
@@ -49,6 +52,9 @@ psim_settings <- function(sbase, vparam, icparam) {
     ## filter out unsupported setting pairs
     if (vparam == "pi0" && icparam == "bl") {
         stop("cant run 'pi0' simulations w/ 'bl' format indep covariate")
+    }
+    if (vparam == "allnull" && icparam == "bl") {
+        stop("cant run 'allnull' simulations w/ 'bl' format indep covariate")
     }
     
     ## ##########################################################################
@@ -149,6 +155,27 @@ psim_settings <- function(sbase, vparam, icparam) {
                          "altnoise_chisq_df4_shift3sq" =
[…]

### 2017-11-10 datasets/simulations/psim-submit.R

@@ -14,7 +14,8 @@ params <- list(c("pi0", "uniform"),
                c("esize_random_shift", "uniform"),
                c("esize_random_shift", "bl"),
                c("altnoise", "uniform"),
-               c("altnoise", "bl"))
+               c("altnoise", "bl"),
+               c("allnull", "uniform"))
 
 for (ip in params) {
     cmd <- paste("sbatch",

### 2017-11-10 datasets/microbiome/cdi_schubert.Rmd

@@
   reads <- rowMeans(otu)
+
+  ## TODO: median read depth!
 
@@
 # Read and clean cdi_schubert
+## TODO: less filtering! Keep as many hypotheses open as possible.
 otu_path = paste0(wd, "cdi_schubert_results/RDP/cdi_schubert.otu_table.100.denovo.rdp_assigned")
@@
 var <- "logubiquity"
-g <- ggplot(goodrichdat, aes(x = get(var), y = pval)) + geom_point()
+g <- ggplot(goodrichdat, aes(x = get(var), y = logpval)) + geom_point()
 g
@@
 ```{r}
-g <- ggplot(goodrichdat, aes(x = get(covcol), y = pval)) + geom_point()
+g <- ggplot(goodrichdat, aes(x = get(covcol), y = logpval)) + geom_point()
 g

### 2017-11-10 datasets/microbiome/microbiome-SummarizedBenchmark.Rmd

---
title: "Comparison of FDR methods on GWAS with SummarizedBenchmark"
author: "Keegan Korthauer"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "/n/irizarryfs01_backed_up/kkorthauer/FDR")
```

### Summary

In the previous document `cdi_schubert.Rmd`, we explored two microbiome datasets for FDR
benchmarking. Here we will make use of the new `SummarizedBenchmark` package to
perform the benchmarking in a standardized way.

### Workspace setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
# project directory & data/results folders
setwd("/Users/claire/github/benchmark-fdr/datasets/microbiome/")
datdir <- "/Users/claire/github/benchmark-fdr/datasets/microbiome/DATA/"
resdir <- "/Users/claire/github/benchmark-fdr/datasets/microbiome/RESULTS/"

# wrangling & plotting tools
library(data.table)
#library(readxl)
library(dplyr)
library(ggplot2)
library(magrittr)
library(R.utils)
library(cowplot)

# benchmark methods
library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)
library(FDRreg)

# comparison tools/functions
library(SummarizedBenchmark)  
sourceDirectory("/Users/claire/github/benchmark-fdr/datasets/R/")

## set up parallel backend 
library(BiocParallel)
cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
multicoreParam <- MulticoreParam(workers = cores)
```

Let's just keep going and hopefully it's okay...

### CDI Schubert - diarrhea dataset

Here we download and analyze the CDI Schubert diarrhea (C. diff and non-C.diff diarrhea) dataset. 
We'll download the processed OTU tables from Zenodo and unzip them in the 
`benchmark-fdr/datasets/microbiome/DATA` folder.

#### Data download

```{bash, schubert-download, eval=FALSE}
cd /Users/claire/github/benchmark-fdr/datasets/microbiome/DATA/
curl -O "https://zenodo.org/record/840333/files/cdi_schubert_results.tar.gz"
tar -xzvf cdi_schubert_results.tar.gz
cd ..
```

Next, we'll read in the unzipped OTU table and metadata files into R.

We'll need to manually define which disease labels are of interest and what metadata
column contains the sample IDs (i.e. the IDs in the OTU table).

```{r, schubert-readdata}
otu_path = paste0(datdir, "cdi_schubert_results/RDP/cdi_schubert.otu_table.100.denovo.rdp_assigned")
meta_path = paste0(datdir, "cdi_schubert_results/cdi_schubert.metadata.txt")

# DiseaseState labels to keep
labels <- c("H", "nonCDI", "CDI")

# Metadata column that has sample IDs
col_label <- 'sample_id'

# load OTU table and metadata
otu <- read.table(otu_path)
meta <- read.csv(meta_path, sep='\t')

# Keep only samples with the right DiseaseState metadata
meta <- meta %>% filter(DiseaseState %in% labels)

# Keep only samples with both metadata and 16S data
keep_samples <- intersect(colnames(otu), meta[, 1])
otu <- otu[, keep_samples]
meta <- meta %>% filter(get(col_label) %in% keep_samples)

```

Since we'll be using OTU-wise covariates, we shouldn't need to perform any
filtering/cleaning of the OTUs, apart from removing any that are all zeros. 
(This may happen after removing shallow samples, I think.)
However, let's make sure we get rid of any samples with too few reads. 
We define the minimum number of reads per sample in `sample_reads`.

After we've removed any shallow samples, we'll convert the OTU table to relative
abundances. I'll add a pseudo-count of 1e-6 to any zero entries, to avoid
problems with taking logs.

```{r, schubert-cleandata}
remove_shallow_smpls <- function(df, n_reads) {
  # Removes samples with fewer than n_reads from dataframe df.
  # df has OTUs in rows and samples in columns
  return(df[, colSums(df) > n_reads])
  }

remove_shallow_otus <- function(df, n_reads){
  return(df[rowSums(df) > n_reads, ])
  }

## Remove OTUs with fewere than 10 reads
#otu <- remove_shallow_otus(otu, 10)

# Remove samples with fewer than sample_Reads reads
sample_reads <- 100
otu <- remove_shallow_smpls(otu, sample_reads)

# Update metadata with new samples
meta <- meta %>% filter(get(col_label) %in% colnames(otu))

# Remove empty OTUs
otu <- otu[rowSums(otu) > 0, ]

# Convert to relative abundance
abun_otu <- t(t(otu) / rowSums(t(otu)))

# Add pseudo counts
#minabun <- min(abun_otu[abun_otu > 0]) # Use this to figure out what pseudo-count value to add
zeroabun <- 1e-6
abun_otu <- abun_otu + zeroabun

```

Next, we need to calculate the pvalues, effect size, and standard error for each OTU.
Here, we'll compare diarrhea vs. healthy. Diarrhea will include both CDI and nonCDI
patients. We'll put these results into a dataframe, and label the columns with the 
standardized names for downstream use (`pval`, `SE`, `effect_size`, `test_statistic`).
The test statistic is the one returned by `wilcox.test()`.

Note that the effect here is calculated as logfold change of mean abundance in controls
relative to cases (i.e. `log(mean_abun[controls]/mean_abun[cases])`)

While we're at it, we'll also calculate the mean abundance and ubiquity (detection rate)
of each OTU. Later, we can assign their values to a new column called `ind_covariate` 
for use in downstream steps.

```{r, schubert-pvals}
resfile <- paste0(resdir, "schubert_results_",
                  nrow(abun_otu), "_OTUs.RData")
if (!file.exists(resfile)){
  # Get case and control indices
  case_idx <- meta$DiseaseState %in% c("CDI", "nonCDI")
  ctrl_idx <- meta$DiseaseState %in% c("H")

  # Calculate pvalues, effects, and stderr
  pvals <- c()
  teststats <- c()
  ses <- c()
  effs <- c()
  mean_abuns <- c()
  mean_abuns_present <- c()
  ubis <- c()

  casedf <- abun_otu[, case_idx]
  ctrldf <- abun_otu[, ctrl_idx]
  for (o in rownames(abun_otu)) {
    # wilcoxon p value  
    w <- wilcox.test(casedf[o, ], 
                     ctrldf[o, ])
    p <- w$p.value
    teststat <- w$statistic

    pvals <- c(pvals, p)
    teststats <- c(teststats, teststat)

    # standard error of the OTU abundance, across all samples
    ses <- c(ses, 
             sd(abun_otu[o, ])/sqrt(length(abun_otu[o, ]))
             )

    # mean OTU abundance across all samples (after removing pseudo-count)
    mean_abuns <- c(mean_abuns, 
                    mean(abun_otu[o, ] - zeroabun)
                    )

    # mean OTU abundance across only samples with the OTU present
    mean_abuns_present <- c(mean_abuns_present, 
                            sum(abun_otu[o, ] - zeroabun) / sum(abun_otu[o, ] > zeroabun)
                            )

    # ubiquity of OTU across all samples
    ubis <- c(ubis, 
              sum(abun_otu[o, ] > zeroabun) / length(abun_otu[o, ])
              )

    # effect (logfold difference)
    effs <- c(effs, 
              log(mean(abun_otu[o, ctrl_idx])/mean(abun_otu[o, case_idx]))
              )

  }

  res <- data.frame(otu = rownames(abun_otu), 
                    pval = pvals, test_statistic = teststats,
                    SE = ses, effect_size = effs, 
                    mean_abun = mean_abuns, mean_abun_present = mean_abuns_present,
                    ubiquity = ubis)
  save(res, file=resfile)

}else{
  load(resfile)
}
head(res)
```

Finally, let's try to add phylogeny as covariates. Here we'll have columns for each separate taxonomic level.

```{r, schubert-addphylo}
res <- res %>% separate(otu, 
                        c("kingdom", "phylum", "class", "order", "family", "genus", "species", "denovo"), 
                        sep=";", remove = FALSE)

```

#### Check Covariate Diagnostics

Here we look to see if the covariates do indeed look informative.

##### Ubiquity

```{r, schubert-ubi, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=20)
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

##### Mean abundance (across non-zero samples)

```{r, schubert-abun, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=20)
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

##### Phylogeny?!

Let's look at phylum-level stratification first. A priori, I *might* expect
Proteobacteria to be enriched for low p-values? But I don't know if that's
super legit, and Eric doesn't seem to think that phylogeny will be informative at all...

```{r, schubert-phylo, fig.width=10, fig.height=6}
#strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=20)
#rank_scatter(res, pvalue="pval", covariate="ubiquity")
ggplot(res, aes(x=pval)) + geom_histogram() + facet_wrap(~phylum, scales = "free")
```

Let's use `ubiquity` as our `ind_covariate`

```{r, schubert-labelcovariate}
res <- res %>% mutate(ind_covariate = ubiquity)
```

#### Set up BenchDesign object

First, we'll create an object of `BenchDesign` class to hold the data and 
add the benchmark methods to the `BenchDesign` object.

```{r, schubert-benchdesign}
bd <- initializeBenchDesign()
```

Now, we're ready to construct the `SummarizedBenchmark` object, which will run
the functions specified in each method (these are actually sourced in from the
helper scripts).

```{r, GWAS1-sb, results="hide", message=FALSE, eval=FALSE}
resfile <- paste0(resdir, "schubert_summarizedBenchmark_",
                  nrow(res), ".RData")
duration <- NA
if (!file.exists(resfile)){
  t1 <- proc.time()
  sb <- bd %>% buildBench(data=res, ftCols = "ind_covariate")
                          #parallel=TRUE, BPPARAM=multicoreParam)
  metadata(sb)$data_download_link <- "https://zenodo.org/record/840333/files/cdi_schubert_results.tar.gz"
  save(sb, file=resfile)
  duration <- round((proc.time()-t1)[3]/60,1)
}else{
  load(resfile)
}
```

```{r, echo=FALSE, eval=FALSE}
if(!is.na(duration)){
  message("This step took ", duration, " minutes for ", nrow(res), " OTUs using ",
        cores, " cores.")
}
```

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, schubert-metrics, eval=FALSE}
# rename assay to qvalue
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, schubert-plot, results="hide", width=15, height=15, eval=FALSE}
# plot nrejects by method overall and stratified by covariate
rejections_scatter(sb)

rejection_scatter_bins(sb, covariate="ind_covariate", bins=4)

# upset plot (only plot IHW for the plotted alpha)
rmv.ihw <- which(grepl("ihw", colnames(sb)) &
                              colData(sb)$param.alpha != 0.05)

plotMethodsOverlap(sb[,-rmv.ihw], 
                   alpha=0.05, nsets=ncol(sb)-length(rmv.ihw),
                   order.by="freq", decreasing=TRUE)

# upset plot excluding "unadjusted"
plotMethodsOverlap(sb[,-c(1,rmv.ihw)], 
                   alpha=0.05, nsets=ncol(sb)-length(rmv.ihw)-1,
                   order.by="freq", decreasing=TRUE)
```

And we'll clean up the workspace before moving on to the next dataset.

```{r, schubert-cleanup}
rm(res)
rm(bd)
rm(sb)
rm(pf)
```

### OB Goodrich - obesity dataset

Let's repeat these analyses for the OB Goodrich dataset.

Here we download and analyze the CDI Schubert diarrhea (C. diff and non-C.diff diarrhea) dataset. 
We'll download the processed OTU tables from Zenodo and unzip them in the 
`benchmark-fdr/datasets/microbiome/DATA` folder.

#### Data download

```{bash, goodrich-download, eval=FALSE}
cd /Users/claire/github/benchmark-fdr/datasets/microbiome/DATA/
curl -O "https://zenodo.org/record/840333/files/ob_goodrich_results.tar.gz"
tar -xzvf ob_goodrich_results.tar.gz
cd ..
```

Next, we'll read in the unzipped OTU table and metadata files into R.

We'll need to manually define which disease labels are of interest and what metadata
column contains the sample IDs (i.e. the IDs in the OTU table).

```{r, goodrich-readdata}
otu_path = paste0(datdir, "ob_goodrich_results/RDP/ob_goodrich.otu_table.100.denovo.rdp_assigned")
meta_path = paste0(datdir, "ob_goodrich_results/ob_goodrich.metadata.txt")

# DiseaseState labels to keep
labels <- c("H", "OB")

# Metadata column that has sample IDs
col_label <- "X"

# load OTU table and metadata
otu <- read.table(otu_path)
meta <- read.csv(meta_path, sep='\t')

# Keep only samples with the right DiseaseState metadata
meta <- meta %>% filter(DiseaseState %in% labels)

# Keep only samples with both metadata and 16S data
keep_samples <- intersect(colnames(otu), meta[, 1])
otu <- otu[, keep_samples]
meta <- meta %>% filter(get(col_label) %in% keep_samples)

```

Since we'll be using OTU-wise covariates, we shouldn't need to perform any
filtering/cleaning of the OTUs, apart from removing any that are all zeros. 
(This may happen after removing shallow samples, I think.)
However, let's make sure we get rid of any samples with too few reads. 
We define the minimum number of reads per sample in `sample_reads`.

After we've removed any shallow samples, we'll convert the OTU table to relative
abundances. I'll add a pseudo-count of 1e-6 to any zero entries, to avoid
problems with taking logs.

```{r, goodrich-cleandata}
remove_shallow_smpls <- function(df, n_reads) {
  # Removes samples with fewer than n_reads from dataframe df.
  # df has OTUs in rows and samples in columns
  return(df[, colSums(df) > n_reads])
  }

# Remove samples with fewer than sample_Reads reads
sample_reads <- 100
otu <- remove_shallow_smpls(otu, sample_reads)

# Update metadata with new samples
meta <- meta %>% filter(get(col_label) %in% colnames(otu))

# Remove empty OTUs
otu <- otu[rowSums(otu) > 0, ]

# Convert to relative abundance
abun_otu <- t(t(otu) / rowSums(t(otu)))

# Add pseudo counts
#minabun <- min(abun_otu[abun_otu > 0]) # Use this to figure out what pseudo-count value to add
zeroabun <- 1e-6
abun_otu <- abun_otu + zeroabun

```

Next, we need to calculate the pvalues, effect size, and standard error for each OTU.
Here, we'll compare lean vs. obese. We'll put these results into a dataframe, and label 
the columns with the standardized names for downstream use 
(`pval`, `SE`, `effect_size`, `test_statistic`). The test statistic is the one returned 
by `wilcox.test()`.

Note that the effect here is calculated as logfold change of mean abundance in controls
relative to cases (i.e. `log(mean_abun[controls]/mean_abun[cases])`)

While we're at it, we'll also calculate the mean abundance and ubiquity (detection rate)
of each OTU. Later, we can assign their values to a new column called `ind_covariate` 
for use in downstream steps.

Note that OB Goodrich has ~70,000 OTUs, so this is going to take a while. There are probably
faster ways to write this, but I don't know them.

```{r}
dim(abun_otu)
```

```{r, goodrich-pvals}
resfile <- paste0(resdir, "goodrich_results_",
                  nrow(abun_otu), "_OTUs.RData")
if (!file.exists(resfile)){
  # Get case and control indices
  case_idx <- meta$DiseaseState %in% c("OB")
  ctrl_idx <- meta$DiseaseState %in% c("H")

  # Calculate pvalues, effects, and stderr
  pvals <- c()
  teststats <- c()
  ses <- c()
  effs <- c()
  mean_abuns <- c()
  mean_abuns_present <- c()
  ubis <- c()

  casedf <- abun_otu[, case_idx]
  ctrldf <- abun_otu[, ctrl_idx]
  for (o in rownames(abun_otu)) {
    # wilcoxon p value  
    w <- wilcox.test(casedf[o, ], 
                     ctrldf[o, ])
    p <- w$p.value
    teststat <- w$statistic

    pvals <- c(pvals, p)
    teststats <- c(teststats, teststat)

    # standard error of the OTU abundance, across all samples
    ses <- c(ses, 
             sd(abun_otu[o, ])/sqrt(length(abun_otu[o, ]))
             )

    # mean OTU abundance across all samples (after removing pseudo-count)
    mean_abuns <- c(mean_abuns, 
                    mean(abun_otu[o, ] - zeroabun)
                    )

    # mean OTU abundance across only samples with the OTU present
    mean_abuns_present <- c(mean_abuns_present, 
                            sum(abun_otu[o, ] - zeroabun) / sum(abun_otu[o, ] > zeroabun)
                            )

    # ubiquity of OTU across all samples
    ubis <- c(ubis, 
              sum(abun_otu[o, ] > zeroabun) / length(abun_otu[o, ])
              )

    # effect (logfold difference)
    effs <- c(effs, 
              log(mean(abun_otu[o, ctrl_idx])/mean(abun_otu[o, case_idx]))
              )

  }

  res <- data.frame(otu = rownames(abun_otu), 
                    pval = pvals, test_statistic = teststats,
                    SE = ses, effect_size = effs, 
                    mean_abun = mean_abuns, mean_abun_present = mean_abuns_present,
                    ubiquity = ubis)
  save(res, file=resfile)

}else{
  load(resfile)
}
head(res)
```

Finally, let's try to add phylogeny as covariates. Here we'll have columns for each separate taxonomic level.

```{r, goodrich-addphylo}
res <- res %>% separate(otu, 
                        c("kingdom", "phylum", "class", "order", "family", "genus", "species", "denovo"), 
                        sep=";", remove = FALSE)

```

#### Check Covariate Diagnostics

Here we look to see if the covariates do indeed look informative.

##### Ubiquity

```{r, goodrich-ubi, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=5, binwidth=0.05, numQ=4)
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

##### Mean abundance (across non-zero samples)

```{r, goodrich-abun, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=5, binwidth=0.05, numQ=4)
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

##### Phylogeny?!

Let's look at phylum-level stratification first. A priori, I *might* expect
Proteobacteria to be enriched for low p-values? But I don't know if that's
super legit, and Eric doesn't seem to think that phylogeny will be informative at all...

```{r, goodrich-phylo, fig.width=10, fig.height=6}
#strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=20)
#rank_scatter(res, pvalue="pval", covariate="ubiquity")
ggplot(res, aes(x=pval)) + geom_histogram() + facet_wrap(~phylum, scales = "free")
```

Let's use `ubiquity` as our `ind_covariate`

```{r, goodrich-labelcovariate}
res <- res %>% mutate(ind_covariate = ubiquity)
```

### 2017-11-11 datasets/GSEA/gsea.Rmd
Started exploring sizes of sets

@@
 ```{r}
+plotFDRMethodsOverlap( sGSEA, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+```
 
-plotFDRMethodsOverlap( sGSEA, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+# Gene set size of overlaps
+
+```{r}
+
+getOverlapIndexes <- function( labs=c("lfdr", "ihw-a10"), alpha=0.1 ){
+  rowSums( assays(sGSEA)[["qvalue"]][,labs] < alpha) == length(labs) &
+  !(rowSums( assays(sGSEA)[["qvalue"]][,c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf")] < alpha, na.rm=TRUE ) > (length(labs)))
+}
+
+set1 <- getOverlapIndexes( c("lfdr", "ihw-a10"), alpha=0.1 )
+set2 <- getOverlapIndexes( c("lfdr","ihw-a10","bl-df03","qvalue","bh"), alpha=0.1 )
+set3 <- getOverlapIndexes( c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf"), alpha=0.1 )
+
+table( set1 )
+table( set2 )
 
 ```
+
+# NULL comparisons
+
+```{r, eval=FALSE }
+nms <- names( genes )
+xx <- seq_len(length(nms))
+gns <- 1*( xx %in% sample( xx, 1000, replace=FALSE ) )
+names( gns ) <- nms
+pwf <- nullp( gns, bias.data=geneLength )
+goRes <- goseq( pwf, gene2cat=goSets )
+res <- goRes %>%
+  dplyr:::filter( numDEInCat > 0 ) %>%
+  dplyr:::select( c("over_represented_pvalue", "numInCat") ) %>%
+  dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
+pr <- designBen %>% buildBench( res, ftCols="ind_covariate")
+assayNames(pr) <- "qvalue"
+pr <- addDefaultMetrics(pr)
+
+```

### 2017-12-07 datasets/simulations/analysis-03-plot-tsim-summaries.Rmd
Plot results for 'BL style' sims w/ 20% alt tests

@@
 <!-- ######################################################################## -->
+### Manual Covariate (Step - Less Informative)
+
+```{r}
+f <- "data-tsim/M100/esize_fixed-bl-step-less/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")),
+           param >= 10) %>%
+    mutate(param = paste("effect size =", sprintf("%.1f", param / 10)))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Two-Sample T-Tests w/ Varying Effect Size, Manual Covariate (step - less informative) (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
+### Manual Covariate (Step - More Informative)
+
+```{r}
+f <- "data-tsim/M100/esize_fixed-bl-step-more/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")),
+           param >= 10) %>%
+    mutate(param = paste("effect size =", sprintf("%.1f", param / 10)))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Two-Sample T-Tests w/ Varying Effect Size, Manual Covariate (step - more informative) (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
+### Manual Covariate (Cubic)
+
+```{r}
+f <- "data-tsim/M100/esize_fixed-bl-cubic/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")),
+           param >= 10) %>%
+    mutate(param = paste("effect size =", sprintf("%.1f", param / 10)))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Two-Sample T-Tests w/ Varying Effect Size, Manual Covariate (cubic) (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
 ## Varying Null Proportion
@@
 <!-- ######################################################################## -->
+### Manual Covariate (Step - Less Informative)
+
+```{r}
+f <- "data-tsim/M100/n-bl-step-less/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove rejections: scaling is off from rest (also dup'd w/ rejprop)
+## remove FWER: not informative as a boxplot
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER"))) %>%
+    mutate(param = factor(paste0("n (per group) = ", param),
+                          levels = paste0("n (per group) = ", c(3, 6, 20))))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Two-Sample T-Tests w/ Varying Sample Sizes, Manual Covariate (step - less informative) (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
+### Manual Covariate (Step - More Informative)
+
+```{r}
+f <- "data-tsim/M100/n-bl-step-more/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove rejections: scaling is off from rest (also dup'd w/ rejprop)
+## remove FWER: not informative as a boxplot
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER"))) %>%
+    mutate(param = factor(paste0("n (per group) = ", param),
+                          levels = paste0("n (per group) = ", c(3, 6, 20))))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Two-Sample T-Tests w/ Varying Sample Sizes, Manual Covariate (step - more informative) (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
+### Manual Covariate (Cubic)
+
+```{r}
+f <- "data-tsim/M100/n-bl-cubic/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove rejections: scaling is off from rest (also dup'd w/ rejprop)
+## remove FWER: not informative as a boxplot
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER"))) %>%
+    mutate(param = factor(paste0("n (per group) = ", param),
+                          levels = paste0("n (per group) = ", c(3, 6, 20))))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Two-Sample T-Tests w/ Varying Sample Sizes, Manual Covariate (cubic) (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
 ## Randomly-Sampled Effect Size
@@
 ```
+
+<!-- ######################################################################## -->
+### Manual Covariate (Step - Less Informative)
+
+```{r}
+f <- "data-tsim/M100/esize_random-bl-step-less/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*)-.+", "\\1", param),
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Two-Sample T-Tests w/ Varying Randomly Sampled Effect, Manual Covariate (step - less informative) (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
+### Manual Covariate (Step - More Informative)
+
+```{r}
+f <- "data-tsim/M100/esize_random-bl-step-more/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*)-.+", "\\1", param),
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Two-Sample T-Tests w/ Varying Randomly Sampled Effect, Manual Covariate (step - more informative) (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
+### Manual Covariate (Cubic)
+
+```{r}
+f <- "data-tsim/M100/esize_random-bl-cubic/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*)-.+", "\\1", param),
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Two-Sample T-Tests w/ Varying Randomly Sampled Effect, Manual Covariate (cubic) (100 replications)",
+         tsb, cutoffs)
+```

### 2017-12-08 datasets/GSEA/gsea.Rmd
Start adding mouse datasets

--- 2017-10-30 datasets/GSEA/gsea.Rmd
@@
 gnType <-  genes(EnsDb.Hsapiens.v86, column="gene_biotype")
+
 dsdObject <- dsdObject[rownames(dsdObject) %in% gnType[gnType$gene_biotype == "protein_coding"]$gene_id,]
@@
 ```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
-
 library(ggplot2)
@@
 res$pval <- pmin( res$pval, 1 )
-head(res)
-sGSEA <- designBen %>% buildBench( res, ftCols="ind_covariate")
-elementMetadata( rowData(sGSEA) )
+sGSEA <- designBen %>% buildBench( res )
 assayNames(sGSEA) <- "qvalue"
@@
 stopifnot( all(assays(sGSEA)[["qvalue"]][,"unadjusted"] == pmin(goRes$over_represented_pvalue, 1) ) )
+rowData(sGSEA)$numInCat <- goRes$numInCat
 
-sGSEA <- estimatePerformanceMetrics(sGSEA, addColData=TRUE)
-rejection_scatter_bins( sGSEA, "ind_covariate", threshold=15, bins=4, as_fraction=TRUE )
+rejection_scatter_bins( sGSEA, "numInCat", threshold=15, bins=4, as_fraction=TRUE )
+```
+
+Plotting the methods overlap
+
+```{r}
+
+sGSEASub <- sGSEA[,!apply( is.na( assays(sGSEA)[["qvalue"]] ), 2, all )]
+sGSEASub <- sGSEASub[,-which(grepl("ihw", colData( sGSEASub )$blabel) & colData( sGSEASub )$param.alpha != 0.1)]
+sGSEASub <- sGSEASub[,-which( grepl("bl", colData( sGSEASub)$blabel) & 
+                   colData(sGSEASub)$param.smooth.df != 2L )]
+sGSEASub <- sGSEASub[,!colData(sGSEASub)$blabel %in% "unadjusted"]
+
+plotMethodsOverlap(sGSEASub, order.by = "freq", nsets=10)
 
 ```
+
+# Same analysis but using the HSC-MPP dataset from Nina
+
+```{r}
+dsdMouse <- "dsdHSCMPP.RData"
+if( !file.exists( dsdMouse ) ){
+  download.file("https://www.dropbox.com/s/spo1pifum5k72rn/dseHSCMPP.RData?dl=1", destfile="dsdHSCMPP.RData")
+}
+load(dsdMouse)
+dseHSCMPP <- DESeq( dseHSCMPP)
+
+res <- results( dseHSCMPP, contrast=c("conditions", "HSC", "MPP1") )
+genes <- as.numeric( res$padj < 0.1 )
+names( genes ) <- rownames( res )
+genes[is.na( genes )] <- 0
+
+### preparing gene length info and goSets ###
+library("EnsDb.Mmusculus.v75")
+
+pwf <- nullp(genes,  )
+goRes <- goseq( pwf, gene2cat=goSets )
+
+```

### 2017-12-08 datasets/GSEA/gsea.Rmd

--- 2017-11-11 datasets/GSEA/gsea.Rmd
@@
 
+Plotting the methods overlap
+
+```{r}
+
+sGSEASub <- sGSEA[,!apply( is.na( assays(sGSEA)[["qvalue"]] ), 2, all )]
+sGSEASub <- sGSEASub[,-which(grepl("ihw", colData( sGSEASub )$blabel) & colData( sGSEASub )$param.alpha != 0.1)]
+sGSEASub <- sGSEASub[,-which( grepl("bl", colData( sGSEASub)$blabel) & 
+                   colData(sGSEASub)$param.smooth.df != 2L )]
+sGSEASub <- sGSEASub[,!colData(sGSEASub)$blabel %in% "unadjusted"]
+
+plotMethodsOverlap(sGSEASub, order.by = "freq", nsets=10)
+
+```
+
 # NULL comparisons
@@
 ```
+
+# Same analysis but using the HSC-MPP dataset from Nina
+
+```{r}
+dsdMouse <- "dsdHSCMPP.RData"
+if( !file.exists( dsdMouse ) ){
+  download.file("https://www.dropbox.com/s/spo1pifum5k72rn/dseHSCMPP.RData?dl=1", destfile="dsdHSCMPP.RData")
+}
+load(dsdMouse)
+dseHSCMPP <- DESeq( dseHSCMPP)
+
+res <- results( dseHSCMPP, contrast=c("conditions", "HSC", "MPP1") )
+genes <- as.numeric( res$padj < 0.1 )
+names( genes ) <- rownames( res )
+genes[is.na( genes )] <- 0
+
+### preparing gene length info and goSets ###
+library("EnsDb.Mmusculus.v75")
+
+pwf <- nullp(genes,  )
+goRes <- goseq( pwf, gene2cat=goSets )
+
+```

### 2017-12-08 datasets/simulations/analysis-01-plot.Rmd
Move t-test sim summary parsing from 01 to 03

--- 2017-10-20 datasets/simulations/analysis-01-plot.Rmd
@@
 ```{r, load-libraries, message=FALSE}
-library("dplyr")
-library("ggplot2")
-library("viridis")
 library("data.table")
 library("SummarizedBenchmark")
+library("tidyverse")
+library("magrittr")
+library("viridis")
+library("cowplot")
 ```
@@
 ## list of results from 100 replication experiments
-dl <- list.dirs("data-tsim/M1", recursive=FALSE, full.names=TRUE)
+dl <- c(list.dirs("data-tsim/M1", recursive=FALSE, full.names=TRUE),
+        list.dirs("data-psim/M1", recursive=FALSE, full.names=TRUE))
 
@@
 ## list of results from 100 replication experiments
-dl <- list.dirs("data-tsim/M100", recursive=FALSE, full.names=TRUE)
+dl <- c(list.dirs("data-tsim/M100", recursive=FALSE, full.names=TRUE),
+        list.dirs("data-psim/M100", recursive=FALSE, full.names=TRUE))
 
 for (dd in dl) {
-    fl <- list.files(dd, pattern=".rds", full.names=TRUE)
+    fl <- list.files(dd, pattern="M100\\.rds", full.names=TRUE)
 

### 2017-12-08 datasets/simulations/analysis-02-try-adapt.Rmd

@@
 
-```{r, fig.width = }
+```{r, fig.width = 5, fig.height = 6}
 vec_tib %>%

### 2017-12-08 datasets/GSEA/gsea.Rmd
local changes

--- 2017-11-11 datasets/GSEA/gsea.Rmd
@@
 ```{r plottingResults}
-rejections_scatter( sGSEA, as_fraction=TRUE, supplementary=FALSE)
+rejections_scatter( sGSEA, as_fraction=FALSE, supplementary=FALSE)
 
 sGSEA <- estimatePerformanceMetrics(sGSEA, addColData=TRUE)
-rejection_scatter_bins( sGSEA, "ind_covariate", threshold=15, bins=4, as_fraction=TRUE, supplementary=FALSE)
+
+estimatePerformanceMetrics( sGSEA, alpha=0.1, tidy=TRUE ) %>%
+  filter( blabel %in% c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf") ) %>%
+  select( blabel, value )
+
+rejection_scatter_bins( sGSEA, "ind_covariate", threshold=15, bins=4, as_fraction=FALSE, supplementary=FALSE)
 ```
@@
 
-table( set1 )
-table( set2 )
+sizesDf <- data.frame( 
+  size=c( rowData( sGSEA )$ind_covariate[which(set1)], 
+          rowData( sGSEA )$ind_covariate[which(set2)], 
+          rowData( sGSEA )$ind_covariate[which(set3)] ),
+  label=rep(c("set1", "set2", "set3"), c(sum(set1), sum(set2), sum(set3)) ) )
+
+ggplot( sizesDf, aes( label, log2(size) )) +
+  geom_boxplot()
+
+dim(goRes)
+
+goRes[set1,]
 
@@
 ```{r, eval=FALSE }
+
 nms <- names( genes )
@@
 assayNames(pr) <- "qvalue"
-pr <- addDefaultMetrics(pr)
+pr <- addDefaultMetrics(pr, metrics="rejections")
+
+estimatePerformanceMetrics( pr, alpha=0.1, tidy=TRUE ) %>%
+  filter( blabel %in% c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf") ) %>%
+  select( blabel, value )
 

### 2017-12-08 datasets/GSEA/gsea.Rmd
GSEA

--- 2017-12-08 datasets/GSEA/gsea.Rmd
@@
 gnType <-  genes(EnsDb.Hsapiens.v86, column="gene_biotype")
+
 dsdObject <- dsdObject[rownames(dsdObject) %in% gnType[gnType$gene_biotype == "protein_coding"]$gene_id,]
 dsdObject <- DESeq(dsdObject)
-res <- results( dsdObject, lfcThreshold=1 )
+res <- results( dsdObject, independentFiltering=FALSE )
 
@@
 dplyr:::filter( goRes, numDEInCat > 0, numInCat > 15 ) %>%
-  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=7)
+  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=11)
 
@@
 
-rejection_scatter_bins( sGSEA, "ind_covariate", threshold=15, bins=4, as_fraction=FALSE, supplementary=FALSE)
 ```
@@
 
-getOverlapIndexes <- function( labs=c("lfdr", "ihw-a10"), alpha=0.1 ){
-  rowSums( assays(sGSEA)[["qvalue"]][,labs] < alpha) == length(labs) &
+getOverlapIndexes <- function( labs=c("lfdr", "ihw-a10"), alpha=0.1, sGSEA ){
+  rowSums( assays(sGSEA)[["qvalue"]][,labs,drop=FALSE] < alpha) == length(labs) &
   !(rowSums( assays(sGSEA)[["qvalue"]][,c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf")] < alpha, na.rm=TRUE ) > (length(labs)))
@@
 
-set1 <- getOverlapIndexes( c("lfdr", "ihw-a10"), alpha=0.1 )
-set2 <- getOverlapIndexes( c("lfdr","ihw-a10","bl-df03","qvalue","bh"), alpha=0.1 )
-set3 <- getOverlapIndexes( c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf"), alpha=0.1 )
+set1 <- getOverlapIndexes( c("lfdr", "ihw-a10"), alpha=0.1, sGSEA=sGSEA )
+set2 <- getOverlapIndexes( c("bl-df03"), alpha=0.1, sGSEA=sGSEA )
+set3 <- getOverlapIndexes( c("lfdr","ihw-a10","bl-df03","qvalue","bh"), alpha=0.1, sGSEA=sGSEA )
 
@@
   geom_boxplot()
-
-dim(goRes)
-
-goRes[set1,]
 
@@
 ```
+
+# Same analysis but using the HSC-MPP dataset from Nina
+
+```{r}
+
+dsdMouse <- "dsdHSCMPP.RData"
+if( !file.exists( dsdMouse ) ){
+  download.file("https://www.dropbox.com/s/spo1pifum5k72rn/dseHSCMPP.RData?dl=1", destfile="dsdHSCMPP.RData")
+}
+load(dsdMouse)
+dseHSCMPP <- DESeq( dseHSCMPP)
+
+res <- results( dseHSCMPP, contrast=c("conditions", "HSC", "MPP1"), independentFiltering=FALSE )
+genes <- as.numeric( res$padj < 0.1 )
+names( genes ) <- rownames( res )
+genes[is.na( genes )] <- 0
+table(genes)
+
+### preparing gene length info and goSets ###
+library("EnsDb.Mmusculus.v75")
+txByGene <- transcriptsBy(EnsDb.Mmusculus.v75, "gene")
+geneLength <- sapply( width( txByGene ), median )
+geneLength <- geneLength[names(genes)]
+
+file="goSetsMm.RData"
+if(!file.exists(file)){
+  stopifnot( require( biomaRt ) )
+  mart <- useMart( "ensembl", "mmusculus_gene_ensembl" )
+  goSets <- getBM( c("ensembl_gene_id", "go_id"), mart=mart, filters="ensembl_gene_id", values=rownames(res))
+  goSets <- goSets[!nchar( goSets$go_id ) == 0,]
+  goSets <- with( goSets, split( go_id, ensembl_gene_id ) )
+  save(goSets, file=file)
+}else{
+  load(file)
+}
+
+pwf <- nullp(genes, bias.data=geneLength )
+goRes <- goseq( pwf, gene2cat=goSets )
+
+res <- dplyr:::select( goRes, c("over_represented_pvalue", "numInCat") ) %>%
+  dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
+res$pval <- pmin( res$pval, 1 )
+sGSEAMouse <- designBen %>% buildBench( res, ftCols="ind_covariate")
+assayNames(sGSEAMouse) <- "qvalue"
+sGSEAMouse <- addDefaultMetrics( sGSEAMouse )
+
+rank_scatter( dat=goRes, pval="over_represented_pvalue", 
+              covariate="numInCat", bins=50, funx=log2, 
+              funfill=log10_trans()) +
+    ylim(0, 12) + ggtitle("Over-represented gene sets") +
+    xlab( expression(log[10]~"(# of genes)")) +
+    ylab( expression(-log[10]~"(p-value)") )
+
+library(dplyr)
+
+dplyr:::filter( goRes, numDEInCat > 0, numInCat > 20)  %>%
+  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=7)
+
+rejections_scatter( sGSEAMouse, as_fraction=FALSE, supplementary=FALSE)
+
+plotFDRMethodsOverlap( sGSEAMouse, supplementary=FALSE, order.by="freq")
+
+set1 <- getOverlapIndexes( c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf"), alpha=0.1, sGSEA=sGSEAMouse)
+set2 <- getOverlapIndexes( c("lfdr", "ihw-a10"), alpha=0.1, sGSEA=sGSEAMouse )
+set3 <- getOverlapIndexes( c("bl-df03"), alpha=0.1, sGSEA=sGSEAMouse )
+
+sizesDf <- data.frame( 
+  size=c( rowData( sGSEAMouse )$ind_covariate[which(set1)], 
+          rowData( sGSEAMouse )$ind_covariate[which(set2)], 
+          rowData( sGSEAMouse )$ind_covariate[which(set3)] ),
+  label=rep(c("set1", "set2", "set3"), c(sum(set1), sum(set2), sum(set3)) ) )
+
+ggplot( sizesDf, aes( label, log2(size) )) +
+  geom_boxplot()
+
+```

### 2017-12-09 datasets/GSEA/gsea.Rmd
Added code to plot boxplot of covariates for the overlaps between methods

--- 2017-12-08 datasets/GSEA/gsea.Rmd
@@
 
-any( is.na(genes) )
-stopifnot(!any(is.na(genes)))
 pwf <- nullp(genes, bias.data=geneLength )
@@
 
-dplyr:::filter( goRes, numDEInCat > 0, numInCat > 15 ) %>%
-  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=7)
+under_represented_pvalue
+dplyr:::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
+  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=10)
 
@@
 ```{r plottingResults}
-rejections_scatter( sGSEA, as_fraction=FALSE, supplementary=FALSE)
+rejections_scatter( sGSEA, as_fraction=TRUE, supplementary=FALSE)
 
 sGSEA <- estimatePerformanceMetrics(sGSEA, addColData=TRUE)
-
-estimatePerformanceMetrics( sGSEA, alpha=0.1, tidy=TRUE ) %>%
-  filter( blabel %in% c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf") ) %>%
-  select( blabel, value )
-
-rejection_scatter_bins( sGSEA, "ind_covariate", threshold=15, bins=4, as_fraction=FALSE, supplementary=FALSE)
 ```
@@
 
-getOverlapIndexes <- function( labs=c("lfdr", "ihw-a10"), alpha=0.1 ){
-  rowSums( assays(sGSEA)[["qvalue"]][,labs] < alpha) == length(labs) &
-  !(rowSums( assays(sGSEA)[["qvalue"]][,c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf")] < alpha, na.rm=TRUE ) > (length(labs)))
-}
-
-set1 <- getOverlapIndexes( c("lfdr", "ihw-a10"), alpha=0.1 )
-set2 <- getOverlapIndexes( c("lfdr","ihw-a10","bl-df03","qvalue","bh"), alpha=0.1 )
-set3 <- getOverlapIndexes( c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf"), alpha=0.1 )
-
-sizesDf <- data.frame( 
-  size=c( rowData( sGSEA )$ind_covariate[which(set1)], 
-          rowData( sGSEA )$ind_covariate[which(set2)], 
-          rowData( sGSEA )$ind_covariate[which(set3)] ),
-  label=rep(c("set1", "set2", "set3"), c(sum(set1), sum(set2), sum(set3)) ) )
-
-ggplot( sizesDf, aes( label, log2(size) )) +
-  geom_boxplot()
-
-dim(goRes)
-
-goRes[set1,]
-
-```
-
-Plotting the methods overlap
-
-```{r}
-
-sGSEASub <- sGSEA[,!apply( is.na( assays(sGSEA)[["qvalue"]] ), 2, all )]
-sGSEASub <- sGSEASub[,-which(grepl("ihw", colData( sGSEASub )$blabel) & colData( sGSEASub )$param.alpha != 0.1)]
-sGSEASub <- sGSEASub[,-which( grepl("bl", colData( sGSEASub)$blabel) & 
-                   colData(sGSEASub)$param.smooth.df != 2L )]
-sGSEASub <- sGSEASub[,!colData(sGSEASub)$blabel %in% "unadjusted"]
-
-plotMethodsOverlap(sGSEASub, order.by = "freq", nsets=10)
+methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
+plotCovariateBoxplots( sGSEA, alpha=0.1, nsets=6, methods=methods, trans=log2 )
 
@@
 ```{r, eval=FALSE }
-
 nms <- names( genes )
@@
 assayNames(pr) <- "qvalue"
-pr <- addDefaultMetrics(pr, metrics="rejections")
-
-estimatePerformanceMetrics( pr, alpha=0.1, tidy=TRUE ) %>%
-  filter( blabel %in% c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf") ) %>%
-  select( blabel, value )
+pr <- addDefaultMetrics(pr)
 

### 2017-12-09 datasets/R/rejections_scatter.R

@@ -129,3 +129,52 @@ plotFDRMethodsOverlap <- function( object, supplementary=TRUE, alpha=0.1, ... ){
   colnames(object) <- colData( object )$blabel
   plotMethodsOverlap( object, alpha=alpha, ... )
 }
+
+
+#' Boxplots of covariates for rejections unique to methods subsets
+#'   
+#' @param object A SummarizedBenchmark object
+#' @param alpha An alpha value.
+#' @param nsets Number of method sets to plot. Method sets are ordered decreasingly according 
+#' to the number of rejections specific to the sets of methods.
+#' @param methods A character vector of methods to consider (must be a level of colData(object)$blabel).
+#' @param trans A function to transform the y-axis of the boxplot
+#'        
+#' @return an ggplot2 boxplot
+#' 
+#' @author Alejandro Reyes
+plotCovariateBoxplots <- function( object, alpha, nsets=8, methods=NULL, 
+                                   maxNum=length(methods), 
+                                   trans=function(x){x} ){
+  stopifnot( is(object, "SummarizedBenchmark") )
+  if( is.null( methods ) ){
+    methods <- colData(object)$blabel
+  }
+  mat <- ( assays( object )[["qvalue"]][,methods] < alpha )
+  covariate <- mcols( object )$ind_covariate
+  allLengths <- lapply( seq_len( maxNum ), function(n){
+    combs <- combn( methods, n )
+    indCov <- lapply( seq_len( ncol( combs ) ), function(x){
+      detectedInComb <- rowSums( mat[,combs[,x], drop=FALSE], na.rm=TRUE ) == n & 
+        !(rowSums( mat[,methods,drop=FALSE], na.rm=TRUE ) > n)
+      covariate[detectedInComb]
+    } )
+    combs <- apply( combs, 2, function(x){ paste( gsub("-", "", x), collapse="-" ) } )
+    names(indCov) <- combs
+    indCov[sapply( indCov, length ) > 0]
+  } )
+  allLengths <- unlist( allLengths, recursive=FALSE )
+  allLenghts <- allLengths[order( sapply( allLengths, length ), decreasing = TRUE)]
+  allLenghts <- head( allLenghts, nsets )
+  sizesAll <- unlist( allLenghts )
+  dfCov <- data.frame(
+    covariate = sizesAll,
+    methods=factor( rep( names( allLenghts ), sapply( allLenghts, length ) ), 
+                    levels=names(allLenghts)  ) )
[…]

### 2017-12-09 datasets/GSEA/gsea.Rmd

--- 2017-12-08 datasets/GSEA/gsea.Rmd
@@
     ylab( expression(-log[10]~"(p-value)") ) 
-
 ```
@@
 
-dplyr:::filter( goRes, numDEInCat > 0, numInCat > 15 ) %>%
-  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=11)
+under_represented_pvalue
+dplyr:::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
+  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=10)
 
@@
 
-estimatePerformanceMetrics( sGSEA, alpha=0.1, tidy=TRUE ) %>%
-  filter( blabel %in% c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf") ) %>%
-  select( blabel, value )
-
 ```
@@
 
-getOverlapIndexes <- function( labs=c("lfdr", "ihw-a10"), alpha=0.1, sGSEA ){
-  rowSums( assays(sGSEA)[["qvalue"]][,labs,drop=FALSE] < alpha) == length(labs) &
-  !(rowSums( assays(sGSEA)[["qvalue"]][,c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf")] < alpha, na.rm=TRUE ) > (length(labs)))
-}
-
-set1 <- getOverlapIndexes( c("lfdr", "ihw-a10"), alpha=0.1, sGSEA=sGSEA )
-set2 <- getOverlapIndexes( c("bl-df03"), alpha=0.1, sGSEA=sGSEA )
-set3 <- getOverlapIndexes( c("lfdr","ihw-a10","bl-df03","qvalue","bh"), alpha=0.1, sGSEA=sGSEA )
-
-sizesDf <- data.frame( 
-  size=c( rowData( sGSEA )$ind_covariate[which(set1)], 
-          rowData( sGSEA )$ind_covariate[which(set2)], 
-          rowData( sGSEA )$ind_covariate[which(set3)] ),
-  label=rep(c("set1", "set2", "set3"), c(sum(set1), sum(set2), sum(set3)) ) )
-
-ggplot( sizesDf, aes( label, log2(size) )) +
-  geom_boxplot()
+methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
+plotCovariateBoxplots( sGSEA, alpha=0.1, nsets=6, methods=methods, trans=log2 )
 

### 2017-12-09 datasets/GSEA/gsea.Rmd
Generalized plotCovariateBoxplot for any SB object

@@
     ylab( expression(-log[10]~"(p-value)") )
+
 ```
@@
 
-under_represented_pvalue
 dplyr:::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
@@
 
-set1 <- getOverlapIndexes( c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf"), alpha=0.1, sGSEA=sGSEAMouse)
-set2 <- getOverlapIndexes( c("lfdr", "ihw-a10"), alpha=0.1, sGSEA=sGSEAMouse )
-set3 <- getOverlapIndexes( c("bl-df03"), alpha=0.1, sGSEA=sGSEAMouse )
+plotCovariateBoxplots( sGSEAMouse, alpha=0.1, nsets=4, methods=methods, trans=log2, maxNum=5)
 
-sizesDf <- data.frame( 
-  size=c( rowData( sGSEAMouse )$ind_covariate[which(set1)], 
-          rowData( sGSEAMouse )$ind_covariate[which(set2)], 
-          rowData( sGSEAMouse )$ind_covariate[which(set3)] ),
-  label=rep(c("set1", "set2", "set3"), c(sum(set1), sum(set2), sum(set3)) ) )
-
-ggplot( sizesDf, aes( label, log2(size) )) +
-  geom_boxplot()
+plotCovariateBoxplots( sGSEA, alpha=0.1, nsets=7, methods=methods, trans=log2, maxNum=5)
 

### 2017-12-09 datasets/RNA-Seq/rna-seq-benchmarkFDR.Rmd
Adding RNA-Seq R Markdown file

---
title: "R Notebook"
output: html_notebook
---
```{r}
# Load packageas and source benchmark FDR
library(data.table)
library(readxl)
library(dplyr)
library(ggplot2)
library(magrittr)
library(R.utils)
library(cowplot)

# benchmark methods
library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)
library(FDRreg)

# comparison tools/functions
library(SummarizedBenchmark)
sourceDirectory("~/Downloads/benchmark-fdr/datasets/R/")

# set up parallel backend
library(BiocParallel)
cores <- 4
multicoreParam <- MulticoreParam(workers = cores)

```

```{r}
#Load expression data
load('~/Downloads/dsdObjects.RData')

#Filter to include samples from only 2 tissuesdsd
filteredDsd1 <- dsd1[,colData(dsd1)$tissue %in% c('Brain - Nucleus accumbens (basal ganglia)','Brain - Putamen (basal ganglia)')]
filteredDsd1$tissue <- droplevels(filteredDsd1$tissue)

#Buid DESeq object and run DESeq2
dds <- DESeq(filteredDsd1, parallel = T)
res <- results(dds, independentFiltering = F)

#Build input data frame for summarized benchmark
#Important to not change col names since SB is strict about that
geneExp <- data.frame(geneName=rownames(res), pval=res$pvalue, SE=res$lfcSE, 
                      ind_covariate = res$baseMean, effect_size=res$log2FoldChange, test_statistic=res$stat)

```

```{r}
strat_hist(geneExp, pvalue="pval", covariate="ind_covariate", maxy=10)
```

```{r}
rank_scatter(geneExp, pvalue="pval", covariate="ind_covariate")
```

```{r}
bd <- initializeBenchDesign()
sb <- bd %>% buildBench(data=geneExp, parallel = T)

assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)

```

```{r}
rejections_scatter(sb, supplementary=FALSE)
```

```{r}
plotFDRMethodsOverlap(sb, alpha=0.05, nsets=ncol(sb), order.by="freq", decreasing=TRUE, supplementary=FALSE)
```

### 2017-12-10 datasets/simulations/analysis-03-plot-tsim-summaries.Rmd
Update general plots of tsim/psim results

@@
     if (!file.exists(file.path(dd, "results-performanceMetrics-alpha05.rds"))) {
+        print(paste0("running: ", basename(dd))) 
         fl <- list.files(dd, pattern=".rds", full.names=TRUE)
@@
         tsbc <- as.tibble(bind_rows(tsbl, .id = "param"))
+        print(paste0("saving: ", basename(dd))) 
         saveRDS(tsbc, file.path(dd, "results-performanceMetrics-alpha05.rds"))
+        print(paste0("done: ", basename(dd))) 
+    } else {
+        print(paste0("skipping: ", basename(dd)))
     }
@@
 For simulation settings where the effect size (expected group differece) is not varying, a fixed difference of `1.5` was used for all alternative hypotheses.
-For simulation settings with the uniform (un-informative) or SE informative covariate, if the proportion of null hypotheses is not varying, `80%` of hypotheses were simulated under the null.
+For simulation settings with the uniform (uninformative) or SE (informative) covariate, if the proportion of null hypotheses is not varying, `80%` of hypotheses were simulated under the null.
 For simulation settings with the "Boca-Leek" style informative covariate, a functional relationship was first defined between the probability of a test being null and the "informative covariate".
@@
 The form of the relationship used here was taken from Setting II of the Boca-Leek simulations.
+The marginal proportion of null hypothses is approximately `92.6%`.
+For the Manual Covariates, as with the Boca-Leek style covariate, a functional relationship was defined between the randomly sampled covariate and the probability of a hypothesis being null or alternative.
+Three functional relationships were tested: "step (less)", "step (more)", and "cubic".
+The first two specified a piecewise constant relationship of varying informativeness (more or less), and the third was a (continuous) scaled and shifted cubic function.
+The three variations on the manual covariate were chosen to have a marginal proportion of null hypotheses of `80%`.
+By definition, the ASH method makes use of the standard error (SE) for each hypothesis test, and does not accept any other additional information.
+The results for the ASH method are therefore **always** based on using the SE as the "informative covariate."
 
@@
                takes_cov = blabel %in%
-                   c("IHW", "ASH", "Boca-Leek", "FDRreg (theoretical)", "FDRreg (empirical)")) %>%
+                   c("IHW", "ASH", "Boca-Leek", "local FDR", 
+                     "FDRreg (theoretical)", "FDRreg (empirical)")) %>%
         mutate(blabel = factor(blabel, levels = lab_order))

### 2017-12-10 datasets/simulations/analysis-04-plot-psim-summaries.Rmd
Update general plots of tsim/psim results

@@
 This approach to simulating data follows similar simulations carried out in the Boca-Leek, IHW ("size investing strategy" simulations), and ASH papers.
-Different simulation settings were considered to evaluate the impact of varying effect sizes, effect size distributions, effect size sampling distributions, proportions of null hypotheses, and presence or absence of an informative covariate.
-More details on the simulation settings considered are included in the header of the `psim-core-settings.R` script.
+Different simulation settings were considered to evaluate the impact of varying effect sizes, effect size distributions, sampling distributions, proportions of null hypotheses, and presence or absence of an informative covariate.
+By definition, the ASH method makes use of the standard error (SE) for each hypothesis test, and does not accept any other additional information.
+The results for the ASH method are therefore always based on using the SE as the "informative covariate." More details on the simulation settings considered are included in the header of the `psim-core-settings.R` script.
 Each setting was replicated 100 times.
 
-Since these simulations do not rely on underlying simulated data, but rather simulate summary statistics directly, there is no way of calculating a corresponding standard error (SE) for each hypothesis test.
-*As a result, the ASH method, which requires the SE for each test, is not included in these simulations.*
+Since these simulations do not rely on underlying simulated data, but rather simulate summary statistics directly, there is no way of calculating a corresponding SE for each hypothesis test.
+*As such, the ASH method, which requires the SE for each test, is run using the true (typically unknown) SE underlying the simulation.* Therefore, the ASH results must be interpreted with caution.
 
@@
                "IHW", "Boca-Leek", "FDRreg (theoretical)", "FDRreg (empirical)",
-               "unadjusted")
+               "ASH", "unadjusted")
 ```
@@
     tsb %>%
-        mutate(blabel = fct_recode(f = blabel, "IHW" = "ihw-a05",
+        mutate(blabel = fct_recode(f = blabel, "IHW" = "ihw-a05", "ASH" = "ashs",
                                    "Boca-Leek" = "bl-df03", "local FDR" = "lfdr",
@@
                takes_cov = blabel %in%
-                   c("IHW", "Boca-Leek", "FDRreg (theoretical)", "FDRreg (empirical)")) %>%
+                   c("IHW", "ASH", "Boca-Leek", "local FDR",
+                     "FDRreg (theoretical)", "FDRreg (empirical)")) %>%
         mutate(blabel = factor(blabel, levels = lab_order))
@@
 <!-- ######################################################################## -->
+### Manual Covariate (Step - Less Informative)
+
+```{r}
+f <- "data-psim/M100/esize_fixed-bl-step-less/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")),
+           param >= 10) %>%
+    mutate(param = paste("effect size =", sprintf("%.1f", param / 10)))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Direct Parameter Simulation w/ Varying Effect Size, Manual Covariate (step - less informative) (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
+### Manual Covariate (Step - More Informative)
+
+```{r}
+f <- "data-psim/M100/esize_fixed-bl-step-more/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")),
+           param >= 10) %>%
+    mutate(param = paste("effect size =", sprintf("%.1f", param / 10)))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Direct Parameter Simulation w/ Varying Effect Size, Manual Covariate (step - more informative) (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
+### Manual Covariate (Cubic)
+
+```{r}
+f <- "data-psim/M100/esize_fixed-bl-cubic/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*?)-.*", "\\1", param) %>% as.numeric,
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")),
+           param >= 10) %>%
+    mutate(param = paste("effect size =", sprintf("%.1f", param / 10)))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Direct Parameter Simulation w/ Varying Effect Size, Manual Covariate (cubic) (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
 ## Varying Null Proportion
@@
 ```{r}
-## older run with alternative tests at 1.5.
-f <- "data-psim/M100/_pi0-uniform-delta15/results-performanceMetrics-alpha05.rds"
+f <- "data-psim/M100/pi0-uniform/results-performanceMetrics-alpha05.rds"
 tsb <- readRDS(f)
@@
 <!-- ######################################################################## -->
+### Manual Covariate (Step - Less Informative)
+
+```{r}
+f <- "data-psim/M100/esize_random_ua-bl-step-less/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*)-.+", "\\1", param),
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Direct Parameter Simulation w/ Varying Randomly Sampled UA Effect, Manual Covariate (step - less informative) (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
+### Manual Covariate (Step - More Informative)
+
+```{r}
+f <- "data-psim/M100/esize_random_ua-bl-step-more/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*)-.+", "\\1", param),
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Direct Parameter Simulation w/ Varying Randomly Sampled UA Effect, Manual Covariate (step - more informative) (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
+### Manual Covariate (Cubic)
+
+```{r}
+f <- "data-psim/M100/esize_random_ua-bl-cubic/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*)-.+", "\\1", param),
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Direct Parameter Simulation w/ Varying Randomly Sampled UA Effect, Manual Covariate (cubic) (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
 ## Randomly-Sampled Mean-Shift Effect Size
@@
 gen_plot("Direct Parameter Simulation w/ Varying Randomly Sampled Mean Shifted Effect, Boca-Leek Style Covariate (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
+### Manual Covariate (Step - Less Informative)
+
+```{r}
+f <- "data-psim/M100/esize_random_shift-bl-step-less/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*)-.+", "\\1", param),
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Direct Parameter Simulation w/ Varying Randomly Sampled Mean Shifted Effect, Manual Covariate (step - less informative) (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
+### Manual Covariate (Step - More Informative)
+
+```{r}
+f <- "data-psim/M100/esize_random_shift-bl-step-more/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*)-.+", "\\1", param),
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Direct Parameter Simulation w/ Varying Randomly Sampled Mean Shifted Effect, Manual Covariate (step - more informative) (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
+### Manual Covariate (Cubic)
+
+```{r}
+f <- "data-psim/M100/esize_random_shift-bl-cubic/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*)-.+", "\\1", param),
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Direct Parameter Simulation w/ Varying Randomly Sampled Mean Shifted Effect, Manual Covariate (cubic) (100 replications)",
          tsb, cutoffs)
@@
 ```
+
+<!-- ######################################################################## -->
+### Manual Covariate (Step - Less Informative)
+
+```{r}
+f <- "data-psim/M100/altnoise-bl-step-less/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*)-.+", "\\1", param),
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Direct Parameter Simulation w/ Varying Sampling Distribution, Manual Covariate (step - less informative) (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
+### Manual Covariate (Step - More Informative)
+
+```{r}
+f <- "data-psim/M100/altnoise-bl-step-more/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*)-.+", "\\1", param),
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Direct Parameter Simulation w/ Varying Sampling Distribution, Manual Covariate (step - more informative) (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
+### Manual Covariate (Cubic)
+
+```{r}
+f <- "data-psim/M100/altnoise-bl-cubic/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*)-.+", "\\1", param),
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.20))
+
+## summarize results in plot
+gen_plot("Direct Parameter Simulation w/ Varying Sampling Distribution, Manual Covariate (cubic) (100 replications)",
+         tsb, cutoffs)
+```
+
+<!-- ######################################################################## -->
+## Varying Parameter Sampling Distribution (all null)
+<!-- ######################################################################## -->
+
+Similar simulation settings as above, where various sampling distributions are considered, but with `pi0 = 1`, i.e. all null hypotheses.
+
+<!-- ######################################################################## -->
+### Non-Informative Covariate
+
+```{r}
+f <- "data-psim/M100/allnull-uniform/results-performanceMetrics-alpha05.rds"
+tsb <- readRDS(f)
+
+outp <- gsub("data", "plots", dirname(f))
+dir.create(outp, showWarnings=FALSE, recursive=TRUE)
+
+## remove non-percentage metric (rejections),
+## subset on settings (effect < 1.0 not really informative)
+tsb %<>%
+    mutate(param = gsub(".*?-.*?_(.*)-.+", "\\1", param),
+           performanceMetric =
+               factor(performanceMetric,
+                      levels = c("rejections", "rejectprop", "TPR",
+                                 "TNR", "FPR", "FNR", "FWER"))) %>%
+    filter(!(performanceMetric %in% c("rejections", "FWER")))
+
+## relabel methods for easier reading
+tsb %<>% re_blabel
+
+## expected thresholds based on simulation setting
+cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
+                                                 levels=levels(tsb$performanceMetric)),
+                      expected = c(0.05, 0.00))
+
+## summarize results in plot
+gen_plot("Direct Parameter Simulation w/ Varying Sampling Dist (all null), Non-Informative Covariate (100 replications)",
+         tsb, cutoffs)
+```

### 2017-12-14 datasets/microbiome/microbiome-SummarizedBenchmark.Rmd
Explore weird microbiome results + add CRC Baxter

@@
 ---
-title: "Comparison of FDR methods on GWAS with SummarizedBenchmark"
-author: "Keegan Korthauer"
+title: "Comparison of FDR methods on 16S microbiome data with SummarizedBenchmark"
+author: "Claire Duvallet"
 date: '`r format(Sys.Date(), "%B %d, %Y")`'
@@
 ```{r, schubert-ubi, fig.width=10, fig.height=3.5}
-strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=20)
+strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=20, binwidth=0.05)
 rank_scatter(res, pvalue="pval", covariate="ubiquity")
@@
 ```{r, schubert-abun, fig.width=10, fig.height=3.5}
-strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=20)
-rank_scatter(res, pvalue="pval", covariate="ubiquity")
+strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
+rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
 ```
@@
 First, we'll create an object of `BenchDesign` class to hold the data and 
-add the benchmark methods to the `BenchDesign` object.
+add the benchmark methods to the `BenchDesign` object. We'll do this for a few
+values of nmids, to explore its effects on the outputs.
 
-```{r, schubert-benchdesign}
-bd <- initializeBenchDesign()
-```
-
-Now, we're ready to construct the `SummarizedBenchmark` object, which will run
+Then, we'll construct the `SummarizedBenchmark` object, which will run
 the functions specified in each method (these are actually sourced in from the
@@
 
-```{r, GWAS1-sb, results="hide", message=FALSE, eval=FALSE}
+```{r, schubert-sb, results="hide", message=FALSE, eval=FALSE}
+for (nmids in c(5, 50, 150)){
+  print(nmids)
+  bd <- initializeBenchDesign(nmids=nmids)
+
   resfile <- paste0(resdir, "schubert_summarizedBenchmark_",
-                  nrow(res), ".RData")
+                    nrow(res), "_nmids_", nmids, ".RData")
   duration <- NA
@@
   }
+}
+
+```
+
+Just for future runs, where I load the file instead of re-calculating it, here are the errors I got:
+
+```
+[1] 5
+Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !NaNs produced<simpleError in if (sum(D^2) > qchisq(0.9, nmids - 2 - df)) {    warning(paste0("f(z) misfit = ", round(D, 1), ".
+Rerun with increased df."))}: missing value where TRUE/FALSE needed>
+
+[1] 50
+Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !f(z) misfit = -21.6.
+Rerun with increased df.f(z) misfit = -22.5.
+Rerun with increased df.f(z) misfit = -10.
+Rerun with increased df.f(z) misfit = 32.
+Rerun with increased df.f(z) misfit = 14.
+Rerun with increased df.f(z) misfit = -32.2.
+Rerun with increased df.f(z) misfit = 6.9.
+Rerun with increased df.f(z) misfit = -1.3.
+Rerun with increased df.f(z) misfit = 0.7.
+Rerun with increased df.f(z) misfit = 3.4.
+Rerun with increased df.f(z) misfit = 5.5.
+Rerun with increased df.f(z) misfit = 0.1.
+Rerun with increased df.f(z) misfit = -0.9.
+Rerun with increased df.f(z) misfit = -1.5.
+Rerun with increased df.f(z) misfit = -1.9.
+Rerun with increased df.f(z) misfit = -0.8.
+Rerun with increased df.f(z) misfit = -2.4.
+Rerun with increased df.f(z) misfit = 1.7.
+Rerun with increased df.f(z) misfit = 0.5.
+Rerun with increased df.f(z) misfit = 2.
+Rerun with increased df.f(z) misfit = 0.1.
+Rerun with increased df.f(z) misfit = 1.4.
+Rerun with increased df.f(z) misfit = 0.2.
+Reru...
+<truncated>
+
+[1] 150
+Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !f(z) misfit = 83.4.
+Rerun with increased df.f(z) misfit = -8.5.
+Rerun with increased df.f(z) misfit = -13.6.
+Rerun with increased df.f(z) misfit = -17.1.
+Rerun with increased df.f(z) misfit = -6.4.
+Rerun with increased df.f(z) misfit = -19.5.
+Rerun with increased df.f(z) misfit = -17.8.
+Rerun with increased df.f(z) misfit = -3.
+Rerun with increased df.f(z) misfit = -21.1.
+Rerun with increased df.f(z) misfit = 28.6.
+Rerun with increased df.f(z) misfit = -26.
+Rerun with increased df.f(z) misfit = 59.3.
+Rerun with increased df.f(z) misfit = -17.1.
+Rerun with increased df.f(z) misfit = 12.1.
+Rerun with increased df.f(z) misfit = -28.3.
+Rerun with increased df.f(z) misfit = -27.6.
+Rerun with increased df.f(z) misfit = 83.7.
+Rerun with increased df.f(z) misfit = -10.5.
+Rerun with increased df.f(z) misfit = -22.6.
+Rerun with increased df.f(z) misfit = -23.3.
+Rerun with increased df.f(z) misfit = 37.3.
+Rerun with increased df.f(z) misfit = -18.
+Rerun with increased df.f(z...
+<truncated>>
 ```
@@
 
+Debugging note: to look at the matrix of qvalues, run `assays(sb)[["qvalue"]]`
+
 ```{r, schubert-plot, results="hide", width=15, height=15, eval=FALSE}
+# don't plot scott theoretical for now
+rmv.scott <- which(grepl("scott", colnames(sb)))
+
 # plot nrejects by method overall and stratified by covariate
-rejections_scatter(sb)
+rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
 
-rejection_scatter_bins(sb, covariate="ind_covariate", bins=4)
+rejection_scatter_bins(sb, covariate="ind_covariate", bins=4, supplementary=FALSE)
 
@@
                               colData(sb)$param.alpha != 0.05)
+rmv.cols <- c(rmv.ihw, rmv.scott)
 
-plotMethodsOverlap(sb[,-rmv.ihw], 
-                   alpha=0.05, nsets=ncol(sb)-length(rmv.ihw),
+plotMethodsOverlap(sb[,-rmv.cols], 
+                   alpha=0.05, nsets=ncol(sb)-length(rmv.cols),
                    order.by="freq", decreasing=TRUE)
@@
 # upset plot excluding "unadjusted"
-plotMethodsOverlap(sb[,-c(1,rmv.ihw)], 
-                   alpha=0.05, nsets=ncol(sb)-length(rmv.ihw)-1,
+plotMethodsOverlap(sb[,-c(1,rmv.cols)], 
+                   alpha=0.05, nsets=ncol(sb)-length(rmv.cols)-1,
                    order.by="freq", decreasing=TRUE)
 ```
+
+Hm, now the code runs. However, there are clearly still some issues:
+- ashs rejects all hypotheses (all q-values are essentially 0).
+- lfdr and scott-empirical are all NaN (I think this is likely related to the df error)
+
+```{r, eval=FALSE}
+assays(sb)[["qvalue"]][ , 15] %>% max()
+sum(is.na(assays(sb)[["qvalue"]][ , 22]))
+sum(is.na(assays(sb)[["qvalue"]][ , 20]))
+```
+
+Rather than de-bug this dataset (which might be weird), let's try instead to work with the obesity dataset and see if I get the same problems.
 
@@
 strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=5, binwidth=0.05, numQ=4)
-rank_scatter(res, pvalue="pval", covariate="ubiquity")
+rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
 ```
@@
 ```
+
+#### Set up BenchDesign object
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object.
+
+Next, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, goodrich-sb, results="hide", message=FALSE, eval=FALSE}
+for (nmids in c(5, 50, 150)){
+  print(nmids)
+  bd <- initializeBenchDesign(nmids=nmids)
+
+  resfile <- paste0(resdir, "goodrich_summarizedBenchmark_",
+                    nrow(res), "_nmids_", nmids, ".RData")
+  duration <- NA
+  if (!file.exists(resfile)){
+    t1 <- proc.time()
+    sb <- bd %>% buildBench(data=res, ftCols = "ind_covariate")
+                            #parallel=TRUE, BPPARAM=multicoreParam)
+    metadata(sb)$data_download_link <- "https://zenodo.org/record/840333/files/cdi_schubert_results.tar.gz"
+    save(sb, file=resfile)
+    duration <- round((proc.time()-t1)[3]/60,1)
+  }else{
+    load(resfile)
+  }
+}
+```
+
+The outputs of this call are:
+
+```
+[1] 5
+Squarem-1
+Objective fn: 715262
+Objective fn: 104929  Extrapolation: 0  Steplength: 1
+Objective fn: 95918.9  Extrapolation: 1  Steplength: 2.44204
+Objective fn: 95614.3  Extrapolation: 1  Steplength: 2.72536
+Objective fn: 95583.3  Extrapolation: 1  Steplength: 2.8568
+Objective fn: 95578.1  Extrapolation: 1  Steplength: 3.54373
+Objective fn: 95576.3  Extrapolation: 1  Steplength: 3.52004
+Objective fn: 95575.4  Extrapolation: 1  Steplength: 4
+.
+.
+.
+NaNs produced<simpleError in if (sum(D^2) > qchisq(0.9, nmids - 2 - df)) {    warning(paste0("f(z) misfit = ", round(D, 1), ".  Rerun with increased df."))}: missing value where TRUE/FALSE needed>
+
+[1] 50
+Squarem-1
+Objective fn: 715262
+Objective fn: 104929  Extrapolation: 0  Steplength: 1
+Objective fn: 95918.9  Extrapolation: 1  Steplength: 2.44204
+Objective fn: 95614.3  Extrapolation: 1  Steplength: 2.72536
+Objective fn: 95583.3  Extrapolation: 1  Steplength: 2.8568
+Objective fn: 95578.1  Extrapolation: 1  Steplength: 3.54373
+Objective fn: 95576.3  Extrapolation: 1  Steplength: 3.52004
+Objective fn: 95575.4  Extrapolation: 1  Steplength: 4
+Objective fn: 95574.9  Extrapolation: 1  Steplength: 3.36769
+Objective fn: 95574.5  Extrapolation: 1  Steplength: 5.09115
+Objective fn: 95574.3  Extrapolation: 1  Steplength: 2.52066
+Objective fn: 95574  Extrapolation: 1  Steplength: 13.3966
+.
+.
+.
+Objective fn: 95572.9  Extrapolation: 1  Steplength: 1.61619
+f(z) misfit = -8.7.
+Rerun with increased df.f(z) misfit = 33.3.
+Rerun with increased df.f(z) misfit = -9.4.
+Rerun with increased df.f(z) misfit = -37.3.
+Rerun with increased df.f(z) misfit = 49.2.
+Rerun with increased df.f(z) misfit = -36.9.
+Rerun with increased df.f(z) misfit = 10.6.
+Rerun with increased df.f(z) misfit = 13.3.
+Rerun with increased df.f(z) misfit = -18.3.
+Rerun with increased df.f(z) misfit = 20.1.
+Rerun with increased df.f(z) misfit = -21.6.
+Rerun with increased df.f(z) misfit = 23.4.
+Rerun with increased df.f(z) misfit = -27.
+Rerun with increased df.f(z) misfit = 17.2.
+Rerun with increased df.f(z) misfit = -8.2.
+Rerun with increased df.f(z) misfit = 4.3.
+Rerun with increased df.f(z) misfit = 3.4.
+Rerun with increased df.f(z) misfit = 2.4.
+Rerun with increased df.f(z) misfit = -2.3.
+Rerun with increased df.f(z) misfit = -10.7.
+Rerun with increased df.f(z) misfit = -2.4.
+Rerun with increased df.f(z) misfit = 15.4.
+Rerun with increased df.f(z) misf...
+<truncated>
+
+[1] 150
+Squarem-1
+Objective fn: 715262
+Objective fn: 104929  Extrapolation: 0  Steplength: 1
+Objective fn: 95918.9  Extrapolation: 1  Steplength: 2.44204
+Objective fn: 95614.3  Extrapolation: 1  Steplength: 2.72536
+Objective fn: 95583.3  Extrapolation: 1  Steplength: 2.8568
+Objective fn: 95578.1  Extrapolation: 1  Steplength: 3.54373
+Objective fn: 95576.3  Extrapolation: 1  Steplength: 3.52004
+Objective fn: 95575.4  Extrapolation: 1  Steplength: 4
+Objective fn: 95574.9  Extrapolation: 1  Steplength: 3.36769
+Objective fn: 95574.5  Extrapolation: 1  Steplength: 5.09115
+Objective fn: 95574.3  Extrapolation: 1  Steplength: 2.52066
+Objective fn: 95574  Extrapolation: 1  Steplength: 13.3966
+.
+.
+.
+f(z) misfit = -20.2.
+Rerun with increased df.f(z) misfit = 13.6.
+Rerun with increased df.f(z) misfit = -3.8.
+Rerun with increased df.f(z) misfit = 9.
+Rerun with increased df.f(z) misfit = -21.2.
+Rerun with increased df.f(z) misfit = 28.8.
+Rerun with increased df.f(z) misfit = 41.9.
+Rerun with increased df.f(z) misfit = -13.2.
+Rerun with increased df.f(z) misfit = -12.8.
+Rerun with increased df.f(z) misfit = 15.2.
+Rerun with increased df.f(z) misfit = -18.1.
+Rerun with increased df.f(z) misfit = -19.1.
+Rerun with increased df.f(z) misfit = -18.6.
+Rerun with increased df.f(z) misfit = -26.
+Rerun with increased df.f(z) misfit = -7.2.
+Rerun with increased df.f(z) misfit = 67.1.
+Rerun with increased df.f(z) misfit = 27.2.
+Rerun with increased df.f(z) misfit = -16.4.
+Rerun with increased df.f(z) misfit = -21.1.
+Rerun with increased df.f(z) misfit = -25.9.
+Rerun with increased df.f(z) misfit = -7.5.
+Rerun with increased df.f(z) misfit = 10.1.
+Rerun with increased df.f(z)...
+<truncated>
+```
+
+```{r, echo=FALSE, eval=FALSE}
+if(!is.na(duration)){
+  message("This step took ", duration, " minutes for ", nrow(res), " OTUs using ",
+        cores, " cores.")
+}
+```
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, goodrich-metrics, eval=FALSE}
+# rename assay to qvalue
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+Debugging note: to look at the matrix of qvalues, run `assays(sb)[["qvalue"]]`
+
+```{r, goodrich-plot, results="hide", width=15, height=15, eval=FALSE}
+
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
+
+rejection_scatter_bins(sb, covariate="ind_covariate", bins=4, supplementary=FALSE)
+
+# upset plot (only plot IHW for the plotted alpha)
+rmv.ihw <- which(grepl("ihw", colnames(sb)) &
+                              colData(sb)$param.alpha != 0.05)
+
+plotMethodsOverlap(sb[,-rmv.ihw], 
+                   alpha=0.05, nsets=ncol(sb)-length(rmv.ihw),
+                   order.by="freq", decreasing=TRUE)
+
+# upset plot excluding "unadjusted"
+plotMethodsOverlap(sb[,-c(1,rmv.ihw)], 
+                   alpha=0.05, nsets=ncol(sb)-length(rmv.ihw)-1,
+                   order.by="freq", decreasing=TRUE)
+```
+
+Hm, looks like all of the methods run without giving NaN's with nmids = 150 or lower...
+But the results still aren't right.
+
+```{r, eval=FALSE}
+assays(sb)[["qvalue"]][, "ashs"] %>% max()
+sum(is.na(assays(sb)[["qvalue"]][, "scott-empirical"]))
+sum(is.na(assays(sb)[["qvalue"]][, "lfdr"]))
+
+```
+
+### CRC Baxter - colorectal cancer
+
+Next, I'll look at the CRC Baxter dataset.
+For colorectal cancer, we do expect some amount of truly differentially abundant OTUs, but not as many as in diarrhea.
+This dataset will hopefully provide an intermediate non-extreme case study.
+We'll download the processed OTU tables from Zenodo and unzip them in the 
+`benchmark-fdr/datasets/microbiome/DATA` folder.
+
+#### Data download
+
+```{bash, baxter-download, eval=FALSE}
+cd /Users/claire/github/benchmark-fdr/datasets/microbiome/DATA/
+curl -O "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
+tar -xzvf crc_baxter_results.tar.gz
+cd ..
+```
+
+Next, we'll read in the unzipped OTU table and metadata files into R.
+
+We'll need to manually define which disease labels are of interest and what metadata
+column contains the sample IDs (i.e. the IDs in the OTU table).
+
+```{r, baxter-readdata}
+otu_path = paste0(datdir, "crc_baxter_results/RDP/crc_baxter.otu_table.100.denovo.rdp_assigned")
+meta_path = paste0(datdir, "crc_baxter_results/crc_baxter.metadata.txt")
+
+# DiseaseState labels to keep
+labels <- c("H", "CRC")
+
+# Metadata column that has sample IDs
+col_label <- "Sample_Name_s"
+
+# load OTU table and metadata
+otu <- read.table(otu_path)
+meta <- read.csv(meta_path, sep='\t')
+
+# Keep only samples with the right DiseaseState metadata
+meta <- meta %>% filter(DiseaseState %in% labels)
+
+# Add "X" in front of sample IDs because of how R read in the OTU table
+meta[, 1] <- sub("^", "X", meta[, 1])
+
+# Keep only samples with both metadata and 16S data
+keep_samples <- intersect(colnames(otu), meta[, 1])
+otu <- otu[, keep_samples]
+meta <- meta %>% filter(get(col_label) %in% keep_samples)
+
+```
+
+Since we'll be using OTU-wise covariates, we shouldn't need to perform any
+filtering/cleaning of the OTUs, apart from removing any that are all zeros. 
+(This may happen after removing shallow samples, I think.)
+However, let's make sure we get rid of any samples with too few reads. 
+We define the minimum number of reads per sample in `sample_reads`.
+
+After we've removed any shallow samples, we'll convert the OTU table to relative
+abundances. I'll add a pseudo-count of 1e-7 to any zero entries, to avoid
+problems with taking logs.
+
+```{r, baxter-cleandata}
+remove_shallow_smpls <- function(df, n_reads) {
+  # Removes samples with fewer than n_reads from dataframe df.
+  # df has OTUs in rows and samples in columns
+  return(df[, colSums(df) > n_reads])
+  }
+
+remove_shallow_otus <- function(df, n_reads){
+  return(df[rowSums(df) > n_reads, ])
+  }
+
+## Remove OTUs with fewere than 10 reads
+#otu <- remove_shallow_otus(otu, 10)
+
+# Remove samples with fewer than sample_Reads reads
+sample_reads <- 100
+otu <- remove_shallow_smpls(otu, sample_reads)
+
+# Update metadata with new samples
+meta <- meta %>% filter(get(col_label) %in% colnames(otu))
+
+# Remove empty OTUs
+otu <- otu[rowSums(otu) > 0, ]
+
+# Convert to relative abundance
+abun_otu <- t(t(otu) / rowSums(t(otu)))
+
+# Add pseudo counts
+#minabun <- min(abun_otu[abun_otu > 0]) # Use this to figure out what pseudo-count value to add
+zeroabun <- 1e-7
+abun_otu <- abun_otu + zeroabun
+
+```
+
+Next, we need to calculate the pvalues, effect size, and standard error for each OTU.
+Here, we'll compare CRC vs. healthy. We won't consider the nonCRC adenoma patients.
+We'll put these results into a dataframe, and label the columns with the 
+standardized names for downstream use (`pval`, `SE`, `effect_size`, `test_statistic`).
+The test statistic is the one returned by `wilcox.test()`.
+
+Note that the effect here is calculated as logfold change of mean abundance in controls
+relative to cases (i.e. `log(mean_abun[controls]/mean_abun[cases])`)
+
+While we're at it, we'll also calculate the mean abundance and ubiquity (detection rate)
+of each OTU. Later, we can assign their values to a new column called `ind_covariate` 
+for use in downstream steps.
+
+```{r, baxter-pvals}
+resfile <- paste0(resdir, "baxter_results_",
+                  nrow(abun_otu), "_OTUs.RData")
+if (!file.exists(resfile)){
+  # Get case and control indices
+  case_idx <- meta$DiseaseState %in% c("CRC")
+  ctrl_idx <- meta$DiseaseState %in% c("H")
+
+  # Calculate pvalues, effects, and stderr
+  pvals <- c()
+  teststats <- c()
+  ses <- c()
+  effs <- c()
+  mean_abuns <- c()
+  mean_abuns_present <- c()
+  ubis <- c()
+
+  casedf <- abun_otu[, case_idx]
+  ctrldf <- abun_otu[, ctrl_idx]
+  for (o in rownames(abun_otu)) {
+    # wilcoxon p value  
+    w <- wilcox.test(casedf[o, ], 
+                     ctrldf[o, ])
+    p <- w$p.value
+    teststat <- w$statistic
+
+    pvals <- c(pvals, p)
+    teststats <- c(teststats, teststat)
+
+    # standard error of the OTU abundance, across all samples
+    ses <- c(ses, 
+             sd(abun_otu[o, ])/sqrt(length(abun_otu[o, ]))
+             )
+
+    # mean OTU abundance across all samples (after removing pseudo-count)
+    mean_abuns <- c(mean_abuns, 
+                    mean(abun_otu[o, ] - zeroabun)
+                    )
+
+    # mean OTU abundance across only samples with the OTU present
+    mean_abuns_present <- c(mean_abuns_present, 
+                            sum(abun_otu[o, ] - zeroabun) / sum(abun_otu[o, ] > zeroabun)
+                            )
+
+    # ubiquity of OTU across all samples
+    ubis <- c(ubis, 
+              sum(abun_otu[o, ] > zeroabun) / length(abun_otu[o, ])
+              )
+
+    # effect (logfold difference)
+    effs <- c(effs, 
+              log(mean(abun_otu[o, ctrl_idx])/mean(abun_otu[o, case_idx]))
+              )
+
+  }
+
+  res <- data.frame(otu = rownames(abun_otu), 
+                    pval = pvals, test_statistic = teststats,
+                    SE = ses, effect_size = effs, 
+                    mean_abun = mean_abuns, mean_abun_present = mean_abuns_present,
+                    ubiquity = ubis)
+  save(res, file=resfile)
+
+}else{
+  load(resfile)
+}
+head(res)
+```
+
+Finally, let's try to add phylogeny as covariates. Here we'll have columns for each separate taxonomic level.
+
+```{r, baxter-addphylo}
+res <- res %>% separate(otu, 
+                        c("kingdom", "phylum", "class", "order", "family", "genus", "species", "denovo"), 
+                        sep=";", remove = FALSE)
+
+```
+
+#### Check Covariate Diagnostics
+
+Here we look to see if the covariates do indeed look informative.
+
+##### Ubiquity
+
+```{r, baxter-ubi, fig.width=10, fig.height=3.5}
+strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
+rank_scatter(res, pvalue="pval", covariate="ubiquity")
+```
+
+Something weird is happening with p=0.20 and p=0.40 - maybe this has something to do with ties?
+Either way, ubiquity looks to be a bit informative (from the scatter plot, but not really the histograms...)
+
+##### Mean abundance (across non-zero samples)
+
+```{r, baxter-abun, fig.width=10, fig.height=3.5}
+strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
+rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
+```
+
+##### Phylogeny?!
+
+Let's look at phylum-level stratification first. A priori, I *might* expect
+Proteobacteria to be enriched for low p-values? But I don't know if that's
+super legit, and Eric doesn't seem to think that phylogeny will be informative at all...
+
+```{r, baxter-phylo, fig.width=10, fig.height=6}
+#strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=20)
+#rank_scatter(res, pvalue="pval", covariate="ubiquity")
+ggplot(res, aes(x=pval)) + geom_histogram() + facet_wrap(~phylum, scales = "free")
+```
+
+Not really informative either...
+
+Let's use `ubiquity` as our `ind_covariate`
+
+```{r, baxter-labelcovariate}
+res <- res %>% mutate(ind_covariate = ubiquity)
+```
+
+#### Set up BenchDesign object
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We'll do this for a few
+values of nmids, to explore its effects on the outputs.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, baxter-sb, results="hide", message=FALSE, eval=FALSE}
+for (nmids in c(5, 50, 150)){
+  print(nmids)
+  bd <- initializeBenchDesign(nmids=nmids)
+
+  resfile <- paste0(resdir, "baxter_summarizedBenchmark_",
+                    nrow(res), "_nmids_", nmids, ".RData")
+  duration <- NA
+  if (!file.exists(resfile)){
+    t1 <- proc.time()
+    sb <- bd %>% buildBench(data=res, ftCols = "ind_covariate")
+                            #parallel=TRUE, BPPARAM=multicoreParam)
+    metadata(sb)$data_download_link <- "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
+    save(sb, file=resfile)
+    duration <- round((proc.time()-t1)[3]/60,1)
+  }else{
+    load(resfile)
+  }
+}
+
+```
+
+Just for future runs, where I load the file instead of re-calculating it, here are the errors I got:
+
+```
+[1] 5
+Squarem-1
+Objective fn: 944053
+Objective fn: 111059  Extrapolation: 0  Steplength: 1
+Objective fn: 92567.2  Extrapolation: 1  Steplength: 2.85015
+Objective fn: 91136.8  Extrapolation: 1  Steplength: 2.70993
+Objective fn: 90930.1  Extrapolation: 1  Steplength: 4
+.
+.
+.
+Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !NaNs produced<simpleError in if (sum(D^2) > qchisq(0.9, nmids - 2 - df)) {    warning(paste0("f(z) misfit = ", round(D, 1), ".
+Rerun with increased df."))}: missing value where TRUE/FALSE needed>
+
+[1] 50
+Squarem-1
+Objective fn: 944053
+Objective fn: 111059  Extrapolation: 0  Steplength: 1
+Objective fn: 92567.2  Extrapolation: 1  Steplength: 2.85015
+Objective fn: 91136.8  Extrapolation: 1  Steplength: 2.70993
+Objective fn: 90930.1  Extrapolation: 1  Steplength: 4
+.
+.
+.
+Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !f(z) misfit = -33.4.
+Rerun with increased df.f(z) misfit = -40.5.
+Rerun with increased df.f(z) misfit = 207.5.
+Rerun with increased df.f(z) misfit = -33.
+Rerun with increased df.f(z) misfit = -13.
+Rerun with increased df.f(z) misfit = -35.
+Rerun with increased df.f(z) misfit = -44.7.
+Rerun with increased df.f(z) misfit = -38.4.
+Rerun with increased df.f(z) misfit = -35.2.
+Rerun with increased df.f(z) misfit = -59.5.
+Rerun with increased df.f(z) misfit = 275.
+Rerun with increased df.f(z) misfit = -30.1.
+Rerun with increased df.f(z) misfit = -74.7.
+Rerun with increased df.f(z) misfit = -87.9.
+Rerun with increased df.f(z) misfit = -85.1.
+Rerun with increased df.f(z) misfit = 268.2.
+Rerun with increased df.f(z) misfit = -56.9.
+Rerun with increased df.f(z) misfit = -39.4.
+Rerun with increased df.f(z) misfit = -8.8.
+Rerun with increased df.f(z) misfit = -35.6.
+Rerun with increased df.f(z) misfit = -30.2.
+Rerun with increased df.f(z) misfit = -14.3.
+Rerun with increased ...
+<truncated>
+
+[1] 150
+Squarem-1
+Objective fn: 944053
+Objective fn: 111059  Extrapolation: 0  Steplength: 1
+Objective fn: 92567.2  Extrapolation: 1  Steplength: 2.85015
+Objective fn: 91136.8  Extrapolation: 1  Steplength: 2.70993
+Objective fn: 90930.1  Extrapolation: 1  Steplength: 4
+Objective fn: 90882.9  Extrapolation: 0  Steplength: 1
+.
+.
+.
+Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !f(z) misfit = -15.9.
+Rerun with increased df.f(z) misfit = 8.3.
+Rerun with increased df.f(z) misfit = -10.8.
+Rerun with increased df.f(z) misfit = -18.1.
+Rerun with increased df.f(z) misfit = -24.6.
+Rerun with increased df.f(z) misfit = -27.1.
+Rerun with increased df.f(z) misfit = -24.6.
+Rerun with increased df.f(z) misfit = -18.8.
+Rerun with increased df.f(z) misfit = -21.4.
+Rerun with increased df.f(z) misfit = 279.5.
+Rerun with increased df.f(z) misfit = 83.1.
+Rerun with increased df.f(z) misfit = -27.4.
+Rerun with increased df.f(z) misfit = -28.3.
+Rerun with increased df.f(z) misfit = -7.4.
+Rerun with increased df.f(z) misfit = 11.6.
+Rerun with increased df.f(z) misfit = -18.2.
+Rerun with increased df.f(z) misfit = -23.
+Rerun with increased df.f(z) misfit = -16.7.
+Rerun with increased df.f(z) misfit = -24.1.
+Rerun with increased df.f(z) misfit = -23.2.
+Rerun with increased df.f(z) misfit = -24.4.
+Rerun with increased df.f(z) misfit = -26.6.
+Rerun with increase...
+<truncated>>
+```
+
+```{r, echo=FALSE, eval=FALSE}
+if(!is.na(duration)){
+  message("This step took ", duration, " minutes for ", nrow(res), " OTUs using ",
+        cores, " cores.")
+}
+```
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, baxter-metrics, eval=FALSE}
+# rename assay to qvalue
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+Debugging note: to look at the matrix of qvalues, run `assays(sb)[["qvalue"]]`
+
+```{r, baxter-plot, results="hide", width=15, height=15, eval=FALSE}
+# don't plot scott theoretical for now
+rmv.scott <- which(grepl("scott", colnames(sb)))
+
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
+
+rejection_scatter_bins(sb, covariate="ind_covariate", bins=4, supplementary=FALSE)
+
+# upset plot (only plot IHW for the plotted alpha)
+rmv.ihw <- which(grepl("ihw", colnames(sb)) &
+                              colData(sb)$param.alpha != 0.05)
+rmv.cols <- c(rmv.ihw, rmv.scott)
+
+plotMethodsOverlap(sb[,-rmv.cols], 
+                   alpha=0.05, nsets=ncol(sb)-length(rmv.cols),
+                   order.by="freq", decreasing=TRUE)
+
+# upset plot excluding "unadjusted"
+plotMethodsOverlap(sb[,-c(1,rmv.cols)], 
+                   alpha=0.05, nsets=ncol(sb)-length(rmv.cols)-1,
+                   order.by="freq", decreasing=TRUE)
+```
+
+```{r, eval=FALSE}
+assays(sb)[["qvalue"]][, "ashs"] %>% max()
+sum(is.na(assays(sb)[["qvalue"]][, "scott-empirical"]))
+sum(is.na(assays(sb)[["qvalue"]][, "lfdr"]))
+
+```

### 2017-12-14 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
First success at efficient LD pruning

@@
 ```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
-# project directory & data/results folders
+# project directory & data/results folders - the datdir needs to match where
+#  files are downloaded with the first few bash chunks
 #setwd("/n/irizarryfs01_backed_up/kkorthauer/FDR")
-datdir <- "/n/irizarryfs01_backed_up/kkorthauer/FDR/DATA/GWAS/"
-resdir <- "/n/irizarryfs01_backed_up/kkorthauer/FDR/RESULTS/GWAS/"
+datdir <- "DATA/GWAS/"
+resdir <- "RESULTS/GWAS/"
 
@@
 library(readxl)
+library(readr)
 library(dplyr)
@@
 ```{bash, GWAS1-download, eval=FALSE}
-cd /n/irizarryfs01_backed_up/kkorthauer/FDR/DATA/GWAS/
-curl -O "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
+curl -o DATA/GWAS "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
 ```
@@
 ```{bash, GWAS1-unzip, eval=FALSE}
-unzip BMI.SNPadjSMK.zip
-rm BMI.SNPadjSMK.zip
-rm BMI.SNPadjSMK.Women.EuropeanOnly.txt
-rm BMI.SNPadjSMK.Women.AllAncestry.txt
-rm BMI.SNPadjSMK.Men.EuropeanOnly.txt
-rm BMI.SNPadjSMK.Men.AllAncestry.txt
-rm BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt
+unzip DATA/GWAS/BMI.SNPadjSMK.zip
+rm DATA/GWAS/BMI.SNPadjSMK.zip
+rm DATA/GWAS/BMI.SNPadjSMK.Women.EuropeanOnly.txt
+rm DATA/GWAS/BMI.SNPadjSMK.Women.AllAncestry.txt
+rm DATA/GWAS/BMI.SNPadjSMK.Men.EuropeanOnly.txt
+rm DATA/GWAS/BMI.SNPadjSMK.Men.AllAncestry.txt
+rm DATA/GWAS/BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt
+```
+
+We also download the 1000 genomes CEU reference population data, which will be used to 
+determine which SNPs are LD buddies for the purposes of obtaining an independent set of SNPs.
+
+```{bash, GWAS1-refdata, eval=FALSE}
+# download CEU 1000 genomes SNP reference data 
+wget 'http://neurogenetics.qimrberghofer.edu.au/iSECA/1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01.zip' -P DATA/GWAS/
+
+# unzip reference data
+unzip DATA/GWAS/1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01.zip -d DATA/GWAS
 ```
@@
 for `r nrow(bmi)` SNPs.
+
+#### Prune SNPs that are in Linkage Disequilibrium
+
+Since nearby SNPs in linkage disequilibrium may not constitute separate discoveries, we carry out 
+a pruning step. This commonly done by examining correlation of genotypes among nearby markers, but
+since we don't have the genotype data, we'll use the linkage disequilibrium results of common SNPs and 
+previously observed LD blocks. Specifically, we'll use the phase 3 data from 1000 genomes, restricted
+to the CEU population (since this most closely matches our population of European ancestry). This 
+reference data was downloaded in a previous code chunk. Here we read in the list of more than
+9 million SNPs with data in 1000 genomes and create a list of SNPs that overlap our set. We'll write 
+this list to a file that matches PLINK format.
+
+```{r, GWAS1-snplist, eval=TRUE}
+# load reference data 
+onekg <- fread(paste0(datdir, "1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01.bim"))$V2
+
+# construct input list of SNPs that overlap - 97.5% overlap
+bmi$BP <- as.numeric(unlist(lapply(strsplit(bmi$markername, ":"), function(x) x[[2]])))
+write_delim(bmi %>% dplyr::rename(CHR=chromosome, SNP=rs_id, A1=allele_1, A2=allele_2) %>%
+              mutate(P=0.5, F_A=Freq_Allele1_HapMapCEU, F_U=F_A, CHISQ=1, OR=1) %>% 
+              select(CHR, SNP, BP, A1, F_A, F_U, A2, CHISQ, P, OR) %>%
+              filter(SNP %in% onekg), 
+            path=paste0(datdir, "rslist.assoc"), delim="\t")
+
+ol <- sum(bmi$rs_id %in% onekg)/nrow(bmi)
+rm(onekg)
+```
+
+It turns out that `r signif(ol*100, 3)` percent of SNPs in our set are included in the 1000 genomes 
+reference data. Next we'll use the PLINK software to read in our list of SNPs and the 1000 genomes
+linkage disequilibrium data to 'clump' the SNPs into independent sets of SNPs that have LD correlation (r squared)
+less than 0.2 with all neighbors within 250 kilobases of distance. We write the results to a file called `plink_clump.clumped`. We have to jump out of R here since PLINK is a command line tool.
+
+```{bash, clump, eval=FALSE}
+# use PLINK clumping tool to obtain a set of independent SNPs by LD
+module load plink
+
+plink \
+--bfile DATA/GWAS/1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01 \
+--clump DATA/GWAS/rslist.assoc \
+--clump-field P \
+--clump-p1 0.9999 \
+--clump-p2 0.9999 \
+--clump-r2 0.2 \
+--clump-kb 250 \
+--out DATA/GWAS/plink_clump
+
+```
+
+Next, we'll read in the clumped results and subset our data by our list of independent SNPs.
+
+```{r, GWAS1-prune}
+clump <- fread(paste0(datdir, "plink_clump.clumped"))
+
+# subset to include only the SNPs in the ld.list
+bmi <- bmi[bmi$rs_id %in% clump$SNP,]
+
+rm(clump)
+```
+
+After pruning, we are left with for `r nrow(bmi)` approximately independent SNPs.
 
@@
                       supplementary=FALSE)
+
+# properties of method-exclusive discoveries
+methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
+mcols(sb)$ind_covariate <- mcols(sb)$ind_covar_N
+plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=5)
+
 ```
@@
 supplementary=FALSE)
+
+# properties of method-exclusive discoveries
+methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
+mcols(sb)$ind_covariate <- abs(mcols(sb)$ind_covar_AF-0.5)
+plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=5)
+
 ```
@@
 rm(sb)
-rm(pf)
 ```

### 2017-12-14 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
First success at efficient LD pruning

@@
 ```{r, GWAS1-diag, fig.width=10, fig.height=3.5}
-strat_hist(bmi, pvalue="pval", covariate="ind_covar_N", maxy=3)
+strat_hist(bmi, pvalue="pval", covariate="ind_covar_N", maxy=2)
 rank_scatter(bmi, pvalue="pval", covariate="ind_covar_N")
@@
 ```{r, GWAS1-diag2, fig.width=10, fig.height=3.5}
-strat_hist(bmi, pvalue="pval", covariate="ind_covar_AF", maxy=3)
+strat_hist(bmi, pvalue="pval", covariate="ind_covar_AF", maxy=2)
 rank_scatter(bmi, pvalue="pval", covariate="ind_covar_AF")
@@
 mcols(sb)$ind_covariate <- abs(mcols(sb)$ind_covar_AF-0.5)
-plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=5)
+plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods, maxNum=5)
 

### 2017-12-18 datasets/microbiome/microbiome-SummarizedBenchmark.Rmd
Remove ash from being calculated

@@
 
-```{r, schubert-readdata}
+```{r, schubert-readdata, eval=TRUE}
 otu_path = paste0(datdir, "cdi_schubert_results/RDP/cdi_schubert.otu_table.100.denovo.rdp_assigned")
@@
 
-```{r, schubert-cleandata}
+```{r, schubert-cleandata, eval=TRUE}
 remove_shallow_smpls <- function(df, n_reads) {
@@
 #minabun <- min(abun_otu[abun_otu > 0]) # Use this to figure out what pseudo-count value to add
-zeroabun <- 1e-6
+zeroabun <- 0
 abun_otu <- abun_otu + zeroabun
@@
 
-```{r, schubert-pvals}
+```{r, schubert-pvals, eval=TRUE}
 resfile <- paste0(resdir, "schubert_results_",
@@
   res <- data.frame(otu = rownames(abun_otu), 
-                    pval = pvals, test_statistic = teststats,
+                    pval = pvals, wilcox_teststat = teststats,
                     SE = ses, effect_size = effs, 
@@
                     ubiquity = ubis)
+  res <- res %>%
+    mutate(test_statistic = qnorm(exp(log(pval)-log(2)), lower.tail=FALSE) * sign(effect_size))
+
   save(res, file=resfile)
@@
 
-```{r, schubert-addphylo}
+```{r, schubert-addphylo, eval=TRUE}
 res <- res %>% separate(otu, 
@@
 
-```{r, schubert-ubi, fig.width=10, fig.height=3.5}
+```{r, schubert-ubi, fig.width=10, fig.height=3.5, eval=TRUE}
 strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=20, binwidth=0.05)
@@
 
-```{r, schubert-abun, fig.width=10, fig.height=3.5}
+```{r, schubert-abun, fig.width=10, fig.height=3.5, eval=TRUE}
 strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
@@
 
-```{r, schubert-phylo, fig.width=10, fig.height=6}
+```{r, schubert-phylo, fig.width=10, fig.height=6, eval=TRUE}
 #strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=20)
@@
 
-```{r, schubert-labelcovariate}
+```{r, schubert-labelcovariate, eval=TRUE}
 res <- res %>% mutate(ind_covariate = ubiquity)
+```
+
+Also (hacky), remove test_statistic so that ash doesn't run.
+
+```{r, schubert-rm-ash, eval=TRUE}
+res <- res %>% 
+  mutate(test_statistic = NA) %>%
+  mutate(effect_size = NA)
 ```
@@
 
-```{r, schubert-sb, results="hide", message=FALSE, eval=FALSE}
+```{r, schubert-sb, results="hide", message=FALSE, eval=TRUE}
 for (nmids in c(5, 50, 150)){
@@
 
-```{r, schubert-metrics, eval=FALSE}
+```{r, schubert-metrics, eval=TRUE}
 # rename assay to qvalue
@@
 
-```{r, schubert-plot, results="hide", width=15, height=15, eval=FALSE}
-# don't plot scott theoretical for now
-rmv.scott <- which(grepl("scott", colnames(sb)))
+```{r, schubert-plot, width=15, height=15, eval=TRUE}
 
 # plot nrejects by method overall and stratified by covariate
-rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
-
-rejection_scatter_bins(sb, covariate="ind_covariate", bins=4, supplementary=FALSE)
-
-# upset plot (only plot IHW for the plotted alpha)
-rmv.ihw <- which(grepl("ihw", colnames(sb)) &
-                              colData(sb)$param.alpha != 0.05)
-rmv.cols <- c(rmv.ihw, rmv.scott)
-
-plotMethodsOverlap(sb[,-rmv.cols], 
-                   alpha=0.05, nsets=ncol(sb)-length(rmv.cols),
-                   order.by="freq", decreasing=TRUE)
-
-# upset plot excluding "unadjusted"
-plotMethodsOverlap(sb[,-c(1,rmv.cols)], 
-                   alpha=0.05, nsets=ncol(sb)-length(rmv.cols)-1,
-                   order.by="freq", decreasing=TRUE)
+rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+
 ```
@@
 ```{r, eval=FALSE}
-assays(sb)[["qvalue"]][ , 15] %>% max()
-sum(is.na(assays(sb)[["qvalue"]][ , 22]))
-sum(is.na(assays(sb)[["qvalue"]][ , 20]))
-```
-
-Rather than de-bug this dataset (which might be weird), let's try instead to work with the obesity dataset and see if I get the same problems.
+assays(sb)[["qvalue"]]["ashs"] %>% max()
+sum(is.na(assays(sb)[["qvalue"]]["lfdr"]))
+sum(is.na(assays(sb)[["qvalue"]]["scott-empirical"]))
+sum(is.na(assays(sb)[["qvalue"]]["scott-theoretical"]))
+```
+
+Plotting methods are giving errors for some reason. Let's try to use Alejandro's code instead.
 
@@
 
-```{r, schubert-cleanup}
+```{r, schubert-cleanup, eval=FALSE}
 rm(res)
@@
 
+Also (hacky), remove test_statistic so that ash doesn't run.
+
+```{r, goodrich-rm-ash}
+res <- res %>% 
+  mutate(test_statistic = NA) %>%
+  mutate(effect_size = NA)
+```
+
 #### Set up BenchDesign object
@@
 
-```{r, goodrich-sb, results="hide", message=FALSE, eval=FALSE}
+```{r, goodrich-sb, results="hide", message=FALSE, eval=TRUE}
 for (nmids in c(5, 50, 150)){
@@
 
-```{r, goodrich-metrics, eval=FALSE}
+```{r, goodrich-metrics, eval=TRUE}
 # rename assay to qvalue
@@
 
-```{r, goodrich-plot, results="hide", width=15, height=15, eval=FALSE}
-
+```{r, goodrich-plot, width=15, height=15, eval=TRUE}
 # plot nrejects by method overall and stratified by covariate
-rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
-
-rejection_scatter_bins(sb, covariate="ind_covariate", bins=4, supplementary=FALSE)
-
-# upset plot (only plot IHW for the plotted alpha)
-rmv.ihw <- which(grepl("ihw", colnames(sb)) &
-                              colData(sb)$param.alpha != 0.05)
-
-plotMethodsOverlap(sb[,-rmv.ihw], 
-                   alpha=0.05, nsets=ncol(sb)-length(rmv.ihw),
-                   order.by="freq", decreasing=TRUE)
-
-# upset plot excluding "unadjusted"
-plotMethodsOverlap(sb[,-c(1,rmv.ihw)], 
-                   alpha=0.05, nsets=ncol(sb)-length(rmv.ihw)-1,
-                   order.by="freq", decreasing=TRUE)
-```
-
-Hm, looks like all of the methods run without giving NaN's with nmids = 150 or lower...
-But the results still aren't right.
+rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+#plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE )
+```
+
+There is no upsetR plot for OB Goodrich for now because only lfdr is returning any rejections.
 
@@
 #minabun <- min(abun_otu[abun_otu > 0]) # Use this to figure out what pseudo-count value to add
-zeroabun <- 1e-7
+zeroabun <- 0
 abun_otu <- abun_otu + zeroabun
@@
 
+Also (hacky), remove test_statistic so that ash doesn't run.
+
+```{r, baxter-rm-ash}
+res <- res %>% 
+  mutate(test_statistic = NA) %>%
+  mutate(effect_size = NA)
+```
+
 #### Set up BenchDesign object
@@
 
-```{r, baxter-sb, results="hide", message=FALSE, eval=FALSE}
+```{r, baxter-sb, results="hide", message=FALSE, eval=TRUE}
 for (nmids in c(5, 50, 150)){
@@
 
-```{r, baxter-metrics, eval=FALSE}
+```{r, baxter-metrics, eval=TRUE}
 # rename assay to qvalue
@@
 
-```{r, baxter-plot, results="hide", width=15, height=15, eval=FALSE}
-# don't plot scott theoretical for now
-rmv.scott <- which(grepl("scott", colnames(sb)))
-
+```{r, baxter-plot, width=15, height=15, eval=TRUE}
 # plot nrejects by method overall and stratified by covariate
-rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
-
-rejection_scatter_bins(sb, covariate="ind_covariate", bins=4, supplementary=FALSE)
-
-# upset plot (only plot IHW for the plotted alpha)
-rmv.ihw <- which(grepl("ihw", colnames(sb)) &
-                              colData(sb)$param.alpha != 0.05)
-rmv.cols <- c(rmv.ihw, rmv.scott)
-
-plotMethodsOverlap(sb[,-rmv.cols], 
-                   alpha=0.05, nsets=ncol(sb)-length(rmv.cols),
-                   order.by="freq", decreasing=TRUE)
-
-# upset plot excluding "unadjusted"
-plotMethodsOverlap(sb[,-c(1,rmv.cols)], 
-                   alpha=0.05, nsets=ncol(sb)-length(rmv.cols)-1,
-                   order.by="freq", decreasing=TRUE)
+rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+#plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
 ```

### 2017-12-18 datasets/RNA-Seq/rna-seq-benchmarkFDR.Rmd
filter geneExp object to solve the weird hill dist

@@
 library(SummarizedBenchmark)
-sourceDirectory("~/Downloads/benchmark-fdr/datasets/R/")
+sourceDirectory("~/Documents/Research/benchmark-fdr/datasets/R/")
 
@@
 #Load expression data
-load('~/Downloads/dsdObjects.RData')
+load('~/Documents/Research/benchmark-fdr/datasets/RNA-Seq/dsdObjects.RData')
 
@@
 #Buid DESeq object and run DESeq2
-dds <- DESeq(filteredDsd1, parallel = T)
+dds <- DESeq(filteredDsd1, parallel = T) # This step can take a few minutes since we have many samples
 res <- results(dds, independentFiltering = F)
@@
 #Important to not change col names since SB is strict about that
-geneExp <- data.frame(geneName=rownames(res), pval=res$pvalue, SE=res$lfcSE, 
-                      ind_covariate = res$baseMean, effect_size=res$log2FoldChange, test_statistic=res$stat)
+geneExp <- tbl_df(data.frame(geneName=rownames(res), pval=res$pvalue, SE=res$lfcSE, 
+                      ind_covariate = res$baseMean, effect_size=res$log2FoldChange, test_statistic=res$stat))
+
+geneExp <- geneExp %>% na.omit() %>% dplyr::filter(ind_covariate>=1)
 
@@
 ```{r}
-strat_hist(geneExp, pvalue="pval", covariate="ind_covariate", maxy=10)
+strat_hist(geneExp, pvalue="pval", covariate="ind_covariate", maxy=25)
 ```

### 2017-12-19 datasets/ChIP-seq/ChIP-seq.Rmd
ChIP-seq update

@@
 author: "Mingxiang Teng"
-date: "9/14/2017"
+date: '`r format(Sys.Date(), "%B %d, %Y")`'
 output: html_document
@@
 knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_knit$set(root.dir = "~/workspace/cowork/fdr/")
 ```
@@
 For FDR benchmark using ChIP-seq datasets, we may focus on peak calling and
-differential peaks. Peak calling is the most common way that people analyze 
-ChIP-seq since this technique was invented, though some groups started to care
-about differential peaks in some complex experiment designs. For peak calling,
-peak callers always report effect size (they call "enrichment signal") 
-and p-value, but standard error. I think it is too complicate if we dig into
-their code for generating standard error. For differential peaks, classical
-pipeline would call peaks first, then recruit existing algorithms (*e.g.* DESeq2)
-to test the covariant of interest, in a way that all effect size, p-value and 
-standard error can be generated. However, this pipeline probably gives 
-very similar patterns (distributions) of effect size / p-value to those
-based on differential gene expression analysis. Thus, differential peak 
-analysis might not be useful in our FDR benchmark.
+differential peaks. Peak calling has been the most common procedure to analyze 
+ChIP-seq since this technique was invented, though people started to look
+into differential peaks in some complex experiment designs. In peak calling,
+callers always report effect size (they call "enrichment signal") 
+and p-value, but standard error. It is complicate to hack the original
+code for re-generating standard errors. In differential peaks, classical
+pipelines firstly call peaks, and recruit existing algorithms (*i.e.* edgeR/DESeq2)
+to test the covariant of interest later, in a way that the effect size, p-value and 
+standard error can be generated at the same time.
 
-### Oestrogen receptor binding case
+Here, we focus on test the differential peaks. It is noted that by using DESeq2
+to test the differential peaks, we have introduced similarity into testing results to that
+we used for RNA-seq differential analysis. Nevertheless, benchmarking on ChIP-seq still provides
+insights with a unique data structure which has been widely utilized in modern genomics.
 
-This dataset have plenty of replicates so that we can focus on both peak
-calling and differential peaks. And it is the first dataset people used to 
-illustrate differential peak detection, where the tools DiffBind was 
-introduced.
+### Testing differential ChIP-seq signals between two cell lines (K562 *vs.* Gm12878)
+These two cell lines hold the largest sequencing datasets on ENCODE project. 
+We chose histone H3K4Me3 ChIP-seq data for our analysis here.
 
-We can download the processed data from GEO.
+#### Data Download
 
-```{bash, chipseq-download, eval=FALSE}
-curl -O "https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE32222&format=file"
-tar -xvf "?acc=GSE32222&format=file"
-rm "?acc=GSE32222&format=file"
+We download the bam files directly from UCSC ENCODE protal.
+
+```{bash, encode-download, eval=FALSE}
+wget -N -P data http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeBroadHistone/wgEncodeBroadHistoneGm12878H3k4me3StdAlnRep1.bam
+wget -N -P data http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeBroadHistone/wgEncodeBroadHistoneGm12878H3k4me3StdAlnRep1.bam.bai
+wget -N -P data http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeBroadHistone/wgEncodeBroadHistoneGm12878H3k04me3StdAlnRep2V2.bam
+wget -N -P data http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeBroadHistone/wgEncodeBroadHistoneGm12878H3k04me3StdAlnRep2V2.bam.bai
+wget -N -P data http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeBroadHistone/wgEncodeBroadHistoneK562H3k4me3StdAlnRep1.bam
+wget -N -P data http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeBroadHistone/wgEncodeBroadHistoneK562H3k4me3StdAlnRep1.bam.bai
+wget -N -P data http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeBroadHistone/wgEncodeBroadHistoneK562H3k4me3StdAlnRep2.bam
+wget -N -P data http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeBroadHistone/wgEncodeBroadHistoneK562H3k4me3StdAlnRep2.bam.bai
+wget -N -P data http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwHistone/wgEncodeUwHistoneGm12878H3k4me3StdAlnRep1.bam
+wget -N -P data http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwHistone/wgEncodeUwHistoneGm12878H3k4me3StdAlnRep1.bam.bai
+wget -N -P data http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwHistone/wgEncodeUwHistoneGm12878H3k4me3StdAlnRep2.bam
+wget -N -P data http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwHistone/wgEncodeUwHistoneGm12878H3k4me3StdAlnRep2.bam.bai
+wget -N -P data http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwHistone/wgEncodeUwHistoneK562H3k4me3StdAlnRep1.bam
+wget -N -P data http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwHistone/wgEncodeUwHistoneK562H3k4me3StdAlnRep1.bam.bai
+wget -N -P data http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwHistone/wgEncodeUwHistoneK562H3k4me3StdAlnRep2.bam
+wget -N -P data http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwHistone/wgEncodeUwHistoneK562H3k4me3StdAlnRep2.bam.bai
 ```
 
-For one sample, extract the peak calling results.
+#### Workspace Setup
 
-```{bash, chipseq-unzip1, eval=FALSE}
-gunzip "GSM798383_SLX-1201.250.s_4_SLX-1202.250.s_1_peaks.txt.gz"
-rm *peaks.txt.gz
+```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
+datdir <- "data"
+resdir <- "results"
+# wrangling & plotting tools
+library(data.table)
+library(readxl)
+library(readr)
+library(dplyr)
+library(ggplot2)
+library(magrittr)
+library(R.utils)
+library(cowplot)
+# benchmark methods
+library(IHW)
+library(ashr)
+library(qvalue)
+library(swfdr)
+library(fdrtool)
+library(FDRreg)
+# comparison tools/functions
+library(SummarizedBenchmark)
+library(R.utils)
+sourceDirectory("benchmark-fdr/datasets/R/")
+# set up parallel backend
+library(BiocParallel)
+cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
+multicoreParam <- MulticoreParam(workers = cores)
+library(GenomicRanges)
+library(TxDb.Hsapiens.UCSC.hg19.knownGene)
+prom <- promoters(genes(TxDb.Hsapiens.UCSC.hg19.knownGene))
+seqlevels(prom,force=T) <- seqlevels(prom)[!grepl("_", seqlevels(prom))]
+bamfiles <- list.files(datdir,pattern='bam$',full=T)
+labs <- rep(c('Broad','UW'),each=4)
+cells <- rep(rep(c('GM12878','K562'),each=2),2)
+meta <- data.frame(cellline=cells,lab=labs,file=bamfiles)
 ```
 
-And load in to R.
-```{r, chipseq-read1}
-peaks <- read.table("~/GSM798383_SLX-1201.250.s_4_SLX-1202.250.s_1_peaks.txt", header=TRUE, sep='\t')
-dim(peaks)
-head(peaks)
+#### Generating Benchmarking Inputs
+
+```{r count, results='hide', message=FALSE, warning=FALSE}
+if(file.exists('h3k4me3.rda')){
+    load('h3k4me3.rda')
+}else{
+    library(Rsubread)	
+    anno <- data.frame(GeneID=seq_len(length(prom)),Chr=seqnames(prom), Start=start(prom),
+                   End=end(prom), Strand=strand(prom))
+    rc <- featureCounts(files=bamfiles,annot.ext=anno,allowMultiOverlap=TRUE,minOverlap=50,readExtension3=150,
+                    ignoreDup=TRUE)$counts
+    library(DESeq2)
+    dds <- DESeqDataSetFromMatrix(countData=rc,colData=as.data.frame(meta),design= ~ lab + cellline)
+    dds <- DESeq(dds,fitType="mean")
+    dat <- as.data.frame(results(dds))
+    colnames(dat) <- c('ind_covariate','effect_size','SE','stat','pval','padj')
+    save(rc,prom,deseq,dat,file='h3k4me3.rda')
+}
+dim(dat)
+head(dat)
+dat <- dat[!is.na(dat[,2]),]
+```
+#### Check Covariate Diagnostics
+
+```{r, chipseq-diag, fig.width=10, fig.height=3.5}
+strat_hist(dat, pvalue="pval", covariate="ind_covariate", maxy=2)
+rank_scatter(dat, pvalue="pval", covariate="ind_covariate")
 ```
 
-Here, "X.10.log10.pvalue." is the -log10P, and fold_enrichment is the effect size.
+#### Setup BenchDesign object
+
+```{r, chipseq-benchdesign}
+bd <- initializeBenchDesign()
+```
+
+#### Run buildBench for the coverage covariate and examine results
+
+Now, we're ready to construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+                                        helper scripts).
+
+```{r, chipseq-sb, results="hide", message=FALSE}
+resN <- paste0(resdir, "/chipseq_summarizedBenchmark_",
+               nrow(dat), ".RData")
+
+duration <- NA
+if (!file.exists(resN)){
+  t1 <- proc.time()
+  sb <- bd %>% buildBench(data=(dat %>% mutate(test_statistic = qnorm(exp(log(pval)-log(2)), lower.tail=FALSE) * sign(effect_size))), 
+                          ftCols = c("ind_covariate"),
+                          parallel=TRUE, BPPARAM=multicoreParam)
+  save(sb, file=resN)
+  duration <- round((proc.time()-t1)[3]/60,1)
+}else{
+  load(resN)
+}
+```
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, chipseq-metrics}
+# rename assay to qvalue
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, chipseq-plot, results="hide", width=15, height=15}
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter(sb,supplementary=FALSE)
+
+rejection_scatter_bins(sb, covariate="ind_covariate", bins=4,
+                       supplementary=FALSE)
+
+# upset plot 
+plotFDRMethodsOverlap(sb, 
+                      alpha=0.05, nsets=ncol(sb),
+                      order.by="freq", decreasing=TRUE,
+                      supplementary=FALSE)
+
+# properties of method-exclusive discoveries
+methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
+plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=5)
+rm(bmi)
+rm(bd)
+rm(sb)
+
+```

### 2017-12-19 datasets/ChIP-seq/ChIP-seq.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+#SBATCH -J knitChIPseq                 # A single job name for the array
+#SBATCH -n 8                        # Number of cores
+#SBATCH -N 1                        # All cores on one machine
+#SBATCH -p irizarry,serial_requeue  # Partition
+#SBATCH --mem 60000                 # Memory request 
+#SBATCH -t 0-5:00                   # Maximum execution time (D-HH:MM)
+#SBATCH -o render-%j.out  # Standard output
+#SBATCH -e render-%j.err  # Standard error
+ 
+module load rstudio
+export SLURM_NTASKS
+export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/rstudio/0.98.1103-fasrc01/bin/pandoc/"
+
+# change filename to Rmd to be knitted. ncores is passed in through environment
+# variable SLURM_NTASKS
+R -e "rmarkdown::render('~/workspace/cowork/fdr/benchmark-fdr/datasets/ChIP-seq/ChIP-seq.Rmd')"

### 2017-12-20 datasets/ChIP-seq/ChIP-seq.Rmd
ChIP-seq

@@
 ### Testing differential ChIP-seq signals between two cell lines (K562 *vs.* Gm12878)
-These two cell lines hold the largest sequencing datasets on ENCODE project. 
+
+These two cell lines hold the largest sequencing datasets in ENCODE project. 
 We chose histone H3K4Me3 ChIP-seq data for our analysis here.
@@
 
-We download the bam files directly from UCSC ENCODE protal.
+We downloaded the bam files directly from UCSC ENCODE protal.
 
@@
 #### Workspace Setup
+
+Here, we first load in all R-packages needed in the analysis. As 
+mentioned above, classical differential analysis pipeline would 
+involve peak calling as the first step. In that case, differential 
+testing is performed on a filtered set of genomic regions after 
+peak calling. However, this will result in limited null tests since 
+regions are chosen with significance at least in one condition. 
+To address this issue, we selected ChIP-seq data of H3K4Me3 which is 
+an active marker of gene expression. We focus our analysis on 
+promoter regions of all genes for this histone marker. This helps 
+avoid limited null tests with peak calling by not losing
+most interesting regions for this histone marker.
 
@@
 
+Basically, we count sequencing reads for all promoter regions here.
+We chose the average coverage across conditions and replicates as the covariate.
+
 ```{r count, results='hide', message=FALSE, warning=FALSE}
@@
 
+Here, we can see that the covariant variable is actually informative.
+As we know, ChIP-seq data always involves two sets of regions: 
+signal and background. Sometimes, there are a few regions with 
+signal coverage between signal and background. In this example, 
+we see that p-values are differently distributed under low, 
+median and high coverage.
+
 ```{r, chipseq-diag, fig.width=10, fig.height=3.5}
@@
 Now, we're ready to construct the `SummarizedBenchmark` object, which will run
-the functions specified in each method (these are actually sourced in from the
-                                        helper scripts).
+the functions specified in each method.
 

### 2017-12-22 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
restrict max number of methods in boxplot to 3

@@
 # properties of method-exclusive discoveries
-methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
+methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
+             "scott-empirical", "scott-theoretical")
 mcols(sb)$ind_covariate <- mcols(sb)$ind_covar_N
-plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=5)
+plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3)
 
@@
 # properties of method-exclusive discoveries
-methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
+methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
+             "scott-empirical", "scott-theoretical")
 mcols(sb)$ind_covariate <- abs(mcols(sb)$ind_covar_AF-0.5)
-plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods, maxNum=5)
+plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods, maxNum=3)
 

### 2018-01-05 datasets/RNA-Seq/polyester-benchmarkFDR.Rmd
add initial polyester RNA-Seq simulation R Markdown

---
title: "Simulation study using RNA-Seq counts with Polyester"
author: "Rafalab Journal Club Members"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
        code_folding: hide

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages needed to simulate RNA-seq counts with polyester
library(zebrafishRNASeq) 
library(genefilter)
library(limma)
library(polyester)

# Differential testing
library(DESeq2)

# Load supporting packages for data analyses
library(data.table)
library(magrittr)
library(dplyr)
library(ggplot2)
library(R.utils)
library(cowplot)

# benchmark methods
library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)
library(FDRreg)

# comparison tools/functions
library(SummarizedBenchmark)

# load helper functions
working_dir <- list(project_home = normalizePath(getwd()))
sourceDirectory(file.path(working_dir$project_home,"datasets/R/"))

# set up parallel backend
library(BiocParallel)
cores <- 4
multicoreParam <- MulticoreParam(workers = cores)
```

### Simulate RNA-Seq counts using polyester

We use the [polyester](https://bioconductor.org/packages/release/bioc/html/polyester.html) Bioconductor package to simulate RNA-Seq reads. We start with the [zebrafish](https://bioconductor.org/packages/release/data/experiment/html/zebrafishRNASeq.html) Bioconductor data package.

```{r simulate-data}
data(zfGenes)
filter <- apply(zfGenes, 1, function(x) length(x[x>5])>=2)
counts <- zfGenes[filter,]

## Estimate the zero inflated negative binomial parameters
params = get_params(counts)

plot(rowMeans(log(counts+1)), rowVars(log(counts+1)), 
     pch=19, main="zebrafish data w/fit")
lines(params$fit, col = 2)

## Create some data from the model
dat0 = create_read_numbers(params$mu, params$fit, params$p0,
                           m=dim(counts)[1], n=dim(counts)[2])
```

### Build DESeq2 object and run DESeq2

```{r run-deseq2}
group = factor(rep(c(0,1), each=3))
design = model.matrix(~group)

## Buid DESeq object and run DESeq2
dds <- DESeqDataSetFromMatrix(countData=dat0, colData = DataFrame(group), design=~group)
dds <- DESeq(dds) 
res <- results(dds, independentFiltering = F)
```

### Create SummarizedBenchmark object

```{r create-summarizedbenchmark-object}
# Build input data frame for SummarizedBenchmark object
# Important to not change col names since SB is strict about that
geneExp <- tbl_df(data.frame(geneName=rownames(zfGenes)[filter], pval=res$pvalue, SE=res$lfcSE, 
                      ind_covariate = res$baseMean, effect_size=res$log2FoldChange, test_statistic=res$stat))

geneExp <- geneExp %>% na.omit() %>% dplyr::filter(ind_covariate>=1)
```

```{r plot-strat-hist}
strat_hist(geneExp, pvalue="pval", covariate="ind_covariate", maxy=2)
```

```{r plot-rank-scatter}
rank_scatter(geneExp, pvalue="pval", covariate="ind_covariate")
```

```{r}
bd <- initializeBenchDesign()
sb <- bd %>% buildBench(data=geneExp, parallel = T)

assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)

```

```{r}
rejections_scatter(sb, supplementary=FALSE)
```

```{r}
plotFDRMethodsOverlap(sb, alpha=0.05, nsets=ncol(sb), order.by="freq", decreasing=TRUE, supplementary=FALSE)
```

### 2018-01-09 datasets/RNA-Seq/yeast48-benchmarkFDR.Rmd
add null yeast48 comparison

---
title: "Yeast RNA-seq 48 sample comparison"
author: "Keegan Korthauer"
date: "1/9/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction

This is an analysis of the yeast data with 48 biological replicates in each of 
two conditions (analyzed in this [publication](https://www.ncbi.nlm.nih.gov/pubmed/26206307/)). 
We chose this experiment because of the large number of biological replicates, which
will allow us to (1) implement null comparisons on random subsets of samples within
one condition, (2) compare FDR approaches on their ability to discover 'true positive' 
DE genes using subsets of the full dataset, where 'true positive' is defined as a 
gene detected as DE using all 48 samples in each condition, and (3) start with a null
comparison and add in artificial differences to a subset of genes to define
'true positives'.

Processed count table is made available by the authors in their paper codebase 
[GitHub repository](https://github.com/bartongroup/profDGE48).

### Set up workspace

```{r, results='hide'}
# Load packages and source benchmark FDR
library(data.table)
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(magrittr)
library(R.utils)
library(cowplot)
library(purrr)
library(DESeq2)

# benchmark methods
library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)
library(FDRreg)

# comparison tools/functions
library(SummarizedBenchmark)
sourceDirectory("../R/")

# set up parallel backend
library(BiocParallel)
cores <- 4
multicoreParam <- MulticoreParam(workers = cores)
```

### Data download

First, we download the processed count data from GitHub. There is one file for the 
Snf2 condition and one file for the wild type condition. The Snf2 condition is a yeast
strain that has the transcriptional regulator gene Snf2 knocked out. Each of these
files is a compressed `tar.gz` archive that contains a 
single bam file for each replicate in each condition, which we'll place in a subdirectory
called `data`.

```{r, download}
system("wget https://github.com/bartongroup/profDGE48/raw/master/Preprocessed_data/Snf2_countdata.tar.gz")
system("wget https://github.com/bartongroup/profDGE48/raw/master/Preprocessed_data/WT_countdata.tar.gz")
system("mkdir yeast48-data")
system("tar -xvf WT_countdata.tar.gz -C ./yeast48-data")
system("tar -xvf Snf2_countdata.tar.gz -C ./yeast48-data")
system("rm *.tar.gz")
```

Each of the data files contains two columns, one with a gene/feature name, and one
with the count value.

### Read data into R and create a count table

Here we make use of the map and map2 functions in the `purrr` package, to swiftly 
apply the `read_tsv` function from `readr` to read in all of the 96 sample tables, 
as well as add in the sample name (derived from the file name) to each subtable.
Finally, the `reduce` function is used to join all the replicates together.

```{r, readin}
files <- dir(path = "./data/", pattern = "*.bam.gbgout", full.names = TRUE)
sample_names <- sapply(strsplit(dir(path = "./data/", pattern = "*.bam.gbgout"), "_MID"),
                       function(x) x[[1]])

counts <- files %>%
  map(read_tsv, col_names = FALSE) %>% # read in all the files individually
  map2(sample_names, ~ dplyr:::rename(.x, !! .y := X2, feature = X1) ) %>% # add sample names
  purrr::reduce(left_join, by = "feature") # reduce with rbind into one dataframe
```

### Analysis with DESeq2

Now, we're ready to construct a DESeq2 object. First we pull out the feature names and
add them as rownames for the count table, and next we construct a column data object
that houses the sample names, replicate numbers, and condition factor (WT versus
Snf2 knockout).

```{r, deseq}
feats <- (counts %>% select(1))$feature
counts <- as.matrix(counts %>% select(-1))
rownames(counts) <- feats

coldat <- tibble(sample=colnames(counts)) %>% 
  separate(sample, sep="_", into=c("condition", "replicate"), remove=FALSE) %>%
  mutate(condition = factor(condition))

dds <- DESeqDataSetFromMatrix(countData = counts,
                              colData = coldat,
                              design= ~ condition)

# results on full set
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients
resFULL <- results(dds, name="condition_WT_vs_Snf2")

head(resFULL)

sum(resFULL$padj < 0.05, na.rm=T, independentFiltering = F)
```

There are many thousand genes detected using the full set.

### Null Comparison

```{r, null}
# select a random subset of 20 WT samples
set.seed(85873)
dds_null <- dds[,colData(dds)$condition == "WT"]
dds_null <- dds_null[,sample(1:ncol(dds_null), 20)]

# add a fake condition column to coldat
colData(dds_null)$fake <-  factor(c(rep("WT", 10), rep("TRT", 10))[sample(1:20, 20)])
design(dds_null) <- ~fake

# results on full set
dds_null <- DESeq(dds_null)
resultsNames(dds_null) # lists the coefficients
resNULL <- results(dds_null, name="fake_WT_vs_TRT", independentFiltering = F)

head(resNULL)

sum(resNULL$padj < 0.05, na.rm=T)
```

### FDR Benchmark

Build input data frame for summarized benchmark.

```{r}
geneExp <- tbl_df(data.frame(geneName=rownames(resNULL), pval=resNULL$pvalue, SE=resNULL$lfcSE, 
                      ind_covariate = resNULL$baseMean, effect_size=resNULL$log2FoldChange, 
                      test_statistic=resNULL$stat))

geneExp <- geneExp %>% na.omit() %>% dplyr::filter(ind_covariate>=1)

```

Look at covariate diagnostic plots.

```{r}
strat_hist(geneExp, pvalue="pval", covariate="ind_covariate", maxy=2)
```

```{r}
rank_scatter(geneExp, pvalue="pval", covariate="ind_covariate")
```

Run benchmark methods.

```{r}
bd <- initializeBenchDesign()
sb <- bd %>% buildBench(data=geneExp, parallel = T)

assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Plot results.

```{r}
rejections_scatter(sb, supplementary=FALSE)
```

```{r}
plotFDRMethodsOverlap(sb, alpha=0.05, nsets=ncol(sb), order.by="freq", decreasing=TRUE, supplementary=FALSE)
```

### Session information

```{r}
sessionInfo()
```

### 2018-01-10 datasets/RNA-Seq/polyester-benchmarkFDR.Rmd
update polyester RNA-Seq simulations with null and non-null comparisons

@@
 title: "Simulation study using RNA-Seq counts with Polyester"
-author: "Rafalab Journal Club Members"
+author: "Stephanie Hicks"
 date: "`r Sys.Date()`"
@@
         number_sections: true
-        code_folding: hide
 
@@
 knitr::opts_chunk$set(echo = TRUE)
+```
 
+Set up workspace 
+```{r workspace-setup, results='hide', message=FALSE, warning=FALSE}
 # Load packages needed to simulate RNA-seq counts with polyester
@@
 # load helper functions
-working_dir <- list(project_home = normalizePath(getwd()))
+working_dir <- list(project_home = "/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/shicks/projects/benchmark-fdr")
 sourceDirectory(file.path(working_dir$project_home,"datasets/R/"))
@@
 
-### Simulate RNA-Seq counts using polyester
+# Simulate RNA-Seq counts using polyester
 
 We use the [polyester](https://bioconductor.org/packages/release/bioc/html/polyester.html) Bioconductor package to simulate RNA-Seq reads.
-We start with the [zebrafish](https://bioconductor.org/packages/release/data/experiment/html/zebrafishRNASeq.html) Bioconductor data package.
+The package requires a count matrix as input which estimates the mean variance relationship.
+Here we start with the [zebrafish](https://bioconductor.org/packages/release/data/experiment/html/zebrafishRNASeq.html) Bioconductor data package.
 
-```{r simulate-data}
+## Estimate mean and variance relationship
+
+```{r estimate-parameters-NB}
 data(zfGenes)
+dim(zfGenes)
+
 filter <- apply(zfGenes, 1, function(x) length(x[x>5])>=2)
-counts <- zfGenes[filter,]
+counts <- as.matrix(zfGenes[filter,])
 
@@
 params = get_params(counts)
-
-plot(rowMeans(log(counts+1)), rowVars(log(counts+1)), 
-     pch=19, main="zebrafish data w/fit")
-lines(params$fit, col = 2)
-
-## Create some data from the model
-dat0 = create_read_numbers(params$mu, params$fit, params$p0,
-                           m=dim(counts)[1], n=dim(counts)[2])
 ```
 
-### Build DESeq2 object and run DESeq2
+## Simulate null data
 
-```{r run-deseq2}
-group = factor(rep(c(0,1), each=3))
+Create some null data from the NB model. 10 samples total. 
+```{r simulate-null-data}
+dat_null = create_read_numbers(params$mu, params$fit, params$p0,
+                           m=dim(counts)[1], n=10)
+head(dat_null)
+```
+
+## Simulate non-null data
+
+Create some non-null data from NB model. Simulate 2000 differentially expressed genes. 10 samples total. 5 in each group. 
+```{r simulate-alt-data}
+nonnullgroup = rep(c(-1,1),each=5)
+mod = model.matrix(~-1 + nonnullgroup)
+coeffs = cbind(c(rnorm(2000), rep(0,dim(counts)[1]-2000)))
+dat_alt = create_read_numbers(params$mu, params$fit,
+                           params$p0, m=dim(counts)[1], n=10,
+                           beta=coeffs, mod=mod)
+head(dat_alt)
+```
+
+# Differential testing
+
+Next we build the DESeq2 objects and run DESeq2
+
+```{r run-deseq2, result='hide', message=FALSE}
+group = factor(rep(c(0,1), each=5))
 design = model.matrix(~group)
 
-## Buid DESeq object and run DESeq2
-dds <- DESeqDataSetFromMatrix(countData=dat0, colData = DataFrame(group), design=~group)
-dds <- DESeq(dds) 
-res <- results(dds, independentFiltering = F)
+dds_null <- DESeqDataSetFromMatrix(countData=dat_null, colData = DataFrame(group), design=~group)
+dds_null <- DESeq(dds_null) 
+res_null <- results(dds_null, independentFiltering = F)
+
+dds_alt <- DESeqDataSetFromMatrix(countData=dat_alt, colData = DataFrame(group), design=~group)
+dds_alt <- DESeq(dds_alt) 
+res_alt <- results(dds_alt, independentFiltering = F)
 ```
 
-### Create SummarizedBenchmark object
+# Benchmarking
+
+## Create SummarizedBenchmark objects
+
+Create SummarizedBenchmark objects for null and non-null datasets
 
@@
 # Important to not change col names since SB is strict about that
-geneExp <- tbl_df(data.frame(geneName=rownames(zfGenes)[filter], pval=res$pvalue, SE=res$lfcSE, 
-                      ind_covariate = res$baseMean, effect_size=res$log2FoldChange, test_statistic=res$stat))
+gene_exp_null <- tbl_df(data.frame(geneName=rownames(zfGenes)[filter], 
+                                   pval=res_null$pvalue, SE=res_null$lfcSE, 
+                                   ind_covariate = res_null$baseMean, 
+                                   effect_size=res_null$log2FoldChange, 
+                                   test_statistic=res_null$stat))
 
-geneExp <- geneExp %>% na.omit() %>% dplyr::filter(ind_covariate>=1)
-```
+gene_exp_alt <- tbl_df(data.frame(geneName=rownames(zfGenes)[filter], 
+                                   pval=res_alt$pvalue, SE=res_alt$lfcSE, 
+                                   ind_covariate = res_alt$baseMean, 
+                                   effect_size=res_alt$log2FoldChange, 
+                                   test_statistic=res_alt$stat))
 
-```{r plot-strat-hist}
-strat_hist(geneExp, pvalue="pval", covariate="ind_covariate", maxy=2)
-```
-
-```{r plot-rank-scatter}
-rank_scatter(geneExp, pvalue="pval", covariate="ind_covariate")
-```
-
-```{r}
-bd <- initializeBenchDesign()
-sb <- bd %>% buildBench(data=geneExp, parallel = T)
-
-assayNames(sb) <- "qvalue"
-sb <- addDefaultMetrics(sb)
+gene_exp_null <- gene_exp_null %>% na.omit() %>% dplyr::filter(ind_covariate>=1)
+gene_exp_alt  <- gene_exp_alt  %>% na.omit() %>% dplyr::filter(ind_covariate>=1)
 
@@
 
-```{r}
-rejections_scatter(sb, supplementary=FALSE)
+## Plots (sanity checks)
+
+### Stratified histograms
+
+Null data 
+```{r plot-strat-hist-null}
+strat_hist(gene_exp_null, pvalue="pval", covariate="ind_covariate", maxy=5)
 ```
 
+Non-null data with 2000 DE genes
+```{r plot-strat-hist-alt}
+strat_hist(gene_exp_alt, pvalue="pval", covariate="ind_covariate", maxy=5)
+```
+
+### Rank scatter plots
+
+Null data
+```{r plot-rank-scatter-null}
+rank_scatter(gene_exp_null, pvalue="pval", covariate="ind_covariate")
+```
+
+Non-null data with 2000 DE genes
+```{r plot-rank-scatter-alt}
+rank_scatter(gene_exp_alt, pvalue="pval", covariate="ind_covariate")
+```
+
+## Run Benchmark
+
+Null data
+```{r run-benchmark-null}
+bd_null <- initializeBenchDesign()
+sb_null <- bd_null %>% buildBench(data=gene_exp_null, parallel = T)
+
+assayNames(sb_null) <- "qvalue"
+sb_null <- addDefaultMetrics(sb_null)
+```
+
+Non-null data with 2000 DE genes
+```{r run-benchmark-alt}
+bd_alt <- initializeBenchDesign()
+sb_alt <- bd_alt %>% buildBench(data=gene_exp_alt, parallel = T)
+
+assayNames(sb_alt) <- "qvalue"
+sb_alt <- addDefaultMetrics(sb_alt)
+```
+
+# Results
+
+## Number of rejections
+
+Null data
+```{r rejection-scatter-null}
+rejections_scatter(sb_null, supplementary=FALSE)
+```
+
+Non-null data with 2000 DE genes
+```{r rejection-scatter-alt}
+rejections_scatter(sb_alt, supplementary=FALSE)
+```
+
+## Overlaps
+
+Null data
+```{r overlap-null}
+plotFDRMethodsOverlap(sb_null, alpha=0.05, nsets=ncol(sb_null), 
+                      order.by="freq", decreasing=TRUE, supplementary=FALSE)
+```
+
+Non-null data with 2000 DE genes
+```{r overlap-alt}
+plotFDRMethodsOverlap(sb_alt, alpha=0.05, nsets=ncol(sb_null),
+                      order.by="freq", decreasing=TRUE, supplementary=FALSE)
+```
+
+# Session information
+
 ```{r}
-plotFDRMethodsOverlap(sb, alpha=0.05, nsets=ncol(sb), order.by="freq", decreasing=TRUE, supplementary=FALSE)
+sessionInfo()
 ```

### 2018-01-10 datasets/RNA-Seq/polyester-benchmarkFDR.sh

@@ -1,0 +1,12 @@
+#!/bin/bash
+
+#SBATCH -n 4 # Number of cores
+#SBATCH -t 0-01:00 # Runtime in D-HH:MM
+#SBATCH -p irizarry,serial_requeue
+#SBATCH -e polyester-benchmarkFDR.err
+#SBATCH -o polyester-benchmarkFDR.out
+#SBATCH --mem=20GB # Memory pool for all cores (see also --mem-per-cpu)
+
+module load centos6/pandoc-1.12.3.3
+cd /n/irizarryfs01_backed_up/shicks/projects/benchmark-fdr
+Rscript -e 'library(knitr); rmarkdown::render("datasets/RNA-Seq/polyester-benchmarkFDR.Rmd", "html_document")' 

### 2018-01-10 datasets/RNA-Seq/yeast48-benchmarkFDR.Rmd
Add non-null comparisons

@@
 author: "Keegan Korthauer"
-date: "1/9/2018"
-output: html_document
+output: 
+    html_document:
+        toc: true
+        toc_float: true
+        highlight: tango
+        number_sections: true
 ---
@@
 
-### Introduction
+## Introduction
 
@@
 
-### Set up workspace
+## Set up workspace
 
-```{r, results='hide'}
+```{r, results='hide', message=FALSE}
 # Load packages and source benchmark FDR
@@
 library(DESeq2)
+library(truncnorm)
 
@@
 
-### Data download
+## Data download
 
@@
 
-```{r, readin}
-files <- dir(path = "./data/", pattern = "*.bam.gbgout", full.names = TRUE)
-sample_names <- sapply(strsplit(dir(path = "./data/", pattern = "*.bam.gbgout"), "_MID"),
+```{r, readin, results='hide', message=FALSE}
+files <- dir(path = "./yeast48-data/", pattern = "*.bam.gbgout", full.names = TRUE)
+sample_names <- sapply(strsplit(dir(path = "./yeast48-data/", pattern = "*.bam.gbgout"), "_MID"),
                        function(x) x[[1]])
@@
 
-### Analysis with DESeq2
+## Analysis with DESeq2 & FDR Benchmarking
 
@@
 Snf2 knockout).
+
+### Full analysis (48 vs 48 non-null)
 
@@
   mutate(condition = factor(condition))
+
+# filter low count genes
+counts <- counts[rowMeans(counts) > 1,]
 
@@
 
-There are many thousand genes detected using the full set.
-
-### Null Comparison
-
-```{r, null}
-# select a random subset of 20 WT samples
-set.seed(85873)
-dds_null <- dds[,colData(dds)$condition == "WT"]
-dds_null <- dds_null[,sample(1:ncol(dds_null), 20)]
-
-# add a fake condition column to coldat
-colData(dds_null)$fake <-  factor(c(rep("WT", 10), rep("TRT", 10))[sample(1:20, 20)])
-design(dds_null) <- ~fake
-
-# results on full set
-dds_null <- DESeq(dds_null)
-resultsNames(dds_null) # lists the coefficients
-resNULL <- results(dds_null, name="fake_WT_vs_TRT", independentFiltering = F)
-
-head(resNULL)
-
-sum(resNULL$padj < 0.05, na.rm=T)
-```
-
-### FDR Benchmark
-
-Build input data frame for summarized benchmark.
+There are many thousand genes detected using the full set. Next, we'll build input data frame for summarized benchmark.
 
 ```{r}
-geneExp <- tbl_df(data.frame(geneName=rownames(resNULL), pval=resNULL$pvalue, SE=resNULL$lfcSE, 
-                      ind_covariate = resNULL$baseMean, effect_size=resNULL$log2FoldChange, 
-                      test_statistic=resNULL$stat))
+geneExp <- tbl_df(data.frame(geneName=rownames(resFULL), pval=resFULL$pvalue, SE=resFULL$lfcSE, 
+                      ind_covariate = resFULL$baseMean, effect_size=resFULL$log2FoldChange, 
+                      test_statistic=resFULL$stat))
 
@@
 
-```{r}
-strat_hist(geneExp, pvalue="pval", covariate="ind_covariate", maxy=2)
+```{r, width=15, height=15}
+strat_hist(geneExp, pvalue="pval", covariate="ind_covariate", maxy=32)
 ```
@@
 plotFDRMethodsOverlap(sb, alpha=0.05, nsets=ncol(sb), order.by="freq", decreasing=TRUE, supplementary=FALSE)
+
+```
+
+To do: compare this to results using a subset of samples from each condition.
+
+### Null Comparison (10 versus 10)
+
+Here we'll repeat the above, but for a null comparison 10 versus 10 Snf2 samples,
+where the groups are selected randomly.
+
+```{r}
+# select a random subset of 20 WT samples
+set.seed(8373)
+dds_null <- dds[,colData(dds)$condition == "Snf2"]
+dds_null <- dds_null[,sample(1:ncol(dds_null), 20)]
+
+# add a fake condition column to coldat
+colData(dds_null)$fake <-  factor(c(rep("A", 10), rep("B", 10))[sample(1:20, 20)])
+design(dds_null) <- ~fake
+
+# results 
+dds_null <- DESeq(dds_null)
+resultsNames(dds_null) # lists the coefficients
+resNULL <- results(dds_null, name="fake_B_vs_A", independentFiltering = F)
+
+head(resNULL)
+
+sum(resNULL$padj < 0.05, na.rm=T)
+```
+
+Build input data frame for summarized benchmark.
+
+```{r}
+geneExp <- tbl_df(data.frame(geneName=rownames(resNULL), pval=resNULL$pvalue, SE=resNULL$lfcSE, 
+                      ind_covariate = resNULL$baseMean, effect_size=resNULL$log2FoldChange, 
+                      test_statistic=resNULL$stat))
+
+geneExp <- geneExp %>% na.omit() %>% dplyr::filter(ind_covariate>=0)
+
+```
+
+Look at covariate diagnostic plots.
+
+```{r, width=15, height=15}
+strat_hist(geneExp, pvalue="pval", covariate="ind_covariate", maxy=2)
+```
+
+```{r}
+rank_scatter(geneExp, pvalue="pval", covariate="ind_covariate")
+```
+
+Run benchmark methods.
+
+```{r}
+bd <- initializeBenchDesign()
+sb <- bd %>% buildBench(data=geneExp, parallel = T)
+
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Plot results.
+
+```{r}
+rejections_scatter(sb, supplementary=FALSE)
+```
+
+### Non-null Comparison (10 versus 10 null with spike-in)
+
+```{r}
+nDE <- 500
+
+# pick random set of nDE genes to add signal to
+DE <- sample(1:nrow(dds_null), nDE)
+truth <- rep(FALSE, nrow(dds_null))
+truth[DE] <- TRUE
+names(truth) <- rownames(dds_null)
+
+# randomly sample a log2FC from truncated N(0,0.5) (greater than 0.5) 
+log2FC <- rep(0, nrow(dds_null))
+log2FC[DE] <- rtruncnorm(nDE, 0.5, sd=0.5) * sample(c(1,-1), size=nDE, replace=TRUE)
+counts_new <- counts(dds_null)
+counts_new[,colData(dds_null)$fake=="A"] <- counts(dds_null)[,colData(dds_null)$fake=="A"] * 2^log2FC
+counts_new <- apply(counts_new, 2, as.integer)
+counts(dds_null) <- counts_new
+
+# results 
+dds_null <- DESeq(dds_null)
+resultsNames(dds_null) # lists the coefficients
+resNULL <- results(dds_null, name="fake_B_vs_A", independentFiltering = F)
+
+head(resNULL)
+
+sum(resNULL$padj < 0.05, na.rm=T)
+```
+
+Build input data frame for summarized benchmark.
+
+```{r}
+geneExp <- tbl_df(data.frame(geneName=rownames(resNULL), pval=resNULL$pvalue, SE=resNULL$lfcSE, 
+                      ind_covariate = resNULL$baseMean, effect_size=resNULL$log2FoldChange, 
+                      test_statistic=resNULL$stat,
+                      qvalue=truth))
+
+geneExp <- geneExp %>% na.omit() %>% dplyr::filter(ind_covariate>=0)
+
+```
+
+Look at covariate diagnostic plots.
+
+```{r, width=15, height=15}
+strat_hist(geneExp, pvalue="pval", covariate="ind_covariate", maxy=4)
+```
+
+```{r}
+rank_scatter(geneExp, pvalue="pval", covariate="ind_covariate")
+```
+
+Run benchmark methods and examine ROC curve.
+
+```{r}
+bd <- initializeBenchDesign()
+sb <- bd %>% buildBench(data=geneExp, parallel = T, truthCols = "qvalue")
+
+sb <- addDefaultMetrics(sb)
+pf <- estimatePerformanceMetrics(sb, alpha = c(0.001, 0.01, 0.025, 0.05, 0.1, 0.2), tidy=TRUE) 
+tpr <- pf %>% filter(performanceMetric == "TPR") %>% 
+  mutate(TPR=value) %>% 
+  select(blabel, alpha, TPR)
+fpr <- pf %>% filter(performanceMetric == "FPR") %>% 
+  mutate(FPR=value) %>% 
+  select(blabel, alpha, FPR)
+roc <- left_join(tpr, fpr) %>%
+  filter(blabel %in% c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
+             "scott-empirical", "scott-theoretical"))
+
+ggplot(data=roc,aes(FPR, TPR, color=blabel)) +
+  geom_point() +
+  geom_line() +
+  xlim(0,0.2)
+```
+
+Plot results.
+
+```{r}
+rejections_scatter(sb, supplementary=FALSE)
+```
+
+```{r}
+plotFDRMethodsOverlap(sb, alpha=0.05, nsets=ncol(sb), order.by="freq", decreasing=TRUE, supplementary=FALSE)
+
+```
+
+### Null Comparison (5 versus 5)
+
+Again we'll look at a null comparison of wild type samples in artifical groups, but
+for a smaller sample size (5 versus 5).
+
+```{r}
+# select a random subset of 10 WT samples
+dds_null <- dds[,colData(dds)$condition == "Snf2"]
+dds_null <- dds_null[,sample(1:ncol(dds_null), 10)]
+
+# add a fake condition column to coldat
+colData(dds_null)$fake <-  factor(c(rep("A", 5), rep("B", 5))[sample(1:10, 10)])
+design(dds_null) <- ~fake
+
+# results
+dds_null <- DESeq(dds_null)
+resultsNames(dds_null) # lists the coefficients
+resNULL <- results(dds_null, name="fake_B_vs_A", independentFiltering = F)
+
+head(resNULL)
+
+sum(resNULL$padj < 0.05, na.rm=T)
+```
+
+Build input data frame for summarized benchmark.
+
+```{r}
+geneExp <- tbl_df(data.frame(geneName=rownames(resNULL), pval=resNULL$pvalue, SE=resNULL$lfcSE, 
+                      ind_covariate = resNULL$baseMean, effect_size=resNULL$log2FoldChange, 
+                      test_statistic=resNULL$stat))
+
+geneExp <- geneExp %>% na.omit() %>% dplyr::filter(ind_covariate>=1)
+
+```
+
+Look at covariate diagnostic plots.
+
+```{r, width=15, height=15}
+strat_hist(geneExp, pvalue="pval", covariate="ind_covariate", maxy=3)
+```
+
+```{r}
+rank_scatter(geneExp, pvalue="pval", covariate="ind_covariate")
+```
+
+Run benchmark methods.
+
+```{r}
+bd <- initializeBenchDesign()
+sb <- bd %>% buildBench(data=geneExp, parallel = T)
+
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Plot results.
+
+```{r}
+rejections_scatter(sb, supplementary=FALSE)
+```
+
+```{r}
+#plotFDRMethodsOverlap(sb, alpha=0.05, nsets=ncol(sb), order.by="freq", decreasing=TRUE, supplementary=FALSE)
+```
+
+### Non-null Comparison (5 versus 5 null with spike-in)
+
+```{r}
+nDE <- 500
+
+# pick random set of nDE genes to add signal to
+DE <- sample(1:nrow(dds_null), nDE)
+truth <- rep(FALSE, nrow(dds_null))
+truth[DE] <- TRUE
+names(truth) <- rownames(dds_null)
+
+# randomly sample a log2FC from truncated N(0,0.5) (greater than 0.5) 
+log2FC <- rep(0, nrow(dds_null))
+log2FC[DE] <- rtruncnorm(nDE, 0.5, sd=0.5) * sample(c(1,-1), size=nDE, replace=TRUE)
+counts_new <- counts(dds_null)
+counts_new[,colData(dds_null)$fake=="A"] <- counts(dds_null)[,colData(dds_null)$fake=="A"] * 2^log2FC
+counts_new <- apply(counts_new, 2, as.integer)
+counts(dds_null) <- counts_new
+
+# results 
+dds_null <- DESeq(dds_null)
+resultsNames(dds_null) # lists the coefficients
+resNULL <- results(dds_null, name="fake_B_vs_A", independentFiltering = F)
+
+head(resNULL)
+
+sum(resNULL$padj < 0.05, na.rm=T)
+```
+
+Build input data frame for summarized benchmark.
+
+```{r}
+geneExp <- tbl_df(data.frame(geneName=rownames(resNULL), pval=resNULL$pvalue, SE=resNULL$lfcSE, 
+                      ind_covariate = resNULL$baseMean, effect_size=resNULL$log2FoldChange, 
+                      test_statistic=resNULL$stat,
+                      qvalue=truth))
+
+geneExp <- geneExp %>% na.omit() %>% dplyr::filter(ind_covariate>=1)
+
+```
+
+Look at covariate diagnostic plots.
+
+```{r,width=15, height=15}
+strat_hist(geneExp, pvalue="pval", covariate="ind_covariate", maxy=4)
+```
+
+```{r}
+rank_scatter(geneExp, pvalue="pval", covariate="ind_covariate")
+```
+
+Run benchmark methods and examine ROC curve.
+
+```{r}
+bd <- initializeBenchDesign()
+sb <- bd %>% buildBench(data=geneExp, parallel = T, truthCols = "qvalue")
+
+sb <- addDefaultMetrics(sb)
+pf <- estimatePerformanceMetrics(sb, alpha = c(0.001, 0.01, 0.025, 0.05, 0.1, 0.2), tidy=TRUE) 
+tpr <- pf %>% filter(performanceMetric == "TPR") %>% 
+  mutate(TPR=value) %>% 
+  select(blabel, alpha, TPR)
+fpr <- pf %>% filter(performanceMetric == "FPR") %>% 
+  mutate(FPR=value) %>% 
+  select(blabel, alpha, FPR)
+roc <- left_join(tpr, fpr) %>%
+  filter(blabel %in% c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
+             "scott-empirical", "scott-theoretical"))
+
+ggplot(data=roc,aes(FPR, TPR, color=blabel)) +
+  geom_point() +
+  geom_line() + 
+  xlim(0,0.2)
+```
+
+Plot results.
+
+```{r}
+rejections_scatter(sb, supplementary=FALSE)
+```
+
+```{r}
+plotFDRMethodsOverlap(sb, alpha=0.05, nsets=ncol(sb), order.by="freq", decreasing=TRUE, supplementary=FALSE)
 ```

### 2018-01-12 datasets/simulations/analysis-05-plot-psim-roc-upsetr.Rmd
Add average UpSet code

---
title: "Benchmarking FDR Methods - Simulation Summarized ROC / UpsetR Plots"
author: "Rafalab Journal Club Members"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    highlight: tango
    number_sections: true
    code_folding: hide
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, fig.width = 12, fig.height = 10)
```

<!-- ######################################################################## -->
## Workspace Setup
<!-- ######################################################################## -->

```{r, load-libraries, message=FALSE}
library("SummarizedBenchmark")
library("tidyverse")
```

Let's just start with a single set of simulations. These simulations include 20,000 test statistics simulated from a normal distribution, with 90% simulated from a standard normal, and 10% simulated from a mean-shifted normal distribution. The different settings correspond to different mean shifts for the alternative tests. For each setting, the simulation results are summarized over 100 replications.

```{r}
fl <- list.files("data-psim/M100/esize_fixed-bl-cubic-90-0.8",
                 pattern="M100\\.rds", full.names=TRUE)
```

<!-- ######################################################################## -->
## Average UpSet Plot Across Replicates
<!-- ######################################################################## -->

First, we consider generating UpSet plots for the average behavior over the 100 replications.

```{r}
## loop over simulation settings
for (f in fl) {

    ## read in list of SummarizedBenchmark objects (100 replications) 
    res <- readRDS(f)

    ## determine significant hits at 0.10 cutoff
    hits_tabs <- lapply(res, function(x) {
        as.tibble(cbind((assay(x, "qvalue") < 0.10) + 0,
                        truth = rowData(x)$qvalue)) %>%
            select(-starts_with("ihw-a0"),
                   -`bl-df02`,  -`bl-df04`,  -`bl-df05`) %>%
            rename(ihw = `ihw-a10`, bl = `bl-df03`,
                   scott_e = `scott-empirical`,
                   scott_t = `scott-theoretical`) %>%
            as.data.frame
    })

    ## count up frequencies
    freq_tabs <- lapply(hits_tabs,
                        function(x) {
                            UpSetR:::Counter(x, 11, 1, names(x), nintersections = 2^11,
                                             mbar_color = "gray23",
                                             order_mat = "degree", aggregate = "degree",
                                             cut = NULL, empty_intersects = TRUE,
                                             decrease = TRUE)
                        })

    ## merge all frequencies into single table
    freq_tab <- freq_tabs %>%
        mapply(function(itab, idx) { as.tibble(itab) %>%
                                         select(-x, -color) %>%
                                         rename(!!(paste0("freq.", idx)) := freq) },
               itab = ., idx = 1:length(freq_tabs),
               SIMPLIFY = FALSE) %>%
        reduce(left_join, by = c("unadjusted", "bonf", "bh", "qvalue", "ihw",
                                 "ashs", "bl", "lfdr", "scott_t", "scott_e", "truth"))

    ## summarize across 100 replications of each setting
    freq_tab <- freq_tab %>%
        gather(repl, cnt, starts_with("freq")) %>%
        group_by(unadjusted, bonf, bh, qvalue, ihw,
                 ashs, bl, lfdr, scott_t, scott_e, truth) %>%
        summarize(freq_mean = round(mean(cnt)),
                  freq_50p = round(median(cnt)),
                  freq_10p = round(quantile(cnt, .1)),
                  freq_90p = round(quantile(cnt, .9))) %>%
        ungroup %>%
        mutate(setname = paste(ifelse(unadjusted, "unadjusted", NA),
                               ifelse(bonf, "bonf", NA),
                               ifelse(bh, "bh", NA),
                               ifelse(qvalue, "qvalue", NA),
                               ifelse(ihw, "ihw", NA),
                               ifelse(ashs, "ashs", NA),
                               ifelse(bl, "bl", NA),
                               ifelse(lfdr, "lfdr", NA),
                               ifelse(scott_t, "scott_t", NA),
                               ifelse(scott_e, "scott_e", NA),
                               ifelse(scott_e, "truth", NA),
                               sep = '&'),
               setname = gsub("NA", "", setname) %>%
                   gsub("&{2,}", "&", .) %>%
                   gsub("^&", "", .) %>%
                   gsub("&$", "", .)) %>%
        select(setname, freq_mean,
               freq_50p, freq_10p, freq_90p)

    ## convert to vector to pass to UpSetR package
    freq_list <- freq_tab$freq_mean
    names(freq_list) <- freq_tab$setname

    ## make simple upset plot - save as png
    outdir <- gsub("data-", "plot-", dirname(f))
    dir.create(outdir, showWarnings = FALSE, recursive = TRUE)
    png(file.path(outdir, gsub("\\.rds", "-mean-upset.png", basename(f))),
        width = 10, height= 5, unit = "in", res = 300)

    upset(fromExpression(freq_list),
          nsets = 11,
          mb.ratio = c(0.55, 0.45),
          order.by = "freq",
          decreasing = TRUE,
          set.metadata = list(data = data.frame(sets = c("unadjusted", "bonf", "bh", "qvalue", "ihw",
                                                         "ashs", "bl", "lfdr", "scott_t", "scott_e", "truth"),
                                                isTruth = c(rep(FALSE, 10), TRUE)),
                              plots = list(
                                  list(type = "matrix_rows", 
                                       column = "isTruth",
                                       colors = c("TRUE" = "blue", "FALSE" = "gray"), 
                                       alpha = 0.2))))
    dev.off()
}
```

<!-- ######################################################################## -->
## Average ROC Plot Across Replicates
<!-- ######################################################################## -->

Next, we consider drawing average ROC curves for each setting...

### 2018-01-18 datasets/RNA-Seq/yeast48-benchmarkFDR.Rmd
Add more diagnosticplots, more realistic fold change simulation

@@
 
-## Introduction
+# Introduction
 
@@
 
-## Set up workspace
+# Set up workspace
 
@@
 
-## Data download
+# Data download
 
@@
 
-### Read data into R and create a count table
+## Read data into R and create a count table
 
@@
 
-## Analysis with DESeq2 & FDR Benchmarking
+# Analysis of full 48v48 with DESeq2 (non-null)
 
@@
 
-### Full analysis (48 vs 48 non-null)
+## Set up DESeq2 object
 
@@
 
+## Check assumptions
+
 There are many thousand genes detected using the full set. Next, we'll build input data frame for summarized benchmark.
@@
                       ind_covariate = resFULL$baseMean, effect_size=resFULL$log2FoldChange, 
-                      test_statistic=resFULL$stat))
+                      test_statistic=resFULL$stat,
+                      pzero=rowSums(counts(dds)==0)/ncol(counts(dds))  ))
 
-geneExp <- geneExp %>% na.omit() %>% dplyr::filter(ind_covariate>=1)
+# filter NAs and those with less than 50% expressed 
+geneExp <- geneExp %>% na.omit() %>% dplyr::filter(pzero < 0.5)
 
+```
+Check unimodal assumption (for ash).
+
+```{r}
+ggplot(data=geneExp, aes(test_statistic)) +
+  geom_histogram(bins=30)
 ```
@@
 ```
+
+## FDR benchmarking
 
@@
 
-### Null Comparison (10 versus 10)
+# Null Comparison (10 versus 10)
 
@@
 where the groups are selected randomly.
+
+## Run DESeq2 on subset
 
@@
                       ind_covariate = resNULL$baseMean, effect_size=resNULL$log2FoldChange, 
-                      test_statistic=resNULL$stat))
+                      test_statistic=resNULL$stat,
+                      pzero=rowSums(counts(dds)==0)/ncol(counts(dds))  ))
 
-geneExp <- geneExp %>% na.omit() %>% dplyr::filter(ind_covariate>=0)
+# filter NAs and those with less than 50% expressed 
+geneExp <- geneExp %>% na.omit() %>% dplyr::filter(pzero < 0.5)
 
+```
+
+## Check assumptions
+
+Check unimodal assumption (for ash).
+
+```{r}
+ggplot(data=geneExp, aes(test_statistic)) +
+  geom_histogram(bins=30)
 ```
@@
 ```
+
+## FDR Benchmark
 
@@
 
-### Non-null Comparison (10 versus 10 null with spike-in)
+# Non-null Comparison (10 versus 10 null with spike-in)
 
+Here we'll repeat the above, but adding in some DE genes so that we can evaluate the
+true positive and false positive rates of each method.
+
+## Run DESeq2 on subset
 ```{r}
@@
 
-# randomly sample a log2FC from truncated N(0,0.5) (greater than 0.5) 
+# randomly sample a log2FC from original FCs (without regard to DE)
 log2FC <- rep(0, nrow(dds_null))
-log2FC[DE] <- rtruncnorm(nDE, 0.5, sd=0.5) * sample(c(1,-1), size=nDE, replace=TRUE)
+log2FC[DE] <- resFULL$log2FoldChange[DE]
+
 counts_new <- counts(dds_null)
@@
                       test_statistic=resNULL$stat,
-                      qvalue=truth))
+                      qvalue=truth,
+                      pzero=rowSums(counts(dds)==0)/ncol(counts(dds))  ))
 
-geneExp <- geneExp %>% na.omit() %>% dplyr::filter(ind_covariate>=0)
+# filter NAs and those with less than 50% expressed 
+geneExp <- geneExp %>% na.omit() %>% dplyr::filter(pzero < 0.5)
 
+```
+## Check assumptions
+
+Check unimodal assumption (for ash).
+
+```{r}
+ggplot(data=geneExp, aes(test_statistic)) +
+  geom_histogram(bins=30)
 ```
@@
 ```
+
+## FDR benchmark
 
@@
 
+# fpr vs tpr
 ggplot(data=roc,aes(FPR, TPR, color=blabel)) +
@@
   xlim(0,0.2)
+
+# alpha vs tpr
+ggplot(data=roc,aes(alpha, TPR, color=blabel)) +
+  geom_point() +
+  geom_line() + 
+  xlim(0,0.2)
+
+# alpha vs fpr
+ggplot(data=roc,aes(alpha, FPR, color=blabel)) +
+  geom_point() +
+  geom_line() + 
+  xlim(0,0.2)+
+  geom_abline(slope=1, linetype="dashed", color="grey")
 ```
@@
 
-### Null Comparison (5 versus 5)
+# Null Comparison (5 versus 5)
 
@@
 
+## Run DESeq2 on subset
 ```{r}
@@
                       ind_covariate = resNULL$baseMean, effect_size=resNULL$log2FoldChange, 
-                      test_statistic=resNULL$stat))
+                      test_statistic=resNULL$stat,
+                      pzero=rowSums(counts(dds)==0)/ncol(counts(dds))  ))
 
-geneExp <- geneExp %>% na.omit() %>% dplyr::filter(ind_covariate>=1)
+# filter NAs and those with less than 50% expressed 
+geneExp <- geneExp %>% na.omit() %>% dplyr::filter(pzero < 0.5)
+```
 
+## Check assumptions
+
+Check unimodal assumption (for ash).
+
+```{r}
+ggplot(data=geneExp, aes(test_statistic)) +
+  geom_histogram(bins=30)
 ```
@@
 
-### Non-null Comparison (5 versus 5 null with spike-in)
+# Non-null Comparison (5 versus 5 null with spike-in)
+
+Here we'll repeat the above, but adding in some DE genes so that we can evaluate the
+true positive and false positive rates of each method.
+
+## Run DESeq2 on subset
 
@@
                       test_statistic=resNULL$stat,
-                      qvalue=truth))
+                      qvalue=truth,
+                      pzero=rowSums(counts(dds)==0)/ncol(counts(dds))  ))
 
-geneExp <- geneExp %>% na.omit() %>% dplyr::filter(ind_covariate>=1)
+# filter NAs and those with less than 50% expressed 
+geneExp <- geneExp %>% na.omit() %>% dplyr::filter(pzero < 0.5)
 
+```
+
+## Check assumptions
+
+Check unimodal assumption (for ash).
+
+```{r}
+ggplot(data=geneExp, aes(test_statistic)) +
+  geom_histogram(bins=30)
 ```
@@
   xlim(0,0.2)
+
+# alpha vs tpr
+ggplot(data=roc,aes(alpha, TPR, color=blabel)) +
+  geom_point() +
+  geom_line() + 
+  xlim(0,0.2)
+
+# alpha vs fpr
+ggplot(data=roc,aes(alpha, FPR, color=blabel)) +
+  geom_point() +
+  geom_line() + 
+  xlim(0,0.2) +
+  geom_abline(slope=1, linetype="dashed", color="grey")
 ```
@@
 
-### Session information
+# Session information
 

### 2018-01-19 datasets/microbiome/microbiome-SummarizedBenchmark.Rmd
Update microbiome results.
Filter OTUs by presence in 1% of samples. Do genus-level analysis
for Baxter and Goodrich.

@@
 
-## Remove OTUs with fewere than 10 reads
-#otu <- remove_shallow_otus(otu, 10)
+## Remove OTUs with fewer than 10 reads
+otu <- remove_shallow_otus(otu, 10)
 
@@
 
-##### Phylogeny?!
+##### Phylogeny
 
@@
 
-```{r, schubert-phylo, fig.width=10, fig.height=6, eval=TRUE}
+```{r, schubert-phylo, fig.width=10, fig.height=6, eval=FALSE}
 #strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=20)
@@
 
-# Remove samples with fewer than sample_Reads reads
+remove_shallow_otus <- function(df, n_reads){
+  return(df[rowSums(df) > n_reads, ])
+}
+
+remove_rare_otus <- function(df, perc_samples){
+  return(df[rowSums(df > 0) / dim(df)[2] > perc_samples, ])
+}
+
+## Remove OTUs with fewer than 10 reads
+otu <- remove_shallow_otus(otu, 10)
+print(dim(otu))
+## Remove OTUs in fewer than 1% of samples
+otu <- remove_rare_otus(otu, 0.01)
+print(dim(otu))
+
+# Remove samples with fewer than sample_reads reads
 sample_reads <- 100
@@
 
-```{r, goodrich-phylo, fig.width=10, fig.height=6}
+```{r, goodrich-phylo, fig.width=10, fig.height=6, eval=FALSE}
 #strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=20)
@@
 
-### CRC Baxter - colorectal cancer
-
-Next, I'll look at the CRC Baxter dataset.
-For colorectal cancer, we do expect some amount of truly differentially abundant OTUs, but not as many as in diarrhea.
-This dataset will hopefully provide an intermediate non-extreme case study.
-We'll download the processed OTU tables from Zenodo and unzip them in the 
-`benchmark-fdr/datasets/microbiome/DATA` folder.
-
-#### Data download
-
-```{bash, baxter-download, eval=FALSE}
-cd /Users/claire/github/benchmark-fdr/datasets/microbiome/DATA/
-curl -O "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
-tar -xzvf crc_baxter_results.tar.gz
-cd ..
-```
-
-Next, we'll read in the unzipped OTU table and metadata files into R.
-
-We'll need to manually define which disease labels are of interest and what metadata
-column contains the sample IDs (i.e. the IDs in the OTU table).
-
-```{r, baxter-readdata}
-otu_path = paste0(datdir, "crc_baxter_results/RDP/crc_baxter.otu_table.100.denovo.rdp_assigned")
-meta_path = paste0(datdir, "crc_baxter_results/crc_baxter.metadata.txt")
-
-# DiseaseState labels to keep
-labels <- c("H", "CRC")
-
-# Metadata column that has sample IDs
-col_label <- "Sample_Name_s"
-
-# load OTU table and metadata
-otu <- read.table(otu_path)
-meta <- read.csv(meta_path, sep='\t')
-
-# Keep only samples with the right DiseaseState metadata
-meta <- meta %>% filter(DiseaseState %in% labels)
-
-# Add "X" in front of sample IDs because of how R read in the OTU table
-meta[, 1] <- sub("^", "X", meta[, 1])
-
-# Keep only samples with both metadata and 16S data
-keep_samples <- intersect(colnames(otu), meta[, 1])
-otu <- otu[, keep_samples]
-meta <- meta %>% filter(get(col_label) %in% keep_samples)
-
-```
-
-Since we'll be using OTU-wise covariates, we shouldn't need to perform any
-filtering/cleaning of the OTUs, apart from removing any that are all zeros. 
-(This may happen after removing shallow samples, I think.)
-However, let's make sure we get rid of any samples with too few reads. 
-We define the minimum number of reads per sample in `sample_reads`.
-
-After we've removed any shallow samples, we'll convert the OTU table to relative
-abundances. I'll add a pseudo-count of 1e-7 to any zero entries, to avoid
-problems with taking logs.
-
-```{r, baxter-cleandata}
-remove_shallow_smpls <- function(df, n_reads) {
-  # Removes samples with fewer than n_reads from dataframe df.
-  # df has OTUs in rows and samples in columns
-  return(df[, colSums(df) > n_reads])
-  }
-
-remove_shallow_otus <- function(df, n_reads){
-  return(df[rowSums(df) > n_reads, ])
-  }
-
-## Remove OTUs with fewere than 10 reads
-#otu <- remove_shallow_otus(otu, 10)
-
-# Remove samples with fewer than sample_Reads reads
-sample_reads <- 100
-otu <- remove_shallow_smpls(otu, sample_reads)
-
-# Update metadata with new samples
-meta <- meta %>% filter(get(col_label) %in% colnames(otu))
-
-# Remove empty OTUs
-otu <- otu[rowSums(otu) > 0, ]
-
-# Convert to relative abundance
-abun_otu <- t(t(otu) / rowSums(t(otu)))
-
-# Add pseudo counts
-#minabun <- min(abun_otu[abun_otu > 0]) # Use this to figure out what pseudo-count value to add
-zeroabun <- 0
-abun_otu <- abun_otu + zeroabun
-
-```
-
-Next, we need to calculate the pvalues, effect size, and standard error for each OTU.
-Here, we'll compare CRC vs. healthy. We won't consider the nonCRC adenoma patients.
-We'll put these results into a dataframe, and label the columns with the 
-standardized names for downstream use (`pval`, `SE`, `effect_size`, `test_statistic`).
-The test statistic is the one returned by `wilcox.test()`.
-
-Note that the effect here is calculated as logfold change of mean abundance in controls
-relative to cases (i.e. `log(mean_abun[controls]/mean_abun[cases])`)
-
-While we're at it, we'll also calculate the mean abundance and ubiquity (detection rate)
-of each OTU. Later, we can assign their values to a new column called `ind_covariate` 
-for use in downstream steps.
-
-```{r, baxter-pvals}
-resfile <- paste0(resdir, "baxter_results_",
-                  nrow(abun_otu), "_OTUs.RData")
+## Goodrich, genus-level analysis
+
+```{r, goodrich-genus}
+## Collapse baxter OTU table to genus level
+
+# Add column with otu names
+genus_df <- as.data.frame(abun_otu)
+genus_df$otu <- rownames(genus_df)
+# Melt into tidy format
+genus_df <- genus_df %>% melt(id.var="otu", variable.name="sample", value.name="abun")
+# Split into separate taxonomic levels
+genus_df <- genus_df %>% 
+  separate(otu, c("kingdom", "phylum", "class", "order", "family", "genus", "species", "denovo"), 
+           sep=";", remove = FALSE)
+
+# Get rid of unannoated genera, and sum abundances for genera
+genus_df <- genus_df %>% 
+  filter(genus != "g__") %>% 
+  group_by(genus, sample) %>% 
+  summarise(total_abun = sum(abun))
+
+# Convert back to longform
+genus_df <- genus_df %>% spread(genus, total_abun) %>% as.data.frame
+rownames(genus_df) <- genus_df$sample
+genus_df <- genus_df %>% select(-sample) %>% t
+
+# And re-order columns to match metadata
+genus_df <- genus_df[, match(meta$X, colnames(genus_df))]
+```
+
+```{r, goodrich-genus-pvals}
+resfile <- paste0(resdir, "goodrich_results_",
+                  nrow(genus_df), "_genera.RData")
 if (!file.exists(resfile)){
   # Get case and control indices
-  case_idx <- meta$DiseaseState %in% c("CRC")
+  case_idx <- meta$DiseaseState %in% c("OB")
   ctrl_idx <- meta$DiseaseState %in% c("H")
@@
 
+  casedf <- genus_df[, case_idx]
+  ctrldf <- genus_df[, ctrl_idx]
+  for (o in rownames(genus_df)) {
+    # wilcoxon p value  
+    w <- wilcox.test(casedf[o, ], 
+                     ctrldf[o, ])
+    p <- w$p.value
+    teststat <- w$statistic
+
+    pvals <- c(pvals, p)
+    teststats <- c(teststats, teststat)
+
+    # standard error of the OTU abundance, across all samples
+    ses <- c(ses, 
+             sd(genus_df[o, ])/sqrt(length(genus_df[o, ]))
+             )
+
+    # mean OTU abundance across all samples (after removing pseudo-count)
+    mean_abuns <- c(mean_abuns, 
+                    mean(genus_df[o, ] - zeroabun)
+                    )
+
+    # mean OTU abundance across only samples with the OTU present
+    mean_abuns_present <- c(mean_abuns_present, 
+                            sum(genus_df[o, ] - zeroabun) / sum(genus_df[o, ] > zeroabun)
+                            )
+
+    # ubiquity of OTU across all samples
+    ubis <- c(ubis, 
+              sum(genus_df[o, ] > zeroabun) / length(genus_df[o, ])
+              )
+
+    # effect (logfold difference)
+    effs <- c(effs, 
+              log(mean(genus_df[o, ctrl_idx])/mean(genus_df[o, case_idx]))
+              )
+
+  }
+
+  res <- data.frame(otu = rownames(genus_df), 
+                    pval = pvals, test_statistic = teststats,
+                    SE = ses, effect_size = effs, 
+                    mean_abun = mean_abuns, mean_abun_present = mean_abuns_present,
+                    ubiquity = ubis)
+  res <- res %>%
+    mutate(test_statistic = qnorm(exp(log(pval)-log(2)), lower.tail=FALSE) * sign(effect_size))
+
+  save(res, file=resfile)
+
+}else{
+  load(resfile)
+}
+```
+
+#### Check Covariate Diagnostics
+
+Here we look to see if the covariates do indeed look informative.
+
+##### Ubiquity
+
+```{r, goodrich-genus-ubi, fig.width=10, fig.height=3.5}
+strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
+rank_scatter(res, pvalue="pval", covariate="ubiquity")
+```
+
+Something weird is happening with p=0.20 and p=0.40 - maybe this has something to do with ties?
+Either way, ubiquity looks to be a bit informative (from the scatter plot, but not really the histograms...)
+
+##### Mean abundance (across non-zero samples)
+
+```{r, goodrich-genus-abun, fig.width=10, fig.height=3.5}
+strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
+rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
+```
+
+Let's use `ubiquity` as our `ind_covariate`
+
+```{r, goodrich-genus-labelcovariate}
+res <- res %>% mutate(ind_covariate = ubiquity)
+```
+
+Also (hacky), remove test_statistic so that ash doesn't run.
+
+```{r, goodrich-genus-rm-ash}
+res <- res %>% 
+  mutate(test_statistic = NA) %>%
+  mutate(effect_size = NA)
+```
+
+#### Set up BenchDesign object
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We'll do this for a few
+values of nmids, to explore its effects on the outputs.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, goodrich-genus-sb, results="hide", message=FALSE, eval=TRUE}
+for (nmids in c(5, 50, 150)){
+  print(nmids)
+  bd <- initializeBenchDesign(nmids=nmids)
+
+  resfile <- paste0(resdir, "goodrich_summarizedBenchmark_",
+                    nrow(res), "_genera_nmids_", nmids, ".RData")
+  duration <- NA
+  if (!file.exists(resfile)){
+    t1 <- proc.time()
+    sb <- bd %>% buildBench(data=res, ftCols = "ind_covariate")
+                            #parallel=TRUE, BPPARAM=multicoreParam)
+    metadata(sb)$data_download_link <- "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
+    save(sb, file=resfile)
+    duration <- round((proc.time()-t1)[3]/60,1)
+  }else{
+    load(resfile)
+  }
+}
+
+```
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, goodrich-genus-metrics, eval=TRUE}
+# rename assay to qvalue
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+Debugging note: to look at the matrix of qvalues, run `assays(sb)[["qvalue"]]`
+
+```{r, goodrich-genus-plot, width=15, height=15, eval=TRUE}
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+#plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+```
+
+Note: Benjamini-Hochberg (bh) overlaps exactly with the IHW results.
+
+```{r, goodrich-genus-overlap}
+plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+```
+
+### CRC Baxter - colorectal cancer
+
+Next, I'll look at the CRC Baxter dataset.
+For colorectal cancer, we do expect some amount of truly differentially abundant OTUs, but not as many as in diarrhea.
+This dataset will hopefully provide an intermediate non-extreme case study.
+We'll download the processed OTU tables from Zenodo and unzip them in the 
+`benchmark-fdr/datasets/microbiome/DATA` folder.
+
+#### Data download
+
+```{bash, baxter-download, eval=FALSE}
+cd /Users/claire/github/benchmark-fdr/datasets/microbiome/DATA/
+curl -O "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
+tar -xzvf crc_baxter_results.tar.gz
+cd ..
+```
+
+Next, we'll read in the unzipped OTU table and metadata files into R.
+
+We'll need to manually define which disease labels are of interest and what metadata
+column contains the sample IDs (i.e. the IDs in the OTU table).
+
+```{r, baxter-readdata}
+otu_path = paste0(datdir, "crc_baxter_results/RDP/crc_baxter.otu_table.100.denovo.rdp_assigned")
+meta_path = paste0(datdir, "crc_baxter_results/crc_baxter.metadata.txt")
+
+# DiseaseState labels to keep
+labels <- c("H", "CRC")
+
+# Metadata column that has sample IDs
+col_label <- "Sample_Name_s"
+
+# load OTU table and metadata
+otu <- read.table(otu_path)
+meta <- read.csv(meta_path, sep='\t')
+
+# Keep only samples with the right DiseaseState metadata
+meta <- meta %>% filter(DiseaseState %in% labels)
+
+# Add "X" in front of sample IDs because of how R read in the OTU table
+meta[, 1] <- sub("^", "X", meta[, 1])
+
+# Keep only samples with both metadata and 16S data
+keep_samples <- intersect(colnames(otu), meta[, 1])
+otu <- otu[, keep_samples]
+meta <- meta %>% filter(get(col_label) %in% keep_samples)
+
+```
+
+A brief aside: what's the count distribution of these OTUs?
+
+```{r, eval=FALSE}
+otu %>% colSums %>% hist()
+```
+
+Since we'll be using OTU-wise covariates, we shouldn't need to perform any
+filtering/cleaning of the OTUs, apart from removing any that are all zeros. 
+(This may happen after removing shallow samples, I think.)
+However, let's make sure we get rid of any samples with too few reads. 
+We define the minimum number of reads per sample in `sample_reads`.
+
+After we've removed any shallow samples, we'll convert the OTU table to relative
+abundances. I'll add a pseudo-count of 1e-7 to any zero entries, to avoid
+problems with taking logs.
+
+```{r, baxter-cleandata}
+remove_shallow_smpls <- function(df, n_reads) {
+  # Removes samples with fewer than n_reads from dataframe df.
+  # df has OTUs in rows and samples in columns
+  return(df[, colSums(df) >= n_reads])
+  }
+
+remove_shallow_otus <- function(df, n_reads){
+  return(df[rowSums(df) > n_reads, ])
+  }
+
+remove_rare_otus <- function(df, perc_samples){
+  return(df[rowSums(df > 0) / dim(df)[2] > perc_samples, ])
+}
+
+## Remove OTUs with fewer than 10 reads
+otu <- remove_shallow_otus(otu, 10)
+print(dim(otu))
+## Remove OTUs in fewer than 1% of samples
+otu <- remove_rare_otus(otu, 0.01)
+print(dim(otu))
+
+# Remove samples with fewer than sample_Reads reads
+sample_reads <- 100
+otu <- remove_shallow_smpls(otu, sample_reads)
+print(dim(otu))
+
+# Update metadata with new samples
+meta <- meta %>% filter(get(col_label) %in% colnames(otu))
+
+# Remove empty OTUs
+otu <- otu[rowSums(otu) > 0, ]
+print(dim(otu))
+
+# Convert to relative abundance
+abun_otu <- t(t(otu) / rowSums(t(otu)))
+
+# Add pseudo counts
+#minabun <- min(abun_otu[abun_otu > 0]) # Use this to figure out what pseudo-count value to add
+zeroabun <- 0
+abun_otu <- abun_otu + zeroabun
+
+```
+
+Next, we need to calculate the pvalues, effect size, and standard error for each OTU.
+Here, we'll compare CRC vs. healthy. We won't consider the nonCRC adenoma patients.
+We'll put these results into a dataframe, and label the columns with the 
+standardized names for downstream use (`pval`, `SE`, `effect_size`, `test_statistic`).
+The test statistic is the one returned by `wilcox.test()`.
+
+Note that the effect here is calculated as logfold change of mean abundance in controls
+relative to cases (i.e. `log(mean_abun[controls]/mean_abun[cases])`)
+
+While we're at it, we'll also calculate the mean abundance and ubiquity (detection rate)
+of each OTU. Later, we can assign their values to a new column called `ind_covariate` 
+for use in downstream steps.
+
+```{r, baxter-pvals}
+resfile <- paste0(resdir, "baxter_results_",
+                  nrow(abun_otu), "_OTUs.RData")
+if (!file.exists(resfile)){
+  # Get case and control indices
+  case_idx <- meta$DiseaseState %in% c("CRC")
+  ctrl_idx <- meta$DiseaseState %in% c("H")
+
+  # Calculate pvalues, effects, and stderr
+  pvals <- c()
+  teststats <- c()
+  ses <- c()
+  effs <- c()
+  mean_abuns <- c()
+  mean_abuns_present <- c()
+  ubis <- c()
+
   casedf <- abun_otu[, case_idx]
@@
 
-```{r, baxter-phylo, fig.width=10, fig.height=6}
+```{r, baxter-phylo, fig.width=10, fig.height=6, eval=FALSE}
 #strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=20)
@@
 
-```{r, echo=FALSE, eval=FALSE}
-if(!is.na(duration)){
-  message("This step took ", duration, " minutes for ", nrow(res), " OTUs using ",
-        cores, " cores.")
-}
-```
-
 Next, we'll add the default performance metric for q-value assays. First, we have
@@
 ```
+
+## Baxter, genus-level analysis
+
+```{r, baxter-genus}
+## Collapse baxter OTU table to genus level
+
+# Add column with otu names
+genus_df <- as.data.frame(abun_otu)
+genus_df$otu <- rownames(genus_df)
+# Melt into tidy format
+genus_df <- genus_df %>% melt(id.var="otu", variable.name="sample", value.name="abun")
+# Split into separate taxonomic levels
+genus_df <- genus_df %>% 
+  separate(otu, c("kingdom", "phylum", "class", "order", "family", "genus", "species", "denovo"), 
+           sep=";", remove = FALSE)
+
+# Get rid of unannoated genera, and sum abundances for genera
+genus_df <- genus_df %>% 
+  filter(genus != "g__") %>% 
+  group_by(genus, sample) %>% 
+  summarise(total_abun = sum(abun))
+
+# Convert back to longform
+genus_df <- genus_df %>% spread(genus, total_abun) %>% as.data.frame
+rownames(genus_df) <- genus_df$sample
+genus_df <- genus_df %>% select(-sample) %>% t
+
+# And re-order columns to match metadata
+genus_df <- genus_df[, match(meta$Sample_Name_s, colnames(genus_df))]
+```
+
+```{r, baxter-genus-pvals}
+resfile <- paste0(resdir, "baxter_results_",
+                  nrow(genus_df), "_genera.RData")
+if (!file.exists(resfile)){
+  # Get case and control indices
+  case_idx <- meta$DiseaseState %in% c("CRC")
+  ctrl_idx <- meta$DiseaseState %in% c("H")
+
+  # Calculate pvalues, effects, and stderr
+  pvals <- c()
+  teststats <- c()
+  ses <- c()
+  effs <- c()
+  mean_abuns <- c()
+  mean_abuns_present <- c()
+  ubis <- c()
+
+  casedf <- genus_df[, case_idx]
+  ctrldf <- genus_df[, ctrl_idx]
+  for (o in rownames(genus_df)) {
+    # wilcoxon p value  
+    w <- wilcox.test(casedf[o, ], 
+                     ctrldf[o, ])
+    #w <- kruskal.test(list(casedf[o, ], ctrldf[o, ]))
+    p <- w$p.value
+    teststat <- w$statistic
+
+    pvals <- c(pvals, p)
+    teststats <- c(teststats, teststat)
+
+    # standard error of the OTU abundance, across all samples
+    ses <- c(ses, 
+             sd(genus_df[o, ])/sqrt(length(genus_df[o, ]))
+             )
+
+    # mean OTU abundance across all samples (after removing pseudo-count)
+    mean_abuns <- c(mean_abuns, 
+                    mean(genus_df[o, ] - zeroabun)
+                    )
+
+    # mean OTU abundance across only samples with the OTU present
+    mean_abuns_present <- c(mean_abuns_present, 
+                            sum(genus_df[o, ] - zeroabun) / sum(genus_df[o, ] > zeroabun)
+                            )
+
+    # ubiquity of OTU across all samples
+    ubis <- c(ubis, 
+              sum(genus_df[o, ] > zeroabun) / length(genus_df[o, ])
+              )
+
+    # effect (logfold difference)
+    effs <- c(effs, 
+              log(mean(genus_df[o, ctrl_idx])/mean(genus_df[o, case_idx]))
+              )
+
+  }
+
+  res <- data.frame(otu = rownames(genus_df), 
+                    pval = pvals, test_statistic = teststats,
+                    SE = ses, effect_size = effs, 
+                    mean_abun = mean_abuns, mean_abun_present = mean_abuns_present,
+                    ubiquity = ubis)
+  res <- res %>%
+    mutate(test_statistic = qnorm(exp(log(pval)-log(2)), lower.tail=FALSE) * sign(effect_size))
+
+  save(res, file=resfile)
+
+}else{
+  load(resfile)
+}
+```
+
+#### Check Covariate Diagnostics
+
+Here we look to see if the covariates do indeed look informative.
+
+##### Ubiquity
+
+```{r, baxter-genus-ubi, fig.width=10, fig.height=3.5}
+strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
+rank_scatter(res, pvalue="pval", covariate="ubiquity")
+```
+
+Something weird is happening with p=0.20 and p=0.40 - maybe this has something to do with ties?
+Either way, ubiquity looks to be a bit informative (from the scatter plot, but not really the histograms...)
+
+##### Mean abundance (across non-zero samples)
+
+```{r, baxter-genus-abun, fig.width=10, fig.height=3.5}
+strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
+rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
+```
+
+Let's use `ubiquity` as our `ind_covariate`
+
+```{r, baxter-genus-labelcovariate}
+res <- res %>% mutate(ind_covariate = ubiquity)
+```
+
+Also (hacky), remove test_statistic so that ash doesn't run.
+
+```{r, baxter-genus-rm-ash}
+res <- res %>% 
+  mutate(test_statistic = NA) %>%
+  mutate(effect_size = NA)
+```
+
+#### Set up BenchDesign object
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We'll do this for a few
+values of nmids, to explore its effects on the outputs.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, baxter-genus-sb, results="hide", message=FALSE, eval=TRUE}
+for (nmids in c(5, 50, 150)){
+  print(nmids)
+  bd <- initializeBenchDesign(nmids=nmids)
+
+  resfile <- paste0(resdir, "baxter_summarizedBenchmark_",
+                    nrow(res), "_genera_nmids_", nmids, ".RData")
+  duration <- NA
+  if (!file.exists(resfile)){
+    t1 <- proc.time()
+    sb <- bd %>% buildBench(data=res, ftCols = "ind_covariate")
+                            #parallel=TRUE, BPPARAM=multicoreParam)
+    metadata(sb)$data_download_link <- "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
+    save(sb, file=resfile)
+    duration <- round((proc.time()-t1)[3]/60,1)
+  }else{
+    load(resfile)
+  }
+}
+
+```
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, baxter-genus-metrics, eval=TRUE}
+# rename assay to qvalue
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+Debugging note: to look at the matrix of qvalues, run `assays(sb)[["qvalue"]]`
+
+```{r, baxter-genus-plot, width=15, height=15, eval=TRUE}
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+#plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+```
+
+Note: Benjamini-Hochberg (bh) overlaps exactly with the IHW results.
+
+```{r, baxter-genus-overlap}
+plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+```

### 2018-01-19 datasets/microbiome/microbiome-SummarizedBenchmark.Rmd
Add covariate boxplot

@@
 
+```{r, schubert-genus-covariates}
+methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
+plotCovariateBoxplots( sb, alpha=0.1, nsets=6, methods=methods)
+```
+
 ```{r, eval=FALSE}
@@
 ```
+```{r, goodrich-genus-covariates}
+methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
+plotCovariateBoxplots( sb, alpha=0.1, nsets=6, methods=methods)
+```
 
@@
 ```
+```{r}
+methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
+plotCovariateBoxplots( sb, alpha=0.1, nsets=6, methods=methods)
+```

### 2018-01-19 datasets/simulations/analysis-05-plot-psim-roc-upsetr.Rmd
Create function for generating aggregated UpSet plots

@@
 
-```{r}
-## loop over simulation settings
-for (f in fl) {
-
-    ## read in list of SummarizedBenchmark objects (100 replications) 
-    res <- readRDS(f)
-
-    ## determine significant hits at 0.10 cutoff
-    hits_tabs <- lapply(res, function(x) {
-        as.tibble(cbind((assay(x, "qvalue") < 0.10) + 0,
-                        truth = rowData(x)$qvalue)) %>%
-            select(-starts_with("ihw-a0"),
-                   -`bl-df02`,  -`bl-df04`,  -`bl-df05`) %>%
-            rename(ihw = `ihw-a10`, bl = `bl-df03`,
-                   scott_e = `scott-empirical`,
-                   scott_t = `scott-theoretical`) %>%
-            as.data.frame
-    })
-
-    ## count up frequencies
-    freq_tabs <- lapply(hits_tabs,
-                        function(x) {
-                            UpSetR:::Counter(x, 11, 1, names(x), nintersections = 2^11,
-                                             mbar_color = "gray23",
-                                             order_mat = "degree", aggregate = "degree",
-                                             cut = NULL, empty_intersects = TRUE,
-                                             decrease = TRUE)
-                        })
-
-    ## merge all frequencies into single table
-    freq_tab <- freq_tabs %>%
-        mapply(function(itab, idx) { as.tibble(itab) %>%
-                                         select(-x, -color) %>%
-                                         rename(!!(paste0("freq.", idx)) := freq) },
-               itab = ., idx = 1:length(freq_tabs),
-               SIMPLIFY = FALSE) %>%
-        reduce(left_join, by = c("unadjusted", "bonf", "bh", "qvalue", "ihw",
-                                 "ashs", "bl", "lfdr", "scott_t", "scott_e", "truth"))
-
-    ## summarize across 100 replications of each setting
-    freq_tab <- freq_tab %>%
-        gather(repl, cnt, starts_with("freq")) %>%
-        group_by(unadjusted, bonf, bh, qvalue, ihw,
-                 ashs, bl, lfdr, scott_t, scott_e, truth) %>%
-        summarize(freq_mean = round(mean(cnt)),
-                  freq_50p = round(median(cnt)),
-                  freq_10p = round(quantile(cnt, .1)),
-                  freq_90p = round(quantile(cnt, .9))) %>%
-        ungroup %>%
-        mutate(setname = paste(ifelse(unadjusted, "unadjusted", NA),
-                               ifelse(bonf, "bonf", NA),
-                               ifelse(bh, "bh", NA),
-                               ifelse(qvalue, "qvalue", NA),
-                               ifelse(ihw, "ihw", NA),
-                               ifelse(ashs, "ashs", NA),
-                               ifelse(bl, "bl", NA),
-                               ifelse(lfdr, "lfdr", NA),
-                               ifelse(scott_t, "scott_t", NA),
-                               ifelse(scott_e, "scott_e", NA),
-                               ifelse(scott_e, "truth", NA),
-                               sep = '&'),
-               setname = gsub("NA", "", setname) %>%
-                   gsub("&{2,}", "&", .) %>%
-                   gsub("^&", "", .) %>%
-                   gsub("&$", "", .)) %>%
-        select(setname, freq_mean,
-               freq_50p, freq_10p, freq_90p)
-
-    ## convert to vector to pass to UpSetR package
-    freq_list <- freq_tab$freq_mean
-    names(freq_list) <- freq_tab$setname
-
-    ## make simple upset plot - save as png
-    outdir <- gsub("data-", "plot-", dirname(f))
-    dir.create(outdir, showWarnings = FALSE, recursive = TRUE)
-    png(file.path(outdir, gsub("\\.rds", "-mean-upset.png", basename(f))),
-        width = 10, height= 5, unit = "in", res = 300)
-
-    upset(fromExpression(freq_list),
-          nsets = 11,
-          mb.ratio = c(0.55, 0.45),
-          order.by = "freq",
-          decreasing = TRUE,
-          set.metadata = list(data = data.frame(sets = c("unadjusted", "bonf", "bh", "qvalue", "ihw",
-                                                         "ashs", "bl", "lfdr", "scott_t", "scott_e", "truth"),
-                                                isTruth = c(rep(FALSE, 10), TRUE)),
-                              plots = list(
-                                  list(type = "matrix_rows", 
-                                       column = "isTruth",
-                                       colors = c("TRUE" = "blue", "FALSE" = "gray"), 
-                                       alpha = 0.2))))
-    dev.off()
-}
-```
-
 <!-- ######################################################################## -->

### 2018-01-19 datasets/R/plotsim_aggupset.R

@@ -1,0 +1,165 @@
+#' Aggregated UpSet Plot
+#'
+#' Generate an UpSet plot showing the (rounded) median overlap between methods
+#' across a collection of SummarizedBenchmark objects, e.g. corresponding
+#' to simulation replicates.
+#' 
+#' @param res list of SummarizedBenchmark objects to be combined in the plot.
+#' @param alpha significance threshold to use for distinguishing significant
+#'        and non-significant tests.
+#' @param supplementary logical whether plot is for supplementary materials.
+#'        (default = FALSE)
+#' @param return_list logical whether frequency list should be returned instead
+#'        of upset plot. The returned list can be used to generate the upset plot
+#'        using `upset(fromExpression(freq_list)`. This can be useful if the user
+#'        wants to experiment with upset plot styles. (default = FALSE)
+#' 
+#' @return
+#' an upset plot if `return_list = FALSE` (default), else a list of frequencies that
+#' can be used to generate the same upset plot.
+#'
+#' @details
+#' Note: this can get incredibly slow if the number of methods being compared is
+#' large since the possible number of overlaps grows exponentially. Anecdotally,
+#' in simulations comparing 9 methods (+ truth) with 20,000 tests takes approximately
+#' 15 to 20 seconds for 20 replications, and 40 to 45 seconds with 100 replications.  
+#' 
+#' @import dplyr magrittr
+#' @author Patrick Kimes
+aggupset <- function(res, alpha, supplementary = FALSE, return_list = FALSE) { 
+
+    ## find significant hits at alpha cutoff for all replicates
+    hits_tabs <- lapply(res, sb2hits, a = alpha, s = supplementary)
+
+    ## count up frequencies in each intersection
+    n_cols <- unique(sapply(hits_tabs, ncol))
+    if (length(n_cols) > 1) {
+        stop("not all SummarizedBenchmarks have the same set of methods")
+    }
+    freq_tabs <- lapply(hits_tabs, hits2freq, nm = n_cols)
+
+    ## merge all freqs into single table - first rename 'freq' columns to 'freq.i' (i = 1..100)
+    method_names <- setdiff(names(freq_tabs[[1]]), "freq")
+    freq_tab <- mapply(function(itab, idx) { dplyr::rename(itab, !!(paste0("freq.", idx)) := freq) },
+                       itab = freq_tabs, idx = 1:length(freq_tabs),
+                       SIMPLIFY = FALSE) %>%
[…]

### 2018-01-19 datasets/simulations/analysis-05-psim-subsetonly.Rmd
Update psim analysis

---
title: "Benchmarking FDR Methods - Simulation Summarized ROC / UpsetR Plots"
author: "Rafalab Journal Club Members"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    highlight: tango
    number_sections: true
    code_folding: hide
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, fig.width = 14, fig.height = 6)
```

<!-- ######################################################################## -->
## Workspace Setup
<!-- ######################################################################## -->

```{r, load-libraries, message=FALSE}
library("SummarizedBenchmark")
library("tidyverse")
```

```{r, helper-functions, message=FALSE}
source("../R/plotsim_aggupset.R")
```

In contrast to the t-test simulations, underlying data was not simulated in this series of simulations. Instead, for each simulation, a collection (20,000) of test statistics
were directly sampled from univariate distributions (Normal, non-central-t, or non-central-chi-squared). Test statistics corresponding to null tests are sampled with non-centrality
parameter zero, and non-null tests are sampled with some non-zero non-centrality parameter, determined based on the simulation setting. In these simulations, the magnitude of
the non-centrality parameter for each test can be understood as the effect size or signal in that test. The univariate distributions (Normal, t, chi-squared) are the variyng
sampling distributions that may be encountered under different testing scenarios (e.g. Student's t-test or Pearson's goodness-of-fit).

Since ASH requires standard errors, the true standard errors is provided to the method.

All simulation settings were carried out with an independent covariate. Both settings where the covariate was informative and uninformative were considered. Additionally, several different informative covariates were considered. Following the approach of Boca and Leek (2015) and Chen, Robinson and Storey (2017), we sampled the independent covariate for each test uniformly from the interval [0, 1]. Whether the covariate was informative or uninformative was determined by a functional relationship between the value of the covariate and the probability of the corresponding test being null. The following relationships were considered:

- pi0(x) = pi0 (`uniform-80`)
- pi0(x) = pi0 (`uniform-90`)
- pi0(x) = pi0 (`uniform-95`)
- pi0(x) = 0.25*I(x < 0.25) + ... (`step-more`, 80% pi0)
- pi0(x) = 0.25*I(x < 0.25) + ... (`step-90`, 90% pi0)
- pi0(x) = 0.25*I(x < 0.25) + ... (`step-95`, 95% pi0)
- pi0(x) = 1 - x^3 ... (`cubic`, 80% pi0)
- pi0(x) = 1 - x^3 ... (`cubic-90`, 90% pi0)
- pi0(x) = 1 - x^3 ... (`cubic-95`, 95% pi0)

Here, we only consider a subset of the entire set of simulation settings. Namely, we look at: [1] null simulations, [2] non-null simulations with an informative covariate (cubic), and [3] non-null simulations with an uninformative covariate. For each of the 3 simulation sets, we further consider three different sampling distributions of the test statistics (Normal, t, chi-squared).

Performance is evaluated with: the false positive rate (FPR), true positive rate (TPR), and method overlaps (upset plots). For null simulations, since FPR and TPR are useless, we instead plot the proportion of rejected tests.

<!-- ######################################################################## -->
## Local Helper Functions
<!-- ######################################################################## -->

```{r}
preclean <- function(res, merge_ihw = TRUE, supplementary = FALSE) {
    ## set of method names
    method_names <- unique(res$blabel)
    ## remove unadjusted p-values - too extreme
    if ("unadjusted" %in% method_names) {
        res <- res %>%
            filter(blabel != "unadjusted")
    }
    ## drop ihw not matching 'alpha'
    if (merge_ihw) {
        res <- res %>%
            mutate(ihw_alpha = gsub("ihw-a(.*)", "\\1", blabel),
                   ihw_alpha = ifelse(grepl("^ihw-a", blabel), ihw_alpha, alpha*100),
                   ihw_alpha = as.numeric(ihw_alpha) / 100) %>%
            filter(abs(alpha - ihw_alpha) < 1e-6) %>%
            mutate(blabel = ifelse(grepl("^ihw-a", blabel), "ihw", blabel))
    }
    ## drop excess Boca-Leek results
    if(!supplementary) {
        methods_bl <- grep("^bl-", method_names, value = TRUE)
        res <- res %>%
            filter(!(blabel %in% setdiff(methods_bl, "bl-df03")))
    }
    return(res)
}

myplot <- function(res, metrics = c("TPR", "FPR"), alpha_max = 1) {
    gp <- res %>%
        filter(performanceMetric %in% metrics) %>%
        filter (alpha <= alpha_max) %>%
        group_by(alpha, blabel, performanceMetric) %>%
        summarize(value50 = median(value, na.rm = TRUE),
                  value10 = quantile(value, 0.10, na.rm = TRUE),
                  value90 = quantile(value, 0.90, na.rm = TRUE)) %>%
        ungroup %>%
        ggplot(aes(x = alpha, y = value50, ymin = value10, ymax = value90,
                   group = blabel, color = blabel))
    if ("FPR" %in% metrics) {
        gp <- gp +
            geom_abline(aes(intercept = y, slope = s), alpha = 1/2,
                        data = data.frame(y = 0, s = 1, performanceMetric = "FPR"))
    }
    gp <- gp +
        geom_line(alpha = 1/4,
                  position = position_dodge(width = 0.003)) +
        geom_errorbar(width = .01,
                      position = position_dodge(width = 0.003)) + 
        geom_point(alpha = 1/2, position = position_dodge(width = 0.003)) +
        facet_wrap(~ performanceMetric) +
        scale_color_discrete("Method") + 
        ylab("performance across replicates") +
        xlab("alpha cutoff") +
        theme_bw()
    gp 
}
```

<!-- ######################################################################## -->
## Null Simulations
<!-- ######################################################################## -->

First, we look at the null simulation settings to assess which methods control false discoveries when no truly significant tests are present. The TPR is not reported here because there are no "true positivies" in the null setting.

- *null* (pi0 = 100) simulations across *noise distributions*
  1. *null* (pi0 = 100) simulations across *Gaussian*
  2. *null* (pi0 = 100) simulations across *t*
  3. *null* (pi0 = 100) simulations across *chi-squared*

### Normal

```{r, fig.width = 7, fig.height = 6}
## common suffix 
rds_name <- "M100/allnull-uniform/results-altnoise_normal_null-M100.rds"

## read in performance metrics, plot metric
mets <- readRDS(file.path("data-psim-parsed", rds_name))
mets <- preclean(mets)
myplot(mets, metrics = "rejectprop") + 
    ggtitle("Distribution: normal, pi0: 100% (100 replications)")
```

### t (10 dof)

```{r, fig.width = 7, fig.height = 6}
## common suffix 
rds_name <- "M100/allnull-uniform/results-altnoise_t_df10_null-M100.rds"

## read in performance metrics, plot metric
mets <- readRDS(file.path("data-psim-parsed", rds_name))
mets <- preclean(mets)
myplot(mets, metrics = "rejectprop") +
    ggtitle("Distribution: t (10 dof), pi0: 100% (100 replications)")
```

### chi-sq (4 dof)

```{r, fig.width = 7, fig.height = 6}
## common suffix 
rds_name <- "M100/allnull-uniform/results-altnoise_chisq_df4_null-M100.rds"

## read in performance metrics, plot metric
mets <- readRDS(file.path("data-psim-parsed", rds_name))
mets <- preclean(mets)
myplot(mets, metrics = "rejectprop") +
    ggtitle("Distribution: chi-squared (4 dof), pi0: 100% (100 replications)")
```

<!-- ######################################################################## -->
## Informative Covariate Simulations
<!-- ######################################################################## -->

Next, we consider simulations where a truly informative covariate is present.

- *cubic-informative*, *non-null* (pi0 = 80,90,95) simulations across *noise distributions*
  1. *cubic-informative*, *non-null* (pi0 = 80,90,95) simulations across *Gaussian*
  2. *cubic-informative*, *non-null* (pi0 = 80,90,95) simulations across *t*
  3. *cubic-informative*, *non-null* (pi0 = 80,90,95) simulations across *chi-squared*

### Normal

```{r}
## common suffix 
rds_name <- "M100/esize_fixed-bl-cubic-90-0.8/results-alteff_30-M100.rds"

## read in performance metrics, plot metric
mets <- readRDS(file.path("data-psim-parsed", rds_name))
mets <- preclean(mets)
myplot(mets) + 
    ggtitle("Distribution: normal, pi0: 90% (100 replications)")

## read in SB objects, plot aggregate upset plot
res <- readRDS(file.path("data-psim", rds_name))
aggupset(res, alpha = 0.05)
```

Also looking at varying pi0s (only showing FPR/TPR).

```{r}
"M100/esize_fixed-bl-cubic/results-alteff_30-M100.rds" %>%
    file.path("data-psim-parsed", .) %>%
    readRDS() %>%
    preclean() %>%
    myplot() + 
    ggtitle("Distribution: normal, pi0: 80% (100 replications)")

"M100/esize_fixed-bl-cubic-95-0.8/results-alteff_30-M100.rds" %>%
    file.path("data-psim-parsed", .) %>%
    readRDS() %>%
    preclean() %>%
    myplot() + 
    ggtitle("Distribution: normal, pi0: 95% (100 replications)")
```

### t (10 dof)

```{r}
rds_name <- "M100/altnoise-bl-cubic-90-0.8/results-altnoise_t_df10_bimodal-M100.rds"

## read in performance metrics, plot metric
mets <- readRDS(file.path("data-psim-parsed", rds_name))
mets <- preclean(mets)
myplot(mets) +
    ggtitle("Distribution: t (10 dof), pi0: 90% (100 replications)")

## read in SB objects, plot aggregate upset plot
res <- readRDS(file.path("data-psim", rds_name))
aggupset(res, alpha = 0.05)
```

Also looking at varying pi0s (only showing FPR/TPR).

```{r}
"M100/altnoise-bl-cubic/results-altnoise_t_df10_bimodal-M100.rds" %>%
    file.path("data-psim-parsed", .) %>%
    readRDS() %>%
    preclean() %>%
    myplot() + 
    ggtitle("Distribution: t (10 dof), pi0: 80% (100 replications)")

"M100/altnoise-bl-cubic-95-0.8/results-altnoise_t_df10_bimodal-M100.rds" %>%
    file.path("data-psim-parsed", .) %>%
    readRDS() %>%
    preclean() %>%
    myplot() + 
    ggtitle("Distribution: t (10 dof), pi0: 95% (100 replications)")
```

### chi-sq (4 dof)

```{r}
## common suffix 
rds_name <- "M100/altnoise-bl-cubic-90-0.8/results-altnoise_chisq_df4_shift3sq-M100.rds"

## read in performance metrics, plot metric
mets <- readRDS(file.path("data-psim-parsed", rds_name))
mets <- preclean(mets)
myplot(mets) +
    ggtitle("Distribution: chi-squared (4 dof), pi0: 90% (100 replications)")

## read in SB objects, plot aggregate upset plot
res <- readRDS(file.path("data-psim", rds_name))
aggupset(res, alpha = 0.05)
```

Also looking at varying pi0s (only showing FPR/TPR).

```{r}
"M100/altnoise-bl-cubic/results-altnoise_chisq_df4_shift3sq-M100.rds" %>%
    file.path("data-psim-parsed", .) %>%
    readRDS() %>%
    preclean() %>%
    myplot() + 
    ggtitle("Distribution: chi-squared (4 dof), pi0: 80% (100 replications)")

"M100/altnoise-bl-cubic-95-0.8/results-altnoise_chisq_df4_shift3sq-M100.rds" %>%
    file.path("data-psim-parsed", .) %>%
    readRDS() %>%
    preclean() %>%
    myplot() + 
    ggtitle("Distribution: chi-squared (4 dof), pi0: 95% (100 replications)")
```

<!-- ######################################################################## -->
## Uninformative Covariate Simulations
<!-- ######################################################################## -->

Finally, we consider simulations with non-null (alternative) tests are present.

- *non-informative*, *non-null* (pi0 = 80,90,95) simulations across *noise distributions*
  1. *uninformative*, *non-null* (pi0 = 80,90,95) simulations across *Gaussian*
  2. *uninformative*, *non-null* (pi0 = 80,90,95) simulations across *t*
  3. *uninformative*, *non-null* (pi0 = 80,90,95) simulations across *chi-squared*

### Normal

```{r}
## common suffix 
rds_name <- "M100/esize_fixed-uniform-0.9/results-alteff_30-M100.rds"

## read in performance metrics, plot metric
mets <- readRDS(file.path("data-psim-parsed", rds_name))
mets <- preclean(mets)
myplot(mets) + 
    ggtitle("Distribution: normal, pi0: 90% (100 replications)")

## read in SB objects, plot aggregate upset plot
res <- readRDS(file.path("data-psim", rds_name))
aggupset(res, alpha = 0.05)
```

Also looking at varying pi0s (only showing FPR/TPR).

```{r}
"M100/esize_fixed-uniform/results-alteff_30-M100.rds" %>%
    file.path("data-psim-parsed", .) %>%
    readRDS() %>%
    preclean() %>%
    myplot() + 
    ggtitle("Distribution: normal, pi0: 80% (100 replications)")

"M100/esize_fixed-bl-uniform-0.95/results-alteff_30-M100.rds" %>%
    file.path("data-psim-parsed", .) %>%
    readRDS() %>%
    preclean() %>%
    myplot() + 
    ggtitle("Distribution: normal, pi0: 95% (100 replications)")

"M100/esize_fixed-bl-uniform-0.99/results-alteff_30-M100.rds" %>%
    file.path("data-psim-parsed", .) %>%
    readRDS() %>%
    preclean() %>%
    myplot() + 
    ggtitle("Distribution: normal, pi0: 99% (100 replications)")
```

### t (10 dof)

```{r}
rds_name <- "M100/altnoise-uniform/results-altnoise_t_df10_bimodal-M100.rds"

## read in performance metrics, plot metric
mets <- readRDS(file.path("data-psim-parsed", rds_name))
mets <- preclean(mets)
myplot(mets) +
    ggtitle("Distribution: t (10 dof), pi0: 80% (100 replications)")

## read in SB objects, plot aggregate upset plot
res <- readRDS(file.path("data-psim", rds_name))
aggupset(res, alpha = 0.05)
```

*Need to run with different pi0 values.*

### chi-sq (4 dof)

```{r}
rds_name <- "M100/altnoise-uniform/results-altnoise_chisq_df4_shift3sq-M100.rds"

## read in performance metrics, plot metric
mets <- readRDS(file.path("data-psim-parsed", rds_name))
mets <- preclean(mets)
myplot(mets) +
    ggtitle("Distribution: chi-squared (4 dof), pi0: 80% (100 replications)")

## read in SB objects, plot aggregate upset plot
res <- readRDS(file.path("data-psim", rds_name))
aggupset(res, alpha = 0.05)
```

*Need to run with different pi0 values.*

### 2018-01-19 datasets/simulations/analysis-05-psim-subsetonly.Rmd
Fix typo

@@
 
-"M100/esize_fixed-bl-uniform-0.95/results-alteff_30-M100.rds" %>%
+"M100/esize_fixed-uniform-0.95/results-alteff_30-M100.rds" %>%
     file.path("data-psim-parsed", .) %>%
@@
 
-"M100/esize_fixed-bl-uniform-0.99/results-alteff_30-M100.rds" %>%
+"M100/esize_fixed-uniform-0.99/results-alteff_30-M100.rds" %>%
     file.path("data-psim-parsed", .) %>%

### 2018-02-02 datasets/RNA-Seq/yeast48-replicates-benchmarkFDR.Rmd
Add yeast simulation with replicates

---
title: "Yeast RNA-seq 48 sample simulation study"
author: "Keegan Korthauer"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This is an analysis of the yeast data with 48 biological replicates in each of 
two conditions (analyzed in this [publication](https://www.ncbi.nlm.nih.gov/pubmed/26206307/)). 
We chose this experiment because of the large number of biological replicates, which
will allow us to (1) implement null comparisons on random subsets of samples within
one condition, (2) compare FDR approaches on their ability to discover 'true positive' 
DE genes using subsets of the full dataset, where 'true positive' is defined as a 
gene detected as DE using all 48 samples in each condition, and (3) start with a null
comparison and add in artificial differences to a subset of genes to define
'true positives'.

Processed count table is made available by the authors in their paper codebase 
[GitHub repository](https://github.com/bartongroup/profDGE48).

In this Rmd we will carry out the steps in `yeast48-benchmarkFDR.Rmd` for multiple
replicates, and plot results averaging over the replications.

# Set up workspace

```{r, results='hide', message=FALSE}
# Load packages and source benchmark FDR
library(data.table)
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(magrittr)
library(R.utils)
library(cowplot)
library(purrr)
library(DESeq2)
library(tibble)
library(ggthemes)

# benchmark methods
library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)
library(FDRreg)

# comparison tools/functions
library(SummarizedBenchmark)
sourceDirectory("../R/")

# set up parallel backend
library(parallel)
cores <- 20
```

# Data download

First, we download the processed count data from GitHub. There is one file for the 
Snf2 condition and one file for the wild type condition. The Snf2 condition is a yeast
strain that has the transcriptional regulator gene Snf2 knocked out. Each of these
files is a compressed `tar.gz` archive that contains a 
single bam file for each replicate in each condition, which we'll place in a subdirectory
called `data`.

```{r, download}
system("wget https://github.com/bartongroup/profDGE48/raw/master/Preprocessed_data/Snf2_countdata.tar.gz")
system("wget https://github.com/bartongroup/profDGE48/raw/master/Preprocessed_data/WT_countdata.tar.gz")
system("mkdir yeast48-data")
system("tar -xvf WT_countdata.tar.gz -C ./yeast48-data")
system("tar -xvf Snf2_countdata.tar.gz -C ./yeast48-data")
system("rm *.tar.gz")
```

Each of the data files contains two columns, one with a gene/feature name, and one
with the count value.

## Read data into R and create a count table

Here we make use of the map and map2 functions in the `purrr` package, to swiftly 
apply the `read_tsv` function from `readr` to read in all of the 96 sample tables, 
as well as add in the sample name (derived from the file name) to each subtable.
Finally, the `reduce` function is used to join all the replicates together.

```{r, readin, results='hide', message=FALSE}
files <- dir(path = "./yeast48-data/", pattern = "*.bam.gbgout", full.names = TRUE)
sample_names <- sapply(strsplit(dir(path = "./yeast48-data/", pattern = "*.bam.gbgout"), "_MID"),
                       function(x) x[[1]])

counts <- files %>%
  purrr::map(read_tsv, col_names = FALSE) %>% # read in all the files individually
  purrr::map2(sample_names, ~ dplyr:::rename(.x, !! .y := X2, feature = X1) ) %>% # add sample names
  purrr::reduce(left_join, by = "feature") # reduce with rbind into one dataframe
```

# Analysis of full 48v48 (non-null)

Here we'll carry out an analysis of the full dataset, comparing the controls to the
knowckout samples. The fold changes observed in this comparison will be used when 
generating the non-null simulated data in the following sections.

We're ready to construct a DESeq2 object. First we pull out the feature names and
add them as rownames for the count table, and next we construct a column data object
that houses the sample names, replicate numbers, and condition factor (WT versus
Snf2 knockout).

## Set up DESeq2 object

```{r, deseq}
feats <- (counts %>% select(1))$feature
counts <- as.matrix(counts %>% select(-1))
rownames(counts) <- feats

coldat <- tibble(sample=colnames(counts)) %>% 
  separate(sample, sep="_", into=c("condition", "replicate"), remove=FALSE) %>%
  mutate(condition = factor(condition))

# filter low count genes
counts <- counts[rowMeans(counts) > 1,]

dds <- DESeqDataSetFromMatrix(countData = counts,
                              colData = coldat,
                              design= ~ condition)

# results on full set
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients
resFULL <- results(dds, name="condition_WT_vs_Snf2")

head(resFULL)

sum(resFULL$padj < 0.05, na.rm=T, independentFiltering = F)
```

## Check assumptions

There are many thousand genes detected using the full set. Next, we'll build input data frame for summarized benchmark.

```{r}
geneExp <- tbl_df(data.frame(geneName=rownames(resFULL), 
                             pval=resFULL$pvalue, 
                             SE=resFULL$lfcSE,
                             ind_covariate = resFULL$baseMean, 
                             effect_size=resFULL$log2FoldChange, 
                             test_statistic=resFULL$stat,
                             pzero=rowSums(counts(dds)==0)/ncol(counts(dds))))

# filter NAs and those with less than 50% expressed 
geneExp <- geneExp %>% na.omit() %>% dplyr::filter(pzero < 0.5)
```

Check unimodal assumption (for ash).

```{r}
ggplot(data=geneExp, aes(test_statistic)) +
  geom_histogram(bins=30)
```

Look at covariate diagnostic plots.

```{r, width=15, height=15}
strat_hist(geneExp, pvalue="pval", covariate="ind_covariate", maxy=32)
```

```{r}
rank_scatter(geneExp, pvalue="pval", covariate="ind_covariate")
```

## FDR benchmarking

Run benchmark methods.

```{r}
bd <- initializeBenchDesign()
sb <- bd %>% buildBench(data=geneExp, parallel = F)

assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Plot results.

```{r}
rejections_scatter(sb, supplementary=FALSE)
```

```{r}
plotFDRMethodsOverlap(sb, alpha=0.05, nsets=ncol(sb), order.by="freq", decreasing=TRUE, supplementary=FALSE)

```

# Simulation set up

Next, we'll analyze random splits of one condition, both with and without the 
addition of simulated DE genes. Here we'll create a function that we can use to 
run one replicate given sample size and number of DE gene settings. This will be
looped over many replications and results averaged over them.

```{r}
#' @param X simnumber
#' @param seed random seed
#' @param dds is the full DESeq2 dataset
#' @param resFULL is the results object from running DESeq2 on dds (used in
#'   to obtain the fold change effect sizes when adding non-null genes)
#' @param sampleSize is the number of samples in each condition
#' @param nDE is the number of DE genes
#' @param bd is the bench design object
simulateOneSplit <- function(X, rseed, dds, resFULL, nDE, sampleSize, bd){

  # set random seed
  set.seed(as.numeric(X)*as.numeric(rseed))

  # select a random subset of 20 WT samples
  dds_test <- dds[,colData(dds)$condition == "Snf2"]
  dds_test <- dds_test[,sample(1:ncol(dds_test), sampleSize*2)]

  # add a fake condition column to coldat
  colData(dds_test)$fake <-  factor(c(rep("A", sampleSize), 
                                      rep("B", sampleSize))[sample(1:(sampleSize*2), 
                                                                   sampleSize*2)])
  design(dds_test) <- ~fake
  truth <- rep(FALSE, nrow(dds_test))

  if(nDE > 0){
    # pick random set of nDE genes to add signal to
    DE <- sample(1:nrow(dds_test), nDE)
    truth <- rep(FALSE, nrow(dds_test))
    truth[DE] <- TRUE
    names(truth) <- rownames(dds_test)

    # randomly sample a log2FC from original FCs (without regard to DE)
    log2FC <- rep(0, nrow(dds_test))
    log2FC[DE] <- resFULL$log2FoldChange[DE]

    counts_new <- counts(dds_test)
    counts_new[,colData(dds_test)$fake=="A"] <- counts(dds_test)[,colData(dds_test)$fake=="A"] * 2^log2FC
    counts_new <- apply(counts_new, 2, as.integer)
    counts(dds_test) <- counts_new
  }

  dds_test <- DESeq(dds_test)
  resultsNames(dds_test) # lists the coefficients
  resTEST <- results(dds_test, name="fake_B_vs_A", independentFiltering = F)

  geneExp <- tbl_df(data.frame(geneName=rownames(resTEST), 
                               pval=resTEST$pvalue, 
                               SE=resTEST$lfcSE, 
                               ind_covariate = resTEST$baseMean,
                               effect_size = resTEST$log2FoldChange, 
                               test_statistic = resTEST$stat,
                               qvalue = truth,
                               pzero = rowSums(counts(dds_test)==0)/
                                       ncol(counts(dds_test))  ))

  # filter NAs and those with less than 50% expressed 
  geneExp <- geneExp %>% na.omit() %>% dplyr::filter(pzero < 0.5)

  sb <- bd %>% buildBench(data=geneExp, parallel = F, truthCols = "qvalue")

  return(sb)
}

```

We'll also set some parameters that will be common to all simulations. These
include the number of replications, the bench design object, the set of 
methods to exclude in the results plots, and the alpha cutoff level to 
be used when plotting the aggregated Upset results.

```{r}
B <- 100
bd <- initializeBenchDesign() # only needs to be done once
excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
ualpha <- 0.05
```

Here's a helper function to return the number of methods with rejections at
a particular alpha level (this helps us determine whether or not to plot the
aggregated upset plot - if there aren't at least 2 methods it will throw an
error, which is a problem for the null simulations).

```{r}
#' @param res standardized metric data.table generated using
#'        standardize_results.
#' @param alpha alpha cutoff
#' @param filterSet which methods to exclude from consideration 
numberMethodsReject <- function(res, alphacutoff, filterSet){
  res <- res %>% 
    filter(is.na(param.alpha) | (param.alpha == alphacutoff)) %>%
    filter(!(blabel %in% filterSet)) %>%
    filter(alpha == alphacutoff) %>%
    filter(performanceMetric == "rejections") %>%
    select(blabel, performanceMetric, value) %>%
    group_by(blabel) %>%
    summarize(mean_value = mean(value)) %>%
    filter(mean_value > 0)
  return(nrow(res))
}

```

# Sim N5: Null Comparison 5v5

Here we'll repeat the above, but for a null comparison 5 versus 5 Snf2 samples,
where the groups are selected randomly. This will be done for 100 random splits.

## Generate a list of SB results objects

```{r, results='hide', message=FALSE}
rseed <- 225
sampleSize <- 5
nDE <- 0

null5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd,
                   mc.cores=cores)
```

## Plot average results over replications

Plot results.

```{r}
res5 <- plotsim_standardize(null5, alpha = seq(0.01, 0.20, 0.01))

# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
plotsim_average(res5, met="rejections",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE) 
plotsim_average(res5, met="TNR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE)

if (numberMethodsReject(res5, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
  aggupset(null5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
}else{
  message("Not enough methods found rejections at alpha ", ualpha, 
          "; skipping upset plot")
}

```

# Sim N10: Null Comparison 10v10

Here we'll repeat the above, but for a null comparison 10 versus 10 Snf2 samples,
where the groups are selected randomly. This will be done for 100 random splits.

## Generate a list of SB results objects

```{r, results='hide', message=FALSE}
rseed <- 837
sampleSize <- 10
nDE <- 0

null10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd,
                   mc.cores=cores)
```

## Plot average results over replications

Plot results.

```{r}
res10 <- plotsim_standardize(null10, alpha = seq(0.01, 0.20, 0.01))

# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
plotsim_average(res10, met="rejections",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE) 
plotsim_average(res10, met="TNR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE)

if (numberMethodsReject(res10, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
  aggupset(null10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
}else{
  message("Not enough methods found rejections at alpha ", ualpha, 
          "; skipping upset plot")
}

```

# Sim D5: DE Comparison 5v5

Here we'll repeat the above, but for a DE comparison 5 versus 5 Snf2 samples,
where the groups are selected randomly and 500 DE genes are added. 
This will be done for 100 random splits.

## Generate a list of SB results objects

```{r, results='hide', message=FALSE}
rseed <- 198
sampleSize <- 5
nDE <- 500

de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd,
                   mc.cores=cores)
```

## Plot average results over replications

Plot results.

```{r}
res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.20, 0.01))

# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
plotsim_average(res5d, met="rejections",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE) 
plotsim_average(res5d, met="FPR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE) 
plotsim_average(res5d, met="TPR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE)

if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
  aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
}else{
  message("Not enough methods found rejections at alpha ", ualpha, 
          "; skipping upset plot")
}

```

# Sim D10: DE Comparison 10v10

Here we'll repeat the above, but for a DE comparison 10 versus 10 Snf2 samples,
where the groups are selected randomly and 500 DE genes are added.
This will be done for 100 random splits.

## Generate a list of SB results objects

```{r, results='hide', message=FALSE}
rseed <- 961
sampleSize <- 10
nDE <- 500

de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd,
                   mc.cores=cores)
```

## Plot average results over replications

Plot results.

```{r}
res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.20, 0.01))

# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
plotsim_average(res10d, met="rejections",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE) 
plotsim_average(res10d, met="FPR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE) 
plotsim_average(res10d, met="TPR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE)

if (numberMethodsReject(res10d, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
  aggupset(de10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
}else{
  message("Not enough methods found rejections at alpha ", ualpha, 
          "; skipping upset plot")
}

```

# Save results objects

```{r}
system("mkdir -p yeast48-results")
saveRDS(null5, file="yeast48-results/null5.rds")
saveRDS(null10, file="yeast48-results/null10.rds")
saveRDS(de5, file="yeast48-results/de5.rds")
saveRDS(de10, file="yeast48-results/de10.rds")
```

# Session information

```{r}
sessionInfo()
```

### 2018-02-09 datasets/scrnaseq/scRNAseq-latest.Rmd
Add files via upload

---
title: "scRNAseq"
author: "Ayshwarya Subramanian"
date: "9/15/2017"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE)

```
## scRNAseq datasets

For the application of differential gene expression, we look at scRNAseq and bulk RNAseq data. Here we discuss the analyses of scRNAseq.

```{r echo=FALSE, results='asis'}
library(knitr)
df=data.frame(datasets=c("GSE63818-GPL16791","GSE45719"),organism=c("human","mouse"),technology=c("smart-seq","smart-seq"))
kable(df,caption="scRNAseq Datasets")
df1=data.frame(DE_method=c("Wilcoxon-rank sum","ScDD","MAST"),input=c("tpm","count","tpm"),output=c("p-val","p-val","p-val"))
kable(df1,caption="DE methods")

```

We selected two datasets, one each from mouse and human spp, from the `Conquer` (Soneson et al., 2017) scRNAseq data collection. `Conquer` provides uniformly processed gene expression data summarized to both count (aggregated from transcript members), and length-scaled TPMs at the gene level. The authors use `salmon` to quantify raw reads and `tximport` to summarize to gene level. The data comes from Smartseq, a plate based sequencing method which results in deeper sequencing of reads at the cost of numbers of cells.

We also select two droplet-based datasets from the `10x` website:4k PBMC and 4k T cells, both from healthy donors.

## Data Download

We first download the data in our working directory. First, let's download the Human data:GSE63818-GPL16791

```{bash, humandata-download, eval=FALSE}
# all data and results are on the wd: /net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/fdrbenchmark-scrnaseq

#these look well separated on the tsne plots on the conquer manuscript supplement
wget http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE63818-GPL16791.rds
# qc report: 
wget http://imlspenticton.uzh.ch/robinson_lab/conquer/report-multiqc/GSE63818-GPL16791_multiqc_report.html
# scater report
wget http://imlspenticton.uzh.ch/robinson_lab/conquer/report-scater/GSE63818-GPL16791_scater.html
```

Next, we download the Mouse data:GSE45719

```{bash, mousedata-download, eval=FALSE}
#another well-separated dataset on tsne: GSE60749−GPL13112
wget http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE45719.rds
```

We directly download the cell-ranger processed filtered expression matrices [4k PBMC cells](https://support.10xgenomics.com/single-cell-gene-expression/datasets/2.1.0/pbmc4k) and [4k T-cells](https://support.10xgenomics.com/single-cell-gene-expression/datasets/2.1.0/t_4k) from the 10x website.

## R workspace setup

The downloaded plate data are stored in R objects called `MultiAssayExperiment`; this package requires BioConductor release 3.5 and R version 3.4 or greater for installation.

```{r, workspace-setup, eval=FALSE}
#installation
# source("https://bioconductor.org/biocLite.R")
# biocLite("swfdr")
# biocLite("IHW")
# biocLite("qvalue")
# install.packages("ashr")
# install.packages("fdrtool")
# biocLite("edgeR")
# source("https://bioconductor.org/biocLite.R")
# biocLite("scran")
#  Samseq installation
# source("https://bioconductor.org/biocLite.R")
# biocLite("impute") 
# install.packages("samr")
# source("https://bioconductor.org/biocLite.R")
#biocLite("MultiAssayExperiment")
#source("https://bioconductor.org/biocLite.R")
#biocLite("SummarizedExperiment")
#biocLite("BiocGenerics")
#biocLite("S4Vectors")
#biocLite("BiocParallel")
#install_github("areyesq89/SummarizedBenchmark")
#biocLite("EBSeq")
#biocLite("scDD")
#source("https://bioconductor.org/biocLite.R")
#biocLite("SingleCellExperiment")
#source("https://bioconductor.org/biocLite.R")
#biocLite("MAST")

# packages for data analyses
library(MultiAssayExperiment)
library(dplyr)
library(ggplot2)
library(cowplot)
library(UpSetR)
library(hexbin)

# packages for differential expression testing
library(impute)
#library(samr)
library(scDD)
library(EBSeq)
library(edgeR)
library(MAST)

# benchmark methods
library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)

# summarizebenchmark
library(SummarizedBenchmark)
library(SingleCellExperiment)
source("/Users/subraman/Documents/research/fdr_jc/code/fdr-benchmark-scrnaseq-helper.R")

```

## Data processing and filtering

For data processing, we follow practices and recommendations from Soneson et al., 2017. Briefly, we limit ourselves to 2-group comparisons, and  select the maximal size defined as the smallest number of cells in either group as the group-level sample size for the DE analyses.

First, we get the human data for the two groups of interest.

We look at results with and without prefiltering genes (keeping genes with TPM > 1 in atleast 25% of cells)

```{r, dataload, eval=FALSE}
setwd("/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/fdrbenchmark-scrnaseq")
setwd("/Users/subraman/Documents/research/fdr_jc/data/")
#load compressed mae
humandata=readRDS("GSE63818-GPL16791.rds")
#get gene-level data
genes=experiments(humandata)[1]
#TPMS
#get tpms
tpms=assays(genes$gene)[1]
#get tpm dataframe
human_tpms=tpms$TPM
# get column data/metadata for sample
cdata=colData(humandata)
cdata=as.data.frame(cdata)
# select groups of interest
g1 = cdata%>%subset(source_name_ch1=="Somatic Cells" & characteristics_ch1=="developmental stage: 7 week gestation")
g2 = cdata%>%subset(source_name_ch1=="Primordial Germ Cells" & characteristics_ch1=="developmental stage: 7 week gestation")
# select maximal sample size
sample_size = c(dim(g1)[1],dim(g2)[1])
ssize = which.min(sample_size)
g1_ids=sample.int(dim(g1)[1], size = sample_size[ssize], replace = FALSE)
g2_ids=sample.int(dim(g2)[1], size = sample_size[ssize], replace = FALSE)
# get group data of maximal size
g1_data=human_tpms[,colnames(human_tpms)%in%g1$geo_accession[g1_ids]]
g2_data=human_tpms[,colnames(human_tpms)%in%g2$geo_accession[g2_ids]]
#save data
saveRDS(g1_data,"h_g1_data.rds")
saveRDS(g2_data,"h_g2_data.rds")
#COUNTS
#get counts
count=assays(genes$gene)[2]
#get count dataframe
human_count=count$count
# get column data/metadata for sample
cdata=colData(humandata)
cdata=as.data.frame(cdata)
# get group data of maximal size
g1_cdata=human_count[,colnames(human_count)%in%g1$geo_accession[g1_ids]]
g2_cdata=human_count[,colnames(human_count)%in%g2$geo_accession[g2_ids]]
#save data
saveRDS(g1_cdata,"h_g1_cdata.rds")
saveRDS(g2_cdata,"h_g2_cdata.rds")

```

For the human data, each group has 26 cells and 65218 genes. We merge the data and remove all genes which have 0s in all samples

```{r, nofilter, eval=FALSE}
g_data=cbind(g1_data,g2_data)
g_data_sub=g_data[rowSums(g_data==0)!=52,] # 0.25*52=13
saveRDS(g_data_sub,"h_g_nofilter_subdata.rds") # removing rows with all zeros reduces the number of genes to 33169

g_cdata=cbind(g1_cdata,g2_cdata)
g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=52,]
saveRDS(g_cdata_sub,"h_g_nofilter_subcdata.rds") # removing rows with all zeros reduces the number of genes to 33169

saveRDS(g1_ids,"g1_ids.rds")
saveRDS(g2_ids,"g2_ids.rds")
```

```{r,mouse}
##mouse data
mousedata=readRDS("GSE45719.rds")
#get gene-level data
genes=experiments(mousedata)[1]
#TPMS
#get tpms
tpms=assays(genes$gene)[1]
#get tpm dataframe
mouse_tpms=tpms$TPM
# get column data/metadata for sample
cdata=colData(mousedata)
cdata=as.data.frame(cdata)
# select groups of interest
g1 = cdata%>%subset(source_name_ch1=="16-cell stage blastomere")
g2 = cdata%>%subset(source_name_ch1=="Mid blastocyst cell (92-94h post-fertilization)")
# select maximal sample size
sample_size = c(dim(g1)[1],dim(g2)[1])
ssize = which.min(sample_size)
g1_ids=sample.int(dim(g1)[1], size = sample_size[ssize], replace = FALSE)
g2_ids=sample.int(dim(g2)[1], size = sample_size[ssize], replace = FALSE)
# get group data of maximal size
g1_data=mouse_tpms[,colnames(mouse_tpms)%in%g1$geo_accession[g1_ids]]
g2_data=mouse_tpms[,colnames(mouse_tpms)%in%g2$geo_accession[g2_ids]]
#save data
saveRDS(g1_data,"m_g1_data.rds")
saveRDS(g2_data,"m_g2_data.rds")
#COUNTS
#get counts
count=assays(genes$gene)[2]
#get count dataframe
mouse_count=count$count
# get column data/metadata for sample
cdata=colData(mousedata)
cdata=as.data.frame(cdata)
# get group data of maximal size
g1_cdata=mouse_count[,colnames(mouse_count)%in%g1$geo_accession[g1_ids]]
g2_cdata=mouse_count[,colnames(mouse_count)%in%g2$geo_accession[g2_ids]]
#save data
saveRDS(g1_cdata,"m_g1_cdata.rds")
saveRDS(g2_cdata,"m_g2_cdata.rds")

```

For the mouse data, each group has 50 cells and 45686 genes. We merge the data and remove all genes which have 0s in all samples

```{r, nofilter, eval=FALSE}
g_data=cbind(g1_data,g2_data)
g_data_sub=g_data[rowSums(g_data==0)!=dim(g_data)[2],] # 0.25*52=13
saveRDS(g_data_sub,"m_g_nofilter_subdata.rds") # removing rows with all zeros reduces the number of genes to 33169

g_cdata=cbind(g1_cdata,g2_cdata)
g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=dim(g_data)[2],]
saveRDS(g_cdata_sub,"m_g_nofilter_subcdata.rds") # removing rows with all zeros reduces the number of genes to 33169

saveRDS(g1_ids,"m_g1_ids.rds")
saveRDS(g2_ids,"m_g2_ids.rds")
```

## Differential expression testing

We perform differential expression testing using 3 methods on the unfiltered data:

### Wilcoxon rank-sum test

Per `Conquer`, we run TMM normalization to the TPM units before running Wilcoxon.
```{r, wilcoxon, eval=FALSE}
g_data_sub=readRDS("h_g_nofilter_subdata.rds")
groups=c(rep(1,26),rep(2,26))
compute_wilcoxde_sc(g_data_sub,"h_data_wilcox_p.rds",groups)

g_data_sub=readRDS("m_g_nofilter_subdata.rds")
groups=c(rep(1,50),rep(2,50))
compute_wilcoxde_sc(g_data_sub,"m_data_wilcox_p.rds",groups)
## warnings:
## In wilcox.test.default(x = structure(c(0, 0, 14.2107316559397,  ... :
## cannot compute exact p-value with ties

```

### Differences in distributions of gene expression in the 2 conditions

We apply scDD to log-transformed counts as below:

```{r, scdd, eval=FALSE}
# library(SummarizedExperiment)
# condition=c(rep(1,26),rep(2,26))
# rownames(ng_cdata_sub) <- paste0(rownames(ng_cdata_sub), 1:nrow(ng_cdata_sub), sep="")
# colnames(ng_cdata_sub) <- names(condition) <- paste0("Sample",
# 1:ncol(ng_cdata_sub), sep="")
# SDSumExp <- SummarizedExperiment(assays=list("NormCounts"=ng_cdata_sub),
# colData=data.frame(condition))
#saveRDS(SDSumExp,"data_sumexp.rds")

## human data
compute_scdd_sc(g_cdata_sub,"h_sizefactors.rds","h_data_sce.rds","humanplate_scDatEx_ks.rds")
##Notice: 16096 genes have less than 3 nonzero cells per condition.  Skipping these genes.
##Clustering observed expression data for each gene
##Setting up parallel back-end using 6 cores
# perm=1000
# scDatExperm <- scDD(SDSumExp, prior_param=prior_param, param=param, permutations=1000,testZeroes=FALSE)
# saveRDS(scDatExperm,"/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/fdrbenchmark-scrnaseq/scDatExperm.rds")

## mouse data
compute_scdd_sc(g_cdata_sub,"m_sizefactors.rds","m_data_sce.rds","mouseplate_scDatEx_ks.rds")
# Notice: 10246 genes have less than 3 nonzero cells per condition.  Skipping these genes.
# Clustering observed expression data for each gene
# Setting up parallel back-end using 6 cores
# Notice: Number of permutations is set to zero; using 
#             Kolmogorov-Smirnov to test for differences in distributions
#             instead of the Bayes Factor permutation test
# Classifying significant genes into patterns

```

### MAST

```{r, MAST}
names(groups)=colnames(g_cdata_sub)
compute_MASTtpmde(g_data_sub,groups,"mouseplate_masttpm.rds")

g_data_sub=readRDS("h_g_nofilter_subdata.rds")
groups=c(rep(1,26),rep(2,26))
names(groups)=colnames(g_data_sub)
compute_MASTtpmde(g_data_sub,groups,"humanplate_masttpm.rds")
```

## Covariate Diagnostics

We could use a few different covariates: strength of the signal defined as the median gene expression unit for the gene, detection rate or a combination of the two.

```{r,cov,eval=TRUE}
sourceDirectory("/Users/subraman/Documents/research/fdr_jc/src/benchmark-fdr/datasets/R")
source("/Users/subraman/Documents/research/fdr_jc/src/benchmark-fdr/datasets/R/strat_hist.R")
h_cov=compute_covariates_scrnaseq("h_g_nofilter_subdata.rds")
wilcox=readRDS("h_data_wilcox_p.rds")
h_wilcox_cov=data.frame(genes=names(wilcox),pvalue=wilcox)
h_wilcox_cov=h_wilcox_cov%>%left_join(h_cov,by=c("genes"))
dat=data.frame(pvalue=h_wilcox_cov$pvalue,covariate=h_wilcox_cov$median_exp)
p=strat_hist(h_wilcox_cov,pvalue="pvalue",covariate="median_exp",numQ=5)

hmast=readRDS("humanplate_masttpm.rds")
h_mast_cov=
h_scdd_cov=
p=strat_hist()
scdd_m=metadata(scDatEx)
scdd_p_df=data.frame(genes=scdd_m$Genes$gene,scdd_pval=matrix(scdd_m$Genes$nonzero.pvalue))
wilcox_p_df=data.frame(genes=names(wilcox_p),wilcox_pval=matrix(wilcox_p))

```

## Benchmarking covariate adjusted FDR tools for application

```{r, benchmarking, eval=FALSE}
g_cdata_sub_se=apply(g_cdata_sub,1,function(x) sd(x)/sqrt(length(x))) #standard error
g_cdata_sub_es=apply(g_cdata_sub,1,function(x) mean(x[1:26])-mean(x[27:52])) #difference of means

b_dat=data.frame(pval=dat$wilcox_pval, ind_covariate=dat$mean_exp,effect_size=matrix(g_cdata_sub_es),SE=matrix(g_cdata_sub_se))
b_dat=b_dat%>%mutate(zscore = effect_size/SE)
saveRDS(b_dat,"wilcox_meanexp.RDS")
cutoffs <- c(0.01, 0.025, 0.05, 0.10, 0.20)
resfile <- paste0(getwd(), "/scrnaseqhuman_wilcox_meanexp_results_",
                  nrow(b_dat), ".RData")
library(devtools)
source("https://install-github.me/mangothecat/callr")
source("https://install-github.me/r-lib/remotes")
remotes::install_github('jgscott/FDRreg', subdir="R_pkg/")

source("benchmark-fdr/datasets/R/evaluation-helpers_mod.R")
if (!file.exists(resfile)){
  t1 <- proc.time()
  b_dat_results <- run_benchmarks_mod(dat=b_dat, alphas=cutoffs, pvals=TRUE)
  save(b_dat_results, file=resfile)
  message(paste0("Took ", round((proc.time()-t1)[3],1),
                 " seconds for ", nrow(b_dat), " genes"))
}else{
  load(resfile)
}

# plot N rejects vs FDR cutoff

pdf("scrnaseq_wilcox_results.pdf",width=20,height=10)
p <- ggplot(b_dat_results$stats, aes(alpha, n_rejects, color=method)) +
  geom_point() +
  geom_line() +
  xlab("FDR cutoff") + ylab("Number of rejections") +
  ggtitle("scRNAseq with Mean gene expression Covariate, wilcox") +
  theme_classic()
print(p)

# log-scale y axis
print(p + scale_y_continuous(trans="log2") + 
        ylab("Number of rejections (log2 scale)"))
dev.off()

b_dat=data.frame(pval=dat$scdd_pval, ind_covariate=dat$mean_exp,effect_size=matrix(g_cdata_sub_es),SE=matrix(g_cdata_sub_se))
b_dat=b_dat%>%mutate(zscore = effect_size/SE)
saveRDS(b_dat,"wilcox_meanexp.RDS")

resfile <- paste0(getwd(), "/scrnaseqhuman_scdd_meanexp_results_",
                  nrow(b_dat), ".RData")
if (!file.exists(resfile)){
  t1 <- proc.time()
  b_dat_results <- run_benchmarks_mod(dat=b_dat%>%filter(!is.na(pval)), alphas=cutoffs, pvals=TRUE)
  save(b_dat_results, file=resfile)
  message(paste0("Took ", round((proc.time()-t1)[3],1),
                 " seconds for ", nrow(b_dat), " genes"))
}else{
  load(resfile)
}

# plot N rejects vs FDR cutoff

pdf("scrnaseq_scdd_results.pdf",width=20,height=10)
p <- ggplot(b_dat_results$stats, aes(alpha, n_rejects, color=method)) +
  geom_point() +
  geom_line() +
  xlab("FDR cutoff") + ylab("Number of rejections") +
  ggtitle("scRNAseq with Mean gene expression Covariate, scdd") +
  theme_classic()
print(p)

# log-scale y axis
print(p + scale_y_continuous(trans="log2") + 
        ylab("Number of rejections (log2 scale)"))
dev.off()

```

### 2018-02-14 datasets/scrnaseq/scRNAseq-latest.Rmd
updated scRNAseq-latest.Rmd

@@
 #biocLite("MAST")
+#install.packages("FDRreg")
 
@@
 library(fdrtool)
+library(FDRreg)
 
@@
 ```{r, MAST}
+g_data_sub=readRDS("m_g_nofilter_subdata.rds")
+groups=c(rep(1,50),rep(2,50))
 names(groups)=colnames(g_cdata_sub)
@@
 source("/Users/subraman/Documents/research/fdr_jc/src/benchmark-fdr/datasets/R/strat_hist.R")
+
 h_cov=compute_covariates_scrnaseq("h_g_nofilter_subdata.rds")
+
 wilcox=readRDS("h_data_wilcox_p.rds")
-h_wilcox_cov=data.frame(genes=names(wilcox),pvalue=wilcox)
+h_wilcox_cov=data.frame(genes=names(wilcox),pval=wilcox)
 h_wilcox_cov=h_wilcox_cov%>%left_join(h_cov,by=c("genes"))
-dat=data.frame(pvalue=h_wilcox_cov$pvalue,covariate=h_wilcox_cov$median_exp)
-p=strat_hist(h_wilcox_cov,pvalue="pvalue",covariate="median_exp",numQ=5)
+
+plot_cov_diagnosis(h_wilcox_cov,"humanplate_wilcox")
+plot_cov_diagnosis(h_wilcox_cov%>%subset(detection_rate>=0.25),"humanplate_wilcox_filter")
 
 hmast=readRDS("humanplate_masttpm.rds")
-h_mast_cov=
-h_scdd_cov=
-p=strat_hist()
-scdd_m=metadata(scDatEx)
-scdd_p_df=data.frame(genes=scdd_m$Genes$gene,scdd_pval=matrix(scdd_m$Genes$nonzero.pvalue))
-wilcox_p_df=data.frame(genes=names(wilcox_p),wilcox_pval=matrix(wilcox_p))
+h_mast_cov=data.frame(genes=rownames(hmast),pval=hmast$pval)
+h_mast_cov=h_mast_cov%>%left_join(h_cov,by=c("genes"))
+plot_cov_diagnosis(h_mast_cov,"humanplate_mast")
+plot_cov_diagnosis(h_mast_cov%>%subset(detection_rate>=0.25),"humanplate_mast_filter")
+
+h_scdd=readRDS("humanplate_scDatEx_ks.rds")
+h_scdd_cov=data.frame(genes=rownames(metadata(h_scdd)$Genes),pval=metadata(h_scdd)$Genes$nonzero.pvalue)
+h_scdd_cov$genes=h_cov$genes #scdd changes gene names by appending numeric indices 
+h_scdd_cov=h_scdd_cov%>%subset(!is.na(pval))
+h_scdd_cov=h_scdd_cov%>%left_join(h_cov,by=c("genes"))
+plot_cov_diagnosis(h_scdd_cov,"humanplate_scdd")
+plot_cov_diagnosis(h_scdd_cov%>%subset(detection_rate>=0.25),"humanplate_scdd_filter")
+
+m_cov=compute_covariates_scrnaseq("m_g_nofilter_subdata.rds")
+
+wilcox=readRDS("./mouse_plate/m_data_wilcox_p.rds")
+m_wilcox_cov=data.frame(genes=names(wilcox),pval=wilcox)
+m_wilcox_cov=m_wilcox_cov%>%left_join(m_cov,by=c("genes"))
+
+plot_cov_diagnosis(m_wilcox_cov,"mouseplate_wilcox")
+
+m_wilcox_cov=m_wilcox_cov%>%subset(detection_rate>=0.25)
+plot_cov_diagnosis(m_wilcox_cov,"mouseplate_wilcox_filter")
+
+mmast=readRDS("./mouse_plate/mouseplate_masttpm.rds")
+m_mast_cov=data.frame(genes=rownames(mmast),pval=mmast$pval)
+m_mast_cov=m_mast_cov%>%left_join(m_cov,by=c("genes"))
+plot_cov_diagnosis(m_mast_cov,"mouseplate_mast")
+
+m_mast_cov=m_mast_cov%>%subset(detection_rate>=0.25)
+plot_cov_diagnosis(m_mast_cov,"mouseplate_mast_filter")
+
+m_scdd=readRDS("mouseplate_scDatEx_ks.rds")
+m_scdd_cov=data.frame(genes=rownames(metadata(m_scdd)$Genes),pval=metadata(m_scdd)$Genes$nonzero.pvalue)
+m_scdd_cov$genes=m_cov$genes
+m_scdd_cov=m_scdd_cov%>%subset(!is.na(pval))
+m_scdd_cov=m_scdd_cov%>%left_join(m_cov,by=c("genes"))
+plot_cov_diagnosis(m_scdd_cov,"mouseplate_scdd")
+plot_cov_diagnosis(m_scdd_cov%>%subset(detection_rate>=0.25),"mouseplate_scdd_filter")
 
 ```
-
-## Benchmarking covariate adjusted FDR tools for application
-
-```{r, benchmarking, eval=FALSE}
-g_cdata_sub_se=apply(g_cdata_sub,1,function(x) sd(x)/sqrt(length(x))) #standard error
-g_cdata_sub_es=apply(g_cdata_sub,1,function(x) mean(x[1:26])-mean(x[27:52])) #difference of means
-
-b_dat=data.frame(pval=dat$wilcox_pval, ind_covariate=dat$mean_exp,effect_size=matrix(g_cdata_sub_es),SE=matrix(g_cdata_sub_se))
-b_dat=b_dat%>%mutate(zscore = effect_size/SE)
-saveRDS(b_dat,"wilcox_meanexp.RDS")
-cutoffs <- c(0.01, 0.025, 0.05, 0.10, 0.20)
-resfile <- paste0(getwd(), "/scrnaseqhuman_wilcox_meanexp_results_",
-                  nrow(b_dat), ".RData")
-library(devtools)
-source("https://install-github.me/mangothecat/callr")
-source("https://install-github.me/r-lib/remotes")
-remotes::install_github('jgscott/FDRreg', subdir="R_pkg/")
-
-source("benchmark-fdr/datasets/R/evaluation-helpers_mod.R")
-if (!file.exists(resfile)){
-  t1 <- proc.time()
-  b_dat_results <- run_benchmarks_mod(dat=b_dat, alphas=cutoffs, pvals=TRUE)
-  save(b_dat_results, file=resfile)
-  message(paste0("Took ", round((proc.time()-t1)[3],1),
-                 " seconds for ", nrow(b_dat), " genes"))
-}else{
-  load(resfile)
-}
-
-# plot N rejects vs FDR cutoff
-
-pdf("scrnaseq_wilcox_results.pdf",width=20,height=10)
-p <- ggplot(b_dat_results$stats, aes(alpha, n_rejects, color=method)) +
-  geom_point() +
-  geom_line() +
-  xlab("FDR cutoff") + ylab("Number of rejections") +
-  ggtitle("scRNAseq with Mean gene expression Covariate, wilcox") +
-  theme_classic()
-print(p)
-
-# log-scale y axis
-print(p + scale_y_continuous(trans="log2") + 
-        ylab("Number of rejections (log2 scale)"))
-dev.off()
-
-b_dat=data.frame(pval=dat$scdd_pval, ind_covariate=dat$mean_exp,effect_size=matrix(g_cdata_sub_es),SE=matrix(g_cdata_sub_se))
-b_dat=b_dat%>%mutate(zscore = effect_size/SE)
-saveRDS(b_dat,"wilcox_meanexp.RDS")
-
-resfile <- paste0(getwd(), "/scrnaseqhuman_scdd_meanexp_results_",
-                  nrow(b_dat), ".RData")
-if (!file.exists(resfile)){
-  t1 <- proc.time()
-  b_dat_results <- run_benchmarks_mod(dat=b_dat%>%filter(!is.na(pval)), alphas=cutoffs, pvals=TRUE)
-  save(b_dat_results, file=resfile)
-  message(paste0("Took ", round((proc.time()-t1)[3],1),
-                 " seconds for ", nrow(b_dat), " genes"))
-}else{
-  load(resfile)
-}
-
-# plot N rejects vs FDR cutoff
-
-pdf("scrnaseq_scdd_results.pdf",width=20,height=10)
-p <- ggplot(b_dat_results$stats, aes(alpha, n_rejects, color=method)) +
-  geom_point() +
-  geom_line() +
-  xlab("FDR cutoff") + ylab("Number of rejections") +
-  ggtitle("scRNAseq with Mean gene expression Covariate, scdd") +
-  theme_classic()
-print(p)
-
-# log-scale y axis
-print(p + scale_y_continuous(trans="log2") + 
-        ylab("Number of rejections (log2 scale)"))
-dev.off()
-
-```

### 2018-02-14 datasets/RNA-Seq/yeast48-replicates-benchmarkFDR.Rmd
Add log2FC to truth; don't re-run if results file present

@@
 
-  sb <- bd %>% buildBench(data=geneExp, parallel = F, truthCols = "qvalue")
+  sb <- bd %>% buildBench(data=geneExp, parallel = F, 
+                          truthCols = "qvalue")
+  rowData(sb)$log2FC <- geneExp$effect_size
 
@@
 
+system("mkdir -p yeast48-results")
+
+if (!file.exists("yeast48-results/null5.rds")){
   null5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
@@
                    mc.cores=cores)
+  saveRDS(null5, file="yeast48-results/null5.rds")
+}else{
+  null5 <- readRDS(file="yeast48-results/null5.rds")
+}
+
 ```
@@
 
-# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
 plotsim_average(res5, met="rejections",filter_set = excludeSet,
@@
 
+if (!file.exists("yeast48-results/null10.rds")){
   null10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
@@
                    mc.cores=cores)
+  saveRDS(null10, file="yeast48-results/null10.rds")
+}else{
+  null10 <- readRDS(file="yeast48-results/null10.rds")
+}
+
 ```
@@
 
-# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
 plotsim_average(res10, met="rejections",filter_set = excludeSet,
@@
 
+if (!file.exists("yeast48-results/de5.rds")){
   de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
@@
                    mc.cores=cores)
+  saveRDS(de5, file="yeast48-results/de5.rds")
+}else{
+  de5 <- readRDS(file="yeast48-results/de5.rds")
+}
 ```
@@
 ```{r}
+  mat <- ( assays( de5[[1]] )[["qvalue"]] < 0.1 )
+  dfCov <- data.frame(
+    covariate = sizesAll,
+    methods=factor( rep( names( allLenghts ), sapply( allLenghts, length ) ), 
+                    levels=names(allLenghts)  ) )
+  pl <- ggplot( dfCov, aes( methods, trans(covariate), col=methods) ) +
+    geom_boxplot( ) + theme(legend.position="top", 
+                            axis.text.x=element_blank(), 
+                            legend.direction="vertical") +
+    labs(col="", y="Covariate", x="") + guides(color=guide_legend(ncol=2))
+  pl
+}
+
 res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.20, 0.01))
 
-# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
 plotsim_average(res5d, met="rejections",filter_set = excludeSet,
@@
   aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE)
+
 }else{
@@
 
+if(!file.exists("yeast48-results/de10.rds")){
   de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
@@
                    mc.cores=cores)
+  saveRDS(de10, file="yeast48-results/de10.rds")
+}else{
+  de10 <- readRDS(file="yeast48-results/de10.rds")
+}
 ```
@@
 
-# Save results objects
-
-```{r}
-system("mkdir -p yeast48-results")
-saveRDS(null5, file="yeast48-results/null5.rds")
-saveRDS(null10, file="yeast48-results/null10.rds")
-saveRDS(de5, file="yeast48-results/de5.rds")
-saveRDS(de10, file="yeast48-results/de10.rds")
-```
-
 # Session information

### 2018-02-15 datasets/RNA-Seq/yeast48-replicates-benchmarkFDR.Rmd
Add effect size plots; keep ind_cov in ftCols

@@
   sb <- bd %>% buildBench(data=geneExp, parallel = F, 
-                          truthCols = "qvalue")
+                          truthCols = "qvalue",
+                          ftCols = "ind_covariate")
   rowData(sb)$log2FC <- geneExp$effect_size
@@
 
+plotEffectSize(de5)
+plotEffectSize(de5, DE=FALSE)
+
 if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
@@
 
-# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
 plotsim_average(res10d, met="rejections",filter_set = excludeSet,
@@
                 merge_ihw = TRUE, errorBars=TRUE)
+
+plotEffectSize(de10)
+plotEffectSize(de10, DE=FALSE)
 

### 2018-02-15 datasets/RNA-Seq/yeast48-replicates-benchmarkFDR.Rmd
More flexibility to effectSize plotter - can also plot ind_covariate

@@
 
-plotEffectSize(de5)
-plotEffectSize(de5, DE=FALSE)
+plotEffectSize(de5, covname="log2FC")
+plotEffectSize(de5, covname="log2FC", DE=FALSE)
+
+plotEffectSize(de5, covname="ind_covariate")
+plotEffectSize(de5, covname="ind_covariate", DE=FALSE)
 
@@
 
-plotEffectSize(de10)
-plotEffectSize(de10, DE=FALSE)
+plotEffectSize(de10, covname="log2FC")
+plotEffectSize(de10, covname="log2FC", DE=FALSE)
+
+plotEffectSize(de10, covname="ind_covariate")
+plotEffectSize(de10, covname="ind_covariate", DE=FALSE)
 

### 2018-02-15 datasets/R/effectSizePlot.R

@@ -2,7 +2,7 @@
 #' stratified by significance and DE status
 #' 
 #' Created for the yeast48-simulation analysis. Assumes that effect sizes are
-#' in the log2FC slot of rowData. For each method, calculates the median effect
+#' in the rowData. For each method, calculates the median effect
 #' size for genes detected as DE (at specified alpha), separately for true DE 
 #' genes and null genes. The set of genes to be plotted is specified by the DE 
 #' argument.
@@ -12,9 +12,11 @@
 #' @param alpha the alpha cutoff
 #' @param DE whether to plot the effect sizes for DE genes (if false, will plot
 #' the effect sizes for non-DE genes)
+#' @param covname character that specifies column name of rowData to use for 
+#' y-axis (e.g. effect size, or ind_covariate)
 #' 
 #' @return a ggplot object
-plotEffectSize <- function(sbl, alpha=0.05, DE=TRUE){
+plotEffectSize <- function(sbl, alpha=0.05, DE=TRUE, covname){
 
 summarize_one_item <- function(object, alpha){
   object <- object[,!( grepl("^ihw", as.character( colData( object )$blabel ) )
@@ -24,19 +26,19 @@ summarize_one_item <- function(object, alpha){
                             1*(assays(object)[["qvalue"]]) < 0.05))
   colnames(df)[1] <- "truth" 
   df <- df %>%
-    select(truth, log2FC, bonf, bh,
+    select(truth, covname, bonf, bh,
            qvalue, contains("ihw"), ashs, "bl-df03", lfdr, "scott-theoretical", 
            "scott-empirical") %>%
     rename("scott-theoretical"="scott-t",
            "scott-empirical"="scott-e") %>%
-    gather(method, significant, -truth, -log2FC) %>%
+    gather(method, significant, -truth, -covname) %>%
     group_by(method, truth, significant) %>%
-    summarize(mean_effect_size=mean(abs(log2FC)),
-              med_effect_size=median(abs(log2FC)),
-              min_effect_size=min(abs(log2FC)),
-              max_effect_size=max(abs(log2FC)),
-              botQ=quantile(abs(log2FC),0.25),
-              topQ=quantile(abs(log2FC), 0.75)) %>%
+    summarize(mean_effect_size=mean(abs(get(covname))),
+              med_effect_size=median(abs(get(covname))),
+              min_effect_size=min(abs(get(covname))),
[…]

### 2018-02-15 datasets/scrnaseq/scRNAseq-latest.Rmd
AS: cleaned up rmd

@@
 knitr::opts_chunk$set(echo = TRUE,cache=TRUE)
+#knitr::opts_knit$set(root.dir = "/n/irizarryfs01_backed_up/kkorthauer/FDR")
 
 ```
-## scRNAseq datasets
-
-For the application of differential gene expression, we look at scRNAseq and bulk RNAseq data. Here we discuss the analyses of scRNAseq.
-
-```{r echo=FALSE, results='asis'}
-library(knitr)
-df=data.frame(datasets=c("GSE63818-GPL16791","GSE45719"),organism=c("human","mouse"),technology=c("smart-seq","smart-seq"))
-kable(df,caption="scRNAseq Datasets")
-df1=data.frame(DE_method=c("Wilcoxon-rank sum","ScDD","MAST"),input=c("tpm","count","tpm"),output=c("p-val","p-val","p-val"))
-kable(df1,caption="DE methods")
-
-```
-
-We selected two datasets, one each from mouse and human spp, from the `Conquer` (Soneson et al., 2017) scRNAseq data collection.
-`Conquer` provides uniformly processed gene expression data summarized to both count (aggregated from transcript members), and length-scaled TPMs at the gene level.
-The authors use `salmon` to quantify raw reads and `tximport` to summarize to gene level.
-The data comes from Smartseq, a plate based sequencing method which results in deeper sequencing of reads at the cost of numbers of cells.
-
-We also select two droplet-based datasets from the `10x` website:4k PBMC and 4k T cells, both from healthy donors.
-
-## Data Download
-
-We first download the data in our working directory. First, let's download the Human data:GSE63818-GPL16791
-
-```{bash, humandata-download, eval=FALSE}
-# all data and results are on the wd: /net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/fdrbenchmark-scrnaseq
-
-#these look well separated on the tsne plots on the conquer manuscript supplement
-wget http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE63818-GPL16791.rds
-# qc report: 
-wget http://imlspenticton.uzh.ch/robinson_lab/conquer/report-multiqc/GSE63818-GPL16791_multiqc_report.html
-# scater report
-wget http://imlspenticton.uzh.ch/robinson_lab/conquer/report-scater/GSE63818-GPL16791_scater.html
-```
-
-Next, we download the Mouse data:GSE45719
-
-```{bash, mousedata-download, eval=FALSE}
-#another well-separated dataset on tsne: GSE60749−GPL13112
-wget http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE45719.rds
-```
-
-We directly download the cell-ranger processed filtered expression matrices [4k PBMC cells](https://support.10xgenomics.com/single-cell-gene-expression/datasets/2.1.0/pbmc4k) and [4k T-cells](https://support.10xgenomics.com/single-cell-gene-expression/datasets/2.1.0/t_4k) from the 10x website.
-
 ## R workspace setup
@@
 
-```{r, workspace-setup, eval=FALSE}
+```{r, workspace-setup, results='hide', message=FALSE, warning=FALSE}
+datdir <- "data/scrnaseq/"
+resdir <- "results/scrnaseq/"
 #installation
@@
 #install.packages("FDRreg")
+sourceDirectory("benchmark-fdr/datasets/R/")
 
 # packages for data analyses
-library(MultiAssayExperiment)
-library(dplyr)
-library(ggplot2)
-library(cowplot)
-library(UpSetR)
-library(hexbin)
+suppressMessages(library(MultiAssayExperiment))
+suppressMessages(library(dplyr))
+suppressMessages(library(ggplot2))
+suppressMessages(library(cowplot))
+suppressMessages(library(UpSetR))
+suppressMessages(library(hexbin))
 
 # packages for differential expression testing
-library(impute)
-#library(samr)
-library(scDD)
-library(EBSeq)
-library(edgeR)
-library(MAST)
+suppressMessages(library(impute))
+#suppressMessages(library(samr)
+suppressMessages(library(scDD))
+suppressMessages(library(EBSeq))
+suppressMessages(library(edgeR))
+suppressMessages(library(MAST))
 
 # benchmark methods
-library(IHW)
-library(ashr)
-library(qvalue)
-library(swfdr)
-library(fdrtool)
-library(FDRreg)
+suppressMessages(library(IHW))
+suppressMessages(library(ashr))
+suppressMessages(library(qvalue))
+suppressMessages(library(swfdr))
+suppressMessages(library(fdrtool))
+suppressMessages(library(FDRreg))
 
 # summarizebenchmark
-library(SummarizedBenchmark)
-library(SingleCellExperiment)
-source("/Users/subraman/Documents/research/fdr_jc/code/fdr-benchmark-scrnaseq-helper.R")
+suppressMessages(library(SummarizedBenchmark))
+suppressMessages(library(SingleCellExperiment))
 
+# useful libraries
+suppressMessages(library(R.utils))
+suppressMessages(library(knitr))
+library(BiocParallel)
+cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
+multicoreParam <- MulticoreParam(workers = cores)
+sourceDirectory('~/Documents/research/fdr_jc/src/benchmark-fdr/datasets/R/')
+```
+
+## scRNAseq datasets
+
+For the application of differential gene expression, we look at scRNAseq and bulk RNAseq data. Here we discuss the analyses of scRNAseq.
+
+```{r echo=FALSE, results='asis'}
+df=data.frame(datasets=c("GSE63818-GPL16791","GSE45719"),organism=c("human","mouse"),technology=c("smart-seq","smart-seq"))
+kable(df,caption="scRNAseq Datasets")
+df1=data.frame(DE_method=c("Wilcoxon-rank sum","ScDD","MAST"),input=c("tpm","count","tpm"),output=c("p-val","p-val","p-val"))
+kable(df1,caption="DE methods")
+
+```
+
+We selected two datasets, one each from mouse and human spp, from the `Conquer` (Soneson et al., 2017) scRNAseq data collection.
+`Conquer` provides uniformly processed gene expression data summarized to both count (aggregated from transcript members), and length-scaled TPMs at the gene level.
+The authors use `salmon` to quantify raw reads and `tximport` to summarize to gene level.
+The data comes from Smartseq, a plate based sequencing method which results in deeper sequencing of reads at the cost of numbers of cells.
+
+## Data Download
+
+We first download the data in our working directory. First, let's download the Human data:GSE63818-GPL16791
+
+```{bash, humandata-download, eval=FALSE}
+# all data and results are on the wd: /net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/fdrbenchmark-scrnaseq
+
+#these look well separated on the tsne plots on the conquer manuscript supplement
+curl -o data/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE63818-GPL16791.rds
+# qc report: 
+curl -o data/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/report-multiqc/GSE63818-GPL16791_multiqc_report.html
+# scater report
+curl -o data/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/report-scater/GSE63818-GPL16791_scater.html
+```
+
+Next, we download the Mouse data:GSE45719
+
+```{bash, mousedata-download, eval=FALSE}
+#another well-separated dataset on tsne: GSE60749−GPL13112
+curl -o DATA/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE45719.rds
 ```
@@
 Briefly, we limit ourselves to 2-group comparisons, and  select the maximal size defined as the smallest number of cells in either group as the group-level sample size for the DE analyses.
+We do not filter genes as filtering may affect the subsequent false discovery control steps.
 
@@
 
-We look at results with and without prefiltering genes (keeping genes with TPM > 1 in atleast 25% of cells)
-
 ```{r, dataload, eval=FALSE}
-setwd("/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/fdrbenchmark-scrnaseq")
-setwd("/Users/subraman/Documents/research/fdr_jc/data/")
 #load compressed mae
@@
 
-```{r,mouse}
+Next, we repeat the processing steps for the mouse data.
+
+```{r,mouse, eval=FALSE}
 ##mouse data
@@
 
-```{r, nofilter, eval=FALSE}
+```{r, m_nofilter, eval=FALSE}
 g_data=cbind(g1_data,g2_data)
@@
 
-We perform differential expression testing using 3 methods on the unfiltered data:
+We perform differential expression testing using 3 methods on the unfiltered data: non-parametric Wilcoxon, MAST which has a hurdle model to account for zeros and scDD which implicitly models any heterogeneity in expression.
 
@@
 
-Per `Conquer`, we run TMM normalization to the TPM units before running Wilcoxon.
+Per Soneson et al., we run TMM normalization to the TPM units before running Wilcoxon.
+
 ```{r, wilcoxon, eval=FALSE}
@@
 
-We apply scDD to log-transformed counts as below:
+We apply scDD to log-transformed counts as below. Th helper functions set up the Single-Cell experiment object.
 
 ```{r, scdd, eval=FALSE}
-# library(SummarizedExperiment)
-# condition=c(rep(1,26),rep(2,26))
-# rownames(ng_cdata_sub) <- paste0(rownames(ng_cdata_sub), 1:nrow(ng_cdata_sub), sep="")
-# colnames(ng_cdata_sub) <- names(condition) <- paste0("Sample",
-# 1:ncol(ng_cdata_sub), sep="")
-# SDSumExp <- SummarizedExperiment(assays=list("NormCounts"=ng_cdata_sub),
-# colData=data.frame(condition))
-#saveRDS(SDSumExp,"data_sumexp.rds")
 
@@
 # perm=1000
-# scDatExperm <- scDD(SDSumExp, prior_param=prior_param, param=param, permutations=1000,testZeroes=FALSE)
-# saveRDS(scDatExperm,"/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/fdrbenchmark-scrnaseq/scDatExperm.rds")
 
@@
 
-```{r, MAST}
+We run MAST on the TPMs following the approach in Soneson et al.
+
+```{r, MAST,eval=FALSE}
 g_data_sub=readRDS("m_g_nofilter_subdata.rds")
@@
 
-We could use a few different covariates: strength of the signal defined as the median gene expression unit for the gene, detection rate or a combination of the two.
+In scrnaseq data, strength of the signal can be a function of level of gene expression.
+We quantify it as median or mean gene expression,and detection rate, defined as the percentage of cells expressing the gene at a level of TPM>0.
 
-```{r,cov,eval=TRUE}
-sourceDirectory("/Users/subraman/Documents/research/fdr_jc/src/benchmark-fdr/datasets/R")
-source("/Users/subraman/Documents/research/fdr_jc/src/benchmark-fdr/datasets/R/strat_hist.R")
+```{r,cov}
+h_cov=compute_covariates_scrnaseq("h_g_nofilter_subdata.rds")
+m_cov=compute_covariates_scrnaseq("m_g_nofilter_subdata.rds")
+```
 
-h_cov=compute_covariates_scrnaseq("h_g_nofilter_subdata.rds")
+First we will look at the covariates for the Wilcoxon text
 
+```{r, wilcoxcov, fig.width=10, fig.height=3.5}
+print("human_plate")
 wilcox=readRDS("h_data_wilcox_p.rds")
@@
 
-plot_cov_diagnosis(h_wilcox_cov,"humanplate_wilcox")
-plot_cov_diagnosis(h_wilcox_cov%>%subset(detection_rate>=0.25),"humanplate_wilcox_filter")
+print_plot_cov_diagnosis(h_wilcox_cov)
+print("With filter")
+print_plot_cov_diagnosis(h_wilcox_cov%>%subset(detection_rate>=0.05))
 
+print("mouse_plate")
+wilcox=readRDS("m_data_wilcox_p.rds")
+m_wilcox_cov=data.frame(genes=names(wilcox),pval=wilcox)
+m_wilcox_cov=m_wilcox_cov%>%left_join(m_cov,by=c("genes"))
+print_plot_cov_diagnosis(m_wilcox_cov)
+print("With filter")
+print_plot_cov_diagnosis(m_wilcox_cov%>%subset(detection_rate>=0.05))
+
+```
+
+Next, we look at the covariates for MAST
+
+```{r, mastcov, fig.width=10, fig.height=3.5}
+print("human_plate")
 hmast=readRDS("humanplate_masttpm.rds")
@@
 h_mast_cov=h_mast_cov%>%left_join(h_cov,by=c("genes"))
-plot_cov_diagnosis(h_mast_cov,"humanplate_mast")
-plot_cov_diagnosis(h_mast_cov%>%subset(detection_rate>=0.25),"humanplate_mast_filter")
+print_plot_cov_diagnosis(h_mast_cov)
+print("With filter")
+print_plot_cov_diagnosis(h_mast_cov%>%subset(detection_rate>=0.05))
 
+print("mouse_plate")
+mmast=readRDS("mouseplate_masttpm.rds")
+m_mast_cov=data.frame(genes=rownames(mmast),pval=mmast$pval)
+m_mast_cov=m_mast_cov%>%left_join(m_cov,by=c("genes"))
+print_plot_cov_diagnosis(m_mast_cov)
+print("With filter")
+print_plot_cov_diagnosis(m_mast_cov%>%subset(detection_rate>=0.05))
+```
+
+Finally, we look at how informative the covariates are with results from ScDD.
+
+```{r,scddcov, fig.width=10, fig.height=3.5}
+print("human_plate")
 h_scdd=readRDS("humanplate_scDatEx_ks.rds")
@@
 h_scdd_cov=h_scdd_cov%>%left_join(h_cov,by=c("genes"))
-plot_cov_diagnosis(h_scdd_cov,"humanplate_scdd")
-plot_cov_diagnosis(h_scdd_cov%>%subset(detection_rate>=0.25),"humanplate_scdd_filter")
+print_plot_cov_diagnosis(h_scdd_cov)
+print("With filter")
+print_plot_cov_diagnosis(h_scdd_cov%>%subset(detection_rate>=0.05))
 
-m_cov=compute_covariates_scrnaseq("m_g_nofilter_subdata.rds")
-
-wilcox=readRDS("./mouse_plate/m_data_wilcox_p.rds")
-m_wilcox_cov=data.frame(genes=names(wilcox),pval=wilcox)
-m_wilcox_cov=m_wilcox_cov%>%left_join(m_cov,by=c("genes"))
-
-plot_cov_diagnosis(m_wilcox_cov,"mouseplate_wilcox")
-
-m_wilcox_cov=m_wilcox_cov%>%subset(detection_rate>=0.25)
-plot_cov_diagnosis(m_wilcox_cov,"mouseplate_wilcox_filter")
-
-mmast=readRDS("./mouse_plate/mouseplate_masttpm.rds")
-m_mast_cov=data.frame(genes=rownames(mmast),pval=mmast$pval)
-m_mast_cov=m_mast_cov%>%left_join(m_cov,by=c("genes"))
-plot_cov_diagnosis(m_mast_cov,"mouseplate_mast")
-
-m_mast_cov=m_mast_cov%>%subset(detection_rate>=0.25)
-plot_cov_diagnosis(m_mast_cov,"mouseplate_mast_filter")
-
+print("mouse_plate")
 m_scdd=readRDS("mouseplate_scDatEx_ks.rds")
@@
 m_scdd_cov=m_scdd_cov%>%left_join(m_cov,by=c("genes"))
-plot_cov_diagnosis(m_scdd_cov,"mouseplate_scdd")
-plot_cov_diagnosis(m_scdd_cov%>%subset(detection_rate>=0.25),"mouseplate_scdd_filter")
+print_plot_cov_diagnosis(m_scdd_cov)
+print("With filter")
+print_plot_cov_diagnosis(m_scdd_cov%>%subset(detection_rate>=0.05))
 
 ```
+
+### Benchmarking FDR methods
+
+First we will create the benchdesign object for wilcox
+
+```{r, benchmarking, eval=FALSE}
+
+bd <- initializeBenchDesign()
+resN <- paste0(resdir, "scrnaseq_wilcox_summarizedBenchmark_",nrow(m_scdd_cov), ".RData")
+
+duration <- NA
+if (!file.exists(resN)){
+  t1 <- proc.time()
+  new_sb <- bd %>% buildBench(data=(m_scdd_cov%>%select(pval,mean_exp)%>%mutate(effect_size=NA, test_statistic=NA, SE=NA) %>% mutate(ind_covariate=mean_exp)), 
+                          ftCols = c("mean_exp"),
+                          parallel=TRUE, BPPARAM=multicoreParam)
+
+  save(sb, file=resN)
+  duration <- round((proc.time()-t1)[3]/60,1)
+}else{
+  load(resN)
+}
+
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb,metrics=c("rejections"))
+
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter(sb,
+                   supplementary=FALSE)
+
+rejection_scatter_bins(sb, covariate="ind_covar_N", bins=4,
+                       supplementary=FALSE)
+
+# upset plot 
+plotFDRMethodsOverlap(sb, 
+                      alpha=0.05, nsets=ncol(sb),
+                      order.by="freq", decreasing=TRUE,
+                      supplementary=FALSE)
+
+# properties of method-exclusive discoveries
+methods <- c("lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
+mcols(sb)$ind_covariate <- mcols(sb)$ind_covar_N
+plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3)

### 2018-02-15 datasets/R/fdr-benchmark-scrnaseq-helper.R

@@ -1,0 +1,131 @@
+#### helper functions for the scRNAseq analyses
+
+#' compute_wilcoxde_sc
+#'
+#' @param g_data_sub # the scrnaseq dataframe with gene expression in tpm
+#' @param fname #name of RDS file to be saved
+#' @param groups #the conditions to be tested
+#'
+#' @return #none, saves an RDS with computed p-values for each gene along with runtime
+#' @export
+#'
+#' @examples
+compute_wilcoxde_sc <- function(g_data_sub,fname,groups){
+  
+## code borrowed from https://github.com/csoneson/conquer_comparison/blob/master/scripts/apply_Wilcoxon.R
+tmm <- edgeR::calcNormFactors(g_data_sub)
+tpmtmm <- edgeR::cpm(g_data_sub, lib.size = tmm * colSums(g_data_sub))
+idx <- 1:nrow(tpmtmm)
+names(idx) <- rownames(tpmtmm)
+timing <- system.time(wilcox_p <- sapply(idx, function(i) {
+  wilcox.test(tpmtmm[i, ] ~ groups)$p.value
+}))
+saveRDS(wilcox_p,fname)
+}
+
+
+compute_scdd_sc <- function(g_cdata_sub,fname1,fname2,fname3){
+  Sizes=MedianNorm(g_cdata_sub)
+  ng_cdata_sub = Sizes*g_cdata_sub
+  saveRDS(Sizes,fname1)
+  condition=c(rep(1,(dim(g_cdata_sub)[2])/2),rep(2,(dim(g_cdata_sub)[2])/2))
+  rownames(ng_cdata_sub) <- paste0(rownames(ng_cdata_sub), 1:nrow(ng_cdata_sub), sep="")
+  colnames(ng_cdata_sub) <- names(condition) <- paste0("Sample",
+                                                       1:ncol(ng_cdata_sub), sep="")
+  sce <- SingleCellExperiment(assays=list("normcounts"=ng_cdata_sub),
+                              colData=data.frame(condition))
+  saveRDS(sce,fname2)
+  prior_param=list(alpha=0.01, mu0=0, s0=0.01, a0=0.01, b0=0.01)
+  param=BiocParallel::MulticoreParam(workers=8)
+  scDatEx <- scDD(sce, prior_param=prior_param, testZeroes=FALSE)
[…]

### 2018-02-15 datasets/RNA-Seq/yeast48-replicates-benchmarkFDR.Rmd
Remove extraneous code chunk

@@
 ```{r}
-  mat <- ( assays( de5[[1]] )[["qvalue"]] < 0.1 )
-  dfCov <- data.frame(
-    covariate = sizesAll,
-    methods=factor( rep( names( allLenghts ), sapply( allLenghts, length ) ), 
-                    levels=names(allLenghts)  ) )
-  pl <- ggplot( dfCov, aes( methods, trans(covariate), col=methods) ) +
-    geom_boxplot( ) + theme(legend.position="top", 
-                            axis.text.x=element_blank(), 
-                            legend.direction="vertical") +
-    labs(col="", y="Covariate", x="") + guides(color=guide_legend(ncol=2))
-  pl
-}
-
 res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.20, 0.01))

### 2018-02-15 datasets/RNA-Seq/polyester-replicates-benchmarkFDR.Rmd

---
title: "Polyester RNA-seq simulation study"
author: "Stephanie Hicks"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This is an analysis using [polyester](https://bioconductor.org/packages/release/bioc/html/polyester.html) Bioconductor package to simulate RNA-Seq reads. We will  
(1) implement null comparisons on samples simulated using polyester
with no differentially expressed (DE) genes, (2) compare FDR approaches 
on their ability to discover 'true positive' DE genes with log2 fold 
changes simulated from a normal distribution.

In this Rmd, we will perform a Monte Carlo simulation study 
and average over the replicates in plots.

## Set up workspace

```{r workspace-setup, results='hide', message=FALSE, warning=FALSE}
# Load packages needed to simulate RNA-seq counts with polyester
library(zebrafishRNASeq) 
library(genefilter)
library(limma)
library(polyester)

# Differential testing
library(DESeq2)

# Load supporting packages for data analyses
library(data.table)
library(magrittr)
library(dplyr)
library(ggplot2)
library(R.utils)
library(cowplot)
library(tibble)
library(ggthemes)

# benchmark methods
library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)
# Installed BayesLogit from CRAN archive via `R CMD INSTALL BayesLogit_0.6.tar.gz`
# install_github('jgscott/FDRreg', subdir="R_pkg/")
library(FDRreg)

# comparison tools/functions
library(SummarizedBenchmark)

# load helper functions
working_dir <- list(project_home = "/users/shicks1/projects/benchmark-fdr")
sourceDirectory(file.path(working_dir$project_home,"datasets/R/"))

# # set up parallel backend:
library(doParallel)
nCores <- 1
multicoreParam <- MulticoreParam(workers = nCores)
```

## Set up Polyester

We use the [polyester](https://bioconductor.org/packages/release/bioc/html/polyester.html) Bioconductor package to simulate RNA-Seq reads. The package requires a count matrix as input which estimates the mean variance relationship. Here we start with the [zebrafish](https://bioconductor.org/packages/release/data/experiment/html/zebrafishRNASeq.html) Bioconductor data package.

Estimate mean and variance relationship

```{r estimate-parameters-NB}
data(zfGenes)
dim(zfGenes)

filter <- apply(zfGenes, 1, function(x) length(x[x>5])>=2)
counts <- as.matrix(zfGenes[filter,])

## Estimate the zero inflated negative binomial parameters
params = get_params(counts)
```

```{r}
bd <- initializeBenchDesign() # only needs to be done once
```

Try running `buildBench()` one time.
```{r}
sampleSize <- 5
nDE <- 0

groupPolyester = rep(c(-1,1),each=sampleSize)
modPolyester = model.matrix(~-1 + groupPolyester)

# things needed for differential testing using DESeq2
groupDESeq = factor(rep(c(0,1), each=sampleSize))

# NULL simulation
truth <- rep(FALSE, nrow(counts)) 
names(truth) <- rownames(counts)

# log2FC are all 0 in null simulation
log2FC = rep(0,nrow(counts))
dat_alt = create_read_numbers(params$mu, params$fit,
                              params$p0, m=dim(counts)[1], 
                              n=sampleSize*2, beta=cbind(log2FC), 
                              mod=modPolyester)

# run DESeq2
dds_alt <- DESeqDataSetFromMatrix(countData=dat_alt, 
                                  colData = DataFrame(groupDESeq), 
                                  design=~groupDESeq)
dds_alt <- DESeq(dds_alt, parallel = FALSE)
res_alt <- results(dds_alt, independentFiltering = F)

geneExp <- tbl_df(data.frame(geneName=rownames(counts),
                             pval=res_alt$pvalue, SE=res_alt$lfcSE, 
                             ind_covariate = res_alt$baseMean, 
                             effect_size=res_alt$log2FoldChange, 
                             test_statistic=res_alt$stat, qvalue=truth, 
                             pzero = rowSums(counts(dds_alt)==0)/
                                     ncol(counts(dds_alt))  ))

# filter NAs and those with less than 50% expressed 
geneExp  <- geneExp %>% na.omit() %>% dplyr::filter(pzero < 0.5)
```

```{r}
head(geneExp)
```

```{r}
# build Benchmark data
sb <- bd %>% buildBench(data=geneExp, parallel=FALSE, BPPARAM=multicoreParam, 
                        truthCols = "qvalue")
```

# Session information

```{r}
sessionInfo()
```

### 2018-02-15 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
Use download.file instead of bash calls to fetch data

@@
 date: '`r format(Sys.Date(), "%B %d, %Y")`'
-output: html_document
+output: 
+    html_document:
+        toc: true
+        toc_float: true
+        highlight: tango
+        number_sections: true
 ---
@@
 knitr::opts_chunk$set(echo = TRUE)
-knitr::opts_knit$set(root.dir = "/n/irizarryfs01_backed_up/kkorthauer/FDR")
 ```
@@
 ```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
-# project directory & data/results folders - the datdir needs to match where
-#  files are downloaded with the first few bash chunks
-#setwd("/n/irizarryfs01_backed_up/kkorthauer/FDR")
-datdir <- "DATA/GWAS/"
-resdir <- "RESULTS/GWAS/"
+system("mkdir -p DATA")
+system("mkdir -p RESULTS")
+datdir <- "DATA/"
+resdir <- "RESULTS/"
 
@@
 library(SummarizedBenchmark)
-sourceDirectory("benchmark-fdr/datasets/R/")
+sourceDirectory("../R/")
 
@@
 
-```{bash, GWAS1-download, eval=FALSE}
-curl -o DATA/GWAS "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
+```{r, GWAS1-download}
+download.file(url="http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip", 
+             destfile=paste0(datdir,"BMI.SNPadjSMK.zip")) 
 ```
@@
 
-```{bash, GWAS1-unzip, eval=FALSE}
-unzip DATA/GWAS/BMI.SNPadjSMK.zip
-rm DATA/GWAS/BMI.SNPadjSMK.zip
-rm DATA/GWAS/BMI.SNPadjSMK.Women.EuropeanOnly.txt
-rm DATA/GWAS/BMI.SNPadjSMK.Women.AllAncestry.txt
-rm DATA/GWAS/BMI.SNPadjSMK.Men.EuropeanOnly.txt
-rm DATA/GWAS/BMI.SNPadjSMK.Men.AllAncestry.txt
-rm DATA/GWAS/BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt
+```{r, GWAS1-unzip}
+system(paste0("unzip ", datdir, "BMI.SNPadjSMK.zip"))
+system(paste0("rm ", "BMI.SNPadjSMK.Women.EuropeanOnly.txt"))
+system(paste0("rm ", "BMI.SNPadjSMK.Women.AllAncestry.txt"))
+system(paste0("rm ", "BMI.SNPadjSMK.Men.EuropeanOnly.txt"))
+system(paste0("rm ", "BMI.SNPadjSMK.Men.AllAncestry.txt"))
+system(paste0("rm ", "BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt"))
+system(paste0("mv BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt ", datdir))
+system(paste0("rm -f ", datdir, "/*.zip"))
 ```
@@
 
-```{bash, GWAS1-refdata, eval=FALSE}
+```{r, GWAS1-refdata}
+# remove remnants of previous runs, if present
+system(paste0("rm -f ", datdir, "/*CEU_MAF0.01*"))
+
 # download CEU 1000 genomes SNP reference data 
-wget 'http://neurogenetics.qimrberghofer.edu.au/iSECA/1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01.zip' -P DATA/GWAS/
+download.file(url="http://neurogenetics.qimrberghofer.edu.au/iSECA/1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01.zip", 
+              destfile=paste0(datdir, "1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01.zip"))
 
 # unzip reference data
-unzip DATA/GWAS/1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01.zip -d DATA/GWAS
+system(paste0("unzip ", datdir, "1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01.zip -d ", datdir))
+system(paste0("rm ", datdir, "1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01.zip"))
 ```
@@
 
-And we'll clean up the workspace before moving on to the next dataset.
+Clean up the workspace.
 

### 2018-02-15 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
Use download.file instead of bash calls to fetch data

@@
 
-### Summary
+# Summary
 
@@
 
-### Workspace setup
+# Workspace setup
 
@@
 
-### GWAS meta-analysis for BMI (included in Boca-Leek)
+# GWAS meta-analysis for BMI (included in Boca-Leek)
 
@@
 
-#### Data download
+## Data download
 
@@
 
-#### Prune SNPs that are in Linkage Disequilibrium
+## Prune SNPs that are in Linkage Disequilibrium
 
@@
 
-#### Check Covariate Diagnostics
+## Check Covariate Diagnostics
 
@@
 
-#### Set up BenchDesign object
+## Set up BenchDesign object
 
@@
 
-#### Run buildBench for the sample size covariate and examine results
+## Run buildBench for the sample size covariate and examine results
 
@@
 
-#### Run buildBench for the allele frequency covariate and examine results
+## Run buildBench for the allele frequency covariate and examine results
 

### 2018-02-16 datasets/RNA-Seq/polyester-replicates-benchmarkFDR.Rmd
updated polyester sims to B=10 reps; deleted old polyester sims

@@
 # # set up parallel backend:
-library(doParallel)
-nCores <- 1
+library(BiocParallel)
+nCores <- 10
 multicoreParam <- MulticoreParam(workers = nCores)
@@
 
+## Helper functions for simulation
+
+Next, we'll simulate RNA-Seq samples, both with and without the 
+addition of simulated DE genes. Here we'll create a function that we can use to 
+run one replicate given sample size and number of DE gene settings. This will be
+looped over many replications and results averaged over them.
+
 ```{r}
-bd <- initializeBenchDesign() # only needs to be done once
-```
+#' @param counts is the full count dataset
+#' @param paramPolyester is the parameters from running polyester. 
+#' Contains information on the mean-variance relationship used to 
+#' simulate RNA-Seq counts. 
+#' @param sampleSize is the number of samples in each condition
+#' @param nDE is the number of DE genes
+#' @param bd is the bench design object
+simulateOneSplit <- function(counts, paramPolyester, 
+                             nDE, sampleSize, bd, BPPARAM){
 
-Try running `buildBench()` one time.
-```{r}
-sampleSize <- 5
-nDE <- 0
-
+  # things needed to simulate with Polyester
   groupPolyester = rep(c(-1,1),each=sampleSize)
@@
   # NULL simulation
+  if(nDE == 0) {
     truth <- rep(FALSE, nrow(counts)) 
@@
     log2FC = rep(0,nrow(counts))
-dat_alt = create_read_numbers(params$mu, params$fit,
-                              params$p0, m=dim(counts)[1], 
+    dat_alt = create_read_numbers(paramPolyester$mu, paramPolyester$fit,
+                                  paramPolyester$p0, m=dim(counts)[1], 
                                   n=sampleSize*2, beta=cbind(log2FC), 
                                   mod=modPolyester)
+  }
+
+  # DE simulation
+  if(nDE > 0){
+    # pick random set of nDE genes to add signal to
+    DE <- sample(1:nrow(counts), nDE)
+    truth <- rep(FALSE, nrow(counts))
+    truth[DE] <- TRUE
+    names(truth) <- rownames(counts)
+
+    # randomly sample a log2FC from truncated N(0,0.5) (greater than 0.5) 
+    log2FC <- rep(0, nrow(counts))
+    # log2FC[DE] <- rtruncnorm(nDE, 1, sd=.5) * sample(c(1,-1), size=nDE, replace=TRUE)
+    log2FC[DE] <- c(rnorm(nDE, sd = 1.5), rep(0,nrow(counts)-nDE))
+    dat_alt = create_read_numbers(paramPolyester$mu, paramPolyester$fit,
+                                  paramPolyester$p0, m=dim(counts)[1], 
+                                  n=sampleSize*2, beta=cbind(log2FC), 
+                                  mod=modPolyester)
+  }
 
@@
   geneExp  <- geneExp  %>% na.omit() %>% dplyr::filter(pzero < 0.5)
+
+  # built Benchmark data
+  sb <- bd %>% buildBench(data=geneExp, parallel = FALSE, 
+                          BPPARAM = BPPARAM, truthCols = "qvalue", 
+                          ftCols = "ind_covariate")
+  rowData(sb)$log2FC <- geneExp$effect_size
+
+  return(sb)
+}
+
 ```
 
+We'll also set some parameters that will be common to all simulations. These
+include the number of replications, the bench design object, the set of 
+methods to exclude in the results plots, and the alpha cutoff level to 
+be used when plotting the aggregated Upset results.
+
 ```{r}
-head(geneExp)
+B <- 10
+bd <- initializeBenchDesign() # only needs to be done once
+excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
+ualpha <- 0.05
 ```
 
+Here's a helper function to return the number of methods with rejections at
+a particular alpha level (this helps us determine whether or not to plot the
+aggregated upset plot - if there aren't at least 2 methods it will throw an
+error, which is a problem for the null simulations).
+
 ```{r}
-# build Benchmark data
-sb <- bd %>% buildBench(data=geneExp, parallel=FALSE, BPPARAM=multicoreParam, 
-                        truthCols = "qvalue")
+#' @param res standardized metric data.table generated using
+#'        standardize_results.
+#' @param alpha alpha cutoff
+#' @param filterSet which methods to exclude from consideration 
+numberMethodsReject <- function(res, alphacutoff, filterSet){
+  res <- res %>% 
+    filter(is.na(param.alpha) | (param.alpha == alphacutoff)) %>%
+    filter(!(blabel %in% filterSet)) %>%
+    filter(alpha == alphacutoff) %>%
+    filter(performanceMetric == "rejections") %>%
+    select(blabel, performanceMetric, value) %>%
+    group_by(blabel) %>%
+    summarize(mean_value = mean(value)) %>%
+    filter(mean_value > 0)
+  return(nrow(res))
+}
+
+addDefaultMetrics <- function(sb) {
+    sb <- addPerformanceMetric(sb, evalMetric = c("TPR", "FPR", "TNR", "FNR", "rejections"),
+                               assay = "qvalue")
+    return(sb)
+}
+
+```
+
+# Sim N5: Null Comparison 5v5
+
+Here we run a null comparison 5 versus 5 RNA-Seq samples. 
+This will be done for 100 simulations.
+
+## Generate a list of SB results objects
+
+```{r, results='hide', message=FALSE}
+sampleSize <- 5
+nDE <- 0
+
+if (!file.exists(file.path(working_dir, "datasets/RNA-Seq/polyester-results-null5.RDS"))){
+  null5 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
+                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd,
+                   BPPARAM=multicoreParam)
+  saveRDS(null5, file=file.path(working_dir, "datasets/RNA-Seq/polyester-results-null5.RDS"))
+} else {
+  null5 <- readRDS(file=file.path(working_dir, "datasets/RNA-Seq/polyester-results-null5.RDS"))
+}
+```
+
+## Plot average results over replications
+
+Plot results.
+
+```{r}
+
+res5 <- plotsim_standardize(null5, alpha = seq(0.01, 0.20, 0.01))
+
+# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
+plotsim_average(res5, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5, met="TNR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+if (numberMethodsReject(res5, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+  aggupset(null5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+  message("Not enough methods found rejections at alpha ", ualpha, 
+          "; skipping upset plot")
+}
+```
+
+# Sim N10: Null Comparison 10v10
+
+Here we'll repeat the above, but for a null comparison 10 versus 10 samples. 
+This will be done for 100 simulations.
+
+## Generate a list of SB results objects
+
+```{r, results='hide', message=FALSE}
+sampleSize <- 10
+nDE <- 0
+
+if (!file.exists(file.path(working_dir, "datasets/RNA-Seq/polyester-results-null10.RDS"))){
+  null10 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
+                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd,
+                   BPPARAM=multicoreParam)
+  saveRDS(null10, file=file.path(working_dir, "datasets/RNA-Seq/polyester-results-null10.RDS"))
+} else {
+  null10 <- readRDS(file=file.path(working_dir, "datasets/RNA-Seq/polyester-results-null10.RDS"))
+}
+```
+
+## Plot average results over replications
+
+Plot results.
+
+```{r}
+res10 <- plotsim_standardize(null10, alpha = seq(0.01, 0.20, 0.01))
+
+# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
+plotsim_average(res10, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10, met="TNR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+if (numberMethodsReject(res10, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+  aggupset(null10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
+}
+
+```
+
+# Sim D5: DE Comparison 5v5
+
+Here we'll repeat the above, but for a DE comparison 5 versus 5 samples
+and 1500 DE genes are added. This will be done for 100 simulations.
+
+## Generate a list of SB results objects
+
+```{r, results='hide', message=FALSE}
+sampleSize <- 5
+nDE <- 1500
+
+if (!file.exists(file.path(working_dir, "datasets/RNA-Seq/polyester-results-de5.RDS"))){
+  de5 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
+                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd,
+                   BPPARAM=multicoreParam)
+  saveRDS(de5, file=file.path(working_dir, "datasets/RNA-Seq/polyester-results-de5.RDS"))
+} else {
+  de5 <- readRDS(file=file.path(working_dir, "datasets/RNA-Seq/polyester-results-de5.RDS"))
+}
+```
+
+## Plot average results over replications
+
+Plot results.
+
+```{r}
+plotEffectSize <- function(sbl, alpha=0.05, DE=TRUE, covname){
+
+summarize_one_item <- function(object, alpha){
+  object <- object[,!( grepl("^ihw", as.character( colData( object )$blabel ) )
+                       & colData( object )$param.alpha != alpha )]
+
+  df <- as.data.frame(cbind(as.matrix(rowData(object)), 
+                            1*(assays(object)[["qvalue"]]) < 0.05))
+  colnames(df)[1] <- "truth" 
+  df <- df %>%
+    select(truth, covname, bonf, bh,
+                  qvalue, contains("ihw"), ashs, "bl-df03", lfdr, 
+                  "scott-theoretical", "scott-empirical") %>%
+    # rename("scott-theoretical"="scott-t", "scott-empirical"="scott-e") %>%
+    gather(method, significant, -truth, -covname) %>%
+    group_by(method, truth, significant) %>%
+    summarize(mean_effect_size=mean(abs(get(covname))),
+              med_effect_size=median(abs(get(covname))),
+              min_effect_size=min(abs(get(covname))),
+              max_effect_size=max(abs(get(covname))),
+              botQ=quantile(abs(get(covname)),0.25),
+              topQ=quantile(abs(get(covname)), 0.75)) %>% na.omit()
+  df$significant <- as.factor(df$significant)
+  levels(df$significant)=c("Not Significant", "Significant")
+  df$truth <- as.factor(df$truth)
+  levels(df$truth)=c("Not DE", "DE")
+  return(df)
+}
+
+df <- lapply(sbl, summarize_one_item, alpha=alpha)
+df <- bind_rows(df, .id = "rep")
+df <- as.tibble(df)
+
+if (DE){
+  df <- df %>% filter(truth=="DE")
+}else{
+  df <- df %>% filter(truth=="Not DE")
+}
+
+p <- ggplot(df, aes(x = as.factor(method), y=med_effect_size)) +
+  geom_boxplot() +
+  facet_grid(significant~., scales="free_y") +
+  xlab("Method") +
+  ylab(paste0("Median ", covname))
+
+if (DE){
+  p <- p + ggtitle(paste0(covname, " of True DE genes by significance at alpha ", 
+                       alpha))
+}else{
+  p <- p + ggtitle(paste0(covname, " of Non-DE genes by significance at alpha ", 
+                       alpha))
+}
+
+return(p)
+}
+
+res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.20, 0.01))
+
+# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
+plotsim_average(res5d, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5d, met="FPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5d, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+plotEffectSize(de5, covname="log2FC")
+plotEffectSize(de5, covname="log2FC", DE=FALSE)
+
+plotEffectSize(de5, covname="ind_covariate")
+plotEffectSize(de5, covname="ind_covariate", DE=FALSE)
+
+if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+  aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
+}
+
+```
+
+# Sim D10: DE Comparison 10v10
+
+Here we'll repeat the above, but for a DE comparison 10 versus 10 samples
+and 1500 DE genes are added. This will be done for 100 simulations.
+
+## Generate a list of SB results objects
+
+```{r, results='hide', message=FALSE}
+sampleSize <- 10
+nDE <- 1500
+
+if (!file.exists(file.path(working_dir, "datasets/RNA-Seq/polyester-results-de10.RDS"))){
+  de10 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
+                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd,
+                   BPPARAM=multicoreParam)
+  saveRDS(de10, file=file.path(working_dir, "datasets/RNA-Seq/polyester-results-de10.RDS"))
+} else {
+  de10 <- readRDS(file=file.path(working_dir, "datasets/RNA-Seq/polyester-results-de10.RDS"))
+}
+```
+
+## Plot average results over replications
+
+Plot results.
+
+```{r}
+res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.20, 0.01))
+
+# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
+plotsim_average(res10d, met="rejections",filter_set = excludeSet, 
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10d, met="FPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10d, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+plotEffectSize(de5, covname="log2FC")
+plotEffectSize(de5, covname="log2FC", DE=FALSE)
+
+plotEffectSize(de5, covname="ind_covariate")
+plotEffectSize(de5, covname="ind_covariate", DE=FALSE)
+
+if (numberMethodsReject(res10d, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+  aggupset(de10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
+}
 ```

### 2018-02-16 datasets/RNA-Seq/polyester-benchmarkFDR.sh

@@ -1,12 +1,0 @@
-#!/bin/bash
-
-#SBATCH -n 4 # Number of cores
-#SBATCH -t 0-01:00 # Runtime in D-HH:MM
-#SBATCH -p irizarry,serial_requeue
-#SBATCH -e polyester-benchmarkFDR.err
-#SBATCH -o polyester-benchmarkFDR.out
-#SBATCH --mem=20GB # Memory pool for all cores (see also --mem-per-cpu)
-
-module load centos6/pandoc-1.12.3.3
-cd /n/irizarryfs01_backed_up/shicks/projects/benchmark-fdr
-Rscript -e 'library(knitr); rmarkdown::render("datasets/RNA-Seq/polyester-benchmarkFDR.Rmd", "html_document")' 

### 2018-02-16 datasets/RNA-Seq/polyester-replicates-benchmarkFDR.sh

@@ -1,7 +1,0 @@
-#$ -cwd
-#$ -pe local 1
-#$ -l mem_free=2G,h_vmem=2G 
-
-cd /users/shicks1/projects/benchmark-fdr/
-Rscript -e 'library(knitr); rmarkdown::render("datasets/RNA-Seq/polyester-replicates-benchmarkFDR.Rmd", "html_document")' 
-

### 2018-02-16 datasets/RNA-Seq/polyester-replicates-benchmarkFDR.Rmd
changed log2FC sd to 1

@@
     # log2FC[DE] <- rtruncnorm(nDE, 1, sd=.5) * sample(c(1,-1), size=nDE, replace=TRUE)
-    log2FC[DE] <- c(rnorm(nDE, sd = 1.5), rep(0,nrow(counts)-nDE))
+    log2FC[DE] <- c(rnorm(nDE, sd = 1), rep(0,nrow(counts)-nDE))
     dat_alt = create_read_numbers(paramPolyester$mu, paramPolyester$fit,

### 2018-02-16 datasets/RNA-Seq/polyester-replicates-benchmarkFDR.Rmd
fixed typo de5 to de10

@@
 library(BiocParallel)
-nCores <- 10
+nCores <- 1
 multicoreParam <- MulticoreParam(workers = nCores)
@@
 
-plotEffectSize(de5, covname="log2FC")
-plotEffectSize(de5, covname="log2FC", DE=FALSE)
+plotEffectSize(de10, covname="log2FC")
+plotEffectSize(de10, covname="log2FC", DE=FALSE)
 
-plotEffectSize(de5, covname="ind_covariate")
-plotEffectSize(de5, covname="ind_covariate", DE=FALSE)
+plotEffectSize(de10, covname="ind_covariate")
+plotEffectSize(de10, covname="ind_covariate", DE=FALSE)
 

### 2018-02-19 datasets/microbiome/microbiome-schubert.Rmd
Start splitting microbiome analyses

--- 2017-12-18 datasets/microbiome/microbiome-SummarizedBenchmark.Rmd
@@
 date: '`r format(Sys.Date(), "%B %d, %Y")`'
-output: html_document
----
+output: 
+   html_document:
+        toc: true
+        toc_float: true
+        highlight: tango
+        number_sections: true
+----
 
@@
 knitr::opts_chunk$set(echo = TRUE)
-#knitr::opts_knit$set(root.dir = "/n/irizarryfs01_backed_up/kkorthauer/FDR")
 ```
 
-### Summary
+# Summary
+
+Here we download and analyze the CDI Schubert diarrhea (C. diff and non-C. diff diarrhea) dataset. 
+We'll download the processed OTU tables from Zenodo and unzip them in the 
+`benchmark-fdr/datasets/microbiome/DATA` folder.
 
@@
 
-### Workspace setup
+# Workspace setup
 
 ```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
-# project directory & data/results folders
-setwd("/Users/claire/github/benchmark-fdr/datasets/microbiome/")
-datdir <- "/Users/claire/github/benchmark-fdr/datasets/microbiome/DATA/"
-resdir <- "/Users/claire/github/benchmark-fdr/datasets/microbiome/RESULTS/"
-
-# wrangling & plotting tools
-library(data.table)
-#library(readxl)
 library(dplyr)
@@
 library(magrittr)
-library(R.utils)
-library(cowplot)
+library(SummarizedBenchmark)
 
-# benchmark methods
-library(IHW)
-library(ashr)
-library(qvalue)
-library(swfdr)
-library(fdrtool)
-library(FDRreg)
+## load helper functions
+for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
+    source(f)
+}
 
-# comparison tools/functions
-library(SummarizedBenchmark)  
-sourceDirectory("/Users/claire/github/benchmark-fdr/datasets/R/")
-
-## set up parallel backend 
-library(BiocParallel)
-cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
-multicoreParam <- MulticoreParam(workers = cores)
+## project data/results folders
+datdir <- "data"
+resdir <- "results"
+dir.create(datdir, showWarnings = FALSE)
+dir.create(resdir, showWarnings = FALSE)
 ```
 
-Let's just keep going and hopefully it's okay...
+# Data Preparation
 
-### CDI Schubert - diarrhea dataset
+```{r, schubert-download}
+cdi_data <- file.path(datdir, "cdi_schubert_results.tar.gz")
+if (!file.exists(cdi_data)) {
+    download.file("https://zenodo.org/record/840333/files/cdi_schubert_results.tar.gz",
+                  destfile = cdi_data)
+}
 
-Here we download and analyze the CDI Schubert diarrhea (C. diff and non-C.diff diarrhea) dataset. 
-We'll download the processed OTU tables from Zenodo and unzip them in the 
-`benchmark-fdr/datasets/microbiome/DATA` folder.
-
-#### Data download
-
-```{bash, schubert-download, eval=FALSE}
-cd /Users/claire/github/benchmark-fdr/datasets/microbiome/DATA/
-curl -O "https://zenodo.org/record/840333/files/cdi_schubert_results.tar.gz"
-tar -xzvf cdi_schubert_results.tar.gz
-cd ..
+cdi_dir <- file.path(datdir, "cdi_schubert_results")
+if (!file.exists(file.path(cdi_dir, "cdi_schubert.metadata.txt"))) {
+    untar(cdi_data, exdir = datdir)
+}
 ```
@@
 ```{r, schubert-readdata, eval=TRUE}
-otu_path = paste0(datdir, "cdi_schubert_results/RDP/cdi_schubert.otu_table.100.denovo.rdp_assigned")
-meta_path = paste0(datdir, "cdi_schubert_results/cdi_schubert.metadata.txt")
-
-# DiseaseState labels to keep
-labels <- c("H", "nonCDI", "CDI")
-
-# Metadata column that has sample IDs
-col_label <- 'sample_id'
-
-# load OTU table and metadata
-otu <- read.table(otu_path)
-meta <- read.csv(meta_path, sep='\t')
+## load OTU table and metadata
+otu <- read.table(file.path(cdi_dir, "RDP",
+                            "cdi_schubert.otu_table.100.denovo.rdp_assigned"))
+meta <- read.csv(file.path(cdi_dir, "cdi_schubert.metadata.txt"), sep = '\t')
 
 # Keep only samples with the right DiseaseState metadata
-meta <- meta %>% filter(DiseaseState %in% labels)
+meta <- filter(meta, DiseaseState %in% c("H", "nonCDI", "CDI"))
 
 # Keep only samples with both metadata and 16S data
-keep_samples <- intersect(colnames(otu), meta[, 1])
+keep_samples <- intersect(colnames(otu), meta$sample_id)
 otu <- otu[, keep_samples]
-meta <- meta %>% filter(get(col_label) %in% keep_samples)
-
+meta <- filter(meta, 'sample_id' %in% keep_samples)
 ```
@@
 ```{r, schubert-cleandata, eval=TRUE}
-remove_shallow_smpls <- function(df, n_reads) {
-  # Removes samples with fewer than n_reads from dataframe df.
-  # df has OTUs in rows and samples in columns
-  return(df[, colSums(df) > n_reads])
-  }
+min_otu <- 10
+min_sample <- 100
 
-remove_shallow_otus <- function(df, n_reads){
-  return(df[rowSums(df) > n_reads, ])
-  }
+## Remove OTUs w/ <= min reads, samples w/ <= min reads
+otu <- otu[, colSums(otu) > min_otu]
+otu <- otu[rowSums(otu) > min_sample, ]
 
-## Remove OTUs with fewere than 10 reads
-#otu <- remove_shallow_otus(otu, 10)
+## Update metadata with new samples
+meta <- filter(meta, sample_id %in% colnames(otu))
 
-# Remove samples with fewer than sample_Reads reads
-sample_reads <- 100
-otu <- remove_shallow_smpls(otu, sample_reads)
-
-# Update metadata with new samples
-meta <- meta %>% filter(get(col_label) %in% colnames(otu))
-
-# Remove empty OTUs
+## Remove empty OTUs
 otu <- otu[rowSums(otu) > 0, ]
 
-# Convert to relative abundance
+## Convert to relative abundance
 abun_otu <- t(t(otu) / rowSums(t(otu)))
 
-# Add pseudo counts
-#minabun <- min(abun_otu[abun_otu > 0]) # Use this to figure out what pseudo-count value to add
+## Add pseudo counts
 zeroabun <- 0
 abun_otu <- abun_otu + zeroabun
+```
 
-```
+# Data Analysis
+
+## Differential Testing
 
@@
 ```{r, schubert-pvals, eval=TRUE}
-resfile <- paste0(resdir, "schubert_results_",
-                  nrow(abun_otu), "_OTUs.RData")
-if (!file.exists(resfile)){
-  # Get case and control indices
-  case_idx <- meta$DiseaseState %in% c("CDI", "nonCDI")
-  ctrl_idx <- meta$DiseaseState %in% c("H")
-
-  # Calculate pvalues, effects, and stderr
-  pvals <- c()
-  teststats <- c()
-  ses <- c()
-  effs <- c()
-  mean_abuns <- c()
-  mean_abuns_present <- c()
-  ubis <- c()
-
-  casedf <- abun_otu[, case_idx]
-  ctrldf <- abun_otu[, ctrl_idx]
-  for (o in rownames(abun_otu)) {
-    # wilcoxon p value  
-    w <- wilcox.test(casedf[o, ], 
-                     ctrldf[o, ])
-    p <- w$p.value
-    teststat <- w$statistic
-
-    pvals <- c(pvals, p)
-    teststats <- c(teststats, teststat)
-
-    # standard error of the OTU abundance, across all samples
-    ses <- c(ses, 
-             sd(abun_otu[o, ])/sqrt(length(abun_otu[o, ]))
-             )
-
-    # mean OTU abundance across all samples (after removing pseudo-count)
-    mean_abuns <- c(mean_abuns, 
-                    mean(abun_otu[o, ] - zeroabun)
-                    )
-
-    # mean OTU abundance across only samples with the OTU present
-    mean_abuns_present <- c(mean_abuns_present, 
-                            sum(abun_otu[o, ] - zeroabun) / sum(abun_otu[o, ] > zeroabun)
-                            )
-
-    # ubiquity of OTU across all samples
-    ubis <- c(ubis, 
-              sum(abun_otu[o, ] > zeroabun) / length(abun_otu[o, ])
-              )
-
-    # effect (logfold difference)
-    effs <- c(effs, 
-              log(mean(abun_otu[o, ctrl_idx])/mean(abun_otu[o, case_idx]))
-              )
-
-  }
-
-  res <- data.frame(otu = rownames(abun_otu), 
-                    pval = pvals, wilcox_teststat = teststats,
-                    SE = ses, effect_size = effs, 
-                    mean_abun = mean_abuns, mean_abun_present = mean_abuns_present,
-                    ubiquity = ubis)
-  res <- res %>%
-    mutate(test_statistic = qnorm(exp(log(pval)-log(2)), lower.tail=FALSE) * sign(effect_size))
-
+resfile <- file.path(resdir, paste0("schubert_results_", nrow(abun_otu), "_OTUs.rda"))
+if (!file.exists(resfile)) {
+    res <- test_microbiome(abundance = abun_otu, shift = zeroabun,
+                           is_case = meta$DiseaseState %in% c("CDI", "nonCDI"))
     save(res, file=resfile)
-
-}else{
+} else {
   load(resfile)
@@
 ```{r, schubert-addphylo, eval=TRUE}
-res <- res %>% separate(otu, 
-                        c("kingdom", "phylum", "class", "order", "family", "genus", "species", "denovo"), 
-                        sep=";", remove = FALSE)
-
+res <- dplyr::separate(res, otu, 
+                       c("kingdom", "phylum", "class", "order",
+                         "family", "genus", "species", "denovo"), 
+                       sep = ";", remove = FALSE)
 ```
 
-#### Check Covariate Diagnostics
+## Covariate Diagnostics
 
@@
 
-##### Ubiquity
+### Ubiquity
 
@@
 
-##### Mean abundance (across non-zero samples)
+### Mean Abundance (across non-zero samples)
 
@@
 
-##### Phylogeny?!
+### Phylogeny
 
@@
 
-```{r, schubert-phylo, fig.width=10, fig.height=6, eval=TRUE}
-#strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=20)
-#rank_scatter(res, pvalue="pval", covariate="ubiquity")
-ggplot(res, aes(x=pval)) + geom_histogram() + facet_wrap(~phylum, scales = "free")
+```{r, schubert-phylo, fig.width=10, fig.height=6, eval=FALSE}
+ggplot(res, aes(x=pval)) +
+    geom_histogram() +
+    facet_wrap(~phylum, scales = "free")
 ```
 
-Let's use `ubiquity` as our `ind_covariate`
+## Multiple-Testing Correction
+
+Let's use `ubiquity` as our `ind_covariate`.
 
 ```{r, schubert-labelcovariate, eval=TRUE}
-res <- res %>% mutate(ind_covariate = ubiquity)
+res <- dplyr::mutate(res, ind_covariate = ubiquity)
 ```
-
-Also (hacky), remove test_statistic so that ash doesn't run.
-
-```{r, schubert-rm-ash, eval=TRUE}
-res <- res %>% 
-  mutate(test_statistic = NA) %>%
-  mutate(effect_size = NA)
-```
-
-#### Set up BenchDesign object
 
@@
 ```{r, schubert-sb, results="hide", message=FALSE, eval=TRUE}
-for (nmids in c(5, 50, 150)){
+for (nmids in c(5, 50, 150)) {
     print(nmids)
-  bd <- initializeBenchDesign(nmids=nmids)
+    bd <- initializeBenchDesign(nmids = nmids)
+    ## remove ASH from comparison
+    bd <- dropBMethod(bd, "ashs")
 
-  resfile <- paste0(resdir, "schubert_summarizedBenchmark_",
-                    nrow(res), "_nmids_", nmids, ".RData")
+    resfile <- file.path(resdir, paste0("schubert_summarizedBenchmark_",
+                                        nrow(res),  "_nmids_", nmids, ".rda"))
     duration <- NA
-  if (!file.exists(resfile)){
-    t1 <- proc.time()
-    sb <- bd %>% buildBench(data=res, ftCols = "ind_covariate")
-                            #parallel=TRUE, BPPARAM=multicoreParam)
-    metadata(sb)$data_download_link <- "https://zenodo.org/record/840333/files/cdi_schubert_results.tar.gz"
-    save(sb, file=resfile)
-    duration <- round((proc.time()-t1)[3]/60,1)
-  }else{
+    if (!file.exists(resfile)) {
+        sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+        metadata(sb)$data_download_link <-
+                       "https://zenodo.org/record/840333/files/cdi_schubert_results.tar.gz"
+        save(sb, file = resfile)
+    } else {
         load(resfile)
@@
 }
-
 ```
 
-Just for future runs, where I load the file instead of re-calculating it, here are the errors I got:
-
-```
-[1] 5
-Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !NaNs produced<simpleError in if (sum(D^2) > qchisq(0.9, nmids - 2 - df)) {    warning(paste0("f(z) misfit = ", round(D, 1), ".
-Rerun with increased df."))}: missing value where TRUE/FALSE needed>
-
-[1] 50
-Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !f(z) misfit = -21.6.
-Rerun with increased df.f(z) misfit = -22.5.
-Rerun with increased df.f(z) misfit = -10.
-Rerun with increased df.f(z) misfit = 32.
-Rerun with increased df.f(z) misfit = 14.
-Rerun with increased df.f(z) misfit = -32.2.
-Rerun with increased df.f(z) misfit = 6.9.
-Rerun with increased df.f(z) misfit = -1.3.
-Rerun with increased df.f(z) misfit = 0.7.
-Rerun with increased df.f(z) misfit = 3.4.
-Rerun with increased df.f(z) misfit = 5.5.
-Rerun with increased df.f(z) misfit = 0.1.
-Rerun with increased df.f(z) misfit = -0.9.
-Rerun with increased df.f(z) misfit = -1.5.
-Rerun with increased df.f(z) misfit = -1.9.
-Rerun with increased df.f(z) misfit = -0.8.
-Rerun with increased df.f(z) misfit = -2.4.
-Rerun with increased df.f(z) misfit = 1.7.
-Rerun with increased df.f(z) misfit = 0.5.
-Rerun with increased df.f(z) misfit = 2.
-Rerun with increased df.f(z) misfit = 0.1.
-Rerun with increased df.f(z) misfit = 1.4.
-Rerun with increased df.f(z) misfit = 0.2.
-Reru...
-<truncated>
-
-[1] 150
-Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !f(z) misfit = 83.4.
-Rerun with increased df.f(z) misfit = -8.5.
-Rerun with increased df.f(z) misfit = -13.6.
-Rerun with increased df.f(z) misfit = -17.1.
-Rerun with increased df.f(z) misfit = -6.4.
-Rerun with increased df.f(z) misfit = -19.5.
-Rerun with increased df.f(z) misfit = -17.8.
-Rerun with increased df.f(z) misfit = -3.
-Rerun with increased df.f(z) misfit = -21.1.
-Rerun with increased df.f(z) misfit = 28.6.
-Rerun with increased df.f(z) misfit = -26.
-Rerun with increased df.f(z) misfit = 59.3.
-Rerun with increased df.f(z) misfit = -17.1.
-Rerun with increased df.f(z) misfit = 12.1.
-Rerun with increased df.f(z) misfit = -28.3.
-Rerun with increased df.f(z) misfit = -27.6.
-Rerun with increased df.f(z) misfit = 83.7.
-Rerun with increased df.f(z) misfit = -10.5.
-Rerun with increased df.f(z) misfit = -22.6.
-Rerun with increased df.f(z) misfit = -23.3.
-Rerun with increased df.f(z) misfit = 37.3.
-Rerun with increased df.f(z) misfit = -18.
-Rerun with increased df.f(z...
-<truncated>>
-```
-
-```{r, echo=FALSE, eval=FALSE}
-if(!is.na(duration)){
-  message("This step took ", duration, " minutes for ", nrow(res), " OTUs using ",
-        cores, " cores.")
-}
-```
+## Benchmark Metrics
 
@@
 ```{r, schubert-metrics, eval=TRUE}
-# rename assay to qvalue
 assayNames(sb) <- "qvalue"
@@
 
-Debugging note: to look at the matrix of qvalues, run `assays(sb)[["qvalue"]]`
-
 ```{r, schubert-plot, width=15, height=15, eval=TRUE}
-
 # plot nrejects by method overall and stratified by covariate
-rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
 rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
 plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
-
 ```
@@
 - lfdr and scott-empirical are all NaN (I think this is likely related to the df error)
+
+```{r, schubert-genus-covariates}
+methods <- c("lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
+plotCovariateBoxplots(sb, alpha = 0.1, nsets = 6, methods = methods)
+```
 
@@
 
-And we'll clean up the workspace before moving on to the next dataset.
-
-```{r, schubert-cleanup, eval=FALSE}
-rm(res)
-rm(bd)
-rm(sb)
-rm(pf)
-```
-
-### OB Goodrich - obesity dataset
-
-Let's repeat these analyses for the OB Goodrich dataset.
-
-Here we download and analyze the CDI Schubert diarrhea (C. diff and non-C.diff diarrhea) dataset. 
-We'll download the processed OTU tables from Zenodo and unzip them in the 
-`benchmark-fdr/datasets/microbiome/DATA` folder.
-
-#### Data download
-
-```{bash, goodrich-download, eval=FALSE}
-cd /Users/claire/github/benchmark-fdr/datasets/microbiome/DATA/
-curl -O "https://zenodo.org/record/840333/files/ob_goodrich_results.tar.gz"
-tar -xzvf ob_goodrich_results.tar.gz
-cd ..
-```
-
-Next, we'll read in the unzipped OTU table and metadata files into R.
-
-We'll need to manually define which disease labels are of interest and what metadata
-column contains the sample IDs (i.e. the IDs in the OTU table).
-
-```{r, goodrich-readdata}
-otu_path = paste0(datdir, "ob_goodrich_results/RDP/ob_goodrich.otu_table.100.denovo.rdp_assigned")
-meta_path = paste0(datdir, "ob_goodrich_results/ob_goodrich.metadata.txt")
-
-# DiseaseState labels to keep
-labels <- c("H", "OB")
-
-# Metadata column that has sample IDs
-col_label <- "X"
-
-# load OTU table and metadata
-otu <- read.table(otu_path)
-meta <- read.csv(meta_path, sep='\t')
-
-# Keep only samples with the right DiseaseState metadata
-meta <- meta %>% filter(DiseaseState %in% labels)
-
-# Keep only samples with both metadata and 16S data
-keep_samples <- intersect(colnames(otu), meta[, 1])
-otu <- otu[, keep_samples]
-meta <- meta %>% filter(get(col_label) %in% keep_samples)
-
-```
-
-Since we'll be using OTU-wise covariates, we shouldn't need to perform any
-filtering/cleaning of the OTUs, apart from removing any that are all zeros. 
-(This may happen after removing shallow samples, I think.)
-However, let's make sure we get rid of any samples with too few reads. 
-We define the minimum number of reads per sample in `sample_reads`.
-
-After we've removed any shallow samples, we'll convert the OTU table to relative
-abundances. I'll add a pseudo-count of 1e-6 to any zero entries, to avoid
-problems with taking logs.
-
-```{r, goodrich-cleandata}
-remove_shallow_smpls <- function(df, n_reads) {
-  # Removes samples with fewer than n_reads from dataframe df.
-  # df has OTUs in rows and samples in columns
-  return(df[, colSums(df) > n_reads])
-  }
-
-# Remove samples with fewer than sample_Reads reads
-sample_reads <- 100
-otu <- remove_shallow_smpls(otu, sample_reads)
-
-# Update metadata with new samples
-meta <- meta %>% filter(get(col_label) %in% colnames(otu))
-
-# Remove empty OTUs
-otu <- otu[rowSums(otu) > 0, ]
-
-# Convert to relative abundance
-abun_otu <- t(t(otu) / rowSums(t(otu)))
-
-# Add pseudo counts
-#minabun <- min(abun_otu[abun_otu > 0]) # Use this to figure out what pseudo-count value to add
-zeroabun <- 0
-abun_otu <- abun_otu + zeroabun
-
-```
-
-Next, we need to calculate the pvalues, effect size, and standard error for each OTU.
-Here, we'll compare lean vs. obese. We'll put these results into a dataframe, and label 
-the columns with the standardized names for downstream use 
-(`pval`, `SE`, `effect_size`, `test_statistic`). The test statistic is the one returned 
-by `wilcox.test()`.
-
-Note that the effect here is calculated as logfold change of mean abundance in controls
-relative to cases (i.e. `log(mean_abun[controls]/mean_abun[cases])`)
-
-While we're at it, we'll also calculate the mean abundance and ubiquity (detection rate)
-of each OTU. Later, we can assign their values to a new column called `ind_covariate` 
-for use in downstream steps.
-
-Note that OB Goodrich has ~70,000 OTUs, so this is going to take a while. There are probably
-faster ways to write this, but I don't know them.
+# Session Info
 
 ```{r}
-dim(abun_otu)
+sessionInfo()
 ```
-
-```{r, goodrich-pvals}
-resfile <- paste0(resdir, "goodrich_results_",
-                  nrow(abun_otu), "_OTUs.RData")
-if (!file.exists(resfile)){
-  # Get case and control indices
-  case_idx <- meta$DiseaseState %in% c("OB")
-  ctrl_idx <- meta$DiseaseState %in% c("H")
-
-  # Calculate pvalues, effects, and stderr
-  pvals <- c()
-  teststats <- c()
-  ses <- c()
-  effs <- c()
-  mean_abuns <- c()
-  mean_abuns_present <- c()
-  ubis <- c()
-
-  casedf <- abun_otu[, case_idx]
-  ctrldf <- abun_otu[, ctrl_idx]
-  for (o in rownames(abun_otu)) {
-    # wilcoxon p value  
-    w <- wilcox.test(casedf[o, ], 
-                     ctrldf[o, ])
-    p <- w$p.value
-    teststat <- w$statistic
-
-    pvals <- c(pvals, p)
-    teststats <- c(teststats, teststat)
-
-    # standard error of the OTU abundance, across all samples
-    ses <- c(ses, 
-             sd(abun_otu[o, ])/sqrt(length(abun_otu[o, ]))
-             )
-
-    # mean OTU abundance across all samples (after removing pseudo-count)
-    mean_abuns <- c(mean_abuns, 
-                    mean(abun_otu[o, ] - zeroabun)
-                    )
-
-    # mean OTU abundance across only samples with the OTU present
-    mean_abuns_present <- c(mean_abuns_present, 
-                            sum(abun_otu[o, ] - zeroabun) / sum(abun_otu[o, ] > zeroabun)
-                            )
-
-    # ubiquity of OTU across all samples
-    ubis <- c(ubis, 
-              sum(abun_otu[o, ] > zeroabun) / length(abun_otu[o, ])
-              )
-
-    # effect (logfold difference)
-    effs <- c(effs, 
-              log(mean(abun_otu[o, ctrl_idx])/mean(abun_otu[o, case_idx]))
-              )
-
-  }
-
-  res <- data.frame(otu = rownames(abun_otu), 
-                    pval = pvals, test_statistic = teststats,
-                    SE = ses, effect_size = effs, 
-                    mean_abun = mean_abuns, mean_abun_present = mean_abuns_present,
-                    ubiquity = ubis)
-  res <- res %>%
-    mutate(test_statistic = qnorm(exp(log(pval)-log(2)), lower.tail=FALSE) * sign(effect_size))
-
-  save(res, file=resfile)
-
-}else{
-  load(resfile)
-}
-head(res)
-```
-
-Finally, let's try to add phylogeny as covariates. Here we'll have columns for each separate taxonomic level.
-
-```{r, goodrich-addphylo}
-res <- res %>% separate(otu, 
-                        c("kingdom", "phylum", "class", "order", "family", "genus", "species", "denovo"), 
-                        sep=";", remove = FALSE)
-
-```
-
-#### Check Covariate Diagnostics
-
-Here we look to see if the covariates do indeed look informative.
-
-##### Ubiquity
-
-```{r, goodrich-ubi, fig.width=10, fig.height=3.5}
-strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=5, binwidth=0.05, numQ=4)
-rank_scatter(res, pvalue="pval", covariate="ubiquity")
-```
-
-##### Mean abundance (across non-zero samples)
-
-```{r, goodrich-abun, fig.width=10, fig.height=3.5}
-strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=5, binwidth=0.05, numQ=4)
-rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
-```
-
-##### Phylogeny?!
-
-Let's look at phylum-level stratification first. A priori, I *might* expect
-Proteobacteria to be enriched for low p-values? But I don't know if that's
-super legit, and Eric doesn't seem to think that phylogeny will be informative at all...
-
-```{r, goodrich-phylo, fig.width=10, fig.height=6}
-#strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=20)
-#rank_scatter(res, pvalue="pval", covariate="ubiquity")
-ggplot(res, aes(x=pval)) + geom_histogram() + facet_wrap(~phylum, scales = "free")
-```
-
-Let's use `ubiquity` as our `ind_covariate`
-
-```{r, goodrich-labelcovariate}
-res <- res %>% mutate(ind_covariate = ubiquity)
-```
-
-Also (hacky), remove test_statistic so that ash doesn't run.
-
-```{r, goodrich-rm-ash}
-res <- res %>% 
-  mutate(test_statistic = NA) %>%
-  mutate(effect_size = NA)
-```
-
-#### Set up BenchDesign object
-
-First, we'll create an object of `BenchDesign` class to hold the data and 
-add the benchmark methods to the `BenchDesign` object.
-
-Next, we'll construct the `SummarizedBenchmark` object, which will run
-the functions specified in each method (these are actually sourced in from the
-helper scripts).
-
-```{r, goodrich-sb, results="hide", message=FALSE, eval=TRUE}
-for (nmids in c(5, 50, 150)){
-  print(nmids)
-  bd <- initializeBenchDesign(nmids=nmids)
-
-  resfile <- paste0(resdir, "goodrich_summarizedBenchmark_",
-                    nrow(res), "_nmids_", nmids, ".RData")
-  duration <- NA
-  if (!file.exists(resfile)){
-    t1 <- proc.time()
-    sb <- bd %>% buildBench(data=res, ftCols = "ind_covariate")
-                            #parallel=TRUE, BPPARAM=multicoreParam)
-    metadata(sb)$data_download_link <- "https://zenodo.org/record/840333/files/cdi_schubert_results.tar.gz"
-    save(sb, file=resfile)
-    duration <- round((proc.time()-t1)[3]/60,1)
-  }else{
-    load(resfile)
-  }
-}
-```
-
-The outputs of this call are:
-
-```
-[1] 5
-Squarem-1
-Objective fn: 715262
-Objective fn: 104929  Extrapolation: 0  Steplength: 1
-Objective fn: 95918.9  Extrapolation: 1  Steplength: 2.44204
-Objective fn: 95614.3  Extrapolation: 1  Steplength: 2.72536
-Objective fn: 95583.3  Extrapolation: 1  Steplength: 2.8568
-Objective fn: 95578.1  Extrapolation: 1  Steplength: 3.54373
-Objective fn: 95576.3  Extrapolation: 1  Steplength: 3.52004
-Objective fn: 95575.4  Extrapolation: 1  Steplength: 4
-.
-.
-.
-NaNs produced<simpleError in if (sum(D^2) > qchisq(0.9, nmids - 2 - df)) {    warning(paste0("f(z) misfit = ", round(D, 1), ".  Rerun with increased df."))}: missing value where TRUE/FALSE needed>
-
-[1] 50
-Squarem-1
-Objective fn: 715262
-Objective fn: 104929  Extrapolation: 0  Steplength: 1
-Objective fn: 95918.9  Extrapolation: 1  Steplength: 2.44204
-Objective fn: 95614.3  Extrapolation: 1  Steplength: 2.72536
-Objective fn: 95583.3  Extrapolation: 1  Steplength: 2.8568
-Objective fn: 95578.1  Extrapolation: 1  Steplength: 3.54373
-Objective fn: 95576.3  Extrapolation: 1  Steplength: 3.52004
-Objective fn: 95575.4  Extrapolation: 1  Steplength: 4
-Objective fn: 95574.9  Extrapolation: 1  Steplength: 3.36769
-Objective fn: 95574.5  Extrapolation: 1  Steplength: 5.09115
-Objective fn: 95574.3  Extrapolation: 1  Steplength: 2.52066
-Objective fn: 95574  Extrapolation: 1  Steplength: 13.3966
-.
-.
-.
-Objective fn: 95572.9  Extrapolation: 1  Steplength: 1.61619
-f(z) misfit = -8.7.
-Rerun with increased df.f(z) misfit = 33.3.
-Rerun with increased df.f(z) misfit = -9.4.
-Rerun with increased df.f(z) misfit = -37.3.
-Rerun with increased df.f(z) misfit = 49.2.
-Rerun with increased df.f(z) misfit = -36.9.
-Rerun with increased df.f(z) misfit = 10.6.
-Rerun with increased df.f(z) misfit = 13.3.
-Rerun with increased df.f(z) misfit = -18.3.
-Rerun with increased df.f(z) misfit = 20.1.
-Rerun with increased df.f(z) misfit = -21.6.
-Rerun with increased df.f(z) misfit = 23.4.
-Rerun with increased df.f(z) misfit = -27.
-Rerun with increased df.f(z) misfit = 17.2.
-Rerun with increased df.f(z) misfit = -8.2.
-Rerun with increased df.f(z) misfit = 4.3.
-Rerun with increased df.f(z) misfit = 3.4.
-Rerun with increased df.f(z) misfit = 2.4.
-Rerun with increased df.f(z) misfit = -2.3.
-Rerun with increased df.f(z) misfit = -10.7.
-Rerun with increased df.f(z) misfit = -2.4.
-Rerun with increased df.f(z) misfit = 15.4.
-Rerun with increased df.f(z) misf...
-<truncated>
-
-[1] 150
-Squarem-1
-Objective fn: 715262
-Objective fn: 104929  Extrapolation: 0  Steplength: 1
-Objective fn: 95918.9  Extrapolation: 1  Steplength: 2.44204
-Objective fn: 95614.3  Extrapolation: 1  Steplength: 2.72536
-Objective fn: 95583.3  Extrapolation: 1  Steplength: 2.8568
-Objective fn: 95578.1  Extrapolation: 1  Steplength: 3.54373
-Objective fn: 95576.3  Extrapolation: 1  Steplength: 3.52004
-Objective fn: 95575.4  Extrapolation: 1  Steplength: 4
-Objective fn: 95574.9  Extrapolation: 1  Steplength: 3.36769
-Objective fn: 95574.5  Extrapolation: 1  Steplength: 5.09115
-Objective fn: 95574.3  Extrapolation: 1  Steplength: 2.52066
-Objective fn: 95574  Extrapolation: 1  Steplength: 13.3966
-.
-.
-.
-f(z) misfit = -20.2.
-Rerun with increased df.f(z) misfit = 13.6.
-Rerun with increased df.f(z) misfit = -3.8.
-Rerun with increased df.f(z) misfit = 9.
-Rerun with increased df.f(z) misfit = -21.2.
-Rerun with increased df.f(z) misfit = 28.8.
-Rerun with increased df.f(z) misfit = 41.9.
-Rerun with increased df.f(z) misfit = -13.2.
-Rerun with increased df.f(z) misfit = -12.8.
-Rerun with increased df.f(z) misfit = 15.2.
-Rerun with increased df.f(z) misfit = -18.1.
-Rerun with increased df.f(z) misfit = -19.1.
-Rerun with increased df.f(z) misfit = -18.6.
-Rerun with increased df.f(z) misfit = -26.
-Rerun with increased df.f(z) misfit = -7.2.
-Rerun with increased df.f(z) misfit = 67.1.
-Rerun with increased df.f(z) misfit = 27.2.
-Rerun with increased df.f(z) misfit = -16.4.
-Rerun with increased df.f(z) misfit = -21.1.
-Rerun with increased df.f(z) misfit = -25.9.
-Rerun with increased df.f(z) misfit = -7.5.
-Rerun with increased df.f(z) misfit = 10.1.
-Rerun with increased df.f(z)...
-<truncated>
-```
-
-```{r, echo=FALSE, eval=FALSE}
-if(!is.na(duration)){
-  message("This step took ", duration, " minutes for ", nrow(res), " OTUs using ",
-        cores, " cores.")
-}
-```
-
-Next, we'll add the default performance metric for q-value assays. First, we have
-to rename the assay to 'qvalue'.
-
-```{r, goodrich-metrics, eval=TRUE}
-# rename assay to qvalue
-assayNames(sb) <- "qvalue"
-sb <- addDefaultMetrics(sb)
-```
-
-Now, we'll plot the results.
-
-Debugging note: to look at the matrix of qvalues, run `assays(sb)[["qvalue"]]`
-
-```{r, goodrich-plot, width=15, height=15, eval=TRUE}
-# plot nrejects by method overall and stratified by covariate
-rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
-rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
-#plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE )
-```
-
-There is no upsetR plot for OB Goodrich for now because only lfdr is returning any rejections.
-
-```{r, eval=FALSE}
-assays(sb)[["qvalue"]][, "ashs"] %>% max()
-sum(is.na(assays(sb)[["qvalue"]][, "scott-empirical"]))
-sum(is.na(assays(sb)[["qvalue"]][, "lfdr"]))
-
-```
-
-### CRC Baxter - colorectal cancer
-
-Next, I'll look at the CRC Baxter dataset.
-For colorectal cancer, we do expect some amount of truly differentially abundant OTUs, but not as many as in diarrhea.
-This dataset will hopefully provide an intermediate non-extreme case study.
-We'll download the processed OTU tables from Zenodo and unzip them in the 
-`benchmark-fdr/datasets/microbiome/DATA` folder.
-
-#### Data download
-
-```{bash, baxter-download, eval=FALSE}
-cd /Users/claire/github/benchmark-fdr/datasets/microbiome/DATA/
-curl -O "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
-tar -xzvf crc_baxter_results.tar.gz
-cd ..
-```
-
-Next, we'll read in the unzipped OTU table and metadata files into R.
-
-We'll need to manually define which disease labels are of interest and what metadata
-column contains the sample IDs (i.e. the IDs in the OTU table).
-
-```{r, baxter-readdata}
-otu_path = paste0(datdir, "crc_baxter_results/RDP/crc_baxter.otu_table.100.denovo.rdp_assigned")
-meta_path = paste0(datdir, "crc_baxter_results/crc_baxter.metadata.txt")
-
-# DiseaseState labels to keep
-labels <- c("H", "CRC")
-
-# Metadata column that has sample IDs
-col_label <- "Sample_Name_s"
-
-# load OTU table and metadata
-otu <- read.table(otu_path)
-meta <- read.csv(meta_path, sep='\t')
-
-# Keep only samples with the right DiseaseState metadata
-meta <- meta %>% filter(DiseaseState %in% labels)
-
-# Add "X" in front of sample IDs because of how R read in the OTU table
-meta[, 1] <- sub("^", "X", meta[, 1])
-
-# Keep only samples with both metadata and 16S data
-keep_samples <- intersect(colnames(otu), meta[, 1])
-otu <- otu[, keep_samples]
-meta <- meta %>% filter(get(col_label) %in% keep_samples)
-
-```
-
-Since we'll be using OTU-wise covariates, we shouldn't need to perform any
-filtering/cleaning of the OTUs, apart from removing any that are all zeros. 
-(This may happen after removing shallow samples, I think.)
-However, let's make sure we get rid of any samples with too few reads. 
-We define the minimum number of reads per sample in `sample_reads`.
-
-After we've removed any shallow samples, we'll convert the OTU table to relative
-abundances. I'll add a pseudo-count of 1e-7 to any zero entries, to avoid
-problems with taking logs.
-
-```{r, baxter-cleandata}
-remove_shallow_smpls <- function(df, n_reads) {
-  # Removes samples with fewer than n_reads from dataframe df.
-  # df has OTUs in rows and samples in columns
-  return(df[, colSums(df) > n_reads])
-  }
-
-remove_shallow_otus <- function(df, n_reads){
-  return(df[rowSums(df) > n_reads, ])
-  }
-
-## Remove OTUs with fewere than 10 reads
-#otu <- remove_shallow_otus(otu, 10)
-
-# Remove samples with fewer than sample_Reads reads
-sample_reads <- 100
-otu <- remove_shallow_smpls(otu, sample_reads)
-
-# Update metadata with new samples
-meta <- meta %>% filter(get(col_label) %in% colnames(otu))
-
-# Remove empty OTUs
-otu <- otu[rowSums(otu) > 0, ]
-
-# Convert to relative abundance
-abun_otu <- t(t(otu) / rowSums(t(otu)))
-
-# Add pseudo counts
-#minabun <- min(abun_otu[abun_otu > 0]) # Use this to figure out what pseudo-count value to add
-zeroabun <- 0
-abun_otu <- abun_otu + zeroabun
-
-```
-
-Next, we need to calculate the pvalues, effect size, and standard error for each OTU.
-Here, we'll compare CRC vs. healthy. We won't consider the nonCRC adenoma patients.
-We'll put these results into a dataframe, and label the columns with the 
-standardized names for downstream use (`pval`, `SE`, `effect_size`, `test_statistic`).
-The test statistic is the one returned by `wilcox.test()`.
-
-Note that the effect here is calculated as logfold change of mean abundance in controls
-relative to cases (i.e. `log(mean_abun[controls]/mean_abun[cases])`)
-
-While we're at it, we'll also calculate the mean abundance and ubiquity (detection rate)
-of each OTU. Later, we can assign their values to a new column called `ind_covariate` 
-for use in downstream steps.
-
-```{r, baxter-pvals}
-resfile <- paste0(resdir, "baxter_results_",
-                  nrow(abun_otu), "_OTUs.RData")
-if (!file.exists(resfile)){
-  # Get case and control indices
-  case_idx <- meta$DiseaseState %in% c("CRC")
-  ctrl_idx <- meta$DiseaseState %in% c("H")
-
-  # Calculate pvalues, effects, and stderr
-  pvals <- c()
-  teststats <- c()
-  ses <- c()
-  effs <- c()
-  mean_abuns <- c()
-  mean_abuns_present <- c()
-  ubis <- c()
-
-  casedf <- abun_otu[, case_idx]
-  ctrldf <- abun_otu[, ctrl_idx]
-  for (o in rownames(abun_otu)) {
-    # wilcoxon p value  
-    w <- wilcox.test(casedf[o, ], 
-                     ctrldf[o, ])
-    p <- w$p.value
-    teststat <- w$statistic
-
-    pvals <- c(pvals, p)
-    teststats <- c(teststats, teststat)
-
-    # standard error of the OTU abundance, across all samples
-    ses <- c(ses, 
-             sd(abun_otu[o, ])/sqrt(length(abun_otu[o, ]))
-             )
-
-    # mean OTU abundance across all samples (after removing pseudo-count)
-    mean_abuns <- c(mean_abuns, 
-                    mean(abun_otu[o, ] - zeroabun)
-                    )
-
-    # mean OTU abundance across only samples with the OTU present
-    mean_abuns_present <- c(mean_abuns_present, 
-                            sum(abun_otu[o, ] - zeroabun) / sum(abun_otu[o, ] > zeroabun)
-                            )
-
-    # ubiquity of OTU across all samples
-    ubis <- c(ubis, 
-              sum(abun_otu[o, ] > zeroabun) / length(abun_otu[o, ])
-              )
-
-    # effect (logfold difference)
-    effs <- c(effs, 
-              log(mean(abun_otu[o, ctrl_idx])/mean(abun_otu[o, case_idx]))
-              )
-
-  }
-
-  res <- data.frame(otu = rownames(abun_otu), 
-                    pval = pvals, test_statistic = teststats,
-                    SE = ses, effect_size = effs, 
-                    mean_abun = mean_abuns, mean_abun_present = mean_abuns_present,
-                    ubiquity = ubis)
-  res <- res %>%
-    mutate(test_statistic = qnorm(exp(log(pval)-log(2)), lower.tail=FALSE) * sign(effect_size))
-
-  save(res, file=resfile)
-
-}else{
-  load(resfile)
-}
-head(res)
-```
-
-Finally, let's try to add phylogeny as covariates. Here we'll have columns for each separate taxonomic level.
-
-```{r, baxter-addphylo}
-res <- res %>% separate(otu, 
-                        c("kingdom", "phylum", "class", "order", "family", "genus", "species", "denovo"), 
-                        sep=";", remove = FALSE)
-
-```
-
-#### Check Covariate Diagnostics
-
-Here we look to see if the covariates do indeed look informative.
-
-##### Ubiquity
-
-```{r, baxter-ubi, fig.width=10, fig.height=3.5}
-strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
-rank_scatter(res, pvalue="pval", covariate="ubiquity")
-```
-
-Something weird is happening with p=0.20 and p=0.40 - maybe this has something to do with ties?
-Either way, ubiquity looks to be a bit informative (from the scatter plot, but not really the histograms...)
-
-##### Mean abundance (across non-zero samples)
-
-```{r, baxter-abun, fig.width=10, fig.height=3.5}
-strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
-rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
-```
-
-##### Phylogeny?!
-
-Let's look at phylum-level stratification first. A priori, I *might* expect
-Proteobacteria to be enriched for low p-values? But I don't know if that's
-super legit, and Eric doesn't seem to think that phylogeny will be informative at all...
-
-```{r, baxter-phylo, fig.width=10, fig.height=6}
-#strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=20)
-#rank_scatter(res, pvalue="pval", covariate="ubiquity")
-ggplot(res, aes(x=pval)) + geom_histogram() + facet_wrap(~phylum, scales = "free")
-```
-
-Not really informative either...
-
-Let's use `ubiquity` as our `ind_covariate`
-
-```{r, baxter-labelcovariate}
-res <- res %>% mutate(ind_covariate = ubiquity)
-```
-
-Also (hacky), remove test_statistic so that ash doesn't run.
-
-```{r, baxter-rm-ash}
-res <- res %>% 
-  mutate(test_statistic = NA) %>%
-  mutate(effect_size = NA)
-```
-
-#### Set up BenchDesign object
-
-First, we'll create an object of `BenchDesign` class to hold the data and 
-add the benchmark methods to the `BenchDesign` object. We'll do this for a few
-values of nmids, to explore its effects on the outputs.
-
-Then, we'll construct the `SummarizedBenchmark` object, which will run
-the functions specified in each method (these are actually sourced in from the
-helper scripts).
-
-```{r, baxter-sb, results="hide", message=FALSE, eval=TRUE}
-for (nmids in c(5, 50, 150)){
-  print(nmids)
-  bd <- initializeBenchDesign(nmids=nmids)
-
-  resfile <- paste0(resdir, "baxter_summarizedBenchmark_",
-                    nrow(res), "_nmids_", nmids, ".RData")
-  duration <- NA
-  if (!file.exists(resfile)){
-    t1 <- proc.time()
-    sb <- bd %>% buildBench(data=res, ftCols = "ind_covariate")
-                            #parallel=TRUE, BPPARAM=multicoreParam)
-    metadata(sb)$data_download_link <- "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
-    save(sb, file=resfile)
-    duration <- round((proc.time()-t1)[3]/60,1)
-  }else{
-    load(resfile)
-  }
-}
-
-```
-
-Just for future runs, where I load the file instead of re-calculating it, here are the errors I got:
-
-```
-[1] 5
-Squarem-1
-Objective fn: 944053
-Objective fn: 111059  Extrapolation: 0  Steplength: 1
-Objective fn: 92567.2  Extrapolation: 1  Steplength: 2.85015
-Objective fn: 91136.8  Extrapolation: 1  Steplength: 2.70993
-Objective fn: 90930.1  Extrapolation: 1  Steplength: 4
-.
-.
-.
-Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !NaNs produced<simpleError in if (sum(D^2) > qchisq(0.9, nmids - 2 - df)) {    warning(paste0("f(z) misfit = ", round(D, 1), ".
-Rerun with increased df."))}: missing value where TRUE/FALSE needed>
-
-[1] 50
-Squarem-1
-Objective fn: 944053
-Objective fn: 111059  Extrapolation: 0  Steplength: 1
-Objective fn: 92567.2  Extrapolation: 1  Steplength: 2.85015
-Objective fn: 91136.8  Extrapolation: 1  Steplength: 2.70993
-Objective fn: 90930.1  Extrapolation: 1  Steplength: 4
-.
-.
-.
-Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !f(z) misfit = -33.4.
-Rerun with increased df.f(z) misfit = -40.5.
-Rerun with increased df.f(z) misfit = 207.5.
-Rerun with increased df.f(z) misfit = -33.
-Rerun with increased df.f(z) misfit = -13.
-Rerun with increased df.f(z) misfit = -35.
-Rerun with increased df.f(z) misfit = -44.7.
-Rerun with increased df.f(z) misfit = -38.4.
-Rerun with increased df.f(z) misfit = -35.2.
-Rerun with increased df.f(z) misfit = -59.5.
-Rerun with increased df.f(z) misfit = 275.
-Rerun with increased df.f(z) misfit = -30.1.
-Rerun with increased df.f(z) misfit = -74.7.
-Rerun with increased df.f(z) misfit = -87.9.
-Rerun with increased df.f(z) misfit = -85.1.
-Rerun with increased df.f(z) misfit = 268.2.
-Rerun with increased df.f(z) misfit = -56.9.
-Rerun with increased df.f(z) misfit = -39.4.
-Rerun with increased df.f(z) misfit = -8.8.
-Rerun with increased df.f(z) misfit = -35.6.
-Rerun with increased df.f(z) misfit = -30.2.
-Rerun with increased df.f(z) misfit = -14.3.
-Rerun with increased ...
-<truncated>
-
-[1] 150
-Squarem-1
-Objective fn: 944053
-Objective fn: 111059  Extrapolation: 0  Steplength: 1
-Objective fn: 92567.2  Extrapolation: 1  Steplength: 2.85015
-Objective fn: 91136.8  Extrapolation: 1  Steplength: 2.70993
-Objective fn: 90930.1  Extrapolation: 1  Steplength: 4
-Objective fn: 90882.9  Extrapolation: 0  Steplength: 1
-.
-.
-.
-Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !Censored sample for null model estimation has only size 0 !f(z) misfit = -15.9.
-Rerun with increased df.f(z) misfit = 8.3.
-Rerun with increased df.f(z) misfit = -10.8.
-Rerun with increased df.f(z) misfit = -18.1.
-Rerun with increased df.f(z) misfit = -24.6.
-Rerun with increased df.f(z) misfit = -27.1.
-Rerun with increased df.f(z) misfit = -24.6.
-Rerun with increased df.f(z) misfit = -18.8.
-Rerun with increased df.f(z) misfit = -21.4.
-Rerun with increased df.f(z) misfit = 279.5.
-Rerun with increased df.f(z) misfit = 83.1.
-Rerun with increased df.f(z) misfit = -27.4.
-Rerun with increased df.f(z) misfit = -28.3.
-Rerun with increased df.f(z) misfit = -7.4.
-Rerun with increased df.f(z) misfit = 11.6.
-Rerun with increased df.f(z) misfit = -18.2.
-Rerun with increased df.f(z) misfit = -23.
-Rerun with increased df.f(z) misfit = -16.7.
-Rerun with increased df.f(z) misfit = -24.1.
-Rerun with increased df.f(z) misfit = -23.2.
-Rerun with increased df.f(z) misfit = -24.4.
-Rerun with increased df.f(z) misfit = -26.6.
-Rerun with increase...
-<truncated>>
-```
-
-```{r, echo=FALSE, eval=FALSE}
-if(!is.na(duration)){
-  message("This step took ", duration, " minutes for ", nrow(res), " OTUs using ",
-        cores, " cores.")
-}
-```
-
-Next, we'll add the default performance metric for q-value assays. First, we have
-to rename the assay to 'qvalue'.
-
-```{r, baxter-metrics, eval=TRUE}
-# rename assay to qvalue
-assayNames(sb) <- "qvalue"
-sb <- addDefaultMetrics(sb)
-```
-
-Now, we'll plot the results.
-
-Debugging note: to look at the matrix of qvalues, run `assays(sb)[["qvalue"]]`
-
-```{r, baxter-plot, width=15, height=15, eval=TRUE}
-# plot nrejects by method overall and stratified by covariate
-rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
-rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
-#plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
-```
-
-```{r, eval=FALSE}
-assays(sb)[["qvalue"]][, "ashs"] %>% max()
-sum(is.na(assays(sb)[["qvalue"]][, "scott-empirical"]))
-sum(is.na(assays(sb)[["qvalue"]][, "lfdr"]))
-
-```

### 2018-02-19 datasets/R/microbiome-helpers.R

@@ -1,0 +1,73 @@
+#' Differential Abundance Testing in Microbiome Data
+#'
+#' This wrapper function performs differential testing of
+#' OTU (relative) abundances between case and control samples
+#' using the Wilcoxon rank sum test (equiv. Mann-Whitney test)
+#' for each OTU. Results are reported as a data.frame with
+#' one row per OTU.
+#' 
+#' @param abundance matrix of relative abunances with rows and
+#'        columns corresponding to OTUs and samples.
+#' @param shift psuedocount that was added to all relative abundance
+#'        values in \code{abundance} to handle zero counts.
+#' @param is_case logical vector of length equal to the number of
+#'        columns in \code{abundance} specifying whether each sample
+#'        is a case (TRUE) or control (FALSE).
+#'
+#' @return
+#' data.frame of differential testing results.
+#'
+#' @author Claire Duvallet
+test_microbiome <- function(abundance, shift, is_case) {
+
+    casedf <- abundance[, is_case]
+    ctrldf <- abundance[, !is_case]
+
+    ## Calculate pvalues, effects, and stderr
+    res <- data.frame(otu = rownames(abundance), 
+                      pval = NA, wilcox_teststat = NA,
+                      SE = NA, effect_size = NA, 
+                      mean_abun = NA, mean_abun_present = NA,
+                      ubiquity = NA)
+    
+    for (i in seq_len(nrow(abundance))) {
+        ## wilcoxon p value
+        w <- wilcox.test(casedf[i, ], ctrldf[i, ])
+        res$pval[i] <- w$p.value
+        res$wilcox_teststat[i] <- w$statistic
+                
+        ## standard error of the OTU abundance, across all samples
+        res$SE[i] <- sd(abundance[i, ]) / sqrt(length(abundance[i, ]))
+        
+        ## mean OTU abundance across all samples (after removing pseudo-count)
+        res$mean_abun[i] <- mean(abundance[i, ]) - shift
+        
+        ## mean OTU abundance across only samples with the OTU present
+        res$mean_abun_present[i] <-
+            sum(abundance[i, ] - shift) / sum(abundance[i, ] > shift)
+        
+        ## ubiquity of OTU across all samples
+        res$ubiquity[i] <- sum(abundance[i, ] > shift) / length(abundance[i, ])
+        
+        ## effect (logfold difference)
+        res$effect_size[i] <- log( mean(ctrldf[i, ]) / mean(casedf[i, ]) )
[…]

### 2018-02-20 datasets/R/strat_hist.R
Load libraries at top of function definition for strat_hist

@@ -24,6 +24,12 @@
 #' 
 #' @author Keegan Korthauer     
 strat_hist <- function(dat, pvalue, covariate, binwidth=0.025, maxy=3, numQ=3){
+
+  library(cowplot)
+  library(ggplot2)
+  library(magrittr)
+  library(dplyr)
+  
   # check numQ input
   if (numQ > 5){
     stop("Please specify a valid value of numQ")

### 2018-02-20 datasets/microbiome/microbiome-schubert.Rmd
Verify microbiome schubert knits

@@
 otu <- otu[, keep_samples]
-meta <- filter(meta, 'sample_id' %in% keep_samples)
+meta <- filter(meta, sample_id %in% keep_samples)
 ```
@@
 ## Update metadata with new samples
-meta <- filter(meta, sample_id %in% colnames(otu))
+meta <- dplyr::filter(meta, sample_id %in% colnames(otu))
 
@@
 ```{r, schubert-addphylo, eval=TRUE}
-res <- dplyr::separate(res, otu, 
+res <- tidyr::separate(res, otu, 
                        c("kingdom", "phylum", "class", "order",

### 2018-02-20 datasets/microbiome/microbiome-baxter.Rmd
Add more notes to Schubert; get through Baxter

---
title: "Case Study: 16S microbiome data (Baxter et al., 2016)"
author: "Claire Duvallet"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
----

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

Here we download and analyze the Baxter colorectal cancer (CRC) dataset
[(Baxter et al., 2016)](http://dx.doi.org/10.1186/s13073-016-0290-3).
The dataset includes stool samples from 172 patients with no colonic lesions,
and 120 patients with CRC. We'll download the processed OTU tables from
Zenodo and unzip them in the `benchmark-fdr/datasets/microbiome/data` folder.

For colorectal cancer, we do expect some amount of truly differentially abundant OTUs,
but not as many as in diarrhea (e.g. Schubert et al., 2014). This dataset will hopefully
provide an intermediate non-extreme case study.

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(magrittr)
library(SummarizedBenchmark)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "data"
resdir <- "results"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)
```

# Data Preparation

```{r, baxter-download}
crc_data <- file.path(datdir, "crc_baxter_results.tar.gz")
if (!file.exists(crc_data)) {
    download.file("https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz",
                  destfile = crc_data)
}

crc_dir <- file.path(datdir, "crc_baxter_results")
if (!file.exists(file.path(crc_dir, "crc_baxter.metadata.txt"))) {
    untar(crc_data, exdir = datdir)
}
```

Next, we'll read in the unzipped OTU table and metadata files into R.

```{r, baxter-readdata}
## load OTU table and metadata
otu <- read.table(file.path(crc_dir, "RDP",
                            "crc_baxter.otu_table.100.denovo.rdp_assigned"))
meta <- read.csv(file.path(crc_dir, "crc_baxter.metadata.txt"), sep='\t')

## Keep only samples with the right DiseaseState metadata
meta <- dplyr::filter(meta, DiseaseState %in% c("H", "CRC"))

## add "X" in front of sample IDs because of how R read in the OTU table
meta$Sample_Name_s <- paste0("X", meta$Sample_Name_s)

## keep only samples with both metadata and 16S data
keep_samples <- intersect(colnames(otu), meta$Sample_Name_s)
otu <- otu[, keep_samples]
meta <- dplyr::filter(meta, Sample_Name_s %in% keep_samples)
```

A brief aside: what's the count distribution of these OTUs?

```{r, eval=FALSE}
hist(colSums(otu))
```

Since we'll be using OTU-wise covariates, we shouldn't need to perform any
filtering/cleaning of the OTUs, apart from removing any that are all zeros.
(This may happen after removing shallow samples, I think.)
We still apply a minimum threshold of 10 reads per OTU across all samples.
After removing these shallow OTUs, we also get rid of any samples with too few reads. 
We define the minimum number of reads per OTU in `min_otu`, and
the minimum number of reads per sample in `min_sample`.

After we've removed any shallow OTUs and samples, we'll convert the OTU table to
relative abundances.

```{r, baxter-cleandata}
min_otu <- 10
minp_otu <- 0.01
min_sample <- 100

## Remove OTUs w/ <= min reads, w/ <= min prop, samples w/ <= min reads
otu <- otu[rowSums(otu) > min_otu, ]
out <- otu[rowSums(otu > 0) / nrow(otu) > minp_otu, ]
otu <- otu[, colSums(otu) > min_sample]

## Update metadata with new samples
meta <- dplyr::filter(meta, Sample_Name_s %in% colnames(otu))

## Remove empty OTUs
otu <- otu[rowSums(otu) > 0, ]

## Convert to relative abundance
abun_otu <- t(t(otu) / rowSums(t(otu)))

## Add pseudo counts
zeroabun <- 0
abun_otu <- abun_otu + zeroabun
```

# Data Analysis

## Differential Testing

Next, we need to calculate the pvalues, effect size, and standard error for each OTU.
Here, we'll compare CRC vs. healthy. We won't consider the nonCRC adenoma patients.
We'll put these results into a dataframe, and label the columns with the 
standardized names for downstream use (`pval`, `SE`, `effect_size`, `test_statistic`).
The test statistic is the one returned by `wilcox.test()`.

Note that the effect here is calculated as logfold change of mean abundance in controls
relative to cases (i.e. `log(mean_abun[controls]/mean_abun[cases])`)

While we're at it, we'll also calculate the mean abundance and ubiquity (detection rate)
of each OTU. Later, we can assign their values to a new column called `ind_covariate` 
for use in downstream steps.

```{r, baxter-pvals}
resfile <- file.path(resdir, paste0("baxter_results_", nrow(abun_otu), "_OTUs.RData"))
if (!file.exists(resfile)) {
    res <- test_microbiome(abundance = abun_otu, shift = zeroabun,
                           is_case = meta$DiseaseState == "CRC")
    save(res, file = resfile)
} else {
    load(resfile)
}
```

Finally, let's try to add phylogeny as covariates. Here we'll have columns for each separate taxonomic level.

```{r, baxter-addphylo}
res <- tidyr::separate(res, otu, 
                       c("kingdom", "phylum", "class", "order",
                         "family", "genus", "species", "denovo"), 
                       sep=";", remove = FALSE)
```

## Covariate Diagnostics

Here we look to see if the covariates do indeed look informative.

### Ubiquity

```{r, baxter-ubi, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

Something weird is happening with p=0.20 and p=0.40 - maybe this has something to do with ties?
Either way, ubiquity looks to be a bit informative (from the scatter plot, but not really the histograms...)

### Mean Abundance (across non-zero samples)

```{r, baxter-abun, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

### Phylogeny

Let's look at phylum-level stratification first. A priori, I *might* expect
Proteobacteria to be enriched for low p-values? But I don't know if that's
super legit, and Eric doesn't seem to think that phylogeny will be informative at all...

```{r, baxter-phylo, fig.width=10, fig.height=6, eval=FALSE}
ggplot(res, aes(x=pval)) +
    geom_histogram() +
    facet_wrap(~phylum, scales = "free")
```

Not really informative either...

## Multiple-Testing Correction

Let's use `ubiquity` as our `ind_covariate`.

```{r, baxter-labelcovariate}
res <- dplyr::mutate(res, ind_covariate = ubiquity)
```

First, we'll create an object of `BenchDesign` class to hold the data and 
add the benchmark methods to the `BenchDesign` object. We'll do this for a few
values of nmids, to explore its effects on the outputs. We remove ASH from
the default comparison.

Then, we'll construct the `SummarizedBenchmark` object, which will run
the functions specified in each method (these are actually sourced in from the
helper scripts).

```{r, baxter-sb, results="hide", message=FALSE, eval=TRUE}
for (nmids in c(5, 50, 150)){
    print(nmids)
    bd <- initializeBenchDesign(nmids=nmids)
    ## remove ASH from comparison
    bd <- dropBMethod(bd, "ashs")

    resfile <- file.path(resdir, paste0("baxter_summarizedBenchmark_",
                                        nrow(res), "_nmids_", nmids, ".rda"))
    if (!file.exists(resfile)){
        sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
        metadata(sb)$data_download_link <-
                       "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
        save(sb, file = resfile)
    } else {
        load(resfile)
    }
}
```

## Benchmark Metrics

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, baxter-metrics, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, baxter-plot, width=15, height=15, eval=TRUE}
## plot nrejects by method overall and stratified by covariate
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
#plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
```

```{r, eval=FALSE}
assays(sb)[["qvalue"]][, "ashs"] %>% max()
sum(is.na(assays(sb)[["qvalue"]][, "scott-empirical"]))
sum(is.na(assays(sb)[["qvalue"]][, "lfdr"]))
```

# Data Analysis (genus-level)

## Collapse to Genus

```{r, baxter-genus}
## add column with otu names
genus_df <- as.data.frame(abun_otu)
genus_df <- dplyr::as_tibble(genus_df, rownames = "otu")

## just get genus information
genus_df <- dplyr::mutate(genus_df, genus = sapply(strsplit(otu, ";"), `[`, 6))

## gather into tidy format
genus_df <- tidyr::gather(genus_df, key = "sample", value = "abun", -otu, -genus)

## get rid of unannoated genera, and sum abundances for genera
genus_df <- genus_df %>% 
    dplyr::filter(genus != "g__") %>% 
    dplyr::group_by(genus, sample) %>% 
    dplyr::summarise(total_abun = sum(abun)) %>%
    ungroup()

## convert back to longform
genus_df <- tidyr::spread(genus_df, sample, total_abun)
genus_df <- as.data.frame(as.list(dplyr::select(genus_df, -genus)),
                          row.names = genus_df$genus)

## re-order columns to match metadata
genus_df <- genus_df[, match(meta$Sample_Name_s, colnames(genus_df))]

## use matrix
genus <- as.matrix(genus_df)
```

## Differential Testing

```{r, baxter-genus-pvals}
resfile <- file.path(resdir, paste0("baxter_results_", nrow(genus_df), "_genera.RData"))
if (!file.exists(resfile)) {
    res <- test_microbiome(abundance = genus, shift = zeroabun,
                           is_case = meta$DiseaseState == "CRC")
    save(res, file = resfile)
} else {
    load(resfile)
}
```

## Covariate Diagnostics

Here we look to see if the covariates do indeed look informative.

### Ubiquity

```{r, baxter-genus-ubi, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

Something weird is happening with p=0.20 and p=0.40 - maybe this has something to do with ties?
Either way, ubiquity looks to be a bit informative (from the scatter plot, but not really the histograms...)

### Mean Abundance (across non-zero samples)

```{r, baxter-genus-abun, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

## Multiple-Testing Correction

Let's use `ubiquity` as our `ind_covariate`.

```{r, baxter-genus-labelcovariate}
res <- res %>% mutate(ind_covariate = ubiquity)
```

First, we'll create an object of `BenchDesign` class to hold the data and 
add the benchmark methods to the `BenchDesign` object. We'll do this for a few
values of nmids, to explore its effects on the outputs. We remove ASH from
the comparison.

Then, we'll construct the `SummarizedBenchmark` object, which will run
the functions specified in each method (these are actually sourced in from the
helper scripts).

```{r, baxter-genus-sb, results="hide", message=FALSE, eval=TRUE}
for (nmids in c(5, 50, 150)) {
    print(nmids)
    bd <- initializeBenchDesign(nmids=nmids)
    ## remove ASH from comparison
    bd <- dropBMethod(bd, "ashs")

    resfile <- file.path(resdir, paste0("baxter_summarizedBenchmark_", nrow(res),
                                        "_genera_nmids_", nmids, ".rda"))
    if (!file.exists(resfile)) {
        sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
        metadata(sb)$data_download_link <-
                       "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
        save(sb, file = resfile)
    } else {
        load(resfile)
    }
}
```

## Benchmark Metrics

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, baxter-genus-metrics, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, baxter-genus-plot, width=15, height=15, eval=TRUE}
## plot nrejects by method overall and stratified by covariate
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
##plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
```

Note: Benjamini-Hochberg (bh) overlaps exactly with the IHW results.

```{r, baxter-genus-overlap}
plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
```
```{r}
methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
plotCovariateBoxplots( sb, alpha=0.1, nsets=6, methods=methods)
```

# Session Info

```{r}
sessionInfo()
```

### 2018-02-20 datasets/microbiome/microbiome-schubert.Rmd
Add more notes to Schubert; get through Baxter

@@
 ---
-title: "Comparison of FDR methods on 16S microbiome data with SummarizedBenchmark"
+title: "Case Study: 16S microbiome data (Schubert et al., 2014)"
 author: "Claire Duvallet"
@@
 
-Here we download and analyze the CDI Schubert diarrhea (C. diff and non-C. diff diarrhea) dataset. 
-We'll download the processed OTU tables from Zenodo and unzip them in the 
-`benchmark-fdr/datasets/microbiome/DATA` folder.
+Here we download and analyze the Schubert diarrhea dataset [(Schubert et al., 2014)](http://dx.doi.org/10.1128/mBio.01021-14).
+The dataset includes 154 healthy nondiarrheal, 93 Clostridium difficile infection (CDI) associated diarrheal,
+and 89 non-CDI associated diarrheal stool samples. We'll download the processed OTU tables from
+Zenodo and unzip them in the `benchmark-fdr/datasets/microbiome/data` folder.
 
-In the previous document `cdi_schubert.Rmd`, we explored two microbiome datasets for FDR
-benchmarking. Here we will make use of the new `SummarizedBenchmark` package to
-perform the benchmarking in a standardized way.
-
-# Workspace setup
+# Workspace Setup
 
@@
 
-We'll need to manually define which disease labels are of interest and what metadata
-column contains the sample IDs (i.e. the IDs in the OTU table).
-
 ```{r, schubert-readdata, eval=TRUE}
@@
 (This may happen after removing shallow samples, I think.)
-However, let's make sure we get rid of any samples with too few reads. 
-We define the minimum number of reads per sample in `sample_reads`.
+We still apply a minimum threshold of 10 reads per OTU across all samples.
+After removing these shallow OTUs, we also get rid of any samples with too few reads. 
+We define the minimum number of reads per OTU in `min_otu`, and
+the minimum number of reads per sample in `min_sample`.
 
-After we've removed any shallow samples, we'll convert the OTU table to relative
-abundances. I'll add a pseudo-count of 1e-6 to any zero entries, to avoid
-problems with taking logs.
+After we've removed any shallow OTUs and samples, we'll convert the OTU table to
+relative abundances.
 
@@
                            is_case = meta$DiseaseState %in% c("CDI", "nonCDI"))
-    save(res, file=resfile)
+    save(res, file = resfile)
 } else {
@@
 }
-head(res)
 ```
@@
 add the benchmark methods to the `BenchDesign` object. We'll do this for a few
-values of nmids, to explore its effects on the outputs.
+values of nmids, to explore its effects on the outputs. We remove ASH from
+the comparison.
 
@@
                                         nrow(res),  "_nmids_", nmids, ".rda"))
-    duration <- NA
     if (!file.exists(resfile)) {
@@
 ```{r, schubert-plot, width=15, height=15, eval=TRUE}
-# plot nrejects by method overall and stratified by covariate
+## plot nrejects by method overall and stratified by covariate
 rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)

### 2018-02-20 datasets/R/microbiome-helpers.R

@@ -71,3 +71,6 @@ remove_shallow_otus <- function(df, n_reads) {
     return(df[rowSums(df) > n_reads, ])
 }
 
+remove_rare_otus <- function(df, perc_samples){
+  return(df[rowSums(df > 0) / dim(df)[2] > perc_samples, ])
+}

### 2018-02-20 datasets/microbiome/microbiome-baxter.Rmd
Fix yaml header to knit

@@
         number_sections: true
-----
+---
 
@@
 library(dplyr)
+library(tidyr)
 library(ggplot2)

### 2018-02-20 datasets/microbiome/microbiome-goodrich.Rmd
Add Goodrich analysis; fix errors in filtering of other data sets

---
title: "Case Study: 16S microbiome data (Goodrich et al., 2014)"
author: "Claire Duvallet"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

Here we download and analyze the Goodrich obesity (OB) dataset
[(Goodrich et al., 2014)](http://dx.doi.org/10.1016/j.cell.2014.09.053).
The dataset includes fecal samples from 428 individuals with low-to-normal body mass index (BMI < 25),
and 185 obese individuals (BMI > 30). We'll download the processed OTU tables from
Zenodo and unzip them in the `benchmark-fdr/datasets/microbiome/data` folder.

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(magrittr)
library(SummarizedBenchmark)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "data"
resdir <- "results"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)
```

# Data Preparation

```{r, goodrich-download}
ob_data <- file.path(datdir, "ob_goodrich_results.tar.gz")
if (!file.exists(ob_data)) {
    download.file("https://zenodo.org/record/840333/files/ob_goodrich_results.tar.gz",
                  destfile = ob_data)
}

ob_dir <- file.path(datdir, "ob_goodrich_results")
if (!file.exists(file.path(ob_dir, "ob_goodrich.metadata.txt"))) {
    untar(ob_data, exdir = datdir)
}
```

Next, we'll read in the unzipped OTU table and metadata files into R.

We'll need to manually define which disease labels are of interest and what metadata
column contains the sample IDs (i.e. the IDs in the OTU table).

```{r, goodrich-readdata}
## load OTU table and metadata
otu <- read.table(file.path(ob_dir, "RDP",
                            "ob_goodrich.otu_table.100.denovo.rdp_assigned"))
meta <- read.csv(file.path(ob_dir, "ob_goodrich.metadata.txt"), sep='\t')

## Keep only samples with the right DiseaseState metadata
meta <- dplyr::filter(meta, DiseaseState %in% c("H", "OB"))

## keep only samples with both metadata and 16S data
keep_samples <- intersect(colnames(otu), meta$X)
otu <- otu[, keep_samples]
meta <- dplyr::filter(meta, X %in% keep_samples)
```

Since we'll be using OTU-wise covariates, we shouldn't need to perform any
filtering/cleaning of the OTUs, apart from removing any that are all zeros. 
(This may happen after removing shallow samples, I think.)
We still apply a minimum threshold of 10 reads per OTU across all samples.
After removing these shallow OTUs, we also get rid of any samples with too few reads. 
We define the minimum number of reads per OTU in `min_otu`, and
the minimum number of reads per sample in `min_sample`.

After we've removed any shallow samples, we'll convert the OTU table to relative
abundances.

```{r, goodrich-cleandata}
min_otu <- 10
minp_otu <- 0.01
min_sample <- 100

## Remove OTUs w/ <= min reads, w/ <= min prop, samples w/ <= min reads
otu <- otu[rowSums(otu) > min_otu, ]
otu <- otu[rowSums(otu > 0) / ncol(otu) > minp_otu, ]
otu <- otu[, colSums(otu) > min_sample]

# Update metadata with new samples
meta <- dplyr::filter(meta, X %in% colnames(otu))

# Remove empty OTUs
otu <- otu[rowSums(otu) > 0, ]

# Convert to relative abundance
abun_otu <- t(t(otu) / rowSums(t(otu)))

# Add pseudo counts
zeroabun <- 0
abun_otu <- abun_otu + zeroabun
```

# Data Analysis

## Differential Testing

Next, we need to calculate the pvalues, effect size, and standard error for each OTU.
Here, we'll compare lean vs. obese. We'll put these results into a dataframe, and label 
the columns with the standardized names for downstream use 
(`pval`, `SE`, `effect_size`, `test_statistic`). The test statistic is the one returned 
by `wilcox.test()`.

Note that the effect here is calculated as logfold change of mean abundance in controls
relative to cases (i.e. `log(mean_abun[controls]/mean_abun[cases])`).

While we're at it, we'll also calculate the mean abundance and ubiquity (detection rate)
of each OTU. Later, we can assign their values to a new column called `ind_covariate` 
for use in downstream steps.

Note that OB Goodrich has ~70,000 OTUs, so this is going to take a while.

```{r, goodrich-pvals}
resfile <- file.path(resdir, paste0("goodrich_results_", nrow(abun_otu), "_OTUs.rda"))
if (!file.exists(resfile)) {
    res <- test_microbiome(abundance = abun_otu, shift = zeroabun,
                           is_case = meta$DiseaseState == "OB")
    save(res, file = resfile)
} else {
    load(resfile)
}
```

Finally, let's try to add phylogeny as covariates. Here we'll have columns for each separate taxonomic level.

```{r, goodrich-addphylo}
res <- tidyr::separate(res, otu, 
                       c("kingdom", "phylum", "class", "order",
                         "family", "genus", "species", "denovo"), 
                       sep=";", remove = FALSE)
```

## Covariate Diagnostics

Here we look to see if the covariates do indeed look informative.

### Ubiquity

```{r, goodrich-ubi, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=5, binwidth=0.05, numQ=4)
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

### Mean Abundance (across non-zero samples)

```{r, goodrich-abun, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=5, binwidth=0.05, numQ=4)
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

### Phylogeny

Let's look at phylum-level stratification first. A priori, I *might* expect
Proteobacteria to be enriched for low p-values? But I don't know if that's
super legit, and Eric doesn't seem to think that phylogeny will be informative at all...

```{r, goodrich-phylo, fig.width=10, fig.height=6, eval=FALSE}
##strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=20)
##rank_scatter(res, pvalue="pval", covariate="ubiquity")
ggplot(res, aes(x=pval)) + geom_histogram() + facet_wrap(~phylum, scales = "free")
```

## Multiple-Testing Correction

Let's use `ubiquity` as our `ind_covariate`.

```{r, goodrich-labelcovariate}
res <- dplyr::mutate(res, ind_covariate = ubiquity)
```

First, we'll create an object of `BenchDesign` class to hold the data and 
add the benchmark methods to the `BenchDesign` object. We'll do this for a few
values of nmids, to explore its effects on the outputs. We remove ASH from
the default comparison.

Next, we'll construct the `SummarizedBenchmark` object, which will run
the functions specified in each method (these are actually sourced in from the
helper scripts).

```{r, goodrich-sb, results="hide", message=FALSE, eval=TRUE}
for (nmids in c(5, 50, 150)) {
    ## use common BenchDesign w/out ASH
    bd <- initializeBenchDesign(nmids = nmids)
    bd <- dropBMethod(bd, "ashs")

    resfile <- file.path(resdir, paste0("goodrich_summarizedBenchmark_",
                                        nrow(res), "_nmids_", nmids, ".rda"))
    if (!file.exists(resfile)) {
        sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
        metadata(sb)$data_download_link <-
                       "https://zenodo.org/record/840333/files/ob_goodrich_results.tar.gz"
        save(sb, file = resfile)
    } else {
        load(resfile)
    }
}
```

## Benchmark Metrics

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, goodrich-metrics, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, goodrich-plot, width=15, height=15, eval=TRUE}
## plot nrejects by method overall and stratified by covariate
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
##plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE )
```

There is no upsetR plot for OB Goodrich for now because only lfdr is returning any rejections.

```{r, eval=FALSE}
assays(sb)[["qvalue"]][, "ashs"] %>% max()
sum(is.na(assays(sb)[["qvalue"]][, "scott-empirical"]))
sum(is.na(assays(sb)[["qvalue"]][, "lfdr"]))
```

# Data Analysis (genus-level)

## Collapse to Genus

```{r, goodrich-genus}
## add column with otu names
genus_df <- as.data.frame(abun_otu)
genus_df <- dplyr::as_tibble(genus_df, rownames = "otu")

## just get genus information
genus_df <- dplyr::mutate(genus_df, genus = sapply(strsplit(otu, ";"), `[`, 6))

## gather into tidy format
genus_df <- tidyr::gather(genus_df, key = "sample", value = "abun", -otu, -genus)

## get rid of unannoated genera, and sum abundances for genera
genus_df <- genus_df %>% 
    dplyr::filter(genus != "g__") %>% 
    dplyr::group_by(genus, sample) %>% 
    dplyr::summarise(total_abun = sum(abun)) %>%
    ungroup()

## convert back to longform
genus_df <- tidyr::spread(genus_df, sample, total_abun)
genus_df <- as.data.frame(as.list(dplyr::select(genus_df, -genus)),
                          row.names = genus_df$genus)

## re-order columns to match metadata
genus_df <- genus_df[, match(meta$X, colnames(genus_df))]

## use matrix
genus <- as.matrix(genus_df)
```

## Differential Testing

```{r, goodrich-genus-pvals}
resfile <- file.path(resdir, paste0("goodrich_results_", nrow(genus_df), "_genera.rda"))
if (!file.exists(resfile)) {
    res <- test_microbiome(abundance = genus, shift = zeroabun,
                           is_case = meta$DiseaseState == "OB")
    save(res, file = resfile)
} else {
    load(resfile)
}
```

## Covariate Diagnostics

Here we look to see if the covariates do indeed look informative.

### Ubiquity

```{r, goodrich-genus-ubi, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

Something weird is happening with p=0.20 and p=0.40 - maybe this has something to do with ties?
Either way, ubiquity looks to be a bit informative (from the scatter plot, but not really the histograms...)

### Mean Abundance (across non-zero samples)

```{r, goodrich-genus-abun, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

## Multiple-Testing Correction

Let's use `ubiquity` as our `ind_covariate`.

```{r, goodrich-genus-labelcovariate}
res <- dplyr::mutate(res, ind_covariate = ubiquity)
```

#### Set up BenchDesign object

First, we'll create an object of `BenchDesign` class to hold the data and 
add the benchmark methods to the `BenchDesign` object. We'll do this for a few
values of nmids, to explore its effects on the outputs.

Then, we'll construct the `SummarizedBenchmark` object, which will run
the functions specified in each method (these are actually sourced in from the
helper scripts).

```{r, goodrich-genus-sb, results="hide", message=FALSE, eval=TRUE}
for (nmids in c(5, 50, 150)) {
    ## use common BenchDesign w/out ASH
    bd <- initializeBenchDesign(nmids = nmids)
    bd <- dropBMethod(bd, "ashs")

    resfile <- file.path(resdir, paste0("goodrich_summarizedBenchmark_",
                                        nrow(res), "_genera_nmids_", nmids, ".rda"))
    if (!file.exists(resfile)) {
        sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
        metadata(sb)$data_download_link <-
                       "https://zenodo.org/record/840333/files/ob_goodrich_results.tar.gz"
        save(sb, file = resfile)
    } else {
        load(resfile)
    }
}
```

## Benchmark Metrics

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, goodrich-genus-metrics, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, goodrich-genus-plot, width=15, height=15, eval=TRUE}
# plot nrejects by method overall and stratified by covariate
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
#plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
```

Note: Benjamini-Hochberg (bh) overlaps exactly with the IHW results.

```{r, goodrich-genus-overlap}
plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
```
```{r, goodrich-genus-covariates}
methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
plotCovariateBoxplots( sb, alpha=0.1, nsets=6, methods=methods)
```

# Session Info

```{r}
sessionInfo()
```

### 2018-02-20 datasets/microbiome/microbiome-schubert.Rmd
Add Goodrich analysis; fix errors in filtering of other data sets

@@
 ## Remove OTUs w/ <= min reads, samples w/ <= min reads
-otu <- otu[, colSums(otu) > min_otu]
-otu <- otu[rowSums(otu) > min_sample, ]
+otu <- otu[rowSums(otu) > min_otu, ]
+otu <- otu[, colSums(otu) > min_sample]
 
@@
 for (nmids in c(5, 50, 150)) {
-    print(nmids)
+    ## use common BenchDesign w/out ASH
     bd <- initializeBenchDesign(nmids = nmids)
-    ## remove ASH from comparison
     bd <- dropBMethod(bd, "ashs")

### 2018-02-20 datasets/ChIP-seq/ChIP-seq.Rmd
Clean ChIP-seq RMD, match other vignettes

--- 2017-09-15 datasets/ChIP-seq/ChIP-seq.Rmd
@@
 ---
-title: "ChIP-seq datasets for FDR benchmarking"
+title: "Case Study: ChIP-seq Differential Peak Calling"
 author: "Mingxiang Teng"
-date: "9/14/2017"
-output: html_document
+date: '`r format(Sys.Date(), "%B %d, %Y")`'
+output: 
+   html_document:
+        toc: true
+        toc_float: true
+        highlight: tango
+        number_sections: true
 ---
@@
 
-### Summary
+# Summary
 
-For FDR benchmark using ChIP-seq datasets, we may focus on peak calling and
-differential peaks. Peak calling is the most common way that people analyze 
-ChIP-seq since this technique was invented, though some groups started to care
-about differential peaks in some complex experiment designs. For peak calling,
-peak callers always report effect size (they call "enrichment signal") 
-and p-value, but standard error. I think it is too complicate if we dig into
-their code for generating standard error. For differential peaks, classical
-pipeline would call peaks first, then recruit existing algorithms (*e.g.* DESeq2)
-to test the covariant of interest, in a way that all effect size, p-value and 
-standard error can be generated. However, this pipeline probably gives 
-very similar patterns (distributions) of effect size / p-value to those
-based on differential gene expression analysis. Thus, differential peak 
-analysis might not be useful in our FDR benchmark.
+Common analyses in ChIP-seq data include the calling of peaks within individual
+samples, and the testing of differential peaks across conditions. While peak
+calling continues to be the most wide-spread method applied to ChIP-seq data,
+differential testing has gained popularity with the increased availability of
+data from large-scale experiments with more complex designs.
 
-### Oestrogen receptor binding case
+In differential peak calling, classical pipelines often begin by first calling peaks
+in samples, followed by performing differential testing using existing algorithms
+developed for other genomic data tpyes (e.g. RNA-seq). However, with these approaches,
+testing is only performed on a filtered set of genomic regions, namely those with
+significance in at least in one condition, resulting in only a limited number of
+null tests cases.
 
-This dataset have plenty of replicates so that we can focus on both peak
-calling and differential peaks. And it is the first dataset people used to 
-illustrate differential peak detection, where the tools DiffBind was 
-introduced.
+To avoid this limitation, we examine ChIP-seq data of the histone H3K4Me3, an active
+marker of gene expression, and focus our analysis on promoter regions of all genes
+for this histone marker. By doing so, we avoid limiting the set of tests to only
+regions significant in at least one sample, while also keeping most regions of
+interest for this histone marker.
 
-We can download the processed data from GEO.
+Here, we focus on test differential peaks between two cell lines (K562 and Gm12878)
+using publicly available data generated by the ENCODE Project. These two cell lines
+comprise the largest sequencing datasets in the ENCODE Project. As mentioned above,
+we use ChIP-seq data for histone H3K4Me3 in our analysis. Differential peaks are tested
+using DESeq2.
 
-```{bash, chipseq-download, eval=FALSE}
-curl -O "https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE32222&format=file"
-tar -xvf "?acc=GSE32222&format=file"
-rm "?acc=GSE32222&format=file"
+We note that while the algorithm for differential testing, DESeq2, used here is the same
+as that used in our analysis of RNA-seq data, benchmarking on ChIP-seq still provides
+insights for a unique data structure which has been widely utilized in modern genomics.
+
+# Workspace Setup
+
+```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(magrittr)
+library(SummarizedBenchmark)
+library(BiocParallel)
+library(GenomicRanges)
+library(TxDb.Hsapiens.UCSC.hg19.knownGene)
+library(DESeq2)
+library(Rsubread)
+
+## load helper functions
+for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
+    source(f)
+}
+
+## project data/results folders
+datdir <- "data"
+resdir <- "results"
+dir.create(datdir, showWarnings = FALSE)
+dir.create(residr, showWarnings = FALSE)
+
+## set up parallel backend
+cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
+multicoreParam <- MulticoreParam(workers = cores)
 ```
 
-For one sample, extract the peak calling results.
+# Data Preparation
 
-```{bash, chipseq-unzip1, eval=FALSE}
-gunzip "GSM798383_SLX-1201.250.s_4_SLX-1202.250.s_1_peaks.txt.gz"
-rm *peaks.txt.gz
+We downloaded the bam files directly from UCSC ENCODE portal.
+
+```{r, encode-download}
+broad_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeBroadHistone/"
+broad_bams <- c("wgEncodeBroadHistoneGm12878H3k4me3StdAlnRep1.bam",
+                "wgEncodeBroadHistoneGm12878H3k04me3StdAlnRep2V2.bam",
+                "wgEncodeBroadHistoneK562H3k4me3StdAlnRep1.bam",
+                "wgEncodeBroadHistoneK562H3k4me3StdAlnRep2.bam")
+
+for (i_bam in broad_bams) {
+    if (!file.exists(file.path(datdir, i_bam))) {
+        download.file(paste0(broad_url, i_bam),
+                      destfile = file.path(datdir, i_bam))
+    }
+    i_bai <- paste0(i_bam, ".bai")
+    if (!file.exists(file.path(datdir, i_bai))) {
+        download.file(paste0(broad_url, i_bai),
+                      destfile = file.path(datdir, i_bai))
+    }
+}
+
+uw_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwHistone/"
+uw_bams <- c("wgEncodeUwHistoneGm12878H3k4me3StdAlnRep1.bam",
+             "wgEncodeUwHistoneGm12878H3k4me3StdAlnRep2.bam",
+             "wgEncodeUwHistoneK562H3k4me3StdAlnRep1.bam",
+             "wgEncodeUwHistoneK562H3k4me3StdAlnRep2.bam")
+
+for (i_bam in uw_bams) {
+    if (!file.exists(file.path(datdir, i_bam))) {
+        download.file(paste0(uw_url, i_bam),
+                      destfile = file.path(datdir, i_bam))
+    }
+    i_bai <- paste0(i_bam, ".bai")
+    if (!file.exists(file.path(datdir, i_bai))) {
+        download.file(paste0(uw_url, i_bai),
+                      destfile = file.path(datdir, i_bai))
+    }
+}
 ```
 
-And load in to R.
-```{r, chipseq-read1}
-peaks <- read.table("~/GSM798383_SLX-1201.250.s_4_SLX-1202.250.s_1_peaks.txt", header=TRUE, sep='\t')
-dim(peaks)
-head(peaks)
+We determine sample metadata from the file names.
+
+```{r}
+bamfiles <- list.files(datdir, pattern = 'bam$', full = TRUE)
+labs <- gsub("wgEncode(.*)Histone.*", "\\1", basename(bamfiles))
+cells <- gsub("wgEncode.*Histone(.*)H3k.*", "\\1", basename(bamfiles))
+
+meta <- data.frame(cellline = cells, lab = labs, file = bamfiles)
 ```
 
-Here, "X.10.log10.pvalue." is the -log10P, and fold_enrichment is the effect size.
+Next, we identify promotor regions using the UCSC "Known Gene" annotations for
+human genome assembly hg19 (GRCh37).
+
+```{r}
+prom <- promoters(genes(TxDb.Hsapiens.UCSC.hg19.knownGene))
+seqlevels(prom, force = TRUE) <- seqlevels(prom)[!grepl("_", seqlevels(prom))]
+```
+
+# Data Analysis
+
+## Differential Testing
+
+We count sequencing reads for all promoter regions here, and test for differential peaks using these counts.
+
+```{r count, results='hide', message=FALSE, warning=FALSE}
+deseq_res  <- file.path(resdir, "h3k4me3.rda")
+if (file.exists(deseq_res)) {
+    load(deseq_res)
+} else {
+    ## count reads in each promotor region
+    anno <- data.frame(GeneID = seq_len(length(prom)),
+                       Chr = seqnames(prom),
+                       Start = start(prom),
+                       End = end(prom),
+                       Strand = strand(prom))
+    rc <- featureCounts(files = bamfiles,
+                        annot.ext = anno,
+                        allowMultiOverlap = TRUE,
+                        minOverlap = 50,
+                        readExtension3 = 150,
+                        ignoreDup = TRUE)$counts
+
+    ## perform differential testing
+    dds <- DESeqDataSetFromMatrix(countData = rc,
+                                  colData = as.data.frame(meta),
+                                  design= ~ lab + cellline)
+    dds <- DESeq(dds, fitType = "mean")
+    dat <- as.data.frame(results(dds))
+    colnames(dat) <- c('ind_covariate', 'effect_size', 'SE',
+                       'stat', 'pval', 'padj')
+
+    ## compute test-statistic from p-value for ASH
+    dat <- dplyr::mutate(dat,
+                         test_statistic = qnorm(exp(log(pval)-log(2)),
+                                                lower.tail=FALSE),
+                         test_statistic = test_statistic * sign(effect_size))
+
+    save(deseq, dat, file = deseq_res)
+}
+
+## remove regions with effect sizes = NA
+dat <- dat[!is.na(dat[, 2]), ]
+```
+
+## Covariate Diagnostics
+
+Here, we can see that the covariant variable is actually informative.
+
+### Mean Coverage
+
+We consider the average coverage across conditions and replicates as a
+potential covariate of interest. This is motivated by our prior assumption that
+ChIP-seq data always involves two sets of regions: signal and background.
+Sometimes, there are a few regions with signal coverage between signal and background.
+In this example, we see that p-values are differently distributed under low, 
+median and high coverage.
+
+```{r, chipseq-diag, fig.width=10, fig.height=3.5}
+strat_hist(dat, pvalue = "pval", covariate = "ind_covariate", maxy = 2)
+rank_scatter(dat, pvalue = "pval", covariate = "ind_covariate")
+```
+
+## Multiple-Testing Correction
+
+We use the common `BenchDesign` with the set of multiple testing correction
+methods already included.
+
+```{r, chipseq-benchdesign}
+bd <- initializeBenchDesign()
+```
+
+Now, we're ready to construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method.
+
+```{r, chipseq-sb, results="hide", message=FALSE}
+resN <- paste0(resdir, "/chipseq_summarizedBenchmark_", nrow(dat), ".RData")
+if (!file.exists(resN)) {
+    sb <- buildBench(bd, data = dat, ftCols = c("ind_covariate"),
+                     parallel = TRUE, BPPARAM = multicoreParam)
+    save(sb, file = resN)
+} else {
+    load(resN)
+}
+```
+
+## Benchmark Metrics
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, chipseq-metrics}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, chipseq-plot, results="hide", width=15, height=15}
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter(sb, supplementary = FALSE)
+
+rejection_scatter_bins(sb, covariate = "ind_covariate",
+                       bins = 4, supplementary = FALSE)
+
+plotFDRMethodsOverlap(sb,  alpha = 0.05, nsets = ncol(sb),
+                      order.by = "freq", decreasing = TRUE,
+                      supplementary = FALSE)
+
+# properties of method-exclusive discoveries
+methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
+plotCovariateBoxplots(sb, alpha = 0.1, nsets = 6,
+                      methods = methods, trans = log2, maxNum = 5)
+```
+
+# Session Info
+
+```{r}
+sessionInfo()
+```

### 2018-02-20 datasets/ChIP-seq/ChIP-seq.sh

@@ -1,17 +1,16 @@
 #!/bin/bash
-#SBATCH -J knitChIPseq                 # A single job name for the array
-#SBATCH -n 8                        # Number of cores
-#SBATCH -N 1                        # All cores on one machine
-#SBATCH -p irizarry,serial_requeue  # Partition
-#SBATCH --mem 60000                 # Memory request 
-#SBATCH -t 0-5:00                   # Maximum execution time (D-HH:MM)
-#SBATCH -o render-%j.out  # Standard output
-#SBATCH -e render-%j.err  # Standard error
+#SBATCH -J knitChIPseq
+#SBATCH -n 8
+#SBATCH -N 1
+#SBATCH -p irizarry,serial_requeue
+#SBATCH --mem 8GB
+#SBATCH -t 0-05:00
+#SBATCH -o render-%j.out
+#SBATCH -e render-%j.err
  
-module load rstudio
 export SLURM_NTASKS
 export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/rstudio/0.98.1103-fasrc01/bin/pandoc/"
 
 # change filename to Rmd to be knitted. ncores is passed in through environment
 # variable SLURM_NTASKS
-R -e "rmarkdown::render('~/workspace/cowork/fdr/benchmark-fdr/datasets/ChIP-seq/ChIP-seq.Rmd')"
+R -e "rmarkdown::render('ChIP-seq.Rmd')"

### 2018-02-20 datasets/microbiome/microbiome-joint.Rmd
Rename/move microbiome joint analysis

---
title: "Exploration into microbiome qvalues"
author: "Claire Duvallet"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

In this notebook, I want to see what's going on with all the different methods for three microbiome datasets.
It seems that the two datasets I've used so far (ob_goodrich, with ~70,000 OTUs but very few true differentially abundant OTUs, and cdi_schubert, with ~18,000 OTUs and very many true differentially abundant OTUs) are getting weird results in different ways and with different methods.

My hypotheses so far are that:

- ash is rejecting all hypotheses
  - True
- lfdr is also being affected by the `nmids`/`df` [error](https://github.com/pkimes/benchmark-fdr/issues/16#issuecomment-345844180)
  - probably not a problem
- decreasing nmids doesn't fix the error/problem with Scott (FDRreg)
  - True

To look into these, I'll run SummarizedBenchmark for each value of nmids (5, 50, 150) for each dataset, and plot histograms of the resulting q values for each method.

Note that I ran SummarizedBenchmark for different values of nmids in `microbiome-SummarizedBenchmark.Rmd` and saved the results to files.

```{r, load-wkspc, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)

library(SummarizedBenchmark)

setwd("/Users/claire/github/benchmark-fdr/datasets/microbiome/")
datdir <- "/Users/claire/github/benchmark-fdr/datasets/microbiome/DATA/"
resdir <- "/Users/claire/github/benchmark-fdr/datasets/microbiome/RESULTS/"

```

# CDI Schubert

```{r, cdi-schubert}
dflst = list()
i <- 1
for (nmids in c(5, 50, 150)){
  print(nmids)
  resfile <- paste0(resdir, "schubert_summarizedBenchmark_",
                    "18703", "_nmids_", nmids, ".RData")
  load(resfile)

  # Extract qvalues and melt into tidy dataframe
  qvals <- assay(sb) %>% as.data.frame %>% gather("method", "qval")
  qvals$nmids <- nmids

  # TO DO: add/concatenate to other things?
  dflst[[i]] <- qvals
  i <- i + 1
}
allres <- bind_rows(dflst)
```
```{r, nmids-effect}
nmids_qvals <- c()
# First, check which methods are affected by nmids
for (met in unique(allres$method)){
  n5 <- allres %>% filter(method == met) %>% filter(nmids == 5)
  n50 <- allres %>% filter(method == met) %>% filter(nmids == 50)
  n150 <- allres %>% filter(method == met) %>% filter(nmids == 150)
  if (!(n5$qval == n50$qval && n50$qval == n150$qval) | is.na(n5$qval == n50$qval && n50$qval == n150$qval)){
    nmids_qvals <- c(nmids_qvals, met)
  }
}
nmids_qvals
```
## scott and lfdr (are messed up)

LFDR seems to work okay (though its values do depend on nmids), but Scott either returns all NaN's or all zeros...

```{r, plot-qvals}
allres %>% 
  filter(method %in% nmids_qvals) %>% 
  #mutate(logqval = log10(qval)) %>%
  ggplot(aes(x=qval)) + geom_histogram() + facet_grid(nmids ~ method)
```

```{r}
## Look at number of NaN q-values for each method
allres %>% filter(method %in% c(nmids_qvals, "ashs")) %>% group_by(method, nmids) %>% summarize(n_nan = sum(is.na(qval)))

```

### ashs (return all q-values = 0)

And one final note: all ashs qvalues are basically 0...

```{r}
allres %>% filter(method == "ashs") %>% group_by(nmids) %>% summarize(min = min(qval), max = max(qval))
```

### All other methods (seem OK)

Let's also plot the q-value distributions for all methods.

```{r}
allres %>% 
  filter(!method %in% nmids_qvals) %>% 
  filter(!method == "ashs") %>%
  #filter(!method == "unadjusted") %>%
  group_by(method, qval) %>% filter(row_number() == 1) %>%
  ggplot(aes(x=qval)) + geom_histogram() + facet_grid(method ~ .) + theme(strip.text.y = element_text(angle = 0))

```

# OB Goodrich

```{r, ob-goodrich}
dflst = list()
i <- 1
for (nmids in c(5, 50, 150)){
  print(nmids)
  resfile <- paste0(resdir, "goodrich_summarizedBenchmark_",
                    "71710", "_nmids_", nmids, ".RData")
  load(resfile)

  # Extract qvalues and melt into tidy dataframe
  qvals <- assay(sb) %>% as.data.frame %>% gather("method", "qval")
  qvals$nmids <- nmids

  # TO DO: add/concatenate to other things?
  dflst[[i]] <- qvals
  i <- i + 1
}
allres <- bind_rows(dflst)
```
```{r, nmids-effect-goodrich}
nmids_qvals <- c()
# First, check which methods are affected by nmids
for (met in unique(allres$method)){
  n5 <- allres %>% filter(method == met) %>% filter(nmids == 5)
  n50 <- allres %>% filter(method == met) %>% filter(nmids == 50)
  n150 <- allres %>% filter(method == met) %>% filter(nmids == 150)
  if (!(n5$qval == n50$qval && n50$qval == n150$qval) | is.na(n5$qval == n50$qval && n50$qval == n150$qval)){
    nmids_qvals <- c(nmids_qvals, met)
  }
}
nmids_qvals
```

## scott and lfdr (are messed up)

LFDR seems to work okay (though its values do depend on nmids), but Scott either returns all NaN's or all zeros...

```{r, plot-qvals-goodrich}
allres %>% 
  filter(method %in% nmids_qvals) %>% 
  #mutate(logqval = log10(qval)) %>%
  ggplot(aes(x=qval)) + geom_histogram() + facet_grid(nmids ~ method)
```
```{r}
## Look at number of NaN q-values for each method
allres %>% filter(method %in% c(nmids_qvals, "ashs")) %>% group_by(method, nmids) %>% summarize(n_nan = sum(is.na(qval)))

```

### ashs (return all q-values = 0)

And one final note: all ashs qvalues are basically 0...

```{r}
allres %>% filter(method == "ashs") %>% group_by(nmids) %>% summarize(min = min(qval), max = max(qval))
```

### All other methods (seem OK)

Let's also plot the q-value distributions for all methods.

```{r}
allres %>% 
  filter(!method %in% nmids_qvals) %>% 
  filter(!method == "ashs") %>%
  #filter(!method == "unadjusted") %>%
  group_by(method, qval) %>% filter(row_number() == 1) %>%
  ggplot(aes(x=qval)) + geom_histogram() + facet_grid(method ~ .) + theme(strip.text.y = element_text(angle = 0))

```

# CRC Baxter

```{r, crc-baxter}
dflst = list()
i <- 1
for (nmids in c(5, 50, 150)){
  print(nmids)
  resfile <- paste0(resdir, "baxter_summarizedBenchmark_",
                    "91843", "_nmids_", nmids, ".RData")
  load(resfile)

  # Extract qvalues and melt into tidy dataframe
  qvals <- assay(sb) %>% as.data.frame %>% gather("method", "qval")
  qvals$nmids <- nmids

  # TO DO: add/concatenate to other things?
  dflst[[i]] <- qvals
  i <- i + 1
}
allres <- bind_rows(dflst)
```

```{r, nmids-effect-baxter}
nmids_qvals <- c()
# First, check which methods are affected by nmids
for (met in unique(allres$method)){
  n5 <- allres %>% filter(method == met) %>% filter(nmids == 5)
  n50 <- allres %>% filter(method == met) %>% filter(nmids == 50)
  n150 <- allres %>% filter(method == met) %>% filter(nmids == 150)
  if (!(n5$qval == n50$qval && n50$qval == n150$qval) | is.na(n5$qval == n50$qval && n50$qval == n150$qval)){
    nmids_qvals <- c(nmids_qvals, met)
  }
}
nmids_qvals
```

## scott and lfdr

LFDR seems to work okay (though its values do depend on nmids), but Scott either returns all NaN's or all zeros...

```{r, plot-qvals-baxter}
allres %>% 
  filter(method %in% nmids_qvals) %>% 
  #mutate(logqval = log10(qval)) %>%
  ggplot(aes(x=qval)) + geom_histogram() + facet_grid(nmids ~ method)
```
```{r}
## Look at number of NaN q-values for each method
allres %>% filter(method %in% c(nmids_qvals, "ashs")) %>% group_by(method, nmids) %>% summarize(n_nan = sum(is.na(qval)))

```

### ashs

And one final note: all ashs qvalues are basically 0...

```{r}
allres %>% filter(method == "ashs") %>% group_by(nmids) %>% summarize(min = min(qval), max = max(qval))
```

### All other methods

Let's also plot the q-value distributions for all methods.

```{r}
allres %>% 
  filter(!method %in% nmids_qvals) %>% 
  filter(!method == "ashs") %>%
  #filter(!method == "unadjusted") %>%
  group_by(method, qval) %>% filter(row_number() == 1) %>%
  ggplot(aes(x=qval)) + geom_histogram() + facet_grid(method ~ .) + theme(strip.text.y = element_text(angle = 0))

```

# Conclusions:

- not sure what's going on with the scott-empirical and scott-theoretical. Seems like the same df error is being thrown, but it does run to completion *but also* all the q-values are zero (or NaN).
- ashs is also returning every q-value = 0. This is the one that needs stderr, right? Maybe I am calculating this wrong. Should check with Keegan/Patrick.

### 2018-02-20 datasets/R/microbiome-helpers.R

@@ -53,6 +53,7 @@ test_microbiome <- function(abundance, shift, is_case) {
         res$effect_size[i] <- log( mean(ctrldf[i, ]) / mean(casedf[i, ]) )
     }
 
+    ## compute test-statistic from p-value for ASH
     res <- dplyr::mutate(res,
                          test_statistic = qnorm(exp(log(pval) - log(2)),
                                                 lower.tail=FALSE),
@@ -61,7 +62,6 @@ test_microbiome <- function(abundance, shift, is_case) {
 
 
 remove_shallow_smpls <- function(df, n_reads) {
-    ## Removes samples with fewer than n_reads from dataframe df.
     ## df has OTUs in rows and samples in columns
     return(df[, colSums(df) > n_reads])
 }

### 2018-02-20 datasets/ChIP-seq/ChIP-seq.Rmd
Clean up RNA-Seq analysis

@@
 library(dplyr)
-library(tidyr)
 library(ggplot2)
-library(magrittr)
 library(SummarizedBenchmark)

### 2018-02-20 datasets/RNA-Seq/RNA-Seq.Rmd
Clean up RNA-Seq analysis

---
title: "Case Study: RNA-Seq Differential Analysis"
author: "Chinmay Shukla"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

**Add brief summary of analysis goals and data set.**

# Workspace Setup

```{r, workspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(DESeq2)
library(SummarizedBenchmark)
library(BiocParallel)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "data"
resdir <- "results"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)

## set up parallel backend
multicoreParam <- MulticoreParam(workers = 2)
```

# Data Preparation

```{r, dsd-download}
dsd_data <- file.path(datdir, "dsdObjects.RData")
if (!file.exists(dsd_data)) {
    download.file("https://www.dropbox.com/s/8hf0ccl27ircodz/dsdObjects.RData?dl=1",
                  destfile = dsd_data)
}
load(dsd_data)
```

**Add details on content of data and subsetting.**

```{r}
## Filter to include samples from only 2 tissues
tissues <-  c('Brain - Nucleus accumbens (basal ganglia)',
              'Brain - Putamen (basal ganglia)')
filteredDsd1 <- dsd1[, colData(dsd1)$tissue %in% tissues]
filteredDsd1$tissue <- droplevels(filteredDsd1$tissue)
```

# Data Analysis

## Differential Testing

**Add details on how testing is performed.**

```{r}
res_rda <- file.path(resdir, "dsd1-results.rda")
if (!file.exists(res_rda)) {
    ## build DESeq object and run DESeq2
    dds <- DESeq(filteredDsd1, parallel = TRUE, BPPARAM = multicoreParam)
    res <- results(dds, independentFiltering = FALSE)

    save(dds, res, file = res_rda)
} else {
    load(res_rda)
}
```

**Add details on filtering.**

```{r}
## build input data.frame
geneExp <- data.frame(geneName = rownames(res),
                      pval = res$pvalue,
                      SE = res$lfcSE, 
                      ind_covariate = res$baseMean,
                      effect_size = res$log2FoldChange,
                      test_statistic = res$stat)
geneExp <- dplyr::as_tibble(geneExp)

## filter data
geneExp <- na.omit(geneExp)
geneExp <- dplyr::filter(geneExp, ind_covariate >= 1)
```

## Covariate Diagnostics

**Add description of covariates.**

### Mean Coverage

```{r}
strat_hist(geneExp, pvalue = "pval",
           covariate = "ind_covariate", maxy = 25)
```

```{r}
rank_scatter(geneExp, pvalue = "pval", covariate = "ind_covariate")
```

## Multiple-Testing Correction

We use the common `BenchDesign` with the set of multiple testing correction
methods already included.

```{r}
bd <- initializeBenchDesign()
sb <- buildBench(bd, data = geneExp, parallel = TRUE)
```

## Benchmark Metrics

```{r}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

**Add description of results.**

```{r}
rejections_scatter(sb, supplementary = FALSE)
```

```{r}
plotFDRMethodsOverlap(sb, alpha = 0.05, nsets = ncol(sb),
                      order.by = "freq", decreasing = TRUE,
                      supplementary = FALSE)
```

# Session Info

```{r}
sessionInfo()
```

### 2018-02-21 datasets/scrnaseq/scRNAseq-latest.Rmd
AS:added 2 more datasets, helper fxn

@@
 knitr::opts_chunk$set(echo = TRUE,cache=TRUE)
-#knitr::opts_knit$set(root.dir = "/n/irizarryfs01_backed_up/kkorthauer/FDR")
 
@@
 #install.packages("FDRreg")
-sourceDirectory("benchmark-fdr/datasets/R/")
 
@@
 library(BiocParallel)
-cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
+cores = 6
 multicoreParam <- MulticoreParam(workers = cores)
-sourceDirectory('~/Documents/research/fdr_jc/src/benchmark-fdr/datasets/R/')
+sourceDirectory('benchmark-fdr/datasets/R/')
 ```
@@
 ```{r echo=FALSE, results='asis'}
-df=data.frame(datasets=c("GSE63818-GPL16791","GSE45719"),organism=c("human","mouse"),technology=c("smart-seq","smart-seq"))
+df=data.frame(datasets=c("GSE63818-GPL16791","GSE45719"),organism=c("human","human","mouse","mouse"),technology=c("smart-seq","smart-seq","smart-seq2","smart-seq2"))
 kable(df,caption="scRNAseq Datasets")
@@
 curl -o data/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/report-scater/GSE63818-GPL16791_scater.html
+
+## human ds2
+curl -o data/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE84465.rds
 ```
@@
 #save data
-saveRDS(g1_data,"h_g1_data.rds")
-saveRDS(g2_data,"h_g2_data.rds")
+saveRDS(g1_data,"h_g1_data_GSE63818-GPL16791.rds")
+saveRDS(g2_data,"h_g2_data_GSE63818-GPL16791.rds")
 #COUNTS
@@
 #save data
-saveRDS(g1_cdata,"h_g1_cdata.rds")
-saveRDS(g2_cdata,"h_g2_cdata.rds")
+saveRDS(g1_cdata,"h_g1_cdata_GSE63818-GPL16791.rds")
+saveRDS(g2_cdata,"h_g2_cdata_GSE63818-GPL16791.rds")
+
+#load compressed mao
+humandata2=readRDS("GSE84465.rds")
+#get gene-level data
+genes=experiments(humandata2)[1]
+#TPMS
+#get tpms
+tpms=assays(genes$gene)[1]
+#get tpm dataframe
+human_tpms=tpms$TPM
+# get column data/metadata for sample
+cdata=colData(humandata2)
+cdata=as.data.frame(cdata)
+# select groups of interest
+g1 = cdata%>%subset(characteristics_ch1.7=="neoplastic: Neoplastic" & characteristics_ch1.5=="tsne cluster: 11")
+g2 = cdata%>%subset(characteristics_ch1.7=="neoplastic: Regular"& characteristics_ch1.5=="tsne cluster: 8")
+# select maximal sample size
+sample_size = c(dim(g1)[1],dim(g2)[1])
+ssize = which.min(sample_size)
+g1_ids=sample.int(dim(g1)[1], size = sample_size[ssize], replace = FALSE)
+g2_ids=sample.int(dim(g2)[1], size = sample_size[ssize], replace = FALSE)
+# get group data of maximal size
+g1_data=human_tpms[,colnames(human_tpms)%in%g1$geo_accession[g1_ids]]
+g2_data=human_tpms[,colnames(human_tpms)%in%g2$geo_accession[g2_ids]]
+#save data
+saveRDS(g1_data,"h_g1_data_GSE84465.rds")
+saveRDS(g2_data,"h_g2_data_GSE84465rds")
+#COUNTS
+#get counts
+count=assays(genes$gene)[2]
+#get count dataframe
+human_count=count$count
+# get column data/metadata for sample
+cdata=colData(humandata2)
+cdata=as.data.frame(cdata)
+# get group data of maximal size
+g1_cdata=human_count[,colnames(human_count)%in%g1$geo_accession[g1_ids]]
+g2_cdata=human_count[,colnames(human_count)%in%g2$geo_accession[g2_ids]]
+#save data
+saveRDS(g1_cdata,"h_g1_cdata_GSE84465.rds")
+saveRDS(g2_cdata,"h_g2_cdata_GSE84465.rds")
 
@@
 
-For the human data, each group has 26 cells and 65218 genes. We merge the data and remove all genes which have 0s in all samples
+For the human dataset1, each group has 26 cells and 65218 genes. We merge the data and remove all genes which have 0s in all samples
 
-```{r, nofilter, eval=FALSE}
+```{r, nofilter1, eval=FALSE}
 g_data=cbind(g1_data,g2_data)
-g_data_sub=g_data[rowSums(g_data==0)!=52,] # 0.25*52=13
-saveRDS(g_data_sub,"h_g_nofilter_subdata.rds") # removing rows with all zeros reduces the number of genes to 33169
+g_data_sub=g_data[rowSums(g_data==0)!=dim(g_data)[2],] 
+saveRDS(g_data_sub,"h_g_nofilter_subdata_GSE63818-GPL16791.rds") # removing rows with all zeros reduces the number of genes to 33169
 
 g_cdata=cbind(g1_cdata,g2_cdata)
-g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=52,]
-saveRDS(g_cdata_sub,"h_g_nofilter_subcdata.rds") # removing rows with all zeros reduces the number of genes to 33169
+g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=dim(g_cdata)[2],]
+saveRDS(g_cdata_sub,"h_g_nofilter_subcdata_GSE63818-GPL16791.rds") # removing rows with all zeros reduces the number of genes to 33169
 
-saveRDS(g1_ids,"g1_ids.rds")
-saveRDS(g2_ids,"g2_ids.rds")
+saveRDS(g1_ids,"g1_ids_GSE63818-GPL16791.rds")
+saveRDS(g2_ids,"g2_ids_GSE63818-GPL16791.rds")
 ```
 
-Next, we repeat the processing steps for the mouse data.
+For the human dataset2, each group has 695 cells and 65218 genes. We merge the data and remove all genes which have 0s in all samples
+
+```{r, nofilter2, eval=FALSE}
+g_data=cbind(g1_data,g2_data)
+g_data_sub=g_data[rowSums(g_data==0)!=dim(g_data)[2],] 
+saveRDS(g_data_sub,"h_g_nofilter_subdata_GSE84465.rds") # removing rows with all zeros reduces the number of genes to 52942
+
+g_cdata=cbind(g1_cdata,g2_cdata)
+g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=dim(g_cdata)[2],]
+saveRDS(g_cdata_sub,"h_g_nofilter_subcdata_GSE84465.rds") # removing rows with all zeros reduces the number of genes to 52942
+
+saveRDS(g1_ids,"g1_ids_GSE84465.rds")
+saveRDS(g2_ids,"g2_ids_GSE84465.rds")
+```
+
+Next, we repeat the processing steps for the mouse datasets.
 
@@
 #save data
-saveRDS(g1_data,"m_g1_data.rds")
-saveRDS(g2_data,"m_g2_data.rds")
+saveRDS(g1_data,"m_g1_data_GSE45719.rds")
+saveRDS(g2_data,"m_g2_data_GSE45719.rds")
 #COUNTS
@@
 #save data
-saveRDS(g1_cdata,"m_g1_cdata.rds")
-saveRDS(g2_cdata,"m_g2_cdata.rds")
+saveRDS(g1_cdata,"m_g1_cdata_GSE45719.rds")
+saveRDS(g2_cdata,"m_g2_cdata_GSE45719.rds")
+
+mousedata2=readRDS("GSE94383.rds")
+#get gene-level data
+genes=experiments(mousedata2)[1]
+#TPMS
+#get tpms
+tpms=assays(genes$gene)[1]
+#get tpm dataframe
+mouse_tpms=tpms$TPM
+# get column data/metadata for sample
+cdata=colData(mousedata2)
+cdata=as.data.frame(cdata)
+# select groups of interest
+g1 = cdata%>%subset(characteristics_ch1.2=="condition: No stimulation")
+g2 = cdata%>%subset(characteristics_ch1.2=="condition: LPS"&characteristics_ch1.1=="time point: 150 min")
+# select maximal sample size
+sample_size = c(dim(g1)[1],dim(g2)[1])
+ssize = which.min(sample_size)
+g1_ids=sample.int(dim(g1)[1], size = sample_size[ssize], replace = FALSE)
+g2_ids=sample.int(dim(g2)[1], size = sample_size[ssize], replace = FALSE)
+# get group data of maximal size
+g1_data=mouse_tpms[,colnames(mouse_tpms)%in%g1$geo_accession[g1_ids]]
+g2_data=mouse_tpms[,colnames(mouse_tpms)%in%g2$geo_accession[g2_ids]]
+#save data
+saveRDS(g1_data,"m_g1_GSE94383.rds")
+saveRDS(g2_data,"m_g2_GSE94383.rds")
+#COUNTS
+#get counts
+count=assays(genes$gene)[2]
+#get count dataframe
+mouse_count=count$count
+# get column data/metadata for sample
+cdata=colData(mousedata2)
+cdata=as.data.frame(cdata)
+# get group data of maximal size
+g1_cdata=mouse_count[,colnames(mouse_count)%in%g1$geo_accession[g1_ids]]
+g2_cdata=mouse_count[,colnames(mouse_count)%in%g2$geo_accession[g2_ids]]
+#save data
+saveRDS(g1_cdata,"m_g1_c_GSE94383.rds")
+saveRDS(g2_cdata,"m_g2_c_GSE94383.rds")
 
@@
 
-For the mouse data, each group has 50 cells and 45686 genes. We merge the data and remove all genes which have 0s in all samples
+For the mouse dataset1, each group has 50 cells and 45686 genes. We merge the data and remove all genes which have 0s in all samples
 
@@
 g_data_sub=g_data[rowSums(g_data==0)!=dim(g_data)[2],] # 0.25*52=13
-saveRDS(g_data_sub,"m_g_nofilter_subdata.rds") # removing rows with all zeros reduces the number of genes to 33169
+saveRDS(g_data_sub,"m_g_nofilter_subdata_GSE45719.rds") # removing rows with all zeros reduces the number of genes to 33169
 
@@
 g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=dim(g_data)[2],]
-saveRDS(g_cdata_sub,"m_g_nofilter_subcdata.rds") # removing rows with all zeros reduces the number of genes to 33169
+saveRDS(g_cdata_sub,"m_g_nofilter_subcdata_GSE45719.rds") # removing rows with all zeros reduces the number of genes to 33169
 
-saveRDS(g1_ids,"m_g1_ids.rds")
-saveRDS(g2_ids,"m_g2_ids.rds")
+saveRDS(g1_ids,"m_g1_ids_GSE45719.rds")
+saveRDS(g2_ids,"m_g2_ids_GSE45719.rds")
+```
+
+For the mouse dataset2, each group has 202 cells and 25179 genes after filtering for non-detected genes.
+
+```{r, m_nofilter_ds2, eval=FALSE}
+g_data=cbind(g1_data,g2_data)
+g_data_sub=g_data[rowSums(g_data==0)!=dim(g_data)[2],] # 0.25*52=13
+saveRDS(g_data_sub,"m_g_nofilter_subdata_GSE94383.rds") # removing rows with all zeros reduces the number of genes to 25179
+
+g_cdata=cbind(g1_cdata,g2_cdata)
+g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=dim(g_data)[2],]
+saveRDS(g_cdata_sub,"m_g_nofilter_subcdata_GSE94383.rds") # removing rows with all zeros reduces the number of genes to 25179
+
+saveRDS(g1_ids,"m_g1_ids_GSE94383.rds")
+saveRDS(g2_ids,"m_g2_ids_GSE94383.rds")
 ```
@@
 We perform differential expression testing using 3 methods on the unfiltered data: non-parametric Wilcoxon, MAST which has a hurdle model to account for zeros and scDD which implicitly models any heterogeneity in expression.
+We will first assign names for all files
+
+```{r, filenames}
+h_wilcox_dat1="h_data_wilcox_p.rds"
+h_wilcox_dat2="h_data_wilcox_p_GSE84465.rds"
+m_wilcox_dat1="m_data_wilcox_p_GSE45719.rds"
+m_wilcox_dat2="m_data_wilcox_p_GSE94383.rds"
+
+h_scdd_dat1="humanplate_scDatEx_ks_GSE63818-GPL16791.rds"
+h_scdd_dat2="humanplate_scDatEx_ks_GSE84465.rds"
+m_scdd_dat1="mouseplate_scDatEx_ks_GSE45719.rds"
+m_scdd_dat2="mouseplate_scDatEx_ks_GSE94383.rds"
+
+m_mast_dat1="mouseplate_masttpm_GSE45719.rds"
+h_mast_dat1="humanplate_masttpm_GSE63818-GPL16791.rds"
+h_mast_dat2="humanplate_masttpm_GSE84465.rds"
+m_mast_dat2="mouseplate_masttpm_GSE94383.rds"
+
+```
 
@@
 ```{r, wilcoxon, eval=FALSE}
+
 g_data_sub=readRDS("h_g_nofilter_subdata.rds")
-groups=c(rep(1,26),rep(2,26))
-compute_wilcoxde_sc(g_data_sub,"h_data_wilcox_p.rds",groups)
+groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
+compute_wilcoxde_sc(g_data_sub,h_wilcox_dat1,groups)
 
-g_data_sub=readRDS("m_g_nofilter_subdata.rds")
-groups=c(rep(1,50),rep(2,50))
-compute_wilcoxde_sc(g_data_sub,"m_data_wilcox_p.rds",groups)
+g_data_sub=readRDS("h_g_nofilter_subdata_GSE84465.rds")
+groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
+compute_wilcoxde_sc(g_data_sub,h_wilcox_dat2,groups)
+
+g_data_sub=readRDS("m_g_nofilter_subdata_GSE45719.rds")
+groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
+compute_wilcoxde_sc(g_data_sub,m_wilcox_dat1,groups)
+
+g_data_sub=readRDS("m_g_nofilter_subdata_GSE94383.rds")
+groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
+compute_wilcoxde_sc(g_data_sub,m_wilcox_dat2,groups)
 ## warnings:
@@
 ## human data
-compute_scdd_sc(g_cdata_sub,"h_sizefactors.rds","h_data_sce.rds","humanplate_scDatEx_ks.rds")
+g_cdata_sub=readRDS("h_g_nofilter_subcdata_GSE63818-GPL16791.rds")
+compute_scdd_sc(g_cdata_sub,"h_sizefactors_GSE63818-GPL16791.rds","h_data_sce_GSE63818-GPL16791.rds",h_scdd_dat1)
 ##Notice: 16096 genes have less than 3 nonzero cells per condition.  Skipping these genes.
@@
 # perm=1000
+g_cdata_sub=readRDS("h_g_nofilter_subcdata_GSE84465.rds")
+compute_scdd_sc(g_cdata_sub,"h_sizefactors_GSE84465.rds","h_data_sce_GSE84465.rds",h_scdd_dat2)
+# Notice: 22390 genes have less than 3 nonzero cells per condition.  Skipping these genes.
+# Clustering observed expression data for each gene
+# Setting up parallel back-end using 6 cores
 
 ## mouse data
-compute_scdd_sc(g_cdata_sub,"m_sizefactors.rds","m_data_sce.rds","mouseplate_scDatEx_ks.rds")
+g_cdata_sub=readRDS("m_g_nofilter_subcdata_GSE45719.rds")
+compute_scdd_sc(g_cdata_sub,"m_sizefactors_GSE45719.rds","m_data_sce_GSE45719.rds",m_scdd_dat1)
 # Notice: 10246 genes have less than 3 nonzero cells per condition.  Skipping these genes.
@@
 # Classifying significant genes into patterns
+g_cdata_sub=readRDS("m_g_nofilter_subcdata_GSE94383.rds")
+compute_scdd_sc(g_cdata_sub,"m_sizefactors_GSE94383.rds","m_data_sce_GSE94383.rds",m_scdd_dat2)
+# Notice: 8982 genes have less than 3 nonzero cells per condition.  Skipping these genes.
+# Clustering observed expression data for each gene
+# Setting up parallel back-end using 6 cores
 
@@
 ```{r, MAST,eval=FALSE}
-g_data_sub=readRDS("m_g_nofilter_subdata.rds")
-groups=c(rep(1,50),rep(2,50))
+g_data_sub=readRDS("m_g_nofilter_subdata_GSE45719.rds")
+groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
 names(groups)=colnames(g_cdata_sub)
-compute_MASTtpmde(g_data_sub,groups,"mouseplate_masttpm.rds")
+compute_MASTtpmde(g_data_sub,groups,)
 
-g_data_sub=readRDS("h_g_nofilter_subdata.rds")
-groups=c(rep(1,26),rep(2,26))
+g_data_sub=readRDS("h_g_nofilter_subdata_GSE63818-GPL16791.rds")
+groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
 names(groups)=colnames(g_data_sub)
-compute_MASTtpmde(g_data_sub,groups,"humanplate_masttpm.rds")
+compute_MASTtpmde(g_data_sub,groups,h_mast_dat1)
+
+g_data_sub=readRDS("h_g_nofilter_subdata_GSE84465.rds")
+groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
+names(groups)=colnames(g_data_sub)
+compute_MASTtpmde(g_data_sub,groups,h_mast_dat2)
+
+g_data_sub=readRDS("m_g_nofilter_subdata_GSE94383.rds")
+groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
+names(groups)=colnames(g_data_sub)
+compute_MASTtpmde(g_data_sub,groups,m_mast_dat2)
 ```
@@
 m_cov=compute_covariates_scrnaseq("m_g_nofilter_subdata.rds")
+h2_cov=compute_covariates_scrnaseq("h_g_nofilter_subdata_GSE84465.rds")
+m2_cov=compute_covariates_scrnaseq("m_g_nofilter_subdata_GSE94383.rds")
 ```
@@
 
-```{r, wilcoxcov, fig.width=10, fig.height=3.5}
-print("human_plate")
-wilcox=readRDS("h_data_wilcox_p.rds")
-h_wilcox_cov=data.frame(genes=names(wilcox),pval=wilcox)
-h_wilcox_cov=h_wilcox_cov%>%left_join(h_cov,by=c("genes"))
+```{r, wilcoxcov, fig.width=10, fig.height=7}
+print("human_plate_1")
+hwilcox1 <- plot_covariate_diagnostics(paste0(h_wilcox_dat1),"wilcox",h_cov)
+print(hwilcox1$plot)
 
-print_plot_cov_diagnosis(h_wilcox_cov)
-print("With filter")
-print_plot_cov_diagnosis(h_wilcox_cov%>%subset(detection_rate>=0.05))
+print("human_plate_2")
+hwilcox2 <- plot_covariate_diagnostics(paste0(h_wilcox_dat2),"wilcox",h2_cov)
+print(hwilcox2$plot)
 
-print("mouse_plate")
-wilcox=readRDS("m_data_wilcox_p.rds")
-m_wilcox_cov=data.frame(genes=names(wilcox),pval=wilcox)
-m_wilcox_cov=m_wilcox_cov%>%left_join(m_cov,by=c("genes"))
-print_plot_cov_diagnosis(m_wilcox_cov)
-print("With filter")
-print_plot_cov_diagnosis(m_wilcox_cov%>%subset(detection_rate>=0.05))
+print("mouse_plate_1")
+mwilcox1 <- plot_covariate_diagnostics(paste0(m_wilcox_dat1),"wilcox",m_cov)
+print(mwilcox1$plot)
 
+print("mouse_plate_2")
+mwilcox2 <- plot_covariate_diagnostics(paste0(m_wilcox_dat2),"wilcox",m2_cov)
+print(mwilcox2$plot)
 ```
@@
 
-```{r, mastcov, fig.width=10, fig.height=3.5}
-print("human_plate")
-hmast=readRDS("humanplate_masttpm.rds")
-h_mast_cov=data.frame(genes=rownames(hmast),pval=hmast$pval)
-h_mast_cov=h_mast_cov%>%left_join(h_cov,by=c("genes"))
-print_plot_cov_diagnosis(h_mast_cov)
-print("With filter")
-print_plot_cov_diagnosis(h_mast_cov%>%subset(detection_rate>=0.05))
+```{r, mastcov, fig.width=10, fig.height=7}
+print("human_plate_1")
+hmast1 <- plot_covariate_diagnostics(paste0(,h_mast_dat1),"mast",h_cov)
+print(hmast1$plot)
 
-print("mouse_plate")
-mmast=readRDS("mouseplate_masttpm.rds")
-m_mast_cov=data.frame(genes=rownames(mmast),pval=mmast$pval)
-m_mast_cov=m_mast_cov%>%left_join(m_cov,by=c("genes"))
-print_plot_cov_diagnosis(m_mast_cov)
-print("With filter")
-print_plot_cov_diagnosis(m_mast_cov%>%subset(detection_rate>=0.05))
+print("human_plate_2")
+hmast2 <- plot_covariate_diagnostics(paste0(h_mast_dat2),"mast",h2_cov)
+print(hmast2$plot)
+
+print("mouse_plate_1")
+mmast1 <- plot_covariate_diagnostics(paste0(m_mast_dat1),"mast",m_cov)
+print(mmast1$plot)
+
+print("mouse_plate_2")
+mmast2 <- plot_covariate_diagnostics(paste0(m_mast_dat2),"mast",m2_cov)
+print(mmast2$plot)
 ```
@@
 
-```{r,scddcov, fig.width=10, fig.height=3.5}
-print("human_plate")
-h_scdd=readRDS("humanplate_scDatEx_ks.rds")
-h_scdd_cov=data.frame(genes=rownames(metadata(h_scdd)$Genes),pval=metadata(h_scdd)$Genes$nonzero.pvalue)
-h_scdd_cov$genes=h_cov$genes #scdd changes gene names by appending numeric indices 
-h_scdd_cov=h_scdd_cov%>%subset(!is.na(pval))
-h_scdd_cov=h_scdd_cov%>%left_join(h_cov,by=c("genes"))
-print_plot_cov_diagnosis(h_scdd_cov)
-print("With filter")
-print_plot_cov_diagnosis(h_scdd_cov%>%subset(detection_rate>=0.05))
+```{r,scddcov, fig.width=10, fig.height=7}
+print("human_plate_1")
+hscdd1 <- plot_covariate_diagnostics(paste0(h_scdd_dat1),"scdd",h_cov)
+print(hscdd1$plot)
 
-print("mouse_plate")
-m_scdd=readRDS("mouseplate_scDatEx_ks.rds")
-m_scdd_cov=data.frame(genes=rownames(metadata(m_scdd)$Genes),pval=metadata(m_scdd)$Genes$nonzero.pvalue)
-m_scdd_cov$genes=m_cov$genes
-m_scdd_cov=m_scdd_cov%>%subset(!is.na(pval))
-m_scdd_cov=m_scdd_cov%>%left_join(m_cov,by=c("genes"))
-print_plot_cov_diagnosis(m_scdd_cov)
-print("With filter")
-print_plot_cov_diagnosis(m_scdd_cov%>%subset(detection_rate>=0.05))
+print("human_plate_2")
+hscdd2 <- plot_covariate_diagnostics(paste0(h_scdd_dat2),"scdd",h2_cov)
+print(hscdd2$plot)
 
+print("mouse_plate_1")
+mscdd1 <- plot_covariate_diagnostics(paste0(m_scdd_dat1),"scdd",m_cov)
+print(mscdd1$plot)
+
+print("mouse_plate_2")
+mscdd2 <- plot_covariate_diagnostics(paste0(m_scdd_dat2),"scdd",m2_cov)
+print(mscdd2$plot)
 ```
@@
 
-First we will create the benchdesign object for wilcox
+First we will create the benchdesign object for wilcox -- IN PROGRESS
 
@@
   t1 <- proc.time()
-  new_sb <- bd %>% buildBench(data=(m_scdd_cov%>%select(pval,mean_exp)%>%mutate(effect_size=NA, test_statistic=NA, SE=NA) %>% mutate(ind_covariate=mean_exp)), 
+  scrna_sb <- bd %>% buildBench(data=(hscdd1$data%>%select(pval,mean_exp)%>%mutate(effect_size=0, test_statistic=0, SE=0) %>% mutate(ind_covariate=mean_exp)), 
                           ftCols = c("mean_exp"),
@@
 
-  save(sb, file=resN)
+  save(scrna_sb, file=resN)
   duration <- round((proc.time()-t1)[3]/60,1)
@@
 
-assayNames(sb) <- "qvalue"
-sb <- addDefaultMetrics(sb,metrics=c("rejections"))
+assayNames(scrna_sb) <- "qvalue"
+scrna_sb <- addDefaultMetrics(scrna_sb)
 
 # plot nrejects by method overall and stratified by covariate
-rejections_scatter(sb,
-                   supplementary=FALSE)
+rejections_scatter(scrna_sb,supplementary=FALSE)
 
-rejection_scatter_bins(sb, covariate="ind_covar_N", bins=4,
-                       supplementary=FALSE)
+rejection_scatter_bins(scrna_sb, covariate="ind_covar_N", bins=3,supplementary=FALSE)
 
 # upset plot 
-plotFDRMethodsOverlap(sb, 
-                      alpha=0.05, nsets=ncol(sb),
+plotFDRMethodsOverlap(scrna_sb, 
+                      alpha=0.05, nsets=ncol(scrna_sb),
                       order.by="freq", decreasing=TRUE,
@@
 methods <- c("lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
-mcols(sb)$ind_covariate <- mcols(sb)$ind_covar_N
-plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3)
+mcols(scrna_sb)$ind_covariate <- mcols(scrna_sb)$ind_covar_N
+plotCovariateBoxplots(scrna_sb, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3)

### 2018-02-21 datasets/R/fdr-benchmark-scrnaseq-helper.R

@@ -129,3 +129,57 @@ print_plot_cov_diagnosis <- function(cov_data){
   p=rank_scatter(cov_data,pvalue ="pval",covariate="detection_rate")
   print(p)
 }
+
+cowplot_cov_diagnosis <- function(cov_data){
+  #print("median_exp")
+  #p=strat_hist(cov_data,pvalue ="pval",covariate="median_exp",numQ=5,maxy=10)
+  #print(p)
+  
+  p1=strat_hist(cov_data,pvalue ="pval",covariate="median_exp",numQ=3,maxy=10)
+  p2=strat_hist(cov_data,pvalue ="pval",covariate="mean_exp",numQ=3,maxy=10)
+  p3=strat_hist(cov_data,pvalue ="pval",covariate="detection_rate",numQ=3,maxy=10)
+  
+  p4=rank_scatter(cov_data,pvalue ="pval",covariate="median_exp")
+  p5=rank_scatter(cov_data,pvalue ="pval",covariate="mean_exp")
+  p6=rank_scatter(cov_data,pvalue ="pval",covariate="detection_rate")
+  mean_p=plot_grid(p4,p5,p6, ncol = 3,nrow=1) +draw_label("median_exp", x = 0.15, y = 0.97,fontface="bold") +draw_label("mean_exp", x = 0.45, y = 0.97,fontface="bold")+draw_label("det_rate", x = 0.85, y = 0.97,fontface="bold")
+  
+  #print("detection_rate")
+  #p=strat_hist(cov_data,pvalue ="pval",covariate="detection_rate",numQ=5,maxy=10)
+  #print(p)
+  #print("mean_exp")
+  #p=strat_hist(cov_data,pvalue ="pval",covariate="mean_exp",numQ=5,maxy=10)
+  #print(p)
+  
+  #detr_p=plot_grid(p3, p4, labels = c("mean_exp", "mean_exp"), ncol = 1,nrow=2)
+  p<-plot_grid(p1, p2, p3, mean_p, ncol = 1,nrow=4) +draw_label("median_exp", x = 0.5, y = 0.99,fontface="bold") +draw_label("mean_exp", x = 0.5, y = 0.75,fontface="bold")+draw_label("det_rate", x = 0.5, y = 0.5,fontface="bold")
+  return(p)
+  
[…]

### 2018-02-21 datasets/scrnaseq/scRNAseq-latest.Rmd
AS:fixed typos

@@
 ```{r, filenames}
-h_wilcox_dat1="h_data_wilcox_p.rds"
+h_wilcox_dat1="h_data_wilcox_p_GSE63818-GPL16791.rds"
 h_wilcox_dat2="h_data_wilcox_p_GSE84465.rds"
@@
 
-g_data_sub=readRDS("h_g_nofilter_subdata.rds")
+g_data_sub=readRDS("h_g_nofilter_subdata_GSE63818-GPL16791.rds")
 groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
@@
 print("human_plate_1")
-hmast1 <- plot_covariate_diagnostics(paste0(,h_mast_dat1),"mast",h_cov)
+hmast1 <- plot_covariate_diagnostics(paste0(h_mast_dat1),"mast",h_cov)
 print(hmast1$plot)

### 2018-02-22 datasets/R/common-BenchDesign.R
Move around helper functions

@@ -118,3 +118,53 @@ addDefaultMetrics <- function(sb) {
                                assay = "qvalue")
     return(sb)
 }
+
+
+## adapted from IHWpaper::clfdr()
+clfdr_hickswrapper <- function(unadj_p, groups, lfdr_estimation="fdrtool") {
+  
+  ## estimate local fdr within each stratum first
+  lfdr_res <- lfdr_fit(unadj_p, groups, lfdr_estimation=lfdr_estimation)
+  lfdrs <- lfdr_res$lfdr
+  
+  ## now use the rejection rule described in Cai's paper
+  
+  ## Remark:
+  ## When sorting lfdrs, we break ties by pvalues so that in the end within each stratum
+  ## we get monotonic adjusted p-values as a function of the p-values
+  ## This is mainly needed for grenander based lfdrs, with most other
+  ## lfdr estimation methods lfdr ties are not a problem usually
+  
+  o <- order(lfdrs, unadj_p)
+  lfdrs_sorted <- lfdrs[o]
+  fdr_estimate <- cumsum(lfdrs_sorted)/(1:length(unadj_p))
+  adj_p <- rev(cummin(rev(fdr_estimate)))
+  adj_p <- adj_p[order(o)]
+  return(adj_p)
+}
+
+
+## helper function for cai
+#' @importFrom fdrtool fdrtool
+lfdr_fit <- function(unadj_p, group, lfdr_estimation="fdrtool"){
+    
+    pvals_list <- split(unadj_p, group)
+    
+    if (lfdr_estimation == "fdrtool"){
+        lfdr_fun <- function(pv) fdrtool::fdrtool(pv, statistic="pvalue",plot=FALSE,verbose=FALSE)$lfdr
+        
+    } else if (lfdr_estimation == "locfdr"){
+        if (!requireNamespace("locfdr", quietly=TRUE)){
+            stop("locfdr package required for this function to work.")
+        }
+        lfdr_fun <- function(pv) locfdr::locfdr(qnorm(pv), nulltype=0, plot=0)$fdr
+    } else {
+        stop("This lfdr estimation method is not available.")
+    }
+    
+    lfdr_list <- lapply(pvals_list, lfdr_fun)
+    lfdrs <- unsplit(lfdr_list, group)
+    
+    fit_obj <- data.frame(pvalue=unadj_p, lfdr=lfdrs, group=group)
+    fit_obj
+}

### 2018-02-22 datasets/R/du_psim.R

@@ -1,97 +1,0 @@
-#' Simulate Collection of Test Statistics
-#'
-#' Function for simulating a large set of (independent) test results.
-#' Unlike `du_tsim.R` which simulates actual data sets to which two-sample
-#' t-tests are applied, this function directly simulates test statistics
-#' and p-values without any underlying data.
-#' This simulator more closely represents the set of simulations carried
-#' out in the manuscripts of the Boca-Leek, ASH, and IHW methods.
-#' (For IHW, in particular, the simulations illustrating the "size investing strategy"
-#' fall under this class of simulations.)
-#' 
-#' @param m number of hypothesis tests. 
-#' @param pi0 proportion of true null hypotheses, can either be a single numeric value
-#'        between 0 and 1 or a function which takes a numeric vector and returns a
-#'        vector of the same length containing null probabilities between 0 and 1. 
-#' @param icovariate specification for the independent covariate, must be a function
-#'        which takes an integer value and returns a vector of the specified length to
-#'        be used as the independent covariate.
-#' @param tstat expected test statistic under the alternative, can either be a single
-#'        numeric value or a function which takes an integer as input and returns a
-#'        numeric vector of test statistics of the specified length.
-#' @param tstat_dist sampling distribution of test statistic, must be a function
-#'        which takes a vector of test statistics, and returns a list of perturbed
-#'        test statistics of the same length.
-#' @param null_dist distribution under the null used to calculate p-values from the
-#'        test statistic, must be a function which takes the full vector of null and
-#'        alternative test statistics and return the corresponding p-value. 
-#' @param seed integer seed for random number generator, ignored if NULL. (default = NULL) 
-#'
-#' @return
-#' data.frame of test results for `m` simulated data sets, with columns:
-#' * `H`: 0/1 indicator whether data simulated under null (0) or alternative (1)
-#' * `test_statistic`: simulated test statistic (NOT scaled by SE estimate)
-#' * `effect_size`: z-score transform of p-value (not really effect size)
-#' * `pval`: test p-value calculated from test-statistic using `null_dist`
[…]

### 2018-02-22 datasets/R/funcs_pi0.R

@@ -1,102 +1,0 @@
-## ##############################################################################
-## START CODE FROM EXTERNAL SOURCE
-## source: https://github.com/SiminaB/Fdr-regression/blob/71b123f/functions.R
-## license: not listed
-## date: 2017/10/17
-## modified: Patrick Kimes
-## ##############################################################################
-
-##------Functions of covariates-------##
-pi0_f1 <- function(x) {
-    p2 <- -0.2
-    p1 <- 1.2
-    a <- 4/(p1-p2)^2
-    
-    y <- -a*(x-p1)*(x-p2)
-    y[x >= 0.7] <- -a*(0.7-p1)*(0.7-p2)
-    y[x <= (p1+p2)/2] <- 1
-    y
-}
-pi0_f2 <- function(x) {
-    y <- rep(0, length=length(x))
-    y[x >= 0.7] <- -2.5*(x[x >= 0.7]-0.7)^2
-    y
-}
-pi0_f3 <- function(x) {
-    y <- rep(0, length=length(x))
-    y[x < 0.7] <- -(x[x < 0.7]-0.1)^2
-    y[x >= 0.7] <- -(min(x[x >= 0.7])-0.1)^2
-    y[x<=0.1] <- 0
-    y
-}
-
-##smooth function of one covariate for different levels of second covariate
-pi0_smooth2 <- function(x1, x2) {
-    y1 <- pi0_f1(x1)
-    y2 <- pi0_f2(x1)
-    y3 <- pi0_f3(x1)
-    y <- rep(0, length(x1))
-    y[x2 == 1] <- y1[x2 == 1] + y2[x2 == 1] + 0.12*y3[x2 == 1]
-    y[x2 == 2] <- y1[x2 == 2] + 0.5*y2[x2 == 2] + 0.06*y3[x2 == 2]
-    y[x2 == 3] <- y1[x2 == 3] + 0.3*y2[x2 == 3] 
-    y
-}
-
-##smooth function of a single variable
-##note: assumes x in (0, 1)
-pi0_smooth1 <- function(x) {
-    y1 <- pi0_f1(x)
-    y2 <- pi0_f2(x)
-    y3 <- pi0_f3(x)
-    y <- rep(0, length(x))
-    y <- y1 + y2 + 0.12*y3
-    y
-}
-
-## ##############################################################################
-## END CODE FROM EXTERNAL SOURCE
-## ##############################################################################
-
[…]

### 2018-02-22 datasets/GSEA/gsea-mouse.Rmd
Split GSEA analyses

---
title: "Case Study: Gene Set Enrichment Analysis (GSEA), Mouse"
author: "Alejandro Reyes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

**Some sort of summary of the analysis.**
Same analysis but using the HSC-MPP dataset from Nina.  
**Who is Nina?**

# Workspace Setup

```{r}
library(dplyr)
library(ggplot2)
library(scales)
library(DESeq2)
library(EnsDb.Mmusculus.v75)
library(biomaRt)
library(goseq)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "data"
resdir <- "results"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)
```

# Data Preparation

Download *DESeqDataSet* if not present locally.

```{r, message=FALSE}
dsd_data <- "dsdHSCMPP.RData"
if(!file.exists(dsdMouse)) {
    download.file("https://www.dropbox.com/s/spo1pifum5k72rn/dseHSCMPP.RData?dl=1",
                  destfile = dsd_data)
}
load(dsd_data)
```

**Why no filtering on gene_biotype == "protein_coding" for this data?**

# Data Analysis

## Differential Testing

Specify model matrix design and run DESeq2 to get differentially expressed genes.

```{r deseq2, message=FALSE}
dseHSCMPP <- DESeq(dseHSCMPP)
res <- results(dseHSCMPP, contrast = c("conditions", "HSC", "MPP1"),
               independentFiltering = FALSE)
```

## Enrichment Analysis

**Add description of what's going on.**

```{r goseq, message=FALSE}
goset_data <- file.path(datdir, "goSetsMm.RData")
if (!file.exists(goset_data)) {
    mart <- useMart("ensembl", "mmusculus_gene_ensembl")
    goSets <- getBM(c("ensembl_gene_id", "go_id"), mart = mart,
                    filters = "ensembl_gene_id", values = rownames(res))
    goSets <- goSets[!nchar(goSets$go_id) == 0, ]
    goSets <- with(goSets, split(go_id, ensembl_gene_id))
    save(goSets, file = file)
} else {
   load(goset_data)
}

## getting median transcript length
genes <- as.numeric(res$padj < 0.1)
names(genes) <- rownames(res)

## getting median transcript length
txByGene <- transcriptsBy(EnsDb.Mmusculus.v75, "gene")
geneLength <- sapply(width(txByGene), median)
geneLength <- geneLength[names(genes)]
genes[is.na( genes )] <- 0
```

Performing a gene set enrichment analysis with `goseq`.

```{r}
pwf <- nullp(genes, bias.data = geneLength)
goRes <- goseq(pwf, gene2cat = goSets)
```

## Covariate Diagnostics

Here, we can see that the covariant variable is actually informative.

### Gene Set Size

**Add description of independent covariate.**

```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
rank_scatter(dat = goRes, pval = "over_represented_pvalue", 
              covariate = "numInCat", bins = 50, funx = log2, 
              funfill = log10_trans()) +
    ylim(0, 12) +
    ggtitle("Over-represented gene sets") +
    xlab(expression(log[10]~"(# of genes)")) +
    ylab(expression(-log[10]~"(p-value)")) 
```

```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
## filter numInCat below only used for plotting
## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
#dplyr:::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)
dplyr::filter(goRes, numDEInCat > 0, numInCat > 10) %>%
    strat_hist(pval = "over_represented_pvalue", covariate = "numInCat", maxy=10)
```

## Multiple-Testing Correction

Generating the *SummarizedBenchmark* object:

```{r runningBenckmark, message=FALSE}
## rename columns and prepare for benchmarking
res <- dplyr::select(goRes, over_represented_pvalue, numInCat) %>%
    dplyr::rename(pval = over_represented_pvalue,
                  ind_covariate = numInCat) %>%
    dplyr::mutate(pval = pmin(pval, 1))

## generate default BenchDesign
designBen <- initializeBenchDesign()

## run benchmark
sGSEAMouse <- buildBench(designBen, res, ftCols = "ind_covariate")
```

## Benchmark Metrics

```{r}
assayNames(sGSEAMouse) <- "qvalue"
sGSEAMouse <- addDefaultMetrics(sGSEAMouse)
```

```{r plottingResults}
rejections_scatter(sGSEAMouse, as_fraction = FALSE, supplementary = FALSE)
```

```{r}
plotFDRMethodsOverlap(sGSEAMouse, alpha = 0.1, supplementary = FALSE,
                      order.by = "freq", nsets = 100)
```

Gene set size of overlaps.

```{r}
methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
plotCovariateBoxplots(sGSEAMouse, alpha = 0.1, nsets = 4,
                      methods = methods, trans = log2, maxNum = 5)
```

# Session Info

```{r}
sessionInfo()
```

### 2018-02-22 datasets/ChIP-seq/ChIP-seq.Rmd
Fix typo in ChIP-seq case study code

@@
 dir.create(datdir, showWarnings = FALSE)
-dir.create(residr, showWarnings = FALSE)
+dir.create(resdir, showWarnings = FALSE)
 

### 2018-02-22 datasets/scrnaseq/scRNAseq-latest.Rmd
AS: making all paths relative

@@
 ```{r, workspace-setup, results='hide', message=FALSE, warning=FALSE}
-datdir <- "data/scrnaseq/"
-resdir <- "results/scrnaseq/"
+system("mkdir -p DATA")
+system("mkdir -p RESULTS")
+datdir <- "DATA/"
+resdir <- "RESULTS/"
+
 #installation
@@
 suppressMessages(library(knitr))
-library(BiocParallel)
-cores = 6
+suppressMessages(library(BiocParallel))
+core <- as.numeric(Sys.getenv("SLURM_NTASKS"))
 multicoreParam <- MulticoreParam(workers = cores)
-sourceDirectory('benchmark-fdr/datasets/R/')
+
+sourceDirectory("../R/")
 ```
@@
 ```{r echo=FALSE, results='asis'}
-df=data.frame(datasets=c("GSE63818-GPL16791","GSE45719"),organism=c("human","human","mouse","mouse"),technology=c("smart-seq","smart-seq","smart-seq2","smart-seq2"))
+df=data.frame(datasets=c("GSE63818-GPL16791","GSE84465","GSE45719","GSE94383"),organism=c("human","human","mouse","mouse"),technology=c("smart-seq","smart-seq","smart-seq2","smart-seq2"))
 kable(df,caption="scRNAseq Datasets")
@@
 
-We first download the data in our working directory. First, let's download the Human data:GSE63818-GPL16791
+We first download the data in the data directory.`conquer` provides the processed datasets in convenient .rds files. First, let's download the Human datasets.
 
-```{bash, humandata-download, eval=FALSE}
-# all data and results are on the wd: /net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/fdrbenchmark-scrnaseq
+```{bash, humandata-download}
 
-#these look well separated on the tsne plots on the conquer manuscript supplement
-curl -o data/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE63818-GPL16791.rds
-# qc report: 
-curl -o data/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/report-multiqc/GSE63818-GPL16791_multiqc_report.html
-# scater report
-curl -o data/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/report-scater/GSE63818-GPL16791_scater.html
+download.file(url="http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE63818-GPL16791.rds",destfile=paste0(datdir,"GSE63818-GPL16791.rds"))
 
-## human ds2
-curl -o data/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE84465.rds
+download.file(url="http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE84465.rds",destfile=paste0(datdir,"GSE84465.rds"))
+
 ```
@@
 
-```{bash, mousedata-download, eval=FALSE}
-#another well-separated dataset on tsne: GSE60749−GPL13112
-curl -o DATA/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE45719.rds
+```{bash, mousedata-download}
+
+download.file(url="http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE45719.rds",destfile=paste0(datdir,"GSE45719.rds"))
+
+download.file(url="http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE94383.rds",destfile=paste0(datdir,"GSE94383.rds"))
+
 ```
@@
 
-```{r, dataload, eval=FALSE}
+```{r, dataload}
 #load compressed mae
-humandata=readRDS("GSE63818-GPL16791.rds")
+humandata=readRDS(paste0(datdir,"GSE63818-GPL16791.rds"))
 #get gene-level data
@@
 #save data
-saveRDS(g1_data,"h_g1_data_GSE63818-GPL16791.rds")
-saveRDS(g2_data,"h_g2_data_GSE63818-GPL16791.rds")
+saveRDS(g1_data,paste0(datdir,"h_g1_data_GSE63818-GPL16791.rds"))
+saveRDS(g2_data,paste0(datdir,"h_g2_data_GSE63818-GPL16791.rds"))
 #COUNTS
@@
 #save data
-saveRDS(g1_cdata,"h_g1_cdata_GSE63818-GPL16791.rds")
-saveRDS(g2_cdata,"h_g2_cdata_GSE63818-GPL16791.rds")
+saveRDS(g1_cdata,paste0(datdir,"h_g1_cdata_GSE63818-GPL16791.rds"))
+saveRDS(g2_cdata,paste0(datdir,"h_g2_cdata_GSE63818-GPL16791.rds"))
 
 #load compressed mao
-humandata2=readRDS("GSE84465.rds")
+humandata2=readRDS(paste0(datdir,"GSE84465.rds"))
 #get gene-level data
@@
 #save data
-saveRDS(g1_data,"h_g1_data_GSE84465.rds")
-saveRDS(g2_data,"h_g2_data_GSE84465rds")
+saveRDS(g1_data,paste0(datdir,"h_g1_data_GSE84465.rds"))
+saveRDS(g2_data,paste0(datdir,"h_g2_data_GSE84465rds"))
 #COUNTS
@@
 #save data
-saveRDS(g1_cdata,"h_g1_cdata_GSE84465.rds")
-saveRDS(g2_cdata,"h_g2_cdata_GSE84465.rds")
+saveRDS(g1_cdata,paste0(datdir,"h_g1_cdata_GSE84465.rds"))
+saveRDS(g2_cdata,paste0(datdir,"h_g2_cdata_GSE84465.rds"))
 
@@
 
-```{r, nofilter1, eval=FALSE}
+```{r, nofilter1}
 g_data=cbind(g1_data,g2_data)
 g_data_sub=g_data[rowSums(g_data==0)!=dim(g_data)[2],] 
-saveRDS(g_data_sub,"h_g_nofilter_subdata_GSE63818-GPL16791.rds") # removing rows with all zeros reduces the number of genes to 33169
+saveRDS(g_data_sub,paste0(datdir,"h_g_nofilter_subdata_GSE63818-GPL16791.rds")) # removing rows with all zeros reduces the number of genes to 33169
 
@@
 g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=dim(g_cdata)[2],]
-saveRDS(g_cdata_sub,"h_g_nofilter_subcdata_GSE63818-GPL16791.rds") # removing rows with all zeros reduces the number of genes to 33169
+saveRDS(g_cdata_sub,paste0(datdir,"h_g_nofilter_subcdata_GSE63818-GPL16791.rds"))# removing rows with all zeros reduces the number of genes to 33169
 
-saveRDS(g1_ids,"g1_ids_GSE63818-GPL16791.rds")
-saveRDS(g2_ids,"g2_ids_GSE63818-GPL16791.rds")
+saveRDS(g1_ids,paste0(datdir,"g1_ids_GSE63818-GPL16791.rds"))
+saveRDS(g2_ids,paste0(datdir,"g2_ids_GSE63818-GPL16791.rds"))
 ```
@@
 
-```{r, nofilter2, eval=FALSE}
+```{r, nofilter2}
 g_data=cbind(g1_data,g2_data)
 g_data_sub=g_data[rowSums(g_data==0)!=dim(g_data)[2],] 
-saveRDS(g_data_sub,"h_g_nofilter_subdata_GSE84465.rds") # removing rows with all zeros reduces the number of genes to 52942
+saveRDS(g_data_sub,paste0(datdir,"h_g_nofilter_subdata_GSE84465.rds")) # removing rows with all zeros reduces the number of genes to 52942
 
@@
 g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=dim(g_cdata)[2],]
-saveRDS(g_cdata_sub,"h_g_nofilter_subcdata_GSE84465.rds") # removing rows with all zeros reduces the number of genes to 52942
+saveRDS(g_cdata_sub,paste0(datdir,"h_g_nofilter_subcdata_GSE84465.rds")) # removing rows with all zeros reduces the number of genes to 52942
 
-saveRDS(g1_ids,"g1_ids_GSE84465.rds")
-saveRDS(g2_ids,"g2_ids_GSE84465.rds")
+saveRDS(g1_ids,paste0(datdir,"g1_ids_GSE84465.rds"))
+saveRDS(g2_ids,paste0(datdir,"g2_ids_GSE84465.rds"))
 ```
@@
 
-```{r,mouse, eval=FALSE}
+```{r,mouse}
 ##mouse data
-mousedata=readRDS("GSE45719.rds")
+mousedata=readRDS(paste0(datdir,"GSE45719.rds"))
 #get gene-level data
@@
 #save data
-saveRDS(g1_data,"m_g1_data_GSE45719.rds")
-saveRDS(g2_data,"m_g2_data_GSE45719.rds")
+saveRDS(g1_data,paste0(datdir,"m_g1_data_GSE45719.rds"))
+saveRDS(g2_data,paste0(datdir,"m_g2_data_GSE45719.rds"))
 #COUNTS
@@
 #save data
-saveRDS(g1_cdata,"m_g1_cdata_GSE45719.rds")
-saveRDS(g2_cdata,"m_g2_cdata_GSE45719.rds")
+saveRDS(g1_cdata,paste0(datdir,"m_g1_cdata_GSE45719.rds"))
+saveRDS(g2_cdata,paste0(datdir,"m_g2_cdata_GSE45719.rds"))
 
-mousedata2=readRDS("GSE94383.rds")
+mousedata2=readRDS(paste0(datdir,"GSE94383.rds"))
 #get gene-level data
@@
 #save data
-saveRDS(g1_data,"m_g1_GSE94383.rds")
-saveRDS(g2_data,"m_g2_GSE94383.rds")
+saveRDS(g1_data,paste0(datdir,"m_g1_GSE94383.rds"))
+saveRDS(g2_data,paste0(datdir,"m_g2_GSE94383.rds"))
 #COUNTS
@@
 #save data
-saveRDS(g1_cdata,"m_g1_c_GSE94383.rds")
-saveRDS(g2_cdata,"m_g2_c_GSE94383.rds")
+saveRDS(g1_cdata,paste0(datdir,"m_g1_c_GSE94383.rds"))
+saveRDS(g2_cdata,paste0(datdir,"m_g2_c_GSE94383.rds"))
 
@@
 
-```{r, m_nofilter, eval=FALSE}
+```{r, m_nofilter}
 g_data=cbind(g1_data,g2_data)
 g_data_sub=g_data[rowSums(g_data==0)!=dim(g_data)[2],] # 0.25*52=13
-saveRDS(g_data_sub,"m_g_nofilter_subdata_GSE45719.rds") # removing rows with all zeros reduces the number of genes to 33169
+saveRDS(g_data_sub,paste0(datdir,"m_g_nofilter_subdata_GSE45719.rds")) # removing rows with all zeros reduces the number of genes to 33169
 
@@
 g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=dim(g_data)[2],]
-saveRDS(g_cdata_sub,"m_g_nofilter_subcdata_GSE45719.rds") # removing rows with all zeros reduces the number of genes to 33169
+saveRDS(g_cdata_sub,paste0(datdir,"m_g_nofilter_subcdata_GSE45719.rds")) # removing rows with all zeros reduces the number of genes to 33169
 
-saveRDS(g1_ids,"m_g1_ids_GSE45719.rds")
-saveRDS(g2_ids,"m_g2_ids_GSE45719.rds")
+saveRDS(g1_ids,paste0(datdir,"m_g1_ids_GSE45719.rds"))
+saveRDS(g2_ids,paste0(datdir,"m_g2_ids_GSE45719.rds"))
 ```
@@
 
-```{r, m_nofilter_ds2, eval=FALSE}
+```{r, m_nofilter_ds2}
 g_data=cbind(g1_data,g2_data)
 g_data_sub=g_data[rowSums(g_data==0)!=dim(g_data)[2],] # 0.25*52=13
-saveRDS(g_data_sub,"m_g_nofilter_subdata_GSE94383.rds") # removing rows with all zeros reduces the number of genes to 25179
+saveRDS(g_data_sub,paste0(datdir,"m_g_nofilter_subdata_GSE94383.rds")) # removing rows with all zeros reduces the number of genes to 25179
 
@@
 g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=dim(g_data)[2],]
-saveRDS(g_cdata_sub,"m_g_nofilter_subcdata_GSE94383.rds") # removing rows with all zeros reduces the number of genes to 25179
+saveRDS(g_cdata_sub,paste0(datdir,"m_g_nofilter_subcdata_GSE94383.rds")) # removing rows with all zeros reduces the number of genes to 25179
 
-saveRDS(g1_ids,"m_g1_ids_GSE94383.rds")
-saveRDS(g2_ids,"m_g2_ids_GSE94383.rds")
+saveRDS(g1_ids,paste0(datdir,"m_g1_ids_GSE94383.rds"))
+saveRDS(g2_ids,paste0(datdir,"m_g2_ids_GSE94383.rds"))
 ```
@@
 We perform differential expression testing using 3 methods on the unfiltered data: non-parametric Wilcoxon, MAST which has a hurdle model to account for zeros and scDD which implicitly models any heterogeneity in expression.
-We will first assign names for all files
+
+We will first assign names for output files of the DE methods as they will be handy downstream.
 
 ```{r, filenames}
-h_wilcox_dat1="h_data_wilcox_p_GSE63818-GPL16791.rds"
-h_wilcox_dat2="h_data_wilcox_p_GSE84465.rds"
-m_wilcox_dat1="m_data_wilcox_p_GSE45719.rds"
-m_wilcox_dat2="m_data_wilcox_p_GSE94383.rds"
+h_wilcox_dat1=paste0(resdir,"h_data_wilcox_p_GSE63818-GPL16791.rds")
+h_wilcox_dat2=paste0(resdir,"h_data_wilcox_p_GSE84465.rds")
+m_wilcox_dat1=paste0(resdir,"m_data_wilcox_p_GSE45719.rds")
+m_wilcox_dat2=paste0(resdir,"m_data_wilcox_p_GSE94383.rds")
 
-h_scdd_dat1="humanplate_scDatEx_ks_GSE63818-GPL16791.rds"
-h_scdd_dat2="humanplate_scDatEx_ks_GSE84465.rds"
-m_scdd_dat1="mouseplate_scDatEx_ks_GSE45719.rds"
-m_scdd_dat2="mouseplate_scDatEx_ks_GSE94383.rds"
+h_scdd_dat1=paste0(resdir,"humanplate_scDatEx_ks_GSE63818-GPL16791.rds")
+h_scdd_dat2=paste0(resdir,"humanplate_scDatEx_ks_GSE84465.rds")
+m_scdd_dat1=paste0(resdir,"mouseplate_scDatEx_ks_GSE45719.rds")
+m_scdd_dat2=paste0(resdir,"mouseplate_scDatEx_ks_GSE94383.rds")
 
-m_mast_dat1="mouseplate_masttpm_GSE45719.rds"
-h_mast_dat1="humanplate_masttpm_GSE63818-GPL16791.rds"
-h_mast_dat2="humanplate_masttpm_GSE84465.rds"
-m_mast_dat2="mouseplate_masttpm_GSE94383.rds"
+m_mast_dat1=paste0(resdir,"mouseplate_masttpm_GSE45719.rds")
+h_mast_dat1=paste0(resdir,"humanplate_masttpm_GSE63818-GPL16791.rds")
+h_mast_dat2=paste0(resdir,"humanplate_masttpm_GSE84465.rds")
+m_mast_dat2=paste0(resdir,"mouseplate_masttpm_GSE94383.rds")
 
@@
 
-```{r, wilcoxon, eval=FALSE}
+```{r, wilcoxon}
 
-g_data_sub=readRDS("h_g_nofilter_subdata_GSE63818-GPL16791.rds")
+g_data_sub=readRDS(paste0(datdir,"h_g_nofilter_subdata_GSE63818-GPL16791.rds"))
 groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
@@
 
-g_data_sub=readRDS("h_g_nofilter_subdata_GSE84465.rds")
+g_data_sub=readRDS(paste0(datdir,"h_g_nofilter_subdata_GSE84465.rds"))
 groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
@@
 
-g_data_sub=readRDS("m_g_nofilter_subdata_GSE45719.rds")
+g_data_sub=readRDS(paste0(datdir,"m_g_nofilter_subdata_GSE45719.rds"))
 groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
@@
 
-g_data_sub=readRDS("m_g_nofilter_subdata_GSE94383.rds")
+g_data_sub=readRDS(paste0(datdir,"m_g_nofilter_subdata_GSE94383.rds"))
 groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
@@
 
-```{r, scdd, eval=FALSE}
+```{r, scdd}
 
 ## human data
-g_cdata_sub=readRDS("h_g_nofilter_subcdata_GSE63818-GPL16791.rds")
+g_cdata_sub=readRDS(paste0(datdir,"h_g_nofilter_subcdata_GSE63818-GPL16791.rds"))
 compute_scdd_sc(g_cdata_sub,"h_sizefactors_GSE63818-GPL16791.rds","h_data_sce_GSE63818-GPL16791.rds",h_scdd_dat1)
@@
 # perm=1000
-g_cdata_sub=readRDS("h_g_nofilter_subcdata_GSE84465.rds")
+g_cdata_sub=readRDS(paste0(datdir,"h_g_nofilter_subcdata_GSE84465.rds"))
 compute_scdd_sc(g_cdata_sub,"h_sizefactors_GSE84465.rds","h_data_sce_GSE84465.rds",h_scdd_dat2)
@@
 ## mouse data
-g_cdata_sub=readRDS("m_g_nofilter_subcdata_GSE45719.rds")
+g_cdata_sub=readRDS(paste0(datdir,"m_g_nofilter_subcdata_GSE45719.rds"))
 compute_scdd_sc(g_cdata_sub,"m_sizefactors_GSE45719.rds","m_data_sce_GSE45719.rds",m_scdd_dat1)
@@
 # Classifying significant genes into patterns
-g_cdata_sub=readRDS("m_g_nofilter_subcdata_GSE94383.rds")
+g_cdata_sub=readRDS(paste0(datdir,"m_g_nofilter_subcdata_GSE94383.rds"))
 compute_scdd_sc(g_cdata_sub,"m_sizefactors_GSE94383.rds","m_data_sce_GSE94383.rds",m_scdd_dat2)
@@
 ```{r, MAST,eval=FALSE}
-g_data_sub=readRDS("m_g_nofilter_subdata_GSE45719.rds")
+g_data_sub=readRDS(paste0(datdir,"m_g_nofilter_subdata_GSE45719.rds"))
 groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
@@
 
-g_data_sub=readRDS("h_g_nofilter_subdata_GSE63818-GPL16791.rds")
+g_data_sub=readRDS(paste0(datdir,"h_g_nofilter_subdata_GSE63818-GPL16791.rds"))
 groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
@@
 ```{r,cov}
-h_cov=compute_covariates_scrnaseq("h_g_nofilter_subdata.rds")
-m_cov=compute_covariates_scrnaseq("m_g_nofilter_subdata.rds")
-h2_cov=compute_covariates_scrnaseq("h_g_nofilter_subdata_GSE84465.rds")
-m2_cov=compute_covariates_scrnaseq("m_g_nofilter_subdata_GSE94383.rds")
+h_cov=compute_covariates_scrnaseq(paste0(datdir,"h_g_nofilter_subdata_GSE63818-GPL16791.rds"))
+m_cov=compute_covariates_scrnaseq(paste0(datdir,"m_g_nofilter_subdata_GSE45719.rds"))
+h2_cov=compute_covariates_scrnaseq(paste0(datdir,"h_g_nofilter_subdata_GSE84465.rds"))
+m2_cov=compute_covariates_scrnaseq(paste0(datdir,"m_g_nofilter_subdata_GSE94383.rds"))
 ```

### 2018-02-22 datasets/scrnaseq/scRNAseq-latest.Rmd
Try to make reproducible

--- 2018-02-21 datasets/scrnaseq/scRNAseq-latest.Rmd
@@
 resdir <- "results/scrnaseq/"
-#installation
-# source("https://bioconductor.org/biocLite.R")
-# biocLite("swfdr")
-# biocLite("IHW")
-# biocLite("qvalue")
-# install.packages("ashr")
-# install.packages("fdrtool")
-# biocLite("edgeR")
-# source("https://bioconductor.org/biocLite.R")
-# biocLite("scran")
-#  Samseq installation
-# source("https://bioconductor.org/biocLite.R")
-# biocLite("impute") 
-# install.packages("samr")
-# source("https://bioconductor.org/biocLite.R")
-#biocLite("MultiAssayExperiment")
-#source("https://bioconductor.org/biocLite.R")
-#biocLite("SummarizedExperiment")
-#biocLite("BiocGenerics")
-#biocLite("S4Vectors")
-#biocLite("BiocParallel")
-#install_github("areyesq89/SummarizedBenchmark")
-#biocLite("EBSeq")
-#biocLite("scDD")
-#source("https://bioconductor.org/biocLite.R")
-#biocLite("SingleCellExperiment")
-#source("https://bioconductor.org/biocLite.R")
-#biocLite("MAST")
-#install.packages("FDRreg")
+system(paste0("mkdir -p ", datdir))
+system(paste0("mkdir -p ", resdir))
 
 # packages for data analyses
-suppressMessages(library(MultiAssayExperiment))
-suppressMessages(library(dplyr))
-suppressMessages(library(ggplot2))
-suppressMessages(library(cowplot))
-suppressMessages(library(UpSetR))
-suppressMessages(library(hexbin))
+library(MultiAssayExperiment)
+library(dplyr)
+library(ggplot2)
+library(cowplot)
+library(UpSetR)
+library(hexbin)
 
 # packages for differential expression testing
-suppressMessages(library(impute))
-#suppressMessages(library(samr)
-suppressMessages(library(scDD))
-suppressMessages(library(EBSeq))
-suppressMessages(library(edgeR))
-suppressMessages(library(MAST))
-
-# benchmark methods
-suppressMessages(library(IHW))
-suppressMessages(library(ashr))
-suppressMessages(library(qvalue))
-suppressMessages(library(swfdr))
-suppressMessages(library(fdrtool))
-suppressMessages(library(FDRreg))
+library(impute)
+library(scDD)
+library(EBSeq)
+library(edgeR)
+library(MAST)
 
 # summarizebenchmark
-suppressMessages(library(SummarizedBenchmark))
-suppressMessages(library(SingleCellExperiment))
+library(SummarizedBenchmark)
+library(SingleCellExperiment)
 
 # useful libraries
-suppressMessages(library(R.utils))
-suppressMessages(library(knitr))
+library(R.utils)
+library(knitr)
 library(BiocParallel)
@@
 multicoreParam <- MulticoreParam(workers = cores)
-sourceDirectory('benchmark-fdr/datasets/R/')
+sourceDirectory('../R/')
 ```
@@
 
-```{bash, humandata-download, eval=FALSE}
+```{r, humandata-download}
 # all data and results are on the wd: /net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/fdrbenchmark-scrnaseq
@@
 #these look well separated on the tsne plots on the conquer manuscript supplement
-curl -o data/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE63818-GPL16791.rds
-# qc report: 
-curl -o data/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/report-multiqc/GSE63818-GPL16791_multiqc_report.html
-# scater report
-curl -o data/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/report-scater/GSE63818-GPL16791_scater.html
+download.file("http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE63818-GPL16791.rds",
+  destfile=paste0(datdir, "GSE63818-GPL16791.rds"))
 
 ## human ds2
-curl -o data/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE84465.rds
+download.file("http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE84465.rds",
+  destfile=paste0(datdir, "GSE84465.rds"))
 ```
@@
 
-```{bash, mousedata-download, eval=FALSE}
+```{r, mousedata-download}
 #another well-separated dataset on tsne: GSE60749−GPL13112
-curl -o DATA/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE45719.rds
+download.file("http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE45719.rds",
+  destfile=paste0(datdir, "GSE45719.rds"))
 ```
@@
 
-```{r, dataload, eval=FALSE}
+```{r, dataload}
 #load compressed mae
-humandata=readRDS("GSE63818-GPL16791.rds")
+humandata=readRDS(paste0(datdir, "GSE63818-GPL16791.rds"))
 #get gene-level data
@@
 #save data
-saveRDS(g1_cdata,"h_g1_cdata_GSE63818-GPL16791.rds")
-saveRDS(g2_cdata,"h_g2_cdata_GSE63818-GPL16791.rds")
-
+saveRDS(g1_cdata,paste0(datdir,"h_g1_cdata_GSE63818-GPL16791.rds"))
+saveRDS(g2_cdata,paste0(datdir,"h_g2_cdata_GSE63818-GPL16791.rds"))
 #load compressed mao
-humandata2=readRDS("GSE84465.rds")
+humandata2=readRDS(paste0(datdir,"GSE84465.rds"))
 #get gene-level data
@@
 #save data
-saveRDS(g1_data,"h_g1_data_GSE84465.rds")
-saveRDS(g2_data,"h_g2_data_GSE84465rds")
+saveRDS(g1_data, paste0(datdir,"h_g1_data_GSE84465.rds"))
+saveRDS(g2_data, paste0(datdir,"h_g2_data_GSE84465rds"))
 #COUNTS
@@
 #save data
-saveRDS(g1_cdata,"h_g1_cdata_GSE84465.rds")
-saveRDS(g2_cdata,"h_g2_cdata_GSE84465.rds")
+saveRDS(g1_cdata, paste0(datdir,"h_g1_cdata_GSE84465.rds"))
+saveRDS(g2_cdata, paste0(datdir,"h_g2_cdata_GSE84465.rds"))
 
@@
 
-```{r, nofilter1, eval=FALSE}
+```{r, nofilter1}
 g_data=cbind(g1_data,g2_data)
 g_data_sub=g_data[rowSums(g_data==0)!=dim(g_data)[2],] 
-saveRDS(g_data_sub,"h_g_nofilter_subdata_GSE63818-GPL16791.rds") # removing rows with all zeros reduces the number of genes to 33169
+saveRDS(g_data_sub, paste0(datdir, "h_g_nofilter_subdata_GSE63818-GPL16791.rds")) # removing rows with all zeros reduces the number of genes to 33169
 
@@
 
-saveRDS(g1_ids,"g1_ids_GSE63818-GPL16791.rds")
-saveRDS(g2_ids,"g2_ids_GSE63818-GPL16791.rds")
+saveRDS(g1_ids, paste0(datdir, "g1_ids_GSE63818-GPL16791.rds"))
+saveRDS(g2_ids, paste0(datdir, "g2_ids_GSE63818-GPL16791.rds"))
 ```
@@
 
-```{r, nofilter2, eval=FALSE}
+```{r, nofilter2}
 g_data=cbind(g1_data,g2_data)
 g_data_sub=g_data[rowSums(g_data==0)!=dim(g_data)[2],] 
-saveRDS(g_data_sub,"h_g_nofilter_subdata_GSE84465.rds") # removing rows with all zeros reduces the number of genes to 52942
+saveRDS(g_data_sub, paste0(datdir, "h_g_nofilter_subdata_GSE84465.rds")) # removing rows with all zeros reduces the number of genes to 52942
 
@@
 g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=dim(g_cdata)[2],]
-saveRDS(g_cdata_sub,"h_g_nofilter_subcdata_GSE84465.rds") # removing rows with all zeros reduces the number of genes to 52942
+saveRDS(g_cdata_sub, paste0(datdir, "h_g_nofilter_subcdata_GSE84465.rds")) # removing rows with all zeros reduces the number of genes to 52942
 
-saveRDS(g1_ids,"g1_ids_GSE84465.rds")
-saveRDS(g2_ids,"g2_ids_GSE84465.rds")
+saveRDS(g1_ids, paste0(datdir, "g1_ids_GSE84465.rds"))
+saveRDS(g2_ids, paste0(datdir, "g2_ids_GSE84465.rds"))
 ```
@@
 
-```{r,mouse, eval=FALSE}
+```{r,mouse}
 ##mouse data
-mousedata=readRDS("GSE45719.rds")
+mousedata=readRDS(paste0(datdir, "GSE45719.rds"))
 #get gene-level data
@@
 #save data
-saveRDS(g1_data,"m_g1_data_GSE45719.rds")
-saveRDS(g2_data,"m_g2_data_GSE45719.rds")
+saveRDS(g1_data,paste0(datdir, "m_g1_data_GSE45719.rds"))
+saveRDS(g2_data,paste0(datdir, "m_g2_data_GSE45719.rds"))
 #COUNTS
@@
 #save data
-saveRDS(g1_cdata,"m_g1_cdata_GSE45719.rds")
-saveRDS(g2_cdata,"m_g2_cdata_GSE45719.rds")
-
-mousedata2=readRDS("GSE94383.rds")
+saveRDS(g1_cdata,paste0(datdir, "m_g1_cdata_GSE45719.rds"))
+saveRDS(g2_cdata,paste0(datdir, "m_g2_cdata_GSE45719.rds"))
+```
+Comment out this section since GSE94383 was never downloaded, so it won't run.
+```{r, eval=FALSE}
+mousedata2=readRDS(paste0(datdir, "GSE94383.rds"))
 #get gene-level data
@@
 #save data
-saveRDS(g1_data,"m_g1_GSE94383.rds")
-saveRDS(g2_data,"m_g2_GSE94383.rds")
+saveRDS(g1_data,paste0(datdir, "m_g1_GSE94383.rds"))
+saveRDS(g2_data,paste0(datdir, "m_g2_GSE94383.rds"))
 #COUNTS
@@
 #save data
-saveRDS(g1_cdata,"m_g1_c_GSE94383.rds")
-saveRDS(g2_cdata,"m_g2_c_GSE94383.rds")
+saveRDS(g1_cdata,paste0(datdir, "m_g1_c_GSE94383.rds"))
+saveRDS(g2_cdata,paste0(datdir, "m_g2_c_GSE94383.rds"))
 
@@
 
-```{r, m_nofilter, eval=FALSE}
+```{r, m_nofilter}
 g_data=cbind(g1_data,g2_data)
 g_data_sub=g_data[rowSums(g_data==0)!=dim(g_data)[2],] # 0.25*52=13
-saveRDS(g_data_sub,"m_g_nofilter_subdata_GSE45719.rds") # removing rows with all zeros reduces the number of genes to 33169
+saveRDS(g_data_sub,paste0(datdir, "m_g_nofilter_subdata_GSE45719.rds")) # removing rows with all zeros reduces the number of genes to 33169
 
@@
 g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=dim(g_data)[2],]
-saveRDS(g_cdata_sub,"m_g_nofilter_subcdata_GSE45719.rds") # removing rows with all zeros reduces the number of genes to 33169
+saveRDS(g_cdata_sub,paste0(datdir, "m_g_nofilter_subcdata_GSE45719.rds")) # removing rows with all zeros reduces the number of genes to 33169
 
-saveRDS(g1_ids,"m_g1_ids_GSE45719.rds")
-saveRDS(g2_ids,"m_g2_ids_GSE45719.rds")
+saveRDS(g1_ids,paste0(datdir, "m_g1_ids_GSE45719.rds"))
+saveRDS(g2_ids,paste0(datdir, "m_g2_ids_GSE45719.rds"))
 ```
@@
 
+comment out since mouse2 is not downloaded
 ```{r, m_nofilter_ds2, eval=FALSE}
@@
 
-g_data_sub=readRDS("m_g_nofilter_subdata_GSE94383.rds")
-groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
-compute_wilcoxde_sc(g_data_sub,m_wilcox_dat2,groups)
+#g_data_sub=readRDS("m_g_nofilter_subdata_GSE94383.rds")
+#groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
+#compute_wilcoxde_sc(g_data_sub,m_wilcox_dat2,groups)
+
 ## warnings:
@@
 plotCovariateBoxplots(scrna_sb, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3)
+```

### 2018-02-22 datasets/GSEA/gsea-human.Rmd
Rename GSEA analysis Rmd

---
title: "Case Study: Gene Set Enrichment Analysis (GSEA)"
author: "Alejandro Reyes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

**Some sort of summary of the analysis.**

# Workspace Setup

```{r}
library(dplyr)
library(ggplot2)
library(scales)
library(DESeq2)
library(EnsDb.Hsapiens.v86)
library(biomaRt)
library(goseq)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "data"
resdir <- "results"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)
```

# Data Preparation

Download *DESeqDataSet* if not present locally.

```{r, message=FALSE}
dsd_data <- file.path(datdir, "dsd.RData")
if (!file.exists(dsd_data)) {
    download.file("https://www.dropbox.com/s/rdmv3tyqt4g37b0/dsd.RData?dl=1",
                  destfile = dsd_data)
}
load(dsd_data)
```

**What are we doing here?**

```{r}
levels(colData(dsdObject)$individual) <- gsub("GTEX-", "", levels(colData(dsdObject)$individual))
```

We only perform differential tesing on protein coding genes, as specified in Ensembl release 86.

```{r}
gnType <-  genes(EnsDb.Hsapiens.v86, column = "gene_biotype")
dsdObject <- dsdObject[rownames(dsdObject) %in% gnType[gnType$gene_biotype == "protein_coding"]$gene_id, ]
```

# Data Analysis

## Differential Testing

Specify model matrix design and run DESeq2 to get differentially expressed genes.

```{r deseq2, message=FALSE}
dsdObject <- DESeq(dsdObject)
res <- results(dsdObject, independentFiltering = FALSE)
```

## Enrichment Analysis

**Add description of what's going on.**

```{r goseq, message=FALSE}
goset_data <- file.path(datdir, "goSets.RData")
if (!file.exists(goset_data)) {
    mart <- useMart("ensembl", "hsapiens_gene_ensembl")
    goSets <- getBM( c("ensembl_gene_id", "go_id"), mart = mart,
                    filters = "ensembl_gene_id", values = rownames(res))
    goSets <- goSets[!nchar( goSets$go_id ) == 0, ]
    goSets <- with(goSets, split(go_id, ensembl_gene_id))
    save(goSets, file = file)
} else {
    load(goset_data)
}

genes <- as.numeric(res$padj < 0.1)
names(genes) <- rownames(res)

### getting median transcript length
txByGene <- transcriptsBy(EnsDb.Hsapiens.v86, "gene")
geneLength <- sapply( width( txByGene ), median )
geneLength <- geneLength[names(genes)]
genes[is.na( genes )] <- 0
```

Performing a gene set enrichment analysis with `goseq`.

```{r}
pwf <- nullp(genes, bias.data = geneLength)
goRes <- goseq(pwf, gene2cat = goSets)
```

## Covariate Diagnostics

Here, we can see that the covariant variable is actually informative.

### Gene Set Size

**Add description of independent covariate.**

```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
rank_scatter( dat=goRes, pval="over_represented_pvalue", 
              covariate="numInCat", bins=50, funx=log2, 
              funfill=log10_trans()) +
    ylim(0, 12) + ggtitle("Over-represented gene sets") +
    xlab( expression(log[10]~"(# of genes)")) +
    ylab( expression(-log[10]~"(p-value)") )

```

```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
## filter numInCat below only used for plotting
## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
#dplyr:::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)
dplyr:::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=10)
```

## Multiple-Testing Correction

Generating the *SummarizedBenchmark* object:

```{r runningBenckmark, message=FALSE}
## rename columns and prepare for benchmarking
res <- dplyr:::select( goRes, c("over_represented_pvalue", "numInCat") ) %>%
    dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
res$pval <- pmin( res$pval, 1 )

## generate default BenchDesign
designBen <- initializeBenchDesign()

## run benchmark
sGSEA <- buildBench(designBen, res, ftCols="ind_covariate")
```

## Benchmark Metrics

```{r}
assayNames(sGSEA) <- "qvalue"
sGSEA <- addDefaultMetrics(sGSEA)
sGSEA <- estimatePerformanceMetrics(sGSEA, addColData=TRUE)
```

```{r plottingResults}
rejections_scatter(sGSEA, as_fraction=FALSE, supplementary=FALSE)
```

```{r}
plotFDRMethodsOverlap(sGSEA, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100)
```

Gene set size of overlaps.

```{r}
methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
plotCovariateBoxplots(sGSEA, alpha=0.1, nsets=6, methods=methods, trans=log2)
plotCovariateBoxplots(sGSEA, alpha=0.1, nsets=7, methods=methods, trans=log2, maxNum=5)
```

# Data Analysis (Null Comparison)

**Add description.**

## Create Null Data

```{r, eval=FALSE }
nms <- names( genes )
xx <- seq_len(length(nms))
gns <- 1*( xx %in% sample( xx, 1000, replace=FALSE ) )
names( gns ) <- nms
pwf <- nullp( gns, bias.data=geneLength )
goRes <- goseq( pwf, gene2cat=goSets )
res <- goRes %>%
    dplyr:::filter( numDEInCat > 0 ) %>%
    dplyr:::select( c("over_represented_pvalue", "numInCat") ) %>%
    dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
```

## Multiple-Testing Correction

```{r}
pr <- designBen %>% buildBench( res, ftCols="ind_covariate")
assayNames(pr) <- "qvalue"
pr <- addDefaultMetrics(pr, metrics="rejections")
```

## Benchmark Metrics

```{r}
estimatePerformanceMetrics( pr, alpha=0.1, tidy=TRUE ) %>%
    filter( blabel %in% c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf") ) %>%
    select( blabel, value )
```

### 2018-02-22 datasets/RNA-Seq/RNA-Seq.Rmd
Run RNA-Seq case on cluster

@@
 ## set up parallel backend
-multicoreParam <- MulticoreParam(workers = 2)
+cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
+multicoreParam <- MulticoreParam(workers = cores)
 ```
@@
 bd <- initializeBenchDesign()
-sb <- buildBench(bd, data = geneExp, parallel = TRUE)
+sb <- buildBench(bd, data = geneExp)
 ```

### 2018-02-22 datasets/RNA-Seq/RNA-Seq.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+#SBATCH -J knitRNAseq
+#SBATCH -n 8
+#SBATCH -N 1
+#SBATCH -p shared
+#SBATCH --mem 32G
+#SBATCH -t 0-6:00
+#SBATCH -o render-%j.out
+#SBATCH -e render-%j.err
+ 
+##module load rstudio
+export SLURM_NTASKS
+export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/rstudio/0.98.1103-fasrc01/bin/pandoc/"
+
+# change filename to Rmd to be knitted. ncores is passed in through environment
+# variable SLURM_NTASKS
+R -e "rmarkdown::render('RNA-Seq.Rmd')"

### 2018-02-22 datasets/GSEA/gsea.Rmd
Standardize format of gsea

@@
 author: "Alejandro Reyes (on behalf of the Rafalab Benchmarking Consortium)"
-output: html_document
+date: '`r format(Sys.Date(), "%B %d, %Y")`'
+output: 
+    html_document:
+        toc: true
+        toc_float: true
+        highlight: tango
+        number_sections: true
 ---
@@
 
-# Differential expression analysis with DESeq2
+# Summary
 
-Downloading *DESeqDataSet* if not present locally,
+Here we will make use of the `SummarizedBenchmark` package to benchmark two 
+Gene Set Enrichment Analysis (GSEA) case studies.
+
+# Workspace setup
+
+```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
+# data and results directories
+system("mkdir -p DATA")
+system("mkdir -p RESULTS")
+datdir <- "DATA/"
+resdir <- "RESULTS/"
+
+# GSEA analysis tools
+library(DESeq2)
+library(goseq)
+library(EnsDb.Hsapiens.v86)
+library(EnsDb.Mmusculus.v75)
+library(biomaRt)
+
+# wrangling & plotting tools
+library(ggplot2)
+library(cowplot)
+library(scales)
+library(dplyr)
+
+# comparison tools/functions
+library(R.utils)
+sourceDirectory("../R/")
+```
+
+# GTEX expression data
+
+## Differential expression analysis with DESeq2
+
+First we'll download the dataset from GTEX, which is saved as an `RData` file and
+contains a *DESeqDataSet* object.
 
@@
 
-file <- "dsd.RData"
+file <- paste0(datdir, "dsd.RData")
 if(!file.exists( file ) ){
-    download.file("https://www.dropbox.com/s/rdmv3tyqt4g37b0/dsd.RData?dl=1",destfile="dsd.RData")
+    download.file("https://www.dropbox.com/s/rdmv3tyqt4g37b0/dsd.RData?dl=1",
+                  destfile=file)
 }
@@
 
-library(DESeq2)
 levels(colData(dsdObject)$individual) <- gsub("GTEX-", "",levels(colData(dsdObject)$individual))
@@
 
-Specify model matrix design and run DESeq2 to get differentially expressed genes.
+Specify model matrix design and run **DESeq2** to get differentially expressed genes.
 
 ```{r deseq2, message=FALSE}
-
-library(EnsDb.Hsapiens.v86)
 gnType <-  genes(EnsDb.Hsapiens.v86, column="gene_biotype")
@@
 res <- results( dsdObject, independentFiltering=FALSE )
-
 ```
 
-Performing a gene set enrichment analysis with goseq.
+## Gene set enrichment analysis with goseq
 
-```{r goseq, message=FALSE}
+Next we'll use the 
+**biomaRt** package to download GO gene set annotations (for humans since the
+GTEX data contains human samples).
 
-library(goseq)
-
-file="goSets.RData"
+```{r biomart, message=FALSE}
+file <- paste0(datdir, "goSets.RData")
 if(!file.exists(file)){
-  stopifnot( require( biomaRt ) )
   mart <- useMart( "ensembl", "hsapiens_gene_ensembl" )
-  goSets <- getBM( c("ensembl_gene_id", "go_id"), mart=mart, filters="ensembl_gene_id", values=rownames(res))
+  goSets <- getBM( c("ensembl_gene_id", "go_id"), mart=mart, 
+                   filters="ensembl_gene_id", values=rownames(res))
   goSets <- goSets[!nchar( goSets$go_id ) == 0,]
@@
 }
+```
 
+Finally, we'll perform a gene set enrichment analysis with **goseq**. We'll 
+include all genes with a **DESeq2** adjusted p-value less than 0.10, and adjust
+for gene length since not doing so may bias the results toward GO sets with 
+longer genes.
+
+```{r goseq, message=FALSE}
 genes <- as.numeric( res$padj < 0.1 )
@@
 goRes <- goseq( pwf, gene2cat=goSets )
-
 ```
 
+## Check Covariate Diagnostics
+
 ```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
-
-library(ggplot2)
-library(magrittr)
-library(cowplot)
-library(R.utils)
-library(scales)
-sourceDirectory("../R/")
-
 rank_scatter( dat=goRes, pval="over_represented_pvalue", 
@@
 ```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
-
-library(dplyr)
-
 ## filter numInCat below only used for plotting
@@
 ```
+
+## Run benchmarks on set size covariate
 
@@
 
-# Gene set size of overlaps
+## Gene set size of overlaps
 
@@
 
-# NULL comparisons
+## NULL comparisons
 
@@
 ### preparing gene length info and goSets ###
-library("EnsDb.Mmusculus.v75")
 txByGene <- transcriptsBy(EnsDb.Mmusculus.v75, "gene")
@@
 
-library(dplyr)
-
 dplyr:::filter( goRes, numDEInCat > 0, numInCat > 20)  %>%

### 2018-02-23 datasets/GSEA/gsea-human.Rmd
Fix typo in save file name

@@
     goSets <- with(goSets, split(go_id, ensembl_gene_id))
-    save(goSets, file = file)
+    save(goSets, file = goset_data)
 } else {

### 2018-02-23 datasets/GSEA/gsea-human.Rmd
Fix typos

@@
 ## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
-#dplyr:::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)
-dplyr:::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
+#dplyr::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)
+dplyr::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
     strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=10)
@@
 res <- goRes %>%
-    dplyr:::filter( numDEInCat > 0 ) %>%
-    dplyr:::select( c("over_represented_pvalue", "numInCat") ) %>%
-    dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
+    dplyr::filter( numDEInCat > 0 ) %>%
+    dplyr::select( c("over_represented_pvalue", "numInCat") ) %>%
+    dplyr::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
 ```
@@
 ```{r}
-pr <- designBen %>% buildBench( res, ftCols="ind_covariate")
+pr <- buildBench(designBen, res, ftCols = "ind_covariate")
 assayNames(pr) <- "qvalue"
-pr <- addDefaultMetrics(pr, metrics="rejections")
+pr <- addDefaultMetrics(pr)
 ```
@@
 estimatePerformanceMetrics( pr, alpha=0.1, tidy=TRUE ) %>%
-    filter( blabel %in% c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf") ) %>%
-    select( blabel, value )
+    dplyr::filter( blabel %in% c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf") ) %>%
+    dplyr::select( blabel, value )
 ```
+
+# Session Info
+
+```{r}
+sessionInfo()
+```

### 2018-02-23 datasets/GSEA/gsea-mouse.Rmd
Fix typos

@@
 dsd_data <- "dsdHSCMPP.RData"
-if(!file.exists(dsdMouse)) {
+if(!file.exists(dsd_data)) {
     download.file("https://www.dropbox.com/s/spo1pifum5k72rn/dseHSCMPP.RData?dl=1",
@@
 ## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
-#dplyr:::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)
+#dplyr::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)
 dplyr::filter(goRes, numDEInCat > 0, numInCat > 10) %>%

### 2018-02-23 datasets/GSEA/gsea.Rmd
Add details to gsea

@@
 
-## Differential expression analysis with DESeq2
-
 First we'll download the dataset from GTEX, which is saved as an `RData` file and
@@
 
-Specify model matrix design and run **DESeq2** to get differentially expressed genes.
+## Differential expression analysis with DESeq2
+
+Check model matrix design and run **DESeq2** to get differentially expressed genes. 
+The design includes a factor for tissue (here we have Cortex and Cerebellum), 
+and adjusts for individual.
 
 ```{r deseq2, message=FALSE}
+colData(dsdObject)
+design(dsdObject)
+
 gnType <-  genes(EnsDb.Hsapiens.v86, column="gene_biotype")
@@
 dsdObject <- DESeq(dsdObject)
-res <- results( dsdObject, independentFiltering=FALSE )
+res <- results( dsdObject, contrast=c("tissue", "Cortex", "Cerebellum"),
+                independentFiltering=FALSE)
+
 ```
@@
 for gene length since not doing so may bias the results toward GO sets with 
-longer genes.
+longer genes. The bias is seen in the plot below, with gene length bin on the
+x-axis and proportion of genes detected as DE on the y-axis.
 
@@
 
+First we'll look at a scatterplot of the relationship between the pvalue and the number 
+of genes in the gene set.
+
 ```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
@@
 
+Next we'll look at the histogram of p-values stratified by quantiles of the 
+covariate (number of genes in the gene set).
+
 ```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
@@
 ## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
-#dplyr:::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)
 
@@
 
-Generating the *SummarizedBenchmark* object:
+Now we'll generate the *SummarizedBenchmark* object and run the benchmark methods.
+Note that we cannot the ash method since we do not have an effect size for this
+test. We also can't run the scott method since we don't have a test statistic.
 
@@
 
+Finally, we'll visualize the results by looking at the number of rejections at
+several alpha levels.
+
 ```{r plottingResults}
@@
 
+In addition, we'll look at the overlap of gene sets found by each method at the
+alpha cutoff of 0.10.
+
 ```{r}
@@
 
-## Gene set size of overlaps
+Finally, we'll examine whether there are any preferences for methods to detect
+gene sets with certain covariate values.
 
@@
 
-## NULL comparisons
+## Null comparison
+
+Here we'll evaluate specificity of the methods by investigating performance in 
+the null setting. Ideally, no significant gene sets should be found when we 
+perform the analysis on a set of random genes instead of a set of DE genes.
+Here we sample 1000 random genes and perform a GSEA.
 
 ```{r, eval=FALSE }
-
+set.seed(4885)
 nms <- names( genes )
@@
 goRes <- goseq( pwf, gene2cat=goSets )
+```
+
+We'll use the same covariate (gene set size) and run the benchmark methods
+on the GSEA results.
+
+```{r}
 res <- goRes %>%
@@
   dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
+
 pr <- designBen %>% buildBench( res, ftCols="ind_covariate")
@@
 
+Add some description of the HSC-MPP dataset.
+
 ```{r}
-
 dsdMouse <- "dsdHSCMPP.RData"
@@
 load(dsdMouse)
+```
+
+## Differential expression analysis with DESeq2
+
+```{r}
 dseHSCMPP <- DESeq( dseHSCMPP)
@@
 table(genes)
+```
 
+## Gene set enrichment analysis with goseq
+
+```{r}
 ### preparing gene length info and goSets ###
@@
 res$pval <- pmin( res$pval, 1 )
+```
+## Check Covariate Diagnostics
+
+```{r, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
+rank_scatter( dat=goRes, pval="over_represented_pvalue", 
+              covariate="numInCat", bins=50, funx=log2, 
+              funfill=log10_trans()) +
+    ylim(0, 12) + ggtitle("Over-represented gene sets") +
+    xlab( expression(log[10]~"(# of genes)")) +
+    ylab( expression(-log[10]~"(p-value)") )
+
+```
+
+```{r, message=FALSE, fig.width=10, fig.height=3.2}
+## filter numInCat below only used for plotting
+## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
+
+dplyr:::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
+  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=10)
+
+```
+
+## Run benchmarks on set size covariate
+
+```{r}
 sGSEAMouse <- designBen %>% buildBench( res, ftCols="ind_covariate")

### 2018-02-23 datasets/scrnaseq/scRNAseq-keegan.Rmd
Toward standardized, knittable scRNAseq analyses

---
title: "scRNAseq"
author: "Ayshwarya Subramanian"
date: "9/15/2017"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE,cache.lazy = FALSE)

```
## R workspace setup

The downloaded plate data are stored in R objects called `MultiAssayExperiment`; this package requires BioConductor release 3.5 and R version 3.4 or greater for installation.

```{r, workspace-setup, results='hide', message=FALSE, warning=FALSE}
system("mkdir -p DATA")
system("mkdir -p RESULTS")
datdir <- "DATA/"
resdir <- "RESULTS/"

# packages for data analyses
library(MultiAssayExperiment)
library(dplyr)
library(ggplot2)
library(cowplot)
library(UpSetR)
library(hexbin)

# packages for differential expression testing
library(impute)
library(scDD)
library(EBSeq)
library(edgeR)
library(MAST)

# summarizebenchmark
library(SummarizedBenchmark)
library(SingleCellExperiment)

# useful libraries
library(R.utils)
library(knitr)
library(BiocParallel)

cores <- 8
BiocParallel::register(MulticoreParam(workers = cores))

sourceDirectory("../R/")
```

## scRNAseq datasets

For the application of differential gene expression, we look at scRNAseq and bulk RNAseq data. Here we discuss the analyses of scRNAseq.

```{r echo=FALSE, results='asis'}
df=data.frame(datasets=c("GSE63818-GPL16791","GSE84465","GSE45719","GSE94383"),organism=c("human","human","mouse","mouse"),technology=c("smart-seq","smart-seq","smart-seq2","smart-seq2"))
kable(df,caption="scRNAseq Datasets")
df1=data.frame(DE_method=c("Wilcoxon-rank sum","ScDD","MAST"),input=c("tpm","count","tpm"),output=c("p-val","p-val","p-val"))
kable(df1,caption="DE methods")

```

We selected two datasets, one each from mouse and human spp, from the `Conquer` (Soneson et al., 2017) scRNAseq data collection. `Conquer` provides uniformly processed gene expression data summarized to both count (aggregated from transcript members), and length-scaled TPMs at the gene level. The authors use `salmon` to quantify raw reads and `tximport` to summarize to gene level. The data comes from Smartseq, a plate based sequencing method which results in deeper sequencing of reads at the cost of numbers of cells.

## Data Download

We first download the data in the data directory.`conquer` provides the processed datasets in convenient .rds files. First, let's download the Human datasets.

```{r, humandata-download}
#these look well separated on the tsne plots on the conquer manuscript supplement
download.file("http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE63818-GPL16791.rds",
  destfile=paste0(datdir, "GSE63818-GPL16791.rds"))

## human ds2
download.file("http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE84465.rds",
  destfile=paste0(datdir, "GSE84465.rds"))

```

Next, we download the Mouse data:GSE45719

```{r, mousedata-download}
#another well-separated dataset on tsne: GSE60749−GPL13112
download.file("http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE45719.rds",
  destfile=paste0(datdir, "GSE45719.rds"))

download.file(url="http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE94383.rds",
  destfile=paste0(datdir,"GSE94383.rds"))
```

## Data processing and filtering

For data processing, we follow practices and recommendations from Soneson et al., 2017. Briefly, we limit ourselves to 2-group comparisons, and  select the maximal size defined as the smallest number of cells in either group as the group-level sample size for the DE analyses. We do not filter genes as filtering may affect the subsequent false discovery control steps.

First, we get the human data for the two groups of interest.
We'll extract two conditions: Primordial Germ Cells and Somatic cells at 
7 weeks gestation.

```{r, dataload}
# load MultiAssayExperiment
humandata <- readRDS(paste0(datdir,"GSE63818-GPL16791.rds"))

# subset by groups of interest
humandata <- humandata[, (colData(humandata)$source_name_ch1 %in% 
                         c("Somatic Cells", "Primordial Germ Cells")) &
                         (colData(humandata)$characteristics_ch1 == 
                         c("developmental stage: 7 week gestation"))]
cdata=colData(humandata)

# subset by gene expression TPM & count assays
humandata <- experiments(humandata)[[1]]
assays(humandata) <- assays(humandata)[c("TPM", "count")]
colData(humandata) <- cdata

# subset by genes not all zero
humandata <- humandata[rowSums(assays(humandata)[["count"]]==0) < ncol(humandata),]

# look at number of cells in each group
table(colData(humandata)$source_name_ch1)

# save SummarizedExperiment 
saveRDS(humandata, paste0(datdir, "h1_GSE63818-GPL16791.rds"))
```

For the human dataset1 there are `r ncol(humandata)` cells total and 
`r nrow(humandata)` genes.

Now humandata2. We'll extract two conditions: Neoplastic and Regular cells, as
defined by their tsne clusters.

```{r}
# load MultiAssayExperiment
humandata2 <- readRDS(paste0(datdir, "GSE84465.rds"))

# subset by groups of interest
humandata2 <- humandata2[, (colData(humandata2)$characteristics_ch1.7 %in% 
                         c("neoplastic: Neoplastic", "neoplastic: Regular")) &
                         (colData(humandata2)$characteristics_ch1.5 %in% 
                         c("tsne cluster: 11",  "tsne cluster: 8"))]
cdata=colData(humandata2)

# subset by gene expression TPM & count assays
humandata2 <- experiments(humandata2)[[1]]
assays(humandata2) <- assays(humandata2)[c("TPM", "count")]
colData(humandata2) <- cdata

# subset by genes not all zero
humandata2 <- humandata2[rowSums(assays(humandata2)[["count"]]==0) < ncol(humandata2),]

# look at number of cells in each group
table(colData(humandata2)$characteristics_ch1.7)

# save SummarizedExperiment 
saveRDS(humandata2, paste0(datdir, "h2_GSE84465.rds"))
```

For the human dataset2, there are a total of `r nrow(humandata2)` cells and `r ncol(humandata2)` genes.

Next, we repeat the processing steps for the mouse datasets. For the first one,
we'll extract two conditions: 16-cell stage blastomere and mid blastocysts cells.

```{r,mouse}
# load MultiAssayExperiment
mousedata=readRDS(paste0(datdir, "GSE45719.rds"))

# subset by groups of interest
mousedata <- mousedata[, colData(mousedata)$source_name_ch1 %in% 
                         c("16-cell stage blastomere", 
                           "Mid blastocyst cell (92-94h post-fertilization)")]
cdata=colData(mousedata)

# subset by gene expression TPM & count assays
mousedata <- experiments(mousedata)[[1]]
assays(mousedata) <- assays(mousedata)[c("TPM", "count")]
colData(mousedata) <- cdata

# subset by genes not all zero
mousedata <- mousedata[rowSums(assays(mousedata)[["count"]]==0) < ncol(mousedata),]

# look at number of cells in each group
table(as.character(colData(mousedata)$source_name_ch1))

# save SummarizedExperiment 
saveRDS(mousedata, paste0(datdir, "m1_GSE45719.rds"))
```

For the mouse dataset1, we have a total of `r ncol(mousedata)` cells and 
`r nrow(mousedata)` genes.

Now mousedata2. We'll extract two conditions: No stimulation versus LPS at 150 min.

```{r}
# load MultiAssayExperiment
mousedata2=readRDS(paste0(datdir,"GSE94383.rds"))

# subset by groups of interest
mousedata2 <- mousedata2[, (colData(mousedata2)$characteristics_ch1.2 ==
                              "condition: No stimulation") |
                           (colData(mousedata2)$characteristics_ch1.2 ==
                              "condition: LPS" &
                            colData(mousedata2)$characteristics_ch1.1 %in% 
                              "time point: 150 min") ]
cdata=colData(mousedata2)

# subset by gene expression TPM & count assays
mousedata2 <- experiments(mousedata2)[[1]]
assays(mousedata2) <- assays(mousedata2)[c("TPM", "count")]
colData(mousedata2) <- cdata

# subset by genes not all zero
mousedata2 <- mousedata2[rowSums(assays(mousedata2)[["count"]]==0) < ncol(mousedata2),]

# look at number of cells in each group
table(as.character(colData(mousedata2)$characteristics_ch1.2))

# save SummarizedExperiment 
saveRDS(mousedata2, paste0(datdir, "m2_GSE94383.rds"))
```

For the mouse dataset2, we have a total of `r ncol(mousedata2)` cells and 
`r nrow(mousedata2)` genes after filtering for non-detected genes.

## Differential expression testing

** Still need to clean up from here on; nothing below is eval'd**

We perform differential expression testing using 3 methods on the unfiltered data: non-parametric Wilcoxon, MAST which has a hurdle model to account for zeros and scDD which implicitly models any heterogeneity in expression.

We will first assign names for output files of the DE methods as they will be handy downstream.

```{r, filenames, eval=FALSE}
h_wilcox_dat1=paste0(resdir,"h_data_wilcox_p_GSE63818-GPL16791.rds")
h_wilcox_dat2=paste0(resdir,"h_data_wilcox_p_GSE84465.rds")
m_wilcox_dat1=paste0(resdir,"m_data_wilcox_p_GSE45719.rds")
m_wilcox_dat2=paste0(resdir,"m_data_wilcox_p_GSE94383.rds")

h_scdd_dat1=paste0(resdir,"humanplate_scDatEx_ks_GSE63818-GPL16791.rds")
h_scdd_dat2=paste0(resdir,"humanplate_scDatEx_ks_GSE84465.rds")
m_scdd_dat1=paste0(resdir,"mouseplate_scDatEx_ks_GSE45719.rds")
m_scdd_dat2=paste0(resdir,"mouseplate_scDatEx_ks_GSE94383.rds")

m_mast_dat1=paste0(resdir,"mouseplate_masttpm_GSE45719.rds")
h_mast_dat1=paste0(resdir,"humanplate_masttpm_GSE63818-GPL16791.rds")
h_mast_dat2=paste0(resdir,"humanplate_masttpm_GSE84465.rds")
m_mast_dat2=paste0(resdir,"mouseplate_masttpm_GSE94383.rds")

```

### Wilcoxon rank-sum test

Per Soneson et al., we run TMM normalization to the TPM units before running Wilcoxon.

```{r, wilcoxon, eval=FALSE}

g_data_sub=readRDS(paste0(datdir,"h_g_nofilter_subdata_GSE63818-GPL16791.rds"))
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
compute_wilcoxde_sc(g_data_sub,h_wilcox_dat1,groups)

g_data_sub=readRDS(paste0(datdir,"h_g_nofilter_subdata_GSE84465.rds"))
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
compute_wilcoxde_sc(g_data_sub,h_wilcox_dat2,groups)

g_data_sub=readRDS(paste0(datdir,"m_g_nofilter_subdata_GSE45719.rds"))
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
compute_wilcoxde_sc(g_data_sub,m_wilcox_dat1,groups)

g_data_sub=readRDS(paste0(datdir,"m_g_nofilter_subdata_GSE94383.rds"))
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
compute_wilcoxde_sc(g_data_sub,m_wilcox_dat2,groups)

```

### Differences in distributions of gene expression in the 2 conditions

We apply scDD to log-transformed counts as below. Th helper functions set up the Single-Cell experiment object.

```{r, scdd, eval=FALSE}

## human data
g_cdata_sub=readRDS(paste0(datdir,"h_g_nofilter_subcdata_GSE63818-GPL16791.rds"))
compute_scdd_sc(g_cdata_sub, h_scdd_dat1)

g_cdata_sub=readRDS(paste0(datdir,"h_g_nofilter_subcdata_GSE84465.rds"))
compute_scdd_sc(g_cdata_sub, h_scdd_dat2)

## mouse data
g_cdata_sub=readRDS(paste0(datdir,"m_g_nofilter_subcdata_GSE45719.rds"))
compute_scdd_sc(g_cdata_sub, m_scdd_dat1)

g_cdata_sub=readRDS(paste0(datdir,"m_g_nofilter_subcdata_GSE94383.rds"))
compute_scdd_sc(g_cdata_sub, m_scdd_dat2)

```

### MAST

We run MAST on the TPMs following the approach in Soneson et al.

```{r, MAST,eval=FALSE, eval=FALSE}
g_data_sub=readRDS(paste0(datdir,"m_g_nofilter_subdata_GSE45719.rds"))
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
names(groups)=colnames(g_cdata_sub)
compute_MASTtpmde(g_data_sub,groups,)

g_data_sub=readRDS(paste0(datdir,"h_g_nofilter_subdata_GSE63818-GPL16791.rds"))
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
names(groups)=colnames(g_data_sub)
compute_MASTtpmde(g_data_sub,groups,h_mast_dat1)

g_data_sub=readRDS(paste0(datdir,"h_g_nofilter_subdata_GSE84465.rds"))
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
names(groups)=colnames(g_data_sub)
compute_MASTtpmde(g_data_sub,groups,h_mast_dat2)

g_data_sub=readRDS(paste0(datdir,"m_g_nofilter_subdata_GSE94383.rds"))
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
names(groups)=colnames(g_data_sub)
compute_MASTtpmde(g_data_sub,groups,m_mast_dat2)
```

## Covariate Diagnostics

In scrnaseq data, strength of the signal can be a function of level of gene expression. We quantify it as median or mean gene expression,and detection rate, defined as the percentage of cells expressing the gene at a level of TPM>0.

```{r,cov, eval=FALSE}
h_cov=compute_covariates_scrnaseq(paste0(datdir,"h_g_nofilter_subdata_GSE63818-GPL16791.rds"))
m_cov=compute_covariates_scrnaseq(paste0(datdir,"m_g_nofilter_subdata_GSE45719.rds"))
h2_cov=compute_covariates_scrnaseq(paste0(datdir,"h_g_nofilter_subdata_GSE84465.rds"))
m2_cov=compute_covariates_scrnaseq(paste0(datdir,"m_g_nofilter_subdata_GSE94383.rds"))
```

First we will look at the covariates for the Wilcoxon text

```{r, wilcoxcov, fig.width=10, fig.height=7, eval=FALSE}
print("human_plate_1")
hwilcox1 <- plot_covariate_diagnostics(paste0(h_wilcox_dat1),"wilcox",h_cov)
print(hwilcox1$plot)

print("human_plate_2")
hwilcox2 <- plot_covariate_diagnostics(paste0(h_wilcox_dat2),"wilcox",h2_cov)
print(hwilcox2$plot)

print("mouse_plate_1")
mwilcox1 <- plot_covariate_diagnostics(paste0(m_wilcox_dat1),"wilcox",m_cov)
print(mwilcox1$plot)

print("mouse_plate_2")
mwilcox2 <- plot_covariate_diagnostics(paste0(m_wilcox_dat2),"wilcox",m2_cov)
print(mwilcox2$plot)
```

Next, we look at the covariates for MAST

```{r, mastcov, fig.width=10, fig.height=7, eval=FALSE}
print("human_plate_1")
hmast1 <- plot_covariate_diagnostics(paste0(h_mast_dat1),"mast",h_cov)
print(hmast1$plot)

print("human_plate_2")
hmast2 <- plot_covariate_diagnostics(paste0(h_mast_dat2),"mast",h2_cov)
print(hmast2$plot)

print("mouse_plate_1")
mmast1 <- plot_covariate_diagnostics(paste0(m_mast_dat1),"mast",m_cov)
print(mmast1$plot)

print("mouse_plate_2")
mmast2 <- plot_covariate_diagnostics(paste0(m_mast_dat2),"mast",m2_cov)
print(mmast2$plot)
```

Finally, we look at how informative the covariates are with results from ScDD.

```{r,scddcov, fig.width=10, fig.height=7, eval=FALSE}
print("human_plate_1")
hscdd1 <- plot_covariate_diagnostics(paste0(h_scdd_dat1),"scdd",h_cov)
print(hscdd1$plot)

print("human_plate_2")
hscdd2 <- plot_covariate_diagnostics(paste0(h_scdd_dat2),"scdd",h2_cov)
print(hscdd2$plot)

print("mouse_plate_1")
mscdd1 <- plot_covariate_diagnostics(paste0(m_scdd_dat1),"scdd",m_cov)
print(mscdd1$plot)

print("mouse_plate_2")
mscdd2 <- plot_covariate_diagnostics(paste0(m_scdd_dat2),"scdd",m2_cov)
print(mscdd2$plot)
```

### Benchmarking FDR methods

First we will create the benchdesign object for wilcox -- IN PROGRESS

```{r, benchmarking, eval=FALSE, eval=FALSE}

bd <- initializeBenchDesign()
resN <- paste0(resdir, "scrnaseq_wilcox_summarizedBenchmark_",nrow(m_scdd_cov), ".RData")

duration <- NA
if (!file.exists(resN)){
  t1 <- proc.time()
  scrna_sb <- bd %>% buildBench(data=(hscdd1$data%>%select(pval,mean_exp)%>%mutate(effect_size=0, test_statistic=0, SE=0) %>% mutate(ind_covariate=mean_exp)), 
                          ftCols = c("mean_exp"),
                          parallel=TRUE, BPPARAM=multicoreParam)

  save(scrna_sb, file=resN)
  duration <- round((proc.time()-t1)[3]/60,1)
}else{
  load(resN)
}

assayNames(scrna_sb) <- "qvalue"
scrna_sb <- addDefaultMetrics(scrna_sb)

# plot nrejects by method overall and stratified by covariate
rejections_scatter(scrna_sb,supplementary=FALSE)

rejection_scatter_bins(scrna_sb, covariate="ind_covar_N", bins=3,supplementary=FALSE)

# upset plot 
plotFDRMethodsOverlap(scrna_sb, 
                      alpha=0.05, nsets=ncol(scrna_sb),
                      order.by="freq", decreasing=TRUE,
                      supplementary=FALSE)

# properties of method-exclusive discoveries
methods <- c("lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
mcols(scrna_sb)$ind_covariate <- mcols(scrna_sb)$ind_covar_N
plotCovariateBoxplots(scrna_sb, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3)
```

### 2018-02-23 datasets/GSEA/gsea.Rmd
Added source and description of mouse data

@@
 assayNames(pr) <- "qvalue"
-pr <- addDefaultMetrics(pr, metrics="rejections")
+pr <- addDefaultMetrics(pr)
 
@@
 
-# Same analysis but using the HSC-MPP dataset from Nina
+# GSEA of genes differentially expressed upon blood cell differentiation in mouse
 
-Add some description of the HSC-MPP dataset.
+As a second dataset, we will use the mouse data from the paper by [Cabezas-Wallscheid et al.
+(Cell stem Cell, 2014)](https://www.sciencedirect.com/science/article/pii/S1934590914003014?via%3Dihub).
+The dataset consists of RNA-seq data from mouse hematopoietic stem cells and 4 multipotent progenitor lineages.
 
@@
 ## Differential expression analysis with DESeq2
+
+For this analysis, we will test each gene for differences in expression between hematopoietic stem cells and multipotent progenitors (fraction 1).
 

### 2018-02-26 datasets/ChIP-seq/ChIP-seq.Rmd
Make plot sizes and types uniform

@@
 
-```{r, chipseq-diag, fig.width=10, fig.height=3.5}
+```{r, chipseq-diag-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(dat, pvalue = "pval", covariate = "ind_covariate")
+```
+
+```{r, chipseq-diag-hist, fig.width=10, fig.height=3.2}
 strat_hist(dat, pvalue = "pval", covariate = "ind_covariate", maxy = 2)
-rank_scatter(dat, pvalue = "pval", covariate = "ind_covariate")
 ```
@@
 rejections_scatter(sb, supplementary = FALSE)
-
 rejection_scatter_bins(sb, covariate = "ind_covariate",

### 2018-02-26 datasets/GSEA/gsea-mouse.Rmd
Make plot sizes and types uniform

@@
 ```{r, message=FALSE}
-dsd_data <- "dsdHSCMPP.RData"
+dsd_data <- file.path(datdir, "dsdHSCMPP.RData")
 if(!file.exists(dsd_data)) {
@@
 ## run benchmark
+sb_file <- file.path(resdir, "sb-gsea-mouse.rda")
+if (!file.exists(sb_file)) {
     sGSEAMouse <- buildBench(designBen, res, ftCols = "ind_covariate")
+} else {
+    sGSEAMouse <- readRDS(sb_file)
+}
 ```

### 2018-02-26 datasets/microbiome/microbiome-baxter.Rmd
Make plot sizes and types uniform

@@
 otu <- otu[rowSums(otu) > min_otu, ]
-out <- otu[rowSums(otu > 0) / nrow(otu) > minp_otu, ]
+otu <- otu[rowSums(otu > 0) / ncol(otu) > minp_otu, ]
 otu <- otu[, colSums(otu) > min_sample]
@@
 Note that the effect here is calculated as logfold change of mean abundance in controls
-relative to cases (i.e. `log(mean_abun[controls]/mean_abun[cases])`)
+relative to cases (i.e. `log(mean_abun[controls]/mean_abun[cases])`).
 
@@
 ```{r, baxter-pvals}
-resfile <- file.path(resdir, paste0("baxter_results_", nrow(abun_otu), "_OTUs.RData"))
+resfile <- file.path(resdir, paste0("baxter_results_", nrow(abun_otu), "_OTUs.rda"))
 if (!file.exists(resfile)) {
@@
 
-```{r, baxter-ubi, fig.width=10, fig.height=3.5}
+```{r, baxter-ubi-hist, fig.width=10, fig.height=3.2}
 strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
+```
+
+```{r, baxter-ubi-scatter, fig.width=4.5, fig.height=3.5}
 rank_scatter(res, pvalue="pval", covariate="ubiquity")
@@
 
-```{r, baxter-abun, fig.width=10, fig.height=3.5}
+```{r, baxter-abun-hist, fig.width=10, fig.height=3.2}
 strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
+```
+
+```{r, baxter-abun-scatter, fig.width=4.5, fig.height=3.5}
 rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
@@
 rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
-#plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+##plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
 ```
@@
 ```{r, baxter-genus-pvals}
-resfile <- file.path(resdir, paste0("baxter_results_", nrow(genus_df), "_genera.RData"))
+resfile <- file.path(resdir, paste0("baxter_results_", nrow(genus_df), "_genera.rda"))
 if (!file.exists(resfile)) {
@@
 
-```{r, baxter-genus-ubi, fig.width=10, fig.height=3.5}
+```{r, baxter-genus-ubi, fig.width=10, fig.height=3.2}
 strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
+```
+
+```{r, baxter-genus-ubi-scatter, fig.width=4.5, fig.height=3.5}
 rank_scatter(res, pvalue="pval", covariate="ubiquity")
@@
 
-```{r, baxter-genus-abun, fig.width=10, fig.height=3.5}
+```{r, baxter-genus-abun, fig.width=10, fig.height=3.2}
 strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
+```
+
+```{r, baxter-genus-abun-scatter, fig.width=4.5, fig.height=3.5}
 rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
@@
 ```{r, baxter-genus-labelcovariate}
-res <- res %>% mutate(ind_covariate = ubiquity)
+res <- dplyr::mutate(res, ind_covariate = ubiquity)
 ```
@@
 for (nmids in c(5, 50, 150)) {
-    print(nmids)
+    ## use common BenchDesign w/out ASH
     bd <- initializeBenchDesign(nmids=nmids)
-    ## remove ASH from comparison
     bd <- dropBMethod(bd, "ashs")

### 2018-02-26 datasets/microbiome/microbiome-goodrich.Rmd
Make plot sizes and types uniform

@@
 
-```{r, goodrich-ubi, fig.width=10, fig.height=3.5}
+```{r, goodrich-ubi-hist, fig.width=10, fig.height=3.2}
 strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=5, binwidth=0.05, numQ=4)
+```
+
+```{r, goodrich-ubi-scatter, fig.width=4.5, fig.height=3.5}
 rank_scatter(res, pvalue="pval", covariate="ubiquity")
@@
 
-```{r, goodrich-abun, fig.width=10, fig.height=3.5}
+```{r, goodrich-abun-hist, fig.width=10, fig.height=3.2}
 strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=5, binwidth=0.05, numQ=4)
+```
+
+```{r, goodrich-abun-scatter, fig.width=4.5, fig.height=3.5}
 rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
@@
 
-```{r, goodrich-genus-ubi, fig.width=10, fig.height=3.5}
+```{r, goodrich-genus-ubi, fig.width=10, fig.height=3.2}
 strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
+```
+
+```{r, goodrich-genus-ubi-scatter, fig.width=4.5, fig.height=3.5}
 rank_scatter(res, pvalue="pval", covariate="ubiquity")
@@
 
-```{r, goodrich-genus-abun, fig.width=10, fig.height=3.5}
+```{r, goodrich-genus-abun, fig.width=10, fig.height=3.2}
 strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
+```
+
+```{r, goodrich-genus-abun-scatter, fig.width=4.5, fig.height=3.5}
 rank_scatter(res, pvalue="pval", covariate="mean_abun_present")

### 2018-02-26 datasets/microbiome/microbiome-schubert.Rmd
Make plot sizes and types uniform

@@
 
-```{r, schubert-ubi, fig.width=10, fig.height=3.5, eval=TRUE}
+```{r, schubert-ubi, fig.width=10, fig.height=3.2, eval=TRUE}
 strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=20, binwidth=0.05)
+```
+
+```{r, schubert-ubi-scatter, fig.width=4.5, fig.height=3.5}
 rank_scatter(res, pvalue="pval", covariate="ubiquity")
@@
 
-```{r, schubert-abun, fig.width=10, fig.height=3.5, eval=TRUE}
+```{r, schubert-abun, fig.width=10, fig.height=3.2, eval=TRUE}
 strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
+```
+
+```{r, schubert-abun-scatter, fig.width=4.5, fig.height=3.5}
 rank_scatter(res, pvalue="pval", covariate="mean_abun_present")

### 2018-02-26 datasets/ChIP-seq/ChIP-seq.sh
Add submit scripts for all cases (for now)

@@ -3,7 +3,7 @@
 #SBATCH -n 8
 #SBATCH -N 1
 #SBATCH -p irizarry,serial_requeue
-#SBATCH --mem 8GB
+#SBATCH --mem 16G
 #SBATCH -t 0-05:00
 #SBATCH -o render-%j.out
 #SBATCH -e render-%j.err

### 2018-02-26 datasets/GSEA/gsea.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+#SBATCH -J knitGSEA
+#SBATCH -n 2
+#SBATCH -N 1
+#SBATCH -p shared
+#SBATCH --mem 32G
+#SBATCH -t 0-6:00
+#SBATCH -o render-%j.out
+#SBATCH -e render-%j.err
+ 
+##module load rstudio
+export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/rstudio/0.98.1103-fasrc01/bin/pandoc/"
+
+R -e "rmarkdown::render('gsea-human.Rmd')"
+R -e "rmarkdown::render('gsea-mouse.Rmd')"

### 2018-02-26 datasets/microbiome/microbiome.sh

@@ -1,0 +1,16 @@
+#!/bin/bash
+#SBATCH -J knit16s
+#SBATCH -n 4
+#SBATCH -N 1
+#SBATCH -p shared
+#SBATCH --mem 32G
+#SBATCH -t 0-6:00
+#SBATCH -o render-%j.out
+#SBATCH -e render-%j.err
+ 
+##module load rstudio
+export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/rstudio/0.98.1103-fasrc01/bin/pandoc/"
+
+R -e "rmarkdown::render('microbiome-baxter.Rmd')"
+R -e "rmarkdown::render('microbiome-goodrich.Rmd')"
+R -e "rmarkdown::render('microbiome-schubert.Rmd')"

### 2018-02-26 datasets/scrnaseq/scRNAseq-keegan.Rmd

@@
 For data processing, we follow practices and recommendations from Soneson et al., 2017.
-Briefly, we limit ourselves to 2-group comparisons, and  select the maximal size defined as the smallest number of cells in either group as the group-level sample size for the DE analyses.
+Briefly, we limit ourselves to 2-group comparisons.
 We do not filter genes as filtering may affect the subsequent false discovery control steps.
@@
 cdata=colData(humandata)
+cdata$group <- cdata$source_name_ch1
 
@@
 # look at number of cells in each group
-table(colData(humandata)$source_name_ch1)
+table(colData(humandata)$group)
 
@@
 cdata=colData(humandata2)
+cdata$group <- cdata$characteristics_ch1.7
 
@@
 # look at number of cells in each group
-table(colData(humandata2)$characteristics_ch1.7)
+table(colData(humandata2)$group)
 
@@
 cdata=colData(mousedata)
+cdata$group <- cdata$source_name_ch1
 
@@
 # look at number of cells in each group
-table(as.character(colData(mousedata)$source_name_ch1))
+table(as.character(colData(mousedata)$group))
 
@@
 cdata=colData(mousedata2)
+cdata$group <- cdata$characteristics_ch1.2
 
@@
 # look at number of cells in each group
-table(as.character(colData(mousedata2)$characteristics_ch1.2))
+table(as.character(colData(mousedata2)$group))
 
@@
 
-** Still need to clean up from here on; nothing below is eval'd**
-
 We perform differential expression testing using 3 methods on the unfiltered data: non-parametric Wilcoxon, MAST which has a hurdle model to account for zeros and scDD which implicitly models any heterogeneity in expression.
-
-We will first assign names for output files of the DE methods as they will be handy downstream.
-
-```{r, filenames, eval=FALSE}
-h_wilcox_dat1=paste0(resdir,"h_data_wilcox_p_GSE63818-GPL16791.rds")
-h_wilcox_dat2=paste0(resdir,"h_data_wilcox_p_GSE84465.rds")
-m_wilcox_dat1=paste0(resdir,"m_data_wilcox_p_GSE45719.rds")
-m_wilcox_dat2=paste0(resdir,"m_data_wilcox_p_GSE94383.rds")
-
-h_scdd_dat1=paste0(resdir,"humanplate_scDatEx_ks_GSE63818-GPL16791.rds")
-h_scdd_dat2=paste0(resdir,"humanplate_scDatEx_ks_GSE84465.rds")
-m_scdd_dat1=paste0(resdir,"mouseplate_scDatEx_ks_GSE45719.rds")
-m_scdd_dat2=paste0(resdir,"mouseplate_scDatEx_ks_GSE94383.rds")
-
-m_mast_dat1=paste0(resdir,"mouseplate_masttpm_GSE45719.rds")
-h_mast_dat1=paste0(resdir,"humanplate_masttpm_GSE63818-GPL16791.rds")
-h_mast_dat2=paste0(resdir,"humanplate_masttpm_GSE84465.rds")
-m_mast_dat2=paste0(resdir,"mouseplate_masttpm_GSE94383.rds")
-
-```
 
@@
 
-```{r, wilcoxon, eval=FALSE}
+```{r, wilcoxon}
+# function to add wilcox p-values
+compute_wilcoxde_sc <- function(dat){
+  expr <- assays(dat)[["TPM"]] 
+  tmm <- edgeR::calcNormFactors(expr)
+  expr <- edgeR::cpm(expr, lib.size = tmm * colSums(expr))
+  rowData(dat)$wilcox_p <- sapply(seq_len(nrow(expr)), 
+    function(i) {
+      wilcox.test(expr[i, ] ~ colData(dat)$group)$p.value
+    })
+  return(dat)
+}
 
-g_data_sub=readRDS(paste0(datdir,"h_g_nofilter_subdata_GSE63818-GPL16791.rds"))
-groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
-compute_wilcoxde_sc(g_data_sub,h_wilcox_dat1,groups)
+humandata <- compute_wilcoxde_sc(humandata)
 
-g_data_sub=readRDS(paste0(datdir,"h_g_nofilter_subdata_GSE84465.rds"))
-groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
-compute_wilcoxde_sc(g_data_sub,h_wilcox_dat2,groups)
+humandata2 <- compute_wilcoxde_sc(humandata2)
 
-g_data_sub=readRDS(paste0(datdir,"m_g_nofilter_subdata_GSE45719.rds"))
-groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
-compute_wilcoxde_sc(g_data_sub,m_wilcox_dat1,groups)
+mousedata <- compute_wilcoxde_sc(mousedata)
 
-g_data_sub=readRDS(paste0(datdir,"m_g_nofilter_subdata_GSE94383.rds"))
-groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
-compute_wilcoxde_sc(g_data_sub,m_wilcox_dat2,groups)
-
+mousedata2 <- compute_wilcoxde_sc(mousedata2)
 ```
@@
 ### Differences in distributions of gene expression in the 2 conditions
-
+** Still need to clean up from here on; nothing below is eval'd**
 We apply scDD to log-transformed counts as below. Th helper functions set up the Single-Cell experiment object.

### 2018-02-26 datasets/scrnaseq/scRNAseq-latest.Rmd

--- 2018-02-22 datasets/scrnaseq/scRNAseq-latest.Rmd
@@
 
-#installation
-# source("https://bioconductor.org/biocLite.R")
-# biocLite("swfdr")
-# biocLite("IHW")
-# biocLite("qvalue")
-# install.packages("ashr")
-# install.packages("fdrtool")
-# biocLite("edgeR")
-# source("https://bioconductor.org/biocLite.R")
-# biocLite("scran")
-#  Samseq installation
-# source("https://bioconductor.org/biocLite.R")
-# biocLite("impute") 
-# install.packages("samr")
-# source("https://bioconductor.org/biocLite.R")
-#biocLite("MultiAssayExperiment")
-#source("https://bioconductor.org/biocLite.R")
-#biocLite("SummarizedExperiment")
-#biocLite("BiocGenerics")
-#biocLite("S4Vectors")
-#biocLite("BiocParallel")
-#install_github("areyesq89/SummarizedBenchmark")
-#biocLite("EBSeq")
-#biocLite("scDD")
-#source("https://bioconductor.org/biocLite.R")
-#biocLite("SingleCellExperiment")
-#source("https://bioconductor.org/biocLite.R")
-#biocLite("MAST")
-#install.packages("FDRreg")
-
 # packages for data analyses
-suppressMessages(library(MultiAssayExperiment))
-suppressMessages(library(dplyr))
-suppressMessages(library(ggplot2))
-suppressMessages(library(cowplot))
-suppressMessages(library(UpSetR))
-suppressMessages(library(hexbin))
+library(MultiAssayExperiment)
+library(dplyr)
+library(ggplot2)
+library(cowplot)
+library(UpSetR)
+library(hexbin)
 
 # packages for differential expression testing
-suppressMessages(library(impute))
-#suppressMessages(library(samr)
-suppressMessages(library(scDD))
-suppressMessages(library(EBSeq))
-suppressMessages(library(edgeR))
-suppressMessages(library(MAST))
-
-# benchmark methods
-suppressMessages(library(IHW))
-suppressMessages(library(ashr))
-suppressMessages(library(qvalue))
-suppressMessages(library(swfdr))
-suppressMessages(library(fdrtool))
-suppressMessages(library(FDRreg))
+library(impute)
+library(scDD)
+library(EBSeq)
+library(edgeR)
+library(MAST)
 
 # summarizebenchmark
-suppressMessages(library(SummarizedBenchmark))
-suppressMessages(library(SingleCellExperiment))
+library(SummarizedBenchmark)
+library(SingleCellExperiment)
 
 # useful libraries
-suppressMessages(library(R.utils))
-suppressMessages(library(knitr))
-suppressMessages(library(BiocParallel))
-core <- as.numeric(Sys.getenv("SLURM_NTASKS"))
+library(R.utils)
+library(knitr)
+library(BiocParallel)
+
+cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
 multicoreParam <- MulticoreParam(workers = cores)
@@
 
-```{bash, humandata-download}
+```{r, humandata-download}
+# all data and results are on the wd: /net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/fdrbenchmark-scrnaseq
 
-download.file(url="http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE63818-GPL16791.rds",destfile=paste0(datdir,"GSE63818-GPL16791.rds"))
+#these look well separated on the tsne plots on the conquer manuscript supplement
+download.file("http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE63818-GPL16791.rds",
+  destfile=paste0(datdir, "GSE63818-GPL16791.rds"))
 
-download.file(url="http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE84465.rds",destfile=paste0(datdir,"GSE84465.rds"))
+## human ds2
+download.file("http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE84465.rds",
+  destfile=paste0(datdir, "GSE84465.rds"))
 
@@
 
-```{bash, mousedata-download}
+```{r, mousedata-download}
+#another well-separated dataset on tsne: GSE60749−GPL13112
+download.file("http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE45719.rds",
+  destfile=paste0(datdir, "GSE45719.rds"))
 
-download.file(url="http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE45719.rds",destfile=paste0(datdir,"GSE45719.rds"))
-
-download.file(url="http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE94383.rds",destfile=paste0(datdir,"GSE94383.rds"))
-
+download.file(url="http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE94383.rds",
+  destfile=paste0(datdir,"GSE94383.rds"))
 ```
@@
 g_data_sub=g_data[rowSums(g_data==0)!=dim(g_data)[2],] 
-saveRDS(g_data_sub,paste0(datdir,"h_g_nofilter_subdata_GSE63818-GPL16791.rds")) # removing rows with all zeros reduces the number of genes to 33169
+saveRDS(g_data_sub, paste0(datdir, "h_g_nofilter_subdata_GSE63818-GPL16791.rds")) # removing rows with all zeros reduces the number of genes to 33169
 
@@
 
-saveRDS(g1_ids,paste0(datdir,"g1_ids_GSE63818-GPL16791.rds"))
-saveRDS(g2_ids,paste0(datdir,"g2_ids_GSE63818-GPL16791.rds"))
+saveRDS(g1_ids, paste0(datdir, "g1_ids_GSE63818-GPL16791.rds"))
+saveRDS(g2_ids, paste0(datdir, "g2_ids_GSE63818-GPL16791.rds"))
 ```
@@
 g_data_sub=g_data[rowSums(g_data==0)!=dim(g_data)[2],]
-saveRDS(g_data_sub,paste0(datdir,"h_g_nofilter_subdata_GSE84465.rds")) # removing rows with all zeros reduces the number of genes to 52942
+
+saveRDS(g_data_sub, paste0(datdir, "h_g_nofilter_subdata_GSE84465.rds")) # removing rows with all zeros reduces the number of genes to 52942
 
@@
 g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=dim(g_cdata)[2],]
-saveRDS(g_cdata_sub,paste0(datdir,"h_g_nofilter_subcdata_GSE84465.rds")) # removing rows with all zeros reduces the number of genes to 52942
+saveRDS(g_cdata_sub, paste0(datdir, "h_g_nofilter_subcdata_GSE84465.rds")) # removing rows with all zeros reduces the number of genes to 52942
 
-saveRDS(g1_ids,paste0(datdir,"g1_ids_GSE84465.rds"))
-saveRDS(g2_ids,paste0(datdir,"g2_ids_GSE84465.rds"))
+saveRDS(g1_ids, paste0(datdir, "g1_ids_GSE84465.rds"))
+saveRDS(g2_ids, paste0(datdir, "g2_ids_GSE84465.rds"))
 ```
@@
 ##mouse data
-mousedata=readRDS(paste0(datdir,"GSE45719.rds"))
+mousedata=readRDS(paste0(datdir, "GSE45719.rds"))
+
 #get gene-level data
@@
 #save data
-saveRDS(g1_data,paste0(datdir,"m_g1_data_GSE45719.rds"))
-saveRDS(g2_data,paste0(datdir,"m_g2_data_GSE45719.rds"))
+saveRDS(g1_data,paste0(datdir, "m_g1_data_GSE45719.rds"))
+saveRDS(g2_data,paste0(datdir, "m_g2_data_GSE45719.rds"))
 #COUNTS

### 2018-02-26 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd
Follow standard case study formatting

@@
 ---
-title: "Comparison of FDR methods on GWAS with SummarizedBenchmark"
+title: "Case study: Genome-Wide Association Analyses (GWAS)"
 author: "Keegan Korthauer"
@@
 
-In the previous document `GWAS.Rmd`, we identified three GWAS datasets for FDR
-benchmarking. Here we will make use of the new `SummarizedBenchmark` package to
-perform the benchmarking in a standardized way.
+Here we will make use of the `SummarizedBenchmark` package to benchmark a 
+Genome-Wide Association Study (GWAS) dataset. We will examine a GWAS 
+meta-analysis for Body Mass Index (BMI), which was also included in the 
+Boca-Leek manuscript.
 
@@
 
+This analysis requires that the software PLINK (version 1.9) 
+is installed on your system.
+PLINK is freely available for download 
+[here](http://zzz.bwh.harvard.edu/plink/plink2.shtml).
+
 ```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
+# data and results directories
 system("mkdir -p DATA")
@@
 library(ggplot2)
-library(magrittr)
-library(R.utils)
 library(cowplot)
 
-# benchmark methods
-library(IHW)
-library(ashr)
-library(qvalue)
-library(swfdr)
-library(fdrtool)
-library(FDRreg)
-
 # comparison tools/functions
-library(SummarizedBenchmark)
+library(R.utils)
 sourceDirectory("../R/")
@@
 
-# GWAS meta-analysis for BMI (included in Boca-Leek)
+# Data preparation
 
-Here we download and analyze the GWAS BMI dataset analyzed in the Boca-Leek paper.
+We'll first download the GWAS summary statistics dataset from the 
+GIANT consortium data portal.
 
@@
 # use PLINK clumping tool to obtain a set of independent SNPs by LD
-module load plink
-
 plink \
---bfile DATA/GWAS/1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01 \
---clump DATA/GWAS/rslist.assoc \
+--bfile DATA/1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01 \
+--clump DATA/rslist.assoc \
 --clump-field P \
@@
 --clump-kb 250 \
---out DATA/GWAS/plink_clump
+--out DATA/plink_clump
 
@@
 
-## Check Covariate Diagnostics
+# Data Analysis
+
+## Differential Testing
+Since we do not have access to the individual genotype information and since
+our data already contains the results of performing hypothesis tests of
+association of BMI with each SNP (including p-values and effect sizes), we 
+don't need to carry out any additional testing. We move on to checking the
+covariate diagnostics and adjusting for multiple comparisons.
+
+## Covariate Diagnostics
 
@@
 
-```{r, GWAS1-diag, fig.width=10, fig.height=3.5}
+### Covariate one: Sample Size
+
+```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
+rank_scatter(bmi, pvalue="pval", covariate="ind_covar_N") + 
+  ggtitle("Covariate 1: Sample Size")
+```
+
+```{r, fig.width=10, fig.height=3.2, message=FALSE}
 strat_hist(bmi, pvalue="pval", covariate="ind_covar_N", maxy=2) 
-rank_scatter(bmi, pvalue="pval", covariate="ind_covar_N")
 ```
+
+### Covariate two: Minor Allele Frequency
 
@@
 
-```{r, GWAS1-diag2, fig.width=10, fig.height=3.5}
-strat_hist(bmi, pvalue="pval", covariate="ind_covar_AF", maxy=2)
-rank_scatter(bmi, pvalue="pval", covariate="ind_covar_AF")
+```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
+rank_scatter(bmi, pvalue="pval", covariate="ind_covar_AF") +
+  ggtitle("Covariate 1: Minor allele frequency")
 ```
 
-## Set up BenchDesign object
+```{r, fig.width=10, fig.height=3.2, message=FALSE}
+strat_hist(bmi, pvalue="pval", covariate="ind_covar_AF", maxy=2)
+```
+
+## Multiple-Testing Correction
 
@@
 
-```{r, GWAS1-benchdesign}
+```{r, GWAS1-benchdesign, message=FALSE}
 bd <- initializeBenchDesign()
 ```
-
-## Run buildBench for the sample size covariate and examine results
 
@@
                                         helper scripts).
+
+### Covariate one: Sample size
+
+First we'll include the sample size covariate.
 
@@
   t1 <- proc.time()
-  sb <- bd %>% buildBench(data=(bmi %>% mutate(ind_covariate=ind_covar_N)), 
+  sbN <- bd %>% buildBench(data=(bmi %>% mutate(ind_covariate=ind_covar_N)), 
                           ftCols = c("ind_covar_N"),
                           parallel=TRUE, BPPARAM=multicoreParam)
-  metadata(sb)$data_download_link <- "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
-  save(sb, file=resN)
+  metadata(sbN)$data_download_link <- 
+    "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
+  saveRDS(sbN, file=resN)
   duration <- round((proc.time()-t1)[3]/60,1)
 }else{
-  load(resN)
+  sbN <- readRDS(resN)
 }
@@
 
-Next, we'll add the default performance metric for q-value assays. First, we have
-to rename the assay to 'qvalue'.
+### Covariate two: Minor Allele Frequency
 
-```{r, GWAS1-metrics}
-# rename assay to qvalue
-assayNames(sb) <- "qvalue"
-sb <- addDefaultMetrics(sb)
-```
-
-Now, we'll plot the results.
-
-```{r, GWAS1-plot, results="hide", width=15, height=15}
-# plot nrejects by method overall and stratified by covariate
-rejections_scatter(sb,
-                   supplementary=FALSE)
-
-rejection_scatter_bins(sb, covariate="ind_covar_N", bins=4,
-                       supplementary=FALSE)
-
-# upset plot 
-plotFDRMethodsOverlap(sb, 
-                      alpha=0.05, nsets=ncol(sb),
-                      order.by="freq", decreasing=TRUE,
-                      supplementary=FALSE)
-
-# properties of method-exclusive discoveries
-methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
-             "scott-empirical", "scott-theoretical")
-mcols(sb)$ind_covariate <- mcols(sb)$ind_covar_N
-plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3)
-
-```
-
-## Run buildBench for the allele frequency covariate and examine results
-
-Now, we'll repeat the previous steps for the other covariate (AF):
+Now, we'll repeat the multiple testing correction using the other covariate (AF):
 
 ```{r, GWAS1-sb2, results="hide", message=FALSE}
-bd <- initializeBenchDesign()
 resAF <- paste0(resdir, "bmi_AF_summarizedBenchmark_",
@@
   t1 <- proc.time()
-sb <- bd %>% buildBench(data=(bmi %>% mutate(ind_covariate=ind_covar_AF) %>%
+  sbAF <- bd %>% buildBench(data=(bmi %>% mutate(ind_covariate=ind_covar_AF) %>%
                                     filter(!is.na(ind_covariate))), 
@@
                             parallel=TRUE, BPPARAM=multicoreParam)
-metadata(sb)$data_download_link <- "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
-save(sb, file=resAF)
+  metadata(sbAF)$data_download_link <-    
+    "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
+  saveRDS(sbAF, file=resAF)
   duration <- round((proc.time()-t1)[3]/60,1)
 }else{
-load(resAF)
+  sbAF <- readRDS(resAF)
 }
@@
 
-Next, we'll add the default performance metric for q-value assays. First, we have
+## Benchmark Metrics
+
+Next, we'll add the default performance metric for q-value assays and 
+plot the results. We'll start with covariate one.
+
+### Covariate one: Sample Size
+
+First, we have
 to rename the assay to 'qvalue'.
 
-```{r, GWAS1-metrics2}
+```{r, GWAS1-metrics}
 # rename assay to qvalue
-assayNames(sb) <- "qvalue"
-sb <- addDefaultMetrics(sb)
+assayNames(sbN) <- "qvalue"
+sbN <- addDefaultMetrics(sbN)
 ```
@@
 
-```{r, GWAS1-plot2, results="hide", width=15, height=15}
+```{r, GWAS1-plot, results="hide"}
 # plot nrejects by method overall and stratified by covariate
-rejections_scatter(sb,
-supplementary=FALSE)
+rejections_scatter(sbN,
+                   supplementary=FALSE) +
+  ggtitle("Covariate 1: Sample size")
 
-rejection_scatter_bins(sb, covariate="ind_covar_AF", bins=4,
-supplementary=FALSE)
+rejection_scatter_bins(sbN, covariate="ind_covar_N", bins=4,
+                       supplementary=FALSE) +
+  ggtitle("Covariate 1: Sample size")
 
 # upset plot 
-plotFDRMethodsOverlap(sb, 
+plotFDRMethodsOverlap(sbN, 
                       alpha=0.05, nsets=ncol(sb),
@@
              "scott-empirical", "scott-theoretical")
-mcols(sb)$ind_covariate <- abs(mcols(sb)$ind_covar_AF-0.5)
-plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods, maxNum=3)
+mcols(sbN)$ind_covariate <- mcols(sbN)$ind_covar_N
+plotCovariateBoxplots(sbN, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3) +
+  ggtitle("Covariate 1: Sample size")
 
@@
 
-Clean up the workspace.
+### Covariate two: Minor Allele Frequency
 
-```{r, GWAS1-cleanup}
-rm(bmi)
-rm(bd)
-rm(sb)
+Next, we'll look at the performance metrics for the other covariate (minor
+allele frequency).
+
+```{r, GWAS1-metrics2}
+# rename assay to qvalue
+assayNames(sbAF) <- "qvalue"
+sbAF <- addDefaultMetrics(sbAF)
 ```
+
+Now, we'll plot the results.
+
+```{r, GWAS1-plot2, results="hide"}
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter(sbAF, supplementary=FALSE) +
+  ggtitle("Covariate 2: Minor allele frequency")
+
+rejection_scatter_bins(sbAF, covariate="ind_covar_AF", bins=4,
+                       supplementary=FALSE) +
+  ggtitle("Covariate 2: Minor allele frequency")
+
+# upset plot 
+plotFDRMethodsOverlap(sbAF, 
+                      alpha=0.05, nsets=ncol(sbAF),
+                      order.by="freq", decreasing=TRUE,
+                      supplementary=FALSE)
+
+# properties of method-exclusive discoveries
+methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
+             "scott-empirical", "scott-theoretical")
+mcols(sbAF)$ind_covariate <- abs(mcols(sbAF)$ind_covar_AF-0.5)
+plotCovariateBoxplots(sbAF, alpha=0.1, nsets=6, methods=methods, maxNum=3)+
+  ggtitle("Covariate 2: Minor allele frequency")
+
+```
+
+# Session Information
+
+```{r}
+sessionInfo()
+```

### 2018-02-26 datasets/scrnaseq/scRNAseq-keegan.Rmd
Add checks for precomputed objects; Add scDD evaluation

@@
 title: "scRNAseq"
-author: "Ayshwarya Subramanian"
-date: "9/15/2017"
-output: html_document
+author: "Ayshwarya Subramanian and Keegan Korthauer"
+date: '`r format(Sys.Date(), "%B %d, %Y")`'
+output: 
+   html_document:
+        toc: true
+        toc_float: true
+        highlight: tango
+        number_sections: true
 ---
+
 ```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE,cache=TRUE,cache.lazy = FALSE)
+```
+# Summary
 
+For the application of differential gene expression, we look at scRNAseq and bulk RNAseq data. Here we discuss the analyses of scRNAseq.
+
+```{r echo=FALSE, results='asis'}
+library(knitr)
+df=data.frame(datasets=c("GSE63818-GPL16791","GSE84465","GSE45719","GSE94383"),
+              organism=c("human","human","mouse","mouse"),
+              technology=c("smart-seq","smart-seq","smart-seq2","smart-seq2"),
+              comparison=c("Somatic vs Primordial Germ Cells", 
+                           "Neoplastic vs Regular",
+                           "16-cell stage blastomere vs Mid blastocyst cell",
+                           "LPS vs No stimulation"))
+kable(df,caption="scRNAseq Datasets")
+df1=data.frame(DE_method=c("Wilcoxon-rank sum","ScDD","MAST"),
+               input=c("tpm","count","tpm"),
+               output=c("p-val","p-val","p-val"))
+kable(df1,caption="DE methods")
 ```
-## R workspace setup
+
+We selected four datasets, two each from mouse and human. 
+These datasets come from the `Conquer` (Soneson et al., 2017) scRNAseq data collection.
+`Conquer` provides uniformly processed gene expression data summarized to both count (aggregated from transcript members), and length-scaled TPMs at the gene level.
+The authors used `salmon` to quantify raw reads and `tximport` to summarize to gene level.
+The data comes from Smartseq, a plate based sequencing method which results in deeper sequencing of reads at the cost of numbers of cells.
+
+Here we are interested in performing differential expression analyses to determine
+which
+
+# Workspace setup
 
@@
 library(R.utils)
-library(knitr)
 library(BiocParallel)
@@
 
-## scRNAseq datasets
+# Data preparation
 
-For the application of differential gene expression, we look at scRNAseq and bulk RNAseq data. Here we discuss the analyses of scRNAseq.
-
-```{r echo=FALSE, results='asis'}
-df=data.frame(datasets=c("GSE63818-GPL16791","GSE84465","GSE45719","GSE94383"),organism=c("human","human","mouse","mouse"),technology=c("smart-seq","smart-seq","smart-seq2","smart-seq2"))
-kable(df,caption="scRNAseq Datasets")
-df1=data.frame(DE_method=c("Wilcoxon-rank sum","ScDD","MAST"),input=c("tpm","count","tpm"),output=c("p-val","p-val","p-val"))
-kable(df1,caption="DE methods")
-
-```
-
-We selected two datasets, one each from mouse and human spp, from the `Conquer` (Soneson et al., 2017) scRNAseq data collection.
-`Conquer` provides uniformly processed gene expression data summarized to both count (aggregated from transcript members), and length-scaled TPMs at the gene level.
-The authors use `salmon` to quantify raw reads and `tximport` to summarize to gene level.
-The data comes from Smartseq, a plate based sequencing method which results in deeper sequencing of reads at the cost of numbers of cells.
-
-## Data Download
-
-We first download the data in the data directory.`conquer` provides the processed datasets in convenient .rds files. First, let's download the Human datasets.
+We first download the data in the data directory.`conquer` provides the processed datasets in convenient .rds files. First, let's download the Human datasets: GSE63818-GPL16791 and
+GSE84465.
 
@@
 
-Next, we download the Mouse data:GSE45719
+Next, we download the Mouse data: GSE45719 and GSE94383
 
@@
 ```{r, dataload}
+datfile_h1 <- paste0(datdir, "h1_GSE63818-GPL16791.rds")
+
+if (!file.exists(datfile_h1)){
+
    # load MultiAssayExperiment
@@
   # save SummarizedExperiment 
-saveRDS(humandata, paste0(datdir, "h1_GSE63818-GPL16791.rds"))
+  saveRDS(humandata, datfile_h1)
+}else{
+  humandata <- readRDS(datfile_h1)
+}
 ```
@@
 ```{r}
+datfile_h2 <- paste0(datdir, "h2_GSE84465.rds")
+
+if (!file.exists(datfile_h2)){
+
   # load MultiAssayExperiment
@@
   # save SummarizedExperiment 
-saveRDS(humandata2, paste0(datdir, "h2_GSE84465.rds"))
+  saveRDS(humandata2, datfile_h2)
+}else{
+  humandata2 <- readRDS(datfile_h2)
+}
 ```
@@
 ```{r,mouse}
+datfile_m1 <- paste0(datdir, "m1_GSE45719.rds")
+
+if (!file.exists(datfile_m1)){
+
   # load MultiAssayExperiment
@@
   # save SummarizedExperiment 
-saveRDS(mousedata, paste0(datdir, "m1_GSE45719.rds"))
+  saveRDS(mousedata, datfile_m1)
+}else{
+  mousedata <- readRDS(datfile_m1)
+}
 ```
@@
 ```{r}
+datfile_m2 <- paste0(datdir, "m2_GSE94383.rds")
+
+if (!file.exists(datfile_m2)){
+
   # load MultiAssayExperiment
@@
   # save SummarizedExperiment 
-saveRDS(mousedata2, paste0(datdir, "m2_GSE94383.rds"))
+  saveRDS(mousedata2, datfile_m2)
+}else{
+  mousedata2 <- readRDS(datfile_m2) 
+}
 ```
@@
 compute_wilcoxde_sc <- function(dat){
+  # check if already computed
+  if (! "wilcox_p" %in% colnames(rowData(dat))){
     expr <- assays(dat)[["TPM"]] 
@@
       })
+  }else{
+     message("Wilcox p-values already computed.")
+  }
   return(dat)
@@
 mousedata2 <- compute_wilcoxde_sc(mousedata2)
+
+# save results
+saveRDS(humandata, datfile_h1)
+saveRDS(humandata2, datfile_h2)
+saveRDS(mousedata, datfile_m1)
+saveRDS(mousedata2, datfile_m2)
 ```
@@
 ### Differences in distributions of gene expression in the 2 conditions
-** Still need to clean up from here on; nothing below is eval'd**
-We apply scDD to log-transformed counts as below. Th helper functions set up the Single-Cell experiment object.
+
+We apply scDD to normalized log-transformed counts.
 
 ```{r, scdd, eval=FALSE}
+# function to add scDD p-values
+compute_scdd_sc <- function(dat){
+  # check if already computed
+  if (! "scdd_p" %in% colnames(rowData(dat))){
+    sce <- SingleCellExperiment(assays=list(count=assays(dat)[["count"]]),
+                                colData=data.frame(colData(dat)))
+    # scran normalization
+    libsize <- scran::computeSumFactors(assay(sce, "count"))
+    normcounts(sce) <- GetNormalizedMat(assay(sce, "count"), libsize)
+
+    rowData(dat)$scdd_p <- 
+        results(scDD(sce, 
+             condition = "group",
+             prior_param = list(alpha=0.01, mu0=0, s0=0.01, a0=0.01, b0=0.01), 
+             testZeroes = FALSE))$nonzero.pvalue
+  }else{
+     message("scDD p-values already computed.")
+  }
+  return(dat)
+}
 
 ## human data
-g_cdata_sub=readRDS(paste0(datdir,"h_g_nofilter_subcdata_GSE63818-GPL16791.rds"))
-compute_scdd_sc(g_cdata_sub, h_scdd_dat1)
 
-g_cdata_sub=readRDS(paste0(datdir,"h_g_nofilter_subcdata_GSE84465.rds"))
-compute_scdd_sc(g_cdata_sub, h_scdd_dat2)
+compute_scdd_sc(humandata)
 
-## mouse data
-g_cdata_sub=readRDS(paste0(datdir,"m_g_nofilter_subcdata_GSE45719.rds"))
-compute_scdd_sc(g_cdata_sub, m_scdd_dat1)
+compute_scdd_sc(humandata2)
 
-g_cdata_sub=readRDS(paste0(datdir,"m_g_nofilter_subcdata_GSE94383.rds"))
-compute_scdd_sc(g_cdata_sub, m_scdd_dat2)
+compute_scdd_sc(mousedata)
+
+compute_scdd_sc(mousedata2)
 
@@
 ### MAST
+** Still need to clean up from here on; nothing below is eval'd**
 

### 2018-02-26 datasets/scrnaseq/scRNAseq-keegan.Rmd
Add MAST

@@
 
-```{r, scdd, eval=FALSE}
+```{r, scdd}
 # function to add scDD p-values
@@
 
+# save results
+saveRDS(humandata, datfile_h1)
+saveRDS(humandata2, datfile_h2)
+saveRDS(mousedata, datfile_m1)
+saveRDS(mousedata2, datfile_m2)
 ```
@@
 ### MAST
-** Still need to clean up from here on; nothing below is eval'd**
 
@@
 
-```{r, MAST,eval=FALSE, eval=FALSE}
-g_data_sub=readRDS(paste0(datdir,"m_g_nofilter_subdata_GSE45719.rds"))
-groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
-names(groups)=colnames(g_cdata_sub)
-compute_MASTtpmde(g_data_sub,groups,)
+```{r, MAST}
+# function to add scDD p-values
+compute_MAST_sc <- function(dat){
+  # check if already computed
+  if (! "mast_p" %in% colnames(rowData(dat))){
+    sca <- FromMatrix(exprsArray = log2(assays(dat)[["TPM"]] + 1), 
+                    cData = data.frame(wellKey=colnames(dat), colData(dat)))
+    zlmdata <- zlm.SingleCellAssay(~group, sca)
 
-g_data_sub=readRDS(paste0(datdir,"h_g_nofilter_subdata_GSE63818-GPL16791.rds"))
-groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
-names(groups)=colnames(g_data_sub)
-compute_MASTtpmde(g_data_sub,groups,h_mast_dat1)
+    rowData(dat)$mast_p <- 
+        lrTest(zlmdata, "group")[, "hurdle", "Pr(>Chisq)"]
+  }else{
+     message("scDD p-values already computed.")
+  }
+  return(dat)
+}
 
-g_data_sub=readRDS(paste0(datdir,"h_g_nofilter_subdata_GSE84465.rds"))
-groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
-names(groups)=colnames(g_data_sub)
-compute_MASTtpmde(g_data_sub,groups,h_mast_dat2)
+compute_MAST_sc(humandata)
 
-g_data_sub=readRDS(paste0(datdir,"m_g_nofilter_subdata_GSE94383.rds"))
-groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
-names(groups)=colnames(g_data_sub)
-compute_MASTtpmde(g_data_sub,groups,m_mast_dat2)
+compute_MAST_sc(humandata2)
+
+compute_MAST_sc(mousedata)
+
+compute_MAST_sc(mousedata2)
+
+# save results
+saveRDS(humandata, datfile_h1)
+saveRDS(humandata2, datfile_h2)
+saveRDS(mousedata, datfile_m1)
+saveRDS(mousedata2, datfile_m2)
 ```
@@
 ## Covariate Diagnostics
+** Still need to clean up from here on; nothing below is eval'd**
 

### 2018-02-26 datasets/microbiome/microbiome-goodrich.Rmd
Claire's edits to Patrick's edits

@@
 [(Goodrich et al., 2014)](http://dx.doi.org/10.1016/j.cell.2014.09.053).
-The dataset includes fecal samples from 428 individuals with low-to-normal body mass index (BMI < 25),
+The dataset includes fecal samples from 428 individuals with lean body mass index (BMI < 25),
 and 185 obese individuals (BMI > 30). We'll download the processed OTU tables from
 Zenodo and unzip them in the `benchmark-fdr/datasets/microbiome/data` folder.
+
+We expect few or no truly differentially abundant OTUs ([Finucane et al., 2014](http://dx.doi.org/10.1371/journal.pone.0084689) and [Walters et al., 2014](https://dx.doi.org/10.1016%2Fj.febslet.2014.09.039)), so this study provides a real-life "null" example.
 

### 2018-02-26 datasets/microbiome/microbiome-schubert.Rmd
Claire's edits to Patrick's edits

@@
 Zenodo and unzip them in the `benchmark-fdr/datasets/microbiome/data` folder.
+
+Diarrhea comes with a broad community re-structuring, so we expect there to be many truly differentially abundant OTUs in this dataset.
 
@@
 min_otu <- 10
+minp_otu <- 0.01
 min_sample <- 100
 
-## Remove OTUs w/ <= min reads, samples w/ <= min reads
+## Remove OTUs w/ <= min reads, w/ <= min prop, samples w/ <= min reads
 otu <- otu[rowSums(otu) > min_otu, ]
+otu <- otu[rowSums(otu > 0) / ncol(otu) > minp_otu, ]
 otu <- otu[, colSums(otu) > min_sample]

### 2018-02-26 datasets/R/helpers-microbiome.R

@@ -6,7 +6,7 @@
 #' for each OTU. Results are reported as a data.frame with
 #' one row per OTU.
 #' 
-#' @param abundance matrix of relative abunances with rows and
+#' @param abundance matrix of relative abundances with rows and
 #'        columns corresponding to OTUs and samples.
 #' @param shift psuedocount that was added to all relative abundance
 #'        values in \code{abundance} to handle zero counts.

### 2018-02-27 datasets/scrnaseq/scRNAseq-keegan.Rmd
Organize sc analyses into one dataset/method per Rmd

@@
 Here we are interested in performing differential expression analyses to determine
-which
+which genes have different expression among the two biological groups in each comparison
 
@@
 
-cores <- 8
+cores <- 6
 BiocParallel::register(MulticoreParam(workers = cores))
@@
 Per Soneson et al., we run TMM normalization to the TPM units before running Wilcoxon.
+The p-value results are added to the `rowData` slot of the `SummarizedExperiment` 
+data objects to ensure that gene metadata stays intact.
 
@@
 We apply scDD to normalized log-transformed counts. 
+The p-value results are added to the `rowData` slot of the `SummarizedExperiment` 
+data objects to ensure that gene metadata stays intact.
 
@@
 ## Covariate Diagnostics
-** Still need to clean up from here on; nothing below is eval'd**
 
-In scrnaseq data, strength of the signal can be a function of level of gene expression.
-We quantify it as median or mean gene expression,and detection rate, defined as the percentage of cells expressing the gene at a level of TPM>0.
+In scrnaseq data, strength of the signal can be a affected by the of level of gene expression. 
+We will explore two potential covariates related to expression level: mean expression 
+and detection rate (defined as the proportion of cells expressing the gene at a nonzero level). 
+These will also be added to the `rowData` slot of the `SummarizedExperiments`.
 
-```{r,cov, eval=FALSE}
-h_cov=compute_covariates_scrnaseq(paste0(datdir,"h_g_nofilter_subdata_GSE63818-GPL16791.rds"))
-m_cov=compute_covariates_scrnaseq(paste0(datdir,"m_g_nofilter_subdata_GSE45719.rds"))
-h2_cov=compute_covariates_scrnaseq(paste0(datdir,"h_g_nofilter_subdata_GSE84465.rds"))
-m2_cov=compute_covariates_scrnaseq(paste0(datdir,"m_g_nofilter_subdata_GSE94383.rds"))
+```{r,cov}
+add_covariates_scrnaseq <- function(dat){
+  rowData(dat)$meanExp <- rowMeans(assays(dat)[["count"]])
+  rowData(dat)$detection <- rowMeans(assays(dat)[["count"]]==0)
+  return(dat)
+}
+
+humandata <- add_covariates_scrnaseq(humandata)
+humandata2 <- add_covariates_scrnaseq(humandata2)
+mousedata <- add_covariates_scrnaseq(mousedata)
+mousedata2 <- add_covariates_scrnaseq(mousedata2)
+
+# save results
+saveRDS(humandata, datfile_h1)
+saveRDS(humandata2, datfile_h2)
+saveRDS(mousedata, datfile_m1)
+saveRDS(mousedata2, datfile_m2)
 ```
 
-First we will look at the covariates for the Wilcoxon text
+### Covariate one: Mean Expression
 
-```{r, wilcoxcov, fig.width=10, fig.height=7, eval=FALSE}
-print("human_plate_1")
-hwilcox1 <- plot_covariate_diagnostics(paste0(h_wilcox_dat1),"wilcox",h_cov)
-print(hwilcox1$plot)
+For each covariate, we'll examine the plots for each of the three analysis methods.
 
-print("human_plate_2")
-hwilcox2 <- plot_covariate_diagnostics(paste0(h_wilcox_dat2),"wilcox",h2_cov)
-print(hwilcox2$plot)
+#### Wilcox
 
-print("mouse_plate_1")
-mwilcox1 <- plot_covariate_diagnostics(paste0(m_wilcox_dat1),"wilcox",m_cov)
-print(mwilcox1$plot)
-
-print("mouse_plate_2")
-mwilcox2 <- plot_covariate_diagnostics(paste0(m_wilcox_dat2),"wilcox",m2_cov)
-print(mwilcox2$plot)
+```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
+rank_scatter(data.frame(rowData(humandata)), 
+             pvalue="wilcox_p", covariate="meanExp") + 
+  ggtitle("Wilcox, Covariate 1: Mean Expression")
+rank_scatter(data.frame(rowData(humandata2)), 
+             pvalue="wilcox_p", covariate="meanExp") + 
+  ggtitle("Wilcox, Covariate 1: Mean Expression")
+rank_scatter(data.frame(rowData(mousedata)), 
+             pvalue="wilcox_p", covariate="meanExp") + 
+  ggtitle("Wilcox, Covariate 1: Mean Expression")
+rank_scatter(data.frame(rowData(mousedata2)), 
+             pvalue="wilcox_p", covariate="meanExp") + 
+  ggtitle("Wilcox, Covariate 1: Mean Expression")
 ```
 
-Next, we look at the covariates for MAST
-
-```{r, mastcov, fig.width=10, fig.height=7, eval=FALSE}
-print("human_plate_1")
-hmast1 <- plot_covariate_diagnostics(paste0(h_mast_dat1),"mast",h_cov)
-print(hmast1$plot)
-
-print("human_plate_2")
-hmast2 <- plot_covariate_diagnostics(paste0(h_mast_dat2),"mast",h2_cov)
-print(hmast2$plot)
-
-print("mouse_plate_1")
-mmast1 <- plot_covariate_diagnostics(paste0(m_mast_dat1),"mast",m_cov)
-print(mmast1$plot)
-
-print("mouse_plate_2")
-mmast2 <- plot_covariate_diagnostics(paste0(m_mast_dat2),"mast",m2_cov)
-print(mmast2$plot)
+```{r, fig.width=10, fig.height=3.2, message=FALSE}
+strat_hist(data.frame(rowData(humandata)), 
+           pvalue="wilcox_p", covariate="meanExp", maxy=25) 
+strat_hist(data.frame(rowData(humandata2)), 
+           pvalue="wilcox_p", covariate="meanExp", maxy=25) 
+strat_hist(data.frame(rowData(mousedata)), 
+           pvalue="wilcox_p", covariate="meanExp", maxy=25) 
+strat_hist(data.frame(rowData(mousedata2)), 
+           pvalue="wilcox_p", covariate="meanExp", maxy=25) 
 ```
 
-Finally, we look at how informative the covariates are with results from ScDD.
+#### scDD
 
-```{r,scddcov, fig.width=10, fig.height=7, eval=FALSE}
-print("human_plate_1")
-hscdd1 <- plot_covariate_diagnostics(paste0(h_scdd_dat1),"scdd",h_cov)
-print(hscdd1$plot)
+```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
+rank_scatter(data.frame(rowData(humandata)), 
+             pvalue="scdd_p", covariate="meanExp") + 
+  ggtitle("scDD, Covariate 1: Mean Expression")
+rank_scatter(data.frame(rowData(humandata2)), 
+             pvalue="scdd_p", covariate="meanExp") + 
+  ggtitle("scDD, Covariate 1: Mean Expression")
+rank_scatter(data.frame(rowData(mousedata)), 
+             pvalue="scdd_p", covariate="meanExp") + 
+  ggtitle("scDD, Covariate 1: Mean Expression")
+rank_scatter(data.frame(rowData(mousedata2)), 
+             pvalue="scdd_p", covariate="meanExp") + 
+  ggtitle("scDD, Covariate 1: Mean Expression")
+```
 
-print("human_plate_2")
-hscdd2 <- plot_covariate_diagnostics(paste0(h_scdd_dat2),"scdd",h2_cov)
-print(hscdd2$plot)
+```{r, fig.width=10, fig.height=3.2, message=FALSE}
+strat_hist(data.frame(rowData(humandata)), 
+           pvalue="scdd_p", covariate="meanExp", maxy=25) 
+strat_hist(data.frame(rowData(humandata2)), 
+           pvalue="scdd_p", covariate="meanExp", maxy=25) 
+strat_hist(data.frame(rowData(mousedata)), 
+           pvalue="scdd_p", covariate="meanExp", maxy=25) 
+strat_hist(data.frame(rowData(mousedata2)), 
+           pvalue="scdd_p", covariate="meanExp", maxy=25) 
+```
 
-print("mouse_plate_1")
-mscdd1 <- plot_covariate_diagnostics(paste0(m_scdd_dat1),"scdd",m_cov)
-print(mscdd1$plot)
+#### MAST
 
-print("mouse_plate_2")
-mscdd2 <- plot_covariate_diagnostics(paste0(m_scdd_dat2),"scdd",m2_cov)
-print(mscdd2$plot)
+```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
+rank_scatter(data.frame(rowData(humandata)), 
+             pvalue="mast_p", covariate="meanExp") + 
+  ggtitle("MAST, Covariate 1: Mean Expression")
+rank_scatter(data.frame(rowData(humandata2)), 
+             pvalue="mast_p", covariate="meanExp") + 
+  ggtitle("MAST, Covariate 1: Mean Expression")
+rank_scatter(data.frame(rowData(mousedata)), 
+             pvalue="mast_p", covariate="meanExp") + 
+  ggtitle("MAST, Covariate 1: Mean Expression")
+rank_scatter(data.frame(rowData(mousedata2)), 
+             pvalue="mast_p", covariate="meanExp") + 
+  ggtitle("MAST, Covariate 1: Mean Expression")
+```
+
+```{r, fig.width=10, fig.height=3.2, message=FALSE}
+strat_hist(data.frame(rowData(humandata)), 
+           pvalue="mast_p", covariate="meanExp", maxy=25) 
+strat_hist(data.frame(rowData(humandata2)), 
+           pvalue="mast_p", covariate="meanExp", maxy=25) 
+strat_hist(data.frame(rowData(mousedata)), 
+           pvalue="mast_p", covariate="meanExp", maxy=25) 
+strat_hist(data.frame(rowData(mousedata2)), 
+           pvalue="mast_p", covariate="meanExp", maxy=25) 
+```
+
+The mean expression appears to be informative and approximately satisfies the 
+assumptions
+
+### Covariate two: Detection Rate
+
+Next we look at the detection rate covariate.
+
+#### Wilcox
+
+```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
+rank_scatter(data.frame(rowData(humandata)), 
+             pvalue="wilcox_p", covariate="detection") + 
+  ggtitle("Wilcox, Covariate 2: Detection Rate")
+rank_scatter(data.frame(rowData(humandata2)), 
+             pvalue="wilcox_p", covariate="detection") + 
+  ggtitle("Wilcox, Covariate 2: Detection Rate")
+rank_scatter(data.frame(rowData(mousedata)), 
+             pvalue="wilcox_p", covariate="detection") + 
+  ggtitle("Wilcox, Covariate 2: Detection Rate")
+rank_scatter(data.frame(rowData(mousedata2)), 
+             pvalue="wilcox_p", covariate="detection") + 
+  ggtitle("Wilcox, Covariate 2: Detection Rate")
+```
+
+```{r, fig.width=10, fig.height=3.2, message=FALSE}
+strat_hist(data.frame(rowData(humandata)), 
+           pvalue="wilcox_p", covariate="detection", maxy=25) 
+strat_hist(data.frame(rowData(humandata2)), 
+           pvalue="wilcox_p", covariate="detection", maxy=25) 
+strat_hist(data.frame(rowData(mousedata)), 
+           pvalue="wilcox_p", covariate="detection", maxy=25) 
+strat_hist(data.frame(rowData(mousedata2)), 
+           pvalue="wilcox_p", covariate="detection", maxy=25) 
+```
+
+#### scDD
+
+```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
+rank_scatter(data.frame(rowData(humandata)), 
+             pvalue="scdd_p", covariate="detection") + 
+  ggtitle("scDD, Covariate 2: Detection Rate")
+rank_scatter(data.frame(rowData(humandata2)), 
+             pvalue="scdd_p", covariate="detection") + 
+  ggtitle("scDD, Covariate 2: Detection Rate")
+rank_scatter(data.frame(rowData(mousedata)), 
+             pvalue="scdd_p", covariate="detection") + 
+  ggtitle("scDD, Covariate 2: Detection Rate")
+rank_scatter(data.frame(rowData(mousedata2)), 
+             pvalue="scdd_p", covariate="detection") + 
+  ggtitle("scDD, Covariate 2: Detection Rate")
+```
+
+```{r, fig.width=10, fig.height=3.2, message=FALSE}
+strat_hist(data.frame(rowData(humandata)), 
+           pvalue="scdd_p", covariate="detection", maxy=25) 
+strat_hist(data.frame(rowData(humandata2)), 
+           pvalue="scdd_p", covariate="detection", maxy=25) 
+strat_hist(data.frame(rowData(mousedata)), 
+           pvalue="scdd_p", covariate="detection", maxy=25) 
+strat_hist(data.frame(rowData(mousedata2)), 
+           pvalue="scdd_p", covariate="detection", maxy=25) 
+```
+
+#### MAST
+
+```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
+rank_scatter(data.frame(rowData(humandata)), 
+             pvalue="mast_p", covariate="detection") + 
+  ggtitle("MAST, Covariate 2: Detection Rate")
+rank_scatter(data.frame(rowData(humandata2)), 
+             pvalue="mast_p", covariate="detection") + 
+  ggtitle("MAST, Covariate 2: Detection Rate")
+rank_scatter(data.frame(rowData(mousedata)), 
+             pvalue="mast_p", covariate="detection") + 
+  ggtitle("MAST, Covariate 2: Detection Rate")
+rank_scatter(data.frame(rowData(mousedata2)), 
+             pvalue="mast_p", covariate="detection") + 
+  ggtitle("MAST, Covariate 2: Detection Rate")
+```
+
+```{r, fig.width=10, fig.height=3.2, message=FALSE}
+strat_hist(data.frame(rowData(humandata)), 
+           pvalue="mast_p", covariate="detection", maxy=25) 
+strat_hist(data.frame(rowData(humandata2)), 
+           pvalue="mast_p", covariate="detection", maxy=25) 
+strat_hist(data.frame(rowData(mousedata)), 
+           pvalue="mast_p", covariate="detection", maxy=25) 
+strat_hist(data.frame(rowData(mousedata2)), 
+           pvalue="mast_p", covariate="detection", maxy=25) 
 ```
@@
 
-First we will create the benchdesign object for wilcox -- IN PROGRESS
+** Still need to clean up from here on; nothing below is eval'd**
 
-```{r, benchmarking, eval=FALSE, eval=FALSE}
+We will skip running the Wilcox method for now.
+First we will adjust for multiple comparisons using the scDD results
+
+```{r, benchmarking, eval=FALSE}
 

### 2018-02-27 datasets/scrnaseq/scrnaseq-human-mast.Rmd
Organize sc analyses into one dataset/method per Rmd

--- 2018-02-26 datasets/scrnaseq/scRNAseq-keegan.Rmd
@@
 ---
-title: "scRNAseq"
+title: "Case Study: scRNA-seq (Human data, MAST method)"
 author: "Ayshwarya Subramanian and Keegan Korthauer"
@@
 library(knitr)
-df=data.frame(datasets=c("GSE63818-GPL16791","GSE84465","GSE45719","GSE94383"),
-              organism=c("human","human","mouse","mouse"),
-              technology=c("smart-seq","smart-seq","smart-seq2","smart-seq2"),
-              comparison=c("Somatic vs Primordial Germ Cells", 
-                           "Neoplastic vs Regular",
-                           "16-cell stage blastomere vs Mid blastocyst cell",
+df=data.frame(datasets=c("GSE84465","GSE94383"),
+              organism=c("human","mouse"),
+              technology=c("smart-seq","smart-seq2"),
+              comparison=c("Neoplastic vs Regular",
                            "LPS vs No stimulation"))
 kable(df,caption="scRNAseq Datasets")
-df1=data.frame(DE_method=c("Wilcoxon-rank sum","ScDD","MAST"),
-               input=c("tpm","count","tpm"),
-               output=c("p-val","p-val","p-val"))
+df1=data.frame(DE_method=c("ScDD","MAST"),
+               input=c("count","tpm"),
+               output=c("p-val","p-val"))
 kable(df1,caption="DE methods")
@@
 
-We selected four datasets, two each from mouse and human. 
+We selected two datasets, one each from mouse and human. 
 These datasets come from the `Conquer` (Soneson et al., 2017) scRNAseq data collection.
@@
 Here we are interested in performing differential expression analyses to determine
-which
+which genes have different expression among the two biological groups in each comparison.
+We also apply two different analysis methods: MAST and scDD. In this vignette, we will
+explore the human dataset analysed with MAST.
 
@@
 
-We first download the data in the data directory.`conquer` provides the processed datasets in convenient .rds files. First, let's download the Human datasets: GSE63818-GPL16791 and
+We first download the data in the data directory.`conquer` provides the processed datasets in convenient .rds files. First, we download the Human dataset:
 GSE84465.
@@
 ```{r, humandata-download}
-#these look well separated on the tsne plots on the conquer manuscript supplement
-download.file("http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE63818-GPL16791.rds",
-  destfile=paste0(datdir, "GSE63818-GPL16791.rds"))
-
-## human ds2
 download.file("http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE84465.rds",
@@
 
-```
-
-Next, we download the Mouse data: GSE45719 and GSE94383
-
-```{r, mousedata-download}
-#another well-separated dataset on tsne: GSE60749−GPL13112
-download.file("http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE45719.rds",
-  destfile=paste0(datdir, "GSE45719.rds"))
-
-download.file(url="http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE94383.rds",
-  destfile=paste0(datdir,"GSE94383.rds"))
 ```
@@
 
-First, we get the human data for the two groups of interest.
-We'll extract two conditions: Primordial Germ Cells and Somatic cells at 
-7 weeks gestation.
+We'll extract two conditions: Neoplastic and Regular cells, as
+defined by their tsne clusters.
 
-```{r, dataload}
-datfile_h1 <- paste0(datdir, "h1_GSE63818-GPL16791.rds")
+```{r}
+datfile <- paste0(datdir, "human_GSE84465.rds")
 
-if (!file.exists(datfile_h1)){
+if (!file.exists(datfile)){
 
   # load MultiAssayExperiment
-  humandata <- readRDS(paste0(datdir,"GSE63818-GPL16791.rds"))
+  humandata <- readRDS(paste0(datdir, "GSE84465.rds"))
 
   # subset by groups of interest
-  humandata <- humandata[, (colData(humandata)$source_name_ch1 %in% 
-                         c("Somatic Cells", "Primordial Germ Cells")) &
-                         (colData(humandata)$characteristics_ch1 == 
-                         c("developmental stage: 7 week gestation"))]
+  humandata <- humandata2[, (colData(humandata)$characteristics_ch1.7 %in% 
+                         c("neoplastic: Neoplastic", "neoplastic: Regular")) &
+                         (colData(humandata)$characteristics_ch1.5 %in% 
+                         c("tsne cluster: 11",  "tsne cluster: 8"))]
   cdata=colData(humandata)
-  cdata$group <- cdata$source_name_ch1
+  cdata$group <- cdata$characteristics_ch1.7
 
@@
   # save SummarizedExperiment 
-  saveRDS(humandata, datfile_h1)
+  saveRDS(humandata, datfile)
 }else{
-  humandata <- readRDS(datfile_h1)
+  humandata <- readRDS(datfile)
 }
@@
 
-For the human dataset1 there are `r ncol(humandata)` cells total and 
-`r nrow(humandata)` genes.
-
-Now humandata2. We'll extract two conditions: Neoplastic and Regular cells, as
-defined by their tsne clusters.
-
-```{r}
-datfile_h2 <- paste0(datdir, "h2_GSE84465.rds")
-
-if (!file.exists(datfile_h2)){
-
-  # load MultiAssayExperiment
-  humandata2 <- readRDS(paste0(datdir, "GSE84465.rds"))
-
-  # subset by groups of interest
-  humandata2 <- humandata2[, (colData(humandata2)$characteristics_ch1.7 %in% 
-                         c("neoplastic: Neoplastic", "neoplastic: Regular")) &
-                         (colData(humandata2)$characteristics_ch1.5 %in% 
-                         c("tsne cluster: 11",  "tsne cluster: 8"))]
-  cdata=colData(humandata2)
-  cdata$group <- cdata$characteristics_ch1.7
-
-  # subset by gene expression TPM & count assays
-  humandata2 <- experiments(humandata2)[[1]]
-  assays(humandata2) <- assays(humandata2)[c("TPM", "count")]
-  colData(humandata2) <- cdata
-
-  # subset by genes not all zero
-  humandata2 <- humandata2[rowSums(assays(humandata2)[["count"]]==0) < ncol(humandata2),]
-
-  # look at number of cells in each group
-  table(colData(humandata2)$group)
-
-  # save SummarizedExperiment 
-  saveRDS(humandata2, datfile_h2)
-}else{
-  humandata2 <- readRDS(datfile_h2)
-}
-```
-
-For the human dataset2, there are a total of `r nrow(humandata2)` cells and `r ncol(humandata2)` genes.
-
-Next, we repeat the processing steps for the mouse datasets. For the first one,
-we'll extract two conditions: 16-cell stage blastomere and mid blastocysts cells.
-
-```{r,mouse}
-datfile_m1 <- paste0(datdir, "m1_GSE45719.rds")
-
-if (!file.exists(datfile_m1)){
-
-  # load MultiAssayExperiment
-  mousedata=readRDS(paste0(datdir, "GSE45719.rds"))
-
-  # subset by groups of interest
-  mousedata <- mousedata[, colData(mousedata)$source_name_ch1 %in% 
-                         c("16-cell stage blastomere", 
-                           "Mid blastocyst cell (92-94h post-fertilization)")]
-  cdata=colData(mousedata)
-  cdata$group <- cdata$source_name_ch1
-
-  # subset by gene expression TPM & count assays
-  mousedata <- experiments(mousedata)[[1]]
-  assays(mousedata) <- assays(mousedata)[c("TPM", "count")]
-  colData(mousedata) <- cdata
-
-  # subset by genes not all zero
-  mousedata <- mousedata[rowSums(assays(mousedata)[["count"]]==0) < ncol(mousedata),]
-
-  # look at number of cells in each group
-  table(as.character(colData(mousedata)$group))
-
-  # save SummarizedExperiment 
-  saveRDS(mousedata, datfile_m1)
-}else{
-  mousedata <- readRDS(datfile_m1)
-}
-```
-
-For the mouse dataset1, we have a total of `r ncol(mousedata)` cells and 
-`r nrow(mousedata)` genes.
-
-Now mousedata2. We'll extract two conditions: No stimulation versus LPS at 150 min.
-
-```{r}
-datfile_m2 <- paste0(datdir, "m2_GSE94383.rds")
-
-if (!file.exists(datfile_m2)){
-
-  # load MultiAssayExperiment
-  mousedata2=readRDS(paste0(datdir,"GSE94383.rds"))
-
-  # subset by groups of interest
-  mousedata2 <- mousedata2[, (colData(mousedata2)$characteristics_ch1.2 ==
-                              "condition: No stimulation") |
-                           (colData(mousedata2)$characteristics_ch1.2 ==
-                              "condition: LPS" &
-                            colData(mousedata2)$characteristics_ch1.1 %in% 
-                              "time point: 150 min") ]
-  cdata=colData(mousedata2)
-  cdata$group <- cdata$characteristics_ch1.2
-
-  # subset by gene expression TPM & count assays
-  mousedata2 <- experiments(mousedata2)[[1]]
-  assays(mousedata2) <- assays(mousedata2)[c("TPM", "count")]
-  colData(mousedata2) <- cdata
-
-  # subset by genes not all zero
-  mousedata2 <- mousedata2[rowSums(assays(mousedata2)[["count"]]==0) < ncol(mousedata2),]
-
-  # look at number of cells in each group
-  table(as.character(colData(mousedata2)$group))
-
-  # save SummarizedExperiment 
-  saveRDS(mousedata2, datfile_m2)
-}else{
-  mousedata2 <- readRDS(datfile_m2) 
-}
-```
-
-For the mouse dataset2, we have a total of `r ncol(mousedata2)` cells and 
-`r nrow(mousedata2)` genes after filtering for non-detected genes.
+For the human dataset, there are a total of `r nrow(humandata2)` cells and `r ncol(humandata2)` genes.
 
@@
 
-We perform differential expression testing using 3 methods on the unfiltered data: non-parametric Wilcoxon, MAST which has a hurdle model to account for zeros and scDD which implicitly models any heterogeneity in expression.
+We perform differential expression testing using 
+MAST which has a hurdle model to account for zeroes.
 
-### Wilcoxon rank-sum test
-
-Per Soneson et al., we run TMM normalization to the TPM units before running Wilcoxon.
-
-```{r, wilcoxon}
-# function to add wilcox p-values
-compute_wilcoxde_sc <- function(dat){
-  # check if already computed
-  if (! "wilcox_p" %in% colnames(rowData(dat))){
-    expr <- assays(dat)[["TPM"]] 
-    tmm <- edgeR::calcNormFactors(expr)
-    expr <- edgeR::cpm(expr, lib.size = tmm * colSums(expr))
-    rowData(dat)$wilcox_p <- sapply(seq_len(nrow(expr)), 
-      function(i) {
-        wilcox.test(expr[i, ] ~ colData(dat)$group)$p.value
-      })
-  }else{
-     message("Wilcox p-values already computed.")
-  }
-  return(dat)
-}
-
-humandata <- compute_wilcoxde_sc(humandata)
-
-humandata2 <- compute_wilcoxde_sc(humandata2)
-
-mousedata <- compute_wilcoxde_sc(mousedata)
-
-mousedata2 <- compute_wilcoxde_sc(mousedata2)
-
-# save results
-saveRDS(humandata, datfile_h1)
-saveRDS(humandata2, datfile_h2)
-saveRDS(mousedata, datfile_m1)
-saveRDS(mousedata2, datfile_m2)
-```
-
-### Differences in distributions of gene expression in the 2 conditions
-
-We apply scDD to normalized log-transformed counts.
-
-```{r, scdd}
-# function to add scDD p-values
-compute_scdd_sc <- function(dat){
-  # check if already computed
-  if (! "scdd_p" %in% colnames(rowData(dat))){
-    sce <- SingleCellExperiment(assays=list(count=assays(dat)[["count"]]),
-                                colData=data.frame(colData(dat)))
-    # scran normalization
-    libsize <- scran::computeSumFactors(assay(sce, "count"))
-    normcounts(sce) <- GetNormalizedMat(assay(sce, "count"), libsize)
-
-    rowData(dat)$scdd_p <- 
-        results(scDD(sce, 
-             condition = "group",
-             prior_param = list(alpha=0.01, mu0=0, s0=0.01, a0=0.01, b0=0.01), 
-             testZeroes = FALSE))$nonzero.pvalue
-  }else{
-     message("scDD p-values already computed.")
-  }
-  return(dat)
-}
-
-## human data
-
-compute_scdd_sc(humandata)
-
-compute_scdd_sc(humandata2)
-
-compute_scdd_sc(mousedata)
-
-compute_scdd_sc(mousedata2)
-
-# save results
-saveRDS(humandata, datfile_h1)
-saveRDS(humandata2, datfile_h2)
-saveRDS(mousedata, datfile_m1)
-saveRDS(mousedata2, datfile_m2)
-```
-
-### MAST
-
-We run MAST on the TPMs following the approach in Soneson et al.
+We apply MAST to the TPMs following the approach in Soneson et al.
 
 ```{r, MAST}
-# function to add scDD p-values
+# function to add MAST p-values
 compute_MAST_sc <- function(dat){
@@
   }else{
-     message("scDD p-values already computed.")
+     message("MAST p-values already computed.")
   }
@@
 
-compute_MAST_sc(humandata2)
-
-compute_MAST_sc(mousedata)
-
-compute_MAST_sc(mousedata2)
-
 # save results
-saveRDS(humandata, datfile_h1)
-saveRDS(humandata2, datfile_h2)
-saveRDS(mousedata, datfile_m1)
-saveRDS(mousedata2, datfile_m2)
+saveRDS(humandata, datfile)
 ```
@@
 ## Covariate Diagnostics
-** Still need to clean up from here on; nothing below is eval'd**
 
-In scrnaseq data, strength of the signal can be a function of level of gene expression.
-We quantify it as median or mean gene expression,and detection rate, defined as the percentage of cells expressing the gene at a level of TPM>0.
+In scrnaseq data, strength of the signal can be a affected by the of level of gene expression. 
+We will explore two potential covariates related to expression level: mean expression 
+and detection rate (defined as the proportion of cells expressing the gene at a nonzero level). 
+These will also be added to the `rowData` slot of the `SummarizedExperiments`.
 
-```{r,cov, eval=FALSE}
-h_cov=compute_covariates_scrnaseq(paste0(datdir,"h_g_nofilter_subdata_GSE63818-GPL16791.rds"))
-m_cov=compute_covariates_scrnaseq(paste0(datdir,"m_g_nofilter_subdata_GSE45719.rds"))
-h2_cov=compute_covariates_scrnaseq(paste0(datdir,"h_g_nofilter_subdata_GSE84465.rds"))
-m2_cov=compute_covariates_scrnaseq(paste0(datdir,"m_g_nofilter_subdata_GSE94383.rds"))
+```{r,cov}
+add_covariates_scrnaseq <- function(dat){
+  rowData(dat)$meanExp <- rowMeans(assays(dat)[["count"]])
+  rowData(dat)$detection <- rowMeans(assays(dat)[["count"]]==0)
+  return(dat)
+}
+
+humandata <- add_covariates_scrnaseq(humandata)
+
+# save results
+saveRDS(humandata, datfile)
 ```
 
-First we will look at the covariates for the Wilcoxon text
+### Covariate one: Mean Expression
 
-```{r, wilcoxcov, fig.width=10, fig.height=7, eval=FALSE}
-print("human_plate_1")
-hwilcox1 <- plot_covariate_diagnostics(paste0(h_wilcox_dat1),"wilcox",h_cov)
-print(hwilcox1$plot)
+For each covariate, we'll examine the plots for each of the two analysis methods.
 
-print("human_plate_2")
-hwilcox2 <- plot_covariate_diagnostics(paste0(h_wilcox_dat2),"wilcox",h2_cov)
-print(hwilcox2$plot)
-
-print("mouse_plate_1")
-mwilcox1 <- plot_covariate_diagnostics(paste0(m_wilcox_dat1),"wilcox",m_cov)
-print(mwilcox1$plot)
-
-print("mouse_plate_2")
-mwilcox2 <- plot_covariate_diagnostics(paste0(m_wilcox_dat2),"wilcox",m2_cov)
-print(mwilcox2$plot)
+```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
+rank_scatter(data.frame(rowData(humandata)), 
+             pvalue="mast_p", covariate="meanExp") + 
+  ggtitle("MAST, Covariate 1: Mean Expression")
 ```
 
-Next, we look at the covariates for MAST
-
-```{r, mastcov, fig.width=10, fig.height=7, eval=FALSE}
-print("human_plate_1")
-hmast1 <- plot_covariate_diagnostics(paste0(h_mast_dat1),"mast",h_cov)
-print(hmast1$plot)
-
-print("human_plate_2")
-hmast2 <- plot_covariate_diagnostics(paste0(h_mast_dat2),"mast",h2_cov)
-print(hmast2$plot)
-
-print("mouse_plate_1")
-mmast1 <- plot_covariate_diagnostics(paste0(m_mast_dat1),"mast",m_cov)
-print(mmast1$plot)
-
-print("mouse_plate_2")
-mmast2 <- plot_covariate_diagnostics(paste0(m_mast_dat2),"mast",m2_cov)
-print(mmast2$plot)
+```{r, fig.width=10, fig.height=3.2, message=FALSE}
+strat_hist(data.frame(rowData(humandata)), 
+           pvalue="mast_p", covariate="meanExp", maxy=25) 
 ```
 
-Finally, we look at how informative the covariates are with results from ScDD.
+The mean expression appears to be informative and approximately satisfies the 
+assumptions
 
-```{r,scddcov, fig.width=10, fig.height=7, eval=FALSE}
-print("human_plate_1")
-hscdd1 <- plot_covariate_diagnostics(paste0(h_scdd_dat1),"scdd",h_cov)
-print(hscdd1$plot)
+### Covariate two: Detection Rate
 
-print("human_plate_2")
-hscdd2 <- plot_covariate_diagnostics(paste0(h_scdd_dat2),"scdd",h2_cov)
-print(hscdd2$plot)
+Next we look at the detection rate covariate.
 
-print("mouse_plate_1")
-mscdd1 <- plot_covariate_diagnostics(paste0(m_scdd_dat1),"scdd",m_cov)
-print(mscdd1$plot)
+```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
+rank_scatter(data.frame(rowData(humandata)), 
+             pvalue="mast_p", covariate="detection") + 
+  ggtitle("MAST, Covariate 2: Detection Rate")
+```
 
-print("mouse_plate_2")
-mscdd2 <- plot_covariate_diagnostics(paste0(m_scdd_dat2),"scdd",m2_cov)
-print(mscdd2$plot)
+```{r, fig.width=10, fig.height=3.2, message=FALSE}
+strat_hist(data.frame(rowData(humandata)), 
+           pvalue="mast_p", covariate="detection", maxy=25) 
 ```
@@
 
-First we will create the benchdesign object for wilcox -- IN PROGRESS
+** Still need to clean up from here on; nothing below is eval'd**
 
-```{r, benchmarking, eval=FALSE, eval=FALSE}
+```{r, benchmarking, eval=FALSE}
 

### 2018-02-27 datasets/scrnaseq/scrnaseq-human-scdd.Rmd
Organize sc analyses into one dataset/method per Rmd

--- 2018-02-27 datasets/scrnaseq/scrnaseq-human-mast.Rmd
@@
 ---
-title: "Case Study: scRNA-seq (Human data, MAST method)"
+title: "Case Study: scRNA-seq (Human data, scDD method)"
 author: "Ayshwarya Subramanian and Keegan Korthauer"
@@
 We also apply two different analysis methods: MAST and scDD. In this vignette, we will
-explore the human dataset analysed with MAST.
+explore the human dataset analysed with scDD.
 
@@
 
-We perform differential expression testing using 
-MAST which has a hurdle model to account for zeroes.
+We perform differential expression testing using scDD which implicitly models any heterogeneity in expression.
 
-We apply MAST to the TPMs following the approach in Soneson et al.
+We apply scDD to normalized log-transformed counts. 
+The p-value results are added to the `rowData` slot of the `SummarizedExperiment` 
+data objects to ensure that gene metadata stays intact.
 
-```{r, MAST}
-# function to add MAST p-values
-compute_MAST_sc <- function(dat){
+```{r, scdd}
+# function to add scDD p-values
+compute_scdd_sc <- function(dat){
   # check if already computed
-  if (! "mast_p" %in% colnames(rowData(dat))){
-    sca <- FromMatrix(exprsArray = log2(assays(dat)[["TPM"]] + 1), 
-                    cData = data.frame(wellKey=colnames(dat), colData(dat)))
-    zlmdata <- zlm.SingleCellAssay(~group, sca)
+  if (! "scdd_p" %in% colnames(rowData(dat))){
+    sce <- SingleCellExperiment(assays=list(count=assays(dat)[["count"]]),
+                                colData=data.frame(colData(dat)))
+    # scran normalization
+    libsize <- scran::computeSumFactors(assay(sce, "count"))
+    normcounts(sce) <- GetNormalizedMat(assay(sce, "count"), libsize)
 
-    rowData(dat)$mast_p <- 
-        lrTest(zlmdata, "group")[, "hurdle", "Pr(>Chisq)"]
+    rowData(dat)$scdd_p <- 
+        results(scDD(sce, 
+             condition = "group",
+             prior_param = list(alpha=0.01, mu0=0, s0=0.01, a0=0.01, b0=0.01), 
+             testZeroes = FALSE))$nonzero.pvalue
   }else{
-     message("MAST p-values already computed.")
+     message("scDD p-values already computed.")
   }
@@
 
-compute_MAST_sc(humandata)
+compute_scdd_sc(humandata)
 
 # save results
-saveRDS(humandata, datfile)
+saveRDS(humandata, datfile_h1)
 ```
@@
 # save results
-saveRDS(humandata, datfile)
+saveRDS(humandata, datfile_h1)
 ```
@@
 rank_scatter(data.frame(rowData(humandata)), 
-             pvalue="mast_p", covariate="meanExp") + 
-  ggtitle("MAST, Covariate 1: Mean Expression")
+             pvalue="scdd_p", covariate="meanExp") + 
+  ggtitle("scDD, Covariate 1: Mean Expression")
 ```
@@
 strat_hist(data.frame(rowData(humandata)), 
-           pvalue="mast_p", covariate="meanExp", maxy=25) 
+           pvalue="scdd_p", covariate="meanExp", maxy=25) 
 ```
@@
 rank_scatter(data.frame(rowData(humandata)), 
-             pvalue="mast_p", covariate="detection") + 
-  ggtitle("MAST, Covariate 2: Detection Rate")
+             pvalue="scdd_p", covariate="detection") + 
+  ggtitle("scDD, Covariate 2: Detection Rate")
 ```
@@
 strat_hist(data.frame(rowData(humandata)), 
-           pvalue="mast_p", covariate="detection", maxy=25) 
+           pvalue="scdd_p", covariate="detection", maxy=25) 
 ```
@@
 ** Still need to clean up from here on; nothing below is eval'd**
+
+First we will adjust for multiple comparisons using the scDD results
 

### 2018-02-27 datasets/scrnaseq/scrnaseq-mouse-mast.Rmd
Organize sc analyses into one dataset/method per Rmd

--- 2018-02-27 datasets/scrnaseq/scrnaseq-human-mast.Rmd
@@
 ---
-title: "Case Study: scRNA-seq (Human data, MAST method)"
+title: "Case Study: scRNA-seq (Mouse data, MAST method)"
 author: "Ayshwarya Subramanian and Keegan Korthauer"
@@
 We also apply two different analysis methods: MAST and scDD. In this vignette, we will
-explore the human dataset analysed with MAST.
+explore the mouse dataset analysed with MAST.
 
@@
 
-We first download the data in the data directory.`conquer` provides the processed datasets in convenient .rds files. First, we download the Human dataset:
-GSE84465.
+We first download the data in the data directory.`conquer` provides the processed datasets in convenient .rds files. We download the Mouse data from GSE94383
 
-```{r, humandata-download}
-download.file("http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE84465.rds",
-  destfile=paste0(datdir, "GSE84465.rds"))
-
+```{r, mousedata-download}
+download.file(url="http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE94383.rds",
+  destfile=paste0(datdir,"GSE94383.rds"))
 ```
@@
 
-We'll extract two conditions: Neoplastic and Regular cells, as
-defined by their tsne clusters.
+We'll extract two conditions: No stimulation versus LPS at 150 min.
 
 ```{r}
-datfile <- paste0(datdir, "human_GSE84465.rds")
+datfile <- paste0(datdir, "mouse_GSE94383.rds")
 
@@
   # load MultiAssayExperiment
-  humandata <- readRDS(paste0(datdir, "GSE84465.rds"))
+  mousedata=readRDS(paste0(datdir,"GSE94383.rds"))
 
   # subset by groups of interest
-  humandata <- humandata2[, (colData(humandata)$characteristics_ch1.7 %in% 
-                         c("neoplastic: Neoplastic", "neoplastic: Regular")) &
-                         (colData(humandata)$characteristics_ch1.5 %in% 
-                         c("tsne cluster: 11",  "tsne cluster: 8"))]
-  cdata=colData(humandata)
-  cdata$group <- cdata$characteristics_ch1.7
+  mousedata <- mousedata[, (colData(mousedata)$characteristics_ch1.2 ==
+                              "condition: No stimulation") |
+                           (colData(mousedata)$characteristics_ch1.2 ==
+                              "condition: LPS" &
+                            colData(mousedata)$characteristics_ch1.1 %in% 
+                              "time point: 150 min") ]
+  cdata=colData(mousedata)
+  cdata$group <- cdata$characteristics_ch1.2
 
   # subset by gene expression TPM & count assays
-  humandata <- experiments(humandata)[[1]]
-  assays(humandata) <- assays(humandata)[c("TPM", "count")]
-  colData(humandata) <- cdata
+  mousedata <- experiments(mousedata)[[1]]
+  assays(mousedata) <- assays(mousedata)[c("TPM", "count")]
+  colData(mousedata) <- cdata
 
   # subset by genes not all zero
-  humandata <- humandata[rowSums(assays(humandata)[["count"]]==0) < ncol(humandata),]
+  mousedata <- mousedata2[rowSums(assays(mousedata)[["count"]]==0) < ncol(mousedata),]
 
   # look at number of cells in each group
-  table(colData(humandata)$group)
+  table(as.character(colData(mousedata2)$group))
 
   # save SummarizedExperiment 
-  saveRDS(humandata, datfile)
+  saveRDS(mousedata, datfile)
 }else{
-  humandata <- readRDS(datfile)
+  mousedata <- readRDS(datfile) 
 }
@@
 
-For the human dataset, there are a total of `r nrow(humandata2)` cells and `r ncol(humandata2)` genes.
+For the mouse dataset2, we have a total of `r ncol(mousedata2)` cells and 
+`r nrow(mousedata2)` genes after filtering for non-detected genes.
 
@@
 
-compute_MAST_sc(humandata)
+compute_MAST_sc(mousedata)
 
 # save results
-saveRDS(humandata, datfile)
+saveRDS(mousedata, datfile)
 ```
@@
 
-humandata <- add_covariates_scrnaseq(humandata)
+mousedata <- add_covariates_scrnaseq(mousedata)
 
 # save results
-saveRDS(humandata, datfile)
+saveRDS(mousedata, datfile)
 ```
@@
 ```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
-rank_scatter(data.frame(rowData(humandata)), 
+rank_scatter(data.frame(rowData(mousedata)), 
              pvalue="mast_p", covariate="meanExp") + 
@@
 ```{r, fig.width=10, fig.height=3.2, message=FALSE}
-strat_hist(data.frame(rowData(humandata)), 
+strat_hist(data.frame(rowData(mousedata)), 
            pvalue="mast_p", covariate="meanExp", maxy=25) 
@@
 ```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
-rank_scatter(data.frame(rowData(humandata)), 
+rank_scatter(data.frame(rowData(mousedata)), 
              pvalue="mast_p", covariate="detection") + 
@@
 ```{r, fig.width=10, fig.height=3.2, message=FALSE}
-strat_hist(data.frame(rowData(humandata)), 
+strat_hist(data.frame(rowData(mousedata)), 
            pvalue="mast_p", covariate="detection", maxy=25) 

### 2018-02-27 datasets/scrnaseq/scrnaseq-mouse-scdd.Rmd
Organize sc analyses into one dataset/method per Rmd

--- 2018-02-27 datasets/scrnaseq/scrnaseq-mouse-mast.Rmd
@@
 ---
-title: "Case Study: scRNA-seq (Mouse data, MAST method)"
+title: "Case Study: scRNA-seq (Mouse data, scDD method)"
 author: "Ayshwarya Subramanian and Keegan Korthauer"
@@
 We also apply two different analysis methods: MAST and scDD. In this vignette, we will
-explore the mouse dataset analysed with MAST.
+explore the mouse dataset analysed with scDD.
 
@@
 
-We perform differential expression testing using 
-MAST which has a hurdle model to account for zeroes.
+We perform differential expression testing using scDD which implicitly models any heterogeneity in expression.
 
-We apply MAST to the TPMs following the approach in Soneson et al.
+### scDD: Differences in distributions of gene expression in the 2 conditions
 
-```{r, MAST}
-# function to add MAST p-values
-compute_MAST_sc <- function(dat){
+We apply scDD to normalized log-transformed counts. 
+The p-value results are added to the `rowData` slot of the `SummarizedExperiment` 
+data objects to ensure that gene metadata stays intact.
+
+```{r, scdd}
+# function to add scDD p-values
+compute_scdd_sc <- function(dat){
   # check if already computed
-  if (! "mast_p" %in% colnames(rowData(dat))){
-    sca <- FromMatrix(exprsArray = log2(assays(dat)[["TPM"]] + 1), 
-                    cData = data.frame(wellKey=colnames(dat), colData(dat)))
-    zlmdata <- zlm.SingleCellAssay(~group, sca)
+  if (! "scdd_p" %in% colnames(rowData(dat))){
+    sce <- SingleCellExperiment(assays=list(count=assays(dat)[["count"]]),
+                                colData=data.frame(colData(dat)))
+    # scran normalization
+    libsize <- scran::computeSumFactors(assay(sce, "count"))
+    normcounts(sce) <- GetNormalizedMat(assay(sce, "count"), libsize)
 
-    rowData(dat)$mast_p <- 
-        lrTest(zlmdata, "group")[, "hurdle", "Pr(>Chisq)"]
+    rowData(dat)$scdd_p <- 
+        results(scDD(sce, 
+             condition = "group",
+             prior_param = list(alpha=0.01, mu0=0, s0=0.01, a0=0.01, b0=0.01), 
+             testZeroes = FALSE))$nonzero.pvalue
   }else{
-     message("MAST p-values already computed.")
+     message("scDD p-values already computed.")
   }
@@
 
-compute_MAST_sc(mousedata)
+compute_scdd_sc(mousedata)
 
 # save results
-saveRDS(mousedata, datfile)
+saveRDS(mousedata, datfile_m1)
 ```
@@
 # save results
-saveRDS(mousedata, datfile)
+saveRDS(mousedata, datfile_m1)
 ```
@@
 
-For each covariate, we'll examine the plots for each of the two analysis methods.
+For each covariate, we'll examine the plots for each of the three analysis methods.
 
@@
 rank_scatter(data.frame(rowData(mousedata)), 
-             pvalue="mast_p", covariate="meanExp") + 
-  ggtitle("MAST, Covariate 1: Mean Expression")
+             pvalue="scdd_p", covariate="meanExp") + 
+  ggtitle("scDD, Covariate 1: Mean Expression")
 ```
@@
 strat_hist(data.frame(rowData(mousedata)), 
-           pvalue="mast_p", covariate="meanExp", maxy=25) 
+           pvalue="scdd_p", covariate="meanExp", maxy=25) 
 ```
@@
 rank_scatter(data.frame(rowData(mousedata)), 
-             pvalue="mast_p", covariate="detection") + 
-  ggtitle("MAST, Covariate 2: Detection Rate")
+             pvalue="scdd_p", covariate="detection") + 
+  ggtitle("scDD, Covariate 2: Detection Rate")
 ```
@@
 strat_hist(data.frame(rowData(mousedata)), 
-           pvalue="mast_p", covariate="detection", maxy=25) 
+           pvalue="scdd_p", covariate="detection", maxy=25) 
 ```

### 2018-02-27 datasets/scrnaseq/scrnaseq-knit.sh

@@ -1,0 +1,20 @@
+#!/bin/bash
+#SBATCH -J knit-sc4                  # A single job name for the array
+#SBATCH -n 6                         # Number of cores
+#SBATCH -N 1                         # All cores on one machine
+#SBATCH -p irizarry,serial_requeue   # Partition
+#SBATCH --mem 200G                   # Memory request 
+#SBATCH -t 0-20:00                   # Maximum execution time (D-HH:MM)
+#SBATCH -o render-%j.out
+#SBATCH -e render-%j.err
+ 
+module load rstudio
+export SLURM_NTASKS
+export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/rstudio/0.98.1103-fasrc01/bin/pandoc/"
+
+# change filename to Rmd to be knitted. ncores is passed in through environment
+# variable SLURM_NTASKS
+R -e "rmarkdown::render('scrnaseq-human-scdd.Rmd')"
+R -e "rmarkdown::render('scrnaseq-human-mast.Rmd')"
+R -e "rmarkdown::render('scrnaseq-mouse-scdd.Rmd')"
+R -e "rmarkdown::render('scrnaseq-mouse-mast.Rmd')"

### 2018-02-27 datasets/scrnaseq/scrnaseq-human-mast.Rmd
Fix typo in dataset name

@@
   # subset by groups of interest
-  humandata <- humandata2[, (colData(humandata)$characteristics_ch1.7 %in% 
+  humandata <- humandata[, (colData(humandata)$characteristics_ch1.7 %in% 
                          c("neoplastic: Neoplastic", "neoplastic: Regular")) &
@@
 
-For the human dataset, there are a total of `r nrow(humandata2)` cells and `r ncol(humandata2)` genes.
+For the human dataset, there are a total of `r nrow(humandata)` cells and `r ncol(humandata)` genes.
 

### 2018-02-27 datasets/scrnaseq/scrnaseq-mouse-mast.Rmd
Fix typo in dataset name

@@
   # subset by genes not all zero
-  mousedata <- mousedata2[rowSums(assays(mousedata)[["count"]]==0) < ncol(mousedata),]
+  mousedata <- mousedata[rowSums(assays(mousedata)[["count"]]==0) < ncol(mousedata),]
 
   # look at number of cells in each group
-  table(as.character(colData(mousedata2)$group))
+  table(as.character(colData(mousedata)$group))
 
@@
 
-For the mouse dataset2, we have a total of `r ncol(mousedata2)` cells and 
-`r nrow(mousedata2)` genes after filtering for non-detected genes.
+For the mouse dataset2, we have a total of `r ncol(mousedata)` cells and 
+`r nrow(mousedata)` genes after filtering for non-detected genes.
 

### 2018-02-27 datasets/scrnaseq/scrnaseq-human-mast.Rmd
Add summarized benchmark eval; standardize format

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE,cache=TRUE,cache.lazy = FALSE)
+knitr::opts_chunk$set(echo = TRUE)
 ```
@@
 We selected two datasets, one each from mouse and human. 
-These datasets come from the `Conquer` (Soneson et al., 2017) scRNAseq data collection.
+These datasets come from the `Conquer` (Soneson et al., 2018) scRNAseq data collection.
 `Conquer` provides uniformly processed gene expression data summarized to both count (aggregated from transcript members), and length-scaled TPMs at the gene level.
@@
 
+## Data download
+
 We first download the data in the data directory.`conquer` provides the processed datasets in convenient .rds files. First, we download the Human dataset:
@@
 We do not filter genes as filtering may affect the subsequent false discovery control steps.
+We only filter out genes that are expressed in
+fewer than 1% of cells. 
+We will filter out cells with extremely low mapping rate
+(below 20%), as these represent poor quality cells that may end up having negative
+size factor estimates during normalization. For the same reason, we also filter cells
+which have extremely low proportion of genes detected (less than 5%).
 
@@
 ```{r}
-datfile <- paste0(datdir, "human_GSE84465.rds")
+datfile <- paste0(datdir, "human_GSE84465_mast.rds")
 
@@
   humandata <- readRDS(paste0(datdir, "GSE84465.rds"))
+
+  # exclude cells with extremely low mapping or feature rate (poor quality cells)
+  lowMap <- which(metadata(humandata)$salmon_summary$percent_mapped < 20 |
+                  colMeans(assay(experiments(humandata)[[1]], "count") > 0) < 0.05)
+  if (length(lowMap) > 0){
+    humandata <- humandata[, -lowMap]
+    message("Removed ", length(lowMap), " cells with mapping rate below 20%",
+            " or fewer than 5% of genes detected.")
+  }
 
@@
 
-  # subset by genes not all zero
-  humandata <- humandata[rowSums(assays(humandata)[["count"]]==0) < ncol(humandata),]
+  # subset by genes with detection rate > 1%
+  humandata <- humandata[rowMeans(assay(humandata, "count") > 0) > 0.01, ]
 
-  # look at number of cells in each group
-  table(colData(humandata)$group)
+  # remove spike-in controls 
+  humandata <- humandata[!grepl("ERCC-", rownames(humandata)),]
 
@@
 }
+
+# look at number of cells in each group
+table(colData(humandata)$group)
 ```
 
-For the human dataset, there are a total of `r nrow(humandata)` cells and `r ncol(humandata)` genes.
+For the human dataset, there are a total of `r ncol(humandata)` cells and `r nrow(humandata)` genes.
 
-## Differential expression testing
+# Data analysis
+
+## Differential testing
 
@@
   if (! "mast_p" %in% colnames(rowData(dat))){
-    sca <- FromMatrix(exprsArray = log2(assays(dat)[["TPM"]] + 1), 
+    sca <- FromMatrix(exprsArray = log2(assay(dat, "TPM") + 1), 
                     cData = data.frame(wellKey=colnames(dat), colData(dat)))
-    zlmdata <- zlm.SingleCellAssay(~group, sca)
+    zlmdata <- zlm(~group, sca)
 
@@
 
-compute_MAST_sc(humandata)
+humandata <- compute_MAST_sc(humandata)
 
@@
 add_covariates_scrnaseq <- function(dat){
-  rowData(dat)$meanExp <- rowMeans(assays(dat)[["count"]])
-  rowData(dat)$detection <- rowMeans(assays(dat)[["count"]]==0)
+  rowData(dat)$meanExp <- rowMeans(assay(dat, "count"))
+  rowData(dat)$detection <- rowMeans(assay(dat, "count") > 0)
   return(dat)
@@
 
-For each covariate, we'll examine the plots for each of the two analysis methods.
+For each covariate, we'll examine the covariate diagnostic plots.
 
@@
 strat_hist(data.frame(rowData(humandata)), 
-           pvalue="mast_p", covariate="meanExp", maxy=25) 
+           pvalue="mast_p", covariate="meanExp", maxy=35,
+           main = "MAST, Covariate 1: Mean Expression")
 ```
@@
 strat_hist(data.frame(rowData(humandata)), 
-           pvalue="mast_p", covariate="detection", maxy=25) 
+           pvalue="mast_p", covariate="detection", maxy=35,
+           main = "MAST, Covariate 2: Detection Rate") 
 ```
 
-### Benchmarking FDR methods
+## Multiple-Testing Correction
 
-** Still need to clean up from here on; nothing below is eval'd**
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object.
 
-```{r, benchmarking, eval=FALSE}
+```{r, benchdesign, message=FALSE}
+bd <- initializeBenchDesign()
+```
 
-bd <- initializeBenchDesign()
-resN <- paste0(resdir, "scrnaseq_wilcox_summarizedBenchmark_",nrow(m_scdd_cov), ".RData")
+Now, we're ready to construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+                                        helper scripts).
 
-duration <- NA
-if (!file.exists(resN)){
-  t1 <- proc.time()
-  scrna_sb <- bd %>% buildBench(data=(hscdd1$data%>%select(pval,mean_exp)%>%mutate(effect_size=0, test_statistic=0, SE=0) %>% mutate(ind_covariate=mean_exp)), 
-                          ftCols = c("mean_exp"),
-                          parallel=TRUE, BPPARAM=multicoreParam)
+### Covariate one: Mean expression
 
-  save(scrna_sb, file=resN)
-  duration <- round((proc.time()-t1)[3]/60,1)
+First we'll include the mean expression covariate.
+
+```{r, sb1, results="hide", message=FALSE}
+res1 <- paste0(resdir, "sb_GSE84465_mast_mean.rds")
+
+if (!file.exists(res1)){
+  sb1 <- bd %>% buildBench(data.frame(rowData(humandata)) %>% 
+                                   na.omit() %>%
+                                   mutate(pval=mast_p,
+                                          ind_covariate=meanExp), 
+                          ftCols = c("meanExp"),
+                          parallel=TRUE)
+
+  saveRDS(sb1, file=res1)
 }else{
-  load(resN)
+  sb1 <- readRDS(res1)
 }
+```
 
-assayNames(scrna_sb) <- "qvalue"
-scrna_sb <- addDefaultMetrics(scrna_sb)
+### Covariate two: Detection Rate
 
+Now, we'll repeat the multiple testing correction using the other covariate (detection rate):
+
+```{r, sb2, results="hide", message=FALSE}
+res2 <- paste0(resdir, "sb_GSE84465_mast_det.rds")
+
+if (!file.exists(res2)){
+  sb2 <- bd %>% buildBench(data.frame(rowData(humandata)) %>% 
+                                   na.omit() %>%
+                                   mutate(pval=mast_p,
+                                          ind_covariate=detection), 
+                          ftCols = c("detection"),
+                          parallel=TRUE)
+
+  saveRDS(sb2, file=res2)
+}else{
+  sb2 <- readRDS(res2)
+}
+```
+
+## Benchmark Metrics
+
+Next, we'll add the default performance metric for q-value assays and 
+plot the results. We'll start with covariate one.
+
+### Covariate one: Mean Expression
+
+First, we have
+to rename the assay to 'qvalue'.
+
+```{r, metrics1}
+# rename assay to qvalue
+assayNames(sb1) <- "qvalue"
+sb1 <- addDefaultMetrics(sb1)
+```
+
+Now, we'll plot the results.
+
+```{r, plot1, results="hide"}
 # plot nrejects by method overall and stratified by covariate
-rejections_scatter(scrna_sb,supplementary=FALSE)
+rejections_scatter(sb1,
+                   supplementary=FALSE) +
+  ggtitle("MAST, Covariate 1: Mean Expression")
 
-rejection_scatter_bins(scrna_sb, covariate="ind_covar_N", bins=3,supplementary=FALSE)
+rejection_scatter_bins(sb1, covariate="meanExp", bins=4,
+                       supplementary=FALSE) +
+  ggtitle("MAST, Covariate 1: Mean Expression")
 
 # upset plot 
-plotFDRMethodsOverlap(scrna_sb, 
-                      alpha=0.05, nsets=ncol(scrna_sb),
+plotFDRMethodsOverlap(sb1, 
+                      alpha=0.05, nsets=ncol(sb1),
                       order.by="freq", decreasing=TRUE,
@@
 # properties of method-exclusive discoveries
-methods <- c("lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
-mcols(scrna_sb)$ind_covariate <- mcols(scrna_sb)$ind_covar_N
-plotCovariateBoxplots(scrna_sb, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3)
+methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
+             "scott-empirical", "scott-theoretical")
+mcols(sb1)$ind_covariate <- mcols(sb1)$meanExp
+plotCovariateBoxplots(sb1, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3) +
+  ggtitle("MAST, Covariate 1: Mean Expression")
+
 ```
+
+### Covariate two: Detection Rate
+
+Next, we'll look at the performance metrics for the other covariate (detection rate).
+
+```{r, metrics2}
+# rename assay to qvalue
+assayNames(sb2) <- "qvalue"
+sb2 <- addDefaultMetrics(sb2)
+```
+
+Now, we'll plot the results.
+
+```{r, plot2, results="hide"}
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter(sb2, supplementary=FALSE) +
+  ggtitle("MAST, Covariate 2: Detection Rate")
+
+rejection_scatter_bins(sb2, covariate="detection", bins=4,
+                       supplementary=FALSE) +
+  ggtitle("MAST, Covariate 2: Detection Rate")
+
+# upset plot 
+plotFDRMethodsOverlap(sb2, 
+                      alpha=0.05, nsets=ncol(sb2),
+                      order.by="freq", decreasing=TRUE,
+                      supplementary=FALSE)
+
+# properties of method-exclusive discoveries
+methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
+             "scott-empirical", "scott-theoretical")
+mcols(sb2)$ind_covariate <- mcols(sb2)$detection
+plotCovariateBoxplots(sb2, alpha=0.1, nsets=6, methods=methods, maxNum=3)+
+  ggtitle("MAST, Covariate 2: Detection Rate")
+
+```
+
+# Session Information
+
+```{r}
+sessionInfo()
+```

### 2018-02-27 datasets/scrnaseq/scrnaseq-human-scdd.Rmd
Add summarized benchmark eval; standardize format

@@
 
-cores <- 6
+cores <- 4
 BiocParallel::register(MulticoreParam(workers = cores))
@@
 # Data preparation
+
+## Data download
 
@@
 ```{r}
-datfile <- paste0(datdir, "human_GSE84465.rds")
+datfile <- paste0(datdir, "human_GSE84465_scdd.rds")
 
@@
 
-  # subset by genes not all zero
-  humandata <- humandata[rowSums(assays(humandata)[["count"]]==0) < ncol(humandata),]
+  # subset by genes with detection rate > 1%
+  humandata <- humandata[rowMeans(assay(humandata, "count") > 0) > 0.01, ]
 
-  # look at number of cells in each group
-  table(colData(humandata)$group)
+  # remove spike-in controls
+  humandata <- humandata[!grepl("ERCC-", rownames(humandata)),]
 
@@
   if (! "scdd_p" %in% colnames(rowData(dat))){
-    sce <- SingleCellExperiment(assays=list(count=assays(dat)[["count"]]),
+    sce <- SingleCellExperiment(assays=list(count=assay(dat, "count")),
                                 colData=data.frame(colData(dat)))
@@
 
-    rowData(dat)$scdd_p <- 
-        results(scDD(sce, 
+    res <- results(scDD(sce, 
              condition = "group",
              prior_param = list(alpha=0.01, mu0=0, s0=0.01, a0=0.01, b0=0.01), 
-             testZeroes = FALSE))$nonzero.pvalue
+             testZeroes = TRUE))
+    rowData(dat)$scdd_p <- res$nonzero.pvalue
+    rowData(dat)$scdd_p_combined <- res$combined.pvalue
   }else{
@@
 
-compute_scdd_sc(humandata)
+humandata <- compute_scdd_sc(humandata)
 
 # save results
-saveRDS(humandata, datfile_h1)
+saveRDS(humandata, datfile)
 ```
@@
 # save results
-saveRDS(humandata, datfile_h1)
+saveRDS(humandata, datfile)
 ```
@@
 strat_hist(data.frame(rowData(humandata)), 
-           pvalue="scdd_p", covariate="meanExp", maxy=25) 
+           pvalue="scdd_p", covariate="meanExp", maxy=35,
+           main = "scDD, Covariate 1: Mean Expression") 
 ```
@@
 strat_hist(data.frame(rowData(humandata)), 
-           pvalue="scdd_p", covariate="detection", maxy=25) 
+           pvalue="scdd_p", covariate="detection", maxy=35,
+           main = "scDD, Covariate 2: Detection Rate") 
+```
+## Multiple-Testing Correction
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object.
+
+```{r, benchdesign, message=FALSE}
+bd <- initializeBenchDesign()
 ```
 
-### Benchmarking FDR methods
+Now, we're ready to construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+                                        helper scripts).
 
-** Still need to clean up from here on; nothing below is eval'd**
+### Covariate one: Mean expression
 
-First we will adjust for multiple comparisons using the scDD results
+First we'll include the mean expression covariate.
 
-```{r, benchmarking, eval=FALSE}
+```{r, sb1, results="hide", message=FALSE}
+res1 <- paste0(resdir, "sb_GSE84465_scdd_mean.rds")
 
-bd <- initializeBenchDesign()
-resN <- paste0(resdir, "scrnaseq_wilcox_summarizedBenchmark_",nrow(m_scdd_cov), ".RData")
+if (!file.exists(res1)){
+  sb1 <- bd %>% buildBench(data.frame(rowData(humandata)) %>% 
+                                   na.omit() %>%
+                                   mutate(pval=scdd_p,
+                                          ind_covariate=meanExp), 
+                          ftCols = c("meanExp"),
+                          parallel=TRUE)
 
-duration <- NA
-if (!file.exists(resN)){
-  t1 <- proc.time()
-  scrna_sb <- bd %>% buildBench(data=(hscdd1$data%>%select(pval,mean_exp)%>%mutate(effect_size=0, test_statistic=0, SE=0) %>% mutate(ind_covariate=mean_exp)), 
-                          ftCols = c("mean_exp"),
-                          parallel=TRUE, BPPARAM=multicoreParam)
+  saveRDS(sb1, file=res1)
+}else{
+  sb1 <- readRDS(res1)
+}
+```
 
-  save(scrna_sb, file=resN)
-  duration <- round((proc.time()-t1)[3]/60,1)
+### Covariate two: Detection Rate
+
+Now, we'll repeat the multiple testing correction using the other covariate (detection rate):
+
+```{r, sb2, results="hide", message=FALSE}
+res2 <- paste0(resdir, "sb_GSE84465_scdd_det.rds")
+
+if (!file.exists(res2)){
+  sb2 <- bd %>% buildBench(data.frame(rowData(humandata)) %>% 
+                                   na.omit() %>%
+                                   mutate(pval=scdd_p,
+                                          ind_covariate=detection), 
+                          ftCols = c("detection"),
+                          parallel=TRUE)
+
+  saveRDS(sb2, file=res2)
 }else{
-  load(resN)
+  sb2 <- readRDS(res2)
 }
+```
 
-assayNames(scrna_sb) <- "qvalue"
-scrna_sb <- addDefaultMetrics(scrna_sb)
+## Benchmark Metrics
 
+Next, we'll add the default performance metric for q-value assays and 
+plot the results. We'll start with covariate one.
+
+### Covariate one: Mean Expression
+
+First, we have
+to rename the assay to 'qvalue'.
+
+```{r, metrics1}
+# rename assay to qvalue
+assayNames(sb1) <- "qvalue"
+sb1 <- addDefaultMetrics(sb1)
+```
+
+Now, we'll plot the results.
+
+```{r, plot1, results="hide"}
 # plot nrejects by method overall and stratified by covariate
-rejections_scatter(scrna_sb,supplementary=FALSE)
+rejections_scatter(sb1,
+                   supplementary=FALSE) +
+  ggtitle("scDD, Covariate 1: Mean Expression")
 
-rejection_scatter_bins(scrna_sb, covariate="ind_covar_N", bins=3,supplementary=FALSE)
+rejection_scatter_bins(sb1, covariate="meanExp", bins=4,
+                       supplementary=FALSE) +
+  ggtitle("scDD, Covariate 1: Mean Expression")
 
 # upset plot 
-plotFDRMethodsOverlap(scrna_sb, 
-                      alpha=0.05, nsets=ncol(scrna_sb),
+plotFDRMethodsOverlap(sb1, 
+                      alpha=0.05, nsets=ncol(sb1),
                       order.by="freq", decreasing=TRUE,
@@
 # properties of method-exclusive discoveries
-methods <- c("lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
-mcols(scrna_sb)$ind_covariate <- mcols(scrna_sb)$ind_covar_N
-plotCovariateBoxplots(scrna_sb, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3)
+methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
+             "scott-empirical", "scott-theoretical")
+mcols(sb1)$ind_covariate <- mcols(sb1)$meanExp
+plotCovariateBoxplots(sb1, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3) +
+  ggtitle("scDD, Covariate 1: Mean Expression")
+
 ```
+
+### Covariate two: Detection Rate
+
+Next, we'll look at the performance metrics for the other covariate (detection rate).
+
+```{r, metrics2}
+# rename assay to qvalue
+assayNames(sb2) <- "qvalue"
+sb2 <- addDefaultMetrics(sb2)
+```
+
+Now, we'll plot the results.
+
+```{r, plot2, results="hide"}
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter(sb2, supplementary=FALSE) +
+  ggtitle("scDD, Covariate 2: Detection Rate")
+
+rejection_scatter_bins(sb2, covariate="detection", bins=4,
+                       supplementary=FALSE) +
+  ggtitle("scDD, Covariate 2: Detection Rate")
+
+# upset plot 
+plotFDRMethodsOverlap(sb2, 
+                      alpha=0.05, nsets=ncol(sb2),
+                      order.by="freq", decreasing=TRUE,
+                      supplementary=FALSE)
+
+# properties of method-exclusive discoveries
+methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
+             "scott-empirical", "scott-theoretical")
+mcols(sb2)$ind_covariate <- mcols(sb2)$detection
+plotCovariateBoxplots(sb2, alpha=0.1, nsets=6, methods=methods, maxNum=3)+
+  ggtitle("scDD, Covariate 2: Detection Rate")
+
+```
+
+# Session Information
+
+```{r}
+sessionInfo()
+```

### 2018-02-27 datasets/scrnaseq/scrnaseq-mouse-mast.Rmd
Add summarized benchmark eval; standardize format

@@
 
-We first download the data in the data directory.`conquer` provides the processed datasets in convenient .rds files. We download the Mouse data from GSE94383
+## Data download
+
+We first download the data in the data directory. `conquer` provides the processed datasets in convenient .rds files. We download the Mouse data from GSE94383
 
@@
 ```{r}
-datfile <- paste0(datdir, "mouse_GSE94383.rds")
+datfile <- paste0(datdir, "mouse_GSE94383_mast.rds")
 
@@
   mousedata=readRDS(paste0(datdir,"GSE94383.rds"))
+
+  # exclude cells with extremely low mapping or feature rate (poor quality cells)
+  lowMap <- which(metadata(mousedata)$salmon_summary$percent_mapped < 20 |
+                  colMeans(assay(experiments(mousedata)[[1]], "count") > 0) < 0.05)
+  if (length(lowMap) > 0){
+    mousedata <- mousedata[, -lowMap]
+    message("Removed ", length(lowMap), " cells with mapping rate below 20%",
+            " or fewer than 5% of genes detected.")
+  }
 
@@
 
-  # subset by genes not all zero
-  mousedata <- mousedata[rowSums(assays(mousedata)[["count"]]==0) < ncol(mousedata),]
+  # subset by genes with detection rate > 1%
+  mousedata <- mousedata[rowMeans(assay(mousedata, "count") > 0) > 0.01, ]
 
-  # look at number of cells in each group
-  table(as.character(colData(mousedata)$group))
+  # remove spike-in controls
+  mousedata <- mousedata[!grepl("ERCC-", rownames(mousedata)),]
 
@@
 }
+
+# look at number of cells in each group
+table(as.character(colData(mousedata)$group))
 ```
@@
 For the mouse dataset2, we have a total of `r ncol(mousedata)` cells and 
-`r nrow(mousedata)` genes after filtering for non-detected genes.
+`r nrow(mousedata)` genes.
 
-## Differential expression testing
+# Data analysis
+
+## Differential testing
 
@@
 
-compute_MAST_sc(mousedata)
+mousedata <- compute_MAST_sc(mousedata)
 
@@
 strat_hist(data.frame(rowData(mousedata)),
-           pvalue="mast_p", covariate="meanExp", maxy=25) 
+           pvalue="mast_p", covariate="meanExp", maxy=16,
+           main = "MAST, Covariate 1: Mean Expression")
 ```
@@
 strat_hist(data.frame(rowData(mousedata)), 
-           pvalue="mast_p", covariate="detection", maxy=25) 
+           pvalue="mast_p", covariate="detection", maxy=16,
+           main = "MAST, Covariate 2: Detection Rate")
 ```
 
-### Benchmarking FDR methods
+## Multiple-Testing Correction
 
-** Still need to clean up from here on; nothing below is eval'd**
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object.
 
-```{r, benchmarking, eval=FALSE}
+```{r, benchdesign, message=FALSE}
+bd <- initializeBenchDesign()
+```
 
-bd <- initializeBenchDesign()
-resN <- paste0(resdir, "scrnaseq_wilcox_summarizedBenchmark_",nrow(m_scdd_cov), ".RData")
+Now, we're ready to construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+                                        helper scripts).
 
-duration <- NA
-if (!file.exists(resN)){
-  t1 <- proc.time()
-  scrna_sb <- bd %>% buildBench(data=(hscdd1$data%>%select(pval,mean_exp)%>%mutate(effect_size=0, test_statistic=0, SE=0) %>% mutate(ind_covariate=mean_exp)), 
-                          ftCols = c("mean_exp"),
-                          parallel=TRUE, BPPARAM=multicoreParam)
+### Covariate one: Mean expression
 
-  save(scrna_sb, file=resN)
-  duration <- round((proc.time()-t1)[3]/60,1)
+First we'll include the mean expression covariate.
+
+```{r, sb1, results="hide", message=FALSE}
+res1 <- paste0(resdir, "sb_GSE94383_mast_mean.rds")
+
+if (!file.exists(res1)){
+  sb1 <- bd %>% buildBench(data.frame(rowData(mousedata)) %>% 
+                                   na.omit() %>%
+                                   mutate(pval=mast_p,
+                                          ind_covariate=meanExp), 
+                          ftCols = c("meanExp"),
+                          parallel=TRUE)
+
+  saveRDS(sb1, file=res1)
 }else{
-  load(resN)
+  sb1 <- readRDS(res1)
 }
+```
 
-assayNames(scrna_sb) <- "qvalue"
-scrna_sb <- addDefaultMetrics(scrna_sb)
+### Covariate two: Detection Rate
 
+Now, we'll repeat the multiple testing correction using the other covariate (detection rate):
+
+```{r, sb2, results="hide", message=FALSE}
+res2 <- paste0(resdir, "sb_GSE94383_mast_det.rds")
+
+if (!file.exists(res2)){
+  sb2 <- bd %>% buildBench(data.frame(rowData(mousedata)) %>% 
+                                   na.omit() %>%
+                                   mutate(pval=mast_p,
+                                          ind_covariate=detection), 
+                          ftCols = c("detection"),
+                          parallel=TRUE)
+
+  saveRDS(sb2, file=res2)
+}else{
+  sb2 <- readRDS(res2)
+}
+```
+
+## Benchmark Metrics
+
+Next, we'll add the default performance metric for q-value assays and 
+plot the results. We'll start with covariate one.
+
+### Covariate one: Mean Expression
+
+First, we have
+to rename the assay to 'qvalue'.
+
+```{r, metrics1}
+# rename assay to qvalue
+assayNames(sb1) <- "qvalue"
+sb1 <- addDefaultMetrics(sb1)
+```
+
+Now, we'll plot the results.
+
+```{r, plot1, results="hide"}
 # plot nrejects by method overall and stratified by covariate
-rejections_scatter(scrna_sb,supplementary=FALSE)
+rejections_scatter(sb1,
+                   supplementary=FALSE) +
+  ggtitle("MAST, Covariate 1: Mean Expression")
 
-rejection_scatter_bins(scrna_sb, covariate="ind_covar_N", bins=3,supplementary=FALSE)
+rejection_scatter_bins(sb1, covariate="meanExp", bins=4,
+                       supplementary=FALSE) +
+  ggtitle("MAST, Covariate 1: Mean Expression")
 
 # upset plot 
-plotFDRMethodsOverlap(scrna_sb, 
-                      alpha=0.05, nsets=ncol(scrna_sb),
+plotFDRMethodsOverlap(sb1, 
+                      alpha=0.05, nsets=ncol(sb1),
                       order.by="freq", decreasing=TRUE,

### 2018-02-27 datasets/scrnaseq/scrnaseq-mouse-scdd.Rmd
Add summarized benchmark eval; standardize format

@@
 
-For data processing, we follow practices and recommendations from Soneson et al., 2017.
-Briefly, we limit ourselves to 2-group comparisons.
-We do not filter genes as filtering may affect the subsequent false discovery control steps.
+For data processing, we follow practices and recommendations from Soneson et al., 2018. Briefly, we limit ourselves to 2-group comparisons. We only filter out genes that are expressed in 
+fewer than 1% of cells. 
+We will filter out cells with extremely low mapping rate
+(below 20%), as these represent poor quality cells that may end up having negative
+size factor estimates during normalization. For the same reason, we also filter cells
+which have extremely low proportion of genes detected (less than 5%).
 
@@
 ```{r}
-datfile <- paste0(datdir, "mouse_GSE94383.rds")
+datfile <- paste0(datdir, "mouse_GSE94383_scdd.rds")
 
@@
   mousedata=readRDS(paste0(datdir,"GSE94383.rds"))
+
+  # exclude cells with extremely low mapping rate or too few expressed features
+  # (poor quality cells)
+
+  # exclude cells with extremely low mapping or feature rate (poor quality cells)
+  lowMap <- which(metadata(mousedata)$salmon_summary$percent_mapped < 20 |
+                  colMeans(assay(experiments(mousedata)[[1]], "count") > 0) < 0.05)
+  if (length(lowMap) > 0){
+    mousedata <- mousedata[, -lowMap]
+    message("Removed ", length(lowMap), " cells with mapping rate below 20%",
+            " or fewer than 5% of genes detected.")
+  }
 
@@
 For the mouse dataset2, we have a total of `r ncol(mousedata)` cells and 
-`r nrow(mousedata)` genes after filtering for non-detected genes.
+`r nrow(mousedata)` genes.
 
-## Differential expression testing
+# Data analysis
+
+## Differential testing
 
 We perform differential expression testing using scDD which implicitly models any heterogeneity in expression.
-
-### scDD: Differences in distributions of gene expression in the 2 conditions
 
@@
 
-compute_scdd_sc(mousedata)
+mousedata <- compute_scdd_sc(mousedata)
 
 # save results
-saveRDS(mousedata, datfile_m1)
+saveRDS(mousedata, datfile)
 ```
@@
 # save results
-saveRDS(mousedata, datfile_m1)
+saveRDS(mousedata, datfile)
 ```
@@
 
-For each covariate, we'll examine the plots for each of the three analysis methods.
+For each covariate, we'll examine the covariate diagnostic plots.
 
@@
 strat_hist(data.frame(rowData(mousedata)), 
-           pvalue="scdd_p", covariate="meanExp", maxy=25) 
+           pvalue="scdd_p", covariate="meanExp", maxy=16,
+           main = "scDD, Covariate 1: Mean Expression") 
 ```
@@
 strat_hist(data.frame(rowData(mousedata)), 
-           pvalue="scdd_p", covariate="detection", maxy=25) 
+           pvalue="scdd_p", covariate="detection", maxy=16,
+           main = "scDD, Covariate 2: Detection Rate") 
 ```
 
-### Benchmarking FDR methods
+## Multiple-Testing Correction
 
-** Still need to clean up from here on; nothing below is eval'd**
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object.
 
-First we will adjust for multiple comparisons using the scDD results
+```{r, benchdesign, message=FALSE}
+bd <- initializeBenchDesign()
+```
 
-```{r, benchmarking, eval=FALSE}
+Now, we're ready to construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+                                        helper scripts).
 
-bd <- initializeBenchDesign()
-resN <- paste0(resdir, "scrnaseq_wilcox_summarizedBenchmark_",nrow(m_scdd_cov), ".RData")
+### Covariate one: Mean expression
 
-duration <- NA
-if (!file.exists(resN)){
-  t1 <- proc.time()
-  scrna_sb <- bd %>% buildBench(data=(hscdd1$data%>%select(pval,mean_exp)%>%mutate(effect_size=0, test_statistic=0, SE=0) %>% mutate(ind_covariate=mean_exp)), 
-                          ftCols = c("mean_exp"),
-                          parallel=TRUE, BPPARAM=multicoreParam)
+First we'll include the mean expression covariate.
 
-  save(scrna_sb, file=resN)
-  duration <- round((proc.time()-t1)[3]/60,1)
+```{r, sb1, results="hide", message=FALSE}
+res1 <- paste0(resdir, "sb_GSE94383_scdd_mean.rds")
+
+if (!file.exists(res1)){
+  sb1 <- bd %>% buildBench(data.frame(rowData(mousedata)) %>% 
+                                   na.omit() %>%
+                                   mutate(pval=scdd_p,
+                                          ind_covariate=meanExp), 
+                          ftCols = c("meanExp"),
+                          parallel=TRUE)
+
+  saveRDS(sb1, file=res1)
 }else{
-  load(resN)
+  sb1 <- readRDS(res1)
 }
+```
 
-assayNames(scrna_sb) <- "qvalue"
-scrna_sb <- addDefaultMetrics(scrna_sb)
+### Covariate two: Detection Rate
 
+Now, we'll repeat the multiple testing correction using the other covariate (detection rate):
+
+```{r, sb2, results="hide", message=FALSE}
+res2 <- paste0(resdir, "sb_GSE94383_scdd_det.rds")
+
+if (!file.exists(res2)){
+  sb2 <- bd %>% buildBench(data.frame(rowData(mousedata)) %>% 
+                                   na.omit() %>%
+                                   mutate(pval=scdd_p,
+                                          ind_covariate=detection), 
+                          ftCols = c("detection"),
+                          parallel=TRUE)
+
+  saveRDS(sb2, file=res2)
+}else{
+  sb2 <- readRDS(res2)
+}
+```
+
+## Benchmark Metrics
+
+Next, we'll add the default performance metric for q-value assays and 
+plot the results. We'll start with covariate one.
+
+### Covariate one: Mean Expression
+
+First, we have
+to rename the assay to 'qvalue'.
+
+```{r, metrics1}
+# rename assay to qvalue
+assayNames(sb1) <- "qvalue"
+sb1 <- addDefaultMetrics(sb1)
+```
+
+Now, we'll plot the results.
+
+```{r, plot1, results="hide"}
 # plot nrejects by method overall and stratified by covariate
-rejections_scatter(scrna_sb,supplementary=FALSE)
+rejections_scatter(sb1,
+                   supplementary=FALSE) +
+  ggtitle("scDD, Covariate 1: Mean Expression")
 
-rejection_scatter_bins(scrna_sb, covariate="ind_covar_N", bins=3,supplementary=FALSE)
+rejection_scatter_bins(sb1, covariate="meanExp", bins=4,
+                       supplementary=FALSE) +
+  ggtitle("scDD, Covariate 1: Mean Expression")
 
 # upset plot 
-plotFDRMethodsOverlap(scrna_sb, 
-                      alpha=0.05, nsets=ncol(scrna_sb),
+plotFDRMethodsOverlap(sb1, 
+                      alpha=0.05, nsets=ncol(sb1),
                       order.by="freq", decreasing=TRUE,

### 2018-02-27 datasets/ChIP-seq/ChIPseq-peaks.Rmd

--- 2018-02-22 datasets/ChIP-seq/ChIP-seq.Rmd
@@
 ---
-title: "Case Study: ChIP-seq Differential Peak Calling"
-author: "Mingxiang Teng"
+title: "Case Study: ChIP-seq Differential Peak Testing (de novo regions)"
+author: "Patrick Kimes, Mingxiang Teng"
 date: '`r format(Sys.Date(), "%B %d, %Y")`'
@@
 
-Common analyses in ChIP-seq data include the calling of peaks within individual
-samples, and the testing of differential peaks across conditions. While peak
-calling continues to be the most wide-spread method applied to ChIP-seq data,
-differential testing has gained popularity with the increased availability of
-data from large-scale experiments with more complex designs.
+In this case study, we perform differential peak calling on ChIP-seq data for histone H3K4Me3
+for samples from two cell lines (K562 and Gm12878) using publicly available data generated
+by the ENCODE Project. Peak calls were generated by the ENCODE Project pipeline.
+Differential testing was performed using both _DiffBind_ (Stark and Brown, 2011) and
+_csaw_ (Lun and Smyth, 2016).
 
-In differential peak calling, classical pipelines often begin by first calling peaks
-in samples, followed by performing differential testing using existing algorithms
-developed for other genomic data tpyes (e.g. RNA-seq). However, with these approaches,
-testing is only performed on a filtered set of genomic regions, namely those with
-significance in at least in one condition, resulting in only a limited number of
-null tests cases.
-
-To avoid this limitation, we examine ChIP-seq data of the histone H3K4Me3, an active
-marker of gene expression, and focus our analysis on promoter regions of all genes
-for this histone marker. By doing so, we avoid limiting the set of tests to only
-regions significant in at least one sample, while also keeping most regions of
-interest for this histone marker.
-
-Here, we focus on test differential peaks between two cell lines (K562 and Gm12878)
-using publicly available data generated by the ENCODE Project. These two cell lines
-comprise the largest sequencing datasets in the ENCODE Project. As mentioned above,
-we use ChIP-seq data for histone H3K4Me3 in our analysis. Differential peaks are tested
-using DESeq2.
-
-We note that while the algorithm for differential testing, DESeq2, used here is the same
-as that used in our analysis of RNA-seq data, benchmarking on ChIP-seq still provides
-insights for a unique data structure which has been widely utilized in modern genomics.
+An alternative analysis using the same ChIP-seq data, but using pre-defined regions rather than
+peak calls to test for differential binding, is performed in the accompanying `ChIPseq-promotors.Rmd` case study.
 
@@
 
-We downloaded the bam files directly from UCSC ENCODE portal.
+We download the peak calls from the ENCODE Project site.
 
-```{r, encode-download}
-broad_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeBroadHistone/"
-broad_bams <- c("wgEncodeBroadHistoneGm12878H3k4me3StdAlnRep1.bam",
-                "wgEncodeBroadHistoneGm12878H3k04me3StdAlnRep2V2.bam",
-                "wgEncodeBroadHistoneK562H3k4me3StdAlnRep1.bam",
-                "wgEncodeBroadHistoneK562H3k4me3StdAlnRep2.bam")
-
-for (i_bam in broad_bams) {
-    if (!file.exists(file.path(datdir, i_bam))) {
-        download.file(paste0(broad_url, i_bam),
-                      destfile = file.path(datdir, i_bam))
-    }
-    i_bai <- paste0(i_bam, ".bai")
-    if (!file.exists(file.path(datdir, i_bai))) {
-        download.file(paste0(broad_url, i_bai),
-                      destfile = file.path(datdir, i_bai))
-    }
-}
-
-uw_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwHistone/"
-uw_bams <- c("wgEncodeUwHistoneGm12878H3k4me3StdAlnRep1.bam",
-             "wgEncodeUwHistoneGm12878H3k4me3StdAlnRep2.bam",
-             "wgEncodeUwHistoneK562H3k4me3StdAlnRep1.bam",
-             "wgEncodeUwHistoneK562H3k4me3StdAlnRep2.bam")
-
-for (i_bam in uw_bams) {
-    if (!file.exists(file.path(datdir, i_bam))) {
-        download.file(paste0(uw_url, i_bam),
-                      destfile = file.path(datdir, i_bam))
-    }
-    i_bai <- paste0(i_bam, ".bai")
-    if (!file.exists(file.path(datdir, i_bai))) {
-        download.file(paste0(uw_url, i_bai),
-                      destfile = file.path(datdir, i_bai))
-    }
-}
 ```
-
-We determine sample metadata from the file names.
-
-```{r}
-bamfiles <- list.files(datdir, pattern = 'bam$', full = TRUE)
-labs <- gsub("wgEncode(.*)Histone.*", "\\1", basename(bamfiles))
-cells <- gsub("wgEncode.*Histone(.*)H3k.*", "\\1", basename(bamfiles))
-
-meta <- data.frame(cellline = cells, lab = labs, file = bamfiles)
-```
-
-Next, we identify promotor regions using the UCSC "Known Gene" annotations for
-human genome assembly hg19 (GRCh37).
-
-```{r}
-prom <- promoters(genes(TxDb.Hsapiens.UCSC.hg19.knownGene))
-seqlevels(prom, force = TRUE) <- seqlevels(prom)[!grepl("_", seqlevels(prom))]
+"https://www.encodeproject.org/files/ENCFF392CDW/@@download/ENCFF392CDW.bed.gz"
+"https://www.encodeproject.org/files/ENCFF621QMO/@@download/ENCFF621QMO.bed.gz"
+"https://www.encodeproject.org/files/ENCFF500EUY/@@download/ENCFF500EUY.bed.gz"
+"https://www.encodeproject.org/files/ENCFF779FUO/@@download/ENCFF779FUO.bed.gz"
 ```
@@
 
-We count sequencing reads for all promoter regions here, and test for differential peaks using these counts.
+First, we use _DiffBind_ to test for differential binding.
 
 ```{r count, results='hide', message=FALSE, warning=FALSE}
-deseq_res  <- file.path(resdir, "h3k4me3.rda")
-if (file.exists(deseq_res)) {
-    load(deseq_res)
+diffbind_res  <- file.path(resdir, "h3k4me3-peaks-diffbind.rda")
+if (file.exists(diffbind_res)) {
+    load(diffbind_res)
 } else {
-    ## count reads in each promotor region
-    anno <- data.frame(GeneID = seq_len(length(prom)),
-                       Chr = seqnames(prom),
-                       Start = start(prom),
-                       End = end(prom),
-                       Strand = strand(prom))
-    rc <- featureCounts(files = bamfiles,
-                        annot.ext = anno,
-                        allowMultiOverlap = TRUE,
-                        minOverlap = 50,
-                        readExtension3 = 150,
-                        ignoreDup = TRUE)$counts
-
-    ## perform differential testing
-    dds <- DESeqDataSetFromMatrix(countData = rc,
-                                  colData = as.data.frame(meta),
-                                  design= ~ lab + cellline)
-    dds <- DESeq(dds, fitType = "mean")
-    dat <- as.data.frame(results(dds))
-    colnames(dat) <- c('ind_covariate', 'effect_size', 'SE',
-                       'stat', 'pval', 'padj')
-
-    ## compute test-statistic from p-value for ASH
-    dat <- dplyr::mutate(dat,
-                         test_statistic = qnorm(exp(log(pval)-log(2)),
-                                                lower.tail=FALSE),
-                         test_statistic = test_statistic * sign(effect_size))
-
-    save(deseq, dat, file = deseq_res)
+    ## run diffbind
+    ## save(res, file = diffbind_res)
 }
-
-## remove regions with effect sizes = NA
-dat <- dat[!is.na(dat[, 2]), ]
 ```
@@
 
-### Mean Coverage
-
-We consider the average coverage across conditions and replicates as a
-potential covariate of interest. This is motivated by our prior assumption that
-ChIP-seq data always involves two sets of regions: signal and background.
-Sometimes, there are a few regions with signal coverage between signal and background.
-In this example, we see that p-values are differently distributed under low, 
-median and high coverage.
-
-```{r, chipseq-diag, fig.width=10, fig.height=3.5}
-strat_hist(dat, pvalue = "pval", covariate = "ind_covariate", maxy = 2)
-rank_scatter(dat, pvalue = "pval", covariate = "ind_covariate")
-```
+### Covariate One
 
@@
 ```{r, chipseq-sb, results="hide", message=FALSE}
-resN <- paste0(resdir, "/chipseq_summarizedBenchmark_", nrow(dat), ".RData")
-if (!file.exists(resN)) {
+bench_res  <- file.path(resdir, "h3k4me3-peaks-sb.rda")
+if (file.exists(bench_res)) {
+    load(bench_res)
+} else {
     sb <- buildBench(bd, data = dat, ftCols = c("ind_covariate"),
                      parallel = TRUE, BPPARAM = multicoreParam)
-    save(sb, file = resN)
-} else {
-    load(resN)
+    save(sb, file = bench_res)
 }
@@
 
-```{r, chipseq-plot, results="hide", width=15, height=15}
-# plot nrejects by method overall and stratified by covariate
+```{r, chipseq-plot, results="hide"}
 rejections_scatter(sb, supplementary = FALSE)
-
 rejection_scatter_bins(sb, covariate = "ind_covariate",
@@
 
-# properties of method-exclusive discoveries
 methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")

### 2018-02-27 datasets/ChIP-seq/ChIPseq-promotors.Rmd

--- 2018-02-26 datasets/ChIP-seq/ChIP-seq.Rmd
@@
 ---
-title: "Case Study: ChIP-seq Differential Peak Calling"
+title: "Case Study: ChIP-seq Differential Peak Calling (predefined regions)"
 author: "Mingxiang Teng"
@@
 
-Common analyses in ChIP-seq data include the calling of peaks within individual
-samples, and the testing of differential peaks across conditions. While peak
-calling continues to be the most wide-spread method applied to ChIP-seq data,
-differential testing has gained popularity with the increased availability of
-data from large-scale experiments with more complex designs.
+In differential peak calling, classical pipelines often begin by first calling peaks
+in samples, followed by performing differential testing using one of a few methods,
+e.g. _csaw_ or _DiffBind_. However, with these approaches, testing is only performed on a
+filtered set of genomic regions, namely those with significance in at least one condition,
+resulting in a potentially limited number of null tests cases. An analysis using this
+classical approach is performed in the accompanying `ChIPseq-peaks.Rmd` case study.
 
-In differential peak calling, classical pipelines often begin by first calling peaks
-in samples, followed by performing differential testing using existing algorithms
-developed for other genomic data tpyes (e.g. RNA-seq). However, with these approaches,
-testing is only performed on a filtered set of genomic regions, namely those with
-significance in at least in one condition, resulting in only a limited number of
-null tests cases.
-
-To avoid this limitation, we examine ChIP-seq data of the histone H3K4Me3, an active
-marker of gene expression, and focus our analysis on promoter regions of all genes
-for this histone marker. By doing so, we avoid limiting the set of tests to only
-regions significant in at least one sample, while also keeping most regions of
+To avoid this limitation, in this case study we examine ChIP-seq data of the histone
+H3K4Me3, an active marker of gene expression, and focus our analysis on promoter regions
+of all genes for this histone marker. By doing so, we avoid limiting the set of tests
+to only regions significant in at least one sample, while also keeping most regions of
 interest for this histone marker.
 
-Here, we focus on test differential peaks between two cell lines (K562 and Gm12878)
-using publicly available data generated by the ENCODE Project. These two cell lines
-comprise the largest sequencing datasets in the ENCODE Project. As mentioned above,
-we use ChIP-seq data for histone H3K4Me3 in our analysis. Differential peaks are tested
-using DESeq2.
-
-We note that while the algorithm for differential testing, DESeq2, used here is the same
-as that used in our analysis of RNA-seq data, benchmarking on ChIP-seq still provides
-insights for a unique data structure which has been widely utilized in modern genomics.
+As in `ChIPseq-peaks.Rmd`, we test for differential peaks using publicly available ChIP-seq
+data for histone H3K4Me3 from two cell lines (K562 and Gm12878). Data are obtained from the
+ENCODE Project. Differential peaks are tested using DESeq2.
 
@@
 ```{r count, results='hide', message=FALSE, warning=FALSE}
-deseq_res  <- file.path(resdir, "h3k4me3.rda")
+deseq_res  <- file.path(resdir, "h3k4me3-promotors-deseq.rda")
 if (file.exists(deseq_res)) {
@@
 ```{r, chipseq-sb, results="hide", message=FALSE}
-resN <- paste0(resdir, "/chipseq_summarizedBenchmark_", nrow(dat), ".RData")
-if (!file.exists(resN)) {
+bench_res  <- file.path(resdir, "h3k4me3-promotors-sb.rda")
+if (file.exists(bench_res)) {
+    load(bench_res)
+} else {
     sb <- buildBench(bd, data = dat, ftCols = c("ind_covariate"),
                      parallel = TRUE, BPPARAM = multicoreParam)
-    save(sb, file = resN)
-} else {
-    load(resN)
+    save(sb, file = bench_res)
 }
@@
 
-```{r, chipseq-plot, results="hide", width=15, height=15}
-# plot nrejects by method overall and stratified by covariate
+```{r, chipseq-plot, results="hide"}
 rejections_scatter(sb, supplementary = FALSE)

### 2018-03-06 datasets/RNA-Seq/yeast-simulation.Rmd
Use R utils functions instead of system commands

--- 2018-02-15 datasets/RNA-Seq/yeast48-replicates-benchmarkFDR.Rmd
@@
 
+# set up data / results directories
+datdir <- "yeast-data"
+resdir <- "yeast-results"
+dir.create(datdir, showWarnings = FALSE)
+dir.create(resdir, showWarnings = FALSE)
+
 # set up parallel backend
@@
 ```{r, download}
-system("wget https://github.com/bartongroup/profDGE48/raw/master/Preprocessed_data/Snf2_countdata.tar.gz")
-system("wget https://github.com/bartongroup/profDGE48/raw/master/Preprocessed_data/WT_countdata.tar.gz")
-system("mkdir yeast48-data")
-system("tar -xvf WT_countdata.tar.gz -C ./yeast48-data")
-system("tar -xvf Snf2_countdata.tar.gz -C ./yeast48-data")
-system("rm *.tar.gz")
+download.file(url = "https://github.com/bartongroup/profDGE48/raw/master/Preprocessed_data/Snf2_countdata.tar.gz",
+              destfile = paste0(datdir, "/Snf2_countdata.tar.gz"))
+download.file(url = "https://github.com/bartongroup/profDGE48/raw/master/Preprocessed_data/WT_countdata.tar.gz", 
+              destfile = paste0(datdir, "/WT_countdata.tar.gz"))
+
+gunzip(paste0(datdir, "/Snf2_countdata.tar.gz")) 
+gunzip(paste0(datdir, "/WT_countdata.tar.gz"))
+
+untar(paste0(datdir, "/Snf2_countdata.tar"), exdir = datdir) 
+untar(paste0(datdir, "/WT_countdata.tar"), exdir = datdir)
+
+file.remove(paste0(datdir, "/Snf2_countdata.tar"), 
+            paste0(datdir, "/WT_countdata.tar"))
 ```
@@
 ```{r, readin, results='hide', message=FALSE}
-files <- dir(path = "./yeast48-data/", pattern = "*.bam.gbgout", full.names = TRUE)
-sample_names <- sapply(strsplit(dir(path = "./yeast48-data/", pattern = "*.bam.gbgout"), "_MID"),
+files <- dir(path = datdir, pattern = "*.bam.gbgout", full.names = TRUE)
+sample_names <- sapply(strsplit(dir(path = datdir, pattern = "*.bam.gbgout"), "_MID"),
                        function(x) x[[1]])
@@
 nDE <- 0
+resfile <- paste0(resdir, "/null5.rds")
 
-system("mkdir -p yeast48-results")
-
-if (!file.exists("yeast48-results/null5.rds")){
+if (!file.exists(resfile)){
   null5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
@@
                    mc.cores=cores)
-  saveRDS(null5, file="yeast48-results/null5.rds")
+  saveRDS(null5, file=resfile)
 }else{
-  null5 <- readRDS(file="yeast48-results/null5.rds")
+  null5 <- readRDS(file=resfile)
 }
@@
 nDE <- 0
+resfile <- paste0(resdir, "/null10.rds")
 
-if (!file.exists("yeast48-results/null10.rds")){
+if (!file.exists(resfile)){
   null10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
@@
                    mc.cores=cores)
-  saveRDS(null10, file="yeast48-results/null10.rds")
+  saveRDS(null10, file=resfile)
 }else{
-  null10 <- readRDS(file="yeast48-results/null10.rds")
+  null10 <- readRDS(file=resfile)
 }
@@
 nDE <- 500
+resfile <- paste0(resdir, "/de5.rds")
 
-if (!file.exists("yeast48-results/de5.rds")){
+if (!file.exists(resfile)){
   de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
@@
                    mc.cores=cores)
-  saveRDS(de5, file="yeast48-results/de5.rds")
+  saveRDS(de5, file=resfile)
 }else{
-  de5 <- readRDS(file="yeast48-results/de5.rds")
+  de5 <- readRDS(file=resfile)
 }
@@
                 merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res5d, met="FPR",filter_set = excludeSet,
+plotsim_average(res5d, met="FDR",filter_set = excludeSet,
                 merge_ihw = TRUE, errorBars=TRUE) 
@@
 nDE <- 500
+resfile <- paste0(resdir, "/de10.rds")
 
-if(!file.exists("yeast48-results/de10.rds")){
+if(!file.exists(resfile)){
   de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
@@
                    mc.cores=cores)
-  saveRDS(de10, file="yeast48-results/de10.rds")
+  saveRDS(de10, file=resfile)
 }else{
-  de10 <- readRDS(file="yeast48-results/de10.rds")
+  de10 <- readRDS(file=resfile)
 }
@@
                 merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res10d, met="FPR",filter_set = excludeSet,
+plotsim_average(res10d, met="FDR",filter_set = excludeSet,
                 merge_ihw = TRUE, errorBars=TRUE) 

### 2018-03-08 datasets/GWAS/GWAS.Rmd
Add covariate and effect size heatmaps

--- 2017-09-21 datasets/GWAS/GWAS.Rmd
@@
 ---
-title: "GWAS datasets for FDR benchmarking"
+title: "Case study: Genome-Wide Association Analyses (GWAS)"
 author: "Keegan Korthauer"
-date: "9/21/2017"
-output: html_document
+date: '`r format(Sys.Date(), "%B %d, %Y")`'
+output:
+   html_document:
+        toc: true
+        toc_float: true
+        highlight: tango
+        number_sections: true
 ---
@@
 
-### Summary
+# Summary
 
-Here are the GWAS datasets we will use to benchmark the FDR methods. For each one
-we provide the script to download the data and read into R. We'd like to verify 
-that each dataset contains the desired inputs:
+Here we will make use of the `SummarizedBenchmark` package to benchmark a 
+Genome-Wide Association Study (GWAS) dataset. We will examine a GWAS 
+meta-analysis for Body Mass Index (BMI), which was also included in the 
+Boca-Leek manuscript.
 
-- p-value
-- effect size
-- standard error
-- additional covariates (optional)
+# Workspace setup
 
-Note that while running all the bash code in this document will download the data, 
-some of it doesn't seem to work properly with knitting (e.g. the ftp download 
-in the second dataset), and I'm not committing 
-the data itself to this repo. Run the bash code in the terminal separately to 
-download the data yourself, and see the knitted `.html` document if you do not 
-wish to download the datasets but still wish to see the output.
+This analysis requires that the software PLINK (version 1.9) 
+is installed on your system.
+PLINK is freely available for download 
+[here](http://zzz.bwh.harvard.edu/plink/plink2.shtml).
 
-We'll also run the benchmark methods on each dataset and report the number of 
-discoveries for each one. At first, we'll demo on just a __**small subset of 
-SNPs**__, since the computation is time-intensive. After that is working 
-smoothly, we'll scale it up to the full set of SNPs.
-
-### Workspace setup
-
-To run the benchmark methods on the GWAS datasets, we'll use some of the helper 
-and wrapper functions from the simulated data analysis. We also create and source
-an analgous version of the `sim_runner` function to apply to the real data.
-
-```{r, wkspace-setup}
-# Note that swfdr package depends on R version 3.4 or higher
-setwd("/n/irizarryfs01_backed_up/kkorthauer/FDR")
-source("benchmark-fdr/simulation_studies/R/simulation-helpers.R")
-source("benchmark-fdr/datasets/R/evaluation-helpers.R")
-datdir <- "/n/irizarryfs01_backed_up/kkorthauer/FDR/DATA/GWAS/"
-resdir <- "/n/irizarryfs01_backed_up/kkorthauer/FDR/RESULTS/GWAS/"
+```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
+# data and results directories
+system("mkdir -p DATA")
+system("mkdir -p RESULTS")
+datdir <- "DATA/"
+resdir <- "RESULTS/"
 
@@
 library(readxl)
+library(readr)
 library(dplyr)
 library(ggplot2)
+library(cowplot)
 
-# benchmark methods
-library(IHW)
-library(ashr)
-library(qvalue)
-library(swfdr)
-library(fdrtool)
+# comparison tools/functions
+library(R.utils)
+sourceDirectory("../R/")
+
+# set up parallel backend
+library(BiocParallel)
+cores <- 8
+multicoreParam <- MulticoreParam(workers = cores)
 ```
 
-### GWAS meta-analysis for BMI (included in Boca-Leek)
+# Data preparation
 
-Here we download and analyze the GWAS BMI dataset analyzed in the Boca-Leek paper.
+We'll first download the GWAS summary statistics dataset from the 
+GIANT consortium data portal.
 
-#### Data download
+## Data download
 
-```{bash, GWAS1-download, eval=FALSE}
-curl -O "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
+After downloading, we unzip the file, and delete the files we won't use (they provide results
+subsetted by ancestry and sex), keeping only the file with European ancestry and
+both sexes `BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt`. Note that
+the Boca-Leek paper also used the 'EuropeanOnly' subset, likely because using 
+all ancestries would require that we assess the impact of population stratification
+and adjust for it if present. For simplicity, we follow suit and use the homogeneous 
+subset to avoid the impact of population stratification on our results.
+
+```{r, GWAS1-download}
+if(!file.exists(paste0(datdir,"BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt"))){
+  download.file(url="http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip", 
+             destfile=paste0(datdir,"BMI.SNPadjSMK.zip")) 
+  unzip(paste0(datdir,"BMI.SNPadjSMK.zip"), exdir=datdir)
+  file.remove(paste0(datdir,"BMI.SNPadjSMK.zip"))
+
+  dfiles <- list.files(path=datdir, pattern="BMI.SNPadjSMK.*.txt", 
+                     full.names = TRUE)
+  dfiles <- dfiles[!grepl("BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt", dfiles)]
+  file.remove(dfiles)
+}
 ```
 
-Next we unzip the file, and delete the files we won't use (they provide results
-subsetted by ancestry and sex), keeping only the file with all ancestries and
-both sexes `BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt`. Note that it seems that
-the Boca-Leek paper actually used the 'EuropeanOnly' subset, but it's not clear
-why (to me), and there's no reason we need to do the same.
+We also download the 1000 genomes CEU reference population data, which will be used to 
+determine which SNPs are LD buddies for the purposes of obtaining an independent set of SNPs.
 
-```{bash, GWAS1-unzip, eval=FALSE}
-unzip BMI.SNPadjSMK.zip
-rm BMI.SNPadjSMK.zip
-rm BMI.SNPadjSMK.Women.EuropeanOnly.txt
-rm BMI.SNPadjSMK.Women.AllAncestry.txt
-rm BMI.SNPadjSMK.Men.EuropeanOnly.txt
-rm BMI.SNPadjSMK.Men.AllAncestry.txt
-rm BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt
+```{r, GWAS1-refdata}
+reffile <- paste0(datdir, "1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01")
+if (!file.exists(paste0(reffile, ".fam"))){
+  download.file(url="http://neurogenetics.qimrberghofer.edu.au/iSECA/1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01.zip", 
+              destfile=paste0(reffile, ".zip"))
+  unzip(paste0(reffile, ".zip"))
+  file.remove(paste0(reffile, ".zip"))
+}
 ```
@@
 
-for 2,458,133 SNPs.
+for `r nrow(bmi)` SNPs.
 
-#### Running benchmark methods
+## Prune SNPs that are in Linkage Disequilibrium
 
-First, we'll standardize the data frame to use the wrapper function. For now, we'll
-assume that there is only a single 'independent covariate' for those methods that 
-utilize this, and we'll use the sample size (number of subjects with that SNP
-measured). **Still to-do: consider multiple covariates for methods that allow this.**
+Since nearby SNPs in linkage disequilibrium may not constitute separate discoveries, we carry out 
+a pruning step. This commonly done by examining correlation of genotypes among nearby markers, but
+since we don't have the genotype data, we'll use the linkage disequilibrium results of common SNPs and 
+previously observed LD blocks. Specifically, we'll use the phase 3 data from 1000 genomes, restricted
+to the CEU population (since this most closely matches our population of European ancestry). This 
+reference data was downloaded in a previous code chunk. Here we read in the list of more than
+9 million SNPs with data in 1000 genomes and create a list of SNPs that overlap our set. We'll write 
+this list to a file that matches PLINK format.
 
-```{r, GWAS1-dat}
+```{r, GWAS1-snplist, eval=TRUE}
+# load reference data 
+onekg <- fread(paste0(datdir, "1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01.bim"))$V2
+
+# construct input list of SNPs that overlap - 97.5% overlap
+bmi$BP <- as.numeric(unlist(lapply(strsplit(bmi$markername, ":"), function(x) x[[2]])))
+write_delim(bmi %>% dplyr::rename(CHR=chromosome, SNP=rs_id, A1=allele_1, A2=allele_2) %>%
+              mutate(P=0.5, F_A=Freq_Allele1_HapMapCEU, F_U=F_A, CHISQ=1, OR=1) %>% 
+              select(CHR, SNP, BP, A1, F_A, F_U, A2, CHISQ, P, OR) %>%
+              filter(SNP %in% onekg), 
+            path=paste0(datdir, "rslist.assoc"), delim="\t")
+
+ol <- sum(bmi$rs_id %in% onekg)/nrow(bmi)
+rm(onekg)
+```
+
+It turns out that `r signif(ol*100, 3)` percent of SNPs in our set are included in the 1000 genomes 
+reference data. Next we'll use the PLINK software to read in our list of SNPs and the 1000 genomes
+linkage disequilibrium data to 'clump' the SNPs into independent sets of SNPs that have LD correlation (r squared)
+less than 0.2 with all neighbors within 250 kilobases of distance. We write the results to a file called `plink_clump.clumped`. We have to jump out of R here since PLINK is a command line tool.
+
+```{bash, clump, eval=FALSE}
+# use PLINK clumping tool to obtain a set of independent SNPs by LD
+plink \
+--bfile DATA/1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01 \
+--clump DATA/rslist.assoc \
+--clump-field P \
+--clump-p1 0.9999 \
+--clump-p2 0.9999 \
+--clump-r2 0.2 \
+--clump-kb 250 \
+--out DATA/plink_clump
+```
+
+Next, we'll read in the clumped results and subset our data by our list of independent SNPs.
+
+```{r, GWAS1-prune}
+clump <- fread(paste0(datdir, "plink_clump.clumped"))
+
+# subset to include only the SNPs in the ld.list
+bmi <- bmi[bmi$rs_id %in% clump$SNP,]
+
+rm(clump)
+```
+
+After pruning, we are left with for `r nrow(bmi)` approximately independent SNPs.
+
+Before moving on, we'll rename the relevant columns so that they will use common terms
+across the different datasets. We'll also add a 'test_statistic' column.
+
+```{r, GWAS1-reformat}
 bmi <- bmi %>% 
-        rename(pval = p_value,
+  dplyr::rename(pval = p_value,
                 SE = stderr,
                 effect_size = effect,
-               ind_covariate = N) %>%
-        mutate(zscore = effect_size/SE)
-
-# Demo on a small subset of SNPs first (remove this later)
-bmi <- bmi[sample(1:nrow(bmi), 100000),]
+                ind_covar_N = N,
+                ind_covar_AF = Freq_Allele1_HapMapCEU) %>%
+  mutate(test_statistic = qnorm(exp(log(pval)-log(2)), lower.tail=FALSE) * sign(effect_size)) %>%
+  select(pval, SE, effect_size, ind_covar_N, ind_covar_AF, test_statistic) 
 ```
 
-Now, we'll run the `run_benchmarks` helper function on the `bmi` data frame for
-several alpha values and plot the results. This is fairly computationally intensive 
-for some of the benchmark methods (since there are more than 2.5 million tests)
-and best run on a machine with lots of memory.
+# Data Analysis
 
-```{r, GWAS1-run, results="hide"}
-cutoffs <- c(0.01, 0.025, 0.05, 0.10, 0.20)
-resfile <- paste0(resdir, "bmi_results_",
-                  nrow(bmi), ".RData")
-if (!file.exists(resfile)){
-  t1 <- proc.time()
-  bmi_results <- run_benchmarks(dat=bmi, alphas=cutoffs)
-  save(bmi_results, file=resfile)
-  message(paste0("Took ", round((proc.time()-t1)[3],1),
-                 " seconds for ", nrow(bmi), " SNPs."))
-}else{
-  load(resfile)
-}
+## Differential Testing
+Since we do not have access to the individual genotype information and since
+our data already contains the results of performing hypothesis tests of
+association of BMI with each SNP (including p-values and effect sizes), we 
+don't need to carry out any additional testing. We move on to checking the
+covariate diagnostics and adjusting for multiple comparisons.
 
-# plot N rejects vs FDR cutoff
-p <- ggplot(bmi_results, aes(alpha, n_rejects, color=method)) +
-  geom_point() +
-  geom_line() +
-  xlab("FDR cutoff") + ylab("Number of rejections") +
-  ggtitle("BMI GWAS Metanalysis with Sample Size Covariate") +
-  theme_classic()
-print(p)
+## Covariate Diagnostics
 
-# log-scale y axis
-print(p + scale_y_continuous(trans="log2") + 
-        ylab("Number of rejections (log2 scale)"))
+Here we look to see if the covariates do indeed look informative. First we look at 
+the sample size covariate.
+
+### Covariate one: Sample Size
+
+```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
+rank_scatter(bmi, pvalue="pval", covariate="ind_covar_N") + 
+  ggtitle("Covariate 1: Sample Size")
 ```
 
-And we'll clean up the workspace before moving on to the next dataset.
-
-```{r, GWAS1-cleanup}
-rm(bmi)
-rm(bmi_results)
+```{r, fig.width=10, fig.height=3.2, message=FALSE}
+strat_hist(bmi, pvalue="pval", covariate="ind_covar_N", maxy=2) 
 ```
 
-### GWAS for Asthma
+### Covariate two: Minor Allele Frequency
 
-[This study by Moffatt et al. (2010, NEJM)](http://www.nejm.org/doi/full/10.1056/NEJMoa0906312?query=TOC) performed a large
-GWAS on  10,365 people with asthma and 16,110 controls 23 individual studies.
+Next we look at the allele frequency covariate.
 
-#### Data Download
-
-Let's download the data from the Gabriel site. The second file contains metadata
-for the columns in the main data file.
-
-```{bash, GWAS2-download, eval=FALSE}
-curl -O "https://beaune.cng.fr/gabriel/gabriel_results.zip"
-curl -O "https://beaune.cng.fr/gabriel/gabriel_results_description.xls"
+```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
+rank_scatter(bmi, pvalue="pval", covariate="ind_covar_AF") +
+  ggtitle("Covariate 1: Minor allele frequency")
 ```
 
-Next we unzip the file and remove the zipped copy.
-
-```{bash, GWAS2-unzip, eval=FALSE}
-unzip gabriel_results.zip
-rm gabriel_results.zip
+```{r, fig.width=10, fig.height=3.2, message=FALSE}
+strat_hist(bmi, pvalue="pval", covariate="ind_covar_AF", maxy=2)
 ```
 
-Next, we'll read in the unzipped `.txt` file and the results desription `.xls` 
-file into R and verify that it contains
-the necessary inputs to run the FDR benchmark comparisons.
+## Multiple-Testing Correction
 
-```{r, GWAS2-verify-contents}
-# read in meta-data
-meta <- read_xls(paste0(datdir, "gabriel_results_description.xls"))
-head(meta)
-dim(meta)
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object.
 
-# read in data
-asthma <- fread(paste0(datdir, 
-                "gabriel_asthma_meta-analysis_36studies_format_repository_NEJM.txt"),
-                header=TRUE)
-dim(asthma)
-
-# verify meta-data rows = data cols
-nrow(meta) == ncol(asthma)
-
-# remove study-specific columns
-empty <- which(!is.na(meta$description))
-asthma <- asthma %>%
-            select(empty)
-meta <- meta %>% 
-            slice(empty)
-
-# remove SNPs with missing effect size/pvalues
-asthma <- asthma %>% 
-            filter(!is.na(P_fix) & !is.na(theta_fix))
-
-dim(asthma)
-head(asthma)
+```{r, GWAS1-benchdesign, message=FALSE}
+bd <- initializeBenchDesign()
 ```
 
-It looks like we have:
+Now, we're ready to construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+                                        helper scripts).
 
-- `P_fix` and `P_ran`:p-values for fixed and random effect models
-- `theta_fix` and `theta_ran`:effect size for fixed and random effect models
-- `se_theta_fix` and `se_theta_ran`:standard error for fixed and random effect models
-- additional covariates: 
-    - `no_studies_meta`: Number of studies included in the meta-analysis with this SNP
-    - `freq_all_1_max` and `freq_all_1_min`: Maximum and Minimum Allele Frequency of Allele 1 across samples
+### Covariate one: Sample size
 
-for 546,182 SNPs.
+First we'll include the sample size covariate.
 
-#### Running benchmark methods
+```{r, GWAS1-sb, results="hide", message=FALSE}
+resN <- paste0(resdir, "bmi_N_summarizedBenchmark_",
+               nrow(bmi), ".RData")
 
-First, we'll standardize the data frame to use the wrapper function. For now, we'll
-assume that there is only a single 'independent covariate' for those methods that 
-utilize this, and we'll use the sample size (number of studies with that SNP
-measured). We'll also focus on the fixed effects tests first.
-**Still to-do: consider multiple covariates for methods that allow this.**
-
-```{r, GWAS2-dat}
-asthma <- asthma %>% 
-        rename(pval = P_fix,
-               SE = se_theta_fix,
-               effect_size = theta_fix,
-               ind_covariate = no_studies_meta) %>%
-        mutate(zscore = effect_size/SE)
-
-# Demo on a small subset of SNPs first (remove this later)
-asthma <- asthma[sample(1:nrow(asthma), 100000),]
+duration <- NA
+if (!file.exists(resN)){
+  t1 <- proc.time()
+  sbN <- bd %>% buildBench(data=(bmi %>% mutate(ind_covariate=ind_covar_N)), 
+                          ftCols = c("ind_covar_N"),
+                          parallel=TRUE, BPPARAM=multicoreParam)
+  metadata(sbN)$data_download_link <- 
+    "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
+  saveRDS(sbN, file=resN)
+  duration <- round((proc.time()-t1)[3]/60,1)
+}else{
+  sbN <- readRDS(resN)
+}
 ```
 
-Now, we'll run the `run_benchmarks` helper function on the `asthma` data frame for
-several alpha values and plot the results. This is fairly computationally intensive 
-for some of the benchmark methods
-(since there are more than 500 thousand tests)
-and best run on a machine with lots of memory.
-
-```{r, GWAS2-run, results="hide"}
-# use same cutoffs as for the bmi study (defined above)
-resfile <- paste0(resdir, "asthma_results_",
-                  nrow(asthma), ".RData")
-if (!file.exists(resfile)){
-  t1 <- proc.time()
-  asthma_results <- run_benchmarks(dat=asthma, alphas=cutoffs)
-  save(asthma_results, file=resfile)
-  message(paste0("Took ", round((proc.time()-t1)[3],1),
-                 " seconds for ", nrow(asthma), " SNPs."))
-}else{
-  load(resfile)
+```{r, echo=FALSE}
+if(!is.na(duration)){
+  message("This step took ", duration, " minutes for ", nrow(bmi), " SNPs using ",
+          cores, " cores.")
 }
-
-# plot N rejects vs FDR cutoff
-p <- ggplot(asthma_results, aes(alpha, n_rejects, color=method)) +
-  geom_point() +
-  geom_line() +
-  xlab("FDR cutoff") + ylab("Number of rejections") +
-  ggtitle("Asthma GWAS Metanalysis with Sample Size Covariate") +
-  theme_classic()
-print(p)
-
-# log-scale y axis
-print(p + scale_y_continuous(trans="log2") + 
-        ylab("Number of rejections (log2 scale)"))
 ```
 
-And we'll clean up the workspace before moving on to the next dataset.
+### Covariate two: Minor Allele Frequency
 
-```{r, GWAS2-cleanup}
-rm(asthma)
-rm(asthma_results)
+Now, we'll repeat the multiple testing correction using the other covariate (AF):
+
+```{r, GWAS1-sb2, results="hide", message=FALSE}
+resAF <- paste0(resdir, "bmi_AF_summarizedBenchmark_",
+nrow(bmi), ".RData")
+
+duration <- NA
+if (!file.exists(resAF)){
+  t1 <- proc.time()
+  sbAF <- bd %>% buildBench(data=(bmi %>% mutate(ind_covariate=ind_covar_AF) %>%
+                                    filter(!is.na(ind_covariate))), 
+                            ftCols = c("ind_covar_AF"),
+                            parallel=TRUE, BPPARAM=multicoreParam)
+  metadata(sbAF)$data_download_link <-    
+    "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
+  saveRDS(sbAF, file=resAF)
+  duration <- round((proc.time()-t1)[3]/60,1)
+}else{
+  sbAF <- readRDS(resAF)
+}
 ```
 
-### GWAS for Lipid Levels
-
-[This study by Willer et al. (2013, Nature Genetics)](http://www.nature.com/ng/journal/v45/n11/full/ng.2797.html) performed a large
-GWAS on 188,577 and measured blood lipid levels (as well as several cardiovascular traits).
-
-#### Data Download
-
-Let's download the data from the University of Michigan site. The download page 
-has a handy [metadata table](http://csg.sph.umich.edu//abecasis/public/lipids2013/)
-that describes the contents of the columns in the results files.
-
-There are several types of results files to download (in terms of which phenotype to analyze):
-
-- LDL Cholesterol
-- HDL Cholesterol
-- Triglycerides
-- Total Cholesterol
-
-As well as two types of analyses:
-
-- Joint analyses of GWAS and Metabochip
-- analysis of Metabochip
-
-We would like to look at the joint analysis of GWAS + Metabochip since we'd like 
-to look at all available SNPs (not just those that have previously shown a possible
-association with CVD). We could choose any or all of the traits, but I'll go ahead
-and proceed with Triglycerides since I'm curious about the role of genetic variation 
-on Triglyceride levels.
-
-```{bash, GWAS3-download, eval=FALSE}
-curl -O "http://csg.sph.umich.edu/abecasis/public/lipids2013/jointGwasMc_TG.txt.gz"
+```{r, echo=FALSE}
+if(!is.na(duration)){
+  message("This step took ", duration, " minutes for ", nrow(bmi), " SNPs using ",
+  cores, " cores.")
+}
 ```
 
-Next we uncompress the file.
+## Benchmark Metrics
 
-```{bash, GWAS3-unzip, eval=FALSE}
-gunzip jointGwasMc_TG.txt.gz 
+Next, we'll add the default performance metric for q-value assays and 
+plot the results. We'll start with covariate one.
+
+### Covariate one: Sample Size
+
+First, we have
+to rename the assay to 'qvalue'.
+
+```{r, GWAS1-metrics}
+# rename assay to qvalue
+assayNames(sbN) <- "qvalue"
+sbN <- addDefaultMetrics(sbN)
 ```
 
-Next, we'll read in the uncompressed `.txt` file into R and verify that it contains
-the necessary inputs to run the FDR benchmark comparisons.
+Now, we'll plot the results.
 
-```{r, GWAS3-verify-contents}
-# read in data
-trig <- fread(paste0(datdir, "jointGwasMc_TG.txt"), header=TRUE)
-dim(trig)
-head(trig)
-```
+```{r, GWAS1-plot, results="hide"}
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter(sbN,
+                   supplementary=FALSE) +
+  ggtitle("Covariate 1: Sample size")
 
-It looks like we have:
+rejection_scatter_bins(sbN, covariate="ind_covar_N", bins=4,
+                       supplementary=FALSE) +
+  ggtitle("Covariate 1: Sample size")
 
-- `P_value`:p-value
-- `beta`:effect size
-- `SE`:standard error
-- additional covariates: 
-    - `N`: The number of individuals analyzed for this marker.
-    - `Freq.A1.1000G.EUR`: Frequency of allele A1 from 1000G EUR sample.
+# upset plot 
+plotFDRMethodsOverlap(sbN, 
+                      alpha=0.05, nsets=ncol(sb),
+                      order.by="freq", decreasing=TRUE,
+                      supplementary=FALSE)
 
-for 2,439,432 SNPs.
-
-#### Running benchmark methods
-
-First, we'll standardize the data frame to use the wrapper function. For now, we'll
-assume that there is only a single 'independent covariate' for those methods that 
-utilize this, and we'll use the sample size (number of individuals with that SNP
-measured). 
-**Still to-do: consider multiple covariates for methods that allow this.**
-
-Note that for this dataset, all reported effect sizes are positive. This is because
-the alleles (A1 and A2) have been ordered such that the direction of effect are 
-consistent. Since Scott assumes that the Z-scores are standard normal, we can't 
-apply this method to a set of all positive Z-scores. To recover the direction of 
-effect, we use the `Freq.A1.1000G.EUR` (allele frequency of A1) variable,
-and use it to determine whether
-`A1` (allele 1) is the reference allele or not. If the allele frequence of A1 is
-greater than 0.5, we consider it the reference allele. Then we'll use the 
-comparison of reference to minor as the positive direction. That is,
-if A2 was reference,
-then effect direction is negative. Otherwise it is positive. Since some of the
-SNPs are missing the reference allele information, we cannot determine their 
-direction of effect and they will be removed.
-
-```{r, GWAS3-dat}
-trig <- trig %>% 
-        rename(pval = "P-value",
-               SE = se,
-               effect_size = beta,
-               ind_covariate = N)%>%
-        mutate(zscore = effect_size/SE)
-
-# add effect size
-trig <- trig %>% 
-       mutate(A2.reference = Freq.A1.1000G.EUR < 0.5) %>%
-       mutate(zscore = zscore * (1-2*A2.reference))
-
-# remove SNPs with missing values for zscore (couldn't determine direction of 
-# effect)
-trig <- trig %>% 
-  filter(!is.na(zscore))
-
-# Demo on a small subset of SNPs first (remove this later)
-trig <- trig[sample(1:nrow(trig), 100000),]
-```
-
-Now, we'll run the `run_benchmarks` helper function on the `trig` data frame for
-several alpha values and plot the results. This is fairly computationally intensive 
-for some of the benchmark methods
-(since there are more than 500 thousand tests)
-and best run on a machine with lots of memory.
-
-```{r, GWAS3-run, results="hide"}
-# use same cutoffs as for the bmi study (defined above)
-resfile <- paste0(resdir, "trig_results_",
-                  nrow(trig), ".RData")
-if (!file.exists(resfile)){
-  t1 <- proc.time()
-  trig_results <- run_benchmarks(dat=trig, alphas=cutoffs)
-  save(trig_results, file=resfile)
-  message(paste0("Took ", round((proc.time()-t1)[3],1),
-                 " seconds for ", nrow(trig), " SNPs."))
-}else{
-  load(resfile)
-}
-
-# plot N rejects vs FDR cutoff
-p <- ggplot(trig_results, aes(alpha, n_rejects, color=method)) +
-  geom_point() +
-  geom_line() +
-  xlab("FDR cutoff") + ylab("Number of rejections") +
-  ggtitle("Triglyceride GWAS Metanalysis with Sample Size Covariate") +
-  theme_classic()
-print(p)
-
-# log-scale y axis
-print(p + scale_y_continuous(trans="log2") + 
-        ylab("Number of rejections (log2 scale)"))
+# properties of method-exclusive discoveries
+methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
+             "scott-empirical", "scott-theoretical")
+mcols(sbN)$ind_covariate <- mcols(sbN)$ind_covar_N
+plotCovariateBoxplots(sbN, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3) +
+  ggtitle("Covariate 1: Sample size")
 
@@
 
-And we'll clean up the workspace before moving on to the next dataset.
+```{r, fig.width=8, fig.height=3.5}
+mcols(sbN)$tstat <- bmi %>% filter(!is.na(ind_covar_N)) %>% pull(effect_size)
+covariateHeatmap(sbN, alpha=0.05, covname="tstat")
+covariateHeatmap(sbN, alpha=0.05, covname="ind_covar_N")
+```
 
-```{r, GWAS3-cleanup}
-rm(trig)
-rm(trig_results)
+### Covariate two: Minor Allele Frequency
+
+Next, we'll look at the performance metrics for the other covariate (minor
+allele frequency).
+
+```{r, GWAS1-metrics2}
+# rename assay to qvalue
+assayNames(sbAF) <- "qvalue"
+sbAF <- addDefaultMetrics(sbAF)
 ```
+
+Now, we'll plot the results.
+
+```{r, GWAS1-plot2, results="hide"}
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter(sbAF, supplementary=FALSE) +
+  ggtitle("Covariate 2: Minor allele frequency")
+
+rejection_scatter_bins(sbAF, covariate="ind_covar_AF", bins=4,
+                       supplementary=FALSE) +
+  ggtitle("Covariate 2: Minor allele frequency")
+
+# upset plot 
+plotFDRMethodsOverlap(sbAF, 
+                      alpha=0.05, nsets=ncol(sbAF),
+                      order.by="freq", decreasing=TRUE,
+                      supplementary=FALSE)
+
+# properties of method-exclusive discoveries
+methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
+             "scott-empirical", "scott-theoretical")
+mcols(sbAF)$ind_covariate <- abs(mcols(sbAF)$ind_covar_AF-0.5)
+plotCovariateBoxplots(sbAF, alpha=0.1, nsets=6, methods=methods, maxNum=3)+
+  ggtitle("Covariate 2: Minor allele frequency")
+```
+
+```{r, fig.width=8, fig.height=3.5}
+mcols(sbAF)$tstat <- bmi %>% filter(!is.na(ind_covar_AF)) %>% pull(effect_size)
+covariateHeatmap(sbAF, alpha=0.05, covname="tstat")
+covariateHeatmap(sbAF, alpha=0.05, covname="ind_covar_AF")
+```
+
+# Session Information
+
+```{r}
+sessionInfo()
+```

### 2018-03-08 datasets/GWAS/GWAS-knit.sh

@@ -9,10 +9,9 @@
 #SBATCH -e SLURM/render-%j.err  # Standard error
  
 module load plink
-module load rstudio
 export SLURM_NTASKS
 export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/rstudio/0.98.1103-fasrc01/bin/pandoc/"
 
-# change filename to Rmd to be knitted. ncores is passed in through environment
-# variable SLURM_NTASKS
-R -e "rmarkdown::render('GWAS.Rmd')"
+# change filename to Rmd to be knitted. 
+# make sure ncores in Rmd matches -n batch param above
+R -e "rmarkdown::render('GWAS.Rmd', clean = FALSE)"

### 2018-03-08 datasets/scrnaseq/scrnaseq-mouse-scdd.Rmd
Add covariate heatmaps. Use mean nonzero expression as covariate

@@
 
-cores <- 4
+cores <- 8
 BiocParallel::register(MulticoreParam(workers = cores))
@@
 In scrnaseq data, strength of the signal can be a affected by the of level of gene expression. 
-We will explore two potential covariates related to expression level: mean expression 
+We will explore two potential covariates related to expression level: mean nonzero expression 
 and detection rate (defined as the proportion of cells expressing the gene at a nonzero level). 
@@
 add_covariates_scrnaseq <- function(dat){
-  rowData(dat)$meanExp <- rowMeans(assay(dat, "count"))
+  rowData(dat)$meanExp <- apply(assay(dat, "count"), 1, 
+                                function(x) mean(x[x>0]))
   rowData(dat)$detection <- rowMeans(assay(dat, "count") > 0)
@@
 
-### Covariate one: Mean Expression
+### Covariate one: Mean Nonzero Expression
 
@@
              pvalue="scdd_p", covariate="meanExp") + 
-  ggtitle("scDD, Covariate 1: Mean Expression")
+  ggtitle("scDD, Covariate 1: Mean Nonzero Expression")
 ```
@@
            pvalue="scdd_p", covariate="meanExp", maxy=16,
-           main = "scDD, Covariate 1: Mean Expression") 
+           main = "scDD, Covariate 1: Mean Nonzero Expression") 
 ```
 
-The mean expression appears to be informative and approximately satisfies the 
+The mean nonzero expression appears to be informative and approximately satisfies the 
 assumptions
@@
 
-First we'll include the mean expression covariate.
+First we'll include the mean nonzero expression covariate.
 
@@
                    supplementary=FALSE) +
-  ggtitle("scDD, Covariate 1: Mean Expression")
+  ggtitle("scDD, Covariate 1: Mean Nonzero Expression")
 
@@
                        supplementary=FALSE) +
-  ggtitle("scDD, Covariate 1: Mean Expression")
+  ggtitle("scDD, Covariate 1: Mean Nonzero Expression")
 
@@
 plotCovariateBoxplots(sb1, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3) +
-  ggtitle("scDD, Covariate 1: Mean Expression")
+  ggtitle("scDD, Covariate 1: Mean Nonzero Expression")
 
+```
+
+```{r, fig.width=8, fig.height=3.5}
+covariateHeatmap(sb1, alpha=0.05, covname="ind_covariate") +
+  ggtitle("scDD, Covariate 1: Mean Nonzero Expression")
 ```
@@
 mcols(sb2)$ind_covariate <- mcols(sb2)$detection
-plotCovariateBoxplots(sb2, alpha=0.1, nsets=6, methods=methods, maxNum=3)+
+plotCovariateBoxplots(sb2, alpha=0.1, nsets=6, methods=methods, maxNum=3) +
   ggtitle("scDD, Covariate 2: Detection Rate")
 
+```
+```{r, fig.width=8, fig.height=3.5}
+covariateHeatmap(sb2, alpha=0.05, covname="ind_covariate") +
+  ggtitle("scDD, Covariate 2: Detection Rate")
 ```

### 2018-03-08 datasets/scrnaseq/scrnaseq-knit.sh

@@ -1,20 +1,19 @@
 #!/bin/bash
-#SBATCH -J knit-sc4                  # A single job name for the array
-#SBATCH -n 6                         # Number of cores
-#SBATCH -N 1                         # All cores on one machine
-#SBATCH -p irizarry,serial_requeue   # Partition
-#SBATCH --mem 200G                   # Memory request 
-#SBATCH -t 0-20:00                   # Maximum execution time (D-HH:MM)
+#SBATCH -J knit-sc                  # A single job name for the array
+#SBATCH -n 8                        # Number of cores
+#SBATCH -N 1                        # All cores on one machine
+#SBATCH -p shared                   # Partition
+#SBATCH --mem 125G                  # Memory request 
+#SBATCH -t 0-6:00                   # Maximum execution time (D-HH:MM)
 #SBATCH -o render-%j.out
 #SBATCH -e render-%j.err
  
-module load rstudio
 export SLURM_NTASKS
 export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/rstudio/0.98.1103-fasrc01/bin/pandoc/"
 
-# change filename to Rmd to be knitted. ncores is passed in through environment
-# variable SLURM_NTASKS
-R -e "rmarkdown::render('scrnaseq-human-scdd.Rmd')"
-R -e "rmarkdown::render('scrnaseq-human-mast.Rmd')"
-R -e "rmarkdown::render('scrnaseq-mouse-scdd.Rmd')"
-R -e "rmarkdown::render('scrnaseq-mouse-mast.Rmd')"
+# change filename to Rmd to be knitted
+# make sure number of cores used in Rmd matches sbatch param -n
+R -e "rmarkdown::render('scrnaseq-mouse-scdd.Rmd', clean=FALSE)"
+R -e "rmarkdown::render('scrnaseq-mouse-mast.Rmd', clean=FALSE)"
+R -e "rmarkdown::render('scrnaseq-human-scdd.Rmd', clean=FALSE)"
+R -e "rmarkdown::render('scrnaseq-human-mast.Rmd', clean=FALSE)"

### 2018-03-08 datasets/ChIP-seq/ChIPseq-csaw.Rmd
Start working on csaw, diffbind analysis

---
title: "Case Study: ChIP-seq Differential Peak Calling (csaw)"
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

In this case study, we perform differential peak calling on ChIP-seq data of the histone
H3K4Me3 for samples from two cell lines (K562 and Gm12878) using publicly available data
generated by the ENCODE Project. The same data set is used for all ChIP-seq differential
testing case studies.

The _csaw_ package provides a powerful and flexible approach to differential peak calling
based on counting reads in sliding windows across the genome. Unlike the _DiffBind_ approach
which restricts testing to peaks called from the same samples, the sliding window approach
is able to detect differential peaks over sharper or wider regions, without being constrained
by the original peak boundaries obtained from the peak calling algorithm, e.g. _MACS_.
Differential peaks are called at each window using _edgeR_ and p-values are combined within
adjacent windows (clusters).

For this analysis, the default _csaw_ settings are used as described in the _csaw_ "User Guide."

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(SummarizedBenchmark)
library(BiocParallel)
library(rtracklayer)
library(Rsamtools)
library(csaw)
library(edgeR)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "data"
resdir <- "results"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)

## intermediary files we create below
count_file <- file.path(resdir, "h3k4me3-csaw-counts.rds")
result_file <- file.path(resdir, "h3k4me3-csaw-results.rds")
bench_file <- file.path(resdir, "h3k4me3-csaw-benchmark.rds")

## set up parallel backend
cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
multicoreParam <- MulticoreParam(workers = cores)
```

# Data Preparation

We download the bam files directly from the UCSC ENCODE portal.

```{r, encode-download}
broad_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeBroadHistone/"
broad_bams <- c("wgEncodeBroadHistoneGm12878H3k4me3StdAlnRep1.bam",
                "wgEncodeBroadHistoneGm12878H3k04me3StdAlnRep2V2.bam",
                "wgEncodeBroadHistoneK562H3k4me3StdAlnRep1.bam",
                "wgEncodeBroadHistoneK562H3k4me3StdAlnRep2.bam")

for (i_bam in broad_bams) {
    if (!file.exists(file.path(datdir, i_bam))) {
        download.file(paste0(broad_url, i_bam),
                      destfile = file.path(datdir, i_bam))
    }
    i_bai <- paste0(i_bam, ".bai")
    if (!file.exists(file.path(datdir, i_bai))) {
        download.file(paste0(broad_url, i_bai),
                      destfile = file.path(datdir, i_bai))
    }
}

uw_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwHistone/"
uw_bams <- c("wgEncodeUwHistoneGm12878H3k4me3StdAlnRep1.bam",
             "wgEncodeUwHistoneGm12878H3k4me3StdAlnRep2.bam",
             "wgEncodeUwHistoneK562H3k4me3StdAlnRep1.bam",
             "wgEncodeUwHistoneK562H3k4me3StdAlnRep2.bam")

for (i_bam in uw_bams) {
    if (!file.exists(file.path(datdir, i_bam))) {
        download.file(paste0(uw_url, i_bam),
                      destfile = file.path(datdir, i_bam))
    }
    i_bai <- paste0(i_bam, ".bai")
    if (!file.exists(file.path(datdir, i_bai))) {
        download.file(paste0(uw_url, i_bai),
                      destfile = file.path(datdir, i_bai))
    }
}
```

We determine sample metadata from the file names.

```{r}
bamfiles <- file.path(datdir, c(broad_bams, uw_bams))
labs <- gsub("wgEncode(.*)Histone.*", "\\1", basename(bamfiles))
cells <- gsub("wgEncode.*Histone(.*)H3k.*", "\\1", basename(bamfiles))

meta <- data.frame(cellline = cells, lab = labs, file = bamfiles,
                   sorted = gsub("\\.bam$", ".sorted.bam", bamfiles),
                   stringsAsFactors = FALSE)
```

We also download blacklisted regions for hg19/GRCh37 from ENCODE (https://www.encodeproject.org/annotations/ENCSR636HFF/).

```{r}
blacklist_url <- "https://www.encodeproject.org/files/ENCFF001TDO/@@download/ENCFF001TDO.bed.gz"
if (!file.exists(file.path(datdir, basename(blacklist_url)))) {
    download.file(blacklist_url, destfile = file.path(datdir, basename(blacklist_url)))
}

bl <- import(file.path(datdir, basename(blacklist_url)))
```

Standard set of parameters are defined for the analysis. Only the canonical set of chromosomes are used in the analysis.

```{r}
std_chr <- paste0("chr", c(1:19, "X", "Y"))
param <- readParam(minq = 50, discard = bl, restrict = std_chr)
```

## Read Counting

We count reads within sliding windows across the genome.

```{r count, results = 'hide', message = FALSE, warning = FALSE}
if (file.exists(count_file)) {
    filtered_cnts <- readRDS(count_file)
} else {
    ## sort bam files
    for (i in 1:nrow(meta)) {
        if (!file.exists(meta$sorted[i])) {
            sortBam(meta$file[i], gsub("\\.bam$", "", meta$sorted[i]))
        }
    }

    ## mark duplicates w/ picard
    temp_bam <- "temp_dups.bam"
    temp_file <- "temp_mets.txt"
    temp_dir <- "data/temp_dups"
    dir.create(temp_dir)
    for (i_bam in meta$sorted) {
        code <- paste0("java -jar ${PICARD_TOOLS_HOME}/picard.jar ",
                       "MarkDuplicates I=%s O=%s M=%s ", 
                       "TMP_DIR=%s AS=true REMOVE_DUPLICATES=false ",
                       "VALIDATION_STRINGENCY=SILENT")
        code <- sprintf(code, i_bam, temp_bam, temp_file, temp_dir)
        code <- system(code)
        stopifnot(code == 0L)
        file.rename(temp_bam, i_bam)
    }
    unlink(temp_file)
    unlink(temp_dir, recursive = TRUE)

    ## index bam files
    indexBam(meta$sorted)

    ## use correlateReads to determine fragment length (remove dups)
    x <- correlateReads(meta$sorted, param = reform(param, dedup = TRUE))
    frag_len <- which.max(x) - 1

    ## count reads in sliding windows (keep dups)
    win_cnts <- windowCounts(meta$sorted, param = param,
                             width = 150, ext = frag_len)

    ## filter windows by abundance
    bins <- windowCounts(meta$file, bin = TRUE, width = 2000, param = param)
    filter_stat <- filterWindows(win_cnts, bins, type = "global")
    keep <- filter_stat$filter > log2(3)
    filtered_cnts <- win_cnts[keep, ]

    saveRDS(filtered_cnts, file = count_file)
}
```

# Data Analysis

## Differential Testing

We use _edgeR_ to test for differential binding.

```{r test, results='hide', message=FALSE, warning=FALSE}
if (file.exists(result_file)) {
    res_ranges <- readRDS(result_file)
} else {

    ## normalize for library-specific trended biases
    offsets <- normOffsets(filtered_cnts, type = "loess")

    ## create model design
    design <- model.matrix(~ lab + cellline,
                           data = as.data.frame(meta))

    ## run quasi-likelihood F-test
    y <- asDGEList(filtered_cnts)
    y$offset <- offsets
    y <- estimateDisp(y, design)
    fit <- glmQLFit(y, design, robust = TRUE)
    res <- glmQLFTest(fit)

    ## merge p-values across regions
    merged <- mergeWindows(rowRanges(filtered_cnts), tol = 100, max.width = 5000)
    tab_comb <- combineTests(merged$id, res$table)
    tab_best <- getBestTest(merged$id, res$table)

    ## save results
    res_ranges <- merged$region
    elementMetadata(res_ranges) <-
        data.frame(tab_comb,
                   best_pos = mid(ranges(rowRanges(filtered_cnts[tab_best$best]))),
                   best_logFC = tab_best$logFC)

    saveRDS(res_ranges, file = result_file)
}
```

# Session Info

```{r}
sessionInfo()
```

### 2018-03-08 datasets/ChIP-seq/ChIPseq-diffbind.Rmd
Start working on csaw, diffbind analysis

---
title: "Case Study: ChIP-seq Differential Peak Calling (DiffBind)"
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

In this case study, we perform differential peak calling on ChIP-seq data of the histone
H3K4Me3 for samples from two cell lines (K562 and Gm12878) using publicly available data
generated by the ENCODE Project. The same data set is used for all ChIP-seq differential
testing case studies.

Here, we use a popular differential peak calling algorithm, _DiffBind_ (Stark and Brown, 2011).
Rather than using pre-defined regions (e.g. promoters) or scanning the entire genome
(see _csaw_ case study), the _DiffBind_ approach starts from peak calls obtained using a
separate peak calling algorithm, e.g. _MACS_. Differential peaks are called among the
combined set of significant peaks across samples using methods originally developed for
differential expression testing, namely _DESeq2_ or _edgeR_.

For this analysis, peak calls are obtained from the ENCODE Project pipeline, and
testing is performed using _DiffBind_ and _DESeq2_.

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(SummarizedBenchmark)
library(BiocParallel)
library(DiffBind)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "data"
resdir <- "results"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)

## intermediary files we create below
count_file <- file.path(resdir, "h3k4me3-diffbind-counts.rds")
result_file <- file.path(resdir, "h3k4me3-diffbind-results.rds")
bench_file <- file.path(resdir, "h3k4me3-diffbind-benchmark.rds")

## set up parallel backend
cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
multicoreParam <- MulticoreParam(workers = cores)
```

# Data Preparation

We downloaded the peak calls directly from the UCSC ENCODE portal.

```{r}
broad_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeBroadHistone/"
broad_peaks <- c("wgEncodeBroadHistoneGm12878H3k4me3StdPkRep1.narrowPeak.gz",
                 "wgEncodeBroadHistoneGm12878H3k04me3StdPkRep2V2.narrowPeak.gz",
                 "wgEncodeBroadHistoneK562H3k4me3StdPkRep1.narrowPeak.gz",
                 "wgEncodeBroadHistoneK562H3k4me3StdPkRep2.narrowPeak.gz")

for (i_peak in broad_peaks) {
    if (!file.exists(file.path(datdir, i_peak))) {
        download.file(paste0(broad_url, i_peak),
                      destfile = file.path(datdir, i_peak))
    }
}

uw_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwHistone/"
uw_peaks <- c("wgEncodeUwHistoneGm12878H3k4me3StdPkRep1.narrowPeak.gz",
              "wgEncodeUwHistoneGm12878H3k4me3StdPkRep2.narrowPeak.gz",
              "wgEncodeUwHistoneK562H3k4me3StdPkRep1.narrowPeak.gz",
              "wgEncodeUwHistoneK562H3k4me3StdPkRep2.narrowPeak.gz")

for (i_peak in uw_peaks) {
    if (!file.exists(file.path(datdir, i_peak))) {
        download.file(paste0(uw_url, i_peak),
                      destfile = file.path(datdir, i_peak))
    }
}
```

We determine sample metadata from the file names.

```{r}
peakfiles <- file.path(datdir, c(broad_peaks, uw_peaks))
labs <- gsub("wgEncode(.*)Histone.*", "\\1", basename(peakfiles))
cells <- gsub("wgEncode.*Histone(.*)H3k.*", "\\1", basename(peakfiles))

meta <- data.frame(cellline = cells, lab = labs, file = peakfiles,
                   stringsAsFactors = FALSE)
```

## Read Counting

We determine the combined set of peak regions and count reads for each file.

```{r}
if (file.exists(count_file)) {
    cnts <- readRDS(count_file)
} else {
    ## run diffbind
    ## saveRDS(cnts, file = count_file)
}
```

# Data Analysis

## Differential Testing

First, we use _DiffBind_ to test for differential binding.

```{r count, results='hide', message=FALSE, warning=FALSE}
if (file.exists(result_file)) {
    res <- readRDS(result_file)
} else {
    ## run diffbind
    ## saveRDS(res, file = result_file)
}
```

# Session Info

```{r}
sessionInfo()
```

### 2018-03-08 datasets/ChIP-seq/ChIPseq-promoters.Rmd
Start working on csaw, diffbind analysis

--- 2018-02-27 datasets/ChIP-seq/ChIPseq-promotors.Rmd
@@
 ---
-title: "Case Study: ChIP-seq Differential Peak Calling (predefined regions)"
-author: "Mingxiang Teng"
+title: "Case Study: ChIP-seq Differential Peak Calling (pre-defined regions)"
+author: "Mingxiang Teng, Patrick Kimes"
 date: '`r format(Sys.Date(), "%B %d, %Y")`'
@@
 
-In differential peak calling, classical pipelines often begin by first calling peaks
-in samples, followed by performing differential testing using one of a few methods,
-e.g. _csaw_ or _DiffBind_. However, with these approaches, testing is only performed on a
-filtered set of genomic regions, namely those with significance in at least one condition,
-resulting in a potentially limited number of null tests cases. An analysis using this
-classical approach is performed in the accompanying `ChIPseq-peaks.Rmd` case study.
+In this case study, we perform differential peak calling on ChIP-seq data of the histone
+H3K4Me3 for samples from two cell lines (K562 and Gm12878) using publicly available data
+generated by the ENCODE Project. The same data set is used for all ChIP-seq differential
+testing case studies.
 
-To avoid this limitation, in this case study we examine ChIP-seq data of the histone
-H3K4Me3, an active marker of gene expression, and focus our analysis on promoter regions
-of all genes for this histone marker. By doing so, we avoid limiting the set of tests
-to only regions significant in at least one sample, while also keeping most regions of
-interest for this histone marker.
-
-As in `ChIPseq-peaks.Rmd`, we test for differential peaks using publicly available ChIP-seq
-data for histone H3K4Me3 from two cell lines (K562 and Gm12878). Data are obtained from the
-ENCODE Project. Differential peaks are tested using DESeq2.
+While several approaches have been proposed for differential peak calling, in this case
+study we examine a simple approach of testing for differential peaks in the promoter 
+regions of all genes. While using pre-defined regions can reduce the power of the test to
+detect differences in non-promoter regions, since H3K4Me3 is an active marker of gene
+expression, restricting tests to promoter regions is not unreasonable. Differential peaks
+are tested using DESeq2.
 
@@
 dir.create(resdir, showWarnings = FALSE)
+
+## intermediary files we create below
+count_file <- file.path(resdir, "h3k4me3-promotors-counts.rda")
+result_file <- file.path(resdir, "h3k4me3-promotors-results.rda")
+bench_file <- file.path(resdir, "h3k4me3-promotors-benchmark.rda")
 
@@
 
-# Data Analysis
+## Read Counting
 
-## Differential Testing
-
-We count sequencing reads for all promoter regions here, and test for differential peaks using these counts.
+We count reads for all promoter regions.
 
 ```{r count, results='hide', message=FALSE, warning=FALSE}
-deseq_res  <- file.path(resdir, "h3k4me3-promotors-deseq.rda")
-if (file.exists(deseq_res)) {
-    load(deseq_res)
+if (file.exists(count_file)) {
+    load(count_file)
 } else {
-    ## count reads in each promotor region
     anno <- data.frame(GeneID = seq_len(length(prom)),
@@
                         ignoreDup = TRUE)$counts
+    save(rc, file = count_file)
+}
+```
 
-    ## perform differential testing
+# Data Analysis
+
+## Differential Testing
+
+We test for differential peaks at each of the .
+
+```{r count, results='hide', message=FALSE, warning=FALSE}
+if (file.exists(result_file)) {
+    load(result_file)
+} else {
     dds <- DESeqDataSetFromMatrix(countData = rc,
                                   colData = as.data.frame(meta),
-                                  design= ~ lab + cellline)
+                                  design = ~ lab + cellline)
     dds <- DESeq(dds, fitType = "mean")
@@
                        'stat', 'pval', 'padj')
-
     ## compute test-statistic from p-value for ASH
@@
                          test_statistic = test_statistic * sign(effect_size))
-
-    save(deseq, dat, file = deseq_res)
+    ## remove regions with effect sizes = NA
+    dat <- dat[!is.na(dat$effect_size), ]
+    save(dat, file = result_file)
 }
-
-## remove regions with effect sizes = NA
-dat <- dat[!is.na(dat[, 2]), ]
 ```
@@
 
-Here, we can see that the covariant variable is actually informative.
+Here, we can check diagnostic plots to assess whether candidate covariates are actually informative.
 
@@
 ```{r, chipseq-sb, results="hide", message=FALSE}
-bench_res  <- file.path(resdir, "h3k4me3-promotors-sb.rda")
-if (file.exists(bench_res)) {
-    load(bench_res)
+if (file.exists(bench_file)) {
+    load(bench_file)
 } else {
@@
                      parallel = TRUE, BPPARAM = multicoreParam)
-    save(sb, file = bench_res)
+    save(sb, file = bench_file)
 }

### 2018-03-08 datasets/microbiome/microbiome-baxter.Rmd
Run only single nmids in microbiome; save to rds (not rdata)

@@
 dir.create(resdir, showWarnings = FALSE)
+
+crc_data <- file.path(datdir, "crc_baxter_results.tar.gz")
+crc_dir <- file.path(datdir, "crc_baxter_results")
+
+otu_result_file <- file.path(resdir, "baxter-otus-results.rds")
+otu_bench_file <- file.path(resdir, "baxter-otus-benchmark.rds")
+gns_result_file <- file.path(resdir, "baxter-genus-results.rds")
+gns_bench_file <- file.path(resdir, "baxter-genus-benchmark.rds")
 ```
@@
 ```{r, baxter-download}
-crc_data <- file.path(datdir, "crc_baxter_results.tar.gz")
 if (!file.exists(crc_data)) {
@@
 
-crc_dir <- file.path(datdir, "crc_baxter_results")
 if (!file.exists(file.path(crc_dir, "crc_baxter.metadata.txt"))) {
@@
 ```{r, baxter-pvals}
-resfile <- file.path(resdir, paste0("baxter_results_", nrow(abun_otu), "_OTUs.rda"))
-if (!file.exists(resfile)) {
+if (!file.exists(otu_result_file)) {
     res <- test_microbiome(abundance = abun_otu, shift = zeroabun,
                            is_case = meta$DiseaseState == "CRC")
-    save(res, file = resfile)
+    saveRDS(res, file = otu_result_file)
 } else {
-    load(resfile)
+    res <- readRDS(otu_result_file)
 }
@@
                          "family", "genus", "species", "denovo"), 
-                       sep=";", remove = FALSE)
+                       sep = ";", remove = FALSE)
 ```
@@
 
-Not really informative either...
+Not really informative either.
 
@@
 ```{r, baxter-sb, results="hide", message=FALSE, eval=TRUE}
-for (nmids in c(5, 50, 150)){
-    print(nmids)
-    bd <- initializeBenchDesign(nmids=nmids)
-    ## remove ASH from comparison
+if (!file.exists(otu_bench_file)) {
+    nmids <- 150
+    bd <- initializeBenchDesign(nmids = nmids)
     bd <- dropBMethod(bd, "ashs")
-
-    resfile <- file.path(resdir, paste0("baxter_summarizedBenchmark_",
-                                        nrow(res), "_nmids_", nmids, ".rda"))
-    if (!file.exists(resfile)){
     sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
@@
                    "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
-        save(sb, file = resfile)
+    metadata(sb)$nmids <- nmids
+    saveRDS(sb, file = otu_bench_file)
 } else {
-        load(resfile)
-    }
+    sb <- readRDS(otu_bench_file)
 }
@@
 ```{r, baxter-plot, width=15, height=15, eval=TRUE}
-## plot nrejects by method overall and stratified by covariate
 rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
 rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
-##plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
 ```
@@
 ```{r, baxter-genus-pvals}
-resfile <- file.path(resdir, paste0("baxter_results_", nrow(genus_df), "_genera.rda"))
-if (!file.exists(resfile)) {
+if (!file.exists(gns_result_file)) {
     res <- test_microbiome(abundance = genus, shift = zeroabun,
                            is_case = meta$DiseaseState == "CRC")
-    save(res, file = resfile)
+    saveRDS(res, file = gns_result_file)
 } else {
-    load(resfile)
+    res <- readRDS(gns_result_file)
 }
@@
 ```{r, baxter-genus-sb, results="hide", message=FALSE, eval=TRUE}
-for (nmids in c(5, 50, 150)) {
-    ## use common BenchDesign w/out ASH
-    bd <- initializeBenchDesign(nmids=nmids)
+if (!file.exists(gns_bench_file)) {
+    nmids <- 150
+    bd <- initializeBenchDesign(nmids = nmids)
     bd <- dropBMethod(bd, "ashs")
-
-    resfile <- file.path(resdir, paste0("baxter_summarizedBenchmark_", nrow(res),
-                                        "_genera_nmids_", nmids, ".rda"))
-    if (!file.exists(resfile)) {
     sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
@@
                    "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
-        save(sb, file = resfile)
+    metadata(sb)$nmids <- nmids
+    saveRDS(sb, file = gns_bench_file)
 } else {
-        load(resfile)
-    }
+    sb <- readRDS(gns_bench_file)
 }
@@
 ```{r, baxter-genus-plot, width=15, height=15, eval=TRUE}
-## plot nrejects by method overall and stratified by covariate
-rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
 rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
-##plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
 ```
@@
 methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
-plotCovariateBoxplots( sb, alpha=0.1, nsets=6, methods=methods)
+plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods)
 ```

### 2018-03-08 datasets/microbiome/microbiome-schubert.Rmd
Run only single nmids in microbiome; save to rds (not rdata)

@@
 dir.create(resdir, showWarnings = FALSE)
+
+cdi_data <- file.path(datdir, "cdi_schubert_results.tar.gz")
+cdi_dir <- file.path(datdir, "cdi_schubert_results")
+
+otu_result_file <- file.path(resdir, "schubert-otus-results.rds")
+otu_bench_file <- file.path(resdir, "schubert-otus-benchmark.rds")
 ```
@@
 ```{r, schubert-download}
-cdi_data <- file.path(datdir, "cdi_schubert_results.tar.gz")
 if (!file.exists(cdi_data)) {
@@
 
-cdi_dir <- file.path(datdir, "cdi_schubert_results")
 if (!file.exists(file.path(cdi_dir, "cdi_schubert.metadata.txt"))) {
@@
 ```{r, schubert-pvals, eval=TRUE}
-resfile <- file.path(resdir, paste0("schubert_results_", nrow(abun_otu), "_OTUs.rda"))
-if (!file.exists(resfile)) {
+if (!file.exists(otu_result_file)) {
     res <- test_microbiome(abundance = abun_otu, shift = zeroabun,
                            is_case = meta$DiseaseState %in% c("CDI", "nonCDI"))
-    save(res, file = resfile)
+    saveRDS(res, file = otu_result_file)
 } else {
-  load(resfile)
+    res <- readRDS(otu_result_file)
 }
@@
 ```{r, schubert-sb, results="hide", message=FALSE, eval=TRUE}
-for (nmids in c(5, 50, 150)) {
-    ## use common BenchDesign w/out ASH
+if (!file.exists(otu_bench_file)) {
+    nmids <- 150
     bd <- initializeBenchDesign(nmids = nmids)
     bd <- dropBMethod(bd, "ashs")
-
-    resfile <- file.path(resdir, paste0("schubert_summarizedBenchmark_",
-                                        nrow(res),  "_nmids_", nmids, ".rda"))
-    if (!file.exists(resfile)) {
     sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
@@
                    "https://zenodo.org/record/840333/files/cdi_schubert_results.tar.gz"
-        save(sb, file = resfile)
+    saveRDS(sb, file = otu_bench_file)
 } else {
-        load(resfile)
-    }
+    sb <- readRDS(otu_bench_file)
 }
@@
 ```{r, schubert-plot, width=15, height=15, eval=TRUE}
-## plot nrejects by method overall and stratified by covariate
 rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)

### 2018-03-08 datasets/ChIP-seq/ChIPseq-csaw.Rmd
Fix typo in var name for counting

@@
     ## filter windows by abundance
-    bins <- windowCounts(meta$file, bin = TRUE, width = 2000, param = param)
+    bins <- windowCounts(meta$sorted, bin = TRUE, width = 2000, param = param)
     filter_stat <- filterWindows(win_cnts, bins, type = "global")

### 2018-03-09 datasets/ChIP-seq/ChIPseq-csaw.Rmd
Drop ENCODE BAMs; align FASTQs

--- 2018-03-08 datasets/ChIP-seq/ChIPseq-csaw.Rmd
@@
 library(Rsamtools)
+library(Rsubread)
 library(csaw)
@@
 
-We download the bam files directly from the UCSC ENCODE portal.
+We download the fastq files directly from the UCSC ENCODE portal and align them against the reference.
 
-```{r, encode-download}
+```{r, download-fastqs}
 broad_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeBroadHistone/"
-broad_bams <- c("wgEncodeBroadHistoneGm12878H3k4me3StdAlnRep1.bam",
-                "wgEncodeBroadHistoneGm12878H3k04me3StdAlnRep2V2.bam",
-                "wgEncodeBroadHistoneK562H3k4me3StdAlnRep1.bam",
-                "wgEncodeBroadHistoneK562H3k4me3StdAlnRep2.bam")
+broad_fastqs <- c("wgEncodeBroadHistoneGm12878H3k4me3StdRawDataRep1.fastq.gz",
+                  "wgEncodeBroadHistoneGm12878H3k04me3StdRawDataRep2V2.fastq.gz",
+                  "wgEncodeBroadHistoneK562H3k4me3StdRawDataRep1.fastq.gz",
+                  "wgEncodeBroadHistoneK562H3k4me3StdRawDataRep2.fastq.gz")
 
-for (i_bam in broad_bams) {
-    if (!file.exists(file.path(datdir, i_bam))) {
-        download.file(paste0(broad_url, i_bam),
-                      destfile = file.path(datdir, i_bam))
-    }
-    i_bai <- paste0(i_bam, ".bai")
-    if (!file.exists(file.path(datdir, i_bai))) {
-        download.file(paste0(broad_url, i_bai),
-                      destfile = file.path(datdir, i_bai))
+for (i_fq in broad_fastqs) {
+    if (!file.exists(file.path(datdir, i_fq))) {
+        download.file(paste0(broad_url, i_fq),
+                      destfile = file.path(datdir, i_fq))
     }
@@
 uw_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwHistone/"
-uw_bams <- c("wgEncodeUwHistoneGm12878H3k4me3StdAlnRep1.bam",
-             "wgEncodeUwHistoneGm12878H3k4me3StdAlnRep2.bam",
-             "wgEncodeUwHistoneK562H3k4me3StdAlnRep1.bam",
-             "wgEncodeUwHistoneK562H3k4me3StdAlnRep2.bam")
+uw_fastqs <- c("wgEncodeUwHistoneGm12878H3k4me3StdRawDataRep1.fastq.gz",
+               "wgEncodeUwHistoneGm12878H3k4me3StdRawDataRep2.fastq.gz",
+               "wgEncodeUwHistoneK562H3k4me3StdRawDataRep1.fastq.gz",
+               "wgEncodeUwHistoneK562H3k4me3StdRawDataRep2.fastq.gz")
 
-for (i_bam in uw_bams) {
-    if (!file.exists(file.path(datdir, i_bam))) {
-        download.file(paste0(uw_url, i_bam),
-                      destfile = file.path(datdir, i_bam))
+for (i_fq in uw_fastqs) {
+    if (!file.exists(file.path(datdir, i_fq))) {
+        download.file(paste0(uw_url, i_fq),
+                      destfile = file.path(datdir, i_fq))
     }
-    i_bai <- paste0(i_bam, ".bai")
-    if (!file.exists(file.path(datdir, i_bai))) {
-        download.file(paste0(uw_url, i_bai),
-                      destfile = file.path(datdir, i_bai))
 }
+```
+
+We also use the GRCh38 reference genome used for ENCODE3 analysis. If the reference index is
+not already available, we must first download the reference file. Note that this is a different
+reference than what was originally used for the data.
+
+```{r, download-ref}
+ref_url <- paste0("https://www.encodeproject.org/files/",
+                  "GRCh38_no_alt_analysis_set_GCA_000001405.15/@@download/",
+                  "GRCh38_no_alt_analysis_set_GCA_000001405.15.fasta.gz")
+ref_fastagz <- file.path(datdir, basename(ref_url))
+ref_fasta <- gsub("\\.gz$", "", ref_fastagz)
+
+if (!file.exists(file.path(datdir, "ref_index.00.b.tab"))) {
+    if (!file.exists(ref_fasta)) {
+        download.file(ref_url, destfile = ref_fastagz)
+        system(paste("gunzip", ref_fastagz))
+    }
+    buildindex(basename = file.path(datdir, "ref_index"),
+               reference = ref_fasta)
 }
@@
 ```{r}
-bamfiles <- file.path(datdir, c(broad_bams, uw_bams))
-labs <- gsub("wgEncode(.*)Histone.*", "\\1", basename(bamfiles))
-cells <- gsub("wgEncode.*Histone(.*)H3k.*", "\\1", basename(bamfiles))
+fqfiles <- file.path(datdir, c(broad_fastqs, uw_fastqs))
+labs <- gsub("wgEncode(.*)Histone.*", "\\1", basename(fqfiles))
+cells <- gsub("wgEncode.*Histone(.*)H3k.*", "\\1", basename(fqfiles))
 
-meta <- data.frame(cellline = cells, lab = labs, file = bamfiles,
-                   sorted = gsub("\\.bam$", ".sorted.bam", bamfiles),
+meta <- data.frame(cellline = cells, lab = labs,
+                   fqfile = fqfiles,
+                   bamfile = gsub("\\.fastq.gz", ".bam", fqfiles),
+                   sortedfile = gsub("\\.fastq.gz", ".sorted.bam", fqfiles),
                    stringsAsFactors = FALSE)
@@
 
-We also download blacklisted regions for hg19/GRCh37 from ENCODE (https://www.encodeproject.org/annotations/ENCSR636HFF/).
+We also download blacklisted regions for GRCh38 from ENCODE (https://www.encodeproject.org/annotations/ENCSR636HFF/).
 
 ```{r}
-blacklist_url <- "https://www.encodeproject.org/files/ENCFF001TDO/@@download/ENCFF001TDO.bed.gz"
+blacklist_url <- paste0("http://mitra.stanford.edu/kundaje/akundaje/",
+                        "release/blacklists/hg38-human/hg38.blacklist.bed.gz")
 if (!file.exists(file.path(datdir, basename(blacklist_url)))) {
@@
 } else {
+    ## align reads to reference
+    unaligned <- !file.exists(meta$bamfile)
+    if (any(unaligned)) {
+        align(index = file.path(datdir, "ref_index"),
+              readfile1 = meta$fqfile[unaligned],
+              TH1 = 2, type = 1,
+              output_file = meta$bamfile[unaligned])
+    }
+
     ## sort bam files
     for (i in 1:nrow(meta)) {
-        if (!file.exists(meta$sorted[i])) {
-            sortBam(meta$file[i], gsub("\\.bam$", "", meta$sorted[i]))
+        if (!file.exists(meta$sortedfile[i])) {
+            sortBam(meta$bamfile[i], gsub("\\.bam$", "", meta$sortedfile[i]))
         }
@@
     dir.create(temp_dir)
-    for (i_bam in meta$sorted) {
+    for (i_bam in meta$sortedfile) {
         code <- paste0("java -jar ${PICARD_TOOLS_HOME}/picard.jar ",
@@
     ## index bam files
-    indexBam(meta$sorted)
+    indexBam(meta$sortedfile)
 
     ## use correlateReads to determine fragment length (remove dups)
-    x <- correlateReads(meta$sorted, param = reform(param, dedup = TRUE))
+    x <- correlateReads(meta$sortedfile, param = reform(param, dedup = TRUE))
     frag_len <- which.max(x) - 1
@@
     ## count reads in sliding windows (keep dups)
-    win_cnts <- windowCounts(meta$sorted, param = param,
+    win_cnts <- windowCounts(meta$sortedfile, param = param,
                              width = 150, ext = frag_len)
@@
     ## filter windows by abundance
-    bins <- windowCounts(meta$file, bin = TRUE, width = 2000, param = param)
+    bins <- windowCounts(meta$sortedfile, bin = TRUE, width = 2000, param = param)
     filter_stat <- filterWindows(win_cnts, bins, type = "global")

### 2018-03-09 datasets/ChIP-seq/ChIPseq-csaw.Rmd
Specify phred+64 when aligning UW samples

@@
 } else {
-    ## align reads to reference
-    unaligned <- !file.exists(meta$bamfile)
-    if (any(unaligned)) {
+    ## Broad samples have quality score w/ phred offset +33 (standard)
+    b_meta <- meta[meta$lab == "Broad", ]
+    b_unaligned <- !file.exists(b_meta$bamfile)
+    if (any(b_unaligned)) {
         align(index = file.path(datdir, "ref_index"),
-              readfile1 = meta$fqfile[unaligned],
-              TH1 = 2, type = 1,
-              output_file = meta$bamfile[unaligned])
+              readfile1 = b_meta$fqfile[b_unaligned],
+              TH1 = 2, type = 1, phredOffset = 33,
+              output_file = b_meta$bamfile[b_unaligned])
+    }
+
+    ## UW samples have quality score w/ phred offset +64 (old Illumina scale)
+    u_meta <- meta[meta$lab == "Uw", ]
+    u_unaligned <- !file.exists(u_meta$bamfile)
+    if (any(u_unaligned)) {
+        align(index = file.path(datdir, "ref_index"),
+              readfile1 = u_meta$fqfile[u_unaligned],
+              TH1 = 2, type = 1, phredOffset = 64,
+              output_file = u_meta$bamfile[u_unaligned])
     }

### 2018-03-11 datasets/ChIP-seq/ChIPseq-csaw.Rmd

@@
 In this case study, we perform differential peak calling on ChIP-seq data of the histone
-H3K4Me3 for samples from two cell lines (K562 and Gm12878) using publicly available data
+H3K4Me3 for samples from two cell lines (K562 and GM12878) using publicly available data
 generated by the ENCODE Project. The same data set is used for all ChIP-seq differential
@@
 std_chr <- paste0("chr", c(1:19, "X", "Y"))
-param <- readParam(minq = 50, discard = bl, restrict = std_chr)
+param <- readParam(minq = 15, discard = bl, restrict = std_chr)
 ```
@@
     ## mark duplicates w/ picard
-    temp_bam <- "temp_dups.bam"
-    temp_file <- "temp_mets.txt"
-    temp_dir <- "data/temp_dups"
+    temp_bam <- file.path(datdir, "temp_dups.bam")
+    temp_file <- file.path(datdir, "temp_mets.txt")
+    temp_dir <- file.path(datdir, "temp_dups")
     dir.create(temp_dir)
@@
     ## use correlateReads to determine fragment length (remove dups)
-    x <- correlateReads(meta$sortedfile, param = reform(param, dedup = TRUE))
-    frag_len <- which.max(x) - 1
+    b_x <- correlateReads(b_meta$sortedfile, param = reform(param, dedup = TRUE))
+    u_x <- correlateReads(u_meta$sortedfile, param = reform(param, dedup = TRUE))
+    b_frag_len <- which.max(b_x) - 1
+    u_frag_len <- which.max(u_x) - 1
 
     ## count reads in sliding windows (keep dups)
-    win_cnts <- windowCounts(meta$sortedfile, param = param,
-                             width = 150, ext = frag_len)
+    win_cnts <- windowCounts(meta$sortedfile, param = param, width = 150,
+                             ext = list(rep(c(b_frag_len, u_frag_len), each = 4), 150))
 
@@
 } else {
-
     ## normalize for library-specific trended biases

### 2018-03-11 datasets/GSEA/gsea-human.Rmd
Make plot sizes and types uniform

@@
 ---
-title: "Case Study: Gene Set Enrichment Analysis (GSEA)"
+title: "Case Study: Gene Set Enrichment Analysis (Human Data Set)"
 author: "Alejandro Reyes"
@@
 dir.create(resdir, showWarnings = FALSE)
+
+## intermediary files we create below
+count_file <- file.path(datdir, "human-counts.rds")
+deseq_file <- file.path(datdir, "human-deseq.rds")
+goset_file <- file.path(datdir, "human-gosets.rds")
+result_file <- file.path(resdir, "human-results.rds")
+bench_file <- file.path(resdir, "human-benchmark.rds")
 ```
@@
 ```{r, message=FALSE}
-dsd_data <- file.path(datdir, "dsd.RData")
-if (!file.exists(dsd_data)) {
-    download.file("https://www.dropbox.com/s/rdmv3tyqt4g37b0/dsd.RData?dl=1",
-                  destfile = dsd_data)
+if (!file.exists(count_file)) {
+    ## download.file("", destfile = count_file)
 }
-load(dsd_data)
+dsdObject <- readRDS(count_file)
 ```
@@
 gnType <- genes(EnsDb.Hsapiens.v86, column = "gene_biotype")
-dsdObject <- dsdObject[rownames(dsdObject) %in% gnType[gnType$gene_biotype == "protein_coding"]$gene_id, ]
+protGenes <- gnType$gene_id[gnType$gene_biotype == "protein_coding"]
+dsdObject <- dsdObject[rownames(dsdObject) %in% protGenes, ]
 ```
@@
 
-## Differential Testing
-
-Specify model matrix design and run DESeq2 to get differentially expressed genes.
-
-```{r deseq2, message=FALSE}
-dsdObject <- DESeq(dsdObject)
-res <- results(dsdObject, independentFiltering = FALSE)
-```
-
 ## Enrichment Analysis
@@
 **Add description of what's going on.**
+We first test for differentially expressed genes using DESeq2 and use genes with adjusted p-value less than 0.1.
 
-```{r goseq, message=FALSE}
-goset_data <- file.path(datdir, "goSets.RData")
-if (!file.exists(goset_data)) {
-    mart <- useMart("ensembl", "hsapiens_gene_ensembl")
-    goSets <- getBM( c("ensembl_gene_id", "go_id"), mart = mart,
-                    filters = "ensembl_gene_id", values = rownames(res))
-    goSets <- goSets[!nchar( goSets$go_id ) == 0, ]
-    goSets <- with(goSets, split(go_id, ensembl_gene_id))
-    save(goSets, file = goset_data)
+```{r deseq2, message=FALSE}
+if (!file.exists(deseq_file)) {
+    dsdObject <- DESeq(dsdObject)
+    res <- results(dsdObject, independentFiltering = FALSE)
+    saveRDS(res, file = deseq_file)
 } else {
-    load(goset_data)
+    res <- readRDS(deseq_file)
 }
@@
 names(genes) <- rownames(res)
-
-### getting median transcript length
-txByGene <- transcriptsBy(EnsDb.Hsapiens.v86, "gene")
-geneLength <- sapply( width( txByGene ), median )
-geneLength <- geneLength[names(genes)]
-genes[is.na( genes )] <- 0
 ```
 
-Performing a gene set enrichment analysis with `goseq`.
+We next determine the gene sets associated with each gene.
+
+```{r goseq, message=FALSE}
+if (!file.exists(goset_file)) {
+    mart <- useMart("ensembl", "hsapiens_gene_ensembl")
+    goSets <- getBM(c("ensembl_gene_id", "go_id"), mart = mart,
+                    filters = "ensembl_gene_id", values = rownames(res))
+    goSets <- goSets[!nchar( goSets$go_id ) == 0, ]
+    goSets <- with(goSets, split(go_id, ensembl_gene_id))
+    saveRDS(goSets, file = goset_file)
+} else {
+    goSets <- readRDS(goset_file)
+}
+```
+
+Now we are ready to perform gene set enrichment analysis using `goseq`.
 
 ```{r}
+if (!file.exists(result_file)) {
+    ## get median transcript length for bias adjustment
+    txByGene <- transcriptsBy(EnsDb.Hsapiens.v86, "gene")
+    geneLength <- sapply(width(txByGene), median)
+    geneLength <- geneLength[names(genes)]
+    genes[is.na(genes)] <- 0
+
+    ## perform gsea
     pwf <- nullp(genes, bias.data = geneLength)
     goRes <- goseq(pwf, gene2cat = goSets)
+
+    saveRDS(goRes, file = result_file)
+} else {
+    goRes <- readRDS(result_file)
+}
 ```
@@
 ```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
-rank_scatter( dat=goRes, pval="over_represented_pvalue", 
-              covariate="numInCat", bins=50, funx=log2, 
-              funfill=log10_trans()) +
-    ylim(0, 12) + ggtitle("Over-represented gene sets") +
-    xlab( expression(log[10]~"(# of genes)")) +
-    ylab( expression(-log[10]~"(p-value)") )
-
+rank_scatter(dat = goRes, pval = "over_represented_pvalue", 
+              covariate = "numInCat", bins = 50, funx = log2, 
+              funfill = log10_trans()) +
+    ylim(0, 12) +
+    ggtitle("Over-represented gene sets") +
+    xlab(expression(log[10]~"(# of genes)")) +
+    ylab(expression(-log[10]~"(p-value)") ) 
 ```
@@
 ## rename columns and prepare for benchmarking
-res <- dplyr:::select( goRes, c("over_represented_pvalue", "numInCat") ) %>%
-    dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
-res$pval <- pmin( res$pval, 1 )
+res <- dplyr:::select(goRes, c("over_represented_pvalue", "numInCat")) %>%
+    dplyr:::rename(pval = over_represented_pvalue, ind_covariate = numInCat)
+res$pval <- pmin(res$pval, 1)
 
 ## generate default BenchDesign
-designBen <- initializeBenchDesign()
+bd <- initializeBenchDesign()
 
 ## run benchmark
-sGSEA <- buildBench(designBen, res, ftCols="ind_covariate")
+if (!file.exists(bench_file)) {
+    sGSEA <- buildBench(bd, res, ftCols = "ind_covariate")
+    saveRDS(sGSEA, file = bench_file)
+} else {
+    sGSEA <- readRDS(bench_file)
+}
 ```
@@
 ```{r}
-methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
+methods <- c("lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
 plotCovariateBoxplots(sGSEA, alpha=0.1, nsets=6, methods=methods, trans=log2)
@@
 ```{r}
-pr <- buildBench(designBen, res, ftCols = "ind_covariate")
+pr <- buildBench(bd, res, ftCols = "ind_covariate")
 assayNames(pr) <- "qvalue"

### 2018-03-11 datasets/GSEA/gsea-mouse.Rmd

--- 2018-02-23 datasets/GSEA/gsea-mouse.Rmd
@@
 ---
-title: "Case Study: Gene Set Enrichment Analysis (GSEA), Mouse"
+title: "Case Study: Gene Set Enrichment Analysis (Mouse Data Set)"
 author: "Alejandro Reyes"
@@
 dir.create(resdir, showWarnings = FALSE)
+
+## intermediary files we create below
+count_file <- file.path(datdir, "mouse-counts.rds")
+deseq_file <- file.path(datdir, "mouse-deseq.rds")
+goset_file <- file.path(datdir, "mouse-gosets.rds")
+result_file <- file.path(resdir, "mouse-results.rds")
+bench_file <- file.path(resdir, "mouse-benchmark.rds")
 ```
@@
 ```{r, message=FALSE}
-dsd_data <- "dsdHSCMPP.RData"
-if(!file.exists(dsd_data)) {
-    download.file("https://www.dropbox.com/s/spo1pifum5k72rn/dseHSCMPP.RData?dl=1",
-                  destfile = dsd_data)
+if (!file.exists(count_file)) {
+    ## download.file("", destfile = count_file)
 }
-load(dsd_data)
+dseHSCMPP <- readRDS(count_file)
 ```
@@
 
-## Differential Testing
+## Enrichment Analysis
 
-Specify model matrix design and run DESeq2 to get differentially expressed genes.
+**Add description of what's going on.**
+We first test for differentially expressed genes using DESeq2 and use genes with adjusted p-value less than 0.1.
 
 ```{r deseq2, message=FALSE}
+if (!file.exists(deseq_file)) {
     dseHSCMPP <- DESeq(dseHSCMPP)
@@
                    independentFiltering = FALSE)
+    saveRDS(res, file = deseq_file)
+} else {
+    res <- readRDS(deseq_file)
+}
+
+genes <- as.numeric(res$padj < 0.1)
+names(genes) <- rownames(res)
 ```
 
-## Enrichment Analysis
-
-**Add description of what's going on.**
+We next determine the gene sets associated with each gene.
 
 ```{r goseq, message=FALSE}
-goset_data <- file.path(datdir, "goSetsMm.RData")
-if (!file.exists(goset_data)) {
+if (!file.exists(goset_file)) {
     mart <- useMart("ensembl", "mmusculus_gene_ensembl")
@@
     goSets <- with(goSets, split(go_id, ensembl_gene_id))
-    save(goSets, file = goset_data)
+    saveRDS(goSets, file = goset_file)
 } else {
-   load(goset_data)
+   goSets <- readRDS(goset_file)
 }
+```
 
-## getting median transcript length
-genes <- as.numeric(res$padj < 0.1)
-names(genes) <- rownames(res)
+Now we are ready to perform gene set enrichment analysis using `goseq`.
 
+```{r}
+if (!file.exists(result_file)) {
     ## getting median transcript length
@@
     geneLength <- geneLength[names(genes)]
-genes[is.na( genes )] <- 0
-```
+    genes[is.na(genes)] <- 0
 
-Performing a gene set enrichment analysis with `goseq`.
-
-```{r}
+    ## perform gsea
     pwf <- nullp(genes, bias.data = geneLength)
     goRes <- goseq(pwf, gene2cat = goSets)
+
+    saveRDS(goRes, file = result_file)
+} else {
+    goRes <- readRDS(result_file)
+}
 ```
@@
 ## generate default BenchDesign
-designBen <- initializeBenchDesign()
+bd <- initializeBenchDesign()
 
 ## run benchmark
-sGSEAMouse <- buildBench(designBen, res, ftCols = "ind_covariate")
+if (!file.exists(bench_file)) {
+    sGSEAMouse <- buildBench(bd, res, ftCols = "ind_covariate")
+    saveRDS(sGSEAMouse, file = bench_file)
+} else {
+    sGSEAMouse <- readRDS(bench_file)
+}
 ```

### 2018-03-11 datasets/RNA-Seq/RNA-Seq.Rmd

--- 2018-02-20 datasets/RNA-Seq/RNA-Seq.Rmd
@@
 ## set up parallel backend
-multicoreParam <- MulticoreParam(workers = 2)
+cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
+##multicoreParam <- MulticoreParam(workers = cores)
+multicoreParam <- SerialParam()
 ```
@@
 
-```{r}
+```{r, meancov-diag-scatter, fig.width=4.5, fig.height=3.5}
+gp <- rank_scatter(geneExp, pvalue = "pval", covariate = "ind_covariate") +
+    ggtitle("Mean coverage as independent covariate") +
+    xlab("Mean Expression")
+ggsave("temp.png", gp, width = 4.5, height = 3.5)
+```
+
+```{r, meancov-diag-hist, fig.width=10, fig.height=3.2}
 strat_hist(geneExp, pvalue = "pval",
            covariate = "ind_covariate", maxy = 25)
-```
-
-```{r}
-rank_scatter(geneExp, pvalue = "pval", covariate = "ind_covariate")
 ```
@@
 bd <- initializeBenchDesign()
-sb <- buildBench(bd, data = geneExp, parallel = TRUE)
+sb <- buildBench(bd, data = geneExp, ftCols = "ind_covariate")
 ```
@@
 
-```{r}
+```{r, rnaseq-scatter}
 rejections_scatter(sb, supplementary = FALSE)
+rejection_scatter_bins(sb, covariate = "ind_covariate",
+                       bins = 4, supplementary = FALSE)
 ```
 
-```{r}
+```{r, rnaseq-overlap}
 plotFDRMethodsOverlap(sb, alpha = 0.05, nsets = ncol(sb),
@@
                       supplementary = FALSE)
+```
+
+```{r, rnaseq-boxplots}
+## properties of method-exclusive discoveries
+methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
+plotCovariateBoxplots(sb, alpha = 0.1, nsets = 6,
+                      methods = methods, trans = log2, maxNum = 5)
 ```

### 2018-03-11 datasets/RNA-Seq/yeast-simulation.Rmd

@@
 ```{r}
-res5 <- plotsim_standardize(null5, alpha = seq(0.01, 0.20, 0.01))
+res5 <- plotsim_standardize(null5, alpha = seq(0.01, 0.10, 0.01))
 
@@
                 merge_ihw = TRUE, errorBars=TRUE)
+
+covariateHeatmap(null5, alpha=0.05, covname="log2FC")
+covariateHeatmap(null5, alpha=0.05, covname="ind_covariate")
 
@@
 ```{r}
-res10 <- plotsim_standardize(null10, alpha = seq(0.01, 0.20, 0.01))
+res10 <- plotsim_standardize(null10, alpha = seq(0.01, 0.10, 0.01))
 
@@
                 merge_ihw = TRUE, errorBars=TRUE)
+
+covariateHeatmap(null10, alpha=0.05, covname="log2FC")
+covariateHeatmap(null10, alpha=0.05, covname="ind_covariate")
 
@@
 ```{r}
-res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.20, 0.01))
+res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
 
@@
 
-plotEffectSize(de5, covname="log2FC")
-plotEffectSize(de5, covname="log2FC", DE=FALSE)
-
-plotEffectSize(de5, covname="ind_covariate")
-plotEffectSize(de5, covname="ind_covariate", DE=FALSE)
+covariateHeatmap(de5, alpha=0.05, covname="log2FC")
+covariateHeatmap(de5, alpha=0.05, covname="ind_covariate")
 
@@
 ```{r}
-res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.20, 0.01))
+res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.10, 0.01))
 
@@
 
-plotEffectSize(de10, covname="log2FC")
-plotEffectSize(de10, covname="log2FC", DE=FALSE)
-
-plotEffectSize(de10, covname="ind_covariate")
-plotEffectSize(de10, covname="ind_covariate", DE=FALSE)
+covariateHeatmap(de10, alpha=0.05, covname="log2FC")
+covariateHeatmap(de10, alpha=0.05, covname="ind_covariate")
 

### 2018-03-11 datasets/microbiome/microbiome-goodrich.Rmd

@@
 dir.create(resdir, showWarnings = FALSE)
+
+ob_data <- file.path(datdir, "ob_goodrich_results.tar.gz")
+ob_dir <- file.path(datdir, "ob_goodrich_results")
+
+otu_result_file <- file.path(resdir, "goodrich-otus-results.rds")
+otu_bench_file <- file.path(resdir, "goodrich-otus-benchmark.rds")
+gns_result_file <- file.path(resdir, "goodrich-genus-results.rds")
+gns_bench_file <- file.path(resdir, "goodrich-genus-benchmark.rds")
 ```
@@
 ```{r, goodrich-download}
-ob_data <- file.path(datdir, "ob_goodrich_results.tar.gz")
 if (!file.exists(ob_data)) {
@@
 
-ob_dir <- file.path(datdir, "ob_goodrich_results")
 if (!file.exists(file.path(ob_dir, "ob_goodrich.metadata.txt"))) {
@@
 ```{r, goodrich-pvals}
-resfile <- file.path(resdir, paste0("goodrich_results_", nrow(abun_otu), "_OTUs.rda"))
-if (!file.exists(resfile)) {
+if (!file.exists(otu_result_file)) {
     res <- test_microbiome(abundance = abun_otu, shift = zeroabun,
                            is_case = meta$DiseaseState == "OB")
-    save(res, file = resfile)
+    saveRDS(res, file = otu_result_file)
 } else {
-    load(resfile)
+    res <- readRDS(otu_result_file)
 }
@@
 ```{r, goodrich-phylo, fig.width=10, fig.height=6, eval=FALSE}
-##strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=20)
-##rank_scatter(res, pvalue="pval", covariate="ubiquity")
-ggplot(res, aes(x=pval)) + geom_histogram() + facet_wrap(~phylum, scales = "free")
+ggplot(res, aes(x=pval)) +
+    geom_histogram() +
+    facet_wrap(~phylum, scales = "free")
 ```
@@
 ```{r, goodrich-sb, results="hide", message=FALSE, eval=TRUE}
-for (nmids in c(5, 50, 150)) {
-    ## use common BenchDesign w/out ASH
+if (!file.exists(otu_bench_file)) {
+    nmids <- 150
     bd <- initializeBenchDesign(nmids = nmids)
     bd <- dropBMethod(bd, "ashs")
-
-    resfile <- file.path(resdir, paste0("goodrich_summarizedBenchmark_",
-                                        nrow(res), "_nmids_", nmids, ".rda"))
-    if (!file.exists(resfile)) {
     sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
@@
                    "https://zenodo.org/record/840333/files/ob_goodrich_results.tar.gz"
-        save(sb, file = resfile)
+    metadata(sb)$nmids <- nmids
+    saveRDS(sb, file = otu_bench_file)
 } else {
-        load(resfile)
-    }
+    sb <- readRDS(otu_bench_file)
 }
@@
 ```{r, goodrich-genus-pvals}
-resfile <- file.path(resdir, paste0("goodrich_results_", nrow(genus_df), "_genera.rda"))
-if (!file.exists(resfile)) {
+if (!file.exists(gns_result_file)) {
     res <- test_microbiome(abundance = genus, shift = zeroabun,
                            is_case = meta$DiseaseState == "OB")
-    save(res, file = resfile)
+    saveRDS(res, file = gns_result_file)
 } else {
-    load(resfile)
+    res <- readRDS(gns_result_file)
 }
@@
 
-#### Set up BenchDesign object
-
 First, we'll create an object of `BenchDesign` class to hold the data and 
@@
 ```{r, goodrich-genus-sb, results="hide", message=FALSE, eval=TRUE}
-for (nmids in c(5, 50, 150)) {
-    ## use common BenchDesign w/out ASH
+if (!file.exists(gns_bench_file)) {
+    nmids <- 150
     bd <- initializeBenchDesign(nmids = nmids)
     bd <- dropBMethod(bd, "ashs")
-
-    resfile <- file.path(resdir, paste0("goodrich_summarizedBenchmark_",
-                                        nrow(res), "_genera_nmids_", nmids, ".rda"))
-    if (!file.exists(resfile)) {
     sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
@@
                    "https://zenodo.org/record/840333/files/ob_goodrich_results.tar.gz"
-        save(sb, file = resfile)
+    metadata(sb)$nmids <- nmids
+    saveRDS(sb, file = gns_bench_file)
 } else {
-        load(resfile)
-    }
+    sb <- readRDS(gns_bench_file)
 }
@@
 ```{r, goodrich-genus-plot, width=15, height=15, eval=TRUE}
-# plot nrejects by method overall and stratified by covariate
-rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
 rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
-#plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
 ```
@@
 ```
+
 ```{r, goodrich-genus-covariates}
 methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
-plotCovariateBoxplots( sb, alpha=0.1, nsets=6, methods=methods)
+plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods)
 ```

### 2018-03-11 datasets/scrnaseq/scrnaseq-human-mast.Rmd

@@
 
-cores <- 6
+cores <- 8
 BiocParallel::register(MulticoreParam(workers = cores))
@@
   destfile=paste0(datdir, "GSE84465.rds"))
-
 ```
@@
              pvalue="mast_p", covariate="meanExp") + 
-  ggtitle("MAST, Covariate 1: Mean Expression")
+  ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
 ```
@@
            pvalue="mast_p", covariate="meanExp", maxy=35,
-           main = "MAST, Covariate 1: Mean Expression")
+           main = "MAST, Covariate 1: Mean Nonzero Expression")
 ```
 
-The mean expression appears to be informative and approximately satisfies the 
+The mean nonzero expression appears to be informative and approximately satisfies the 
 assumptions
@@
 
-### Covariate one: Mean expression
+### Covariate one: mean nonzero expression
 
-First we'll include the mean expression covariate.
+First we'll include the mean nonzero expression covariate.
 
@@
 
-### Covariate one: Mean Expression
+### Covariate one: mean nonzero expression
 
@@
                    supplementary=FALSE) +
-  ggtitle("MAST, Covariate 1: Mean Expression")
+  ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
 
@@
                        supplementary=FALSE) +
-  ggtitle("MAST, Covariate 1: Mean Expression")
+  ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
 
@@
 plotCovariateBoxplots(sb1, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3) +
-  ggtitle("MAST, Covariate 1: Mean Expression")
+  ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
 
+```
+
+```{r, fig.width=8, fig.height=3.5}
+covariateHeatmap(sb1, alpha=0.05, covname="ind_covariate") +
+  ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
 ```
@@
 
+```{r, fig.width=8, fig.height=3.5}
+covariateHeatmap(sb2, alpha=0.05, covname="ind_covariate") +
+  ggtitle("MAST, Covariate 2: Detection Rate")
+```
+
 # Session Information

### 2018-03-11 datasets/scrnaseq/scrnaseq-human-scdd.Rmd

@@
            pvalue="scdd_p", covariate="meanExp", maxy=35,
-           main = "scDD, Covariate 1: Mean Expression") 
+           main = "scDD, Covariate 1: Mean Nonzero Expression") 
 ```
 
-The mean expression appears to be informative and approximately satisfies the 
+The mean nonzero expression appears to be informative and approximately satisfies the 
 assumptions
@@
 mcols(sb2)$ind_covariate <- mcols(sb2)$detection
-plotCovariateBoxplots(sb2, alpha=0.1, nsets=6, methods=methods, maxNum=3)+
+plotCovariateBoxplots(sb2, alpha=0.1, nsets=6, methods=methods, maxNum=3) +
   ggtitle("scDD, Covariate 2: Detection Rate")
 
+```
+
+```{r, fig.width=8, fig.height=3.5}
+covariateHeatmap(sb2, alpha=0.05, covname="ind_covariate") +
+  ggtitle("scDD, Covariate 2: Detection Rate")
 ```

### 2018-03-11 datasets/scrnaseq/scrnaseq-mouse-mast.Rmd

@@
            pvalue="mast_p", covariate="meanExp", maxy=16,
-           main = "MAST, Covariate 1: Mean Expression")
+           main = "MAST, Covariate 1: Mean Nonzero Expression")
 ```
 
-The mean expression appears to be informative and approximately satisfies the 
+The mean nonzero expression appears to be informative and approximately satisfies the 
 assumptions
@@
 mcols(sb2)$ind_covariate <- mcols(sb2)$detection
-plotCovariateBoxplots(sb2, alpha=0.1, nsets=6, methods=methods, maxNum=3)+
+plotCovariateBoxplots(sb2, alpha=0.1, nsets=6, methods=methods, maxNum=3) +
   ggtitle("MAST, Covariate 2: Detection Rate")
 
+```
+
+```{r, fig.width=8, fig.height=3.5}
+covariateHeatmap(sb2, alpha=0.05, covname="ind_covariate") +
+  ggtitle("MAST, Covariate 2: Detection Rate")
 ```

### 2018-03-12 datasets/ChIP-seq/ChIPseq-diffbind.Rmd
Run macs2 from Rmd

@@
 
-For this analysis, peak calls are obtained from the ENCODE Project pipeline, and
-testing is performed using _DiffBind_ and _DESeq2_.
+For this analysis, peak calls are made using MACS2 on reads aligned using the _Rsubread_ package,
+and testing is performed using _DiffBind_ and _DESeq2_.
 
@@
 
-We downloaded the peak calls directly from the UCSC ENCODE portal.
-
-```{r}
-broad_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeBroadHistone/"
-broad_peaks <- c("wgEncodeBroadHistoneGm12878H3k4me3StdPkRep1.narrowPeak.gz",
-                 "wgEncodeBroadHistoneGm12878H3k04me3StdPkRep2V2.narrowPeak.gz",
-                 "wgEncodeBroadHistoneK562H3k4me3StdPkRep1.narrowPeak.gz",
-                 "wgEncodeBroadHistoneK562H3k4me3StdPkRep2.narrowPeak.gz")
-
-for (i_peak in broad_peaks) {
-    if (!file.exists(file.path(datdir, i_peak))) {
-        download.file(paste0(broad_url, i_peak),
-                      destfile = file.path(datdir, i_peak))
-    }
-}
-
-uw_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwHistone/"
-uw_peaks <- c("wgEncodeUwHistoneGm12878H3k4me3StdPkRep1.narrowPeak.gz",
-              "wgEncodeUwHistoneGm12878H3k4me3StdPkRep2.narrowPeak.gz",
-              "wgEncodeUwHistoneK562H3k4me3StdPkRep1.narrowPeak.gz",
-              "wgEncodeUwHistoneK562H3k4me3StdPkRep2.narrowPeak.gz")
-
-for (i_peak in uw_peaks) {
-    if (!file.exists(file.path(datdir, i_peak))) {
-        download.file(paste0(uw_url, i_peak),
-                      destfile = file.path(datdir, i_peak))
-    }
-}
-```
-
 We determine sample metadata from the file names.
@@
 ```{r}
-peakfiles <- file.path(datdir, c(broad_peaks, uw_peaks))
-labs <- gsub("wgEncode(.*)Histone.*", "\\1", basename(peakfiles))
-cells <- gsub("wgEncode.*Histone(.*)H3k.*", "\\1", basename(peakfiles))
+## create metadata table
+sortedfiles <- list.files(datdir, pattern = "\\.sorted\\.bam$", full.names = TRUE)
+labs <- gsub("wgEncode(.*)Histone.*", "\\1", basename(sortedfiles))
+cells <- gsub("wgEncode.*Histone(.*)H3k.*", "\\1", basename(sortedfiles))
+meta <- data.frame(cellline = cells, lab = labs, sortedfile = sortedfiles,
+                   peakfile = gsub("\\.bam$", "_peaks.narrowPeak", sortedfiles),
+                   stringsAsFactors = FALSE)
+```
 
-meta <- data.frame(cellline = cells, lab = labs, file = peakfiles,
-                   stringsAsFactors = FALSE)
+We call peaks using MACS using reads aligned in `ChIPseq-csaw.Rmd`.
+
+```{r}
+for (i in seq_len(nrow(meta))) {
+    i_bam <- meta$sortedfile[i]
+    i_peak <- meta$peakfile[i]
+    if (!file.exists(i_peak)) {
+        code <- sprintf("macs2 callpeak -t %s -f BAM -g hs -n %s -q 0.1",
+                        i_bam, gsub("\\.bam", "", i_bam))
+        code <- system(code)
+        stopifnot(code == 0L)
+    }
+}
 ```

### 2018-03-12 datasets/RNA-Seq/RNA-Seq.Rmd
Update RNA-seq analysis using smaller rds; uniform file names

@@
 title: "Case Study: RNA-Seq Differential Analysis"
-author: "Chinmay Shukla"
+author: "Chinmay Shukla, Patrick Kimes"
 date: '`r format(Sys.Date(), "%B %d, %Y")`'
@@
 
-**Add brief summary of analysis goals and data set.**
+**Add brief summary of analysis goals and details of data set.**
+This data set includes 30 samples each from two regions of the human basal ganglia,
+the nucleus accumbens and the putamen, obtained from the GTEx project.
+Gene-level read counts were obtined by **fill**...
 
@@
 
+## intermediary files we create below
+count_file <- file.path(datdir, "brain-counts.rds")
+result_file <- file.path(resdir, "brain-results.rds")
+bench_file <- file.path(resdir, "brain-benchmark.rds")
+
 ## set up parallel backend
 cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
-##multicoreParam <- MulticoreParam(workers = cores)
 multicoreParam <- SerialParam()
@@
 
+The pre-computed set of samples can
+
 ```{r, dsd-download}
-dsd_data <- file.path(datdir, "dsdObjects.RData")
-if (!file.exists(dsd_data)) {
-    download.file("https://www.dropbox.com/s/8hf0ccl27ircodz/dsdObjects.RData?dl=1",
-                  destfile = dsd_data)
+if (!file.exists(count_file)) {
+    download.file("", destfile = count_file)
 }
-load(dsd_data)
-```
-
-**Add details on content of data and subsetting.**
-
-```{r}
-## Filter to include samples from only 2 tissues
-tissues <-  c('Brain - Nucleus accumbens (basal ganglia)',
-              'Brain - Putamen (basal ganglia)')
-filteredDsd1 <- dsd1[, colData(dsd1)$tissue %in% tissues]
-filteredDsd1$tissue <- droplevels(filteredDsd1$tissue)
+dsd <- readRDS(count_file)
 ```
@@
 ```{r}
-res_rda <- file.path(resdir, "dsd1-results.rda")
-if (!file.exists(res_rda)) {
+if (!file.exists(result_file)) {
     ## build DESeq object and run DESeq2
-    dds <- DESeq(filteredDsd1, parallel = TRUE, BPPARAM = multicoreParam)
+    dds <- DESeq(dsd, parallel = TRUE, BPPARAM = multicoreParam)
     res <- results(dds, independentFiltering = FALSE)
 
-    save(dds, res, file = res_rda)
+    saveRDS(res, file = result_file)
 } else {
-    load(res_rda)
+    res <- readRDS(result_file)
 }
@@
 ```{r}
+if (!file.exists(bench_file)) {
     bd <- initializeBenchDesign()
     sb <- buildBench(bd, data = geneExp, ftCols = "ind_covariate")
+    saveRDS(sb, file = bench_file)
+} else {
+    sb <- readRDS(bench_file)
+}
 ```

### 2018-03-13 datasets/RNA-Seq/yeast-simulation.Rmd
Add covariate line plot

@@
 
-covariateHeatmap(null5, alpha=0.05, covname="log2FC")
-covariateHeatmap(null5, alpha=0.05, covname="ind_covariate")
+covariateHeatmap(null5, alpha=0.05, covname="log2FC", trans="log1p")
+covariateHeatmap(null5, alpha=0.05, covname="ind_covariate", trans="log1p")
+
+covariateHeatmap(null5, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p", linePlot = TRUE)
+covariateHeatmap(null5, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 linePlot = TRUE, trans="log1p")
 
@@
 
-covariateHeatmap(null10, alpha=0.05, covname="log2FC")
-covariateHeatmap(null10, alpha=0.05, covname="ind_covariate")
+covariateHeatmap(null10, alpha=0.05, covname="log2FC", trans="log1p")
+covariateHeatmap(null10, alpha=0.05, covname="ind_covariate", trans="log1p")
+
+covariateHeatmap(null10, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p", linePlot = TRUE)
+covariateHeatmap(null10, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 linePlot = TRUE, trans="log1p")
 
@@
 
-covariateHeatmap(de5, alpha=0.05, covname="log2FC")
+covariateHeatmap(de5, alpha=0.05, covname="log2FC", trans="log1p")
 covariateHeatmap(de5, alpha=0.05, covname="ind_covariate")
+
+covariateHeatmap(de5, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p", linePlot = TRUE)
+covariateHeatmap(de5, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 linePlot = TRUE)
 
@@
 
-covariateHeatmap(de10, alpha=0.05, covname="log2FC")
+covariateHeatmap(de10, alpha=0.05, covname="log2FC", trans="log1p")
 covariateHeatmap(de10, alpha=0.05, covname="ind_covariate")
+
+covariateHeatmap(de10, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p", linePlot = TRUE)
+covariateHeatmap(de10, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 linePlot = TRUE)
 

### 2018-03-13 datasets/R/effectSizePlot.R

@@ -132,9 +132,11 @@ covariateHeatmap <- function(sbl, alpha=0.05, nbins = 50,
     df <- lapply(sbl, summarize_one_item, alpha=alpha, nbins=nbins)
     df <- bind_rows(df, .id = "rep")
     df <- as.tibble(df) %>%
-      group_by(method, bin) %>%
-      summarize(nsig = mean(nsig / tot)*100) %>%
-      na.omit()
+        mutate(prop = nsig / tot) %>%
+        group_by(method, bin) %>%
+        summarize(nsig = mean(prop)*100,
+                  se = sd(prop * 100) / sqrt(n())) %>%
+        na.omit()
   }else if ("SummarizedBenchmark" %in% class(sbl)){
     df <- summarize_one_item(sbl, alpha=alpha, nbins=nbins) %>%
       mutate(nsig = nsig/tot*100)
@@ -143,20 +145,45 @@ covariateHeatmap <- function(sbl, alpha=0.05, nbins = 50,
          "a list of SummarizedBenchmark objects.")
   }
   
-  p <- ggplot(df, aes(x = as.factor(method), y = bin/nbins)) +
-         geom_raster(aes(fill = nsig)) +
-    xlab("Method") +
-    scale_y_continuous(labels = scales::percent) +
-    ylab(paste0(covname, " percentile")) +
-    labs(fill="Mean % Significant")
-  
-  mpoint <- median(df$nsig)
-  
-  if(!is.null(trans)){
-    p <- p + scale_fill_gradientn(colors=RColorBrewer::brewer.pal(8, "BuGn"), 
-                                  trans=trans, na.value = "white")
+  if (linePlot){
+    p <- ggplot(df, aes(x = bin/nbins, y = nsig, color = method)) +
+      geom_line(alpha = 3/4) +
+      geom_errorbar(aes(ymin = nsig - se, ymax = nsig + se),
+                    width=0.01, alpha=1/3,) +
+      ylab("Mean % Significant") +
+      scale_x_continuous(labels = scales::percent) +
+      xlab(paste0(covname, " percentile")) +
+      labs(color="Method") + 
+      viridis::scale_color_viridis("Method", discrete = TRUE,
+                                   guide = guide_legend(ncol = 2))
+    
+    if (!is.list(sbl)){
+      p <- p + ylab("% Significant")
+    }
+    
+    if(!is.null(trans)){
+      p <- p + scale_y_continuous(trans=trans)
+    }
[…]

### 2018-03-13 datasets/ChIP-seq/ChIPseq-promoters.Rmd

@@
     dds <- DESeq(dds, fitType = "mean")
-    dat <- as.data.frame(results(dds))
+    dat <- as.data.frame(results(dds, independentFiltering  = FALSE))
     colnames(dat) <- c('ind_covariate', 'effect_size', 'SE',

### 2018-03-13 datasets/ChIP-seq/ChIPseq-promoters.Rmd
chipseq-promoter tests: use indep filtering, save as RDS

--- 2018-03-08 datasets/ChIP-seq/ChIPseq-promoters.Rmd
@@
 ## intermediary files we create below
-count_file <- file.path(resdir, "h3k4me3-promotors-counts.rda")
-result_file <- file.path(resdir, "h3k4me3-promotors-results.rda")
-bench_file <- file.path(resdir, "h3k4me3-promotors-benchmark.rda")
+count_file <- file.path(resdir, "h3k4me3-promotors-counts.rds")
+result_file <- file.path(resdir, "h3k4me3-promotors-results.rds")
+bench_file <- file.path(resdir, "h3k4me3-promotors-benchmark.rds")
 
@@
 if (file.exists(count_file)) {
-    load(count_file)
+    rc <- readRDS(count_file)
 } else {
@@
                         ignoreDup = TRUE)$counts
-    save(rc, file = count_file)
+    saveRDS(rc, file = count_file)
 }
@@
 
-```{r count, results='hide', message=FALSE, warning=FALSE}
+```{r chipseq-deseq, results='hide', message=FALSE, warning=FALSE}
 if (file.exists(result_file)) {
-    load(result_file)
+    dat <- readRDS(result_file)
 } else {
@@
     dat <- dat[!is.na(dat$effect_size), ]
-    save(dat, file = result_file)
+    saveRDS(dat, file = result_file)
 }
@@
 We use the common `BenchDesign` with the set of multiple testing correction
-methods already included.
-
-```{r, chipseq-benchdesign}
-bd <- initializeBenchDesign()
-```
-
-Now, we're ready to construct the `SummarizedBenchmark` object, which will run
-the functions specified in each method.
+methods already included. Now, we're ready to construct the `SummarizedBenchmark`
+object, which will run the functions specified in each method.
 
@@
 if (file.exists(bench_file)) {
-    load(bench_file)
+    sb <- readRDS(bench_file)
 } else {
+    bd <- initializeBenchDesign()
     sb <- buildBench(bd, data = dat, ftCols = c("ind_covariate"),
                      parallel = TRUE, BPPARAM = multicoreParam)
-    save(sb, file = bench_file)
+    saveRDS(sb, file = bench_file)
 }

### 2018-03-13 datasets/ChIP-seq/ChIPseq.sh

@@ -13,4 +13,4 @@ export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/rstudio/0.98.1103-fasrc01/bin/pan
 
 # change filename to Rmd to be knitted. ncores is passed in through environment
 # variable SLURM_NTASKS
-R -e "rmarkdown::render('ChIP-seq.Rmd')"
+R -e "rmarkdown::render('ChIPseq-promoters.Rmd')"

### 2018-03-13 datasets/RNA-Seq/RNA-Seq.Rmd
Don't save plot as png in Rmd

@@
 ```{r, meancov-diag-scatter, fig.width=4.5, fig.height=3.5}
-gp <- rank_scatter(geneExp, pvalue = "pval", covariate = "ind_covariate") +
+rank_scatter(geneExp, pvalue = "pval", covariate = "ind_covariate") +
     ggtitle("Mean coverage as independent covariate") +
     xlab("Mean Expression")
-ggsave("temp.png", gp, width = 4.5, height = 3.5)
 ```

### 2018-03-14 datasets/R/effectSizePlot.R
Allow line plot for single-replicate (case studies)

@@ -148,8 +148,6 @@ covariateHeatmap <- function(sbl, alpha=0.05, nbins = 50,
   if (linePlot){
     p <- ggplot(df, aes(x = bin/nbins, y = nsig, color = method)) +
       geom_line(alpha = 3/4) +
-      geom_errorbar(aes(ymin = nsig - se, ymax = nsig + se),
-                    width=0.01, alpha=1/3,) +
       ylab("Mean % Significant") +
       scale_x_continuous(labels = scales::percent) +
       xlab(paste0(covname, " percentile")) +
@@ -158,7 +156,10 @@ covariateHeatmap <- function(sbl, alpha=0.05, nbins = 50,
                                    guide = guide_legend(ncol = 2))
     
     if (!is.list(sbl)){
-      p <- p + ylab("% Significant")
+      p <- p + ylab("% Significant") 
+    }else{
+      p <- p + geom_errorbar(aes(ymin = nsig - se, ymax = nsig + se),
+                    width=0.01, alpha=1/3)
     }
     
     if(!is.null(trans)){

### 2018-03-14 datasets/GWAS/GWAS.Rmd
Add covariate line plots

@@
 mcols(sbN)$tstat <- bmi %>% filter(!is.na(ind_covar_N)) %>% pull(effect_size)
-covariateHeatmap(sbN, alpha=0.05, covname="tstat")
-covariateHeatmap(sbN, alpha=0.05, covname="ind_covar_N")
+covariateHeatmap(sbN, alpha=0.05, covname="tstat", trans = "log1p")
+covariateHeatmap(sbN, alpha=0.05, covname="ind_covar_N", trans = "log1p")
+
+covariateHeatmap(sbN, alpha=0.05, covname="tstat", nbins=25, 
+                 linePlot = TRUE, trans="log1p")
+covariateHeatmap(sbN, alpha=0.05, covname="ind_covar_N", nbins=25, 
+                 linePlot = TRUE, trans="log1p")
 ```
@@
 mcols(sbAF)$tstat <- bmi %>% filter(!is.na(ind_covar_AF)) %>% pull(effect_size)
-covariateHeatmap(sbAF, alpha=0.05, covname="tstat")
+covariateHeatmap(sbAF, alpha=0.05, covname="tstat", trans = "log1p")
 covariateHeatmap(sbAF, alpha=0.05, covname="ind_covar_AF")
+
+covariateHeatmap(sbAF, alpha=0.05, covname="tstat", nbins=25, 
+                 linePlot = TRUE, trans = "log1p")
+covariateHeatmap(sbAF, alpha=0.05, covname="ind_covar_AF", nbins=25, 
+                 linePlot = TRUE)
 ```

### 2018-03-14 datasets/scrnaseq/scrnaseq-human-mast.Rmd
Add covariate line plots

@@
 ```{r, humandata-download}
+datfile <- paste0(datdir, "GSE84465.rds")
+
+if (!file.exists(datfile)){
 download.file("http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE84465.rds",
-  destfile=paste0(datdir, "GSE84465.rds"))
+  destfile=datfile)
+}
 ```
@@
   ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
+
+covariateHeatmap(sb1, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 linePlot = TRUE) +
+  ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
 ```
@@
   ggtitle("MAST, Covariate 2: Detection Rate")
+
+covariateHeatmap(sb2, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 linePlot = TRUE) +
+  ggtitle("MAST, Covariate 2: Detection Rate")
 ```

### 2018-03-14 datasets/scrnaseq/scrnaseq-human-scdd.Rmd
Add covariate line plots

@@
 ```{r, humandata-download}
+datfile <- paste0(datdir, "GSE84465.rds")
+
+if (!file.exists(datfile)){
   download.file("http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE84465.rds",
-  destfile=paste0(datdir, "GSE84465.rds"))
-
+    destfile=datfile)
+}
 ```
@@
   ggtitle("scDD, Covariate 1: Mean Nonzero Expression")
+
+covariateHeatmap(sb1, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 linePlot = TRUE) +
+  ggtitle("scDD, Covariate 1: Mean Nonzero Expression")
 ```
@@
   ggtitle("scDD, Covariate 2: Detection Rate")
+
+covariateHeatmap(sb2, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 linePlot = TRUE) +
+  ggtitle("scDD, Covariate 2: Detection Rate")
 ```

### 2018-03-14 datasets/scrnaseq/scrnaseq-mouse-mast.Rmd
Add covariate line plots

@@
 ```{r, mousedata-download}
+datfile <- paste0(datdir,"GSE94383.rds")
+
+if (!file.exists(datfile)){
   download.file(url="http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE94383.rds",
-  destfile=paste0(datdir,"GSE94383.rds"))
+    destfile=datfile)
+}
 ```
@@
   ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
+
+covariateHeatmap(sb1, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 linePlot = TRUE, trans = "log1p") +
+  ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
 ```

### 2018-03-14 datasets/scrnaseq/scrnaseq-mouse-scdd.Rmd
Add covariate line plots

@@
   ggtitle("scDD, Covariate 1: Mean Nonzero Expression")
+
+covariateHeatmap(sb1, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 linePlot = TRUE, trans = "log1p") +
+  ggtitle("scDD, Covariate 1: Mean Nonzero Expression")
 ```

### 2018-03-14 datasets/GWAS/GWAS-knit.sh

@@ -2,12 +2,12 @@
 #SBATCH -J knitGWAS                 # A single job name for the array
 #SBATCH -n 8                        # Number of cores
 #SBATCH -N 1                        # All cores on one machine
-#SBATCH -p irizarry,serial_requeue  # Partition
+#SBATCH -p shared                   # Partition
 #SBATCH --mem 60000                 # Memory request 
 #SBATCH -t 0-2:00                   # Maximum execution time (D-HH:MM)
-#SBATCH -o SLURM/render-%j.out  # Standard output
-#SBATCH -e SLURM/render-%j.err  # Standard error
- 
+#SBATCH -o render-%j.out  # Standard output
+#SBATCH -e render-%j.err  # Standard error
+
 module load plink
 export SLURM_NTASKS
 export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/rstudio/0.98.1103-fasrc01/bin/pandoc/"

### 2018-03-14 datasets/scrnaseq/scrnaseq-human-mast.Rmd
Don't use phenotypes derived from tSNE clusters

@@
 
-We'll extract two conditions: Neoplastic and Regular cells, as
-defined by their tsne clusters.
+This dataset (from PMC5810554) contains cells from four Glioblastoma patients, 
+sampled from both the tumor core as well as the tissue peripheral to the tumor. 
+We'll compare the Immune cells in the tumor vs periphery.
 
@@
   # subset by groups of interest
-  humandata <- humandata[, (colData(humandata)$characteristics_ch1.7 %in% 
-                         c("neoplastic: Neoplastic", "neoplastic: Regular")) &
-                         (colData(humandata)$characteristics_ch1.5 %in% 
-                         c("tsne cluster: 11",  "tsne cluster: 8"))]
+  humandata <- humandata[, (colData(humandata)$tissue.ch1 %in% 
+                         c(" Tumor", " Periphery")) &
+                         colData(humandata)$cell.type.ch1 == " Immune cell" ]
   cdata=colData(humandata)
-  cdata$group <- cdata$characteristics_ch1.7
+  cdata$group <- cdata$tissue.ch1
 

### 2018-03-14 datasets/scrnaseq/scrnaseq-human-mast.Rmd
Change human dataset comparison
Previous comparison used very different cell types, resulting in almost everything
being significant. Here we compare the neoplastic cells found in glioblastoma
tumors versus the surrounding tissue (periphery).

@@
 sampled from both the tumor core as well as the tissue peripheral to the tumor. 
-We'll compare the Immune cells in the tumor vs periphery.
+We'll compare the Neoplastic cells in the tumor core vs periphery. We'll exclude 
+the two samples where only a few neoplastic cells were found in the periphery.
 
@@
   # subset by groups of interest
-  humandata <- humandata[, (colData(humandata)$tissue.ch1 %in% 
-                         c(" Tumor", " Periphery")) &
-                         colData(humandata)$cell.type.ch1 == " Immune cell" ]
+  humandata <- humandata[, colData(humandata)$patient.id.ch1 %in% 
+                           c(" BT_S4", " BT_S2") &
+                           colData(humandata)$cell.type.ch1 == " Neoplastic" ]
   cdata=colData(humandata)
@@
 strat_hist(data.frame(rowData(humandata)), 
-           pvalue="mast_p", covariate="meanExp", maxy=35,
+           pvalue="mast_p", covariate="meanExp", maxy=12,
            main = "MAST, Covariate 1: Mean Nonzero Expression")
@@
 strat_hist(data.frame(rowData(humandata)), 
-           pvalue="mast_p", covariate="detection", maxy=35,
+           pvalue="mast_p", covariate="detection", maxy=12,
            main = "MAST, Covariate 2: Detection Rate") 

### 2018-03-14 datasets/scrnaseq/scrnaseq-human-scdd.Rmd
Change human dataset comparison
Previous comparison used very different cell types, resulting in almost everything
being significant. Here we compare the neoplastic cells found in glioblastoma
tumors versus the surrounding tissue (periphery).

@@
 strat_hist(data.frame(rowData(humandata)), 
-           pvalue="scdd_p", covariate="meanExp", maxy=35,
+           pvalue="scdd_p", covariate="meanExp", maxy=12,
            main = "scDD, Covariate 1: Mean Nonzero Expression") 
@@
 strat_hist(data.frame(rowData(humandata)), 
-           pvalue="scdd_p", covariate="detection", maxy=35,
+           pvalue="scdd_p", covariate="detection", maxy=12,
            main = "scDD, Covariate 2: Detection Rate") 

### 2018-03-14 datasets/scrnaseq/scrnaseq-knit.sh

@@ -13,7 +13,8 @@ export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/rstudio/0.98.1103-fasrc01/bin/pan
 
 # change filename to Rmd to be knitted
 # make sure number of cores used in Rmd matches sbatch param -n
-R -e "rmarkdown::render('scrnaseq-mouse-scdd.Rmd', clean=FALSE)"
 R -e "rmarkdown::render('scrnaseq-mouse-mast.Rmd', clean=FALSE)"
-R -e "rmarkdown::render('scrnaseq-human-scdd.Rmd', clean=FALSE)"
+R -e "rmarkdown::render('scrnaseq-mouse-scdd.Rmd', clean=FALSE)"
 R -e "rmarkdown::render('scrnaseq-human-mast.Rmd', clean=FALSE)"
+R -e "rmarkdown::render('scrnaseq-human-scdd.Rmd', clean=FALSE)"
+

### 2018-03-15 datasets/RNA-Seq/RNAseq-Applications.Rmd
AS:Adding new studies for RNAseq case study

---
title: "RNAseq-other studies"
author: "Ayshwarya Subramanian"
date: "03/09/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache = TRUE,fig.width=12, fig.height=8)
#source('http://bioconductor.org/biocLite.R')
#biocLite('recount')
library(dplyr)
library(Matrix)
library(readr)
library(ggplot2)
library(cowplot)
library(reshape2)
library(data.table)
library(recount)
```

## Data from Recount
We select uniformly processed RNAseq datasets from the (Recount2)[https://jhubiostatistics.shinyapps.io/recount/] data repository. The data are available in the form of Ranged Summarized Experiments. We select 2 human datasets.

We first download Dataset 1, which looks at bulk RNA-seq from bloood and immune cells from healthy individuals and individuals with Type-1 Diabetes (and other disease phenotypes), before and 24h after treatment with IFN-beta. We select two-groups: Diabetes and healthy. The presence of multiple groups allows for complex designs for differential expression testing.

```{r, dat1}
url1 <- download_study('SRP045500')
## Load the data
load(file.path('SRP045500', 'rse_gene.Rdata'))
## Scale counts, may not be needed if using counts directly
dat1 <- scale_counts(rse_gene)
sample_info <- data.frame(run = colData(dat1)$run,
    group = ifelse(grepl('Type 1 Diabetes', as.data.frame(colData(dat1))$characteristics), 'Diabetes', ifelse(grepl('Healthy Control', as.data.frame(colData(dat1))$characteristics),'Healthy','Other')), sequencing_depth = colData(dat1)$reads_downloaded)
## Add sample information for DE analysis
colData(dat1)$group <- sample_info$group

```

We next download Dataset 2, which looks at bulk RNA-seq from hematopoietic stem and progenitor cells from healthy individuals and individuals at risk for MDS(Myelodysplastic syndromes).

```{r, dat2}
url2 <- download_study('SRP062025')
## Load the data
load(file.path('SRP062025', 'rse_gene.Rdata'))
## Scale counts, may not be needed if using counts directly
dat2 <- scale_counts(rse_gene)
sample_info <- data.frame(run = colData(dat1)$run,
    group = ifelse(grepl('MDS cell phenotype', as.data.frame(colData(dat1))$characteristics), 'MDS', 'Normal'),sequencing_depth = colData(dat2)$reads_downloaded)
## Add sample information for DE analysis
colData(dat2)$group <- sample_info$group
```

## Differential Expression testing

Now, we perform Differential Expression analysis using DESeq2.

```{r,de}
## Perform differential expression analysis with DESeq2
library('DESeq2')

## Specify design and switch to DESeq2 format
dds <- DESeqDataSet(dat1[,!(colData(dat1)$group=="Other")], ~ reads_downloaded + group)
## Perform DE analysis
dds <- DESeq(dds)
res <- results(dds)

```

### 2018-03-15 datasets/RNA-Seq/RNA-Seq.Rmd
Added missing documentation and data for RNA-seq differential expression of GTEx data.

@@
 
-**Add brief summary of analysis goals and details of data set.**
-This data set includes 30 samples each from two regions of the human basal ganglia,
-the nucleus accumbens and the putamen, obtained from the GTEx project.
-Gene-level read counts were obtined by **fill**...
+The objective of this document is benchmark methods to control FDR in the context of differential gene expression.
+
+The first dataset consists of 10 samples each from two regions of the human basal ganglia, the nucleus accumbens and the putamen, obtained from the GTEx project.
+Shortly, samples were downloaded using the \emph{Short Read Archive Toolkit} and mapped to the human reference genome version GRCh38 using \emph{STAR v2.4.2a}.
+\emph{htseq-count} was used to tabulate the number of uniquely mapping reads for each gene.
 
@@
 
-The pre-computed set of samples can
+We use script `downloadAndSubset.R` to pre-process the data.
+This script downloads a DESeqDataSet object containing gene level counts for a subset of the GTEx data.
+Then, it further subsets it keeping samples from only females and two cell types,  nucleus accumbens and the putamen.
 
@@
 if (!file.exists(count_file)) {
-    download.file("", destfile = count_file)
+    source("downloadAndSubset.R")
 }
@@
 
-**Add details on how testing is performed.**
+We use DESeq2 to test each gene for differential gene expression between the two cell types.
+We set the parameter `independentFiltering=FALSE` to skip the independent filtering step.
+This step would be redundant with some of the FDR control methods that use gene expression as an independent covariate to gain power.
 
@@
     res <- results(dds, independentFiltering = FALSE)
-
     saveRDS(res, file = result_file)
@@
 
-**Add details on filtering.**
+ We reformat DESeqResuls object into a normal data.frame for further analysis.
 
@@
 geneExp <- na.omit(geneExp)
-geneExp <- dplyr::filter(geneExp, ind_covariate >= 1)
 ```
@@
 
-**Add description of covariates.**
+In RNA-seq differential expression analysis, it is very well established that the mean expression is an informative covariate that is independent under the null and that can be used to increase power while keeping FDR control (DESeq2 paper).
 
@@
 
+We noticed some discreteness in the distribution of p-values towards values close to 1 that was particularly pronounced in the first covariate bin.
+The overall distribution of p-values could violate some of the assumptions of the FDR control methods.
+
 ```{r, meancov-diag-hist, fig.width=10, fig.height=3.2}
+strat_hist(geneExp, pvalue = "pval",
+           covariate = "ind_covariate", maxy = 25)
+```
+
+These discreteness, however, corresponded to genes with very low expression values.
+For example, there were `r nrow( dplyr::filter(geneExp, ind_covariate <= 1) )` genes with less than an average of 1 read across the 20 samples.
+The distribution of p-values looked much better when removing this very lowly expressed genes.
+
+```{r}
+geneExp <- dplyr::filter(geneExp, ind_covariate >= 1)
 strat_hist(geneExp, pvalue = "pval",
@@
 sb <- addDefaultMetrics(sb)
+estimatePerformanceMetrics( sb, tidy=TRUE ) %>%
+  filter( performanceMetric == "rejections" ) %>%
+  select( blabel, performanceMetric, alpha, value ) %>%
+  arrange( desc(value) )
 ```
 
-**Add description of results.**
+ash was the method that rejected a larger number of hypothesis, followed by lfdr and scott-theorethical.
 

### 2018-03-15 datasets/RNA-Seq/RNA-Seq.Rmd
Wording.

@@
 We use DESeq2 to test each gene for differential gene expression between the two cell types.
-We set the parameter `independentFiltering=FALSE` to skip the independent filtering step.
-This step would be redundant with some of the FDR control methods that use gene expression as an independent covariate to gain power.
+We set the parameter `independentFiltering=FALSE` to skip the independent filtering step, as this step would be redundant with some of the FDR control methods that use gene expression as an independent covariate to increase power.
 

### 2018-03-15 datasets/RNA-Seq/RNAseq-Applications.Rmd

@@
 ## Data from Recount
-We select uniformly processed RNAseq datasets from the (Recount2)[https://jhubiostatistics.shinyapps.io/recount/] data repository.
+We select uniformly processed RNAseq datasets from the [Recount2](https://jhubiostatistics.shinyapps.io/recount/) data repository.
 The data are available in the form of Ranged Summarized Experiments.

### 2018-03-15 datasets/RNA-Seq/RNA-Seq.Rmd
Added 2nd dataset.

@@
         number_sections: true
+bibliography: bibliography.bib
 ---
@@
 
-# Data Preparation
+# Dataset 1
+
+## Data Preparation
 
@@
 
-# Data Analysis
+## Data Analysis
 
-## Differential Testing
+### Differential Testing
 
@@
 
-## Covariate Diagnostics
+### Covariate Diagnostics
 
@@
 
-### Mean Coverage
+### Mean Counts
 
@@
 
-## Multiple-Testing Correction
+### Multiple-Testing Correction
 
@@
 
-## Benchmark Metrics
+### Benchmark Metrics
 
@@
 
+# Dataset 2
+
+As a second dataset, we will test for differences in gene expression upon the knockout of the microRNA mir-200c [@Kim_2013].
+The raw fastq files can be found under the accession number `SRP030475`.
+As the number of samples is limited the experiment might be underpowered, as in most RNA-seq analysis.
+This is an experimental scenario that could benefit from power gaining from using modern FDR control methods.
+
+## Data Preparation
+
+We will download the pre-processed gene level counts available through recount2.
+
+```{r}
+download_study('SRP030475')
+load(file.path('SRP030475', 'rse_gene.Rdata'))
+dsd <- scale_counts(rse_gene)
+```
+
+We then subset for samples containing the control samples and the samples where mir200c was knocked down.
+Recount2 downloads data as a RangeSummarizedExperiment object, so we convert this into a DESeqDataSet object.
+
+```{r}
+dsd <- dsd[,grepl("WT|200c", colData(dsd)$title)]
+colData(dsd)$mir200c <- factor(ifelse( grepl( "WT", colData(dsd)$title), "WT", "KO"))
+dsd <- as( dsd, "DESeqDataSet" )
+storage.mode(assays(dsd)[["counts"]]) <- "integer"
+```
+
+## Data Analysis
+
+### Differential testing
+
+Then, we set the design parameter to test for differences in expression upon mir200c knockout and run DESeq2. Similarly to the previous dataset, we set the parameter `independentFiltering=FALSE`.
+
+```{r}
+design(dsd) <- ~ mir200c
+dsd <- DESeq(dsd)
+res <- results(dsd, independentFiltering=FALSE) %>% 
+  as.data.frame(  ) %>%
+  na.omit() %>% 
+  select( pvalue, baseMean, log2FoldChange, lfcSE, stat ) %>%
+  rename( pval=pvalue, ind_covariate=baseMean, 
+          effect_size=log2FoldChange, SE=lfcSE, 
+          test_statistic=stat ) 
+head( res )
+```
+
+### Mean Counts
+
+As with the previous example, the mean counts is an informative covariate.
+
+```{r, meancov-diag-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter( res, pvalue = "pval", covariate = "ind_covariate") +
+    ggtitle("Mean coverage as independent covariate") +
+    xlab("Mean Expression")
+```
+
+Similarly to the first dataset, keeping all the tests results in a strange discretness. This is removed once we filter very lowly expressed genes.
+
+```{r, meancov-diag-hist, fig.width=10, fig.height=3.2}
+strat_hist( res, pvalue = "pval",
+           covariate = "ind_covariate", maxy = 7.5)
+res <- filter(res, ind_covariate > 1 )
+strat_hist( res, pvalue = "pval",
+           covariate = "ind_covariate", maxy = 3)
+```
+
+### Multiple-Testing Correction
+
+We use the common `BenchDesign` with the set of multiple testing correction
+methods already included.
+
+```{r}
+bd <- initializeBenchDesign()
+sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+```
+
+### Benchmark Metrics
+
+```{r}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+estimatePerformanceMetrics( sb, tidy=TRUE ) %>%
+  filter( performanceMetric == "rejections" ) %>%
+  select( blabel, performanceMetric, alpha, value ) %>%
+  arrange( desc(value) )
+```
+
+```{r}
+rejections_scatter(sb, supplementary = FALSE)
+rejection_scatter_bins(sb, covariate = "ind_covariate",
+                       bins = 4, supplementary = FALSE)
+```
+
+```{r}
+plotFDRMethodsOverlap(sb, alpha = 0.05, nsets = ncol(sb),
+                      order.by = "freq", decreasing = TRUE,
+                      supplementary = FALSE)
+```
+
+```{r}
+methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf", "scott-empirical")
+plotCovariateBoxplots(sb, alpha = 0.1, nsets = 6,
+                      methods = methods, trans = log2, maxNum = 5)
+```
+
 # Session Info

### 2018-03-16 datasets/RNA-Seq/RNA-Seq.Rmd
Separated RNA-seq case study vignettes

--- 2018-03-15 datasets/RNA-Seq/RNA-Seq.Rmd
@@
 ---
-title: "Case Study: RNA-Seq Differential Analysis"
+title: "Case Study: RNA-Seq Differential Analysis (basal ganglia)"
 author: "Chinmay Shukla, Patrick Kimes"
@@
 
-### Mean Coverage
+## Mean Counts
 

### 2018-03-16 datasets/RNA-Seq/RNAseq-Applications.Rmd
Separated RNA-seq case study vignettes

@@
 colData(dat1)$group <- sample_info$group
-
+ncol(dat1)
 ```

### 2018-03-16 datasets/RNA-Seq/RNA-Seq.Rmd
Download only the relevant subset of GTEx.

@@
 if (!file.exists(count_file)) {
-    source("downloadAndSubset.R")
+    download.file("https://www.dropbox.com/s/pptxqnpsmdkyhcf/brain-counts.rds?dl=1", destfile=count_file)
 }

### 2018-03-16 datasets/RNA-Seq/RNA-Seq-mir200c.Rmd
Added separate vignette with mir200c analysis.

---
title: "Case Study: RNA-Seq Differential Analysis (mir200c knockdown)"
author: "Alejandro Reyes"
date: "3/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

As a second RNA-seq dataset, we will test for differences in gene expression upon the knockout of the microRNA mir-200c [@Kim_2013].  The raw fastq files can be found under the accession number `SRP030475`. As the number of samples is limited the experiment might be underpowered, as in most RNA-seq analysis. This is an experimental scenario that could benefit from power gaining from using modern FDR control methods.

# Workspace Setup

```{r, workspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(DESeq2)
library(SummarizedBenchmark)
library(BiocParallel)
library(recount)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "data"
resdir <- "results"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)

## intermediary files we create below
count_path <- file.path(datdir, "SRP030475")
count_file <- file.path(count_path, 'rse_gene.Rdata')
bench_file <- file.path(resdir, "mir200c-benchmark.rds")

## set up parallel backend
cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
multicoreParam <- SerialParam()
```

# Data Preparation

We will download the pre-processed gene level counts available through recount2.

```{r}
if( !file.exists( count_file ) ){
    download_study('SRP030475', outdir=count_path)
}

load(count_file)
dsd <- scale_counts(rse_gene)
```

We next subset for samples containing the control samples and the samples where mir200c was knocked down. Recount2 downloads data as a RangeSummarizedExperiment object, so we convert this into a DESeqDataSet object.

```{r}
dsd <- dsd[,grepl("WT|200c", colData(dsd)$title)]
colData(dsd)$mir200c <- factor(ifelse( grepl( "WT", colData(dsd)$title), "WT", "KO"))
dsd <- as( dsd, "DESeqDataSet" )
storage.mode(assays(dsd)[["counts"]]) <- "integer"
```

# Data Analysis

## Differential testing

Then, we set the design parameter to test for differences in expression upon mir200c knockout and run DESeq2. Similarly to the previous dataset, we set the parameter `independentFiltering=FALSE`.

```{r}
design(dsd) <- ~ mir200c
dsd <- DESeq(dsd)
res <- results(dsd, independentFiltering=FALSE) %>% 
  as.data.frame(  ) %>%
  na.omit() %>% 
  dplyr::select( pvalue, baseMean, log2FoldChange, lfcSE, stat ) %>%
  dplyr::rename( pval=pvalue, ind_covariate=baseMean, 
          effect_size=log2FoldChange, SE=lfcSE, 
          test_statistic=stat ) 
head( res )
```

## Mean Counts

As with the previous example, the mean counts is an informative covariate.

```{r, meancov-diag-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter( res, pvalue = "pval", covariate = "ind_covariate") +
    ggtitle("Mean coverage as independent covariate") +
    xlab("Mean Expression")
```

Similarly to the first dataset, keeping all the tests results in a strange discretness. This is removed once we filter very lowly expressed genes. For the first covariate bin, however, there is a strange behaviour in which the distribution seems a bit skewed towards larger p-values.

```{r, meancov-diag-hist, fig.width=10, fig.height=3.2}
strat_hist( res, pvalue = "pval",
           covariate = "ind_covariate", maxy = 7.5)
res <- filter(res, ind_covariate > 1 )
strat_hist( res, pvalue = "pval",
           covariate = "ind_covariate", maxy = 3)
```

## Multiple-Testing Correction

We use the common `BenchDesign` with the set of multiple testing correction methods already included.

```{r}
if (!file.exists(bench_file)) {
    bd <- initializeBenchDesign(nmids=10000)
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    saveRDS(sb, file = bench_file)
} else {
    sb <- readRDS(bench_file)
}
```

There are some warnings from both `BH` and `scott-empirical`. However, they do return results. I tried to increase the `nmids` parameter but did not help.

```{r}
head( assays( sb )[["bench"]] )
```

## Benchmark Metrics

Scott empirical rejects many more hypothesis than the rest of the methods, followed by lfdr and IHW.

```{r}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
estimatePerformanceMetrics( sb, tidy=TRUE ) %>%
  filter( performanceMetric == "rejections", alpha == 0.1 ) %>%
  select( blabel, performanceMetric, alpha, value ) %>%
  arrange( desc(value) )
```

```{r}
rejections_scatter(sb, supplementary = FALSE)
rejection_scatter_bins(sb, covariate = "ind_covariate",
                       bins = 4, supplementary = FALSE)
```

```{r}
plotFDRMethodsOverlap(sb, alpha = 0.05, nsets = ncol(sb),
                      order.by = "freq", decreasing = TRUE,
                      supplementary = FALSE)
```

```{r}
methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf", "scott-empirical")
plotCovariateBoxplots(sb, alpha = 0.1, nsets = 6,
                      methods = methods, trans = log2, maxNum = 5)
```

# Session Info

```{r}
sessionInfo()
```

### 2018-03-16 datasets/microbiome/microbiome-papa.Rmd

--- 2018-03-08 datasets/microbiome/microbiome-baxter.Rmd
@@
 ---
-title: "Case Study: 16S microbiome data (Baxter et al., 2016)"
+title: "Case Study: 16S microbiome data (Papa et al., 2012)"
 author: "Claire Duvallet"
@@
 
-Here we download and analyze the Baxter colorectal cancer (CRC) dataset
-[(Baxter et al., 2016)](http://dx.doi.org/10.1186/s13073-016-0290-3).
-The dataset includes stool samples from 172 patients with no colonic lesions,
-and 120 patients with CRC. We'll download the processed OTU tables from
+Here we download and analyze the Papa inflammatory bowel disease (IBD) dataset
+[(Papa et al., 2012)](https://doi.org/10.1371/journal.pone.0039242).
+The dataset includes stool samples from 24 patients nonIBD controls (i.e. patients
+with GI symptoms but not IBD), 23 patients with Crohn's disease and 43 with
+ulcerative colitis (both categorized as inflammatory bowel disease, IBD).
+We'll download the processed OTU tables from
 Zenodo and unzip them in the `benchmark-fdr/datasets/microbiome/data` folder.
 
-For colorectal cancer, we do expect some amount of truly differentially abundant OTUs,
+For inflammatory bowel disease, we do expect some amount of truly differentially abundant OTUs,
 but not as many as in diarrhea (e.g. Schubert et al., 2014). This dataset will hopefully
@@
 
-crc_data <- file.path(datdir, "crc_baxter_results.tar.gz")
-crc_dir <- file.path(datdir, "crc_baxter_results")
+ibd_data <- file.path(datdir, "ibd_alm_results.tar.gz")
+ibd_dir <- file.path(datdir, "ibd_alm_results")
 
-otu_result_file <- file.path(resdir, "baxter-otus-results.rds")
-otu_bench_file <- file.path(resdir, "baxter-otus-benchmark.rds")
-gns_result_file <- file.path(resdir, "baxter-genus-results.rds")
-gns_bench_file <- file.path(resdir, "baxter-genus-benchmark.rds")
+otu_result_file <- file.path(resdir, "papa-otus-results.rds")
+otu_bench_file <- file.path(resdir, "papa-otus-benchmark.rds")
+gns_result_file <- file.path(resdir, "papa-genus-results.rds")
+gns_bench_file <- file.path(resdir, "papa-genus-benchmark.rds")
 ```
@@
 
-```{r, baxter-download}
-if (!file.exists(crc_data)) {
-    download.file("https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz",
-                  destfile = crc_data)
+```{r, papa-download}
+if (!file.exists(ibd_data)) {
+    download.file("https://zenodo.org/record/840333/files/ibd_alm_results.tar.gz",
+                  destfile = ibd_data)
 }
 
-if (!file.exists(file.path(crc_dir, "crc_baxter.metadata.txt"))) {
-    untar(crc_data, exdir = datdir)
+if (!file.exists(file.path(ibd_dir, "ibd_alm.metadata.txt"))) {
+    untar(ibd_data, exdir = datdir)
 }
@@
 
-```{r, baxter-readdata}
+```{r, papa-readdata}
 ## load OTU table and metadata
-otu <- read.table(file.path(crc_dir, "RDP",
-                            "crc_baxter.otu_table.100.denovo.rdp_assigned"))
-meta <- read.csv(file.path(crc_dir, "crc_baxter.metadata.txt"), sep='\t')
+otu <- read.table(file.path(ibd_dir, "RDP",
+                            "ibd_alm.otu_table.100.denovo.rdp_assigned"))
+meta <- read.csv(file.path(ibd_dir, "ibd_alm.metadata.txt"), sep='\t')
 
 ## Keep only samples with the right DiseaseState metadata
-meta <- dplyr::filter(meta, DiseaseState %in% c("H", "CRC"))
+meta <- dplyr::filter(meta, DiseaseState %in% c("nonIBD", "UC", "CD"))
 
-## add "X" in front of sample IDs because of how R read in the OTU table
-meta$Sample_Name_s <- paste0("X", meta$Sample_Name_s)
+### add "X" in front of sample IDs because of how R read in the OTU table
+meta$X <- paste0("X", meta$X)
 
 ## keep only samples with both metadata and 16S data
-keep_samples <- intersect(colnames(otu), meta$Sample_Name_s)
+keep_samples <- intersect(colnames(otu), meta$X)
 otu <- otu[, keep_samples]
-meta <- dplyr::filter(meta, Sample_Name_s %in% keep_samples)
+meta <- dplyr::filter(meta, X %in% keep_samples)
 ```
@@
 
-Since we'll be using OTU-wise covariates, we shouldn't need to perform any
-filtering/cleaning of the OTUs, apart from removing any that are all zeros.
+Even though we'll be using OTU-wise covariates, we still need to do some
+filtering/cleaning of the OTUs so that the differential abundance test doesn't get wonky.
 (This may happen after removing shallow samples, I think.)
@@
 
-```{r, baxter-cleandata}
+```{r, papa-cleandata}
 min_otu <- 10
@@
 ## Update metadata with new samples
-meta <- dplyr::filter(meta, Sample_Name_s %in% colnames(otu))
+meta <- dplyr::filter(meta, X %in% colnames(otu))
 
@@
 Next, we need to calculate the pvalues, effect size, and standard error for each OTU.
-Here, we'll compare CRC vs. healthy. We won't consider the nonCRC adenoma patients.
+Here, we'll compare UC/CD vs. healthy. 
 We'll put these results into a dataframe, and label the columns with the 
@@
 
-```{r, baxter-pvals}
+```{r, papa-pvals}
 if (!file.exists(otu_result_file)) {
     res <- test_microbiome(abundance = abun_otu, shift = zeroabun,
-                           is_case = meta$DiseaseState == "CRC")
+                           is_case = meta$DiseaseState %in% c("CD", "UC"))
     saveRDS(res, file = otu_result_file)
@@
 
-```{r, baxter-addphylo}
+```{r, papa-addphylo}
 res <- tidyr::separate(res, otu, 
@@
 
-```{r, baxter-ubi-hist, fig.width=10, fig.height=3.2}
+```{r, papa-ubi-hist, fig.width=10, fig.height=3.2}
 strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
@@
 
-```{r, baxter-ubi-scatter, fig.width=4.5, fig.height=3.5}
+```{r, papa-ubi-scatter, fig.width=4.5, fig.height=3.5}
 rank_scatter(res, pvalue="pval", covariate="ubiquity")
@@
 
-Something weird is happening with p=0.20 and p=0.40 - maybe this has something to do with ties?
-Either way, ubiquity looks to be a bit informative (from the scatter plot, but not really the histograms...)
-
 ### Mean Abundance (across non-zero samples)
 
-```{r, baxter-abun-hist, fig.width=10, fig.height=3.2}
+```{r, papa-abun-hist, fig.width=10, fig.height=3.2}
 strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
@@
 
-```{r, baxter-abun-scatter, fig.width=4.5, fig.height=3.5}
+```{r, papa-abun-scatter, fig.width=4.5, fig.height=3.5}
 rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
@@
 
-```{r, baxter-phylo, fig.width=10, fig.height=6, eval=FALSE}
+```{r, papa-phylo, fig.width=10, fig.height=6, eval=FALSE}
 ggplot(res, aes(x=pval)) +
@@
 
-```{r, baxter-labelcovariate}
+```{r, papa-labelcovariate}
 res <- dplyr::mutate(res, ind_covariate = ubiquity)
@@
 
-```{r, baxter-sb, results="hide", message=FALSE, eval=TRUE}
+```{r, papa-sb, results="hide", message=FALSE, eval=TRUE}
 if (!file.exists(otu_bench_file)) {
@@
     metadata(sb)$data_download_link <-
-                   "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
+                   "https://zenodo.org/record/840333/files/ibd_alm_results.tar.gz"
     metadata(sb)$nmids <- nmids
@@
 
-```{r, baxter-metrics, eval=TRUE}
+```{r, papa-metrics, eval=TRUE}
 assayNames(sb) <- "qvalue"
@@
 
-```{r, baxter-plot, width=15, height=15, eval=TRUE}
+```{r, papa-plot, width=15, height=15, eval=TRUE}
 rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
@@
 
-```{r, baxter-genus}
+```{r, papa-genus}
 ## add column with otu names
@@
 ## re-order columns to match metadata
-genus_df <- genus_df[, match(meta$Sample_Name_s, colnames(genus_df))]
+genus_df <- genus_df[, match(meta$X, colnames(genus_df))]
 
@@
 
-```{r, baxter-genus-pvals}
+```{r, papa-genus-pvals}
 if (!file.exists(gns_result_file)) {
     res <- test_microbiome(abundance = genus, shift = zeroabun,
-                           is_case = meta$DiseaseState == "CRC")
+                           is_case = meta$DiseaseState %in% c("CD", "UC"))
     saveRDS(res, file = gns_result_file)
@@
 
-```{r, baxter-genus-ubi, fig.width=10, fig.height=3.2}
+```{r, papa-genus-ubi, fig.width=10, fig.height=3.2}
 strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
@@
 
-```{r, baxter-genus-ubi-scatter, fig.width=4.5, fig.height=3.5}
+```{r, papa-genus-ubi-scatter, fig.width=4.5, fig.height=3.5}
 rank_scatter(res, pvalue="pval", covariate="ubiquity")
@@
 
-```{r, baxter-genus-abun, fig.width=10, fig.height=3.2}
+```{r, papa-genus-abun, fig.width=10, fig.height=3.2}
 strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
@@
 
-```{r, baxter-genus-abun-scatter, fig.width=4.5, fig.height=3.5}
+```{r, papa-genus-abun-scatter, fig.width=4.5, fig.height=3.5}
 rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
@@
 
-```{r, baxter-genus-labelcovariate}
+```{r, papa-genus-labelcovariate}
 res <- dplyr::mutate(res, ind_covariate = ubiquity)
@@
 
-```{r, baxter-genus-sb, results="hide", message=FALSE, eval=TRUE}
+```{r, papa-genus-sb, results="hide", message=FALSE, eval=TRUE}
 if (!file.exists(gns_bench_file)) {
@@
 
-```{r, baxter-genus-metrics, eval=TRUE}
+```{r, papa-genus-metrics, eval=TRUE}
 assayNames(sb) <- "qvalue"
@@
 
-```{r, baxter-genus-plot, width=15, height=15, eval=TRUE}
+```{r, papa-genus-plot, width=15, height=15, eval=TRUE}
 rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
@@
 
-Note: Benjamini-Hochberg (bh) overlaps exactly with the IHW results.
-
-```{r, baxter-genus-overlap}
+```{r, papa-genus-overlap}
 plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )

### 2018-03-16 datasets/RNA-Seq/RNA-Seq.Rmd
typo

@@
 title: "Case Study: RNA-Seq Differential Analysis (basal ganglia)"
-author: "Chinmay Shukla, Patrick Kimes"
+author: "Chinmay Shukla, Alejandro Reyes, Patrick Kimes"
 date: '`r format(Sys.Date(), "%B %d, %Y")`'

### 2018-03-21 datasets/ChIP-seq/ChIPseq-promoters-h3k4me3.Rmd
Rename chipseq rmds, drop diffbind

--- 2018-03-13 datasets/ChIP-seq/ChIPseq-promoters.Rmd
@@
 prom <- promoters(genes(TxDb.Hsapiens.UCSC.hg19.knownGene))
-seqlevels(prom, force = TRUE) <- seqlevels(prom)[!grepl("_", seqlevels(prom))]
+prom <- keepSeqlevels(prom, paste0("chr", c(1:22, "X", "Y", "M")),
+                      pruning.mode = "coarse")
 ```
@@
 ```{r, chipseq-diag-hist, fig.width=10, fig.height=3.2}
-strat_hist(dat, pvalue = "pval", covariate = "ind_covariate", maxy = 2)
+strat_hist(dat, pvalue = "pval", covariate = "ind_covariate",
+           maxy = 10)
 ```

### 2018-03-22 datasets/ChIP-seq/ChIPseq-h3k4me3-csaw.Rmd
Rename chipseq, rnaseq Rmd files

--- 2018-03-11 datasets/ChIP-seq/ChIPseq-csaw.Rmd
@@
 count_file <- file.path(resdir, "h3k4me3-csaw-counts.rds")
+filtered_file <- file.path(resdir, "h3k4me3-csaw-counts-filtered.rds")
 result_file <- file.path(resdir, "h3k4me3-csaw-results.rds")
@@
 std_chr <- paste0("chr", c(1:19, "X", "Y"))
-param <- readParam(minq = 15, discard = bl, restrict = std_chr)
+param <- readParam(minq = 20, discard = bl, restrict = std_chr)
 ```
@@
 if (file.exists(count_file)) {
-    filtered_cnts <- readRDS(count_file)
+    win_cnts <- readRDS(count_file)
 } else {
@@
     ## index bam files
+    if (any(!file.exists(paste0(meta$sortedfile, ".bai")))) {
         indexBam(meta$sortedfile)
+    }
 
@@
 
+    ## save unfiltered counts
+    saveRDS(win_cnts, file = count_file)
+}
+```
+
+We can apply prefiltering on windows with low abundance as described in Lun and Smyth (2016).
+
+```{r}
+if (file.exists(filtered_file)) {
+    filtered_cnts <- readRDS(filtered_file)
+} else {
     ## filter windows by abundance
@@
     filter_stat <- filterWindows(win_cnts, bins, type = "global")
-    keep <- filter_stat$filter > log2(3)
+    keep <- filter_stat$filter > log2(2)
     filtered_cnts <- win_cnts[keep, ]
 
-    saveRDS(filtered_cnts, file = count_file)
+    ## save filtered counts
+    saveRDS(filtered_cnts, file = filtered_file)
 }
@@
 
-We use _edgeR_ to test for differential binding.
+We use _edgeR_ to test for differential binding with the filtered data.
 
@@
 }
+
+## covert to df for downstream analysis
+res_df <- as.data.frame(res_ranges)
+```
+
+## Covariate Diagnostics
+
+### Number of Windows
+
+```{r, nwin-diag-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(res_df, pvalue = "PValue", covariate = "nWindows") +
+    ggtitle("Number of windows as independent covariate") +
+    xlab("Number of Windows")
+```
+
+```{r, nwin-diag-hist, fig.width=10, fig.height=3.2}
+strat_hist(res_df, pvalue = "PValue", covariate = "nWindows", maxy = 35, numQ = 5)
+```
+
+### Region Width
+
+```{r, width-diag-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(res_df, pvalue = "PValue", covariate = "width") +
+    ggtitle("Region width as independent covariate") +
+    xlab("Region Width")
+```
+
+```{r, meancov-diag-hist, fig.width=10, fig.height=3.2}
+strat_hist(res_df, pvalue = "PValue", covariate = "width", maxy = 35, numQ = 5)
+```
+
+## Multiple-Testing Correction
+
+We use the common `BenchDesign` with the set of multiple testing correction
+methods already included. We use the width of the regions as the independent
+covariate.
+
+```{r}
+if (!file.exists(bench_file)) {
+    res_df$ind_covariate <- res_df$width
+    res_df$pval <- res_df$PValue
+
+    bd <- initializeBenchDesign()
+    sb <- buildBench(bd, data = res_df, ftCols = "ind_covariate")
+    saveRDS(sb, file = bench_file)
+} else {
+    sb <- readRDS(bench_file)
+}
 ```

### 2018-03-22 datasets/RNA-Seq/RNAseq-gtex.Rmd
Rename chipseq, rnaseq Rmd files

--- 2018-03-16 datasets/RNA-Seq/RNA-Seq.Rmd
@@
 
-We use script `downloadAndSubset.R` to pre-process the data.
-This script downloads a DESeqDataSet object containing gene level counts for a subset of the GTEx data.
-Then, it further subsets it keeping samples from only females and two cell types,  nucleus accumbens and the putamen.
+We download a DESeqDataSet object containing gene level counts for a subset of the GTEx data. The data set includes  females and two cell types, nucleus accumbens and the putamen.
 
@@
 if (!file.exists(count_file)) {
-    download.file("https://www.dropbox.com/s/pptxqnpsmdkyhcf/brain-counts.rds?dl=1", destfile=count_file)
+    download.file("https://www.dropbox.com/s/pptxqnpsmdkyhcf/brain-counts.rds?dl=1",
+                  destfile = count_file)
 }

### 2018-03-22 datasets/GSEA/gsea-human.Rmd
Download GSEA input data from pk dropbox

@@
 
-Download *DESeqDataSet* if not present locally.
+**Some sort of summary of the data set.**  
+We download a *DESeqDataSet* if not present locally.
 
@@
 if (!file.exists(count_file)) {
-    ## download.file("", destfile = count_file)
+    download.file("https://www.dropbox.com/s/itrs2a6cnmurnck/human-counts.rds?dl=1",
+                  destfile = count_file)
 }
 dsdObject <- readRDS(count_file)
-```
-
-**What are we doing here?**
-
-```{r}
-levels(colData(dsdObject)$individual) <- gsub("GTEX-", "", levels(colData(dsdObject)$individual))
 ```

### 2018-03-22 datasets/GSEA/gsea-mouse.Rmd
Download GSEA input data from pk dropbox

@@
 if (!file.exists(count_file)) {
-    ## download.file("", destfile = count_file)
+    download.file("https://www.dropbox.com/s/jq6giqeclxcdypq/mouse-counts.rds?dl=1",
+                  destfile = count_file)
 }

### 2018-03-22 datasets/GSEA/gsea-human.Rmd
Update link to dropbox in gsea rmd again

@@
 if (!file.exists(count_file)) {
-    download.file("https://www.dropbox.com/s/itrs2a6cnmurnck/human-counts.rds?dl=1",
+    download.file("https://www.dropbox.com/s/itrs2a6cnmurnck/gsea-human-counts.rds?dl=1",
                   destfile = count_file)

### 2018-03-22 datasets/GSEA/gsea-mouse.Rmd
Update link to dropbox in gsea rmd again

--- 2018-03-11 datasets/GSEA/gsea-mouse.Rmd
@@
 if (!file.exists(count_file)) {
-    ## download.file("", destfile = count_file)
+    download.file("https://www.dropbox.com/s/jq6giqeclxcdypq/gsea-mouse-counts.rds?dl=1",
+                  destfile = count_file)
 }
@@
 if (!file.exists(goset_file)) {
-    mart <- useMart("ensembl", "mmusculus_gene_ensembl")
+    mart <- useMart("ensembl", "mmusculus_gene_ ensembl")
     goSets <- getBM(c("ensembl_gene_id", "go_id"), mart = mart,

### 2018-03-22 datasets/RNA-Seq/downloadAndSubset.R
Remove extra files from RNAseq dir

@@ -1,16 +1,0 @@
-
-download.file("https://www.dropbox.com/s/8hf0ccl27ircodz/dsdObjects.RData?dl=1",
-              destfile="data/dsdObjects.RData")
-load("data/dsdObjects.RData")
-dsd1 <- dsd1[,colData( dsd1 )$tissue %in% 
-  c("Brain - Nucleus accumbens (basal ganglia)", "Brain - Putamen (basal ganglia)")]
-colData(dsd1)$tissue <- droplevels( colData( dsd1 )$tissue )
-colData(dsd1)$individual <- droplevels( colData( dsd1 )$individual )
-dsd1 <- dsd1[,colData(dsd1)$sex == "female"]
-colData(dsd1)$sex <- droplevels( colData( dsd1 )$sex )
-
-colData(dsd1)$tissue <- factor(gsub(" |Brain -|\\(basal ganglia\\)", "", 
-                                    as.character( colData(dsd1)$tissue )))
-design(dsd1) <- ~ tissue
-saveRDS(dsd1, file="data/brain-counts.rds" )
-

### 2018-03-22 datasets/RNA-Seq/RNAseq-gtex.Rmd
Make RNAseq analyses more uniform

@@
 if (!file.exists(count_file)) {
-    download.file("https://www.dropbox.com/s/pptxqnpsmdkyhcf/brain-counts.rds?dl=1",
+    download.file("https://www.dropbox.com/s/9g06cx9vyonhdao/rnaseq-brain-counts.rds?dl=1",
                   destfile = count_file)
@@
 ```{r}
-if (!file.exists(result_file)) {
-    ## build DESeq object and run DESeq2
+if (file.exists(result_file)) {
+    res <- readRDS(result_file)
+} else {
     dds <- DESeq(dsd, parallel = TRUE, BPPARAM = multicoreParam)
     res <- results(dds, independentFiltering = FALSE)
+    res <- results(dds, independentFiltering = FALSE) %>% 
+        as.data.frame() %>%
+        na.omit() %>% 
+        dplyr::select(pvalue, baseMean, log2FoldChange, lfcSE, stat) %>%
+        dplyr::rename(pval = pvalue,
+                      ind_covariate = baseMean, 
+                      effect_size = log2FoldChange,
+                      SE = lfcSE, 
+                      test_statistic = stat)
     saveRDS(res, file = result_file)
-} else {
-    res <- readRDS(result_file)
 }
-```
-
- We reformat DESeqResuls object into a normal data.frame for further analysis.
-
-```{r}
-## build input data.frame
-geneExp <- data.frame(geneName = rownames(res),
-                      pval = res$pvalue,
-                      SE = res$lfcSE, 
-                      ind_covariate = res$baseMean,
-                      effect_size = res$log2FoldChange,
-                      test_statistic = res$stat)
-geneExp <- dplyr::as_tibble(geneExp)
-
-## filter data
-geneExp <- na.omit(geneExp)
 ```
@@
 
-In RNA-seq differential expression analysis, it is very well established that the mean expression is an informative covariate that is independent under the null and that can be used to increase power while keeping FDR control (DESeq2 paper).
+In RNA-seq differential expression analysis, it is very well established that the mean expression is an informative covariate that is independent under the null and that can be used to increase power while keeping FDR control.
 
-## Mean Counts
+### Mean Counts
 
@@
 These discreteness, however, corresponded to genes with very low expression values.
-For example, there were `r nrow( dplyr::filter(geneExp, ind_covariate <= 1) )` genes with less than an average of 1 read across the 20 samples.
+For example, there were `r nrow(dplyr::filter(geneExp, ind_covariate <= 1))` genes with less than an average of 1 read across the 20 samples.
 The distribution of p-values looked much better when removing this very lowly expressed genes.
@@
 ```{r}
-if (!file.exists(bench_file)) {
+if (file.exists(bench_file)) {
+    sb <- readRDS(bench_file)
+} else {
     bd <- initializeBenchDesign()
@@
     saveRDS(sb, file = bench_file)
-} else {
-    sb <- readRDS(bench_file)
 }
@@
 sb <- addDefaultMetrics(sb)
-estimatePerformanceMetrics( sb, tidy=TRUE ) %>%
-  filter( performanceMetric == "rejections" ) %>%
-  select( blabel, performanceMetric, alpha, value ) %>%
-  arrange( desc(value) )
+estimatePerformanceMetrics(sb, tidy = TRUE) %>%
+    filter(performanceMetric == "rejections") %>%
+    select(blabel, performanceMetric, alpha, value) %>%
+    arrange(desc(value))
 ```

### 2018-03-22 datasets/RNA-Seq/RNAseq-mir200c.Rmd
Make RNAseq analyses more uniform

--- 2018-03-16 datasets/RNA-Seq/RNA-Seq-mir200c.Rmd
@@
 
-As a second RNA-seq dataset, we will test for differences in gene expression upon the knockout of the microRNA mir-200c [@Kim_2013].
+As a second RNA-seq dataset, we will test for differences in gene expression upon the knockout of the microRNA mir-200c [(Kim et al., 2013)](http://dx.doi.org/10.1038/nsmb.2701).
 The raw fastq files can be found under the accession number `SRP030475`.
@@
 ## intermediary files we create below
-count_path <- file.path(datdir, "SRP030475")
-count_file <- file.path(count_path, 'rse_gene.Rdata')
+count_file <- file.path(datdir, "rse_gene.Rdata")
+result_file <- file.path(resdir, "mir200c-results.rds")
 bench_file <- file.path(resdir, "mir200c-benchmark.rds")
@@
 ```{r}
-if( !file.exists( count_file ) ){
-    download_study('SRP030475', outdir=count_path)
+if (!file.exists(count_file)) {
+    download_study('SRP030475', outdir = datdir)
 }
-
 load(count_file)
@@
 ```{r}
-dsd <- dsd[,grepl("WT|200c", colData(dsd)$title)]
-colData(dsd)$mir200c <- factor(ifelse( grepl( "WT", colData(dsd)$title), "WT", "KO"))
-dsd <- as( dsd, "DESeqDataSet" )
+dsd <- dsd[, grepl("WT|200c", colData(dsd)$title)]
+colData(dsd)$mir200c <- factor(ifelse(grepl("WT", colData(dsd)$title), "WT", "KO"))
+dsd <- as(dsd, "DESeqDataSet")
 storage.mode(assays(dsd)[["counts"]]) <- "integer"
@@
 
-## Differential testing
+## Differential Testing
 
@@
 ```{r}
+if (file.exists(result_file)) {
+    res <- readRDS(result_file)
+} else {
     design(dsd) <- ~ mir200c
-dsd <- DESeq(dsd)
-res <- results(dsd, independentFiltering=FALSE) %>% 
-  as.data.frame(  ) %>%
+    dds <- DESeq(dsd)
+    res <- results(dds, independentFiltering = FALSE) %>% 
+        as.data.frame() %>%
         na.omit() %>% 
-  dplyr::select( pvalue, baseMean, log2FoldChange, lfcSE, stat ) %>%
-  dplyr::rename( pval=pvalue, ind_covariate=baseMean, 
-          effect_size=log2FoldChange, SE=lfcSE, 
-          test_statistic=stat ) 
-head( res )
+        dplyr::select(pvalue, baseMean, log2FoldChange, lfcSE, stat) %>%
+        dplyr::rename(pval = pvalue,
+                      ind_covariate = baseMean, 
+                      effect_size = log2FoldChange,
+                      SE = lfcSE, 
+                      test_statistic = stat)
+    saveRDS(res, file = result_file)
+}
 ```
 
-## Mean Counts
+## Covariate Diagnostics
 
-As with the previous example, the mean counts is an informative covariate.
+As with the GTEx example, the mean counts is an informative covariate.
+
+### Mean Counts
 
 ```{r, meancov-diag-scatter, fig.width=4.5, fig.height=3.5}
-rank_scatter( res, pvalue = "pval", covariate = "ind_covariate") +
+rank_scatter(res, pvalue = "pval", covariate = "ind_covariate") +
     ggtitle("Mean coverage as independent covariate") +
@@
 
-Similarly to the first dataset, keeping all the tests results in a strange discretness.
+Similarly to the GTEx dataset, keeping all the tests results in a strange discretness.
 This is removed once we filter very lowly expressed genes.
@@
 ```{r, meancov-diag-hist, fig.width=10, fig.height=3.2}
-strat_hist( res, pvalue = "pval",
+strat_hist(res, pvalue = "pval",
            covariate = "ind_covariate", maxy = 7.5)
-res <- filter(res, ind_covariate > 1 )
-strat_hist( res, pvalue = "pval",
+res <- filter(res, ind_covariate > 1)
+strat_hist(res, pvalue = "pval",
            covariate = "ind_covariate", maxy = 3)
@@
 ```{r}
-if (!file.exists(bench_file)) {
-    bd <- initializeBenchDesign(nmids=10000)
+if (file.exists(bench_file)) {
+    sb <- readRDS(bench_file)
+} else {
+    bd <- initializeBenchDesign(nmids = 10000)
     sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
     saveRDS(sb, file = bench_file)
-} else {
-    sb <- readRDS(bench_file)
 }
@@
 ```{r}
-head( assays( sb )[["bench"]] )
+head(assays(sb)[["bench"]])
 ```
@@
 sb <- addDefaultMetrics(sb)
-estimatePerformanceMetrics( sb, tidy=TRUE ) %>%
-  filter( performanceMetric == "rejections", alpha == 0.1 ) %>%
-  select( blabel, performanceMetric, alpha, value ) %>%
-  arrange( desc(value) )
+estimatePerformanceMetrics(sb, tidy = TRUE) %>%
+    filter(performanceMetric == "rejections", alpha == 0.1) %>%
+    select(blabel, performanceMetric, alpha, value) %>%
+    arrange(desc(value))
 ```

### 2018-03-22 datasets/RNA-Seq/RNAseq.sh
Update RNAseq submit sh to point to new filenames

@@ -14,4 +14,5 @@ export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/rstudio/0.98.1103-fasrc01/bin/pan
 
 # change filename to Rmd to be knitted. ncores is passed in through environment
 # variable SLURM_NTASKS
-R -e "rmarkdown::render('RNA-Seq.Rmd')"
+R -e "rmarkdown::render('RNAseq-gtex.Rmd')"
+R -e "rmarkdown::render('RNAseq-mir200c.Rmd')"

### 2018-03-22 datasets/RNA-Seq/RNAseq-gtex.Rmd
Fix typos in gtex rna-seq analysis var names

@@
 ```{r, meancov-diag-scatter, fig.width=4.5, fig.height=3.5}
-rank_scatter(geneExp, pvalue = "pval", covariate = "ind_covariate") +
+rank_scatter(res, pvalue = "pval", covariate = "ind_covariate") +
     ggtitle("Mean coverage as independent covariate") +
@@
 ```{r, meancov-diag-hist, fig.width=10, fig.height=3.2}
-strat_hist(geneExp, pvalue = "pval",
+strat_hist(res, pvalue = "pval",
            covariate = "ind_covariate", maxy = 25)
@@
 These discreteness, however, corresponded to genes with very low expression values.
-For example, there were `r nrow(dplyr::filter(geneExp, ind_covariate <= 1))` genes with less than an average of 1 read across the 20 samples.
+For example, there were `r nrow(dplyr::filter(res, ind_covariate <= 1))` genes with less than an average of 1 read across the 20 samples.
 The distribution of p-values looked much better when removing this very lowly expressed genes.
@@
 ```{r}
-geneExp <- dplyr::filter(geneExp, ind_covariate >= 1)
-strat_hist(geneExp, pvalue = "pval",
+res <- dplyr::filter(res, ind_covariate >= 1)
+strat_hist(res, pvalue = "pval",
            covariate = "ind_covariate", maxy = 25)
@@
     bd <- initializeBenchDesign()
-    sb <- buildBench(bd, data = geneExp, ftCols = "ind_covariate")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
     saveRDS(sb, file = bench_file)

### 2018-03-22 datasets/RNA-Seq/RNAseq-mir200c.Rmd
Fix typos in gtex rna-seq analysis var names

@@
 
-Similarly to the GTEx dataset, keeping all the tests results in a strange discretness.
+Similarly to the GTEx dataset, keeping all the tests results in a strange discreteness.
 This is removed once we filter very lowly expressed genes.
@@
            covariate = "ind_covariate", maxy = 7.5)
+
 res <- filter(res, ind_covariate > 1)

### 2018-03-22 datasets/RNA-Seq/RNAseq-gtex.Rmd
Remove unneeded line of code from gtex rmd

@@
     dds <- DESeq(dsd, parallel = TRUE, BPPARAM = multicoreParam)
-    res <- results(dds, independentFiltering = FALSE)
     res <- results(dds, independentFiltering = FALSE) %>% 

### 2018-03-22 datasets/RNA-Seq/RNAseq-gtex.Rmd
Remove bib from RNAseq-gtex yaml

@@
         number_sections: true
-bibliography: bibliography.bib
 ---

### 2018-03-22 datasets/GWAS/GWAS.Rmd
Fix a couple typos in RNAseq Rmds

@@
 ```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
-# data and results directories
-system("mkdir -p DATA")
-system("mkdir -p RESULTS")
-datdir <- "DATA/"
-resdir <- "RESULTS/"
-
-# wrangling & plotting tools
 library(data.table)
@@
 library(cowplot)
+library(BiocParallel)
 
-# comparison tools/functions
-library(R.utils)
-sourceDirectory("../R/")
+## load helper functions
+for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
+    source(f)
+}
+
+# data and results directories
+datdir <- "data"
+resdir <- "results"
+dir.create(datdir, showWarnings = FALSE)
+dir.create(resdir, showWarnings = FALSE)
 
 # set up parallel backend
-library(BiocParallel)
 cores <- 8
@@
 ```{r, GWAS1-download}
-if(!file.exists(paste0(datdir,"BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt"))){
-  download.file(url="http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip", 
-             destfile=paste0(datdir,"BMI.SNPadjSMK.zip")) 
-  unzip(paste0(datdir,"BMI.SNPadjSMK.zip"), exdir=datdir)
-  file.remove(paste0(datdir,"BMI.SNPadjSMK.zip"))
+if (!file.exists(file.path(datdir, "BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt"))) {
+  download.file(url = "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip", 
+             destfile = file.path(datdir, "BMI.SNPadjSMK.zip")) 
+  unzip(file.path(datdir, "BMI.SNPadjSMK.zip"), exdir = datdir)
+  file.remove(file.path(datdir,"BMI.SNPadjSMK.zip"))
 
-  dfiles <- list.files(path=datdir, pattern="BMI.SNPadjSMK.*.txt", 
+  dfiles <- list.files(path = datdir, pattern = "BMI.SNPadjSMK.*.txt", 
                        full.names = TRUE)
@@
 ```{r, GWAS1-refdata}
-reffile <- paste0(datdir, "1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01")
-if (!file.exists(paste0(reffile, ".fam"))){
-  download.file(url="http://neurogenetics.qimrberghofer.edu.au/iSECA/1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01.zip", 
-              destfile=paste0(reffile, ".zip"))
-  unzip(paste0(reffile, ".zip"))
+reffile <- file.path(datdir, "1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01")
+if (!file.exists(paste0(reffile, ".fam"))) {
+    download.file("http://neurogenetics.qimrberghofer.edu.au/iSECA/1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01.zip", 
+                  destfile = paste0(reffile, ".zip"))
+    unzip(paste0(reffile, ".zip"), exdir = datdir)
     file.remove(paste0(reffile, ".zip"))
@@
 ```{r, GWAS1-verify-contents}
-bmi <- fread(paste0(datdir, "BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt"),
-             header=TRUE)
+bmi <- fread(file.path(datdir, "BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt"),
+             header = TRUE)
 dim(bmi)
@@
 # load reference data 
-onekg <- fread(paste0(datdir, "1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01.bim"))$V2
+onekg <- fread(file.path(datdir,  "1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01.bim"))$V2
 
@@
 bmi$BP <- as.numeric(unlist(lapply(strsplit(bmi$markername, ":"), function(x) x[[2]])))
-write_delim(bmi %>% dplyr::rename(CHR=chromosome, SNP=rs_id, A1=allele_1, A2=allele_2) %>%
+write_delim(bmi %>%
+            dplyr::rename(CHR=chromosome, SNP=rs_id, A1=allele_1, A2=allele_2) %>%
             mutate(P=0.5, F_A=Freq_Allele1_HapMapCEU, F_U=F_A, CHISQ=1, OR=1) %>% 
@@
             filter(SNP %in% onekg), 
-            path=paste0(datdir, "rslist.assoc"), delim="\t")
+            path=file.path(datdir,  "rslist.assoc"), delim="\t")
 
@@
 
-```{bash, clump, eval=FALSE}
-# use PLINK clumping tool to obtain a set of independent SNPs by LD
-plink \
---bfile DATA/1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01 \
---clump DATA/rslist.assoc \
---clump-field P \
---clump-p1 0.9999 \
---clump-p2 0.9999 \
---clump-r2 0.2 \
---clump-kb 250 \
---out DATA/plink_clump
+```{r, clump}
+## use PLINK clumping tool to obtain a set of independent SNPs by LD
+if (!dir.exists(plink_clump)) {
+    system(paste("plink",
+                 "--bfile DATA/1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01",
+                 "--clump DATA/rslist.assoc",
+                 "--clump-field P",
+                 "--clump-p1 0.9999",
+                 "--clump-p2 0.9999",
+                 "--clump-r2 0.2",
+                 "--clump-kb 250",
+                 "--out DATA/plink_clump"))
+}
 ```
@@
 ```{r, GWAS1-prune}
-clump <- fread(paste0(datdir, "plink_clump.clumped"))
+clump <- fread(file.path(datdir, "plink_clump.clumped"))
 
 # subset to include only the SNPs in the ld.list
-bmi <- bmi[bmi$rs_id %in% clump$SNP,]
+bmi <- bmi[bmi$rs_id %in% clump$SNP, ]
 
@@
 duration <- NA
-if (!file.exists(resN)){
+if (!file.exists(resN)) {
   t1 <- proc.time()
@@
 ```{r, echo=FALSE}
-if(!is.na(duration)){
+if (!is.na(duration)) {
   message("This step took ", duration, " minutes for ", nrow(bmi), " SNPs using ",
@@
 # plot nrejects by method overall and stratified by covariate
-rejections_scatter(sbN,
-                   supplementary=FALSE) +
+rejections_scatter(sbN, supplementary=FALSE) +
     ggtitle("Covariate 1: Sample size")
@@
     ggtitle("Covariate 1: Sample size")
-
 ```

### 2018-03-22 datasets/RNA-Seq/RNAseq-gtex.Rmd
Fix a couple typos in RNAseq Rmds

@@
 
-We download a DESeqDataSet object containing gene level counts for a subset of the GTEx data. The data set includes females and two cell types, nucleus accumbens and the putamen.
+We download a DESeqDataSet object containing gene level counts for GTEx samples from the nucleus accumbens and the putamen.
 
@@
 
-We use DESeq2 to test each gene for differential gene expression between the two cell types.
+We use DESeq2 to test for differential gene expression between the two cell types.
 We set the parameter `independentFiltering=FALSE` to skip the independent filtering step, as this step would be redundant with some of the FDR control methods that use gene expression as an independent covariate to increase power.
@@
 
-These discreteness, however, corresponded to genes with very low expression values.
+This discreteness, however, corresponded to genes with very low expression values.
 For example, there were `r nrow(dplyr::filter(res, ind_covariate <= 1))` genes with less than an average of 1 read across the 20 samples.
-The distribution of p-values looked much better when removing this very lowly expressed genes.
+The distribution of p-values looked much better when removing these very lowly expressed genes.
 
@@
 
-ash was the method that rejected a larger number of hypothesis, followed by lfdr and scott-theorethical.
+ash was the method that rejected the largest number of hypotheses, followed by lfdr and scott-theoretical.
 
@@
 ```{r, rnaseq-boxplots}
-## properties of method-exclusive discoveries
 methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")

### 2018-03-22 datasets/RNA-Seq/RNAseq-mir200c.Rmd
Fix a couple typos in RNAseq Rmds

@@
 As the number of samples is limited the experiment might be underpowered, as in most RNA-seq analysis.
-This is an experimental scenario that could benefit from power gaining from using modern FDR control methods.
+This is an experimental scenario that could benefit from power gained using modern FDR control methods.
 
@@
 We next subset for samples containing the control samples and the samples where mir200c was knocked down.
-Recount2 downloads data as a RangeSummarizedExperiment object, so we convert this into a DESeqDataSet object.
+recount2 downloads data as a RangeSummarizedExperiment object, so we convert this into a DESeqDataSet object.
 
@@
 
-As with the GTEx example, the mean counts is an informative covariate.
+As with the GTEx example, the mean counts is used as the informative covariate.
 
@@
 
-Similarly to the GTEx dataset, keeping all the tests results in a strange discreteness.
+Similar to the GTEx dataset, keeping all the tests results in a strange discreteness.
 This is removed once we filter very lowly expressed genes.
@@
 
-There are some warnings from both `BH` and `scott-empirical`. However, they do return results. I tried to increase the `nmids` parameter but did not help.
+There are some warnings from both `BH` and `scott-empirical`. However, they do return results.
+I tried to increase the `nmids` parameter for Scott's FDR Regression but this does not appear to make a difference.
 
@@
 
-Scott empirical rejects many more hypothesis than the rest of the methods, followed by lfdr and IHW.
+`scott-empirical` rejects many more hypothesis than the rest of the methods, followed by `lfdr` and `ihw`.
 

### 2018-03-22 datasets/GWAS/GWAS.Rmd
Fix pk typos in plink system call

@@
 ## use PLINK clumping tool to obtain a set of independent SNPs by LD
-if (!dir.exists(plink_clump)) {
-    system(paste("plink",
-                 "--bfile DATA/1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01",
-                 "--clump DATA/rslist.assoc",
+if (!file.exists(file.path(datdir, "plink_clump.clumped"))) {
+    code <- paste("plink",
+                  "--bfile", file.path(datdir, "1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01"),
+                  "--clump", file.path(datdir, "rslist.assoc"),
                   "--clump-field P",
@@
                   "--clump-kb 250",
-                 "--out DATA/plink_clump"))
+                  "--out", file.path(datdir, "plink_clump"))
+    system(code)
 }

### 2018-03-22 datasets/GWAS/GWAS.Rmd
Change results output extension to rds

@@
 ```{r, GWAS1-sb, results="hide", message=FALSE}
-resN <- paste0(resdir, "bmi_N_summarizedBenchmark_",
-               nrow(bmi), ".RData")
+resN <- file.path(resdir, paste0("bmi_N_summarizedBenchmark_", nrow(bmi), ".rds"))
 
@@
     duration <- round((proc.time()-t1)[3]/60,1)
-}else{
+} else {
     sbN <- readRDS(resN)
@@
 ```{r, GWAS1-sb2, results="hide", message=FALSE}
-resAF <- paste0(resdir, "bmi_AF_summarizedBenchmark_",
-nrow(bmi), ".RData")
+resAF <- file.path(resdir, paste0("bmi_AF_summarizedBenchmark_", nrow(bmi), ".rds"))
 
 duration <- NA
-if (!file.exists(resAF)){
+if (!file.exists(resAF)) {
     t1 <- proc.time()
@@
                      "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
-  saveRDS(sbAF, file=resAF)
-  duration <- round((proc.time()-t1)[3]/60,1)
-}else{
+    saveRDS(sbAF, file = resAF)
+    duration <- round((proc.time()-t1)[3]/60, 1)
+} else {
     sbAF <- readRDS(resAF)
@@
 ```{r, echo=FALSE}
-if(!is.na(duration)){
+if (!is.na(duration)) {
     message("This step took ", duration, " minutes for ", nrow(bmi), " SNPs using ",

### 2018-03-23 datasets/ChIP-seq/ChIPseq-h3k4me3-csaw.Rmd
Add missing metrics output to csaw rmd

@@
 ```{r, nwin-diag-hist, fig.width=10, fig.height=3.2}
-strat_hist(res_df, pvalue = "PValue", covariate = "nWindows", maxy = 35, numQ = 5)
+strat_hist(res_df, pvalue = "PValue", covariate = "nWindows", maxy = 35, numQ = 3)
 ```
@@
 ```{r, meancov-diag-hist, fig.width=10, fig.height=3.2}
-strat_hist(res_df, pvalue = "PValue", covariate = "width", maxy = 35, numQ = 5)
+strat_hist(res_df, pvalue = "PValue", covariate = "width", maxy = 35, numQ = 3)
 ```
@@
 
+## Benchmark Metrics
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, chipseq-metrics}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, chipseq-plot, results="hide"}
+rejections_scatter(sb, supplementary = FALSE)
+rejection_scatter_bins(sb, covariate = "ind_covariate",
+                       bins = 4, supplementary = FALSE)
+
+plotFDRMethodsOverlap(sb, alpha = 0.05, nsets = ncol(sb),
+                      order.by = "freq", decreasing = TRUE,
+                      supplementary = FALSE)
+
+methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
+plotCovariateBoxplots(sb, alpha = 0.1, nsets = 6,
+                      methods = methods, trans = log2, maxNum = 5)
+```
+
 # Session Info

### 2018-03-23 datasets/RNA-Seq/yeast-simulation.Rmd

@@
 
-In this Rmd we will carry out the steps in `yeast48-benchmarkFDR.Rmd` for multiple
+In this Rmd we will carry out simulations for multiple
 replicates, and plot results averaging over the replications.
@@
 single bam file for each replicate in each condition, which we'll place in a subdirectory
-called `data`.
+called `r datdir`.
 
@@
 Here we'll carry out an analysis of the full dataset, comparing the controls to the
-knowckout samples. The fold changes observed in this comparison will be used when 
+knockout samples. The fold changes observed in this comparison will be used when 
 generating the non-null simulated data in the following sections.
@@
 
-There are many thousand genes detected using the full set. Next, we'll build input data frame for summarized benchmark.
+There are several thousand genes detected using the full set. Next, we'll build input data frame for summarized benchmark.
 

### 2018-03-23 datasets/simulations/parameter-simulation.Rmd

--- 2018-01-19 datasets/simulations/analysis-05-psim-subsetonly.Rmd
@@
 ---
-title: "Benchmarking FDR Methods - Simulation Summarized ROC / UpsetR Plots"
-author: "Rafalab Journal Club Members"
-date: "`r Sys.Date()`"
+title: "Simulation Study"
+author: "Patrick Kimes"
+date: '`r format(Sys.Date(), "%B %d, %Y")`'
 output: 
@@
 
-<!-- ######################################################################## -->
-## Workspace Setup
-<!-- ######################################################################## -->
+# Summary
 
-```{r, load-libraries, message=FALSE}
-library("SummarizedBenchmark")
-library("tidyverse")
-```
-
-```{r, helper-functions, message=FALSE}
-source("../R/plotsim_aggupset.R")
-```
-
-In contrast to the t-test simulations, underlying data was not simulated in this series of simulations. Instead, for each simulation, a collection (20,000) of test statistics
-were directly sampled from univariate distributions (Normal, non-central-t, or non-central-chi-squared). Test statistics corresponding to null tests are sampled with non-centrality
-parameter zero, and non-null tests are sampled with some non-zero non-centrality parameter, determined based on the simulation setting. In these simulations, the magnitude of
-the non-centrality parameter for each test can be understood as the effect size or signal in that test. The univariate distributions (Normal, t, chi-squared) are the variyng
-sampling distributions that may be encountered under different testing scenarios (e.g. Student's t-test or Pearson's goodness-of-fit).
+For each simulation, a collection (20,000) of test statistics were directly
+sampled from univariate distributions (Normal, non-central-t, or
+non-central-chi-squared). Test statistics corresponding to null tests are
+sampled with non-centrality parameter zero, and non-null tests are sampled
+with some non-zero non-centrality parameter, determined based on the
+simulation setting. In these simulations, the magnitude of the non-centrality
+parameter for each test can be understood as the effect size or signal in
+that test. The univariate distributions (Normal, t, chi-squared) are the variyng
+sampling distributions that may be encountered under different testing scenarios
+(e.g. Student's t-test or Pearson's goodness-of-fit).
 
@@
 
-All simulation settings were carried out with an independent covariate.
-Both settings where the covariate was informative and uninformative were considered.
-Additionally, several different informative covariates were considered.
-Following the approach of Boca and Leek (2015) and Chen, Robinson and Storey (2017), we sampled the independent covariate for each test uniformly from the interval [0, 1].
-Whether the covariate was informative or uninformative was determined by a functional relationship between the value of the covariate and the probability of the corresponding test being null.
-The following relationships were considered:
+All simulation settings were carried out with an independent covariate. Both
+settings where the covariate was informative and uninformative were considered.
+Additionally, several different informative covariates were considered. Following
+the approach of Boca and Leek (2015) and Chen, Robinson and Storey (2017), we
+sampled the independent covariate for each test uniformly from the interval [0, 1].
+Whether the covariate was informative or uninformative was determined by a functional
+relationship between the value of the covariate and the probability of the
+corresponding test being null. The following relationships were considered:
 
@@
 Here, we only consider a subset of the entire set of simulation settings.
-Namely, we look at: [1] null simulations, [2] non-null simulations with an informative covariate (cubic), and [3] non-null simulations with an uninformative covariate.
-For each of the 3 simulation sets, we further consider three different sampling distributions of the test statistics (Normal, t, chi-squared).
+Namely, we look at: [1] null simulations, [2] non-null simulations with an
+informative covariate (cubic), and [3] non-null simulations with an uninformative
+covariate. For each of the 3 simulation sets, we further consider three different
+sampling distributions of the test statistics (Normal, t, chi-squared).
 
-Performance is evaluated with: the false positive rate (FPR), true positive rate (TPR), and method overlaps (upset plots).
-For null simulations, since FPR and TPR are useless, we instead plot the proportion of rejected tests.
+Performance is evaluated with: the false positive rate (FPR), true positive rate (TPR),
+and method overlaps (upset plots). For null simulations, since FPR and TPR are useless,
+we instead plot the proportion of rejected tests.
 
-<!-- ######################################################################## -->
-## Local Helper Functions
-<!-- ######################################################################## -->
+# Workspace Setup
 
-```{r}
+```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
+library("SummarizedBenchmark")
+library("tidyverse")
+
+source(file.path("..", "R", "plotsim_aggupset.R"))
+```
+
+# Helper Functions
+
+```{r helper-preclean}
 preclean <- function(res, merge_ihw = TRUE, supplementary = FALSE) {
@@
 }
+```
 
+```{r helper-myplot}
 myplot <- function(res, metrics = c("TPR", "FPR"), alpha_max = 1) {
@@
 
-<!-- ######################################################################## -->
-## Null Simulations
-<!-- ######################################################################## -->
+# Null Simulations
 
-First, we look at the null simulation settings to assess which methods control false discoveries when no truly significant tests are present.
-The TPR is not reported here because there are no "true positivies" in the null setting.
+First, we look at the null simulation settings to assess which methods control
+false discoveries when no truly significant tests are present. The TPR is not
+reported here because there are no "true positivies" in the null setting.
 
@@
 
-### Normal
+## Normal
 
@@
 
-### t (10 dof)
+## t (10 dof)
 
@@
 
-### chi-sq (4 dof)
+## chi-sq (4 dof)
 
@@
 
-<!-- ######################################################################## -->
-## Informative Covariate Simulations
-<!-- ######################################################################## -->
+# Informative Covariate Simulations
 
@@
 
-<!-- ######################################################################## -->
-## Uninformative Covariate Simulations
-<!-- ######################################################################## -->
+# Uninformative Covariate Simulations
 

### 2018-03-26 datasets/GSEA/gsea-human.Rmd

--- 2018-03-22 datasets/GSEA/gsea-human.Rmd
@@
 
-**Some sort of summary of the analysis.**
+The objective of this document is to test different multiple testing methods in the context of Gene Set Enrichment Analysis (GSEA).
+In order to get a list of genes, we will use RNA-seq data consisting on cortex and cerebellum samples of a subset of the GTEx data.
 
@@
 
-**Some sort of summary of the data set.**  
-We download a *DESeqDataSet* if not present locally.
+The dataset consists of 10 samples (6 cerebellum and 4 cortex samples), all from male individuals.
+We download a *DESeqDataSet* if not present locally and modify the design to test for differences in gene expression between cerebellums and cortex.
 
@@
 dsdObject <- readRDS(count_file)
+design( dsdObject ) <- ~ tissue
 ```
 
-We only perform differential tesing on protein coding genes, as specified in Ensembl release 86.
+To keep running times short, we only perform differential tesing on protein coding genes, as specified in Ensembl release 86.
 
@@
 
-**Add description of what's going on.**
-We first test for differentially expressed genes using DESeq2 and use genes with adjusted p-value less than 0.1.
+Next, we run DESeq2 to retrieve a list of differentially expressed genes at a FDR of 10%.
 
@@
 
-We next determine the gene sets associated with each gene.
+Next we’ll use the biomaRt package to download and associate GO annotations for each gene.
 
@@
 
-Now we are ready to perform gene set enrichment analysis using `goseq`.
+Finally, we’ll perform a gene set enrichment analysis with goseq.
+goseq performs GSEA and adjusts for gene length biases.
+The bias is represented in the plot below, with gene length bin on the x-axis and proportion of genes detected to be differentially expressed on the y-axis.
 
@@
 
-**Add description of independent covariate.**
+We will explore whether the size of the gene set can be used as a covariate for modern multiple-testing correction methods in the context of GSEA.
+In the plot below, the log10 of the p-values is plotted as a function of the size of the gene set.
+There is a pattern in which gene sets with a higher number of genes tend to have smaller p-values, which is indicative that gene set size is an informative covariate.
 
@@
 
+We can also explore if the covariate seems to be independent under the null.
+
 ```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
@@
 
-Generating the *SummarizedBenchmark* object:
+We then execute the benchDesign and generate a *SummarizedBenchmark* object containing multiple-testing corrections using several methods.
 
@@
 
-# Data Analysis (Null Comparison)
-
-**Add description.**
-
-## Create Null Data
-
-```{r, eval=FALSE }
-nms <- names( genes )
-xx <- seq_len(length(nms))
-gns <- 1*( xx %in% sample( xx, 1000, replace=FALSE ) )
-names( gns ) <- nms
-pwf <- nullp( gns, bias.data=geneLength )
-goRes <- goseq( pwf, gene2cat=goSets )
-res <- goRes %>%
-    dplyr::filter( numDEInCat > 0 ) %>%
-    dplyr::select( c("over_represented_pvalue", "numInCat") ) %>%
-    dplyr::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
-```
-
-## Multiple-Testing Correction
-
-```{r}
-pr <- buildBench(bd, res, ftCols = "ind_covariate")
-assayNames(pr) <- "qvalue"
-pr <- addDefaultMetrics(pr)
-```
-
 ## Benchmark Metrics

### 2018-03-26 datasets/GSEA/gsea-mouse.Rmd

@@
 
-**Some sort of summary of the analysis.**
-Same analysis but using the HSC-MPP dataset from Nina.  
-**Who is Nina?**
+The objective of this vignette is to test different multiple testing methods in the context of Gene Set Enrichment Analysis (GSEA).
+To do this, we will use mouse data from the paper by [Cabezas-Wallscheid et al.
+(Cell stem Cell, 2014)](https://www.sciencedirect.com/science/article/pii/S1934590914003014?via%3Dihub).
+The dataset consists of RNA-seq data from mouse hematopoietic stem cells and 4 multipotent progenitor lineages.
+We will test for difference in gene expression between two different cell fractions and perform GSEA in the resulting list of genes.
 
@@
 
-Download *DESeqDataSet* if not present locally.
+The data has been preprocessed and saved as a DESeqDataset object. The following lines of code download this *DESeqDataSet* if it is not present locally.
 
@@
 
-**Why no filtering on gene_biotype == "protein_coding" for this data?**
+**Why no filtering on gene_biotype == "protein_coding" for this data?
+AR: I think I did this in the other vignette to keep running times short.
+This shouldn't change the result since anyway protein coding genes are those with GO categories and genes without GO assignments are ignored.**
 
@@
 
-**Add description of what's going on.**
-We first test for differentially expressed genes using DESeq2 and use genes with adjusted p-value less than 0.1.
+In order to get a list of genes to test for GSEA, we will test each gene for differential gene expression between hematopoietic stem cells and multipotent progenitors (fraction 1).
+To do this, we will run DESeq2 to retrieve genes differentially expressed at a FDR of 10%.
 
@@
 
-We next determine the gene sets associated with each gene.
+We next use biomaRt to get the relations between GO categories and genes.
 
@@
 
-Here, we can see that the covariant variable is actually informative.
+Here, we can see that the size of the gene set is actually informative.
 
@@
 
-**Add description of independent covariate.**
+In the following plot, we explore the relationship between the p-value and the gene set size. We can see that this covariate is actually informative.
 
@@
 ```
+
+And at the same time, it seems to be fair to assume that it is independent under the null hypothesis.
 

### 2018-03-27 datasets/scrnaseq/scrnaseq-human-mast.Rmd
Standardize scrnaseq analyses
Change paste0 to file.path, source in R scripts without using the
R.utils package, remove unnecessary libraries.

@@
               technology=c("smart-seq","smart-seq2"),
-              comparison=c("Neoplastic vs Regular",
+              comparison=c("Tumor core vs Periphery",
                            "LPS vs No stimulation"))
 kable(df,caption="scRNAseq Datasets")
-df1=data.frame(DE_method=c("ScDD","MAST"),
+df1=data.frame(DE_method=c("scDD","MAST"),
                input=c("count","tpm"),
@@
 ```{r, workspace-setup, results='hide', message=FALSE, warning=FALSE}
-system("mkdir -p DATA")
-system("mkdir -p RESULTS")
-datdir <- "DATA/"
-resdir <- "RESULTS/"
-
-# packages for data analyses
 library(MultiAssayExperiment)
+library(SingleCellExperiment)
 library(dplyr)
@@
 library(cowplot)
-library(UpSetR)
-library(hexbin)
-
-# packages for differential expression testing
 library(impute)
-library(scDD)
-library(EBSeq)
 library(edgeR)
 library(MAST)
-
-# summarizebenchmark
-library(SummarizedBenchmark)
-library(SingleCellExperiment)
-
-# useful libraries
-library(R.utils)
 library(BiocParallel)
 
+## load helper functions
+for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
+    source(f)
+}
+
+# data and results directories
+datdir <- "data"
+resdir <- "results"
+dir.create(datdir, showWarnings = FALSE)
+dir.create(resdir, showWarnings = FALSE)
+
+# set up parallel backend
 cores <- 8
 BiocParallel::register(MulticoreParam(workers = cores))
-
-sourceDirectory("../R/")
 ```
@@
 ```{r, humandata-download}
-datfile <- paste0(datdir, "GSE84465.rds")
+datfile <- file.path(datdir, "GSE84465.rds")
 
@@
 ```{r}
-datfile <- paste0(datdir, "human_GSE84465_mast.rds")
+datfile <- file.path(datdir, "human_GSE84465_mast.rds")
 
@@
   # load MultiAssayExperiment
-  humandata <- readRDS(paste0(datdir, "GSE84465.rds"))
+  humandata <- readRDS(file.path(datdir, "GSE84465.rds"))
 
@@
 
-  # subset by genes with detection rate > 1%
-  humandata <- humandata[rowMeans(assay(humandata, "count") > 0) > 0.01, ]
+  # subset by genes with detection rate > 5% in at least one condition
+  levs <- unique(colData(humandata)$group)
+  humandata <- humandata[rowMeans(assay(humandata, "count")[, 
+                            colData(humandata)$group == levs[1]] > 0) > 0.05 |
+                         rowMeans(assay(humandata, "count")[, 
+                            colData(humandata)$group == levs[2]] > 0) > 0.05, ]
 
@@
 ```{r, sb1, results="hide", message=FALSE}
-res1 <- paste0(resdir, "sb_GSE84465_mast_mean.rds")
+res1 <- file.path(resdir, "sb_GSE84465_mast_mean.rds")
 
@@
 ```{r, sb2, results="hide", message=FALSE}
-res2 <- paste0(resdir, "sb_GSE84465_mast_det.rds")
+res2 <- file.path(resdir, "sb_GSE84465_mast_det.rds")
 

### 2018-03-27 datasets/scrnaseq/scrnaseq-human-scdd.Rmd
Standardize scrnaseq analyses
Change paste0 to file.path, source in R scripts without using the
R.utils package, remove unnecessary libraries.

@@
 
+Note that this Rmd also relies on scDD version 1.3.4 or higher (install from 
+Bioconductor devel, or from the master repo on GitHub kdkorthauer/scDD), since 
+additional functionality to skip computation of the classification of genes into 
+patterns was implemented.
+
 ```{r, workspace-setup, results='hide', message=FALSE, warning=FALSE}
-system("mkdir -p DATA")
-system("mkdir -p RESULTS")
-datdir <- "DATA/"
-resdir <- "RESULTS/"
-
-# packages for data analyses
 library(MultiAssayExperiment)
+library(SingleCellExperiment)
 library(dplyr)
@@
 library(cowplot)
-library(UpSetR)
-library(hexbin)
-
-# packages for differential expression testing
 library(impute)
@@
 library(EBSeq)
-library(edgeR)
-library(MAST)
-
-# summarizebenchmark
-library(SummarizedBenchmark)
-library(SingleCellExperiment)
-
-# useful libraries
-library(R.utils)
 library(BiocParallel)
 
+if( !packageVersion("scDD") >= '1.3.4'){
+  stop("This code relies on scDD version 1.3.4 or higher - see note above")
+}
+
+## load helper functions
+for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
+    source(f)
+}
+
+# data and results directories
+datdir <- "data"
+resdir <- "results"
+dir.create(datdir, showWarnings = FALSE)
+dir.create(resdir, showWarnings = FALSE)
+
+# set up parallel backend
 cores <- 8
 BiocParallel::register(MulticoreParam(workers = cores))
-
-sourceDirectory("../R/")
 ```
@@
 ```{r}
-datfile <- paste0(datdir, "human_GSE84465_scdd.rds")
+datfile <- file.path(datdir, "human_GSE84465_scdd.rds")
 
@@
              prior_param = list(alpha=0.01, mu0=0, s0=0.01, a0=0.01, b0=0.01), 
-             testZeroes = TRUE))
-    rowData(dat)$scdd_p <- res$nonzero.pvalue
+             testZeroes = TRUE,
+             categorize = FALSE))
+    rowData(dat)$scdd_p_nonzero <- res$nonzero.pvalue
+    rowData(dat)$scdd_p_zero <- res$zero.pvalue
     rowData(dat)$scdd_p_combined <- res$combined.pvalue
@@
 rank_scatter(data.frame(rowData(humandata)), 
-             pvalue="scdd_p", covariate="meanExp") + 
+             pvalue="scdd_p_combined", covariate="meanExp") + 
   ggtitle("scDD, Covariate 1: Mean Nonzero Expression")
@@
 strat_hist(data.frame(rowData(humandata)), 
-           pvalue="scdd_p", covariate="meanExp", maxy=12,
+           pvalue="scdd_p_combined", covariate="meanExp", maxy=12,
            main = "scDD, Covariate 1: Mean Nonzero Expression") 
@@
 rank_scatter(data.frame(rowData(humandata)), 
-             pvalue="scdd_p", covariate="detection") + 
+             pvalue="scdd_p_combined", covariate="detection") + 
   ggtitle("scDD, Covariate 2: Detection Rate")
@@
 strat_hist(data.frame(rowData(humandata)), 
-           pvalue="scdd_p", covariate="detection", maxy=12,
+           pvalue="scdd_p_combined", covariate="detection", maxy=12,
            main = "scDD, Covariate 2: Detection Rate") 
@@
 ```{r, sb1, results="hide", message=FALSE}
-res1 <- paste0(resdir, "sb_GSE84465_scdd_mean.rds")
+res1 <- file.path(resdir, "sb_GSE84465_scdd_mean.rds")
 
@@
                                    na.omit() %>%
-                                   mutate(pval=scdd_p,
+                                   mutate(pval=scdd_p_combined,
                                           ind_covariate=meanExp), 
@@
 ```{r, sb2, results="hide", message=FALSE}
-res2 <- paste0(resdir, "sb_GSE84465_scdd_det.rds")
+res2 <- file.path(resdir, "sb_GSE84465_scdd_det.rds")
 
@@
                                    na.omit() %>%
-                                   mutate(pval=scdd_p,
+                                   mutate(pval=scdd_p_combined,
                                           ind_covariate=detection), 

### 2018-03-27 datasets/scrnaseq/scrnaseq-mouse-mast.Rmd
Standardize scrnaseq analyses
Change paste0 to file.path, source in R scripts without using the
R.utils package, remove unnecessary libraries.

@@
 ```{r, mousedata-download}
-datfile <- paste0(datdir,"GSE94383.rds")
+datfile <- file.path(datdir,"GSE94383.rds")
 
@@
 
+*Add More detail on the experiments here*
 We'll extract two conditions: No stimulation versus LPS at 150 min.
@@
 ```{r}
-datfile <- paste0(datdir, "mouse_GSE94383_mast.rds")
+datfile <- file.path(datdir, "mouse_GSE94383_mast.rds")
 
@@
   # load MultiAssayExperiment
-  mousedata=readRDS(paste0(datdir,"GSE94383.rds"))
+  mousedata=readRDS(file.path(datdir,"GSE94383.rds"))
 
@@
 
-  # subset by genes with detection rate > 1%
-  mousedata <- mousedata[rowMeans(assay(mousedata, "count") > 0) > 0.01, ]
+  # subset by genes with detection rate > 5% in at least one condition
+  levs <- unique(colData(mousedata)$group)
+  mousedata <- mousedata[rowMeans(assay(mousedata, "count")[, 
+                            colData(mousedata)$group == levs[1]] > 0) > 0.05 |
+                         rowMeans(assay(mousedata, "count")[, 
+                            colData(mousedata)$group == levs[2]] > 0) > 0.05, ]
 
@@
 strat_hist(data.frame(rowData(mousedata)),
-           pvalue="mast_p", covariate="meanExp", maxy=16,
+           pvalue="mast_p", covariate="meanExp", maxy=20,
            main = "MAST, Covariate 1: Mean Nonzero Expression")
@@
 strat_hist(data.frame(rowData(mousedata)), 
-           pvalue="mast_p", covariate="detection", maxy=16,
+           pvalue="mast_p", covariate="detection", maxy=20,
            main = "MAST, Covariate 2: Detection Rate")
@@
 ```{r, sb1, results="hide", message=FALSE}
-res1 <- paste0(resdir, "sb_GSE94383_mast_mean.rds")
+res1 <- file.path(resdir, "sb_GSE94383_mast_mean.rds")
 
@@
 ```{r, sb2, results="hide", message=FALSE}
-res2 <- paste0(resdir, "sb_GSE94383_mast_det.rds")
+res2 <- file.path(resdir, "sb_GSE94383_mast_det.rds")
 

### 2018-03-27 datasets/scrnaseq/scrnaseq-mouse-scdd.Rmd
Standardize scrnaseq analyses
Change paste0 to file.path, source in R scripts without using the
R.utils package, remove unnecessary libraries.

@@
 ```{r}
-datfile <- paste0(datdir, "mouse_GSE94383_scdd.rds")
+datfile <- file.path(datdir, "mouse_GSE94383_scdd.rds")
 
@@
 rank_scatter(data.frame(rowData(mousedata)), 
-             pvalue="scdd_p", covariate="meanExp") + 
+             pvalue="scdd_p_combined", covariate="meanExp") + 
   ggtitle("scDD, Covariate 1: Mean Nonzero Expression")
@@
 strat_hist(data.frame(rowData(mousedata)), 
-           pvalue="scdd_p", covariate="meanExp", maxy=16,
+           pvalue="scdd_p_combined", covariate="meanExp", maxy=20,
            main = "scDD, Covariate 1: Mean Nonzero Expression") 
@@
 rank_scatter(data.frame(rowData(mousedata)), 
-             pvalue="scdd_p", covariate="detection") + 
+             pvalue="scdd_p_combined", covariate="detection") + 
   ggtitle("scDD, Covariate 2: Detection Rate")
@@
 strat_hist(data.frame(rowData(mousedata)), 
-           pvalue="scdd_p", covariate="detection", maxy=16,
+           pvalue="scdd_p_combined", covariate="detection", maxy=20,
            main = "scDD, Covariate 2: Detection Rate") 
@@
 ```{r, sb1, results="hide", message=FALSE}
-res1 <- paste0(resdir, "sb_GSE94383_scdd_mean.rds")
+res1 <- file.path(resdir, "sb_GSE94383_scdd_mean.rds")
 
@@
 ```{r, sb2, results="hide", message=FALSE}
-res2 <- paste0(resdir, "sb_GSE94383_scdd_det.rds")
+res2 <- file.path(resdir, "sb_GSE94383_scdd_det.rds")
 

### 2018-03-27 datasets/scrnaseq/scrnaseq-knit.sh

@@ -1,20 +1,21 @@
 #!/bin/bash
-#SBATCH -J knit-sc                  # A single job name for the array
-#SBATCH -n 8                        # Number of cores
-#SBATCH -N 1                        # All cores on one machine
-#SBATCH -p shared                   # Partition
-#SBATCH --mem 125G                  # Memory request 
-#SBATCH -t 0-6:00                   # Maximum execution time (D-HH:MM)
+#SBATCH -J knit-sc
+#SBATCH -n 8
+#SBATCH -N 1
+#SBATCH -p shared
+#SBATCH --mem 125G
+#SBATCH -t 0-6:00
 #SBATCH -o render-%j.out
 #SBATCH -e render-%j.err
  
-export SLURM_NTASKS
 export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/rstudio/0.98.1103-fasrc01/bin/pandoc/"
 
 # change filename to Rmd to be knitted
 # make sure number of cores used in Rmd matches sbatch param -n
-R -e "rmarkdown::render('scrnaseq-mouse-mast.Rmd', clean=FALSE)"
+R -e "rmarkdown::render('scrnaseq-human-scdd.Rmd', clean=FALSE)"
 R -e "rmarkdown::render('scrnaseq-mouse-scdd.Rmd', clean=FALSE)"
 R -e "rmarkdown::render('scrnaseq-human-mast.Rmd', clean=FALSE)"
-R -e "rmarkdown::render('scrnaseq-human-scdd.Rmd', clean=FALSE)"
+R -e "rmarkdown::render('scrnaseq-mouse-mast.Rmd', clean=FALSE)"
+
+
 

### 2018-03-27 datasets/GWAS/GWAS.Rmd
Remove covariate heatmaps and boxplots

@@
                       supplementary=FALSE) 
-
-# properties of method-exclusive discoveries
-methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
-             "scott-empirical", "scott-theoretical")
-mcols(sbN)$ind_covariate <- mcols(sbN)$ind_covar_N
-plotCovariateBoxplots(sbN, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3) +
-    ggtitle("Covariate 1: Sample size")
 ```
@@
 ```{r, fig.width=8, fig.height=3.5}
+mcols(sbN)$ind_covariate <- mcols(sbN)$ind_covar_N
 mcols(sbN)$tstat <- bmi %>% filter(!is.na(ind_covar_N)) %>% pull(effect_size)
-covariateHeatmap(sbN, alpha=0.05, covname="tstat", trans = "log1p")
-covariateHeatmap(sbN, alpha=0.05, covname="ind_covar_N", trans = "log1p")
-
 covariateHeatmap(sbN, alpha=0.05, covname="tstat", nbins=25, 
@@
                       supplementary=FALSE)
-
-# properties of method-exclusive discoveries
-methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
-             "scott-empirical", "scott-theoretical")
-mcols(sbAF)$ind_covariate <- abs(mcols(sbAF)$ind_covar_AF-0.5)
-plotCovariateBoxplots(sbAF, alpha=0.1, nsets=6, methods=methods, maxNum=3)+
-  ggtitle("Covariate 2: Minor allele frequency")
 ```
@@
 ```{r, fig.width=8, fig.height=3.5}
+mcols(sbAF)$ind_covariate <- abs(mcols(sbAF)$ind_covar_AF-0.5)
 mcols(sbAF)$tstat <- bmi %>% filter(!is.na(ind_covar_AF)) %>% pull(effect_size)
-covariateHeatmap(sbAF, alpha=0.05, covname="tstat", trans = "log1p")
-covariateHeatmap(sbAF, alpha=0.05, covname="ind_covar_AF")
-
 covariateHeatmap(sbAF, alpha=0.05, covname="tstat", nbins=25, 

### 2018-03-27 datasets/scrnaseq/scrnaseq-human-mast.Rmd
Remove covariate heatmaps and boxplots

@@
 
-# properties of method-exclusive discoveries
-methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
-             "scott-empirical", "scott-theoretical")
-mcols(sb1)$ind_covariate <- mcols(sb1)$meanExp
-plotCovariateBoxplots(sb1, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3) +
-  ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
-
 ```
@@
 ```{r, fig.width=8, fig.height=3.5}
-covariateHeatmap(sb1, alpha=0.05, covname="ind_covariate") +
-  ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
-
+mcols(sb1)$ind_covariate <- mcols(sb1)$meanExp
 covariateHeatmap(sb1, alpha=0.05, covname="ind_covariate", nbins=25, 
@@
 
-# properties of method-exclusive discoveries
-methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
-             "scott-empirical", "scott-theoretical")
-mcols(sb2)$ind_covariate <- mcols(sb2)$detection
-plotCovariateBoxplots(sb2, alpha=0.1, nsets=6, methods=methods, maxNum=3)+
-  ggtitle("MAST, Covariate 2: Detection Rate")
-
 ```
@@
 ```{r, fig.width=8, fig.height=3.5}
-covariateHeatmap(sb2, alpha=0.05, covname="ind_covariate") +
-  ggtitle("MAST, Covariate 2: Detection Rate")
-
+mcols(sb2)$ind_covariate <- mcols(sb2)$detection
 covariateHeatmap(sb2, alpha=0.05, covname="ind_covariate", nbins=25, 

### 2018-03-27 datasets/scrnaseq/scrnaseq-human-scdd.Rmd
Remove covariate heatmaps and boxplots

@@
                       supplementary=FALSE) 
-
-# properties of method-exclusive discoveries
-methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
-             "scott-empirical", "scott-theoretical")
-mcols(sb1)$ind_covariate <- mcols(sb1)$meanExp
-plotCovariateBoxplots(sb1, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3) +
-  ggtitle("scDD, Covariate 1: Mean Nonzero Expression")
-
 ```
@@
 ```{r, fig.width=8, fig.height=3.5}
-covariateHeatmap(sb1, alpha=0.05, covname="ind_covariate") +
-  ggtitle("scDD, Covariate 1: Mean Nonzero Expression")
-
+mcols(sb1)$ind_covariate <- mcols(sb1)$meanExp
 covariateHeatmap(sb1, alpha=0.05, covname="ind_covariate", nbins=25, 
@@
                       supplementary=FALSE)
-
-# properties of method-exclusive discoveries
-methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
-             "scott-empirical", "scott-theoretical")
-mcols(sb2)$ind_covariate <- mcols(sb2)$detection
-plotCovariateBoxplots(sb2, alpha=0.1, nsets=6, methods=methods, maxNum=3) +
-  ggtitle("scDD, Covariate 2: Detection Rate")
-
 ```
@@
 ```{r, fig.width=8, fig.height=3.5}
-covariateHeatmap(sb2, alpha=0.05, covname="ind_covariate") +
-  ggtitle("scDD, Covariate 2: Detection Rate")
-
+mcols(sb2)$ind_covariate <- mcols(sb2)$detection
 covariateHeatmap(sb2, alpha=0.05, covname="ind_covariate", nbins=25, 

### 2018-03-27 datasets/scrnaseq/scrnaseq-mouse-mast.Rmd
Remove covariate heatmaps and boxplots

@@
                       supplementary=FALSE) 
-
-# properties of method-exclusive discoveries
-methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
-             "scott-empirical", "scott-theoretical")
-mcols(sb1)$ind_covariate <- mcols(sb1)$meanExp
-plotCovariateBoxplots(sb1, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3) +
-  ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
-
 ```
@@
                       supplementary=FALSE)
-
-# properties of method-exclusive discoveries
-methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
-             "scott-empirical", "scott-theoretical")
-mcols(sb2)$ind_covariate <- mcols(sb2)$detection
-plotCovariateBoxplots(sb2, alpha=0.1, nsets=6, methods=methods, maxNum=3) +
-  ggtitle("MAST, Covariate 2: Detection Rate")
-
 ```

### 2018-03-27 datasets/GWAS/GWAS.Rmd
Change covariateHeatmap function name to covariateLinePlot

@@
 mcols(sbN)$tstat <- bmi %>% filter(!is.na(ind_covar_N)) %>% pull(effect_size)
-covariateHeatmap(sbN, alpha=0.05, covname="tstat", nbins=25, 
+covariateLinePlot(sbN, alpha=0.05, covname="tstat", nbins=25, 
                  linePlot = TRUE, trans="log1p")
-covariateHeatmap(sbN, alpha=0.05, covname="ind_covar_N", nbins=25, 
+covariateLinePlot(sbN, alpha=0.05, covname="ind_covar_N", nbins=25, 
                  linePlot = TRUE, trans="log1p")
@@
 mcols(sbAF)$tstat <- bmi %>% filter(!is.na(ind_covar_AF)) %>% pull(effect_size)
-covariateHeatmap(sbAF, alpha=0.05, covname="tstat", nbins=25, 
+covariateLinePlot(sbAF, alpha=0.05, covname="tstat", nbins=25, 
                  linePlot = TRUE, trans = "log1p")
-covariateHeatmap(sbAF, alpha=0.05, covname="ind_covar_AF", nbins=25, 
+covariateLinePlot(sbAF, alpha=0.05, covname="ind_covar_AF", nbins=25, 
                  linePlot = TRUE)

### 2018-03-27 datasets/scrnaseq/scrnaseq-human-mast.Rmd
Change covariateHeatmap function name to covariateLinePlot

@@
 mcols(sb1)$ind_covariate <- mcols(sb1)$meanExp
-covariateHeatmap(sb1, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 linePlot = TRUE) +
+covariateLinePlot(sb1, alpha=0.05, covname="ind_covariate", nbins=25) +
   ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
@@
 mcols(sb2)$ind_covariate <- mcols(sb2)$detection
-covariateHeatmap(sb2, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 linePlot = TRUE) +
+covariateLinePlot(sb2, alpha=0.05, covname="ind_covariate", nbins=25) +
   ggtitle("MAST, Covariate 2: Detection Rate")

### 2018-03-27 datasets/scrnaseq/scrnaseq-human-scdd.Rmd
Change covariateHeatmap function name to covariateLinePlot

@@
 mcols(sb1)$ind_covariate <- mcols(sb1)$meanExp
-covariateHeatmap(sb1, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 linePlot = TRUE) +
+covariateLinePlot(sb1, alpha=0.05, covname="ind_covariate", nbins=25) +
   ggtitle("scDD, Covariate 1: Mean Nonzero Expression")
@@
 mcols(sb2)$ind_covariate <- mcols(sb2)$detection
-covariateHeatmap(sb2, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 linePlot = TRUE) +
+covariateLinePlot(sb2, alpha=0.05, covname="ind_covariate", nbins=25) +
   ggtitle("scDD, Covariate 2: Detection Rate")

### 2018-03-27 datasets/scrnaseq/scrnaseq-mouse-mast.Rmd
Change covariateHeatmap function name to covariateLinePlot

@@
 mcols(sb1)$ind_covariate <- mcols(sb1)$meanExp
-covariateHeatmap(sb1, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 linePlot = TRUE, trans = "log1p") +
+covariateLinePlot(sb1, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans = "log1p") +
   ggtitle("MAST, Covariate 1: Mean Nonzero Expression")

### 2018-03-27 datasets/scrnaseq/scrnaseq-mouse-scdd.Rmd
Change covariateHeatmap function name to covariateLinePlot

@@
 mcols(sb1)$ind_covariate <- mcols(sb1)$meanExp
-covariateHeatmap(sb1, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 linePlot = TRUE, trans = "log1p") +
+covariateLinePlot(sb1, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans = "log1p") +
   ggtitle("scDD, Covariate 1: Mean Nonzero Expression")
@@
 
-# properties of method-exclusive discoveries
-methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
-             "scott-empirical", "scott-theoretical")
-mcols(sb2)$ind_covariate <- mcols(sb2)$detection
-plotCovariateBoxplots(sb2, alpha=0.1, nsets=6, methods=methods, maxNum=3) +
-  ggtitle("scDD, Covariate 2: Detection Rate")
-
 ```
 ```{r, fig.width=8, fig.height=3.5}
-covariateHeatmap(sb2, alpha=0.05, covname="ind_covariate") +
-  ggtitle("scDD, Covariate 2: Detection Rate")
-
-covariateHeatmap(sb2, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 linePlot = TRUE) +
+mcols(sb2)$ind_covariate <- mcols(sb2)$detection
+covariateLinePlot(sb2, alpha=0.05, covname="ind_covariate", nbins=25) +
   ggtitle("scDD, Covariate 2: Detection Rate")

### 2018-03-27 datasets/R/effectSizePlot.R

@@ -75,7 +75,7 @@ return(p)
 }
 
 
-#' Function to create heatmap of percentage significance for different covariate 
+#' Function to create lineplot of percentage significance for different covariate 
 #' value bins.
 #' 
 #' Assumes that covariates are found 
@@ -94,16 +94,18 @@ return(p)
 #' @param covname character that specifies column name of rowData to use for 
 #' y-axis (e.g. effect size, or ind_covariate)
 #' @param nbins positive inter value indicating how many bins to divide the 
-#' covariate values into for the heatmap rows. Defaults to 50.
+#' covariate values into for the x-axis (lineplot) or rows (heatmap). 
+#' Defaults to 50.
 #' @param trans character indicating a transformation to apply to the scale
-#' of fill (e.g. "log1p"). The default NULL means no transformation. 
+#' of the y-axis (lineplot) or fill (heatmap). For example, "log1p" applies 
+#' the log(x+1) transform. The default NULL means no transformation. 
 #' @param linePlot logical whether to plot a line plot (smoothed density) 
-#' instead of heatmap (default is FALSE).
+#' (default is TRUE). If false, will plot a heatmap.
 #' 
 #' @return a ggplot object
-covariateHeatmap <- function(sbl, alpha=0.05, nbins = 50, 
+covariateLinePlot <- function(sbl, alpha=0.05, nbins = 50, 
                              covname, trans = NULL, 
-                             linePlot = FALSE){
+                             linePlot = TRUE){
   
   summarize_one_item <- function(object, alpha, nbins){
     object <- object[,!( grepl("^ihw", as.character( colData( object )$blabel ))

### 2018-03-27 datasets/RNA-Seq/yeast-simulation.Rmd
Standardize and update covariate plots
Change paste0 to file.path, rename covariate Heatmap function to
covariateLinePlot.

@@
 library(magrittr)
-library(R.utils)
 library(cowplot)
@@
 
-# benchmark methods
-library(IHW)
-library(ashr)
-library(qvalue)
-library(swfdr)
-library(fdrtool)
-library(FDRreg)
-
-# comparison tools/functions
-library(SummarizedBenchmark)
-sourceDirectory("../R/")
+## load helper functions
+for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
+    source(f)
+}
 
@@
 download.file(url = "https://github.com/bartongroup/profDGE48/raw/master/Preprocessed_data/Snf2_countdata.tar.gz",
-              destfile = paste0(datdir, "/Snf2_countdata.tar.gz"))
+              destfile = file.path(datdir, "Snf2_countdata.tar.gz"))
 download.file(url = "https://github.com/bartongroup/profDGE48/raw/master/Preprocessed_data/WT_countdata.tar.gz", 
-              destfile = paste0(datdir, "/WT_countdata.tar.gz"))
+              destfile = file.path(datdir, "WT_countdata.tar.gz"))
 
-gunzip(paste0(datdir, "/Snf2_countdata.tar.gz")) 
-gunzip(paste0(datdir, "/WT_countdata.tar.gz"))
+gunzip(file.path(datdir, "Snf2_countdata.tar.gz")) 
+gunzip(file.path(datdir, "WT_countdata.tar.gz"))
 
-untar(paste0(datdir, "/Snf2_countdata.tar"), exdir = datdir) 
-untar(paste0(datdir, "/WT_countdata.tar"), exdir = datdir)
+untar(file.path(datdir, "Snf2_countdata.tar"), exdir = datdir) 
+untar(file.path(datdir, "WT_countdata.tar"), exdir = datdir)
 
-file.remove(paste0(datdir, "/Snf2_countdata.tar"), 
-            paste0(datdir, "/WT_countdata.tar"))
+file.remove(file.path(datdir, "Snf2_countdata.tar"), 
+            file.path(datdir, "WT_countdata.tar"))
 ```
@@
 nDE <- 0
-resfile <- paste0(resdir, "/null5.rds")
+resfile <- file.path(resdir, "null5.rds")
 
@@
 
-covariateHeatmap(null5, alpha=0.05, covname="log2FC", trans="log1p")
-covariateHeatmap(null5, alpha=0.05, covname="ind_covariate", trans="log1p")
-
-covariateHeatmap(null5, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p", linePlot = TRUE)
-covariateHeatmap(null5, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 linePlot = TRUE, trans="log1p")
+covariateLinePlot(null5, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(null5, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
 
@@
 nDE <- 0
-resfile <- paste0(resdir, "/null10.rds")
+resfile <- file.path(resdir, "null10.rds")
 
@@
 
-covariateHeatmap(null10, alpha=0.05, covname="log2FC", trans="log1p")
-covariateHeatmap(null10, alpha=0.05, covname="ind_covariate", trans="log1p")
-
-covariateHeatmap(null10, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p", linePlot = TRUE)
-covariateHeatmap(null10, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 linePlot = TRUE, trans="log1p")
+covariateLinePlot(null10, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(null10, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
 
@@
 nDE <- 500
-resfile <- paste0(resdir, "/de5.rds")
+resfile <- file.path(resdir, "de5.rds")
 
@@
 
-covariateHeatmap(de5, alpha=0.05, covname="log2FC", trans="log1p")
-covariateHeatmap(de5, alpha=0.05, covname="ind_covariate")
-
-covariateHeatmap(de5, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p", linePlot = TRUE)
-covariateHeatmap(de5, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 linePlot = TRUE)
+covariateLinePlot(de5, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(de5, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
 
@@
 nDE <- 500
-resfile <- paste0(resdir, "/de10.rds")
+resfile <- file.path(resdir, "de10.rds")
 
@@
 
-covariateHeatmap(de10, alpha=0.05, covname="log2FC", trans="log1p")
-covariateHeatmap(de10, alpha=0.05, covname="ind_covariate")
-
-covariateHeatmap(de10, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p", linePlot = TRUE)
-covariateHeatmap(de10, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 linePlot = TRUE)
+covariateLinePlot(de10, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(de10, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
 

### 2018-03-27 datasets/scrnaseq/scrnaseq-human-mast.Rmd
Update headers, simplify p-val variable naming

@@
 strat_hist(data.frame(rowData(humandata)), 
-           pvalue="mast_p", covariate="meanExp", maxy=12,
+           pvalue="mast_p", covariate="meanExp", maxy=16,
            main = "MAST, Covariate 1: Mean Nonzero Expression")
@@
 strat_hist(data.frame(rowData(humandata)), 
-           pvalue="mast_p", covariate="detection", maxy=12,
+           pvalue="mast_p", covariate="detection", maxy=16,
            main = "MAST, Covariate 2: Detection Rate") 
@@
 
-### Covariate one: mean nonzero expression
+### Covariate one: Mean nonzero expression
 
@@
 mcols(sb1)$ind_covariate <- mcols(sb1)$meanExp
-covariateLinePlot(sb1, alpha=0.05, covname="ind_covariate", nbins=25) +
+covariateLinePlot(sb1, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans = "log1p") +
   ggtitle("MAST, Covariate 1: Mean Nonzero Expression")

### 2018-03-27 datasets/scrnaseq/scrnaseq-human-scdd.Rmd
Update headers, simplify p-val variable naming

@@
              categorize = FALSE))
-    rowData(dat)$scdd_p_nonzero <- res$nonzero.pvalue
-    rowData(dat)$scdd_p_zero <- res$zero.pvalue
-    rowData(dat)$scdd_p_combined <- res$combined.pvalue
+    rowData(dat)$scdd_p <- res$combined.pvalue
   }else{
@@
 rank_scatter(data.frame(rowData(humandata)), 
-             pvalue="scdd_p_combined", covariate="meanExp") + 
+             pvalue="scdd_p", covariate="meanExp") + 
   ggtitle("scDD, Covariate 1: Mean Nonzero Expression")
@@
 strat_hist(data.frame(rowData(humandata)), 
-           pvalue="scdd_p_combined", covariate="meanExp", maxy=12,
+           pvalue="scdd_p", covariate="meanExp", maxy=9,
            main = "scDD, Covariate 1: Mean Nonzero Expression") 
@@
 rank_scatter(data.frame(rowData(humandata)), 
-             pvalue="scdd_p_combined", covariate="detection") + 
+             pvalue="scdd_p", covariate="detection") + 
   ggtitle("scDD, Covariate 2: Detection Rate")
@@
 strat_hist(data.frame(rowData(humandata)), 
-           pvalue="scdd_p_combined", covariate="detection", maxy=12,
+           pvalue="scdd_p", covariate="detection", maxy=9,
            main = "scDD, Covariate 2: Detection Rate") 
@@
 
-### Covariate one: Mean expression
+### Covariate one: Mean nonzero expression
 
@@
   sb1 <- bd %>% buildBench(data.frame(rowData(humandata)) %>% 
-                                   na.omit() %>%
-                                   mutate(pval=scdd_p_combined,
+                                   filter(!is.na(scdd_p)) %>%
+                                   mutate(pval=scdd_p,
                                           ind_covariate=meanExp), 
@@
   sb2 <- bd %>% buildBench(data.frame(rowData(humandata)) %>% 
-                                   na.omit() %>%
-                                   mutate(pval=scdd_p_combined,
+                                   filter(!is.na(scdd_p)) %>%
+                                   mutate(pval=scdd_p,
                                           ind_covariate=detection), 
@@
 mcols(sb1)$ind_covariate <- mcols(sb1)$meanExp
-covariateLinePlot(sb1, alpha=0.05, covname="ind_covariate", nbins=25) +
+covariateLinePlot(sb1, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans = "log1p") +
   ggtitle("scDD, Covariate 1: Mean Nonzero Expression")

### 2018-03-27 datasets/scrnaseq/scrnaseq-mouse-scdd.Rmd
Update headers, simplify p-val variable naming

@@
 rank_scatter(data.frame(rowData(mousedata)), 
-             pvalue="scdd_p_combined", covariate="meanExp") + 
+             pvalue="scdd_p", covariate="meanExp") + 
   ggtitle("scDD, Covariate 1: Mean Nonzero Expression")
@@
 strat_hist(data.frame(rowData(mousedata)), 
-           pvalue="scdd_p_combined", covariate="meanExp", maxy=20,
+           pvalue="scdd_p", covariate="meanExp", maxy=20,
            main = "scDD, Covariate 1: Mean Nonzero Expression") 
@@
 rank_scatter(data.frame(rowData(mousedata)), 
-             pvalue="scdd_p_combined", covariate="detection") + 
+             pvalue="scdd_p", covariate="detection") + 
   ggtitle("scDD, Covariate 2: Detection Rate")
@@
 strat_hist(data.frame(rowData(mousedata)), 
-           pvalue="scdd_p_combined", covariate="detection", maxy=20,
+           pvalue="scdd_p", covariate="detection", maxy=20,
            main = "scDD, Covariate 2: Detection Rate") 
@@
   sb1 <- bd %>% buildBench(data.frame(rowData(mousedata)) %>% 
-                                   na.omit() %>%
-                                   mutate(pval=scdd_p_combined,
+                                   filter(!is.na(scdd_p)) %>%
+                                   mutate(pval=scdd_p,
                                           ind_covariate=meanExp), 
@@
   sb2 <- bd %>% buildBench(data.frame(rowData(mousedata)) %>% 
-                                   na.omit() %>%
-                                   mutate(pval=scdd_p_combined,
+                                   filter(!is.na(scdd_p)) %>%
+                                   mutate(pval=scdd_p,
                                           ind_covariate=detection), 

### 2018-03-27 datasets/scrnaseq/scrnaseq-knit.sh

@@ -4,7 +4,7 @@
 #SBATCH -N 1
 #SBATCH -p shared
 #SBATCH --mem 125G
-#SBATCH -t 0-6:00
+#SBATCH -t 0-4:00
 #SBATCH -o render-%j.out
 #SBATCH -e render-%j.err
  

### 2018-03-27 datasets/GWAS/GWAS.Rmd
Add plots for ash assumptions and covariate comparisons

--- 2018-03-27 datasets/GWAS/GWAS.Rmd
@@
 
+## Check Assumptions for ash
+
+We'll create a plot to examine the distribution of effect sizes, since
+the ash method assumes that the distribution of true (unobserved) effect
+sizes is unimodal.
+
+```{r}
+ggplot(data=bmi, aes(effect_size)) +
+  geom_histogram(bins=30)
+```
+
+We'll also explore how the standard error (used by ash) 
+correlates with the independent covariates (used by methods that incorporate 
+covariates), in order to get an idea of how these pieces of information relate
+to one another.
+
+```{r}
+ggplot(data=bmi, aes(x = ind_covar_N, y = SE)) +
+  geom_hex(bins=125) +
+  xlab("Covariate: Sample Size")
+
+ggplot(data=bmi, aes(x = ind_covar_AF, y = SE)) +
+  geom_hex(bins = 125)+
+  xlab("Covariate: Minor allele frequency") 
+```
+
 ## Covariate Diagnostics
@@
 ```{r, fig.width=8, fig.height=3.5}
-mcols(sbN)$ind_covariate <- mcols(sbN)$ind_covar_N
-mcols(sbN)$tstat <- bmi %>% filter(!is.na(ind_covar_N)) %>% pull(effect_size)
-covariateHeatmap(sbN, alpha=0.05, covname="tstat", nbins=25, 
-                 linePlot = TRUE, trans="log1p")
-covariateHeatmap(sbN, alpha=0.05, covname="ind_covar_N", nbins=25, 
-                 linePlot = TRUE, trans="log1p")
+mcols(sbN)$effect_size <- bmi %>% filter(!is.na(ind_covar_N)) %>% pull(effect_size)
+covariateLinePlot(sbN, alpha=0.05, covname="effect_size", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(sbN, alpha=0.05, covname="ind_covar_N", nbins=25, 
+                 trans="log1p")
 ```
@@
 ```{r, fig.width=8, fig.height=3.5}
-mcols(sbAF)$ind_covariate <- abs(mcols(sbAF)$ind_covar_AF-0.5)
-mcols(sbAF)$tstat <- bmi %>% filter(!is.na(ind_covar_AF)) %>% pull(effect_size)
-covariateHeatmap(sbAF, alpha=0.05, covname="tstat", nbins=25, 
-                 linePlot = TRUE, trans = "log1p")
-covariateHeatmap(sbAF, alpha=0.05, covname="ind_covar_AF", nbins=25, 
-                 linePlot = TRUE)
+mcols(sbAF)$effect_size <- bmi %>% filter(!is.na(ind_covar_AF)) %>% pull(effect_size)
+covariateLinePlot(sbAF, alpha=0.05, covname="tstat", nbins=25, 
+                 trans = "log1p")
+covariateLinePlot(sbAF, alpha=0.05, covname="ind_covar_AF", nbins=25)
 ```

### 2018-03-27 datasets/RNA-Seq/yeast-simulation.Rmd
Add plots for ash assumptions and covariate comparisons

@@
 library(ggthemes)
+library(R.utils)
 
@@
 
-Check unimodal assumption (for ash).
+We'll create a plot to examine the distribution of effect sizes, since
+the ash method assumes that the distribution of true (unobserved) effect
+sizes is unimodal.
 
 ```{r}
-ggplot(data=geneExp, aes(test_statistic)) +
+ggplot(data=geneExp, aes(effect_size)) +
   geom_histogram(bins=30)
+```
+
+We'll also explore how the standard error (used by ash) 
+correlates with the independent covariate (used by methods that incorporate 
+covariates), in order to get an idea of how these pieces of information relate
+to one another.
+
+```{r}
+ggplot(data=geneExp, aes(x = ind_covariate, y = SE)) +
+  geom_hex(bins = 100) +
+  scale_x_continuous(trans="log10") +
+  xlab("Covariate: Mean gene expression") 
 ```

### 2018-03-27 datasets/R/effectSizePlot.R
Declare namespace of rename

@@ -118,8 +118,8 @@ covariateLinePlot <- function(sbl, alpha=0.05, nbins = 50,
       select(truth, covname, bonf, bh,
              qvalue, contains("ihw"), ashs, "bl-df03", lfdr, "scott-theoretical", 
              "scott-empirical") %>%
-      rename("scott-t" = "scott-theoretical") %>%
-      rename("scott-e" = "scott-empirical") %>%
+      dplyr::rename("scott-t" = "scott-theoretical") %>%
+      dplyr::rename("scott-e" = "scott-empirical") %>%
       mutate(bin = ntile(abs(get(covname)), nbins)) %>%
       gather(method, significant, -covname, -bin) %>%
       group_by(method, bin) %>%

### 2018-03-27 datasets/GWAS/GWAS.Rmd
Fix variable typo in plotting call; remove SLURM command in submit script

@@
 mcols(sbAF)$effect_size <- bmi %>% filter(!is.na(ind_covar_AF)) %>% pull(effect_size)
-covariateLinePlot(sbAF, alpha=0.05, covname="tstat", nbins=25, 
+covariateLinePlot(sbAF, alpha=0.05, covname="effect_size", nbins=25, 
                  trans = "log1p")

### 2018-03-27 datasets/GWAS/GWAS-knit.sh

@@ -9,7 +9,6 @@
 #SBATCH -e render-%j.err  # Standard error
 
 module load plink
-export SLURM_NTASKS
 export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/rstudio/0.98.1103-fasrc01/bin/pandoc/"
 
 # change filename to Rmd to be knitted. 

### 2018-03-28 datasets/microbiome/microbiome-enigma.Rmd
Add ENIGMA data and analysis

---
title: "Case Study: 16S microbiome data (ENIGMA)"
author: "Claire Duvallet"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

Here we analyze a subset of the ENIGMA data ([Smith, M. and Rocha, A. et al 2015](http://dx.doi.org/10.1128/mBio.00326-15)).
The dataset includes water samples from 67 wells with well-characterized biochemical metadata (i.e. pH, heavy metal concentrations, etc).
We acquired the data from a collaborator and have put it in the `benchmark-fdr/datasets/microbiome/DATA/enigma` folder.

Note that I've uploaded this to our repo as a gzipped file, make sure to gunzip the OTU table before knitting this.

We did some preliminary correlation analysis and found that the metadata variables show a range of correlation with different OTUs: some metadata variables don't seem correlated with any OTUs, whereas others seem correlated with basically the entire community.
We'll pick three of these metadata variables to illustrate our FDR correction methods in cases where we expect different numbers of true significant associations.

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(magrittr)
library(SummarizedBenchmark)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "DATA"
resdir <- "RESULTS"
enigma_dir <- "enigma"

# From their paper, we expect SO4 to have little correlation, pH
# to have high correlations, and Al to be in the middle
so4_result_file <- file.path(resdir, "enigma-so4-otus-results.rds")
so4_bench_file <- file.path(resdir, "enigma-so4-otus-benchmark.rds")

ph_result_file <- file.path(resdir, "enigma-ph-otus-results.rds")
ph_bench_file <- file.path(resdir, "enigma-ph-otus-benchmark.rds")

al_result_file <- file.path(resdir, "enigma-al-otus-results.rds")
al_bench_file <- file.path(resdir, "enigma-al-otus-benchmark.rds")

```

# Data Preparation

```{r, enigma-readdata}
## load OTU table and metadata
# Note to others: not sure why read.table breaks on this file.
otu <- read.table(file.path(datdir, enigma_dir, 
                            "otu_table_resampled_updated_r.txt"), 
                  sep='\t', header = TRUE, row.names = 1)
meta <- read.csv(file.path(datdir, enigma_dir, "metadata.txt"), sep='\t')

# Select only the 0.02 um samples
meta <- meta[endsWith(as.character(meta$Sample), "02"), ]
# Make metadata sample ID match OTU table
meta$Sample <- gsub("_", "\\.", as.character(meta$Sample))
meta$Sample <- substr(meta$Sample, 1, nchar(meta$Sample)-3)

## keep only samples with both metadata and 16S data
keep_samples <- intersect(colnames(otu), meta$Sample)
otu <- otu[, keep_samples]
meta <- dplyr::filter(meta, Sample %in% keep_samples)

## Make sure samples in metadata and OTU table match
meta <- meta[match(colnames(otu), meta$Sample), ]
```

We'll do a little filtering here before calculating correlations.
We'll apply a minimum threshold of 10 reads per OTU across all samples.
After removing these shallow OTUs, we also get rid of any samples with too few reads. 
We define the minimum number of reads per OTU in `min_otu`, and
the minimum number of reads per sample in `min_sample`.

After we've removed any shallow OTUs and samples, we'll convert the OTU table to
relative abundances.

```{r, enigma-cleandata}
min_otu <- 10
minp_otu <- 0.01
min_sample <- 100

## Remove OTUs w/ <= min reads, w/ <= min prop, samples w/ <= min reads
otu <- otu[rowSums(otu) > min_otu, ]
otu <- otu[rowSums(otu > 0) / ncol(otu) > minp_otu, ]
otu <- otu[, colSums(otu) > min_sample]

## Update metadata with new samples
meta <- dplyr::filter(meta, Sample %in% colnames(otu))

## Remove empty OTUs
otu <- otu[rowSums(otu) > 0, ]

## Convert to relative abundance
abun_otu <- t(t(otu) / rowSums(t(otu)))

## Add pseudo counts
zeroabun <- 0
abun_otu <- abun_otu + zeroabun
```

# Data Analysis

## pH

Next, we need to calculate the pvalues for correlations between abundances of each OTU and the metadata variables of interest. (i.e. we're correlating each OTU's abundances across wells with the metadata variable across wells).
Let's start with "pH".
We'll store the pvalue results in a dataframe.

While we're at it, we'll also calculate the mean abundance and ubiquity (detection rate)
of each OTU. Later, we can assign their values to a new column called `ind_covariate` 
for use in downstream steps.

```{r, ph-pvals}
if (!file.exists(ph_result_file)) {
    res <- test_microbiome_corr(abundance = abun_otu, meta_col = meta$pH,
                                shift = zeroabun)
    saveRDS(res, file = ph_result_file)
} else {
    res <- readRDS(ph_result_file)
}
```

## Covariate Diagnostics

Here we look to see if the covariates do indeed look informative.

### Ubiquity

```{r, ph-ubi-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
```

```{r, ph-ubi-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

### Mean Abundance (across non-zero samples)

```{r, ph-abun-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
```

```{r, ph-abun-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

## Multiple-Testing Correction

Let's use `ubiquity` as our `ind_covariate`.

```{r, enigma-labelcovariate}
res <- dplyr::mutate(res, ind_covariate = ubiquity)
```

First, we'll create an object of `BenchDesign` class to hold the data and 
add the benchmark methods to the `BenchDesign` object. We'll do this for a few
values of nmids, to explore its effects on the outputs. We remove ASH from
the default comparison.

Then, we'll construct the `SummarizedBenchmark` object, which will run
the functions specified in each method (these are actually sourced in from the
helper scripts).

```{r, ph-sb, results="hide", message=FALSE, eval=TRUE}
if (!file.exists(ph_bench_file)) {
    nmids <- 150
    bd <- initializeBenchDesign(nmids = nmids)
    bd <- dropBMethod(bd, "ashs")
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    metadata(sb)$nmids <- nmids
    saveRDS(sb, file = ph_bench_file)
} else {
    sb <- readRDS(ph_bench_file)
}
```

## Benchmark Metrics

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, ph-metrics, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, ph-plot, width=15, height=15, eval=TRUE}
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
```

## SO4

Next, we'll look at SO4 levels

```{r, so4-pvals}
if (!file.exists(so4_result_file)) {
    res <- test_microbiome_corr(abundance = abun_otu, meta_col = meta$SO4_mgL,
                                shift = zeroabun)
    saveRDS(res, file = so4_result_file)
} else {
    res <- readRDS(so4_result_file)
}
```

## Covariate Diagnostics

Here we look to see if the covariates do indeed look informative.

### Ubiquity

```{r, so4-ubi-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
```

```{r, so4-ubi-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

### Mean Abundance (across non-zero samples)

```{r, so4-abun-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
```

```{r, so4-abun-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

## Multiple-Testing Correction

Let's use `ubiquity` as our `ind_covariate`.

```{r, so4-labelcovariate}
res <- dplyr::mutate(res, ind_covariate = ubiquity)
```

First, we'll create an object of `BenchDesign` class to hold the data and 
add the benchmark methods to the `BenchDesign` object. We'll do this for a few
values of nmids, to explore its effects on the outputs. We remove ASH from
the default comparison.

Then, we'll construct the `SummarizedBenchmark` object, which will run
the functions specified in each method (these are actually sourced in from the
helper scripts).

```{r, so4-sb, results="hide", message=FALSE, eval=TRUE}
if (!file.exists(so4_bench_file)) {
    nmids <- 150
    bd <- initializeBenchDesign(nmids = nmids)
    bd <- dropBMethod(bd, "ashs")
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    metadata(sb)$nmids <- nmids
    saveRDS(sb, file = so4_bench_file)
} else {
    sb <- readRDS(so4_bench_file)
}
```

## Benchmark Metrics

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, so4-metrics, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, so4-plot, width=15, height=15, eval=TRUE}
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
```

## Al

Finally, we'll look at Al levels

```{r, al-pvals}
if (!file.exists(al_result_file)) {
    res <- test_microbiome_corr(abundance = abun_otu, meta_col = meta$Al_mgL,
                                shift = zeroabun)
    saveRDS(res, file = al_result_file)
} else {
    res <- readRDS(al_result_file)
}
```

## Covariate Diagnostics

Here we look to see if the covariates do indeed look informative.

### Ubiquity

```{r, al-ubi-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
```

```{r, al-ubi-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

### Mean Abundance (across non-zero samples)

```{r, al-abun-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
```

```{r, al-abun-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

## Multiple-Testing Correction

Let's use `ubiquity` as our `ind_covariate`.

```{r, al-labelcovariate}
res <- dplyr::mutate(res, ind_covariate = ubiquity)
```

First, we'll create an object of `BenchDesign` class to hold the data and 
add the benchmark methods to the `BenchDesign` object. We'll do this for a few
values of nmids, to explore its effects on the outputs. We remove ASH from
the default comparison.

Then, we'll construct the `SummarizedBenchmark` object, which will run
the functions specified in each method (these are actually sourced in from the
helper scripts).

```{r, al-sb, results="hide", message=FALSE, eval=TRUE}
if (!file.exists(al_bench_file)) {
    nmids <- 150
    bd <- initializeBenchDesign(nmids = nmids)
    bd <- dropBMethod(bd, "ashs")
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    metadata(sb)$nmids <- nmids
    saveRDS(sb, file = al_bench_file)
} else {
    sb <- readRDS(al_bench_file)
}
```

## Benchmark Metrics

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, al-metrics, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, al-plot, width=15, height=15, eval=TRUE}
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
```

# Session Info

```{r}
sessionInfo()
```

### 2018-03-28 datasets/R/helpers-microbiome.R

@@ -60,6 +60,44 @@ test_microbiome <- function(abundance, shift, is_case) {
                          test_statistic = test_statistic * sign(effect_size))
 }
 
+test_microbiome_corr <- function(abundance, meta_col, shift) {
+  
+  ## Calculate correlation and pvalues
+  res <- data.frame(otu = rownames(abundance), 
+                    pval = NA, spearman_teststat = NA,
+                    SE = NA, effect_size = NA, 
+                    mean_abun = NA, mean_abun_present = NA,
+                    ubiquity = NA)
+  
+  for (i in seq_len(nrow(abundance))) {
+    ## spearman p value
+    corr <- cor.test(abundance[i, ], meta_col, 
+                     method='spearman', exact = FALSE, alternative = "two.sided")
+    res$pval[i] <- corr$p.value
+    res$spearman_teststat[i] <- corr$statistic
+    
+    ## standard error of the OTU abundance, across all samples
+    res$SE[i] <- sd(abundance[i, ]) / sqrt(length(abundance[i, ]))
+    
+    ## mean OTU abundance across all samples (after removing pseudo-count)
+    res$mean_abun[i] <- mean(abundance[i, ]) - shift
+    
+    ## mean OTU abundance across only samples with the OTU present
+    res$mean_abun_present[i] <-
+      sum(abundance[i, ] - shift) / sum(abundance[i, ] > shift)
+    
+    ## ubiquity of OTU across all samples
+    res$ubiquity[i] <- sum(abundance[i, ] > shift) / length(abundance[i, ])
+    
+  }
+  
+  ## compute test-statistic from p-value for ASH
+  res <- dplyr::mutate(res,
+                       test_statistic = qnorm(exp(log(pval) - log(2)),
+                                              lower.tail=FALSE),
+                       test_statistic = test_statistic * sign(effect_size))
+}
+  
 
 remove_shallow_smpls <- function(df, n_reads) {
     ## df has OTUs in rows and samples in columns

### 2018-03-28 datasets/microbiome/microbiome-baxter.Rmd
Add log(ubiquity) as covariate

@@
 otu_bench_file <- file.path(resdir, "baxter-otus-benchmark.rds")
+otu_bench_file_log_ubiquity <- file.path(resdir, "baxter-otus-log-ubiquity-benchmark.rds")
+
 gns_result_file <- file.path(resdir, "baxter-genus-results.rds")
 gns_bench_file <- file.path(resdir, "baxter-genus-benchmark.rds")
+gns_bench_file_log_ubiquity <- file.path(resdir, "baxter-genus-log-ubiquity-benchmark.rds")
 ```
@@
 
+## Multiple-Testing Correction - log(ubiquity)
+
+Let's use `log(ubiquity)` as our `ind_covariate`.
+
+```{r, baxter-logubi-labelcovariate}
+res <- dplyr::mutate(res, ind_covariate = log10(ubiquity))
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We'll do this for a few
+values of nmids, to explore its effects on the outputs. We remove ASH from
+the default comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, baxter-logubi-sb, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(otu_bench_file_log_ubiquity)) {
+    nmids <- 150
+    bd <- initializeBenchDesign(nmids = nmids)
+    bd <- dropBMethod(bd, "ashs")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    metadata(sb)$data_download_link <-
+                   "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
+    metadata(sb)$nmids <- nmids
+    saveRDS(sb, file = otu_bench_file_log_ubiquity)
+} else {
+    sb <- readRDS(otu_bench_file_log_ubiquity)
+}
+```
+
+## Benchmark Metrics - log(ubiquity)
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, baxter-logubi-metrics, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, baxter-logubi-plot, width=15, height=15, eval=TRUE}
+rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+
+```{r, eval=FALSE}
+assays(sb)[["qvalue"]][, "ashs"] %>% max()
+sum(is.na(assays(sb)[["qvalue"]][, "scott-empirical"]))
+sum(is.na(assays(sb)[["qvalue"]][, "lfdr"]))
+```
+
 # Data Analysis (genus-level)
@@
 
+## Multiple-Testing Correction - log(ubiquity)
+
+Let's use log(`ubiquity`) as our `ind_covariate`.
+
+```{r, baxter-genus-logubi-labelcovariate}
+res <- dplyr::mutate(res, ind_covariate = log10(ubiquity))
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We'll do this for a few
+values of nmids, to explore its effects on the outputs. We remove ASH from
+the comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, baxter-genus-logubi-sb, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(gns_bench_file_log_ubiquity)) {
+    nmids <- 150
+    bd <- initializeBenchDesign(nmids = nmids)
+    bd <- dropBMethod(bd, "ashs")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    metadata(sb)$data_download_link <-
+                   "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
+    metadata(sb)$nmids <- nmids
+    saveRDS(sb, file = gns_bench_file_log_ubiquity)
+} else {
+    sb <- readRDS(gns_bench_file_log_ubiquity)
+}
+```
+
+## Benchmark Metrics - log(ubiquity)
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, baxter-genus-logubi-metrics, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, baxter-genus-logubi-plot, width=15, height=15, eval=TRUE}
+rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+
+Note: Benjamini-Hochberg (bh) overlaps exactly with the IHW results.
+
+```{r, baxter-genus-logubi-overlap}
+plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+```
+```{r}
+methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
+plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods)
+```
+
 # Session Info

### 2018-03-28 datasets/microbiome/microbiome-papa.Rmd
Add log(ubiquity) as covariate

@@
 gns_bench_file <- file.path(resdir, "papa-genus-benchmark.rds")
+gns_bench_file_log_ubi <- file.path(resdir, "papa-genus-log-ubiquity-benchmark.rds")
 ```
@@
 
+## Multiple-Testing Correction - log(ubiquity)
+
+Let's use log10(`ubiquity`) as our `ind_covariate`.
+
+```{r, papa-genus-logubi-labelcovariate}
+res <- dplyr::mutate(res, ind_covariate = log10(ubiquity))
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We'll do this for a few
+values of nmids, to explore its effects on the outputs. We remove ASH from
+the comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, papa-genus-logubi-sb, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(gns_bench_file_log_ubi)) {
+    nmids <- 150
+    bd <- initializeBenchDesign(nmids = nmids)
+    bd <- dropBMethod(bd, "ashs")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    metadata(sb)$data_download_link <-
+                   "https://zenodo.org/record/840333/files/ibd_alm_results.tar.gz"
+    metadata(sb)$nmids <- nmids
+    saveRDS(sb, file = gns_bench_file_log_ubi)
+} else {
+    sb <- readRDS(gns_bench_file_log_ubi)
+}
+```
+
+## Benchmark Metrics - log(ubiquity)
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, papa-genus-logubi-metrics, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, papa-genus-logbi-plot, width=15, height=15, eval=TRUE}
+rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+
+```{r, papa-genus-logubi-overlap}
+plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+```
+```{r}
+methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
+plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods)
+```
+
 # Session Info

### 2018-03-29 datasets/simulations/simulations-null.Rmd
Make some progress on null sims rmd

---
title: "Simulation Study: Null Settings"
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

In this set of simulations, we only consider settings where all tests are null.

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(SummarizedBenchmark)
library(parallel)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
resdir <- "results"
dir.create(resdir, showWarnings = FALSE, recursive = TRUE)

## intermediary files we create below
gauss_file <- file.path(resdir, "null-benchmark-gaussian.rds")
tdist_file <- file.path(resdir, "null-benchmark-t6.rds")
chisq_file <- file.path(resdir, "null-benchmark-chisq3.rds")

## number of cores for parallelization
cores <- 2
B <- 5

## define bechmarking design
bd <- initializeBenchDesign()
```

All simulation settings will share the following parameters.

```{r parameters-shared}
m <- 2000                         # integer: number of hypothesis tests
pi0 <- 1                          # numeric: proportion of null hypotheses
tstat <- function(x) { 0 }        # functional: dist of alternative test stats
icovariate <- runif               # functional: independent covariate
```

# Gaussian Setting

First, we consider the null setting with gaussian noise.

## Data Simulation

```{r parameters-gaussian}
tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
```

We next run the simulations.

```{r run-simulation}
if (file.exists(gauss_file)) {
    res <- readRDS(gauss_file)
} else {
    res <- mclapply(X=1:B, FUN = simIteration, bench = bd, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    mc.cores = cores)
    saveRDS(res, file = gauss_file)
}
```

## Covariate Diagnostics

## Benchmark Metrics

# Student's t Setting
## Data Simulation
## Covariate Diagnostics
## Benchmark Metrics

# Chi-Squared Setting
## Data Simulation
## Covariate Diagnostics
## Benchmark Metrics

# Session Info

```{r}
sessionInfo()
```

### 2018-03-30 datasets/simulations/simulations-null.Rmd
Add full null sim Rmd

@@
 ## number of cores for parallelization
-cores <- 2
-B <- 5
+cores <- 20
+B <- 100
 
@@
 ```{r parameters-shared}
-m <- 2000                         # integer: number of hypothesis tests
+m <- 20000                        # integer: number of hypothesis tests
 pi0 <- 1                          # numeric: proportion of null hypotheses
@@
 icovariate <- runif               # functional: independent covariate
+```
+
+Simulation results will be presented excluding a subset of methods, and
+for certain plots (upset plots), a single alpha cutoff will be used.
+
+```{r}
+excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
+ualpha <- 0.05
 ```
@@
 
-```{r parameters-gaussian}
+```{r gauss-parameters}
 tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
@@
 
-```{r run-simulation}
+```{r gauss-run-simulation}
 if (file.exists(gauss_file)) {
@@
 
+Here, we show the relationship between the independent covariate and p-values for a
+single replication of the experiment.
+
+```{r gauss-one-simulation}
+onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
+```
+
+```{r, gauss-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
+rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
+```
+
+```{r, gauss-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+```
+
 ## Benchmark Metrics
 
+We plot the average results across `r B` simulations.
+
+```{r gauss-metrics-averages, results = "hide"}
+resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(resdf, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE) 
+```
+
+We also take a look at the distribution of rejects for each method as a function of
+the effect size and independent covariate.
+
+```{r gauss-metrics-covlineplot, results = "hide"}
+covariateLinePlot(res, alpha = 0.05, covname = "effect_size", trans = "log1p")
+
+covariateLinePlot(res, alpha = 0.05, covname = "ind_covariate", trans = "log1p")
+```
+
+Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
+the overlap of rejections between methods.
+
+```{r gauss-metrics-upset, results = "hide"}
+if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
+    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+    message("Not enough methods found rejections at alpha ", ualpha, 
+            "; skipping upset plot")
+}
+```
+
 # Student's t Setting
+
+Next, we consider the null setting with t-distributed noise.
+
 ## Data Simulation
+
+```{r chisq4-parameters}
+tstat_dist <- rchisq_perturber(4)  # functional: sampling dist/noise for test stats
+null_dist <- rchisq_2pvaluer(4)    # functional: dist to calc p-values
+```
+
+We next run the simulations.
+
+```{r chisq4-run-simulation}
+if (file.exists(tdist_file)) {
+    res <- readRDS(tdist_file)
+} else {
+    res <- mclapply(X=1:B, FUN = simIteration, bench = bd, m = m,
+                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
+                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    mc.cores = cores)
+    saveRDS(res, file = tdist_file)
+}
+```
+
 ## Covariate Diagnostics
+
+Here, we show the relationship between the independent covariate and p-values for a
+single replication of the experiment.
+
+```{r chisq4-one-simulation}
+onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
+```
+
+```{r, chisq4-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
+rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
+```
+
+```{r, chisq4-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+```
+
 ## Benchmark Metrics
 
+We plot the average results across `r B` simulations.
+
+```{r chisq4-metrics-averages, results = "hide"}
+resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(resdf, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE) 
+```
+
+We also take a look at the distribution of rejects for each method as a function of
+the effect size and independent covariate.
+
+```{r chisq4-metrics-covlineplot, results = "hide"}
+covariateLinePlot(res, alpha = 0.05, covname = "effect_size", trans = "log1p")
+
+covariateLinePlot(res, alpha = 0.05, covname = "ind_covariate", trans = "log1p")
+```
+
+Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
+the overlap of rejections between methods.
+
+```{r chisq4-metrics-upset, results = "hide"}
+if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
+    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+    message("Not enough methods found rejections at alpha ", ualpha, 
+            "; skipping upset plot")
+}
+```
+
 # Chi-Squared Setting
+
+Finally, we consider the null setting with chi-squared distributed noise.
+
 ## Data Simulation
+
+```{r chisq4-parameters}
+tstat_dist <- rt_perturber(5)  # functional: sampling dist/noise for test stats
+null_dist <- rt_2pvaluer(5)    # functional: dist to calc p-values
+```
+
+We next run the simulations.
+
+```{r chisq4-run-simulation}
+if (file.exists(chisq_file)) {
+    res <- readRDS(chisq_file)
+} else {
+    res <- mclapply(X=1:B, FUN = simIteration, bench = bd, m = m,
+                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
+                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    mc.cores = cores)
+    saveRDS(res, file = chisq_file)
+}
+```
+
 ## Covariate Diagnostics
+
+Here, we show the relationship between the independent covariate and p-values for a
+single replication of the experiment.
+
+```{r chisq4-one-simulation}
+onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
+```
+
+```{r, chisq4-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
+rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
+```
+
+```{r, chisq4-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+```
+
 ## Benchmark Metrics
+
+We plot the average results across `r B` simulations.
+
+```{r chisq4-metrics-averages, results = "hide"}
+resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(resdf, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE) 
+```
+
+We also take a look at the distribution of rejects for each method as a function of
+the effect size and independent covariate.
+
+```{r chisq4-metrics-covlineplot, results = "hide"}
+covariateLinePlot(res, alpha = 0.05, covname = "effect_size", trans = "log1p")
+
+covariateLinePlot(res, alpha = 0.05, covname = "ind_covariate", trans = "log1p")
+```
+
+Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
+the overlap of rejections between methods.
+
+```{r chisq4-metrics-upset, results = "hide"}
+if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
+    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+    message("Not enough methods found rejections at alpha ", ualpha, 
+            "; skipping upset plot")
+}
+```
 

### 2018-03-30 datasets/simulations/simulations-null.Rmd
Add sim iter param and seeds for each sim

@@
 null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
+seed <- 608
 ```
@@
 } else {
-    res <- mclapply(X=1:B, FUN = simIteration, bench = bd, m = m,
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bd, m = m,
                     pi0 = pi0, tstat = tstat, icovariate = icovariate,
                     tstat_dist = tstat_dist, null_dist = null_dist,
-                    mc.cores = cores)
+                    seed = seed, mc.cores = cores)
     saveRDS(res, file = gauss_file)
@@
 null_dist <- rchisq_2pvaluer(4)    # functional: dist to calc p-values
+seed <- 1023
 ```
@@
 } else {
-    res <- mclapply(X=1:B, FUN = simIteration, bench = bd, m = m,
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bd, m = m,
                     pi0 = pi0, tstat = tstat, icovariate = icovariate,
                     tstat_dist = tstat_dist, null_dist = null_dist,
-                    mc.cores = cores)
+                    seed = seed, mc.cores = cores)
     saveRDS(res, file = tdist_file)
@@
 null_dist <- rt_2pvaluer(5)    # functional: dist to calc p-values
+seed <- 815
 ```
@@
 } else {
-    res <- mclapply(X=1:B, FUN = simIteration, bench = bd, m = m,
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bd, m = m,
                     pi0 = pi0, tstat = tstat, icovariate = icovariate,
                     tstat_dist = tstat_dist, null_dist = null_dist,
-                    mc.cores = cores)
+                    seed = seed, mc.cores = cores)
     saveRDS(res, file = chisq_file)

### 2018-03-30 datasets/R/helpers-simulations.R

@@ -8,6 +8,7 @@
 #' (For IHW, in particular, the simulations illustrating the "size investing strategy"
 #' fall under this class of simulations.)
 #'
+#' @param X simulation index
 #' @param bench BenchDesign object of methods to run on simulated data set.
 #' @param m number of test statistics.
 #' @param pi0 proportion of null hypotheses. Can either be a single numeric value
@@ -32,7 +33,8 @@
 #'
 #' @return
 #' SummarizedBenchmark generated by calling `buildBench` with the specified
-#' BenchDesign object and a simulated data set. 
+#' BenchDesign object and a simulated data set. If `execute = FALSE`, the simulated
+#' data set is returned.
 #' 
 #' @details
 #' If a function is specified for either of `effect_size` or `icovariate`, a function must also be
@@ -48,9 +50,9 @@
 #' 
 #' @md
 #' @author Patrick Kimes
-simIteration <- function(bench, m, pi0, tstat, tstat_dist, null_dist,
+simIteration <- function(X, bench, m, pi0, tstat, tstat_dist, null_dist,
                          icovariate, execute = TRUE, seed = NULL) {
-    if (!is.null(seed)) { set.seed(seed) }
+    if (!is.null(seed)) { set.seed(seed * X) }
     stopifnot(is.function(tstat))
     stopifnot(is.function(tstat_dist))
     stopifnot(is.function(icovariate))

### 2018-03-30 datasets/simulations/simulations-null.Rmd
Fix mislabeled blocks in null sim rmd

@@
 
-```{r chisq4-parameters}
-tstat_dist <- rchisq_perturber(4)  # functional: sampling dist/noise for test stats
-null_dist <- rchisq_2pvaluer(4)    # functional: dist to calc p-values
-seed <- 1023
+```{r t5-parameters}
+tstat_dist <- rt_perturber(5)  # functional: sampling dist/noise for test stats
+null_dist <- rt_2pvaluer(5)    # functional: dist to calc p-values
+seed <- 815
 ```
@@
 
-```{r chisq4-run-simulation}
+```{r t5-run-simulation}
 if (file.exists(tdist_file)) {
@@
 
-```{r chisq4-one-simulation}
+```{r t5-one-simulation}
 onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
@@
 
-```{r, chisq4-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
+```{r, t5-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
 rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
@@
 
-```{r, chisq4-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
+```{r, t5-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
 strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
@@
 
-```{r chisq4-metrics-averages, results = "hide"}
+```{r t5-metrics-averages, results = "hide"}
 resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
@@
 
-```{r chisq4-metrics-covlineplot, results = "hide"}
+```{r t5-metrics-covlineplot, results = "hide"}
 covariateLinePlot(res, alpha = 0.05, covname = "effect_size", trans = "log1p")
@@
 
-```{r chisq4-metrics-upset, results = "hide"}
+```{r t5-metrics-upset, results = "hide"}
 if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
@@
 ```{r chisq4-parameters}
-tstat_dist <- rt_perturber(5)  # functional: sampling dist/noise for test stats
-null_dist <- rt_2pvaluer(5)    # functional: dist to calc p-values
-seed <- 815
+tstat_dist <- rchisq_perturber(4)  # functional: sampling dist/noise for test stats
+null_dist <- rchisq_2pvaluer(4)    # functional: dist to calc p-values
+seed <- 1023
 ```

### 2018-03-30 datasets/simulations/simulations-null.Rmd
Remove redundant plot in sims
Fix rchisq_2pvaluer to rchisq_pvaluer, 1-sided pval

@@
                 merge_ihw = TRUE, errorBars = TRUE) 
-
-plotsim_average(resdf, met="FDR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE) 
 ```
@@
 tstat_dist <- rchisq_perturber(4)  # functional: sampling dist/noise for test stats
-null_dist <- rchisq_2pvaluer(4)    # functional: dist to calc p-values
+null_dist <- rchisq_pvaluer(4)     # functional: dist to calc p-values
 seed <- 1023

### 2018-03-30 datasets/simulations/simulations-null.Rmd
Remove unnecessary plots from null sims; fix file names

@@
 gauss_file <- file.path(resdir, "null-benchmark-gaussian.rds")
-tdist_file <- file.path(resdir, "null-benchmark-t6.rds")
-chisq_file <- file.path(resdir, "null-benchmark-chisq3.rds")
+tdist_file <- file.path(resdir, "null-benchmark-t5.rds")
+chisq_file <- file.path(resdir, "null-benchmark-chisq4.rds")
 
@@
 plotsim_average(resdf, met="rejections", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
-
-plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                 merge_ihw = TRUE, errorBars = TRUE)
@@
 
-plotsim_average(resdf, met="TNR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
-
 plotsim_average(resdf, met="FDR", filter_set = excludeSet,

### 2018-03-30 datasets/simulations/simulations-noninformative.Rmd
Add noninformative sims

--- 2018-03-30 datasets/simulations/simulations-null.Rmd
@@
 ---
-title: "Simulation Study: Null Settings"
+title: "Simulation Study: Non-Informative Covariate"
 author: "Patrick Kimes"
@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
-
 # Summary
 
-In this set of simulations, we only consider settings where all tests are null.
+In this set of simulations, we consider settings with both null and non-null
+tests with only an non-informative covariate. We include simulation results
+again with Gaussian, t-distributed, and Chi-Squared distributed noise.
 
@@
 ## intermediary files we create below
-gauss_file <- file.path(resdir, "null-benchmark-gaussian.rds")
-tdist_file <- file.path(resdir, "null-benchmark-t5.rds")
-chisq_file <- file.path(resdir, "null-benchmark-chisq4.rds")
+gauss_file <- file.path(resdir, "noninformative-benchmark-gaussian.rds")
+tdist_file <- file.path(resdir, "noninformative-benchmark-t5.rds")
+chisq_file <- file.path(resdir, "noninformative-benchmark-chisq4.rds")
 
@@
 m <- 20000                        # integer: number of hypothesis tests
-pi0 <- 1                          # numeric: proportion of null hypotheses
-tstat <- function(x) { 0 }        # functional: dist of alternative test stats
+pi0 <- 0.90                       # numeric: proportion of null hypotheses
 icovariate <- runif               # functional: independent covariate
@@
 
-First, we consider the null setting with gaussian noise.
+First, we consider the setting with Gaussian noise.
 
@@
 ```{r gauss-parameters}
+tstat <- rnorm_generator(2)       # functional: dist of alternative test stats
 tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
@@
 
-We plot the average results across `r B` simulations.
+We plot the averaged results across `r B` simulations.
 
@@
                 merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE) 
 ```
@@
 ```{r gauss-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = 0.05, covname = "effect_size", trans = "log1p")
+covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
 
-covariateLinePlot(res, alpha = 0.05, covname = "ind_covariate", trans = "log1p")
+covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 
-Next, we consider the null setting with t-distributed noise.
+Next, we consider the setting with t-distributed noise.
 
@@
 ```{r t5-parameters}
+tstat <- rnorm_generator(5)    # functional: dist of alternative test stats
 tstat_dist <- rt_perturber(5)  # functional: sampling dist/noise for test stats
@@
 ```{r t5-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = 0.05, covname = "effect_size", trans = "log1p")
+covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
 
-covariateLinePlot(res, alpha = 0.05, covname = "ind_covariate", trans = "log1p")
+covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 
-Finally, we consider the null setting with chi-squared distributed noise.
+Finally, we consider the setting with chi-squared distributed noise.
 
@@
 ```{r chisq4-parameters}
+tstat <- rnorm_generator(12)       # functional: dist of alternative test stats
 tstat_dist <- rchisq_perturber(4)  # functional: sampling dist/noise for test stats
@@
 ```{r chisq4-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = 0.05, covname = "effect_size", trans = "log1p")
+covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
 
-covariateLinePlot(res, alpha = 0.05, covname = "ind_covariate", trans = "log1p")
+covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 ```
-
-# Session Info
-
-```{r}
-sessionInfo()
-```

### 2018-03-30 datasets/R/helpers-simulations.R
Turn pi0 helpers to function factories

@@ -115,29 +115,31 @@ simIteration <- function(X, bench, m, pi0, tstat, tstat_dist, null_dist,
 ## ##############################################################################
 
 ## step function (4 steps)
-pi0_step <- function(x, pi0) {
+pi0_step <- function(pi0) {
     stopifnot(pi0 < 1, pi0 >= 0.5)
-    pi0 - (1-pi0)/2 +
-        (1-pi0)/4 * (x > 0.25) +
-        (1-pi0)/2 * (x > 0.5) +
-        (1-pi0)/4 * (x > 0.75)
+    function(x) {
+        pi0 - (1-pi0)/2 +
+            (1-pi0)/4 * (x > 0.25) +
+            (1-pi0)/2 * (x > 0.5) +
+            (1-pi0)/4 * (x > 0.75)
+    }
 }
 
 ## shifted/stretched cubic function
-pi0_cubic <- function(x, pi0) {
+pi0_cubic <- function(pi0) {
     stopifnot(pi0 < 1, pi0 >= 2/3)
-    (1-x)^(1/3) * 4*(1-pi0) + 4 * pi0 - 3
+    function(x) { (1-x)^(1/3) * 4*(1-pi0) + 4 * pi0 - 3 }
 }
 
 ## shifted/stretched cosine function (non-monotone)
-pi0_cosine <- function(x, pi0) {
+pi0_cosine <- function(pi0) {
     stopifnot(pi0 < 1, pi0 >= 1/2)
-    (1-pi0) * cos(2*pi*x) + pi0
+    function(x) { (1-pi0) * cos(2*pi*x) + pi0 }
 }
 
-pi0_sine <- function(x, pi0) {
+pi0_sine <- function(pi0) {
     stopifnot(pi0 < 1, pi0 >= 1/2)
-    (1-pi0) * sin(2*pi*x) + pi0
+    function(x) { (1-pi0) * sin(2*pi*x) + pi0 }
 }
 
 

### 2018-03-30 datasets/simulations/simulations-informative-cubic.Rmd
Add two informative covariate sims

--- 2018-03-30 datasets/simulations/simulations-noninformative.Rmd
@@
 ---
-title: "Simulation Study: Non-Informative Covariate"
+title: "Simulation Study: Informative Covariate (cubic)"
 author: "Patrick Kimes"
@@
 In this set of simulations, we consider settings with both null and non-null
-tests with only an non-informative covariate. We include simulation results
+tests with an informative covariate. The covariate is sampled uniformly from
+the interval [0, 1], and the conditional probability of a test being non-null
+is a smooth (cubic) function of the covariate. We include simulation results
 again with Gaussian, t-distributed, and Chi-Squared distributed noise.
@@
 ## intermediary files we create below
-gauss_file <- file.path(resdir, "noninformative-benchmark-gaussian.rds")
-tdist_file <- file.path(resdir, "noninformative-benchmark-t5.rds")
-chisq_file <- file.path(resdir, "noninformative-benchmark-chisq4.rds")
+gauss_file <- file.path(resdir, "informative-cubic-benchmark-gaussian.rds")
+tdist_file <- file.path(resdir, "informative-cubic-benchmark-t5.rds")
+chisq_file <- file.path(resdir, "informative-cubic-benchmark-chisq4.rds")
 
@@
 m <- 20000                        # integer: number of hypothesis tests
-pi0 <- 0.90                       # numeric: proportion of null hypotheses
+pi0 <- pi0_cubic(0.90)            # numeric: proportion of null hypotheses
 icovariate <- runif               # functional: independent covariate

### 2018-03-30 datasets/simulations/simulations-informative-sine.Rmd
Add two informative covariate sims

--- 2018-03-30 datasets/simulations/simulations-informative-cubic.Rmd
@@
 ---
-title: "Simulation Study: Informative Covariate (cubic)"
+title: "Simulation Study: Informative Covariate (sine)"
 author: "Patrick Kimes"
@@
 the interval [0, 1], and the conditional probability of a test being non-null
-is a smooth (cubic) function of the covariate. We include simulation results
+is a smooth (sine) function of the covariate. We include simulation results
 again with Gaussian, t-distributed, and Chi-Squared distributed noise.
@@
 ## intermediary files we create below
-gauss_file <- file.path(resdir, "informative-cubic-benchmark-gaussian.rds")
-tdist_file <- file.path(resdir, "informative-cubic-benchmark-t5.rds")
-chisq_file <- file.path(resdir, "informative-cubic-benchmark-chisq4.rds")
+gauss_file <- file.path(resdir, "informative-sine-benchmark-gaussian.rds")
+tdist_file <- file.path(resdir, "informative-sine-benchmark-t5.rds")
+chisq_file <- file.path(resdir, "informative-sine-benchmark-chisq4.rds")
 
@@
 m <- 20000                        # integer: number of hypothesis tests
-pi0 <- pi0_cubic(0.90)            # numeric: proportion of null hypotheses
+pi0 <- pi0_sine(0.90)             # numeric: proportion of null hypotheses
 icovariate <- runif               # functional: independent covariate

### 2018-04-01 datasets/simulations/simulations-informative-cubic.Rmd
Add step-func cov sim; increase singal size

@@
 ```{r gauss-parameters}
-tstat <- rnorm_generator(2)       # functional: dist of alternative test stats
+tstat <- rnorm_generator(3)       # functional: dist of alternative test stats
 tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
@@
 
-We plot the averaged results across `r B` simulations.
+We plot the averaged results across `r B` replications.
 
@@
 ```{r t5-parameters}
-tstat <- rnorm_generator(5)    # functional: dist of alternative test stats
+tstat <- rnorm_generator(6)    # functional: dist of alternative test stats
 tstat_dist <- rt_perturber(5)  # functional: sampling dist/noise for test stats
@@
 
-plotsim_average(resdf, met="rejections", filter_set = excludeSet,
+plotsim_average(resdf, met = "rejections", filter_set = excludeSet,
                 merge_ihw = TRUE, errorBars = TRUE)
 
-plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+plotsim_average(resdf, met = "FDR", filter_set = excludeSet,
                 merge_ihw = TRUE, errorBars = TRUE)
 
-plotsim_average(resdf, met="TPR", filter_set = excludeSet,
+plotsim_average(resdf, met = "TPR", filter_set = excludeSet,
                 merge_ihw = TRUE, errorBars = TRUE)
 
-plotsim_average(resdf, met="TNR", filter_set = excludeSet,
+plotsim_average(resdf, met = "TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE) 
+```
+
+Since both `scott-empirical` and `scott-theoretical` are not appropriate when
+the test statistics are not approximately normal, we also plot the results without
+these methods.
+
+```{r chisq4-metrics-averages-nofdrreg, results = "hide"}
+plotsim_average(resdf, met = "rejections",
+                filter_set = c(excludeSet, "scott-empirical", "scott-theoretical"),
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met = "FDR",
+                filter_set = c(excludeSet, "scott-empirical", "scott-theoretical"),
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met = "TPR",
+                filter_set = c(excludeSet, "scott-empirical", "scott-theoretical"),
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met = "TNR",
+                filter_set = c(excludeSet, "scott-empirical", "scott-theoretical"),
                 merge_ihw = TRUE, errorBars = TRUE) 

### 2018-04-01 datasets/simulations/simulations-informative-step.Rmd
Add step-func cov sim; increase singal size

--- 2018-04-01 datasets/simulations/simulations-informative-sine.Rmd
@@
 ---
-title: "Simulation Study: Informative Covariate (sine)"
+title: "Simulation Study: Informative Covariate (step)"
 author: "Patrick Kimes"
@@
 the interval [0, 1], and the conditional probability of a test being non-null
-is a smooth (sine) function of the covariate. We include simulation results
+is a non-smooth (monotone) step function of the covariate. We include simulation results
 again with Gaussian, t-distributed, and Chi-Squared distributed noise.
@@
 ## intermediary files we create below
-gauss_file <- file.path(resdir, "informative-sine-benchmark-gaussian.rds")
-tdist_file <- file.path(resdir, "informative-sine-benchmark-t5.rds")
-chisq_file <- file.path(resdir, "informative-sine-benchmark-chisq4.rds")
+gauss_file <- file.path(resdir, "informative-step-benchmark-gaussian.rds")
+tdist_file <- file.path(resdir, "informative-step-benchmark-t5.rds")
+chisq_file <- file.path(resdir, "informative-step-benchmark-chisq4.rds")
 
@@
 m <- 20000                        # integer: number of hypothesis tests
-pi0 <- pi0_sine(0.90)             # numeric: proportion of null hypotheses
+pi0 <- pi0_step(0.90)             # numeric: proportion of null hypotheses
 icovariate <- runif               # functional: independent covariate

### 2018-04-01 datasets/simulations/simulations-varyingeffectsize.Rmd
Add two more sim settings; varying pi0, varying ES

---
title: "Simulation Study: Varying Effect Size"
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

# Summary

In this set of simulations, we consider settings with both null and non-null
tests with varying distribution of effect sizes under the non-null (alternative)
setting. An informative covariate is included in the setting as described in
`simulations-informative-sine.Rmd`.

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(SummarizedBenchmark)
library(parallel)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
resdir <- "results"
dir.create(resdir, showWarnings = FALSE, recursive = TRUE)

## intermediary files we create below
spiky_file <- file.path(resdir, "varyinges-benchmark-spiky.rds")
flattop_file <- file.path(resdir, "varyinges-benchmark-flattop.rds")
skew_file <- file.path(resdir, "varyinges-benchmark-skew.rds")
bimodal_file <- file.path(resdir, "varyinges-benchmark-bimodal.rds")

## number of cores for parallelization
cores <- 20
B <- 100

## define bechmarking design
bd <- initializeBenchDesign()
```

All simulation settings will share the following parameters.

```{r parameters-shared}
m <- 20000                        # integer: number of hypothesis tests
pi0 <- pi0_sine(0.90)             # numeric: proportion of null hypotheses
tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
icovariate <- runif               # functional: independent covariate
```

Simulation results will be presented excluding a subset of methods, and
for certain plots (upset plots), a single alpha cutoff will be used.

```{r}
excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
ualpha <- 0.05
```

# Spiky Setting

First, we consider the setting where the effect sizes under the alternative are
distributed according to a "spiky" unimodal distribution centered around zero, as
defined in the ASH simulations.

## Data Simulation

```{r spiky-parameters}
tstat <- sampler_spiky       # functional: dist of alternative test stats
seed <- 778
```

We next run the simulations.

```{r spiky-run-simulation}
if (file.exists(spiky_file)) {
    res <- readRDS(spiky_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bd, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = spiky_file)
}
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r spiky-one-simulation}
onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, spiky-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, spiky-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r spiky-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r spiky-metrics-covlineplot, results = "hide"}
covariateLinePlot(res, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r spiky-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

# Flat-Top Setting

Next, we consider the setting where the effect sizes under the alternative are
distributed according to a "flat top" unimodal distribution centered around zero, as
defined in the ASH simulations.

## Data Simulation

```{r flattop-parameters}
tstat <- sampler_flat_top       # functional: dist of alternative test stats
seed <- 980
```

We next run the simulations.

```{r flattop-run-simulation}
if (file.exists(flattop_file)) {
    res <- readRDS(flattop_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bd, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = flattop_file)
}
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r flattop-one-simulation}
onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, flattop-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, flattop-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r flattop-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r flattop-metrics-covlineplot, results = "hide"}
covariateLinePlot(res, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r flattop-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

# Skewed Setting

Next, we consider the setting where the effect sizes under the alternative are
distributed according to a skewed unimodal distribution not centered at zero, as
defined in the ASH simulations.

## Data Simulation

```{r skew-parameters}
tstat <- sampler_skew       # functional: dist of alternative test stats
seed <- 206
```

We next run the simulations.

```{r skew-run-simulation}
if (file.exists(skew_file)) {
    res <- readRDS(skew_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bd, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = skew_file)
}
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r skew-one-simulation}
onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, skew-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, skew-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r skew-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r skew-metrics-covlineplot, results = "hide"}
covariateLinePlot(res, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r skew-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

# Bimodal Setting

Finally, we consider the setting where the effect sizes under the alternative are
distributed according to a bimodal distribution (equal mixture of two normal distributions
centered at -2, 2, with variance 1), again, as defined in the ASH simulations.

## Data Simulation

```{r bimodal-parameters}
tstat <- sampler_bimodal       # functional: dist of alternative test stats
seed <- 913
```

We next run the simulations.

```{r bimodal-run-simulation}
if (file.exists(bimodal_file)) {
    res <- readRDS(bimodal_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bd, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = bimodal_file)
}
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r bimodal-one-simulation}
onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, bimodal-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, bimodal-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r bimodal-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r bimodal-metrics-covlineplot, results = "hide"}
covariateLinePlot(res, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r bimodal-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

### 2018-04-01 datasets/simulations/simulations-varyingpi0.Rmd
Add two more sim settings; varying pi0, varying ES

---
title: "Simulation Study: Varying Null Proportion"
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

# Summary

In this set of simulations, we consider settings with both null and non-null
tests with varying proportion of null tests. An informative covariate is included
in the setting as described in `simulations-informative-sine.Rmd`.

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(SummarizedBenchmark)
library(parallel)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
resdir <- "results"
dir.create(resdir, showWarnings = FALSE, recursive = TRUE)

## intermediary files we create below
null80_file <- file.path(resdir, "varyingpi0-benchmark-nullprop80.rds")
null90_file <- file.path(resdir, "varyingpi0-benchmark-nullprop90.rds")
null95_file <- file.path(resdir, "varyingpi0-benchmark-nullprop95.rds")
null99_file <- file.path(resdir, "varyingpi0-benchmark-nullprop99.rds")

## number of cores for parallelization
cores <- 20
B <- 100

## define bechmarking design
bd <- initializeBenchDesign()
```

All simulation settings will share the following parameters.

```{r parameters-shared}
m <- 20000                        # integer: number of hypothesis tests
tstat <- rnorm_generator(3)       # functional: dist of alternative test stats
tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
icovariate <- runif               # functional: independent covariate
```

Simulation results will be presented excluding a subset of methods, and
for certain plots (upset plots), a single alpha cutoff will be used.

```{r}
excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
ualpha <- 0.05
```

# 80% Null Setting

First, we consider the setting where 80% of tests are null.

## Data Simulation

```{r null80-parameters}
pi0 <- pi0_sine(0.80)             # numeric: proportion of null hypotheses
seed <- 608
```

We next run the simulations.

```{r null80-run-simulation}
if (file.exists(null80_file)) {
    res <- readRDS(null80_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bd, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = null80_file)
}
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r null80-one-simulation}
onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, null80-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, null80-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r null80-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r null80-metrics-covlineplot, results = "hide"}
covariateLinePlot(res, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r null80-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

# 90% Null Setting

Next, we consider the setting where 90% of tests are null.

## Data Simulation

```{r null90-parameters}
pi0 <- pi0_sine(0.90)             # numeric: proportion of null hypotheses
seed <- 808
```

We next run the simulations.

```{r null90-run-simulation}
if (file.exists(null90_file)) {
    res <- readRDS(null90_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bd, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = null90_file)
}
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r null90-one-simulation}
onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, null90-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, null90-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r null90-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r null90-metrics-covlineplot, results = "hide"}
covariateLinePlot(res, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r null90-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

# 95% Null Setting

Next, we consider the setting where 95% of tests are null.

## Data Simulation

```{r null95-parameters}
pi0 <- pi0_sine(0.95)             # numeric: proportion of null hypotheses
seed <- 913
```

We next run the simulations.

```{r null95-run-simulation}
if (file.exists(null95_file)) {
    res <- readRDS(null95_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bd, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = null95_file)
}
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r null95-one-simulation}
onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, null95-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, null95-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r null95-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r null95-metrics-covlineplot, results = "hide"}
covariateLinePlot(res, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r null95-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

# 99% Null Setting

Next, we consider the setting where 99% of tests are null.

## Data Simulation

```{r null99-parameters}
pi0 <- pi0_sine(0.99)             # numeric: proportion of null hypotheses
seed <- 2015
```

We next run the simulations.

```{r null99-run-simulation}
if (file.exists(null99_file)) {
    res <- readRDS(null99_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bd, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = null99_file)
}
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r null99-one-simulation}
onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, null99-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, null99-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r null99-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r null99-metrics-covlineplot, results = "hide"}
covariateLinePlot(res, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r null99-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

### 2018-04-01 datasets/GSEA/gsea-human.Rmd
Add covLinePlots to GSEA case studies

@@
 
-The objective of this document is to test different multiple testing methods in the context of Gene Set Enrichment Analysis (GSEA).
-In order to get a list of genes, we will use RNA-seq data consisting on cortex and cerebellum samples of a subset of the GTEx data.
+The objective of this document is to compare different methods for multiple testing correction in the context of Gene Set Enrichment Analysis (GSEA).
+In order to get a list of genes, we will use RNA-seq data consisting of cortex and cerebellum samples of a subset of the GTEx data.
 
@@
 
-Gene set size of overlaps.
-
 ```{r}
-methods <- c("lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
-plotCovariateBoxplots(sGSEA, alpha=0.1, nsets=6, methods=methods, trans=log2)
-plotCovariateBoxplots(sGSEA, alpha=0.1, nsets=7, methods=methods, trans=log2, maxNum=5)
-```
-
-## Benchmark Metrics
-
-```{r}
-estimatePerformanceMetrics( pr, alpha=0.1, tidy=TRUE ) %>%
-    dplyr::filter( blabel %in% c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf") ) %>%
-    dplyr::select( blabel, value )
+covariateLinePlot(sGSEA, alpha = 0.1, covname = "ind_covariate")
 ```

### 2018-04-01 datasets/GSEA/gsea-human.Rmd
Remove ash,fdrreg from GSEA case studies

@@
 
-```{r runningBenckmark, message=FALSE}
+```{r setup-benckmark, message=FALSE}
 ## rename columns and prepare for benchmarking
@@
 bd <- initializeBenchDesign()
+bd <- dropBMethod(bd, "ashs")
+bd <- dropBMethod(bd, "scott-empirical")
+bd <- dropBMethod(bd, "scott-theoretical")
+```
 
-## run benchmark
+We exclude `ashr`, `scott-theoretical` and `scott-empirical` from the analysis because
+the necessary assumptions are not met in the current case study. Namely, effect sizes
+and standard errors are not available for ASH, and the test statistics are not normally
+distributed under the null and alternative, as required by Scott's FDR regression methods.
+
+```{r run-benchmark}
 if (!file.exists(bench_file)) {

### 2018-04-01 datasets/GSEA/gsea-mouse.Rmd
Add covLinePlots to GSEA case studies
Remove ash,fdrreg from GSEA case studies

@@
 
-Gene set size of overlaps.
-
 ```{r}
-methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
-plotCovariateBoxplots(sGSEAMouse, alpha = 0.1, nsets = 4,
-                      methods = methods, trans = log2, maxNum = 5)
+covariateLinePlot(sGSEAMouse, alpha = 0.1, covname = "ind_covariate")
 ```

### 2018-04-01 datasets/R/effectSizePlot.R
Add dplyr:: scope to calls in plotting functions

@@ -26,19 +26,19 @@ summarize_one_item <- function(object, alpha){
                             1*(assays(object)[["qvalue"]]) < 0.05))
   colnames(df)[1] <- "truth" 
   df <- df %>%
-    select(truth, covname, bonf, bh,
-           qvalue, contains("ihw"), ashs, "bl-df03", lfdr, "scott-theoretical", 
-           "scott-empirical") %>%
-    rename("scott-theoretical"="scott-t") %>%
-    rename("scott-empirical"="scott-e") %>%
-    gather(method, significant, -truth, -covname) %>%
-    group_by(method, truth, significant) %>%
-    summarize(mean_effect_size=mean(abs(get(covname))),
-              med_effect_size=median(abs(get(covname))),
-              min_effect_size=min(abs(get(covname))),
-              max_effect_size=max(abs(get(covname))),
-              botQ=quantile(abs(get(covname)),0.25),
-              topQ=quantile(abs(get(covname)), 0.75)) %>%
+    dplyr::select(truth, covname, bonf, bh,
+                  qvalue, contains("ihw"), ashs, "bl-df03", lfdr, "scott-theoretical", 
+                  "scott-empirical") %>%
+    dplyr::rename("scott-theoretical"="scott-t") %>%
+    dplyr::rename("scott-empirical"="scott-e") %>%
+    tidyr::gather(method, significant, -truth, -covname) %>%
+    dplyr::group_by(method, truth, significant) %>%
+    dplyr::summarize(mean_effect_size=mean(abs(get(covname))),
+                     med_effect_size=median(abs(get(covname))),
+                     min_effect_size=min(abs(get(covname))),
+                     max_effect_size=max(abs(get(covname))),
+                     botQ=quantile(abs(get(covname)),0.25),
+                     topQ=quantile(abs(get(covname)), 0.75)) %>%
     na.omit()
   df$significant <- as.factor(df$significant)
   levels(df$significant)=c("Not Significant", "Significant")
@@ -115,16 +115,16 @@ covariateLinePlot <- function(sbl, alpha=0.05, nbins = 50,
                               1*(assays(object)[["qvalue"]]) < 0.05))
     colnames(df)[1] <- "truth" 
     df <- df %>%
-      select(truth, covname, bonf, bh,
-             qvalue, contains("ihw"), ashs, "bl-df03", lfdr, "scott-theoretical", 
[…]

### 2018-04-03 datasets/ChIP-seq/ChIPseq-h3k4me3-csaw.Rmd
Add covLinePlots to ChIPseq case studies

@@
 
-```{r, chipseq-plot, results="hide"}
+```{r, chipseq-scatter, results="hide"}
 rejections_scatter(sb, supplementary = FALSE)
@@
                        bins = 4, supplementary = FALSE)
+```
 
+```{r, chipseq-overlap}
 plotFDRMethodsOverlap(sb, alpha = 0.05, nsets = ncol(sb),
@@
                       supplementary = FALSE)
+```
 
-methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
-plotCovariateBoxplots(sb, alpha = 0.1, nsets = 6,
-                      methods = methods, trans = log2, maxNum = 5)
+```{r, rnaseq-lineplots}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
 ```

### 2018-04-03 datasets/scrnaseq/scrnaseq-human-mast.Rmd
Add citations

@@
         number_sections: true
+bibliography: references.bib
 ---
@@
 We selected two datasets, one each from mouse and human. 
-These datasets come from the `Conquer` (Soneson et al., 2018) scRNAseq data collection.
+These datasets come from the `Conquer` [@soneson2018] scRNAseq data collection.
 `Conquer` provides uniformly processed gene expression data summarized to both count (aggregated from transcript members), and length-scaled TPMs at the gene level.
@@
 which genes have different expression among the two biological groups in each comparison.
-We also apply two different analysis methods: MAST and scDD. In this vignette, we will
-explore the human dataset analysed with MAST.
+We also apply two different analysis methods: MAST [@finak2015] and scDD [@korthauer2016]. 
+In this vignette, we will explore the human dataset analysed with MAST.
 
@@
 
-For data processing, we follow practices and recommendations from Soneson et al., 2017.
-Briefly, we limit ourselves to 2-group comparisons.
-We do not filter genes as filtering may affect the subsequent false discovery control steps.
-We only filter out genes that are expressed in
-fewer than 1% of cells. 
-We will filter out cells with extremely low mapping rate
+For data processing, we follow practices and recommendations from [@lun2016]. 
+Briefly, we perform quality control filtering on cells to ensure
+that technical artifacts do not distort downstream analysis results. In 
+addition, we also filter out very low-abundance genes, since genes that contain
+almost all zero counts do not contain enough information for reliable inference.
+Specifically, we filter out cells with extremely low mapping rate
 (below 20%), as these represent poor quality cells that may end up having negative
@@
 which have extremely low proportion of genes detected (less than 5%).
+We also filter out genes that are expressed in fewer than 5% of cells in both groups.
 
@@
 ```
+
+# References

### 2018-04-03 datasets/scrnaseq/scrnaseq-human-scdd.Rmd
Add citations

@@
 which genes have different expression among the two biological groups in each comparison.
-We also apply two different analysis methods: MAST and scDD. In this vignette, we will
-explore the human dataset analysed with scDD.
+We also apply two different analysis methods: MAST [@finak2015] and scDD [@korthauer2016]. 
+In this vignette, we will explore the human dataset analysed with MAST.
 

### 2018-04-03 datasets/scrnaseq/scrnaseq-mouse-mast.Rmd
Update headers, simplify p-val variable naming
Add citations

@@
 which genes have different expression among the two biological groups in each comparison.
-We also apply two different analysis methods: MAST and scDD. In this vignette, we will
-explore the mouse dataset analysed with MAST.
+We also apply two different analysis methods: MAST [@finak2015] and scDD [@korthauer2016]. 
+In this vignette, we will explore the human dataset analysed with MAST.
 
@@
 which have extremely low proportion of genes detected (less than 5%).
+We also filter out genes that are expressed in fewer than 5% of cells in both groups.
 
 *Add More detail on the experiments here*
+
 We'll extract two conditions: No stimulation versus LPS at 150 min.

### 2018-04-03 datasets/scrnaseq/scrnaseq-mouse-scdd.Rmd
Add citations

@@
 which genes have different expression among the two biological groups in each comparison.
-We also apply two different analysis methods: MAST and scDD. In this vignette, we will
-explore the mouse dataset analysed with scDD.
+We also apply two different analysis methods: MAST [@finak2015] and scDD [@korthauer2016]. 
+In this vignette, we will explore the human dataset analysed with MAST.
 
@@
 
-For data processing, we follow practices and recommendations from Soneson et al., 2018. Briefly, we limit ourselves to 2-group comparisons. We only filter out genes that are expressed in 
-fewer than 1% of cells. 
-We will filter out cells with extremely low mapping rate
+For data processing, we follow practices and recommendations from [@lun2016]. 
+Briefly, we perform quality control filtering on cells to ensure
+that technical artifacts do not distort downstream analysis results. In 
+addition, we also filter out very low-abundance genes, since genes that contain
+almost all zero counts do not contain enough information for reliable inference.
+Specifically, we filter out cells with extremely low mapping rate
 (below 20%), as these represent poor quality cells that may end up having negative

### 2018-04-03 datasets/microbiome/microbiome-enigma.Rmd
Move enigma data to dropbox

@@
 ## project data/results folders
-datdir <- "DATA"
-resdir <- "RESULTS"
-enigma_dir <- "enigma"
+datdir <- "data"
+resdir <- "results"
+enigma_dir <- file.path(datdir, "enigma_results")
 
@@
 al_bench_file <- file.path(resdir, "enigma-al-otus-benchmark.rds")
-
 ```
@@
 
-```{r, enigma-readdata}
+We download the data from Dropbox if not already available.
+
+```{r, enigma-download-data}
+if (!file.exists(file.path(enigma_dir, "enigma.metadata.txt"))) {
+    download.file("https://www.dropbox.com/s/0r1zjp26u10l0p4/enigma.metadata.txt?dl=1",
+                  destfile = file.path(enigma_dir, "enigma.metadata.txt"))
+}
+
+if (!file.exists(file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"))) {
+    download.file("https://www.dropbox.com/s/u9qh86cd7c0vbn9/enigma.otu_table_resampled_updated_r.txt.gz?dl=1",
+                  destfile = file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"))
+}
+```
+
+```{r enigma-readdata}
 ## load OTU table and metadata
-# Note to others: not sure why read.table breaks on this file.
-otu <- read.table(file.path(datdir, enigma_dir, 
-                            "otu_table_resampled_updated_r.txt"), 
+otu <- read.table(file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"), 
                   sep='\t', header = TRUE, row.names = 1)
-meta <- read.csv(file.path(datdir, enigma_dir, "metadata.txt"), sep='\t')
+meta <- read.csv(file.path(enigma_dir, "enigma.metadata.txt"), sep='\t')
 

### 2018-04-03 datasets/scrnaseq/scrnaseq-human-mast.Rmd
Add cellular detection rate to MAST model

@@
 
-We apply MAST to the TPMs following the approach in Soneson et al.
+We apply MAST to the TPMs and adjust for the cellular detection rate, as 
+suggested by the authors of MAST in the manuscript and vignette on Bioconductor.
 
@@
                     cData = data.frame(wellKey=colnames(dat), colData(dat)))
-    zlmdata <- zlm(~group, sca)
+    colData(sca)$cdr <- scale(colSums(assay(sca)>0))
+    zlmdata <- zlm(~ group + cdr, sca)
 

### 2018-04-04 datasets/ChIP-seq/ChIPseq-h3k4me3-csaw.Rmd
Fix error in chr filtering: hs not mm

@@
 The _csaw_ package provides a powerful and flexible approach to differential peak calling
-based on counting reads in sliding windows across the genome. Unlike the _DiffBind_ approach
-which restricts testing to peaks called from the same samples, the sliding window approach
-is able to detect differential peaks over sharper or wider regions, without being constrained
-by the original peak boundaries obtained from the peak calling algorithm, e.g. _MACS_.
+based on counting reads in sliding windows across the genome. The sliding window approach
+allows for the detection of differential peaks over sharper or wider regions, without being
+constrained by the original peak boundaries obtained from the peak calling algorithm, e.g. _MACS_.
 Differential peaks are called at each window using _edgeR_ and p-values are combined within
@@
 
-We download the fastq files directly from the UCSC ENCODE portal and align them against the reference.
+We download the fastq files directly from the UCSC ENCODE portal.
 
@@
 
-We also use the GRCh38 reference genome used for ENCODE3 analysis. If the reference index is
+We use the GRCh38 reference genome used in the ENCODE3 analysis. If the reference index is
 not already available, we must first download the reference file. Note that this is a different
-reference than what was originally used for the data.
+reference than what was originally used for the data in earlier analyses of the ENCODE project.
 
@@
 ```{r}
-std_chr <- paste0("chr", c(1:19, "X", "Y"))
+std_chr <- paste0("chr", c(1:22, "X", "Y"))
 param <- readParam(minq = 20, discard = bl, restrict = std_chr)

### 2018-04-04 datasets/ChIP-seq/ChIPseq-CBP-csaw.Rmd
Add first pass at CBP ChIPseq analysis

--- 2018-04-03 datasets/ChIP-seq/ChIPseq-h3k4me3-csaw.Rmd
@@
 
-In this case study, we perform differential peak calling on ChIP-seq data of the histone
-H3K4Me3 for samples from two cell lines (K562 and GM12878) using publicly available data
-generated by the ENCODE Project. The same data set is used for all ChIP-seq differential
-testing case studies.
-
-The _csaw_ package provides a powerful and flexible approach to differential peak calling
-based on counting reads in sliding windows across the genome. Unlike the _DiffBind_ approach
-which restricts testing to peaks called from the same samples, the sliding window approach
-is able to detect differential peaks over sharper or wider regions, without being constrained
-by the original peak boundaries obtained from the peak calling algorithm, e.g. _MACS_.
-Differential peaks are called at each window using _edgeR_ and p-values are combined within
-adjacent windows (clusters).
-
-For this analysis, the default _csaw_ settings are used as described in the _csaw_ "User Guide."
+In this case study, we perform differential peak calling on ChIP-seq data for a
+trnscription factor, CREB-binding protein (CBP), from Kasper et al. (2014), used in
+Lun et al. (2015) to demonstrate the usage of _csaw_. As described in a separate case
+study analyzing ChIP-seq data for H3K4Me3, _csaw_ is a package for differential peak
+calling based on counting reads in sliding windows across the genome. Code and steps
+for initial processing and analysis of this data with _csaw_ are taken directly
+from Lun et al. (2015).
 
@@
 ## intermediary files we create below
-count_file <- file.path(resdir, "h3k4me3-csaw-counts.rds")
-filtered_file <- file.path(resdir, "h3k4me3-csaw-counts-filtered.rds")
-result_file <- file.path(resdir, "h3k4me3-csaw-results.rds")
-bench_file <- file.path(resdir, "h3k4me3-csaw-benchmark.rds")
+count_file <- file.path(resdir, "cbp-csaw-counts.rds")
+filtered_file <- file.path(resdir, "cbp-csaw-counts-filtered.rds")
+normfacs_file <- file.path(resdir, "cbp-csaw-counts-normfacs.rds")
+result_file <- file.path(resdir, "cbp-csaw-results.rds")
+bench_file <- file.path(resdir, "cbp-csaw-benchmark.rds")
 
@@
 
-We download the fastq files directly from the UCSC ENCODE portal and align them against the reference.
+We download the fastq files directly from the European Nucleotide Archive (ENA).
 
-```{r, download-fastqs}
-broad_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeBroadHistone/"
-broad_fastqs <- c("wgEncodeBroadHistoneGm12878H3k4me3StdRawDataRep1.fastq.gz",
-                  "wgEncodeBroadHistoneGm12878H3k04me3StdRawDataRep2V2.fastq.gz",
-                  "wgEncodeBroadHistoneK562H3k4me3StdRawDataRep1.fastq.gz",
-                  "wgEncodeBroadHistoneK562H3k4me3StdRawDataRep2.fastq.gz")
+```{r download-fastqs}
+fqurls <- c("007/SRR1145787/SRR1145787.fastq.gz",
+            "008/SRR1145788/SRR1145788.fastq.gz",
+            "009/SRR1145789/SRR1145789.fastq.gz",
+            "000/SRR1145790/SRR1145790.fastq.gz")
+fqurls <- paste0("ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR114/", fqurls)
 
-for (i_fq in broad_fastqs) {
-    if (!file.exists(file.path(datdir, i_fq))) {
-        download.file(paste0(broad_url, i_fq),
-                      destfile = file.path(datdir, i_fq))
-    }
-}
-
-uw_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwHistone/"
-uw_fastqs <- c("wgEncodeUwHistoneGm12878H3k4me3StdRawDataRep1.fastq.gz",
-               "wgEncodeUwHistoneGm12878H3k4me3StdRawDataRep2.fastq.gz",
-               "wgEncodeUwHistoneK562H3k4me3StdRawDataRep1.fastq.gz",
-               "wgEncodeUwHistoneK562H3k4me3StdRawDataRep2.fastq.gz")
-
-for (i_fq in uw_fastqs) {
-    if (!file.exists(file.path(datdir, i_fq))) {
-        download.file(paste0(uw_url, i_fq),
-                      destfile = file.path(datdir, i_fq))
+for (i_fq in fqurls) {
+    out_fq <- file.path(datdir, basename(i_fq))
+    if (!file.exists(out_fq)) {
+        download.file(i_fq, destfile = out_fq)
     }
@@
 
-We also use the GRCh38 reference genome used for ENCODE3 analysis. If the reference index is
-not already available, we must first download the reference file. Note that this is a different
-reference than what was originally used for the data.
+We download the mm10 reference genome used for ENCODE3 and build the
+index for alignment if not already available.
 
@@
 ref_url <- paste0("https://www.encodeproject.org/files/",
-                  "GRCh38_no_alt_analysis_set_GCA_000001405.15/@@download/",
-                  "GRCh38_no_alt_analysis_set_GCA_000001405.15.fasta.gz")
+                  "mm10_no_alt_analysis_set_ENCODE/@@download/",
+                  "mm10_no_alt_analysis_set_ENCODE.fasta.gz")
 ref_fastagz <- file.path(datdir, basename(ref_url))
@@
 
-if (!file.exists(file.path(datdir, "ref_index.00.b.tab"))) {
+if (!file.exists(file.path(datdir, "mm_ref_index.00.b.tab"))) {
     if (!file.exists(ref_fasta)) {
@@
     }
-    buildindex(basename = file.path(datdir, "ref_index"),
+    buildindex(basename = file.path(datdir, "mm_ref_index"),
                reference = ref_fasta)
@@
 
-We determine sample metadata from the file names.
+Sample metadata is stored in a data.frame.
 
 ```{r}
-fqfiles <- file.path(datdir, c(broad_fastqs, uw_fastqs))
-labs <- gsub("wgEncode(.*)Histone.*", "\\1", basename(fqfiles))
-cells <- gsub("wgEncode.*Histone(.*)H3k.*", "\\1", basename(fqfiles))
-
-meta <- data.frame(cellline = cells, lab = labs,
-                   fqfile = fqfiles,
-                   bamfile = gsub("\\.fastq.gz", ".bam", fqfiles),
-                   sortedfile = gsub("\\.fastq.gz", ".sorted.bam", fqfiles),
-                   stringsAsFactors = FALSE)
+fqfiles <- file.path(dardir, basename(fqurls)
+meta <- data.frame(genotype = factor(c("wt", "wt", "ko", "ko")),
+                   fqurl = fqurls, fqfile = fqfiles,
+                   bamfile = gsub("\\.fastq\\.gz", ".bam", fqfiles),
+                   sortedfile = gsub("\\.fastq\\.gz", ".sorted.bam", fqfiles))
 ```
 
-We also download blacklisted regions for GRCh38 from ENCODE (https://www.encodeproject.org/annotations/ENCSR636HFF/).
+We also download blacklisted regions for mm10 from ENCODE (https://www.encodeproject.org/annotations/ENCSR636HFF/).
 
 ```{r}
-blacklist_url <- paste0("http://mitra.stanford.edu/kundaje/akundaje/",
-                        "release/blacklists/hg38-human/hg38.blacklist.bed.gz")
+blacklist_url <- "https://www.encodeproject.org/files/ENCFF547MET/@@download/ENCFF547MET.bed.gz"
 if (!file.exists(file.path(datdir, basename(blacklist_url)))) {
@@
 
-Standard set of parameters are defined for the analysis. Only the canonical set of chromosomes are used in the analysis.
+Standard set of parameters are defined for the analysis.
+Only the canonical set of chromosomes are used in the analysis.
 
@@
 } else {
-    ## Broad samples have quality score w/ phred offset +33 (standard)
-    b_meta <- meta[meta$lab == "Broad", ]
-    b_unaligned <- !file.exists(b_meta$bamfile)
-    if (any(b_unaligned)) {
-        align(index = file.path(datdir, "ref_index"),
-              readfile1 = b_meta$fqfile[b_unaligned],
-              TH1 = 2, type = 1, phredOffset = 33,
-              output_file = b_meta$bamfile[b_unaligned])
-    }
-
-    ## UW samples have quality score w/ phred offset +64 (old Illumina scale)
-    u_meta <- meta[meta$lab == "Uw", ]
-    u_unaligned <- !file.exists(u_meta$bamfile)
-    if (any(u_unaligned)) {
-        align(index = file.path(datdir, "ref_index"),
-              readfile1 = u_meta$fqfile[u_unaligned],
-              TH1 = 2, type = 1, phredOffset = 64,
-              output_file = u_meta$bamfile[u_unaligned])
+    unaligned <- !file.exists(meta$bamfile)
+    if (any(unaligned)) {
+        align(index = file.path(datdir, "mm_ref_index"),
+              readfile1 = meta$fqfile[unaligned],
+              type = 1, phredOffset = 64,
+              input_format = "FASTQ",
+              output_file = meta$bamfile[unaligned])
     }
@@
     ## mark duplicates w/ picard
-    temp_bam <- file.path(datdir, "temp_dups.bam")
-    temp_file <- file.path(datdir, "temp_mets.txt")
-    temp_dir <- file.path(datdir, "temp_dups")
+    temp_bam <- file.path(datdir, "temp_dups_mm.bam")
+    temp_file <- file.path(datdir, "temp_mets_mm.txt")
+    temp_dir <- file.path(datdir, "temp_dups_mm")
     dir.create(temp_dir)
@@
     ## use correlateReads to determine fragment length (remove dups)
-    b_x <- correlateReads(b_meta$sortedfile, param = reform(param, dedup = TRUE))
-    u_x <- correlateReads(u_meta$sortedfile, param = reform(param, dedup = TRUE))
-    b_frag_len <- which.max(b_x) - 1
-    u_frag_len <- which.max(u_x) - 1
+    x <- correlateReads(b_meta$sortedfile, param = reform(param, dedup = TRUE))
+    frag_len <- which.max(x) - 1
 
     ## count reads in sliding windows (keep dups)
-    win_cnts <- windowCounts(meta$sortedfile, param = param, width = 150,
-                             ext = list(rep(c(b_frag_len, u_frag_len), each = 4), 150))
+    win_cnts <- windowCounts(meta$sortedfile, param = param, width = 10, ext = frag.len)
 
@@
     filtered_cnts <- readRDS(filtered_file)
+    normfacs <- readRDS(normfacs_file)
 } else {
     ## filter windows by abundance
-    bins <- windowCounts(meta$sortedfile, bin = TRUE, width = 2000, param = param)
+    bins <- windowCounts(meta$sortedfile, bin = TRUE, width = 10000, param = param)
     filter_stat <- filterWindows(win_cnts, bins, type = "global")
-    keep <- filter_stat$filter > log2(2)
-    filtered_cnts <- win_cnts[keep, ]
+    keep <- filter_stat$filter > log2(3)
+    filtered_cnts <- win.data[keep, ]
 
-    ## save filtered counts
+    ## calculate normalizing offsets based on larger windows
+    normfacs <- normOffsets(bins)
+
+    ## save filtered counts, normalizing factors
     saveRDS(filtered_cnts, file = filtered_file)
+    saveRDS(normfacs, file = normfacs_file)
 }
@@
 } else {
-    ## normalize for library-specific trended biases
-    offsets <- normOffsets(filtered_cnts, type = "loess")
-
-    ## create model design
-    design <- model.matrix(~ lab + cellline,
-                           data = as.data.frame(meta))
+    ## create model diesign
+    design <- model.matrix(~ 0 + genotype, data = meta)
+    colnames(design) <- levels(meta$genotype)
 
     ## run quasi-likelihood F-test
-    y <- asDGEList(filtered_cnts)
-    y$offset <- offsets
+    y <- asDGEList(filtered_cnts, norm.factors = normfacs)
     y <- estimateDisp(y, design)
     fit <- glmQLFit(y, design, robust = TRUE)
-    res <- glmQLFTest(fit)
+    res <- glmQLFTest(fit, contrast = makeContrasts(wt-ko, levels = design))
 
@@
     elementMetadata(res_ranges) <-
-        data.frame(tab_comb,
+        data.frame(tabcom,
                    best_pos = mid(ranges(rowRanges(filtered_cnts[tab_best$best]))),
                    best_logFC = tab_best$logFC)
-
     saveRDS(res_ranges, file = result_file)

### 2018-04-05 datasets/RNA-Seq/yeast-simulation.Rmd
Fix typo in empirical comparison

@@
 resultsNames(dds) # lists the coefficients
-resFULL <- results(dds, name="condition_WT_vs_Snf2")
+resFULL <- results(dds, name="condition_WT_vs_Snf2", independentFiltering = F)
 
@@
 
-sum(resFULL$padj < 0.05, na.rm=T, independentFiltering = F)
+sum(resFULL$padj < 0.05, na.rm=T)
 ```

### 2018-04-05 datasets/ChIP-seq/ChIPseq-CBP-csaw.Rmd
Update ChIPseq analyses; only run picard if not indexed

@@
 ```{r}
-fqfiles <- file.path(dardir, basename(fqurls)
+fqfiles <- file.path(datdir, basename(fqurls))
 meta <- data.frame(genotype = factor(c("wt", "wt", "ko", "ko")),
@@
 
-    ## mark duplicates w/ picard
+    ## mark duplicates w/ picard and index bam files
+    if (any(!file.exists(paste0(meta$sortedfile, ".bai")))) {
         temp_bam <- file.path(datdir, "temp_dups_mm.bam")
@@
 
-    ## index bam files
-    if (any(!file.exists(paste0(meta$sortedfile, ".bai")))) {
         indexBam(meta$sortedfile)

### 2018-04-05 datasets/ChIP-seq/ChIPseq-h3k4me3-csaw.Rmd
Update ChIPseq analyses; only run picard if not indexed

@@
 
-    ## mark duplicates w/ picard
+    ## mark duplicates w/ picard and index bam files
+    if (any(!file.exists(paste0(meta$sortedfile, ".bai")))) {
         temp_bam <- file.path(datdir, "temp_dups.bam")

### 2018-04-05 datasets/ChIP-seq/ChIPseq-h3k4me3-promoters.Rmd
Update ChIPseq analyses; only run picard if not indexed

@@
     dds <- DESeq(dds, fitType = "mean")
-    dat <- as.data.frame(results(dds))
+    dat <- as.data.frame(results(dds, independentFiltering = FALSE))
     colnames(dat) <- c('ind_covariate', 'effect_size', 'SE',
@@
                          test_statistic = qnorm(exp(log(pval)-log(2)),
-                                                lower.tail=FALSE),
+                                                lower.tail = FALSE),
                          test_statistic = test_statistic * sign(effect_size))

### 2018-04-06 datasets/ChIP-seq/ChIPseq-CBP-csaw.Rmd
Fix typos in ChIPseq cpb Rmd code causing errors

@@
                    bamfile = gsub("\\.fastq\\.gz", ".bam", fqfiles),
-                   sortedfile = gsub("\\.fastq\\.gz", ".sorted.bam", fqfiles))
+                   sortedfile = gsub("\\.fastq\\.gz", ".sorted.bam", fqfiles),
+                   stringsAsFactors = FALSE)
 ```
@@
     ## use correlateReads to determine fragment length (remove dups)
-    x <- correlateReads(b_meta$sortedfile, param = reform(param, dedup = TRUE))
+    x <- correlateReads(meta$sortedfile, param = reform(param, dedup = TRUE))
     frag_len <- which.max(x) - 1
@@
     ## count reads in sliding windows (keep dups)
-    win_cnts <- windowCounts(meta$sortedfile, param = param, width = 10, ext = frag.len)
+    win_cnts <- windowCounts(meta$sortedfile, param = param, width = 10, ext = frag_len)
 
@@
     keep <- filter_stat$filter > log2(3)
-    filtered_cnts <- win.data[keep, ]
+    filtered_cnts <- win_data[keep, ]
 

### 2018-04-06 datasets/scrnaseq/scrnaseq-human-mast.Rmd
Add more details to dataset description

@@
 
-This dataset (from PMC5810554) contains cells from four Glioblastoma patients, 
+This dataset (from PMC5810554 [@darmanis2017]) contains cells from four Glioblastoma patients, 
 sampled from both the tumor core as well as the tissue peripheral to the tumor. 
-We'll compare the Neoplastic cells in the tumor core vs periphery. We'll exclude 
+We'll compare the Neoplastic cells in the tumor core vs periphery. Here it is of 
+interest to compare the gene expression profiles with cancerous cells that are part
+of the primary tumor with those that have migrated, since these differences are
+possibly related to metastasis. We'll exclude 
 the two samples where only a few neoplastic cells were found in the periphery.

### 2018-04-06 datasets/scrnaseq/scrnaseq-mouse-mast.Rmd
Add more details to dataset description

@@
 
-*Add More detail on the experiments here*
-
-We'll extract two conditions: No stimulation versus LPS at 150 min.
+This dataset (from PMC28396000 [@lane2017]) contains cells from a murine 
+macrophage cell line that were studied for
+the impact of stimulation of the innate immune transcription factor nuclear factor 
+$\kappa$B (NF-$\kappa$B) on gene expression. Specifically, it is of interest to 
+characterize the dynamics of global gene expression on activating immune response.
+The NF-$\kappa$B was stimulated using 
+a lipopolysaccharide (LPS)-dependent subunit fused to a fluorescent protein. 
+Here we extract and compare two conditions and compare unstimulated cells versus
+cells stimulated with LPS after 150 minutes.
 

### 2018-04-06 datasets/ChIP-seq/ChIPseq-CBP-csaw.Rmd
Fix more typos in CBP-csaw analysis; finally knits

@@
     keep <- filter_stat$filter > log2(3)
-    filtered_cnts <- win_data[keep, ]
+    filtered_cnts <- win_cnts[keep, ]
 
@@
     elementMetadata(res_ranges) <-
-        data.frame(tabcom,
+        data.frame(tab_comb,
                    best_pos = mid(ranges(rowRanges(filtered_cnts[tab_best$best]))),

### 2018-04-10 datasets/RNA-Seq/polyester-simulation.Rmd
added covariateLinePlot() functions and converted to relative paths

--- 2018-02-16 datasets/RNA-Seq/polyester-replicates-benchmarkFDR.Rmd
@@
 # comparison tools/functions
+# devtools::install_github("areyesq89/SummarizedBenchmark", ref="fdrbenchmark") # needed b/c BioC devel branch is 
 library(SummarizedBenchmark)
@@
 # load helper functions
-working_dir <- list(project_home = "/users/shicks1/projects/benchmark-fdr")
-sourceDirectory(file.path(working_dir$project_home,"datasets/R/"))
+sourceDirectory(file.path("../R/"))
 
@@
   sb <- bd %>% buildBench(data=geneExp, parallel = FALSE, 
-                          BPPARAM = BPPARAM, truthCols = "qvalue", 
+                          truthCols = "qvalue", 
                           ftCols = "ind_covariate")
@@
 ```{r}
-B <- 10
+B <- 100
 bd <- initializeBenchDesign() # only needs to be done once
@@
 }
-
-addDefaultMetrics <- function(sb) {
-    sb <- addPerformanceMetric(sb, evalMetric = c("TPR", "FPR", "TNR", "FNR", "rejections"),
-                               assay = "qvalue")
-    return(sb)
-}
-
 ```
@@
 
-if (!file.exists(file.path(working_dir, "datasets/RNA-Seq/polyester-results-null5.RDS"))){
+if (!file.exists(file.path("polyester-results-null5.RDS"))){
   null5 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
@@
                    BPPARAM=multicoreParam)
-  saveRDS(null5, file=file.path(working_dir, "datasets/RNA-Seq/polyester-results-null5.RDS"))
+  saveRDS(null5, file=file.path("polyester-results-null5.RDS"))
 } else {
-  null5 <- readRDS(file=file.path(working_dir, "datasets/RNA-Seq/polyester-results-null5.RDS"))
+  null5 <- readRDS(file=file.path("polyester-results-null5.RDS"))
 }
@@
 ```{r}
-
-res5 <- plotsim_standardize(null5, alpha = seq(0.01, 0.20, 0.01))
+res5 <- plotsim_standardize(null5, alpha = seq(0.01, 0.10, 0.01))
 
@@
 
+covariateLinePlot(null5, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(null5, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
 if (numberMethodsReject(res5, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
@@
 }
+
 ```
@@
 
-if (!file.exists(file.path(working_dir, "datasets/RNA-Seq/polyester-results-null10.RDS"))){
+if (!file.exists(file.path("polyester-results-null10.RDS"))){
   null10 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
@@
                    BPPARAM=multicoreParam)
-  saveRDS(null10, file=file.path(working_dir, "datasets/RNA-Seq/polyester-results-null10.RDS"))
+  saveRDS(null10, file=file.path("polyester-results-null10.RDS"))
 } else {
-  null10 <- readRDS(file=file.path(working_dir, "datasets/RNA-Seq/polyester-results-null10.RDS"))
+  null10 <- readRDS(file=file.path("polyester-results-null10.RDS"))
 }
@@
 ```{r}
-res10 <- plotsim_standardize(null10, alpha = seq(0.01, 0.20, 0.01))
+res10 <- plotsim_standardize(null10, alpha = seq(0.01, 0.10, 0.01))
 
-# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
 plotsim_average(res10, met="rejections",filter_set = excludeSet,
@@
                 merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(null10, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(null10, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
 
@@
 
-if (!file.exists(file.path(working_dir, "datasets/RNA-Seq/polyester-results-de5.RDS"))){
+if (!file.exists(file.path("polyester-results-de5.RDS"))){
   de5 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
@@
                    BPPARAM=multicoreParam)
-  saveRDS(de5, file=file.path(working_dir, "datasets/RNA-Seq/polyester-results-de5.RDS"))
+  saveRDS(de5, file=file.path("polyester-results-de5.RDS"))
 } else {
-  de5 <- readRDS(file=file.path(working_dir, "datasets/RNA-Seq/polyester-results-de5.RDS"))
+  de5 <- readRDS(file=file.path("polyester-results-de5.RDS"))
 }
@@
 ```{r}
-plotEffectSize <- function(sbl, alpha=0.05, DE=TRUE, covname){
-
-summarize_one_item <- function(object, alpha){
-  object <- object[,!( grepl("^ihw", as.character( colData( object )$blabel ) )
-                       & colData( object )$param.alpha != alpha )]
-
-  df <- as.data.frame(cbind(as.matrix(rowData(object)), 
-                            1*(assays(object)[["qvalue"]]) < 0.05))
-  colnames(df)[1] <- "truth" 
-  df <- df %>%
-    select(truth, covname, bonf, bh,
-                  qvalue, contains("ihw"), ashs, "bl-df03", lfdr, 
-                  "scott-theoretical", "scott-empirical") %>%
-    # rename("scott-theoretical"="scott-t", "scott-empirical"="scott-e") %>%
-    gather(method, significant, -truth, -covname) %>%
-    group_by(method, truth, significant) %>%
-    summarize(mean_effect_size=mean(abs(get(covname))),
-              med_effect_size=median(abs(get(covname))),
-              min_effect_size=min(abs(get(covname))),
-              max_effect_size=max(abs(get(covname))),
-              botQ=quantile(abs(get(covname)),0.25),
-              topQ=quantile(abs(get(covname)), 0.75)) %>% na.omit()
-  df$significant <- as.factor(df$significant)
-  levels(df$significant)=c("Not Significant", "Significant")
-  df$truth <- as.factor(df$truth)
-  levels(df$truth)=c("Not DE", "DE")
-  return(df)
-}
-
-df <- lapply(sbl, summarize_one_item, alpha=alpha)
-df <- bind_rows(df, .id = "rep")
-df <- as.tibble(df)
-
-if (DE){
-  df <- df %>% filter(truth=="DE")
-}else{
-  df <- df %>% filter(truth=="Not DE")
-}
-
-p <- ggplot(df, aes(x = as.factor(method), y=med_effect_size)) +
-  geom_boxplot() +
-  facet_grid(significant~., scales="free_y") +
-  xlab("Method") +
-  ylab(paste0("Median ", covname))
-
-if (DE){
-  p <- p + ggtitle(paste0(covname, " of True DE genes by significance at alpha ", 
-                       alpha))
-}else{
-  p <- p + ggtitle(paste0(covname, " of Non-DE genes by significance at alpha ", 
-                       alpha))
-}
-
-return(p)
-}
 
@@
 
-# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
 plotsim_average(res5d, met="rejections",filter_set = excludeSet,
@@
 
-plotEffectSize(de5, covname="log2FC")
-plotEffectSize(de5, covname="log2FC", DE=FALSE)
-
-plotEffectSize(de5, covname="ind_covariate")
-plotEffectSize(de5, covname="ind_covariate", DE=FALSE)
+covariateLinePlot(de5, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(de5, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
 
@@
 
-if (!file.exists(file.path(working_dir, "datasets/RNA-Seq/polyester-results-de10.RDS"))){
+if (!file.exists(file.path("polyester-results-de10.RDS"))){
   de10 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
@@
                    BPPARAM=multicoreParam)
-  saveRDS(de10, file=file.path(working_dir, "datasets/RNA-Seq/polyester-results-de10.RDS"))
+  saveRDS(de10, file=file.path("polyester-results-de10.RDS"))
 } else {
-  de10 <- readRDS(file=file.path(working_dir, "datasets/RNA-Seq/polyester-results-de10.RDS"))
+  de10 <- readRDS(file=file.path("polyester-results-de10.RDS"))
 }
@@
 
-# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
 plotsim_average(res10d, met="rejections",filter_set = excludeSet, 
@@
 
-plotEffectSize(de10, covname="log2FC")
-plotEffectSize(de10, covname="log2FC", DE=FALSE)
-
-plotEffectSize(de10, covname="ind_covariate")
-plotEffectSize(de10, covname="ind_covariate", DE=FALSE)
+covariateLinePlot(de10, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(de10, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
 

### 2018-04-13 datasets/scrnaseq/scrnaseq-human-wilcox.Rmd
AS:Adding scRNAseq benchmark analyses for the Wilcoxon Rank-Sum method for the current human+mouse datasets

--- 2018-04-06 datasets/scrnaseq/scrnaseq-human-mast.Rmd
@@
 ---
-title: "Case Study: scRNA-seq (Human data, MAST method)"
+title: "Case Study: scRNA-seq (Human data, Wilcoxon-ranksum method)"
 author: "Ayshwarya Subramanian and Keegan Korthauer"
@@
 kable(df,caption="scRNAseq Datasets")
-df1=data.frame(DE_method=c("scDD","MAST"),
-               input=c("count","tpm"),
-               output=c("p-val","p-val"))
+df1=data.frame(DE_method=c("scDD","MAST","Wilcoxon_Ranksum"),
+               input=c("count","tpm","tpm"),
+               output=c("p-val","p-val","pval"))
 kable(df1,caption="DE methods")
@@
 which genes have different expression among the two biological groups in each comparison.
-We also apply two different analysis methods: MAST [@finak2015] and scDD [@korthauer2016]. 
-In this vignette, we will explore the human dataset analysed with MAST.
+We also apply three different analysis methods: the non-parametric Wilcoxon Rank-sum test, Wilcoxon_Ranksum [@finak2015] and scDD [@korthauer2016]. 
+In this vignette, we will explore the human dataset analysed with the Wilcoxon Rank-sum test.
 
@@
 library(edgeR)
-library(MAST)
 library(BiocParallel)
@@
 ```{r}
-datfile <- file.path(datdir, "human_GSE84465_mast.rds")
+datfile <- file.path(datdir, "human_GSE84465_wilcox.rds")
 
@@
 
-We perform differential expression testing using 
-MAST which has a hurdle model to account for zeroes.
+We perform differential expression testing using the non-parametric Wilcoxon-rank Sum test.
 
-We apply MAST to the TPMs and adjust for the cellular detection rate, as 
-suggested by the authors of MAST in the manuscript and vignette on Bioconductor.
+We use the base R function `wilcox.test()` and run it on the TPMs.
+We normalize the counts using the TMM function as suggested by Soneson et al.
+who did a comprehensive benchmarking of differential expression tests for single-cell RNAseq data.
 
-```{r, MAST}
-# function to add MAST p-values
-compute_MAST_sc <- function(dat){
+```{r, Wilcoxon_Ranksum}
+# function to add Wilcoxon_Ranksum p-values
+# code borrowed from https://github.com/csoneson/conquer_comparison/blob/master/scripts/apply_Wilcoxon.R
+compute_Wilcoxon_Ranksum_sc <- function(dat){  
 # check if already computed
-  if (! "mast_p" %in% colnames(rowData(dat))){
-    sca <- FromMatrix(exprsArray = log2(assay(dat, "TPM") + 1), 
-                    cData = data.frame(wellKey=colnames(dat), colData(dat)))
-    colData(sca)$cdr <- scale(colSums(assay(sca)>0))
-    zlmdata <- zlm(~ group + cdr, sca)
-
-    rowData(dat)$mast_p <- 
-        lrTest(zlmdata, "group")[, "hurdle", "Pr(>Chisq)"]
+if (! "wilcox_p" %in% colnames(rowData(dat))){
+tmm <- edgeR::calcNormFactors(assay(dat, "TPM"))
+tpmtmm <- edgeR::cpm(assay(dat, "TPM"), lib.size = tmm * colSums(assay(dat, "TPM")))
+idx <- 1:nrow(tpmtmm)
+names(idx) <- rownames(tpmtmm)
+rowData(dat)$wilcox_p <- sapply(idx, function(i) {
+  wilcox.test(tpmtmm[i, ] ~ colData(dat)$group)$p.value
+})
 }else{
-     message("MAST p-values already computed.")
+     message("Wilcoxon_Ranksum p-values already computed.")
   }
@@
 
-humandata <- compute_MAST_sc(humandata)
+humandata <- compute_Wilcoxon_Ranksum_sc(humandata)
 
@@
 rank_scatter(data.frame(rowData(humandata)), 
-             pvalue="mast_p", covariate="meanExp") + 
-  ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
+             pvalue="wilcox_p", covariate="meanExp") + 
+  ggtitle("Wilcoxon_Ranksum, Covariate 1: Mean Nonzero Expression")
 ```
@@
 strat_hist(data.frame(rowData(humandata)), 
-           pvalue="mast_p", covariate="meanExp", maxy=16,
-           main = "MAST, Covariate 1: Mean Nonzero Expression")
+           pvalue="wilcox_p", covariate="meanExp", maxy=16,
+           main = "Wilcoxon_Ranksum, Covariate 1: Mean Nonzero Expression")
 ```
@@
 rank_scatter(data.frame(rowData(humandata)), 
-             pvalue="mast_p", covariate="detection") + 
-  ggtitle("MAST, Covariate 2: Detection Rate")
+             pvalue="wilcox_p", covariate="detection") + 
+  ggtitle("Wilcoxon_Ranksum, Covariate 2: Detection Rate")
 ```
@@
 strat_hist(data.frame(rowData(humandata)), 
-           pvalue="mast_p", covariate="detection", maxy=16,
-           main = "MAST, Covariate 2: Detection Rate") 
+           pvalue="wilcox_p", covariate="detection", maxy=16,
+           main = "Wilcoxon_Ranksum, Covariate 2: Detection Rate") 
 ```
@@
 ```{r, sb1, results="hide", message=FALSE}
-res1 <- file.path(resdir, "sb_GSE84465_mast_mean.rds")
+res1 <- file.path(resdir, "sb_GSE84465_Wilcoxon_Ranksum_mean.rds")
 
@@
                                    na.omit() %>%
-                                   mutate(pval=mast_p,
+                                   mutate(pval=wilcox_p,
                                           ind_covariate=meanExp), 
@@
 ```{r, sb2, results="hide", message=FALSE}
-res2 <- file.path(resdir, "sb_GSE84465_mast_det.rds")
+res2 <- file.path(resdir, "sb_GSE84465_Wilcoxon_Ranksum_det.rds")
 
@@
                                    na.omit() %>%
-                                   mutate(pval=mast_p,
+                                   mutate(pval=wilcox_p,
                                           ind_covariate=detection), 
@@
                    supplementary=FALSE) +
-  ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
+  ggtitle("Wilcoxon_Ranksum, Covariate 1: Mean Nonzero Expression")
 
@@
                        supplementary=FALSE) +
-  ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
+  ggtitle("Wilcoxon_Ranksum, Covariate 1: Mean Nonzero Expression")
 
@@
                  trans = "log1p") +
-  ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
+  ggtitle("Wilcoxon_Ranksum, Covariate 1: Mean Nonzero Expression")
 ```
@@
 rejections_scatter(sb2, supplementary=FALSE) +
-  ggtitle("MAST, Covariate 2: Detection Rate")
+  ggtitle("Wilcoxon_Ranksum, Covariate 2: Detection Rate")
 
@@
                        supplementary=FALSE) +
-  ggtitle("MAST, Covariate 2: Detection Rate")
+  ggtitle("Wilcoxon_Ranksum, Covariate 2: Detection Rate")
 
@@
 covariateLinePlot(sb2, alpha=0.05, covname="ind_covariate", nbins=25) +
-  ggtitle("MAST, Covariate 2: Detection Rate")
+  ggtitle("Wilcoxon_Ranksum, Covariate 2: Detection Rate")
 ```

### 2018-04-13 datasets/scrnaseq/scrnaseq-mouse-wilcox.Rmd
AS:Adding scRNAseq benchmark analyses for the Wilcoxon Rank-Sum method for the current human+mouse datasets

--- 2018-04-06 datasets/scrnaseq/scrnaseq-mouse-mast.Rmd
@@
 ---
-title: "Case Study: scRNA-seq (Mouse data, MAST method)"
+title: "Case Study: scRNA-seq (Mouse data, Wilcoxon-ranksum method)"
 author: "Ayshwarya Subramanian and Keegan Korthauer"
@@
 kable(df,caption="scRNAseq Datasets")
-df1=data.frame(DE_method=c("scDD","MAST"),
-               input=c("count","tpm"),
-               output=c("p-val","p-val"))
+df1=data.frame(DE_method=c("scDD","MAST","Wilcoxon-ranksum"),
+               input=c("count","tpm","tpm"),
+               output=c("p-val","p-val","p-val"))
 kable(df1,caption="DE methods")
@@
 which genes have different expression among the two biological groups in each comparison.
-We also apply two different analysis methods: MAST [@finak2015] and scDD [@korthauer2016]. 
-In this vignette, we will explore the human dataset analysed with MAST.
+We also apply two different analysis methods: Wilcoxon-ranksum [@finak2015] and scDD [@korthauer2016]. 
+In this vignette, we will explore the human dataset analysed with Wilcoxon-ranksum.
 
@@
 ```{r}
-datfile <- file.path(datdir, "mouse_GSE94383_mast.rds")
+datfile <- file.path(datdir, "mouse_GSE94383_Wilcoxon-ranksum.rds")
 
@@
 
-We perform differential expression testing using 
-MAST which has a hurdle model to account for zeroes.
+We perform differential expression testing using the non-parametric Wilcoxon-rank Sum test.
 
-We apply MAST to the TPMs and adjust for the cellular detection rate, as 
-suggested by the authors of MAST in the manuscript and vignette on Bioconductor.
+We use the base R function `wilcox.test()` and run it on the TPMs.
+We normalize the counts using the TMM function as suggested by Soneson et al.
+who did a comprehensive benchmarking of differential expression tests for single-cell RNAseq data.
 
-```{r, MAST}
-# function to add MAST p-values
-compute_MAST_sc <- function(dat){
+```{r, Wilcoxon-ranksum}
+# function to add Wilcoxon_Ranksum p-values
+# code borrowed from https://github.com/csoneson/conquer_comparison/blob/master/scripts/apply_Wilcoxon.R
+compute_Wilcoxon_Ranksum_sc <- function(dat){  
 # check if already computed
-  if (! "mast_p" %in% colnames(rowData(dat))){
-    sca <- FromMatrix(exprsArray = log2(assay(dat, "TPM") + 1), 
-                    cData = data.frame(wellKey=colnames(dat), colData(dat)))
-    colData(sca)$cdr <- scale(colSums(assay(sca)>0))
-    zlmdata <- zlm(~ group + cdr, sca)
-
-    rowData(dat)$mast_p <- 
-        lrTest(zlmdata, "group")[, "hurdle", "Pr(>Chisq)"]
+if (! "wilcox_p" %in% colnames(rowData(dat))){
+tmm <- edgeR::calcNormFactors(assay(dat, "TPM"))
+tpmtmm <- edgeR::cpm(assay(dat, "TPM"), lib.size = tmm * colSums(assay(dat, "TPM")))
+idx <- 1:nrow(tpmtmm)
+names(idx) <- rownames(tpmtmm)
+rowData(dat)$wilcox_p <- sapply(idx, function(i) {
+  wilcox.test(tpmtmm[i, ] ~ colData(dat)$group)$p.value
+})
 }else{
-     message("MAST p-values already computed.")
+     message("Wilcoxon_Ranksum p-values already computed.")
   }
@@
 
-mousedata <- compute_MAST_sc(mousedata)
+mousedata <- compute_Wilcoxon_Ranksum_sc(mousedata)
 
@@
 rank_scatter(data.frame(rowData(mousedata)), 
-             pvalue="mast_p", covariate="meanExp") + 
-  ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
+             pvalue="wilcox_p", covariate="meanExp") + 
+  ggtitle("Wilcoxon-ranksum, Covariate 1: Mean Nonzero Expression")
 ```
@@
 strat_hist(data.frame(rowData(mousedata)),
-           pvalue="mast_p", covariate="meanExp", maxy=20,
-           main = "MAST, Covariate 1: Mean Nonzero Expression")
+           pvalue="wilcox_p", covariate="meanExp", maxy=20,
+           main = "Wilcoxon-ranksum, Covariate 1: Mean Nonzero Expression")
 ```
@@
 rank_scatter(data.frame(rowData(mousedata)), 
-             pvalue="mast_p", covariate="detection") + 
-  ggtitle("MAST, Covariate 2: Detection Rate")
+             pvalue="wilcox_p", covariate="detection") + 
+  ggtitle("Wilcoxon-ranksum, Covariate 2: Detection Rate")
 ```
@@
 strat_hist(data.frame(rowData(mousedata)), 
-           pvalue="mast_p", covariate="detection", maxy=20,
-           main = "MAST, Covariate 2: Detection Rate")
+           pvalue="wilcox_p", covariate="detection", maxy=20,
+           main = "Wilcoxon-ranksum, Covariate 2: Detection Rate")
 ```
@@
 ```{r, sb1, results="hide", message=FALSE}
-res1 <- file.path(resdir, "sb_GSE94383_mast_mean.rds")
+res1 <- file.path(resdir, "sb_GSE94383_Wilcoxon-ranksum_mean.rds")
 
@@
 ```{r, sb2, results="hide", message=FALSE}
-res2 <- file.path(resdir, "sb_GSE94383_mast_det.rds")
+res2 <- file.path(resdir, "sb_GSE94383_Wilcoxon-ranksum_det.rds")
 
@@
                    supplementary=FALSE) +
-  ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
+  ggtitle("Wilcoxon-ranksum, Covariate 1: Mean Nonzero Expression")
 
@@
                        supplementary=FALSE) +
-  ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
+  ggtitle("Wilcoxon-ranksum, Covariate 1: Mean Nonzero Expression")
 
@@
                  trans = "log1p") +
-  ggtitle("MAST, Covariate 1: Mean Nonzero Expression")
+  ggtitle("Wilcoxon-ranksum, Covariate 1: Mean Nonzero Expression")
 ```
@@
 rejections_scatter(sb2, supplementary=FALSE) +
-  ggtitle("MAST, Covariate 2: Detection Rate")
+  ggtitle("Wilcoxon-ranksum, Covariate 2: Detection Rate")
 
@@
                        supplementary=FALSE) +
-  ggtitle("MAST, Covariate 2: Detection Rate")
+  ggtitle("Wilcoxon-ranksum, Covariate 2: Detection Rate")
 
@@
 covariateLinePlot(sb2, alpha=0.05, covname="ind_covariate", nbins=25) +
-  ggtitle("MAST, Covariate 2: Detection Rate")
+  ggtitle("Wilcoxon-ranksum, Covariate 2: Detection Rate")
 ```

### 2018-04-18 datasets/R/common-BenchDesign.R
Remove FDRreg from common BenchDesign

@@ -31,8 +31,7 @@
 #' @import IHW ashr qvalue swfdr fdrtool FDRreg magrittr SummarizedBenchmark
 #' 
 #' @author Patrick Kimes
-
-initializeBenchDesign <- function(nmids=150) {
+initializeBenchDesign <- function(nmids = 150) {
   ## ###########################################################################
   ## check for necessary packages and load them if they aren't present
   ## ###########################################################################
@@ -94,22 +93,6 @@ initializeBenchDesign <- function(nmids=150) {
   bd %<>% addBMethod("lfdr",
                      clfdr_hickswrapper,
                      unadj_p = pval, groups = IHW::groups_by_filter(ind_covariate, 20))
-  ## Scott's FDR regression w/ theoretical null
-  bd %<>% addBMethod("scott-theoretical",
-                     FDRreg::FDRreg,
-                     function(x) { x$FDR },
-                     z = qnorm(exp(log(pval) - log(2)), lower.tail=FALSE) * sign(test_statistic),
-                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
-                     nulltype = 'theoretical',
-                     control = list(lambda = 0.01))
-  ## Scott's FDR regression w/ empirical null
-  bd %<>% addBMethod("scott-empirical",
-                     FDRreg::FDRreg,
-                     function(x) { x$FDR },
-                     z = qnorm(exp(log(pval) - log(2)), lower.tail=FALSE) * sign(test_statistic),
-                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
-                     nulltype = 'empirical',
-                     control = list(lambda = 0.01, nmids = nmids))
   return(bd)
 }
 

### 2018-04-18 datasets/simulations/simulations-informative-cubic.Rmd
Manually add FDRreg in all normal/t simulations

@@
 
+As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
+for simulations with Gaussian or t-distributed noise. Again, we include both
+`nulltype = "empirical"` and `nulltype = "theoretical"`.
+
+```{r}
+bdplus <- bd
+bdplus <- addBMethod(bdplus, "scott-theoretical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'theoretical',
+                     control = list(lambda = 0.01, nmids = nmids))
+bdplus <- addBMethod(bdplus, "scott-empirical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'empirical',
+                     control = list(lambda = 0.01, nmids = nmids))
+```
+
 All simulation settings will share the following parameters.
@@
 
-We next run the simulations.
+We next run the simulations (including Scott's FDR Regression).
 
@@
 } else {
-    res <- mclapply(X = 1:B, FUN = simIteration, bench = bd, m = m,
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
                     pi0 = pi0, tstat = tstat, icovariate = icovariate,
@@
 ```{r gauss-one-simulation}
-onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r t5-one-simulation}
-onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 
-We next run the simulations.
+We next run the simulations. We do not include FDR Regression because the test statistics
+are not approximately normally distributed.
 

### 2018-04-18 datasets/simulations/simulations-null.Rmd
Manually add FDRreg in all normal/t simulations

@@
 
+We include Scott's FDR Regression in the analysis. We include both `nulltype = "empirical"` and
+`nulltype = "theoretical"`. In both cases, test statistics are modeled as being
+sampled from mean-shifted normal distributions with equal variance under the null and alternative.
+While `nulltype = "theoretical"` makes the additional assumption that the null is standard normal,
+`nulltype = "empirical"` attempts to estimate the mean and variance under the null empirically.
+While the assumptions of `nulltype = "theoretical"` are met for a subset of the simulations, we
+include `nulltype = "empirical"` as this is not true for all simulations, and not necessarily
+reflective of all real data sets.
+
+```{r}
+bdplus <- bd
+bdplus <- addBMethod(bdplus, "scott-theoretical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'theoretical',
+                     control = list(lambda = 0.01, nmids = nmids))
+bdplus <- addBMethod(bdplus, "scott-empirical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'empirical',
+                     control = list(lambda = 0.01, nmids = nmids))
+```
+
 All simulation settings will share the following parameters.

### 2018-04-18 datasets/simulations/simulations-varyingeffectsize.Rmd
Manually add FDRreg in all normal/t simulations

@@
 
+As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
+for simulations with Gaussian or t-distributed noise. Again, we include both
+`nulltype = "empirical"` and `nulltype = "theoretical"`. Since all settings in this
+series of simulations use test statistics simulated with Gaussian noise, we include
+Scott's FDR Regression in all of the comparisons.
+
+```{r}
+bdplus <- bd
+bdplus <- addBMethod(bdplus, "scott-theoretical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'theoretical',
+                     control = list(lambda = 0.01, nmids = nmids))
+bdplus <- addBMethod(bdplus, "scott-empirical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'empirical',
+                     control = list(lambda = 0.01, nmids = nmids))
+```
+
 All simulation settings will share the following parameters.
@@
 ```{r spiky-one-simulation}
-onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r flattop-one-simulation}
-onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r skew-one-simulation}
-onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r bimodal-one-simulation}
-onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)

### 2018-04-18 datasets/simulations/simulations-varyingpi0.Rmd
Manually add FDRreg in all normal/t simulations

@@
 ```{r null80-one-simulation}
-onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r null90-one-simulation}
-onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r null95-one-simulation}
-onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r null99-one-simulation}
-onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)

### 2018-04-18 datasets/scrnaseq/scrnaseq-mouse-wilcox.Rmd
Fix typos in Wilcox scrnaseq

@@
 which genes have different expression among the two biological groups in each comparison.
-We also apply two different analysis methods: Wilcoxon-ranksum [@finak2015] and scDD [@korthauer2016]. 
-In this vignette, we will explore the human dataset analysed with Wilcoxon-ranksum.
+We also apply three different analysis methods: Wilcoxon-ranksum, MAST [@finak2015],
+and scDD [@korthauer2016]. 
+In this vignette, we will explore the mouse dataset analysed with Wilcoxon-ranksum.
 

### 2018-04-18 datasets/scrnaseq/scrnaseq-knit.sh
Add Wilcox scrnaseq to batch submit file

@@ -17,6 +17,7 @@ R -e "rmarkdown::render('scrnaseq-human-scdd.Rmd', clean=FALSE)"
 R -e "rmarkdown::render('scrnaseq-mouse-scdd.Rmd', clean=FALSE)"
 R -e "rmarkdown::render('scrnaseq-human-mast.Rmd', clean=FALSE)"
 R -e "rmarkdown::render('scrnaseq-mouse-mast.Rmd', clean=FALSE)"
-
+R -e "rmarkdown::render('scrnaseq-human-wilcox.Rmd', clean=FALSE)"
+R -e "rmarkdown::render('scrnaseq-mouse-wilcox.Rmd', clean=FALSE)"
 
 

### 2018-04-18 datasets/GWAS/GWAS.Rmd
Explicitly add Scott to GWAS

@@
 
+We also add in Scott's FDR Regression (both
+`nulltype = "empirical"` and `nulltype = "theoretical"`)
+since our test statistics are t-distributed.
+
+```{r}
+bd <- addBMethod(bd, "scott-theoretical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'theoretical',
+                     control = list(lambda = 0.01))
+bd <- addBMethod(bd, "scott-empirical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'empirical',
+                     control = list(lambda = 0.01))
+```
+
 Now, we're ready to construct the `SummarizedBenchmark` object, which will run

### 2018-04-19 datasets/simulations/simulations-informative-cubic.Rmd
Remove nmids arg for FDR reg from sims

@@
 
+For comparison, we also include the ASH q-values.
+
+```{r}
+bd <- addBMethod(bd, "ashq", ashr::ash, ashr::get_qvalue,
+                 betahat = effect_size, sebetahat = SE)
+```
+
 As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
@@
                      nulltype = 'theoretical',
-                     control = list(lambda = 0.01, nmids = nmids))
+                     control = list(lambda = 0.01))
 bdplus <- addBMethod(bdplus, "scott-empirical",
@@
                      nulltype = 'empirical',
-                     control = list(lambda = 0.01, nmids = nmids))
+                     control = list(lambda = 0.01))
 ```

### 2018-04-19 datasets/simulations/simulations-null.Rmd
Remove nmids arg for FDR reg from sims

@@
 
+For comparison, we also include the ASH q-values.
+
+```{r}
+bd <- addBMethod(bd, "ashq", ashr::ash, ashr::get_qvalue,
+                 betahat = effect_size, sebetahat = SE)
+```
+
 We include Scott's FDR Regression in the analysis. We include both `nulltype = "empirical"` and

### 2018-04-20 datasets/RNA-Seq/RNAseq-gtex.Rmd
Add covLinePlots to RNAseq case studies
Small changes to results print out in RNAseq case studies

@@
 
-```{r}
+```{r, fig.width=10, fig.height=3.2}
 res <- dplyr::filter(res, ind_covariate >= 1)
@@
 sb <- addDefaultMetrics(sb)
-estimatePerformanceMetrics(sb, tidy = TRUE) %>%
+estimatePerformanceMetrics(sb, alpha = 0.05, tidy = TRUE) %>%
     filter(performanceMetric == "rejections") %>%
     select(blabel, performanceMetric, alpha, value) %>%
-    arrange(desc(value))
+    mutate(n = nrow(sb), prop = round(value / n, 3)) %>%
+    arrange(desc(value)) %>%
+    as_tibble() %>%
+    print(n = 40)
 ```
@@
 
-```{r, rnaseq-boxplots}
-methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
-plotCovariateBoxplots(sb, alpha = 0.1, nsets = 6,
-                      methods = methods, trans = log2, maxNum = 5)
+```{r, rnaseq-lineplots}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
 ```

### 2018-04-20 datasets/RNA-Seq/RNAseq-mir200c.Rmd
Add covLinePlots to RNAseq case studies
Small changes to results print out in RNAseq case studies

@@
 date: "3/16/2018"
-output: html_document
+output: 
+    html_document:
+        toc: true
+        toc_float: true
+        highlight: tango
+        number_sections: true
 ---
@@
 } else {
-    bd <- initializeBenchDesign(nmids = 10000)
+    bd <- initializeBenchDesign()
     sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
@@
 sb <- addDefaultMetrics(sb)
-estimatePerformanceMetrics(sb, tidy = TRUE) %>%
-    filter(performanceMetric == "rejections", alpha == 0.1) %>%
+estimatePerformanceMetrics(sb, alpha = 0.05, tidy = TRUE) %>%
+    filter(performanceMetric == "rejections") %>%
     select(blabel, performanceMetric, alpha, value) %>%
-    arrange(desc(value))
+    mutate(n = nrow(sb), prop = round(value / n, 3)) %>%
+    arrange(desc(value)) %>%
+    as_tibble() %>%
+    print(n = 40)
 ```
@@
 ```{r}
-methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf", "scott-empirical")
-plotCovariateBoxplots(sb, alpha = 0.1, nsets = 6,
-                      methods = methods, trans = log2, maxNum = 5)
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
 ```

### 2018-04-20 datasets/simulations/simulations-informative-cubic.Rmd
Update sim params w/ larger signal

@@
 
-For comparison, we also include the ASH q-values.
-
-```{r}
-bd <- addBMethod(bd, "ashq", ashr::ash, ashr::get_qvalue,
-                 betahat = effect_size, sebetahat = SE)
-```
-
 As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
@@
 ```{r gauss-parameters}
-tstat <- rnorm_generator(3)       # functional: dist of alternative test stats
+tstat <- rnorm_generator(4)       # functional: dist of alternative test stats
 tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
@@
 ```{r chisq4-parameters}
-tstat <- rnorm_generator(12)       # functional: dist of alternative test stats
+tstat <- rnorm_generator(15)       # functional: dist of alternative test stats
 tstat_dist <- rchisq_perturber(4)  # functional: sampling dist/noise for test stats

### 2018-04-20 datasets/simulations/simulations-varyingeffectsize.Rmd
Update sim params w/ larger signal

@@
 pi0 <- pi0_sine(0.90)               # numeric: proportion of null hypotheses
-tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
-null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
+tstat_dist <- rnorm_perturber(1/2)  # functional: sampling dist/noise for test stats
+null_dist <- rnorm_2pvaluer(1/2)    # functional: dist to calc p-values
 icovariate <- runif                 # functional: independent covariate

### 2018-04-20 datasets/simulations/simulations-varyingpi0.Rmd
Update sim params w/ larger signal

@@
 m <- 20000                        # integer: number of hypothesis tests
-tstat <- rnorm_generator(3)       # functional: dist of alternative test stats
+tstat <- rnorm_generator(4)       # functional: dist of alternative test stats
 tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats

### 2018-04-23 datasets/microbiome/microbiome-baxter.Rmd
Remove nmids and change ashs to ashq

@@
 First, we'll create an object of `BenchDesign` class to hold the data and 
-add the benchmark methods to the `BenchDesign` object. We'll do this for a few
-values of nmids, to explore its effects on the outputs. We remove ASH from
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
 the default comparison.
@@
 if (!file.exists(otu_bench_file)) {
-    nmids <- 150
-    bd <- initializeBenchDesign(nmids = nmids)
-    bd <- dropBMethod(bd, "ashs")
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
     sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
@@
                    "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
-    metadata(sb)$nmids <- nmids
     saveRDS(sb, file = otu_bench_file)
@@
 ```{r, eval=FALSE}
-assays(sb)[["qvalue"]][, "ashs"] %>% max()
+assays(sb)[["qvalue"]][, "ashq"] %>% max()
 sum(is.na(assays(sb)[["qvalue"]][, "scott-empirical"]))
@@
 if (!file.exists(otu_bench_file_log_ubiquity)) {
-    nmids <- 150
-    bd <- initializeBenchDesign(nmids = nmids)
-    bd <- dropBMethod(bd, "ashs")
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
     sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
@@
                    "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
-    metadata(sb)$nmids <- nmids
     saveRDS(sb, file = otu_bench_file_log_ubiquity)
@@
 First, we'll create an object of `BenchDesign` class to hold the data and 
-add the benchmark methods to the `BenchDesign` object. We'll do this for a few
-values of nmids, to explore its effects on the outputs. We remove ASH from
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
 the comparison.
@@
 if (!file.exists(gns_bench_file)) {
-    nmids <- 150
-    bd <- initializeBenchDesign(nmids = nmids)
-    bd <- dropBMethod(bd, "ashs")
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
     sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
@@
                    "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
-    metadata(sb)$nmids <- nmids
     saveRDS(sb, file = gns_bench_file)
@@
 if (!file.exists(gns_bench_file_log_ubiquity)) {
-    nmids <- 150
-    bd <- initializeBenchDesign(nmids = nmids)
-    bd <- dropBMethod(bd, "ashs")
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
     sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
@@
                    "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
-    metadata(sb)$nmids <- nmids
     saveRDS(sb, file = gns_bench_file_log_ubiquity)

### 2018-04-23 datasets/microbiome/microbiome-enigma.Rmd
Remove nmids and change ashs to ashq

@@
 
-```{r enigma-readdata}
+```{r, enigma-readdata}
 ## load OTU table and metadata
+print(file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"))
 otu <- read.table(file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"), 
@@
 if (!file.exists(ph_bench_file)) {
-    nmids <- 150
-    bd <- initializeBenchDesign(nmids = nmids)
-    bd <- dropBMethod(bd, "ashs")
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
     sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
-    metadata(sb)$nmids <- nmids
     saveRDS(sb, file = ph_bench_file)
@@
 if (!file.exists(so4_bench_file)) {
-    nmids <- 150
-    bd <- initializeBenchDesign(nmids = nmids)
-    bd <- dropBMethod(bd, "ashs")
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
     sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
-    metadata(sb)$nmids <- nmids
     saveRDS(sb, file = so4_bench_file)
@@
 if (!file.exists(al_bench_file)) {
-    nmids <- 150
-    bd <- initializeBenchDesign(nmids = nmids)
-    bd <- dropBMethod(bd, "ashs")
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
     sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
-    metadata(sb)$nmids <- nmids
     saveRDS(sb, file = al_bench_file)

### 2018-04-23 datasets/microbiome/microbiome-goodrich.Rmd
Remove nmids and change ashs to ashq

@@
                    "https://zenodo.org/record/840333/files/ob_goodrich_results.tar.gz"
-    metadata(sb)$nmids <- nmids
     saveRDS(sb, file = otu_bench_file)
@@
 First, we'll create an object of `BenchDesign` class to hold the data and 
-add the benchmark methods to the `BenchDesign` object. We'll do this for a few
-values of nmids, to explore its effects on the outputs.
+add the benchmark methods to the `BenchDesign` object.
 
@@
                    "https://zenodo.org/record/840333/files/ob_goodrich_results.tar.gz"
-    metadata(sb)$nmids <- nmids
     saveRDS(sb, file = gns_bench_file)

### 2018-04-23 datasets/microbiome/microbiome-papa.Rmd
Remove nmids and change ashs to ashq

@@
                    "https://zenodo.org/record/840333/files/ibd_alm_results.tar.gz"
-    metadata(sb)$nmids <- nmids
     saveRDS(sb, file = otu_bench_file)
@@
                    "https://zenodo.org/record/840333/files/ibd_alm_results.tar.gz"
-    metadata(sb)$nmids <- nmids
     saveRDS(sb, file = gns_bench_file)
@@
 if (!file.exists(gns_bench_file_log_ubi)) {
-    nmids <- 150
-    bd <- initializeBenchDesign(nmids = nmids)
-    bd <- dropBMethod(bd, "ashs")
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
     sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
@@
                    "https://zenodo.org/record/840333/files/ibd_alm_results.tar.gz"
-    metadata(sb)$nmids <- nmids
     saveRDS(sb, file = gns_bench_file_log_ubi)

### 2018-04-27 manuscript_figures/mansuscript_figures.Rmd
started Rmd to create manuscript figures

---
title: "Manuscript Figures for Genome Biology"
author: "Rafalab"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Figure 1

This is a schema of Inputs, Assumptions and Outputs for the various FDR methods. Claire is leading this charge.

# Figure 2

This is a faceted figure with results using the yeast data with 48 biological replicates analyzed in this [publication](https://www.ncbi.nlm.nih.gov/pubmed/26206307/). A complete analysis 
can be found in the `/datasets/RNA-Seq/yeast-simulation.Rmd`. This is the 5v5 comparison.

```{r Figure2-load-data}
# Load packages and source benchmark FDR
library(data.table)
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(magrittr)
library(cowplot)
library(purrr)
library(DESeq2)
library(tibble)
library(ggthemes)
library(R.utils)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

# set up data / results directories
resdir <- "yeast-results"
```

```{r Figure2}
rseed <- 198
sampleSize <- 5
nDE <- 500
resfile <- file.path(resdir, "de5.rds")
de5 <- readRDS(file=resfile)
res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))

# Figure 2A: Not all methods control FDR at nominal alpha level
p1 <- plotsim_average(res5d, met="FDR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE)

# Figure 2B: TPR of methods across nominal alpha level
p2 <- plotsim_average(res5d, met="TPR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE)

# Figure 2C: Total number of discoveries varied across method
p3 <- plotsim_average(res5d, met="rejections",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE)

# Figure 2D: Some methods are tailored to detecting smaller effect sizes
p4 <- covariateHeatmap(de5, alpha=0.05, covname="log2FC", trans="log1p")

# maybe 2E? 
# p5 <- aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE)

# Figure 2 (ALL)
plot_grid(p1, p2, p3, p4, ncol = 2, 
          labels = LETTERS[1:4], label_size = 25)
```

# Session information

```{r}
sessionInfo()
```

### 2018-04-30 datasets/R/effectSizePlot.R
Standardize plot appearance, add facet option to simulation evaluation
Standardize line appearance in covariate line plots and simulation
evaluations. In addition, add option to simulation evaluation plots
to facet by methods that use multiple sources of information or
only one source of information (i.e. the p-value).

@@ -163,7 +163,7 @@ covariateLinePlot <- function(sbl, alpha=0.05, nbins = 25,
   
   if (linePlot){
     p <- ggplot(df, aes(x = bin/nbins, y = nsig, color = Method)) +
-      geom_line(alpha = 3/4, aes(linetype=Method)) +
+      geom_line(alpha = 0.85, aes(linetype=Method)) +
       ylab("Mean % Significant") +
       scale_x_continuous(labels = scales::percent) +
       xlab(paste0(covname, " percentile")) +
@@ -174,7 +174,7 @@ covariateLinePlot <- function(sbl, alpha=0.05, nbins = 25,
       p <- p + ylab("% Significant") 
     }else{
       p <- p + geom_errorbar(aes(ymin = nsig - se, ymax = nsig + se),
-                    width=0.01, alpha=1/3)
+                    width=0.02, alpha=0.5)
     }
     
     if(!is.null(trans)){

### 2018-04-30 datasets/R/plotsim_average.R

@@ -21,6 +21,11 @@
 #'        'bh', 'bl', 'ihw', 'lfdr', 'qvalue', 'scott-empirical',
 #'        'scott-theoretical'. (default = FALSE)
 #' @param errorBars logical indicating whether to include error bars 
+#' @param palette data.frame containing the color palette - should contain 
+#'        three columns: 'Method', 'col', and 'lty'
+#' @param facetMethodType logical indicating whether to facet the plot into 
+#'        two panels - methods that use only one piece of information, and 
+#'        those that use more than just the p-value (default = FALSE)
 #'
 #' @return
 #' a ggplot object.
@@ -28,7 +33,7 @@
 #' @author Patrick Kimes & Keegan Korthauer
 plotsim_average <- function(tsb, met, filter_set = NULL, merge_ihw = TRUE,
                             clean_names = FALSE, errorBars=FALSE,
-                            palette = candycols) {
+                            palette = candycols, facetMethodType = FALSE) {
 
     ## cacluate mean per replication
     tsba <- tsb %>%
@@ -79,10 +84,14 @@ plotsim_average <- function(tsb, met, filter_set = NULL, merge_ihw = TRUE,
     lty <- as.character(tsba$lty)
     names(lty) <- as.character(tsba$Method)
     
+    # add method type
+    tsba <- tsba %>%
+      mutate(Type = ifelse(Method %in% c("unadjusted", "bonf", "bh", "qvalue"),
+                           "Univariate (p-value only)", "Multivariate"))
+    
     gp <- tsba %>%
         ggplot(aes(x = alpha, y = value, color = Method)) +
-        geom_line(alpha = 3/4, aes(linetype=Method)) +
-        geom_point(alpha = 1) +
+        geom_line(alpha = 0.85, aes(linetype=Method)) +
         theme_classic() +
         theme(axis.title = element_text(face="bold"),
               plot.title = element_text(face="bold")) +
@@ -92,9 +101,14 @@ plotsim_average <- function(tsb, met, filter_set = NULL, merge_ihw = TRUE,
         ggtitle(paste0("Mean ", met, " Over ", max(tsba$n), " Replications")) +
         scale_color_manual(values = col) +
         scale_linetype_manual(values = lty) 
+    
+    if(facetMethodType){
+      gp <- gp +
+        facet_wrap( ~ Type)
+    }
         
[…]

### 2018-04-30 manuscript_figures/mansuscript_figures.Rmd
Format legends for Fig2; add Fig2

@@
 resdir <- "yeast-results"
+outdir <- "yeast-figs"
+dir.create(outdir, showWarnings = FALSE)
 ```
@@
 ```{r Figure2}
-rseed <- 198
-sampleSize <- 5
-nDE <- 500
+excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
 resfile <- file.path(resdir, "de5.rds")
@@
 p1 <- plotsim_average(res5d, met="FDR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
+                merge_ihw = TRUE, errorBars=TRUE) +
+      theme(legend.position="none")
 
@@
 p2 <- plotsim_average(res5d, met="TPR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
+                merge_ihw = TRUE, errorBars=TRUE) +
+      theme(legend.position="none")
 
@@
 p3 <- plotsim_average(res5d, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
+                merge_ihw = TRUE, errorBars=TRUE) +
+      theme(legend.position="none")
 
 # Figure 2D: Some methods are tailored to detecting smaller effect sizes
-p4 <- covariateHeatmap(de5, alpha=0.05, covname="log2FC", trans="log1p")
+p4 <- covariateLinePlot(de5, alpha=0.05, covname="log2FC", trans="log1p",
+                        nbins=25)
+leg <- get_legend(p4)
+p4 <- p4 + theme(legend.position="none")
 
-# maybe 2E? 
-# p5 <- aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE)
+# maybe 2E? - doesn't return a ggplot object :( 
+# fails to convert using ggplotify package
+aggupset(de5, alpha=0.05, supplementary = FALSE, 
+                       return_list = FALSE)
 
 # Figure 2 (ALL)
-plot_grid(p1, p2, p3, p4, ncol = 2, 
-          labels = LETTERS[1:4], label_size = 25)
+Fig2 <- plot_grid(p1, p2, p3, p4, ncol = 2, 
+          labels = LETTERS[1:4], label_size = 20)
+Fig2 <- plot_grid(Fig2, leg, ncol = 2, rel_widths = c(1, 0.2))
+#Fig2 <- plot_grid(Fig2, p5, ncol=1, rel_heights = c(1, 0.5),
+#                 labels = c("", "E"), label_size = 20)
+Fig2
+ggsave(file.path(outdir, "Figure2.pdf"), width=10, height=8)
+```
+
+# Figure 2
+
+This is a side-by-side figure displaying the TPR by alpha cutoff, faceted by 
+whether the method uses multiple pieces of information or just one (the p-value).
+This figure uses the same data as Figure 1.
+
+```{r, Figure3}
+# Figure 3: TPR of methods using by pieces of information used 
+Fig3 <- plotsim_average(res5d, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE, facetMethodType = TRUE)
+
+Fig3
+ggsave(file.path(outdir, "Figure3.pdf"), width=8, height=4)
+```
+
+# Figure 4
+
+This is a heatmap of methods by case studies with color representing the ranking
+by of number of rejections @ FDR cutoff 0.05.
+
+```{r, Figure 4}
+# TO DO
+```
+
+# Figure 5
+
+This is a multi-panel Figure displaying examples of covariate line plots in the
+case studies.
+
+```{r, Figure 5}
+# TO DO
+```
+
+# Figure 6
+
+This is a multi-panel Figure displaying examples of covariate vs p-value scatterplots
+and/or histograms in the case studies. 
+It will be used to discuss how we evaluated the use of the 
+covariates as independent and informative in the case studies.
+
+```{r, Figure 6}
+# TO DO
 ```

### 2018-04-30 manuscript_figures/mansuscript_figures.Rmd
Add Figure 4 prototype

@@
 This is a heatmap of methods by case studies with color representing the ranking
-by of number of rejections @ FDR cutoff 0.05.
+by of number of rejections @ FDR cutoff 0.10.
 
 ```{r, Figure 4}
-# TO DO
+library(SummarizedBenchmark)
+library(stringr)
+library(viridis)
+
+## load helper functions
+for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
+    source(f)
+}
+
+outdir <- "casestudy-figs"
+dir.create(outdir, showWarnings = FALSE)
+
+# assumes sb objects are in the following location, which contains subfolders
+# for each casestudy (if this isn't true, then parsing the case study and 
+# dataset names later on will be incorrect)
+path <- "../../sb-objects/"
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+alpha <- 0.10
+
+# remove simulation objects
+sim <- unique(c(which(grepl("simulations", objects)),
+               which(grepl("yeast", objects)),
+               which(grepl("polyester", objects))))
+
+if (length(sim) > 0){
+  objects <- objects[-sim]
+}
+
+# create tidy data frame where each row is a method / dataset observation
+# of a rank 
+ranks <- data.frame()
+obj.names <- str_split(objects, pattern = "/")
+casestudy <- unlist(lapply(obj.names, function(x) x[5] ))
+dataset <- unlist(lapply(obj.names, function(x) x[6] ))
+
+for (i in seq_along(objects)){
+  x <- readRDS(objects[i])
+  assayNames(x) <- "qvalue"
+  x <- addDefaultMetrics( x )
+
+  hasResults <- apply(!is.na( assays(x)[["qvalue"]] ), 2, all )
+
+  tmp <- estimatePerformanceMetrics(x, alpha, tidy=TRUE) %>%
+    filter( performanceMetric == "rejections") %>%
+    rename( method = blabel) %>%
+    filter( is.na(param.alpha) | (param.alpha == alpha)) %>%
+    filter( is.na(param.smooth.df) | (param.smooth.df == "3L")) %>%
+    filter( !method == "unadjusted") %>%
+    select( method, value ) %>%
+    rename( nrejects = value) %>%
+    mutate( method = gsub("-df03", "", method)) %>%
+    mutate( method = gsub("-a10", "", method)) %>%
+    na.omit() %>%
+    mutate( rank = rank(nrejects) / n() ) %>%
+    mutate(casestudy = casestudy[i],
+           dataset = dataset[i])
+
+  ranks <- rbind(ranks, tmp)
+}
+
+# average over datasets within case study
+ranks_avg <- ranks %>% 
+  group_by( casestudy, method) %>%
+  summarize( mean_rank = mean(rank, na.rm = TRUE))
+
+# heatmap : rows method, cols casestudy
+Fig4 <- ggplot(ranks_avg, aes(x = casestudy, y = method, fill = mean_rank)) + 
+  geom_raster() + 
+  scale_fill_viridis() +
+  theme_bw() +
+  xlab("Case Study") +
+  ylab("Method") +
+  labs(fill = "Mean rank")
+
+Fig4
+ggsave(file.path(outdir, "Figure4.pdf"), width=6, height=4, dev = pdf)
 ```

### 2018-05-01 manuscript_figures/manuscript_figures.Rmd
Updates to manuscript plotting
Reorder the axes in Figure 4, add submit a submit script, standardize file
locations relative to cwd, add upset plot (Figure 2E) as a panel in Figure
2, add option to only plot a subset of intersections in aggupset plotter.

--- 2018-04-30 manuscript_figures/mansuscript_figures.Rmd
@@
 # Load packages and source benchmark FDR
-library(data.table)
 library(tidyr)
 library(dplyr)
-library(readr)
 library(ggplot2)
@@
 library(cowplot)
-library(purrr)
-library(DESeq2)
 library(tibble)
 library(ggthemes)
-library(R.utils)
+library(grid)
+library(SummarizedBenchmark)
 
 ## load helper functions
-for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
+for (f in list.files("../datasets/R", "\\.(r|R)$", full.names = TRUE)) {
     source(f)
@@
 
-# set up data / results directories
-resdir <- "yeast-results"
-outdir <- "yeast-figs"
+# assumes sb objects are in the following location, which contains subfolders
+# for each casestudy (if this isn't true, then parsing the case study and 
+# dataset names later on will be incorrect)
+path <- "../sb-objects"
+
+# set up results directorie
+outdir <- "./figures"
 dir.create(outdir, showWarnings = FALSE)
+
+# point to where yeast results are (within 'path' defined above)
+resdir <- file.path(path, "RNAseq/yeast-results")
+
+# set alpha cutoff for plots that are fixed at a certain value of alpha
+alpha <- 0.10
 ```
@@
 
-# maybe 2E? - doesn't return a ggplot object :( 
-# fails to convert using ggplotify package
-aggupset(de5, alpha=0.05, supplementary = FALSE, 
-                       return_list = FALSE)
+# Figure 2E - upsetr plot -- this solution currently chops off left margin hist
+aggupset(de5, alpha=alpha, supplementary = FALSE, 
+                       return_list = FALSE, nintersects = 12)
+grid.text("Mean overlap Over 100 Replications", x = 0.6, y = 0.97, 
+          gp = gpar(fontsize = 16, fontface = "bold")) 
+p5 <- grid.grab()
 
@@
 Fig2 <- plot_grid(Fig2, leg, ncol = 2, rel_widths = c(1, 0.2))
-#Fig2 <- plot_grid(Fig2, p5, ncol=1, rel_heights = c(1, 0.5),
-#                 labels = c("", "E"), label_size = 20)
+
+Fig2 <- plot_grid(Fig2, p5, ncol=1, rel_heights = c(1, 0.5),
+               labels = c("", "E"), label_size = 20)
+
 Fig2
-ggsave(file.path(outdir, "Figure2.pdf"), width=10, height=8)
+ggsave(file.path(outdir, "Figure2.pdf"), width=10, height=12)
 ```
@@
 This is a heatmap of methods by case studies with color representing the ranking
-by of number of rejections @ FDR cutoff 0.05.
+by of number of rejections @ FDR cutoff 0.10.
 
 ```{r, Figure 4}
-# TO DO
+library(SummarizedBenchmark)
+library(stringr)
+library(viridis)
+
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+
+# remove simulation objects
+sim <- unique(c(which(grepl("simulations", objects)),
+               which(grepl("yeast", objects)),
+               which(grepl("polyester", objects))))
+
+if (length(sim) > 0){
+  objects <- objects[-sim]
+}
+
+# create tidy data frame where each row is a method / dataset observation
+# of a rank 
+ranks <- data.frame()
+obj.names <- str_split(objects, pattern = "/")
+casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
+dataset <- unlist(lapply(obj.names, function(x) x[4] ))
+
+for (i in seq_along(objects)){
+  x <- readRDS(objects[i])
+  assayNames(x) <- "qvalue"
+  x <- addDefaultMetrics( x )
+
+  hasResults <- apply(!is.na( assays(x)[["qvalue"]] ), 2, all )
+
+  if (sum(hasResults) > 0){
+    tmp <- estimatePerformanceMetrics(x, alpha, tidy=TRUE) %>%
+      filter( performanceMetric == "rejections") %>%
+      dplyr::rename( method = blabel) %>%
+      filter( is.na(param.alpha) | (param.alpha == alpha)) %>%
+      filter( is.na(param.smooth.df) | (param.smooth.df == "3L")) %>%
+      filter( !method == "unadjusted") %>%
+      select( method, value ) %>%
+      dplyr::rename( nrejects = value) %>%
+      mutate( method = gsub("-df03", "", method)) %>%
+      mutate( method = gsub("-a10", "", method)) %>%
+      na.omit() %>%
+      mutate( rank = rank(nrejects) / n() ) %>%
+      mutate(casestudy = casestudy[i],
+             dataset = dataset[i])
+
+    ranks <- rbind(ranks, tmp)
+  }
+}
+
+# average over datasets within case study
+ranks_avg <- ranks %>% 
+  group_by( casestudy, method) %>%
+  summarize( mean_rank = mean(rank, na.rm = TRUE))
+
+case_dat <- ranks_avg %>%
+  group_by( casestudy) %>%
+  summarize( nmethods = n())
+
+method_dat <- ranks_avg %>%
+  group_by( method) %>%
+  summarize( meta_rank = mean(mean_rank))
+
+ranks_avg <- left_join(ranks_avg, method_dat, by ="method")
+ranks_avg <- left_join(ranks_avg, case_dat, by ="casestudy")
+
+# reorder rows and columns
+ranks_avg$casestudy <- factor(ranks_avg$casestudy, 
+                              levels=unique(ranks_avg$casestudy[order(ranks_avg$nmethods)]))
+ranks_avg$method <- factor(ranks_avg$method,
+                           levels=unique(ranks_avg$method[order(ranks_avg$meta_rank)]))
+
+# heatmap : rows method, cols casestudy
+Fig4 <- ggplot(ranks_avg, aes(x = casestudy, y = method, fill = mean_rank)) + 
+  geom_raster() + 
+  scale_fill_viridis() +
+  theme_bw() +
+  xlab("Case Study") +
+  ylab("Method") +
+  labs(fill = "Mean rank")
+
+Fig4
+ggsave(file.path(outdir, "Figure4.pdf"), width=6, height=4, dev = pdf)
 ```

### 2018-05-01 datasets/R/plotsim_aggupset.R

@@ -13,6 +13,8 @@
 #'        of upset plot. The returned list can be used to generate the upset plot
 #'        using `upset(fromExpression(freq_list)`. This can be useful if the user
 #'        wants to experiment with upset plot styles. (default = FALSE)
+#' @param nintersects scalar value representing number of sets to look at.  
+#'        Default is 40 (same as default in UpSetR package).
 #' 
 #' @return
 #' an upset plot if `return_list = FALSE` (default), else a list of frequencies that
@@ -26,7 +28,8 @@
 #' 
 #' @import dplyr magrittr
 #' @author Patrick Kimes
-aggupset <- function(res, alpha, supplementary = FALSE, return_list = FALSE) { 
+aggupset <- function(res, alpha, supplementary = FALSE, return_list = FALSE,
+                     nintersects = 40) { 
 
     ## find significant hits at alpha cutoff for all replicates
     hits_tabs <- lapply(res, sb2hits, a = alpha, s = supplementary)
@@ -90,6 +93,7 @@ aggupset <- function(res, alpha, supplementary = FALSE, return_list = FALSE) {
     ## draw upset plot if frequency list not returned
     upset(fromExpression(freq_list),
           nsets = n_cols,
+          nintersects = nintersects,
           mb.ratio = c(0.55, 0.45),
           order.by = "freq",
           decreasing = TRUE,

### 2018-05-01 manuscript_figures/manuscript_figures.sh

@@ -1,0 +1,18 @@
+#!/bin/bash
+#SBATCH -J figs
+#SBATCH -n 1
+#SBATCH -N 1
+#SBATCH -p shared
+#SBATCH --mem 50G
+#SBATCH -t 0-1:00
+#SBATCH -o render-%j.out
+#SBATCH -e render-%j.err
+
+module load pandoc
+export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/pandoc/2.0.2-fasrc01/bin/pandoc/"
+
+# change filename to Rmd to be knitted
+# make sure number of cores used in Rmd matches sbatch param -n
+R -e "rmarkdown::render('manuscript_figures.Rmd', clean=TRUE)"
+
+

### 2018-05-02 manuscript_figures/manuscript_figures.Rmd
Suggestion for Fig4

@@
 ranks_avg$method <- factor(ranks_avg$method,
-                           levels=unique(ranks_avg$method[order(ranks_avg$meta_rank)]))
+                           levels=rev(unique(ranks_avg$method[order(ranks_avg$meta_rank)])))
 
@@
 ggsave(file.path(outdir, "Figure4.pdf"), width=6, height=4, dev = pdf)
+
+##
+## AR: alternative suggestion to heatmap, feel free to discard! ##
+##
+ranks_avg$method <- factor(ranks_avg$method,
+                           levels=rev(unique(ranks_avg$method[order(ranks_avg$meta_rank)])))
+Fig4b <- ggplot( ranks_avg, aes(x=method, y=mean_rank, group=casestudy, col=casestudy) ) +
+  geom_line() + geom_point() +
+  labs(y="Mean rank", x="", col="Case study") +
+  theme(axis.text.x=element_text(angle=25, vjust=1, hjust=1))
+Fig4b
 ```

### 2018-05-02 manuscript_figures/manuscript_figures.Rmd
Add one-color version of Fig4 heatmap

--- 2018-05-01 manuscript_figures/manuscript_figures.Rmd
@@
 
-# Figure 2
+# Figure 3
 
@@
 Fig4
-ggsave(file.path(outdir, "Figure4.pdf"), width=6, height=4, dev = pdf)
+ggsave(file.path(outdir, "Figure4.pdf"), width=6, height=4)
+
+# heatmap : rows method, cols casestudy- one color scale 
+Fig4alt <- Fig4 +
+  scale_fill_gradient(low = "white", high = "red", na.value=NA) +
+  theme(panel.background = element_rect(colour="grey60", fill="grey60",
+                                        size = 0.5),
+        panel.grid.major = element_line(colour = "grey60"), 
+        panel.grid.minor = element_line(colour = "grey60"))
+
+Fig4alt
+ggsave(file.path(outdir, "Figure4alt.pdf"), width=6, height=4)
+
+##
+## AR: alternative suggestion to heatmap, feel free to discard! ##
+##
+ranks_avg$method <- factor(ranks_avg$method,
+                           levels=rev(unique(ranks_avg$method[order(ranks_avg$meta_rank)])))
+Fig4b <- ggplot( ranks_avg, aes(x=method, y=mean_rank, group=casestudy, col=casestudy) ) +
+  geom_line() + geom_point() +
+  labs(y="Mean rank", x="", col="Case study") +
+  theme(axis.text.x=element_text(angle=25, vjust=1, hjust=1))
+Fig4b
 ```

### 2018-05-10 datasets/GSEA/gsea-human.Rmd
Save GSEA, GWAS, and scRNAseq results to common dir

@@
 resdir <- "results"
+sbdir <- "../../results/GSEA"
 dir.create(datdir, showWarnings = FALSE)
 dir.create(resdir, showWarnings = FALSE)
+dir.create(sbdir, showWarnings = FALSE)
 
@@
 result_file <- file.path(resdir, "human-results.rds")
-bench_file <- file.path(resdir, "human-benchmark.rds")
+bench_file <- file.path(sbdir, "human-benchmark.rds")
 ```
@@
 bd <- initializeBenchDesign()
-bd <- dropBMethod(bd, "ashs")
-bd <- dropBMethod(bd, "scott-empirical")
-bd <- dropBMethod(bd, "scott-theoretical")
 ```
@@
 ```{r}
-covariateLinePlot(sGSEA, alpha = 0.1, covname = "ind_covariate")
+covariateLinePlot(sGSEA, alpha = 0.1, covname = "ind_covariate", trans = "log1p")
 ```

### 2018-05-10 datasets/GSEA/gsea-mouse.Rmd
Save GSEA, GWAS, and scRNAseq results to common dir

@@
 result_file <- file.path(resdir, "mouse-results.rds")
-bench_file <- file.path(resdir, "mouse-benchmark.rds")
+bench_file <- file.path(sbdir, "mouse-benchmark.rds")
 ```
@@
 if (!file.exists(goset_file)) {
-    mart <- useMart("ensembl", "mmusculus_gene_ ensembl")
+    mart <- useMart("ensembl", "mmusculus_gene_ensembl")
     goSets <- getBM(c("ensembl_gene_id", "go_id"), mart = mart,
@@
 ```{r}
-covariateLinePlot(sGSEAMouse, alpha = 0.1, covname = "ind_covariate")
+covariateLinePlot(sGSEAMouse, alpha = 0.1, covname = "ind_covariate", trans="log1p")
 ```

### 2018-05-10 datasets/GWAS/GWAS.Rmd
Save GSEA, GWAS, and scRNAseq results to common dir

@@
 resdir <- "results"
+sbdir <- "../../results/GWAS"
 dir.create(datdir, showWarnings = FALSE)
 dir.create(resdir, showWarnings = FALSE)
+dir.create(sbdir, showWarnings = FALSE)
+
+# results files
+resfile_N <- file.path(sbdir, paste0("bmi-samplesize-benchmark.rds"))
+resfile_AF <- file.path(sbdir, paste0("bmi-maf-benchmark.rds"))
 
@@
 ```{r, GWAS1-sb, results="hide", message=FALSE}
-resN <- file.path(resdir, paste0("bmi_N_summarizedBenchmark_", nrow(bmi), ".rds"))
-
 duration <- NA
-if (!file.exists(resN)) {
+if (!file.exists(resfile_N)) {
     t1 <- proc.time()
@@
                     "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
-    saveRDS(sbN, file=resN)
+    saveRDS(sbN, file = resfile_N)
     duration <- round((proc.time()-t1)[3]/60,1)
 } else {
-    sbN <- readRDS(resN)
+    sbN <- readRDS(resfile_N)
 }
@@
 ```{r, GWAS1-sb2, results="hide", message=FALSE}
-resAF <- file.path(resdir, paste0("bmi_AF_summarizedBenchmark_", nrow(bmi), ".rds"))
-
 duration <- NA
-if (!file.exists(resAF)) {
+if (!file.exists(resfile_AF)) {
     t1 <- proc.time()
@@
                      "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
-    saveRDS(sbAF, file = resAF)
+    saveRDS(sbAF, file = resfile_AF)
     duration <- round((proc.time()-t1)[3]/60, 1)
 } else {
-    sbAF <- readRDS(resAF)
+    sbAF <- readRDS(resfile_AF)
 }

### 2018-05-10 datasets/scrnaseq/scrnaseq-human-mast.Rmd
Save GSEA, GWAS, and scRNAseq results to common dir

@@
 resdir <- "results"
+sbdir <- "../../results/scRNAseq"
 dir.create(datdir, showWarnings = FALSE)
 dir.create(resdir, showWarnings = FALSE)
+dir.create(sbdir, showWarnings = FALSE)
+
+# results files
+resfile_mean <- file.path(sbdir, "human-benchmark-mast-mean.rds")
+resfile_det <- file.path(sbdir, "human-benchmark-mast-det.rds")
 
@@
 ```{r, sb1, results="hide", message=FALSE}
-res1 <- file.path(resdir, "sb_GSE84465_mast_mean.rds")
-
-if (!file.exists(res1)){
+if (!file.exists(resfile_mean)){
   sb1 <- bd %>% buildBench(data.frame(rowData(humandata)) %>% 
@@
 
-  saveRDS(sb1, file=res1)
+  saveRDS(sb1, file = resfile_mean)
 }else{
-  sb1 <- readRDS(res1)
+  sb1 <- readRDS(resfile_mean)
 }
@@
 ```{r, sb2, results="hide", message=FALSE}
-res2 <- file.path(resdir, "sb_GSE84465_mast_det.rds")
-
-if (!file.exists(res2)){
+if (!file.exists(resfile_det)){
   sb2 <- bd %>% buildBench(data.frame(rowData(humandata)) %>% 
@@
 
-  saveRDS(sb2, file=res2)
+  saveRDS(sb2, file = resfile_det)
 }else{
-  sb2 <- readRDS(res2)
+  sb2 <- readRDS(resfile_det)
 }

### 2018-05-10 datasets/scrnaseq/scrnaseq-human-scdd.Rmd
Save GSEA, GWAS, and scRNAseq results to common dir

@@
 resdir <- "results"
+sbdir <- "../../results/scRNAseq"
 dir.create(datdir, showWarnings = FALSE)
 dir.create(resdir, showWarnings = FALSE)
+dir.create(sbdir, showWarnings = FALSE)
+
+# results files
+resfile_mean <- file.path(sbdir, "human-benchmark-scdd-mean.rds")
+resfile_det <- file.path(sbdir, "human-benchmark-scdd-det.rds")
 
@@
 ```{r, sb1, results="hide", message=FALSE}
-res1 <- file.path(resdir, "sb_GSE84465_scdd_mean.rds")
-
-if (!file.exists(res1)){
+if (!file.exists(resfile_mean)){
   sb1 <- bd %>% buildBench(data.frame(rowData(humandata)) %>% 
@@
 ```{r, sb2, results="hide", message=FALSE}
-res2 <- file.path(resdir, "sb_GSE84465_scdd_det.rds")
-
-if (!file.exists(res2)){
+if (!file.exists(resfile_det)){
   sb2 <- bd %>% buildBench(data.frame(rowData(humandata)) %>% 

### 2018-05-10 datasets/scrnaseq/scrnaseq-human-wilcox.Rmd
Fix typos in Wilcox scrnaseq
Save GSEA, GWAS, and scRNAseq results to common dir

@@
 which genes have different expression among the two biological groups in each comparison.
-We also apply three different analysis methods: the non-parametric Wilcoxon Rank-sum test, Wilcoxon_Ranksum [@finak2015] and scDD [@korthauer2016]. 
+We also apply three different analysis methods: the non-parametric Wilcoxon Rank-sum test,
+MAST [@finak2015], and scDD [@korthauer2016]. 
 In this vignette, we will explore the human dataset analysed with the Wilcoxon Rank-sum test.
@@
 resdir <- "results"
+sbdir <- "../../results/scRNAseq"
 dir.create(datdir, showWarnings = FALSE)
 dir.create(resdir, showWarnings = FALSE)
+dir.create(sbdir, showWarnings = FALSE)
+
+# results files
+resfile_mean <- file.path(sbdir, "human-benchmark-wilcox-mean.rds")
+resfile_det <- file.path(sbdir, "human-benchmark-wilcox-det.rds")
 
@@
 ```{r, sb1, results="hide", message=FALSE}
-res1 <- file.path(resdir, "sb_GSE84465_Wilcoxon_Ranksum_mean.rds")
-
-if (!file.exists(res1)){
+if (!file.exists(resfile_mean)){
   sb1 <- bd %>% buildBench(data.frame(rowData(humandata)) %>% 
@@
 ```{r, sb2, results="hide", message=FALSE}
-res2 <- file.path(resdir, "sb_GSE84465_Wilcoxon_Ranksum_det.rds")
-
-if (!file.exists(res2)){
+if (!file.exists(resfile_det)){
   sb2 <- bd %>% buildBench(data.frame(rowData(humandata)) %>% 

### 2018-05-10 datasets/scrnaseq/scrnaseq-mouse-mast.Rmd
Save GSEA, GWAS, and scRNAseq results to common dir

@@
 resdir <- "results"
+sbdir <- "../../results/scRNAseq"
 dir.create(datdir, showWarnings = FALSE)
 dir.create(resdir, showWarnings = FALSE)
+dir.create(sbdir, showWarnings = FALSE)
+
+# results files
+resfile_mean <- file.path(sbdir, "mouse-benchmark-mast-mean.rds")
+resfile_det <- file.path(sbdir, "mouse-benchmark-mast-det.rds")
 
@@
 ```{r, sb1, results="hide", message=FALSE}
-res1 <- file.path(resdir, "sb_GSE94383_mast_mean.rds")
-
-if (!file.exists(res1)){
+if (!file.exists(resfile_mean)){
   sb1 <- bd %>% buildBench(data.frame(rowData(mousedata)) %>% 
@@
 ```{r, sb2, results="hide", message=FALSE}
-res2 <- file.path(resdir, "sb_GSE94383_mast_det.rds")
-
-if (!file.exists(res2)){
+if (!file.exists(resfile_det)){
   sb2 <- bd %>% buildBench(data.frame(rowData(mousedata)) %>% 

### 2018-05-10 datasets/scrnaseq/scrnaseq-mouse-scdd.Rmd
Save GSEA, GWAS, and scRNAseq results to common dir

@@
 resdir <- "results"
+sbdir <- "../../results/scRNAseq"
 dir.create(datdir, showWarnings = FALSE)
 dir.create(resdir, showWarnings = FALSE)
+dir.create(sbdir, showWarnings = FALSE)
+
+# results files
+resfile_mean <- file.path(sbdir, "mouse-benchmark-scdd-mean.rds")
+resfile_det <- file.path(sbdir, "mouse-benchmark-scdd-det.rds")
 
@@
 ```{r, sb1, results="hide", message=FALSE}
-res1 <- file.path(resdir, "sb_GSE94383_scdd_mean.rds")
-
-if (!file.exists(res1)){
+if (!file.exists(resfile_mean)){
   sb1 <- bd %>% buildBench(data.frame(rowData(mousedata)) %>% 
@@
 ```{r, sb2, results="hide", message=FALSE}
-res2 <- file.path(resdir, "sb_GSE94383_scdd_det.rds")
-
-if (!file.exists(res2)){
+if (!file.exists(resfile_det)){
   sb2 <- bd %>% buildBench(data.frame(rowData(mousedata)) %>% 

### 2018-05-10 datasets/scrnaseq/scrnaseq-mouse-wilcox.Rmd
Save GSEA, GWAS, and scRNAseq results to common dir

@@
 resdir <- "results"
+sbdir <- "../../results/scRNAseq"
 dir.create(datdir, showWarnings = FALSE)
 dir.create(resdir, showWarnings = FALSE)
+dir.create(sbdir, showWarnings = FALSE)
+
+# results files
+resfile_mean <- file.path(sbdir, "mouse-benchmark-wilcox-mean.rds")
+resfile_det <- file.path(sbdir, "mouse-benchmark-wilcox-det.rds")
 
@@
 ```{r, sb1, results="hide", message=FALSE}
-res1 <- file.path(resdir, "sb_GSE94383_Wilcoxon-ranksum_mean.rds")
-
-if (!file.exists(res1)){
+if (!file.exists(resfile_mean)){
   sb1 <- bd %>% buildBench(data.frame(rowData(mousedata)) %>% 
@@
 ```{r, sb2, results="hide", message=FALSE}
-res2 <- file.path(resdir, "sb_GSE94383_Wilcoxon-ranksum_det.rds")
-
-if (!file.exists(res2)){
+if (!file.exists(resfile_det)){
   sb2 <- bd %>% buildBench(data.frame(rowData(mousedata)) %>% 

### 2018-05-10 datasets/GSEA/gsea.sh

@@ -2,7 +2,7 @@
 #SBATCH -J knitGSEA
 #SBATCH -n 2
 #SBATCH -N 1
-#SBATCH -p shared
+#SBATCH -p shared,irizarry
 #SBATCH --mem 32G
 #SBATCH -t 0-6:00
 #SBATCH -o render-%j.out

### 2018-05-10 datasets/GWAS/GWAS-knit.sh

@@ -2,9 +2,9 @@
 #SBATCH -J knitGWAS                 # A single job name for the array
 #SBATCH -n 8                        # Number of cores
 #SBATCH -N 1                        # All cores on one machine
-#SBATCH -p shared                   # Partition
+#SBATCH -p irizarry,serial_requeue                   # Partition
 #SBATCH --mem 60000                 # Memory request 
-#SBATCH -t 0-2:00                   # Maximum execution time (D-HH:MM)
+#SBATCH -t 0-8:00                   # Maximum execution time (D-HH:MM)
 #SBATCH -o render-%j.out  # Standard output
 #SBATCH -e render-%j.err  # Standard error
 
@@ -13,4 +13,4 @@ export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/rstudio/0.98.1103-fasrc01/bin/pan
 
 # change filename to Rmd to be knitted. 
 # make sure ncores in Rmd matches -n batch param above
-R -e "rmarkdown::render('GWAS.Rmd', clean = FALSE)"
+R -e "rmarkdown::render('GWAS.Rmd')"

### 2018-05-10 datasets/scrnaseq/scrnaseq-knit.sh

@@ -2,9 +2,9 @@
 #SBATCH -J knit-sc
 #SBATCH -n 8
 #SBATCH -N 1
-#SBATCH -p shared
+#SBATCH -p shared,serial_requeue
 #SBATCH --mem 125G
-#SBATCH -t 0-4:00
+#SBATCH -t 0-12:00
 #SBATCH -o render-%j.out
 #SBATCH -e render-%j.err
 
@@ -13,11 +13,11 @@ export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/pandoc/2.0.2-fasrc01/bin/pandoc/"
 
 # change filename to Rmd to be knitted
 # make sure number of cores used in Rmd matches sbatch param -n
-R -e "rmarkdown::render('scrnaseq-human-scdd.Rmd', clean=FALSE)"
-R -e "rmarkdown::render('scrnaseq-mouse-scdd.Rmd', clean=FALSE)"
-R -e "rmarkdown::render('scrnaseq-human-mast.Rmd', clean=FALSE)"
-R -e "rmarkdown::render('scrnaseq-mouse-mast.Rmd', clean=FALSE)"
-R -e "rmarkdown::render('scrnaseq-human-wilcox.Rmd', clean=FALSE)"
-R -e "rmarkdown::render('scrnaseq-mouse-wilcox.Rmd', clean=FALSE)"
+R -e "rmarkdown::render('scrnaseq-human-scdd.Rmd')"
+R -e "rmarkdown::render('scrnaseq-mouse-scdd.Rmd')"
+R -e "rmarkdown::render('scrnaseq-human-mast.Rmd')"
+R -e "rmarkdown::render('scrnaseq-mouse-mast.Rmd')"
+R -e "rmarkdown::render('scrnaseq-human-wilcox.Rmd')"
+R -e "rmarkdown::render('scrnaseq-mouse-wilcox.Rmd')"
 
 

### 2018-05-10 datasets/ChIP-seq/ChIPseq-CBP-csaw.Rmd
Save ChIPseq, RNAseq, and microbiome results to common dir
Also add covariateLinePlots to microbiome Rmds.

@@
 resdir <- "results"
+sbdir <- "../../results/ChIPseq"
 dir.create(datdir, showWarnings = FALSE)
 dir.create(resdir, showWarnings = FALSE)
+dir.create(sbdir, showWarnings = FALSE)
 
@@
 result_file <- file.path(resdir, "cbp-csaw-results.rds")
-bench_file <- file.path(resdir, "cbp-csaw-benchmark.rds")
+bench_file <- file.path(sbdir, "cbp-csaw-benchmark.rds")
 

### 2018-05-10 datasets/ChIP-seq/ChIPseq-h3k4me3-csaw.Rmd
Save ChIPseq, RNAseq, and microbiome results to common dir
Also add covariateLinePlots to microbiome Rmds.

@@
 result_file <- file.path(resdir, "h3k4me3-csaw-results.rds")
-bench_file <- file.path(resdir, "h3k4me3-csaw-benchmark.rds")
+bench_file <- file.path(sbdir, "h3k4me3-csaw-benchmark.rds")
 

### 2018-05-10 datasets/ChIP-seq/ChIPseq-h3k4me3-promoters.Rmd
Save ChIPseq, RNAseq, and microbiome results to common dir
Also add covariateLinePlots to microbiome Rmds.

@@
 result_file <- file.path(resdir, "h3k4me3-promotors-results.rds")
-bench_file <- file.path(resdir, "h3k4me3-promotors-benchmark.rds")
+bench_file <- file.path(sbdir, "h3k4me3-promotors-benchmark.rds")
 

### 2018-05-10 datasets/RNA-Seq/RNAseq-gtex.Rmd
Save ChIPseq, RNAseq, and microbiome results to common dir
Also add covariateLinePlots to microbiome Rmds.

@@
 resdir <- "results"
+sbdir <- "../../results/RNAseq"
 dir.create(datdir, showWarnings = FALSE)
 dir.create(resdir, showWarnings = FALSE)
+dir.create(sbdir, showWarnings = FALSE)
 
 ## intermediary files we create below
-count_file <- file.path(datdir, "brain-counts.rds")
+count_file <- file.path(resdir, "brain-counts.rds")
 result_file <- file.path(resdir, "brain-results.rds")
-bench_file <- file.path(resdir, "brain-benchmark.rds")
+bench_file <- file.path(sbdir, "brain-benchmark.rds")
 

### 2018-05-10 datasets/RNA-Seq/RNAseq-mir200c.Rmd
Save ChIPseq, RNAseq, and microbiome results to common dir
Also add covariateLinePlots to microbiome Rmds.

@@
 resdir <- "results"
+sbdir <- "../../results/RNAseq"
 dir.create(datdir, showWarnings = FALSE)
@@
 ## intermediary files we create below
-count_file <- file.path(datdir, "rse_gene.Rdata")
+count_file <- file.path(resdir, "rse_gene.Rdata")
 result_file <- file.path(resdir, "mir200c-results.rds")
-bench_file <- file.path(resdir, "mir200c-benchmark.rds")
+bench_file <- file.path(sbdir, "mir200c-benchmark.rds")
 

### 2018-05-10 datasets/RNA-Seq/polyester-simulation.Rmd
Save ChIPseq, RNAseq, and microbiome results to common dir
Also add covariateLinePlots to microbiome Rmds.

@@
 # Load packages needed to simulate RNA-seq counts with polyester
+library(SummarizedBenchmark)
 library(zebrafishRNASeq) 
@@
 library(polyester)
-
-# Differential testing
 library(DESeq2)
-
-# Load supporting packages for data analyses
 library(data.table)
@@
 library(ggplot2)
-library(R.utils)
 library(cowplot)
@@
 
-# benchmark methods
-library(IHW)
-library(ashr)
-library(qvalue)
-library(swfdr)
-library(fdrtool)
-# Installed BayesLogit from CRAN archive via `R CMD INSTALL BayesLogit_0.6.tar.gz`
-# install_github('jgscott/FDRreg', subdir="R_pkg/")
-library(FDRreg)
+# load helper functions
+for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
+    source(f)
+}
 
-# comparison tools/functions
-# devtools::install_github("areyesq89/SummarizedBenchmark", ref="fdrbenchmark") # needed b/c BioC devel branch is 
-library(SummarizedBenchmark)
+# set up results directories 
+resdir <- "../../results/RNAseq"
+dir.create(resdir, showWarnings = FALSE)
 
-# load helper functions
-sourceDirectory(file.path("../R/"))
+# results files that will be generated below
+resfile_n5 <- file.path(resdir, "polyester-results-null5.rds")
+resfile_n10 <- file.path(resdir, "polyester-results-null10.rds")
+resfile_d5 <- file.path(resdir, "polyester-results-de5.rds")
+resfile_d10 <- file.path(resdir, "polyester-results-de10.rds")
 
@@
 
-if (!file.exists(file.path("polyester-results-null5.RDS"))){
+if (!file.exists(resfile_n5)){
   null5 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
@@
                    BPPARAM=multicoreParam)
-  saveRDS(null5, file=file.path("polyester-results-null5.RDS"))
+  saveRDS(null5, file=resfile_n5)
 } else {
-  null5 <- readRDS(file=file.path("polyester-results-null5.RDS"))
+  null5 <- readRDS(file=resfile_n5)
 }
@@
 
-if (!file.exists(file.path("polyester-results-null10.RDS"))){
+if (!file.exists(resfile_n10)){
   null10 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
@@
                    BPPARAM=multicoreParam)
-  saveRDS(null10, file=file.path("polyester-results-null10.RDS"))
+  saveRDS(null10, file=resfile_n10)
 } else {
-  null10 <- readRDS(file=file.path("polyester-results-null10.RDS"))
+  null10 <- readRDS(file=resfile_n10)
 }
@@
 
-if (!file.exists(file.path("polyester-results-de5.RDS"))){
+if (!file.exists(resfile_d5)){
   de5 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
@@
                    BPPARAM=multicoreParam)
-  saveRDS(de5, file=file.path("polyester-results-de5.RDS"))
+  saveRDS(de5, file=resfile_d5)
 } else {
-  de5 <- readRDS(file=file.path("polyester-results-de5.RDS"))
+  de5 <- readRDS(file=resfile_d5)
 }
@@
 
-if (!file.exists(file.path("polyester-results-de10.RDS"))){
+if (!file.exists(resfile_d10)){
   de10 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
@@
                    BPPARAM=multicoreParam)
-  saveRDS(de10, file=file.path("polyester-results-de10.RDS"))
+  saveRDS(de10, file=resfile_d10)
 } else {
-  de10 <- readRDS(file=file.path("polyester-results-de10.RDS"))
+  de10 <- readRDS(file=resfile_d10)
 }

### 2018-05-10 datasets/RNA-Seq/yeast-simulation.Rmd
Save ChIPseq, RNAseq, and microbiome results to common dir
Also add covariateLinePlots to microbiome Rmds.

@@
 # Load packages and source benchmark FDR
+library(SummarizedBenchmark)
 library(data.table)
@@
 datdir <- "yeast-data"
-resdir <- "yeast-results"
+resdir <- "../../results/RNAseq"
 dir.create(datdir, showWarnings = FALSE)
 dir.create(resdir, showWarnings = FALSE)
+
+# results files that will be generated
+resfile_n5 <- file.path(resdir, "yeast-results-null5.rds")
+resfile_n10 <- file.path(resdir, "yeast-results-null10.rds")
+resfile_d5 <- file.path(resdir, "yeast-results-de5.rds")
+resfile_d10 <- file.path(resdir, "yeast-results-de10.rds")
 
@@
 nDE <- 0
-resfile <- file.path(resdir, "null5.rds")
 
-if (!file.exists(resfile)){
+if (!file.exists(resfile_n5)){
   null5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
@@
                    mc.cores=cores)
-  saveRDS(null5, file=resfile)
+  saveRDS(null5, file = resfile_n5)
 }else{
-  null5 <- readRDS(file=resfile)
+  null5 <- readRDS(file = resfile_n5)
 }
@@
 nDE <- 0
-resfile <- file.path(resdir, "null10.rds")
 
-if (!file.exists(resfile)){
+if (!file.exists(resfile_n10)){
   null10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
@@
                    mc.cores=cores)
-  saveRDS(null10, file=resfile)
+  saveRDS(null10, file = resfile_n10)
 }else{
-  null10 <- readRDS(file=resfile)
+  null10 <- readRDS(file = resfile_n10)
 }
@@
 nDE <- 500
-resfile <- file.path(resdir, "de5.rds")
 
-if (!file.exists(resfile)){
+if (!file.exists(resfile_d5)){
   de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
@@
                    mc.cores=cores)
-  saveRDS(de5, file=resfile)
+  saveRDS(de5, file = resfile_d5)
 }else{
-  de5 <- readRDS(file=resfile)
+  de5 <- readRDS(file = resfile_d5)
 }
@@
 nDE <- 500
-resfile <- file.path(resdir, "de10.rds")
 
-if(!file.exists(resfile)){
+if(!file.exists(resfile_d10)){
   de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
@@
                    mc.cores=cores)
-  saveRDS(de10, file=resfile)
+  saveRDS(de10, file = resfile_d10)
 }else{
-  de10 <- readRDS(file=resfile)
+  de10 <- readRDS(file = resfile_d10)
 }

### 2018-05-10 datasets/microbiome/microbiome-baxter.Rmd
Save ChIPseq, RNAseq, and microbiome results to common dir
Also add covariateLinePlots to microbiome Rmds.

@@
 resdir <- "results"
+sbdir <- "../../results/microbiome"
 dir.create(datdir, showWarnings = FALSE)
 dir.create(resdir, showWarnings = FALSE)
+dir.create(sbdir, showWarnings = FALSE)
 
@@
 otu_result_file <- file.path(resdir, "baxter-otus-results.rds")
-otu_bench_file <- file.path(resdir, "baxter-otus-benchmark.rds")
-otu_bench_file_log_ubiquity <- file.path(resdir, "baxter-otus-log-ubiquity-benchmark.rds")
+otu_bench_file <- file.path(sbdir, "baxter-otus-benchmark.rds")
+otu_bench_file_log_ubiquity <- file.path(sbdir, "baxter-otus-log-ubiquity-benchmark.rds")
 
 gns_result_file <- file.path(resdir, "baxter-genus-results.rds")
-gns_bench_file <- file.path(resdir, "baxter-genus-benchmark.rds")
-gns_bench_file_log_ubiquity <- file.path(resdir, "baxter-genus-log-ubiquity-benchmark.rds")
+gns_bench_file <- file.path(sbdir, "baxter-genus-benchmark.rds")
+gns_bench_file_log_ubiquity <- file.path(sbdir, "baxter-genus-log-ubiquity-benchmark.rds")
 ```
@@
 
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
 ## Multiple-Testing Correction - log(ubiquity)
@@
 sum(is.na(assays(sb)[["qvalue"]][, "lfdr"]))
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
 ```
@@
 ```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
 
@@
 
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
 # Session Info

### 2018-05-10 datasets/microbiome/microbiome-enigma.Rmd
Save ChIPseq, RNAseq, and microbiome results to common dir
Also add covariateLinePlots to microbiome Rmds.

@@
 resdir <- "results"
+sbdir <- "../../results/microbiome"
+dir.create(datdir, showWarnings = FALSE)
+dir.create(resdir, showWarnings = FALSE)
+dir.create(sbdir, showWarnings = FALSE)
+
 enigma_dir <- file.path(datdir, "enigma_results")
+dir.create(enigma_dir, showWarnings = FALSE)
 
@@
 so4_result_file <- file.path(resdir, "enigma-so4-otus-results.rds")
-so4_bench_file <- file.path(resdir, "enigma-so4-otus-benchmark.rds")
+so4_bench_file <- file.path(sbdir, "enigma-so4-otus-benchmark.rds")
 
 ph_result_file <- file.path(resdir, "enigma-ph-otus-results.rds")
-ph_bench_file <- file.path(resdir, "enigma-ph-otus-benchmark.rds")
+ph_bench_file <- file.path(sbdir, "enigma-ph-otus-benchmark.rds")
 
 al_result_file <- file.path(resdir, "enigma-al-otus-results.rds")
-al_bench_file <- file.path(resdir, "enigma-al-otus-benchmark.rds")
+al_bench_file <- file.path(sbdir, "enigma-al-otus-benchmark.rds")
 ```
@@
 rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
 ```

### 2018-05-10 datasets/microbiome/microbiome-goodrich.Rmd
Save ChIPseq, RNAseq, and microbiome results to common dir
Also add covariateLinePlots to microbiome Rmds.

@@
 otu_result_file <- file.path(resdir, "goodrich-otus-results.rds")
-otu_bench_file <- file.path(resdir, "goodrich-otus-benchmark.rds")
+otu_bench_file <- file.path(sbdir, "goodrich-otus-benchmark.rds")
 gns_result_file <- file.path(resdir, "goodrich-genus-results.rds")
-gns_bench_file <- file.path(resdir, "goodrich-genus-benchmark.rds")
+gns_bench_file <- file.path(sbdir, "goodrich-genus-benchmark.rds")
 ```

### 2018-05-10 datasets/microbiome/microbiome-papa.Rmd
Save ChIPseq, RNAseq, and microbiome results to common dir
Also add covariateLinePlots to microbiome Rmds.

@@
 otu_result_file <- file.path(resdir, "papa-otus-results.rds")
-otu_bench_file <- file.path(resdir, "papa-otus-benchmark.rds")
+otu_bench_file <- file.path(sbdir, "papa-otus-benchmark.rds")
 gns_result_file <- file.path(resdir, "papa-genus-results.rds")
-gns_bench_file <- file.path(resdir, "papa-genus-benchmark.rds")
+gns_bench_file <- file.path(sbdir, "papa-genus-benchmark.rds")
 gns_bench_file_log_ubi <- file.path(resdir, "papa-genus-log-ubiquity-benchmark.rds")

### 2018-05-10 datasets/microbiome/microbiome-schubert.Rmd
Save ChIPseq, RNAseq, and microbiome results to common dir
Also add covariateLinePlots to microbiome Rmds.

@@
 otu_result_file <- file.path(resdir, "schubert-otus-results.rds")
-otu_bench_file <- file.path(resdir, "schubert-otus-benchmark.rds")
+otu_bench_file <- file.path(sbdir, "schubert-otus-benchmark.rds")
 ```

### 2018-05-10 datasets/ChIP-seq/ChIPseq.sh

@@ -4,7 +4,7 @@
 #SBATCH -N 1
 #SBATCH -p irizarry,serial_requeue
 #SBATCH --mem 16G
-#SBATCH -t 0-05:00
+#SBATCH -t 1-05:00
 #SBATCH -o render-%j.out
 #SBATCH -e render-%j.err
  
@@ -15,3 +15,4 @@ export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/rstudio/0.98.1103-fasrc01/bin/pan
 # variable SLURM_NTASKS
 R -e "rmarkdown::render('ChIPseq-h3k4me3-promoters.Rmd')"
 R -e "rmarkdown::render('ChIPseq-h3k4me3-csaw.Rmd')"
+R -e "rmarkdown::render('ChIPseq-CBP-csaw.Rmd')"

### 2018-05-10 datasets/microbiome/microbiome.sh

@@ -2,9 +2,9 @@
 #SBATCH -J knit16s
 #SBATCH -n 4
 #SBATCH -N 1
-#SBATCH -p shared
+#SBATCH -p shared,irizarry
 #SBATCH --mem 32G
-#SBATCH -t 0-6:00
+#SBATCH -t 0-8:00
 #SBATCH -o render-%j.out
 #SBATCH -e render-%j.err
  
@@ -14,3 +14,5 @@ export RSTUDIO_PANDOC="/n/sw/fasrcsw/apps/Core/rstudio/0.98.1103-fasrc01/bin/pan
 R -e "rmarkdown::render('microbiome-baxter.Rmd')"
 R -e "rmarkdown::render('microbiome-goodrich.Rmd')"
 R -e "rmarkdown::render('microbiome-schubert.Rmd')"
+R -e "rmarkdown::render('microbiome-enigma.Rmd')"
+R -e "rmarkdown::render('microbiome-papa.Rmd')"

### 2018-05-11 manuscript_figures/manuscript_figures.Rmd
Refine method rank heatmap; add case-study specific heatmaps
Method ranking heatmap is now a standalone function so it can
be reused to make average ranks across case studies, datsets,
and/or covariates. The default now colors by mean proportion of
maximum rejections instead of mean rank.

--- 2018-05-01 manuscript_figures/manuscript_figures.Rmd
@@
 # dataset names later on will be incorrect)
-path <- "../sb-objects"
+path <- "../results"
 
@@
 dir.create(outdir, showWarnings = FALSE)
-
-# point to where yeast results are (within 'path' defined above)
-resdir <- file.path(path, "RNAseq/yeast-results")
 
@@
 ```{r Figure2}
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+resfile <- objects[grepl("yeast", objects) & grepl("de5", objects)] 
+de5 <- readRDS(file=resfile)
 excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
-resfile <- file.path(resdir, "de5.rds")
-de5 <- readRDS(file=resfile)
 res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
@@
 library(stringr)
-library(viridis)
 
@@
                which(grepl("polyester", objects))))
-
 if (length(sim) > 0){
@@
 
-# create tidy data frame where each row is a method / dataset observation
-# of a rank 
-ranks <- data.frame()
+# remove microbiome objects from 'null' comparisons
+# removing baxter, goodrich and papa OTU-level (since they mostly found 
+# nothing for all methods and the genus-level analysis is also present)
+null <- which(grepl(c("baxter|goodrich|papa"), objects) & 
+              grepl(c("otu"), objects))
+
+if (length(null) > 0){
+  objects <- objects[-null]
+}
+
+# remove covariates from individual cases studies that weren't informative
+## TODO ##
+
+# grab casestudy names
 obj.names <- str_split(objects, pattern = "/")
@@
 
-for (i in seq_along(objects)){
-  x <- readRDS(objects[i])
-  assayNames(x) <- "qvalue"
-  x <- addDefaultMetrics( x )
+Fig4 <- plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10)
+Fig4
+ggsave(file.path(outdir, "Figure4.pdf"), width=6, height=4)
 
-  hasResults <- apply(!is.na( assays(x)[["qvalue"]] ), 2, all )
+# heatmap : rows method, cols casestudy- one color scale 
+Fig4alt <- plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10, 
+                        colorLow = "white", colorHigh = "blue", 
+                        colorNA = "grey50")
+Fig4alt
+ggsave(file.path(outdir, "Figure4alt.pdf"), width=6, height=4)
 
-  if (sum(hasResults) > 0){
-    tmp <- estimatePerformanceMetrics(x, alpha, tidy=TRUE) %>%
-      filter( performanceMetric == "rejections") %>%
-      dplyr::rename( method = blabel) %>%
-      filter( is.na(param.alpha) | (param.alpha == alpha)) %>%
-      filter( is.na(param.smooth.df) | (param.smooth.df == "3L")) %>%
-      filter( !method == "unadjusted") %>%
-      select( method, value ) %>%
-      dplyr::rename( nrejects = value) %>%
-      mutate( method = gsub("-df03", "", method)) %>%
-      mutate( method = gsub("-a10", "", method)) %>%
-      na.omit() %>%
-      mutate( rank = rank(nrejects) / n() ) %>%
-      mutate(casestudy = casestudy[i],
-             dataset = dataset[i])
+##
+## AR: alternative suggestion to heatmap, feel free to discard! ##
+##
 
-    ranks <- rbind(ranks, tmp)
-  }
-}
-
-# average over datasets within case study
-ranks_avg <- ranks %>% 
-  group_by( casestudy, method) %>%
-  summarize( mean_rank = mean(rank, na.rm = TRUE))
-
-case_dat <- ranks_avg %>%
-  group_by( casestudy) %>%
-  summarize( nmethods = n())
-
-method_dat <- ranks_avg %>%
-  group_by( method) %>%
-  summarize( meta_rank = mean(mean_rank))
-
-ranks_avg <- left_join(ranks_avg, method_dat, by ="method")
-ranks_avg <- left_join(ranks_avg, case_dat, by ="casestudy")
-
-# reorder rows and columns
-ranks_avg$casestudy <- factor(ranks_avg$casestudy, 
-                              levels=unique(ranks_avg$casestudy[order(ranks_avg$nmethods)]))
-ranks_avg$method <- factor(ranks_avg$method,
-                           levels=unique(ranks_avg$method[order(ranks_avg$meta_rank)]))
-
-# heatmap : rows method, cols casestudy
-Fig4 <- ggplot(ranks_avg, aes(x = casestudy, y = method, fill = mean_rank)) + 
-  geom_raster() + 
-  scale_fill_viridis() +
-  theme_bw() +
-  xlab("Case Study") +
-  ylab("Method") +
-  labs(fill = "Mean rank")
-
-Fig4
-ggsave(file.path(outdir, "Figure4.pdf"), width=6, height=4, dev = pdf)
+Fig4b <- plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10,
+                         linePlot = TRUE)
+Fig4b
 ```
@@
 
+# Supplementary Figures
+
+## Within case-study heatmaps of method rankings
+
+Here we generate heatmaps for each case study, where each column represents a different
+dataset. Here we pick one representative covariate and method (if applicable) 
+for each case study. See the individual Rmds for more detailed plots comparing, e.g.
+different covariates on the same dataset.
+
+```{r}
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+
+# grab casestudy names
+obj.names <- str_split(objects, pattern = "/")
+casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
+dataset <- unlist(lapply(obj.names, function(x) x[4] ))
+```
+
+### ChIP-seq
+
+```{r}
+wcs <- which(grepl("ChIPseq", casestudy))
+objs <- objects[wcs]
+labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
+                      function(x) x[1]))
+plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset") + 
+  ggtitle("ChIP-Seq")
+```
+
+### GSEA
+
+```{r}
+#by dataset
+wcs <- which(grepl("GSEA", casestudy))
+objs <- objects[wcs]
+labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
+                      function(x) x[1]))
+plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset") + 
+  ggtitle("GSEA")
+```
+
+### GWAS
+
+```{r}
+# by covariate
+wcs <- which(grepl("GWAS", casestudy))
+objs <- objects[wcs]
+labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-"), 
+                      function(x) x[2]))
+plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Covariate") + 
+  ggtitle("GWAS")
+```
+
+### Microbiome
+
+```{r}
+# by dataset, excluding otu when genus is available
+# excluding log-ubiquity
+# excluding enigma-so4 (no discoveries)
+wcs <- which(grepl("microbiome", casestudy) & 
+              !(grepl(c("baxter|goodrich|papa"), dataset) & 
+              grepl(c("otu"), dataset)) &
+              !grepl("log-ubiquity", dataset) &
+              !grepl("so4", dataset))
+objs <- objects[wcs]
+labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
+                      function(x) x[1]))
+plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset") + 
+  ggtitle("Microbiome")
+```
+
+### RNA-seq
+```{r}
+# by dataset
+wcs <- which(grepl("RNAseq", casestudy) & 
+            !grepl(c("polyester|yeast"), dataset) &
+            !grepl("scRNAseq", casestudy))
+objs <- objects[wcs]
+labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
+                      function(x) x[1]))
+plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset") + 
+  ggtitle("RNA-seq")
+```
+
+### scRNA-seq
+
+```{r}
+wcs <- which(grepl("scRNAseq", casestudy))
+objs <- objects[wcs]
+ds <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark-"), 
+                      function(x) x[1]))
+m0 <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark-"), 
+                      function(x) x[2]))
+meth <- unlist(lapply(str_split(m0, pattern = "-"), 
+                      function(x) x[1]))
+cv <- gsub(".rds", "", unlist(lapply(str_split(m0, pattern = "-"), 
+                      function(x) x[2])))
+
+# by dataset
+plotMethodRanks(objs, colLabels = ds, alpha = 0.10, xlab = "Dataset") + 
+  ggtitle("scRNA-seq")
+
+# by method
+plotMethodRanks(objs, colLabels = meth, alpha = 0.10, xlab = "Method") + 
+  ggtitle("scRNA-seq")
+
+# by covariate
+plotMethodRanks(objs, colLabels = cv, alpha = 0.10, xlab = "Covariate") + 
+  ggtitle("scRNA-seq")
+```
+
+### Yeast Simulation
+
+```{r}
+TODO
+```
+
+### Polyester Simulation
+
+```{r}
+TODO
+```
+
 # Session information

### 2018-05-11 datasets/R/plotMethodRank.R

@@ -1,0 +1,138 @@
+#' Function to create a heatmap of method rankings for different case studies,
+#' datasets, or covariates. 
+#' 
+#' Takes as input a vector of filepaths, each one pointing to a SummarizedBenchmark
+#' results object. Output is a ggplot object with the heatmap. Each file 
+#' specified will represent one column of the heatmap (rows are methods).
+#' 
+#' @param objects a character vector containing the full filepaths to 
+#'  SummarizedBenchmark results objects.
+#' @param colLabels a character vector containing the labels to use for the 
+#'  heatmap columns (same length and order as `objects`). 
+#' @param alpha numeric value indicating which alpha value to use for IHW 
+#'  (needs to be one of the values specified in the bench object). Default 0.10.
+#' @param colorLow character representing the color to be used for the lowest 
+#'  value
+#' @param colorHigh character representing the color to be used for the highest
+#'  value
+#' @param colorNA character representing the color to be used for NA (i.e. 
+#'  for methods that weren't used in the particular benchmark)
+#' @param linePlot logical indicating whether to plot a line plot in lieu of 
+#'  a heatmap  
+#' @param excludeMethods character vector containing the names of methods to
+#'  exclude from the heatmap. Default is to exclude Scott methods which are 
+#'  only present in GWAS.
+#' @param propMaxRejections logical indicating whether the proportion of max
+#'  number of rejections is plotted (instead of the rank). Default is TRUE. 
+#' @param xlab character with the x label (default is "Case Study")   
+
+plotMethodRanks <- function(objects, colLabels, alpha = 0.10, 
+                            colorLow = "navy", colorHigh = "yellow",
+                            colorNA = "white", xlab = "Case Study",
+                            propMaxRejections = TRUE,
+                            linePlot = FALSE,
+                            excludeMethods = c("scott-theoretical",
+                                               "scott-empirical")){
+
+  # create tidy data frame where each row is a method / dataset observation
+  # of a rank 
+  ranks <- data.frame()
+  
+  for (i in seq_along(objects)){
+    x <- readRDS(objects[i])
[…]

### 2018-05-13 datasets/RNA-Seq/RNAseq-mir200c.Rmd
Hyperlink DOIs against preferred resolver

--- 2018-04-20 datasets/RNA-Seq/RNAseq-mir200c.Rmd
@@
 
-As a second RNA-seq dataset, we will test for differences in gene expression upon the knockout of the microRNA mir-200c [(Kim et al., 2013)](http://dx.doi.org/10.1038/nsmb.2701).
+As a second RNA-seq dataset, we will test for differences in gene expression upon the knockout of the microRNA mir-200c [(Kim et al., 2013)](https://doi.org/10.1038/nsmb.2701).
 The raw fastq files can be found under the accession number `SRP030475`.
@@
 resdir <- "results"
+sbdir <- "../../results/RNAseq"
 dir.create(datdir, showWarnings = FALSE)
 dir.create(resdir, showWarnings = FALSE)
+dir.create(sbdir, showWarnings = FALSE)
 
@@
 result_file <- file.path(resdir, "mir200c-results.rds")
-bench_file <- file.path(resdir, "mir200c-benchmark.rds")
+bench_file <- file.path(sbdir, "mir200c-benchmark.rds")
+bench_file_uninf <- file.path(sbdir, "mir200c-uninf-benchmark.rds")
 
@@
 
+Add random (uninformative) covariate.
+
+```{r}
+set.seed(4719)
+res$rand_covar <- rnorm(nrow(res))
+```
+
 ## Covariate Diagnostics
@@
 
-## Multiple-Testing Correction
+### Random
 
-We use the common `BenchDesign` with the set of multiple testing correction methods already included.
+```{r, rand-diag-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(res, pvalue = "pval", covariate = "rand_covar") +
+    ggtitle("Random independent covariate")
+```
+
+```{r, rand-diag-hist, fig.width=10, fig.height=3.2}
+strat_hist(res, pvalue = "pval",
+           covariate = "rand_covar", maxy = 7.5)
+```
+
+## Multiple-Testing Correction - mean
+
+We use the common `BenchDesign` with the set of multiple testing correction 
+methods already included. We also add in Scott's FDR Regression (both
+`nulltype = "empirical"` and `nulltype = "theoretical"`)
+since our test statistics are t-distributed.
 
@@
     bd <- initializeBenchDesign()
+
+    bd <- addBMethod(bd, "fdrreg-t",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'theoretical',
+                     control = list(lambda = 0.01))
+
+   bd <- addBMethod(bd, "fdrreg-e",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'empirical',
+                     control = list(lambda = 0.01))
+
     sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
@@
 
-There are some warnings from both `BH` and `scott-empirical`. However, they do return results.
+There are some warnings from both `BH` and `fdrreg-empirical`. However, they do return results.
 I tried to increase the `nmids` parameter for Scott's FDR Regression but this does not appear to make a difference.
@@
 
-## Benchmark Metrics
+## Benchmark Metrics - mean
 
-`scott-empirical` rejects many more hypothesis than the rest of the methods, followed by `lfdr` and `ihw`.
+`fdrreg-empirical` rejects many more hypothesis than the rest of the methods, followed by `lfdr` and `ihw`.
 
@@
 
+## Multiple-Testing Correction - random
+
+We use the common `BenchDesign` with the set of multiple testing correction 
+methods already included. We also add in Scott's FDR Regression (both
+`nulltype = "empirical"` and `nulltype = "theoretical"`)
+since our test statistics are t-distributed.
+
+```{r}
+if (file.exists(bench_file_uninf)) {
+    sb <- readRDS(bench_file_uninf)
+} else {
+    bd <- initializeBenchDesign()
+
+    bd <- addBMethod(bd, "fdrreg-t",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'theoretical',
+                     control = list(lambda = 0.01))
+
+   bd <- addBMethod(bd, "fdrreg-e",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'empirical',
+                     control = list(lambda = 0.01))
+
+    res <- res %>% dplyr::mutate(ind_covariate = rand_covar)
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    saveRDS(sb, file = bench_file_uninf)
+}
+```
+
+```{r}
+head(assays(sb)[["bench"]])
+```
+
+## Benchmark Metrics - random
+
+```{r}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+estimatePerformanceMetrics(sb, alpha = 0.05, tidy = TRUE) %>%
+    filter(performanceMetric == "rejections") %>%
+    select(blabel, performanceMetric, alpha, value) %>%
+    mutate(n = nrow(sb), prop = round(value / n, 3)) %>%
+    arrange(desc(value)) %>%
+    as_tibble() %>%
+    print(n = 40)
+```
+
+```{r}
+rejections_scatter(sb, supplementary = FALSE)
+rejection_scatter_bins(sb, covariate = "ind_covariate",
+                       bins = 4, supplementary = FALSE)
+```
+
+```{r}
+plotFDRMethodsOverlap(sb, alpha = 0.05, nsets = ncol(sb),
+                      order.by = "freq", decreasing = TRUE,
+                      supplementary = FALSE)
+```
+
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
+# Covariate comparison
+
+Here we compare the method ranks for the different covariates at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(bench_file, bench_file_uninf), 
+                colLabels = c("mean", "uninf"), 
+                alpha = 0.10, xlab = "Comparison")
+```
+
 # Session Info

### 2018-05-13 datasets/microbiome/microbiome-baxter.Rmd
Hyperlink DOIs against preferred resolver

@@
 Here we download and analyze the Baxter colorectal cancer (CRC) dataset
-[(Baxter et al., 2016)](http://dx.doi.org/10.1186/s13073-016-0290-3).
+[(Baxter et al., 2016)](https://doi.org/10.1186/s13073-016-0290-3).
 The dataset includes stool samples from 172 patients with no colonic lesions,
@@
 gns_bench_file_log_ubiquity <- file.path(sbdir, "baxter-genus-log-ubiquity-benchmark.rds")
+gns_bench_file_abun <- file.path(sbdir, "baxter-genus-mean-abun-benchmark.rds")
 ```
@@
 
+## Multiple-Testing Correction - mean abundance
+
+Let's use `mean_abun_present` as our `ind_covariate`.
+
+```{r, baxter-genus-labelcovariate-abun}
+res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, baxter-genus-sb-abun, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(gns_bench_file_abun)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    metadata(sb)$data_download_link <-
+                   "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
+    saveRDS(sb, file = gns_bench_file_abun)
+} else {
+    sb <- readRDS(gns_bench_file_abun)
+}
+```
+## Benchmark Metrics - mean abundance
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, baxter-genus-abun-metrics, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, baxter-genus-abun-plot, width=15, height=15, eval=TRUE}
+rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+
+Note: Benjamini-Hochberg (bh) overlaps exactly with the IHW results.
+
+```{r, baxter-genus-abun-overlap}
+plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
+# OTU/Genus comparison
+
+Here we compare the method ranks for the different comparisons at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(otu_bench_file, otu_bench_file_log_ubiquity, 
+                  gns_bench_file, gns_bench_file_log_ubiquity,
+                  gns_bench_file_abun), 
+                colLabels = c("OTU", "OTU - log", "Genus", "Genus - log",
+                              "Genus - abun"), 
+                alpha = 0.10, xlab = "Comparison")
+```
+
 # Session Info

### 2018-05-13 datasets/microbiome/microbiome-enigma.Rmd
Hyperlink DOIs against preferred resolver

@@
 
-Here we analyze a subset of the ENIGMA data ([Smith, M. and Rocha, A. et al 2015](http://dx.doi.org/10.1128/mBio.00326-15)).
+Here we analyze a subset of the ENIGMA data ([Smith, M. and Rocha, A. et al 2015](https://doi.org/10.1128/mBio.00326-15)).
 The dataset includes water samples from 67 wells with well-characterized biochemical metadata (i.e. pH, heavy metal concentrations, etc).
@@
 so4_bench_file <- file.path(sbdir, "enigma-so4-otus-benchmark.rds")
+so4_bench_file_abun <- file.path(sbdir, "enigma-so4-otus-abun-benchmark.rds")
+so4_bench_file_uninf <- file.path(sbdir, "enigma-so4-otus-uninf-benchmark.rds")
 
@@
 ph_bench_file <- file.path(sbdir, "enigma-ph-otus-benchmark.rds")
+ph_bench_file_abun <- file.path(sbdir, "enigma-ph-otus-abun-benchmark.rds")
+ph_bench_file_uninf <- file.path(sbdir, "enigma-ph-otus-uninf-benchmark.rds")
 
@@
 al_bench_file <- file.path(sbdir, "enigma-al-otus-benchmark.rds")
+al_bench_file_abun <- file.path(sbdir, "enigma-al-otus-abun-benchmark.rds")
+al_bench_file_uninf <- file.path(sbdir, "enigma-al-otus-uninf-benchmark.rds")
 ```
@@
 
-## Covariate Diagnostics
+Add random (uninformative) covariate.
+
+```{r}
+set.seed(44318)
+res$rand_covar <- rnorm(nrow(res))
+```
+
+### Covariate Diagnostics
 
@@
 
-### Ubiquity
+#### Ubiquity
 
@@
 
-### Mean Abundance (across non-zero samples)
+#### Mean Abundance (across non-zero samples)
 
@@
 
-## Multiple-Testing Correction
+#### Random
+
+```{r, ph-rand-hist, fig.width=10, fig.height=3.2}
+strat_hist(res, pvalue="pval", covariate="rand_covar", maxy=8, binwidth=0.05)
+```
+
+```{r, ph-rand-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(res, pvalue="pval", covariate="rand_covar")
+```
+
+### Multiple-Testing Correction - ubiquity
 
@@
 
-## Benchmark Metrics
+### Benchmark Metrics - ubiquity
 
@@
 ```{r, ph-plot, width=15, height=15, eval=TRUE}
+rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
+### Multiple-Testing Correction - mean abundance
+
+Let's use `mean_abun_present` as our `ind_covariate`.
+
+```{r, ph-labelcovariate-abun}
+res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the default comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, ph-sb-abun, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(ph_bench_file_abun)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    saveRDS(sb, file = ph_bench_file_abun)
+} else {
+    sb <- readRDS(ph_bench_file_abun)
+}
+```
+### Benchmark Metrics - mean abundance
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, ph-metrics-abun, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, ph-plot-abun, width=15, height=15, eval=TRUE}
+rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
+### Multiple-Testing Correction - random
+
+Let's use `rand_covar` as our `ind_covariate`.
+
+```{r, ph-labelcovariate-rand}
+res <- dplyr::mutate(res, ind_covariate = rand_covar)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the default comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, ph-sb-rand, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(ph_bench_file_uninf)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    saveRDS(sb, file = ph_bench_file_uninf)
+} else {
+    sb <- readRDS(ph_bench_file_uninf)
+}
+```
+### Benchmark Metrics - random
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, ph-metrics-rand, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, ph-plot-rand, width=15, height=15, eval=TRUE}
 rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
@@
 
-## Covariate Diagnostics
+Add random (uninformative) covariate.
+
+```{r}
+set.seed(27349)
+res$rand_covar <- rnorm(nrow(res))
+```
+
+### Covariate Diagnostics
 
@@
 
-## Multiple-Testing Correction
+#### Random
+
+```{r, so4-rand-hist, fig.width=10, fig.height=3.2}
+strat_hist(res, pvalue="pval", covariate="rand_covar", maxy=8, binwidth=0.05)
+```
+
+```{r, so4-rand-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(res, pvalue="pval", covariate="rand_covar")
+```
+
+### Multiple-Testing Correction - ubiquity
 
@@
 
+### Multiple-Testing Correction - mean abundance
+
+Let's use `mean_abun_present` as our `ind_covariate`.
+
+```{r, so4-labelcovariate-abun}
+res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the default comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, so4-sb-abun, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(so4_bench_file_abun)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    saveRDS(sb, file = so4_bench_file_abun)
+} else {
+    sb <- readRDS(so4_bench_file_abun)
+}
+```
+### Benchmark Metrics - mean abundance
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, so4-metrics-abun, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, so4-plot-abun, width=15, height=15, eval=TRUE}
+rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
+### Multiple-Testing Correction - random
+
+Let's use `rand_covar` as our `ind_covariate`.
+
+```{r, so4-labelcovariate-rand}
+res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the default comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, so4-sb-rand, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(so4_bench_file_uninf)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    saveRDS(sb, file = so4_bench_file_uninf)
+} else {
+    sb <- readRDS(so4_bench_file_uninf)
+}
+```
+### Benchmark Metrics - random
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, so4-metrics-rand, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, so4-plot-rand, width=15, height=15, eval=TRUE}
+rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
 ## Al
@@
 }
+
+# remove OTUs with missing pvals
+res <- res %>% filter(!is.na(pval))
 ```
 
-## Covariate Diagnostics
+Add random (uninformative) covariate.
+
+```{r}
+set.seed(19474)
+res$rand_covar <- rnorm(nrow(res))
+```
+
+### Covariate Diagnostics
 
@@
 
-## Multiple-Testing Correction
+#### Random
+
+```{r, al-rand-hist, fig.width=10, fig.height=3.2}
+strat_hist(res, pvalue="pval", covariate="rand_covar", maxy=8, binwidth=0.05)
+```
+
+```{r, al-rand-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(res, pvalue="pval", covariate="rand_covar")
+```
+
+### Multiple-Testing Correction - ubiquity
 
@@
 
+### Multiple-Testing Correction - mean abundance
+
+Let's use `mean_abun_present` as our `ind_covariate`.
+
+```{r, al-labelcovariate-abun}
+res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the default comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, al-sb-abun, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(al_bench_file_abun)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    saveRDS(sb, file = al_bench_file_abun)
+} else {
+    sb <- readRDS(al_bench_file_abun)
+}
+```
+### Benchmark Metrics - mean abundance
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, al-metrics-abun, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, al-plot-abun, width=15, height=15, eval=TRUE}
+rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
+### Multiple-Testing Correction - random
+
+Let's use `rand_covar` as our `ind_covariate`.
+
+```{r, al-labelcovariate-rand}
+res <- dplyr::mutate(res, ind_covariate = rand_covar)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the default comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, al-sb-rand, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(al_bench_file_uninf)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    saveRDS(sb, file = al_bench_file_uninf)
+} else {
+    sb <- readRDS(al_bench_file_uninf)
+}
+```
+### Benchmark Metrics - random
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, al-metrics-rand, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, al-plot-rand, width=15, height=15, eval=TRUE}
+rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
+# Response/Covariate comparison
+
+Here we compare the method ranks for the different comparisons at alpha = 0.10.
+
+```{r, width = 15, height = 10}
+plotMethodRanks(c(so4_bench_file, so4_bench_file_abun, so4_bench_file_uninf,
+                  ph_bench_file, ph_bench_file_abun, ph_bench_file_uninf,
+                  al_bench_file, al_bench_file_abun, al_bench_file_uninf), 
+                colLabels = c("SO4-ubiquity", "SO4-abun", "SO4-uninf",
+                              "pH-ubiquity", "ph-abun", "pH-uninf",
+                              "Al-ubiquity", "Al-abun", "Al-uninf"),
+                alpha = 0.10, xlab = "Comparison")
+```
+
 # Session Info

### 2018-05-13 datasets/microbiome/microbiome-goodrich.Rmd
Hyperlink DOIs against preferred resolver

@@
 Here we download and analyze the Goodrich obesity (OB) dataset
-[(Goodrich et al., 2014)](http://dx.doi.org/10.1016/j.cell.2014.09.053).
+[(Goodrich et al., 2014)](https://doi.org/10.1016/j.cell.2014.09.053).
 The dataset includes fecal samples from 428 individuals with lean body mass index (BMI < 25),
@@
 
-We expect few or no truly differentially abundant OTUs ([Finucane et al., 2014](http://dx.doi.org/10.1371/journal.pone.0084689) and [Walters et al., 2014](https://dx.doi.org/10.1016%2Fj.febslet.2014.09.039)), so this study provides a real-life "null" example.
+We expect few or no truly differentially abundant OTUs ([Finucane et al., 2014](https://doi.org/10.1371/journal.pone.0084689) and [Walters et al., 2014](https://doi.org/10.1016%2Fj.febslet.2014.09.039)), so this study provides a real-life "null" example.
 
@@
 gns_bench_file <- file.path(sbdir, "goodrich-genus-benchmark.rds")
+gns_bench_file_abun <- file.path(sbdir, "goodrich-genus-abun-benchmark.rds")
+gns_bench_file_uninf <- file.path(sbdir, "goodrich-genus-uninf-benchmark.rds")
 ```
@@
 
-## Multiple-Testing Correction
+## Multiple-Testing Correction - ubiquity
 
@@
 
-## Benchmark Metrics
+## Benchmark Metrics - ubiquity
 
@@
 
+Add random (uninformative) covariate.
+
+```{r}
+set.seed(76291)
+res$rand_covar <- rnorm(nrow(res))
+```
+
 ## Covariate Diagnostics
@@
 
-## Multiple-Testing Correction
+### Random (uninformative)
+
+```{r, goodrich-uninf-hist, fig.width=10, fig.height=3.2}
+strat_hist(res, pvalue="pval", covariate="rand_covar", maxy=5, binwidth=0.05, numQ=4)
+```
+
+```{r, goodrich-uninf-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(res, pvalue="pval", covariate="rand_covar")
+```
+
+## Multiple-Testing Correction - ubiquity
 
@@
 
+## Multiple-Testing Correction - mean abun
+
+Let's use `mean_abun_present` as our `ind_covariate`.
+
+```{r, goodrich-genus-labelcovariate-abun}
+res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, goodrich-genus-sb-abun, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(gns_bench_file_abun)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    metadata(sb)$data_download_link <-
+                   "https://zenodo.org/record/840333/files/ob_goodrich_results.tar.gz"
+    saveRDS(sb, file = gns_bench_file_abun)
+} else {
+    sb <- readRDS(gns_bench_file_abun)
+}
+```
+## Benchmark Metrics - mean abun
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, goodrich-genus-metrics-abun, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, goodrich-genus-plot-abun, width=15, height=15, eval=TRUE}
+rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
+Note: Benjamini-Hochberg (bh) overlaps exactly with the IHW results.
+
+```{r, goodrich-genus-overlap-abun}
+plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+```
+
+```{r, goodrich-genus-covariates-abun}
+methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
+plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods)
+```
+
+## Multiple-Testing Correction - random
+
+Let's use `rand_covar` as our `ind_covariate`.
+
+```{r, goodrich-genus-labelcovariate-uninf}
+res <- dplyr::mutate(res, ind_covariate = rand_covar)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, goodrich-genus-sb-uninf, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(gns_bench_file_uninf)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    metadata(sb)$data_download_link <-
+                   "https://zenodo.org/record/840333/files/ob_goodrich_results.tar.gz"
+    saveRDS(sb, file = gns_bench_file_uninf)
+} else {
+    sb <- readRDS(gns_bench_file_uninf)
+}
+```
+## Benchmark Metrics - random
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, goodrich-genus-metrics-uninf, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, goodrich-genus-plot-uninf, width=15, height=15, eval=TRUE}
+rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
+Note: Benjamini-Hochberg (bh) overlaps exactly with the IHW results.
+
+```{r, goodrich-genus-overlap-uninf}
+plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+```
+
+```{r, goodrich-genus-covariates-uninf}
+methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
+plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods)
+```
+
+# OTU/Genus comparison
+
+Here we compare the method ranks for the different comparisons at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(otu_bench_file, gns_bench_file, gns_bench_file_abun, gns_bench_file_uninf), 
+                colLabels = c("OTU", "Genus-ubiquity", "Genus-abun", "Genus-uninf"), 
+                alpha = 0.10, xlab = "Comparison")
+```
+
 # Session Info

### 2018-05-13 datasets/microbiome/microbiome-schubert.Rmd
Hyperlink DOIs against preferred resolver

@@
 
-Here we download and analyze the Schubert diarrhea dataset [(Schubert et al., 2014)](http://dx.doi.org/10.1128/mBio.01021-14).
+Here we download and analyze the Schubert diarrhea dataset [(Schubert et al., 2014)](https://doi.org/10.1128/mBio.01021-14).
 The dataset includes 154 healthy nondiarrheal, 93 Clostridium difficile infection (CDI) associated diarrheal,
@@
 otu_bench_file <- file.path(sbdir, "schubert-otus-benchmark.rds")
+otu_bench_file_abun <- file.path(sbdir, "schubert-otus-abun-benchmark.rds")
+otu_bench_file_uninf <- file.path(sbdir, "schubert-otus-uninf-benchmark.rds")
 ```
@@
 
+Add random (uninformative) covariate.
+
+```{r}
+set.seed(9226)
+res$rand_covar <- rnorm(nrow(res))
+```
+
 Finally, let's try to add phylogeny as covariates. Here we'll have columns for each separate taxonomic level.
@@
 ```{r, schubert-abun, fig.width=10, fig.height=3.2, eval=TRUE}
-strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
+strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=17, binwidth=0.05)
 ```
@@
 
-## Multiple-Testing Correction
+### Random
+
+```{r, schubert-rand, fig.width=10, fig.height=3.2, eval=TRUE}
+strat_hist(res, pvalue="pval", covariate="rand_covar", maxy=17, binwidth=0.05)
+```
+
+```{r, schubert-rand-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(res, pvalue="pval", covariate="rand_covar")
+```
+
+## Multiple-Testing Correction - ubiquity
 
@@
 
+## Multiple-Testing Correction - mean abundance
+
+Let's use `mean_abun_present` as our `ind_covariate`.
+
+```{r, schubert-labelcovariate-abun, eval=TRUE}
+res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, schubert-sb-abun, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(otu_bench_file_abun)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    metadata(sb)$data_download_link <-
+                   "https://zenodo.org/record/840333/files/cdi_schubert_results.tar.gz"
+    saveRDS(sb, file = otu_bench_file_abun)
+} else {
+    sb <- readRDS(otu_bench_file_abun)
+}
+```
+
+## Benchmark Metrics (mean abun)
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, schubert-metrics-abun, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, schubert-plot-abun, width=15, height=15, eval=TRUE}
+rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
+## Multiple-Testing Correction - random
+
+Let's use `rand_covar` as our `ind_covariate`.
+
+```{r, schubert-labelcovariate-rand, eval=TRUE}
+res <- dplyr::mutate(res, ind_covariate = rand_covar)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, schubert-sb-rand, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(otu_bench_file_uninf)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    metadata(sb)$data_download_link <-
+                   "https://zenodo.org/record/840333/files/cdi_schubert_results.tar.gz"
+    saveRDS(sb, file = otu_bench_file_uninf)
+} else {
+    sb <- readRDS(otu_bench_file_uninf)
+}
+```
+
+## Benchmark Metrics - random
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, schubert-metrics-rand, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, schubert-plot-rand, width=15, height=15, eval=TRUE}
+rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
+# Ubiquity/Abundance comparison
+
+Here we compare the method ranks for the different comparisons at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(otu_bench_file, otu_bench_file_abun, otu_bench_file_uninf), 
+                colLabels = c("OTU-ubiquity","OTU-abun", "OTU-uninf"), 
+                alpha = 0.10, xlab = "Comparison")
+```
+
 # Session Info

### 2018-05-14 datasets/GWAS/GWAS.Rmd
Compute stat from effect/SE instead of pval

@@
                   ind_covar_AF = Freq_Allele1_HapMapCEU) %>%
-    mutate(test_statistic = qnorm(exp(log(pval)-log(2)), lower.tail=FALSE) * sign(effect_size)) %>%
+    mutate(test_statistic = effect_size / SE) %>%
     select(pval, SE, effect_size, ind_covar_N, ind_covar_AF, test_statistic) 

### 2018-05-14 datasets/RNA-Seq/RNAseq-gtex.Rmd
Add Scott methods to RNA-seq case studies/sims
Since DESeq gives a Wald statistic that has a theoretical normal
distribution, the assumptions of the Scott (FDRreg) are met.
This commit adds the Scott method (theoretical and empirical)
back into the RNA-seq case studies as well as the yeast and
polyester simulations.

@@
 We use the common `BenchDesign` with the set of multiple testing correction
-methods already included.
+methods already included. We also add in Scott's FDR Regression (both
+`nulltype = "empirical"` and `nulltype = "theoretical"`)
+since our test statistics are approximately t-distributed.
 
@@
     bd <- initializeBenchDesign()
+
+    bd <- addBMethod(bd, "scott-theoretical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'theoretical',
+                     control = list(lambda = 0.01))
+
+   bd <- addBMethod(bd, "scott-empirical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'empirical',
+                     control = list(lambda = 0.01))
+
     sb <- buildBench(bd, data = res, ftCols = "ind_covariate")

### 2018-05-14 datasets/RNA-Seq/RNAseq-mir200c.Rmd
Add Scott methods to RNA-seq case studies/sims
Since DESeq gives a Wald statistic that has a theoretical normal
distribution, the assumptions of the Scott (FDRreg) are met.
This commit adds the Scott method (theoretical and empirical)
back into the RNA-seq case studies as well as the yeast and
polyester simulations.

--- 2018-04-20 datasets/RNA-Seq/RNAseq-mir200c.Rmd
@@
 result_file <- file.path(resdir, "mir200c-results.rds")
-bench_file <- file.path(resdir, "mir200c-benchmark.rds")
+bench_file <- file.path(sbdir, "mir200c-benchmark.rds")
 
@@
 
-We use the common `BenchDesign` with the set of multiple testing correction methods already included.
+We use the common `BenchDesign` with the set of multiple testing correction 
+methods already included. We also add in Scott's FDR Regression (both
+`nulltype = "empirical"` and `nulltype = "theoretical"`)
+since our test statistics are t-distributed.
 

### 2018-05-14 datasets/RNA-Seq/polyester-simulation.Rmd
Add Scott methods to RNA-seq case studies/sims
Since DESeq gives a Wald statistic that has a theoretical normal
distribution, the assumptions of the Scott (FDRreg) are met.
This commit adds the Scott method (theoretical and empirical)
back into the RNA-seq case studies as well as the yeast and
polyester simulations.

@@
 library(BiocParallel)
-nCores <- 1
-multicoreParam <- MulticoreParam(workers = nCores)
+nCores <- 20
+register(MulticoreParam(workers = nCores))
 ```
@@
 #' @param bd is the bench design object
-simulateOneSplit <- function(counts, paramPolyester, 
-                             nDE, sampleSize, bd, BPPARAM){
+simulateOneSplit <- function(X, counts, paramPolyester, 
+                             nDE, sampleSize, bd){
 
@@
 
+We also add in Scott's FDR Regression (both
+`nulltype = "empirical"` and `nulltype = "theoretical"`)
+since our test statistics are approximately t-distributed.
+
+```{r}
+bd <- addBMethod(bd, "scott-theoretical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'theoretical',
+                     control = list(lambda = 0.01))
+bd <- addBMethod(bd, "scott-empirical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'empirical',
+                     control = list(lambda = 0.01))
+```
+
 Here's a helper function to return the number of methods with rejections at
@@
   null5 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
-                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   BPPARAM=multicoreParam)
+                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd)
   saveRDS(null5, file=resfile_n5)
@@
 }
+
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(null5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
 ```
@@
   null10 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
-                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   BPPARAM=multicoreParam)
+                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd)
   saveRDS(null10, file=resfile_n10)
@@
 }
+
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(null10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
 ```
@@
   de5 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
-                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   BPPARAM=multicoreParam)
+                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd)
   saveRDS(de5, file=resfile_d5)
@@
 }
+
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
 ```
@@
   de10 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
-                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   BPPARAM=multicoreParam)
+                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd)
   saveRDS(de10, file=resfile_d10)
@@
 }
+
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
 ```

### 2018-05-14 datasets/RNA-Seq/yeast-simulation.Rmd
Add Scott methods to RNA-seq case studies/sims
Since DESeq gives a Wald statistic that has a theoretical normal
distribution, the assumptions of the Scott (FDRreg) are met.
This commit adds the Scott method (theoretical and empirical)
back into the RNA-seq case studies as well as the yeast and
polyester simulations.

@@
 # set up parallel backend
-library(parallel)
-cores <- 20
+library(BiocParallel)
+nCores <- 20
+register(MulticoreParam(workers = nCores))
 ```
@@
 bd <- initializeBenchDesign()
-sb <- bd %>% buildBench(data=geneExp, parallel = F)
+sb <- bd %>% buildBench(data=geneExp, parallel = FALSE)
 
@@
 #' @param bd is the bench design object
-simulateOneSplit <- function(X, rseed, dds, resFULL, nDE, sampleSize, bd){
+#' @param BPPARAM is the BiocParallel bpparam argument to pass to DESeq. Since
+#'  parallelization happens by forking multiple instances of this function, we 
+#'  will specify only one worker per DEseq intance.
+simulateOneSplit <- function(X, rseed, dds, resFULL, nDE, sampleSize,
+                             bd){
 
@@
 
-  dds_test <- DESeq(dds_test)
+  dds_test <- DESeq(dds_test, parallel = FALSE)
   resultsNames(dds_test) # lists the coefficients
-  resTEST <- results(dds_test, name="fake_B_vs_A", independentFiltering = F)
+  resTEST <- results(dds_test, name="fake_B_vs_A", independentFiltering = FALSE)
 
@@
 
-  sb <- bd %>% buildBench(data=geneExp, parallel = F, 
+  sb <- bd %>% buildBench(data=geneExp, parallel = FALSE, 
                           truthCols = "qvalue",
                           ftCols = "ind_covariate")
+
+  assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+  pf <- estimatePerformanceMetrics(sb, tidy=TRUE, alpha=seq(0.01, 0.1, 0.01))
+pf %>% filter(performanceMetric %in% c("rejections", "FDR", "TPR")) %>%
+  with(table(is.na(value), blabel, performanceMetric))
+
   rowData(sb)$log2FC <- geneExp$effect_size
@@
 ualpha <- 0.05
+```
+
+We also add in Scott's FDR Regression (both
+`nulltype = "empirical"` and `nulltype = "theoretical"`)
+since our test statistics are approximately t-distributed.
+
+```{r}
+bd <- addBMethod(bd, "scott-theoretical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'theoretical',
+                     control = list(lambda = 0.01))
+bd <- addBMethod(bd, "scott-empirical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'empirical',
+                     control = list(lambda = 0.01))
 ```
@@
 
-```{r, results='hide', message=FALSE}
+```{r n5, results='hide', message=FALSE}
 rseed <- 225
@@
 if (!file.exists(resfile_n5)){
-  null5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
-                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   mc.cores=cores)
+  null5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
+                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd)
   saveRDS(null5, file = resfile_n5)
@@
 
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(null5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
 ```
@@
 
-```{r, results='hide', message=FALSE}
+```{r n10, results='hide', message=FALSE}
 rseed <- 837
@@
 if (!file.exists(resfile_n10)){
-  null10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
-                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   mc.cores=cores)
+  null10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
+                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd)
   saveRDS(null10, file = resfile_n10)
@@
 
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(null10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
 ```
@@
 
-```{r, results='hide', message=FALSE}
+```{r d5, results='hide', message=FALSE}
 rseed <- 198
@@
 if (!file.exists(resfile_d5)){
-  de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
-                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   mc.cores=cores)
+  de5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
+                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd)
   saveRDS(de5, file = resfile_d5)
@@
 
-```{r, results='hide', message=FALSE}
+```{r d10, results='hide', message=FALSE}
 rseed <- 961
@@
 if(!file.exists(resfile_d10)){
-  de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
-                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   mc.cores=cores)
+  de10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
+                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd)
   saveRDS(de10, file = resfile_d10)

### 2018-05-15 datasets/RNA-Seq/polyester-simulation.Rmd
Add method heatmaps to case studies with more than one covariate/comparison

@@
 }
-
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(null5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
 ```
@@
 ```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(null5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
 res5 <- plotsim_standardize(null5, alpha = seq(0.01, 0.10, 0.01))
@@
 }
-
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(null10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
 ```
@@
 ```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(null10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
 res10 <- plotsim_standardize(null10, alpha = seq(0.01, 0.10, 0.01))
@@
 }
-
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(de5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
 ```
@@
 ```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
 
@@
 
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(de10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
 ```
@@
 ```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
 res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.20, 0.01))
@@
 
+# Sample size comparison
+
+Here we compare the method ranks for the different sample sizes at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(resfile_d5, resfile_d10), 
+                colLabels = c("DE 5", "DE 10"), 
+                alpha = 0.10, xlab = "Comparison",
+                excludeMethods = NULL)
+```
+
 # Session information

### 2018-05-15 datasets/RNA-Seq/yeast-simulation.Rmd
Add method heatmaps to case studies with more than one covariate/comparison

@@
 
-Run benchmark methods.
+Build common bench design. We also add in Scott's FDR Regression (both
+`nulltype = "empirical"` and `nulltype = "theoretical"`)
+since our test statistics are approximately t-distributed.
 
@@
 bd <- initializeBenchDesign()
+bd <- addBMethod(bd, "scott-theoretical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'theoretical',
+                     control = list(lambda = 0.01))
+bd <- addBMethod(bd, "scott-empirical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'empirical',
+                     control = list(lambda = 0.01))
+```
+
+Run benchmark methods. 
+```{r}
 sb <- bd %>% buildBench(data=geneExp, parallel = FALSE)
@@
 B <- 100
-bd <- initializeBenchDesign() # only needs to be done once
 excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
 ualpha <- 0.05
-```
-
-We also add in Scott's FDR Regression (both
-`nulltype = "empirical"` and `nulltype = "theoretical"`)
-since our test statistics are approximately t-distributed.
-
-```{r}
-bd <- addBMethod(bd, "scott-theoretical",
-                     FDRreg::FDRreg,
-                     function(x) { x$FDR },
-                     z = test_statistic,
-                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
-                     nulltype = 'theoretical',
-                     control = list(lambda = 0.01))
-bd <- addBMethod(bd, "scott-empirical",
-                     FDRreg::FDRreg,
-                     function(x) { x$FDR },
-                     z = test_statistic,
-                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
-                     nulltype = 'empirical',
-                     control = list(lambda = 0.01))
 ```
@@
 ```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
 res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
@@
 }
-
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(de10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
 ```
@@
 ```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
 res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.10, 0.01))
@@
 
+# Sample size comparison
+
+Here we compare the method ranks for the different sample sizes at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(resfile_d5, resfile_d10), 
+                colLabels = c("DE 5", "DE 10"), 
+                alpha = 0.10, xlab = "Comparison", 
+                excludeMethods = NULL)
+```
+
 # Session information

### 2018-05-15 datasets/microbiome/microbiome-enigma.Rmd
Add method heatmaps to case studies with more than one covariate/comparison

--- 2018-05-10 datasets/microbiome/microbiome-enigma.Rmd
@@
 ```
+# Response comparison
+
+Here we compare the method ranks for the different comparisons at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(so4_bench_file, ph_bench_file, al_bench_file), 
+                colLabels = c("SO4", "pH", "Al"), 
+                alpha = 0.10, xlab = "Comparison")
+```
 

### 2018-05-15 datasets/R/plotMethodRank.R
Set proportion of max rejects to zero if max==0

@@ -54,19 +54,22 @@ plotMethodRanks <- function(objects, colLabels, alpha = 0.10,
       
       if (sum(hasResults) > 0){
         tmp <- estimatePerformanceMetrics(x, alpha, tidy=TRUE) %>%
-          filter( performanceMetric == "rejections") %>%
+          dplyr::filter( performanceMetric == "rejections") %>%
           dplyr::rename( method = blabel) %>%
-          filter( is.na(param.alpha) | (param.alpha == alpha)) %>%
-          filter( is.na(param.smooth.df) | (param.smooth.df == "3L")) %>%
-          filter( !method == "unadjusted") %>%
-          filter( !(method %in% NAmethods)) %>%
+          dplyr::filter( is.na(param.alpha) | (param.alpha == alpha)) %>%
+          dplyr::filter( is.na(param.smooth.df) | (param.smooth.df == "3L")) %>%
+          dplyr::filter( !method == "unadjusted") %>%
+          dplyr::filter( !(method %in% NAmethods)) %>%
           select( method, value ) %>%
           dplyr::rename( nrejects = value) %>%
           mutate( method = gsub("-df03", "", method)) %>%
           mutate( method = gsub("(-a)(.*)", "", method)) %>%
           na.omit() %>%
           mutate( rank = rank(nrejects) / n(),
-                  propMaxRej = nrejects / max(nrejects, na.rm = TRUE)) %>%
+                  propMaxRej = ifelse(nrejects == min(nrejects) & 
+                                      nrejects == max(nrejects) &
+                                      nrejects == 0, 0, 
+                                      nrejects / max(nrejects, na.rm = TRUE))) %>%
           mutate(casestudy = colLabels[i])
         
         ranks <- rbind(ranks, tmp)
@@ -89,12 +92,12 @@ plotMethodRanks <- function(objects, colLabels, alpha = 0.10,
     }
     
   }else{
-    ranks <- tidy_df(objects)
+    ranks <- tidy_df(objects, colLabels)
   }
   
   # exclude methods 
   if(!is.null(excludeMethods)){
-    ranks <- ranks %>% filter( !(method %in% excludeMethods))
+    ranks <- ranks %>% dplyr::filter( !(method %in% excludeMethods))
   }
   
   # add NAs for missing combinations of methods and casestudies
@@ -136,7 +139,7 @@ plotMethodRanks <- function(objects, colLabels, alpha = 0.10,
         theme_bw() +
[…]

### 2018-05-15 manuscript_figures/manuscript_figures.Rmd
Add Scott methods to RNA-seq case studies/sims
Since DESeq gives a Wald statistic that has a theoretical normal
distribution, the assumptions of the Scott (FDRreg) are met.
This commit adds the Scott method (theoretical and empirical)
back into the RNA-seq case studies as well as the yeast and
polyester simulations.
Add Scott to RNA-seq heatmaps; save 2E separately

@@
 # Figure 2D: Some methods are tailored to detecting smaller effect sizes
-p4 <- covariateLinePlot(de5, alpha=0.05, covname="log2FC", trans="log1p",
+p4 <- covariateLinePlot(de5, alpha=0.05, covname="ind_covariate", trans="log1p",
                         nbins=25)
 leg <- get_legend(p4)
-p4 <- p4 + theme(legend.position="none")
+p4 <- p4 + theme(legend.position="none") + 
+  ggtitle("Mean % significant by covariate %ile") + 
+  theme(axis.title = element_text(face="bold", size = 10),
+        plot.title = element_text(face="bold", size = 14)) +
+  xlab("Independent covariate percentile")
 
-# Figure 2E - upsetr plot -- this solution currently chops off left margin hist
+# Figure 2E - upsetr plot - save separately
+pdf(file.path(outdir, "Figure2E.pdf"), width = 10, height = 4, onefile=FALSE)
 aggupset(de5, alpha=alpha, supplementary = FALSE, 
@@
           gp = gpar(fontsize = 16, fontface = "bold"))
-p5 <- grid.grab()
+grid.text("E", x = 0.02, y = 0.97, 
+          gp = gpar(fontsize = 24, fontface = "bold")) 
+dev.off()
 
@@
 Fig2 <- plot_grid(Fig2, leg, ncol = 2, rel_widths = c(1, 0.2))
+Fig2
+ggsave(file.path(outdir, "Figure2A-D.pdf"), width=10, height=8)
 
-Fig2 <- plot_grid(Fig2, p5, ncol=1, rel_heights = c(1, 0.5),
-               labels = c("", "E"), label_size = 20)
-
-Fig2
-ggsave(file.path(outdir, "Figure2.pdf"), width=10, height=12)
 ```
@@
 ggsave(file.path(outdir, "Figure4.pdf"), width=6, height=4)
+
+Fig4 <- plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10,
+                        excludeMethods = NULL)
+Fig4
+ggsave(file.path(outdir, "Figure4_all.pdf"), width=6, height=4)
 
@@
                       function(x) x[2]))
-plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Covariate") + 
+plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Covariate",
+                excludeMethods = NULL) + 
   ggtitle("GWAS")
@@
                       function(x) x[1]))
-plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset") + 
+plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset", 
+                excludeMethods = NULL) + 
   ggtitle("RNA-seq")
@@
 ```{r}
-TODO
+wcs <- which(grepl("RNAseq", casestudy) & 
+            !grepl(c("brain|mir|polyester|null"), dataset) &
+            !grepl("scRNAseq", casestudy))
+objs <- objects[wcs]
+labs <- gsub(".rds", "", 
+             unlist(lapply(str_split(dataset[wcs], pattern = "-results-"), 
+                      function(x) x[2])))
+
+plotMethodRanks(objs, 
+                colLabels = labs, 
+                alpha = 0.10, xlab = "Comparison",
+                excludeMethods = NULL)
 ```
@@
 ```{r}
-TODO
+wcs <- which(grepl("RNAseq", casestudy) & 
+            !grepl(c("brain|mir|yeast|null"), dataset) &
+            !grepl("scRNAseq", casestudy))
+objs <- objects[wcs]
+labs <- gsub(".rds", "", 
+             unlist(lapply(str_split(dataset[wcs], pattern = "-results-"), 
+                      function(x) x[2])))
+
+plotMethodRanks(objs, 
+                colLabels = labs, 
+                alpha = 0.10, xlab = "Comparison",
+                excludeMethods = NULL)
 ```

### 2018-05-21 datasets/ChIP-seq/ChIPseq-h3k4me3-promoters.Rmd
Don't compute test statistic from pvalue in ChIP-seq

@@
     colnames(dat) <- c('ind_covariate', 'effect_size', 'SE',
-                       'stat', 'pval', 'padj')
-    ## compute test-statistic from p-value for ASH
-    dat <- dplyr::mutate(dat,
-                         test_statistic = qnorm(exp(log(pval)-log(2)),
-                                                lower.tail = FALSE),
-                         test_statistic = test_statistic * sign(effect_size))
-    ## remove regions with effect sizes = NA
+                       'test_statistic', 'pval', 'padj')
     dat <- dat[!is.na(dat$effect_size), ]

### 2018-05-21 manuscript_figures/manuscript_figures.Rmd
Heatmap improvements, adding table S3, save case study heatmaps

@@
 
-# remove covariates from individual cases studies that weren't informative
-## TODO ##
+# remove chip-seq promoter analysis (since it is the only one that uses 
+# ash and we also include the csaw analysis on the same dataset).
+promot <- which(grepl("promot", objects))
+if (length(promot) > 0){
+  objects <- objects[-promot]
+}
 
@@
 
+Fig4.rowOrder <- levels(plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10,
+                            tableOnly = TRUE)$method)
 Fig4 <- plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10)
 Fig4
-ggsave(file.path(outdir, "Figure4.pdf"), width=6, height=4)
+ggsave(file.path(outdir, "Figure4a.pdf"), width=6, height=4)
 
-Fig4 <- plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10,
+Fig4All.rowOrder <- levels(plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10, 
+                               tableOnly = TRUE, excludeMethods = NULL)$method)
+Fig4All <- plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10,
                         excludeMethods = NULL)
-Fig4
-ggsave(file.path(outdir, "Figure4_all.pdf"), width=6, height=4)
+Fig4All
+ggsave(file.path(outdir, "Figure4a_all.pdf"), width=6, height=4)
 
-# heatmap : rows method, cols casestudy- one color scale 
-Fig4alt <- plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10, 
-                        colorLow = "white", colorHigh = "blue", 
-                        colorNA = "grey50")
-Fig4alt
-ggsave(file.path(outdir, "Figure4alt.pdf"), width=6, height=4)
+```
 
-##
-## AR: alternative suggestion to heatmap, feel free to discard! ##
-##
+Extract a table of the values in Figure 4.
 
-Fig4b <- plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10,
-                         linePlot = TRUE)
-Fig4b
+```{r}
+# Supplementary table S3 
+library(xtable)
+tabS3 <- plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10,
+                        tableOnly = TRUE)
+tabS3 <- tabS3 %>% 
+  mutate(mean_prop = ifelse(is.na(mean_prop), "-",
+                            paste0(round(mean_prop,2), " (", round(min_prop,2), 
+                            "-", round(max_prop, 2), ")"))) %>%
+  select(casestudy, method, mean_prop) %>%
+  spread(casestudy, mean_prop)
+
+xtable(tabS3)
+```
+
+Create another panel with heatmap of yeast simulation results.
+
+```{r}
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+
+# grab casestudy names
+obj.names <- str_split(objects, pattern = "/")
+casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
+dataset <- unlist(lapply(obj.names, function(x) x[4] ))
+
+# non-null
+wcs <- which(grepl("RNAseq", casestudy) & 
+            !grepl(c("brain|mir|null|polyester"), dataset) &
+            !grepl("scRNAseq", casestudy))
+objs <- objects[wcs]
+labs <- gsub(".rds", "", 
+             unlist(lapply(str_split(dataset[wcs], pattern = "-results-"), 
+                      function(x) x[2])))
+
+pfdr <- plotMethodRanks(objs, 
+                colLabels = c("sim", "sim"), 
+                alpha = 0.10, xlab = "Comparison",
+                colorLow = "yellow", colorHigh = "navy",
+                fill = "FDR",
+                rowOrder = Fig4.rowOrder)
+
+ptpr <- plotMethodRanks(objs, 
+                colLabels = c("sim", "sim"), 
+                alpha = 0.10, xlab = "Comparison",
+                fill = "TPR",
+                rowOrder = Fig4.rowOrder)
+
+plot_grid(Fig4, pfdr + ylab("") + xlab(""), ptpr + ylab("") + xlab(""), 
+          ncol=3, rel_widths = c(1,0.3,0.3))
+ggsave(file.path(outdir, "Figure4.pdf"), width=12, height=5)
+
 ```
@@
 
-### ChIP-seq
+### GWAS
 
 ```{r}
-wcs <- which(grepl("ChIPseq", casestudy))
+# by covariate
+wcs <- which(grepl("GWAS", casestudy))
 objs <- objects[wcs]
-labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
-                      function(x) x[1]))
-plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset") + 
-  ggtitle("ChIP-Seq")
+labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-"), 
+                      function(x) x[2]))
+plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Covariate",
+                excludeMethods = NULL, Nlabel = FALSE) + 
+  ggtitle("GWAS")
+ggsave(file.path(outdir, "FigureS1.pdf"), width=4, height=4)
 ```
@@
                       function(x) x[1]))
-plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset") + 
+plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset", Nlabel = FALSE) + 
   ggtitle("GSEA")
+ggsave(file.path(outdir, "FigureS2.pdf"), width=4, height=4)
 ```
 
-### GWAS
+### ChIP-seq
 
 ```{r}
-# by covariate
-wcs <- which(grepl("GWAS", casestudy))
-objs <- objects[wcs]
-labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-"), 
-                      function(x) x[2]))
-plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Covariate",
-                excludeMethods = NULL) + 
-  ggtitle("GWAS")
-```
-
-### Microbiome
-
-```{r}
-# by dataset, excluding otu when genus is available
-# excluding log-ubiquity
-# excluding enigma-so4 (no discoveries)
-wcs <- which(grepl("microbiome", casestudy) & 
-              !(grepl(c("baxter|goodrich|papa"), dataset) & 
-              grepl(c("otu"), dataset)) &
-              !grepl("log-ubiquity", dataset) &
-              !grepl("so4", dataset))
+wcs <- which(grepl("ChIPseq", casestudy))
 objs <- objects[wcs]
@@
                       function(x) x[1]))
-plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset") + 
-  ggtitle("Microbiome")
+plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset", Nlabel = FALSE) + 
+  ggtitle("ChIP-Seq")
+ggsave(file.path(outdir, "FigureS3.pdf"), width=5, height=4)
 ```
@@
 plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset", 
-                excludeMethods = NULL) + 
+                excludeMethods = NULL, Nlabel = FALSE) + 
   ggtitle("RNA-seq")
+ggsave(file.path(outdir, "FigureS4.pdf"), width=4, height=4)
 ```
@@
 # by dataset
-plotMethodRanks(objs, colLabels = ds, alpha = 0.10, xlab = "Dataset") + 
+pa <- plotMethodRanks(objs, colLabels = ds, alpha = 0.10, xlab = "Dataset") + 
   ggtitle("scRNA-seq")
@@
 # by method
-plotMethodRanks(objs, colLabels = meth, alpha = 0.10, xlab = "Method") + 
+pb <- plotMethodRanks(objs, colLabels = meth, alpha = 0.10, xlab = "Method") + 
   ggtitle("scRNA-seq")
@@
 # by covariate
-plotMethodRanks(objs, colLabels = cv, alpha = 0.10, xlab = "Covariate") + 
+pc <- plotMethodRanks(objs, colLabels = cv, alpha = 0.10, xlab = "Covariate") + 
   ggtitle("scRNA-seq")
+
+plot_grid(pa, pb, pc)
+ggsave(file.path(outdir, "FigureS5.pdf"), width=8, height=5.5)
+```
+
+### Microbiome
+
+```{r}
+# by dataset, excluding otu when genus is available
+# excluding log-ubiquity
+# excluding enigma-so4 (no discoveries)
+wcs <- which(grepl("microbiome", casestudy) & 
+              !(grepl(c("baxter|goodrich|papa"), dataset) & 
+              grepl(c("otu"), dataset)) &
+              !grepl("log-ubiquity", dataset) &
+              !grepl("so4", dataset))
+objs <- objects[wcs]
+labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
+                      function(x) x[1]))
+plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset", Nlabel = FALSE) + 
+  ggtitle("Microbiome")
+ggsave(file.path(outdir, "FigureS6.pdf"), width=7.5, height=4)
 ```
@@
 ```{r}
+# non-null
 wcs <- which(grepl("RNAseq", casestudy) & 
@@
                 excludeMethods = NULL)
+
+plotMethodRanks(objs, 
+                colLabels = labs, 
+                alpha = 0.10, xlab = "Comparison",
+                excludeMethods = NULL, 
+                fill = "FDR")
+
+plotMethodRanks(objs, 
+                colLabels = labs, 
+                alpha = 0.10, xlab = "Comparison",
+                excludeMethods = NULL, 
+                fill = "TPR")
+
+# null
+wcs <- which(grepl("RNAseq", casestudy) & 
+            !grepl(c("brain|mir|polyester|de"), dataset) &
+            !grepl("scRNAseq", casestudy))
+objs <- objects[wcs]
+labs <- gsub(".rds", "", 
+             unlist(lapply(str_split(dataset[wcs], pattern = "-results-"), 
+                      function(x) x[2])))
+
+plotMethodRanks(objs, 
+                colLabels = labs, 
+                alpha = 0.10, xlab = "Comparison",
+                excludeMethods = NULL)
+
 ```
@@
                 excludeMethods = NULL)
+
+plotMethodRanks(objs, 
+                colLabels = labs, 
+                alpha = 0.10, xlab = "Comparison",
+                excludeMethods = NULL, 
+                fill = "FDR")
+
+plotMethodRanks(objs, 
+                colLabels = labs, 
+                alpha = 0.10, xlab = "Comparison",
+                excludeMethods = NULL, 
+                fill = "TPR")
+
+# null
+wcs <- which(grepl("RNAseq", casestudy) & 
+            !grepl(c("brain|mir|yeast|de"), dataset) &
+            !grepl("scRNAseq", casestudy))
+objs <- objects[wcs]
+labs <- gsub(".rds", "", 
+             unlist(lapply(str_split(dataset[wcs], pattern = "-results-"), 
+                      function(x) x[2])))
+
+plotMethodRanks(objs, 
+                colLabels = labs, 
+                alpha = 0.10, xlab = "Comparison",
+                excludeMethods = NULL) 
 ```

### 2018-05-21 datasets/R/plotMethodRank.R

@@ -22,19 +22,30 @@
 #' @param excludeMethods character vector containing the names of methods to
 #'  exclude from the heatmap. Default is to exclude Scott methods which are 
 #'  only present in GWAS.
-#' @param propMaxRejections logical indicating whether the proportion of max
-#'  number of rejections is plotted (instead of the rank). Default is TRUE. 
+#' @param fill character indicating the outcome variable of interest. Default
+#'  is 'propMaxRejections'. For 'FDR' and 'TPR' sb objects need to contain 
+#'  these performance metrics.
 #' @param xlab character with the x label (default is "Case Study")   
+#' @param tableOnly logical whether or not to return a tibble of the summarized
+#'  results instead of a plot.
+#'  @param rowOrder numerical vector with order of rows/methods (default is 
+#'  NULL, which means rows will be ordered by mean fill value).
+#'  @param Nlabel logical whether or not to include the number of datasets 
+#'  averaged over in the x-axis casestudy label. Default is TRUE.
 
 plotMethodRanks <- function(objects, colLabels, alpha = 0.10, 
                             colorLow = "navy", colorHigh = "yellow",
                             colorNA = "white", xlab = "Case Study",
-                            propMaxRejections = TRUE,
+                            fill = c("propMaxRejections", "meanRank", 
+                                     "FDR", "TPR"),
                             linePlot = FALSE,
                             excludeMethods = c("scott-theoretical",
-                                               "scott-empirical")){
-  
-  tidy_df <- function(objects, colLabels){
+                                               "scott-empirical"),
+                            rowOrder = NULL,
+                            tableOnly = FALSE,
+                            Nlabel = TRUE){
+  fill <- match.arg(fill)
+  tidy_df <- function(objects, colLabels, fill){
     # create tidy data frame where each row is a method / dataset observation
     # of a rank 
     ranks <- data.frame()
@@ -52,9 +63,19 @@ plotMethodRanks <- function(objects, colLabels, alpha = 0.10,
       hasResults <- apply(!is.na( assays(x)[["qvalue"]] ), 2, sum)
[…]

### 2018-05-25 datasets/GWAS/GWAS.Rmd
Add method heatmaps to case studies with more than one covariate/comparison
Save effect size in bench object

@@
     sbN <- bd %>% buildBench(data=(bmi %>% mutate(ind_covariate=ind_covar_N)), 
-                             ftCols = c("ind_covar_N"),
+                             ftCols = c("ind_covar_N", "effect_size"),
                              parallel=TRUE, BPPARAM=multicoreParam)
@@
                                     filter(!is.na(ind_covariate))), 
-                              ftCols = c("ind_covar_AF"),
+                              ftCols = c("ind_covar_AF", "effect_size"),
                               parallel=TRUE, BPPARAM=multicoreParam)
@@
 ```{r, fig.width=8, fig.height=3.5}
-mcols(sbN)$effect_size <- bmi %>% filter(!is.na(ind_covar_N)) %>% pull(effect_size)
 covariateLinePlot(sbN, alpha=0.05, covname="effect_size", nbins=25, 
@@
 ```{r, fig.width=8, fig.height=3.5}
-mcols(sbAF)$effect_size <- bmi %>% filter(!is.na(ind_covar_AF)) %>% pull(effect_size)
 covariateLinePlot(sbAF, alpha=0.05, covname="effect_size", nbins=25, 
@@
 covariateLinePlot(sbAF, alpha=0.05, covname="ind_covar_AF", nbins=25)
+```
+
+# Covariate comparison
+
+Here we compare the method ranks for the two covariates at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(resfile_AF, resfile_N), 
+                colLabels = c("MAF", "Sample Size"), 
+                alpha = 0.10, xlab = "Covariate", 
+                excludeMethods = NULL)
 ```

### 2018-05-25 datasets/R/effectSizePlot.R
Allow transformation of x-axis in covariate lineplot

@@ -104,12 +104,13 @@ return(p)
 #' @param trans character indicating a transformation to apply to the scale
 #' of the y-axis (lineplot) or fill (heatmap). For example, "log1p" applies 
 #' the log(x+1) transform. The default NULL means no transformation. 
+#' @param transx same as \code{trans}, but for x-axis. Only for lineplot.
 #' @param linePlot logical whether to plot a line plot (smoothed density) 
 #' (default is TRUE). If false, will plot a heatmap.
 #' 
 #' @return a ggplot object
 covariateLinePlot <- function(sbl, alpha=0.05, nbins = 25, 
-                             covname, trans = NULL, 
+                             covname, trans = NULL, transx = NULL,
                              linePlot = TRUE, palette=candycols){
   
   summarize_one_item <- function(object, alpha, nbins){
@@ -179,6 +180,10 @@ covariateLinePlot <- function(sbl, alpha=0.05, nbins = 25,
     
     if(!is.null(trans)){
       p <- p + scale_y_continuous(trans=trans)
+    }
+    
+    if(!is.null(transx)){
+      p <- p + scale_x_continuous(trans=transx)
     }
   }else{
     p <- ggplot(df, aes(x = as.factor(Method), y = bin/nbins)) +

### 2018-05-25 manuscript_figures/manuscript_figures.Rmd
Add Figure 5, and a supp figure

@@
 # Figure 2D: Some methods are tailored to detecting smaller effect sizes
-p4 <- covariateLinePlot(de5, alpha=0.05, covname="ind_covariate", trans="log1p",
+p4 <- covariateLinePlot(de5, alpha=alpha, covname="log2FC", trans="log1p",
                         nbins=25)
@@
 p4 <- p4 + theme(legend.position="none") + 
-  ggtitle("Mean % significant by covariate %ile") + 
+  ggtitle("Mean % significant by effect size %ile") + 
   theme(axis.title = element_text(face="bold", size = 10),
         plot.title = element_text(face="bold", size = 14)) +
-  xlab("Independent covariate percentile")
+  xlab("Effect size (log2 fold change) percentile")
 
@@
 This is a heatmap of methods by case studies with color representing the ranking
-by of number of rejections @ FDR cutoff 0.10.
+by of number of rejections @ FDR cutoff `r alpha`.
 
@@
 
-Fig4.rowOrder <- levels(plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10,
+Fig4.rowOrder <- levels(plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
                             tableOnly = TRUE)$method)
-Fig4 <- plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10)
+Fig4 <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha)
 Fig4
@@
 
-Fig4All.rowOrder <- levels(plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10, 
+Fig4All.rowOrder <- levels(plotMethodRanks(objects, colLabels = casestudy, alpha = alpha, 
                                tableOnly = TRUE, excludeMethods = NULL)$method)
-Fig4All <- plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10,
+Fig4All <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
                         excludeMethods = NULL)
@@
 library(xtable)
-tabS3 <- plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10,
+tabS3 <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
                         tableOnly = TRUE)
@@
                 colLabels = c("sim", "sim"), 
-                alpha = 0.10, xlab = "Comparison",
+                alpha = alpha, xlab = "Comparison",
                 colorLow = "yellow", colorHigh = "navy",
@@
                 colLabels = c("sim", "sim"), 
-                alpha = 0.10, xlab = "Comparison",
+                alpha = alpha, xlab = "Comparison",
                 fill = "TPR",
@@
 This is a multi-panel Figure displaying examples of covariate line plots in the
-case studies.
+case studies. We'll create a 4-panel figure with the following case studies:
+RNAseq, scRNAseq, GSEA, GWAS.
 
 ```{r, Figure 5}
-# TO DO
+
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+resfile <- objects[grepl("RNAseq", objects) & grepl("brain", objects)] 
+res <- readRDS(file=resfile)
+assayNames(res) <- "qvalue"
+res <- addDefaultMetrics(res)
+pA <- covariateLinePlot(res, alpha=alpha, covname="ind_covariate", trans="log1p",
+                        nbins=25) +
+  ggtitle("RNA-seq case study: Brain") +
+  xlab("Overall mean expression percentile")
+
+leg <-  get_legend(pA)
+pA <- pA + theme(legend.position="none")
+
+resfile <- objects[grepl("RNAseq", objects) & grepl("human", objects) &
+                   grepl("mast", objects) & grepl("det", objects)] 
+res <- readRDS(file=resfile)
+assayNames(res) <- "qvalue"
+res <- addDefaultMetrics(res)
+pB <- covariateLinePlot(res, alpha=alpha, covname="detection", trans="log1p",
+                        nbins=25) +
+  ggtitle("scRNA-seq case study: Human, MAST") +
+  xlab("Detection rate percentile") + 
+  theme(legend.position="none")
+
+resfile <- objects[grepl("GSEA", objects) & grepl("mouse", objects)] 
+res <- readRDS(file=resfile)
+assayNames(res) <- "qvalue"
+res <- addDefaultMetrics(res)
+pC <- covariateLinePlot(res, alpha=alpha, covname="ind_covariate", trans="log1p",
+                        nbins=25) +
+  ggtitle("GSEA case study: Mouse") +
+  xlab("Gene Set size percentile") + 
+  theme(legend.position="none")
+
+resfile <- objects[grepl("GWAS", objects) & grepl("maf", objects)] 
+res <- readRDS(file=resfile)
+assayNames(res) <- "qvalue"
+res <- addDefaultMetrics(res)
+pD <- covariateLinePlot(res, alpha=alpha, covname="ind_covar_AF", trans="log1p",
+                        nbins=25) +
+  ggtitle("GWAS case study")+
+  xlab("Minor Allele Frequency (MAF) percentile") +
+  theme(legend.position="none")
+
+p5 <- plot_grid(pA, pB, pC, pD, ncol=2, labels = LETTERS[1:4])
+p5 <- plot_grid(p5, leg, ncol=2, rel_widths = c(1,0.2))
+ggsave(file.path(outdir, "Figure5.pdf"), width=9, height=7)
+
 ```
@@
 # Supplementary Figures
+
+## Case-study specific plot analagous to Figure 2
+
+Figure 2 panels C-E for a case study. For now picking GWAS.
+
+```{r Figure2.5}
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+resfile <- objects[grepl("GWAS", objects) & grepl("maf", objects)] 
+res <- readRDS(file=resfile)
+assayNames(res) <- "qvalue"
+res <- addDefaultMetrics(res)
+
+# Figure 2C: Total number of discoveries varied across method
+p3 <- rejections_scatter(res, supplementary = FALSE) +
+      theme(legend.position="none")+ 
+  ggtitle("Number of rejections by alpha cutoff") + 
+  xlab("alpha cutoff") +
+  theme(axis.title = element_text(face="bold", size = 10),
+        plot.title = element_text(face="bold", size = 14))
+
+# Figure 2D: Some methods are tailored to detecting smaller effect sizes
+p4 <- covariateLinePlot(res, alpha=alpha, covname="effect_size", trans="log1p",
+                        transx = "exp", nbins=25)
+leg <- get_legend(p4)
+p4 <- p4 + theme(legend.position="none") + 
+  ggtitle("Mean % significant by effect size %ile") + 
+  theme(axis.title = element_text(face="bold", size = 10),
+        plot.title = element_text(face="bold", size = 14)) +
+  xlab("Effect size percentile")
+
+# Figure 2E - upsetr plot - save separately
+pdf(file.path(outdir, "FigureS0C.pdf"), width = 10, height = 4, onefile=FALSE)
+plotFDRMethodsOverlap(res, 
+                      alpha=alpha, nsets=ncol(res),
+                      order.by="freq", decreasing=TRUE,
+                      supplementary=FALSE, nintersects = 18) 
+grid.text("Overlap for GWAS Case study, MAF covariate", x = 0.6, y = 0.97, 
+          gp = gpar(fontsize = 16, fontface = "bold"))
+grid.text("C", x = 0.02, y = 0.97, 
+          gp = gpar(fontsize = 24, fontface = "bold")) 
+dev.off()
+
+# Figure 2 (ALL)
+Fig2 <- plot_grid(p3, p4, ncol = 2, 
+          labels = LETTERS[1:2], label_size = 20)
+Fig2 <- plot_grid(Fig2, leg, ncol = 2, rel_widths = c(1, 0.2))
+Fig2
+ggsave(file.path(outdir, "FigureS0A-B.pdf"), width=10, height=4)
+
+```
 
@@
                       function(x) x[2]))
-plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Covariate",
+plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Covariate",
                 excludeMethods = NULL, Nlabel = FALSE) + 
@@
                       function(x) x[1]))
-plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset", Nlabel = FALSE) + 
+plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", Nlabel = FALSE) + 
   ggtitle("GSEA")
@@
                       function(x) x[1]))
-plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset", Nlabel = FALSE) + 
+plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", Nlabel = FALSE) + 
   ggtitle("ChIP-Seq")
@@
                       function(x) x[1]))
-plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset", 
+plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", 
                 excludeMethods = NULL, Nlabel = FALSE) + 
@@
 # by dataset
-pa <- plotMethodRanks(objs, colLabels = ds, alpha = 0.10, xlab = "Dataset") + 
+pa <- plotMethodRanks(objs, colLabels = ds, alpha = alpha, xlab = "Dataset") + 
   ggtitle("scRNA-seq")
@@
 # by method
-pb <- plotMethodRanks(objs, colLabels = meth, alpha = 0.10, xlab = "Method") + 
+pb <- plotMethodRanks(objs, colLabels = meth, alpha = alpha, xlab = "Method") + 
   ggtitle("scRNA-seq")
@@
 # by covariate
-pc <- plotMethodRanks(objs, colLabels = cv, alpha = 0.10, xlab = "Covariate") + 
+pc <- plotMethodRanks(objs, colLabels = cv, alpha = alpha, xlab = "Covariate") + 
   ggtitle("scRNA-seq")
@@
                       function(x) x[1]))
-plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset", Nlabel = FALSE) + 
+plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", Nlabel = FALSE) + 
   ggtitle("Microbiome")
@@
                 colLabels = labs, 
-                alpha = 0.10, xlab = "Comparison",
+                alpha = alpha, xlab = "Comparison",
                 excludeMethods = NULL)
@@
                 colLabels = labs, 
-                alpha = 0.10, xlab = "Comparison",
+                alpha = alpha, xlab = "Comparison",
                 excludeMethods = NULL, 
@@
                 colLabels = labs, 
-                alpha = 0.10, xlab = "Comparison",
+                alpha = alpha, xlab = "Comparison",
                 excludeMethods = NULL) 

### 2018-05-25 manuscript_figures/manuscript_figures.sh

@@ -2,7 +2,7 @@
 #SBATCH -J figs
 #SBATCH -n 1
 #SBATCH -N 1
-#SBATCH -p shared
+#SBATCH -p irizarry
 #SBATCH --mem 50G
 #SBATCH -t 0-1:00
 #SBATCH -o render-%j.out

### 2018-05-28 manuscript_figures/manuscript_figures.Rmd
Remove all Baxter(CRC) studies from Microbiome results figs

@@
 # nothing for all methods and the genus-level analysis is also present)
-null <- which(grepl(c("baxter|goodrich|papa"), objects) & 
-              grepl(c("otu"), objects))
+null <- which(grepl(c("goodrich|papa"), objects) & 
+              grepl(c("otu"), objects) | 
+              grepl("baxter", objects))
 
@@
 Fig4
-ggsave(file.path(outdir, "Figure4a.pdf"), width=6, height=4)
 
+# not excluding any methods
 Fig4All.rowOrder <- levels(plotMethodRanks(objects, colLabels = casestudy, alpha = alpha, 
@@
 Fig4All
-ggsave(file.path(outdir, "Figure4a_all.pdf"), width=6, height=4)
 
@@
 wcs <- which(grepl("microbiome", casestudy) & 
-              !(grepl(c("baxter|goodrich|papa"), dataset) & 
+              !(grepl(c("goodrich|papa"), dataset) & 
               grepl(c("otu"), dataset)) &
               !grepl("log-ubiquity", dataset) &
-              !grepl("so4", dataset))
+              !grepl("so4", dataset) &
+              !grepl("baxter", dataset) )
 objs <- objects[wcs]

### 2018-05-28 manuscript_figures/manuscript_figures.Rmd
Add separate color scale to sim in Fig4

@@
 Fig4All
-
 ```
@@
                 alpha = alpha, xlab = "Comparison",
-                colorLow = "yellow", colorHigh = "navy",
                 fill = "FDR",
-                rowOrder = Fig4.rowOrder)
+                rowOrder = Fig4.rowOrder) +
+          scale_fill_gradient2(high = scales::muted("red"), 
+                         mid = "white", 
+                         low = scales::muted("blue"), 
+                         midpoint = alpha)
 
@@
                 fill = "TPR",
+                colorLow = scales::muted("blue"), 
+                colorHigh = "white",
                 rowOrder = Fig4.rowOrder)

### 2018-05-28 manuscript_figures/manuscript_figures.Rmd
Decrease whitespace in Fig4; add A/B labels

@@
                 rowOrder = Fig4.rowOrder) +
-          scale_fill_gradient2(high = scales::muted("red"), 
+          scale_fill_gradient2(high = "red", 
                          mid = "white", 
-                         low = scales::muted("blue"), 
+                         low = "blue", 
                          midpoint = alpha)
@@
                 fill = "TPR",
-                colorLow = scales::muted("blue"), 
+                colorLow = "blue", 
                 colorHigh = "white",
@@
 
-plot_grid(Fig4, pfdr + ylab("") + xlab(""), ptpr + ylab("") + xlab(""), 
-          ncol=3, rel_widths = c(1,0.3,0.3))
+ggdraw() +
+  draw_plot(pfdr + ylab("") + xlab(""), 0.57, 0, 0.22, 1) +
+  draw_plot(ptpr + ylab("") + xlab(""), 0.78, 0, 0.22, 1) +
+  draw_plot(Fig4, 0, 0, 0.60, 1) +
+  draw_plot_label(c("A", "B"), c(0, 0.57), c(1, 1), size = 15)
 ggsave(file.path(outdir, "Figure4.pdf"), width=12, height=5)
-
 ```

### 2018-05-28 manuscript_figures/manuscript_figures.Rmd
Add layer to method rank heatmaps with prop total possible

@@
                             tableOnly = TRUE)$method)
-Fig4 <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha)
+Fig4 <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
+                        annotate = "propPossible")
 Fig4
@@
 
-Create another panel with heatmap of yeast simulation results.
+Create more panels with heatmaps of yeast simulation results.
 
@@
                 fill = "FDR",
-                rowOrder = Fig4.rowOrder) +
+                rowOrder = Fig4.rowOrder, 
+                annotate = "FDR") +
           scale_fill_gradient2(high = "red", 
@@
                 colorHigh = "white",
-                rowOrder = Fig4.rowOrder)
+                rowOrder = Fig4.rowOrder,
+                annotate = "TPR")
 
@@
 plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Covariate",
-                excludeMethods = NULL, Nlabel = FALSE) + 
+                excludeMethods = NULL, Nlabel = FALSE, annotate = "propPossible") + 
   ggtitle("GWAS")
@@
                       function(x) x[1]))
-plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", Nlabel = FALSE) + 
+plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", 
+                Nlabel = FALSE, annotate = "propPossible") + 
   ggtitle("GSEA")
@@
                       function(x) x[1]))
-plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", Nlabel = FALSE) + 
+plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", 
+                Nlabel = FALSE, annotate = "propPossible") + 
   ggtitle("ChIP-Seq")
@@
 plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", 
-                excludeMethods = NULL, Nlabel = FALSE) + 
+                excludeMethods = NULL, Nlabel = FALSE, annotate = "propPossible") + 
   ggtitle("RNA-seq")
@@
 # by dataset
-pa <- plotMethodRanks(objs, colLabels = ds, alpha = alpha, xlab = "Dataset") + 
+pa <- plotMethodRanks(objs, colLabels = ds, alpha = alpha, 
+                      xlab = "Dataset", annotate = "propPossible") + 
   ggtitle("scRNA-seq")
@@
 # by method
-pb <- plotMethodRanks(objs, colLabels = meth, alpha = alpha, xlab = "Method") + 
+pb <- plotMethodRanks(objs, colLabels = meth, alpha = alpha,
+                      xlab = "Method", annotate = "propPossible") + 
   ggtitle("scRNA-seq")
@@
 # by covariate
-pc <- plotMethodRanks(objs, colLabels = cv, alpha = alpha, xlab = "Covariate") + 
+pc <- plotMethodRanks(objs, colLabels = cv, alpha = alpha,
+                      xlab = "Covariate", annotate = "propPossible") + 
   ggtitle("scRNA-seq")
@@
                       function(x) x[1]))
-plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", Nlabel = FALSE) + 
+plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", 
+                Nlabel = FALSE, annotate = "propPossible") + 
   ggtitle("Microbiome")

### 2018-05-28 datasets/R/plotMethodRank.R

@@ -34,6 +34,10 @@
 #'  averaged over in the x-axis casestudy label. Default is TRUE. If FALSE
 #'  assumes each column represents one benchmark object (legend labels will not 
 #'  include the word "Mean").
+#' @param annotate character indicating what should be plotted as text labels
+#'  (heatmap plot only). Default is proportion of total possible rejections 
+#'  (propPossible). Could also be another variable (from \code{fill} choices) 
+#'  or NULL (for no text label annotations).
 #' @author Keegan Korthauer  
 
 plotMethodRanks <- function(objects, colLabels, alpha = 0.10, 
@@ -46,9 +50,10 @@ plotMethodRanks <- function(objects, colLabels, alpha = 0.10,
                                                "scott-empirical"),
                             rowOrder = NULL,
                             tableOnly = FALSE,
-                            Nlabel = TRUE){
+                            Nlabel = TRUE,
+                            annotate = "propPossible"){
   fill <- match.arg(fill)
-  tidy_df <- function(objects, colLabels, fill){
+  tidy_df <- function(objects, colLabels, fill, annotate){
     # create tidy data frame where each row is a method / dataset observation
     # of a rank 
     ranks <- data.frame()
@@ -77,6 +82,11 @@ plotMethodRanks <- function(objects, colLabels, alpha = 0.10,
         if (fill %in% c("TPR", "FDR") && !(fill %in% tmp$performanceMetric))
           stop(fill, " is not found in performanceMetrics")
         
+        if (!is.null(annotate)){
+          if (annotate %in% c("TPR", "FDR") && !(annotate %in% tmp$performanceMetric))
+            stop(annotate, " is not found in performanceMetrics")
+        }
+        
         tmp <- tmp %>%
           dplyr::filter( performanceMetric == pmcol) %>%
           dplyr::rename( method = blabel) %>%
@@ -93,7 +103,8 @@ plotMethodRanks <- function(objects, colLabels, alpha = 0.10,
                   propMaxRej = ifelse(nrejects == min(nrejects) & 
                                       nrejects == max(nrejects) &
                                       nrejects == 0, 0, 
-                                      nrejects / max(nrejects, na.rm = TRUE))) %>%
+                                      nrejects / max(nrejects, na.rm = TRUE)),
[…]

### 2018-05-28 datasets/RNA-Seq/polyester-simulation.Rmd
Add uninformative covariate sim to yeast and poly

@@
 
+resfile_n5_uninfCov <- file.path(resdir, "polyester-results-null5-uninfCov.rds")
+resfile_n10_uninfCov <- file.path(resdir, "polyester-results-null10-uninfCov.rds")
+resfile_d5_uninfCov <- file.path(resdir, "polyester-results-de5-uninfCov.rds")
+resfile_d10_uninfCov <- file.path(resdir, "polyester-results-de10-uninfCov.rds")
+
 # # set up parallel backend:
@@
 #' @param bd is the bench design object
+#' @param uninformativeCovariate logical indicating whether to replace overall mean 
+#'  independent covariate with a randomly generated (uninformative) covariate. 
+#'  Default is FALSE.
 simulateOneSplit <- function(X, counts, paramPolyester, 
-                             nDE, sampleSize, bd){
+                             nDE, sampleSize, bd, uninformativeCovariate = FALSE){
 
@@
                                pval=res_alt$pvalue, SE=res_alt$lfcSE, 
-                               ind_covariate = res_alt$baseMean, 
+                               ind_covariate = ifelse(uninformativeCovariate, 
+                                                      rnorm(nrow(res_alt)),
+                                                      res_alt$baseMean), 
                                effect_size=res_alt$log2FoldChange, 
@@
 
+# Uninformative Covariate
+
+## Sim N5: Null Comparison 5v5
+
+Here we run a null comparison 5 versus 5 RNA-Seq samples. 
+This will be done for 100 simulations.
+
+### Generate a list of SB results objects
+
+```{r, results='hide', message=FALSE}
+sampleSize <- 5
+nDE <- 0
+
+if (!file.exists(resfile_n5_uninfCov)){
+  null5 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
+                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, 
+                   bd=bd, uninformativeCovariate = TRUE)
+  saveRDS(null5, file=resfile_n5_uninfCov)
+} else {
+  null5 <- readRDS(file=resfile_n5_uninfCov)
+}
+```
+
+### Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(null5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res5 <- plotsim_standardize(null5, alpha = seq(0.01, 0.10, 0.01))
+
+# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
+plotsim_average(res5, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5, met="TNR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(null5, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(null5, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res5, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+  aggupset(null5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+  message("Not enough methods found rejections at alpha ", ualpha, 
+          "; skipping upset plot")
+}
+
+```
+
+## Sim N10: Null Comparison 10v10
+
+Here we'll repeat the above, but for a null comparison 10 versus 10 samples. 
+This will be done for 100 simulations.
+
+### Generate a list of SB results objects
+
+```{r, results='hide', message=FALSE}
+sampleSize <- 10
+nDE <- 0
+
+if (!file.exists(resfile_n10_uninfCov)){
+  null10 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
+                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, 
+                   bd=bd, uninformativeCovariate = TRUE)
+  saveRDS(null10, file=resfile_n10_uninfCov)
+} else {
+  null10 <- readRDS(file=resfile_n10_uninfCov)
+}
+```
+
+### Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(null10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res10 <- plotsim_standardize(null10, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(res10, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10, met="TNR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(null10, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(null10, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res10, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+  aggupset(null10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
+}
+
+```
+
+## Sim D5: DE Comparison 5v5
+
+Here we'll repeat the above, but for a DE comparison 5 versus 5 samples
+and 1500 DE genes are added. This will be done for 100 simulations.
+
+### Generate a list of SB results objects
+
+```{r, results='hide', message=FALSE}
+sampleSize <- 5
+nDE <- 1500
+
+if (!file.exists(resfile_d5_uninfCov)){
+  de5 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
+                  paramPolyester=params, nDE=nDE, sampleSize=sampleSize, 
+                  bd=bd, uninformativeCovariate = TRUE)
+  saveRDS(de5, file=resfile_d5_uninfCov)
+} else {
+  de5 <- readRDS(file=resfile_d5_uninfCov)
+}
+```
+
+### Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.20, 0.01))
+
+plotsim_average(res5d, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5d, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5d, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(de5, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(de5, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+  aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
+}
+
+```
+
+## Sim D10: DE Comparison 10v10
+
+Here we'll repeat the above, but for a DE comparison 10 versus 10 samples
+and 1500 DE genes are added. This will be done for 100 simulations.
+
+### Generate a list of SB results objects
+
+```{r, results='hide', message=FALSE}
+sampleSize <- 10
+nDE <- 1500
+
+if (!file.exists(resfile_d10_uninfCov)){
+  de10 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
+                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, 
+                   bd=bd, uninformativeCovariate = TRUE)
+  saveRDS(de10, file=resfile_d10_uninfCov)
+} else {
+  de10 <- readRDS(file=resfile_d10_uninfCov)
+}
+
+```
+
+### Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.20, 0.01))
+
+plotsim_average(res10d, met="rejections",filter_set = excludeSet, 
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10d, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10d, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(de10, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(de10, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res10d, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+  aggupset(de10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
+}
+```
+
+## Sample size comparison
+
+Here we compare the method ranks for the different sample sizes at alpha = 0.10.
+
+```{r}
+# uninformative covariate
+plotMethodRanks(c(resfile_d5_uninfCov, resfile_d10_uninfCov), 
+                colLabels = c("DE 5", "DE 10"), 
+                alpha = 0.10, xlab = "Comparison",
+                excludeMethods = NULL)
+```
+
 # Session information

### 2018-05-28 datasets/RNA-Seq/yeast-simulation.Rmd
Add uninformative covariate sim to yeast and poly

@@
 will allow us to (1) implement null comparisons on random subsets of samples within
-one condition, (2) compare FDR approaches on their ability to discover 'true positive' 
-DE genes using subsets of the full dataset, where 'true positive' is defined as a 
-gene detected as DE using all 48 samples in each condition, and (3) start with a null
-comparison and add in artificial differences to a subset of genes to define
-'true positives'.
+one condition, and (2) start with a null comparison and add in artificial differences
+to a subset of genes to define 'true positives'.
+
+We will investigate the overall mean expression as an informative and independent 
+(under the null hypothesis) covariate. 
+We will also investigate the effects of an uninformative covariate by using a 
+randomly generated covariate.
 
@@
 resfile_d10 <- file.path(resdir, "yeast-results-de10.rds")
+
+resfile_n5_uninfCov <- file.path(resdir, "yeast-results-null5-uninfCov.rds")
+resfile_n10_uninfCov <- file.path(resdir, "yeast-results-null10-uninfCov.rds")
+resfile_d5_uninfCov <- file.path(resdir, "yeast-results-de5-uninfCov.rds")
+resfile_d10_uninfCov <- file.path(resdir, "yeast-results-de10-uninfCov.rds")
 
@@
 #'  will specify only one worker per DEseq intance.
+#' @param uninformativeCovariate logical indicating whether to replace overall mean 
+#'  independent covariate with a randomly generated (uninformative) covariate. 
+#'  Default is FALSE.
 simulateOneSplit <- function(X, rseed, dds, resFULL, nDE, sampleSize,
-                             bd){
+                             bd, uninformativeCovariate = FALSE){
 
@@
                                SE=resTEST$lfcSE, 
-                               ind_covariate = resTEST$baseMean,
+                               ind_covariate = ifelse(uninformativeCovariate, 
+                                                      rnorm(nrow(resTEST)),
+                                                      resTEST$baseMean),
                                effect_size = resTEST$log2FoldChange, 
@@
 
+# Uninformative covariate
+
+Here we repeat the previous four sections, using an uninformative covariate 
+instead of the overall mean expression.
+
+## Sim N5: Null Comparison 5v5
+
+Here we'll repeat the above, but for a null comparison 5 versus 5 Snf2 samples,
+where the groups are selected randomly. This will be done for 100 random splits.
+
+### Generate a list of SB results objects
+
+```{r n5-uc, results='hide', message=FALSE}
+rseed <- 225
+sampleSize <- 5
+nDE <- 0
+
+if (!file.exists(resfile_n5_uninfCov)){
+  null5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
+                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd,
+                   uninformativeCovariate = TRUE)
+  saveRDS(null5, file = resfile_n5_uninfCov)
+}else{
+  null5 <- readRDS(file = resfile_n5_uninfCov)
+}
+```
+
+### Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(null5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res5 <- plotsim_standardize(null5, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(res5, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5, met="TNR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(null5, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(null5, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res5, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+  aggupset(null5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+}else{
+  message("Not enough methods found rejections at alpha ", ualpha, 
+          "; skipping upset plot")
+}
+
+```
+
+## Sim N10: Null Comparison 10v10
+
+Here we'll repeat the above, but for a null comparison 10 versus 10 Snf2 samples,
+where the groups are selected randomly. This will be done for 100 random splits.
+
+### Generate a list of SB results objects
+
+```{r n10-uc, results='hide', message=FALSE}
+rseed <- 837
+sampleSize <- 10
+nDE <- 0
+
+if (!file.exists(resfile_n10_uninfCov)){
+  null10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
+                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd,
+                   uninformativeCovariate = TRUE)
+  saveRDS(null10, file = resfile_n10_uninfCov)
+}else{
+  null10 <- readRDS(file = resfile_n10_uninfCov)
+}
+```
+
+### Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(null10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res10 <- plotsim_standardize(null10, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(res10, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10, met="TNR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(null10, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(null10, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res10, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+  aggupset(null10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+}else{
+  message("Not enough methods found rejections at alpha ", ualpha, 
+          "; skipping upset plot")
+}
+
+```
+
+## Sim D5: DE Comparison 5v5
+
+Here we'll repeat the above, but for a DE comparison 5 versus 5 Snf2 samples,
+where the groups are selected randomly and 500 DE genes are added. 
+This will be done for 100 random splits.
+
+### Generate a list of SB results objects
+
+```{r d5-uc, results='hide', message=FALSE}
+rseed <- 198
+sampleSize <- 5
+nDE <- 500
+
+if (!file.exists(resfile_d5_uninfCov)){
+  de5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
+                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd,
+                   uninformativeCovariate = TRUE)
+  saveRDS(de5, file = resfile_d5_uninfCov)
+}else{
+  de5 <- readRDS(file = resfile_d5_uninfCov)
+}
+```
+
+### Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(res5d, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5d, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5d, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(de5, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(de5, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+  aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE)
+
+}else{
+  message("Not enough methods found rejections at alpha ", ualpha, 
+          "; skipping upset plot")
+}
+
+```
+
+## Sim D10: DE Comparison 10v10
+
+Here we'll repeat the above, but for a DE comparison 10 versus 10 Snf2 samples,
+where the groups are selected randomly and 500 DE genes are added.
+This will be done for 100 random splits.
+
+### Generate a list of SB results objects
+
+```{r d10-uc, results='hide', message=FALSE}
+rseed <- 961
+sampleSize <- 10
+nDE <- 500
+
+if(!file.exists(resfile_d10_uninfCov)){
+  de10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
+                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd,
+                   uninformativeCovariate = TRUE)
+  saveRDS(de10, file = resfile_d10_uninfCov)
+}else{
+  de10 <- readRDS(file = resfile_d10_uninfCov)
+}
+```
+
+### Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(res10d, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10d, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10d, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(de10, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(de10, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res10d, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+  aggupset(de10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+}else{
+  message("Not enough methods found rejections at alpha ", ualpha, 
+          "; skipping upset plot")
+}
+
+```
+
+## Sample size comparison
+
+Here we compare the method ranks for the different sample sizes at alpha = 0.10.
+
+```{r}
+# uninformative covariate
+plotMethodRanks(c(resfile_d5_uninfCov, resfile_d10_uninfCov), 
+                colLabels = c("DE 5", "DE 10"), 
+                alpha = 0.10, xlab = "Comparison", 
+                excludeMethods = NULL)
+```
+
 # Session information

### 2018-05-28 manuscript_figures/manuscript_figures.Rmd
Add uninformative covariate sim to yeast and poly

@@
 objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
-resfile <- objects[grepl("yeast", objects) & grepl("de5", objects)] 
+resfile <- objects[grepl("yeast", objects) & 
+                   grepl("de5", objects) & 
+                  !grepl("_uninfCov", objects)] 
 de5 <- readRDS(file=resfile)
@@
             !grepl(c("brain|mir|null|polyester"), dataset) &
-            !grepl("scRNAseq", casestudy))
+            !grepl("scRNAseq", casestudy) &
+            !grepl("_uninfCov", casestudy))
 objs <- objects[wcs]
@@
 wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("brain|mir|polyester|null"), dataset) &
+            !grepl(c("brain|mir|polyester|null|_uninfCov"), dataset) &
             !grepl("scRNAseq", casestudy))
@@
 wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("brain|mir|polyester|de"), dataset) &
+            !grepl(c("brain|mir|polyester|de|_uninfCov"), dataset) &
             !grepl("scRNAseq", casestudy))
@@
 wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("brain|mir|yeast|null"), dataset) &
+            !grepl(c("brain|mir|yeast|null|_uninfCov"), dataset) &
             !grepl("scRNAseq", casestudy))

### 2018-05-31 datasets/RNA-Seq/yeast-simulation.Rmd
Add qqplot diagnostics for FDRreg

@@
 #'  Default is FALSE.
+#' @param qq logical whether to return qqplot of test statistics (for diagnosing
+#'  problems with Scott method)
 simulateOneSplit <- function(X, rseed, dds, resFULL, nDE, sampleSize,
-                             bd, uninformativeCovariate = FALSE){
+                             bd, uninformativeCovariate = FALSE, qq = FALSE){
 
@@
   resTEST <- results(dds_test, name="fake_B_vs_A", independentFiltering = FALSE)
+
+  if(qq){
+    return(ggplot(data.frame(resTEST), aes(sample = stat)) + 
+             geom_qq() +
+             geom_abline())
+  }
 
@@
 }
+
+# qqplots
+plotfile <- file.path(datdir, "null5_qqplots.pdf")
+if (!file.exists(plotfile)){
+  qqs <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
+                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd, 
+                   qq =TRUE)
+  failed <- ifelse(sapply(null5, 
+                          function(x) sum(is.na(assays(x)$qvalue[,22])) > 0),
+                   "Failed", "OK")
+  pdf(plotfile, width=4, height=4)
+  for(i in 1:length(qqs)){
+    print(qqs[[i]] +
+      ggtitle(paste0(failed[i])))
+  }
+  dev.off()
+}
 ```
@@
   null10 <- readRDS(file = resfile_n10)
+}
+
+# qqplots
+plotfile <- file.path(datdir, "null10_qqplots.pdf")
+if (!file.exists(plotfile)){
+  qqs <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
+                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd, 
+                   qq =TRUE)
+  failed <- ifelse(sapply(null5, 
+                          function(x) sum(is.na(assays(x)$qvalue[,22])) > 0),
+                   "Failed", "OK")
+  pdf(plotfile, width=4, height=4)
+  for(i in 1:length(qqs)){
+    print(qqs[[i]] +
+      ggtitle(paste0(failed[i])))
+  }
+  dev.off()
 }
@@
   de5 <- readRDS(file = resfile_d5)
+}
+
+# qqplots
+plotfile <- file.path(datdir, "de5_qqplots.pdf")
+if (!file.exists(plotfile)){
+  qqs <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
+                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd, 
+                   qq =TRUE)
+  failed <- ifelse(sapply(null5, 
+                          function(x) sum(is.na(assays(x)$qvalue[,22])) > 0),
+                   "Failed", "OK")
+  pdf(plotfile, width=4, height=4)
+  for(i in 1:length(qqs)){
+    print(qqs[[i]] +
+      ggtitle(paste0(failed[i])))
+  }
+  dev.off()
 }
@@
   de10 <- readRDS(file = resfile_d10)
+}
+
+# qqplots
+plotfile <- file.path(datdir, "de10_qqplots.pdf")
+if (!file.exists(plotfile)){
+  qqs <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
+                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd, 
+                   qq =TRUE)
+  failed <- ifelse(sapply(null5, 
+                          function(x) sum(is.na(assays(x)$qvalue[,22])) > 0),
+                   "Failed", "OK")
+  pdf(plotfile, width=4, height=4)
+  for(i in 1:length(qqs)){
+    print(qqs[[i]] +
+      ggtitle(paste0(failed[i])))
+  }
+  dev.off()
 }

### 2018-05-31 datasets/RNA-Seq/yeast-simulation.Rmd
Add status label to qqplot diagnostics

@@
                    qq =TRUE)
-  failed <- ifelse(sapply(null5, 
+  failed <- ifelse(sapply(null10, 
                           function(x) sum(is.na(assays(x)$qvalue[,22])) > 0),
@@
                    qq =TRUE)
-  failed <- ifelse(sapply(null5, 
+  failed <- ifelse(sapply(de5, 
                           function(x) sum(is.na(assays(x)$qvalue[,22])) > 0),
@@
                    qq =TRUE)
-  failed <- ifelse(sapply(null5, 
+  failed <- ifelse(sapply(de10, 
                           function(x) sum(is.na(assays(x)$qvalue[,22])) > 0),

### 2018-06-01 manuscript_figures/manuscript_figures.Rmd
Add Figure S7

@@
 
+## Uninformative covariate impact on simulations
+
+Here we'll contrast the results of informative and uninformative covariates 
+on the yeast simulation (null and DE sample size 5).
+
+```{r FigureS7}
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+resfile <- objects[grepl("yeast", objects) & 
+                   grepl("de5|null5", objects)] 
+de5 <- readRDS(file=resfile[grepl("de5", resfile) & !grepl("uninf", resfile)])
+de5_uninf <- readRDS(file=resfile[grepl("de5", resfile) & grepl("uninf", resfile)])
+null5 <- readRDS(file=resfile[grepl("null5", resfile) & !grepl("uninf", resfile)])
+null5_uninf <- readRDS(file=resfile[grepl("null5", resfile) & grepl("uninf", resfile)])
+excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
+
+res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
+res5d_u <- plotsim_standardize(de5_uninf, alpha = seq(0.01, 0.10, 0.01))
+res5n <- plotsim_standardize(null5, alpha = seq(0.01, 0.10, 0.01))
+res5n_u <- plotsim_standardize(null5_uninf, alpha = seq(0.01, 0.10, 0.01))
+
+pa <- plotsim_average(res5n_u, met="TNR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) +
+        expand_limits(y=.7) +
+        xlab("") +
+        ggtitle("")
+
+pb <- plotsim_average(res5n, met="TNR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) +
+        theme(legend.position="none", plot.title = element_text(hjust = 0.5))+
+        expand_limits(y=.7) +
+        xlab("") +
+        ggtitle("")
+
+pc <- plotsim_average(res5d_u, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) +
+        theme(legend.position="none") +
+        expand_limits(y=0.2) +
+        xlab("") +
+        ggtitle("")
+
+pd <- plotsim_average(res5d, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) +
+        theme(legend.position="none") +
+        expand_limits(y=0.2) +
+        xlab("") +
+        ggtitle("")
+
+pe <- plotsim_average(res5d_u, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) +
+        theme(legend.position="none") +
+        expand_limits(y=0.33) +
+        ggtitle("")
+
+pf <- plotsim_average(res5d, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) +
+        theme(legend.position="none") +
+        expand_limits(y=0.33) +
+        ggtitle("")
+
+leg <- get_legend(pa)
+pa <- pa + theme(legend.position="none", plot.title = element_text(hjust = 0.5))
+
+FigT  <- plot_grid(pa, pb, ncol = 2, 
+          #labels = LETTERS[1:2], 
+          label_size = 20)
+
+FigB  <- plot_grid(pc, pd, pe, pf, ncol = 2, 
+          #labels = LETTERS[3:6], 
+          label_size = 20)
+
+Fig <- ggdraw() +
+  draw_plot(FigT + ylab("") + xlab(""), 0, 0.64, 1, 0.3) +
+  draw_plot(FigB + ylab("") + xlab(""), 0.005, 0, 1, 0.6) +
+  draw_plot_label(c("Uninformative Covariate", "Informative Covariate"), 
+                  c(0.14, 0.64), c(1,1), c(0,0), size = 15) +
+  draw_plot_label(c("Null Simulation", "DE Simulation"), 
+                  c(0.36, 0.37), c(0.96,0.62), size = 13)
+
+Fig <- plot_grid(Fig, leg, ncol = 2, rel_widths = c(1, 0.2))
+Fig
+ggsave(file.path(outdir, "FigureS7.pdf"), width=10, height=11)
+
+```
+
 ## Within case-study heatmaps of method rankings

### 2018-06-01 datasets/RNA-Seq/yeast-simulation.Rmd
Exclude bad replicates following Schurch et al

@@
 
+We'll also download the list of 'bad' replicates which were especially poorly 
+correlated to the others, as determined by the authors.
+
+```{r}
+download.file(url = "https://github.com/bartongroup/profDGE48/raw/master/Bad_replicate_identification/exclude.lst", destfile = (file.path(datdir, "badreps.txt")))
+```
+
 ## Read data into R and create a count table
@@
                        function(x) x[[1]])
+badreps <- read_tsv(file.path(datdir, "badreps.txt"), col_names = FALSE)$X1
+badreps <- unlist(lapply(strsplit(badreps, "_MID"), function(x) x[1]))
 
@@
   purrr::map2(sample_names, ~ dplyr:::rename(.x, !! .y := X2, feature = X1) ) %>% # add sample names
-  purrr::reduce(left_join, by = "feature") # reduce with rbind into one dataframe
+  purrr::reduce(left_join, by = "feature") %>% # reduce with rbind into one dataframe
+  dplyr::select(-badreps ) # remove badreps
 ```

### 2018-06-04 manuscript_figures/manuscript_figures.Rmd
Fix typo in grepl commands to subset uninformative covariate sb objs

@@
                    grepl("de5", objects) & 
-                  !grepl("_uninfCov", objects)] 
+                  !grepl("-uninfCov", objects)] 
 de5 <- readRDS(file=resfile)
@@
             !grepl("scRNAseq", casestudy) &
-            !grepl("_uninfCov", casestudy))
+            !grepl("-uninfCov", casestudy))
 objs <- objects[wcs]
@@
 wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("polyester|yeast"), dataset) &
+            !grepl(c("polyester|yeast"), objects) &
             !grepl("scRNAseq", casestudy))
@@
 wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("brain|mir|polyester|null|_uninfCov"), dataset) &
+            !grepl(c("brain|mir|polyester|null|-uninfCov"), objects) &
             !grepl("scRNAseq", casestudy))
@@
 wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("brain|mir|polyester|de|_uninfCov"), dataset) &
+            !grepl(c("brain|mir|polyester|de|-uninfCov"), objects) &
             !grepl("scRNAseq", casestudy))
@@
 wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("brain|mir|yeast|null|_uninfCov"), dataset) &
+            !grepl(c("brain|mir|yeast|null|-uninfCov"), objects) &
             !grepl("scRNAseq", casestudy))
@@
 wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("brain|mir|yeast|de"), dataset) &
+            !grepl(c("brain|mir|yeast|de"), objects) &
             !grepl("scRNAseq", casestudy))

### 2018-06-05 manuscript_figures/manuscript_figures.Rmd
Add alternate supplementary figure

--- 2018-06-01 manuscript_figures/manuscript_figures.Rmd
@@
 
-# set up results directorie
+# set up results directory
 outdir <- "./figures"
@@
                    grepl("de5", objects) & 
-                  !grepl("_uninfCov", objects)] 
+                  !grepl("uninfCov", objects)] 
 de5 <- readRDS(file=resfile)
@@
 
+# Alternative Supplementary Figures
+
+## All Case Studies
+
+This plot is an alternative to Supplementary Figures S1-S6.
+We include all case studies in a single plot.
+
+```{r}
+## read list of case study results
+objects <- list.files(path, recursive=TRUE, pattern="rds", full.names=TRUE)
+
+## exclude yeast and polyester simulations
+objects <- objects[!grepl("yeast", objects) & !grepl("polyester", objects)]
+
+## create shorter names for case studies
+object_labs <- gsub("\\.rds", "", basename(objects))
+object_labs <- gsub("-benchmark", "", object_labs)
+
+## read in case study results (proportion of rejects vs. case study max)
+objdat <- tidy_df(objects, object_labs, "propMaxRejections", NULL)
+objdat <- as_tibble(objdat)
+objdat <- tidyr::complete(objdat, method, casestudy)
+
+## exclude 'enigma-so4-otus' results with no significant calls
+objdat <- dplyr::filter(objdat, casestudy != "enigma-so4-otus")
+
+## create table of case study-to-assay mapping
+studies <- tibble(
+    casestudy = c("cbp-csaw", "h3k4me3-csaw", "human", "mouse", "bmi-maf", "bmi-samplesize",
+                  "enigma-al-otus", "enigma-ph-otus", "enigma-so4-otus", "goodrich-genus",
+                  "papa-genus", "schubert-otus", "brain", "mir200c",
+                  "human-mast-det", "human-mast-mean", "human-scdd-det",
+                  "human-scdd-mean", "human-wilcox-det", "human-wilcox-mean",
+                  "mouse-mast-det", "mouse-mast-mean", "mouse-scdd-det",
+                  "mouse-scdd-mean", "mouse-wilcox-det", "mouse-wilcox-mean"),
+    assay = c(rep("ChIP-seq", 2), rep("GSEA", 2), rep("GWAS", 2), rep("Microbiome", 6),
+              rep("RNA-seq", 2), rep("scRNA-seq", 12)))
+
+## add assay labels to results table
+objdat <- left_join(objdat, studies, by = "casestudy")
+objdat <- dplyr::filter(objdat, !is.na(assay))
+
+## create table of total tests for each case study
+objdat_tab <- dplyr::select(objdat_props, casestudy, nrejects, propPossible, assay)
+objdat_tab <- dplyr::distinct(objdat_tab)
+objdat_tab <- dplyr::mutate(objdat_tab, `Total Tests` = scales::comma(round(nrejects / propPossible)))
+
+## hack to get total counts and heatmap in same plot
+objdat_joint <- bind_rows(list(real = objdat, fake = objdat_tab), .id = "plotrow")
+
+## specify method order
+method_order <- c("bonf", "bh", "ihw", "qvalue", "bl", "lfdr",
+                  "scott-empirical", "scott-theoretical", "ashq")
+objdat_joint <- dplyr::mutate(objdat_joint, method = factor(method, levels = c(method_order, "Total Tests")))
+
+## create table of rejected proportion for top methods per case study
+objdat_props <- dplyr::group_by(objdat, casestudy)
+objdat_props <- dplyr::filter(objdat_props, nrejects == max(nrejects, na.rm = TRUE))
+objdat_props <- dplyr::ungroup(objdat_props)
+objdat_props <- dplyr::mutate(objdat_props,
+                              lab = paste0(nrejects, "\n(", 100*round(propPossible, 4), "%)"))
+objdat_props$plotrow <- "real"
+```
+
+```{r, fig.width = 12, fig.height = 5}
+SFigAlt <- ggplot(objdat_joint, aes(x = casestudy, y = method, fill = propMaxRej)) +
+    scale_x_discrete("Case Study", expand = c(0, 0)) +
+    scale_y_discrete("Method", expand = c(0, 0)) +
+    geom_tile() +
+    geom_text(aes(label = lab), data = objdat_props, size = 1.8, color = "white") +
+    geom_tile(data = filter(objdat_joint, plotrow == "fake"), fill = "white") + 
+    geom_text(aes(label = cnt), data = dplyr::filter(objdat_joint, !is.na(cnt)), size = 2) + 
+    scale_fill_distiller("% Rejected\n(relative to max)",
+                         palette = "Blues", direction = 1, limits = c(0, 1),
+                         labels = scales::percent) +
+    facet_grid(plotrow ~ assay, scales = "free", space = "free") +
+    theme_bw() + 
+    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
+          strip.text.y = element_blank(),
+          legend.position = c(.99, .85), legend.justification = c(1, 1),
+          legend.background = element_rect(fill = scales::alpha("white", 3/4), color = 'black'),
+          legend.direction = "horizontal",
+          legend.text = element_text(size = 7),
+          legend.title = element_text(size = 9),
+          legend.margin = margin(6, 10, 6, 10))
+SFigAlt
+
+ggsave(file.path(outdir, "FigureSAlt.pdf"), gp, width = 12, height = 5)
+```
+
 # Session information

### 2018-06-05 datasets/simulations/simulations-varyingeffectsize-chisq.Rmd
Add unimodal sims with chisq, t noise

--- 2018-04-19 datasets/simulations/simulations-varyingeffectsize.Rmd
@@
 setting. An informative covariate is included in the setting as described in
-`simulations-informative-sine.Rmd`.
+`simulations-informative-sine.Rmd`. This set of simulations is similar to those
+presented in `simulations-varyingeffectsize.Rmd`, however, Chi-Squared
+distributed noise is used to perturb effects. Additionally, since the effect
+must be non-negative (to be a valid value for the non-centrality parameter
+of the a Chi-Squared distribution), the absolute value of all effect sizes
+are used. Note that this modification still maintains a unimodal distribution
+of the effect sizes for all settings except the "bimodal" case below.
 
@@
 ## intermediary files we create below
-spiky_file <- file.path(resdir, "varyinges-benchmark-spiky.rds")
-flattop_file <- file.path(resdir, "varyinges-benchmark-flattop.rds")
-skew_file <- file.path(resdir, "varyinges-benchmark-skew.rds")
-bimodal_file <- file.path(resdir, "varyinges-benchmark-bimodal.rds")
+spiky_file <- file.path(resdir, "varyinges-chisq-benchmark-spiky.rds")
+flattop_file <- file.path(resdir, "varyinges-chisq-benchmark-flattop.rds")
+skew_file <- file.path(resdir, "varyinges-chisq-benchmark-skew.rds")
+bimodal_file <- file.path(resdir, "varyinges-chisq-benchmark-bimodal.rds")
 
@@
 pi0 <- pi0_sine(0.90)               # numeric: proportion of null hypotheses
-tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
-null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
+tstat_dist <- rchisq_perturber(4)   # functional: sampling dist/noise for test stats
+null_dist <- rchisq_pvaluer(4)      # functional: dist to calc p-values
 icovariate <- runif                 # functional: independent covariate
@@
 ```{r spiky-parameters}
-tstat <- sampler_spiky       # functional: dist of alternative test stats
+tstat <- function(n) { abs(sampler_spiky(n)) } # functional: dist of alternative test stats
 seed <- 778
@@
 ```{r flattop-parameters}
-tstat <- sampler_flat_top       # functional: dist of alternative test stats
+tstat <- function(n) { abs(sampler_flat_top(n)) }       # functional: dist of alternative test stats
 seed <- 980
@@
 ```{r skew-parameters}
-tstat <- sampler_skew       # functional: dist of alternative test stats
+tstat <- function(n) { abs(sampler_skew(n)) }       # functional: dist of alternative test stats
 seed <- 206
@@
 ```{r bimodal-parameters}
-tstat <- sampler_bimodal       # functional: dist of alternative test stats
+tstat <- function(n) { abs(sampler_bimodal(n)) }       # functional: dist of alternative test stats
 seed <- 913

### 2018-06-05 datasets/simulations/simulations-varyingeffectsize-t.Rmd
Add unimodal sims with chisq, t noise

--- 2018-04-19 datasets/simulations/simulations-varyingeffectsize.Rmd
@@
 ---
-title: "Simulation Study: Varying Effect Size"
+title: "Simulation Study: Varying Effect Size (t noise)"
 author: "Patrick Kimes"
@@
 setting. An informative covariate is included in the setting as described in
-`simulations-informative-sine.Rmd`.
+`simulations-informative-sine.Rmd`. This set of simulations is similar to those
+presented in `simulations-varyingeffectsize.Rmd`, however, t-distributed noise
+is used to perturb effects.
 
@@
 ## intermediary files we create below
-spiky_file <- file.path(resdir, "varyinges-benchmark-spiky.rds")
-flattop_file <- file.path(resdir, "varyinges-benchmark-flattop.rds")
-skew_file <- file.path(resdir, "varyinges-benchmark-skew.rds")
-bimodal_file <- file.path(resdir, "varyinges-benchmark-bimodal.rds")
+spiky_file <- file.path(resdir, "varyinges-t-benchmark-spiky.rds")
+flattop_file <- file.path(resdir, "varyinges-t-benchmark-flattop.rds")
+skew_file <- file.path(resdir, "varyinges-t-benchmark-skew.rds")
+bimodal_file <- file.path(resdir, "varyinges-t-benchmark-bimodal.rds")
 
@@
 pi0 <- pi0_sine(0.90)               # numeric: proportion of null hypotheses
-tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
-null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
+tstat_dist <- rt_perturber(5)       # functional: sampling dist/noise for test stats
+null_dist <- rt_2pvaluer(5)         # functional: dist to calc p-values
 icovariate <- runif                 # functional: independent covariate

### 2018-06-05 datasets/RNA-Seq/yeast-simulation.Rmd

@@
 #'  Default is FALSE.
-#' @param qq logical whether to return qqplot of test statistics (for diagnosing
-#'  problems with Scott method)
+#' @param pvalHists logical whether to return histograms of pvalues by covariate
+#'  instead of SB objects
 simulateOneSplit <- function(X, rseed, dds, resFULL, nDE, sampleSize,
-                             bd, uninformativeCovariate = FALSE, qq = FALSE){
+                             bd, uninformativeCovariate = FALSE, pvalHists = FALSE){
 
@@
     # randomly sample a log2FC from original FCs (without regard to DE)
+    counts_new <- counts(dds_test)
     log2FC <- rep(0, nrow(dds_test))
@@
 
-    counts_new <- counts(dds_test)
-    counts_new[,colData(dds_test)$fake=="A"] <- counts(dds_test)[,colData(dds_test)$fake=="A"] * 2^log2FC
+    # randomize which condition is shifted up or down
+    ran <- runif(nrow(dds_test)) 
+    multiplier <- ifelse(ran < 0.5, 2^log2FC, 1/(2^log2FC))
+    refcond <- ifelse(ran < 0.5, "A", "B")
+    down <- which(ran < 0.5)
+
+    counts_new[down,colData(dds_test)$fake==unique(refcond[down])] <- 
+        counts(dds_test)[down, colData(dds_test)$fake==unique(refcond[down])] *
+        multiplier[down]
+    counts_new[-down,colData(dds_test)$fake==unique(refcond[-down])] <- 
+        counts(dds_test)[-down, colData(dds_test)$fake==unique(refcond[-down])] *
+        multiplier[-down]
+
     counts_new <- apply(counts_new, 2, as.integer)
+
     counts(dds_test) <- counts_new
   }
+
+  # replace existing size factors 
+  dds_test <- estimateSizeFactors(dds_test)
 
@@
 
-  if(qq){
-    return(ggplot(data.frame(resTEST), aes(sample = stat)) + 
-             geom_qq() +
-             geom_abline())
-  }
-
   geneExp <- tbl_df(data.frame(geneName=rownames(resTEST), 
@@
                                SE=resTEST$lfcSE,                 
-                               ind_covariate = ifelse(uninformativeCovariate, 
-                                                      rnorm(nrow(resTEST)),
-                                                      resTEST$baseMean),
+                               ind_covariate = resTEST$baseMean,
                                effect_size = resTEST$log2FoldChange, 
@@
                                pzero = rowSums(counts(dds_test)==0)/
-                                       ncol(counts(dds_test))  ))
+                                       ncol(counts(dds_test)))) %>%
+                dplyr::filter(!is.na(pval))
 
-  # filter NAs and those with less than 50% expressed 
-  geneExp <- geneExp %>% na.omit() %>% dplyr::filter(pzero < 0.5)
+  if(uninformativeCovariate)
+    geneExp$ind_covariate <- rnorm(nrow(geneExp))
+
+   if(pvalHists){
+    return(strat_hist(geneExp, pvalue="pval", covariate="ind_covariate", maxy =10))
+  }
 
@@
   sb <- addDefaultMetrics(sb)
-  pf <- estimatePerformanceMetrics(sb, tidy=TRUE, alpha=seq(0.01, 0.1, 0.01))
-pf %>% filter(performanceMetric %in% c("rejections", "FDR", "TPR")) %>%
-  with(table(is.na(value), blabel, performanceMetric))
-
   rowData(sb)$log2FC <- geneExp$effect_size
@@
 }
-
-# qqplots
-plotfile <- file.path(datdir, "null5_qqplots.pdf")
-if (!file.exists(plotfile)){
-  qqs <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
-                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd, 
-                   qq =TRUE)
-  failed <- ifelse(sapply(null5, 
-                          function(x) sum(is.na(assays(x)$qvalue[,22])) > 0),
-                   "Failed", "OK")
-  pdf(plotfile, width=4, height=4)
-  for(i in 1:length(qqs)){
-    print(qqs[[i]] +
-      ggtitle(paste0(failed[i])))
-  }
-  dev.off()
-}
 ```
@@
 
-# qqplots
-plotfile <- file.path(datdir, "null10_qqplots.pdf")
-if (!file.exists(plotfile)){
-  qqs <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
-                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd, 
-                   qq =TRUE)
-  failed <- ifelse(sapply(null10, 
-                          function(x) sum(is.na(assays(x)$qvalue[,22])) > 0),
-                   "Failed", "OK")
-  pdf(plotfile, width=4, height=4)
-  for(i in 1:length(qqs)){
-    print(qqs[[i]] +
-      ggtitle(paste0(failed[i])))
-  }
-  dev.off()
-}
 ```
@@
 Here we'll repeat the above, but for a DE comparison 5 versus 5 Snf2 samples,
-where the groups are selected randomly and 500 DE genes are added. 
+where the groups are selected randomly and 2000 DE genes are added. 
 This will be done for 100 random splits.
@@
 sampleSize <- 5
-nDE <- 500
+nDE <- 2000
 
@@
 
-# qqplots
-plotfile <- file.path(datdir, "de5_qqplots.pdf")
+# pvalue histograms
+plotfile <- file.path(datdir, "de5_pvalhists.pdf")
 if (!file.exists(plotfile)){
-  qqs <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
+  hists <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
                    resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd, 
-                   qq =TRUE)
-  failed <- ifelse(sapply(de5, 
-                          function(x) sum(is.na(assays(x)$qvalue[,22])) > 0),
-                   "Failed", "OK")
-  pdf(plotfile, width=4, height=4)
-  for(i in 1:length(qqs)){
-    print(qqs[[i]] +
-      ggtitle(paste0(failed[i])))
+                   pvalHists = TRUE)
+  pdf(plotfile, width=8, height=4)
+  for(i in 1:length(hists)){
+    print(hists[[i]])
   }
@@
 Here we'll repeat the above, but for a DE comparison 10 versus 10 Snf2 samples,
-where the groups are selected randomly and 500 DE genes are added.
+where the groups are selected randomly and 2000 DE genes are added.
 This will be done for 100 random splits.
@@
 sampleSize <- 10
-nDE <- 500
+nDE <- 2000
 
@@
 
-# qqplots
-plotfile <- file.path(datdir, "de10_qqplots.pdf")
+# pvalue histograms
+plotfile <- file.path(datdir, "de10_pvalhists.pdf")
 if (!file.exists(plotfile)){
-  qqs <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
+  hists <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
                    resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd, 
-                   qq =TRUE)
-  failed <- ifelse(sapply(de10, 
-                          function(x) sum(is.na(assays(x)$qvalue[,22])) > 0),
-                   "Failed", "OK")
-  pdf(plotfile, width=4, height=4)
-  for(i in 1:length(qqs)){
-    print(qqs[[i]] +
-      ggtitle(paste0(failed[i])))
+                   pvalHists = TRUE)
+  pdf(plotfile, width=8, height=4)
+  for(i in 1:length(hists)){
+    print(hists[[i]])
   }

### 2018-06-05 manuscript_figures/manuscript_figures.Rmd

@@
                    grepl("de5", objects) & 
+                  !grepl("uninfCov", objects)] 
                   !grepl("uninfCov", objects)] 
@@
             !grepl("scRNAseq", casestudy) &
-            !grepl("_uninfCov", casestudy))
+            !grepl("uninfCov", casestudy))
 objs <- objects[wcs]
@@
 wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("brain|mir|polyester|null|_uninfCov"), dataset) &
+            !grepl(c("brain|mir|polyester|null|uninfCov"), objects) &
             !grepl("scRNAseq", casestudy))
@@
 wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("brain|mir|polyester|de|_uninfCov"), dataset) &
+            !grepl(c("brain|mir|polyester|de|uninfCov"), objects) &
             !grepl("scRNAseq", casestudy))
@@
 wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("brain|mir|yeast|null|_uninfCov"), dataset) &
+            !grepl(c("brain|mir|yeast|null|uninfCov"), objects) &
             !grepl("scRNAseq", casestudy))

### 2018-06-05 datasets/simulations/simulations-informative-cubic.Rmd
Rerun t-dist sims w/ ASH df specified

@@
 
-We next run the simulations (including Scott's FDR Regression).
+For the t-distributed noise setting, we must specify the number of degrees of freedom
+for ASH. We add an additional parameter to the `ashq` method with the corresponding
+degrees of freedom of the noise distribution.
+
+```{r}
+bdplust <- modifyBMethod(bdplus, "ashq", df = 5)
+```
+
+We next run the simulations (including Scott's FDR Regression and ASH with degrees of
+freedom specified).
 
@@
 } else {
-    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplust, m = m,
                     pi0 = pi0, tstat = tstat, icovariate = icovariate,
@@
 
-Since both `scott-empirical` and `scott-theoretical` are not appropriate when
-the test statistics are not approximately normal, we also plot the results without
-these methods.
+Since `ashq` is not appropriate when the test statistics are not approximately normal or t,
+we also plot the results without this method.
 
-```{r chisq4-metrics-averages-nofdrreg, results = "hide"}
-plotsim_average(resdf, met = "rejections",
-                filter_set = c(excludeSet, "scott-empirical", "scott-theoretical"),
+```{r chisq4-metrics-averages-noash, results = "hide"}
+plotsim_average(resdf, met = "rejections", filter_set = c(excludeSet, "ashq"),
                 merge_ihw = TRUE, errorBars = TRUE)
 
-plotsim_average(resdf, met = "FDR",
-                filter_set = c(excludeSet, "scott-empirical", "scott-theoretical"),
+plotsim_average(resdf, met = "FDR", filter_set = c(excludeSet, "ashq"),
                 merge_ihw = TRUE, errorBars = TRUE)
 
-plotsim_average(resdf, met = "TPR",
-                filter_set = c(excludeSet, "scott-empirical", "scott-theoretical"),
+plotsim_average(resdf, met = "TPR", filter_set = c(excludeSet, "ashq"),
                 merge_ihw = TRUE, errorBars = TRUE)
 
-plotsim_average(resdf, met = "TNR",
-                filter_set = c(excludeSet, "scott-empirical", "scott-theoretical"),
+plotsim_average(resdf, met = "TNR", filter_set = c(excludeSet, "ashq"),
                 merge_ihw = TRUE, errorBars = TRUE) 

### 2018-06-06 manuscript_figures/manuscript_figures.Rmd
Fix typo in Fig2

@@
                   !grepl("uninfCov", objects)]
-                  !grepl("uninfCov", objects)] 
+
 de5 <- readRDS(file=resfile)

### 2018-06-06 datasets/simulations/simulations-null.Rmd
Update sim params w/ larger signal

@@
 bd <- initializeBenchDesign()
-```
-
-For comparison, we also include the ASH q-values.
-
-```{r}
-bd <- addBMethod(bd, "ashq", ashr::ash, ashr::get_qvalue,
-                 betahat = effect_size, sebetahat = SE)
 ```
@@
 
+Since `ashq` is not appropriate when the test statistics are not approximately normal or t,
+we also plot the results without this method.
+
+```{r chisq4-metrics-averages-noash, results = "hide"}
+plotsim_average(resdf, met="rejections", filter_set = c(excludeSet, "ashq"),
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = c(excludeSet, "ashq"),
+                merge_ihw = TRUE, errorBars = TRUE) 
+```
+
 We also take a look at the distribution of rejects for each method as a function of

### 2018-06-06 datasets/simulations/simulations-varyingeffectsize-chisq.Rmd

@@
 
-As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
-for simulations with Gaussian or t-distributed noise. Again, we include both
-`nulltype = "empirical"` and `nulltype = "theoretical"`. Since all settings in this
-series of simulations use test statistics simulated with Gaussian noise, we include
-Scott's FDR Regression in all of the comparisons.
-
-```{r}
-bdplus <- bd
-bdplus <- addBMethod(bdplus, "scott-theoretical",
-                     FDRreg::FDRreg,
-                     function(x) { x$FDR },
-                     z = test_statistic,
-                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
-                     nulltype = 'theoretical',
-                     control = list(lambda = 0.01))
-bdplus <- addBMethod(bdplus, "scott-empirical",
-                     FDRreg::FDRreg,
-                     function(x) { x$FDR },
-                     z = test_statistic,
-                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
-                     nulltype = 'empirical',
-                     control = list(lambda = 0.01))
-```
-
 All simulation settings will share the following parameters.
@@
 } else {
-    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bd, m = m,
                     pi0 = pi0, tstat = tstat, icovariate = icovariate,
@@
 ```{r spiky-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r flattop-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r skew-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r bimodal-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)

### 2018-06-06 manuscript_figures/manuscript_figures.Rmd
Fix Typo in Fig2

@@
             !grepl("scRNAseq", casestudy) &
-            !grepl("uninfCov", casestudy))
+            !grepl("uninfCov", dataset))
 objs <- objects[wcs]

### 2018-06-06 manuscript_figures/manuscript_figures.Rmd
Fix code order in alt supplementary figure

@@
 
+## create table of rejected proportion for top methods per case study
+objdat_props <- dplyr::group_by(objdat, casestudy)
+objdat_props <- dplyr::filter(objdat_props, nrejects == max(nrejects, na.rm = TRUE))
+objdat_props <- dplyr::ungroup(objdat_props)
+objdat_props <- dplyr::mutate(objdat_props, lab = paste0(100*round(propPossible, 4), "%"))
+objdat_props$plotrow <- "real"
+
 ## create table of total tests for each case study
@@
 objdat_tab <- dplyr::mutate(objdat_tab, `Total Tests` = scales::comma(round(nrejects / propPossible)))
+objdat_tab <- dplyr::select(objdat_tab, casestudy, assay, `Total Tests`) 
+objdat_tab <- tidyr::gather(objdat_tab, method, cnt, -casestudy, -assay)
 
@@
 objdat_joint <- dplyr::mutate(objdat_joint, method = factor(method, levels = c(method_order, "Total Tests")))
-
-## create table of rejected proportion for top methods per case study
-objdat_props <- dplyr::group_by(objdat, casestudy)
-objdat_props <- dplyr::filter(objdat_props, nrejects == max(nrejects, na.rm = TRUE))
-objdat_props <- dplyr::ungroup(objdat_props)
-objdat_props <- dplyr::mutate(objdat_props,
-                              lab = paste0(nrejects, "\n(", 100*round(propPossible, 4), "%)"))
-objdat_props$plotrow <- "real"
 ```

### 2018-06-06 datasets/simulations/simulations-informative-cubic.Rmd
Add t11 sims for ashq, drop ashq from chisq sim plots

@@
 tdist_file <- file.path(resdir, "informative-cubic-benchmark-t5.rds")
+tdist11_file <- file.path(resdir, "informative-cubic-benchmark-t11.rds")
 chisq_file <- file.path(resdir, "informative-cubic-benchmark-chisq4.rds")
@@
 
-# Student's t Setting
+# Student's t Setting (df = 5)
 
@@
 
+# Student's t Setting (df = 11)
+
+Next, we consider a second setting with t-distributed noise.
+
+## Data Simulation
+
+```{r t11-parameters}
+tstat <- rnorm_generator(6)    # functional: dist of alternative test stats
+tstat_dist <- rt_perturber(11)  # functional: sampling dist/noise for test stats
+null_dist <- rt_2pvaluer(11)    # functional: dist to calc p-values
+seed <- 9158
+```
+
+For the t-distributed noise setting, we must specify the number of degrees of freedom
+for ASH. We add an additional parameter to the `ashq` method with the corresponding
+degrees of freedom of the noise distribution.
+
+```{r}
+bdplust <- modifyBMethod(bdplus, "ashq", df = 11)
+```
+
+We next run the simulations (including Scott's FDR Regression and ASH with degrees of
+freedom specified).
+
+```{r t11-run-simulation}
+if (file.exists(tdist11_file)) {
+    res <- readRDS(tdist11_file)
+} else {
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplust, m = m,
+                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
+                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    seed = seed, mc.cores = cores)
+    saveRDS(res, file = tdist11_file)
+}
+```
+
+## Covariate Diagnostics
+
+Here, we show the relationship between the independent covariate and p-values for a
+single replication of the experiment.
+
+```{r t11-one-simulation}
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
+```
+
+```{r, t11-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
+rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
+```
+
+```{r, t11-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+```
+
+## Benchmark Metrics
+
+We plot the averaged results across `r B` replications.
+
+```{r t11-metrics-averages, results = "hide"}
+resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(resdf, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE) 
+```
+
+We also take a look at the distribution of rejects for each method as a function of
+the effect size and independent covariate.
+
+```{r t11-metrics-covlineplot, results = "hide"}
+covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+```
+
+Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
+the overlap of rejections between methods.
+
+```{r t11-metrics-upset, results = "hide"}
+if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
+    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
+} else {
+    message("Not enough methods found rejections at alpha ", ualpha, 
+            "; skipping upset plot")
+}
+```
+
 # Chi-Squared Setting
@@
 We also take a look at the distribution of rejects for each method as a function of
-the effect size and independent covariate.
+the effect size and independent covariate. Again, we plot these without `ashq`.
 
 ```{r chisq4-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+covariateLinePlot(lapply(res, function(x) { x[, -which(colnames(x) == "ashq")] }),
+                  alpha = ualpha, covname = "effect_size")
 
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(lapply(res, function(x) { x[, -which(colnames(x) == "ashq")] }),
+                  alpha = ualpha, covname = "ind_covariate")
 ```
@@
 Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
-the overlap of rejections between methods.
+the overlap of rejections between methods. Again, without `ashq`.
 
 ```{r chisq4-metrics-upset, results = "hide"}
-if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
-    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
+if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = c(excludeSet, "ashq")) >= 3) {
+    aggupset(lapply(res, function(x) { x[, -which(colnames(x) == "ashq")] }),
+             alpha = ualpha, supplementary = FALSE, return_list = FALSE)
 } else {
@@
 ```
+
+# Session Info
+
+```{r}
+sessionInfo()
+```

### 2018-06-06 datasets/simulations/simulations-informative-sine.Rmd
Add t11 sims for ashq, drop ashq from chisq sim plots

@@
 tdist_file <- file.path(resdir, "informative-sine-benchmark-t5.rds")
+tdist11_file <- file.path(resdir, "informative-cubic-benchmark-t11.rds")
 chisq_file <- file.path(resdir, "informative-sine-benchmark-chisq4.rds")

### 2018-06-06 datasets/simulations/simulations-informative-step.Rmd
Add t11 sims for ashq, drop ashq from chisq sim plots

@@
 tdist_file <- file.path(resdir, "informative-step-benchmark-t5.rds")
+tdist11_file <- file.path(resdir, "informative-step-benchmark-t11.rds")
 chisq_file <- file.path(resdir, "informative-step-benchmark-chisq4.rds")

### 2018-06-06 datasets/simulations/simulations-noninformative.Rmd
Add t11 sims for ashq, drop ashq from chisq sim plots

@@
 tdist_file <- file.path(resdir, "noninformative-benchmark-t5.rds")
+tdist11_file <- file.path(resdir, "noninformative-benchmark-t11.rds")
 chisq_file <- file.path(resdir, "noninformative-benchmark-chisq4.rds")

### 2018-06-06 datasets/simulations/simulations-null.Rmd
Add t11 sims for ashq, drop ashq from chisq sim plots

@@
 tdist_file <- file.path(resdir, "null-benchmark-t5.rds")
+tdist11_file <- file.path(resdir, "null-benchmark-t11.rds")
 chisq_file <- file.path(resdir, "null-benchmark-chisq4.rds")
@@
 
+# Student's t Setting (df = 11)
+
+Next, we consider the null setting with t-distributed noise.
+
+## Data Simulation
+
+```{r t11-parameters}
+tstat_dist <- rt_perturber(11)  # functional: sampling dist/noise for test stats
+null_dist <- rt_2pvaluer(11)    # functional: dist to calc p-values
+seed <- 9158
+```
+
+For the t-distributed noise setting, we must specify the number of degrees of freedom
+for ASH. We add an additional parameter to the `ashq` method with the corresponding
+degrees of freedom of the noise distribution.
+
+```{r}
+bdplust <- modifyBMethod(bdplus, "ashq", df = 11)
+```
+
+We next run the simulations (including Scott's FDR Regression and ASH with degrees of
+freedom specified).
+
+```{r t11-run-simulation}
+if (file.exists(tdist_file)) {
+    res <- readRDS(tdist_file)
+} else {
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplust, m = m,
+                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
+                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    seed = seed, mc.cores = cores)
+    saveRDS(res, file = tdist_file)
+}
+```
+
+## Covariate Diagnostics
+
+Here, we show the relationship between the independent covariate and p-values for a
+single replication of the experiment.
+
+```{r t11-one-simulation}
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
+```
+
+```{r, t11-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
+rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
+```
+
+```{r, t11-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+```
+
+## Benchmark Metrics
+
+We plot the average results across `r B` simulations.
+
+```{r t11-metrics-averages, results = "hide"}
+resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(resdf, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE) 
+```
+
+We also take a look at the distribution of rejects for each method as a function of
+the effect size and independent covariate.
+
+```{r t11-metrics-covlineplot, results = "hide"}
+covariateLinePlot(res, alpha = 0.05, covname = "effect_size", trans = "log1p")
+
+covariateLinePlot(res, alpha = 0.05, covname = "ind_covariate", trans = "log1p")
+```
+
+Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
+the overlap of rejections between methods.
+
+```{r t11-metrics-upset, results = "hide"}
+if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
+    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+    message("Not enough methods found rejections at alpha ", ualpha, 
+            "; skipping upset plot")
+}
+```
+
 # Chi-Squared Setting
@@
 We also take a look at the distribution of rejects for each method as a function of
-the effect size and independent covariate.
+the effect size and independent covariate. Again, we plot these without `ashq`.
 
 ```{r chisq4-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = 0.05, covname = "effect_size", trans = "log1p")
+covariateLinePlot(lapply(res, function(x) { x[, -which(colnames(x) == "ashq")] }),
+                  alpha = ualpha, covname = "effect_size")
 
-covariateLinePlot(res, alpha = 0.05, covname = "ind_covariate", trans = "log1p")
+covariateLinePlot(lapply(res, function(x) { x[, -which(colnames(x) == "ashq")] }),
+                  alpha = ualpha, covname = "ind_covariate")
 ```
@@
 Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
-the overlap of rejections between methods.
+the overlap of rejections between methods. Again, without `ashq`.
 
 ```{r chisq4-metrics-upset, results = "hide"}
-if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
-    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE) 
+if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = c(excludeSet, "ashq")) >= 3) {
+    aggupset(lapply(res, function(x) { x[, -which(colnames(x) == "ashq")] }),
+             alpha = ualpha, supplementary = FALSE, return_list = FALSE)
 } else {

### 2018-06-06 datasets/simulations/simulations-varyingeffectsize-t.Rmd
Add t11 sims for ashq, drop ashq from chisq sim plots

@@
 
+Since all simulation settings in this case study use t-distributed noise, we must specify
+the number of degrees of freedom for ASH. We add an additional parameter to the `ashq`
+method with the corresponding degrees of freedom of the noise distribution.
+
+```{r}
+bdplus <- modifyBMethod(bdplus, "ashq", df = 11)
+```
+
 All simulation settings will share the following parameters.
@@
 pi0 <- pi0_sine(0.90)               # numeric: proportion of null hypotheses
-tstat_dist <- rt_perturber(5)       # functional: sampling dist/noise for test stats
-null_dist <- rt_2pvaluer(5)         # functional: dist to calc p-values
+tstat_dist <- rt_perturber(11)       # functional: sampling dist/noise for test stats
+null_dist <- rt_2pvaluer(11)         # functional: dist to calc p-values
 icovariate <- runif                 # functional: independent covariate

### 2018-06-06 datasets/simulations/simulations-varyingeffectsize-chisq.Rmd
Use cubic indep cov instead of sine

@@
 setting. An informative covariate is included in the setting as described in
-`simulations-informative-sine.Rmd`. This set of simulations is similar to those
+`simulations-informative-cubic.Rmd`. This set of simulations is similar to those
 presented in `simulations-varyingeffectsize.Rmd`, however, Chi-Squared
@@
 m <- 20000                          # integer: number of hypothesis tests
-pi0 <- pi0_sine(0.90)               # numeric: proportion of null hypotheses
+pi0 <- pi0_cubic(0.90)              # numeric: proportion of null hypotheses
 tstat_dist <- rchisq_perturber(4)   # functional: sampling dist/noise for test stats

### 2018-06-06 datasets/simulations/simulations-varyingeffectsize-t.Rmd
Use cubic indep cov instead of sine

@@
 setting. An informative covariate is included in the setting as described in
-`simulations-informative-sine.Rmd`. This set of simulations is similar to those
+`simulations-informative-cubic.Rmd`. This set of simulations is similar to those
 presented in `simulations-varyingeffectsize.Rmd`, however, t-distributed noise
@@
 m <- 20000                          # integer: number of hypothesis tests
-pi0 <- pi0_sine(0.90)               # numeric: proportion of null hypotheses
+pi0 <- pi0_cubic(0.90)              # numeric: proportion of null hypotheses
 tstat_dist <- rt_perturber(11)      # functional: sampling dist/noise for test stats

### 2018-06-06 datasets/simulations/simulations-varyingeffectsize.Rmd
Use cubic indep cov instead of sine

--- 2018-04-19 datasets/simulations/simulations-varyingeffectsize.Rmd
@@
 setting. An informative covariate is included in the setting as described in
-`simulations-informative-sine.Rmd`.
+`simulations-informative-cubic.Rmd`.
 
@@
 m <- 20000                          # integer: number of hypothesis tests
-pi0 <- pi0_sine(0.90)             # numeric: proportion of null hypotheses
-tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
-null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
+pi0 <- pi0_cubic(0.90)              # numeric: proportion of null hypotheses
+tstat_dist <- rnorm_perturber(1/2)  # functional: sampling dist/noise for test stats
+null_dist <- rnorm_2pvaluer(1/2)    # functional: dist to calc p-values
 icovariate <- runif                 # functional: independent covariate

### 2018-06-06 datasets/simulations/simulations-varyingpi0.Rmd
Rerun t-dist sims w/ ASH df specified
Use cubic indep cov instead of sine

@@
 tests with varying proportion of null tests. An informative covariate is included
-in the setting as described in `simulations-informative-sine.Rmd`.
+in the setting as described in `simulations-informative-cubic.Rmd`.
 
@@
 As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
-for simulations with Gaussian or t-distributed noise. Again, we include both
+for simulations with Gaussian noise. Again, we include both
 `nulltype = "empirical"` and `nulltype = "theoretical"`. Since all settings in this
@@
 ```{r null80-parameters}
-pi0 <- pi0_sine(0.80)             # numeric: proportion of null hypotheses
+pi0 <- pi0_cubic(0.80)             # numeric: proportion of null hypotheses
 seed <- 608
@@
 ```{r null90-parameters}
-pi0 <- pi0_sine(0.90)             # numeric: proportion of null hypotheses
+pi0 <- pi0_cubic(0.90)             # numeric: proportion of null hypotheses
 seed <- 808
@@
 ```{r null95-parameters}
-pi0 <- pi0_sine(0.95)             # numeric: proportion of null hypotheses
+pi0 <- pi0_cubic(0.95)             # numeric: proportion of null hypotheses
 seed <- 913
@@
 ```{r null99-parameters}
-pi0 <- pi0_sine(0.99)             # numeric: proportion of null hypotheses
+pi0 <- pi0_cubic(0.99)             # numeric: proportion of null hypotheses
 seed <- 2015

### 2018-06-06 manuscript_figures/manuscript_figures.Rmd
Fix typo in save of Alt Supp Figure
Adjust margins of Fig S7

@@
                 merge_ihw = TRUE, errorBars=TRUE) +
-        expand_limits(y=.7) +
+        expand_limits(y=0.93) +
         xlab("") +
@@
         theme(legend.position="none", plot.title = element_text(hjust = 0.5))+
-        expand_limits(y=.7) +
+        expand_limits(y=.93) +
         xlab("") +
@@
         theme(legend.position="none") +
-        expand_limits(y=0.2) +
+        expand_limits(y=0.18) +
         xlab("") +
@@
         theme(legend.position="none") +
-        expand_limits(y=0.33) +
+        expand_limits(y=c(0.10,0.55)) +
         ggtitle("")
@@
 
-ggsave(file.path(outdir, "FigureSAlt.pdf"), gp, width = 12, height = 5)
+ggsave(file.path(outdir, "FigureSAlt.pdf"), SFigAlt, width = 12, height = 5)
 ```

### 2018-06-07 datasets/simulations/simulations-varyingpi0.Rmd
Add lower pi0 values in simulations

--- 2018-04-20 datasets/simulations/simulations-varyingpi0.Rmd
@@
 ## intermediary files we create below
+null05_file <- file.path(resdir, "varyingpi0-benchmark-nullprop05.rds")
+null10_file <- file.path(resdir, "varyingpi0-benchmark-nullprop10.rds")
+null20_file <- file.path(resdir, "varyingpi0-benchmark-nullprop20.rds")
+null50_file <- file.path(resdir, "varyingpi0-benchmark-nullprop50.rds")
 null80_file <- file.path(resdir, "varyingpi0-benchmark-nullprop80.rds")
@@
 
+# 5% Null Setting
+
+First, we consider the setting where 5% of tests are null.
+
+## Data Simulation
+
+```{r null05-parameters}
+pi0 <- pi0_sine(0.05)             # numeric: proportion of null hypotheses
+seed <- 1010
+```
+
+We next run the simulations.
+
+```{r null05-run-simulation}
+if (file.exists(null05_file)) {
+    res <- readRDS(null05_file)
+} else {
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
+                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
+                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    seed = seed, mc.cores = cores)
+    saveRDS(res, file = null05_file)
+}
+```
+
+## Covariate Diagnostics
+
+Here, we show the relationship between the independent covariate and p-values for a
+single replication of the experiment.
+
+```{r null05-one-simulation}
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
+```
+
+```{r, null05-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
+rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
+```
+
+```{r, null05-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+```
+
+## Benchmark Metrics
+
+We plot the averaged results across `r B` replications.
+
+```{r null05-metrics-averages, results = "hide"}
+resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(resdf, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE) 
+```
+
+We also take a look at the distribution of rejects for each method as a function of
+the effect size and independent covariate.
+
+```{r null05-metrics-covlineplot, results = "hide"}
+covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+```
+
+Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
+the overlap of rejections between methods.
+
+```{r null05-metrics-upset, results = "hide"}
+if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
+    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
+} else {
+    message("Not enough methods found rejections at alpha ", ualpha, 
+            "; skipping upset plot")
+}
+```
+
+# 10% Null Setting
+
+Next, we consider the setting where 10% of tests are null.
+
+## Data Simulation
+
+```{r null10-parameters}
+pi0 <- pi0_sine(0.10)             # numeric: proportion of null hypotheses
+seed <- 1201
+```
+
+We next run the simulations.
+
+```{r null10-run-simulation}
+if (file.exists(null10_file)) {
+    res <- readRDS(null10_file)
+} else {
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
+                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
+                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    seed = seed, mc.cores = cores)
+    saveRDS(res, file = null10_file)
+}
+```
+
+## Covariate Diagnostics
+
+Here, we show the relationship between the independent covariate and p-values for a
+single replication of the experiment.
+
+```{r null10-one-simulation}
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
+```
+
+```{r, null10-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
+rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
+```
+
+```{r, null10-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+```
+
+## Benchmark Metrics
+
+We plot the averaged results across `r B` replications.
+
+```{r null10-metrics-averages, results = "hide"}
+resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(resdf, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE) 
+```
+
+We also take a look at the distribution of rejects for each method as a function of
+the effect size and independent covariate.
+
+```{r null10-metrics-covlineplot, results = "hide"}
+covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+```
+
+Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
+the overlap of rejections between methods.
+
+```{r null10-metrics-upset, results = "hide"}
+if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
+    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
+} else {
+    message("Not enough methods found rejections at alpha ", ualpha, 
+            "; skipping upset plot")
+}
+```
+
+# 20% Null Setting
+
+Next, we consider the setting where 20% of tests are null.
+
+## Data Simulation
+
+```{r null20-parameters}
+pi0 <- pi0_sine(0.20)             # numeric: proportion of null hypotheses
+seed <- 51
+```
+
+We next run the simulations.
+
+```{r null20-run-simulation}
+if (file.exists(null20_file)) {
+    res <- readRDS(null20_file)
+} else {
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
+                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
+                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    seed = seed, mc.cores = cores)
+    saveRDS(res, file = null20_file)
+}
+```
+
+## Covariate Diagnostics
+
+Here, we show the relationship between the independent covariate and p-values for a
+single replication of the experiment.
+
+```{r null20-one-simulation}
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
+```
+
+```{r, null20-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
+rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
+```
+
+```{r, null20-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+```
+
+## Benchmark Metrics
+
+We plot the averaged results across `r B` replications.
+
+```{r null20-metrics-averages, results = "hide"}
+resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(resdf, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE) 
+```
+
+We also take a look at the distribution of rejects for each method as a function of
+the effect size and independent covariate.
+
+```{r null20-metrics-covlineplot, results = "hide"}
+covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+```
+
+Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
+the overlap of rejections between methods.
+
+```{r null20-metrics-upset, results = "hide"}
+if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
+    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
+} else {
+    message("Not enough methods found rejections at alpha ", ualpha, 
+            "; skipping upset plot")
+}
+```
+
+# 50% Null Setting
+
+Next, we consider the setting where 50% of tests are null.
+
+## Data Simulation
+
+```{r null50-parameters}
+pi0 <- pi0_sine(0.50)             # numeric: proportion of null hypotheses
+seed <- 500
+```
+
+We next run the simulations.
+
+```{r null50-run-simulation}
+if (file.exists(null50_file)) {
+    res <- readRDS(null50_file)
+} else {
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
+                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
+                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    seed = seed, mc.cores = cores)
+    saveRDS(res, file = null50_file)
+}
+```
+
+## Covariate Diagnostics
+
+Here, we show the relationship between the independent covariate and p-values for a
+single replication of the experiment.
+
+```{r null50-one-simulation}
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
+```
+
+```{r, null50-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
+rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
+```
+
+```{r, null50-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+```
+
+## Benchmark Metrics
+
+We plot the averaged results across `r B` replications.
+
+```{r null50-metrics-averages, results = "hide"}
+resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(resdf, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE) 
+```
+
+We also take a look at the distribution of rejects for each method as a function of
+the effect size and independent covariate.
+
+```{r null50-metrics-covlineplot, results = "hide"}
+covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+```
+
+Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
+the overlap of rejections between methods.
+
+```{r null50-metrics-upset, results = "hide"}
+if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
+    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
+} else {
+    message("Not enough methods found rejections at alpha ", ualpha, 
+            "; skipping upset plot")
+}
+```
+
 # 80% Null Setting
 
-First, we consider the setting where 80% of tests are null.
+Next, we consider the setting where 80% of tests are null.
 

### 2018-06-07 datasets/simulations/simulations-varyinginformativeness.Rmd
Try simulations w/ varying informativeness

---
title: "Simulation Study: Varying Informativeness"
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

# Summary

In this set of simulations, we consider settings with both null and non-null
tests with varying levels of informativeness of an informative covariate. The
covariate is sampled uniformly from the interval [0, 1], and the conditional
probability of a test being non-null is a step function of the covariate with
two levels (over the intervals [0, 4/5], (4/5, 1]) where the size of the step
is used to indicate the informativeness of the covariate.

For all settings, the marginal proportion of null hypotheses is set to 80%.
A tuning parameter is used to modify the informativeness of the covariate
such that 0 corresponds to an uninformative covariate, and 1 corresponds to a
fully informative covariate (i.e. all tests with covariate values in [0, 4/5]
are null and all tests in (4/5, 1] are non-null).

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(SummarizedBenchmark)
library(parallel)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
resdir <- "results"
dir.create(resdir, showWarnings = FALSE, recursive = TRUE)

## intermediary files we create below
info00_file <- file.path(resdir, "varyinginfo-benchmark-level00.rds")
info20_file <- file.path(resdir, "varyinginfo-benchmark-level20.rds")
info40_file <- file.path(resdir, "varyinginfo-benchmark-level40.rds")
info60_file <- file.path(resdir, "varyinginfo-benchmark-level60.rds")
info80_file <- file.path(resdir, "varyinginfo-benchmark-level80.rds")
info100_file <- file.path(resdir, "varyinginfo-benchmark-level100.rds")

## number of cores for parallelization
cores <- 20
B <- 100

## define bechmarking design
bd <- initializeBenchDesign()
```

As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
for simulations with Gaussian noise. Again, we include both
`nulltype = "empirical"` and `nulltype = "theoretical"`. Since all settings in this
series of simulations use test statistics simulated with Gaussian noise, we include
Scott's FDR Regression in all of the comparisons.

```{r}
bdplus <- bd
bdplus <- addBMethod(bdplus, "scott-theoretical",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'theoretical',
                     control = list(lambda = 0.01))
bdplus <- addBMethod(bdplus, "scott-empirical",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'empirical',
                     control = list(lambda = 0.01))
```

All simulation settings will share the following parameters.

```{r parameters-shared}
m <- 20000                        # integer: number of hypothesis tests
tstat <- rnorm_generator(4)       # functional: dist of alternative test stats
tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
icovariate <- runif               # functional: independent covariate
```

Simulation results will be presented excluding a subset of methods, and
for certain plots (upset plots), a single alpha cutoff will be used.

```{r}
excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
ualpha <- 0.05
```

# Informativeness = 0 Setting

First, we consider the setting where the informativeness is 0.

## Data Simulation

```{r info00-parameters}
pi0 <- pi0_varyinfo80(0.00)             # numeric: proportion of null hypotheses
seed <- 1010
```

We next run the simulations.

```{r info00-run-simulation}
if (file.exists(info00_file)) {
    res <- readRDS(info00_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = info00_file)
}
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r info00-one-simulation}
onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, info00-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, info00-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r info00-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r info00-metrics-covlineplot, results = "hide"}
covariateLinePlot(res, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r info00-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

# Informativeness = 20 Setting

Next, we consider the setting where the informativeness is 0.20.

## Data Simulation

```{r info20-parameters}
pi0 <- pi0_varyinfo80(0.20)             # numeric: proportion of null hypotheses
seed <- 1201
```

We next run the simulations.

```{r info20-run-simulation}
if (file.exists(info20_file)) {
    res <- readRDS(info20_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = info20_file)
}
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r info20-one-simulation}
onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, info20-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, info20-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r info20-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r info20-metrics-covlineplot, results = "hide"}
covariateLinePlot(res, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r info20-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

# Informativeness = 40 Setting

Next, we consider the setting where the informativeness is 0.40.

## Data Simulation

```{r info40-parameters}
pi0 <- pi0_varyinfo80(0.40)             # numeric: proportion of null hypotheses
seed <- 51
```

We next run the simulations.

```{r info40-run-simulation}
if (file.exists(info40_file)) {
    res <- readRDS(info40_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = info40_file)
}
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r info40-one-simulation}
onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, info40-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, info40-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r info40-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r info40-metrics-covlineplot, results = "hide"}
covariateLinePlot(res, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r info40-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

# Informativeness = 60 Setting

Next, we consider the setting where the informativeness is 0.60.

## Data Simulation

```{r info60-parameters}
pi0 <- pi0_varyinfo80(0.60)             # numeric: proportion of null hypotheses
seed <- 500
```

We next run the simulations.

```{r info60-run-simulation}
if (file.exists(info60_file)) {
    res <- readRDS(info60_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = info60_file)
}
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r info60-one-simulation}
onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, info60-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, info60-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r info60-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r info60-metrics-covlineplot, results = "hide"}
covariateLinePlot(res, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r info60-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

# Informativeness = 80 Setting

Next, we consider the setting where the informativeness is 0.80.

## Data Simulation

```{r info80-parameters}
pi0 <- pi0_varyinfo80(0.80)             # numeric: proportion of null hypotheses
seed <- 608
```

We next run the simulations.

```{r info80-run-simulation}
if (file.exists(info80_file)) {
    res <- readRDS(info80_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = info80_file)
}
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r info80-one-simulation}
onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, info80-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, info80-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r info80-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r info80-metrics-covlineplot, results = "hide"}
covariateLinePlot(res, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r info80-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

# Informativeness = 100 Setting

Finally, we consider the setting where the informativeness is 1.00.

## Data Simulation

```{r info100-parameters}
pi0 <- pi0_varyinfo80(0.100)             # numeric: proportion of null hypotheses
seed <- 808
```

We next run the simulations.

```{r info100-run-simulation}
if (file.exists(info100_file)) {
    res <- readRDS(info100_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = info100_file)
}
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r info100-one-simulation}
onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, info100-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, info100-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r info100-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r info100-metrics-covlineplot, results = "hide"}
covariateLinePlot(res, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r info100-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

### 2018-06-07 datasets/R/helpers-simulations.R

@@ -152,6 +152,7 @@ pi0_cosine <- function(pi0) {
     }
 }
 
+## shifted/stretched sine function (non-monotone)
 pi0_sine <- function(pi0) {
     stopifnot(pi0 < 1L, pi0 >= 0L)
     if (pi0 >= 0.5) {
@@ -161,6 +162,11 @@ pi0_sine <- function(pi0) {
     }
 }
 
+## simple step function to test different levels of "informativeness" with pi0 = 80%
+pi0_varyinfo80 <- function(zz) {
+    stopifnot(zz <= 1L, zz >= 0L)
+    function(x) { ifelse(x > 4/5, 0.8 * (1 - zz), 0.2 * (4 + zz)) }
+}
 
 ## ##############################################################################
 ## CODE DERIVED FROM EXTERNAL SOURCE

### 2018-06-07 datasets/simulations/simulations-informative-sine.Rmd
Fix typo in results file name

@@
 tdist_file <- file.path(resdir, "informative-sine-benchmark-t5.rds")
-tdist11_file <- file.path(resdir, "informative-cubic-benchmark-t11.rds")
+tdist11_file <- file.path(resdir, "informative-sine-benchmark-t11.rds")
 chisq_file <- file.path(resdir, "informative-sine-benchmark-chisq4.rds")

### 2018-06-08 datasets/simulations/simulations-varyinginformativeness-lowsignal.Rmd
Fix typo in 100% informative sims; add lower signal sims

--- 2018-06-07 datasets/simulations/simulations-varyinginformativeness.Rmd
@@
 ## intermediary files we create below
-info00_file <- file.path(resdir, "varyinginfo-benchmark-level00.rds")
-info20_file <- file.path(resdir, "varyinginfo-benchmark-level20.rds")
-info40_file <- file.path(resdir, "varyinginfo-benchmark-level40.rds")
-info60_file <- file.path(resdir, "varyinginfo-benchmark-level60.rds")
-info80_file <- file.path(resdir, "varyinginfo-benchmark-level80.rds")
-info100_file <- file.path(resdir, "varyinginfo-benchmark-level100.rds")
+info00_file <- file.path(resdir, "varyinginfo-lowsignal-benchmark-level00.rds")
+info20_file <- file.path(resdir, "varyinginfo-lowsignal-benchmark-level20.rds")
+info40_file <- file.path(resdir, "varyinginfo-lowsignal-benchmark-level40.rds")
+info60_file <- file.path(resdir, "varyinginfo-lowsignal-benchmark-level60.rds")
+info80_file <- file.path(resdir, "varyinginfo-lowsignal-benchmark-level80.rds")
+info100_file <- file.path(resdir, "varyinginfo-lowsignal-benchmark-level100.rds")
 
@@
 m <- 20000                        # integer: number of hypothesis tests
-tstat <- rnorm_generator(4)       # functional: dist of alternative test stats
+tstat <- rnorm_generator(2)       # functional: dist of alternative test stats
 tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
@@
 ```{r info100-parameters}
-pi0 <- pi0_varyinfo80(0.100)             # numeric: proportion of null hypotheses
+pi0 <- pi0_varyinfo80(1.00)             # numeric: proportion of null hypotheses
 seed <- 808

### 2018-06-08 datasets/simulations/simulations-varyingpi0-lowsignal.Rmd
Fix typo in 100% informative sims; add lower signal sims

--- 2018-06-07 datasets/simulations/simulations-varyingpi0.Rmd
@@
 ---
-title: "Simulation Study: Varying Null Proportion"
+title: "Simulation Study: Varying Null Proportion (lower signal)"
 author: "Patrick Kimes"
@@
 ## intermediary files we create below
-null05_file <- file.path(resdir, "varyingpi0-benchmark-nullprop05.rds")
-null10_file <- file.path(resdir, "varyingpi0-benchmark-nullprop10.rds")
-null20_file <- file.path(resdir, "varyingpi0-benchmark-nullprop20.rds")
-null50_file <- file.path(resdir, "varyingpi0-benchmark-nullprop50.rds")
-null80_file <- file.path(resdir, "varyingpi0-benchmark-nullprop80.rds")
-null90_file <- file.path(resdir, "varyingpi0-benchmark-nullprop90.rds")
-null95_file <- file.path(resdir, "varyingpi0-benchmark-nullprop95.rds")
-null99_file <- file.path(resdir, "varyingpi0-benchmark-nullprop99.rds")
+null05_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop05.rds")
+null10_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop10.rds")
+null20_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop20.rds")
+null50_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop50.rds")
+null80_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop80.rds")
+null90_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop90.rds")
+null95_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop95.rds")
+null99_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop99.rds")
 

### 2018-06-08 manuscript_figures/manuscript_figures_sims.Rmd
Add possible supplement sim drafts

---
title: "Simulation Figures for Genome Biology (simulations)"
author: "Rafalab"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Figure2-load-data}
# Load packages and source benchmark FDR
library(tidyr)
library(dplyr)
library(ggplot2)
library(magrittr)
library(cowplot)
library(tibble)
library(ggthemes)
library(grid)
library(SummarizedBenchmark)

## load helper functions
for (f in list.files("../datasets/R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

# assumes sb objects are in the following location
path <- file.path("..", "datasets", "simulations", "results")

# set up results directory
outdir <- "./figures"
dir.create(outdir, showWarnings = FALSE)

# set alpha cutoff for plots that are fixed at a certain value of alpha
alpha <- 0.10

methodset <- c("bonf", "bh", "ihw-a10", "qvalue", "bl-df03", "lfdr",
               "scott-empirical", "scott-theoretical", "ashq")
```

# Data Preparation

First, we load summaries for all simulations results.

```{r}
summary_file <- paste0("allres_alpha", 100*alpha, ".rds")

if (file.exists(summary_file)) {
    res <- readRDS(summary_file)
} else {
    res_files <- list.files(path, "\\.rds$", full.names = TRUE)
    res <- lapply(res_files, function(x) {
        zz <- readRDS(x)
        zz <- plotsim_standardize(zz, alpha = alpha)
        zz <- dplyr::select(zz, rep, blabel, key, alpha, value)

        zz <- dplyr::group_by(zz, blabel, key, alpha)
        zz <- dplyr::summarize(zz,
                               med_value = median(value, na.rm = TRUE),
                               mean_value = mean(value, na.rm = TRUE),
                               nNA = sum(is.na(value)))
        dplyr::ungroup(zz)
    })
    names(res) <- gsub("\\.rds$", "", basename(res_files))
    saveRDS(res, summary_file)
}

## merge all simulation summaries
res <- bind_rows(res, .id = "setting")

## create shorter simulation setting labels
res <- dplyr::mutate(res, dist = gsub(".*?-benchmark-(.*)", "\\1", setting),
                     setting = gsub("(.*?)-benchmark-.*", "\\1", setting))

## only use subset of methods for plotting
res <- dplyr::filter(res, blabel %in% methodset)
res <- dplyr::rename(res, Method = blabel)

## remove suffixes for unused tuning parameters
res <- dplyr::mutate(res, Method = gsub("-df03", "", Method),
                     Method = gsub("(-a)(.*)", "", Method))
methodset <- gsub("-df03", "", methodset)
methodset <- gsub("(-a)(.*)", "", methodset)

## add candy color scheme from datasets/R/colorPalette.R
res <- dplyr::left_join(res, candycols, by = "Method")

## treat methods as factors for forced ordering
res <- dplyr::mutate(res, Method = factor(Method, levels = methodset))
```

# Null Simulations

We subset on simulation results that look at informative and non-informative covariates across
different noise distributions.

```{r}
res_nullset <- dplyr::filter(res, setting == "null")
res_nullset <- dplyr::mutate(res_nullset, dist = factor(dist, levels = c("gaussian", "t5", "chisq4")))
res_nullset <- tidyr::complete(res_nullset, setting, Method, dist, key)
```

For null simulations, rather than the false discovery rate, we take a look at the number
false positives since the FDR will be 1 if any rejections are made.

In these plots, `ashq`, `scott-theoretical` and `scott-theoretical` are excluded for the Chi-Squared
simulation settings.

```{r, fig.width = 8, fig.height = 3}
dplyr::filter(res_nullset, key == "rejectprop") %>%
    filter(! (Method == "ashq" & dist == "chisq4"),
           ! (grepl("scott", Method) & dist == "chisq4")) %>%
    ggplot(aes(x = Method, y = mean_value)) +
    geom_bar(color = "black", fill = "gray", stat = "identity") +
    facet_grid(. ~ dist) +
    coord_flip() +
    scale_y_continuous("Average % Rejected Tests", breaks = seq(0, 1, .01), labels = scales::percent) +
    theme_bw() +
    ggtitle("Rejected proportion of in null simulations (average over 100 replications)") 
```

Additionally, we take a look at the FWER (the proportion of simulation replicates out of 100
with at least one rejection).

```{r, fig.width = 8, fig.height = 3}
gp <- dplyr::filter(res_nullset, key == "FWER") %>%
    filter(! (Method == "ashq" & dist == "chisq4"),
           ! (grepl("scott", Method) & dist == "chisq4")) %>%
    ggplot(aes(x = Method, y = mean_value)) +
    geom_bar(color = "black", fill = "gray", stat = "identity") +
    facet_grid(. ~ dist) +
    coord_flip() +
    scale_y_continuous("FWER", breaks = seq(0, 1, .2), labels = scales::percent) +
    theme_bw() +
    ggtitle("Proportion of replications with at least one false positive (over 100 replications)")

ggsave(file.path(outdir, "FigureSsim-null.pdf"), gp, width = 8, height = 3)
```

# Informative and Non-Informative Covariates

We subset on simulation results that look at informative and non-informative covariates across
different noise distributions.

```{r}
res_covset <- dplyr::filter(res, grepl("informative", setting))
res_covset <- dplyr::mutate(res_covset, dist = factor(dist,
                                                      levels = c("gaussian", "t11", "t5", "chisq4"),
                                                      labels = c(expression("Noise: " * N(0, 1)),
                                                                 expression("Noise: " * t[11]),
                                                                 expression("Noise: " * t[5]),
                                                                 expression("Noise: " * {chi^2}[4]))))
res_covset <- dplyr::mutate(res_covset, setting = factor(setting,
                                                         levels = c("noninformative", "informative-step",
                                                                    "informative-cubic", "informative-sine")))
res_covset <- tidyr::complete(res_covset, setting, Method, dist, key)
```

First, we look at the false discovery rate across the 100 replications.
We create the plot with a discrete color scale.

```{r, fig.width = 8, fig.height = 6}
fdr_bins <- c(0, .05, .07, .09, .11, .13, .15, .2, 1)
dplyr::filter(res_covset, key == "FDR") %>%
    dplyr::mutate(val = cut(mean_value, breaks = fdr_bins, include.lowest = TRUE)) %>%
    ggplot() +
    geom_tile(aes(x = setting, y = Method, fill = val)) +
    scale_x_discrete("Setting", expand = c(0, 0)) + 
    scale_y_discrete(expand = c(0, 0)) + 
    scale_fill_manual("FDR", values = RColorBrewer::brewer.pal(9, "PRGn")[2:9],
                      na.value = "gray50", drop = FALSE) +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
    guides(fill = guide_legend(keywidth = 2, keyheight = 1)) +
    facet_grid(. ~ dist, scales = "free", space = "free", labeller= label_parsed) +
    ggtitle(paste0("Mean FDR across 100 replications (alpha = ", alpha, ")"))
```

We can create this plot with a continuous color scale as well.
However, it is incredibly hard to distinguish values if we look at the entire range of
possible values, so we will clamp them down to be between 0 and 0.2 (values above 0.2 will be
set to 0.2).

```{r, fig.width = 8, fig.height = 6}
gp_fdr <- dplyr::filter(res_covset, key == "FDR", setting != "null") %>%
    ggplot() +
    geom_tile(aes(x = setting, y = Method, fill = mean_value)) +
    scale_x_discrete("Setting", expand = c(0, 0)) + 
    scale_y_discrete(expand = c(0, 0)) + 
    scale_fill_gradient2("FDR", low = RColorBrewer::brewer.pal(6, "PRGn")[1],
                         high = RColorBrewer::brewer.pal(6, "PRGn")[6],
                         midpoint = alpha, na.value = "gray50", limits = c(0, .2),
                         oob = scales::squish) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
    facet_grid(. ~ dist, scales = "free", space = "free", labeller = label_parsed) + 
    ggtitle(paste0("Mean FDR across 100 replications (alpha = ", alpha, ")"))
gp_fdr
```

We can also look at the true positive rates.

```{r, fig.width = 8, fig.height = 6}
gp_tpr <- dplyr::filter(res_covset, key == "TPR") %>%
    ggplot() +
    geom_tile(aes(x = setting, y = Method, fill = mean_value)) +
    scale_x_discrete("Setting", expand = c(0, 0)) + 
    scale_y_discrete(expand = c(0, 0)) + 
    scale_fill_distiller("TPR", palette = "Blues", limits = c(0, 1), direction = 1) +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
    facet_grid(. ~ dist, scales = "free", space = "free", labeller = label_parsed) + 
    ggtitle(paste0("Mean TPR across 100 replications (alpha = ", alpha, ")"))
gp_tpr
```

Alternatively, rather than using heatmaps, we can use scatterplots to look at these values.

```{r, fig.width = 7, fig.height = 6}
## looking at FDR vs TPR
col <- as.character(res_covset$col)
names(col) <- as.character(res_covset$Method)

dplyr::select(res_covset, Method, setting, dist, key, mean_value, col, lty) %>%
    dplyr::filter(key %in% c("FDR", "TPR")) %>%
    tidyr::spread(key, mean_value) %>%
    ggplot(aes(x = FDR, y = TPR, color = Method)) +
    geom_point(alpha = 1/2) +
    scale_color_manual(values = col) +
    facet_grid(dist ~ setting, labeller = labeller(dist = label_parsed)) +
    theme_bw()

## looking at noninformative vs. informative (cubic)
dplyr::select(res_covset, Method, setting, dist, key, mean_value) %>%
    dplyr::filter(setting %in% c("noninformative", "informative-cubic"),
                  key %in% c("TPR"),
                  ! Method %in% c("ashq", "bonf")) %>%
    tidyr::spread(setting, mean_value) %>%
    ggplot(aes(x = noninformative, y = `informative-cubic`, color = Method)) +
    geom_point(alpha = 1/2) +
    scale_color_manual(values = col) +
    facet_wrap( ~ dist, nrow = 1, scales = "free", labeller = labeller(dist = label_parsed)) +
    theme_bw() +
    scale_x_continuous(labels = scales::percent) + 
    scale_y_continuous(labels = scales::percent) + 
    geom_abline(color = "blue", lty = 2, alpha = 1/3) +
    ggtitle("Mean TPR for Informative and Noninformative Simulations")

## looking at noninformative vs. informative (cubic)
pfdr <- dplyr::select(res_covset, Method, setting, dist, key, mean_value) %>%
    dplyr::filter(setting %in% c("noninformative", "informative-cubic"),
                  key %in% c("FDR")) %>%
    tidyr::spread(setting, mean_value) %>%
    ggplot(aes(x = noninformative, y = `informative-cubic`, color = Method)) +
    geom_abline(color = "blue", lty = 2, alpha = 1/3) +
    geom_vline(xintercept = 0.10, color = "red", lty = 2, alpha = 1/3) +
    geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) +
    geom_point(alpha = 1/2) +
    scale_color_manual(values = col) +
    facet_wrap( ~ dist, nrow = 2, scales = "free", labeller = labeller(dist = label_parsed)) +
    theme_bw() +
    scale_x_continuous("FDR w/ Noninformative Covariate", labels = scales::percent, limits = c(0, 1)) + 
    scale_y_continuous("FDR w/ Informative Covariate (cubic)", labels = scales::percent, limits = c(0, 1)) + 
    ggtitle("Mean FDR for Informative and Noninformative Simulations")

## looking at noninformative vs. informative (cubic)
ptpr <- dplyr::select(res_covset, Method, setting, dist, key, mean_value) %>%
    dplyr::filter(setting %in% c("noninformative", "informative-cubic"),
                  key %in% c("TPR")) %>%
    tidyr::spread(setting, mean_value) %>%
    ggplot(aes(x = noninformative, y = `informative-cubic`, color = Method)) +
    geom_abline(color = "blue", lty = 2, alpha = 1/3) +
    geom_vline(xintercept = 0.10, color = "red", lty = 2, alpha = 1/3) +
    geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) +
    geom_point(alpha = 1/2) +
    scale_color_manual(values = col) +
    facet_wrap( ~ dist, nrow = 2, scales = "free", labeller = labeller(dist = label_parsed)) +
    theme_bw() +
    scale_x_continuous("TPR w/ Noninformative Covariate", labels = scales::percent, limits = c(0, 1)) + 
    scale_y_continuous("TPR w/ Informative Covariate (cubic)", labels = scales::percent, limits = c(0, 1)) + 
    ggtitle("Mean TPR for Informative and Noninformative Simulations") 
```

We can plot both FDR and TPR together.

```{r, fig.width = 11, fig.height = 4.5}
plot_grid(pfdr, ptpr)

gp <- plot_grid(pfdr + coord_cartesian(c(0, .2), c(0, .2)) + guides(color = FALSE, linetype = FALSE),
                ptpr + coord_cartesian(c(.6, 1), c(.6, 1)), rel_widths = c(1, 1.4), labels = LETTERS[1:2])
gp

ggsave(file.path(outdir, "FigureSsim-noise.pdf"), gp, width = 11, height = 4.5)
```

# Varying Null Proportions

Next, we subset on simulation results that look at varying the null proportions.

```{r}
res_pi0set <- dplyr::filter(res, grepl("varyingpi0", setting))
res_pi0set <- dplyr::rename(res_pi0set, pi0 = dist)
res_pi0set <- dplyr::mutate(res_pi0set, pi0 = as.numeric(gsub("nullprop", "", pi0)))
res_pi0set <- tidyr::complete(res_pi0set, setting, Method, pi0, key)
res_pi0set <- dplyr::mutate(res_pi0set, setting = factor(setting, levels = c("varyingpi0", "varyingpi0-lowsignal"),
                                                         labels = c("high signal", "low signal")))
```

We take a look at the TPR and FDR as function of the proportion of null hypotheses.

```{r, fig.width = 8, fig.height = 10}
## looking at FDR vs TPR
col <- as.character(res_pi0set$col)
names(col) <- as.character(res_pi0set$Method)
lty <- as.character(res_pi0set$lty)
names(lty) <- as.character(res_pi0set$Method)

pfdr <- dplyr::select(res_pi0set, Method, setting, pi0, key, mean_value, col, lty) %>%
    dplyr::filter(key == "FDR") %>%
    ggplot(aes(x = pi0 / 100, y = mean_value, color = Method, lty = Method)) +
    geom_line(alpha = 1/6, position = position_dodge(width = .01)) +
    geom_point(alpha = 1/2, position = position_dodge(width = .01)) +
    scale_color_manual(values = col) +
    scale_linetype_manual(values = lty) + 
    geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) + 
    scale_x_continuous("Null Proportion", labels = scales::percent, breaks = seq(0, 1, .1)) + 
    scale_y_continuous("FDR", labels = scales::percent) + 
    theme_bw() +
    facet_grid(setting ~ .) + 
    ggtitle("Average FDR Across Varying Null Proportion")
pfdr

pfdr + coord_cartesian(ylim = c(0, .25))

ptpr <- dplyr::select(res_pi0set, Method, setting, pi0, key, mean_value, col, lty) %>%
    dplyr::filter(key == "TPR") %>%
    ggplot(aes(x = pi0 / 100, y = mean_value, color = Method, lty = Method)) +
    geom_line(alpha = 1/6, position = position_dodge(width = .01)) +
    geom_point(alpha = 1/2, position = position_dodge(width = .01)) +
    scale_color_manual(values = col) +
    scale_linetype_manual(values = lty) + 
    scale_x_continuous("Null Proportion", labels = scales::percent, breaks = seq(0, 1, .1)) + 
    scale_y_continuous("TPR", labels = scales::percent) + 
    theme_bw() +
    facet_grid(setting ~ .) + 
    ggtitle("Average TPR Across Varying Null Proportion")
ptpr
```

Again, we can plot both FDR and TPR together.

```{r, fig.width = 10, fig.height = 5}
gp <- plot_grid(pfdr + coord_cartesian(ylim = c(0, .25)) + guides(color = FALSE, linetype = FALSE),
                ptpr, rel_widths = c(1, 1.25), labels = LETTERS[1:2])
gp
ggsave(file.path(outdir, "FigureSsim-nullprop.pdf"), gp, width = 10, height = 5)
```

# Unimodal Effect Sizes

Next, we subset on simulation results that look at varying unimodal simulation settings
that were used in the simulations of the ASH paper.

```{r}
res_unimodal <- dplyr::filter(res, grepl("varyinges", setting), dist != "bimodal")
res_unimodal <- dplyr::rename(res_unimodal, signal = dist)
res_unimodal <- tidyr::separate(res_unimodal, setting, c("setting", "dist"), by = "-", fill = "right")
res_unimodal$dist[is.na(res_unimodal$dist)] <- "gaussian"
res_unimodal <- tidyr::complete(res_unimodal, setting, Method, dist, signal, key)
res_unimodal <- dplyr::mutate(res_unimodal, dist = factor(dist,
                                                          levels = c("gaussian", "t", "chisq"),
                                                          labels = c(expression("Noise: " * N(0, 1)),
                                                                 expression("Noise: " * t[11]),
                                                                 expression("Noise: " * {chi^2}[4]))))
```

We take a look at the TPR and FDR as function of the informativeness.

```{r, fig.width = 8, fig.height = 5}
## looking at FDR vs TPR
col <- as.character(res_unimodal$col)
names(col) <- as.character(res_unimodal$Method)
lty <- as.character(res_unimodal$lty)
names(lty) <- as.character(res_unimodal$Method)

uniplot <- dplyr::select(res_unimodal, Method, dist, signal, key, mean_value, col, lty) %>%
    dplyr::filter(key %in% c("FDR", "TPR")) %>%
    tidyr::spread(key, mean_value) %>%
    ggplot(aes(x = TPR, y = FDR, color = Method)) +
    geom_point(alpha = 1/2) +
    scale_color_manual(values = col) +
    geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) + 
    scale_x_continuous("TPR", labels = scales::percent, breaks = seq(0, 1, .2)) + 
    scale_y_continuous("FDR", labels = scales::percent, breaks = seq(0, 1, .2)) + 
    theme_bw() +
    facet_grid(signal ~ dist, labeller = labeller(dist = label_parsed)) + 
    ggtitle("Average FDR and TPR Across Unimodal Settings")
uniplot

ggsave(file.path(outdir, "FigureSsim-unimodal.pdf"), uniplot, width = 8, height = 5)
```

# Varying Informativeness

Next, we subset on simulation results that look at varying informativeness of the
informative covariate.

```{r}
res_infoset <- dplyr::filter(res, grepl("varyinginfo", setting))
res_infoset <- dplyr::rename(res_infoset, level = dist)
res_infoset <- dplyr::mutate(res_infoset, level = as.numeric(gsub("level", "", level)))
res_infoset <- tidyr::complete(res_infoset, setting, Method, level, key)
res_infoset <- dplyr::mutate(res_infoset, setting = factor(setting, levels = c("varyinginfo", "varyinginfo-lowsignal"),
                                                         labels = c("high signal", "low signal")))
```

We take a look at the TPR and FDR as function of the informativeness.

```{r, fig.width = 8, fig.height = 10}
## looking at FDR vs TPR
col <- as.character(res_infoset$col)
names(col) <- as.character(res_infoset$Method)
lty <- as.character(res_infoset$lty)
names(lty) <- as.character(res_infoset$Method)

pfdr <- dplyr::select(res_infoset, Method, setting, level, key, mean_value, col, lty) %>%
    dplyr::filter(key == "FDR") %>%
    ggplot(aes(x = level / 100, y = mean_value, color = Method, lty = Method)) +
    geom_line(alpha = 1/6, position = position_dodge(width = .01)) +
    geom_point(alpha = 1/2, position = position_dodge(width = .01)) +
    scale_color_manual(values = col) +
    scale_linetype_manual(values = lty) + 
    geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) + 
    scale_x_continuous("Informativeness", labels = scales::percent, breaks = seq(0, 1, .1)) + 
    scale_y_continuous("FDR", labels = scales::percent) + 
    theme_bw() +
    facet_grid(setting ~ .) + 
    ggtitle("Average FDR Across Varying Informativeness")
pfdr

ptpr <- dplyr::select(res_infoset, Method, setting, level, key, mean_value, col, lty) %>%
    dplyr::filter(key == "TPR") %>%
    ggplot(aes(x = level / 100, y = mean_value, color = Method, lty = Method)) +
    geom_line(alpha = 1/6, position = position_dodge(width = .01)) +
    geom_point(alpha = 1/2, position = position_dodge(width = .01)) +
    scale_color_manual(values = col) +
    scale_linetype_manual(values = lty) + 
    scale_x_continuous("Informativeness", labels = scales::percent, breaks = seq(0, 1, .1)) + 
    scale_y_continuous("TPR", labels = scales::percent) + 
    theme_bw() +
    facet_grid(setting ~ ., scales = "free_y") + expand_limits(y = 0) +
    ggtitle("Average TPR Across Varying Informativeness")
ptpr
```

Again, we can plot both FDR and TPR together.

```{r, fig.width = 10, fig.height = 5}
gp <- plot_grid(pfdr + guides(color = FALSE, linetype = FALSE), ptpr,
          rel_widths = c(1, 1.25), labels = LETTERS[1:2])
gp

ggsave(file.path(outdir, "FigureSsim-informativeness.pdf"), gp, width = 10, height = 5)
```

### 2018-06-09 datasets/simulations/simulations-varyingpi0-lowsignal.Rmd
Add more levels for varying pi0 simulations

@@
 ---
-title: "Simulation Study: Varying Null Proportion (lower signal)"
+title: "Simulation Study: Varying Null Proportion"
 author: "Patrick Kimes"
@@
 in the setting as described in `simulations-informative-sine.Rmd`.
+The simulations are identical to those performed in `simulations-varyingpi0.Rmd`
+except that the strength of the "singal" (the effect size for non-null tests)
+is weaker (smaller).
 
@@
 null20_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop20.rds")
+null30_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop30.rds")
+null40_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop40.rds")
 null50_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop50.rds")
+null60_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop60.rds")
+null70_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop70.rds")
 null80_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop80.rds")
@@
 
+# 30% Null Setting
+
+Next, we consider the setting where 30% of tests are null.
+
+## Data Simulation
+
+```{r null30-parameters}
+pi0 <- pi0_sine(0.30)             # numeric: proportion of null hypotheses
+seed <- 511
+```
+
+We next run the simulations.
+
+```{r null30-run-simulation}
+if (file.exists(null30_file)) {
+    res <- readRDS(null30_file)
+} else {
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
+                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
+                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    seed = seed, mc.cores = cores)
+    saveRDS(res, file = null30_file)
+}
+```
+
+## Covariate Diagnostics
+
+Here, we show the relationship between the independent covariate and p-values for a
+single replication of the experiment.
+
+```{r null30-one-simulation}
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
+```
+
+```{r, null30-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
+rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
+```
+
+```{r, null30-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+```
+
+## Benchmark Metrics
+
+We plot the averaged results across `r B` replications.
+
+```{r null30-metrics-averages, results = "hide"}
+resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(resdf, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE) 
+```
+
+We also take a look at the distribution of rejects for each method as a function of
+the effect size and independent covariate.
+
+```{r null30-metrics-covlineplot, results = "hide"}
+covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+```
+
+Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
+the overlap of rejections between methods.
+
+```{r null30-metrics-upset, results = "hide"}
+if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
+    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
+} else {
+    message("Not enough methods found rejections at alpha ", ualpha, 
+            "; skipping upset plot")
+}
+```
+
+# 40% Null Setting
+
+Next, we consider the setting where 40% of tests are null.
+
+## Data Simulation
+
+```{r null40-parameters}
+pi0 <- pi0_sine(0.40)             # numeric: proportion of null hypotheses
+seed <- 515
+```
+
+We next run the simulations.
+
+```{r null40-run-simulation}
+if (file.exists(null40_file)) {
+    res <- readRDS(null40_file)
+} else {
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
+                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
+                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    seed = seed, mc.cores = cores)
+    saveRDS(res, file = null40_file)
+}
+```
+
+## Covariate Diagnostics
+
+Here, we show the relationship between the independent covariate and p-values for a
+single replication of the experiment.
+
+```{r null40-one-simulation}
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
+```
+
+```{r, null40-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
+rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
+```
+
+```{r, null40-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+```
+
+## Benchmark Metrics
+
+We plot the averaged results across `r B` replications.
+
+```{r null40-metrics-averages, results = "hide"}
+resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(resdf, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE) 
+```
+
+We also take a look at the distribution of rejects for each method as a function of
+the effect size and independent covariate.
+
+```{r null40-metrics-covlineplot, results = "hide"}
+covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+```
+
+Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
+the overlap of rejections between methods.
+
+```{r null40-metrics-upset, results = "hide"}
+if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
+    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
+} else {
+    message("Not enough methods found rejections at alpha ", ualpha, 
+            "; skipping upset plot")
+}
+```
+
 # 50% Null Setting
@@
 ```{r null50-metrics-upset, results = "hide"}
+if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
+    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
+} else {
+    message("Not enough methods found rejections at alpha ", ualpha, 
+            "; skipping upset plot")
+}
+```
+
+# 60% Null Setting
+
+Next, we consider the setting where 60% of tests are null.
+
+## Data Simulation
+
+```{r null60-parameters}
+pi0 <- pi0_sine(0.60)             # numeric: proportion of null hypotheses
+seed <- 1502
+```
+
+We next run the simulations.
+
+```{r null60-run-simulation}
+if (file.exists(null60_file)) {
+    res <- readRDS(null60_file)
+} else {
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
+                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
+                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    seed = seed, mc.cores = cores)
+    saveRDS(res, file = null60_file)
+}
+```
+
+## Covariate Diagnostics
+
+Here, we show the relationship between the independent covariate and p-values for a
+single replication of the experiment.
+
+```{r null60-one-simulation}
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
+```
+
+```{r, null60-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
+rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
+```
+
+```{r, null60-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+```
+
+## Benchmark Metrics
+
+We plot the averaged results across `r B` replications.
+
+```{r null60-metrics-averages, results = "hide"}
+resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(resdf, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE) 
+```
+
+We also take a look at the distribution of rejects for each method as a function of
+the effect size and independent covariate.
+
+```{r null60-metrics-covlineplot, results = "hide"}
+covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+```
+
+Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
+the overlap of rejections between methods.
+
+```{r null60-metrics-upset, results = "hide"}
+if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
+    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
+} else {
+    message("Not enough methods found rejections at alpha ", ualpha, 
+            "; skipping upset plot")
+}
+```
+
+# 70% Null Setting
+
+Next, we consider the setting where 70% of tests are null.
+
+## Data Simulation
+
+```{r null70-parameters}
+pi0 <- pi0_sine(0.70)             # numeric: proportion of null hypotheses
+seed <- 1722
+```
+
+We next run the simulations.
+
+```{r null70-run-simulation}
+if (file.exists(null70_file)) {
+    res <- readRDS(null70_file)
+} else {
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
+                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
+                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    seed = seed, mc.cores = cores)
+    saveRDS(res, file = null70_file)
+}
+```
+
+## Covariate Diagnostics
+
+Here, we show the relationship between the independent covariate and p-values for a
+single replication of the experiment.
+
+```{r null70-one-simulation}
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
+```
+
+```{r, null70-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
+rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
+```
+
+```{r, null70-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+```
+
+## Benchmark Metrics
+
+We plot the averaged results across `r B` replications.
+
+```{r null70-metrics-averages, results = "hide"}
+resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(resdf, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE) 
+```
+
+We also take a look at the distribution of rejects for each method as a function of
+the effect size and independent covariate.
+
+```{r null70-metrics-covlineplot, results = "hide"}
+covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+```
+
+Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
+the overlap of rejections between methods.
+
+```{r null70-metrics-upset, results = "hide"}
 if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {

### 2018-06-09 datasets/simulations/simulations-varyingpi0.Rmd
Add more levels for varying pi0 simulations

@@
 null20_file <- file.path(resdir, "varyingpi0-benchmark-nullprop20.rds")
+null30_file <- file.path(resdir, "varyingpi0-benchmark-nullprop30.rds")
+null40_file <- file.path(resdir, "varyingpi0-benchmark-nullprop40.rds")
 null50_file <- file.path(resdir, "varyingpi0-benchmark-nullprop50.rds")
+null60_file <- file.path(resdir, "varyingpi0-benchmark-nullprop60.rds")
+null70_file <- file.path(resdir, "varyingpi0-benchmark-nullprop70.rds")
 null80_file <- file.path(resdir, "varyingpi0-benchmark-nullprop80.rds")

### 2018-06-09 manuscript_figures/manuscript_figures_sims.Rmd
Add TNR plot for null; calc more stats; match other plot styles

@@
         zz <- dplyr::summarize(zz,
+                               se = sd(value, na.rm = TRUE) / sqrt(sum(!is.na(value))),
+                               q10 = quantile(value, .1, na.rm = TRUE),
+                               q90 = quantile(value, .9, na.rm = TRUE),
                                med_value = median(value, na.rm = TRUE),
@@
 res_nullset <- tidyr::complete(res_nullset, setting, Method, dist, key)
+
+col <- as.character(res_nullset$col)
+names(col) <- as.character(res_nullset$Method)
 ```
@@
 
-```{r, fig.width = 8, fig.height = 3}
+```{r, fig.width = 8, fig.height = 5}
 dplyr::filter(res_nullset, key == "rejectprop") %>%
@@
            ! (grepl("scott", Method) & dist == "chisq4")) %>%
-    ggplot(aes(x = Method, y = mean_value)) +
-    geom_bar(color = "black", fill = "gray", stat = "identity") +
+    ggplot(aes(x = Method, y = mean_value, color = Method)) +
+    geom_point() +
     facet_grid(. ~ dist) +
-    coord_flip() +
+    scale_color_manual(values = col, guide = FALSE) +
     scale_y_continuous("Average % Rejected Tests", breaks = seq(0, 1, .01), labels = scales::percent) +
-    theme_bw() +
+    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
     ggtitle("Rejected proportion of in null simulations (average over 100 replications)") 
@@
 
-```{r, fig.width = 8, fig.height = 3}
-gp <- dplyr::filter(res_nullset, key == "FWER") %>%
+```{r, fig.width = 8, fig.height = 5}
+dplyr::filter(res_nullset, key == "FWER") %>%
     filter(! (Method == "ashq" & dist == "chisq4"),
            ! (grepl("scott", Method) & dist == "chisq4")) %>%
-    ggplot(aes(x = Method, y = mean_value)) +
-    geom_bar(color = "black", fill = "gray", stat = "identity") +
+    ggplot(aes(x = Method, y = mean_value, color = Method)) +
+    geom_point() +
+    scale_color_manual(values = col, guide = FALSE) +
     facet_grid(. ~ dist) +
-    coord_flip() +
     scale_y_continuous("FWER", breaks = seq(0, 1, .2), labels = scales::percent) +
-    theme_bw() +
+    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
     ggtitle("Proportion of replications with at least one false positive (over 100 replications)") 
+```
 
-ggsave(file.path(outdir, "FigureSsim-null.pdf"), gp, width = 8, height = 3)
+Finally, we plot the TNR as was done for the yeast in silico experiments/simulations.
+
+```{r, fig.width = 8, fig.height = 5}
+gp <- dplyr::filter(res_nullset, key == "TNR") %>%
+    filter(! (Method == "ashq" & dist == "chisq4"),
+           ! (grepl("scott", Method) & dist == "chisq4")) %>%
+    ggplot(aes(x = Method, y = mean_value, color = Method)) +
+    geom_point() +
+    facet_grid(. ~ dist) +
+    scale_y_continuous("TPR", labels = scales::percent) +
+    scale_color_manual(values = col, guide = FALSE) +
+    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
+    ggtitle("Average TNR (over 100 replications)") 
+gp
+
+ggsave(file.path(outdir, "FigureSsim-null.pdf"), gp, width = 8, height = 5)
 ```
@@
 
-First, we look at the false discovery rate across the 100 replications.
-We create the plot with a discrete color scale.
+We use scatterplots to look at the FDR and TPR values.
 
-```{r, fig.width = 8, fig.height = 6}
-fdr_bins <- c(0, .05, .07, .09, .11, .13, .15, .2, 1)
-dplyr::filter(res_covset, key == "FDR") %>%
-    dplyr::mutate(val = cut(mean_value, breaks = fdr_bins, include.lowest = TRUE)) %>%
-    ggplot() +
-    geom_tile(aes(x = setting, y = Method, fill = val)) +
-    scale_x_discrete("Setting", expand = c(0, 0)) + 
-    scale_y_discrete(expand = c(0, 0)) + 
-    scale_fill_manual("FDR", values = RColorBrewer::brewer.pal(9, "PRGn")[2:9],
-                      na.value = "gray50", drop = FALSE) +
-    theme_bw() + 
-    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
-    guides(fill = guide_legend(keywidth = 2, keyheight = 1)) +
-    facet_grid(. ~ dist, scales = "free", space = "free", labeller= label_parsed) +
-    ggtitle(paste0("Mean FDR across 100 replications (alpha = ", alpha, ")"))
-```
-
-We can create this plot with a continuous color scale as well.
-However, it is incredibly hard to distinguish values if we look at the entire range of
-possible values, so we will clamp them down to be between 0 and 0.2 (values above 0.2 will be
-set to 0.2).
-
-```{r, fig.width = 8, fig.height = 6}
-gp_fdr <- dplyr::filter(res_covset, key == "FDR", setting != "null") %>%
-    ggplot() +
-    geom_tile(aes(x = setting, y = Method, fill = mean_value)) +
-    scale_x_discrete("Setting", expand = c(0, 0)) + 
-    scale_y_discrete(expand = c(0, 0)) + 
-    scale_fill_gradient2("FDR", low = RColorBrewer::brewer.pal(6, "PRGn")[1],
-                         high = RColorBrewer::brewer.pal(6, "PRGn")[6],
-                         midpoint = alpha, na.value = "gray50", limits = c(0, .2),
-                         oob = scales::squish) +
-    theme_bw() +
-    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
-    facet_grid(. ~ dist, scales = "free", space = "free", labeller = label_parsed) + 
-    ggtitle(paste0("Mean FDR across 100 replications (alpha = ", alpha, ")"))
-gp_fdr
-```
-
-We can also look at the true positive rates.
-
-```{r, fig.width = 8, fig.height = 6}
-gp_tpr <- dplyr::filter(res_covset, key == "TPR") %>%
-    ggplot() +
-    geom_tile(aes(x = setting, y = Method, fill = mean_value)) +
-    scale_x_discrete("Setting", expand = c(0, 0)) + 
-    scale_y_discrete(expand = c(0, 0)) + 
-    scale_fill_distiller("TPR", palette = "Blues", limits = c(0, 1), direction = 1) +
-    theme_bw() + 
-    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
-    facet_grid(. ~ dist, scales = "free", space = "free", labeller = label_parsed) + 
-    ggtitle(paste0("Mean TPR across 100 replications (alpha = ", alpha, ")"))
-gp_tpr
-```
-
-Alternatively, rather than using heatmaps, we can use scatterplots to look at these values.
-
-```{r, fig.width = 7, fig.height = 6}
+```{r, fig.width = 11, fig.height = 4.5}
 ## looking at FDR vs TPR
@@
 names(col) <- as.character(res_covset$Method)
-
-dplyr::select(res_covset, Method, setting, dist, key, mean_value, col, lty) %>%
-    dplyr::filter(key %in% c("FDR", "TPR")) %>%
-    tidyr::spread(key, mean_value) %>%
-    ggplot(aes(x = FDR, y = TPR, color = Method)) +
-    geom_point(alpha = 1/2) +
-    scale_color_manual(values = col) +
-    facet_grid(dist ~ setting, labeller = labeller(dist = label_parsed)) +
-    theme_bw()
 
@@
     ggplot(aes(x = noninformative, y = `informative-cubic`, color = Method)) +
-    geom_point(alpha = 1/2) +
+    geom_point(alpha = 1) +
     scale_color_manual(values = col) +
     facet_wrap( ~ dist, nrow = 1, scales = "free", labeller = labeller(dist = label_parsed)) +
-    theme_bw() +
     scale_x_continuous(labels = scales::percent) + 
@@
     ggtitle("Mean TPR for Informative and Noninformative Simulations") 
+```
 
+```{r}
 ## looking at noninformative vs. informative (cubic)
@@
     geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) +
-    geom_point(alpha = 1/2) +
+    geom_point(alpha = 1) +
     scale_color_manual(values = col) +
     facet_wrap( ~ dist, nrow = 2, scales = "free", labeller = labeller(dist = label_parsed)) +
-    theme_bw() +
     scale_x_continuous("FDR w/ Noninformative Covariate", labels = scales::percent, limits = c(0, 1)) + 
     scale_y_continuous("FDR w/ Informative Covariate (cubic)", labels = scales::percent, limits = c(0, 1)) + 
-    ggtitle("Mean FDR for Informative and Noninformative Simulations")
+    ggtitle("FDR (Informative and Noninformative Sims)")
 
@@
     geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) +
-    geom_point(alpha = 1/2) +
+    geom_point(alpha = 1) +
     scale_color_manual(values = col) +
     facet_wrap( ~ dist, nrow = 2, scales = "free", labeller = labeller(dist = label_parsed)) +
-    theme_bw() +
     scale_x_continuous("TPR w/ Noninformative Covariate", labels = scales::percent, limits = c(0, 1)) + 
     scale_y_continuous("TPR w/ Informative Covariate (cubic)", labels = scales::percent, limits = c(0, 1)) + 
-    ggtitle("Mean TPR for Informative and Noninformative Simulations") 
+    ggtitle("TPR (Informative and Noninformative Sims)") 
 ```
@@
 
-```{r, fig.width = 11, fig.height = 4.5}
-plot_grid(pfdr, ptpr)
+```{r, fig.width = 15, fig.height = 7.5}
+plot_grid(pfdr + guides(color = FALSE, linetype = FALSE), ptpr)
 
@@
 
-ggsave(file.path(outdir, "FigureSsim-noise.pdf"), gp, width = 11, height = 4.5)
+ggsave(file.path(outdir, "FigureSsim-noise.pdf"), gp, width = 15, height = 7.5)
 ```
@@
 
-```{r, fig.width = 8, fig.height = 10}
+```{r}
 ## looking at FDR vs TPR
@@
     geom_line(alpha = 1/6, position = position_dodge(width = .01)) +
-    geom_point(alpha = 1/2, position = position_dodge(width = .01)) +
+    geom_point(alpha = 1, position = position_dodge(width = .01)) +
     scale_color_manual(values = col) +
@@
     scale_y_continuous("FDR", labels = scales::percent) + 
-    theme_bw() +
     facet_grid(setting ~ .) + 
     ggtitle("Average FDR Across Varying Null Proportion")
-pfdr
-
-pfdr + coord_cartesian(ylim = c(0, .25))
 
@@
     scale_y_continuous("TPR", labels = scales::percent) + 
-    theme_bw() +
     facet_grid(setting ~ .) + 
     ggtitle("Average TPR Across Varying Null Proportion")
-ptpr
 ```
@@
 
-```{r, fig.width = 10, fig.height = 5}
+```{r, fig.width = 15, fig.height = 7.5}
+plot_grid(pfdr + guides(color = FALSE, linetype = FALSE),
+          ptpr, rel_widths = c(1, 1.25), labels = LETTERS[1:2])
+
+## trimmed ranges
 gp <- plot_grid(pfdr + coord_cartesian(ylim = c(0, .25)) + guides(color = FALSE, linetype = FALSE),
@@
 gp
-ggsave(file.path(outdir, "FigureSsim-nullprop.pdf"), gp, width = 10, height = 5)
+
+ggsave(file.path(outdir, "FigureSsim-nullprop.pdf"), gp, width = 15, height = 7.5)
 ```
@@
 
-```{r, fig.width = 8, fig.height = 5}
+```{r, fig.width = 12, fig.height = 7.5}
 ## looking at FDR vs TPR
@@
     ggplot(aes(x = TPR, y = FDR, color = Method)) +
-    geom_point(alpha = 1/2) +
+    geom_point(alpha = 1) +
     scale_color_manual(values = col) +
@@
     scale_y_continuous("FDR", labels = scales::percent, breaks = seq(0, 1, .2)) + 
-    theme_bw() +
     facet_grid(signal ~ dist, labeller = labeller(dist = label_parsed)) + 
@@
 
-ggsave(file.path(outdir, "FigureSsim-unimodal.pdf"), uniplot, width = 8, height = 5)
+ggsave(file.path(outdir, "FigureSsim-unimodal.pdf"), uniplot, width = 12, height = 7.5)
 ```
@@
     scale_y_continuous("FDR", labels = scales::percent) + 
-    theme_bw() +
     facet_grid(setting ~ .) + 
@@
     scale_y_continuous("TPR", labels = scales::percent) + 
-    theme_bw() +
     facet_grid(setting ~ ., scales = "free_y") + expand_limits(y = 0) +
@@
 
-```{r, fig.width = 10, fig.height = 5}
+```{r, fig.width = 15, fig.height = 7.5}
 gp <- plot_grid(pfdr + guides(color = FALSE, linetype = FALSE), ptpr,
@@
 
-ggsave(file.path(outdir, "FigureSsim-informativeness.pdf"), gp, width = 10, height = 5)
+ggsave(file.path(outdir, "FigureSsim-informativeness.pdf"), gp, width = 15, height = 7.5)
 ```

### 2018-06-11 manuscript_figures/manuscript_figures_sims.Rmd
Add zoomed/unzoomed noise plots

@@
 ```{r, fig.width = 15, fig.height = 7.5}
-plot_grid(pfdr + guides(color = FALSE, linetype = FALSE), ptpr)
-
-gp <- plot_grid(pfdr + coord_cartesian(c(0, .2), c(0, .2)) + guides(color = FALSE, linetype = FALSE),
-                ptpr + coord_cartesian(c(.6, 1), c(.6, 1)), rel_widths = c(1, 1.4), labels = LETTERS[1:2])
+gp <- plot_grid(pfdr + guides(color = FALSE, linetype = FALSE), ptpr, , rel_widths = c(1, 1.4), labels = LETTERS[1:2])
 gp
 
-ggsave(file.path(outdir, "FigureSsim-noise.pdf"), gp, width = 15, height = 7.5)
+ggsave(file.path(outdir, "FigureSsim-noise.pdf"), gp, width = 11, height = 4.5)
+
+gp_zoom <- plot_grid(pfdr + coord_cartesian(c(0, .2), c(0, .2)) + guides(color = FALSE, linetype = FALSE),
+                     ptpr + coord_cartesian(c(.6, 1), c(.6, 1)), rel_widths = c(1, 1.4), labels = LETTERS[1:2])
+gp_zoom
+
+ggsave(file.path(outdir, "FigureSsim-noise-zoom.pdf"), gp_zoom, width = 11, height = 4.5)
 ```

### 2018-06-11 manuscript_figures/manuscript_figures_sims.Rmd
Try adding SE bars to simulation summary plots

@@
     geom_point() +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se)) + 
     facet_grid(. ~ dist) +
@@
     theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
-    ggtitle("Rejected proportion of in null simulations (average over 100 replications)") 
+    ggtitle("Rejected proportion in null simulations (average over 100 replications)") 
 ```
@@
 ## looking at noninformative vs. informative (cubic)
-dplyr::select(res_covset, Method, setting, dist, key, mean_value) %>%
+dplyr::select(res_covset, Method, setting, dist, key, mean_value, se) %>%
     dplyr::filter(setting %in% c("noninformative", "informative-cubic"),
@@
                   ! Method %in% c("ashq", "bonf")) %>%
-    tidyr::spread(setting, mean_value) %>%
-    ggplot(aes(x = noninformative, y = `informative-cubic`, color = Method)) +
-    geom_point(alpha = 1) +
+    dplyr::mutate(plus1se = mean_value + se, minus1se = mean_value - se) %>%
+    dplyr::select(-se) %>%
+    tidyr::gather(metric, value, mean_value, plus1se, minus1se) %>%
+    tidyr::unite(metric, c("setting", "metric")) %>%
+    tidyr::spread(metric, value) %>%
+    ggplot(aes(x = noninformative_mean_value, y = `informative-cubic_mean_value`, color = Method)) +
+    geom_errorbar(aes(ymin = `informative-cubic_minus1se`, ymax = `informative-cubic_plus1se`), width = 0) +
+    geom_errorbarh(aes(xmin = noninformative_minus1se, xmax = noninformative_plus1se), height = 0) + 
+    ##geom_point(alpha = 1) +
     scale_color_manual(values = col) +
@@
 ```{r, fig.width = 15, fig.height = 7.5}
-gp <- plot_grid(pfdr + guides(color = FALSE, linetype = FALSE), ptpr, , rel_widths = c(1, 1.4), labels = LETTERS[1:2])
+gp <- plot_grid(pfdr + guides(color = FALSE, linetype = FALSE), ptpr,
+                rel_widths = c(1, 1.4), labels = LETTERS[1:2])
 gp
@@
 
-pfdr <- dplyr::select(res_pi0set, Method, setting, pi0, key, mean_value, col, lty) %>%
-    dplyr::filter(key == "FDR") %>%
+## not including SE because it's too small
+pfdr <- dplyr::select(res_pi0set, Method, setting, pi0, key, mean_value, se, col, lty) %>%
+    dplyr::filter(key == "FDR", !is.na(mean_value)) %>%
     ggplot(aes(x = pi0 / 100, y = mean_value, color = Method, lty = Method)) +
     geom_line(alpha = 1/6, position = position_dodge(width = .01)) +
+    ##geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), alpha = 1) +
     geom_point(alpha = 1, position = position_dodge(width = .01)) +
@@
 
-ptpr <- dplyr::select(res_pi0set, Method, setting, pi0, key, mean_value, col, lty) %>%
-    dplyr::filter(key == "TPR") %>%
+## not including SE because it's too small
+ptpr <- dplyr::select(res_pi0set, Method, setting, pi0, key, mean_value, se, col, lty) %>%
+    dplyr::filter(key == "TPR", !is.na(mean_value)) %>%
     ggplot(aes(x = pi0 / 100, y = mean_value, color = Method, lty = Method)) +
     geom_line(alpha = 1/6, position = position_dodge(width = .01)) +
+    ##geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), alpha = 1) +
     geom_point(alpha = 1, position = position_dodge(width = .01)) +
@@
 
-uniplot <- dplyr::select(res_unimodal, Method, dist, signal, key, mean_value, col, lty) %>%
+## We don't include the error bars because the SEs are so small
+uniplot <- dplyr::select(res_unimodal, Method, dist, signal, key, mean_value, se, col, lty) %>%
     dplyr::filter(key %in% c("FDR", "TPR")) %>%
-    tidyr::spread(key, mean_value) %>%
-    ggplot(aes(x = TPR, y = FDR, color = Method)) +
+    dplyr::mutate(plus1se = mean_value + se, minus1se = mean_value - se) %>%
+    dplyr::select(-se) %>%
+    tidyr::gather(metric, value, mean_value, plus1se, minus1se) %>%
+    tidyr::unite(metric, c("key", "metric")) %>%
+    tidyr::spread(metric, value) %>%
+    ggplot(aes(x = TPR_mean_value, y = FDR_mean_value, color = Method)) +
+    ##geom_errorbar(aes(ymin = FDR_minus1se, ymax = FDR_plus1se), width = 0) +
+    ##geom_errorbarh(aes(xmin = TPR_minus1se, xmax = TPR_plus1se), height = 0) + 
     geom_point(alpha = 1) +
@@
 
-pfdr <- dplyr::select(res_infoset, Method, setting, level, key, mean_value, col, lty) %>%
+## not including SE because it's too small
+pfdr <- dplyr::select(res_infoset, Method, setting, level, key, mean_value, se, q10, q90, col, lty) %>%
     dplyr::filter(key == "FDR") %>%
@@
     geom_line(alpha = 1/6, position = position_dodge(width = .01)) +
+    ##geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0) +
     geom_point(alpha = 1, position = position_dodge(width = .01)) +
@@
 
-ptpr <- dplyr::select(res_infoset, Method, setting, level, key, mean_value, col, lty) %>%
+ptpr <- dplyr::select(res_infoset, Method, setting, level, key, mean_value, se, col, lty) %>%
     dplyr::filter(key == "TPR") %>%
@@
     geom_line(alpha = 1/6, position = position_dodge(width = .01)) +
+    ##geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0) +    
     geom_point(alpha = 1, position = position_dodge(width = .01)) +

### 2018-06-22 manuscript_figures/manuscript_figures.Rmd
Add Supp figure to compare ranks in sim vs casestudies

@@
 xtable(tabS3)
+
+# generate detailed case study table for creating sim vs cs ranks boxplots
+casestudy.tab <- plotMethodRanks(objects, colLabels = dataset, alpha = alpha,
+                                 tableOnly = TRUE)
 ```
@@
 ggsave(file.path(outdir, "Figure4.pdf"), width=12, height=5)
+
+# generate detailed case study table for creating sim vs cs ranks boxplots
+yeastsim.tab <- plotMethodRanks(objs, colLabels = labs, alpha = alpha,
+                                 tableOnly = TRUE)
+```
+
+Create boxplot of ranking in sims versus rankings in case studies using the
+`casestudy.tab` and `yeastsim.tab` objects created above.
+
+```{r}
+casestudy.tab <- mutate(casestudy.tab, prop_casestudy = mean_prop) %>%
+  select(casestudy, method, prop_casestudy)
+yeastsim.tab <- yeastsim.tab %>% 
+  mutate(prop_yeastsim = mean_prop) %>% 
+  filter(casestudy == "de5") %>%
+  select(casestudy, method, prop_yeastsim)
+
+rank.tab <- left_join(casestudy.tab, yeastsim.tab, by = "method") %>%
+  left_join(candycols %>% dplyr::mutate(method=Method), by = "method")
+
+col <- as.character(rank.tab$col)
+names(col) <- as.character(rank.tab$method)
+
+rank.tab$method <- factor(rank.tab$method,
+                          levels = yeastsim.tab$method[order(yeastsim.tab$prop_yeastsim)])
+
+ggplot(rank.tab, aes(y = prop_casestudy, x = method, group = method, fill = method)) +
+  geom_boxplot() +
+  scale_fill_manual(values=col) +
+  theme_bw() +
+  xlab("Method (ordered by rejections in Yeast simulation (n=5))") + 
+  ylab("Proportion of maximum rejections in casestudies")
+ggsave(file.path(outdir, "Figure_simVcasestudyRanks.pdf"), width=6, height=4)
 ```

### 2018-06-22 manuscript_figures/manuscript_figures.Rmd
Add Supp figure to compare ranks in sim vs casestudies

@@
 
-# generate detailed case study table for creating sim vs cs ranks boxplots
+# generate detailed yeast sim table for creating sim vs cs ranks boxplots
 yeastsim.tab <- plotMethodRanks(objs, colLabels = labs, alpha = alpha,

### 2018-06-24 datasets/microbiome/microbiome-baxter.Rmd
Add method heatmaps to case studies with more than one covariate/comparison
Add mean abundance covariate

--- 2018-05-10 datasets/microbiome/microbiome-baxter.Rmd
@@
 
+## Multiple-Testing Correction - mean abundance
+
+Let's use `mean_abun_present` as our `ind_covariate`.
+
+```{r, baxter-genus-labelcovariate-abun}
+res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, baxter-genus-sb-abun, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(gns_bench_file_abun)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    metadata(sb)$data_download_link <-
+                   "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
+    saveRDS(sb, file = gns_bench_file_abun)
+} else {
+    sb <- readRDS(gns_bench_file_abun)
+}
+```
+
+# OTU/Genus comparison
+
+Here we compare the method ranks for the different comparisons at alpha = 0.10.
+
+```{r, eval = FALSE}
+plotMethodRanks(c(otu_bench_file, otu_bench_file_log_ubiquity, 
+                  gns_bench_file, gns_bench_file_log_ubiquity), 
+                colLabels = c("OTU", "OTU - log", "Genus", "Genus - log"), 
+                alpha = 0.10, xlab = "Comparison")
+```
+
 # Session Info

### 2018-06-24 datasets/microbiome/microbiome-enigma.Rmd
Add mean abundance covariate

--- 2018-05-10 datasets/microbiome/microbiome-enigma.Rmd
@@
 so4_bench_file <- file.path(sbdir, "enigma-so4-otus-benchmark.rds")
+so4_bench_file_abun <- file.path(sbdir, "enigma-so4-otus-abun-benchmark.rds")
 
@@
 ph_bench_file <- file.path(sbdir, "enigma-ph-otus-benchmark.rds")
+ph_bench_file_abun <- file.path(sbdir, "enigma-ph-otus-abun-benchmark.rds")
 
@@
 al_bench_file <- file.path(sbdir, "enigma-al-otus-benchmark.rds")
+al_bench_file_abun <- file.path(sbdir, "enigma-al-otus-abun-benchmark.rds")
 ```
@@
 
+### Multiple-Testing Correction - mean abundance
+
+Let's use `mean_abun_present` as our `ind_covariate`.
+
+```{r, ph-labelcovariate-abun}
+res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the default comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, ph-sb-abun, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(ph_bench_file_abun)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    saveRDS(sb, file = ph_bench_file_abun)
+} else {
+    sb <- readRDS(ph_bench_file_abun)
+}
+```
+
 ## SO4
@@
 
+### Multiple-Testing Correction - mean abundance
+
+Let's use `mean_abun_present` as our `ind_covariate`.
+
+```{r, so4-labelcovariate-abun}
+res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the default comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, so4-sb-abun, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(so4_bench_file_abun)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    saveRDS(sb, file = so4_bench_file_abun)
+} else {
+    sb <- readRDS(so4_bench_file_abun)
+}
+```
+
 ## Al
@@
 
+### Multiple-Testing Correction - mean abundance
+
+Let's use `mean_abun_present` as our `ind_covariate`.
+
+```{r, al-labelcovariate-abun}
+res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the default comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, al-sb-abun, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(al_bench_file_abun)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    saveRDS(sb, file = al_bench_file_abun)
+} else {
+    sb <- readRDS(al_bench_file_abun)
+}
+```
+
+# Response comparison
+
+Here we compare the method ranks for the different comparisons at alpha = 0.10.
+
+```{r, eval = FALSE}
+plotMethodRanks(c(so4_bench_file, ph_bench_file, al_bench_file), 
+                colLabels = c("SO4", "pH", "Al"), 
+                alpha = 0.10, xlab = "Comparison")
+```
+
 # Session Info

### 2018-06-24 datasets/microbiome/microbiome-goodrich.Rmd
Add method heatmaps to case studies with more than one covariate/comparison
Add mean abundance covariate

--- 2018-05-10 datasets/microbiome/microbiome-goodrich.Rmd
@@
 gns_bench_file <- file.path(sbdir, "goodrich-genus-benchmark.rds")
+gns_bench_file_abun <- file.path(sbdir, "goodrich-genus-abun-benchmark.rds")
 ```
@@
 
+## Multiple-Testing Correction - mean abun
+
+Let's use `mean_abun_present` as our `ind_covariate`.
+
+```{r, goodrich-genus-labelcovariate-abun}
+res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, goodrich-genus-sb-abun, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(gns_bench_file_abun)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    metadata(sb)$data_download_link <-
+                   "https://zenodo.org/record/840333/files/ob_goodrich_results.tar.gz"
+    saveRDS(sb, file = gns_bench_file_abun)
+} else {
+    sb <- readRDS(gns_bench_file_abun)
+}
+```
+
+# OTU/Genus comparison
+
+Here we compare the method ranks for the different comparisons at alpha = 0.10.
+
+```{r, eval = FALSE}
+plotMethodRanks(c(otu_bench_file, gns_bench_file), 
+                colLabels = c("OTU", "Genus"), 
+                alpha = 0.10, xlab = "Comparison")
+```
+
 # Session Info

### 2018-06-24 datasets/microbiome/microbiome-papa.Rmd
Add method heatmaps to case studies with more than one covariate/comparison
Add mean abundance covariate

@@
 gns_bench_file_log_ubi <- file.path(resdir, "papa-genus-log-ubiquity-benchmark.rds")
+gns_bench_file_abun <- file.path(resdir, "papa-genus-abun-benchmark.rds")
 ```
@@
 
+## Multiple-Testing Correction - mean abundance
+
+Let's use `mean_abun_present` as our `ind_covariate`.
+
+```{r, papa-genus-labelcovariate-abun}
+res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, papa-genus-sb-abun, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(gns_bench_file_abun)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    metadata(sb)$data_download_link <-
+                   "https://zenodo.org/record/840333/files/ibd_alm_results.tar.gz"
+    saveRDS(sb, file = gns_bench_file_abun)
+} else {
+    sb <- readRDS(gns_bench_file_abun)
+}
+```
+
+# OTU/Genus comparison
+
+Here we compare the method ranks for the different comparisons at alpha = 0.10.
+
+```{r, eval = FALSE}
+plotMethodRanks(c(otu_bench_file, gns_bench_file, gns_bench_file_log_ubi), 
+                colLabels = c("OTU", "Genus", "Genus - log ubiquity"), 
+                alpha = 0.10, xlab = "Comparison")
+```
+
 # Session Info

### 2018-06-24 datasets/microbiome/microbiome-schubert.Rmd
Add mean abundance covariate

@@
 otu_bench_file <- file.path(sbdir, "schubert-otus-benchmark.rds")
+otu_bench_file_abun <- file.path(sbdir, "schubert-otus-abun-benchmark.rds")
 ```
@@
 
+## Multiple-Testing Correction - mean abundance
+
+Let's use `mean_abun_present` as our `ind_covariate`.
+
+```{r, schubert-labelcovariate-abun, eval=TRUE}
+res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, schubert-sb-abun, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(otu_bench_file_abun)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    metadata(sb)$data_download_link <-
+                   "https://zenodo.org/record/840333/files/cdi_schubert_results.tar.gz"
+    saveRDS(sb, file = otu_bench_file_abun)
+} else {
+    sb <- readRDS(otu_bench_file_abun)
+}
+```
+
 # Session Info

### 2018-06-24 datasets/microbiome/microbiome-enigma.Rmd
Add results plots for mean abunandance covariate

@@
 ```
+### Benchmark Metrics
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, ph-metrics-abun, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, ph-plot-abun, width=15, height=15, eval=TRUE}
+rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
 
@@
 }
+```
+### Benchmark Metrics
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, so4-metrics-abun, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, so4-plot-abun, width=15, height=15, eval=TRUE}
+rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
 ```
@@
 ```
+### Benchmark Metrics
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, al-metrics-abun, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, al-plot-abun, width=15, height=15, eval=TRUE}
+rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
 
@@
 
-```{r, eval = FALSE}
-plotMethodRanks(c(so4_bench_file, ph_bench_file, al_bench_file), 
-                colLabels = c("SO4", "pH", "Al"), 
+```{r}
+plotMethodRanks(c(so4_bench_file, so4_bench_file_abun,
+                  ph_bench_file, ph_bench_file_abun,
+                  al_bench_file, al_bench_file_abun), 
+                colLabels = c("SO4", "SO4-abun", "pH", "ph-abun",
+                              "Al", "Al-abun"), 
                 alpha = 0.10, xlab = "Comparison")

### 2018-06-24 datasets/microbiome/microbiome-goodrich.Rmd
Add results plots for mean abunandance covariate

@@
 ```
+## Benchmark Metrics - mean abun
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, goodrich-genus-metrics-abun, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, goodrich-genus-plot-abun, width=15, height=15, eval=TRUE}
+rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
+Note: Benjamini-Hochberg (bh) overlaps exactly with the IHW results.
+
+```{r, goodrich-genus-overlap-abun}
+plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+```
+
+```{r, goodrich-genus-covariates-abun}
+methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
+plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods)
+```
 
@@
 
-```{r, eval = FALSE}
-plotMethodRanks(c(otu_bench_file, gns_bench_file), 
-                colLabels = c("OTU", "Genus"), 
+```{r}
+plotMethodRanks(c(otu_bench_file, gns_bench_file, gns_bench_file_abun), 
+                colLabels = c("OTU", "Genus", "Genus-abun"), 
                 alpha = 0.10, xlab = "Comparison")

### 2018-06-24 datasets/microbiome/microbiome-papa.Rmd
Add results plots for mean abunandance covariate

@@
 
+## Benchmark Metrics - mean abundance
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, papa-genus-abun-metrics, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, papa-genus-abun-plot, width=15, height=15, eval=TRUE}
+rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+
+```{r, papa-genus-abun-overlap}
+plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+```
+```{r}
+methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
+plotCovariateBoxplots(sb, alpha=0.1, nsets=6, methods=methods)
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
 # OTU/Genus comparison
@@
 
-```{r, eval = FALSE}
-plotMethodRanks(c(otu_bench_file, gns_bench_file, gns_bench_file_log_ubi), 
-                colLabels = c("OTU", "Genus", "Genus - log ubiquity"), 
+```{r}
+plotMethodRanks(c(otu_bench_file, gns_bench_file, 
+                  gns_bench_file_log_ubi,gns_bench_file_abun), 
+                colLabels = c("OTU", "Genus", "Genus-log", "Genus-abun"), 
                 alpha = 0.10, xlab = "Comparison")

### 2018-06-24 datasets/microbiome/microbiome-schubert.Rmd
Add results plots for mean abunandance covariate

@@
 
+## Benchmark Metrics (mean abun)
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, schubert-metrics-abun, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, schubert-plot-abun, width=15, height=15, eval=TRUE}
+rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
+# Ubiquity/Abundance comparison
+
+Here we compare the method ranks for the different comparisons at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(otu_bench_file, otu_bench_file_abun), 
+                colLabels = c("OTU","OTU-abun"), 
+                alpha = 0.10, xlab = "Comparison")
+```
+
 # Session Info

### 2018-06-26 manuscript_figures/manuscript_figures_sims.Rmd
Check sim plots w/ more pi0

@@
 
-methodset <- c("bonf", "bh", "ihw-a10", "qvalue", "bl-df03", "lfdr",
+methodset <- c("bonf", "bh", paste0("ihw-a", round(alpha * 100)),
+               "qvalue", "bl-df03", "lfdr",
                "scott-empirical", "scott-theoretical", "ashq")
@@
 
-## not including SE because it's too small
+## including SE to show variability in FDRR
 pfdr <- dplyr::select(res_pi0set, Method, setting, pi0, key, mean_value, se, col, lty) %>%
@@
     geom_line(alpha = 1/6, position = position_dodge(width = .01)) +
-    ##geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), alpha = 1) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), alpha = 1) +
     geom_point(alpha = 1, position = position_dodge(width = .01)) +
@@
 
-## not including SE because it's too small
+## including SE to show variability in FDRR
 ptpr <- dplyr::select(res_pi0set, Method, setting, pi0, key, mean_value, se, col, lty) %>%
@@
 ```{r, fig.width = 15, fig.height = 7.5}
-plot_grid(pfdr + guides(color = FALSE, linetype = FALSE),
-          ptpr, rel_widths = c(1, 1.25), labels = LETTERS[1:2])
-
-## trimmed ranges
-gp <- plot_grid(pfdr + coord_cartesian(ylim = c(0, .25)) + guides(color = FALSE, linetype = FALSE),
+gp <- plot_grid(pfdr + guides(color = FALSE, linetype = FALSE),
                 ptpr, rel_widths = c(1, 1.25), labels = LETTERS[1:2])
@@
 ggsave(file.path(outdir, "FigureSsim-nullprop.pdf"), gp, width = 15, height = 7.5)
+
+## trimmed ranges
+gp_zoom <- plot_grid(pfdr + coord_cartesian(ylim = c(0, .25)) + guides(color = FALSE, linetype = FALSE),
+                     ptpr, rel_widths = c(1, 1.25), labels = LETTERS[1:2])
+gp_zoom
+
+ggsave(file.path(outdir, "FigureSsim-nullprop-zoom.pdf"), gp_zoom, width = 15, height = 7.5)
 ```

### 2018-07-05 manuscript_figures/manuscript_figures.Rmd
Add points to casestudy vs sim ranks boxplots

@@
 ```{r}
+library(ggbeeswarm)
+
 casestudy.tab <- mutate(casestudy.tab, prop_casestudy = mean_prop) %>%
@@
                           levels = yeastsim.tab$method[order(yeastsim.tab$prop_yeastsim)])
+x <- match(rank.tab$casestudy.x, dataset)
+rank.tab$casestudy <- casestudy[x]
 
-ggplot(rank.tab, aes(y = prop_casestudy, x = method, group = method, fill = method)) +
-  geom_boxplot() +
-  scale_fill_manual(values=col) +
+ggplot(rank.tab , 
+       aes(y = prop_casestudy, x = method, group = method, color = Method)) +
+  geom_boxplot(aes(x = prop_yeastsim, group = as.factor(prop_yeastsim), y = prop_casestudy),
+                   alpha = 0.3, width = 0.04, outlier.shape = NA) +
+  geom_quasirandom(aes(x = prop_yeastsim, y = prop_casestudy, color = Method, shape = casestudy),
+                   alpha = 0.6, cex=3, width = 0.002,  groupOnX = TRUE,
+                   varwidth = TRUE) +
+  scale_color_manual(values=col) +
   theme_bw() +
-  xlab("Method (ordered by rejections in Yeast simulation (n=5))") + 
-  ylab("Proportion of maximum rejections in casestudies")
-ggsave(file.path(outdir, "Figure_simVcasestudyRanks.pdf"), width=6, height=4)
+  xlab("Proportion of maximum rejections in yeast simulation (n=5)") +  
+  ylab("Proportion of maximum rejections in casestudies") +
+  labs(shape = "Case Study") 
+ggsave(file.path(outdir, "Figure_simVcasestudyRanks_points.pdf"), width=6, height=4)
+
 ```

### 2018-07-05 datasets/GWAS/GWAS.Rmd
Add uninformative covariate to GWAS

@@
 resfile_AF <- file.path(sbdir, paste0("bmi-maf-benchmark.rds"))
+resfile_uninf <- file.path(sbdir, paste0("bmi-uninf-benchmark.rds"))
 
@@
 
+Now we'll add an random (uninformative covariate) to compare results with.
+
+```{r}
+bmi <- bmi %>% mutate(ind_covar_uninf = rnorm(nrow(bmi)))
+```
+
 # Data Analysis
@@
   xlab("Covariate: Minor allele frequency")
+
+ggplot(data=bmi, aes(x = ind_covar_AF, y = SE)) +
+  geom_hex(bins = 125)+
+  xlab("Covariate: Random") 
 ```
@@
 strat_hist(bmi, pvalue="pval", covariate="ind_covar_AF", maxy=2)
+```
+
+### Covariate two: Random
+
+Next we look at the random (uninformative) covariate.
+
+```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
+rank_scatter(bmi, pvalue="pval", covariate="ind_covar_uninf") +
+  ggtitle("Covariate 1: Minor allele frequency")
+```
+
+```{r, fig.width=10, fig.height=3.2, message=FALSE}
+strat_hist(bmi, pvalue="pval", covariate="ind_covar_uninf", maxy=2)
 ```
@@
 
+### Covariate three: Random
+
+Now, we'll repeat the multiple testing correction using the other covariate (random):
+
+```{r, GWAS1-sb3, results="hide", message=FALSE}
+duration <- NA
+if (!file.exists(resfile_uninf)) {
+    t1 <- proc.time()
+    sbU <- bd %>% buildBench(data=(bmi %>% mutate(ind_covariate=ind_covar_uninf) %>%
+                                    filter(!is.na(ind_covariate))), 
+                              ftCols = c("ind_covar_uninf", "effect_size"),
+                              parallel=TRUE, BPPARAM=multicoreParam)
+    metadata(sbU)$data_download_link <-    
+                     "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
+    saveRDS(sbU, file = resfile_uninf)
+    duration <- round((proc.time()-t1)[3]/60, 1)
+} else {
+    sbU <- readRDS(resfile_uninf)
+}
+```
+
+```{r, echo=FALSE}
+if (!is.na(duration)) {
+    message("This step took ", duration, " minutes for ", nrow(bmi), " SNPs using ",
+            cores, " cores.")
+}
+```
+
 ## Benchmark Metrics
@@
 
+### Covariate three: Random
+
+Next, we'll look at the performance metrics for the other covariate (random).
+
+```{r, GWAS1-metrics3}
+# rename assay to qvalue
+assayNames(sbU) <- "qvalue"
+sbU <- addDefaultMetrics(sbU)
+```
+
+Now, we'll plot the results.
+
+```{r, GWAS1-plot3, results="hide"}
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter(sbU, supplementary=FALSE) +
+  ggtitle("Covariate 3: Random")
+
+rejection_scatter_bins(sbU, covariate="ind_covar_uninf", bins=4,
+                       supplementary=FALSE) +
+  ggtitle("Covariate 3: Random")
+
+# upset plot 
+plotFDRMethodsOverlap(sbU, 
+                      alpha=0.05, nsets=ncol(sbAF),
+                      order.by="freq", decreasing=TRUE,
+                      supplementary=FALSE)
+```
+
+```{r, fig.width=8, fig.height=3.5}
+covariateLinePlot(sbU, alpha=0.05, covname="effect_size", nbins=25, 
+                 trans = "log1p")
+covariateLinePlot(sbU, alpha=0.05, covname="ind_covar_uninf", nbins=25)
+```
+
 # Covariate comparison
@@
 ```{r}
-plotMethodRanks(c(resfile_AF, resfile_N), 
-                colLabels = c("MAF", "Sample Size"), 
+plotMethodRanks(c(resfile_AF, resfile_N, resfile_uninf), 
+                colLabels = c("MAF", "Sample Size", "Random"), 
                 alpha = 0.10, xlab = "Covariate", 

### 2018-07-05 datasets/ChIP-seq/ChIPseq-CBP-csaw.Rmd
Add random (uninformative) covariate to ChIP-seq

@@
 bench_file <- file.path(sbdir, "cbp-csaw-benchmark.rds")
+bench_file_uninf <- file.path(sbdir, "cbp-csaw-uninf-benchmark.rds")
 
@@
 res_df <- as.data.frame(res_ranges)
+
+## add random covariate 
+set.seed(11245)
+res_df$uninf_covar = rnorm(nrow(res_df))
 ```
@@
 
-```{r, meancov-diag-hist, fig.width=10, fig.height=3.2}
+```{r, width-diag-hist, fig.width=10, fig.height=3.2}
 strat_hist(res_df, pvalue = "PValue", covariate = "width", maxy = 35, numQ = 3)
+```
+
+### Random
+
+```{r, rand-diag-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(res_df, pvalue = "PValue", covariate = "uninf_covar") +
+    ggtitle("Random independent covariate") +
+    xlab("Region Width")
+```
+
+```{r, rand-diag-hist, fig.width=10, fig.height=3.2}
+strat_hist(res_df, pvalue = "PValue", covariate = "uninf_covar", maxy = 35, numQ = 3)
 ```
@@
 
+We'll also compare to the random covariate.
+
+```{r}
+if (!file.exists(bench_file_uninf)) {
+    res_df$ind_covariate <- res_df$uninf_covar
+    res_df$pval <- res_df$PValue
+
+    bd <- initializeBenchDesign()
+    sbU <- buildBench(bd, data = res_df, ftCols = "ind_covariate")
+    saveRDS(sbU, file = bench_file_uninf)
+} else {
+    sbU <- readRDS(bench_file_uninf)
+}
+```
+
 ## Benchmark Metrics
+
+### Region width
 
@@
 
+### Random
+
+We'll do the same for the random (uninformative covariate).
+
+```{r, chipseq-metrics-rand}
+assayNames(sbU) <- "qvalue"
+sbU <- addDefaultMetrics(sbU)
+```
+
+Now, we'll plot the results.
+
+```{r, chipseq-scatter-rand, results="hide"}
+rejections_scatter(sbU, supplementary = FALSE)
+rejection_scatter_bins(sbU, covariate = "ind_covariate",
+                       bins = 4, supplementary = FALSE)
+```
+
+```{r, chipseq-overlap-rand}
+plotFDRMethodsOverlap(sbU, alpha = 0.05, nsets = ncol(sb),
+                      order.by = "freq", decreasing = TRUE,
+                      supplementary = FALSE)
+```
+
+```{r, rnaseq-lineplot-rand}
+covariateLinePlot(sbU, alpha = 0.05, covname = "ind_covariate")
+```
+
+## Covariate comparison
+
+Here we compare the method ranks for the two covariates at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(benchfile, benchfile_uninf), 
+                colLabels = c("Region width", "Random"), 
+                alpha = 0.10, xlab = "Covariate", 
+                excludeMethods = NULL)
+```
+
 # Session Info

### 2018-07-05 datasets/ChIP-seq/ChIPseq-h3k4me3-csaw.Rmd
Add random (uninformative) covariate to ChIP-seq

@@
 bench_file <- file.path(sbdir, "h3k4me3-csaw-benchmark.rds")
+bench_file_uninf <- file.path(sbdir, "h3k4me3-csaw-uninf-benchmark.rds")
 
@@
 res_df <- as.data.frame(res_ranges)
+
+## add random covariate 
+set.seed(7778)
+res_df$uninf_covar = rnorm(nrow(res_df))
 ```
@@
 
-```{r, meancov-diag-hist, fig.width=10, fig.height=3.2}
+```{r, width-diag-hist, fig.width=10, fig.height=3.2}
 strat_hist(res_df, pvalue = "PValue", covariate = "width", maxy = 35, numQ = 3)
+```
+
+### Random
+
+```{r, rand-diag-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(res_df, pvalue = "PValue", covariate = "uninf_covar") +
+    ggtitle("Region width as independent covariate") +
+    xlab("Region Width")
+```
+
+```{r, rand-diag-hist, fig.width=10, fig.height=3.2}
+strat_hist(res_df, pvalue = "PValue", covariate = "uninf_covar", maxy = 35, numQ = 3)
 ```
@@
 
+### Random
+
+We'll do the same for the random (uninformative covariate).
+
+```{r, chipseq-metrics-rand}
+assayNames(sbU) <- "qvalue"
+sbU <- addDefaultMetrics(sbU)
+```
+
+Now, we'll plot the results.
+
+```{r, chipseq-scatter-rand, results="hide"}
+rejections_scatter(sbU, supplementary = FALSE)
+rejection_scatter_bins(sbU, covariate = "ind_covariate",
+                       bins = 4, supplementary = FALSE)
+```
+
+```{r, chipseq-overlap-rand}
+plotFDRMethodsOverlap(sbU, alpha = 0.05, nsets = ncol(sb),
+                      order.by = "freq", decreasing = TRUE,
+                      supplementary = FALSE)
+```
+
+```{r, rnaseq-lineplots-rand}
+covariateLinePlot(sbU, alpha = 0.05, covname = "ind_covariate")
+```
+
+## Covariate comparison
+
+Here we compare the method ranks for the two covariates at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(benchfile, benchfile_uninf), 
+                colLabels = c("Region width", "Random"), 
+                alpha = 0.10, xlab = "Covariate", 
+                excludeMethods = NULL)
+```
+
 # Session Info

### 2018-07-06 datasets/ChIP-seq/ChIPseq-CBP-csaw.Rmd
Fix typo in filename

@@
 ```{r}
-plotMethodRanks(c(benchfile, benchfile_uninf), 
+plotMethodRanks(c(bench_file, bench_file_uninf), 
                 colLabels = c("Region width", "Random"), 

### 2018-07-06 datasets/ChIP-seq/ChIPseq-h3k4me3-csaw.Rmd
Update covariate plot title for random covariate

@@
 rank_scatter(res_df, pvalue = "PValue", covariate = "uninf_covar") +
-    ggtitle("Region width as independent covariate") +
+    ggtitle("Random independent covariate") +
     xlab("Region Width")

### 2018-07-06 datasets/GSEA/gsea-human.Rmd
Add random uninformative covariates to GSEA

@@
 bench_file <- file.path(sbdir, "human-benchmark.rds")
+bench_file_uninf <- file.path(sbdir, "human-uninf-benchmark.rds")
 ```
@@
 }
+
+## Add random (uninformative) covariate
+set.seed(66778)
+goRes$rand_covar <- rnorm(nrow(goRes))
+
 ```
@@
 
+### Random
+
+We will explore whether the random covariate can be used as a covariate for modern multiple-testing correction methods in the context of GSEA.
+In the plot below, the log10 of the p-values is plotted as a function of the
+random covariate. This covariate looks independent of the p-values.
+
+```{r informativeCovariate-rand, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
+rank_scatter(dat = goRes, pval = "over_represented_pvalue", 
+              covariate = "rand_covar", bins = 50, 
+              funfill = log10_trans()) +
+    ylim(0, 12) +
+    ggtitle("Over-represented gene sets") +
+    ylab(expression(-log[10]~"(p-value)") ) 
+```
+
+We can also explore if the covariate seems to be independent under the null.
+
+```{r assumptions-rand, message=FALSE, fig.width=10, fig.height=3.2}
+## filter numInCat below only used for plotting
+## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
+#dplyr::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)
+dplyr::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
+    strat_hist( pval="over_represented_pvalue", covariate="rand_covar", maxy=10)
+```
+
 ## Multiple-Testing Correction
@@
 ## rename columns and prepare for benchmarking
-res <- dplyr:::select(goRes, c("over_represented_pvalue", "numInCat")) %>%
+res <- dplyr:::select(goRes, c("over_represented_pvalue", "numInCat", "rand_covar")) %>%
     dplyr:::rename(pval = over_represented_pvalue, ind_covariate = numInCat)
@@
 
+We'll also compare the results to an uninformative (random) covariate.
+
+```{r run-benchmark-rand}
+if (!file.exists(bench_file_uninf)) {
+  res$ind_covariate <- res$rand_covar
+  sGSEA_rand <- buildBench(bd, res, ftCols = "ind_covariate")
+  saveRDS(sGSEA_rand, file = bench_file_uninf)
+} else {
+  sGSEA_rand <- readRDS(bench_file_uninf)
+}
+```
+
 ## Benchmark Metrics
+
+### Gene Set Size
 
@@
 
+### Random
+
+```{r}
+assayNames(sGSEA_rand) <- "qvalue"
+sGSEA_rand <- addDefaultMetrics(sGSEA_rand)
+sGSEA_rand <- estimatePerformanceMetrics(sGSEA_rand, addColData=TRUE)
+```
+
+```{r plottingResults-rand}
+rejections_scatter(sGSEA_rand, as_fraction=FALSE, supplementary=FALSE)
+```
+
+```{r}
+plotFDRMethodsOverlap(sGSEA_rand, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100)
+```
+
+```{r}
+covariateLinePlot(sGSEA_rand, alpha = 0.1, covname = "ind_covariate", trans = "log1p")
+```
+
+## Covariate comparison
+
+Here we compare the method ranks for the two covariates at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(bench_file, bench_file_uninf), 
+                colLabels = c("Set Size", "Random"), 
+                alpha = 0.10, xlab = "Covariate", 
+                excludeMethods = NULL)
+```
+
 # Session Info

### 2018-07-06 datasets/GSEA/gsea-mouse.Rmd
Add random uninformative covariates to GSEA

@@
 bench_file <- file.path(sbdir, "mouse-benchmark.rds")
+bench_file_uninf <- file.path(sbdir, "mouse-uninf-benchmark.rds")
 ```
@@
 }
+
+## Add random (uninformative) covariate
+set.seed(7476)
+goRes$rand_covar <- rnorm(nrow(goRes))
+
 ```
@@
 ## rename columns and prepare for benchmarking
-res <- dplyr::select(goRes, over_represented_pvalue, numInCat) %>%
+res <- dplyr::select(goRes, over_represented_pvalue, numInCat, rand_covar) %>%
     dplyr::rename(pval = over_represented_pvalue,
@@
 
+We'll also compare the results to an uninformative (random) covariate.
+
+```{r run-benchmark-rand}
+if (!file.exists(bench_file_uninf)) {
+  res$ind_covariate <- res$rand_covar
+  sGSEAMouse_rand <- buildBench(bd, res, ftCols = "ind_covariate")
+  saveRDS(sGSEAMouse_rand, file = bench_file_uninf)
+} else {
+  sGSEAMouse_rand <- readRDS(bench_file_uninf)
+}
+```
+
 ## Benchmark Metrics
+
+### Gene Set Size
 
@@
 
+### Random
+
+```{r}
+assayNames(sGSEAMouse_rand) <- "qvalue"
+sGSEAMouse_rand <- addDefaultMetrics(sGSEAMouse_rand)
+sGSEAMouse_rand <- estimatePerformanceMetrics(sGSEAMouse_rand, addColData=TRUE)
+```
+
+```{r plottingResults-rand}
+rejections_scatter(sGSEAMouse_rand, as_fraction=FALSE, supplementary=FALSE)
+```
+
+```{r}
+plotFDRMethodsOverlap(sGSEAMouse_rand, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100)
+```
+
+```{r}
+covariateLinePlot(sGSEAMouse_rand, alpha = 0.1, covname = "ind_covariate", trans = "log1p")
+```
+
+## Covariate comparison
+
+Here we compare the method ranks for the two covariates at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(bench_file, bench_file_uninf), 
+                colLabels = c("Set Size", "Random"), 
+                alpha = 0.10, xlab = "Covariate", 
+                excludeMethods = NULL)
+```
+
 # Session Info

### 2018-07-06 datasets/R/plotMethodRank.R
Declare namespace of select to sidestep package collisions

@@ -262,7 +262,7 @@ tidy_df <- function(objects, colLabels, fill, annotate, alpha){
                 dplyr::filter( is.na(param.smooth.df) | (param.smooth.df == "3L")) %>%
                 dplyr::filter( !method == "unadjusted") %>%
                 dplyr::filter( !(method %in% NAmethods)) %>%
-                select( method, value ) %>%
+                dplyr::select( method, value ) %>%
                 dplyr::rename( nrejects = value) %>%
                 mutate( method = gsub("-df03", "", method)) %>%
                 mutate( method = gsub("(-a)(.*)", "", method)) %>%

### 2018-07-06 datasets/microbiome/microbiome-enigma.Rmd
Add uninformative covariate analyses to RNAseq

@@
 so4_bench_file_abun <- file.path(sbdir, "enigma-so4-otus-abun-benchmark.rds")
+so4_bench_file_uninf <- file.path(sbdir, "enigma-so4-otus-uninf-benchmark.rds")
 
@@
 ph_bench_file_abun <- file.path(sbdir, "enigma-ph-otus-abun-benchmark.rds")
+ph_bench_file_uninf <- file.path(sbdir, "enigma-ph-otus-uninf-benchmark.rds")
 
@@
 al_bench_file_abun <- file.path(sbdir, "enigma-al-otus-abun-benchmark.rds")
+al_bench_file_uninf <- file.path(sbdir, "enigma-al-otus-uninf-benchmark.rds")
 ```
@@
 
+Add random (uninformative) covariate.
+
+```{r}
+set.seed(44318)
+res$rand_covar <- rnorm(nrow(res))
+```
+
 ### Covariate Diagnostics
@@
 
-### Multiple-Testing Correction
+#### Random
+
+```{r, ph-rand-hist, fig.width=10, fig.height=3.2}
+strat_hist(res, pvalue="pval", covariate="rand_covar", maxy=8, binwidth=0.05)
+```
+
+```{r, ph-rand-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(res, pvalue="pval", covariate="rand_covar")
+```
+
+### Multiple-Testing Correction - ubiquity
 
@@
 
-### Benchmark Metrics
+### Benchmark Metrics - ubiquity
 
@@
 ```
-### Benchmark Metrics
+### Benchmark Metrics - mean abundance
 
@@
 
+### Multiple-Testing Correction - random
+
+Let's use `rand_covar` as our `ind_covariate`.
+
+```{r, ph-labelcovariate-rand}
+res <- dplyr::mutate(res, ind_covariate = rand_covar)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the default comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, ph-sb-rand, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(ph_bench_file_uninf)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    saveRDS(sb, file = ph_bench_file_uninf)
+} else {
+    sb <- readRDS(ph_bench_file_uninf)
+}
+```
+### Benchmark Metrics - random
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, ph-metrics-rand, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, ph-plot-rand, width=15, height=15, eval=TRUE}
+rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
 ## SO4
@@
 }
+```
+
+Add random (uninformative) covariate.
+
+```{r}
+set.seed(27349)
+res$rand_covar <- rnorm(nrow(res))
 ```
@@
 
-### Multiple-Testing Correction
+#### Random
+
+```{r, so4-rand-hist, fig.width=10, fig.height=3.2}
+strat_hist(res, pvalue="pval", covariate="rand_covar", maxy=8, binwidth=0.05)
+```
+
+```{r, so4-rand-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(res, pvalue="pval", covariate="rand_covar")
+```
+
+### Multiple-Testing Correction - ubiquity
 
@@
 
+### Multiple-Testing Correction - random
+
+Let's use `rand_covar` as our `ind_covariate`.
+
+```{r, so4-labelcovariate-rand}
+res <- dplyr::mutate(res, ind_covariate = mean_abun_present)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the default comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, so4-sb-rand, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(so4_bench_file_uninf)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    saveRDS(sb, file = so4_bench_file_uninf)
+} else {
+    sb <- readRDS(so4_bench_file_uninf)
+}
+```
+### Benchmark Metrics - random
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, so4-metrics-rand, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, so4-plot-rand, width=15, height=15, eval=TRUE}
+rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
 ## Al
@@
 }
+```
+
+Add random (uninformative) covariate.
+
+```{r}
+set.seed(19474)
+res$rand_covar <- rnorm(nrow(res))
 ```
@@
 
-### Multiple-Testing Correction
+#### Random
+
+```{r, al-rand-hist, fig.width=10, fig.height=3.2}
+strat_hist(res, pvalue="pval", covariate="rand_covar", maxy=8, binwidth=0.05)
+```
+
+```{r, al-rand-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(res, pvalue="pval", covariate="rand_covar")
+```
+
+### Multiple-Testing Correction - ubiquity
 
@@
 
-# Response comparison
+### Multiple-Testing Correction - random
+
+Let's use `rand_covar` as our `ind_covariate`.
+
+```{r, al-labelcovariate-rand}
+res <- dplyr::mutate(res, ind_covariate = rand_covar)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the default comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, al-sb-rand, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(al_bench_file_uninf)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    saveRDS(sb, file = al_bench_file_uninf)
+} else {
+    sb <- readRDS(al_bench_file_uninf)
+}
+```
+### Benchmark Metrics - random
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, al-metrics-rand, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, al-plot-rand, width=15, height=15, eval=TRUE}
+rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
+# Response/Covariate comparison
 
@@
 
-```{r}
-plotMethodRanks(c(so4_bench_file, so4_bench_file_abun,
-                  ph_bench_file, ph_bench_file_abun,
-                  al_bench_file, al_bench_file_abun), 
-                colLabels = c("SO4", "SO4-abun", "pH", "ph-abun",
-                              "Al", "Al-abun"), 
+```{r, width = 15, height = 10}
+plotMethodRanks(c(so4_bench_file, so4_bench_file_abun, so4_bench_file_uninf,
+                  ph_bench_file, ph_bench_file_abun, ph_bench_file_uninf,
+                  al_bench_file, al_bench_file_abun, al_bench_file_uninf), 
+                colLabels = c("SO4-ubiquity", "SO4-abun", "SO4-uninf",
+                              "pH-ubiquity", "ph-abun", "pH-uninf",
+                              "Al-ubiquity", "Al-abun", "Al-uninf"),
                 alpha = 0.10, xlab = "Comparison")

### 2018-07-06 datasets/microbiome/microbiome-papa.Rmd
Add uninformative covariate analyses to RNAseq

@@
 gns_bench_file <- file.path(sbdir, "papa-genus-benchmark.rds")
-gns_bench_file_log_ubi <- file.path(resdir, "papa-genus-log-ubiquity-benchmark.rds")
-gns_bench_file_abun <- file.path(resdir, "papa-genus-abun-benchmark.rds")
+gns_bench_file_log_ubi <- file.path(sbdir, "papa-genus-log-ubiquity-benchmark.rds")
+gns_bench_file_abun <- file.path(sbdir, "papa-genus-abun-benchmark.rds")
+gns_bench_file_uninf <- file.path(sbdir, "papa-genus-uninf-benchmark.rds")
 ```
@@
 
+Add random (uninformative) covariate to test at the genus level.
+
+```{r}
+set.seed(33128)
+res$rand_covar <- rnorm(nrow(res))
+```
+
 ## Covariate Diagnostics
@@
 
-## Multiple-Testing Correction
+### Random
+
+```{r, papa-genus-rand, fig.width=10, fig.height=3.2}
+strat_hist(res, pvalue="pval", covariate="rand_covar", maxy=8, binwidth=0.05)
+```
+
+```{r, papa-genus-rand-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(res, pvalue="pval", covariate="rand_covar")
+```
+
+## Multiple-Testing Correction - ubiquity
 
@@
 
+## Multiple-Testing Correction - random
+
+Let's use the random (uninformative) covariate as our `ind_covariate`.
+
+```{r, papa-genus-rand-labelcovariate}
+res <- dplyr::mutate(res, ind_covariate = rand_covar)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts). 
+```{r, papa-genus-sb-rand, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(gns_bench_file_uninf)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb_rand <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    metadata(sb_rand)$data_download_link <-
+                   "https://zenodo.org/record/840333/files/ibd_alm_results.tar.gz"
+    saveRDS(sb_rand, file = gns_bench_file_uninf)
+} else {
+    sb_rand <- readRDS(gns_bench_file_uninf)
+}
+```
+
+## Benchmark Metrics - random
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, papa-genus-rand-metrics, eval=TRUE}
+assayNames(sb_rand) <- "qvalue"
+sb_rand <- addDefaultMetrics(sb_rand)
+```
+
+Now, we'll plot the results.
+
+```{r, papa-genus-rand-plot, width=15, height=15, eval=TRUE}
+rejections_scatter(sb_rand, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb_rand, covariate="ind_covariate", supplementary=FALSE)
+```
+
+```{r, papa-genus-rand-overlap}
+plotFDRMethodsOverlap(sb_rand, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+```
+```{r}
+methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
+plotCovariateBoxplots(sb_rand, alpha=0.1, nsets=6, methods=methods)
+```
+```{r}
+covariateLinePlot(sb_rand, alpha = 0.05, covname = "ind_covariate")
+```
+
 # OTU/Genus comparison
@@
 plotMethodRanks(c(otu_bench_file, gns_bench_file, 
-                  gns_bench_file_log_ubi,gns_bench_file_abun), 
-                colLabels = c("OTU", "Genus", "Genus-log", "Genus-abun"), 
+                  gns_bench_file_log_ubi,gns_bench_file_abun,
+                  gns_bench_file_uninf), 
+                colLabels = c("OTU-ubiquity", "Genus-ubiquity", "Genus-log", "Genus-abun", "Genus-uninf"), 
                 alpha = 0.10, xlab = "Comparison")

### 2018-07-06 datasets/microbiome/microbiome-schubert.Rmd
Add uninformative covariate analyses to RNAseq

@@
 otu_bench_file_abun <- file.path(sbdir, "schubert-otus-abun-benchmark.rds")
+otu_bench_file_uninf <- file.path(sbdir, "schubert-otus-uninf-benchmark.rds")
 ```
@@
 
+## Multiple-Testing Correction - random
+
+Let's use `rand_covar` as our `ind_covariate`.
+
+```{r, schubert-labelcovariate-rand, eval=TRUE}
+res <- dplyr::mutate(res, ind_covariate = rand_covar)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, schubert-sb-rand, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(otu_bench_file_uninf)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    metadata(sb)$data_download_link <-
+                   "https://zenodo.org/record/840333/files/cdi_schubert_results.tar.gz"
+    saveRDS(sb, file = otu_bench_file_uninf)
+} else {
+    sb <- readRDS(otu_bench_file_uninf)
+}
+```
+
+## Benchmark Metrics - random
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, schubert-metrics-rand, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, schubert-plot-rand, width=15, height=15, eval=TRUE}
+rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
 # Ubiquity/Abundance comparison
@@
 ```{r}
-plotMethodRanks(c(otu_bench_file, otu_bench_file_abun), 
-                colLabels = c("OTU","OTU-abun"), 
+plotMethodRanks(c(otu_bench_file, otu_bench_file_abun, otu_bench_file_uninf), 
+                colLabels = c("OTU-ubiquity","OTU-abun", "OTU-uninf"), 
                 alpha = 0.10, xlab = "Comparison")

### 2018-07-06 datasets/RNA-Seq/RNAseq-gtex.Rmd
Add uninformative covariate analyses to RNAseq

@@
 bench_file <- file.path(sbdir, "brain-benchmark.rds")
+bench_file_uninf <- file.path(sbdir, "brain-uninf-benchmark.rds")
 
@@
 
+Add random (uninformative) covariate.
+
+```{r}
+set.seed(83750)
+res$rand_covar <- rnorm(nrow(res))
+```
+
 ## Covariate Diagnostics
@@
 
-## Multiple-Testing Correction
+### Random
+
+```{r, rand-diag-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(res, pvalue = "pval", covariate = "rand_covar") +
+    ggtitle("Random independent covariate")
+```
+
+```{r, rand-diag-hist, fig.width=10, fig.height=3.2}
+strat_hist(res, pvalue = "pval",
+           covariate = "rand_covar", maxy = 25)
+```
+
+## Multiple-Testing Correction - mean
 
@@
 
-## Benchmark Metrics
+## Benchmark Metrics - mean
 
@@
 
+## Multiple-Testing Correction - random
+
+We use the common `BenchDesign` with the set of multiple testing correction
+methods already included. We also add in Scott's FDR Regression (both
+`nulltype = "empirical"` and `nulltype = "theoretical"`)
+since our test statistics are approximately t-distributed.
+
+```{r}
+if (file.exists(bench_file_uninf)) {
+    sb <- readRDS(bench_file_uninf)
+} else {
+    bd <- initializeBenchDesign()
+
+    bd <- addBMethod(bd, "scott-theoretical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'theoretical',
+                     control = list(lambda = 0.01))
+
+   bd <- addBMethod(bd, "scott-empirical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'empirical',
+                     control = list(lambda = 0.01))
+
+    res <- res %>% dplyr::mutate(ind_covariate = rand_covar)
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    saveRDS(sb, file = bench_file_uninf)
+}
+```
+
+## Benchmark Metrics - random
+
+```{r}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+estimatePerformanceMetrics(sb, alpha = 0.05, tidy = TRUE) %>%
+    filter(performanceMetric == "rejections") %>%
+    select(blabel, performanceMetric, alpha, value) %>%
+    mutate(n = nrow(sb), prop = round(value / n, 3)) %>%
+    arrange(desc(value)) %>%
+    as_tibble() %>%
+    print(n = 40)
+```
+
+```{r, rnaseq-scatter-rand}
+rejections_scatter(sb, supplementary = FALSE)
+rejection_scatter_bins(sb, covariate = "ind_covariate",
+                       bins = 4, supplementary = FALSE)
+```
+
+```{r, rnaseq-overlap-rand}
+plotFDRMethodsOverlap(sb, alpha = 0.05, nsets = ncol(sb),
+                      order.by = "freq", decreasing = TRUE,
+                      supplementary = FALSE)
+```
+
+```{r, rnaseq-lineplots-rand}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
+# Covariate comparison
+
+Here we compare the method ranks for the different covariates at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(bench_file, bench_file_uninf), 
+                colLabels = c("mean", "uninf"), 
+                alpha = 0.10, xlab = "Comparison")
+```
+
 # Session Info

### 2018-07-06 datasets/RNA-Seq/RNAseq-mir200c.Rmd
Add uninformative covariate analyses to RNAseq

@@
 bench_file <- file.path(sbdir, "mir200c-benchmark.rds")
+bench_file_uninf <- file.path(sbdir, "mir200c-uninf-benchmark.rds")
 
@@
 
-## Multiple-Testing Correction
+### Random
+
+```{r, rand-diag-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(res, pvalue = "pval", covariate = "rand_covar") +
+    ggtitle("Random independent covariate")
+```
+
+```{r, rand-diag-hist, fig.width=10, fig.height=3.2}
+strat_hist(res, pvalue = "pval",
+           covariate = "rand_covar", maxy = 7.5)
+```
+
+## Multiple-Testing Correction - mean
 
@@
 
+## Multiple-Testing Correction - random
+
+We use the common `BenchDesign` with the set of multiple testing correction 
+methods already included. We also add in Scott's FDR Regression (both
+`nulltype = "empirical"` and `nulltype = "theoretical"`)
+since our test statistics are t-distributed.
+
+```{r}
+if (file.exists(bench_file_uninf)) {
+    sb <- readRDS(bench_file_uninf)
+} else {
+    bd <- initializeBenchDesign()
+
+    bd <- addBMethod(bd, "scott-theoretical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'theoretical',
+                     control = list(lambda = 0.01))
+
+   bd <- addBMethod(bd, "scott-empirical",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'empirical',
+                     control = list(lambda = 0.01))
+
+    res <- res %>% dplyr::mutate(ind_covariate = rand_covar)
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    saveRDS(sb, file = bench_file_uninf)
+}
+```
+
+```{r}
+head(assays(sb)[["bench"]])
+```
+
+## Benchmark Metrics - random
+
+```{r}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+estimatePerformanceMetrics(sb, alpha = 0.05, tidy = TRUE) %>%
+    filter(performanceMetric == "rejections") %>%
+    select(blabel, performanceMetric, alpha, value) %>%
+    mutate(n = nrow(sb), prop = round(value / n, 3)) %>%
+    arrange(desc(value)) %>%
+    as_tibble() %>%
+    print(n = 40)
+```
+
+```{r}
+rejections_scatter(sb, supplementary = FALSE)
+rejection_scatter_bins(sb, covariate = "ind_covariate",
+                       bins = 4, supplementary = FALSE)
+```
+
+```{r}
+plotFDRMethodsOverlap(sb, alpha = 0.05, nsets = ncol(sb),
+                      order.by = "freq", decreasing = TRUE,
+                      supplementary = FALSE)
+```
+
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
+# Covariate comparison
+
+Here we compare the method ranks for the different covariates at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(bench_file, bench_file_uninf), 
+                colLabels = c("mean", "uninf"), 
+                alpha = 0.10, xlab = "Comparison")
+```
+
 # Session Info

### 2018-07-06 datasets/scrnaseq/scrnaseq-human-mast.Rmd
Add method heatmaps to case studies with more than one covariate/comparison
Add uninformative covariate analyses to scrnaseq

@@
 resfile_det <- file.path(sbdir, "human-benchmark-mast-det.rds")
+resfile_uninf <- file.path(sbdir, "human-benchmark-mast-uninf.rds")
 
@@
 and detection rate (defined as the proportion of cells expressing the gene at a nonzero level). 
+In addition, we'll add a random (uninformative covariate).
 These will also be added to the `rowData` slot of the `SummarizedExperiments`.
@@
   rowData(dat)$detection <- rowMeans(assay(dat, "count") > 0)
+  rowData(dat)$rand_covar <- rnorm(nrow(dat))
   return(dat)
@@
 
+set.seed(2878)
 humandata <- add_covariates_scrnaseq(humandata)
@@
 
+### Covariate three: Random
+
+Next we look at the random (uninformative) covariate.
+
+```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
+rank_scatter(data.frame(rowData(humandata)), 
+             pvalue="mast_p", covariate="rand_covar") + 
+  ggtitle("MAST, Covariate 3: Random")
+```
+
+```{r, fig.width=10, fig.height=3.2, message=FALSE}
+strat_hist(data.frame(rowData(humandata)), 
+           pvalue="mast_p", covariate="rand_covar", maxy=16,
+           main = "MAST, Covariate 3: Random") 
+```
+
 ## Multiple-Testing Correction
@@
 
-Now, we'll repeat the multiple testing correction using the other covariate (detection rate):
+Now, we'll repeat the multiple testing correction using the covariate detection rate:
 
@@
   sb2 <- readRDS(resfile_det)
+}
+```
+
+### Covariate two: Detection Rate
+
+Now, we'll repeat the multiple testing correction using the random covariate.
+
+```{r, sb3, results="hide", message=FALSE}
+if (!file.exists(resfile_uninf)){
+  sb3 <- bd %>% buildBench(data.frame(rowData(humandata)) %>% 
+                                   na.omit() %>%
+                                   mutate(pval=mast_p,
+                                          ind_covariate=rand_covar), 
+                          ftCols = c("rand_covar"),
+                          parallel=TRUE)
+
+  saveRDS(sb3, file = resfile_uninf)
+}else{
+  sb3 <- readRDS(resfile_uninf)
 }
@@
 
-Next, we'll look at the performance metrics for the other covariate (detection rate).
+Next, we'll look at the performance metrics for the covariate detection rate.
 
@@
 
+### Covariate three: Random
+
+Next, we'll look at the performance metrics for the random covariate.
+
+```{r, metrics3}
+# rename assay to qvalue
+assayNames(sb3) <- "qvalue"
+sb3 <- addDefaultMetrics(sb3)
+```
+
+Now, we'll plot the results.
+
+```{r, plot3, results="hide"}
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter(sb3, supplementary=FALSE) +
+  ggtitle("MAST, Covariate 3: Random")
+
+rejection_scatter_bins(sb3, covariate="rand_covar", bins=4,
+                       supplementary=FALSE) +
+  ggtitle("MAST, Covariate 3: Random")
+
+# upset plot 
+plotFDRMethodsOverlap(sb3, 
+                      alpha=0.05, nsets=ncol(sb3),
+                      order.by="freq", decreasing=TRUE,
+                      supplementary=FALSE)
+
+```
+
+```{r, fig.width=8, fig.height=3.5}
+mcols(sb3)$ind_covariate <- mcols(sb3)$rand_covar
+covariateLinePlot(sb3, alpha=0.05, covname="ind_covariate", nbins=25) +
+  ggtitle("MAST, Covariate 3: Random")
+```
+
+# Covariate comparison
+
+Here we compare the method ranks for the two covariates at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(resfile_mean, resfile_det, resfile_uninf), 
+                colLabels = c("Mean nonzero", "Detection rate", "Uninf"), 
+                alpha = 0.10, xlab = "Covariate")
+```
+
 # Session Information

### 2018-07-06 datasets/scrnaseq/scrnaseq-human-scdd.Rmd
Add method heatmaps to case studies with more than one covariate/comparison
Add uninformative covariate analyses to scrnaseq

@@
 resfile_det <- file.path(sbdir, "human-benchmark-scdd-det.rds")
+resfile_uninf <- file.path(sbdir, "human-benchmark-scdd-uninf.rds")
 
@@
 
+set.seed(8723)
 humandata <- add_covariates_scrnaseq(humandata)
@@
 ```
+
+### Covariate three: Random
+
+Next we look at the random covariate.
+
+```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
+rank_scatter(data.frame(rowData(humandata)), 
+             pvalue="scdd_p", covariate="rand_covar") + 
+  ggtitle("scDD, Covariate 3: Random")
+```
+
+```{r, fig.width=10, fig.height=3.2, message=FALSE}
+strat_hist(data.frame(rowData(humandata)), 
+           pvalue="scdd_p", covariate="rand_covar", maxy=9,
+           main = "scDD, Covariate 3: Random") 
+```
+
 ## Multiple-Testing Correction
@@
   sb2 <- readRDS(resfile_det)
+}
+```
+
+### Covariate three: Random
+
+Now, we'll repeat the multiple testing correction using the random covariate:
+
+```{r, sb3, results="hide", message=FALSE}
+if (!file.exists(resfile_uninf)){
+  sb3 <- bd %>% buildBench(data.frame(rowData(humandata)) %>% 
+                                   filter(!is.na(scdd_p)) %>%
+                                   mutate(pval=scdd_p,
+                                          ind_covariate=rand_covar), 
+                          ftCols = c("rand_covar"),
+                          parallel=TRUE)
+
+  saveRDS(sb3, file = resfile_uninf)
+}else{
+  sb3 <- readRDS(resfile_uninf)
 }
@@
 
+### Covariate three: Random
+
+Next, we'll look at the performance metrics for the random covariate.
+
+```{r, metrics3}
+# rename assay to qvalue
+assayNames(sb3) <- "qvalue"
+sb3 <- addDefaultMetrics(sb3)
+```
+
+Now, we'll plot the results.
+
+```{r, plot3, results="hide"}
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter(sb3, supplementary=FALSE) +
+  ggtitle("scDD, Covariate 3: Random")
+
+rejection_scatter_bins(sb3, covariate="rand_covar", bins=4,
+                       supplementary=FALSE) +
+  ggtitle("scDD, Covariate 3: Random")
+
+# upset plot 
+plotFDRMethodsOverlap(sb3, 
+                      alpha=0.05, nsets=ncol(sb3),
+                      order.by="freq", decreasing=TRUE,
+                      supplementary=FALSE)
+```
+
+```{r, fig.width=8, fig.height=3.5}
+mcols(sb3)$ind_covariate <- mcols(sb3)$rand_covar
+covariateLinePlot(sb3, alpha=0.05, covname="ind_covariate", nbins=25) +
+  ggtitle("scDD, Covariate 3: Random")
+```
+
+# Covariate comparison
+
+Here we compare the method ranks for the two covariates at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(resfile_mean, resfile_det, resfile_uninf), 
+                colLabels = c("Mean nonzero", "Detection rate", "Uninf"), 
+                alpha = 0.10, xlab = "Covariate")
+```
+
 # Session Information

### 2018-07-06 datasets/scrnaseq/scrnaseq-human-wilcox.Rmd
Add method heatmaps to case studies with more than one covariate/comparison
Add uninformative covariate analyses to scrnaseq

@@
 resfile_det <- file.path(sbdir, "human-benchmark-wilcox-det.rds")
+resfile_uninf <- file.path(sbdir, "human-benchmark-wilcox-uninf.rds")
 
@@
 and detection rate (defined as the proportion of cells expressing the gene at a nonzero level).
+In addition, we'll add a random (uninformative covariate).
 These will also be added to the `rowData` slot of the `SummarizedExperiments`.
@@
 
+set.seed(1339)
 humandata <- add_covariates_scrnaseq(humandata)
@@
 
+### Covariate two: Detection Rate
+
+Next we look at the random covariate.
+
+```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
+rank_scatter(data.frame(rowData(humandata)), 
+             pvalue="wilcox_p", covariate="rand_covar") + 
+  ggtitle("Wilcoxon_Ranksum, Covariate 3: Random")
+```
+
+```{r, fig.width=10, fig.height=3.2, message=FALSE}
+strat_hist(data.frame(rowData(humandata)), 
+           pvalue="wilcox_p", covariate="rand_covar", maxy=16,
+           main = "Wilcoxon_Ranksum, Covariate 3: Random") 
+```
+
 ## Multiple-Testing Correction
@@
   sb2 <- readRDS(resfile_det)
+}
+```
+
+### Covariate three: Random
+
+Now, we'll repeat the multiple testing correction using the random covariate:
+
+```{r, sb3, results="hide", message=FALSE}
+if (!file.exists(resfile_uninf)){
+  sb3 <- bd %>% buildBench(data.frame(rowData(humandata)) %>% 
+                                   na.omit() %>%
+                                   mutate(pval=wilcox_p,
+                                          ind_covariate=rand_covar), 
+                          ftCols = c("rand_covar"),
+                          parallel=TRUE)
+
+  saveRDS(sb3, file = resfile_uninf)
+}else{
+  sb3 <- readRDS(resfile_uninf)
 }
@@
 
+### Covariate three: Random
+
+Next, we'll look at the performance metrics for the rancom covariate.
+
+```{r, metrics3}
+# rename assay to qvalue
+assayNames(sb3) <- "qvalue"
+sb3 <- addDefaultMetrics(sb3)
+```
+
+Now, we'll plot the results.
+
+```{r, plot3, results="hide"}
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter(sb3, supplementary=FALSE) +
+  ggtitle("Wilcoxon_Ranksum, Covariate 3: Random")
+
+rejection_scatter_bins(sb3, covariate="rand_covar", bins=4,
+                       supplementary=FALSE) +
+  ggtitle("Wilcoxon_Ranksum, Covariate 3: Random")
+
+# upset plot 
+plotFDRMethodsOverlap(sb3, 
+                      alpha=0.05, nsets=ncol(sb3),
+                      order.by="freq", decreasing=TRUE,
+                      supplementary=FALSE)
+
+```
+
+```{r, fig.width=8, fig.height=3.5}
+mcols(sb3)$ind_covariate <- mcols(sb3)$rand_covar
+covariateLinePlot(sb3, alpha=0.05, covname="ind_covariate", nbins=25) +
+  ggtitle("Wilcoxon_Ranksum, Covariate 3: Random")
+```
+
+# Covariate comparison
+
+Here we compare the method ranks for the two covariates at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(resfile_mean, resfile_det, resfile_uninf), 
+                colLabels = c("Mean nonzero", "Detection rate", "Uninf"), 
+                alpha = 0.10, xlab = "Covariate")
+```
+
 # Session Information

### 2018-07-06 datasets/scrnaseq/scrnaseq-mouse-mast.Rmd
Add method heatmaps to case studies with more than one covariate/comparison
Add uninformative covariate analyses to scrnaseq

@@
 resfile_det <- file.path(sbdir, "mouse-benchmark-mast-det.rds")
+resfile_uninf <- file.path(sbdir, "mouse-benchmark-mast-uninf.rds")
 
@@
 
+set.seed(6823)
 mousedata <- add_covariates_scrnaseq(mousedata)
@@
 
+### Covariate three: Random
+
+Next we look at the random covariate.
+
+```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
+rank_scatter(data.frame(rowData(mousedata)), 
+             pvalue="mast_p", covariate="rand_covar") + 
+  ggtitle("MAST, Covariate 3: Random")
+```
+
+```{r, fig.width=10, fig.height=3.2, message=FALSE}
+strat_hist(data.frame(rowData(mousedata)), 
+           pvalue="mast_p", covariate="rand_covar", maxy=20,
+           main = "MAST, Covariate 3: Random")
+```
+
 ## Multiple-Testing Correction
@@
 
-Now, we'll repeat the multiple testing correction using the other covariate (detection rate):
+Now, we'll repeat the multiple testing correction using the detection rate covariate:
 
@@
   sb2 <- readRDS(resfile_det)
+}
+```
+
+### Covariate three: Random
+
+Now, we'll repeat the multiple testing correction using the random covariate:
+
+```{r, sb3, results="hide", message=FALSE}
+if (!file.exists(resfile_uninf)){
+  sb3 <- bd %>% buildBench(data.frame(rowData(mousedata)) %>% 
+                                   na.omit() %>%
+                                   mutate(pval=mast_p,
+                                          ind_covariate=rand_covar), 
+                          ftCols = c("rand_covar"),
+                          parallel=TRUE)
+
+  saveRDS(sb3, file = resfile_uninf)
+}else{
+  sb3 <- readRDS(resfile_uninf)
 }
@@
 
-Next, we'll look at the performance metrics for the other covariate (detection rate).
+Next, we'll look at the performance metrics for the detection rate covariate.
 
@@
 
+### Covariate three: Random
+
+Next, we'll look at the performance metrics for the random covariate.
+
+```{r, metrics3}
+# rename assay to qvalue
+assayNames(sb3) <- "qvalue"
+sb3 <- addDefaultMetrics(sb3)
+```
+
+Now, we'll plot the results.
+
+```{r, plot3, results="hide"}
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter(sb3, supplementary=FALSE) +
+  ggtitle("MAST, Covariate 3: Random")
+
+rejection_scatter_bins(sb3, covariate="rand_covar", bins=4,
+                       supplementary=FALSE) +
+  ggtitle("MAST, Covariate 3: Random")
+
+# upset plot 
+plotFDRMethodsOverlap(sb3, 
+                      alpha=0.05, nsets=ncol(sb3),
+                      order.by="freq", decreasing=TRUE,
+                      supplementary=FALSE)
+```
+
+```{r, fig.width=8, fig.height=3.5}
+mcols(sb3)$ind_covariate <- mcols(sb3)$rand_covar
+covariateLinePlot(sb3, alpha=0.05, covname="ind_covariate", nbins=25) +
+  ggtitle("MAST, Covariate 3: Random")
+```
+
+# Covariate comparison
+
+Here we compare the method ranks for the two covariates at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(resfile_mean, resfile_det, resfile_uninf), 
+                colLabels = c("Mean nonzero", "Detection rate", "Uninf"), 
+                alpha = 0.10, xlab = "Covariate")
+```
+
 # Session Information

### 2018-07-06 datasets/scrnaseq/scrnaseq-mouse-scdd.Rmd
Add method heatmaps to case studies with more than one covariate/comparison
Add uninformative covariate analyses to scrnaseq

@@
 resfile_det <- file.path(sbdir, "mouse-benchmark-scdd-det.rds")
+resfile_uninf <- file.path(sbdir, "mouse-benchmark-scdd-uninf.rds")
 
@@
 
+set.seed(99764)
 mousedata <- add_covariates_scrnaseq(mousedata)
@@
 
+### Covariate three: Random
+
+Next we look at the random rate covariate.
+
+```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
+rank_scatter(data.frame(rowData(mousedata)), 
+             pvalue="scdd_p", covariate="rand_covar") + 
+  ggtitle("scDD, Covariate 3: Random")
+```
+
+```{r, fig.width=10, fig.height=3.2, message=FALSE}
+strat_hist(data.frame(rowData(mousedata)), 
+           pvalue="scdd_p", covariate="rand_covar", maxy=20,
+           main = "scDD, Covariate 3: Random") 
+```
+
 ## Multiple-Testing Correction
@@
   sb2 <- readRDS(resfile_det)
+}
+```
+
+### Covariate three: Random
+
+Now, we'll repeat the multiple testing correction using the random covariate:
+
+```{r, sb3, results="hide", message=FALSE}
+if (!file.exists(resfile_uninf)){
+  sb3 <- bd %>% buildBench(data.frame(rowData(mousedata)) %>% 
+                                   filter(!is.na(scdd_p)) %>%
+                                   mutate(pval=scdd_p,
+                                          ind_covariate=rand_covar), 
+                          ftCols = c("rand_covar"),
+                          parallel=TRUE)
+
+  saveRDS(sb3, file = resfile_uninf)
+}else{
+  sb3 <- readRDS(resfile_uninf)
 }
@@
 
+### Covariate three: Random
+
+Next, we'll look at the performance metrics for the random covariate.
+
+```{r, metrics3}
+# rename assay to qvalue
+assayNames(sb3) <- "qvalue"
+sb3 <- addDefaultMetrics(sb3)
+```
+
+Now, we'll plot the results.
+
+```{r, plot3, results="hide"}
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter(sb3, supplementary=FALSE) +
+  ggtitle("scDD, Covariate 3: Random")
+
+rejection_scatter_bins(sb3, covariate="rand_covar", bins=4,
+                       supplementary=FALSE) +
+  ggtitle("scDD, Covariate 3: Random")
+
+# upset plot 
+plotFDRMethodsOverlap(sb3, 
+                      alpha=0.05, nsets=ncol(sb3),
+                      order.by="freq", decreasing=TRUE,
+                      supplementary=FALSE)
+
+```
+```{r, fig.width=8, fig.height=3.5}
+mcols(sb3)$ind_covariate <- mcols(sb3)$rand_covar
+covariateLinePlot(sb3, alpha=0.05, covname="ind_covariate", nbins=25) +
+  ggtitle("scDD, Covariate 3: Random")
+```
+
+# Covariate comparison
+
+Here we compare the method ranks for the two covariates at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(resfile_mean, resfile_det, resfile_uninf), 
+                colLabels = c("Mean nonzero", "Detection rate", "Uninf"), 
+                alpha = 0.10, xlab = "Covariate")
+```
+
 # Session Information

### 2018-07-06 datasets/scrnaseq/scrnaseq-mouse-wilcox.Rmd
Add method heatmaps to case studies with more than one covariate/comparison
Add uninformative covariate analyses to scrnaseq

@@
 resfile_det <- file.path(sbdir, "mouse-benchmark-wilcox-det.rds")
+resfile_uninf <- file.path(sbdir, "mouse-benchmark-wilcox-uninf.rds")
 
@@
 
+set.seed(13048)
 mousedata <- add_covariates_scrnaseq(mousedata)
@@
 
+### Covariate three: Random
+
+Next we look at the random covariate.
+
+```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
+rank_scatter(data.frame(rowData(mousedata)), 
+             pvalue="wilcox_p", covariate="rand_covar") + 
+  ggtitle("Wilcoxon-ranksum, Covariate 3: Random")
+```
+
+```{r, fig.width=10, fig.height=3.2, message=FALSE}
+strat_hist(data.frame(rowData(mousedata)), 
+           pvalue="wilcox_p", covariate="rand_covar", maxy=20,
+           main = "Wilcoxon-ranksum, Covariate 3: Random")
+```
+
 ## Multiple-Testing Correction
@@
   sb2 <- readRDS(resfile_det)
+}
+```
+
+### Covariate three: Random
+
+Now, we'll repeat the multiple testing correction using the random covariate:
+
+```{r, sb3, results="hide", message=FALSE}
+if (!file.exists(resfile_uninf)){
+  sb3 <- bd %>% buildBench(data.frame(rowData(mousedata)) %>% 
+                                   na.omit() %>%
+                                   mutate(pval=wilcox_p,
+                                          ind_covariate=rand_covar), 
+                          ftCols = c("rand_covar"),
+                          parallel=TRUE)
+
+  saveRDS(sb3, file = resfile_uninf)
+}else{
+  sb3 <- readRDS(resfile_uninf)
 }
@@
 
+### Covariate three: Random
+
+Next, we'll look at the performance metrics for the random covariate.
+
+```{r, metrics3}
+# rename assay to qvalue
+assayNames(sb3) <- "qvalue"
+sb3 <- addDefaultMetrics(sb3)
+```
+
+Now, we'll plot the results.
+
+```{r, plot3, results="hide"}
+# plot nrejects by method overall and stratified by covariate
+rejections_scatter(sb3, supplementary=FALSE) +
+  ggtitle("Wilcoxon-ranksum, Covariate 3: Random")
+
+rejection_scatter_bins(sb3, covariate="rand_covar", bins=4,
+                       supplementary=FALSE) +
+  ggtitle("Wilcoxon-ranksum, Covariate 3: Random")
+
+# upset plot 
+plotFDRMethodsOverlap(sb3, 
+                      alpha=0.05, nsets=ncol(sb3),
+                      order.by="freq", decreasing=TRUE,
+                      supplementary=FALSE)
+```
+
+```{r, fig.width=8, fig.height=3.5}
+mcols(sb3)$ind_covariate <- mcols(sb3)$rand_covar
+covariateLinePlot(sb3, alpha=0.05, covname="ind_covariate", nbins=25) +
+  ggtitle("Wilcoxon-ranksum, Covariate 3: Random")
+```
+
+# Covariate comparison
+
+Here we compare the method ranks for the two covariates at alpha = 0.10.
+
+```{r}
+plotMethodRanks(c(resfile_mean, resfile_det, resfile_uninf), 
+                colLabels = c("Mean nonzero", "Detection rate", "Uninf"), 
+                alpha = 0.10, xlab = "Covariate")
+```
+
 # Session Information

### 2018-07-06 datasets/microbiome/microbiome-enigma.Rmd
Remove NA pvalues in microbiome-enigma Al

@@
 }
+
+# remove OTUs with missing pvals
+res <- res %>% filter(!is.na(pval))
 ```

### 2018-07-06 manuscript_figures/manuscript_figures.Rmd
Incorporate uninformative sb objects; standardize heatmap styles

@@
                    grepl("de5", objects) & 
-                  !grepl("uninfCov", objects)]
+                  !grepl("uninf", objects)]
 
@@
 # remove microbiome objects from 'null' comparisons
-# removing baxter, goodrich and papa OTU-level (since they mostly found 
+# removing baxter, goodrich, papa OTU-level, and So4-enigma (since they mostly found 
 # nothing for all methods and the genus-level analysis is also present)
+# Also remove'log' ubiquity since redundant
 null <- which(grepl(c("goodrich|papa"), objects) & 
               grepl(c("otu"), objects) | 
-              grepl("baxter", objects))
+              grepl("baxter", objects) |
+              grepl("log", objects) | 
+              grepl("so4", objects))
 
@@
 
+# remove random covariate analyses
+uninf <- which(grepl("uninf", objects))
+if (length(uninf) > 0){
+  objects <- objects[-uninf]
+}
+
 # grab casestudy names
@@
 
-Fig4.rowOrder <- levels(plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
-                            tableOnly = TRUE)$method)
+Fig4.rowOrder <- c("bonf", "bh", "ihw", "qvalue", "bl", "lfdr",
+                   "ashq")
 Fig4 <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
@@
 # not excluding any methods
-Fig4All.rowOrder <- levels(plotMethodRanks(objects, colLabels = casestudy, alpha = alpha, 
-                               tableOnly = TRUE, excludeMethods = NULL)$method)
+Fig4All.rowOrder <- c("bonf", "bh", "ihw", "qvalue", "bl", "lfdr",
+                  "scott-empirical", "scott-theoretical", "ashq")
 Fig4All <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
-                        excludeMethods = NULL)
+                        excludeMethods = NULL, rowOrder = Fig4All.rowOrder)
 Fig4All
@@
             !grepl("scRNAseq", casestudy) &
-            !grepl("uninfCov", dataset))
+            !grepl("uninf", dataset))
 objs <- objects[wcs]
@@
                 fill = "FDR",
-                rowOrder = Fig4.rowOrder, 
-                annotate = "FDR") +
+                rowOrder = Fig4All.rowOrder, 
+                annotate = "FDR", excludeMethods = NULL) +
           scale_fill_gradient2(high = "red", 
@@
                          low = "blue", 
-                         midpoint = alpha)
+                         midpoint = alpha) +
+  labs(fill = "Mean FDR")
 
@@
                 fill = "TPR",
-                colorLow = "blue", 
-                colorHigh = "white",
-                rowOrder = Fig4.rowOrder,
-                annotate = "TPR")
+                rowOrder = Fig4All.rowOrder,
+                annotate = "TPR", excludeMethods = NULL)
 
@@
   draw_plot(ptpr + ylab("") + xlab(""), 0.78, 0, 0.22, 1) +
-  draw_plot(Fig4, 0, 0, 0.60, 1) +
+  draw_plot(Fig4All, 0, 0, 0.60, 1) +
   draw_plot_label(c("A", "B"), c(0, 0.57), c(1, 1), size = 15)
-ggsave(file.path(outdir, "Figure4.pdf"), width=12, height=5)
+ggsave(file.path(outdir, "Figure4.pdf"), width=14, height=5)
 
@@
 ggplot(rank.tab , 
+       aes(y = prop_casestudy, x = method, group = method, fill = Method)) +
+  geom_boxplot(alpha=0.7) +
+  scale_fill_manual(values=col) +
+  theme_bw() +
+  xlab("Method (ordered by rejections in yeast simulation (n=5))") +  
+  ylab("Proportion of maximum rejections in casestudies")
+ggsave(file.path(outdir, "Figure_simVcasestudyRanks.pdf"), width=6, height=4)
+
+ggplot(rank.tab , 
        aes(y = prop_casestudy, x = method, group = method, color = Method)) +
@@
 objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
-resfile <- objects[grepl("RNAseq", objects) & grepl("brain", objects)] 
+resfile <- objects[grepl("RNAseq", objects) & grepl("brain", objects) & !grepl("uninf", objects)] 
 res <- readRDS(file=resfile)
@@
 resfile <- objects[grepl("RNAseq", objects) & grepl("human", objects) &
-                   grepl("mast", objects) & grepl("det", objects)] 
+                   grepl("mast", objects) & grepl("det", objects) & !grepl("uninf", objects)] 
 res <- readRDS(file=resfile)
@@
 
-resfile <- objects[grepl("GSEA", objects) & grepl("mouse", objects)] 
+resfile <- objects[grepl("GSEA", objects) & grepl("mouse", objects) & !grepl("uninf", objects)] 
 res <- readRDS(file=resfile)
@@
 p5 <- plot_grid(p5, leg, ncol=2, rel_widths = c(1,0.2))
+p5
 ggsave(file.path(outdir, "Figure5.pdf"), width=9, height=7)
@@
                 excludeMethods = NULL, Nlabel = FALSE, annotate = "propPossible") + 
-  ggtitle("GWAS")
-ggsave(file.path(outdir, "FigureS1.pdf"), width=4, height=4)
+  ggtitle("GWAS") + 
+  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
+ggsave(file.path(outdir, "FigureS1.pdf"), width=5, height=4.5)
 ```
@@
                 Nlabel = FALSE, annotate = "propPossible") + 
-  ggtitle("GSEA")
-ggsave(file.path(outdir, "FigureS2.pdf"), width=4, height=4)
+  ggtitle("GSEA") + 
+  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
+ggsave(file.path(outdir, "FigureS2.pdf"), width=6, height=4.5)
 ```
@@
                 Nlabel = FALSE, annotate = "propPossible") + 
-  ggtitle("ChIP-Seq")
-ggsave(file.path(outdir, "FigureS3.pdf"), width=5, height=4)
+  ggtitle("ChIP-Seq") + 
+  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
+ggsave(file.path(outdir, "FigureS3.pdf"), width=5, height=5)
 ```
@@
                 excludeMethods = NULL, Nlabel = FALSE, annotate = "propPossible") + 
-  ggtitle("RNA-seq")
-ggsave(file.path(outdir, "FigureS4.pdf"), width=4, height=4)
+  ggtitle("RNA-seq") + 
+  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
+ggsave(file.path(outdir, "FigureS4.pdf"), width=6, height=5)
 ```
@@
 
+uninf <- which(grepl("uninf", objs))
+
 # by dataset
-pa <- plotMethodRanks(objs, colLabels = ds, alpha = alpha, 
+pa <- plotMethodRanks(objs[-uninf], colLabels = ds[-uninf], alpha = alpha, 
                       xlab = "Dataset", annotate = "propPossible") + 
-  ggtitle("scRNA-seq")
+  ggtitle("scRNA-seq")+ 
+  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
 
 # by method
-pb <- plotMethodRanks(objs, colLabels = meth, alpha = alpha,
+pb <- plotMethodRanks(objs[-uninf], colLabels = meth[-uninf], alpha = alpha,
                       xlab = "Method", annotate = "propPossible") + 
-  ggtitle("scRNA-seq")
+  ggtitle("scRNA-seq")+ 
+  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
 
@@
                       xlab = "Covariate", annotate = "propPossible") + 
-  ggtitle("scRNA-seq")
+  ggtitle("scRNA-seq")+ 
+  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
 
-plot_grid(pa, pb, pc)
-ggsave(file.path(outdir, "FigureS5.pdf"), width=8, height=5.5)
+plot_grid(pa, pb, pc, ncol=3)
+ggsave(file.path(outdir, "FigureS5.pdf"), width=15, height=6)
 ```
@@
                 Nlabel = FALSE, annotate = "propPossible") + 
-  ggtitle("Microbiome")
-ggsave(file.path(outdir, "FigureS6.pdf"), width=7.5, height=4)
+  ggtitle("Microbiome")+ 
+  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
+ggsave(file.path(outdir, "FigureS6.pdf"), width=14, height=6.5)
 ```
@@
 wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("brain|mir|polyester|null|uninfCov"), objects) &
+            !grepl(c("brain|mir|polyester|null"), objects) &
             !grepl("scRNAseq", casestudy))
@@
                 alpha = alpha, xlab = "Comparison",
-                excludeMethods = NULL)
+                excludeMethods = NULL,
+                annotate = "propPossible")
 
@@
                 excludeMethods = NULL, 
-                fill = "FDR")
+                fill = "FDR",
+                annotate = "propPossible")
 
@@
                 excludeMethods = NULL, 
-                fill = "TPR")
+                fill = "TPR",
+                annotate = "propPossible")
 
@@
 wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("brain|mir|polyester|de|uninfCov"), objects) &
+            !grepl(c("brain|mir|polyester|de"), objects) &
             !grepl("scRNAseq", casestudy))
@@
 wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("brain|mir|yeast|null|uninfCov"), objects) &
+            !grepl(c("brain|mir|yeast|null"), objects) &
             !grepl("scRNAseq", casestudy))
@@
 
-## exclude yeast and polyester simulations
-objects <- objects[!grepl("yeast", objects) & !grepl("polyester", objects)]
+## exclude yeast and polyester simulations, as well as uninformative covariate
+## analyses
+objects <- objects[!grepl("yeast", objects) & !grepl("polyester", objects) &
+                   !grepl("uninf", objects)]
 
@@
 ## read in case study results (proportion of rejects vs. case study max)
-objdat <- tidy_df(objects, object_labs, "propMaxRejections", NULL)
+objdat <- tidy_df(objects=objects, colLabels=object_labs, 
+                  fill="propMaxRejections", annotate=NULL,
+                  alpha =0.10)
 objdat <- as_tibble(objdat)

### 2018-07-06 datasets/R/plotMethodRank.R

@@ -11,12 +11,6 @@
 #'  heatmap columns (same length and order as `objects`). 
 #' @param alpha numeric value indicating which alpha value to use for IHW 
 #'  (needs to be one of the values specified in the bench object). Default 0.10.
-#' @param colorLow character representing the color to be used for the lowest 
-#'  value
-#' @param colorHigh character representing the color to be used for the highest
-#'  value
-#' @param colorNA character representing the color to be used for NA (i.e. 
-#'  for methods that weren't used in the particular benchmark)
 #' @param linePlot logical indicating whether to plot a line plot in lieu of 
 #'  a heatmap  
 #' @param excludeMethods character vector containing the names of methods to
@@ -41,9 +35,9 @@
 #' @author Keegan Korthauer  
 
 plotMethodRanks <- function(objects, colLabels, alpha = 0.10, 
-                            colorLow = "navy", colorHigh = "yellow",
-                            colorNA = "white", xlab = "Case Study",
+                            xlab = "Case Study",
                             fill = c("propMaxRejections", "meanRank", 
+                                     "meanRejections",
                                      "FDR", "TPR"),
                             linePlot = FALSE,
                             excludeMethods = c("scott-theoretical",
@@ -87,6 +81,14 @@ plotMethodRanks <- function(objects, colLabels, alpha = 0.10,
     annot <- annotate
   }
   
+  if (fill == "propMaxRejections"){
+    annot2 <- "mean_prop"
+  }else if(fill == "meanRank"){
+    annot2 <- "mean_rank"
+  }else{
+    annot2 <- "mean_nrej"
+  }
+  
   if(is.null(annot))
     annot <- "propPossible"
   ranks_avg <- ranks %>% 
@@ -98,6 +100,14 @@ plotMethodRanks <- function(objects, colLabels, alpha = 0.10,
                max_prop = max(propMaxRej, na.rm = TRUE),
                topLayer = mean(get(annot), na.rm = TRUE),
                nsets = sum(!is.na(nrejects)))
+  
+  # replace all but max value of annotate variable with NA (within casestudy)
+  maxvals <- ranks_avg %>% 
+    group_by(casestudy) %>%
+    summarize(maxann = max(get(annot2), na.rm=TRUE))
+  ranks_avg <- left_join(ranks_avg, maxvals, by = "casestudy")
+  ranks_avg <- ranks_avg %>%
[…]

### 2018-07-09 manuscript_figures/manuscript_figures_sims.Rmd
Change informativeness and pi0 sims plots to match Fig2 style

@@
 ```{r, fig.width = 8, fig.height = 5}
-dplyr::filter(res_nullset, key == "FWER") %>%
+gp_nullA <- dplyr::filter(res_nullset, key == "FWER") %>%
     filter(! (Method == "ashq" & dist == "chisq4"),
@@
     ggtitle("Proportion of replications with at least one false positive (over 100 replications)")
+gp_nullA
 ```
@@
 ```{r, fig.width = 8, fig.height = 5}
-gp <- dplyr::filter(res_nullset, key == "TNR") %>%
+gp_nullB <- dplyr::filter(res_nullset, key == "TNR") %>%
     filter(! (Method == "ashq" & dist == "chisq4"),
@@
     facet_grid(. ~ dist) +
-    scale_y_continuous("TPR", labels = scales::percent) +
+    scale_y_continuous("TNR", labels = scales::percent) +
     scale_color_manual(values = col, guide = FALSE) +
     theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
-    ggtitle("Average TNR (over 100 replications)") 
-gp
+    ggtitle("TNR in Null Simulations (over 100 replications)") 
+gp_nullB
 
+ggsave(file.path(outdir, "FigureSsim-nullA.pdf"), gp_nullA, width = 8, height = 5)
+ggsave(file.path(outdir, "FigureSsim-nullB.pdf"), gp_nullB, width = 8, height = 5)
+
+gp_zoom <- plot_grid(pfdr + coord_cartesian(c(0, .2), c(0, .2)) + guides(color = FALSE, linetype = FALSE),
+                     ptpr + coord_cartesian(c(.6, 1), c(.6, 1)), rel_widths = c(1, 1.4), labels = LETTERS[1:2])
 ggsave(file.path(outdir, "FigureSsim-null.pdf"), gp, width = 8, height = 5)
+
 ```
@@
 
+We notice that under certain settings, the mean and SE FDR values are `NA`s for some methods under some
+settings. We take a closer look at these cases.
+
+```{r}
+dplyr::select(res_pi0set, setting, Method, pi0, mean_value, key) %>%
+    dplyr::filter(key %in% c("rejections", "FDR")) %>%
+        tidyr::spread(key, mean_value) %>%
+        dplyr::filter(is.na(FDR))
+```
+
+All cases correspond to methods and settings where the method rejected 0 tests over all replications.
+We deal with these cases by setting these values to zero for plotting purposes.
+
+```{r}
+res_pi0set$se[is.na(res_pi0set$mean_value) & res_pi0set$key == "FDR"] <- 0
+res_pi0set$mean_value[is.na(res_pi0set$mean_value) & res_pi0set$key == "FDR"] <- 0
+```
+
 We take a look at the TPR and FDR as function of the proportion of null hypotheses.
@@
     ggplot(aes(x = pi0 / 100, y = mean_value, color = Method, lty = Method)) +
-    geom_line(alpha = 1/6, position = position_dodge(width = .01)) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), alpha = 1) +
-    geom_point(alpha = 1, position = position_dodge(width = .01)) +
+    geom_line(alpha = 0.85, position = position_dodge(width = .01)) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width=0.025, alpha=0.5) +
     scale_color_manual(values = col) +
@@
     facet_grid(setting ~ .) + 
-    ggtitle("Average FDR Across Varying Null Proportion")
+    ggtitle("Mean FDR Over 100 Replications") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
 
@@
     facet_grid(setting ~ .) + 
-    ggtitle("Average TPR Across Varying Null Proportion")
+    ggtitle("Mean TPR Over 100 Replications") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
 ```
@@
 
-ggsave(file.path(outdir, "FigureSsim-nullprop.pdf"), gp, width = 15, height = 7.5)
+ggsave(file.path(outdir, "FigureSsim-nullprop.pdf"), gp, width = 10, height = 5)
 
@@
 
-ggsave(file.path(outdir, "FigureSsim-nullprop-zoom.pdf"), gp_zoom, width = 15, height = 7.5)
+ggsave(file.path(outdir, "FigureSsim-nullprop-zoom.pdf"), gp_zoom, width = 10, height = 5)
 ```
@@
 
+## including SE to show variability in FDRR
+pfdr <- dplyr::select(res_pi0set, Method, setting, pi0, key, mean_value, se, col, lty) %>%
+    dplyr::filter(key == "FDR", !is.na(mean_value)) %>%
+    ggplot(aes(x = pi0 / 100, y = mean_value, color = Method, lty = Method)) +
+    geom_line(alpha = 0.85, position = position_dodge(width = .01)) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width=0.025, alpha=0.5) +
+    scale_color_manual(values = col) +
+    scale_linetype_manual(values = lty) + 
+    geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) + 
+    scale_x_continuous("Null Proportion", labels = scales::percent, breaks = seq(0, 1, .1)) + 
+    scale_y_continuous("FDR", labels = scales::percent) + 
+    facet_grid(setting ~ .) + 
+    ggtitle("Mean FDR Over 100 Replications") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
 ## not including SE because it's too small
@@
     ggplot(aes(x = level / 100, y = mean_value, color = Method, lty = Method)) +
-    geom_line(alpha = 1/6, position = position_dodge(width = .01)) +
-    ##geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0) +
-    geom_point(alpha = 1, position = position_dodge(width = .01)) +
+    geom_line(alpha = 0.85, position = position_dodge(width = .01)) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width=0.025, alpha=0.5) +
     scale_color_manual(values = col) +
@@
     facet_grid(setting ~ .) + 
-    ggtitle("Average FDR Across Varying Informativeness")
+    ggtitle("Mean FDR Over 100 Replications") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
 pfdr
@@
     ggplot(aes(x = level / 100, y = mean_value, color = Method, lty = Method)) +
-    geom_line(alpha = 1/6, position = position_dodge(width = .01)) +
-    ##geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0) +    
-    geom_point(alpha = 1, position = position_dodge(width = .01)) +
+    geom_line(alpha = 0.85, position = position_dodge(width = .01)) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width=0.025, alpha=0.5) +
     scale_color_manual(values = col) +
@@
     facet_grid(setting ~ ., scales = "free_y") + expand_limits(y = 0) +
-    ggtitle("Average TPR Across Varying Informativeness")
+    ggtitle("Mean TPR Over 100 Replications") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
 ptpr
@@
 gp <- plot_grid(pfdr + guides(color = FALSE, linetype = FALSE), ptpr,
-          rel_widths = c(1, 1.25), labels = LETTERS[1:2])
+          rel_widths = c(1, 1.25), labels = LETTERS[3:4])
 gp
 
-ggsave(file.path(outdir, "FigureSsim-informativeness.pdf"), gp, width = 15, height = 7.5)
+ggsave(file.path(outdir, "FigureSsim-informativeness.pdf"), gp, width = 10, height = 5)
 ```

### 2018-07-09 manuscript_figures/manuscript_figures_sims.Rmd
Fix info, pi0 sim plots and clean up into single figure

@@
 
-# Varying Null Proportions
-
-Next, we subset on simulation results that look at varying the null proportions.
-
-```{r}
-res_pi0set <- dplyr::filter(res, grepl("varyingpi0", setting))
-res_pi0set <- dplyr::rename(res_pi0set, pi0 = dist)
-res_pi0set <- dplyr::mutate(res_pi0set, pi0 = as.numeric(gsub("nullprop", "", pi0)))
-res_pi0set <- tidyr::complete(res_pi0set, setting, Method, pi0, key)
-res_pi0set <- dplyr::mutate(res_pi0set, setting = factor(setting, levels = c("varyingpi0", "varyingpi0-lowsignal"),
-                                                         labels = c("high signal", "low signal")))
-```
-
-We notice that under certain settings, the mean and SE FDR values are `NA`s for some methods under some
-settings. We take a closer look at these cases.
-
-```{r}
-dplyr::select(res_pi0set, setting, Method, pi0, mean_value, key) %>%
-    dplyr::filter(key %in% c("rejections", "FDR")) %>%
-        tidyr::spread(key, mean_value) %>%
-        dplyr::filter(is.na(FDR))
-```
-
-All cases correspond to methods and settings where the method rejected 0 tests over all replications.
-We deal with these cases by setting these values to zero for plotting purposes.
-
-```{r}
-res_pi0set$se[is.na(res_pi0set$mean_value) & res_pi0set$key == "FDR"] <- 0
-res_pi0set$mean_value[is.na(res_pi0set$mean_value) & res_pi0set$key == "FDR"] <- 0
-```
-
-We take a look at the TPR and FDR as function of the proportion of null hypotheses.
-
-```{r}
-## looking at FDR vs TPR
-col <- as.character(res_pi0set$col)
-names(col) <- as.character(res_pi0set$Method)
-lty <- as.character(res_pi0set$lty)
-names(lty) <- as.character(res_pi0set$Method)
-
-## including SE to show variability in FDRR
-pfdr <- dplyr::select(res_pi0set, Method, setting, pi0, key, mean_value, se, col, lty) %>%
-    dplyr::filter(key == "FDR", !is.na(mean_value)) %>%
-    ggplot(aes(x = pi0 / 100, y = mean_value, color = Method, lty = Method)) +
-    geom_line(alpha = 0.85, position = position_dodge(width = .01)) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width=0.025, alpha=0.5) +
-    scale_color_manual(values = col) +
-    scale_linetype_manual(values = lty) + 
-    geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) + 
-    scale_x_continuous("Null Proportion", labels = scales::percent, breaks = seq(0, 1, .1)) + 
-    scale_y_continuous("FDR", labels = scales::percent) + 
-    facet_grid(setting ~ .) + 
-    ggtitle("Mean FDR Over 100 Replications") +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-## including SE to show variability in FDRR
-ptpr <- dplyr::select(res_pi0set, Method, setting, pi0, key, mean_value, se, col, lty) %>%
-    dplyr::filter(key == "TPR", !is.na(mean_value)) %>%
-    ggplot(aes(x = pi0 / 100, y = mean_value, color = Method, lty = Method)) +
-    geom_line(alpha = 0.85, position = position_dodge(width = .01)) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width=0.025, alpha=0.5) +
-    scale_color_manual(values = col) +
-    scale_linetype_manual(values = lty) + 
-    scale_x_continuous("Null Proportion", labels = scales::percent, breaks = seq(0, 1, .1)) + 
-    scale_y_continuous("TPR", labels = scales::percent) + 
-    facet_grid(setting ~ .) + 
-    ggtitle("Mean TPR Over 100 Replications") +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-```
-
-Again, we can plot both FDR and TPR together.
-
-```{r, fig.width = 15, fig.height = 7.5}
-gp <- plot_grid(pfdr + guides(color = FALSE, linetype = FALSE),
-                ptpr, rel_widths = c(1, 1.25), labels = LETTERS[1:2])
-gp
-
-ggsave(file.path(outdir, "FigureSsim-nullprop.pdf"), gp, width = 10, height = 5)
-
-## trimmed ranges
-gp_zoom <- plot_grid(pfdr + coord_cartesian(ylim = c(0, .25)) + guides(color = FALSE, linetype = FALSE),
-                     ptpr, rel_widths = c(1, 1.25), labels = LETTERS[1:2])
-gp_zoom
-
-ggsave(file.path(outdir, "FigureSsim-nullprop-zoom.pdf"), gp_zoom, width = 10, height = 5)
-```
-
 # Unimodal Effect Sizes
@@
 
+# Varying Null Proportions
+
+Next, we subset on simulation results that look at varying the null proportions.
+
+```{r}
+res_pi0set <- dplyr::filter(res, setting == "varyingpi0-lowsignal")
+res_pi0set <- dplyr::rename(res_pi0set, pi0 = dist)
+res_pi0set <- dplyr::mutate(res_pi0set, pi0 = as.numeric(gsub("nullprop", "", pi0)))
+res_pi0set <- tidyr::complete(res_pi0set, setting, Method, pi0, key)
+```
+
+We notice that under certain settings, the mean and SE FDR values are `NA`s for some methods under some
+settings. We take a closer look at these cases.
+
+```{r}
+dplyr::select(res_pi0set, setting, Method, pi0, mean_value, key) %>%
+    dplyr::filter(key %in% c("rejections", "FDR")) %>%
+        tidyr::spread(key, mean_value) %>%
+        dplyr::filter(is.na(FDR))
+```
+
+All cases correspond to methods and settings where the method rejected 0 tests over all replications.
+We deal with these cases by setting these values to zero for plotting purposes.
+
+```{r}
+res_pi0set$se[is.na(res_pi0set$mean_value) & res_pi0set$key == "FDR"] <- 0
+res_pi0set$mean_value[is.na(res_pi0set$mean_value) & res_pi0set$key == "FDR"] <- 0
+```
+
+We take a look at the TPR and FDR as function of the proportion of null hypotheses.
+
+```{r}
+## looking at FDR vs TPR
+col <- as.character(res_pi0set$col)
+names(col) <- as.character(res_pi0set$Method)
+lty <- as.character(res_pi0set$lty)
+names(lty) <- as.character(res_pi0set$Method)
+
+## including SE to show variability in FDRR
+pi0_fdr <- dplyr::select(res_pi0set, Method, setting, pi0, key, mean_value, se, col, lty) %>%
+    dplyr::filter(key == "FDR", !is.na(mean_value)) %>%
+    ggplot(aes(x = pi0 / 100, y = mean_value, color = Method, lty = Method)) +
+    geom_line(alpha = 0.85, position = position_dodge(width = .01)) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width=0.025, alpha=0.5) +
+    scale_color_manual(values = col) +
+    scale_linetype_manual(values = lty) + 
+    geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) + 
+    scale_x_continuous("Null Proportion", labels = scales::percent, breaks = seq(0, 1, .1)) + 
+    scale_y_continuous("FDR", labels = scales::percent) + 
+    ggtitle("Mean FDR Over 100 Replications") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+## including SE to show variability in FDRR
+pi0_tpr <- dplyr::select(res_pi0set, Method, setting, pi0, key, mean_value, se, col, lty) %>%
+    dplyr::filter(key == "TPR", !is.na(mean_value)) %>%
+    ggplot(aes(x = pi0 / 100, y = mean_value, color = Method, lty = Method)) +
+    geom_line(alpha = 0.85, position = position_dodge(width = .01)) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width=0.025, alpha=0.5) +
+    scale_color_manual(values = col) +
+    scale_linetype_manual(values = lty) + 
+    scale_x_continuous("Null Proportion", labels = scales::percent, breaks = seq(0, 1, .1)) + 
+    scale_y_continuous("TPR", labels = scales::percent) + 
+    ##facet_grid(setting ~ .) + 
+    ggtitle("Mean TPR Over 100 Replications") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+```
+
+Again, we can plot both FDR and TPR together.
+
+```{r, fig.width = 15, fig.height = 7.5}
+gp <- plot_grid(pi0_fdr + guides(color = FALSE, linetype = FALSE),
+                pi0_tpr + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:2])
+gp <- plot_grid(gp, get_legend(pi0_tpr), rel_widths = c(1, .125))
+gp
+
+ggsave(file.path(outdir, "FigureSsim-nullprop.pdf"), gp, width = 10, height = 5)
+```
+
 # Varying Informativeness
@@
 ```{r}
-res_infoset <- dplyr::filter(res, grepl("varyinginfo", setting))
+res_infoset <- dplyr::filter(res, setting == "varyinginfo-lowsignal")
 res_infoset <- dplyr::rename(res_infoset, level = dist)
@@
 res_infoset <- tidyr::complete(res_infoset, setting, Method, level, key)
-res_infoset <- dplyr::mutate(res_infoset, setting = factor(setting, levels = c("varyinginfo", "varyinginfo-lowsignal"),
-                                                         labels = c("high signal", "low signal")))
 ```
@@
 
-## including SE to show variability in FDRR
-pfdr <- dplyr::select(res_pi0set, Method, setting, pi0, key, mean_value, se, col, lty) %>%
-    dplyr::filter(key == "FDR", !is.na(mean_value)) %>%
-    ggplot(aes(x = pi0 / 100, y = mean_value, color = Method, lty = Method)) +
-    geom_line(alpha = 0.85, position = position_dodge(width = .01)) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width=0.025, alpha=0.5) +
-    scale_color_manual(values = col) +
-    scale_linetype_manual(values = lty) + 
-    geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) + 
-    scale_x_continuous("Null Proportion", labels = scales::percent, breaks = seq(0, 1, .1)) + 
-    scale_y_continuous("FDR", labels = scales::percent) + 
-    facet_grid(setting ~ .) + 
-    ggtitle("Mean FDR Over 100 Replications") +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
 ## not including SE because it's too small
-pfdr <- dplyr::select(res_infoset, Method, setting, level, key, mean_value, se, q10, q90, col, lty) %>%
+info_fdr <- dplyr::select(res_infoset, Method, setting, level, key, mean_value, se, q10, q90, col, lty) %>%
     dplyr::filter(key == "FDR") %>%
@@
     scale_y_continuous("FDR", labels = scales::percent) + 
-    facet_grid(setting ~ .) + 
     ggtitle("Mean FDR Over 100 Replications") +
@@
           plot.title = element_text(face="bold", size = 14))
-pfdr
 
-ptpr <- dplyr::select(res_infoset, Method, setting, level, key, mean_value, se, col, lty) %>%
+info_tpr <- dplyr::select(res_infoset, Method, setting, level, key, mean_value, se, col, lty) %>%
     dplyr::filter(key == "TPR") %>%
@@
     scale_y_continuous("TPR", labels = scales::percent) + 
-    facet_grid(setting ~ ., scales = "free_y") + expand_limits(y = 0) +
+    expand_limits(y = 0) +
     ggtitle("Mean TPR Over 100 Replications") +
@@
           plot.title = element_text(face="bold", size = 14))
-ptpr
 ```
@@
 ```{r, fig.width = 15, fig.height = 7.5}
-gp <- plot_grid(pfdr + guides(color = FALSE, linetype = FALSE), ptpr,
-          rel_widths = c(1, 1.25), labels = LETTERS[3:4])
+gp <- plot_grid(info_fdr + guides(color = FALSE, linetype = FALSE),
+                info_tpr + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:2])
+gp <- plot_grid(gp, get_legend(info_tpr), rel_widths = c(1, .125))
 gp
@@
 ```
+
+We also create a single plot with simulation results for both varying pi0 and varying
+informativeness settings.
+
+```{r}
+gp <- plot_grid(pi0_fdr + guides(color = FALSE, linetype = FALSE),
+                pi0_tpr + guides(color = FALSE, linetype = FALSE),
+                info_fdr + guides(color = FALSE, linetype = FALSE),
+                info_tpr + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:4])
+gp <- plot_grid(gp, get_legend(info_tpr), rel_widths = c(1, .2))
+gp
+
+ggsave(file.path(outdir, "FigureAlt-Sims.pdf"), gp, width = 10, height = 8)
+```

### 2018-07-11 datasets/microbiome/microbiome-goodrich.Rmd
Add uninf covariate for goodrich

--- 2018-05-13 datasets/microbiome/microbiome-goodrich.Rmd
@@
 Here we download and analyze the Goodrich obesity (OB) dataset
-[(Goodrich et al., 2014)](https://doi.org/10.1016/j.cell.2014.09.053).
+[(Goodrich et al., 2014)](http://dx.doi.org/10.1016/j.cell.2014.09.053).
 The dataset includes fecal samples from 428 individuals with lean body mass index (BMI < 25),
@@
 
-We expect few or no truly differentially abundant OTUs ([Finucane et al., 2014](https://doi.org/10.1371/journal.pone.0084689) and [Walters et al., 2014](https://doi.org/10.1016%2Fj.febslet.2014.09.039)), so this study provides a real-life "null" example.
+We expect few or no truly differentially abundant OTUs ([Finucane et al., 2014](http://dx.doi.org/10.1371/journal.pone.0084689) and [Walters et al., 2014](https://dx.doi.org/10.1016%2Fj.febslet.2014.09.039)), so this study provides a real-life "null" example.
 

### 2018-07-11 manuscript_figures/manuscript_figures.Rmd
Add summary figure for inf covar in casestudies

@@
                 annotate = "FDR", excludeMethods = NULL) +
-          scale_fill_gradient2(high = "red", 
+          scale_fill_gradient2(high = "darkred", 
                          mid = "white", 
-                         low = "blue", 
+                         low = "darkblue", 
                          midpoint = alpha) +
@@
 
+# Informativeness of covariates in case studies
+
+Here we create a heatmap similar to figure 4, but instead of containing the mean
+proportion of maximum rejections, it plots the mean difference in the number of 
+rejections using an informative covariate and the number of rejections using a 
+random uninformative covariate, expressed as a proportion of the number of rejections
+using the uninformative covariate. This is only created for those methods that
+use an informative covariate.
+
+```{r}
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+
+# remove simulation objects
+sim <- unique(c(which(grepl("simulations", objects)),
+               which(grepl("yeast", objects)),
+               which(grepl("polyester", objects))))
+if (length(sim) > 0){
+  objects <- objects[-sim]
+}
+
+# remove microbiome objects from 'null' comparisons
+# removing baxter, goodrich, papa OTU-level, and So4-enigma (since they mostly found 
+# nothing for all methods and the genus-level analysis is also present)
+# Also remove'log' ubiquity since redundant
+null <- which(grepl(c("goodrich|papa"), objects) & 
+              grepl(c("otu"), objects) | 
+              grepl("baxter", objects) |
+              grepl("log", objects) | 
+              grepl("so4", objects))
+
+if (length(null) > 0){
+  objects <- objects[-null]
+}
+
+# remove chip-seq promoter analysis (since it is the only one that uses 
+# ash and we also include the csaw analysis on the same dataset).
+promot <- which(grepl("promot", objects))
+if (length(promot) > 0){
+  objects <- objects[-promot]
+}
+
+# methods that use a covariate
+cov.methods <- c("ihw", "bl", "lfdr", "scott-empirical", "scott-theoretical")
+
+# regexp to construct master data table separating case study datasets, methods
+# and covariates
+obj.names <- str_split(objects, pattern = "/")
+casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
+filename <- gsub(".rds", "", unlist(lapply(obj.names, function(x) x[4] )))
+dataset <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[1]]))
+extD <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[min(length(x), 2)]]))
+extD <- gsub("samplesize", "n", extD)
+extM <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[min(length(x), 3)]]))
+extC <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[min(length(x), 4)]]))
+inform <-  ifelse(grepl("uninf", objects), "uninf", "inf")
+cov <- inform
+df <- data.frame(filename, casestudy, dataset, inform, cov, stringsAsFactors = FALSE) %>%
+  mutate(dataset = ifelse(grepl("enigma", dataset), paste(dataset, extD, sep="-"),
+                          dataset)) %>%  
+  mutate(dataset = ifelse(casestudy == "scRNAseq", paste(dataset, extM, sep="-"),
+                          dataset)) %>%
+  mutate(cov = ifelse(casestudy == "scRNAseq" & inform != "uninf", extC,  cov)) %>%
+  mutate(cov = ifelse(casestudy == "ChIPseq" & inform != "uninf", "region-width",  cov)) %>%
+  mutate(cov = ifelse(casestudy == "GSEA" & inform != "uninf", "set-size",  cov)) %>%
+  mutate(cov = ifelse(casestudy == "GWAS" & inform != "uninf", extD,  cov)) %>%
+  mutate(cov = ifelse(casestudy == "microbiome" & inform != "uninf", 
+                      ifelse(grepl("abun", filename), "abundance", "ubiquity"),  cov))%>%
+  mutate(cov = ifelse(casestudy == "RNAseq" & inform != "uninf", "mean",  cov))
+
+tab <- tidy_df(objects=objects, colLabels=filename, 
+              fill="meanRank", annotate=NULL, alpha =0.10) %>%
+  rename(casestudy = "filename") %>%
+  mutate(filename = gsub(".rds", "", filename))
+
+dftab <- left_join(tab, df, by="filename") %>%
+  filter(method %in% cov.methods) %>%
+  select(filename, casestudy, dataset, method, cov, inform, nrejects) %>%
+  mutate(id = paste(casestudy, dataset, method, sep = "-"))
+
+uninftab <- dftab %>% 
+  filter(inform == "uninf") %>%
+  select(id, nrejects) %>%
+  rename(nrejects = "uninf")
+
+inftab <- dftab %>%
+  filter(inform == "inf") %>%
+  select(-inform)
+
+rowOrder <- c("ihw", "bl", "lfdr", "scott-empirical", "scott-theoretical")
+
+dftab <- left_join(inftab, uninftab, by = "id") %>%
+  mutate(diff_inf = nrejects - uninf) %>%
+  mutate(diff_inf_percent = diff_inf / nrejects) %>%
+  mutate(method = factor(method, levels = rowOrder)) %>%
+  group_by(casestudy, cov) %>%
+  complete(method, dataset)
+
+Fig_impactofinfcov <- ggplot(dftab, aes(x = dataset, y = method, fill = diff_inf_percent)) + 
+        geom_raster() + 
+        scale_fill_gradientn("% Increase\nin rejections",
+                             labels = scales::percent, trans="log1p",
+                             colours = c("darkblue","white","darkred"), 
+                             values = scales::rescale(c(-.1,0,1)),
+                             guide = "colorbar", limits=c(-.05,0.99)) +
+        theme_bw() +
+        xlab("Case Study") +
+        ylab("Method") +
+        facet_grid( ~ casestudy + cov, scales = "free", space = "free") +
+        ggtitle("Impact of informative covariate on number of rejections") +
+        theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
+              legend.position = c(0.985, 0.91), legend.justification = c(1, 1),
+              legend.background = element_rect(fill = scales::alpha("white", 3/4), color = 'black'),
+              legend.direction = "horizontal",
+              legend.text = element_text(size = 7),
+              legend.title = element_text(size = 9),
+              legend.margin = margin(6, 10, 6, 10))
+
+Fig_impactofinfcov
+ggsave(file.path(outdir, "Figure_impactofinfcov.pdf"), width = 15, height = 4.25)
+```
+
 # Session information

### 2018-07-11 manuscript_figures/manuscript_figures.Rmd
Tweak color scale

@@
                              colours = c("darkblue","white","darkred"), 
-                             values = scales::rescale(c(-.1,0,1)),
-                             guide = "colorbar", limits=c(-.05,0.99)) +
+                             values = scales::rescale(c(-.05,0,1)),
+                             guide = "colorbar", limits=c(-.05,1)) +
         theme_bw() +

### 2018-07-12 manuscript_figures/manuscript_figures.Rmd
Legend-fu on Figure 4 to reduce white space

@@
 
-```{r Figure2}
+```{r Figure2, fig.width = 10, fig.height = 8}
 objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
@@
 
-```{r, Figure3}
+```{r, Figure3, fig.width = 8, fig.height = 4}
 # Figure 3: TPR of methods using by pieces of information used 
@@
 
-```{r, Figure 4}
+```{r, Figure 4, fig.width = 9, fig.height = 7}
 library(SummarizedBenchmark)
@@
 
-Fig4.rowOrder <- c("bonf", "bh", "ihw", "qvalue", "bl", "lfdr",
-                   "ashq")
-Fig4 <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
-                        annotate = "propPossible")
-Fig4
-
 # not excluding any methods
@@
 Fig4All <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
-                        excludeMethods = NULL, rowOrder = Fig4All.rowOrder)
+                        excludeMethods = NULL, rowOrder = Fig4All.rowOrder) +
+        theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
+              legend.position = c(0.96, 0.93), legend.justification = c(1, 1),
+              legend.background = element_rect(fill = scales::alpha("white", 3/4), color = 'black'),
+              legend.direction = "horizontal",
+              legend.text = element_text(size = 7),
+              legend.title = element_text(size = 9),
+              legend.margin = margin(6, 10, 6, 10))
+
 Fig4All
@@
 
-```{r}
+```{r, fig.width = 14, fig.height = 5}
 objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
@@
                          midpoint = alpha) +
-  labs(fill = "Mean FDR")
+          labs(fill = "Mean FDR") +
+          theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
+              legend.justification = c(1, 0.5),
+              legend.text = element_text(size = 7),
+              legend.title = element_text(size = 8),
+              legend.margin = margin(1, 1, 1, 1)) +
+          ylab("") + xlab("") +  
+          theme(plot.margin=unit(c(5.5, 5.5, 30, 5.5), "points"))
 
@@
                 rowOrder = Fig4All.rowOrder,
-                annotate = "TPR", excludeMethods = NULL)
+                annotate = "TPR", excludeMethods = NULL) +
+          theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
+              legend.justification = c(1, 0.5),
+              legend.text = element_text(size = 7),
+              legend.title = element_text(size = 8),
+              legend.margin = margin(1, 1, 1, 1)) +
+          ylab("") + xlab("") +  
+          theme(plot.margin=unit(c(5.5, 5.5, 30, 5.5), "points"))
 
 ggdraw() +
-  draw_plot(pfdr + ylab("") + xlab(""), 0.57, 0, 0.22, 1) +
-  draw_plot(ptpr + ylab("") + xlab(""), 0.78, 0, 0.22, 1) +
-  draw_plot(Fig4All, 0, 0, 0.60, 1) +
-  draw_plot_label(c("A", "B"), c(0, 0.57), c(1, 1), size = 15)
-ggsave(file.path(outdir, "Figure4.pdf"), width=14, height=5)
+  draw_plot(pfdr, 0.51, 0, 0.24, 1) +
+  draw_plot(ptpr, 0.75, 0, 0.24, 1) +
+  draw_plot(Fig4All, 0, 0, 0.5, 1) +
+  draw_plot_label(c("A", "B"), c(0, 0.51), c(1, 1), size = 15) +
+  draw_plot_label("Simulation", 0.72, 0.06, size =11, fontface = "plain")
+ggsave(file.path(outdir, "Figure4.pdf"), width=11, height=5)
 
@@
 
-```{r}
+```{r, fig.width = 6, fig.height = 4}
 library(ggbeeswarm)
@@
 
-```{r, Figure 5}
+```{r, Figure 5, fig.width = 9, fig.height = 7}
 
@@
 
-```{r Figure2.5}
+```{r Figure2.5,  fig.width = 10, fig.height = 4}
 objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
@@
 
-```{r FigureS7}
+```{r FigureS7,  fig.width = 10, fig.height = 7}
 objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
@@
 
-```{r}
+```{r, fig.width = 5, fig.height = 4.5}
 # by covariate
@@
 
-```{r}
+```{r,  fig.width = 6, fig.height = 4.5}
 #by dataset
@@
 
-```{r}
+```{r, fig.width = 5, fig.height = 5}
 wcs <- which(grepl("ChIPseq", casestudy))
@@
 ### RNA-seq
-```{r}
+```{r, fig.width = 6, fig.height = 5}
 # by dataset
@@
 
-```{r}
+```{r, fig.width = 15, fig.height = 6}
 wcs <- which(grepl("scRNAseq", casestudy))
@@
 
-```{r}
+```{r, fig.width = 14, fig.height = 6.5}
 # by dataset, excluding otu when genus is available
@@
 
-```{r}
+```{r, fig.width = 15, fig.height = 4.25}
 objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )

### 2018-07-12 datasets/GSEA/gsea-human.Rmd
Expand y-axis on pval histograms

@@
 dplyr::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
-    strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=10)
+    strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=18)
 ```

### 2018-07-12 datasets/GSEA/gsea-mouse.Rmd
Expand y-axis on pval histograms

@@
 dplyr::filter(goRes, numDEInCat > 0, numInCat > 10) %>%
-    strat_hist(pval = "over_represented_pvalue", covariate = "numInCat", maxy=10)
+    strat_hist(pval = "over_represented_pvalue", covariate = "numInCat", maxy=35)
 ```

### 2018-07-12 datasets/R/strat_hist.R
Make hist plotter more flexible

@@ -19,13 +19,14 @@
 #'  split on the covariate. Will result in `numQ` + 1 histograms generated. This
 #'  must be less than 6.
 #' @param main a character string that will be used as the main plot title
+#' @param axislabs a logical whether to include x and y axis labels on each panel
 #'        
 #' @return a cowplot object 
 #' 
 #' 
 #' @author Keegan Korthauer     
 strat_hist <- function(dat, pvalue, covariate, binwidth=0.025, maxy=3, numQ=3,
-                       main = ""){
+                       main = "", axislabs = TRUE){
 
   library(cowplot)
   library(ggplot2)
@@ -42,32 +43,39 @@ strat_hist <- function(dat, pvalue, covariate, binwidth=0.025, maxy=3, numQ=3,
     stop("pvalue and covariate must be variables that define column names in dat")
   }
   
-  plotOne <- function(d, pvalue, covariate, title=""){
-    ggplot(d, aes(x=get(pvalue))) + 
+  plotOne <- function(d, pvalue, covariate, title="", axislabs){
+    p<- ggplot(d, aes(x=get(pvalue))) + 
       geom_histogram(binwidth = binwidth, boundary = 0, 
                    colour="grey", fill="lightgrey") +
       aes(y=..density..)+
       theme_classic() +
-      theme(axis.title = element_text(face="bold"),
-          plot.title = element_text(face="bold")) +
-      scale_x_continuous(expand = c(0.02, 0)) + 
+      theme(plot.title = element_text(size = 11),
+            axis.title = element_text(size = 11)) +
+      xlim(0,1) +
       scale_y_continuous(expand = c(0.02, 0)) +
       coord_cartesian(ylim=c(0, maxy)) + 
-      xlab("p-value") +
-      ylab("Density") +
-      ggtitle(title)
+      xlab("") +
+      ylab("") +
+      ggtitle(title) 
+    
+    if (axislabs)
+      p <- p + xlab("p-value") + ylab("Density")
+    
+    p
   }
   
   gglist <- vector("list", numQ+1)
   
   for (q in 1:(numQ+1)){
     if(q==(numQ+1)){
-      gglist[[q]] <- plotOne( dat, pvalue=pvalue, covariate=covariate, title="All" )
+      gglist[[q]] <- plotOne( dat, pvalue=pvalue, covariate=covariate, title="All",
+                              axislabs = axislabs)
     }else{
       dat.strat <- dat %>% 
[…]

### 2018-07-12 manuscript_figures/manuscript_figures.Rmd
Add figure 6 - pvalue histograms

@@
 
+# RNAseq
 objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
@@
 
+# scRNAseq
 resfile <- objects[grepl("RNAseq", objects) & grepl("human", objects) &
@@
 
+# GSEA
 resfile <- objects[grepl("GSEA", objects) & grepl("mouse", objects) & !grepl("uninf", objects)] 
@@
 
+# GWAS
 resfile <- objects[grepl("GWAS", objects) & grepl("maf", objects)] 
@@
                         nbins=25) +
-  ggtitle("GWAS case study")+
+  ggtitle("GWAS case study") +
   xlab("Minor Allele Frequency (MAF) percentile") +
@@
 
+We'll do so for the same 4 case studies as in figure 5.
+
 ```{r, Figure 6}
-# TO DO
+# RNAseq 
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+resfile <- objects[grepl("RNAseq", objects) & grepl("brain", objects) & !grepl("uninf", objects)] 
+sb <- readRDS(file=resfile)
+res <- data.frame(rowData(sb)) %>%
+  dplyr::mutate(pval = assays(sb)$bench[,"unadjusted"])
+pa <- strat_hist(res, pvalue = "pval",
+           covariate = "ind_covariate", maxy = 10, axislabs = FALSE) +
+  draw_plot_label("  RNA-seq case study: Brain", 0.5, 1, hjust = 0.5,fontface = "bold", 
+                  size = 12) +
+  draw_plot_label("Overall mean expression pvalue", size = 11, 0.5, 0.14, fontface = "plain",
+                  hjust =0.5) +
+  draw_plot_label("Density", size = 11, 0, 0.36, fontface = "plain", angle=90)
+
+# scRNAseq
+resfile <- objects[grepl("RNAseq", objects) & grepl("human", objects) &
+                   grepl("mast", objects) & grepl("det", objects) & !grepl("uninf", objects)] 
+sb <- readRDS(file=resfile)
+
+res <- data.frame(rowData(sb)) %>%
+  dplyr::mutate(pval = assays(sb)$bench[,"unadjusted"]) 
+pb <- strat_hist(res, pvalue = "pval",
+           covariate = "detection", maxy = 12, axislabs = FALSE)+
+  draw_plot_label("  scRNA-seq case study: Human, MAST",0.5, 1, hjust = 0.5, fontface = "bold",
+                  size = 12 ) +
+  draw_plot_label("Detection rate pvalue", size = 11, 0.5, 0.14, fontface = "plain",
+                  hjust =0.5) +
+  draw_plot_label("Density", size = 11, 0, 0.36, fontface = "plain", angle=90)
+
+# GSEA
+resfile <- objects[grepl("GSEA", objects) & grepl("mouse", objects) & !grepl("uninf", objects)] 
+sb <- readRDS(file=resfile)
+res <- data.frame(rowData(sb)) %>%
+  dplyr::mutate(pval = assays(sb)$bench[,"unadjusted"])
+
+pc <- strat_hist(res, pvalue = "pval",
+           covariate = "ind_covariate", maxy = 35, axislabs = FALSE) +
+  draw_plot_label("  GSEA case study: Mouse", 0.5, 1, hjust = 0.5, fontface = "bold" , size = 12) +
+  draw_plot_label("Gene Set size pvalue", size = 11, 0.5, 0.14, fontface = "plain",
+                  hjust =0.5) +
+  draw_plot_label("Density", size = 11, 0, 0.36, fontface = "plain", angle=90)
+
+# GWAS
+resfile <- objects[grepl("GWAS", objects) & grepl("maf", objects)] 
+sb <- readRDS(file=resfile)
+res <- data.frame(rowData(sb)) %>%
+  dplyr::mutate(pval = assays(sb)$bench[,"unadjusted"])
+
+pd <- strat_hist(res, pvalue = "pval",
+           covariate = "ind_covar_AF", maxy = 3, axislabs = FALSE) +
+  draw_plot_label("GWAS case study", 0.5, 1, fontface = "bold", hjust = 0.5, size = 12) +
+  draw_plot_label("Minor Allele Frequency (MAF) pvalue", size = 11, 0.5, 0.14, fontface = "plain", 
+                  hjust = 0.5) +
+  draw_plot_label("Density", size = 11, 0, 0.36, fontface = "plain", angle=90)
+
+p6 <- plot_grid(pa, pb, pc, pd, ncol=1, labels = LETTERS[1:4])
+p6
+ggsave(file.path(outdir, "Figure6.pdf"), width=7, height=7)
+
 ```

### 2018-07-13 manuscript_figures/manuscript_figures.Rmd
Switch from geom_raster to tile to avoid Preview smearing colors

@@
 Fig_impactofinfcov <- ggplot(dftab, aes(x = dataset, y = method, fill = diff_inf_percent)) + 
-        geom_raster() + 
+        geom_tile() + 
         scale_fill_gradientn("% Increase\nin rejections",

### 2018-07-13 datasets/R/plotMethodRank.R
Switch from geom_raster to tile to avoid Preview smearing colors

@@ -163,7 +163,7 @@ plotMethodRanks <- function(objects, colLabels, alpha = 0.10,
       
       # heatmap : rows method, cols casestudy
       Fig <- ggplot(ranks_avg, aes(x = studyname, y = method, fill = mean_prop)) + 
-        geom_raster() + 
+        geom_tile() + 
         scale_fill_distiller("% Rejected\n(relative to max)",
                              palette = "Blues", direction = 1, limits = c(0, 1),
                              labels = scales::percent) +
@@ -182,7 +182,7 @@ plotMethodRanks <- function(objects, colLabels, alpha = 0.10,
                                  levels=unique(ranks_avg$method[order(ranks_avg$meta_rank)]))
       
       Fig <- ggplot(ranks_avg, aes(x = studyname, y = method, fill = mean_rank)) + 
-        geom_raster() + 
+        geom_tile() + 
         scale_fill_distiller("Mean Rank", palette = "Blues", direction = 1) +
         theme_bw() +
         xlab(xlab) +
@@ -198,7 +198,7 @@ plotMethodRanks <- function(objects, colLabels, alpha = 0.10,
                                  levels=unique(ranks_avg$method[order(ranks_avg$meta_nrej)]))
       
       Fig <- ggplot(ranks_avg, aes(x = studyname, y = method, fill = mean_nrej)) + 
-        geom_raster() + 
+        geom_tile() + 
         scale_fill_distiller(paste0("Mean ", fill), palette = "Blues", direction = 1) +
         theme_bw() +
         xlab(xlab) +

### 2018-07-13 manuscript_figures/manuscript_figures.Rmd
Annotate Fig-infCovariate in case study with max vals

@@
 
+maxVals <- dftab %>% 
+  group_by(dataset,cov) %>%
+  summarize(top = max(diff_inf_percent, na.rm=TRUE))
+
+dftab <- left_join(dftab, maxVals, by = c("dataset", "cov")) %>%
+  mutate(top = ifelse(diff_inf_percent == top, diff_inf_percent, NA))
+
 Fig_impactofinfcov <- ggplot(dftab, aes(x = dataset, y = method, fill = diff_inf_percent)) + 
@@
               legend.title = element_text(size = 9),
-              legend.margin = margin(6, 10, 6, 10))
+              legend.margin = margin(6, 10, 6, 10)) +
+        geom_text(aes(label = ifelse(is.na(top), "", scales::percent(top))),
+                  color = "black", size = 2)
 

### 2018-07-16 datasets/simulations/simulations-informative-cubic.Rmd
Lower signal strength in sims to get TPR closer to real yeast sims

@@
 ```{r gauss-parameters}
-tstat <- rnorm_generator(4)       # functional: dist of alternative test stats
+tstat <- rnorm_generator(3)       # functional: dist of alternative test stats
 tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
@@
 ```{r t11-parameters}
-tstat <- rnorm_generator(6)    # functional: dist of alternative test stats
+tstat <- rnorm_generator(3)     # functional: dist of alternative test stats
 tstat_dist <- rt_perturber(11)  # functional: sampling dist/noise for test stats

### 2018-07-17 datasets/R/plotMethodRank.R
Add replicate variable to tidy_df()

@@ -285,6 +285,10 @@ tidy_df <- function(objects, colLabels, fill, annotate, alpha){
                                            nrejects / max(nrejects, na.rm = TRUE)),
                        propPossible = nrejects / nrow(x)) %>%
                 mutate(casestudy = colLabels[i])
+            if ( class(objects[[i]]) == "character") {
+             tmp <- tmp %>%
+               mutate(replicate = i)
+            }
             
             ranks <- rbind(ranks, tmp)
         }

### 2018-07-17 datasets/RNA-Seq/polyester-simulation.Rmd
Use the same dataset for informative and uninformative covariates within replicate

--- 2018-05-15 datasets/RNA-Seq/polyester-simulation.Rmd
@@
 #' @param bd is the bench design object
-simulateOneSplit <- function(X, counts, paramPolyester, 
-                             nDE, sampleSize, bd){
+#' @param uninformativeCovariate logical indicating whether to replace overall mean 
+#'  independent covariate with a randomly generated (uninformative) covariate. 
+#'  Default is FALSE.
+#' @param rseed random seed
+simulateOneSplit <- function(X, rseed, counts, paramPolyester, 
+                             nDE, sampleSize, bd, uninformativeCovariate = FALSE){
+  if(uninformativeCovariate)
+    uninf_covar <- rnorm(nrow(counts))
+
+  # set random seed
+  set.seed(as.numeric(X)*as.numeric(rseed))
 
@@
     log2FC <- rep(0, nrow(counts))
-    # log2FC[DE] <- rtruncnorm(nDE, 1, sd=.5) * sample(c(1,-1), size=nDE, replace=TRUE)
-    log2FC[DE] <- c(rnorm(nDE, sd = 1), rep(0,nrow(counts)-nDE))
+    log2FC[DE] <- rnorm(nDE, sd = 1)
     dat_alt = create_read_numbers(paramPolyester$mu, paramPolyester$fit,
@@
                                pval=res_alt$pvalue, SE=res_alt$lfcSE, 
-                               ind_covariate = res_alt$baseMean, 
+                               ind_covariate = ifelse(uninformativeCovariate, 
+                                                      uninf_covar,
+                                                      res_alt$baseMean), 
                                effect_size=res_alt$log2FoldChange, 
@@
 nDE <- 0
+rseed <- 225
 
 if (!file.exists(resfile_n5)){
-  null5 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
+  null5 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, counts=counts,
                    paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd)
@@
 nDE <- 0
+rseed <- 837
 
 if (!file.exists(resfile_n10)){
-  null10 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
+  null10 <- bplapply(X=1:B, rseed=837, FUN=simulateOneSplit, counts=counts,
                    paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd)
@@
 nDE <- 1500
+rseed <- 198
 
 if (!file.exists(resfile_d5)){
-  de5 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
+  de5 <- bplapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, counts=counts,
                    paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd)
@@
 nDE <- 1500
+rseed <- 961
 
 if (!file.exists(resfile_d10)){
-  de10 <- bplapply(X=1:B, FUN=simulateOneSplit, counts=counts,
+  de10 <- bplapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, counts=counts,
                    paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd)
@@
 
+# Uninformative Covariate
+
+## Sim N5: Null Comparison 5v5
+
+Here we run a null comparison 5 versus 5 RNA-Seq samples. 
+This will be done for 100 simulations.
+
+### Generate a list of SB results objects
+
+```{r, results='hide', message=FALSE}
+sampleSize <- 5
+nDE <- 0
+rseed <- 225
+
+if (!file.exists(resfile_n5_uninfCov)){
+  null5 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, counts=counts,
+                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, 
+                   bd=bd, uninformativeCovariate = TRUE)
+  saveRDS(null5, file=resfile_n5_uninfCov)
+} else {
+  null5 <- readRDS(file=resfile_n5_uninfCov)
+}
+```
+
+### Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(null5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res5 <- plotsim_standardize(null5, alpha = seq(0.01, 0.10, 0.01))
+
+# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
+plotsim_average(res5, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5, met="TNR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(null5, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(null5, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res5, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+  aggupset(null5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+  message("Not enough methods found rejections at alpha ", ualpha, 
+          "; skipping upset plot")
+}
+
+```
+
+## Sim N10: Null Comparison 10v10
+
+Here we'll repeat the above, but for a null comparison 10 versus 10 samples. 
+This will be done for 100 simulations.
+
+### Generate a list of SB results objects
+
+```{r, results='hide', message=FALSE}
+sampleSize <- 10
+nDE <- 0
+rseed <- 837
+
+if (!file.exists(resfile_n10_uninfCov)){
+  null10 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, counts=counts,
+                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, 
+                   bd=bd, uninformativeCovariate = TRUE)
+  saveRDS(null10, file=resfile_n10_uninfCov)
+} else {
+  null10 <- readRDS(file=resfile_n10_uninfCov)
+}
+```
+
+### Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(null10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res10 <- plotsim_standardize(null10, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(res10, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10, met="TNR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(null10, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(null10, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res10, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+  aggupset(null10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
+}
+
+```
+
+## Sim D5: DE Comparison 5v5
+
+Here we'll repeat the above, but for a DE comparison 5 versus 5 samples
+and 1500 DE genes are added. This will be done for 100 simulations.
+
+### Generate a list of SB results objects
+
+```{r, results='hide', message=FALSE}
+sampleSize <- 5
+nDE <- 1500
+rseed <- 198
+
+if (!file.exists(resfile_d5_uninfCov)){
+  de5 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, counts=counts,
+                  paramPolyester=params, nDE=nDE, sampleSize=sampleSize, 
+                  bd=bd, uninformativeCovariate = TRUE)
+  saveRDS(de5, file=resfile_d5_uninfCov)
+} else {
+  de5 <- readRDS(file=resfile_d5_uninfCov)
+}
+```
+
+### Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.20, 0.01))
+
+plotsim_average(res5d, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5d, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5d, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(de5, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(de5, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+  aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
+}
+
+```
+
+## Sim D10: DE Comparison 10v10
+
+Here we'll repeat the above, but for a DE comparison 10 versus 10 samples
+and 1500 DE genes are added. This will be done for 100 simulations.
+
+### Generate a list of SB results objects
+
+```{r, results='hide', message=FALSE}
+sampleSize <- 10
+nDE <- 1500
+rseed <- 961
+
+if (!file.exists(resfile_d10_uninfCov)){
+  de10 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, counts=counts,
+                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, 
+                   bd=bd, uninformativeCovariate = TRUE)
+  saveRDS(de10, file=resfile_d10_uninfCov)
+} else {
+  de10 <- readRDS(file=resfile_d10_uninfCov)
+}
+
+```
+
+### Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.20, 0.01))
+
+plotsim_average(res10d, met="rejections",filter_set = excludeSet, 
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10d, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10d, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(de10, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(de10, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res10d, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+  aggupset(de10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
+}
+```
+
+## Sample size comparison
+
+Here we compare the method ranks for the different sample sizes at alpha = 0.10.
+
+```{r}
+# uninformative covariate
+plotMethodRanks(c(resfile_d5_uninfCov, resfile_d10_uninfCov), 
+                colLabels = c("DE 5", "DE 10"), 
+                alpha = 0.10, xlab = "Comparison",
+                excludeMethods = NULL)
+```
+
 # Session information

### 2018-07-17 datasets/RNA-Seq/yeast-simulation.Rmd
Use the same dataset for informative and uninformative covariates within replicate

@@
                              bd, uninformativeCovariate = FALSE, pvalHists = FALSE){
+  if(uninformativeCovariate)
+    ind_covar <- rnorm(nrow(dds))
 
@@
                                SE=resTEST$lfcSE,                 
-                               ind_covariate = resTEST$baseMean,
+                               ind_covariate = ifelse(uninformativeCovariate, 
+                                                      ind_covar,
+                                                      resTEST$baseMean),
                                effect_size = resTEST$log2FoldChange, 
@@
                 dplyr::filter(!is.na(pval))
-
-  if(uninformativeCovariate)
-    geneExp$ind_covariate <- rnorm(nrow(geneExp))
 

### 2018-07-17 manuscript_figures/manuscript_figures.Rmd
Add draft prototype figure of uninformative covariate impact in Yeast and Polyester *within replicate*

@@
 
+# Informativeness of covariates in Yeast simulation
+
+As an alternative to the line plot above (Figure S7), we make a heatmap plot that
+displays the mean difference within simulation replicate.
+
+```{r}
+
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+resfile <- objects[grepl("yeast", objects) ]
+
+# methods that use a covariate
+cov.methods <- c("ihw", "bl", "lfdr", "scott-empirical", "scott-theoretical")
+
+# regexp to construct master data table separating case study datasets, methods
+# and covariates
+obj.names <- str_split(resfile, pattern = "/")
+casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
+filename <- gsub(".rds", "", unlist(lapply(obj.names, function(x) x[4] )))
+dataset <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[3]]))
+inform <-  ifelse(grepl("uninf", resfile), "uninf", "inf")
+type <-  ifelse(grepl("null", resfile), "null", "de")
+size <-  ifelse(grepl("10", resfile), "10", "5")
+
+tab <- data.frame()
+for (l in seq_along(resfile)){
+  x <- readRDS(resfile[l])
+
+  tmp2 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
+                 fill =  "FDR", annotate = NULL, alpha = 0.10) %>%
+    mutate(fdr = nrejects) %>%
+    select(method, casestudy, replicate, fdr)
+  tmp3 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
+                  fill =  "TNR", annotate = NULL, alpha = 0.10) %>%
+    mutate(tnr = nrejects) %>%
+  select(method, casestudy, replicate, tnr)
+
+  if (type[l] != "null"){
+    tmp1 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
+                    fill =  "TPR", annotate = NULL, alpha = 0.10) %>%
+      mutate(tpr = mean(nrejects)) %>%
+    select(method, casestudy, replicate, tpr)
+
+    tmp <- left_join(tmp1, tmp2, by = c("method", "casestudy", "replicate")) %>%
+      left_join(tmp3, by = c("method", "casestudy", "replicate")) %>%
+      mutate(inform = inform[l],
+             type = type[l],
+             size = size[l])
+  }else{
+    tmp <- left_join(tmp2, tmp3, by = c("method", "casestudy", "replicate")) %>%
+      mutate(inform = inform[l],
+             type = type[l],
+             size = size[l])
+  }
+
+  tab <- bind_rows(tab, tmp)
+}
+
+# take difference within replicate, then average differences across
+# type and size and method
+
+dftab <- tab %>%
+  filter(method %in% cov.methods) %>%
+  dplyr::mutate(id = paste(type, size, method, replicate, sep = "-"))
+
+uninftab <- dftab %>% 
+  filter(inform == "uninf") %>%
+  select(id, fdr, tpr, tnr) %>%
+  rename(fdr_uninf = fdr,
+         tnr_uninf = tnr,
+         tpr_uninf = tpr)
+
+inftab <- dftab %>%
+  filter(inform == "inf") %>%
+  select(-inform)
+
+rowOrder <- c("ihw", "bl", "lfdr", "scott-empirical", "scott-theoretical")
+
+dftab <- left_join(inftab, uninftab, by = "id") %>%
+  mutate(tpr_diff = tpr-tpr_uninf,
+         fdr_diff = fdr-fdr_uninf,
+         tnr_diff = tnr-tnr_uninf) %>%
+  group_by(type, size, method) %>%
+  summarize(tpr_diff = mean(tpr_diff),
+            fdr_diff = mean(fdr_diff),
+            tnr_diff = mean(tnr_diff)) %>%
+  mutate(method = factor(method, levels = rowOrder))
+
+plot_fun <- function(var){
+
+  plot  <- ggplot(dftab %>% filter(!is.na(get(var))), 
+                  aes(x = interaction(type, size), 
+                                        y = method)) + 
+    geom_tile(aes(fill=get(var))) + 
+    scale_fill_distiller(paste0("Change\nin ", toupper(var)),
+                         palette = "Blues", direction = 1,
+                         labels = scales::percent) +
+    theme_bw() +
+    xlab("Experiment") +
+    ylab("Method") +
+    facet_grid( ~ type + size, scales = "free", space = "free") +
+    ggtitle(paste0("Impact of informative covariate on ", toupper(var))) +
+    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
+        legend.justification = c(1, 1),
+        legend.background = element_rect(fill = scales::alpha("white", 3/4), color = 'black'),
+        legend.text = element_text(size = 7),
+        legend.title = element_text(size = 9),
+        legend.margin = margin(6, 10, 6, 10)) +
+    geom_text(aes(label = scales::percent(get(var))),
+            color = "black", size = 3)
+
+  return(plot)
+}
+
+plot_fun("tpr_diff")
+plot_fun("fdr_diff")
+plot_fun("tnr_diff")
+```
+
 # Session information

### 2018-07-17 datasets/simulations/simulations-informative-cubic.Rmd
Specify 'mode' parameter for ASH in chi-sq simulations

@@
 
+For the chi-squared distributed noise setting, we must change the "mode" setting for ASH
+from the default of 0 to `"estimate"` since the mode of null and alternative test statistics
+are no longer centered at 0. While both approximate Normality and unimodality of effects are
+violated in this simulation setting, by allowing the mode to be estimated, rather than forced
+to 0, should return more comparable results.
+
+```{r}
+bdchi <- modifyBMethod(bd, "ashq", mode = "empirical")
+```
+
 We next run the simulations. We do not include FDR Regression because the test statistics
@@
 } else {
-    res <- mclapply(X = 1:B, FUN = simIteration, bench = bd, m = m,
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdchi, m = m,
                     pi0 = pi0, tstat = tstat, icovariate = icovariate,
@@
 ```{r chisq4-one-simulation}
-onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdchi, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)

### 2018-07-17 datasets/simulations/simulations-varyingeffectsize-chisq.Rmd
Include Session Info at bottom of simulations
Specify 'mode' parameter for ASH in chi-sq simulations

@@
 bd <- initializeBenchDesign()
+```
+
+For the chi-squared distributed noise setting, we must change the "mode" setting for ASH
+from the default of 0 to `"estimate"` since the mode of null and alternative test statistics
+are no longer centered at 0. While both approximate Normality and unimodality of effects are
+violated in this simulation setting, by allowing the mode to be estimated, rather than forced
+to 0, should return more comparable results.
+
+```{r}
+bd <- modifyBMethod(bd, "ashq", mode = "empirical")
 ```

### 2018-07-18 datasets/RNA-Seq/polyester-simulation.Rmd
Standardize number of DE genes

@@
                                pval=res_alt$pvalue, SE=res_alt$lfcSE, 
-                               ind_covariate = ifelse(uninformativeCovariate, 
-                                                      uninf_covar,
-                                                      res_alt$baseMean), 
+                               ind_covariate = res_alt$baseMean, 
                                effect_size=res_alt$log2FoldChange, 
@@
                                  ncol(counts(dds_alt))  ))
+
+  if (uninformativeCovariate)
+    geneExp <- mutate(geneExp, ind_covariate = uninf_covar)
 
@@
 Here we'll repeat the above, but for a DE comparison 5 versus 5 samples
-and 1500 DE genes are added. This will be done for 100 simulations.
+and 2000 DE genes are added. This will be done for 100 simulations.
 
@@
 sampleSize <- 5
-nDE <- 1500
+nDE <- 2000
 rseed <- 198
@@
 Here we'll repeat the above, but for a DE comparison 10 versus 10 samples
-and 1500 DE genes are added. This will be done for 100 simulations.
+and 2000 DE genes are added. This will be done for 100 simulations.
 
@@
 sampleSize <- 10
-nDE <- 1500
+nDE <- 2000
 rseed <- 961

### 2018-07-18 manuscript_figures/manuscript_figures.Rmd
Update figure on impact of covariate in yeast/polyester

--- 2018-07-13 manuscript_figures/manuscript_figures.Rmd
@@
 
+# Informativeness of covariates in Yeast & Polyester simulation
+
+As an alternative to the line plot above (Figure S7), we make a heatmap plot that
+displays the mean difference within simulation replicate.
+
+```{r}
+
+for (org in c("yeast", "polyester")){
+
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+resfile <- objects[grepl(org, objects) ]
+
+# methods that use a covariate
+cov.methods <- c("ihw", "bl", "lfdr", "scott-empirical", "scott-theoretical")
+
+# regexp to construct master data table separating case study datasets, methods
+# and covariates
+obj.names <- str_split(resfile, pattern = "/")
+casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
+filename <- gsub(".rds", "", unlist(lapply(obj.names, function(x) x[4] )))
+dataset <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[3]]))
+inform <-  ifelse(grepl("uninf", resfile), "uninf", "inf")
+type <-  ifelse(grepl("null", resfile), "null", "de")
+size <-  ifelse(grepl("10", resfile), "10", "5")
+
+tab <- data.frame()
+for (l in seq_along(resfile)){
+  x <- readRDS(resfile[l])
+
+  tmp2 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
+                 fill =  "FDR", annotate = NULL, alpha = 0.10) %>%
+    mutate(fdr = nrejects) %>%
+    select(method, casestudy, replicate, fdr)
+  tmp3 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
+                  fill =  "TNR", annotate = NULL, alpha = 0.10) %>%
+    mutate(tnr = nrejects) %>%
+  select(method, casestudy, replicate, tnr)
+
+  if (type[l] != "null"){
+    tmp1 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
+                    fill =  "TPR", annotate = NULL, alpha = 0.10) %>%
+      mutate(tpr = nrejects) %>%
+    select(method, casestudy, replicate, tpr)
+
+    tmp <- left_join(tmp1, tmp2, by = c("method", "casestudy", "replicate")) %>%
+      left_join(tmp3, by = c("method", "casestudy", "replicate")) %>%
+      mutate(inform = inform[l],
+             type = type[l],
+             size = size[l])
+  }else{
+    tmp <- left_join(tmp2, tmp3, by = c("method", "casestudy", "replicate")) %>%
+      mutate(inform = inform[l],
+             type = type[l],
+             size = size[l])
+  }
+
+  tab <- bind_rows(tab, tmp)
+}
+
+# take difference within replicate, then average differences across
+# type and size and method
+
+dftab <- tab %>%
+  filter(method %in% cov.methods) %>%
+  dplyr::mutate(id = paste(type, size, method, replicate, sep = "-"))
+
+uninftab <- dftab %>% 
+  filter(inform == "uninf") %>%
+  select(id, fdr, tpr, tnr) %>%
+  rename(fdr = "fdr_uninf",
+         tnr = "tnr_uninf",
+         tpr = "tpr_uninf")
+
+inftab <- dftab %>%
+  filter(inform == "inf") %>%
+  select(-inform)
+
+rowOrder <- c("ihw", "bl", "lfdr", "scott-empirical", "scott-theoretical")
+
+dftab <- left_join(inftab, uninftab, by = "id") %>%
+  mutate(tpr_diff = tpr-tpr_uninf,
+         fdr_diff = fdr-fdr_uninf,
+         tnr_diff = tnr-tnr_uninf) %>%
+  group_by(type, size, method) %>%
+  summarize(tpr_se = sqrt(sd(tpr_diff)/n()),
+            fdr_se = sqrt(sd(fdr_diff)/n()),
+            tnr_se = sqrt(sd(tnr_diff)/n()),
+            tpr_diff = mean(tpr_diff),
+            fdr_diff = mean(fdr_diff),
+            tnr_diff = mean(tnr_diff)) %>%
+  mutate(method = factor(method, levels = rowOrder)) %>%
+  mutate(fdr_diff = ifelse(type == "null", NA, fdr_diff),
+         tpr_diff = ifelse(type == "null", NA, tpr_diff),
+         fdr_se = ifelse(type == "null", NA, fdr_se),
+         tpr_se = ifelse(type == "null", NA, tpr_se))
+
+plot_fun <- function(var){
+
+  plot  <- ggplot(dftab %>% filter(!is.na(get(var))), 
+                  aes(x = interaction(type, size), 
+                                        y = method)) + 
+    geom_tile(aes(fill=get(var))) + 
+    scale_fill_gradient2(paste0("Change\nin ", toupper(gsub("_diff", "", var))),
+                         low = "red", midpoint = 0, mid = "white", high = "blue",
+                         labels = scales::percent) +
+    theme_bw() +
+    xlab("Experiment") +
+    ylab("Method") +
+    facet_grid( ~ type + size, scales = "free", space = "free") +
+    ggtitle(paste0("Impact of informative covariate on ", toupper(gsub("_diff", "", var)))) +
+    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
+    geom_text(aes(label = scales::percent(get(var))),
+            color = "black", size = 3)
+
+  return(plot)
+}
+
+# arrange in multipanel figure?
+plot_fun("tpr_diff")
+ggsave(file.path(outdir, paste0("Figure_infCov_tpr_", org, ".pdf")), height = 5, width = 5)
+
+plot_fun("fdr_diff")
+ggsave(file.path(outdir, paste0("Figure_infCov_fpr_", org, ".pdf")), height = 5, width = 5)
+
+plot_fun("tnr_diff")
+ggsave(file.path(outdir, paste0("Figure_infCov_tnr_", org, ".pdf")), height = 5, width = 7)
+
+}
+
+```
+
 # Session information

### 2018-07-19 datasets/RNA-Seq/yeast-simulation.Rmd
Adjust nmids and densknots for scott-empirical

--- 2018-06-05 datasets/RNA-Seq/yeast-simulation.Rmd
@@
                      nulltype = 'empirical',
-                     control = list(lambda = 0.01))
+                     control = list(lambda = 0.01, nmids = 300, densknots = 20))
 ```
@@
                              bd, uninformativeCovariate = FALSE, pvalHists = FALSE){
+  if(uninformativeCovariate)
+    uninf_covar <- rnorm(nrow(dds))
 
@@
                                pzero = rowSums(counts(dds_test)==0)/
-                                       ncol(counts(dds_test)))) %>%
-                dplyr::filter(!is.na(pval))
+                                       ncol(counts(dds_test))))
 
-  if(uninformativeCovariate)
-    geneExp$ind_covariate <- rnorm(nrow(geneExp))
+  if (uninformativeCovariate)
+    geneExp <- mutate(geneExp, ind_covariate = uninf_covar)
+
+  geneExp <-  geneExp %>% dplyr::filter(!is.na(pval)) %>% dplyr::filter(pzero < 0.5)
 

### 2018-07-19 manuscript_figures/manuscript_figures_sims.Rmd
Redo all simulation figures to better match other figures

--- 2018-06-26 manuscript_figures/manuscript_figures_sims.Rmd
@@
 
-# set alpha cutoff for plots that are fixed at a certain value of alpha
-alpha <- 0.10
-
-methodset <- c("bonf", "bh", paste0("ihw-a", round(alpha * 100)),
-               "qvalue", "bl-df03", "lfdr",
+methodset <- c("bonf", "bh", "ihw", "qvalue", "bl-df03", "lfdr",
                "scott-empirical", "scott-theoretical", "ashq")
@@
 
-First, we load summaries for all simulations results.
+First, we load summaries for all simulations results at significance thresholds
+between 0.01 and 0.10. This process can take a long time (approximately 40 minutes
+per alpha cutoff) because of the large number of simulation settings that must be
+read in and summarized. Ideally, this initial step should be run externally separate
+from this Rmd (by default, the following code chunk will not be run).
 
-```{r}
+```{r, eval = FALSE}
+for (ia in 1:10) {
     summary_file <- paste0("allres_alpha", 100*alpha, ".rds")
-
-if (file.exists(summary_file)) {
-    res <- readRDS(summary_file)
-} else {
+    if (!file.exists(summary_file)) {
         res_files <- list.files(path, "\\.rds$", full.names = TRUE)
@@
             zz <- dplyr::select(zz, rep, blabel, key, alpha, value)
-
             zz <- dplyr::group_by(zz, blabel, key, alpha)
@@
     }
+}
+```
 
-## merge all simulation summaries
-res <- bind_rows(res, .id = "setting")
+Assuming that the summaries have been saved for all simulation settings at
+the various FDR thresholds using the code above, we load the summarized results
+and merge them into a single results table.
 
-## create shorter simulation setting labels
-res <- dplyr::mutate(res, dist = gsub(".*?-benchmark-(.*)", "\\1", setting),
+```{r}
+res <- vector("list", 10)
+for (ia in 1:10) {
+    imethodset <- methodset
+    imethodset[which(imethodset == "ihw")] <- paste0("ihw-a", sprintf("%02d", ia))
+    ires <- readRDS(paste0("allres_alpha", ia, ".rds"))
+    ires <- bind_rows(ires, .id = "setting")
+    ires <- dplyr::mutate(ires, dist = gsub(".*?-benchmark-(.*)", "\\1", setting),
                          setting = gsub("(.*?)-benchmark-.*", "\\1", setting))
+    ires <- dplyr::filter(ires, blabel %in% imethodset)
+    ires <- dplyr::mutate(ires, blabel = gsub("(-a)(.*)", "", blabel))
+    res[[ia]] <- ires
+}
+res <- bind_rows(res)
+res <- dplyr::rename(res, Method = blabel)
+res <- dplyr::mutate(res, Method = gsub("-df03", "", Method))
+res <- dplyr::mutate(res, Method = factor(Method, levels = gsub("-df03", "", methodset)))
+```
 
-## only use subset of methods for plotting
-res <- dplyr::filter(res, blabel %in% methodset)
-res <- dplyr::rename(res, Method = blabel)
+We clean up the annotations in the table for some of the more complex simulation settings.
 
-## remove suffixes for unused tuning parameters
-res <- dplyr::mutate(res, Method = gsub("-df03", "", Method),
-                     Method = gsub("(-a)(.*)", "", Method))
-methodset <- gsub("-df03", "", methodset)
-methodset <- gsub("(-a)(.*)", "", methodset)
+```{r}
+res <- dplyr::mutate(res, inform = ifelse(grepl("varyinginfo", setting), dist, NA))
+res <- dplyr::mutate(res, inform = as.numeric(gsub("level", "", inform)))
+res <- dplyr::mutate(res, dist = ifelse(grepl("varyinginfo", setting), "gaussian", dist))
 
-## add candy color scheme from datasets/R/colorPalette.R
-res <- dplyr::left_join(res, candycols, by = "Method")
+res <- dplyr::mutate(res, pi0 = ifelse(grepl("varyingpi0", setting), dist, 90))
+res <- dplyr::mutate(res, pi0 = as.numeric(gsub("nullprop", "", pi0)))
+res <- dplyr::mutate(res, dist = ifelse(grepl("varyingpi0", setting), "gaussian", dist))
 
-## treat methods as factors for forced ordering
-res <- dplyr::mutate(res, Method = factor(Method, levels = methodset))
+res <- dplyr::mutate(res, signal = ifelse(grepl("varyinges", setting), dist, NA))
+res <- dplyr::mutate(res, dist = ifelse(grepl("^varyinges$", setting), "gaussian", dist))
+res <- dplyr::mutate(res, dist = ifelse(grepl("^varyinges-t$", setting), "t11", dist))
+res <- dplyr::mutate(res, dist = ifelse(grepl("^varyinges-chisq$", setting), "chisq4", dist))
+res <- dplyr::mutate(res, setting = ifelse(grepl("varyinges", setting), "varyinges", setting))
 ```
+
+We also replace the sampling noise labels with more plotting-friendly labels.
+
+```{r}
+res <- dplyr::mutate(res, dist = factor(dist, levels = c("gaussian", "t11", "t5", "chisq4"),
+                                        labels = c(expression("Noise: " * N(0, 1)),
+                                                   expression("Noise: " * t[11]),
+                                                   expression("Noise: " * t[5]),
+                                                   expression("Noise: " * {chi^2}[4]))))
+```
+
+We use the standardize candy color scheme and line types for the plots.
+
+```{r}
+col <- as.character(candycols$col)
+names(col) <- as.character(candycols$Method)
+lty <- as.character(candycols$lty)
+names(lty) <- as.character(candycols$Method)
+```
+
+There are several different simulation settings that can be investigated.
+
+```{r}
+dplyr::distinct(res, setting)
+```
+
+Namely, simulation settings were run with to investigate the impact of different
+informative covariates, noninformative covariates, varying null proportions,
+varying covariate informativeness, and varying effect size distributions (including
+unimodal effects as described in the ASH manuscript). Many of these simulation
+settings were also run with different sampling noise distributions, including
+Normal/Gaussian, t with 11 degrees of freedom, t with 5 degrees of freedom, and
+Chi-squared with 4 degrees of freedom.
 
@@
 
-```{r}
-res_nullset <- dplyr::filter(res, setting == "null")
-res_nullset <- dplyr::mutate(res_nullset, dist = factor(dist, levels = c("gaussian", "t5", "chisq4")))
+For null simulations, rather than the false discovery rate, we take a look at the true negative
+rate (equivalently, the proportion of false positives) since the FDR will be 1 if any rejections
+are made. Additionally, we take a look at the FWER (the proportion of simulation replicates out
+of 100 with at least one rejection). In these plots, `ashq`, `scott-theoretical` and `scott-theoretical`
+are excluded for the Chi-Squared simulation settings.
+
+```{r, fig.width = 18, fig.height = 8}
+res_nullset <- dplyr::filter(res, setting == "null",
+                             !(Method == "ashq" & grepl("chi\\^2", dist)),
+                             !(grepl("scott", Method) & grepl("chi\\^2", dist)))
 res_nullset <- tidyr::complete(res_nullset, setting, Method, dist, key)
 
-col <- as.character(res_nullset$col)
-names(col) <- as.character(res_nullset$Method)
+null_tnr <- dplyr::filter(res_nullset, key == "TNR") %>%
+    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
+    scale_y_continuous("TNR", labels = scales::percent) +
+    ggtitle("Mean TNR Over 100 Replications") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+null_fwer <- dplyr::filter(res_nullset, key == "FWER") %>%
+    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
+    scale_y_continuous("FWER", labels = scales::percent) +
+    ggtitle("FWER Over 100 Replications") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+gp <- plot_grid(null_tnr + guides(color = FALSE, linetype = FALSE),
+                null_fwer + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
+                labels = LETTERS[1:2], ncol = 1)
+gp <- plot_grid(gp, get_legend(null_tnr), rel_widths = c(1, 1/16))
+gp
 ```
 
-For null simulations, rather than the false discovery rate, we take a look at the number
-false positives since the FDR will be 1 if any rejections are made.
+# Sampling Distributions
 
-In these plots, `ashq`, `scott-theoretical` and `scott-theoretical` are excluded for the Chi-Squared
-simulation settings.
+We just take a look at different noise distributions for a single informative covariate.
 
-```{r, fig.width = 8, fig.height = 5}
-dplyr::filter(res_nullset, key == "rejectprop") %>%
-    filter(! (Method == "ashq" & dist == "chisq4"),
-           ! (grepl("scott", Method) & dist == "chisq4")) %>%
-    ggplot(aes(x = Method, y = mean_value, color = Method)) +
-    geom_point() +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se)) + 
-    facet_grid(. ~ dist) +
-    scale_color_manual(values = col, guide = FALSE) +
-    scale_y_continuous("Average % Rejected Tests", breaks = seq(0, 1, .01), labels = scales::percent) +
-    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
-    ggtitle("Rejected proportion in null simulations (average over 100 replications)") 
+```{r, fig.width = 18, fig.height = 8}
+res_noise <- dplyr::filter(res, setting == "informative-cubic")
+res_noise <- tidyr::complete(res_noise, setting, Method, dist, key)
+
+noise_fdr <- dplyr::filter(res_noise, key == "FDR") %>%
+    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
+    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
+    scale_y_continuous("FDP", labels = scales::percent) +
+    ggtitle("Mean FDP Over 100 Replications") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+noise_tpr <- dplyr::filter(res_noise, key == "TPR") %>%
+    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
+    scale_y_continuous("TPR", labels = scales::percent) +
+    ggtitle("Mean TPR Over 100 Replications") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+gp <- plot_grid(noise_fdr + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
+                noise_tpr + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:2], ncol = 1)
+gp <- plot_grid(gp, get_legend(noise_tpr), rel_widths = c(1, 1/16))
+gp
 ```
 
-Additionally, we take a look at the FWER (the proportion of simulation replicates out of 100
-with at least one rejection).
+# Informative Covariates
 
-```{r, fig.width = 8, fig.height = 5}
-dplyr::filter(res_nullset, key == "FWER") %>%
-    filter(! (Method == "ashq" & dist == "chisq4"),
-           ! (grepl("scott", Method) & dist == "chisq4")) %>%
-    ggplot(aes(x = Method, y = mean_value, color = Method)) +
-    geom_point() +
-    scale_color_manual(values = col, guide = FALSE) +
-    facet_grid(. ~ dist) +
-    scale_y_continuous("FWER", breaks = seq(0, 1, .2), labels = scales::percent) +
-    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
-    ggtitle("Proportion of replications with at least one false positive (over 100 replications)") 
+We also take a look across different informative covariates.
+
+```{r, fig.width = 18, fig.height = 8}
+res_icov <- dplyr::filter(res, grepl("^informative", setting), grepl("N\\(0, 1\\)", dist))
+res_icov <- tidyr::complete(res_icov, setting, Method, dist, key)
+
+icov_fdr <- dplyr::filter(res_icov, key == "FDR") %>%
+    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
+    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
+    scale_y_continuous("FDP", labels = scales::percent) +
+    ggtitle("Mean FDP Over 100 Replications") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+icov_tpr <- dplyr::filter(res_icov, key == "TPR") %>%
+    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
+    scale_y_continuous("TPR", labels = scales::percent) +
+    ggtitle("Mean TPR Over 100 Replications") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+gp <- plot_grid(icov_fdr + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
+                icov_tpr + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:2], ncol = 1)
+gp <- plot_grid(gp, get_legend(icov_tpr), rel_widths = c(1, 1/16))
+gp
 ```
 
-Finally, we plot the TNR as was done for the yeast in silico experiments/simulations.
+The general trends are identical across informative covariates considered (cubic: a monotonic relationship,
+sine: a non-monotone relationship, and step: a discrete relationship).
 
-```{r, fig.width = 8, fig.height = 5}
-gp <- dplyr::filter(res_nullset, key == "TNR") %>%
-    filter(! (Method == "ashq" & dist == "chisq4"),
-           ! (grepl("scott", Method) & dist == "chisq4")) %>%
-    ggplot(aes(x = Method, y = mean_value, color = Method)) +
-    geom_point() +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se)) + 
-    facet_grid(. ~ dist) +
+# Noninformative Covariates
+
+We take a look at noninformative covariates on their own.
+
+```{r, fig.width = 18, fig.height = 8}
+res_noni <- dplyr::filter(res, grepl("noninformative", setting))
+res_noni <- tidyr::complete(res_noni, setting, Method, dist, key)
+
+noni_fdr <- dplyr::filter(res_noni, key == "FDR") %>%
+    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
+    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
+    scale_y_continuous("FDP", labels = scales::percent) +
+    ggtitle("Mean FDP Over 100 Replications (noninformative covariates)") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+noni_tpr <- dplyr::filter(res_noni, key == "TPR") %>%
+    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
     scale_y_continuous("TPR", labels = scales::percent) +
-    scale_color_manual(values = col, guide = FALSE) +
-    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
-    ggtitle("Average TNR (over 100 replications)") 
+    ggtitle("Mean TPR Over 100 Replications (noninformative covariates)") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+gp <- plot_grid(noni_fdr + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
+                noni_tpr + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:2], ncol = 1)
+gp <- plot_grid(gp, get_legend(noni_tpr), rel_widths = c(1, 1/16))
 gp
-
-ggsave(file.path(outdir, "FigureSsim-null.pdf"), gp, width = 8, height = 5)
 ```
 
-# Informative and Non-Informative Covariates
+We next take a look at noninformative covariates against informative covariates.
 
-We subset on simulation results that look at informative and non-informative covariates across
-different noise distributions.
+# Noninformative vs. Informative
 
-```{r}
-res_covset <- dplyr::filter(res, grepl("informative", setting))
-res_covset <- dplyr::mutate(res_covset, dist = factor(dist,
-                                                      levels = c("gaussian", "t11", "t5", "chisq4"),
-                                                      labels = c(expression("Noise: " * N(0, 1)),
-                                                                 expression("Noise: " * t[11]),
-                                                                 expression("Noise: " * t[5]),
-                                                                 expression("Noise: " * {chi^2}[4]))))
-res_covset <- dplyr::mutate(res_covset, setting = factor(setting,
-                                                         levels = c("noninformative", "informative-step",
-                                                                    "informative-cubic", "informative-sine")))
-res_covset <- tidyr::complete(res_covset, setting, Method, dist, key)
+```{r, fig.width = 10, fig.height = 4}
+res_compc <- dplyr::filter(res, grepl("informative-cubic", setting) |
+                                grepl("noninformative", setting),
+                           grepl("N\\(0, 1\\)", dist))
+res_compc <- tidyr::complete(res_compc, setting, Method, dist, key)
+res_compc <- dplyr::mutate(res_compc, lab = round(100*alpha))
+res_compc <- dplyr::mutate(res_compc, lab = ifelse(lab %in% c(1, 10), lab, NA))
+
+compc_fdr <- dplyr::filter(res_compc, key == "FDR") %>%
+    dplyr::select(setting, Method, dist, alpha, lab, mean_value) %>%
+    tidyr::spread(setting, mean_value) %>%
+    ggplot(aes(x = noninformative, y = `informative-cubic`, label = lab,
+               color = Method, linetype = Method)) +
+    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
+    geom_hline(yintercept = c(.01, .1), lty = 2, color = "blue", alpha = 1/2) + 
+    geom_vline(xintercept = c(.01, .1), lty = 2, color = "blue", alpha = 1/2) + 
+    geom_line(alpha = 0.85) +
+    geom_text(size = 3) + 
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("noninformative (FDP)", breaks = seq(0, 1, .02), labels = scales::percent) + 
+    scale_y_continuous("informative (FDP)", breaks = seq(0, 1, .02), labels = scales::percent) + 
+    ggtitle("Mean FDP Over 100 Replications") +
+    expand_limits(x = 0, y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+compc_tpr <- dplyr::filter(res_compc, key == "TPR") %>%
+    dplyr::select(setting, Method, dist, alpha, lab, mean_value) %>%
+    tidyr::spread(setting, mean_value) %>%
+    ggplot(aes(x = noninformative, y = `informative-cubic`, label = lab,
+               color = Method, linetype = Method)) +
+    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
+    geom_line(alpha = 0.85) +
+    geom_text(size = 3) + 
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("noninformative (TPR)", labels = scales::percent) + 
+    scale_y_continuous("informative (TPR)", labels = scales::percent) + 
+    ggtitle("Mean TPR Over 100 Replications") +
+    expand_limits(x = 0, y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+gp <- plot_grid(compc_fdr + guides(color = FALSE, linetype = FALSE),
+                compc_tpr + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:2], nrow = 1)
+gp <- plot_grid(gp, get_legend(compc_tpr), rel_widths = c(1, 1/8))
+gp
 ```
 
-We use scatterplots to look at the FDR and TPR values.
+We can also plot the differences.
 
-```{r, fig.width = 11, fig.height = 4.5}
-## looking at FDR vs TPR
-col <- as.character(res_covset$col)
-names(col) <- as.character(res_covset$Method)
+```{r, fig.width = 10, fig.height = 4}
+compc2_fdr <- dplyr::filter(res_compc, key == "FDR") %>%
+    dplyr::select(setting, Method, dist, alpha, lab, mean_value) %>%
+    tidyr::spread(setting, mean_value) %>%
+    dplyr::mutate(gain = `informative-cubic` - noninformative) %>%
+    ggplot(aes(x = alpha, y = gain, color = Method, linetype = Method)) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    geom_line(alpha = 0.85) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
+    scale_y_continuous("change in FDP", breaks = seq(-1, 1, .001), labels = scales::percent) + 
+    ggtitle("Mean Change in FDP Over 100 Replications") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
 
-## looking at noninformative vs. informative (cubic)
-dplyr::select(res_covset, Method, setting, dist, key, mean_value, se) %>%
-    dplyr::filter(setting %in% c("noninformative", "informative-cubic"),
-                  key %in% c("TPR"),
-                  ! Method %in% c("ashq", "bonf")) %>%
-    dplyr::mutate(plus1se = mean_value + se, minus1se = mean_value - se) %>%
-    dplyr::select(-se) %>%
-    tidyr::gather(metric, value, mean_value, plus1se, minus1se) %>%
-    tidyr::unite(metric, c("setting", "metric")) %>%
-    tidyr::spread(metric, value) %>%
-    ggplot(aes(x = noninformative_mean_value, y = `informative-cubic_mean_value`, color = Method)) +
-    geom_errorbar(aes(ymin = `informative-cubic_minus1se`, ymax = `informative-cubic_plus1se`), width = 0) +
-    geom_errorbarh(aes(xmin = noninformative_minus1se, xmax = noninformative_plus1se), height = 0) + 
-    ##geom_point(alpha = 1) +
+compc2_tpr <- dplyr::filter(res_compc, key == "TPR") %>%
+    dplyr::select(setting, Method, dist, alpha, lab, mean_value) %>%
+    tidyr::spread(setting, mean_value) %>%
+    dplyr::mutate(gain = `informative-cubic` - noninformative) %>%
+    ggplot(aes(x = alpha, y = gain, color = Method, linetype = Method)) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    geom_line(alpha = 0.85) +
+    scale_linetype_manual(values = lty) + 
     scale_color_manual(values = col) +
-    facet_wrap( ~ dist, nrow = 1, scales = "free", labeller = labeller(dist = label_parsed)) +
-    scale_x_continuous(labels = scales::percent) + 
-    scale_y_continuous(labels = scales::percent) + 
-    geom_abline(color = "blue", lty = 2, alpha = 1/3) +
-    ggtitle("Mean TPR for Informative and Noninformative Simulations") 
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
+    scale_y_continuous("change in TPR", labels = scales::percent) + 
+    ggtitle("Mean Change in TPR Over 100 Replications") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+gp <- plot_grid(compc2_fdr + guides(color = FALSE, linetype = FALSE),
+                compc2_tpr + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:2], nrow = 1)
+gp <- plot_grid(gp, get_legend(compc2_tpr), rel_widths = c(1, 1/8))
+gp
 ```
 
-```{r}
-## looking at noninformative vs. informative (cubic)
-pfdr <- dplyr::select(res_covset, Method, setting, dist, key, mean_value) %>%
-    dplyr::filter(setting %in% c("noninformative", "informative-cubic"),
-                  key %in% c("FDR")) %>%
-    tidyr::spread(setting, mean_value) %>%
-    ggplot(aes(x = noninformative, y = `informative-cubic`, color = Method)) +
-    geom_abline(color = "blue", lty = 2, alpha = 1/3) +
-    geom_vline(xintercept = 0.10, color = "red", lty = 2, alpha = 1/3) +
-    geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) +
-    geom_point(alpha = 1) +
+# Null Proportions
+
+We take a look at the impact of the proportion of null hypotheses.
+
+```{r, fig.width = 10, fig.height = 8}
+res_pi0 <- dplyr::filter(res, setting == "varyingpi0-lowsignal", alpha %in% c(0.05, 0.10))
+res_pi0 <- tidyr::complete(res_pi0, setting, Method, alpha, pi0, key)
+
+pi0_fdr <- dplyr::filter(res_pi0, key == "FDR") %>%
+    ggplot(aes(x = pi0, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 2.5, alpha=0.5) +
+    geom_hline(aes(yintercept = alpha), lty = 2, color = "blue", alpha = 1/2) + 
+    facet_grid(. ~ alpha, scales = "free_x", labeller = label_both) +
+    scale_linetype_manual(values = lty) + 
     scale_color_manual(values = col) +
-    facet_wrap( ~ dist, nrow = 2, scales = "free", labeller = labeller(dist = label_parsed)) +
-    scale_x_continuous("FDR w/ Noninformative Covariate", labels = scales::percent, limits = c(0, 1)) + 
-    scale_y_continuous("FDR w/ Informative Covariate (cubic)", labels = scales::percent, limits = c(0, 1)) + 
-    ggtitle("FDR (Informative and Noninformative Sims)")
+    scale_x_continuous("pi0", breaks = seq(0, 100, 10)) +
+    scale_y_continuous("FDP", labels = scales::percent) +
+    ggtitle("Mean FDP Over 100 Replications") +
+    expand_limits(x =0, y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
 
-## looking at noninformative vs. informative (cubic)
-ptpr <- dplyr::select(res_covset, Method, setting, dist, key, mean_value) %>%
-    dplyr::filter(setting %in% c("noninformative", "informative-cubic"),
-                  key %in% c("TPR")) %>%
-    tidyr::spread(setting, mean_value) %>%
-    ggplot(aes(x = noninformative, y = `informative-cubic`, color = Method)) +
-    geom_abline(color = "blue", lty = 2, alpha = 1/3) +
-    geom_vline(xintercept = 0.10, color = "red", lty = 2, alpha = 1/3) +
-    geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) +
-    geom_point(alpha = 1) +
+pi0_tpr <- dplyr::filter(res_pi0, key == "TPR") %>%
+    ggplot(aes(x = pi0, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 2.5, alpha=0.5) +
+    facet_grid(. ~ alpha, scales = "free_x", labeller = label_both) +
+    scale_linetype_manual(values = lty) + 
     scale_color_manual(values = col) +
-    facet_wrap( ~ dist, nrow = 2, scales = "free", labeller = labeller(dist = label_parsed)) +
-    scale_x_continuous("TPR w/ Noninformative Covariate", labels = scales::percent, limits = c(0, 1)) + 
-    scale_y_continuous("TPR w/ Informative Covariate (cubic)", labels = scales::percent, limits = c(0, 1)) + 
-    ggtitle("TPR (Informative and Noninformative Sims)") 
+    scale_x_continuous("pi0", breaks = seq(0, 100, 10)) +
+    scale_y_continuous("TPR", labels = scales::percent) +
+    ggtitle("Mean TPR Over 100 Replications") +
+    expand_limits(x =0, y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+gp <- plot_grid(pi0_fdr + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
+                pi0_tpr + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:2], ncol = 1)
+gp <- plot_grid(gp, get_legend(pi0_tpr), rel_widths = c(1, 1/8))
+gp
 ```
 
-We can plot both FDR and TPR together.
+# Informativeness
 
-```{r, fig.width = 15, fig.height = 7.5}
-gp <- plot_grid(pfdr + guides(color = FALSE, linetype = FALSE), ptpr,
-                rel_widths = c(1, 1.4), labels = LETTERS[1:2])
+We take a look at the impact of the informativeness of a covariate.
+
+```{r, fig.width = 10, fig.height = 8}
+res_info <- dplyr::filter(res, setting == "varyinginfo-lowsignal", alpha %in% c(0.05, 0.10))
+res_info <- tidyr::complete(res_info, setting, Method, alpha, inform, key)
+
+info_fdr <- dplyr::filter(res_info, key == "FDR") %>%
+    ggplot(aes(x = inform, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 2.5, alpha=0.5) +
+    geom_hline(aes(yintercept = alpha), lty = 2, color = "blue", alpha = 1/2) + 
+    facet_grid(. ~ alpha, scales = "free_x", labeller = label_both) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
+    scale_y_continuous("FDP", labels = scales::percent) +
+    ggtitle("Mean FDP Over 100 Replications") +
+    expand_limits(x =0, y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+info_tpr <- dplyr::filter(res_info, key == "TPR") %>%
+    ggplot(aes(x = inform, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 2.5, alpha=0.5) +
+    facet_grid(. ~ alpha, scales = "free_x", labeller = label_both) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
+    scale_y_continuous("TPR", labels = scales::percent) +
+    ggtitle("Mean TPR Over 100 Replications") +
+    expand_limits(x =0, y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+gp <- plot_grid(info_fdr + guides(color = FALSE, linetype = FALSE),
+                info_tpr + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:2], ncol = 1)
+gp <- plot_grid(gp, get_legend(info_tpr), rel_widths = c(1, 1/8))
 gp
-
-ggsave(file.path(outdir, "FigureSsim-noise.pdf"), gp, width = 11, height = 4.5)
-
-gp_zoom <- plot_grid(pfdr + coord_cartesian(c(0, .2), c(0, .2)) + guides(color = FALSE, linetype = FALSE),
-                     ptpr + coord_cartesian(c(.6, 1), c(.6, 1)), rel_widths = c(1, 1.4), labels = LETTERS[1:2])
-gp_zoom
-
-ggsave(file.path(outdir, "FigureSsim-noise-zoom.pdf"), gp_zoom, width = 11, height = 4.5)
 ```
 
-# Varying Null Proportions
+# Unimodal Effects
 
-Next, we subset on simulation results that look at varying the null proportions.
+```{r, fig.width = 18, fig.height = 12}
+res_uni <- dplyr::filter(res, setting == "varyinges")
+res_uni <- dplyr::mutate(res_uni, dist = droplevels(dist))
+res_uni <- tidyr::complete(res_uni, Method, signal, dist, key)
 
-```{r}
-res_pi0set <- dplyr::filter(res, grepl("varyingpi0", setting))
-res_pi0set <- dplyr::rename(res_pi0set, pi0 = dist)
-res_pi0set <- dplyr::mutate(res_pi0set, pi0 = as.numeric(gsub("nullprop", "", pi0)))
-res_pi0set <- tidyr::complete(res_pi0set, setting, Method, pi0, key)
-res_pi0set <- dplyr::mutate(res_pi0set, setting = factor(setting, levels = c("varyingpi0", "varyingpi0-lowsignal"),
-                                                         labels = c("high signal", "low signal")))
+dplyr::filter(res_uni, key == "FDR") %>%
+    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
+    facet_grid(dist ~ signal, scales = "free_x", labeller = label_parsed) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
+    scale_y_continuous("FDP", labels = scales::percent) +
+    ggtitle("Mean FDP Over 100 Replications") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+dplyr::filter(res_uni, key == "TPR") %>%
+    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    facet_grid(dist ~ signal, scales = "free", labeller = label_parsed) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
+    scale_y_continuous("TPR", labels = scales::percent) +
+    ggtitle("Mean TPR Over 100 Replications") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
 ```
 
-We take a look at the TPR and FDR as function of the proportion of null hypotheses.
+We replot this only looking at Normal sampling noise.
 
-```{r}
-## looking at FDR vs TPR
-col <- as.character(res_pi0set$col)
-names(col) <- as.character(res_pi0set$Method)
-lty <- as.character(res_pi0set$lty)
-names(lty) <- as.character(res_pi0set$Method)
+```{r, fig.width = 18, fig.height = 8}
+res_uni <- dplyr::filter(res_uni, grepl("N\\(0, 1\\)", dist))
 
-## including SE to show variability in FDRR
-pfdr <- dplyr::select(res_pi0set, Method, setting, pi0, key, mean_value, se, col, lty) %>%
-    dplyr::filter(key == "FDR", !is.na(mean_value)) %>%
-    ggplot(aes(x = pi0 / 100, y = mean_value, color = Method, lty = Method)) +
-    geom_line(alpha = 1/6, position = position_dodge(width = .01)) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), alpha = 1) +
-    geom_point(alpha = 1, position = position_dodge(width = .01)) +
+uni_fdr <- dplyr::filter(res_uni, key == "FDR") %>%
+    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
+    facet_wrap(~ signal, scales = "free_x", label = label_both, nrow = 1) +
+    scale_linetype_manual(values = lty) + 
     scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
+    scale_y_continuous("FDP", labels = scales::percent) +
+    ggtitle("Mean FDP Over 100 Replications") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+uni_tpr <- dplyr::filter(res_uni, key == "TPR") %>%
+    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    facet_wrap(~ signal, scales = "free", label = label_both, nrow = 1) +
     scale_linetype_manual(values = lty) + 
-    geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) + 
-    scale_x_continuous("Null Proportion", labels = scales::percent, breaks = seq(0, 1, .1)) + 
-    scale_y_continuous("FDR", labels = scales::percent) + 
-    facet_grid(setting ~ .) + 
-    ggtitle("Average FDR Across Varying Null Proportion")
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
+    scale_y_continuous("TPR", labels = scales::percent) +
+    ggtitle("Mean TPR Over 100 Replications") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
 
-## including SE to show variability in FDRR
-ptpr <- dplyr::select(res_pi0set, Method, setting, pi0, key, mean_value, se, col, lty) %>%
-    dplyr::filter(key == "TPR", !is.na(mean_value)) %>%
-    ggplot(aes(x = pi0 / 100, y = mean_value, color = Method, lty = Method)) +
-    geom_line(alpha = 1/6, position = position_dodge(width = .01)) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), alpha = 1) +
-    geom_point(alpha = 1, position = position_dodge(width = .01)) +
-    scale_color_manual(values = col) +
-    scale_linetype_manual(values = lty) + 
-    scale_x_continuous("Null Proportion", labels = scales::percent, breaks = seq(0, 1, .1)) + 
-    scale_y_continuous("TPR", labels = scales::percent) + 
-    facet_grid(setting ~ .) + 
-    ggtitle("Average TPR Across Varying Null Proportion")
+gp <- plot_grid(uni_fdr + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
+                uni_tpr + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:2], ncol = 1)
+gp <- plot_grid(gp, get_legend(uni_tpr), rel_widths = c(1, 1/16))
+gp
 ```
-
-Again, we can plot both FDR and TPR together.
-
-```{r, fig.width = 15, fig.height = 7.5}
-gp <- plot_grid(pfdr + guides(color = FALSE, linetype = FALSE),
-                ptpr, rel_widths = c(1, 1.25), labels = LETTERS[1:2])
-gp
-
-ggsave(file.path(outdir, "FigureSsim-nullprop.pdf"), gp, width = 15, height = 7.5)
-
-## trimmed ranges
-gp_zoom <- plot_grid(pfdr + coord_cartesian(ylim = c(0, .25)) + guides(color = FALSE, linetype = FALSE),
-                     ptpr, rel_widths = c(1, 1.25), labels = LETTERS[1:2])
-gp_zoom
-
-ggsave(file.path(outdir, "FigureSsim-nullprop-zoom.pdf"), gp_zoom, width = 15, height = 7.5)
-```
-
-# Unimodal Effect Sizes
-
-Next, we subset on simulation results that look at varying unimodal simulation settings
-that were used in the simulations of the ASH paper.
-
-```{r}
-res_unimodal <- dplyr::filter(res, grepl("varyinges", setting), dist != "bimodal")
-res_unimodal <- dplyr::rename(res_unimodal, signal = dist)
-res_unimodal <- tidyr::separate(res_unimodal, setting, c("setting", "dist"), by = "-", fill = "right")
-res_unimodal$dist[is.na(res_unimodal$dist)] <- "gaussian"
-res_unimodal <- tidyr::complete(res_unimodal, setting, Method, dist, signal, key)
-res_unimodal <- dplyr::mutate(res_unimodal, dist = factor(dist,
-                                                          levels = c("gaussian", "t", "chisq"),
-                                                          labels = c(expression("Noise: " * N(0, 1)),
-                                                                 expression("Noise: " * t[11]),
-                                                                 expression("Noise: " * {chi^2}[4]))))
-```
-
-We take a look at the TPR and FDR as function of the informativeness.
-
-```{r, fig.width = 12, fig.height = 7.5}
-## looking at FDR vs TPR
-col <- as.character(res_unimodal$col)
-names(col) <- as.character(res_unimodal$Method)
-lty <- as.character(res_unimodal$lty)
-names(lty) <- as.character(res_unimodal$Method)
-
-## We don't include the error bars because the SEs are so small
-uniplot <- dplyr::select(res_unimodal, Method, dist, signal, key, mean_value, se, col, lty) %>%
-    dplyr::filter(key %in% c("FDR", "TPR")) %>%
-    dplyr::mutate(plus1se = mean_value + se, minus1se = mean_value - se) %>%
-    dplyr::select(-se) %>%
-    tidyr::gather(metric, value, mean_value, plus1se, minus1se) %>%
-    tidyr::unite(metric, c("key", "metric")) %>%
-    tidyr::spread(metric, value) %>%
-    ggplot(aes(x = TPR_mean_value, y = FDR_mean_value, color = Method)) +
-    ##geom_errorbar(aes(ymin = FDR_minus1se, ymax = FDR_plus1se), width = 0) +
-    ##geom_errorbarh(aes(xmin = TPR_minus1se, xmax = TPR_plus1se), height = 0) + 
-    geom_point(alpha = 1) +
-    scale_color_manual(values = col) +
-    geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) + 
-    scale_x_continuous("TPR", labels = scales::percent, breaks = seq(0, 1, .2)) + 
-    scale_y_continuous("FDR", labels = scales::percent, breaks = seq(0, 1, .2)) + 
-    facet_grid(signal ~ dist, labeller = labeller(dist = label_parsed)) + 
-    ggtitle("Average FDR and TPR Across Unimodal Settings")
-uniplot
-
-ggsave(file.path(outdir, "FigureSsim-unimodal.pdf"), uniplot, width = 12, height = 7.5)
-```
-
-# Varying Informativeness
-
-Next, we subset on simulation results that look at varying informativeness of the
-informative covariate.
-
-```{r}
-res_infoset <- dplyr::filter(res, grepl("varyinginfo", setting))
-res_infoset <- dplyr::rename(res_infoset, level = dist)
-res_infoset <- dplyr::mutate(res_infoset, level = as.numeric(gsub("level", "", level)))
-res_infoset <- tidyr::complete(res_infoset, setting, Method, level, key)
-res_infoset <- dplyr::mutate(res_infoset, setting = factor(setting, levels = c("varyinginfo", "varyinginfo-lowsignal"),
-                                                         labels = c("high signal", "low signal")))
-```
-
-We take a look at the TPR and FDR as function of the informativeness.
-
-```{r, fig.width = 8, fig.height = 10}
-## looking at FDR vs TPR
-col <- as.character(res_infoset$col)
-names(col) <- as.character(res_infoset$Method)
-lty <- as.character(res_infoset$lty)
-names(lty) <- as.character(res_infoset$Method)
-
-## not including SE because it's too small
-pfdr <- dplyr::select(res_infoset, Method, setting, level, key, mean_value, se, q10, q90, col, lty) %>%
-    dplyr::filter(key == "FDR") %>%
-    ggplot(aes(x = level / 100, y = mean_value, color = Method, lty = Method)) +
-    geom_line(alpha = 1/6, position = position_dodge(width = .01)) +
-    ##geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0) +
-    geom_point(alpha = 1, position = position_dodge(width = .01)) +
-    scale_color_manual(values = col) +
-    scale_linetype_manual(values = lty) + 
-    geom_hline(yintercept = 0.10, color = "red", lty = 2, alpha = 1/3) + 
-    scale_x_continuous("Informativeness", labels = scales::percent, breaks = seq(0, 1, .1)) + 
-    scale_y_continuous("FDR", labels = scales::percent) + 
-    facet_grid(setting ~ .) + 
-    ggtitle("Average FDR Across Varying Informativeness")
-pfdr
-
-ptpr <- dplyr::select(res_infoset, Method, setting, level, key, mean_value, se, col, lty) %>%
-    dplyr::filter(key == "TPR") %>%
-    ggplot(aes(x = level / 100, y = mean_value, color = Method, lty = Method)) +
-    geom_line(alpha = 1/6, position = position_dodge(width = .01)) +
-    ##geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0) +    
-    geom_point(alpha = 1, position = position_dodge(width = .01)) +
-    scale_color_manual(values = col) +
-    scale_linetype_manual(values = lty) + 
-    scale_x_continuous("Informativeness", labels = scales::percent, breaks = seq(0, 1, .1)) + 
-    scale_y_continuous("TPR", labels = scales::percent) + 
-    facet_grid(setting ~ ., scales = "free_y") + expand_limits(y = 0) +
-    ggtitle("Average TPR Across Varying Informativeness")
-ptpr
-```
-
-Again, we can plot both FDR and TPR together.
-
-```{r, fig.width = 15, fig.height = 7.5}
-gp <- plot_grid(pfdr + guides(color = FALSE, linetype = FALSE), ptpr,
-          rel_widths = c(1, 1.25), labels = LETTERS[1:2])
-gp
-
-ggsave(file.path(outdir, "FigureSsim-informativeness.pdf"), gp, width = 15, height = 7.5)
-```

### 2018-07-20 datasets/RNA-Seq/polyester-simulation.Rmd
Use yeast data to estimate polyester params; don't pass ref data to each fork in bplapply

@@
 library(ggthemes)
+library(readr)
 
@@
 # set up results directories 
+datdir <- "yeast-data"
 resdir <- "../../results/RNAseq"
+dir.create(datdir, showWarnings = FALSE)
 dir.create(resdir, showWarnings = FALSE)
@@
 # results files that will be generated below
-resfile_n5 <- file.path(resdir, "polyester-results-null5.rds")
-resfile_n10 <- file.path(resdir, "polyester-results-null10.rds")
-resfile_d5 <- file.path(resdir, "polyester-results-de5.rds")
-resfile_d10 <- file.path(resdir, "polyester-results-de10.rds")
+resfile_n5 <- file.path(resdir, "polyesterY-results-null5.rds")
+resfile_n10 <- file.path(resdir, "polyesterY-results-null10.rds")
+resfile_d5 <- file.path(resdir, "polyesterY-results-de5.rds")
+resfile_d10 <- file.path(resdir, "polyesterY-results-de10.rds")
 
-resfile_n5_uninfCov <- file.path(resdir, "polyester-results-null5-uninfCov.rds")
-resfile_n10_uninfCov <- file.path(resdir, "polyester-results-null10-uninfCov.rds")
-resfile_d5_uninfCov <- file.path(resdir, "polyester-results-de5-uninfCov.rds")
-resfile_d10_uninfCov <- file.path(resdir, "polyester-results-de10-uninfCov.rds")
+resfile_n5_uninfCov <- file.path(resdir, "polyesterY-results-null5-uninfCov.rds")
+resfile_n10_uninfCov <- file.path(resdir, "polyesterY-results-null10-uninfCov.rds")
+resfile_d5_uninfCov <- file.path(resdir, "polyesterY-results-de5-uninfCov.rds")
+resfile_d10_uninfCov <- file.path(resdir, "polyesterY-results-de10-uninfCov.rds")
 
@@
 ## Estimate the zero inflated negative binomial parameters
-params = get_params(counts)
+paramPolyester = get_params(counts)
+```
+
+Try using the yeast data instead - want all samples of the same condition, and the
+zebrafish dataset only has 3v3.
+
+```{r, readin, results='hide', message=FALSE}
+files <- dir(path = datdir, pattern = "*.bam.gbgout", full.names = TRUE)
+sample_names <- sapply(strsplit(dir(path = datdir, pattern = "*.bam.gbgout"), "_MID"),
+                       function(x) x[[1]])
+badreps <- read_tsv(file.path(datdir, "badreps.txt"), col_names = FALSE)$X1
+badreps <- unlist(lapply(strsplit(badreps, "_MID"), function(x) x[1]))
+
+counts <- files %>%
+  purrr::map(read_tsv, col_names = FALSE) %>% # read in all the files individually
+  purrr::map2(sample_names, ~ dplyr:::rename(.x, !! .y := X2, feature = X1) ) %>% # add sample names
+  purrr::reduce(left_join, by = "feature") %>% # reduce with rbind into one dataframe
+  dplyr::select(-badreps ) %>% # remove badreps
+  data.frame() %>%
+  dplyr::select(contains("WT"))  %>%
+  as.matrix
+
+# filter low count genes
+counts <- counts[rowMeans(counts) > 1,]
+
+## Estimate the zero inflated negative binomial parameters
+paramPolyester = get_params(counts)
 ```
@@
 ```{r}
-#' @param counts is the full count dataset
-#' @param paramPolyester is the parameters from running polyester. 
-#' Contains information on the mean-variance relationship used to 
-#' simulate RNA-Seq counts. 
 #' @param sampleSize is the number of samples in each condition
@@
 #' @param rseed random seed
-simulateOneSplit <- function(X, rseed, counts, paramPolyester, 
+simulateOneSplit <- function(X, rseed, 
                              nDE, sampleSize, bd, uninformativeCovariate = FALSE){
@@
 
-  geneExp <- tbl_df(data.frame(geneName=rownames(counts),
-                               pval=res_alt$pvalue, SE=res_alt$lfcSE, 
+  geneExp <- tbl_df(data.frame(pval=res_alt$pvalue, SE=res_alt$lfcSE, 
                                ind_covariate = res_alt$baseMean, 
@@
 ```{r}
+# To be included in the upset agg plot, method needs to have found on average
+# at least one rejection per replicate. To create an upset plot, require that
+# at least two methods rejected at this threshold.
 #' @param res standardized metric data.table generated using
@@
     summarize(mean_value = mean(value)) %>%
-    filter(mean_value > 0)
+    filter(mean_value > 1)
   return(nrow(res))
@@
 if (!file.exists(resfile_n5)){
-  null5 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, counts=counts,
-                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd)
+  null5 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit,  
+                    nDE=nDE, sampleSize=sampleSize, bd=bd)
   saveRDS(null5, file=resfile_n5)
@@
 
-if (numberMethodsReject(res5, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+if (numberMethodsReject(res5, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
   aggupset(null5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
@@
 if (!file.exists(resfile_n10)){
-  null10 <- bplapply(X=1:B, rseed=837, FUN=simulateOneSplit, counts=counts,
-                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd)
+  null10 <- bplapply(X=1:B, rseed=837, FUN=simulateOneSplit,
+                     nDE=nDE, sampleSize=sampleSize, bd=bd)
   saveRDS(null10, file=resfile_n10)
@@
 
-if (numberMethodsReject(res10, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+if (numberMethodsReject(res10, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
   aggupset(null10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
@@
 if (!file.exists(resfile_d5)){
-  de5 <- bplapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, counts=counts,
-                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd)
+  de5 <- bplapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, 
+                  nDE=nDE, sampleSize=sampleSize, bd=bd)
   saveRDS(de5, file=resfile_d5)
@@
 
-if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
   aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
@@
 if (!file.exists(resfile_d10)){
-  de10 <- bplapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, counts=counts,
-                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, bd=bd)
+  de10 <- bplapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, 
+                   nDE=nDE, sampleSize=sampleSize, bd=bd)
   saveRDS(de10, file=resfile_d10)
@@
 
-if (numberMethodsReject(res10d, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+if (numberMethodsReject(res10d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
   aggupset(de10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
@@
 if (!file.exists(resfile_n5_uninfCov)){
-  null5 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, counts=counts,
-                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, 
+  null5 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
+                    nDE=nDE, sampleSize=sampleSize, 
                    bd=bd, uninformativeCovariate = TRUE)
@@
 if (!file.exists(resfile_n10_uninfCov)){
-  null10 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, counts=counts,
-                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, 
+  null10 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
+                     nDE=nDE, sampleSize=sampleSize, 
                    bd=bd, uninformativeCovariate = TRUE)
@@
 if (!file.exists(resfile_d5_uninfCov)){
-  de5 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, counts=counts,
-                  paramPolyester=params, nDE=nDE, sampleSize=sampleSize, 
+  de5 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
+                  nDE=nDE, sampleSize=sampleSize, 
                   bd=bd, uninformativeCovariate = TRUE)
@@
 if (!file.exists(resfile_d10_uninfCov)){
-  de10 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, counts=counts,
-                   paramPolyester=params, nDE=nDE, sampleSize=sampleSize, 
+  de10 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
+                   nDE=nDE, sampleSize=sampleSize, 
                    bd=bd, uninformativeCovariate = TRUE)

### 2018-07-20 datasets/RNA-Seq/yeast-simulation.Rmd
Use yeast data to estimate polyester params; don't pass ref data to each fork in bplapply

@@
 #' @param seed random seed
-#' @param dds is the full DESeq2 dataset
-#' @param resFULL is the results object from running DESeq2 on dds (used in
-#'   to obtain the fold change effect sizes when adding non-null genes)
 #' @param sampleSize is the number of samples in each condition
@@
 #'  instead of SB objects
-simulateOneSplit <- function(X, rseed, dds, resFULL, nDE, sampleSize,
+simulateOneSplit <- function(X, rseed, nDE, sampleSize,
                              bd, uninformativeCovariate = FALSE, pvalHists = FALSE){
@@
 if (!file.exists(resfile_n5)){
-  null5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
-                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd)
+  null5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+                    nDE=nDE, sampleSize=sampleSize, bd=bd)
   saveRDS(null5, file = resfile_n5)
@@
 if (!file.exists(resfile_n10)){
-  null10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
-                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd)
+  null10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+                    nDE=nDE, sampleSize=sampleSize, bd=bd)
   saveRDS(null10, file = resfile_n10)
@@
 if (!file.exists(resfile_d5)){
-  de5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
-                  resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd)
+  de5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+                  nDE=nDE, sampleSize=sampleSize, bd=bd)
   saveRDS(de5, file = resfile_d5)
@@
 if (!file.exists(plotfile)){
-  hists <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
-                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd, 
+  hists <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+                   nDE=nDE, sampleSize=sampleSize, bd=bd, 
                    pvalHists = TRUE)
@@
 
-if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 3){
+if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
   aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE)
@@
 if(!file.exists(resfile_d10)){
-  de10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
-                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd)
+  de10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+                  nDE=nDE, sampleSize=sampleSize, bd=bd)
   saveRDS(de10, file = resfile_d10)
@@
 if (!file.exists(resfile_n5_uninfCov)){
-  null5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
-                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd,
+  null5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed,
+                   nDE=nDE, sampleSize=sampleSize, bd=bd,
                    uninformativeCovariate = TRUE)
@@
 if (!file.exists(resfile_n10_uninfCov)){
-  null10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
-                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd,
+  null10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+                   nDE=nDE, sampleSize=sampleSize, bd=bd,
                    uninformativeCovariate = TRUE)
@@
 if (!file.exists(resfile_d5_uninfCov)){
-  de5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
-                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd,
+  de5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+                  nDE=nDE, sampleSize=sampleSize, bd=bd,
                   uninformativeCovariate = TRUE)
@@
 if(!file.exists(resfile_d10_uninfCov)){
-  de10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, dds=dds, 
-                   resFULL=resFULL, nDE=nDE, sampleSize=sampleSize, bd=bd,
+  de10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+                   nDE=nDE, sampleSize=sampleSize, bd=bd,
                    uninformativeCovariate = TRUE)

### 2018-07-23 datasets/simulations/simulations-informative-cosine.Rmd
Add cosine informative covariate sim

--- 2018-07-17 datasets/simulations/simulations-informative-step.Rmd
@@
 ---
-title: "Simulation Study: Informative Covariate (step)"
+title: "Simulation Study: Informative Covariate (cosine)"
 author: "Patrick Kimes"
@@
 the interval [0, 1], and the conditional probability of a test being non-null
-is a non-smooth (monotone) step function of the covariate. We include simulation results
+is a smooth (cosine) function of the covariate. We include simulation results
 again with Gaussian, t-distributed, and Chi-Squared distributed noise.
@@
 ## intermediary files we create below
-gauss_file <- file.path(resdir, "informative-step-benchmark-gaussian.rds")
-tdist_file <- file.path(resdir, "informative-step-benchmark-t5.rds")
-tdist11_file <- file.path(resdir, "informative-step-benchmark-t11.rds")
-chisq_file <- file.path(resdir, "informative-step-benchmark-chisq4.rds")
+gauss_file <- file.path(resdir, "informative-cosine-benchmark-gaussian.rds")
+tdist_file <- file.path(resdir, "informative-cosine-benchmark-t5.rds")
+tdist11_file <- file.path(resdir, "informative-cosine-benchmark-t11.rds")
+chisq_file <- file.path(resdir, "informative-cosine-benchmark-chisq4.rds")
 
@@
 m <- 20000                        # integer: number of hypothesis tests
-pi0 <- pi0_step(0.90)             # numeric: proportion of null hypotheses
+pi0 <- pi0_cosine(0.90)           # numeric: proportion of null hypotheses
 icovariate <- runif               # functional: independent covariate

### 2018-07-23 datasets/simulations/simulations-null.Rmd
Fix t11 null sim output file names

@@
 ```{r t11-run-simulation}
-if (file.exists(tdist_file)) {
-    res <- readRDS(tdist_file)
+if (file.exists(tdist11_file)) {
+    res <- readRDS(tdist11_file)
 } else {
@@
                     seed = seed, mc.cores = cores)
-    saveRDS(res, file = tdist_file)
+    saveRDS(res, file = tdist11_file)
 }

### 2018-07-23 manuscript_figures/manuscript_figures_sims.Rmd
Look at higher FDR cutoffs for ASH unimodal sims

@@
 ```{r, eval = FALSE}
-for (ia in 1:10) {
-    summary_file <- paste0("allres_alpha", 100*alpha, ".rds")
+for (ia in c(1:10, 20, 30, 40, 50)) {
+    summary_file <- paste0("allres_alpha", ia, ".rds")
     if (!file.exists(summary_file)) {
@@
             zz <- readRDS(x)
-            zz <- plotsim_standardize(zz, alpha = alpha)
+            zz <- plotsim_standardize(zz, alpha = ia / 100L)
             zz <- dplyr::select(zz, rep, blabel, key, alpha, value)
@@
 ```{r}
-res <- vector("list", 10)
-for (ia in 1:10) {
+alphas <- 1:10
+res <- vector("list", length(alphas))
+for (ia in seq_len(length(alphas))) {
     imethodset <- methodset
-    imethodset[which(imethodset == "ihw")] <- paste0("ihw-a", sprintf("%02d", ia))
-    ires <- readRDS(paste0("allres_alpha", ia, ".rds"))
+    imethodset[which(imethodset == "ihw")] <- paste0("ihw-a", sprintf("%02d", alphas[ia]))
+    ires <- readRDS(paste0("allres_alpha", alphas[ia], ".rds"))
     ires <- bind_rows(ires, .id = "setting")
@@
     scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
-    scale_y_continuous("FDP", labels = scales::percent) +
-    ggtitle("Mean FDP Over 100 Replications") +
+    scale_y_continuous("FDR", labels = scales::percent) +
+    ggtitle("Mean FDR Over 100 Replications") +
     expand_limits(y = 0) +
@@
 
+gp <- plot_grid(noise_fdr + guides(color = FALSE, linetype = FALSE),
+                noise_tpr + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:2], ncol = 1)
+gp <- plot_grid(gp, get_legend(noise_tpr), rel_widths = c(1, 1/16))
+gp
+
 gp <- plot_grid(noise_fdr + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
@@
 
-```{r, fig.width = 18, fig.height = 8}
+```{r, fig.width = 14, fig.height = 8}
 res_icov <- dplyr::filter(res, grepl("^informative", setting), grepl("N\\(0, 1\\)", dist))
@@
 
-gp <- plot_grid(icov_fdr + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
+gp <- plot_grid(icov_fdr + guides(color = FALSE, linetype = FALSE),
                 icov_tpr + guides(color = FALSE, linetype = FALSE),
                 labels = LETTERS[1:2], ncol = 1)
-gp <- plot_grid(gp, get_legend(icov_tpr), rel_widths = c(1, 1/16))
+gp <- plot_grid(gp, get_legend(icov_tpr), rel_widths = c(1, 1/12))
 gp
@@
     scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
-    scale_y_continuous("FDP", labels = scales::percent) +
-    ggtitle("Mean FDP Over 100 Replications (noninformative covariates)") +
+    scale_y_continuous("FDR", labels = scales::percent) +
+    ggtitle("Mean FDR Over 100 Replications (noninformative covariates)") +
     expand_limits(y = 0) +
@@
           plot.title = element_text(face="bold", size = 14))
+
+gp <- plot_grid(noni_fdr + guides(color = FALSE, linetype = FALSE),
+                noni_tpr + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:2], ncol = 1)
+gp <- plot_grid(gp, get_legend(noni_tpr), rel_widths = c(1, 1/16))
+gp
 
@@
     scale_color_manual(values = col) +
-    scale_x_continuous("noninformative (FDP)", breaks = seq(0, 1, .02), labels = scales::percent) + 
-    scale_y_continuous("informative (FDP)", breaks = seq(0, 1, .02), labels = scales::percent) + 
-    ggtitle("Mean FDP Over 100 Replications") +
+    scale_x_continuous("noninformative (FDR)", breaks = seq(0, 1, .02), labels = scales::percent) + 
+    scale_y_continuous("informative (FDR)", breaks = seq(0, 1, .02), labels = scales::percent) + 
+    ggtitle("Mean FDR Over 100 Replications") +
     expand_limits(x = 0, y = 0) +
@@
     scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
-    scale_y_continuous("change in FDP", breaks = seq(-1, 1, .001), labels = scales::percent) + 
-    ggtitle("Mean Change in FDP Over 100 Replications") +
+    scale_y_continuous("change in FDR", breaks = seq(-1, 1, .001), labels = scales::percent) + 
+    ggtitle("Mean Change in FDR Over 100 Replications") +
     theme_classic() +
@@
     scale_x_continuous("pi0", breaks = seq(0, 100, 10)) +
-    scale_y_continuous("FDP", labels = scales::percent) +
-    ggtitle("Mean FDP Over 100 Replications") +
+    scale_y_continuous("FDR", labels = scales::percent) +
+    ggtitle("Mean FDR Over 100 Replications") +
     expand_limits(x =0, y = 0) +
@@
     scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
-    scale_y_continuous("FDP", labels = scales::percent) +
-    ggtitle("Mean FDP Over 100 Replications") +
+    scale_y_continuous("FDR", labels = scales::percent) +
+    ggtitle("Mean FDR Over 100 Replications") +
     expand_limits(x =0, y = 0) +
@@
 ```
+
+# ASH unimodality
+
+```{r}
+alphas50 <- c(5, 10, 20, 30, 40, 50)
+res_fdr50 <- vector("list", length(alphas50))
+for (ia in seq_len(length(alphas50))) {
+    imethodset <- methodset
+    imethodset[which(imethodset == "ihw")] <- "ihw-a10"
+    ires <- readRDS(paste0("allres_alpha", alphas50[ia], ".rds"))
+    ires <- bind_rows(ires, .id = "setting")
+    ires <- dplyr::mutate(ires, dist = gsub(".*?-benchmark-(.*)", "\\1", setting),
+                         setting = gsub("(.*?)-benchmark-.*", "\\1", setting))
+    ires <- dplyr::filter(ires, blabel %in% imethodset)
+    ires <- dplyr::mutate(ires, blabel = gsub("(-a)(.*)", "", blabel))
+    res_fdr50[[ia]] <- ires
+}
+res_fdr50 <- bind_rows(res_fdr50)
+res_fdr50 <- dplyr::rename(res_fdr50, Method = blabel)
+res_fdr50 <- dplyr::mutate(res_fdr50, Method = gsub("-df03", "", Method))
+res_fdr50 <- dplyr::mutate(res_fdr50, Method = factor(Method, levels = gsub("-df03", "", methodset)))
+```
+
+We clean up the annotations in the table for some of the more complex simulation settings.
+
+```{r}
+res_fdr50 <- dplyr::mutate(res_fdr50, inform = ifelse(grepl("varyinginfo", setting), dist, NA))
+res_fdr50 <- dplyr::mutate(res_fdr50, inform = as.numeric(gsub("level", "", inform)))
+res_fdr50 <- dplyr::mutate(res_fdr50, dist = ifelse(grepl("varyinginfo", setting), "gaussian", dist))
+
+res_fdr50 <- dplyr::mutate(res_fdr50, pi0 = ifelse(grepl("varyingpi0", setting), dist, 90))
+res_fdr50 <- dplyr::mutate(res_fdr50, pi0 = as.numeric(gsub("nullprop", "", pi0)))
+res_fdr50 <- dplyr::mutate(res_fdr50, dist = ifelse(grepl("varyingpi0", setting), "gaussian", dist))
+
+res_fdr50 <- dplyr::mutate(res_fdr50, signal = ifelse(grepl("varyinges", setting), dist, NA))
+res_fdr50 <- dplyr::mutate(res_fdr50, dist = ifelse(grepl("^varyinges$", setting), "gaussian", dist))
+res_fdr50 <- dplyr::mutate(res_fdr50, dist = ifelse(grepl("^varyinges-t$", setting), "t11", dist))
+res_fdr50 <- dplyr::mutate(res_fdr50, dist = ifelse(grepl("^varyinges-chisq$", setting), "chisq4", dist))
+res_fdr50 <- dplyr::mutate(res_fdr50, setting = ifelse(grepl("varyinges", setting), "varyinges", setting))
+```
+
+We also replace the sampling noise labels with more plotting-friendly labels.
+
+```{r}
+res_fdr50 <- dplyr::mutate(res_fdr50, dist = factor(dist, levels = c("gaussian", "t11", "t5", "chisq4"),
+                                                    labels = c(expression("Noise: " * N(0, 1)),
+                                                               expression("Noise: " * t[11]),
+                                                               expression("Noise: " * t[5]),
+                                                               expression("Noise: " * {chi^2}[4]))))
+```
+
+We take a look at the unimodal set of simulations with Normal sampling noise.
+
+```{r, fig.width = 18, fig.height = 8}
+res50_uni <- dplyr::filter(res_fdr50, setting == "varyinges", grepl("N\\(0, 1\\)", dist))
+res50_uni <- dplyr::mutate(res50_uni, dist = droplevels(dist))
+res50_uni <- tidyr::complete(res50_uni, Method, signal, dist, key)
+
+uni50_fdr <- dplyr::filter(res50_uni, key == "FDR") %>%
+    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.025, alpha=0.5) +
+    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
+    facet_wrap(~ signal, scales = "free_x", label = label_both, nrow = 1) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .1)) +
+    scale_y_continuous("FDR", labels = scales::percent) +
+    ggtitle("Mean FDR Over 100 Replications") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+uni50_tpr <- dplyr::filter(res50_uni, key == "TPR") %>%
+    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.025, alpha=0.5) +
+    facet_wrap(~ signal, scales = "free", label = label_both, nrow = 1) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .1)) +
+    scale_y_continuous("TPR", labels = scales::percent) +
+    ggtitle("Mean TPR Over 100 Replications") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+gp <- plot_grid(uni50_fdr + guides(color = FALSE, linetype = FALSE),
+                uni50_tpr + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:2], ncol = 1)
+gp <- plot_grid(gp, get_legend(uni50_tpr), rel_widths = c(1, 1/16))
+gp
+```
+
+# Figure Sim1
+
+We combine some of the results from above into a single figure.
+
+```{r, fig.width = 10, fig.height = 11}
+## Figure Sim1, PanelA: FDR under different sampling distributions
+
+## Figure Sim1, PanelB: informativeness, FDR at alpha = 0.10
+fs1b <- dplyr::filter(res_info, key == "FDR", alpha == 0.10) %>%
+    ggplot(aes(x = inform, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 2.5, alpha=0.5) +
+    geom_hline(aes(yintercept = alpha), lty = 2, color = "blue", alpha = 1/2) + 
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
+    scale_y_continuous("FDR", labels = scales::percent) +
+    ggtitle("Mean FDR Over 100 Replications") +
+    expand_limits(x =0, y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+## Figure Sim1, PanelC: informativeness, TPR at alpha = 0.10
+fs1c <- dplyr::filter(res_info, key == "TPR", alpha == 0.10) %>%
+    ggplot(aes(x = inform, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 2.5, alpha=0.5) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
+    scale_y_continuous("TPR", labels = scales::percent) +
+    ggtitle("Mean TPR Over 100 Replications") +
+    expand_limits(x =0, y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+## Figure Sim1, PanelD: varying null proportions, FDR at alpha = 0.10
+fs1d <- dplyr::filter(res_pi0, key == "FDR", alpha == 0.10) %>%
+    ggplot(aes(x = pi0, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 2.5, alpha=0.5) +
+    geom_hline(aes(yintercept = alpha), lty = 2, color = "blue", alpha = 1/2) + 
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("pi0", breaks = seq(0, 100, 10)) +
+    scale_y_continuous("FDR", labels = scales::percent) +
+    ggtitle("Mean FDR Over 100 Replications") +
+    expand_limits(x =0, y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+## Figure Sim1, PanelE: varying null proportions, TPR at alpha = 0.10
+fs1e <- dplyr::filter(res_pi0, key == "TPR", alpha == 0.10) %>%
+    ggplot(aes(x = pi0, y = mean_value, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 2.5, alpha=0.5) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("pi0", breaks = seq(0, 100, 10)) +
+    scale_y_continuous("TPR", labels = scales::percent) +
+    ggtitle("Mean TPR Over 100 Replications") +
+    expand_limits(x =0, y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+gp <- plot_grid(fs1b + guides(color = FALSE, linetype = FALSE),
+                fs1c + guides(color = FALSE, linetype = FALSE),
+                fs1d + guides(color = FALSE, linetype = FALSE),
+                fs1e + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[2:5], ncol = 2)
+gp <- plot_grid(gp, get_legend(fs1b), rel_widths = c(1, 1/8))
+plot_grid(noise_fdr + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
+          gp, ncol = 1, rel_heights = c(5/12, 1),
+          labels = c(LETTERS[1], NA))
+```

### 2018-07-23 datasets/RNA-Seq/polyester-simulation.Rmd
Add flexibility to simulate arbitrary number of genes

@@
 library(SummarizedBenchmark)
-library(zebrafishRNASeq) 
 library(genefilter)
@@
 # results files that will be generated below
-resfile_n5 <- file.path(resdir, "polyesterY-results-null5.rds")
-resfile_n10 <- file.path(resdir, "polyesterY-results-null10.rds")
-resfile_d5 <- file.path(resdir, "polyesterY-results-de5.rds")
-resfile_d10 <- file.path(resdir, "polyesterY-results-de10.rds")
+resfile_n5 <- file.path(resdir, "polyester-results-null5.rds")
+resfile_n10 <- file.path(resdir, "polyester-results-null10.rds")
+resfile_d5 <- file.path(resdir, "polyester-results-de5.rds")
+resfile_d10 <- file.path(resdir, "polyester-results-de10.rds")
 
-resfile_n5_uninfCov <- file.path(resdir, "polyesterY-results-null5-uninfCov.rds")
-resfile_n10_uninfCov <- file.path(resdir, "polyesterY-results-null10-uninfCov.rds")
-resfile_d5_uninfCov <- file.path(resdir, "polyesterY-results-de5-uninfCov.rds")
-resfile_d10_uninfCov <- file.path(resdir, "polyesterY-results-de10-uninfCov.rds")
+resfile_n5_uninfCov <- file.path(resdir, "polyester-results-null5-uninfCov.rds")
+resfile_n10_uninfCov <- file.path(resdir, "polyester-results-null10-uninfCov.rds")
+resfile_d5_uninfCov <- file.path(resdir, "polyester-results-de5-uninfCov.rds")
+resfile_d10_uninfCov <- file.path(resdir, "polyester-results-de10-uninfCov.rds")
 
@@
 The package requires a count matrix as input which estimates the mean variance relationship.
-Here we start with the [zebrafish](https://bioconductor.org/packages/release/data/experiment/html/zebrafishRNASeq.html) Bioconductor data package.
-
-Estimate mean and variance relationship
-
-```{r estimate-parameters-NB}
-data(zfGenes)
-dim(zfGenes)
-
-filter <- apply(zfGenes, 1, function(x) length(x[x>5])>=2)
-counts <- as.matrix(zfGenes[filter,])
-
-## Estimate the zero inflated negative binomial parameters
-paramPolyester = get_params(counts)
-```
-
-Try using the yeast data instead - want all samples of the same condition, and the
-zebrafish dataset only has 3v3.
+Here we start with the yeast data with 48 biological replicates in each of
+two conditions (analyzed in this [publication](https://www.ncbi.nlm.nih.gov/pubmed/26206307/)). 
+We'll use the samples passing QC in the original study that belong to the WT group to 
+estimate mean and variance relationship. We assume the data has already been downloaded
+(this is carried out in the `yeast-simulation.Rmd` vignette).
 
@@
                              nDE, sampleSize, bd, uninformativeCovariate = FALSE){
-  if(uninformativeCovariate)
-    uninf_covar <- rnorm(nrow(counts))
+  ngenes <- 10000
+
+  uninf_covar <- rnorm(ngenes)
 
@@
 
-  # NULL simulation
-  if(nDE == 0) {
-    truth <- rep(FALSE, nrow(counts)) 
-    names(truth) <- rownames(counts)
+  gs <- sample(1:nrow(counts), ngenes, replace = TRUE)
+  mu <- paramPolyester$mu[gs]
+  p0 <- paramPolyester$p0[gs]
 
-    # log2FC are all 0 in null simulation
-    log2FC = rep(0,nrow(counts))
-    dat_alt = create_read_numbers(paramPolyester$mu, paramPolyester$fit,
-                                  paramPolyester$p0, m=dim(counts)[1], 
+  # Null genes
+  log2FC = rep(0, ngenes)
+  dat_alt = create_read_numbers(mu,
+                                paramPolyester$fit,
+                                p0,
                                 n=sampleSize*2, beta=cbind(log2FC), 
                                 mod=modPolyester)
-  }
+
+  # filter genes with mostly zeroes
+  pzero = rowSums(dat_alt==0)/ncol(dat_alt)
+  dat_alt <- dat_alt[pzero < 0.5,]
+  truth <- rep(FALSE, nrow(dat_alt))
+  log2FC <- log2FC[pzero < 0.5]
+  mu <- mu[pzero < 0.5]
+  p0 <- p0[pzero < 0.5]
+  uninf_covar <- uninf_covar[pzero < 0.5]
 
@@
     # pick random set of nDE genes to add signal to
-    DE <- sample(1:nrow(counts), nDE)
-    truth <- rep(FALSE, nrow(counts))
+    DE <- sample(1:nrow(dat_alt), nDE)
     truth[DE] <- TRUE
-    names(truth) <- rownames(counts)
 
-    # randomly sample a log2FC from truncated N(0,0.5) (greater than 0.5) 
-    log2FC <- rep(0, nrow(counts))
+    # randomly sample a log2FC from N(0,1) 
     log2FC[DE] <- rnorm(nDE, sd = 1)
-    dat_alt = create_read_numbers(paramPolyester$mu, paramPolyester$fit,
-                                  paramPolyester$p0, m=dim(counts)[1], 
-                                  n=sampleSize*2, beta=cbind(log2FC), 
+    dat_alt[DE,] = create_read_numbers(mu[DE], paramPolyester$fit, p0[DE], 
+                                  n=sampleSize*2, beta=cbind(log2FC[DE]), 
                                   mod=modPolyester)
@@
                                effect_size=res_alt$log2FoldChange, 
-                               test_statistic=res_alt$stat, qvalue=truth, 
-                               pzero = rowSums(counts(dds_alt)==0)/
-                                 ncol(counts(dds_alt))  ))
+                               test_statistic=res_alt$stat, qvalue=truth))
 
@@
 
-  # filter NAs and those with less than 50% expressed 
-  geneExp <- geneExp %>% na.omit() %>% dplyr::filter(pzero < 0.5)
+  # filter NAs 
+  geneExp <- geneExp %>% na.omit()
 
@@
                           ftCols = "ind_covariate")
+  sb <- addDefaultMetrics(sb)
   rowData(sb)$log2FC <- geneExp$effect_size

### 2018-07-23 datasets/RNA-Seq/yeast-simulation.Rmd
Reshuffle sample labels if PC1 or PC2 show significant differences in null

@@
 Finally, the `reduce` function is used to join all the replicates together.
+We will remove samples that failed QC in the original study.
 
@@
 
-```{r, deseq}
+First we set up the DESeq2 object.
+
+```{r, deseqsetup}
 feats <- (counts %>% select(1))$feature
@@
                               design= ~ condition)
+```
 
+## Run DESeq2
+
+Next we run DESeq2.
+
+```{r}
 # results on full set
@@
                      nulltype = 'empirical',
-                     control = list(lambda = 0.01, nmids = 300, densknots = 20))
+                     control = list(lambda = 0.01))
 ```
@@
                              bd, uninformativeCovariate = FALSE, pvalHists = FALSE){
-  if(uninformativeCovariate)
+
   uninf_covar <- rnorm(nrow(dds))
@@
   design(dds_test) <- ~fake
+
+  pzero = rowSums(counts(dds_test)==0)/ncol(counts(dds_test))
+  dds_test <- dds_test[pzero < 0.5,]
+  uninf_covar <- uninf_covar[pzero < 0.5]
+
   truth <- rep(FALSE, nrow(dds_test))
+
+  # make sure null comparison is truly null: if PC1 or PC2 sig different, 
+  # reshuffle sample labels
+  dds_test <- estimateSizeFactors(dds_test)
+  x <- t(counts(dds_test, normalize=TRUE))
+  pc <- prcomp(log(x + 0.5), scale.=TRUE)
+
+  a1 <- pc$x[colData(dds_test)$fake=="A",1]
+  b1 <- pc$x[colData(dds_test)$fake=="B",1] 
+  a2 <- pc$x[colData(dds_test)$fake=="A",2]
+  b2 <- pc$x[colData(dds_test)$fake=="B",2] 
+  p1 <- wilcox.test(a1, b1)$p.value
+  p2 <- wilcox.test(a2, b2)$p.value
+  tries <- 0
+
+  while(p1 < 0.10 || p2 < 0.10 && tries < 5){
+    colData(dds_test)$fake <- sample(colData(dds_test)$fake, ncol(dds_test))
+    x <- t(counts(dds_test, normalize=TRUE))
+    pc <- prcomp(log(x + 0.5), scale.=TRUE)
+
+    a1 <- pc$x[colData(dds_test)$fake=="A",1]
+    b1 <- pc$x[colData(dds_test)$fake=="B",1] 
+    a2 <- pc$x[colData(dds_test)$fake=="A",2]
+    b2 <- pc$x[colData(dds_test)$fake=="B",2] 
+    p1 <- t.test(a1, b1)$p.value
+    p2 <- t.test(a2, b2)$p.value
+
+    tries <- tries + 1
+  }
 
@@
     DE <- sample(1:nrow(dds_test), nDE)
-    truth <- rep(FALSE, nrow(dds_test))
     truth[DE] <- TRUE
-    names(truth) <- rownames(dds_test)
 
@@
                                test_statistic = resTEST$stat,
-                               qvalue = truth,
-                               pzero = rowSums(counts(dds_test)==0)/
-                                       ncol(counts(dds_test))))
+                               qvalue = truth))
 
@@
 
-  geneExp <-  geneExp %>% dplyr::filter(!is.na(pval)) %>% dplyr::filter(pzero < 0.5)
+  geneExp <-  geneExp %>% dplyr::filter(!is.na(pval))
 

### 2018-07-23 datasets/GWAS/GWAS.Rmd
Set random seed for uninformative covariate
Change 'scott' labels to 'fdrreg'

@@
 ```{r}
+set.seed(39580)
 bmi <- bmi %>% mutate(ind_covar_uninf = rnorm(nrow(bmi)))
@@
 ```{r}
-bd <- addBMethod(bd, "scott-theoretical",
+bd <- addBMethod(bd, "fdrreg-t",
                      FDRreg::FDRreg,
@@
                      control = list(lambda = 0.01))
-bd <- addBMethod(bd, "scott-empirical",
+bd <- addBMethod(bd, "fdrreg-e",
                      FDRreg::FDRreg,

### 2018-07-23 datasets/RNA-Seq/RNAseq-gtex.Rmd
Change 'scott' labels to 'fdrreg'

@@
 
-    bd <- addBMethod(bd, "scott-theoretical",
+    bd <- addBMethod(bd, "fdrreg-t",
                      FDRreg::FDRreg,
@@
 
-   bd <- addBMethod(bd, "scott-empirical",
+   bd <- addBMethod(bd, "fdrreg-e",
                      FDRreg::FDRreg,
@@
 
-ash was the method that rejected the largest number of hypotheses, followed by lfdr and scott-theoretical.
+ash was the method that rejected the largest number of hypotheses, followed by lfdr and fdrreg-theoretical.
 

### 2018-07-23 datasets/RNA-Seq/RNAseq-mir200c.Rmd
Change 'scott' labels to 'fdrreg'

--- 2018-05-13 datasets/RNA-Seq/RNAseq-mir200c.Rmd
@@
 
-As a second RNA-seq dataset, we will test for differences in gene expression upon the knockout of the microRNA mir-200c [(Kim et al., 2013)](https://doi.org/10.1038/nsmb.2701).
+As a second RNA-seq dataset, we will test for differences in gene expression upon the knockout of the microRNA mir-200c [(Kim et al., 2013)](http://dx.doi.org/10.1038/nsmb.2701).
 The raw fastq files can be found under the accession number `SRP030475`.

### 2018-07-23 datasets/RNA-Seq/yeast-simulation.Rmd
Change 'scott' labels to 'fdrreg'

@@
 bd <- initializeBenchDesign()
-bd <- addBMethod(bd, "scott-theoretical",
+bd <- addBMethod(bd, "fdrreg-t",
                      FDRreg::FDRreg,

### 2018-07-23 datasets/simulations/simulations-informative-cosine.Rmd
Change 'scott' labels to 'fdrreg'

@@
 bdplus <- bd
-bdplus <- addBMethod(bdplus, "scott-theoretical",
+bdplus <- addBMethod(bdplus, "fdrreg-t",
                      FDRreg::FDRreg,
@@
                      control = list(lambda = 0.01))
-bdplus <- addBMethod(bdplus, "scott-empirical",
+bdplus <- addBMethod(bdplus, "fdrreg-e",
                      FDRreg::FDRreg,

### 2018-07-23 manuscript_figures/manuscript_figures.Rmd
Change 'scott' labels to 'fdrreg'

@@
 Fig4All.rowOrder <- c("bonf", "bh", "ihw", "qvalue", "bl", "lfdr",
-                  "scott-empirical", "scott-theoretical", "ashq")
+                  "fdrreg-e", "fdrreg-t", "ashq")
 Fig4All <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
@@
 method_order <- c("bonf", "bh", "ihw", "qvalue", "bl", "lfdr",
-                  "scott-empirical", "scott-theoretical", "ashq")
+                  "fdrreg-e", "fdrreg-t", "ashq")
 objdat_joint <- dplyr::mutate(objdat_joint, method = factor(method, levels = c(method_order, "Total Tests")))
@@
 # methods that use a covariate
-cov.methods <- c("ihw", "bl", "lfdr", "scott-empirical", "scott-theoretical")
+cov.methods <- c("ihw", "bl", "lfdr", "fdrreg-e", "fdrreg-t")
 
@@
 
-rowOrder <- c("ihw", "bl", "lfdr", "scott-empirical", "scott-theoretical")
+rowOrder <- c("ihw", "bl", "lfdr", "fdrreg-e", "fdrreg-t")
 

### 2018-07-23 manuscript_figures/manuscript_figures_sims.Rmd
Change 'scott' labels to 'fdrreg'

@@
 methodset <- c("bonf", "bh", "ihw", "qvalue", "bl-df03", "lfdr",
-               "scott-empirical", "scott-theoretical", "ashq")
+               "fdrreg-e", "fdrreg-t", "ashq")
 ```
@@
 are made. Additionally, we take a look at the FWER (the proportion of simulation replicates out
-of 100 with at least one rejection). In these plots, `ashq`, `scott-theoretical` and `scott-theoretical`
+of 100 with at least one rejection). In these plots, `ashq`, `fdrreg-e` and `fdrreg-t`
 are excluded for the Chi-Squared simulation settings.
@@
                              !(Method == "ashq" & grepl("chi\\^2", dist)),
-                             !(grepl("scott", Method) & grepl("chi\\^2", dist)))
+                             !(grepl("fdrreg", Method) & grepl("chi\\^2", dist)))
 res_nullset <- tidyr::complete(res_nullset, setting, Method, dist, key)

### 2018-07-23 datasets/R/colorPalette.R

@@ -16,7 +16,7 @@
 
 candycols <- data.frame(Method = c("bonf", "bh", "qvalue", "ashs", "ashq", 
                                    "ihw", "bl", "lfdr", 
-                                   "scott-empirical", "scott-theoretical"),
+                                   "fdrreg-t", "fdrreg-e"),
                         col = c("black", "black", "gray45", "hotpink", "red3",
                                   "green3", "darkorange1", "dodgerblue3", 
                                   "purple4", "mediumpurple3"),

### 2018-07-23 datasets/R/effectSizePlot.R

@@ -28,13 +28,7 @@ summarize_one_item <- function(object, alpha){
   df <- df %>%
     dplyr::select(truth, covname, bonf, bh,
                   qvalue, contains("ihw"), starts_with("ash"), "bl-df03", lfdr,
-                  matches("scott"))
-  if ("scott-theoretical" %in% names(df)) {
-    df <- dplyr::rename(df, "scott-t" = "scott-theoretical")
-  }
-  if ("scott-empirical" %in% names(df)) {
-    df <- dplyr::rename(df, "scott-e" = "scott-empirical")
-  }
+                  matches("fdrreg"))
   df <- df %>%
     tidyr::gather(method, significant, -truth, -covname) %>%
     dplyr::group_by(method, truth, significant) %>%
@@ -124,8 +118,7 @@ covariateLinePlot <- function(sbl, alpha=0.05, nbins = 25,
       dplyr::select(truth, covname, bonf, bh,
                     qvalue, contains("ihw"), contains("ash"),
                     "bl-df03", lfdr, 
-                    contains("scott-theoretical"), 
-                    contains("scott-empirical")) %>%
+                    contains("fdrreg")) %>%
       dplyr::mutate(bin = ntile(abs(get(covname)), nbins)) %>%
       tidyr::gather(method, significant, -covname, -bin) %>%
       dplyr::group_by(method, bin) %>%

### 2018-07-23 datasets/R/plotMethodRank.R

@@ -14,7 +14,7 @@
 #' @param linePlot logical indicating whether to plot a line plot in lieu of 
 #'  a heatmap  
 #' @param excludeMethods character vector containing the names of methods to
-#'  exclude from the heatmap. Default is to exclude Scott methods which are 
+#'  exclude from the heatmap. Default is to exclude fdrreg methods which are 
 #'  only present in GWAS.
 #' @param fill character indicating the outcome variable of interest. Default
 #'  is 'propMaxRejections'. For 'FDR' and 'TPR' sb objects need to contain 
@@ -40,8 +40,8 @@ plotMethodRanks <- function(objects, colLabels, alpha = 0.10,
                                      "meanRejections",
                                      "FDR", "TPR", "TNR"),
                             linePlot = FALSE,
-                            excludeMethods = c("scott-theoretical",
-                                               "scott-empirical"),
+                            excludeMethods = c("fdrreg-t",
+                                               "fdrreg-e"),
                             rowOrder = NULL,
                             tableOnly = FALSE,
                             Nlabel = TRUE,

### 2018-07-23 datasets/R/plotsim_aggupset.R

@@ -140,13 +140,6 @@ sb2hits <- function(x, a, s) {
             dplyr::select(-`bl-df02`,  -`bl-df04`,  -`bl-df05`) %>%
             dplyr::rename(bl = `bl-df03`)
     }
-    ## rename scott methods for easier parsing later, drop unadjusted p-vals from comparison
-    if ("scott-theoretical" %in% names(ht)) {
-        ht <- dplyr::rename(ht, "scott-t" = "scott-theoretical")
-    }
-    if ("scott-empirical" %in% names(ht)) {
-        ht <- dplyr::rename(ht, "scott-e" = "scott-empirical")
-    }
     ht <- dplyr::select(ht, -unadjusted)
     as.data.frame(ht)
 }

### 2018-07-23 datasets/R/plotsim_average.R

@@ -18,8 +18,8 @@
 #' @param clean_names logical whether to clean-up method names in 'blabel'
 #'        column. The tsb table must only contain one column with each of the
 #'        following labels, else method labels will not be changed: 'ashq',
-#'        'bh', 'bl', 'ihw', 'lfdr', 'qvalue', 'scott-empirical',
-#'        'scott-theoretical'. (default = FALSE)
+#'        'bh', 'bl', 'ihw', 'lfdr', 'qvalue', 'fdrreg-e',
+#'        'fdrreg-t'. (default = FALSE)
 #' @param errorBars logical indicating whether to include error bars 
 #' @param palette data.frame containing the color palette - should contain 
 #'        three columns: 'Method', 'col', and 'lty'
@@ -64,8 +64,8 @@ plotsim_average <- function(tsb, met, filter_set = NULL, merge_ihw = TRUE,
         ulabs <- unique(tsba$blabel)
         vlabs <- c('ashq', 'bh', 'bl', 'ihw', 'lfdr', 'qvalue')
         clabs <- c("ASH q-value", "Benjamini-Hochberg", "Boca-Leek", "IHW", "local FDR", "Storey's q-value")
-        if (any(grepl("scott", ulabs))) {
-            vlabs <- c(vlabs, 'scott-empirical', 'scott-theoretical')
+        if (any(grepl("fdrreg", ulabs))) {
+            vlabs <- c(vlabs, 'fdrreg-e', 'fdrreg-t')
             clabs <- c(clabs, "FDRreg (emp)", "FDRreg (theor)")
         }
         lcnts <- sapply(vlabs, function(x) { grep(paste0("^", x), ulabs, value=TRUE) })

### 2018-07-23 datasets/R/plotsim_single.R

@@ -15,8 +15,8 @@
 #' @param clean_names logical whether to clean-up method names. The sba table
 #'        must only contain one column for each of the following labels,
 #'        else method labels will not be changed: 'ashq',
-#'        'bh', 'bl', 'ihw', 'lfdr', 'qvalue', 'scott-empirical',
-#'        'scott-theoretical'. (default = FALSE)
+#'        'bh', 'bl', 'ihw', 'lfdr', 'qvalue', 'fdrreg-e',
+#'        'fdrreg-t'. (default = FALSE)
 #'
 #' @return
 #' a ggplot object.
@@ -42,8 +42,8 @@ plotsim_single <- function(sba, df, strat = "effect_size", clean_names = FALSE) 
         ulabs <- unique(sba$method)
         vlabs <- c('ashq', 'bh', 'bl', 'ihw', 'lfdr', 'qvalue')
         clabs <- c("ASH q-value", "Benjamini-Hochberg", "Boca-Leek", "IHW", "local FDR", "Storey's q-value")
-        if (any(grepl("scott", ulabs))) {
-            vlabs <- c(vlabs, 'scott-empirical', 'scott-theoretical')
+        if (any(grepl("fdrreg", ulabs))) {
+            vlabs <- c(vlabs, 'fdrreg-e', 'fdrreg-t')
             clabs <- c(clabs, "FDRreg (emp)", "FDRreg (theor)")
         }
         lcnts <- sapply(vlabs, function(x) { grep(paste0("^", x), ulabs, value=TRUE) })

### 2018-07-23 datasets/GSEA/gsea-human.Rmd
Add filter to GSEA multiple testing

@@
 names(genes) <- rownames(res)
+sum(genes, na.rm=TRUE)
 ```
@@
 
+We filter out really small gene sets and those gene sets which have no DE genes
+in them (many of these are set to 1).
+
+```{r}
+goRes <- goRes %>%
+  dplyr::filter(numDEInCat > 0, numInCat > 10 )
+```
+
 ## Covariate Diagnostics
@@
 ```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
-## filter numInCat below only used for plotting
-## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
-#dplyr::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)
-dplyr::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
-    strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=18)
+strat_hist(goRes, pval="over_represented_pvalue", covariate="numInCat", maxy=7)
 ```
@@
 ```{r assumptions-rand, message=FALSE, fig.width=10, fig.height=3.2}
-## filter numInCat below only used for plotting
-## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
-#dplyr::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)
-dplyr::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
-    strat_hist( pval="over_represented_pvalue", covariate="rand_covar", maxy=10)
+strat_hist(goRes, pval="over_represented_pvalue", covariate="rand_covar", maxy=7)
 ```
@@
 
-We exclude `ashr`, `scott-theoretical` and `scott-empirical` from the analysis because
+We don't include `ashq`, `fdrreg-e` and `fdrreg-t` from the analysis because
 the necessary assumptions are not met in the current case study. Namely, effect sizes

### 2018-07-23 datasets/GSEA/gsea-mouse.Rmd
Add filter to GSEA multiple testing

@@
 ```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
-## filter numInCat below only used for plotting
-## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
-#dplyr::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)
-dplyr::filter(goRes, numDEInCat > 0, numInCat > 10) %>%
-    strat_hist(pval = "over_represented_pvalue", covariate = "numInCat", maxy=35)
+strat_hist(goRes, pval = "over_represented_pvalue", covariate = "numInCat", maxy=7)
 ```
@@
 ```{r assumptions-rand, message=FALSE, fig.width=10, fig.height=3.2}
-## filter numInCat below only used for plotting
-## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
-#dplyr::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)
-dplyr::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
-    strat_hist( pval="over_represented_pvalue", covariate="rand_covar", maxy=10)
+strat_hist(goRes, pval="over_represented_pvalue", covariate="rand_covar", maxy=10)
 ```

### 2018-07-23 manuscript_figures/manuscript_figures.Rmd
Remove points from case study vs sim boxplot fig

@@
                    alpha = 0.3, width = 0.04, outlier.shape = NA) +
-  geom_quasirandom(aes(x = prop_yeastsim, y = prop_casestudy, color = Method, shape = casestudy),
-                   alpha = 0.6, cex=3, width = 0.002,  groupOnX = TRUE,
-                   varwidth = TRUE) +
   scale_color_manual(values=col) +

### 2018-07-23 manuscript_figures/manuscript_figures.Rmd
Add impact of covariates averaged over datasets in casestudies

@@
                       ifelse(grepl("abun", filename), "abundance", "ubiquity"),  cov))%>%
-  mutate(cov = ifelse(casestudy == "RNAseq" & inform != "uninf", "mean",  cov))
+  mutate(cov = ifelse(casestudy == "RNAseq" & inform != "uninf", "mean",  cov)) %>%  
+  mutate(cov = ifelse(cov == "det", "detection-rate", cov))
 
@@
               fill="meanRank", annotate=NULL, alpha =0.10) %>%
-  rename(casestudy = "filename") %>%
+  rename(filename = "casestudy") %>%
+  #rename(casestudy = "filename") %>%
   mutate(filename = gsub(".rds", "", filename))
@@
 dftab <- left_join(tab, df, by="filename") %>%
-  filter(method %in% cov.methods) %>%
-  select(filename, casestudy, dataset, method, cov, inform, nrejects) %>%
-  mutate(id = paste(casestudy, dataset, method, sep = "-"))
+  dplyr::filter(method %in% cov.methods) %>%
+  dplyr::select(filename, casestudy, dataset, method, cov, inform, nrejects) %>%
+  dplyr::mutate(id = paste(casestudy, dataset, method, sep = "-"))
 
 uninftab <- dftab %>% 
-  filter(inform == "uninf") %>%
-  select(id, nrejects) %>%
-  rename(nrejects = "uninf")
+  dplyr::filter(inform == "uninf") %>%
+  dplyr::select(id, nrejects) %>%
+  dplyr::rename("uninf" = "nrejects")
+  #dplyr::rename(nrejects = "uninf")
 
 inftab <- dftab %>%
-  filter(inform == "inf") %>%
-  select(-inform)
+  dplyr::filter(inform == "inf") %>%
+  dplyr::select(-inform)
 
@@
 dftab <- left_join(inftab, uninftab, by = "id") %>%
-  mutate(diff_inf = nrejects - uninf) %>%
-  mutate(diff_inf_percent = diff_inf / nrejects) %>%
-  mutate(method = factor(method, levels = rowOrder)) %>%
-  group_by(casestudy, cov) %>%
+  dplyr::mutate(diff_inf = nrejects - uninf) %>%
+  dplyr::mutate(diff_inf_percent = diff_inf / nrejects) %>%
+  dplyr::mutate(method = factor(method, levels = rowOrder)) %>%
+  dplyr::group_by(casestudy, cov) %>%
   complete(method, dataset)
@@
 
+Make the Figure above, but average over all datasets within a case study.
+
+```{r}
+dftab <- dftab %>%
+  group_by(casestudy, method) %>%
+  summarize(diff_inf_percent = mean(diff_inf_percent, na.rm = TRUE))
+
+maxVals <- dftab %>% 
+  group_by(casestudy) %>%
+  summarize(top = max(diff_inf_percent, na.rm=TRUE))
+
+dftab <- left_join(dftab, maxVals, by = c("casestudy")) %>%
+  mutate(top = ifelse(diff_inf_percent == top, diff_inf_percent, NA))
+
+Fig_impactofinfcov <- ggplot(dftab, aes(x = casestudy, y = method, fill = diff_inf_percent)) + 
+        geom_tile() + 
+        scale_fill_gradientn("% Increase\nin rejections",
+                             labels = scales::percent, trans="log1p",
+                             colours = c("darkblue","white","darkred"), 
+                             values = scales::rescale(c(-.05,0,1)),
+                             guide = "colorbar", limits=c(-.05,1)) +
+        theme_bw() +
+        xlab("Case Study") +
+        ylab("Method") +
+        ggtitle("Impact of informative covariate on number of rejections") +
+        theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
+        geom_text(aes(label = ifelse(is.na(top), "", scales::percent(top))),
+                  color = "black", size = 2)
+
+Fig_impactofinfcov
+ggsave(file.path(outdir, "Figure_impactofinfcov_avg.pdf"), width = 5.75, height = 4)
+```
+
 # Informativeness of covariates in Yeast & Polyester simulation

### 2018-07-23 manuscript_figures/manuscript_figures.Rmd
Adjust y-axis of figure 6c after filtering GSEA tests

@@
 pc <- strat_hist(res, pvalue = "pval",
-           covariate = "ind_covariate", maxy = 35, axislabs = FALSE) +
+           covariate = "ind_covariate", maxy = 7, axislabs = FALSE) +
   draw_plot_label("  GSEA case study: Mouse", 0.5, 1, hjust = 0.5, fontface = "bold" , size = 12) +

### 2018-07-24 datasets/RNA-Seq/polyester-simulation.Rmd
Weight DE sampling by informative covariate; switch to mclapply

@@
 # # set up parallel backend:
-library(BiocParallel)
+library(parallel)
 nCores <- 20
-register(MulticoreParam(workers = nCores))
 ```
@@
 #' @param rseed random seed
+#' @param ngenes the number of genes to simulate. If this number is greater than 
+#'  the number of genes in the count matrix, then the polyester parameters will be
+#'  drawn with replacement from the available pool in order to obtain the requested
+#'  number of genes.
 simulateOneSplit <- function(X, rseed, 
-                             nDE, sampleSize, bd, uninformativeCovariate = FALSE){
-  ngenes <- 10000
-
+                             nDE, sampleSize, bd, uninformativeCovariate = FALSE,
+                             ngenes = nrow(counts)){
   uninf_covar <- rnorm(ngenes)
@@
 
-  gs <- sample(1:nrow(counts), ngenes, replace = TRUE)
+  gs <- sample(1:nrow(counts), ngenes, replace = as.logical(ngenes > nrow(counts)))
   mu <- paramPolyester$mu[gs]
@@
   if(nDE > 0){
-    # pick random set of nDE genes to add signal to
-    DE <- sample(1:nrow(dat_alt), nDE)
+    wts <- rep(1, nrow(dds_test))
+    if(!uninformativeCovariate)
+      wts <- log10(rowMeans(counts)+1)
+
+    DE <- sample(1:nrow(dat_alt), nDE, prob = wts)
     truth[DE] <- TRUE
@@
 if (!file.exists(resfile_n5)){
-  null5 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit,  
-                    nDE=nDE, sampleSize=sampleSize, bd=bd)
+  null5 <- mclapply(X=1:B, rseed=rseed, FUN=simulateOneSplit,  
+                    nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores = nCores)
   saveRDS(null5, file=resfile_n5)
@@
 if (!file.exists(resfile_n10)){
-  null10 <- bplapply(X=1:B, rseed=837, FUN=simulateOneSplit,
-                     nDE=nDE, sampleSize=sampleSize, bd=bd)
+  null10 <- mclapply(X=1:B, rseed=837, FUN=simulateOneSplit,
+                     nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
   saveRDS(null10, file=resfile_n10)
@@
 if (!file.exists(resfile_d5)){
-  de5 <- bplapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, 
-                  nDE=nDE, sampleSize=sampleSize, bd=bd)
+  de5 <- mclapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, 
+                  nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
   saveRDS(de5, file=resfile_d5)
@@
 if (!file.exists(resfile_d10)){
-  de10 <- bplapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, 
-                   nDE=nDE, sampleSize=sampleSize, bd=bd)
+  de10 <- mclapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, 
+                   nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores = nCores)
   saveRDS(de10, file=resfile_d10)
@@
 if (!file.exists(resfile_n5_uninfCov)){
-  null5 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
+  null5 <- mclapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
                     nDE=nDE, sampleSize=sampleSize, 
-                   bd=bd, uninformativeCovariate = TRUE)
+                   bd=bd, uninformativeCovariate = TRUE, mc.cores=nCores)
   saveRDS(null5, file=resfile_n5_uninfCov)
@@
 if (!file.exists(resfile_n10_uninfCov)){
-  null10 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
+  null10 <- mclapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
                      nDE=nDE, sampleSize=sampleSize, 
-                   bd=bd, uninformativeCovariate = TRUE)
+                   bd=bd, uninformativeCovariate = TRUE, mc.cores=nCores)
   saveRDS(null10, file=resfile_n10_uninfCov)
@@
 if (!file.exists(resfile_d5_uninfCov)){
-  de5 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
+  de5 <- mclapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
                   nDE=nDE, sampleSize=sampleSize, 
-                  bd=bd, uninformativeCovariate = TRUE)
+                  bd=bd, uninformativeCovariate = TRUE, mc.cores=nCores)
   saveRDS(de5, file=resfile_d5_uninfCov)
@@
 if (!file.exists(resfile_d10_uninfCov)){
-  de10 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
+  de10 <- mclapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
                    nDE=nDE, sampleSize=sampleSize, 
-                   bd=bd, uninformativeCovariate = TRUE)
+                   bd=bd, uninformativeCovariate = TRUE, mc.cores = nCores)
   saveRDS(de10, file=resfile_d10_uninfCov)

### 2018-07-24 datasets/RNA-Seq/yeast-simulation.Rmd
Weight DE sampling by informative covariate; switch to mclapply

@@
 # set up parallel backend
-library(BiocParallel)
+library(parallel)
 nCores <- 20
-register(MulticoreParam(workers = nCores))
 ```
@@
   # select a random subset of 20 WT samples
-  dds_test <- dds[,colData(dds)$condition == "Snf2"]
-  dds_test <- dds_test[,sample(1:ncol(dds_test), sampleSize*2)]
+  dds_test <- dds[,sample(1:ncol(dds), sampleSize*2)]
 
@@
     # pick random set of nDE genes to add signal to
-    DE <- sample(1:nrow(dds_test), nDE)
+    wts <- rep(1, nrow(dds_test))
+    if(!uninformativeCovariate)
+      wts <- log10(rowMeans(counts(dds_test, normalize = TRUE))+1)
+    DE <- sample(1:nrow(dds_test), nDE, prob = wts)
     truth[DE] <- TRUE
@@
     log2FC <- rep(0, nrow(dds_test))
-    log2FC[DE] <- resFULL$log2FoldChange[DE]
+    log2FC[DE] <- resFULL$log2FoldChange[pzero < 0.5][DE]
 
@@
   dds_test <- DESeq(dds_test, parallel = FALSE)
-  resultsNames(dds_test) # lists the coefficients
   resTEST <- results(dds_test, name="fake_B_vs_A", independentFiltering = FALSE)
@@
 ualpha <- 0.05
+
+# only keep one condition for subsetting in simulations that follow 
+dds <- dds[,colData(dds)$condition == "Snf2"]
 ```
@@
 if (!file.exists(resfile_n5)){
-  null5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
-                    nDE=nDE, sampleSize=sampleSize, bd=bd)
+  null5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+                    nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
   saveRDS(null5, file = resfile_n5)
@@
 if (!file.exists(resfile_n10)){
-  null10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
-                    nDE=nDE, sampleSize=sampleSize, bd=bd)
+  null10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+                    nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
   saveRDS(null10, file = resfile_n10)
@@
 if (!file.exists(resfile_d5)){
-  de5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
-                  nDE=nDE, sampleSize=sampleSize, bd=bd)
+  de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+                  nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
   saveRDS(de5, file = resfile_d5)
@@
 if (!file.exists(plotfile)){
-  hists <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+  hists <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                    nDE=nDE, sampleSize=sampleSize, bd=bd, 
-                   pvalHists = TRUE)
+                   pvalHists = TRUE, mc.cores = nCores)
   pdf(plotfile, width=8, height=4)
@@
 if(!file.exists(resfile_d10)){
-  de10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
-                  nDE=nDE, sampleSize=sampleSize, bd=bd)
+  de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+                  nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
   saveRDS(de10, file = resfile_d10)
@@
 if (!file.exists(resfile_n5_uninfCov)){
-  null5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed,
+  null5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed,
                    nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   uninformativeCovariate = TRUE)
+                   uninformativeCovariate = TRUE, mc.cores = nCores)
   saveRDS(null5, file = resfile_n5_uninfCov)
@@
 if (!file.exists(resfile_n10_uninfCov)){
-  null10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+  null10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                    nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   uninformativeCovariate = TRUE)
+                   uninformativeCovariate = TRUE, mc.cores=nCores)
   saveRDS(null10, file = resfile_n10_uninfCov)
@@
 if (!file.exists(resfile_d5_uninfCov)){
-  de5 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+  de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd,
-                  uninformativeCovariate = TRUE)
+                  uninformativeCovariate = TRUE, mc.cores=nCores)
   saveRDS(de5, file = resfile_d5_uninfCov)
@@
 if(!file.exists(resfile_d10_uninfCov)){
-  de10 <- bplapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+  de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                    nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   uninformativeCovariate = TRUE)
+                   uninformativeCovariate = TRUE, mc.cores=nCores)
   saveRDS(de10, file = resfile_d10_uninfCov)

### 2018-07-24 datasets/simulations/simulations-informative-cosine.Rmd

@@
 }
+res_i <- lapply(res, `[[`, "informative")
+res_u <- lapply(res, `[[`, "uninformative")
 ```
@@
 ```{r gauss-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r gauss-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
 
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
-    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
+    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
 } else {
@@
 ```{r t5-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r t5-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
 
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 ```{r t11-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r t11-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
 
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 ```{r chisq4-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r chisq4-metrics-covlineplot, results = "hide"}
-covariateLinePlot(lapply(res, function(x) { x[, -which(colnames(x) == "ashq")] }),
+covariateLinePlot(lapply(res_i, function(x) { x[, -which(colnames(x) == "ashq")] }),
                   alpha = ualpha, covname = "effect_size")
 
-covariateLinePlot(lapply(res, function(x) { x[, -which(colnames(x) == "ashq")] }),
+covariateLinePlot(lapply(res_i, function(x) { x[, -which(colnames(x) == "ashq")] }),
                   alpha = ualpha, covname = "ind_covariate")
@@
 if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = c(excludeSet, "ashq")) >= 3) {
-    aggupset(lapply(res, function(x) { x[, -which(colnames(x) == "ashq")] }),
+    aggupset(lapply(res_i, function(x) { x[, -which(colnames(x) == "ashq")] }),
              alpha = ualpha, supplementary = FALSE, return_list = FALSE)

### 2018-07-24 datasets/simulations/simulations-varyingntests.Rmd

---
title: "Simulation Study: Varying Number of Tests"
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

# Summary

In this set of simulations, we consider settings with varying number of hypothesis
tests. Since many newer FDR controlling approaches require fitting some model to
the independent covariates, they may be sensitive to lower number of tests. An
informative covariate is included in the setting as described in
`simulations-informative-sine.Rmd`.

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(SummarizedBenchmark)
library(parallel)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
resdir <- "results"
dir.create(resdir, showWarnings = FALSE, recursive = TRUE)

## intermediary files we create below
n100_file <- file.path(resdir, "varyingntests-lowsignal-benchmark-n100.rds")
n500_file <- file.path(resdir, "varyingntests-lowsignal-benchmark-n500.rds")
n1000_file <- file.path(resdir, "varyingntests-lowsignal-benchmark-n1000.rds")
n5000_file <- file.path(resdir, "varyingntests-lowsignal-benchmark-n5000.rds")
n10000_file <- file.path(resdir, "varyingntests-lowsignal-benchmark-n10000.rds")
n50000_file <- file.path(resdir, "varyingntests-lowsignal-benchmark-n50000.rds")

## number of cores for parallelization
cores <- 20
B <- 100

## define bechmarking design
bd <- initializeBenchDesign()
```

As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
for simulations with Gaussian noise. Again, we include both
`nulltype = "empirical"` and `nulltype = "theoretical"`. Since all settings in this
series of simulations use test statistics simulated with Gaussian noise, we include
Scott's FDR Regression in all of the comparisons.

```{r}
bdplus <- bd
bdplus <- addBMethod(bdplus, "scott-theoretical",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'theoretical',
                     control = list(lambda = 0.01))
bdplus <- addBMethod(bdplus, "scott-empirical",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'empirical',
                     control = list(lambda = 0.01))
```

All simulation settings will share the following parameters.

```{r parameters-shared}
tstat <- rnorm_generator(2)       # functional: dist of alternative test stats
tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
icovariate <- runif               # functional: independent covariate
pi0 <- pi0_sine(0.90)             # numeric: proportion of null hypotheses
```

Simulation results will be presented excluding a subset of methods, and
for certain plots (upset plots), a single alpha cutoff will be used.

```{r}
excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
ualpha <- 0.05
```

# 100 Tests

First, we consider the setting where 100 hypotheses are tested.

## Data Simulation

```{r n100-parameters}
m <- 100                        # integer: number of hypothesis tests
seed <- 1010
```

We next run the simulations.

```{r n100-run-simulation}
if (file.exists(n100_file)) {
    res <- readRDS(n100_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = n100_file)
}
res_i <- lapply(res, `[[`, "informative")
res_u <- lapply(res, `[[`, "uninformative")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r n100-one-simulation}
onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, n100-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, n100-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r n100-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r n100-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r n100-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

# 500 Tests

Next, we consider the setting where 500 hypotheses are tested.

## Data Simulation

```{r n500-parameters}
m <- 500                        # integer: number of hypothesis tests
seed <- 5050
```

We next run the simulations.

```{r n500-run-simulation}
if (file.exists(n500_file)) {
    res <- readRDS(n500_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = n500_file)
}
res_i <- lapply(res, `[[`, "informative")
res_u <- lapply(res, `[[`, "uninformative")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r n500-one-simulation}
onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, n500-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, n500-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r n500-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r n500-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r n500-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

# 1000 Tests

Next, we consider the setting where 1000 hypotheses are tested.

## Data Simulation

```{r n1000-parameters}
m <- 1000                        # integer: number of hypothesis tests
seed <- 10001
```

We next run the simulations.

```{r n1000-run-simulation}
if (file.exists(n1000_file)) {
    res <- readRDS(n1000_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = n1000_file)
}
res_i <- lapply(res, `[[`, "informative")
res_u <- lapply(res, `[[`, "uninformative")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r n1000-one-simulation}
onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, n1000-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, n1000-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r n1000-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r n1000-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r n1000-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

# 5000 Tests

Next, we consider the setting where 5000 hypotheses are tested.

## Data Simulation

```{r n5000-parameters}
m <- 5000                        # integer: number of hypothesis tests
seed <- 50005
```

We next run the simulations.

```{r n5000-run-simulation}
if (file.exists(n5000_file)) {
    res <- readRDS(n5000_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = n5000_file)
}
res_i <- lapply(res, `[[`, "informative")
res_u <- lapply(res, `[[`, "uninformative")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r n5000-one-simulation}
onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, n5000-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, n5000-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r n5000-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r n5000-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r n5000-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

# 10000 Tests

Next, we consider the setting where 10000 hypotheses are tested.

## Data Simulation

```{r n10000-parameters}
m <- 10000                        # integer: number of hypothesis tests
seed <- 1515
```

We next run the simulations.

```{r n10000-run-simulation}
if (file.exists(n10000_file)) {
    res <- readRDS(n10000_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = n10000_file)
}
res_i <- lapply(res, `[[`, "informative")
res_u <- lapply(res, `[[`, "uninformative")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r n10000-one-simulation}
onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, n10000-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, n10000-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r n10000-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r n10000-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r n10000-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

# 50000 Tests

Next, we consider the setting where 50000 hypotheses are tested.

## Data Simulation

```{r n50000-parameters}
m <- 50000                        # integer: number of hypothesis tests
seed <- 5555
```

We next run the simulations.

```{r n50000-run-simulation}
if (file.exists(n50000_file)) {
    res <- readRDS(n50000_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = n50000_file)
}
res_i <- lapply(res, `[[`, "informative")
res_u <- lapply(res, `[[`, "uninformative")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r n50000-one-simulation}
onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, n50000-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, n50000-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r n50000-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r n50000-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r n50000-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

# Session Info

```{r}
sessionInfo()
```

### 2018-07-25 datasets/RNA-Seq/polyester-simulation.Rmd
Fix typo introduced in previous commit
Make weighting scheme the same in informative and uninformative sims

@@
   if(nDE > 0){
-    wts <- rep(1, nrow(dds_test))
-    if(!uninformativeCovariate)
-      wts <- log10(rowMeans(counts)+1)
-
+    wts <- log10(rowMeans(dat_alt)+1)
     DE <- sample(1:nrow(dat_alt), nDE, prob = wts)

### 2018-07-25 datasets/RNA-Seq/yeast-simulation.Rmd
Make weighting scheme the same in informative and uninformative sims

@@
     # pick random set of nDE genes to add signal to
-    wts <- rep(1, nrow(dds_test))
-    if(!uninformativeCovariate)
     wts <- log10(rowMeans(counts(dds_test, normalize = TRUE))+1)

### 2018-07-25 manuscript_figures/manuscript_figures.Rmd
Add informative covariate impact line plot; remove unnecessary figs

@@
         theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
-              legend.position = c(0.96, 0.93), legend.justification = c(1, 1),
+              legend.position = c(0.98, 0.93), legend.justification = c(1, 1),
               legend.background = element_rect(fill = scales::alpha("white", 3/4), color = 'black'),
@@
 
-xtable(tabS3)
+print(xtable(tabS3), include.rownames=FALSE)
 
@@
                          mid = "white", 
-                         low = "darkblue", 
+                         low = "#2171B5", 
                          midpoint = alpha) +
@@
 ggdraw() +
-  draw_plot(pfdr, 0.51, 0, 0.24, 1) +
-  draw_plot(ptpr, 0.75, 0, 0.24, 1) +
-  draw_plot(Fig4All, 0, 0, 0.5, 1) +
-  draw_plot_label(c("A", "B"), c(0, 0.51), c(1, 1), size = 15) +
-  draw_plot_label("Simulation", 0.72, 0.06, size =11, fontface = "plain")
-ggsave(file.path(outdir, "Figure4.pdf"), width=11, height=5)
+  draw_plot(pfdr, 0.53, 0, 0.23, 1) +
+  draw_plot(ptpr, 0.76, 0, 0.23, 1) +
+  draw_plot(Fig4All, 0, 0, 0.52, 1) +
+  draw_plot_label(c("A", "B"), c(0, 0.53), c(1, 1), size = 15) +
+  draw_plot_label("Simulation", 0.69, 0.06, size =11, fontface = "plain")
+ggsave(file.path(outdir, "Figure4.pdf"), width=9, height=5)
 
@@
 ```{r, fig.width = 6, fig.height = 4}
-library(ggbeeswarm)
-
 casestudy.tab <- mutate(casestudy.tab, prop_casestudy = mean_prop) %>%
@@
 
-ggplot(rank.tab , 
-       aes(y = prop_casestudy, x = method, group = method, fill = Method)) +
-  geom_boxplot(alpha=0.7) +
-  scale_fill_manual(values=col) +
-  theme_bw() +
-  xlab("Method (ordered by rejections in yeast simulation (n=5))") +  
-  ylab("Proportion of maximum rejections in casestudies")
-ggsave(file.path(outdir, "Figure_simVcasestudyRanks.pdf"), width=6, height=4)
-
-ggplot(rank.tab , 
-       aes(y = prop_casestudy, x = method, group = method, color = Method)) +
-  geom_boxplot(aes(x = prop_yeastsim, group = as.factor(prop_yeastsim), y = prop_casestudy),
-                   alpha = 0.3, width = 0.04, outlier.shape = NA) +
+ggplot(rank.tab %>% filter(method != "bonf"), 
+       aes(y = prop_casestudy, x = prop_yeastsim, group = prop_yeastsim, color=Method)) +
+  geom_boxplot(aes(x = prop_yeastsim, y = prop_casestudy),
+                   alpha = 0.3, position = position_dodge(width = 0),
+                   width = 0.02) +
   scale_color_manual(values=col) +
   theme_bw() +
-  xlab("Proportion of maximum rejections in yeast simulation (n=5)") +  
+  xlab("Proportion of maximum rejections in yeast simulation (n=5s)") +  
   ylab("Proportion of maximum rejections in casestudies") +
   labs(shape = "Case Study") 
-ggsave(file.path(outdir, "Figure_simVcasestudyRanks_points.pdf"), width=6, height=4)
+ggsave(file.path(outdir, "Figure_simVcasestudyRanks_scaled.pdf"), width=6, height=4)
 
@@
 
-## Case-study specific plot analagous to Figure 2
-
-Figure 2 panels C-E for a case study. For now picking GWAS.
-
-```{r Figure2.5,  fig.width = 10, fig.height = 4}
+## Uninformative covariate impact on simulations
+
+Here we'll contrast the results of informative and uninformative covariates 
+on the yeast and polyester simulation (DE sample size 5). We'll plot
+differences in TPR and FDR due to informative covariate by alpha level.
+
+```{r, fig.width = 10, fig.height = 4}
+for (org in c("yeast", "polyester")){
+
   objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
-resfile <- objects[grepl("GWAS", objects) & grepl("maf", objects)] 
-res <- readRDS(file=resfile)
-assayNames(res) <- "qvalue"
-res <- addDefaultMetrics(res)
-
-# Figure 2C: Total number of discoveries varied across method
-p3 <- rejections_scatter(res, supplementary = FALSE) +
-      theme(legend.position="none")+ 
-  ggtitle("Number of rejections by alpha cutoff") + 
-  xlab("alpha cutoff") +
+  resfile <- objects[grepl(org, objects) ] 
+  resfile <- resfile[grepl("5", resfile) & !grepl("null", resfile)]
+
+  # regexp to construct master data table separating case study datasets, methods
+  # and covariates
+  obj.names <- str_split(resfile, pattern = "/")
+  casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
+  filename <- gsub(".rds", "", unlist(lapply(obj.names, function(x) x[4] )))
+  dataset <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[3]]))
+  inform <-  ifelse(grepl("uninf", resfile), "uninf", "inf")
+  type <-  ifelse(grepl("null", resfile), "null", "de")
+  size <-  ifelse(grepl("10", resfile), "10", "5")
+  aseq <- seq(0.01, 0.10, by = 0.01)
+
+  tab <- data.frame()
+  for (l in seq_along(resfile)){
+    x <- readRDS(resfile[l])
+
+    tmp2 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
+                    fill =  "FDR", annotate = NULL, 
+                    alpha = aseq) %>%
+      mutate(fdr = nrejects) %>%
+      select(method, casestudy, replicate, fdr, alpha)
+    tmp3 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
+                    fill =  "TNR", annotate = NULL, 
+                    alpha = aseq) %>%
+      mutate(tnr = nrejects) %>%
+      select(method, casestudy, replicate, tnr, alpha)
+
+    if (type[l] != "null"){
+      tmp1 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
+                      fill =  "TPR", annotate = NULL, 
+                      alpha = aseq) %>%
+        mutate(tpr = nrejects) %>%
+        select(method, casestudy, replicate, tpr, alpha)
+
+      tmp <- left_join(tmp1, tmp2, by = c("method", "casestudy", "replicate", "alpha")) %>%
+        left_join(tmp3, by = c("method", "casestudy", "replicate", "alpha")) %>%
+        mutate(inform = inform[l],
+               type = type[l],
+               size = size[l])
+    }else{
+      tmp <- left_join(tmp2, tmp3, by = c("method", "casestudy", "replicate", "alpha")) %>%
+        mutate(inform = inform[l],
+               type = type[l],
+               size = size[l])
+    }
+
+    tab <- bind_rows(tab, tmp)
+  }
+
+myspread <- function(df, key, value) {
+    # quote key
+    keyq <- rlang::enquo(key)
+    # break value vector into quotes
+    valueq <- rlang::enquo(value)
+    s <- rlang::quos(!!valueq)
+    df %>% gather(variable, value, !!!s) %>%
+        unite(temp, !!keyq, variable) %>%
+        spread(temp, value)
+}
+
+dftab <- tab %>%
+  select(-casestudy) %>%
+  myspread(inform, c(tpr, fdr, tnr))
+
+dftab <- dftab %>%
+  mutate(tpr_diff = inf_tpr - uninf_tpr,
+         fdr_diff = inf_fdr - uninf_fdr,
+         tnr_diff = inf_tnr - uninf_tnr) %>%
+  group_by(type, size, method, alpha) %>%
+  summarize(tpr_se = sd(tpr_diff)/n(),
+            fdr_se = sd(fdr_diff)/n(),
+            tnr_se = sd(tnr_diff)/n(),
+            tpr_diff = mean(tpr_diff),
+            fdr_diff = mean(fdr_diff),
+            tnr_diff = mean(tnr_diff))
+
+col <- as.character(candycols$col)
+names(col) <- as.character(candycols$Method)
+lty <- as.character(candycols$lty)
+names(lty) <- as.character(candycols$Method)
+
+lineplot_fdr <- ggplot(dftab,
+                       aes(x = alpha, y = fdr_diff, color = method, linetype = method)) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    geom_line(alpha = 0.85) +
+    geom_errorbar(width=.003, alpha = 0.5, aes(ymin = fdr_diff - fdr_se, 
+                                ymax = fdr_diff + fdr_se)) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
+    scale_y_continuous(expression(paste("Mean ", Delta, "FDR (informative-uninformative)")), 
+                       breaks = seq(-1, 1, .005), labels = scales::percent) + 
+    ggtitle("FDR difference with informative covariate") +
+    theme_classic() +
     theme(axis.title = element_text(face="bold", size = 10),
-        plot.title = element_text(face="bold", size = 14))
-
-# Figure 2D: Some methods are tailored to detecting smaller effect sizes
-p4 <- covariateLinePlot(res, alpha=alpha, covname="effect_size", trans="log1p",
-                        transx = "exp", nbins=25)
-leg <- get_legend(p4)
-p4 <- p4 + theme(legend.position="none") + 
-  ggtitle("Mean % significant by effect size %ile") + 
+          plot.title = element_text(face="bold", size = 11, hjust = 0.8))
+
+lineplot_tpr <- ggplot(dftab, aes(x = alpha, y = tpr_diff, color = method, linetype = method)) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    geom_errorbar(width=.003, alpha = 0.5, aes(ymin = tpr_diff - tpr_se, 
+                                ymax = tpr_diff + tpr_se)) +
+    geom_line(alpha = 0.85) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
+    scale_y_continuous(expression(paste("Mean ", Delta, "TPR (informative-uninformative)")), 
+                       labels = scales::percent) + 
+    ggtitle("TPR difference with informative covariate") +
+    theme_classic() +
     theme(axis.title = element_text(face="bold", size = 10),
-        plot.title = element_text(face="bold", size = 14)) +
-  xlab("Effect size percentile")
-
-# Figure 2E - upsetr plot - save separately
-pdf(file.path(outdir, "FigureS0C.pdf"), width = 10, height = 4, onefile=FALSE)
-plotFDRMethodsOverlap(res, 
-                      alpha=alpha, nsets=ncol(res),
-                      order.by="freq", decreasing=TRUE,
-                      supplementary=FALSE, nintersects = 18) 
-grid.text("Overlap for GWAS Case study, MAF covariate", x = 0.6, y = 0.97, 
-          gp = gpar(fontsize = 16, fontface = "bold"))
-grid.text("C", x = 0.02, y = 0.97, 
-          gp = gpar(fontsize = 24, fontface = "bold")) 
-dev.off()
-
-# Figure 2 (ALL)
-Fig2 <- plot_grid(p3, p4, ncol = 2, 
-          labels = LETTERS[1:2], label_size = 20)
-Fig2 <- plot_grid(Fig2, leg, ncol = 2, rel_widths = c(1, 0.2))
-Fig2
-ggsave(file.path(outdir, "FigureS0A-B.pdf"), width=10, height=4)
-
-```
-
-## Uninformative covariate impact on simulations
-
-Here we'll contrast the results of informative and uninformative covariates 
-on the yeast simulation (null and DE sample size 5).
-
-```{r FigureS7,  fig.width = 10, fig.height = 7}
-objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
-resfile <- objects[grepl("yeast", objects) & 
-                   grepl("de5|null5", objects)] 
-de5 <- readRDS(file=resfile[grepl("de5", resfile) & !grepl("uninf", resfile)])
-de5_uninf <- readRDS(file=resfile[grepl("de5", resfile) & grepl("uninf", resfile)])
-null5 <- readRDS(file=resfile[grepl("null5", resfile) & !grepl("uninf", resfile)])
-null5_uninf <- readRDS(file=resfile[grepl("null5", resfile) & grepl("uninf", resfile)])
-excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
-
-res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
-res5d_u <- plotsim_standardize(de5_uninf, alpha = seq(0.01, 0.10, 0.01))
-res5n <- plotsim_standardize(null5, alpha = seq(0.01, 0.10, 0.01))
-res5n_u <- plotsim_standardize(null5_uninf, alpha = seq(0.01, 0.10, 0.01))
-
-pa <- plotsim_average(res5n_u, met="TNR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) +
-        expand_limits(y=0.93) +
-        xlab("") +
-        ggtitle("")
-
-pb <- plotsim_average(res5n, met="TNR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) +
-        theme(legend.position="none", plot.title = element_text(hjust = 0.5))+
-        expand_limits(y=.93) +
-        xlab("") +
-        ggtitle("")
-
-pc <- plotsim_average(res5d_u, met="FDR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) +
-        theme(legend.position="none") +
-        expand_limits(y=0.18) +
-        xlab("") +
-        ggtitle("")
-
-pd <- plotsim_average(res5d, met="FDR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) +
-        theme(legend.position="none") +
-        expand_limits(y=0.18) +
-        xlab("") +
-        ggtitle("")
-
-pe <- plotsim_average(res5d_u, met="TPR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) +
-        theme(legend.position="none") +
-        expand_limits(y=c(0.10,0.55)) +
-        ggtitle("")
-
-pf <- plotsim_average(res5d, met="TPR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) +
-        theme(legend.position="none") +
-        expand_limits(y=c(0.10,0.55)) +
-        ggtitle("")
-
-leg <- get_legend(pa)
-pa <- pa + theme(legend.position="none", plot.title = element_text(hjust = 0.5))
-
-FigT  <- plot_grid(pa, pb, ncol = 2, 
-          #labels = LETTERS[1:2], 
-          label_size = 20)
-
-FigB  <- plot_grid(pc, pd, pe, pf, ncol = 2, 
-          #labels = LETTERS[3:6], 
-          label_size = 20)
-
-Fig <- ggdraw() +
-  draw_plot(FigT + ylab("") + xlab(""), 0, 0.64, 1, 0.3) +
-  draw_plot(FigB + ylab("") + xlab(""), 0.005, 0, 1, 0.6) +
-  draw_plot_label(c("Uninformative Covariate", "Informative Covariate"), 
-                  c(0.14, 0.64), c(1,1), c(0,0), size = 15) +
-  draw_plot_label(c("Null Simulation", "DE Simulation"), 
-                  c(0.36, 0.37), c(0.96,0.62), size = 13)
-
-Fig <- plot_grid(Fig, leg, ncol = 2, rel_widths = c(1, 0.2))
-Fig
-ggsave(file.path(outdir, "FigureS7.pdf"), width=10, height=11)
-
-```
-
-## Within case-study heatmaps of method rankings
-
-Here we generate heatmaps for each case study, where each column represents a different
-dataset. Here we pick one representative covariate and method (if applicable) 
-for each case study. See the individual Rmds for more detailed plots comparing, e.g.
-different covariates on the same dataset.
-
-```{r}
-objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
-
-# grab casestudy names
-obj.names <- str_split(objects, pattern = "/")
-casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
-dataset <- unlist(lapply(obj.names, function(x) x[4] ))
-```
-
-### GWAS
-
-```{r, fig.width = 5, fig.height = 4.5}
-# by covariate
-wcs <- which(grepl("GWAS", casestudy))
-objs <- objects[wcs]
-labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-"), 
-                      function(x) x[2]))
-plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Covariate",
-                excludeMethods = NULL, Nlabel = FALSE, annotate = "propPossible") + 
-  ggtitle("GWAS") + 
-  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
-ggsave(file.path(outdir, "FigureS1.pdf"), width=5, height=4.5)
-```
-
-### GSEA
-
-```{r,  fig.width = 6, fig.height = 4.5}
-#by dataset
-wcs <- which(grepl("GSEA", casestudy))
-objs <- objects[wcs]
-labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
-                      function(x) x[1]))
-plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", 
-                Nlabel = FALSE, annotate = "propPossible") + 
-  ggtitle("GSEA") + 
-  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
-ggsave(file.path(outdir, "FigureS2.pdf"), width=6, height=4.5)
-```
-
-### ChIP-seq
-
-```{r, fig.width = 5, fig.height = 5}
-wcs <- which(grepl("ChIPseq", casestudy))
-objs <- objects[wcs]
-labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
-                      function(x) x[1]))
-plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", 
-                Nlabel = FALSE, annotate = "propPossible") + 
-  ggtitle("ChIP-Seq") + 
-  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
-ggsave(file.path(outdir, "FigureS3.pdf"), width=5, height=5)
-```
-
-### RNA-seq
-```{r, fig.width = 6, fig.height = 5}
-# by dataset
-wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("polyester|yeast"), objects) &
-            !grepl("scRNAseq", casestudy))
-objs <- objects[wcs]
-labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
-                      function(x) x[1]))
-plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", 
-                excludeMethods = NULL, Nlabel = FALSE, annotate = "propPossible") + 
-  ggtitle("RNA-seq") + 
-  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
-ggsave(file.path(outdir, "FigureS4.pdf"), width=6, height=5)
-```
-
-### scRNA-seq
-
-```{r, fig.width = 15, fig.height = 6}
-wcs <- which(grepl("scRNAseq", casestudy))
-objs <- objects[wcs]
-ds <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark-"), 
-                      function(x) x[1]))
-m0 <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark-"), 
-                      function(x) x[2]))
-meth <- unlist(lapply(str_split(m0, pattern = "-"), 
-                      function(x) x[1]))
-cv <- gsub(".rds", "", unlist(lapply(str_split(m0, pattern = "-"), 
-                      function(x) x[2])))
-
-uninf <- which(grepl("uninf", objs))
-
-# by dataset
-pa <- plotMethodRanks(objs[-uninf], colLabels = ds[-uninf], alpha = alpha, 
-                      xlab = "Dataset", annotate = "propPossible") + 
-  ggtitle("scRNA-seq")+ 
-  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
-
-# by method
-pb <- plotMethodRanks(objs[-uninf], colLabels = meth[-uninf], alpha = alpha,
-                      xlab = "Method", annotate = "propPossible") + 
-  ggtitle("scRNA-seq")+ 
-  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
-
-# by covariate
-pc <- plotMethodRanks(objs, colLabels = cv, alpha = alpha,
-                      xlab = "Covariate", annotate = "propPossible") + 
-  ggtitle("scRNA-seq")+ 
-  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
-
-plot_grid(pa, pb, pc, ncol=3)
-ggsave(file.path(outdir, "FigureS5.pdf"), width=15, height=6)
-```
-
-### Microbiome
-
-```{r, fig.width = 14, fig.height = 6.5}
-# by dataset, excluding otu when genus is available
-# excluding log-ubiquity
-# excluding enigma-so4 (no discoveries)
-wcs <- which(grepl("microbiome", casestudy) & 
-              !(grepl(c("goodrich|papa"), dataset) & 
-              grepl(c("otu"), dataset)) &
-              !grepl("log-ubiquity", dataset) &
-              !grepl("so4", dataset) &
-              !grepl("baxter", dataset) )
-objs <- objects[wcs]
-labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
-                      function(x) x[1]))
-plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", 
-                Nlabel = FALSE, annotate = "propPossible") + 
-  ggtitle("Microbiome")+ 
-  theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1))
-ggsave(file.path(outdir, "FigureS6.pdf"), width=14, height=6.5)
-```
-
-### Yeast Simulation
-
-```{r}
-# non-null
-wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("brain|mir|polyester|null"), objects) &
-            !grepl("scRNAseq", casestudy))
-objs <- objects[wcs]
-labs <- gsub(".rds", "", 
-             unlist(lapply(str_split(dataset[wcs], pattern = "-results-"), 
-                      function(x) x[2])))
-
-plotMethodRanks(objs, 
-                colLabels = labs, 
-                alpha = alpha, xlab = "Comparison",
-                excludeMethods = NULL,
-                annotate = "propPossible")
-
-plotMethodRanks(objs, 
-                colLabels = labs, 
-                alpha = alpha, xlab = "Comparison",
-                excludeMethods = NULL, 
-                fill = "FDR",
-                annotate = "propPossible")
-
-plotMethodRanks(objs, 
-                colLabels = labs, 
-                alpha = alpha, xlab = "Comparison",
-                excludeMethods = NULL, 
-                fill = "TPR",
-                annotate = "propPossible")
-
-# null
-wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("brain|mir|polyester|de"), objects) &
-            !grepl("scRNAseq", casestudy))
-objs <- objects[wcs]
-labs <- gsub(".rds", "", 
-             unlist(lapply(str_split(dataset[wcs], pattern = "-results-"), 
-                      function(x) x[2])))
-
-plotMethodRanks(objs, 
-                colLabels = labs, 
-                alpha = alpha, xlab = "Comparison",
-                excludeMethods = NULL,
-                annotate = "propPossible")
-
-```
-
-### Polyester Simulation
-
-```{r}
-wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("brain|mir|yeast|null"), objects) &
-            !grepl("scRNAseq", casestudy))
-objs <- objects[wcs]
-labs <- gsub(".rds", "", 
-             unlist(lapply(str_split(dataset[wcs], pattern = "-results-"), 
-                      function(x) x[2])))
-
-plotMethodRanks(objs, 
-                colLabels = labs, 
-                alpha = alpha, xlab = "Comparison",
-                excludeMethods = NULL)
-
-plotMethodRanks(objs, 
-                colLabels = labs, 
-                alpha = alpha, xlab = "Comparison",
-                excludeMethods = NULL, 
-                fill = "FDR")
-
-plotMethodRanks(objs, 
-                colLabels = labs, 
-                alpha = alpha, xlab = "Comparison",
-                excludeMethods = NULL, 
-                fill = "TPR")
-
-# null
-wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("brain|mir|yeast|de"), objects) &
-            !grepl("scRNAseq", casestudy))
-objs <- objects[wcs]
-labs <- gsub(".rds", "", 
-             unlist(lapply(str_split(dataset[wcs], pattern = "-results-"), 
-                      function(x) x[2])))
-
-plotMethodRanks(objs, 
-                colLabels = labs, 
-                alpha = alpha, xlab = "Comparison",
-                excludeMethods = NULL) 
-```
-
-# Alternative Supplementary Figures
-
-## All Case Studies
-
-This plot is an alternative to Supplementary Figures S1-S6.
-We include all case studies in a single plot.
+          plot.title = element_text(face="bold", size = 11))
+
+  gp <- plot_grid(lineplot_fdr + guides(color = FALSE, linetype = FALSE),
+                  lineplot_tpr + guides(color = FALSE, linetype = FALSE),
+                  labels = LETTERS[1:2], nrow = 1)
+  gp <- plot_grid(gp, get_legend(lineplot_fdr), rel_widths = c(1, 1/8))
+  gp
+  ggsave(file.path(outdir, paste0("Figure_infCov_", org, ".pdf")), height = 3.5, width = 8)
+}
+```
+
+## Case study rankings
+
+This is a figure similar to Figure 4, except that we 
+include all case studies in a single plot.
 
@@
 
-# Informativeness of covariates in case studies
+## Informativeness of covariates in case studies
 
@@
 
-# Informativeness of covariates in Yeast & Polyester simulation
-
-As an alternative to the line plot above (Figure S7), we make a heatmap plot that
-displays the mean difference within simulation replicate.
-
-```{r}
-
-for (org in c("yeast", "polyester")){
-
-objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
-resfile <- objects[grepl(org, objects) ]
-
-# methods that use a covariate
-cov.methods <- c("ihw", "bl", "lfdr", "fdrreg-e", "fdrreg-t")
-
-# regexp to construct master data table separating case study datasets, methods
-# and covariates
-obj.names <- str_split(resfile, pattern = "/")
-casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
-filename <- gsub(".rds", "", unlist(lapply(obj.names, function(x) x[4] )))
-dataset <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[3]]))
-inform <-  ifelse(grepl("uninf", resfile), "uninf", "inf")
-type <-  ifelse(grepl("null", resfile), "null", "de")
-size <-  ifelse(grepl("10", resfile), "10", "5")
-
-tab <- data.frame()
-for (l in seq_along(resfile)){
-  x <- readRDS(resfile[l])
-
-  tmp2 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
-                 fill =  "FDR", annotate = NULL, alpha = 0.10) %>%
-    mutate(fdr = nrejects) %>%
-    select(method, casestudy, replicate, fdr)
-  tmp3 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
-                  fill =  "TNR", annotate = NULL, alpha = 0.10) %>%
-    mutate(tnr = nrejects) %>%
-  select(method, casestudy, replicate, tnr)
-
-  if (type[l] != "null"){
-    tmp1 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
-                    fill =  "TPR", annotate = NULL, alpha = 0.10) %>%
-      mutate(tpr = nrejects) %>%
-    select(method, casestudy, replicate, tpr)
-
-    tmp <- left_join(tmp1, tmp2, by = c("method", "casestudy", "replicate")) %>%
-      left_join(tmp3, by = c("method", "casestudy", "replicate")) %>%
-      mutate(inform = inform[l],
-             type = type[l],
-             size = size[l])
-  }else{
-    tmp <- left_join(tmp2, tmp3, by = c("method", "casestudy", "replicate")) %>%
-      mutate(inform = inform[l],
-             type = type[l],
-             size = size[l])
-  }
-
-  tab <- bind_rows(tab, tmp)
-}
-
-# take difference within replicate, then average differences across
-# type and size and method
-
-dftab <- tab %>%
-  filter(method %in% cov.methods) %>%
-  dplyr::mutate(id = paste(type, size, method, replicate, sep = "-"))
-
-uninftab <- dftab %>% 
-  filter(inform == "uninf") %>%
-  select(id, fdr, tpr, tnr) %>%
-  rename(fdr = "fdr_uninf",
-         tnr = "tnr_uninf",
-         tpr = "tpr_uninf")
-
-inftab <- dftab %>%
-  filter(inform == "inf") %>%
-  select(-inform)
-
-rowOrder <- c("ihw", "bl", "lfdr", "fdrreg-e", "fdrreg-t")
-
-dftab <- left_join(inftab, uninftab, by = "id") %>%
-  mutate(tpr_diff = tpr-tpr_uninf,
-         fdr_diff = fdr-fdr_uninf,
-         tnr_diff = tnr-tnr_uninf) %>%
-  group_by(type, size, method) %>%
-  summarize(tpr_se = sqrt(sd(tpr_diff)/n()),
-            fdr_se = sqrt(sd(fdr_diff)/n()),
-            tnr_se = sqrt(sd(tnr_diff)/n()),
-            tpr_diff = mean(tpr_diff),
-            fdr_diff = mean(fdr_diff),
-            tnr_diff = mean(tnr_diff)) %>%
-  mutate(method = factor(method, levels = rowOrder)) %>%
-  mutate(fdr_diff = ifelse(type == "null", NA, fdr_diff),
-         tpr_diff = ifelse(type == "null", NA, tpr_diff),
-         fdr_se = ifelse(type == "null", NA, fdr_se),
-         tpr_se = ifelse(type == "null", NA, tpr_se))
-
-plot_fun <- function(var){
-
-  plot  <- ggplot(dftab %>% filter(!is.na(get(var))), 
-                  aes(x = interaction(type, size), 
-                                        y = method)) + 
-    geom_tile(aes(fill=get(var))) + 
-    scale_fill_gradient2(paste0("Change\nin ", toupper(gsub("_diff", "", var))),
-                         low = "red", midpoint = 0, mid = "white", high = "blue",
-                         labels = scales::percent) +
-    theme_bw() +
-    xlab("Experiment") +
-    ylab("Method") +
-    facet_grid( ~ type + size, scales = "free", space = "free") +
-    ggtitle(paste0("Impact of informative covariate on ", toupper(gsub("_diff", "", var)))) +
-    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
-    geom_text(aes(label = scales::percent(get(var))),
-            color = "black", size = 3)
-
-  return(plot)
-}
-
-# arrange in multipanel figure?
-plot_fun("tpr_diff")
-ggsave(file.path(outdir, paste0("Figure_infCov_tpr_", org, ".pdf")), height = 5, width = 5)
-
-plot_fun("fdr_diff")
-ggsave(file.path(outdir, paste0("Figure_infCov_fpr_", org, ".pdf")), height = 5, width = 5)
-
-plot_fun("tnr_diff")
-ggsave(file.path(outdir, paste0("Figure_infCov_tnr_", org, ".pdf")), height = 5, width = 7)
-
-}
-
-```
-
 # Session information

### 2018-07-25 datasets/RNA-Seq/polyester-simulation.Rmd
Normalize counts before estimating polyester params; adjust beta coefficients

@@
 
+# normalize
+dds <- DESeqDataSetFromMatrix(countData = counts, 
+                              colData = tibble(sample=colnames(counts)),
+                              design = ~1)
+dds <- estimateSizeFactors(dds)
+counts <- counts(dds, normalize=TRUE)
+
 ## Estimate the zero inflated negative binomial parameters
@@
   if(nDE > 0){
-    wts <- log10(rowMeans(dat_alt)+1)
+    wts <- rowMeans(dat_alt)
     DE <- sample(1:nrow(dat_alt), nDE, prob = wts)
@@
     dat_alt[DE,] = create_read_numbers(mu[DE], paramPolyester$fit, p0[DE], 
-                                  n=sampleSize*2, beta=cbind(log2FC[DE]), 
+                                  n=sampleSize*2, beta=cbind(log2FC[DE]/2), 
                                   mod=modPolyester)

### 2018-07-25 datasets/RNA-Seq/yeast-simulation.Rmd
Use raw scale informative weights

@@
     # pick random set of nDE genes to add signal to
-    wts <- log10(rowMeans(counts(dds_test, normalize = TRUE))+1)
+    wts <- rowMeans(counts(dds_test, normalize = TRUE))
     DE <- sample(1:nrow(dds_test), nDE, prob = wts)

### 2018-07-26 datasets/R/plotsim_average.R
Add option to change plotsim style for diff plots

@@ -26,6 +26,8 @@
 #' @param facetMethodType logical indicating whether to facet the plot into 
 #'        two panels - methods that use only one piece of information, and 
 #'        those that use more than just the p-value (default = FALSE)
+#' @param diffplot logical indicating whether 'value' in tsb is a difference
+#'        between informative and uninformative covariates. (default = FALSE)
 #'
 #' @return
 #' a ggplot object.
@@ -33,7 +35,8 @@
 #' @author Patrick Kimes & Keegan Korthauer
 plotsim_average <- function(tsb, met, filter_set = NULL, merge_ihw = TRUE,
                             clean_names = FALSE, errorBars=FALSE,
-                            palette = candycols, facetMethodType = FALSE) {
+                            palette = candycols, facetMethodType = FALSE,
+                            diffplot = FALSE) {
 
     ## cacluate mean per replication
     tsba <- tsb %>%
@@ -103,7 +106,9 @@ plotsim_average <- function(tsb, met, filter_set = NULL, merge_ihw = TRUE,
         expand_limits(x = 0) +
         scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.01)) +
         ylab(met) +
-        ggtitle(paste0("Mean ", met, " Over ", max(tsba$n), " Replications")) +
+        ggtitle(ifelse(diffplot,
+                       paste0("Mean Difference ", met, " Over ", max(tsba$n), " Replications"),
+                       paste0("Mean ", met, " Over ", max(tsba$n), " Replications"))) +
         scale_color_manual(values = col) +
         scale_linetype_manual(values = lty) 
     
@@ -120,22 +125,31 @@ plotsim_average <- function(tsb, met, filter_set = NULL, merge_ihw = TRUE,
     ## use percentage on y-axis labels when appropriate
     if (met %in% c("FDR", "FNR", "TPR", "TNR", "FWER", "rejectprop")) {
         gp <- gp +
-            scale_y_continuous(met, labels=scales::percent)
+            scale_y_continuous(ifelse(diffplot, paste(met, "(informative - uninformative)"), met),
+                               labels=scales::percent)
+    } else if (diffplot) {
+        gp <- gp + ylab(paste(met, "(informative - uninformative)")) 
     }
 
-    ## include 0% or 100% in plotting range depending on metric
[…]

### 2018-07-26 datasets/simulations/simulations-informative-cosine.Rmd
Add difference plots to cosine simulations

@@
 
+We also compare the simulation results with and without an informative covariate.
+
+```{r gauss-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key, performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key, performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key", "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+```
+
 # Student's t Setting (df = 5)
@@
 
+We also compare the simulation results with and without an informative covariate.
+
+```{r t5-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key, performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key, performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key", "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+```
+
 # Student's t Setting (df = 11)
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r t11-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key, performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key, performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key", "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 
+We also compare the simulation results with and without an informative covariate.
+
+```{r chisq4-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key, performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key, performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key", "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+```
+
 # Session Info

### 2018-07-26 datasets/simulations/simulations-informative-sine.Rmd
Add info vs uninfo comparisons to all sims

@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r gauss-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r t5-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r t11-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 
+We also compare the simulation results with and without an informative covariate.
+
+```{r chisq4-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+```
+
 # Session Info

### 2018-07-26 datasets/simulations/simulations-null.Rmd
Add info vs uninfo comparisons to all sims

@@
 ```{r gauss-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = 0.05, covname = "effect_size", trans = "log1p")
+covariateLinePlot(res_i, alpha = 0.05, covname = "effect_size", trans = "log1p")
 
-covariateLinePlot(res, alpha = 0.05, covname = "ind_covariate", trans = "log1p")
+covariateLinePlot(res_i, alpha = 0.05, covname = "ind_covariate", trans = "log1p")
 ```
@@
 ```{r t5-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = 0.05, covname = "effect_size", trans = "log1p")
+covariateLinePlot(res_i, alpha = 0.05, covname = "effect_size", trans = "log1p")
 
-covariateLinePlot(res, alpha = 0.05, covname = "ind_covariate", trans = "log1p")
+covariateLinePlot(res_i, alpha = 0.05, covname = "ind_covariate", trans = "log1p")
 ```
@@
 if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
-    aggupset(res, alpha = ualpha, supplementary = FALSE, return_list = FALSE) 
+    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE) 
 } else {
@@
 ```{r t11-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = 0.05, covname = "effect_size", trans = "log1p")
+covariateLinePlot(res_i, alpha = 0.05, covname = "effect_size", trans = "log1p")
 
-covariateLinePlot(res, alpha = 0.05, covname = "ind_covariate", trans = "log1p")
+covariateLinePlot(res_i, alpha = 0.05, covname = "ind_covariate", trans = "log1p")
 ```

### 2018-07-26 datasets/simulations/simulations-varyingpi0.Rmd
Add info vs uninfo comparisons to all sims

--- 2018-06-07 datasets/simulations/simulations-varyingpi0.Rmd
@@
 ## intermediary files we create below
-null05_file <- file.path(resdir, "varyingpi0-benchmark-nullprop05.rds")
-null10_file <- file.path(resdir, "varyingpi0-benchmark-nullprop10.rds")
-null20_file <- file.path(resdir, "varyingpi0-benchmark-nullprop20.rds")
-null50_file <- file.path(resdir, "varyingpi0-benchmark-nullprop50.rds")
-null80_file <- file.path(resdir, "varyingpi0-benchmark-nullprop80.rds")
-null90_file <- file.path(resdir, "varyingpi0-benchmark-nullprop90.rds")
-null95_file <- file.path(resdir, "varyingpi0-benchmark-nullprop95.rds")
-null99_file <- file.path(resdir, "varyingpi0-benchmark-nullprop99.rds")
+null05_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop05.rds")
+null10_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop10.rds")
+null20_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop20.rds")
+null30_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop30.rds")
+null40_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop40.rds")
+null50_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop50.rds")
+null60_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop60.rds")
+null70_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop70.rds")
+null80_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop80.rds")
+null90_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop90.rds")
+null95_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop95.rds")
+null99_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop99.rds")
 
@@
 ```{r null05-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r null05-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
-
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r null05-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 ```{r null10-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r null10-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
-
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r null10-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 ```{r null20-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r null20-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
-
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r null20-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+```
+
+# 30% Null Setting
+
+Next, we consider the setting where 30% of tests are null.
+
+## Data Simulation
+
+```{r null30-parameters}
+pi0 <- pi0_sine(0.30)             # numeric: proportion of null hypotheses
+seed <- 511
+```
+
+We next run the simulations.
+
+```{r null30-run-simulation}
+if (file.exists(null30_file)) {
+    res <- readRDS(null30_file)
+} else {
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
+                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
+                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    seed = seed, mc.cores = cores)
+    saveRDS(res, file = null30_file)
+}
+res_i <- lapply(res, `[[`, "informative")
+res_u <- lapply(res, `[[`, "uninformative")
+```
+
+## Covariate Diagnostics
+
+Here, we show the relationship between the independent covariate and p-values for a
+single replication of the experiment.
+
+```{r null30-one-simulation}
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
+```
+
+```{r, null30-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
+rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
+```
+
+```{r, null30-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+```
+
+## Benchmark Metrics
+
+We plot the averaged results across `r B` replications.
+
+```{r null30-metrics-averages, results = "hide"}
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(resdf, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE) 
+```
+
+We also take a look at the distribution of rejects for each method as a function of
+the effect size and independent covariate.
+
+```{r null30-metrics-covlineplot, results = "hide"}
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
+```
+
+Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
+the overlap of rejections between methods.
+
+```{r null30-metrics-upset, results = "hide"}
+if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
+    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
+} else {
+    message("Not enough methods found rejections at alpha ", ualpha, 
+            "; skipping upset plot")
+}
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r null30-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+```
+
+# 40% Null Setting
+
+Next, we consider the setting where 40% of tests are null.
+
+## Data Simulation
+
+```{r null40-parameters}
+pi0 <- pi0_sine(0.40)             # numeric: proportion of null hypotheses
+seed <- 515
+```
+
+We next run the simulations.
+
+```{r null40-run-simulation}
+if (file.exists(null40_file)) {
+    res <- readRDS(null40_file)
+} else {
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
+                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
+                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    seed = seed, mc.cores = cores)
+    saveRDS(res, file = null40_file)
+}
+res_i <- lapply(res, `[[`, "informative")
+res_u <- lapply(res, `[[`, "uninformative")
+```
+
+## Covariate Diagnostics
+
+Here, we show the relationship between the independent covariate and p-values for a
+single replication of the experiment.
+
+```{r null40-one-simulation}
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
+```
+
+```{r, null40-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
+rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
+```
+
+```{r, null40-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+```
+
+## Benchmark Metrics
+
+We plot the averaged results across `r B` replications.
+
+```{r null40-metrics-averages, results = "hide"}
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(resdf, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE) 
+```
+
+We also take a look at the distribution of rejects for each method as a function of
+the effect size and independent covariate.
+
+```{r null40-metrics-covlineplot, results = "hide"}
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
+```
+
+Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
+the overlap of rejections between methods.
+
+```{r null40-metrics-upset, results = "hide"}
+if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
+    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
+} else {
+    message("Not enough methods found rejections at alpha ", ualpha, 
+            "; skipping upset plot")
+}
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r null40-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 ```{r null50-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r null50-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
-
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r null50-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+```
+
+# 60% Null Setting
+
+Next, we consider the setting where 60% of tests are null.
+
+## Data Simulation
+
+```{r null60-parameters}
+pi0 <- pi0_sine(0.60)             # numeric: proportion of null hypotheses
+seed <- 1502
+```
+
+We next run the simulations.
+
+```{r null60-run-simulation}
+if (file.exists(null60_file)) {
+    res <- readRDS(null60_file)
+} else {
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
+                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
+                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    seed = seed, mc.cores = cores)
+    saveRDS(res, file = null60_file)
+}
+res_i <- lapply(res, `[[`, "informative")
+res_u <- lapply(res, `[[`, "uninformative")
+```
+
+## Covariate Diagnostics
+
+Here, we show the relationship between the independent covariate and p-values for a
+single replication of the experiment.
+
+```{r null60-one-simulation}
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
+```
+
+```{r, null60-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
+rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
+```
+
+```{r, null60-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+```
+
+## Benchmark Metrics
+
+We plot the averaged results across `r B` replications.
+
+```{r null60-metrics-averages, results = "hide"}
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(resdf, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE) 
+```
+
+We also take a look at the distribution of rejects for each method as a function of
+the effect size and independent covariate.
+
+```{r null60-metrics-covlineplot, results = "hide"}
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
+```
+
+Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
+the overlap of rejections between methods.
+
+```{r null60-metrics-upset, results = "hide"}
+if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
+    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
+} else {
+    message("Not enough methods found rejections at alpha ", ualpha, 
+            "; skipping upset plot")
+}
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r null60-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+```
+
+# 70% Null Setting
+
+Next, we consider the setting where 70% of tests are null.
+
+## Data Simulation
+
+```{r null70-parameters}
+pi0 <- pi0_sine(0.70)             # numeric: proportion of null hypotheses
+seed <- 1722
+```
+
+We next run the simulations.
+
+```{r null70-run-simulation}
+if (file.exists(null70_file)) {
+    res <- readRDS(null70_file)
+} else {
+    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
+                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
+                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    seed = seed, mc.cores = cores)
+    saveRDS(res, file = null70_file)
+}
+res_i <- lapply(res, `[[`, "informative")
+res_u <- lapply(res, `[[`, "uninformative")
+```
+
+## Covariate Diagnostics
+
+Here, we show the relationship between the independent covariate and p-values for a
+single replication of the experiment.
+
+```{r null70-one-simulation}
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
+```
+
+```{r, null70-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
+rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
+```
+
+```{r, null70-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+```
+
+## Benchmark Metrics
+
+We plot the averaged results across `r B` replications.
+
+```{r null70-metrics-averages, results = "hide"}
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(resdf, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE)
+
+plotsim_average(resdf, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE) 
+```
+
+We also take a look at the distribution of rejects for each method as a function of
+the effect size and independent covariate.
+
+```{r null70-metrics-covlineplot, results = "hide"}
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
+```
+
+Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
+the overlap of rejections between methods.
+
+```{r null70-metrics-upset, results = "hide"}
+if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
+    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
+} else {
+    message("Not enough methods found rejections at alpha ", ualpha, 
+            "; skipping upset plot")
+}
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r null70-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 ```{r null80-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r null80-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
-
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r null80-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 ```{r null90-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r null90-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
-
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r null90-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 ```{r null95-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r null95-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
-
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r null95-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 ```{r null99-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r null99-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
-
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
+
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 ```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r null99-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+```
+
+# Session Info
+
+```{r}
+sessionInfo()
+```

### 2018-07-26 manuscript_figures/manuscript_figures.Rmd
Fix syntax in dplyr::rename

@@
               fill="meanRank", annotate=NULL, alpha =0.10) %>%
-  rename(filename = "casestudy") %>%
-  #rename(casestudy = "filename") %>%
+  rename(casestudy = "filename") %>%
   mutate(filename = gsub(".rds", "", filename))

### 2018-07-26 datasets/simulations/simulations-varyingeffectsize-chisq.Rmd

@@
 ```{r spiky-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r spiky-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
 
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r spiky-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 ```{r flattop-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r flattop-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
 
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r flattop-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 ```{r skew-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r skew-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
 
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r skew-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 ```{r bimodal-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r bimodal-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
 
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r bimodal-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```

### 2018-07-26 datasets/simulations/simulations-varyinginformativeness.Rmd

@@
 ```{r info00-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r info00-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
 
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r info00-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 ```{r info20-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r info20-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
 
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r info20-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 ```{r info40-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r info40-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
 
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r info40-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 ```{r info60-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r info60-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
 
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r info60-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 ```{r info80-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r info80-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
 
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r info80-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 ```{r info100-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res, alpha = seq(0.01, 0.10, 0.01))
+resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
 
@@
 ```{r info100-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res, alpha = ualpha, covname = "effect_size")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
 
-covariateLinePlot(res, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
 ```
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r info100-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```

### 2018-07-26 datasets/simulations/simulations-varyingntests.Rmd

@@
 ```{r parameters-shared}
-tstat <- rnorm_generator(2)       # functional: dist of alternative test stats
+tstat <- rnorm_generator(3)       # functional: dist of alternative test stats
 tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
@@
 
+We also compare the simulation results with and without an informative covariate.
+
+```{r n100-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+```
+
 # 500 Tests
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r n500-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 
+We also compare the simulation results with and without an informative covariate.
+
+```{r n1000-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+```
+
 # 5000 Tests
@@
 }
+```
+
+We also compare the simulation results with and without an informative covariate.
+
+```{r n5000-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
 ```
@@
 
+We also compare the simulation results with and without an informative covariate.
+
+```{r n10000-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+```
+
 # 50000 Tests
@@
 
+We also compare the simulation results with and without an informative covariate.
+
+```{r n50000-metrics-differences, results = "hide"}
+resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
+
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
+                            suffix = c(".info", ".uninfo"))
+resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
+
+plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+
+plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+```
+
 # Session Info

### 2018-07-27 manuscript_figures/manuscript_figures_sims.Rmd
Update sim figures with difference plots

--- 2018-07-19 manuscript_figures/manuscript_figures_sims.Rmd
@@
 methodset <- c("bonf", "bh", "ihw", "qvalue", "bl-df03", "lfdr",
-               "scott-empirical", "scott-theoretical", "ashq")
+               "fdrreg-e", "fdrreg-t", "ashq", "adapt-glm")
 ```
@@
             zz <- readRDS(x)
-            zz <- plotsim_standardize(zz, alpha = alpha)
-            zz <- dplyr::select(zz, rep, blabel, key, alpha, value)
-            zz <- dplyr::group_by(zz, blabel, key, alpha)
-            zz <- dplyr::summarize(zz,
-                                   se = sd(value, na.rm = TRUE) / sqrt(sum(!is.na(value))),
-                                   q10 = quantile(value, .1, na.rm = TRUE),
-                                   q90 = quantile(value, .9, na.rm = TRUE),
-                                   med_value = median(value, na.rm = TRUE),
-                                   mean_value = mean(value, na.rm = TRUE),
-                                   nNA = sum(is.na(value)))
-            dplyr::ungroup(zz)
+            zz_i <- lapply(zz, `[[`, "informative")
+            zz_u <- lapply(zz, `[[`, "uninformative")
+            zz_i <- plotsim_standardize(zz_i, alpha = ia / 100L)
+            zz_i <- dplyr::select(zz_i, rep, blabel, param.alpha, key, performanceMetric, alpha, value)
+            zz_u <- plotsim_standardize(zz_u, alpha = ia / 100L)
+            zz_u <- dplyr::select(zz_u, rep, blabel, param.alpha, key, performanceMetric, alpha, value)
+            dplyr::full_join(zz_i, zz_u,
+                             by = c("rep", "blabel", "param.alpha", "key", "performanceMetric", "alpha"),
+                             suffix = c(".info", ".uninfo"))
         })
@@
 res <- bind_rows(res)
+res <- dplyr::select(res, -param.alpha)
 res <- dplyr::rename(res, Method = blabel)
@@
 
+We compute summary metrics for each method, setting, and cutoff.
+
+```{r}
+resmet <- dplyr::group_by(res, setting, dist, Method, key, performanceMetric, alpha)
+resmet <- dplyr::summarize(resmet,
+                           mean.info = mean(value.info, na.rm = TRUE),
+                           se.info = sd(value.info, na.rm = TRUE) / sqrt(sum(!is.na(value.info))),
+                           nNA.info = sum(is.na(value.info)),
+                           mean.diff = mean(value.info - value.uninfo, na.rm = TRUE),
+                           se.diff = sd(value.info - value.uninfo, na.rm = TRUE) /
+                               sqrt(sum(!is.na(value.info - value.uninfo))))
+resmet <- dplyr::ungroup(resmet)
+```
+
 We clean up the annotations in the table for some of the more complex simulation settings.
@@
 ```{r}
-res <- dplyr::mutate(res, inform = ifelse(grepl("varyinginfo", setting), dist, NA))
-res <- dplyr::mutate(res, inform = as.numeric(gsub("level", "", inform)))
-res <- dplyr::mutate(res, dist = ifelse(grepl("varyinginfo", setting), "gaussian", dist))
+resmet <- dplyr::mutate(resmet, inform = ifelse(grepl("varyinginfo", setting), dist, NA))
+resmet <- dplyr::mutate(resmet, inform = as.numeric(gsub("level", "", inform)))
+resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("varyinginfo", setting), "gaussian", dist))
 
-res <- dplyr::mutate(res, pi0 = ifelse(grepl("varyingpi0", setting), dist, 90))
-res <- dplyr::mutate(res, pi0 = as.numeric(gsub("nullprop", "", pi0)))
-res <- dplyr::mutate(res, dist = ifelse(grepl("varyingpi0", setting), "gaussian", dist))
+resmet <- dplyr::mutate(resmet, pi0 = ifelse(grepl("varyingpi0", setting), dist, 90))
+resmet <- dplyr::mutate(resmet, pi0 = as.numeric(gsub("nullprop", "", pi0)))
+resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("varyingpi0", setting), "gaussian", dist))
 
-res <- dplyr::mutate(res, signal = ifelse(grepl("varyinges", setting), dist, NA))
-res <- dplyr::mutate(res, dist = ifelse(grepl("^varyinges$", setting), "gaussian", dist))
-res <- dplyr::mutate(res, dist = ifelse(grepl("^varyinges-t$", setting), "t11", dist))
-res <- dplyr::mutate(res, dist = ifelse(grepl("^varyinges-chisq$", setting), "chisq4", dist))
-res <- dplyr::mutate(res, setting = ifelse(grepl("varyinges", setting), "varyinges", setting))
+resmet <- dplyr::mutate(resmet, signal = ifelse(grepl("varyinges", setting), dist, NA))
+resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyinges$", setting), "gaussian", dist))
+resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyinges-t$", setting), "t11", dist))
+resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyinges-chisq$", setting), "chisq4", dist))
+resmet <- dplyr::mutate(resmet, setting = ifelse(grepl("varyinges", setting), "varyinges", setting))
 ```
@@
 ```{r}
-res <- dplyr::mutate(res, dist = factor(dist, levels = c("gaussian", "t11", "t5", "chisq4"),
+resmet <- dplyr::mutate(resmet, dist = factor(dist, levels = c("gaussian", "t11", "t5", "chisq4"),
                                               labels = c(expression("Noise: " * N(0, 1)),
@@
 ```{r}
-dplyr::distinct(res, setting)
+dplyr::distinct(resmet, setting)
 ```
@@
 ```{r, fig.width = 18, fig.height = 8}
-res_nullset <- dplyr::filter(res, setting == "null",
+res_nullset <- dplyr::filter(resmet, setting == "null",
                              !(Method == "ashq" & grepl("chi\\^2", dist)),
-                             !(grepl("scott", Method) & grepl("chi\\^2", dist)))
+                             !(grepl("fdrreg", Method) & grepl("chi\\^2", dist)))
 res_nullset <- tidyr::complete(res_nullset, setting, Method, dist, key)
@@
 null_tnr <- dplyr::filter(res_nullset, key == "TNR") %>%
-    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
     geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
     facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
@@
 null_fwer <- dplyr::filter(res_nullset, key == "FWER") %>%
-    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
     geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
     facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
@@
 
+We also look at the difference between informative and uninformative covariates.
+
+```{r, fig.width = 18, fig.height = 8}
+null_tnr_diff <- dplyr::filter(res_nullset, key == "TNR") %>%
+    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
+    scale_y_continuous("TNR\n(informative - uninformative)", labels = scales::percent) +
+    ggtitle("Mean TNR Difference Over 100 Replications") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+null_fwer_diff <- dplyr::filter(res_nullset, key == "FWER") %>%
+    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
+    scale_y_continuous("FWER\n(informative - uninformative)", labels = scales::percent) +
+    ggtitle("FWER Difference Over 100 Replications") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+gp <- plot_grid(null_tnr_diff + guides(color = FALSE, linetype = FALSE),
+                null_fwer_diff + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:2], ncol = 1)
+gp <- plot_grid(gp, get_legend(null_tnr_diff), rel_widths = c(1, 1/16))
+gp
+```
+
 # Sampling Distributions
@@
 ```{r, fig.width = 18, fig.height = 8}
-res_noise <- dplyr::filter(res, setting == "informative-cubic")
+res_noise <- dplyr::filter(resmet, setting == "informative-cubic")
 res_noise <- tidyr::complete(res_noise, setting, Method, dist, key)
@@
 noise_fdr <- dplyr::filter(res_noise, key == "FDR") %>%
-    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
     geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
     geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
@@
 noise_tpr <- dplyr::filter(res_noise, key == "TPR") %>%
-    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
     geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
     facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
@@
 gp <- plot_grid(gp, get_legend(noise_tpr), rel_widths = c(1, 1/16))
+gp
+```
+
+We also look at the difference between informative and uninformative covariates.
+
+```{r, fig.width = 18, fig.height = 8}
+noise_fdr_diff <- dplyr::filter(res_noise, key == "FDR") %>%
+    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
+    scale_y_continuous("FDR\n(informative - uninformative)", labels = scales::percent) +
+    ggtitle("Mean FDR Difference Over 100 Replications") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+noise_tpr_diff <- dplyr::filter(res_noise, key == "TPR") %>%
+    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
+    scale_y_continuous("TPR\n(informative - uninformative)", labels = scales::percent) +
+    ggtitle("Mean TPR Difference Over 100 Replications") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+gp <- plot_grid(noise_fdr_diff + guides(color = FALSE, linetype = FALSE),
+                noise_tpr_diff + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:2], ncol = 1)
+gp <- plot_grid(gp, get_legend(noise_tpr_diff), rel_widths = c(1, 1/16))
 gp
@@
 ```{r, fig.width = 18, fig.height = 8}
-res_icov <- dplyr::filter(res, grepl("^informative", setting), grepl("N\\(0, 1\\)", dist))
+res_icov <- dplyr::filter(resmet, grepl("^informative", setting), grepl("N\\(0, 1\\)", dist))
 res_icov <- tidyr::complete(res_icov, setting, Method, dist, key)
@@
 icov_fdr <- dplyr::filter(res_icov, key == "FDR") %>%
-    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
     geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
     geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
@@
 icov_tpr <- dplyr::filter(res_icov, key == "TPR") %>%
-    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
     geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
     facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
@@
 
-The general trends are identical across informative covariates considered (cubic: a monotonic relationship,
-sine: a non-monotone relationship, and step: a discrete relationship).
+We also look at the difference between informative and uninformative covariates.
+
+```{r, fig.width = 18, fig.height = 8}
+icov_fdr_diff <- dplyr::filter(res_icov, key == "FDR") %>%
+    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
+    scale_y_continuous("FDR\n(informative - uninformative)", labels = scales::percent) +
+    ggtitle("Mean FDR Difference Over 100 Replications") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+icov_tpr_diff <- dplyr::filter(res_icov, key == "TPR") %>%
+    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
+    scale_y_continuous("TPR\n(informative - uninformative)", labels = scales::percent) +
+    ggtitle("Mean TPR Difference Over 100 Replications") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 14))
+
+gp <- plot_grid(icov_fdr_diff + guides(color = FALSE, linetype = FALSE),
+                icov_tpr_diff + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:2], ncol = 1)
+gp <- plot_grid(gp, get_legend(icov_tpr_diff), rel_widths = c(1, 1/12))
+gp
+```
 
@@
 ```{r, fig.width = 18, fig.height = 8}
-res_noni <- dplyr::filter(res, grepl("noninformative", setting))
-res_noni <- tidyr::complete(res_noni, setting, Method, dist, key)
+res_noni <- dplyr::filter(res, grepl("informative-cubic", setting))
+res_noni <- dplyr::group_by(res_noni, setting, dist, Method, key, performanceMetric, alpha)
+res_noni <- dplyr::summarize(res_noni,
+                             mean.uninfo = mean(value.uninfo, na.rm = TRUE),
+                             se.uninfo = sd(value.uninfo, na.rm = TRUE) /
+                                 sqrt(sum(!is.na(value.uninfo))))
+res_noni <- dplyr::ungroup(res_noni)
 
 noni_fdr <- dplyr::filter(res_noni, key == "FDR") %>%
-    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    ggplot(aes(x = alpha, y = mean.uninfo, color = Method)) +
     geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    geom_errorbar(aes(ymin = mean.uninfo - se.uninfo, ymax = mean.uninfo + se.uninfo), width = 0.0025, alpha=0.5) +
     geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
@@
 noni_tpr <- dplyr::filter(res_noni, key == "TPR") %>%
-    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+    ggplot(aes(x = alpha, y = mean.uninfo, color = Method)) +
     geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
+    geom_errorbar(aes(ymin = mean.uninfo - se.uninfo, ymax = mean.uninfo + se.uninfo), width = 0.0025, alpha=0.5) +
     facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
@@
 
+gp <- plot_grid(noni_fdr + guides(color = FALSE, linetype = FALSE),
+                noni_tpr + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:2], ncol = 1)
+gp <- plot_grid(gp, get_legend(noni_tpr), rel_widths = c(1, 1/16))
+gp
+
 gp <- plot_grid(noni_fdr + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
@@
 
-We next take a look at noninformative covariates against informative covariates.
-
-# Noninformative vs. Informative
-
-```{r, fig.width = 10, fig.height = 4}
-res_compc <- dplyr::filter(res, grepl("informative-cubic", setting) |
-                                grepl("noninformative", setting),
-                           grepl("N\\(0, 1\\)", dist))
-res_compc <- tidyr::complete(res_compc, setting, Method, dist, key)
-res_compc <- dplyr::mutate(res_compc, lab = round(100*alpha))
-res_compc <- dplyr::mutate(res_compc, lab = ifelse(lab %in% c(1, 10), lab, NA))
-
-compc_fdr <- dplyr::filter(res_compc, key == "FDR") %>%
-    dplyr::select(setting, Method, dist, alpha, lab, mean_value) %>%
-    tidyr::spread(setting, mean_value) %>%
-    ggplot(aes(x = noninformative, y = `informative-cubic`, label = lab,
-               color = Method, linetype = Method)) +
-    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
-    geom_hline(yintercept = c(.01, .1), lty = 2, color = "blue", alpha = 1/2) + 
-    geom_vline(xintercept = c(.01, .1), lty = 2, color = "blue", alpha = 1/2) + 
-    geom_line(alpha = 0.85) +
-    geom_text(size = 3) + 
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("noninformative (FDP)", breaks = seq(0, 1, .02), labels = scales::percent) + 
-    scale_y_continuous("informative (FDP)", breaks = seq(0, 1, .02), labels = scales::percent) + 
-    ggtitle("Mean FDP Over 100 Replications") +
-    expand_limits(x = 0, y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-compc_tpr <- dplyr::filter(res_compc, key == "TPR") %>%
-    dplyr::select(setting, Method, dist, alpha, lab, mean_value) %>%
-    tidyr::spread(setting, mean_value) %>%
-    ggplot(aes(x = noninformative, y = `informative-cubic`, label = lab,
-               color = Method, linetype = Method)) +
-    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
-    geom_line(alpha = 0.85) +
-    geom_text(size = 3) + 
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("noninformative (TPR)", labels = scales::percent) + 
-    scale_y_continuous("informative (TPR)", labels = scales::percent) + 
-    ggtitle("Mean TPR Over 100 Replications") +
-    expand_limits(x = 0, y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-gp <- plot_grid(compc_fdr + guides(color = FALSE, linetype = FALSE),
-                compc_tpr + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:2], nrow = 1)
-gp <- plot_grid(gp, get_legend(compc_tpr), rel_widths = c(1, 1/8))
-gp
-```
-
-We can also plot the differences.
-
-```{r, fig.width = 10, fig.height = 4}
-compc2_fdr <- dplyr::filter(res_compc, key == "FDR") %>%
-    dplyr::select(setting, Method, dist, alpha, lab, mean_value) %>%
-    tidyr::spread(setting, mean_value) %>%
-    dplyr::mutate(gain = `informative-cubic` - noninformative) %>%
-    ggplot(aes(x = alpha, y = gain, color = Method, linetype = Method)) +
-    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
-    geom_line(alpha = 0.85) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
-    scale_y_continuous("change in FDP", breaks = seq(-1, 1, .001), labels = scales::percent) + 
-    ggtitle("Mean Change in FDP Over 100 Replications") +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-compc2_tpr <- dplyr::filter(res_compc, key == "TPR") %>%
-    dplyr::select(setting, Method, dist, alpha, lab, mean_value) %>%
-    tidyr::spread(setting, mean_value) %>%
-    dplyr::mutate(gain = `informative-cubic` - noninformative) %>%
-    ggplot(aes(x = alpha, y = gain, color = Method, linetype = Method)) +
-    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
-    geom_line(alpha = 0.85) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
-    scale_y_continuous("change in TPR", labels = scales::percent) + 
-    ggtitle("Mean Change in TPR Over 100 Replications") +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-gp <- plot_grid(compc2_fdr + guides(color = FALSE, linetype = FALSE),
-                compc2_tpr + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:2], nrow = 1)
-gp <- plot_grid(gp, get_legend(compc2_tpr), rel_widths = c(1, 1/8))
-gp
-```
-
 # Null Proportions
@@
 
-```{r, fig.width = 10, fig.height = 8}
-res_pi0 <- dplyr::filter(res, setting == "varyingpi0-lowsignal", alpha %in% c(0.05, 0.10))
-res_pi0 <- tidyr::complete(res_pi0, setting, Method, alpha, pi0, key)
+```{r, fig.width = 10, fig.height = 4}
+res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0-lowsignal",
+                         alpha == 0.05)
+res_pi0 <- dplyr::mutate(res_pi0,
+                         mean.info = ifelse(is.nan(mean.info) & key == "FDR", 0, mean.info))
 
 pi0_fdr <- dplyr::filter(res_pi0, key == "FDR") %>%
-    ggplot(aes(x = pi0, y = mean_value, color = Method)) +
+    ggplot(aes(x = pi0, y = mean.info, color = Method)) +
     geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 2.5, alpha=0.5) +
+    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
     geom_hline(aes(yintercept = alpha), lty = 2, color = "blue", alpha = 1/2) + 
-    facet_grid(. ~ alpha, scales = "free_x", labeller = label_both) +
     scale_linetype_manual(values = lty) + 
@@
     scale_x_continuous("pi0", breaks = seq(0, 100, 10)) +
-    scale_y_continuous("FDP", labels = scales::percent) +
-    ggtitle("Mean FDP Over 100 Replications") +
+    scale_y_continuous("FDR", labels = scales::percent) +
+    ggtitle("Mean FDR at alpha = 0.05 Over 100 Replications") +
     expand_limits(x =0, y = 0) +
@@
 pi0_tpr <- dplyr::filter(res_pi0, key == "TPR") %>%
-    ggplot(aes(x = pi0, y = mean_value, color = Method)) +
+    ggplot(aes(x = pi0, y = mean.info, color = Method)) +
     geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 2.5, alpha=0.5) +
-    facet_grid(. ~ alpha, scales = "free_x", labeller = label_both) +
+    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
     scale_linetype_manual(values = lty) + 
@@
     scale_y_continuous("TPR", labels = scales::percent) +
-    ggtitle("Mean TPR Over 100 Replications") +
+    ggtitle("Mean TPR at alpha = 0.05 Over 100 Replications") +
     expand_limits(x =0, y = 0) +
@@
 
-gp <- plot_grid(pi0_fdr + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
+gp <- plot_grid(pi0_fdr + guides(color = FALSE, linetype = FALSE),
                 pi0_tpr + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:2], ncol = 1)
+                labels = LETTERS[1:2], ncol = 2)
 gp <- plot_grid(gp, get_legend(pi0_tpr), rel_widths = c(1, 1/8))
@@
 
-```{r, fig.width = 10, fig.height = 8}
-res_info <- dplyr::filter(res, setting == "varyinginfo-lowsignal", alpha %in% c(0.05, 0.10))
-res_info <- tidyr::complete(res_info, setting, Method, alpha, inform, key)
+```{r, fig.width = 10, fig.height = 4}
+res_info <- dplyr::filter(resmet, setting == "varyinginfo-lowsignal", alpha == 0.05)
 
 info_fdr <- dplyr::filter(res_info, key == "FDR") %>%
-    ggplot(aes(x = inform, y = mean_value, color = Method)) +
+    ggplot(aes(x = inform, y = mean.info, color = Method)) +
     geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 2.5, alpha=0.5) +
+    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
     geom_hline(aes(yintercept = alpha), lty = 2, color = "blue", alpha = 1/2) + 
-    facet_grid(. ~ alpha, scales = "free_x", labeller = label_both) +
     scale_linetype_manual(values = lty) + 
@@
     scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
-    scale_y_continuous("FDP", labels = scales::percent) +
-    ggtitle("Mean FDP Over 100 Replications") +
+    scale_y_continuous("FDR", labels = scales::percent) +
+    ggtitle("Mean FDR at alpha = 0.05 Over 100 Replications") +
     expand_limits(x =0, y = 0) +
@@
 info_tpr <- dplyr::filter(res_info, key == "TPR") %>%
-    ggplot(aes(x = inform, y = mean_value, color = Method)) +
+    ggplot(aes(x = inform, y = mean.info, color = Method)) +
     geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 2.5, alpha=0.5) +
-    facet_grid(. ~ alpha, scales = "free_x", labeller = label_both) +
+    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
     scale_linetype_manual(values = lty) + 
@@
                 info_tpr + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:2], ncol = 1)
+                labels = LETTERS[1:2], ncol = 2)
 gp <- plot_grid(gp, get_legend(info_tpr), rel_widths = c(1, 1/8))
@@
 
-# Unimodal Effects
+We also look at the difference between informative and uninformative covariates.
 
-```{r, fig.width = 18, fig.height = 12}
-res_uni <- dplyr::filter(res, setting == "varyinges")
-res_uni <- dplyr::mutate(res_uni, dist = droplevels(dist))
-res_uni <- tidyr::complete(res_uni, Method, signal, dist, key)
-
-dplyr::filter(res_uni, key == "FDR") %>%
-    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+```{r, fig.width = 10, fig.height = 4}
+info_fdr_diff <- dplyr::filter(res_info, key == "FDR") %>%
+    ggplot(aes(x = inform, y = mean.diff, color = Method)) +
     geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
-    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
-    facet_grid(dist ~ signal, scales = "free_x", labeller = label_parsed) +
+    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 2.5, alpha=0.5) +
+    geom_hline(aes(yintercept = 0), lty = 2, color = "blue", alpha = 1/2) + 
     scale_linetype_manual(values = lty) + 
     scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
-    scale_y_continuous("FDP", labels = scales::percent) +
-    ggtitle("Mean FDP Over 100 Replications") +
-    expand_limits(y = 0) +
+    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
+    scale_y_continuous("FDR", labels = scales::percent) +
+    ggtitle("Mean FDR Difference at alpha = 0.05 Over 100 Replications") +
+    expand_limits(x =0, y = 0) +
     theme_classic() +
@@
 
-dplyr::filter(res_uni, key == "TPR") %>%
-    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
+info_tpr_diff <- dplyr::filter(res_info, key == "TPR") %>%
+    ggplot(aes(x = inform, y = mean.diff, color = Method)) +
     geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
-    facet_grid(dist ~ signal, scales = "free", labeller = label_parsed) +
+    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 2.5, alpha=0.5) +
+    geom_hline(aes(yintercept = 0), lty = 2, color = "blue", alpha = 1/2) + 
     scale_linetype_manual(values = lty) + 
     scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
+    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
     scale_y_continuous("TPR", labels = scales::percent) +
-    ggtitle("Mean TPR Over 100 Replications") +
-    expand_limits(y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-```
-
-We replot this only looking at Normal sampling noise.
-
-```{r, fig.width = 18, fig.height = 8}
-res_uni <- dplyr::filter(res_uni, grepl("N\\(0, 1\\)", dist))
-
-uni_fdr <- dplyr::filter(res_uni, key == "FDR") %>%
-    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
-    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
-    facet_wrap(~ signal, scales = "free_x", label = label_both, nrow = 1) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
-    scale_y_continuous("FDP", labels = scales::percent) +
-    ggtitle("Mean FDP Over 100 Replications") +
-    expand_limits(y = 0) +
+    ggtitle("Mean TPR Difference at alpha = 0.05 Over 100 Replications") +
+    expand_limits(x =0, y = 0) +
     theme_classic() +
@@
 
-uni_tpr <- dplyr::filter(res_uni, key == "TPR") %>%
-    ggplot(aes(x = alpha, y = mean_value, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.0025, alpha=0.5) +
-    facet_wrap(~ signal, scales = "free", label = label_both, nrow = 1) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
-    scale_y_continuous("TPR", labels = scales::percent) +
-    ggtitle("Mean TPR Over 100 Replications") +
-    expand_limits(y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-gp <- plot_grid(uni_fdr + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
-                uni_tpr + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:2], ncol = 1)
-gp <- plot_grid(gp, get_legend(uni_tpr), rel_widths = c(1, 1/16))
+gp <- plot_grid(info_fdr_diff + guides(color = FALSE, linetype = FALSE),
+                info_tpr_diff + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:2], ncol = 2)
+gp <- plot_grid(gp, get_legend(info_tpr_diff), rel_widths = c(1, 1/8))
 gp
 ```
+
+# Figure Sim1
+
+We combine some of the results from above into a single figure.
+
+```{r, fig.width = 10, fig.height = 11}
+```

### 2018-08-02 datasets/RNA-Seq/polyester-simulation.Rmd
Factor model matrix for polyester; polyester assumes effect sizes are log fold changes

@@
   # things needed to simulate with Polyester
-  groupPolyester = rep(c(-1,1),each=sampleSize)
+  groupPolyester = rep(c(0,1),each=sampleSize)
   modPolyester = model.matrix(~-1 + groupPolyester)
@@
     log2FC[DE] <- rnorm(nDE, sd = 1)
+    # polyester assumes FC is on the log scale
     dat_alt[DE,] = create_read_numbers(mu[DE], paramPolyester$fit, p0[DE], 
-                                  n=sampleSize*2, beta=cbind(log2FC[DE]/2), 
+                                  n=sampleSize*2, beta=cbind(log(2^log2FC[DE])), 
                                   mod=modPolyester)

### 2018-08-05 datasets/simulations/simulations-informative-cubic.Rmd
Update sim parameter names in vignettes to match new sim function

@@
 ```{r gauss-parameters}
-tstat <- rnorm_generator(3)       # functional: dist of alternative test stats
-tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
+es_dist <- rnorm_generator(3)       # functional: dist of alternative test stats
+ts_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
 null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
@@
     res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
-                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
-                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    pi0 = pi0, es_dist = es_dist, icovariate = icovariate,
+                    ts_dist = ts_dist, null_dist = null_dist,
                     seed = seed, mc.cores = cores)
@@
 ```{r gauss-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, gauss-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r t5-parameters}
-tstat <- rnorm_generator(6)    # functional: dist of alternative test stats
-tstat_dist <- rt_perturber(5)  # functional: sampling dist/noise for test stats
+es_dist <- rnorm_generator(6)    # functional: dist of alternative test stats
+ts_dist <- rt_perturber(5)  # functional: sampling dist/noise for test stats
 null_dist <- rt_2pvaluer(5)    # functional: dist to calc p-values
@@
     res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplust, m = m,
-                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
-                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    pi0 = pi0, es_dist = es_dist, icovariate = icovariate,
+                    ts_dist = ts_dist, null_dist = null_dist,
                     seed = seed, mc.cores = cores)
@@
 ```{r t5-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, t5-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r t11-parameters}
-tstat <- rnorm_generator(3)     # functional: dist of alternative test stats
-tstat_dist <- rt_perturber(11)  # functional: sampling dist/noise for test stats
+es_dist <- rnorm_generator(3)     # functional: dist of alternative test stats
+ts_dist <- rt_perturber(11)  # functional: sampling dist/noise for test stats
 null_dist <- rt_2pvaluer(11)    # functional: dist to calc p-values
@@
 ```{r t11-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, t11-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r chisq4-parameters}
-tstat <- rnorm_generator(15)       # functional: dist of alternative test stats
-tstat_dist <- rchisq_perturber(4)  # functional: sampling dist/noise for test stats
+es_dist <- rnorm_generator(15)       # functional: dist of alternative test stats
+ts_dist <- rchisq_perturber(4)  # functional: sampling dist/noise for test stats
 null_dist <- rchisq_pvaluer(4)     # functional: dist to calc p-values
@@
     res <- mclapply(X = 1:B, FUN = simIteration, bench = bdchi, m = m,
-                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
-                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    pi0 = pi0, es_dist = es_dist, icovariate = icovariate,
+                    ts_dist = ts_dist, null_dist = null_dist,
                     seed = seed, mc.cores = cores)
@@
 ```{r chisq4-one-simulation}
-onerun <- simIteration(bdchi, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdchi, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, chisq4-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```

### 2018-08-05 datasets/simulations/simulations-varyingeffectsize-chisq.Rmd
Update sim parameter names in vignettes to match new sim function

@@
 pi0 <- pi0_cubic(0.90)              # numeric: proportion of null hypotheses
-tstat_dist <- rchisq_perturber(4)   # functional: sampling dist/noise for test stats
+ts_dist <- rchisq_perturber(4)   # functional: sampling dist/noise for test stats
 null_dist <- rchisq_pvaluer(4)      # functional: dist to calc p-values
@@
 ```{r spiky-parameters}
-tstat <- function(n) { abs(sampler_spiky(n)) } # functional: dist of alternative test stats
+es_dist <- function(n) { abs(sampler_spiky(n)) } # functional: dist of alternative test stats
 seed <- 778
@@
     res <- mclapply(X = 1:B, FUN = simIteration, bench = bd, m = m,
-                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
-                    tstat_dist = tstat_dist, null_dist = null_dist,
+                    pi0 = pi0, es_dist = es_dist, icovariate = icovariate,
+                    ts_dist = ts_dist, null_dist = null_dist,
                     seed = seed, mc.cores = cores)
@@
 ```{r spiky-one-simulation}
-onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bd, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, spiky-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r flattop-parameters}
-tstat <- function(n) { abs(sampler_flat_top(n)) }       # functional: dist of alternative test stats
+es_dist <- function(n) { abs(sampler_flat_top(n)) }       # functional: dist of alternative test stats
 seed <- 980
@@
 ```{r flattop-one-simulation}
-onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bd, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, flattop-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r skew-parameters}
-tstat <- function(n) { abs(sampler_skew(n)) }       # functional: dist of alternative test stats
+es_dist <- function(n) { abs(sampler_skew(n)) }       # functional: dist of alternative test stats
 seed <- 206
@@
 ```{r skew-one-simulation}
-onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bd, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, skew-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r bimodal-parameters}
-tstat <- function(n) { abs(sampler_bimodal(n)) }       # functional: dist of alternative test stats
+es_dist <- function(n) { abs(sampler_bimodal(n)) }       # functional: dist of alternative test stats
 seed <- 913
@@
 ```{r bimodal-one-simulation}
-onerun <- simIteration(bd, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bd, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, bimodal-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```

### 2018-08-05 datasets/simulations/simulations-varyingeffectsize-t.Rmd
Update sim parameter names in vignettes to match new sim function

@@
 pi0 <- pi0_cubic(0.90)              # numeric: proportion of null hypotheses
-tstat_dist <- rt_perturber(11)      # functional: sampling dist/noise for test stats
+ts_dist <- rt_perturber(11)      # functional: sampling dist/noise for test stats
 null_dist <- rt_2pvaluer(11)        # functional: dist to calc p-values
@@
 ```{r spiky-parameters}
-tstat <- sampler_spiky       # functional: dist of alternative test stats
+es_dist <- sampler_spiky       # functional: dist of alternative test stats
 seed <- 778
@@
 ```{r spiky-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r flattop-parameters}
-tstat <- sampler_flat_top       # functional: dist of alternative test stats
+es_dist <- sampler_flat_top       # functional: dist of alternative test stats
 seed <- 980
@@
 ```{r flattop-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r skew-parameters}
-tstat <- sampler_skew       # functional: dist of alternative test stats
+es_dist <- sampler_skew       # functional: dist of alternative test stats
 seed <- 206
@@
 ```{r skew-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r bimodal-parameters}
-tstat <- sampler_bimodal       # functional: dist of alternative test stats
+es_dist <- sampler_bimodal       # functional: dist of alternative test stats
 seed <- 913
@@
 ```{r bimodal-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)

### 2018-08-05 datasets/simulations/simulations-varyingeffectsize.Rmd
Update sim parameter names in vignettes to match new sim function

@@
 pi0 <- pi0_cubic(0.90)              # numeric: proportion of null hypotheses
-tstat_dist <- rnorm_perturber(1/2)  # functional: sampling dist/noise for test stats
+ts_dist <- rnorm_perturber(1/2)  # functional: sampling dist/noise for test stats
 null_dist <- rnorm_2pvaluer(1/2)    # functional: dist to calc p-values

### 2018-08-06 manuscript_figures/manuscript_figures.Rmd
Add adapt to manuscript figs

@@
 # not excluding any methods
-Fig4All.rowOrder <- c("bonf", "bh", "ihw", "qvalue", "bl", "lfdr",
+Fig4All.rowOrder <- c("bonf", "bh", "ihw", "qvalue", "bl", "adapt-glm", "lfdr",
                      "fdrreg-e", "fdrreg-t", "ashq")
@@
         theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
-              legend.position = c(0.98, 0.93), legend.justification = c(1, 1),
+              legend.position = c(0.98, 0.91), legend.justification = c(1, 1),
               legend.background = element_rect(fill = scales::alpha("white", 3/4), color = 'black'),
@@
 ## specify method order
-method_order <- c("bonf", "bh", "ihw", "qvalue", "bl", "lfdr",
+method_order <- c("bonf", "bh", "ihw", "qvalue", "bl", "adapt-glm", "lfdr",
                   "fdrreg-e", "fdrreg-t", "ashq")
@@
 # methods that use a covariate
-cov.methods <- c("ihw", "bl", "lfdr", "fdrreg-e", "fdrreg-t")
+cov.methods <- c("ihw", "bl", "adapt-glm", "lfdr", "fdrreg-e", "fdrreg-t")
 
@@
 
-rowOrder <- c("ihw", "bl", "lfdr", "fdrreg-e", "fdrreg-t")
+rowOrder <- c("ihw", "bl", "adapt-glm", "lfdr", "fdrreg-e", "fdrreg-t")
 

### 2018-08-06 datasets/simulations/simulations-informative-cosine.Rmd

@@
 
-resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key, performanceMetric, alpha, value),
-                            select(resdfu, rep, blabel, param.alpha, key, performanceMetric, alpha, value),
-                            by = c("rep", "blabel", "param.alpha", "key", "performanceMetric", "alpha"),
+resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            select(resdfu, rep, blabel, param.alpha, key,
+                                   performanceMetric, alpha, value),
+                            by = c("rep", "blabel", "param.alpha", "key",
+                                   "performanceMetric", "alpha"),
                             suffix = c(".info", ".uninfo"))

### 2018-08-06 datasets/simulations/simulations-null.Rmd

@@
 pi0 <- 1                          # numeric: proportion of null hypotheses
-tstat <- function(x) { 0 }        # functional: dist of alternative test stats
+es_dist <- function(x) { 0 }        # functional: dist of alternative test stats
 icovariate <- runif               # functional: independent covariate
@@
 ```{r gauss-parameters}
-tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
+ts_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
 null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
@@
 ```{r t5-parameters}
-tstat_dist <- rt_perturber(5)  # functional: sampling dist/noise for test stats
+ts_dist <- rt_perturber(5)  # functional: sampling dist/noise for test stats
 null_dist <- rt_2pvaluer(5)    # functional: dist to calc p-values
@@
 ```{r t11-parameters}
-tstat_dist <- rt_perturber(11)  # functional: sampling dist/noise for test stats
+ts_dist <- rt_perturber(11)  # functional: sampling dist/noise for test stats
 null_dist <- rt_2pvaluer(11)    # functional: dist to calc p-values
@@
 ```{r chisq4-parameters}
-tstat_dist <- rchisq_perturber(4)  # functional: sampling dist/noise for test stats
+ts_dist <- rchisq_perturber(4)  # functional: sampling dist/noise for test stats
 null_dist <- rchisq_pvaluer(4)     # functional: dist to calc p-values

### 2018-08-06 datasets/simulations/simulations-varyinginformativeness.Rmd

--- 2018-07-26 datasets/simulations/simulations-varyinginformativeness.Rmd
@@
 m <- 20000                        # integer: number of hypothesis tests
-tstat <- rnorm_generator(2)       # functional: dist of alternative test stats
-tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
+es_dist <- rnorm_generator(2)       # functional: dist of alternative test stats
+ts_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
 null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
@@
 ```{r info00-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, info00-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r info20-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, info20-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r info40-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, info40-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r info60-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, info60-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r info80-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, info80-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r info100-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, info100-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```

### 2018-08-06 datasets/simulations/simulations-varyingntests.Rmd

@@
 ```{r parameters-shared}
-tstat <- rnorm_generator(3)       # functional: dist of alternative test stats
-tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
+es_dist <- rnorm_generator(3)       # functional: dist of alternative test stats
+ts_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
 null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
@@
 ```{r n100-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, n100-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r n500-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, n500-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r n1000-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, n1000-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r n5000-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, n5000-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r n10000-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, n10000-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r n50000-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, n50000-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```

### 2018-08-06 datasets/simulations/simulations-varyingpi0.Rmd

@@
 ```{r null05-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, null05-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r null10-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, null10-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r null20-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, null20-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r null30-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, null30-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r null40-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, null40-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r null50-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, null50-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r null60-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, null60-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r null70-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, null70-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r null80-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, null80-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r null90-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, null90-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r null95-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, null95-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```
@@
 ```{r null99-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
+onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 ```{r, null99-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
+strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
 ```

### 2018-08-06 manuscript_figures/manuscript_figures.Rmd
Adjust margins of Figure 4

@@
         theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
-              legend.position = c(0.98, 0.91), legend.justification = c(1, 1),
+              legend.position = c(0.97, 0.92), legend.justification = c(1, 1),
               legend.background = element_rect(fill = scales::alpha("white", 3/4), color = 'black'),
@@
 ggdraw() +
-  draw_plot(pfdr, 0.53, 0, 0.23, 1) +
-  draw_plot(ptpr, 0.76, 0, 0.23, 1) +
+  draw_plot(pfdr, 0.53, 0, 0.24, 1) +
+  draw_plot(ptpr, 0.77, 0, 0.24, 1) +
   draw_plot(Fig4All, 0, 0, 0.52, 1) +
@@
   draw_plot_label("Simulation", 0.69, 0.06, size =11, fontface = "plain")
-ggsave(file.path(outdir, "Figure4.pdf"), width=9, height=5)
+ggsave(file.path(outdir, "Figure4.pdf"), width=9.5, height=5)
 

### 2018-08-06 manuscript_figures/manuscript_figures.Rmd
Remove Fig 3; adjust legend pos. in inf. cov. plot; fix typo in xlab of rankings boxplot

@@
 
-This is a side-by-side figure displaying the TPR by alpha cutoff, faceted by 
-whether the method uses multiple pieces of information or just one (the p-value).
-This figure uses the same data as Figure 1.
-
-```{r, Figure3, fig.width = 8, fig.height = 4}
-# Figure 3: TPR of methods using by pieces of information used 
-Fig3 <- plotsim_average(res5d, met="TPR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE, facetMethodType = TRUE)
-
-Fig3
-ggsave(file.path(outdir, "Figure3.pdf"), width=8, height=4)
-```
+Replaced with a pure simulation figure - see `manuscript_figures_sims.Rmd`.
 
@@
   theme_bw() +
-  xlab("Proportion of maximum rejections in yeast simulation (n=5s)") +  
+  xlab("Proportion of maximum rejections in yeast simulation (n=5)") +  
   ylab("Proportion of maximum rejections in casestudies") +
@@
   dplyr::rename("uninf" = "nrejects")
-  #dplyr::rename(nrejects = "uninf")
 
@@
         theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
-              legend.position = c(0.985, 0.91), legend.justification = c(1, 1),
+              legend.position = c(0.985, 0.93), legend.justification = c(1, 1),
               legend.background = element_rect(fill = scales::alpha("white", 3/4), color = 'black'),

### 2018-08-06 datasets/simulations/simulations-varyingeffectsize.Rmd

@@
 pi0 <- pi0_cubic(0.90)              # numeric: proportion of null hypotheses
-ts_dist <- rnorm_perturber(1/2)  # functional: sampling dist/noise for test stats
-null_dist <- rnorm_2pvaluer(1/2)    # functional: dist to calc p-values
+ts_dist <- rnorm_perturber(1)       # functional: sampling dist/noise for test stats
+null_dist <- rnorm_2pvaluer(1)      # functional: dist to calc p-values
 icovariate <- runif                 # functional: independent covariate

### 2018-08-09 datasets/simulations/simulations-varyinginformativeness.Rmd
Remove 'lowsignal' from file names, not running 2 signal levels

@@
 ## intermediary files we create below
-info00_file <- file.path(resdir, "varyinginfo-lowsignal-benchmark-level00.rds")
-info20_file <- file.path(resdir, "varyinginfo-lowsignal-benchmark-level20.rds")
-info40_file <- file.path(resdir, "varyinginfo-lowsignal-benchmark-level40.rds")
-info60_file <- file.path(resdir, "varyinginfo-lowsignal-benchmark-level60.rds")
-info80_file <- file.path(resdir, "varyinginfo-lowsignal-benchmark-level80.rds")
-info100_file <- file.path(resdir, "varyinginfo-lowsignal-benchmark-level100.rds")
+info00_file <- file.path(resdir, "varyinginfo-benchmark-level00.rds")
+info20_file <- file.path(resdir, "varyinginfo-benchmark-level20.rds")
+info40_file <- file.path(resdir, "varyinginfo-benchmark-level40.rds")
+info60_file <- file.path(resdir, "varyinginfo-benchmark-level60.rds")
+info80_file <- file.path(resdir, "varyinginfo-benchmark-level80.rds")
+info100_file <- file.path(resdir, "varyinginfo-benchmark-level100.rds")
 

### 2018-08-09 datasets/simulations/simulations-varyingntests.Rmd
Remove 'lowsignal' from file names, not running 2 signal levels

@@
 ## intermediary files we create below
-n100_file <- file.path(resdir, "varyingntests-lowsignal-benchmark-n100.rds")
-n500_file <- file.path(resdir, "varyingntests-lowsignal-benchmark-n500.rds")
-n1000_file <- file.path(resdir, "varyingntests-lowsignal-benchmark-n1000.rds")
-n5000_file <- file.path(resdir, "varyingntests-lowsignal-benchmark-n5000.rds")
-n10000_file <- file.path(resdir, "varyingntests-lowsignal-benchmark-n10000.rds")
-n50000_file <- file.path(resdir, "varyingntests-lowsignal-benchmark-n50000.rds")
+n100_file <- file.path(resdir, "varyingntests-benchmark-n100.rds")
+n500_file <- file.path(resdir, "varyingntests-benchmark-n500.rds")
+n1000_file <- file.path(resdir, "varyingntests-benchmark-n1000.rds")
+n5000_file <- file.path(resdir, "varyingntests-benchmark-n5000.rds")
+n10000_file <- file.path(resdir, "varyingntests-benchmark-n10000.rds")
+n50000_file <- file.path(resdir, "varyingntests-benchmark-n50000.rds")
 

### 2018-08-09 datasets/simulations/simulations-varyingpi0.Rmd
Remove 'lowsignal' from file names, not running 2 signal levels

@@
 ## intermediary files we create below
-null05_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop05.rds")
-null10_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop10.rds")
-null20_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop20.rds")
-null30_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop30.rds")
-null40_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop40.rds")
-null50_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop50.rds")
-null60_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop60.rds")
-null70_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop70.rds")
-null80_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop80.rds")
-null90_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop90.rds")
-null95_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop95.rds")
-null99_file <- file.path(resdir, "varyingpi0-lowsignal-benchmark-nullprop99.rds")
+null05_file <- file.path(resdir, "varyingpi0-benchmark-nullprop05.rds")
+null10_file <- file.path(resdir, "varyingpi0-benchmark-nullprop10.rds")
+null20_file <- file.path(resdir, "varyingpi0-benchmark-nullprop20.rds")
+null30_file <- file.path(resdir, "varyingpi0-benchmark-nullprop30.rds")
+null40_file <- file.path(resdir, "varyingpi0-benchmark-nullprop40.rds")
+null50_file <- file.path(resdir, "varyingpi0-benchmark-nullprop50.rds")
+null60_file <- file.path(resdir, "varyingpi0-benchmark-nullprop60.rds")
+null70_file <- file.path(resdir, "varyingpi0-benchmark-nullprop70.rds")
+null80_file <- file.path(resdir, "varyingpi0-benchmark-nullprop80.rds")
+null90_file <- file.path(resdir, "varyingpi0-benchmark-nullprop90.rds")
+null95_file <- file.path(resdir, "varyingpi0-benchmark-nullprop95.rds")
+null99_file <- file.path(resdir, "varyingpi0-benchmark-nullprop99.rds")
 

### 2018-08-09 datasets/simulations/simulations-summary.Rmd
Move simulation summary plots to dataset dir; not yet manuscript figures

---
title: "Simulation Summary Figures (simulations)"
author: "Rafalab"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Figure2-load-data}
# Load packages and source benchmark FDR
library(tidyr)
library(dplyr)
library(ggplot2)
library(magrittr)
library(cowplot)
library(tibble)
library(ggthemes)
library(grid)
library(SummarizedBenchmark)

## load helper functions
for (f in list.files("../datasets/R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

# assumes sb objects are in the following location
path <- file.path("..", "datasets", "simulations", "results")

# set up results directory
outdir <- "./figures"
dir.create(outdir, showWarnings = FALSE)

methodset <- c("bonf", "bh", "ihw", "qvalue", "bl-df03", "lfdr",
               "fdrreg-e", "fdrreg-t", "ashq", "adapt-glm")
```

# Data Preparation

First, we load summaries for all simulations results at significance thresholds
between 0.01 and 0.10. This process can take a long time (approximately 40 minutes
per alpha cutoff) because of the large number of simulation settings that must be
read in and summarized. Ideally, this initial step should be run externally separate
from this Rmd (by default, the following code chunk will not be run).

```{r, eval = FALSE}
for (ia in c(1:10, 20, 30, 40, 50)) {
    summary_file <- file.path("results", paste0("summary_alpha", ia, ".rds"))
    if (!file.exists(summary_file)) {
        res_files <- list.files(path, "\\.rds$", full.names = TRUE)
        res <- lapply(res_files, function(x) {
            zz <- readRDS(x)
            zz_i <- lapply(zz, `[[`, "informative")
            zz_u <- lapply(zz, `[[`, "uninformative")
            zz_i <- plotsim_standardize(zz_i, alpha = ia / 100L)
            zz_i <- dplyr::select(zz_i, rep, blabel, param.alpha, key, performanceMetric, alpha, value)
            zz_u <- plotsim_standardize(zz_u, alpha = ia / 100L)
            zz_u <- dplyr::select(zz_u, rep, blabel, param.alpha, key, performanceMetric, alpha, value)
            dplyr::full_join(zz_i, zz_u,
                             by = c("rep", "blabel", "param.alpha", "key", "performanceMetric", "alpha"),
                             suffix = c(".info", ".uninfo"))
        })
        names(res) <- gsub("\\.rds$", "", basename(res_files))
        saveRDS(res, summary_file)
    }
}
```

Assuming that the summaries have been saved for all simulation settings at
the various FDR thresholds using the code above, we load the summarized results
and merge them into a single results table.

```{r}
alphas <- 1:10
res <- vector("list", length(alphas))
for (ia in seq_len(length(alphas))) {
    imethodset <- methodset
    imethodset[which(imethodset == "ihw")] <- paste0("ihw-a", sprintf("%02d", alphas[ia]))
    ires <- readRDS(file.path("results", paste0("summary_alpha", alphas[ia], ".rds")))
    ires <- bind_rows(ires, .id = "setting")
    ires <- dplyr::mutate(ires, dist = gsub(".*?-benchmark-(.*)", "\\1", setting),
                         setting = gsub("(.*?)-benchmark-.*", "\\1", setting))
    ires <- dplyr::filter(ires, blabel %in% imethodset)
    ires <- dplyr::mutate(ires, blabel = gsub("(-a)(.*)", "", blabel))
    res[[ia]] <- ires
}
res <- bind_rows(res)
res <- dplyr::select(res, -param.alpha)
res <- dplyr::rename(res, Method = blabel)
res <- dplyr::mutate(res, Method = gsub("-df03", "", Method))
res <- dplyr::mutate(res, Method = factor(Method, levels = gsub("-df03", "", methodset)))
```

We compute summary metrics for each method, setting, and cutoff.

```{r}
resmet <- dplyr::mutate(res, value.info = ifelse(is.na(value.info) & grepl("R$", key), 0, value.info),
                        value.uninfo = ifelse(is.na(value.uninfo) & grepl("R$", key), 0, value.uninfo))
resmet <- dplyr::group_by(resmet, setting, dist, Method, key, performanceMetric, alpha)
resmet <- dplyr::summarize(resmet,
                           mean.info = mean(value.info, na.rm = TRUE),
                           se.info = sd(value.info, na.rm = TRUE) / sqrt(sum(!is.na(value.info))),
                           nNA.info = sum(is.na(value.info)),
                           mean.diff = mean(value.info - value.uninfo, na.rm = TRUE),
                           se.diff = sd(value.info - value.uninfo, na.rm = TRUE) /
                               sqrt(sum(!is.na(value.info - value.uninfo))))
resmet <- dplyr::ungroup(resmet)
```

We clean up the annotations in the table for some of the more complex simulation settings.

```{r}
resmet <- dplyr::mutate(resmet, inform = ifelse(grepl("varyinginfo", setting), dist, NA))
resmet <- dplyr::mutate(resmet, inform = as.numeric(gsub("level", "", inform)))
resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("varyinginfo", setting), "gaussian", dist))

resmet <- dplyr::mutate(resmet, pi0 = ifelse(grepl("varyingpi0", setting), dist, 90))
resmet <- dplyr::mutate(resmet, pi0 = as.numeric(gsub("nullprop", "", pi0)))
resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("varyingpi0", setting), "gaussian", dist))

resmet <- dplyr::mutate(resmet, signal = ifelse(grepl("varyinges", setting), dist, NA))
resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyinges$", setting), "gaussian", dist))
resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyinges-t$", setting), "t11", dist))
resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyinges-chisq$", setting), "chisq4", dist))
resmet <- dplyr::mutate(resmet, setting = ifelse(grepl("varyinges", setting), "varyinges", setting))
```

We also replace the sampling noise labels with more plotting-friendly labels.

```{r}
resmet <- dplyr::mutate(resmet, dist = factor(dist, levels = c("gaussian", "t11", "t5", "chisq4"),
                                              labels = c(expression("Noise: " * N(0, 1)),
                                                         expression("Noise: " * t[11]),
                                                         expression("Noise: " * t[5]),
                                                         expression("Noise: " * {chi^2}[4]))))
```

We use the standardize candy color scheme and line types for the plots.

```{r}
col <- as.character(candycols$col)
names(col) <- as.character(candycols$Method)
lty <- as.character(candycols$lty)
names(lty) <- as.character(candycols$Method)
```

There are several different simulation settings that can be investigated.

```{r}
dplyr::distinct(resmet, setting)
```

Namely, simulation settings were run with to investigate the impact of different
informative covariates, noninformative covariates, varying null proportions,
varying covariate informativeness, and varying effect size distributions (including
unimodal effects as described in the ASH manuscript). Many of these simulation
settings were also run with different sampling noise distributions, including
Normal/Gaussian, t with 11 degrees of freedom, t with 5 degrees of freedom, and
Chi-squared with 4 degrees of freedom.

Results are reproted for the Chi-Squared simulation settings without `ashq`,
`fdrreg-e` and `fdrreg-t`, since the assumptions of these methods are clearly
violated under these settings, and it is neither appropriate nor fair to include
them in the comparisons.

```{r}
resmet <- dplyr::filter(resmet, !( grepl("chi", dist) & Method == "ashq" ))
resmet <- dplyr::filter(resmet, !( grepl("chi", dist) & grepl("fdrreg", Method) ))
```

# Null Simulations

We subset on simulation results that look at informative and non-informative covariates across
different noise distributions.

For null simulations, rather than the false discovery rate, we take a look at the true negative
rate (equivalently, the proportion of false positives) since the FDR will be 1 if any rejections
are made. Additionally, we take a look at the FWER (the proportion of simulation replicates out
of 100 with at least one rejection).

```{r, fig.width = 18, fig.height = 8}
res_nullset <- dplyr::filter(resmet, setting == "null")

null_tnr <- dplyr::filter(res_nullset, key == "TNR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TNR", labels = scales::percent) +
    ggtitle("Mean TNR Over 100 Replications") +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

null_fwer <- dplyr::filter(res_nullset, key == "FWER") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FWER", labels = scales::percent) +
    ggtitle("FWER Over 100 Replications") +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(null_tnr + guides(color = FALSE, linetype = FALSE),
                null_fwer + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(null_tnr), rel_widths = c(1, 1/16))
gp
```

We also look at the difference between informative and uninformative covariates.

```{r, fig.width = 18, fig.height = 8}
null_tnr_diff <- dplyr::filter(res_nullset, key == "TNR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TNR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean TNR Difference Over 100 Replications") +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

null_fwer_diff <- dplyr::filter(res_nullset, key == "FWER") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FWER\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("FWER Difference Over 100 Replications") +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(null_tnr_diff + guides(color = FALSE, linetype = FALSE),
                null_fwer_diff + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(null_tnr_diff), rel_widths = c(1, 1/16))
gp
```

# Sampling Distributions

We just take a look at different noise distributions for a single informative covariate.
We take a look at the false discovery proportion controlled along a grid of nominal alpha cutoffs
between 0.01 and 0.10 for each method, as well as the true positive rate for each method along
the set of cutoffs.

```{r, fig.width = 18, fig.height = 8}
res_noise <- dplyr::filter(resmet, setting == "informative-cosine")

noise_fdr <- dplyr::filter(res_noise, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR", labels = scales::percent) +
    ggtitle("Mean FDR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

noise_tpr <- dplyr::filter(res_noise, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR", labels = scales::percent) +
    ggtitle("Mean TPR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(noise_fdr + guides(color = FALSE, linetype = FALSE),
                noise_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(noise_tpr), rel_widths = c(1, 1/16))
gp

gp <- plot_grid(noise_fdr + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
                noise_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(noise_tpr), rel_widths = c(1, 1/16))
gp
```

We also look at the difference between informative and uninformative covariates.

```{r, fig.width = 18, fig.height = 8}
noise_fdr_diff <- dplyr::filter(res_noise, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean FDR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

noise_tpr_diff <- dplyr::filter(res_noise, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean TPR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(noise_fdr_diff + guides(color = FALSE, linetype = FALSE),
                noise_tpr_diff + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(noise_tpr_diff), rel_widths = c(1, 1/16))
gp
```

# Informative Covariates

We also take a look across different informative covariates under just the normal
sampling noise setting. As before, we take a look at the false discovery proportion
controlled along a grid of nominal alpha cutoffs between 0.01 and 0.10 for each method,
as well as the true positive rate for each method along the set of cutoffs.

```{r, fig.width = 18, fig.height = 8}
res_icov <- dplyr::filter(resmet, grepl("^informative", setting), grepl("N\\(0, 1\\)", dist))

icov_fdr <- dplyr::filter(res_icov, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR", labels = scales::percent) +
    ggtitle("Mean FDR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

icov_tpr <- dplyr::filter(res_icov, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR", labels = scales::percent) +
    ggtitle("Mean TPR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(icov_fdr + guides(color = FALSE, linetype = FALSE),
                icov_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(icov_tpr), rel_widths = c(1, 1/12))
gp
```

We also look at the difference between informative and uninformative covariates.

```{r, fig.width = 18, fig.height = 8}
icov_fdr_diff <- dplyr::filter(res_icov, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean FDR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

icov_tpr_diff <- dplyr::filter(res_icov, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean TPR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(icov_fdr_diff + guides(color = FALSE, linetype = FALSE),
                icov_tpr_diff + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(icov_tpr_diff), rel_widths = c(1, 1/12))
gp
```

# Null Proportions

We take a look at the impact of the proportion of null hypotheses on the performance
of the different methods under just the normal sampling noise setting. We plot the
false discovery proportion and true positive rate at a nominal alpha cutoff of 0.05
for each method. The sine informative covariate was used in these simulation settings.

```{r, fig.width = 10, fig.height = 4}
res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0-lowsignal", alpha == 0.05)

pi0_fdr <- dplyr::filter(res_pi0, key == "FDR") %>%
    ggplot(aes(x = pi0, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
    geom_hline(aes(yintercept = alpha), lty = 2, color = "blue", alpha = 1/2) + 
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("pi0", breaks = seq(0, 100, 10)) +
    scale_y_continuous("FDR", labels = scales::percent) +
    ggtitle("Mean FDR at alpha = 0.05 Over 100 Replications") +
    expand_limits(x =0, y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

pi0_tpr <- dplyr::filter(res_pi0, key == "TPR") %>%
    ggplot(aes(x = pi0, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("pi0", breaks = seq(0, 100, 10)) +
    scale_y_continuous("TPR", labels = scales::percent) +
    ggtitle("Mean TPR at alpha = 0.05 Over 100 Replications") +
    expand_limits(x =0, y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(pi0_fdr + guides(color = FALSE, linetype = FALSE),
                pi0_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 2)
gp <- plot_grid(gp, get_legend(pi0_tpr), rel_widths = c(1, 1/8))
gp
```

# Informativeness

We take a look at the impact of the informativeness of a covariate on the performance
of the different methods under just the normal sampling noise setting. Here, the informative
covariate is sampled uniformly between 0 and 1 and the probability of the corresponding test
being null is one value if the covariate is less than 0.8, and another value if the covariate
is above 0.8. The "informativeness" is measured by the difference between these two probabilities,
with informativeness 0 and 1 corresponding to no difference and maximal difference (i.e. 0 and 1),
respectively. We plot the false discovery proportion and true positive rate at a nominal alpha
cutoff of 0.05 for each method.

```{r, fig.width = 10, fig.height = 4}
res_info <- dplyr::filter(resmet, setting == "varyinginfo-lowsignal", alpha == 0.05)

info_fdr <- dplyr::filter(res_info, key == "FDR") %>%
    ggplot(aes(x = inform, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
    geom_hline(aes(yintercept = alpha), lty = 2, color = "blue", alpha = 1/2) + 
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
    scale_y_continuous("FDR", labels = scales::percent) +
    ggtitle("Mean FDR at alpha = 0.05 Over 100 Replications") +
    expand_limits(x =0, y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

info_tpr <- dplyr::filter(res_info, key == "TPR") %>%
    ggplot(aes(x = inform, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
    scale_y_continuous("TPR", labels = scales::percent) +
    ggtitle("Mean TPR at alpha = 0.05 Over 100 Replications") +
    expand_limits(x =0, y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(info_fdr + guides(color = FALSE, linetype = FALSE),
                info_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 2)
gp <- plot_grid(gp, get_legend(info_tpr), rel_widths = c(1, 1/8))
gp
```

We also look at the difference between informative and uninformative covariates.

```{r, fig.width = 10, fig.height = 4}
info_fdr_diff <- dplyr::filter(res_info, key == "FDR") %>%
    ggplot(aes(x = inform, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 2.5, alpha=0.5) +
    geom_hline(aes(yintercept = 0), lty = 2, color = "blue", alpha = 1/2) + 
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
    scale_y_continuous("FDR", labels = scales::percent) +
    ggtitle("Mean FDR Difference at alpha = 0.05 Over 100 Replications") +
    expand_limits(x =0, y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

info_tpr_diff <- dplyr::filter(res_info, key == "TPR") %>%
    ggplot(aes(x = inform, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 2.5, alpha=0.5) +
    geom_hline(aes(yintercept = 0), lty = 2, color = "blue", alpha = 1/2) + 
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
    scale_y_continuous("TPR", labels = scales::percent) +
    ggtitle("Mean TPR Difference at alpha = 0.05 Over 100 Replications") +
    expand_limits(x =0, y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(info_fdr_diff + guides(color = FALSE, linetype = FALSE),
                info_tpr_diff + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 2)
gp <- plot_grid(gp, get_legend(info_tpr_diff), rel_widths = c(1, 1/8))
gp
```

# Unimodal Assumption

We also take a look at the results for simulation settings where the effect sizes follow a unimodal
distribution (null and alternative, combined) as proposed in the ASH paper. We implement some of the
unimodal effect size distributions described in the paper, "flattop", "skew", and "spiky", as well
as a single bimodal setting also described in the paper, "bimodal". As before, we plot the the
false discovery proportion and true positive rate for each method along a grid of nominal FDR cutoffs
between 0.01 and 0.10. The cubic informative covariate was used in these simulation settings.

## Normal Noise

First, we look at the unimodal setting under Normal sampling noise.

```{r, fig.width = 18, fig.height = 8}
res_ua <- dplyr::filter(resmet, setting == "varyinges", grepl("N\\(0, 1\\)", dist))

noise_fdr <- dplyr::filter(res_ua, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ signal, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR", labels = scales::percent) +
    ggtitle("Mean FDR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

noise_tpr <- dplyr::filter(res_ua, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    facet_grid(. ~ signal, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR", labels = scales::percent) +
    ggtitle("Mean TPR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(noise_fdr + guides(color = FALSE, linetype = FALSE),
                noise_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(noise_tpr), rel_widths = c(1, 1/16))
gp
```

We also look at the difference between informative and uninformative covariates.

```{r, fig.width = 18, fig.height = 8}
noise_fdr_diff <- dplyr::filter(res_ua, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ signal, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean FDR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

noise_tpr_diff <- dplyr::filter(res_ua, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ signal, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean TPR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(noise_fdr_diff + guides(color = FALSE, linetype = FALSE),
                noise_tpr_diff + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(noise_tpr_diff), rel_widths = c(1, 1/16))
gp
```

## t Noise

Next, we look at the unimodal settings under t (11 degrees of freedom) distributed
sampling noise.

```{r, fig.width = 18, fig.height = 8}
res_ua <- dplyr::filter(resmet, setting == "varyinges", grepl("t\\[11\\]", dist))

noise_fdr <- dplyr::filter(res_ua, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ signal, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR", labels = scales::percent) +
    ggtitle("Mean FDR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

noise_tpr <- dplyr::filter(res_ua, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    facet_grid(. ~ signal, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR", labels = scales::percent) +
    ggtitle("Mean TPR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(noise_fdr + guides(color = FALSE, linetype = FALSE),
                noise_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(noise_tpr), rel_widths = c(1, 1/16))
gp
```

We also look at the difference between informative and uninformative covariates.

```{r, fig.width = 18, fig.height = 8}
noise_fdr_diff <- dplyr::filter(res_ua, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ signal, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean FDR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

noise_tpr_diff <- dplyr::filter(res_ua, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ signal, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean TPR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(noise_fdr_diff + guides(color = FALSE, linetype = FALSE),
                noise_tpr_diff + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(noise_tpr_diff), rel_widths = c(1, 1/16))
gp
```

### 2018-08-09 datasets/simulations/simulations-summary.Rmd
Fix links/names broken after moving analysis/changing res names

@@
 ---
-title: "Simulation Summary Figures (simulations)"
-author: "Rafalab"
+title: "Simulation Summary Figures"
+author: "Patrick Kimes"
 output: 
@@
 ## load helper functions
-for (f in list.files("../datasets/R", "\\.(r|R)$", full.names = TRUE)) {
+for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
     source(f)
@@
 ```{r, fig.width = 10, fig.height = 4}
-res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0-lowsignal", alpha == 0.05)
+res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0", alpha == 0.05)
 
@@
 ```{r, fig.width = 10, fig.height = 4}
-res_info <- dplyr::filter(resmet, setting == "varyinginfo-lowsignal", alpha == 0.05)
+res_info <- dplyr::filter(resmet, setting == "varyinginfo", alpha == 0.05)
 

### 2018-08-09 datasets/simulations/simulations-informative-cosine.Rmd
Add plots of covariate vs FDR to sims

@@
 
+We also look at the FDR as a function of the independent covariate.
+
+```{r gauss-metrics-covlineplotFDR, results = "hide"}
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR")
+```
+
 Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
@@
 
+We also look at the FDR as a function of the independent covariate.
+
+```{r t5-metrics-covlineplotFDR, results = "hide"}
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR")
+```
+
 Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
@@
 covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
+```
+
+We also look at the FDR as a function of the independent covariate.
+
+```{r t11-metrics-covlineplotFDR, results = "hide"}
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR")
 ```
@@
 
+We also look at the FDR as a function of the independent covariate.
+
+```{r chisq4-metrics-covlineplotFDR, results = "hide"}
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR")
+```
+
 Finally, (if enough methods produce rejections at `r ualpha`) we take a look at

### 2018-08-09 datasets/simulations/simulations-summary.Rmd
Plot sampling dist effect with cubic covariate to see bl
Change sim plot panel letters

@@
                 null_fwer_diff + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:2], ncol = 1)
+                labels = LETTERS[3:4], ncol = 1)
 gp <- plot_grid(gp, get_legend(null_tnr_diff), rel_widths = c(1, 1/16))
@@
 ```{r, fig.width = 18, fig.height = 8}
-res_noise <- dplyr::filter(resmet, setting == "informative-cosine")
+res_noise <- dplyr::filter(resmet, setting == "informative-cubic")
 
@@
                 noise_tpr_diff + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:2], ncol = 1)
+                labels = LETTERS[3:4], ncol = 1)
 gp <- plot_grid(gp, get_legend(noise_tpr_diff), rel_widths = c(1, 1/16))
@@
                 icov_tpr_diff + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:2], ncol = 1)
+                labels = LETTERS[3:4], ncol = 1)
 gp <- plot_grid(gp, get_legend(icov_tpr_diff), rel_widths = c(1, 1/12))
@@
                 info_tpr_diff + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:2], ncol = 2)
+                labels = LETTERS[3:4], ncol = 2)
 gp <- plot_grid(gp, get_legend(info_tpr_diff), rel_widths = c(1, 1/8))

### 2018-08-10 datasets/RNA-Seq/yeast-simulation-bimodalalternative.Rmd
Add yeast simulation with bimodal alternative (FDRreg-empirical assumption)

--- 2018-07-25 datasets/RNA-Seq/yeast-simulation.Rmd
@@
 ---
-title: "Yeast RNA-seq 48 sample simulation study"
+title: "Yeast RNA-seq 48 sample simulation study - II"
 author: "Keegan Korthauer"
@@
 In this Rmd we will carry out simulations for multiple
-replicates, and plot results averaging over the replications.
+replicates, and plot results averaging over the replications. In contrast to
+`yeast-simulation.Rmd` we will use an alternate simulation framework for generating
+the effect sizes of the DE genes. Here we will use the fold changes of only genes
+that were DE in a comparison of wild type versus knockout (with sample size 10).
+Since this results in a distribution of effect sizes (log2 fold changes) under the
+alternative that does not have a mode at zero, the assumptions for ashq are violated.
+For this reason, we do not include this method in the benchmark comparisons.
 
@@
 # results files that will be generated
-resfile_n5 <- file.path(resdir, "yeast-results-null5.rds")
-resfile_n10 <- file.path(resdir, "yeast-results-null10.rds")
-resfile_d5 <- file.path(resdir, "yeast-results-de5.rds")
-resfile_d10 <- file.path(resdir, "yeast-results-de10.rds")
+resfile_n5 <- file.path(resdir, "yeastII-results-null5.rds")
+resfile_n10 <- file.path(resdir, "yeastII-results-null10.rds")
+resfile_d5 <- file.path(resdir, "yeastII-results-de5.rds")
+resfile_d10 <- file.path(resdir, "yeastII-results-de10.rds")
 
-resfile_n5_uninfCov <- file.path(resdir, "yeast-results-null5-uninfCov.rds")
-resfile_n10_uninfCov <- file.path(resdir, "yeast-results-null10-uninfCov.rds")
-resfile_d5_uninfCov <- file.path(resdir, "yeast-results-de5-uninfCov.rds")
-resfile_d10_uninfCov <- file.path(resdir, "yeast-results-de10-uninfCov.rds")
+resfile_n5_uninfCov <- file.path(resdir, "yeastII-results-null5-uninfCov.rds")
+resfile_n10_uninfCov <- file.path(resdir, "yeastII-results-null10-uninfCov.rds")
+resfile_d5_uninfCov <- file.path(resdir, "yeastII-results-de5-uninfCov.rds")
+resfile_d10_uninfCov <- file.path(resdir, "yeastII-results-de10-uninfCov.rds")
 
@@
 
-# Analysis of full 48v48 (non-null)
+# Empirical analysis of 10 vs 10 (non-null)
 
-Here we'll carry out an analysis of the full dataset, comparing the controls to the
+Here we'll carry out an analysis of a subset of the full dataset, comparing the controls to the
 knockout samples. The fold changes observed in this comparison will be used when 
-generating the non-null simulated data in the following sections.
+generating the non-null simulated data in the following sections. Here we use a subset
+of the samples (10 samples in each group) to mimic the sample sizes we use in 
+simulation.
 
@@
 
-dds <- DESeqDataSetFromMatrix(countData = counts,
+dds_full <- DESeqDataSetFromMatrix(countData = counts,
                               colData = coldat,
@@
 
-Next we run DESeq2.
+Next we run DESeq2, and subset results on genes with FDR < 0.05.
 
@@
 # results on full set
-dds <- DESeq(dds)
+set.seed(9384)
+subset.wt <- dds_full[,colData(dds_full)$condition == "WT"]
+subset.sn <- dds_full[,colData(dds_full)$condition == "Snf2"]
+subset <- cbind(subset.wt[, sample(1:ncol(subset.wt), 10)],
+                subset.sn[, sample(1:ncol(subset.sn), 10)])
+
+dds <- DESeq(subset)
 resultsNames(dds) # lists the coefficients
@@
 sum(resFULL$padj < 0.05, na.rm=T)
+```
+
+Check densities of the test statistic and effect sizes for significant DE genes.
+
+```{r}
+data.frame(resFULL[resFULL$padj < 0.05,]) %>% 
+  ggplot(aes(x = stat)) +
+  geom_density() +
+  xlab("Test Statistic")
+
+data.frame(resFULL[resFULL$padj < 0.05,]) %>% 
+  ggplot(aes(x = log2FoldChange)) +
+  geom_density() +
+  xlab("Effect Size (log2 Fold Change") 
 ```
@@
                              test_statistic=resFULL$stat,
-                             pzero=rowSums(counts(dds)==0)/ncol(counts(dds))))
+                             pzero=rowSums(counts(subset)==0)/ncol(counts(subset))))
 
@@
 
+Subset on only significant DE genes (FDR < 0.05) to draw fold changes from.
+
+```{r}
+resFULL <- resFULL[resFULL$padj < 0.05,]
+```
+
 ## FDR benchmarking
@@
 `nulltype = "empirical"` and `nulltype = "theoretical"`)
-since our test statistics are approximately t-distributed.
+since our test statistics are approximately t-distributed. We also 
+drop the ashq method since its assumptions are violated here.
 
@@
                      control = list(lambda = 0.01))
+bd <- dropBMethod(bd, "ashq")
 ```
@@
   # select a random subset of 20 WT samples
-  dds_test <- dds[,sample(1:ncol(dds), sampleSize*2)]
+  dds_test <- dds_full[,sample(1:ncol(dds_full), sampleSize*2)]
 
@@
     log2FC <- rep(0, nrow(dds_test))
-    log2FC[DE] <- resFULL$log2FoldChange[pzero < 0.5][DE]
+    log2FC[DE] <- resFULL$log2FoldChange[sample(1:nrow(resFULL), nDE, 
+                                                replace = TRUE)]
+    log2FC[DE] <- pmax(1, abs(rnorm(nDE, 2)))
 
@@
     ran <- runif(nrow(dds_test)) 
-    multiplier <- ifelse(ran < 0.5, 2^log2FC, 1/(2^log2FC))
     refcond <- ifelse(ran < 0.5, "A", "B")
@@
         counts(dds_test)[down, colData(dds_test)$fake==unique(refcond[down])] *
-        multiplier[down]
+        2^log2FC
     counts_new[-down,colData(dds_test)$fake==unique(refcond[-down])] <- 
         counts(dds_test)[-down, colData(dds_test)$fake==unique(refcond[-down])] *
-        multiplier[-down]
-
+        2^log2FC
     counts_new <- apply(counts_new, 2, as.integer)
@@
 # only keep one condition for subsetting in simulations that follow 
-dds <- dds[,colData(dds)$condition == "Snf2"]
+dds_full <- dds_full[,colData(dds_full)$condition == "Snf2"]
 ```

### 2018-08-10 datasets/RNA-Seq/yeast-simulation.Rmd
Exclude FDRreg-empirical from yeast sims where alternative has mode at 0

@@
 In this Rmd we will carry out simulations for multiple
-replicates, and plot results averaging over the replications.
+replicates, and plot results averaging over the replications. Here we draw effect
+sizes (log2 fold changes) from the observed fold changes in the full comparison of the 
+two conditions. Since this results in a distribution of effect sizes and test
+statistics under the alternative that has a mode at zero, the assumptions for 
+FDRreg-empirical are violated. For this reason, we do not include this method in
+the benchmark comparisons.
 
@@
 
+Check densities of the test statistic and effect sizes.
+
+```{r}
+data.frame(resFULL) %>% 
+  ggplot(aes(x = stat)) +
+  geom_density() +
+  xlab("Test Statistic")
+
+data.frame(resFULL) %>% 
+  ggplot(aes(x = log2FoldChange)) +
+  geom_density() +
+  xlab("Effect Size (log2 Fold Change") 
+```
+
 ## Check assumptions
@@
 
-Build common bench design. We also add in Scott's FDR Regression (both
-`nulltype = "empirical"` and `nulltype = "theoretical"`)
+Build common bench design. We also add in Scott's FDR Regression (only
+`nulltype = "theoretical"` since assumptions of the empirical method are violated)
 since our test statistics are approximately t-distributed.
@@
                      nulltype = 'theoretical',
-                     control = list(lambda = 0.01))
-bd <- addBMethod(bd, "fdrreg-e",
-                     FDRreg::FDRreg,
-                     function(x) { x$FDR },
-                     z = test_statistic,
-                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
-                     nulltype = 'empirical',
                      control = list(lambda = 0.01))
@@
         counts(dds_test)[down, colData(dds_test)$fake==unique(refcond[down])] *
-        multiplier[down]
+        2^log2FC
     counts_new[-down,colData(dds_test)$fake==unique(refcond[-down])] <- 
         counts(dds_test)[-down, colData(dds_test)$fake==unique(refcond[-down])] *
-        multiplier[-down]
+        2^log2FC
 

### 2018-08-14 datasets/R/common-BenchDesign.R
Return error for lfdr if fewer than 200 tests

@@ -114,6 +114,11 @@ addDefaultMetrics <- function(sb) {
 ## adapted from IHWpaper::clfdr()
 clfdr_hickswrapper <- function(unadj_p, groups, lfdr_estimation="fdrtool") {
   
+  # Exclude this method if there are fewer than 200 tests (fdrtool throws 
+  # a warning in this case)
+  if(length(unadj_p) < 200)
+    stop("Not enough tests to apply this method. Require at least 200.")
+  
   ## estimate local fdr within each stratum first
   lfdr_res <- lfdr_fit(unadj_p, groups, lfdr_estimation=lfdr_estimation)
   lfdrs <- lfdr_res$lfdr

### 2018-08-14 manuscript_figures/manuscript_figures.Rmd
Update manuscript figures to distinguish between yeast sim settings

@@
 objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
-resfile <- objects[grepl("yeast", objects) & 
+resfile <- objects[grepl("yeast-", objects) & 
                    grepl("de5", objects) & 
@@
 ```{r, fig.width = 10, fig.height = 4}
-for (org in c("yeast", "polyester")){
+for (org in c("yeast-", "polyester")){
 
@@
 sim <- unique(c(which(grepl("simulations", objects)),
-               which(grepl("yeast", objects)),
+               which(grepl("yeast-", objects)),
                which(grepl("polyester", objects))))

### 2018-08-14 datasets/simulations/simulations-summary.Rmd

@@
 resmet <- dplyr::mutate(resmet, setting = ifelse(grepl("varyinges", setting), "varyinges", setting))
+
+resmet <- dplyr::mutate(resmet, pi0 = ifelse(grepl("varyingntests", setting), 90, pi0))
+resmet <- dplyr::mutate(resmet, ntests = ifelse(grepl("varyingntests", setting), dist, "n20000"))
+resmet <- dplyr::mutate(resmet, ntests = as.numeric(gsub("n", "", ntests)))
+resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("varyingntests", setting), "gaussian", dist))
 ```
@@
 ```
+
+# Varying Test Count
+
+We take a look at a sequence of simulation settings where only the total number of hypthesis tests
+is varied. We take a look at the false discovery proportion and true positive rate for each method
+at a nomial alpha cutoff of 0.05 as the number of tests is changed.
+
+```{r, fig.width = 10, fig.height = 4}
+res_ntests <- dplyr::filter(resmet, setting == "varyingntests", alpha == 0.05)
+
+ntests_fdr <- dplyr::filter(res_ntests, key == "FDR") %>%
+    ggplot(aes(x = as.factor(ntests), y = mean.info, color = Method, group = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.1, alpha=0.5) +
+    geom_hline(yintercept = 0.05, lty = 2, color = "blue", alpha = 1/2) + 
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    xlab("Number of Tests") +
+    scale_y_continuous("FDR", labels = scales::percent) +
+    ggtitle("Mean FDR Over 100 Replications") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14))
+
+ntests_tpr <- dplyr::filter(res_ntests, key == "TPR") %>%
+    ggplot(aes(x = as.factor(ntests), y = mean.info, color = Method, group = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.1, alpha=0.5) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    xlab("Number of Tests") + 
+    scale_y_continuous("TPR", labels = scales::percent) +
+    ggtitle("Mean TPR Over 100 Replications") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14))
+
+gp <- plot_grid(ntests_fdr + guides(color = FALSE, linetype = FALSE),
+                ntests_tpr + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:2], ncol = 2)
+gp <- plot_grid(gp, get_legend(ntests_tpr), rel_widths = c(1, 1/8))
+gp
+```
+
+We also look at the difference between informative and uninformative covariates.
+
+```{r, fig.width = 10, fig.height = 4}
+ntests_fdr_diff <- dplyr::filter(res_ntests, key == "FDR") %>%
+    ggplot(aes(x = as.factor(ntests), y = mean.diff, color = Method, group = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.1, alpha=0.5) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    xlab("Number of Tests") + 
+    scale_y_continuous("FDR\n(informative - uninformative)", labels = scales::percent) +
+    ggtitle("Mean FDR Difference Over 100 Replications") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14))
+
+ntests_tpr_diff <- dplyr::filter(res_ntests, key == "TPR") %>%
+    ggplot(aes(x = as.factor(ntests), y = mean.diff, color = Method, group = Method)) +
+    geom_line(aes(linetype = Method), alpha = 0.85) +
+    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.1, alpha=0.5) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    xlab("Number of Tests") + 
+    scale_y_continuous("TPR\n(informative - uninformative)", labels = scales::percent) +
+    ggtitle("Mean TPR Difference Over 100 Replications") +
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14))
+
+gp <- plot_grid(ntests_fdr_diff + guides(color = FALSE, linetype = FALSE),
+                ntests_tpr_diff + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[3:4], ncol = 2)
+gp <- plot_grid(gp, get_legend(ntests_tpr_diff), rel_widths = c(1, 1/8))
+gp
+```

### 2018-08-15 datasets/RNA-Seq/yeast-simulation.Rmd
More realistic effect sizes in yeast (drawn from similar sample sizes)

--- 2018-07-25 datasets/RNA-Seq/yeast-simulation.Rmd
@@
 In this Rmd we will carry out simulations for multiple
-replicates, and plot results averaging over the replications.
+replicates, and plot results averaging over the replications. Here we draw effect
+sizes (log2 fold changes) from the observed fold changes in the full comparison of the 
+two conditions. Since this results in a distribution of effect sizes and test
+statistics under the alternative that has a mode at zero, the assumptions for 
+FDRreg-empirical are violated. However, since this assumption is impossible to check
+outside of a simulation setting, we still include this method in
+the benchmark comparisons.
 
@@
 
-# Analysis of full 48v48 (non-null)
-
-Here we'll carry out an analysis of the full dataset, comparing the controls to the
+# Empirical analysis of 10 vs 10 (non-null)
+
+Here we'll carry out an analysis of a subset of the full dataset, comparing the controls to the
 knockout samples. The fold changes observed in this comparison will be used when 
-generating the non-null simulated data in the following sections.
+generating the non-null simulated data in the following sections. Here we use a subset
+of the samples (10 samples in each group) to mimic the sample sizes we use in 
+simulation.
 
@@
 # results on full set
-dds <- DESeq(dds)
+set.seed(9384)
+subset.wt <- dds_full[,colData(dds_full)$condition == "WT"]
+subset.sn <- dds_full[,colData(dds_full)$condition == "Snf2"]
+subset <- cbind(subset.wt[, sample(1:ncol(subset.wt), 10)],
+                subset.sn[, sample(1:ncol(subset.sn), 10)])
+
+dds <- DESeq(subset)
 resultsNames(dds) # lists the coefficients
-resFULL <- results(dds, name="condition_WT_vs_Snf2", independentFiltering = F)
-
-head(resFULL)
-
-sum(resFULL$padj < 0.05, na.rm=T)
+res_10 <- results(dds, name="condition_WT_vs_Snf2", independentFiltering = F)
+
+head(res_10)
+
+sum(res_10$padj < 0.05, na.rm=T)
+```
+
+Check densities of the test statistic and effect sizes for significant DE genes.
+
+```{r}
+data.frame(res_10) %>% 
+  ggplot(aes(x = stat)) +
+  geom_density() +
+  xlab("Test Statistic")
+
+data.frame(res_10) %>% 
+  ggplot(aes(x = log2FoldChange)) +
+  geom_density() +
+  xlab("Effect Size (log2 Fold Change)") 
 ```
@@
 ```{r}
-geneExp <- tbl_df(data.frame(geneName=rownames(resFULL), 
-                             pval=resFULL$pvalue, 
-                             SE=resFULL$lfcSE,
-                             ind_covariate = resFULL$baseMean, 
-                             effect_size=resFULL$log2FoldChange, 
-                             test_statistic=resFULL$stat,
-                             pzero=rowSums(counts(dds)==0)/ncol(counts(dds))))
+geneExp <- tbl_df(data.frame(geneName=rownames(res_10), 
+                             pval=res_10$pvalue, 
+                             SE=res_10$lfcSE,
+                             ind_covariate = res_10$baseMean, 
+                             effect_size=res_10$log2FoldChange, 
+                             test_statistic=res_10$stat,
+                             pzero=rowSums(counts(subset)==0)/ncol(counts(subset))))
 
@@
 ```{r, width=15, height=15}
-strat_hist(geneExp, pvalue="pval", covariate="ind_covariate", maxy=32)
+strat_hist(geneExp, pvalue="pval", covariate="ind_covariate", maxy=30)
 ```
@@
 
-Build common bench design. We also add in Scott's FDR Regression (both
-`nulltype = "empirical"` and `nulltype = "theoretical"`)
+Build common bench design. We also add in Scott's FDR Regression 
 since our test statistics are approximately t-distributed. 
+Note that the assumption of FDRreg-empirical that the distribution of non-null test 
+statistics does not have a significant mass at zero is violated here, but
+we still include fdrreg-e in the evaluation since this assumption is impossible to check
+outside of a simulation setting (since it is not possible to know which tests are 
+non-null). Thus, we'd like to evaluate how results change 
+when the assumption is violated, since it's plausible that it is violated in the 
+case studies.
 
@@
 
+# Empirical analysis of 5 vs 5 (non-null)
+
+Here we'll carry out an analysis of a subset of the full dataset, comparing the controls to the
+knockout samples. The fold changes observed in this comparison will be used when 
+generating the non-null simulated data in the following sections. Here we use a subset
+of the samples (5 samples in each group) to mimic the sample sizes we use in 
+simulation.
+
+## Run DESeq2
+
+First we run DESeq2.
+
+```{r}
+# results on full set
+set.seed(728)
+subset.wt <- dds_full[,colData(dds_full)$condition == "WT"]
+subset.sn <- dds_full[,colData(dds_full)$condition == "Snf2"]
+subset <- cbind(subset.wt[, sample(1:ncol(subset.wt), 5)],
+                subset.sn[, sample(1:ncol(subset.sn), 5)])
+
+dds <- DESeq(subset)
+resultsNames(dds) # lists the coefficients
+res_5 <- results(dds, name="condition_WT_vs_Snf2", independentFiltering = F)
+
+head(res_5)
+
+sum(res_5$padj < 0.05, na.rm=T)
+```
+
+Check densities of the test statistic and effect sizes for significant DE genes.
+
+```{r}
+data.frame(res_5) %>% 
+  ggplot(aes(x = stat)) +
+  geom_density() +
+  xlab("Test Statistic")
+
+data.frame(res_5) %>% 
+  ggplot(aes(x = log2FoldChange)) +
+  geom_density() +
+  xlab("Effect Size (log2 Fold Change)") 
+```
+
+## Check assumptions
+
+There are several thousand genes detected using the full set. Next, we'll build input data frame for summarized benchmark.
+
+```{r}
+geneExp <- tbl_df(data.frame(geneName=rownames(res_5), 
+                             pval=res_5$pvalue, 
+                             SE=res_5$lfcSE,
+                             ind_covariate = res_5$baseMean, 
+                             effect_size=res_5$log2FoldChange, 
+                             test_statistic=res_5$stat,
+                             pzero=rowSums(counts(subset)==0)/ncol(counts(subset))))
+
+# filter NAs and those with less than 50% expressed 
+geneExp <- geneExp %>% na.omit() %>% dplyr::filter(pzero < 0.5)
+```
+
+We'll create a plot to examine the distribution of effect sizes, since
+the ash method assumes that the distribution of true (unobserved) effect
+sizes is unimodal.
+
+```{r}
+ggplot(data=geneExp, aes(effect_size)) +
+  geom_histogram(bins=30)
+```
+
+We'll also explore how the standard error (used by ash) 
+correlates with the independent covariate (used by methods that incorporate 
+covariates), in order to get an idea of how these pieces of information relate
+to one another.
+
+```{r}
+ggplot(data=geneExp, aes(x = ind_covariate, y = SE)) +
+  geom_hex(bins = 100) +
+  scale_x_continuous(trans="log10") +
+  xlab("Covariate: Mean gene expression") 
+```
+
+Look at covariate diagnostic plots.
+
+```{r, width=15, height=15}
+strat_hist(geneExp, pvalue="pval", covariate="ind_covariate", maxy=23)
+```
+
+```{r}
+rank_scatter(geneExp, pvalue="pval", covariate="ind_covariate")
+```
+
+## FDR benchmarking
+
+Run benchmark methods. 
+```{r}
+sb <- bd %>% buildBench(data=geneExp, parallel = FALSE)
+
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Plot results.
+
+```{r}
+rejections_scatter(sb, supplementary=FALSE)
+```
+
+```{r}
+plotFDRMethodsOverlap(sb, alpha=0.05, nsets=ncol(sb), order.by="freq", decreasing=TRUE, supplementary=FALSE)
+
+```
+
 # Simulation set up
@@
 
-  uninf_covar <- rnorm(nrow(dds))
+  uninf_covar <- rnorm(nrow(dds_full))
 
@@
     log2FC <- rep(0, nrow(dds_test))
-    log2FC[DE] <- resFULL$log2FoldChange[pzero < 0.5][DE]
+
+    if(sampleSize == 5){
+      log2FC[DE] <- res_5$log2FoldChange[pzero < 0.5][DE]
+    }else if (sampleSize == 10){
+      log2FC[DE] <- res_10$log2FoldChange[pzero < 0.5][DE]
+    }else{
+      stop("Only sample sizes 5 and 10 are currently supported with pre-",
+           "computed fold changes for sampling from.")
+    }
 
@@
         counts(dds_test)[down, colData(dds_test)$fake==unique(refcond[down])] *
-        multiplier[down]
+        2^log2FC[down]
     counts_new[-down,colData(dds_test)$fake==unique(refcond[-down])] <- 
         counts(dds_test)[-down, colData(dds_test)$fake==unique(refcond[-down])] *
-        multiplier[-down]
+        2^log2FC[-down]
 

### 2018-08-15 datasets/RNA-Seq/yeast-simulation.Rmd
Plot distribution of effect sizes of DE genes

@@
 ```{r}
-data.frame(res_10) %>% 
+data.frame(res_10[res_10$padj < 0.5]) %>% 
   ggplot(aes(x = stat)) +
@@
 
-data.frame(res_10) %>% 
+data.frame(res_10[res_10$padj < 0.5]) %>% 
   ggplot(aes(x = log2FoldChange)) +
@@
 ```{r}
-data.frame(res_5) %>% 
+data.frame(res_5[res_5$padj < 0.5]) %>% 
   ggplot(aes(x = stat)) +
@@
 
-data.frame(res_5) %>% 
+data.frame(res_5[res_5$padj < 0.5]) %>% 
   ggplot(aes(x = log2FoldChange)) +

### 2018-08-15 datasets/RNA-Seq/yeast-simulation-bimodalalternative.Rmd
Sample observed test statistics using bimodal sampling weights

@@
 alternative that does not have a mode at zero, the assumptions for ashq are violated.
-For this reason, we do not include this method in the benchmark comparisons.
+However, since this assumption is impossible to check, we still include this 
+method in the benchmark comparisons.
 
@@
 resultsNames(dds) # lists the coefficients
-resFULL <- results(dds, name="condition_WT_vs_Snf2", independentFiltering = F)
-
-head(resFULL)
-
-sum(resFULL$padj < 0.05, na.rm=T)
+res_10 <- results(dds, name="condition_WT_vs_Snf2", independentFiltering = F)
+
+head(res_10)
+
+sum(res_10$padj < 0.05, na.rm=T)
 ```
@@
 ```{r}
-data.frame(resFULL[resFULL$padj < 0.05,]) %>% 
+data.frame(res_10[res_10$padj < 0.5]) %>% 
   ggplot(aes(x = stat)) +
@@
 
-data.frame(resFULL[resFULL$padj < 0.05,]) %>% 
+data.frame(res_10[res_10$padj < 0.5]) %>% 
   ggplot(aes(x = log2FoldChange)) +
   geom_density() +
-  xlab("Effect Size (log2 Fold Change") 
+  xlab("Effect Size (log2 Fold Change)") 
 ```
@@
 ```{r}
-geneExp <- tbl_df(data.frame(geneName=rownames(resFULL), 
-                             pval=resFULL$pvalue, 
-                             SE=resFULL$lfcSE,
-                             ind_covariate = resFULL$baseMean, 
-                             effect_size=resFULL$log2FoldChange, 
-                             test_statistic=resFULL$stat,
+geneExp <- tbl_df(data.frame(geneName=rownames(res_10), 
+                             pval=res_10$pvalue, 
+                             SE=res_10$lfcSE,
+                             ind_covariate = res_10$baseMean, 
+                             effect_size=res_10$log2FoldChange, 
+                             test_statistic=res_10$stat,
                              pzero=rowSums(counts(subset)==0)/ncol(counts(subset))))
@@
 
-Subset on only significant DE genes (FDR < 0.05) to draw fold changes from.
-
-```{r}
-resFULL <- resFULL[resFULL$padj < 0.05,]
-```
-
 ## FDR benchmarking
 
-Build common bench design. We also add in Scott's FDR Regression (both
-`nulltype = "empirical"` and `nulltype = "theoretical"`)
-since our test statistics are approximately t-distributed. We also 
-drop the ashq method since its assumptions are violated here.
+Build common bench design. We also add in Scott's FDR Regression 
+since our test statistics are approximately t-distributed. 
+Note that the assumption of FDRreg-empirical that the distribution of non-null test 
+statistics does not have a significant mass at zero is violated here, but
+we still include fdrreg-e in the evaluation since this assumption is impossible to check
+outside of a simulation setting (since it is not possible to know which tests are 
+non-null). Thus, we'd like to evaluate how results change 
+when the assumption is violated, since it's plausible that it is violated in the 
+case studies.
 
@@
                      control = list(lambda = 0.01))
-bd <- dropBMethod(bd, "ashq")
 ```
@@
 
+# Empirical analysis of 5 vs 5 (non-null)
+
+Here we'll carry out an analysis of a subset of the full dataset, comparing the controls to the
+knockout samples. The fold changes observed in this comparison will be used when 
+generating the non-null simulated data in the following sections. Here we use a subset
+of the samples (5 samples in each group) to mimic the sample sizes we use in 
+simulation.
+
+## Run DESeq2
+
+First we run DESeq2, and subset results on genes with FDR < 0.05.
+
+```{r}
+# results on full set
+set.seed(728)
+subset.wt <- dds_full[,colData(dds_full)$condition == "WT"]
+subset.sn <- dds_full[,colData(dds_full)$condition == "Snf2"]
+subset <- cbind(subset.wt[, sample(1:ncol(subset.wt), 5)],
+                subset.sn[, sample(1:ncol(subset.sn), 5)])
+
+dds <- DESeq(subset)
+resultsNames(dds) # lists the coefficients
+res_5 <- results(dds, name="condition_WT_vs_Snf2", independentFiltering = F)
+
+head(res_5)
+
+sum(res_5$padj < 0.05, na.rm=T)
+```
+
+Check densities of the test statistic and effect sizes for significant DE genes.
+
+```{r}
+data.frame(res_5[res_5$padj < 0.5]) %>% 
+  ggplot(aes(x = stat)) +
+  geom_density() +
+  xlab("Test Statistic")
+
+data.frame(res_5[res_5$padj < 0.5]) %>% 
+  ggplot(aes(x = log2FoldChange)) +
+  geom_density() +
+  xlab("Effect Size (log2 Fold Change)") 
+```
+
+## Check assumptions
+
+There are several thousand genes detected using the full set. Next, we'll build input data frame for summarized benchmark.
+
+```{r}
+geneExp <- tbl_df(data.frame(geneName=rownames(res_5), 
+                             pval=res_5$pvalue, 
+                             SE=res_5$lfcSE,
+                             ind_covariate = res_5$baseMean, 
+                             effect_size=res_5$log2FoldChange, 
+                             test_statistic=res_5$stat,
+                             pzero=rowSums(counts(subset)==0)/ncol(counts(subset))))
+
+# filter NAs and those with less than 50% expressed 
+geneExp <- geneExp %>% na.omit() %>% dplyr::filter(pzero < 0.5)
+```
+
+We'll create a plot to examine the distribution of effect sizes, since
+the ash method assumes that the distribution of true (unobserved) effect
+sizes is unimodal.
+
+```{r}
+ggplot(data=geneExp, aes(effect_size)) +
+  geom_histogram(bins=30)
+```
+
+We'll also explore how the standard error (used by ash) 
+correlates with the independent covariate (used by methods that incorporate 
+covariates), in order to get an idea of how these pieces of information relate
+to one another.
+
+```{r}
+ggplot(data=geneExp, aes(x = ind_covariate, y = SE)) +
+  geom_hex(bins = 100) +
+  scale_x_continuous(trans="log10") +
+  xlab("Covariate: Mean gene expression") 
+```
+
+Look at covariate diagnostic plots.
+
+```{r, width=15, height=15}
+strat_hist(geneExp, pvalue="pval", covariate="ind_covariate", maxy=23)
+```
+
+```{r}
+rank_scatter(geneExp, pvalue="pval", covariate="ind_covariate")
+```
+
+## FDR benchmarking
+
+Run benchmark methods. 
+```{r}
+sb <- bd %>% buildBench(data=geneExp, parallel = FALSE)
+
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Plot results.
+
+```{r}
+rejections_scatter(sb, supplementary=FALSE)
+```
+
+```{r}
+plotFDRMethodsOverlap(sb, alpha=0.05, nsets=ncol(sb), order.by="freq", decreasing=TRUE, supplementary=FALSE)
+
+```
+
 # Simulation set up
+
+First, here's a function that will generate sampling weights for test statistics
+of DE genes that downweights observations near zero, as well as extremely large values.
+
+```{r}
+stat_wt <- function(z){
+  pos <- pmax(0,dgamma(z,4.5, 1-1e-4))
+  neg <- pmax(0,dgamma(-z,4.5, 1-1e-4))
+
+  res <- pos
+  res[z < 0] <- neg[z < 0]
+  return(res)
+}
+
+# Plot weights by test statistic value
+df = data.frame(z = seq(-10,10, by = 0.1))
+df$w = stat_wt(df$z)
+ggplot(data=df, aes(x=z, y=w)) +
+  geom_line() +
+  ylab("Sampling Weight") +
+  xlab("Test statistic value")
+```
 
@@
 
-    # randomly sample a log2FC from original FCs (without regard to DE)
+    # randomly sample a test statistic from original test statistics, with probability
+    # weights that downweight observations near zero 
     counts_new <- counts(dds_test)
-    log2FC <- rep(0, nrow(dds_test))
-    log2FC[DE] <- resFULL$log2FoldChange[sample(1:nrow(resFULL), nDE, 
-                                                replace = TRUE)]
-    log2FC[DE] <- pmax(1, abs(rnorm(nDE, 2)))
+    stat <- rep(0, nrow(dds_test))
+
+    # assume fixed standard error (median observed) and calculate FC from stat
+    if(sampleSize == 5){
+      stat[DE] <- sample(res_5$stat, nDE, prob = stat_wt(res_5$stat))
+      log2FC <- stat * median(res_5$lfcSE)
+    }else if (sampleSize == 10){
+      stat[DE] <- sample(res_10$stat, nDE, prob = stat_wt(res_10$stat))
+      log2FC <- stat * median(res_10$lfcSE)
+    }else{
+      stop("Only sample sizes 5 and 10 are currently supported with pre-",
+           "computed fold changes for sampling from.")
+    }
 
@@
         counts(dds_test)[down, colData(dds_test)$fake==unique(refcond[down])] *
-        2^log2FC
+        2^log2FC[down]
     counts_new[-down,colData(dds_test)$fake==unique(refcond[-down])] <- 
         counts(dds_test)[-down, colData(dds_test)$fake==unique(refcond[-down])] *
-        2^log2FC
+        2^log2FC[-down]
     counts_new <- apply(counts_new, 2, as.integer)

### 2018-08-15 datasets/RNA-Seq/yeast-simulation-bimodalalternative.Rmd
Add usetting for unweighted by covariate

@@
 ---
-title: "Yeast RNA-seq 48 sample simulation study - II"
+title: "Yeast RNA-seq 48 sample simulation study: Bimodal alternative, 2000 DE genes"
 author: "Keegan Korthauer"
@@
 method in the benchmark comparisons.
+
+In this setting, we assume 500 DE genes and a bimodal alternative.
 
@@
 resfile_d10 <- file.path(resdir, "yeastII-results-de10.rds")
+resfile_d5_uw <- file.path(resdir, "yeastIIU-results-de5.rds")
+resfile_d10_uw <- file.path(resdir, "yeastIIU-results-de10.rds")
 
@@
 resfile_d10_uninfCov <- file.path(resdir, "yeastII-results-de10-uninfCov.rds")
+resfile_d5_uninfCov_uw <- file.path(resdir, "yeastIIU-results-de5-uninfCov.rds")
+resfile_d10_uninfCov_uw <- file.path(resdir, "yeastIIU-results-de10-uninfCov.rds")
 
@@
 ```{r}
-data.frame(res_10[res_10$padj < 0.5]) %>% 
+data.frame(res_10[res_10$padj < 0.05 & !is.na(res_10$padj),]) %>% 
   ggplot(aes(x = stat)) +
@@
 
-data.frame(res_10[res_10$padj < 0.5]) %>% 
+data.frame(res_10[res_10$padj < 0.05 & !is.na(res_10$padj),]) %>% 
   ggplot(aes(x = log2FoldChange)) +
@@
 ```{r}
-data.frame(res_5[res_5$padj < 0.5]) %>% 
+data.frame(res_5[res_5$padj < 0.05 & !is.na(res_5$padj),]) %>% 
   ggplot(aes(x = stat)) +
@@
 
-data.frame(res_5[res_5$padj < 0.5]) %>% 
+data.frame(res_5[res_5$padj < 0.05 & !is.na(res_5$padj),]) %>% 
   ggplot(aes(x = log2FoldChange)) +
@@
 simulateOneSplit <- function(X, rseed, nDE, sampleSize,
-                             bd, uninformativeCovariate = FALSE, pvalHists = FALSE){
+                             bd, uninformativeCovariate = FALSE, pvalHists = FALSE,
+                             weightByCovariate = TRUE){
 
@@
     # pick random set of nDE genes to add signal to
+    if(weightByCovariate){
       wts <- rowMeans(counts(dds_test, normalize = TRUE))
+    }else{
+      wts <- rep(1, nrow(dds_test))
+    }
     DE <- sample(1:nrow(dds_test), nDE, prob = wts)
@@
 
-# Sample size comparison
-
-Here we compare the method ranks for the different sample sizes at alpha = 0.10.
-
-```{r}
-plotMethodRanks(c(resfile_d5, resfile_d10), 
-                colLabels = c("DE 5", "DE 10"), 
-                alpha = 0.10, xlab = "Comparison", 
-                excludeMethods = NULL)
+# Sim D5: Unweighted 5v5
+
+Here we'll repeat the above, but not explicitly weighting the DE gene selection by
+the informative covariate.
+
+## Generate a list of SB results objects
+
+```{r d5-uw, results='hide', message=FALSE}
+rseed <- 445
+sampleSize <- 5
+nDE <- 2000
+
+if (!file.exists(resfile_d5_uw)){
+  de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+                  nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores,
+                  weightByCovariate = FALSE)
+  saveRDS(de5, file = resfile_d5_uw)
+}else{
+  de5 <- readRDS(file = resfile_d5_uw)
+}
+```
+
+## Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(res5d, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5d, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5d, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(de5, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(de5, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
+  aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE)
+
+}else{
+  message("Not enough methods found rejections at alpha ", ualpha, 
+          "; skipping upset plot")
+}
+
+```
+
+# Sim D10: Unweighted 10v10
+
+Here we'll repeat the above, but not explicitly weighting the DE gene selection by
+the informative covariate.
+
+## Generate a list of SB results objects
+
+```{r d10-uw, results='hide', message=FALSE}
+rseed <- 868
+sampleSize <- 10
+nDE <- 2000
+
+if(!file.exists(resfile_d10_uw)){
+  de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+                  nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores,
+                  weightByCovariate = FALSE)
+  saveRDS(de10, file = resfile_d10_uw)
+}else{
+  de10 <- readRDS(file = resfile_d10_uw)
+}
+```
+
+## Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(res10d, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10d, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10d, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(de10, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(de10, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res10d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
+  aggupset(de10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+}else{
+  message("Not enough methods found rejections at alpha ", ualpha, 
+          "; skipping upset plot")
+}
+
 ```
@@
 
-## Sample size comparison
-
-Here we compare the method ranks for the different sample sizes at alpha = 0.10.
-
-```{r}
-# uninformative covariate
-plotMethodRanks(c(resfile_d5_uninfCov, resfile_d10_uninfCov), 
-                colLabels = c("DE 5", "DE 10"), 
+## Sim D5: Unweighted 5v5
+
+Here we'll repeat the above, but not explicitly weighting the DE gene selection by
+the informative covariate.
+This will be done for 100 random splits.
+
+### Generate a list of SB results objects
+
+```{r d5-uc-uw, results='hide', message=FALSE}
+rseed <- 445
+sampleSize <- 5
+nDE <- 2000
+
+if (!file.exists(resfile_d5_uninfCov_uw)){
+  de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+                  nDE=nDE, sampleSize=sampleSize, bd=bd,
+                  uninformativeCovariate = TRUE, mc.cores=nCores)
+  saveRDS(de5, file = resfile_d5_uninfCov_uw)
+}else{
+  de5 <- readRDS(file = resfile_d5_uninfCov_uw)
+}
+```
+
+### Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(res5d, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5d, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5d, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(de5, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(de5, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
+  aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE)
+
+}else{
+  message("Not enough methods found rejections at alpha ", ualpha, 
+          "; skipping upset plot")
+}
+
+```
+
+## Sim D10: Unweighted 10v10
+
+Here we'll repeat the above, but not explicitly weighting the DE gene selection by
+the informative covariate.
+
+### Generate a list of SB results objects
+
+```{r d10-uc-uw, results='hide', message=FALSE}
+rseed <- 868
+sampleSize <- 10
+nDE <- 2000
+
+if(!file.exists(resfile_d10_uninfCov_uw)){
+  de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
+                   nDE=nDE, sampleSize=sampleSize, bd=bd,
+                   uninformativeCovariate = TRUE, mc.cores=nCores)
+  saveRDS(de10, file = resfile_d10_uninfCov_uw)
+}else{
+  de10 <- readRDS(file = resfile_d10_uninfCov_uw)
+}
+```
+
+### Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(res10d, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10d, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10d, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(de10, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(de10, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res10d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
+  aggupset(de10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+}else{
+  message("Not enough methods found rejections at alpha ", ualpha, 
+          "; skipping upset plot")
+}
+
+```
+
+# Simulation setting comparison
+
+Here we compare the method ranks for the different sample sizes, weighting, and
+covariate settings at alpha = 0.10.
+
+```{r, fig.width=18, fig.height=12}
+plotMethodRanks(c(resfile_d5, resfile_d5_uw, resfile_d10, resfile_d10_uw,
+                  resfile_d5_uninfCov, resfile_d5_uninfCov_uw, resfile_d10_uninfCov, resfile_d10_uninfCov_uw), 
+                colLabels = c("5", "5 UW", "10", "10 UW",
+                              "5 UI", "5 UI UW", "10 UI", "10 UI UW"), 
                 alpha = 0.10, xlab = "Comparison", 

### 2018-08-15 datasets/RNA-Seq/yeast-simulation.Rmd
Add usetting for unweighted by covariate

--- 2018-08-15 datasets/RNA-Seq/yeast-simulation.Rmd
@@
 ---
-title: "Yeast RNA-seq 48 sample simulation study"
+title: "Yeast RNA-seq 48 sample simulation study: Unimodal alternative, 2000 DE genes"
 author: "Keegan Korthauer"
@@
 the benchmark comparisons.
+
+In this setting, we assume 2000 DE genes and a unimodal alternative.
 
@@
 resfile_d10 <- file.path(resdir, "yeast-results-de10.rds")
+resfile_d5_uw <- file.path(resdir, "yeastU-results-de5.rds")
+resfile_d10_uw <- file.path(resdir, "yeastU-results-de10.rds")
 
@@
 resfile_d10_uninfCov <- file.path(resdir, "yeast-results-de10-uninfCov.rds")
+resfile_d5_uninfCov_uw <- file.path(resdir, "yeastU-results-de5-uninfCov.rds")
+resfile_d10_uninfCov_uw <- file.path(resdir, "yeastU-results-de10-uninfCov.rds")
 
@@
 ```{r}
-data.frame(res_10) %>% 
+data.frame(res_10[res_10$padj < 0.05 & !is.na(res_10$padj),]) %>% 
   ggplot(aes(x = stat)) +
@@
 
-data.frame(res_10) %>% 
+data.frame(res_10[res_10$padj < 0.05 & !is.na(res_10$padj),]) %>% 
   ggplot(aes(x = log2FoldChange)) +
@@
 ```{r}
-data.frame(res_5) %>% 
+data.frame(res_5[res_5$padj < 0.05 & !is.na(res_5$padj),]) %>% 
   ggplot(aes(x = stat)) +
@@
 
-data.frame(res_5) %>% 
+data.frame(res_5[res_5$padj < 0.05 & !is.na(res_5$padj),]) %>% 
   ggplot(aes(x = log2FoldChange)) +

### 2018-08-15 datasets/RNA-Seq/yeast-simulation-bimodalalternative-highpi0.Rmd
Add setting for high pi0

--- 2018-08-15 datasets/RNA-Seq/yeast-simulation-bimodalalternative.Rmd
@@
 ---
-title: "Yeast RNA-seq 48 sample simulation study: Bimodal alternative, 2000 DE genes"
+title: "Yeast RNA-seq 48 sample simulation study: bimodal alternative, 500 DE genes"
 author: "Keegan Korthauer"
@@
 Here we'll repeat the above, but for a DE comparison 5 versus 5 Snf2 samples,
-where the groups are selected randomly and 2000 DE genes are added. 
+where the groups are selected randomly and 500 DE genes are added. 
 This will be done for 100 random splits.
@@
 sampleSize <- 5
-nDE <- 2000
+nDE <- 500
 
@@
 Here we'll repeat the above, but for a DE comparison 10 versus 10 Snf2 samples,
-where the groups are selected randomly and 2000 DE genes are added.
+where the groups are selected randomly and 500 DE genes are added.
 This will be done for 100 random splits.
@@
 sampleSize <- 10
-nDE <- 2000
+nDE <- 500
 

### 2018-08-15 datasets/RNA-Seq/yeast-simulation-highpi0.Rmd
Add setting for high pi0

--- 2018-08-15 datasets/RNA-Seq/yeast-simulation.Rmd
@@
 ---
-title: "Yeast RNA-seq 48 sample simulation study: Unimodal alternative, 2000 DE genes"
+title: "Yeast RNA-seq 48 sample simulation study: Unimodal alternative, 500 DE genes"
 author: "Keegan Korthauer"
@@
 
-In this setting, we assume 2000 DE genes and a unimodal alternative.
+In this setting, we assume 500 DE genes and a unimodal alternative.
 
@@
 
-# results files that will be generated
+# files generated by yeast-simulation.Rmd
 resfile_n5 <- file.path(resdir, "yeast-results-null5.rds")
 resfile_n10 <- file.path(resdir, "yeast-results-null10.rds")
+resfile_n5_uninfCov <- file.path(resdir, "yeast-results-null5-uninfCov.rds")
+resfile_n10_uninfCov <- file.path(resdir, "yeast-results-null10-uninfCov.rds")
+
+# results files that will be generated
 resfile_d5 <- file.path(resdir, "yeast-results-de5.rds")
@@
 
-resfile_n5_uninfCov <- file.path(resdir, "yeast-results-null5-uninfCov.rds")
-resfile_n10_uninfCov <- file.path(resdir, "yeast-results-null10-uninfCov.rds")
 resfile_d5_uninfCov <- file.path(resdir, "yeast-results-de5-uninfCov.rds")
@@
 sampleSize <- 5
-nDE <- 2000
-
-if (!file.exists(resfile_d5_uninfCov_uw)){
+nDE <- 500
+
+if(!file.exists(resfile_d5_uninfCov_uw)){
   de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
@@
 
-if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
+if(numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
   aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE)

### 2018-08-15 datasets/RNA-Seq/yeast-simulation-bimodalalternative-highpi0.Rmd
Don't rerun null sims with new DE sample size

@@
 
-# results files that will be generated
+# results files already generated by yeast-simulation-bimodalalternative.Rmd
 resfile_n5 <- file.path(resdir, "yeastII-results-null5.rds")
 resfile_n10 <- file.path(resdir, "yeastII-results-null10.rds")
-resfile_d5 <- file.path(resdir, "yeastII-results-de5.rds")
-resfile_d10 <- file.path(resdir, "yeastII-results-de10.rds")
-resfile_d5_uw <- file.path(resdir, "yeastIIU-results-de5.rds")
-resfile_d10_uw <- file.path(resdir, "yeastIIU-results-de10.rds")
-
 resfile_n5_uninfCov <- file.path(resdir, "yeastII-results-null5-uninfCov.rds")
 resfile_n10_uninfCov <- file.path(resdir, "yeastII-results-null10-uninfCov.rds")
-resfile_d5_uninfCov <- file.path(resdir, "yeastII-results-de5-uninfCov.rds")
-resfile_d10_uninfCov <- file.path(resdir, "yeastII-results-de10-uninfCov.rds")
-resfile_d5_uninfCov_uw <- file.path(resdir, "yeastIIU-results-de5-uninfCov.rds")
-resfile_d10_uninfCov_uw <- file.path(resdir, "yeastIIU-results-de10-uninfCov.rds")
+
+# results files that will be generated
+resfile_d5 <- file.path(resdir, "yeastIIH-results-de5.rds")
+resfile_d10 <- file.path(resdir, "yeastIIH-results-de10.rds")
+resfile_d5_uw <- file.path(resdir, "yeastIIHU-results-de5.rds")
+resfile_d10_uw <- file.path(resdir, "yeastIIHU-results-de10.rds")
+
+resfile_d5_uninfCov <- file.path(resdir, "yeastIIH-results-de5-uninfCov.rds")
+resfile_d10_uninfCov <- file.path(resdir, "yeastIIH-results-de10-uninfCov.rds")
+resfile_d5_uninfCov_uw <- file.path(resdir, "yeastIIHU-results-de5-uninfCov.rds")
+resfile_d10_uninfCov_uw <- file.path(resdir, "yeastIIHU-results-de10-uninfCov.rds")
 
@@
 
-# Sim N5: Null Comparison 5v5
-
-Here we'll repeat the above, but for a null comparison 5 versus 5 Snf2 samples,
-where the groups are selected randomly. This will be done for 100 random splits.
-
-## Generate a list of SB results objects
-
-```{r n5, results='hide', message=FALSE}
-rseed <- 225
-sampleSize <- 5
-nDE <- 0
-
-if (!file.exists(resfile_n5)){
-  null5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
-                    nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
-  saveRDS(null5, file = resfile_n5)
-}else{
-  null5 <- readRDS(file = resfile_n5)
-}
-```
-
-## Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(null5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res5 <- plotsim_standardize(null5, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(res5, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res5, met="TNR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(null5, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(null5, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res5, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(null5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
-}else{
-  message("Not enough methods found rejections at alpha ", ualpha, 
-          "; skipping upset plot")
-}
-
-```
-
-# Sim N10: Null Comparison 10v10
-
-Here we'll repeat the above, but for a null comparison 10 versus 10 Snf2 samples,
-where the groups are selected randomly. This will be done for 100 random splits.
-
-## Generate a list of SB results objects
-
-```{r n10, results='hide', message=FALSE}
-rseed <- 837
-sampleSize <- 10
-nDE <- 0
-
-if (!file.exists(resfile_n10)){
-  null10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
-                    nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
-  saveRDS(null10, file = resfile_n10)
-}else{
-  null10 <- readRDS(file = resfile_n10)
-}
-
-```
-
-## Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(null10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res10 <- plotsim_standardize(null10, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(res10, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res10, met="TNR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(null10, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(null10, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res10, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(null10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
-}else{
-  message("Not enough methods found rejections at alpha ", ualpha, 
-          "; skipping upset plot")
-}
-
-```
-
 # Sim D5: DE Comparison 5v5
@@
 
-## Sim N5: Null Comparison 5v5
-
-Here we'll repeat the above, but for a null comparison 5 versus 5 Snf2 samples,
-where the groups are selected randomly. This will be done for 100 random splits.
-
-### Generate a list of SB results objects
-
-```{r n5-uc, results='hide', message=FALSE}
-rseed <- 225
-sampleSize <- 5
-nDE <- 0
-
-if (!file.exists(resfile_n5_uninfCov)){
-  null5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed,
-                   nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   uninformativeCovariate = TRUE, mc.cores = nCores)
-  saveRDS(null5, file = resfile_n5_uninfCov)
-}else{
-  null5 <- readRDS(file = resfile_n5_uninfCov)
-}
-```
-
-### Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(null5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res5 <- plotsim_standardize(null5, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(res5, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res5, met="TNR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(null5, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(null5, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res5, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(null5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
-}else{
-  message("Not enough methods found rejections at alpha ", ualpha, 
-          "; skipping upset plot")
-}
-
-```
-
-## Sim N10: Null Comparison 10v10
-
-Here we'll repeat the above, but for a null comparison 10 versus 10 Snf2 samples,
-where the groups are selected randomly. This will be done for 100 random splits.
-
-### Generate a list of SB results objects
-
-```{r n10-uc, results='hide', message=FALSE}
-rseed <- 837
-sampleSize <- 10
-nDE <- 0
-
-if (!file.exists(resfile_n10_uninfCov)){
-  null10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
-                   nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   uninformativeCovariate = TRUE, mc.cores=nCores)
-  saveRDS(null10, file = resfile_n10_uninfCov)
-}else{
-  null10 <- readRDS(file = resfile_n10_uninfCov)
-}
-```
-
-### Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(null10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res10 <- plotsim_standardize(null10, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(res10, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res10, met="TNR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(null10, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(null10, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res10, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(null10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
-}else{
-  message("Not enough methods found rejections at alpha ", ualpha, 
-          "; skipping upset plot")
-}
-
-```
-
 ## Sim D5: DE Comparison 5v5

### 2018-08-15 datasets/RNA-Seq/yeast-simulation-highpi0.Rmd
Don't rerun null sims with new DE sample size

@@
 # results files that will be generated
-resfile_d5 <- file.path(resdir, "yeast-results-de5.rds")
-resfile_d10 <- file.path(resdir, "yeast-results-de10.rds")
-resfile_d5_uw <- file.path(resdir, "yeastU-results-de5.rds")
-resfile_d10_uw <- file.path(resdir, "yeastU-results-de10.rds")
-
-resfile_d5_uninfCov <- file.path(resdir, "yeast-results-de5-uninfCov.rds")
-resfile_d10_uninfCov <- file.path(resdir, "yeast-results-de10-uninfCov.rds")
-resfile_d5_uninfCov_uw <- file.path(resdir, "yeastU-results-de5-uninfCov.rds")
-resfile_d10_uninfCov_uw <- file.path(resdir, "yeastU-results-de10-uninfCov.rds")
+resfile_d5 <- file.path(resdir, "yeastH-results-de5.rds")
+resfile_d10 <- file.path(resdir, "yeastH-results-de10.rds")
+resfile_d5_uw <- file.path(resdir, "yeastHU-results-de5.rds")
+resfile_d10_uw <- file.path(resdir, "yeastHU-results-de10.rds")
+
+resfile_d5_uninfCov <- file.path(resdir, "yeastH-results-de5-uninfCov.rds")
+resfile_d10_uninfCov <- file.path(resdir, "yeastH-results-de10-uninfCov.rds")
+resfile_d5_uninfCov_uw <- file.path(resdir, "yeastHU-results-de5-uninfCov.rds")
+resfile_d10_uninfCov_uw <- file.path(resdir, "yeastHU-results-de10-uninfCov.rds")
 

### 2018-08-15 datasets/RNA-Seq/yeast-simulation-bimodalalternative-highpi0.Rmd
Don't repeat null settings for varying alternative settings

@@
 dir.create(resdir, showWarnings = FALSE)
-
-# results files already generated by yeast-simulation-bimodalalternative.Rmd
-resfile_n5 <- file.path(resdir, "yeastII-results-null5.rds")
-resfile_n10 <- file.path(resdir, "yeastII-results-null10.rds")
-resfile_n5_uninfCov <- file.path(resdir, "yeastII-results-null5-uninfCov.rds")
-resfile_n10_uninfCov <- file.path(resdir, "yeastII-results-null10-uninfCov.rds")
 

### 2018-08-15 datasets/RNA-Seq/yeast-simulation-bimodalalternative.Rmd
Don't repeat null settings for varying alternative settings

@@
 # results files that will be generated
-resfile_n5 <- file.path(resdir, "yeastII-results-null5.rds")
-resfile_n10 <- file.path(resdir, "yeastII-results-null10.rds")
 resfile_d5 <- file.path(resdir, "yeastII-results-de5.rds")
@@
 
-resfile_n5_uninfCov <- file.path(resdir, "yeastII-results-null5-uninfCov.rds")
-resfile_n10_uninfCov <- file.path(resdir, "yeastII-results-null10-uninfCov.rds")
 resfile_d5_uninfCov <- file.path(resdir, "yeastII-results-de5-uninfCov.rds")
@@
 
-# Sim N5: Null Comparison 5v5
-
-Here we'll repeat the above, but for a null comparison 5 versus 5 Snf2 samples,
-where the groups are selected randomly. This will be done for 100 random splits.
-
-## Generate a list of SB results objects
-
-```{r n5, results='hide', message=FALSE}
-rseed <- 225
-sampleSize <- 5
-nDE <- 0
-
-if (!file.exists(resfile_n5)){
-  null5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
-                    nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
-  saveRDS(null5, file = resfile_n5)
-}else{
-  null5 <- readRDS(file = resfile_n5)
-}
-```
-
-## Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(null5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res5 <- plotsim_standardize(null5, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(res5, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res5, met="TNR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(null5, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(null5, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res5, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(null5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
-}else{
-  message("Not enough methods found rejections at alpha ", ualpha, 
-          "; skipping upset plot")
-}
-
-```
-
-# Sim N10: Null Comparison 10v10
-
-Here we'll repeat the above, but for a null comparison 10 versus 10 Snf2 samples,
-where the groups are selected randomly. This will be done for 100 random splits.
-
-## Generate a list of SB results objects
-
-```{r n10, results='hide', message=FALSE}
-rseed <- 837
-sampleSize <- 10
-nDE <- 0
-
-if (!file.exists(resfile_n10)){
-  null10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
-                    nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
-  saveRDS(null10, file = resfile_n10)
-}else{
-  null10 <- readRDS(file = resfile_n10)
-}
-
-```
-
-## Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(null10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res10 <- plotsim_standardize(null10, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(res10, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res10, met="TNR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(null10, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(null10, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res10, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(null10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
-}else{
-  message("Not enough methods found rejections at alpha ", ualpha, 
-          "; skipping upset plot")
-}
-
-```
-
 # Sim D5: DE Comparison 5v5
 
-Here we'll repeat the above, but for a DE comparison 5 versus 5 Snf2 samples,
+Here we'll carry out a DE comparison 5 versus 5 Snf2 samples,
 where the groups are selected randomly and 2000 DE genes are added. 
-This will be done for 100 random splits.
+This will be done for 100 random splits. Note that we do not need to run null
+comparisons, since these are done in `yeast-simulation.Rmd`, and the only
+difference here is that we use a different under the alternative.
 
@@
 
-## Sim N5: Null Comparison 5v5
-
-Here we'll repeat the above, but for a null comparison 5 versus 5 Snf2 samples,
-where the groups are selected randomly. This will be done for 100 random splits.
-
-### Generate a list of SB results objects
-
-```{r n5-uc, results='hide', message=FALSE}
-rseed <- 225
-sampleSize <- 5
-nDE <- 0
-
-if (!file.exists(resfile_n5_uninfCov)){
-  null5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed,
-                   nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   uninformativeCovariate = TRUE, mc.cores = nCores)
-  saveRDS(null5, file = resfile_n5_uninfCov)
-}else{
-  null5 <- readRDS(file = resfile_n5_uninfCov)
-}
-```
-
-### Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(null5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res5 <- plotsim_standardize(null5, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(res5, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res5, met="TNR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(null5, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(null5, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res5, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(null5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
-}else{
-  message("Not enough methods found rejections at alpha ", ualpha, 
-          "; skipping upset plot")
-}
-
-```
-
-## Sim N10: Null Comparison 10v10
-
-Here we'll repeat the above, but for a null comparison 10 versus 10 Snf2 samples,
-where the groups are selected randomly. This will be done for 100 random splits.
-
-### Generate a list of SB results objects
-
-```{r n10-uc, results='hide', message=FALSE}
-rseed <- 837
-sampleSize <- 10
-nDE <- 0
-
-if (!file.exists(resfile_n10_uninfCov)){
-  null10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
-                   nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   uninformativeCovariate = TRUE, mc.cores=nCores)
-  saveRDS(null10, file = resfile_n10_uninfCov)
-}else{
-  null10 <- readRDS(file = resfile_n10_uninfCov)
-}
-```
-
-### Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(null10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res10 <- plotsim_standardize(null10, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(res10, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res10, met="TNR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(null10, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(null10, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res10, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(null10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
-}else{
-  message("Not enough methods found rejections at alpha ", ualpha, 
-          "; skipping upset plot")
-}
-
-```
-
 ## Sim D5: DE Comparison 5v5
 
-Here we'll repeat the above, but for a DE comparison 5 versus 5 Snf2 samples,
+Here we'll carry out a DE comparison 5 versus 5 Snf2 samples,
 where the groups are selected randomly and 2000 DE genes are added. 
-This will be done for 100 random splits.
+This will be done for 100 random splits. Note that we do not need to run null
+comparisons, since these are done in `yeast-simulation.Rmd`, and the only
+difference here is that we use a different under the alternative.
 

### 2018-08-15 datasets/RNA-Seq/yeast-simulation-highpi0.Rmd
Don't repeat null settings for varying alternative settings

@@
 dir.create(resdir, showWarnings = FALSE)
-
-# files generated by yeast-simulation.Rmd
-resfile_n5 <- file.path(resdir, "yeast-results-null5.rds")
-resfile_n10 <- file.path(resdir, "yeast-results-null10.rds")
-resfile_n5_uninfCov <- file.path(resdir, "yeast-results-null5-uninfCov.rds")
-resfile_n10_uninfCov <- file.path(resdir, "yeast-results-null10-uninfCov.rds")
 

### 2018-08-16 manuscript_figures/manuscript_figures.Rmd
Update filename parsing of yeast sb objs after expansion of simulations

@@
 wcs <- which(grepl("RNAseq", casestudy) & 
-            !grepl(c("brain|mir|null|polyester"), dataset) &
+            grepl(c("yeast-"), dataset) &
+            grepl("de", dataset) &
             !grepl("scRNAseq", casestudy) &

### 2018-08-16 datasets/R/rejections_scatter_bins.R
Don't plot rejections as zero for method with all NAs

@@ -27,6 +27,10 @@ rejection_scatter_bins <- function( sb, covariate, threshold=NULL, bins= 4, ncol
   if( !is.null(threshold) ){
     sb <- sb[rowData(sb)[[covariate]] > threshold,]
   }
+  if(sum(colSums(is.na(assays(sb)$qvalue)) == nrow(sb)) > 0){
+    miss <- which(colSums(is.na(assays(sb)$qvalue)) == nrow(sb)) 
+    sb <- sb[,-miss]
+  }
   rowData(sb)$bin <- cut(rank( rowData(sb)[[covariate]], ties="first"),
                          quantile( rank(rowData(sb)[[covariate]], ties="first"), seq(0, 1, length.out=bins+1) ),
                          include.lowest=TRUE)

### 2018-08-16 manuscript_figures/manuscript_figures.Rmd
Make max FDR label visible in Fig 4

@@
                 rowOrder = Fig4All.rowOrder, 
-                annotate = "FDR", excludeMethods = NULL) +
+                annotate = "FDR", excludeMethods = NULL,
+                maxColor = "black") +
           scale_fill_gradient2(high = "darkred", 
@@
 sim <- unique(c(which(grepl("simulations", objects)),
-               which(grepl("yeast-", objects)),
+               which(grepl("yeast", objects)),
                which(grepl("polyester", objects))))

### 2018-08-16 datasets/RNA-Seq/yeast-simulation-bimodalalternative-highpi0.Rmd
Add unweighted and uninformative covariate setting

@@
                   nDE=nDE, sampleSize=sampleSize, bd=bd,
-                  uninformativeCovariate = TRUE, mc.cores=nCores)
+                  uninformativeCovariate = TRUE, 
+                  weightByCovariate = FALSE,
+                  mc.cores=nCores)
   saveRDS(de5, file = resfile_d5_uninfCov_uw)
@@
                    nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   uninformativeCovariate = TRUE, mc.cores=nCores)
+                   uninformativeCovariate = TRUE, 
+                   weightByCovariate = FALSE,
+                   mc.cores=nCores)
   saveRDS(de10, file = resfile_d10_uninfCov_uw)

### 2018-08-16 datasets/RNA-Seq/yeast-simulation-highpi0.Rmd
Add unweighted and uninformative covariate setting

@@
                   nDE=nDE, sampleSize=sampleSize, bd=bd,
-                  uninformativeCovariate = TRUE, mc.cores=nCores)
+                  uninformativeCovariate = TRUE, 
+                  weightByCovariate=FALSE,
+                  mc.cores=nCores)
   saveRDS(de5, file = resfile_d5_uninfCov_uw)

### 2018-08-16 manuscript_figures/manuscript_figures.Rmd
Rename Figure of informative covariate impact in yeast

@@
   gp
-  ggsave(file.path(outdir, paste0("Figure_infCov_", org, ".pdf")), height = 3.5, width = 8)
+  ggsave(file.path(outdir, paste0("Figure_infCov_", gsub("-", "", org), ".pdf")), 
+                   height = 3.5, width = 8)
 }

### 2018-08-23 datasets/ChIP-seq/ChIPseq-h3k4me3-promoters.Rmd
Fix typo in ChIPseq; adjust pvalue histograms

@@
 
-We test for differential peaks at each of the .
+We test for differential binding at each of the promoters.
 
@@
 strat_hist(dat, pvalue = "pval", covariate = "ind_covariate",
-           maxy = 10)
+           maxy = 30)
 ```

### 2018-08-28 datasets/GSEA/fgsea-human.Rmd
Add GSEA analysis; remove benchmarking of goSeq analysis
Adding GSEA analysis (by fgsea package) which implements a FCS
(Functional Class Scoring) method. Unlike goSeq, it does not
require an arbitrary threshold to define DE genes. Instead it
uses the rankings from a DE analysis, and checks for enrichment
of these rankings for all genes. Here the gene set size appears
to be both informative and independent under the null, since
we no longer have to perform selection/filtering of the p-values.

Also removing the benchmarking from the goSeq analyses. This is
because we see that (and logically in

--- 2018-07-23 datasets/GSEA/gsea-human.Rmd
@@
 title: "Case Study: Gene Set Enrichment Analysis (Human Data Set)"
-author: "Alejandro Reyes"
+author: "Alejandro Reyes and Keegan Korthauer"
 date: '`r format(Sys.Date(), "%B %d, %Y")`'
@@
 
+Here we use the `fgsea` Bioconductor package to implement the GSEA method. This is 
+a Functional Class Scoring approach, which does not require setting an arbitrary threshold
+for Differential Expression, but instead relies on the gene's rank (here we rank by
+DESeq2 test statistic).
+
 # Workspace Setup
@@
 library(biomaRt)
-library(goseq)
+library(fgsea)
 
@@
 
-Finally, we’ll perform a gene set enrichment analysis with goseq.
-goseq performs GSEA and adjusts for gene length biases.
-The bias is represented in the plot below, with gene length bin on the x-axis and proportion of genes detected to be differentially expressed on the y-axis.
+Now we use the `fgsea` package to perform the gene set enrichment analysis and 
+obtain a enrichment p-value for each pathway.
 
 ```{r}
+# invert the list so each item is a pathway instead of a gene
+goSets <- split(rep(names(goSets), lengths(goSets)), unlist(goSets))
+stats <- res$stat
+names(stats) <- rownames(res)
+stats <- stats[!is.na(stats)]
+
 if (!file.exists(result_file)) {
-    ## get median transcript length for bias adjustment
-    txByGene <- transcriptsBy(EnsDb.Hsapiens.v86, "gene")
-    geneLength <- sapply(width(txByGene), median)
-    geneLength <- geneLength[names(genes)]
-    genes[is.na(genes)] <- 0
-
-    ## perform gsea
-    pwf <- nullp(genes, bias.data = geneLength)
-    goRes <- goseq(pwf, gene2cat = goSets)
-
+  goRes <- fgsea(goSets, 
+                  stats, 
+                  nperm=10000, 
+                  maxSize=500,
+                  minSize=5)  
   saveRDS(goRes, file = result_file)
@@
 }
+```
 
+Add a random (uninformative covariate) to the dataset.
+
+```{r}
 ## Add random (uninformative) covariate
@@
 
-We filter out really small gene sets and those gene sets which have no DE genes
-in them (many of these are set to 1).
-
-```{r}
-goRes <- goRes %>%
-  dplyr::filter(numDEInCat > 0, numInCat > 10 )
-```
-
 ## Covariate Diagnostics
 
-Here, we can see that the covariant variable is actually informative.
+Here, we want to check whether the size of the gene set is actually informative and 
+independent under the null.
 
@@
 ```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
-rank_scatter(dat = goRes, pval = "over_represented_pvalue", 
-              covariate = "numInCat", bins = 50, funx = log2, 
+rank_scatter(dat = goRes, pval = "pval", 
+              covariate = "size", bins = 50, funx = log2, 
               funfill = log10_trans()) +
-    ylim(0, 12) +
-    ggtitle("Over-represented gene sets") +
+    ggtitle("Enriched gene sets") +
     xlab(expression(log[10]~"(# of genes)")) +
@@
 ```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
-strat_hist(goRes, pval="over_represented_pvalue", covariate="numInCat", maxy=7)
+strat_hist(goRes, pval="pval", covariate="size", maxy=11)
 ```
@@
 ```{r informativeCovariate-rand, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
-rank_scatter(dat = goRes, pval = "over_represented_pvalue", 
+rank_scatter(dat = goRes, pval = "pval", 
               covariate = "rand_covar", bins = 50, 
               funfill = log10_trans()) +
-    ylim(0, 12) +
-    ggtitle("Over-represented gene sets") +
+    ggtitle("Enriched gene sets") +
     ylab(expression(-log[10]~"(p-value)") ) 
@@
 ```{r assumptions-rand, message=FALSE, fig.width=10, fig.height=3.2}
-strat_hist(goRes, pval="over_represented_pvalue", covariate="rand_covar", maxy=7)
+strat_hist(goRes, pval="pval", covariate="rand_covar", maxy=11)
 ```
@@
 ## rename columns and prepare for benchmarking
-res <- dplyr:::select(goRes, c("over_represented_pvalue", "numInCat", "rand_covar")) %>%
-    dplyr:::rename(pval = over_represented_pvalue, ind_covariate = numInCat)
-res$pval <- pmin(res$pval, 1)
+res <- dplyr:::select(goRes, c("pval", "size", "rand_covar")) %>%
+    dplyr:::rename(ind_covariate = size)
 

### 2018-08-28 datasets/GSEA/fgsea-mouse.Rmd
Add GSEA analysis; remove benchmarking of goSeq analysis
Adding GSEA analysis (by fgsea package) which implements a FCS
(Functional Class Scoring) method. Unlike goSeq, it does not
require an arbitrary threshold to define DE genes. Instead it
uses the rankings from a DE analysis, and checks for enrichment
of these rankings for all genes. Here the gene set size appears
to be both informative and independent under the null, since
we no longer have to perform selection/filtering of the p-values.

Also removing the benchmarking from the goSeq analyses. This is
because we see that (and logically in

--- 2018-07-23 datasets/GSEA/gsea-mouse.Rmd
@@
 title: "Case Study: Gene Set Enrichment Analysis (Mouse Data Set)"
-author: "Alejandro Reyes"
+author: "Alejandro Reyes and Keegan Korthauer"
 date: '`r format(Sys.Date(), "%B %d, %Y")`'
@@
 library(EnsDb.Mmusculus.v75)
-library(biomaRt)
-library(goseq)
+library(fgsea)
 
@@
 
-**Why no filtering on gene_biotype == "protein_coding" for this data?
-AR: I think I did this in the other vignette to keep running times short.
-This shouldn't change the result since anyway protein coding genes are those with GO categories and genes without GO assignments are ignored.**
-
 # Data Analysis
@@
 
-In order to get a list of genes to test for GSEA, we will test each gene for differential gene expression between hematopoietic stem cells and multipotent progenitors (fraction 1).
-To do this, we will run DESeq2 to retrieve genes differentially expressed at a FDR of 10%.
+In order to rank the list of genes for GSEA, we will test each gene for 
+differential gene expression between hematopoietic stem cells and multipotent 
+progenitors (fraction 1). To do this, we will run DESeq2 to obtain a statistic 
+for differential expression.
 
@@
 
-```{r goseq, message=FALSE}
+```{r gocat, message=FALSE}
 if (!file.exists(goset_file)) {
@@
 
-Now we are ready to perform gene set enrichment analysis using `goseq`.
+Now we use the `fgsea` package to perform the gene set enrichment analysis and 
+obtain a enrichment p-value for each pathway.
 
 ```{r}
+# invert the list so each item is a pathway instead of a gene
+goSets <- split(rep(names(goSets), lengths(goSets)), unlist(goSets))
+stats <- res$stat
+names(stats) <- rownames(res)
+stats <- stats[!is.na(stats)]
+
 if (!file.exists(result_file)) {
-    ## getting median transcript length
-    txByGene <- transcriptsBy(EnsDb.Mmusculus.v75, "gene")
-    geneLength <- sapply(width(txByGene), median)
-    geneLength <- geneLength[names(genes)]
-    genes[is.na(genes)] <- 0
-
-    ## perform gsea
-    pwf <- nullp(genes, bias.data = geneLength)
-    goRes <- goseq(pwf, gene2cat = goSets)
-
+  goRes <- fgsea(goSets, 
+                  stats, 
+                  nperm=10000, 
+                  maxSize=500,
+                  minSize=5)  
   saveRDS(goRes, file = result_file)
@@
 
-We filter out really small gene sets and those gene sets which have no DE genes
-in them (many of these are set to 1).
-
-```{r}
-goRes <- goRes %>%
-  dplyr::filter(numDEInCat > 0, numInCat > 10 )
-```
-
 ## Covariate Diagnostics
 
-Here, we can see that the size of the gene set is actually informative.
+Here, we want to check whether the size of the gene set is actually informative and 
+independent under the null.
 
@@
 ```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
-rank_scatter(dat = goRes, pval = "over_represented_pvalue", 
-             covariate = "numInCat", bins = 50, funx = log2, 
+rank_scatter(dat = goRes, pval = "pval", 
+             covariate = "size", bins = 50, funx = log2, 
              funfill = log10_trans()) +
-    ylim(0, 12) +
-    ggtitle("Over-represented gene sets") +
+    ggtitle("Enriched gene sets") +
     xlab(expression(log[10]~"(# of genes)")) +
@@
 ```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
-strat_hist(goRes, pval = "over_represented_pvalue", covariate = "numInCat", maxy=7)
+strat_hist(goRes, pval = "pval", covariate = "size", maxy=12)
 ```
@@
 ```{r assumptions-rand, message=FALSE, fig.width=10, fig.height=3.2}
-strat_hist(goRes, pval="over_represented_pvalue", covariate="rand_covar", maxy=10)
+strat_hist(goRes, pval="pval", covariate="rand_covar", maxy=10)
 ```
@@
 ## rename columns and prepare for benchmarking
-res <- dplyr::select(goRes, over_represented_pvalue, numInCat, rand_covar) %>%
-    dplyr::rename(pval = over_represented_pvalue,
-                  ind_covariate = numInCat) %>%
-    dplyr::mutate(pval = pmin(pval, 1))
+res <- dplyr::select(goRes, pval, size, rand_covar) %>%
+    dplyr::rename(pval = pval,
+                  ind_covariate = size)
 

### 2018-08-28 datasets/GSEA/goseq-human.Rmd
Add GSEA analysis; remove benchmarking of goSeq analysis
Adding GSEA analysis (by fgsea package) which implements a FCS
(Functional Class Scoring) method. Unlike goSeq, it does not
require an arbitrary threshold to define DE genes. Instead it
uses the rankings from a DE analysis, and checks for enrichment
of these rankings for all genes. Here the gene set size appears
to be both informative and independent under the null, since
we no longer have to perform selection/filtering of the p-values.

Also removing the benchmarking from the goSeq analyses. This is
because we see that (and logically in

--- 2018-07-23 datasets/GSEA/gsea-human.Rmd
@@
 In order to get a list of genes, we will use RNA-seq data consisting of cortex and cerebellum samples of a subset of the GTEx data.
+
+Here we use the `goseq` Bioconductor package to implement the gene set analysis. This is 
+an Over-Representation Analysis which does requires setting an arbitrary threshold
+for Differential Expression. The test concerns whether any of the GO sets are enriched
+for DE genes.
 
@@
 goset_file <- file.path(datdir, "human-gosets.rds")
-result_file <- file.path(resdir, "human-results.rds")
-bench_file <- file.path(sbdir, "human-benchmark.rds")
-bench_file_uninf <- file.path(sbdir, "human-uninf-benchmark.rds")
+result_file <- file.path(resdir, "human-results-goseq.rds")
 ```
@@
 goRes <- goRes %>%
-  dplyr::filter(numDEInCat > 0, numInCat > 10 )
+  dplyr::filter(numDEInCat > 0, numInCat > 5 )
 ```
@@
 
-Here, we can see that the covariant variable is actually informative.
+Here, we want to check whether the size of the gene set is actually informative and 
+independent under the null.
 
@@
 
-We can also explore if the covariate seems to be independent under the null.
+But unfortunately the distribution 
+p-values is quite different for the different covariate strata. This is not surprising, as
+we expect that smaller gene sets that have at least one DE gene in them will be skewed 
+toward smaller p-values (as compared to larger gene sets that have at least one DE gene), 
+simply because a single DE gene represents a larger proportion in the smaller sets.
+If we don't condition on sets that have at least one DE gene, however, the distribution 
+of p-values is spiked at 1 since goseq assigns all such sets to a p-value of 1.
 
@@
 
-### Random
-
-We will explore whether the random covariate can be used as a covariate for modern multiple-testing correction methods in the context of GSEA.
-In the plot below, the log10 of the p-values is plotted as a function of the
-random covariate. This covariate looks independent of the p-values.
-
-```{r informativeCovariate-rand, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
-rank_scatter(dat = goRes, pval = "over_represented_pvalue", 
-              covariate = "rand_covar", bins = 50, 
-              funfill = log10_trans()) +
-    ylim(0, 12) +
-    ggtitle("Over-represented gene sets") +
-    ylab(expression(-log[10]~"(p-value)") ) 
-```
-
-We can also explore if the covariate seems to be independent under the null.
-
-```{r assumptions-rand, message=FALSE, fig.width=10, fig.height=3.2}
-strat_hist(goRes, pval="over_represented_pvalue", covariate="rand_covar", maxy=7)
-```
-
-## Multiple-Testing Correction
-
-We then execute the benchDesign and generate a *SummarizedBenchmark* object containing multiple-testing corrections using several methods.
-
-```{r setup-benckmark, message=FALSE}
-## rename columns and prepare for benchmarking
-res <- dplyr:::select(goRes, c("over_represented_pvalue", "numInCat", "rand_covar")) %>%
-    dplyr:::rename(pval = over_represented_pvalue, ind_covariate = numInCat)
-res$pval <- pmin(res$pval, 1)
-
-## generate default BenchDesign
-bd <- initializeBenchDesign()
-```
-
-We don't include `ashq`, `fdrreg-e` and `fdrreg-t` from the analysis because
-the necessary assumptions are not met in the current case study. Namely, effect sizes
-and standard errors are not available for ASH, and the test statistics are not normally
-distributed under the null and alternative, as required by Scott's FDR regression methods.
-
-```{r run-benchmark}
-if (!file.exists(bench_file)) {
-  sGSEA <- buildBench(bd, res, ftCols = "ind_covariate")
-  saveRDS(sGSEA, file = bench_file)
-} else {
-  sGSEA <- readRDS(bench_file)
-}
-```
-
-We'll also compare the results to an uninformative (random) covariate.
-
-```{r run-benchmark-rand}
-if (!file.exists(bench_file_uninf)) {
-  res$ind_covariate <- res$rand_covar
-  sGSEA_rand <- buildBench(bd, res, ftCols = "ind_covariate")
-  saveRDS(sGSEA_rand, file = bench_file_uninf)
-} else {
-  sGSEA_rand <- readRDS(bench_file_uninf)
-}
-```
-
-## Benchmark Metrics
-
-### Gene Set Size
-
-```{r}
-assayNames(sGSEA) <- "qvalue"
-sGSEA <- addDefaultMetrics(sGSEA)
-sGSEA <- estimatePerformanceMetrics(sGSEA, addColData=TRUE)
-```
-
-```{r plottingResults}
-rejections_scatter(sGSEA, as_fraction=FALSE, supplementary=FALSE)
-```
-
-```{r}
-plotFDRMethodsOverlap(sGSEA, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100)
-```
-
-```{r}
-covariateLinePlot(sGSEA, alpha = 0.1, covname = "ind_covariate", trans = "log1p")
-```
-
-### Random
-
-```{r}
-assayNames(sGSEA_rand) <- "qvalue"
-sGSEA_rand <- addDefaultMetrics(sGSEA_rand)
-sGSEA_rand <- estimatePerformanceMetrics(sGSEA_rand, addColData=TRUE)
-```
-
-```{r plottingResults-rand}
-rejections_scatter(sGSEA_rand, as_fraction=FALSE, supplementary=FALSE)
-```
-
-```{r}
-plotFDRMethodsOverlap(sGSEA_rand, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100)
-```
-
-```{r}
-covariateLinePlot(sGSEA_rand, alpha = 0.1, covname = "ind_covariate", trans = "log1p")
-```
-
-## Covariate comparison
-
-Here we compare the method ranks for the two covariates at alpha = 0.10.
-
-```{r}
-plotMethodRanks(c(bench_file, bench_file_uninf), 
-                colLabels = c("Set Size", "Random"), 
-                alpha = 0.10, xlab = "Covariate", 
-                excludeMethods = NULL)
-```
+This suggests that
+the covariate is *not* independent under the null hypothesis, so the assumptions
+of many of the methods (which use an independent covariate) are not satisfied. 
+Therefore we will not proceed with the benchmarking using these results.
 

### 2018-08-28 datasets/GSEA/goseq-mouse.Rmd
Add GSEA analysis; remove benchmarking of goSeq analysis
Adding GSEA analysis (by fgsea package) which implements a FCS
(Functional Class Scoring) method. Unlike goSeq, it does not
require an arbitrary threshold to define DE genes. Instead it
uses the rankings from a DE analysis, and checks for enrichment
of these rankings for all genes. Here the gene set size appears
to be both informative and independent under the null, since
we no longer have to perform selection/filtering of the p-values.

Also removing the benchmarking from the goSeq analyses. This is
because we see that (and logically in

--- 2018-07-23 datasets/GSEA/gsea-mouse.Rmd
@@
 We will test for difference in gene expression between two different cell fractions and perform GSEA in the resulting list of genes.
+
+Here we use the `goseq` Bioconductor package to implement the gene set analysis. This is 
+an Over-Representation Analysis which does requires setting an arbitrary threshold
+for Differential Expression. The test concerns whether any of the GO sets are enriched
+for DE genes.
 
@@
 goset_file <- file.path(datdir, "mouse-gosets.rds")
-result_file <- file.path(resdir, "mouse-results.rds")
-bench_file <- file.path(sbdir, "mouse-benchmark.rds")
-bench_file_uninf <- file.path(sbdir, "mouse-uninf-benchmark.rds")
+result_file <- file.path(resdir, "mouse-results-goseq.rds")
 ```
@@
 
-Here, we can see that the size of the gene set is actually informative.
+Here, we want to check whether the size of the gene set is actually informative and 
+independent under the null.
 
@@
 
-And at the same time, it seems to be fair to assume that it is independent under the null hypothesis.
+But unfortunately the distribution 
+p-values is quite different for the different covariate strata. This is not surprising, as
+we expect that smaller gene sets that have at least one DE gene in them will be skewed 
+toward smaller p-values (as compared to larger gene sets that have at least one DE gene), 
+simply because a single DE gene represents a larger proportion in the smaller sets.
+If we don't condition on sets that have at least one DE gene, however, the distribution 
+of p-values is spiked at 1 since goseq assigns all such sets to a p-value of 1.
 
@@
 
-### Random
-
-We will explore whether the random covariate can be used as a covariate for modern multiple-testing correction methods in the context of GSEA.
-In the plot below, the log10 of the p-values is plotted as a function of the
-random covariate. This covariate looks independent of the p-values.
-
-```{r informativeCovariate-rand, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
-rank_scatter(dat = goRes, pval = "over_represented_pvalue", 
-              covariate = "rand_covar", bins = 50, 
-              funfill = log10_trans()) +
-    ylim(0, 12) +
-    ggtitle("Over-represented gene sets") +
-    ylab(expression(-log[10]~"(p-value)") ) 
-```
-
-We can also explore if the covariate seems to be independent under the null.
-
-```{r assumptions-rand, message=FALSE, fig.width=10, fig.height=3.2}
-strat_hist(goRes, pval="over_represented_pvalue", covariate="rand_covar", maxy=10)
-```
-
-## Multiple-Testing Correction
-
-Generating the *SummarizedBenchmark* object:
-
-```{r setup-benckmark, message=FALSE}
-## rename columns and prepare for benchmarking
-res <- dplyr::select(goRes, over_represented_pvalue, numInCat, rand_covar) %>%
-    dplyr::rename(pval = over_represented_pvalue,
-                  ind_covariate = numInCat) %>%
-    dplyr::mutate(pval = pmin(pval, 1))
-
-## generate default BenchDesign
-bd <- initializeBenchDesign()
-```
-
-We don't include `ashq`, `fdrreg-e` and `fdrreg-t` from the analysis because
-the necessary assumptions are not met in the current case study. Namely, effect sizes
-and standard errors are not available for ASH, and the test statistics are not normally
-distributed under the null and alternative, as required by Scott's FDR regression methods.
-
-```{r run-benchmark}
-if (!file.exists(bench_file)) {
-    sGSEAMouse <- buildBench(bd, res, ftCols = "ind_covariate")
-    saveRDS(sGSEAMouse, file = bench_file)
-} else {
-    sGSEAMouse <- readRDS(bench_file)
-}
-```
-
-We'll also compare the results to an uninformative (random) covariate.
-
-```{r run-benchmark-rand}
-if (!file.exists(bench_file_uninf)) {
-  res$ind_covariate <- res$rand_covar
-  sGSEAMouse_rand <- buildBench(bd, res, ftCols = "ind_covariate")
-  saveRDS(sGSEAMouse_rand, file = bench_file_uninf)
-} else {
-  sGSEAMouse_rand <- readRDS(bench_file_uninf)
-}
-```
-
-## Benchmark Metrics
-
-### Gene Set Size
-
-```{r}
-assayNames(sGSEAMouse) <- "qvalue"
-sGSEAMouse <- addDefaultMetrics(sGSEAMouse)
-```
-
-```{r plottingResults}
-rejections_scatter(sGSEAMouse, as_fraction = FALSE, supplementary = FALSE)
-```
-
-```{r}
-plotFDRMethodsOverlap(sGSEAMouse, alpha = 0.1, supplementary = FALSE,
-                      order.by = "freq", nsets = 100)
-```
-
-```{r}
-covariateLinePlot(sGSEAMouse, alpha = 0.1, covname = "ind_covariate", trans="log1p")
-```
-
-### Random
-
-```{r}
-assayNames(sGSEAMouse_rand) <- "qvalue"
-sGSEAMouse_rand <- addDefaultMetrics(sGSEAMouse_rand)
-sGSEAMouse_rand <- estimatePerformanceMetrics(sGSEAMouse_rand, addColData=TRUE)
-```
-
-```{r plottingResults-rand}
-rejections_scatter(sGSEAMouse_rand, as_fraction=FALSE, supplementary=FALSE)
-```
-
-```{r}
-plotFDRMethodsOverlap(sGSEAMouse_rand, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100)
-```
-
-```{r}
-covariateLinePlot(sGSEAMouse_rand, alpha = 0.1, covname = "ind_covariate", trans = "log1p")
-```
-
-## Covariate comparison
-
-Here we compare the method ranks for the two covariates at alpha = 0.10.
-
-```{r}
-plotMethodRanks(c(bench_file, bench_file_uninf), 
-                colLabels = c("Set Size", "Random"), 
-                alpha = 0.10, xlab = "Covariate", 
-                excludeMethods = NULL)
-```
+This suggests that
+the covariate is *not* independent under the null hypothesis, so the assumptions
+of many of the methods (which use an independent covariate) are not satisfied. 
+Therefore we will not proceed with the benchmarking using these results.
 

### 2018-08-29 manuscript_figures/manuscript_figures.Rmd
Adjust y-axis limits in Figure 6

@@
 # set alpha cutoff for plots that are fixed at a certain value of alpha
-alpha <- 0.10
+alpha <- 0.05
 ```
@@
 pa <- strat_hist(res, pvalue = "pval",
-           covariate = "ind_covariate", maxy = 10, axislabs = FALSE) +
+           covariate = "ind_covariate", maxy = 8.75, axislabs = FALSE) +
   draw_plot_label("  RNA-seq case study: Brain", 0.5, 1, hjust = 0.5,fontface = "bold", 
@@
 pc <- strat_hist(res, pvalue = "pval",
-           covariate = "ind_covariate", maxy = 7, axislabs = FALSE) +
+           covariate = "ind_covariate", maxy = 8.5, axislabs = FALSE) +
   draw_plot_label("  GSEA case study: Mouse", 0.5, 1, hjust = 0.5, fontface = "bold" , size = 12) +

### 2018-08-29 datasets/ChIP-seq/ChIPseq-h3k4me3-promoters.Rmd
Update ChIPseq promoter analysis
Add FDRreg to the promoter ChIPseq analysis since the test statistic
(DESeq2) is normally distributed. In addition, don't use list.files
to grab a list of the bam files when constructing metadata (since
if other ChIP analyses have been run, this will grab all a vector
of all the bam files used in all analyses).

@@
 ## intermediary files we create below
-count_file <- file.path(resdir, "h3k4me3-promotors-counts.rds")
-result_file <- file.path(resdir, "h3k4me3-promotors-results.rds")
-bench_file <- file.path(sbdir, "h3k4me3-promotors-benchmark.rds")
+count_file <- file.path(resdir, "h3k4me3-promoters-counts.rds")
+result_file <- file.path(resdir, "h3k4me3-promoters-results.rds")
+bench_file <- file.path(sbdir, "h3k4me3-promoters-benchmark.rds")
 
@@
 ```{r}
-bamfiles <- list.files(datdir, pattern = 'bam$', full = TRUE)
+bamfiles <- c(broad_bams, uw_bams)
 labs <- gsub("wgEncode(.*)Histone.*", "\\1", basename(bamfiles))
@@
                        Strand = strand(prom))
-    rc <- featureCounts(files = bamfiles,
+    rc <- featureCounts(files = file.path(datdir, bamfiles),
                         annot.ext = anno,
@@
 object, which will run the functions specified in each method. 
+We also add in Scott's FDR Regression (both
+`nulltype = "empirical"` and `nulltype = "theoretical"`)
+since our test statistics are normally-distributed.
 
@@
     bd <- initializeBenchDesign()
+    bd <- addBMethod(bd, "fdrreg-t",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'theoretical',
+                     control = list(lambda = 0.01))
+    bd <- addBMethod(bd, "fdrreg-e",
+                     FDRreg::FDRreg,
+                     function(x) { x$FDR },
+                     z = test_statistic,
+                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
+                     nulltype = 'empirical',
+                     control = list(lambda = 0.01))
+
     sb <- buildBench(bd, data = dat, ftCols = c("ind_covariate"),

### 2018-08-29 datasets/ChIP-seq/ChIPseq-CBP-csaw.Rmd
Fix typo in ChIPseq; adjust pvalue histograms
Update ChiPseq csaw analyses
Add mean coverage (across all samples) as an informative covariate.
In addition, update call to csaw::normOffsets to reflect changes in
current release version that change the default return object to a
SE.

--- 2018-07-05 datasets/ChIP-seq/ChIPseq-CBP-csaw.Rmd
@@
 library(edgeR)
+library(GenomicAlignments)
 
@@
 bench_file <- file.path(sbdir, "cbp-csaw-benchmark.rds")
+bench_file_cov <- file.path(sbdir, "cbp-csaw-cov-benchmark.rds")
 bench_file_uninf <- file.path(sbdir, "cbp-csaw-uninf-benchmark.rds")
@@
     ## calculate normalizing offsets based on larger windows
-    normfacs <- normOffsets(bins)
+    normfacs <- normOffsets(bins, se.out = FALSE)
 
@@
                    best_logFC = tab_best$logFC)
+
+    ## get overall mean counts for each window
+    merged_cnts <- summarizeOverlaps(res_ranges, BamFileList(meta$sortedfile))
+    res_ranges$meancnt <- rowMeans(assays(merged_cnts)$counts)
+
     saveRDS(res_ranges, file = result_file)
@@
 ```{r, nwin-diag-hist, fig.width=10, fig.height=3.2}
-strat_hist(res_df, pvalue = "PValue", covariate = "nWindows", maxy = 35, numQ = 3)
+strat_hist(res_df, pvalue = "PValue", covariate = "nWindows", maxy = 15)
 ```
@@
 ```{r, width-diag-hist, fig.width=10, fig.height=3.2}
-strat_hist(res_df, pvalue = "PValue", covariate = "width", maxy = 35, numQ = 3)
+strat_hist(res_df, pvalue = "PValue", covariate = "width", maxy = 15)
+```
+
+### Mean Coverage
+
+```{r, cov-diag-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(res_df, pvalue = "PValue", covariate = "meancnt") +
+    ggtitle("Mean coverage as independent covariate") +
+    xlab("Mean coverage")
+```
+
+```{r, cov-diag-hist, fig.width=10, fig.height=3.2}
+strat_hist(res_df, pvalue = "PValue", covariate = "meancnt", maxy = 15)
 ```
@@
 ```{r, rand-diag-hist, fig.width=10, fig.height=3.2}
-strat_hist(res_df, pvalue = "PValue", covariate = "uninf_covar", maxy = 35, numQ = 3)
+strat_hist(res_df, pvalue = "PValue", covariate = "uninf_covar", maxy = 15)
 ```
@@
 We use the common `BenchDesign` with the set of multiple testing correction
-methods already included. We use the width of the regions as the independent
-covariate.
+methods already included. We investigate both the width of the regions and their
+mean coverage as the independent covariate. We won't assess the number of windows,
+since this is very tightly correlated with the width of the region.
+
+```{r}
+cor(res_df[,c("width", "nWindows", "meancnt")])
+```
+
+First, we'll use the region width covariate.
 
@@
     sb <- readRDS(bench_file)
+}
+```
+
+Next, we'll use the mean coverage covariate.
+
+```{r}
+if (!file.exists(bench_file_cov)) {
+    res_df$ind_covariate <- res_df$meancnt
+    res_df$pval <- res_df$PValue
+
+    bd <- initializeBenchDesign()
+    sbC <- buildBench(bd, data = res_df, ftCols = "ind_covariate")
+    saveRDS(sbC, file = bench_file_cov)
+} else {
+    sbC <- readRDS(bench_file_cov)
 }
@@
 
+### Mean Coverage
+
+We'll do the same for the mean coverage covariate.
+
+```{r, chipseq-metrics-cov}
+assayNames(sbC) <- "qvalue"
+sbC <- addDefaultMetrics(sbC)
+```
+
+Now, we'll plot the results.
+
+```{r, chipseq-scatter-cov, results="hide"}
+rejections_scatter(sbC, supplementary = FALSE)
+rejection_scatter_bins(sbC, covariate = "ind_covariate",
+                       bins = 4, supplementary = FALSE)
+```
+
+```{r, chipseq-overlap-cov}
+plotFDRMethodsOverlap(sbC, alpha = 0.05, nsets = ncol(sbC),
+                      order.by = "freq", decreasing = TRUE,
+                      supplementary = FALSE)
+```
+
+```{r, rnaseq-lineplot-cov}
+covariateLinePlot(sbC, alpha = 0.05, covname = "ind_covariate")
+```
+
 ### Random
@@
 ```{r, chipseq-overlap-rand}
-plotFDRMethodsOverlap(sbU, alpha = 0.05, nsets = ncol(sb),
+plotFDRMethodsOverlap(sbU, alpha = 0.05, nsets = ncol(sbU),
                       order.by = "freq", decreasing = TRUE,
@@
 ```{r}
-plotMethodRanks(c(benchfile, benchfile_uninf), 
-                colLabels = c("Region width", "Random"), 
+plotMethodRanks(c(bench_file, bench_file_cov, bench_file_uninf), 
+                colLabels = c("Region width", "Mean Coverage", "Random"), 
                 alpha = 0.10, xlab = "Covariate", 

### 2018-08-29 datasets/ChIP-seq/ChIPseq-h3k4me3-csaw.Rmd
Update ChiPseq csaw analyses
Add mean coverage (across all samples) as an informative covariate.
In addition, update call to csaw::normOffsets to reflect changes in
current release version that change the default return object to a
SE.

@@
 bench_file <- file.path(sbdir, "h3k4me3-csaw-benchmark.rds")
+bench_file_cov <- file.path(sbdir, "h3k4me3-csaw-cov-benchmark.rds")
 bench_file_uninf <- file.path(sbdir, "h3k4me3-csaw-uninf-benchmark.rds")
@@
     ## normalize for library-specific trended biases
-    offsets <- normOffsets(filtered_cnts, type = "loess")
+    filtered_cnts <- normOffsets(filtered_cnts, type = "loess")
 
@@
     y <- asDGEList(filtered_cnts)
-    y$offset <- offsets
     y <- estimateDisp(y, design)
@@
                    best_logFC = tab_best$logFC)
+
+    ## get overall mean counts for each window
+    merged_cnts <- summarizeOverlaps(res_ranges, BamFileList(meta$sortedfile))
+    res_ranges$meancnt <- rowMeans(assays(merged_cnts)$counts)
 
@@
 ```{r, nwin-diag-hist, fig.width=10, fig.height=3.2}
-strat_hist(res_df, pvalue = "PValue", covariate = "nWindows", maxy = 35, numQ = 3)
+strat_hist(res_df, pvalue = "PValue", covariate = "nWindows", maxy = 35)
 ```
@@
 ```{r, width-diag-hist, fig.width=10, fig.height=3.2}
-strat_hist(res_df, pvalue = "PValue", covariate = "width", maxy = 35, numQ = 3)
+strat_hist(res_df, pvalue = "PValue", covariate = "width", maxy = 35)
+```
+
+### Mean Coverage
+
+```{r, cov-diag-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(res_df, pvalue = "PValue", covariate = "meancnt") +
+    ggtitle("Mean coverage as independent covariate") +
+    xlab("Mean coverage")
+```
+
+```{r, cov-diag-hist, fig.width=10, fig.height=3.2}
+strat_hist(res_df, pvalue = "PValue", covariate = "meancnt", maxy = 35)
 ```
@@
 ```{r, rand-diag-hist, fig.width=10, fig.height=3.2}
-strat_hist(res_df, pvalue = "PValue", covariate = "uninf_covar", maxy = 35, numQ = 3)
+strat_hist(res_df, pvalue = "PValue", covariate = "uninf_covar", maxy = 35)
 ```
@@
 
+### Mean coverage
+
+We'll do the same for the mean coverage covariate.
+
+```{r, chipseq-metrics-cov}
+assayNames(sbC) <- "qvalue"
+sbC <- addDefaultMetrics(sbC)
+```
+
+Now, we'll plot the results.
+
+```{r, chipseq-scatter-cov, results="hide"}
+rejections_scatter(sbC, supplementary = FALSE)
+rejection_scatter_bins(sbC, covariate = "ind_covariate",
+                       bins = 4, supplementary = FALSE)
+```
+
+```{r, chipseq-overlap-cov}
+plotFDRMethodsOverlap(sbC, alpha = 0.05, nsets = ncol(sbC),
+                      order.by = "freq", decreasing = TRUE,
+                      supplementary = FALSE)
+```
+
+```{r, rnaseq-lineplots-cov}
+covariateLinePlot(sbC, alpha = 0.05, covname = "ind_covariate")
+```
+
 ### Random
@@
 ```{r}
-plotMethodRanks(c(bench_file, bench_file_uninf), 
-                colLabels = c("Region width", "Random"), 
+plotMethodRanks(c(bench_file, bench_file_cov, bench_file_uninf), 
+                colLabels = c("Region width", "Mean Coverage", "Random"), 
                 alpha = 0.10, xlab = "Covariate", 

### 2018-08-30 datasets/simulations/simulations-summary.Rmd
Add manuscript Figure3 drafts discussed on 0824

@@
 
-```{r Figure2-load-data}
+```{r load-workspace}
 # Load packages and source benchmark FDR
@@
 }
-
-# assumes sb objects are in the following location
-path <- file.path("..", "datasets", "simulations", "results")
 
@@
 ```{r, eval = FALSE}
-for (ia in c(1:10, 20, 30, 40, 50)) {
-    summary_file <- file.path("results", paste0("summary_alpha", ia, ".rds"))
+for (ia in 1:10) {
+    summary_file <- file.path("results-summary", paste0("summary_alpha", ia, ".rds"))
     if (!file.exists(summary_file)) {
-        res_files <- list.files(path, "\\.rds$", full.names = TRUE)
+        res_files <- list.files("results", "\\.rds$", full.names = TRUE)
         res <- lapply(res_files, function(x) {
@@
     imethodset[which(imethodset == "ihw")] <- paste0("ihw-a", sprintf("%02d", alphas[ia]))
-    ires <- readRDS(file.path("results", paste0("summary_alpha", alphas[ia], ".rds")))
+    ires <- readRDS(file.path("results-summary", paste0("summary_alpha", alphas[ia], ".rds")))
     ires <- bind_rows(ires, .id = "setting")
@@
 resmet <- dplyr::filter(resmet, !( grepl("chi", dist) & grepl("fdrreg", Method) ))
+```
+
+Since the table will be used again for generating manuscript figures, we save it as
+an RDS object for easier future access.
+
+```{r save-resmet}
+resmet_file <- file.path("results-summary", "result-metrics.rds")
+if (!file.exists(resmet_file)) {
+    saveRDS(resmet, file = resmet_file)
+}
 ```

### 2018-08-30 manuscript_figures/manuscript_figures.Rmd
Consider 0 -> 0 as zero change in infCov heatmaps (previously NA)

@@
 
+# calculate percentage diff as percentage of the smallest (nonzero) number 
+# of rejections (i.e. using n rejects with informative or uninf for denominator,
+# depending on which one was smaller while still being greater than zero)
 dftab <- left_join(inftab, uninftab, by = "id") %>%
   dplyr::mutate(diff_inf = nrejects - uninf) %>%
-  dplyr::mutate(diff_inf_percent = diff_inf / nrejects) %>%
+  dplyr::mutate(basenum = ifelse(nrejects == 0 | uninf == 0, 
+                                 pmax(nrejects, uninf),
+                                 pmin(nrejects, uninf))) %>%
+  dplyr::mutate(diff_inf_percent = diff_inf / basenum) %>%
   dplyr::mutate(method = factor(method, levels = rowOrder)) %>%
@@
   mutate(top = ifelse(diff_inf_percent == top, diff_inf_percent, NA))
+
+# if no change, set diff_inf_percent to zero, even if no rejections in either case
+dftab <- dftab %>% 
+  mutate(diff_inf_percent = ifelse(nrejects == uninf, 0, diff_inf_percent))
 

### 2018-08-30 manuscript_figures/manuscript_figures_sims.Rmd

@@
 ---
-title: "Simulation Figures for Genome Biology (simulations)"
+title: "Manuscript Figures for Genome Biology (simulation subset)"
 author: "Rafalab"
@@
 
+Running this
+
 ```{r Figure2-load-data}
-# Load packages and source benchmark FDR
-library(tidyr)
-library(dplyr)
-library(ggplot2)
-library(magrittr)
+## Load packages and source benchmark FDR
+library(tidyverse)
 library(cowplot)
-library(tibble)
 library(ggthemes)
@@
 # assumes sb objects are in the following location
-path <- file.path("..", "datasets", "simulations", "results")
+resmet_file <- file.path("..", "datasets", "simulations", "results-summary", "result-metrics.rds")
 
@@
                "fdrreg-e", "fdrreg-t", "ashq", "adapt-glm")
-```
-
-# Data Preparation
-
-First, we load summaries for all simulations results at significance thresholds
-between 0.01 and 0.10. This process can take a long time (approximately 40 minutes
-per alpha cutoff) because of the large number of simulation settings that must be
-read in and summarized. Ideally, this initial step should be run externally separate
-from this Rmd (by default, the following code chunk will not be run).
-
-```{r, eval = FALSE}
-for (ia in c(1:10, 20, 30, 40, 50)) {
-    summary_file <- paste0("allres_alpha", ia, ".rds")
-    if (!file.exists(summary_file)) {
-        res_files <- list.files(path, "\\.rds$", full.names = TRUE)
-        res <- lapply(res_files, function(x) {
-            zz <- readRDS(x)
-            zz_i <- lapply(zz, `[[`, "informative")
-            zz_u <- lapply(zz, `[[`, "uninformative")
-            zz_i <- plotsim_standardize(zz_i, alpha = ia / 100L)
-            zz_i <- dplyr::select(zz_i, rep, blabel, param.alpha, key, performanceMetric, alpha, value)
-            zz_u <- plotsim_standardize(zz_u, alpha = ia / 100L)
-            zz_u <- dplyr::select(zz_u, rep, blabel, param.alpha, key, performanceMetric, alpha, value)
-            dplyr::full_join(zz_i, zz_u,
-                             by = c("rep", "blabel", "param.alpha", "key", "performanceMetric", "alpha"),
-                             suffix = c(".info", ".uninfo"))
-        })
-        names(res) <- gsub("\\.rds$", "", basename(res_files))
-        saveRDS(res, summary_file)
-    }
-}
-```
-
-Assuming that the summaries have been saved for all simulation settings at
-the various FDR thresholds using the code above, we load the summarized results
-and merge them into a single results table.
-
-```{r}
-alphas <- 1:10
-res <- vector("list", length(alphas))
-for (ia in seq_len(length(alphas))) {
-    imethodset <- methodset
-    imethodset[which(imethodset == "ihw")] <- paste0("ihw-a", sprintf("%02d", alphas[ia]))
-    ires <- readRDS(paste0("allres_alpha", alphas[ia], ".rds"))
-    ires <- bind_rows(ires, .id = "setting")
-    ires <- dplyr::mutate(ires, dist = gsub(".*?-benchmark-(.*)", "\\1", setting),
-                         setting = gsub("(.*?)-benchmark-.*", "\\1", setting))
-    ires <- dplyr::filter(ires, blabel %in% imethodset)
-    ires <- dplyr::mutate(ires, blabel = gsub("(-a)(.*)", "", blabel))
-    res[[ia]] <- ires
-}
-res <- bind_rows(res)
-res <- dplyr::select(res, -param.alpha)
-res <- dplyr::rename(res, Method = blabel)
-res <- dplyr::mutate(res, Method = gsub("-df03", "", Method))
-res <- dplyr::mutate(res, Method = factor(Method, levels = gsub("-df03", "", methodset)))
-```
-
-We compute summary metrics for each method, setting, and cutoff.
-
-```{r}
-resmet <- dplyr::group_by(res, setting, dist, Method, key, performanceMetric, alpha)
-resmet <- dplyr::summarize(resmet,
-                           mean.info = mean(value.info, na.rm = TRUE),
-                           se.info = sd(value.info, na.rm = TRUE) / sqrt(sum(!is.na(value.info))),
-                           nNA.info = sum(is.na(value.info)),
-                           mean.diff = mean(value.info - value.uninfo, na.rm = TRUE),
-                           se.diff = sd(value.info - value.uninfo, na.rm = TRUE) /
-                               sqrt(sum(!is.na(value.info - value.uninfo))))
-resmet <- dplyr::ungroup(resmet)
-```
-
-We clean up the annotations in the table for some of the more complex simulation settings.
-
-```{r}
-resmet <- dplyr::mutate(resmet, inform = ifelse(grepl("varyinginfo", setting), dist, NA))
-resmet <- dplyr::mutate(resmet, inform = as.numeric(gsub("level", "", inform)))
-resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("varyinginfo", setting), "gaussian", dist))
-
-resmet <- dplyr::mutate(resmet, pi0 = ifelse(grepl("varyingpi0", setting), dist, 90))
-resmet <- dplyr::mutate(resmet, pi0 = as.numeric(gsub("nullprop", "", pi0)))
-resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("varyingpi0", setting), "gaussian", dist))
-
-resmet <- dplyr::mutate(resmet, signal = ifelse(grepl("varyinges", setting), dist, NA))
-resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyinges$", setting), "gaussian", dist))
-resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyinges-t$", setting), "t11", dist))
-resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyinges-chisq$", setting), "chisq4", dist))
-resmet <- dplyr::mutate(resmet, setting = ifelse(grepl("varyinges", setting), "varyinges", setting))
-```
-
-We also replace the sampling noise labels with more plotting-friendly labels.
-
-```{r}
-resmet <- dplyr::mutate(resmet, dist = factor(dist, levels = c("gaussian", "t11", "t5", "chisq4"),
-                                              labels = c(expression("Noise: " * N(0, 1)),
-                                                         expression("Noise: " * t[11]),
-                                                         expression("Noise: " * t[5]),
-                                                         expression("Noise: " * {chi^2}[4]))))
 ```
@@
 
-There are several different simulation settings that can be investigated.
+To generate the figures in this document, the simulation results must first be aggregated by running the code at `datasets/simulations/simulations-summary.Rmd`.
 
 ```{r}
-dplyr::distinct(resmet, setting)
+resmet <- readRDS(resmet_file)
 ```
 
-Namely, simulation settings were run with to investigate the impact of different
-informative covariates, noninformative covariates, varying null proportions,
-varying covariate informativeness, and varying effect size distributions (including
-unimodal effects as described in the ASH manuscript). Many of these simulation
-settings were also run with different sampling noise distributions, including
-Normal/Gaussian, t with 11 degrees of freedom, t with 5 degrees of freedom, and
-Chi-squared with 4 degrees of freedom.
+Unfortunately, because the plots in this section illustrate FDR and TPR acorss varying simulation settings rather than varying nominal FDR cutoff, we cannot simply use the `plotsim_average()` function to generate plots.
+Instead, we define the following helper function to consistently generate plots similar to those output by `plotsim_average()` with arbitrary covariate, e.g.
+number of tests or proportion of null hypotheses, as the x-axis.
+The function assumes that the input table only includes FDR or TPR values for each method at a single nominal alpha cutoff.
+In these figures, we only plot FDR and TPR values at the nominal alpha cutoff of 0.05.
 
-# Null Simulations
-
-We subset on simulation results that look at informative and non-informative covariates across
-different noise distributions.
-
-For null simulations, rather than the false discovery rate, we take a look at the true negative
-rate (equivalently, the proportion of false positives) since the FDR will be 1 if any rejections
-are made. Additionally, we take a look at the FWER (the proportion of simulation replicates out
-of 100 with at least one rejection). In these plots, `ashq`, `fdrreg-e` and `fdrreg-t`
-are excluded for the Chi-Squared simulation settings.
-
-```{r, fig.width = 18, fig.height = 8}
-res_nullset <- dplyr::filter(resmet, setting == "null",
-                             !(Method == "ashq" & grepl("chi\\^2", dist)),
-                             !(grepl("fdrreg", Method) & grepl("chi\\^2", dist)))
-res_nullset <- tidyr::complete(res_nullset, setting, Method, dist, key)
-
-null_tnr <- dplyr::filter(res_nullset, key == "TNR") %>%
-    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
+```{r}
+genplot <- function(tab, cov, xt = "", met = c("FDR", "TPR"), ebw = 0.0025) {
+    met <- match.arg(met)
+    cov <- rlang::enquo(cov)
+    gp <- dplyr::filter(tab, key == met) %>%
+        ggplot(aes(x = !!cov, y = mean.info, color = Method, group = Method)) +
         geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
-    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
+        geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = ebw, alpha=0.5) +
         scale_linetype_manual(values = lty) + 
         scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
-    scale_y_continuous("TNR", labels = scales::percent) +
-    ggtitle("Mean TNR Over 100 Replications") +
+        scale_y_continuous(met, labels = scales::percent) +
+        expand_limits(y = 0) +
+        ggtitle(paste("Mean", met)) +
         theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-null_fwer <- dplyr::filter(res_nullset, key == "FWER") %>%
-    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
-    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
-    scale_y_continuous("FWER", labels = scales::percent) +
-    ggtitle("FWER Over 100 Replications") +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-gp <- plot_grid(null_tnr + guides(color = FALSE, linetype = FALSE),
-                null_fwer + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
-                labels = LETTERS[1:2], ncol = 1)
-gp <- plot_grid(gp, get_legend(null_tnr), rel_widths = c(1, 1/16))
+        theme(axis.title = element_text(face = "bold", size = 10),
+              plot.title = element_text(face = "bold", size = 14))
+    if (met == "FDR") {
+        if (rlang::quo_name(cov) == "alpha") {
+            gp <- gp + geom_abline(lty = 2, color = "blue", alpha = 1/2) 
+        } else {
+            gp <- gp + geom_hline(yintercept = 0.05, lty = 2, color = "blue", alpha = 1/2) 
+        }
+    }
     gp
+}
 ```
 
-We also look at the difference between informative and uninformative covariates.
+# Figure 3
 
-```{r, fig.width = 18, fig.height = 8}
-null_tnr_diff <- dplyr::filter(res_nullset, key == "TNR") %>%
-    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
-    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
-    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
-    scale_y_continuous("TNR\n(informative - uninformative)", labels = scales::percent) +
-    ggtitle("Mean TNR Difference Over 100 Replications") +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
+This is a faceted figure with results from the pure simulation studies showing several key results from a subset of simulation settings.
+Namely, the figure includes plots of FDR and TPR values at a nominal FDR cutoff of 0.05 over 100 replications illustrating the impact of (A,B) changing total number of tests, (C,D) changing the proportion of non-null tests, and (E,F) changing the informativeness of a covariate.
 
-null_fwer_diff <- dplyr::filter(res_nullset, key == "FWER") %>%
-    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
-    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
-    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
-    scale_y_continuous("FWER\n(informative - uninformative)", labels = scales::percent) +
-    ggtitle("FWER Difference Over 100 Replications") +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
+```{r, fig.width = 10, fig.height = 12}
+## Figure 3A,B: FDR/TPR of methods across varying number of tests at alpha = 0.05
+res_ntests <- dplyr::filter(resmet, setting == "varyingntests", alpha == 0.05)
+p3a <- genplot(res_ntests, cov = as.factor(ntests), met = "FDR", ebw = 0.15) +
+    xlab("Number of Tests")
+p3b <- genplot(res_ntests, cov = as.factor(ntests), met = "TPR", ebw = 0.15) +
+    xlab("Number of Tests")
 
-gp <- plot_grid(null_tnr_diff + guides(color = FALSE, linetype = FALSE),
-                null_fwer_diff + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:2], ncol = 1)
-gp <- plot_grid(gp, get_legend(null_tnr_diff), rel_widths = c(1, 1/16))
-gp
+## Figure 3C,D: FDR/TPR of methods across varying null proportion at alpha = 0.05
+res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0", alpha == 0.05)
+p3c <- genplot(res_pi0, cov = pi0, met = "FDR", ebw = 2.5) +
+    scale_x_continuous("Proportion of Null Hypotheses",
+                       breaks = c(5, seq(10, 90, by = 10), 95, 99))
+p3d <- genplot(res_pi0, cov = pi0, met = "TPR", ebw = 2.5) +
+    scale_x_continuous("Proportion of Null Hypotheses",
+                       breaks = c(5, seq(10, 90, by = 10), 95, 99))
+
+## Figure 3E,F: FDR/TPR of methods across varying informativeness at alpha = 0.05
+res_info <- dplyr::filter(resmet, setting == "varyinginfo", alpha == 0.05)
+p3e <- genplot(res_info, cov = inform, met = "FDR", ebw = 2.5) +
+    scale_x_continuous("Informativeness", breaks = seq(0, 100, by = 20))
+p3f <- genplot(res_info, cov = inform, met = "TPR", ebw = 2.5) +
+    scale_x_continuous("Informativeness", breaks = seq(0, 100, by = 20))
+
+Fig3 <- plot_grid(p3a + guides(color = FALSE, linetype = FALSE),
+                p3b + guides(color = FALSE, linetype = FALSE),
+                p3c + guides(color = FALSE, linetype = FALSE),
+                p3d + guides(color = FALSE, linetype = FALSE),
+                p3e + guides(color = FALSE, linetype = FALSE),
+                p3f + guides(color = FALSE, linetype = FALSE),
+                labels = LETTERS[1:6], ncol = 2)
+Fig3 <- plot_grid(Fig3, get_legend(p3a), rel_widths = c(1, .2))
+Fig3
+ggsave(file.path(outdir, "Figure3.pdf"), width=10, height=12)
 ```
 
-# Sampling Distributions
+# Figure 3+
 
-We just take a look at different noise distributions for a single informative covariate.
+This is a faceted figure with results from the pure simulation studies showing the behavior of methods across different test statistic distributions.
+Test statistics were simulated from Normal, t with 11 degrees of freedom, t with 5 degrees of freedom, and Chi-squared with 4 degrees of freedom.
 
-```{r, fig.width = 18, fig.height = 8}
+```{r, fig.width = 10, fig.height = 12}
+## Figure 3plus A,B: FDR/TPR of methods across varying test statistics distribution
 res_noise <- dplyr::filter(resmet, setting == "informative-cubic")
-res_noise <- tidyr::complete(res_noise, setting, Method, dist, key)
+p3pa <- genplot(res_noise, cov = alpha, met = "FDR") +
+    facet_grid(. ~ dist, labeller = label_parsed)
+p3pb <- genplot(res_noise, cov = alpha, met = "TPR") +
+    facet_grid(. ~ dist, labeller = label_parsed)
 
-noise_fdr <- dplyr::filter(res_noise, key == "FDR") %>%
-    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
-    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
-    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
-    scale_y_continuous("FDR", labels = scales::percent) +
-    ggtitle("Mean FDR Over 100 Replications") +
-    expand_limits(y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-noise_tpr <- dplyr::filter(res_noise, key == "TPR") %>%
-    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
-    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
-    scale_y_continuous("TPR", labels = scales::percent) +
-    ggtitle("Mean TPR Over 100 Replications") +
-    expand_limits(y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-gp <- plot_grid(noise_fdr + guides(color = FALSE, linetype = FALSE),
-                noise_tpr + guides(color = FALSE, linetype = FALSE),
+Fig3p <- plot_grid(p3pa + guides(color = FALSE, linetype = FALSE),
+                   p3pb + guides(color = FALSE, linetype = FALSE),
                    labels = LETTERS[1:2], ncol = 1)
-gp <- plot_grid(gp, get_legend(noise_tpr), rel_widths = c(1, 1/16))
-gp
-
-gp <- plot_grid(noise_fdr + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
-                noise_tpr + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:2], ncol = 1)
-gp <- plot_grid(gp, get_legend(noise_tpr), rel_widths = c(1, 1/16))
-gp
+Fig3p <- plot_grid(Fig3p, get_legend(p3pa), rel_widths = c(7.5, 1))
+Fig3p
+ggsave(file.path(outdir, "Figure3plus.pdf"), width=14, height=8)
 ```
 
-We also look at the difference between informative and uninformative covariates.
+# Session information
 
-```{r, fig.width = 18, fig.height = 8}
-noise_fdr_diff <- dplyr::filter(res_noise, key == "FDR") %>%
-    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
-    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
-    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
-    scale_y_continuous("FDR\n(informative - uninformative)", labels = scales::percent) +
-    ggtitle("Mean FDR Difference Over 100 Replications") +
-    expand_limits(y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-noise_tpr_diff <- dplyr::filter(res_noise, key == "TPR") %>%
-    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
-    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
-    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
-    scale_y_continuous("TPR\n(informative - uninformative)", labels = scales::percent) +
-    ggtitle("Mean TPR Difference Over 100 Replications") +
-    expand_limits(y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-gp <- plot_grid(noise_fdr_diff + guides(color = FALSE, linetype = FALSE),
-                noise_tpr_diff + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:2], ncol = 1)
-gp <- plot_grid(gp, get_legend(noise_tpr_diff), rel_widths = c(1, 1/16))
-gp
+```{r}
+sessionInfo()
 ```
-
-# Informative Covariates
-
-We also take a look across different informative covariates.
-
-```{r, fig.width = 18, fig.height = 8}
-res_icov <- dplyr::filter(resmet, grepl("^informative", setting), grepl("N\\(0, 1\\)", dist))
-res_icov <- tidyr::complete(res_icov, setting, Method, dist, key)
-
-icov_fdr <- dplyr::filter(res_icov, key == "FDR") %>%
-    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
-    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
-    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
-    scale_y_continuous("FDR", labels = scales::percent) +
-    ggtitle("Mean FDR Over 100 Replications") +
-    expand_limits(y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-icov_tpr <- dplyr::filter(res_icov, key == "TPR") %>%
-    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
-    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
-    scale_y_continuous("TPR", labels = scales::percent) +
-    ggtitle("Mean TPR Over 100 Replications") +
-    expand_limits(y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-gp <- plot_grid(icov_fdr + guides(color = FALSE, linetype = FALSE),
-                icov_tpr + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:2], ncol = 1)
-gp <- plot_grid(gp, get_legend(icov_tpr), rel_widths = c(1, 1/12))
-gp
-```
-
-We also look at the difference between informative and uninformative covariates.
-
-```{r, fig.width = 18, fig.height = 8}
-icov_fdr_diff <- dplyr::filter(res_icov, key == "FDR") %>%
-    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
-    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
-    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
-    scale_y_continuous("FDR\n(informative - uninformative)", labels = scales::percent) +
-    ggtitle("Mean FDR Difference Over 100 Replications") +
-    expand_limits(y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-icov_tpr_diff <- dplyr::filter(res_icov, key == "TPR") %>%
-    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
-    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
-    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
-    scale_y_continuous("TPR\n(informative - uninformative)", labels = scales::percent) +
-    ggtitle("Mean TPR Difference Over 100 Replications") +
-    expand_limits(y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-gp <- plot_grid(icov_fdr_diff + guides(color = FALSE, linetype = FALSE),
-                icov_tpr_diff + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:2], ncol = 1)
-gp <- plot_grid(gp, get_legend(icov_tpr_diff), rel_widths = c(1, 1/12))
-gp
-```
-
-# Noninformative Covariates
-
-We take a look at noninformative covariates on their own.
-
-```{r, fig.width = 18, fig.height = 8}
-res_noni <- dplyr::filter(res, grepl("informative-cubic", setting))
-res_noni <- dplyr::group_by(res_noni, setting, dist, Method, key, performanceMetric, alpha)
-res_noni <- dplyr::summarize(res_noni,
-                             mean.uninfo = mean(value.uninfo, na.rm = TRUE),
-                             se.uninfo = sd(value.uninfo, na.rm = TRUE) /
-                                 sqrt(sum(!is.na(value.uninfo))))
-res_noni <- dplyr::ungroup(res_noni)
-
-noni_fdr <- dplyr::filter(res_noni, key == "FDR") %>%
-    ggplot(aes(x = alpha, y = mean.uninfo, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.uninfo - se.uninfo, ymax = mean.uninfo + se.uninfo), width = 0.0025, alpha=0.5) +
-    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
-    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
-    scale_y_continuous("FDR", labels = scales::percent) +
-    ggtitle("Mean FDR Over 100 Replications (noninformative covariates)") +
-    expand_limits(y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-noni_tpr <- dplyr::filter(res_noni, key == "TPR") %>%
-    ggplot(aes(x = alpha, y = mean.uninfo, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.uninfo - se.uninfo, ymax = mean.uninfo + se.uninfo), width = 0.0025, alpha=0.5) +
-    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
-    scale_y_continuous("TPR", labels = scales::percent) +
-    ggtitle("Mean TPR Over 100 Replications (noninformative covariates)") +
-    expand_limits(y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-gp <- plot_grid(noni_fdr + guides(color = FALSE, linetype = FALSE),
-                noni_tpr + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:2], ncol = 1)
-gp <- plot_grid(gp, get_legend(noni_tpr), rel_widths = c(1, 1/16))
-gp
-
-gp <- plot_grid(noni_fdr + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
-                noni_tpr + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:2], ncol = 1)
-gp <- plot_grid(gp, get_legend(noni_tpr), rel_widths = c(1, 1/16))
-gp
-```
-
-# Null Proportions
-
-We take a look at the impact of the proportion of null hypotheses.
-
-```{r, fig.width = 10, fig.height = 4}
-res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0-lowsignal",
-                         alpha == 0.05)
-res_pi0 <- dplyr::mutate(res_pi0,
-                         mean.info = ifelse(is.nan(mean.info) & key == "FDR", 0, mean.info))
-
-pi0_fdr <- dplyr::filter(res_pi0, key == "FDR") %>%
-    ggplot(aes(x = pi0, y = mean.info, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
-    geom_hline(aes(yintercept = alpha), lty = 2, color = "blue", alpha = 1/2) + 
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("pi0", breaks = seq(0, 100, 10)) +
-    scale_y_continuous("FDR", labels = scales::percent) +
-    ggtitle("Mean FDR at alpha = 0.05 Over 100 Replications") +
-    expand_limits(x =0, y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-pi0_tpr <- dplyr::filter(res_pi0, key == "TPR") %>%
-    ggplot(aes(x = pi0, y = mean.info, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("pi0", breaks = seq(0, 100, 10)) +
-    scale_y_continuous("TPR", labels = scales::percent) +
-    ggtitle("Mean TPR at alpha = 0.05 Over 100 Replications") +
-    expand_limits(x =0, y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-gp <- plot_grid(pi0_fdr + guides(color = FALSE, linetype = FALSE),
-                pi0_tpr + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:2], ncol = 2)
-gp <- plot_grid(gp, get_legend(pi0_tpr), rel_widths = c(1, 1/8))
-gp
-```
-
-# Informativeness
-
-We take a look at the impact of the informativeness of a covariate.
-
-```{r, fig.width = 10, fig.height = 4}
-res_info <- dplyr::filter(resmet, setting == "varyinginfo-lowsignal", alpha == 0.05)
-
-info_fdr <- dplyr::filter(res_info, key == "FDR") %>%
-    ggplot(aes(x = inform, y = mean.info, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
-    geom_hline(aes(yintercept = alpha), lty = 2, color = "blue", alpha = 1/2) + 
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
-    scale_y_continuous("FDR", labels = scales::percent) +
-    ggtitle("Mean FDR at alpha = 0.05 Over 100 Replications") +
-    expand_limits(x =0, y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-info_tpr <- dplyr::filter(res_info, key == "TPR") %>%
-    ggplot(aes(x = inform, y = mean.info, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
-    scale_y_continuous("TPR", labels = scales::percent) +
-    ggtitle("Mean TPR at alpha = 0.05 Over 100 Replications") +
-    expand_limits(x =0, y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-gp <- plot_grid(info_fdr + guides(color = FALSE, linetype = FALSE),
-                info_tpr + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:2], ncol = 2)
-gp <- plot_grid(gp, get_legend(info_tpr), rel_widths = c(1, 1/8))
-gp
-```
-
-We also look at the difference between informative and uninformative covariates.
-
-```{r, fig.width = 10, fig.height = 4}
-info_fdr_diff <- dplyr::filter(res_info, key == "FDR") %>%
-    ggplot(aes(x = inform, y = mean.diff, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 2.5, alpha=0.5) +
-    geom_hline(aes(yintercept = 0), lty = 2, color = "blue", alpha = 1/2) + 
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
-    scale_y_continuous("FDR", labels = scales::percent) +
-    ggtitle("Mean FDR Difference at alpha = 0.05 Over 100 Replications") +
-    expand_limits(x =0, y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-info_tpr_diff <- dplyr::filter(res_info, key == "TPR") %>%
-    ggplot(aes(x = inform, y = mean.diff, color = Method)) +
-    geom_line(aes(linetype = Method), alpha = 0.85) +
-    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 2.5, alpha=0.5) +
-    geom_hline(aes(yintercept = 0), lty = 2, color = "blue", alpha = 1/2) + 
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
-    scale_y_continuous("TPR", labels = scales::percent) +
-    ggtitle("Mean TPR Difference at alpha = 0.05 Over 100 Replications") +
-    expand_limits(x =0, y = 0) +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 14))
-
-gp <- plot_grid(info_fdr_diff + guides(color = FALSE, linetype = FALSE),
-                info_tpr_diff + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:2], ncol = 2)
-gp <- plot_grid(gp, get_legend(info_tpr_diff), rel_widths = c(1, 1/8))
-gp
-```
-
-# Figure Sim1
-
-We combine some of the results from above into a single figure.
-
-```{r, fig.width = 10, fig.height = 11}
-```

### 2018-08-30 manuscript_figures/manuscript_figures.Rmd
Add fig analogous to Figure 4 for impact with informative covariate.
Averages change in n rejects over all datasets / covariates within
a case study, also plots mean change in FDR and TPR in simulation.
Also harmonize heatmaps (order of columns, truncated color scales).

--- 2018-08-29 manuscript_figures/manuscript_figures.Rmd
@@
   mutate(cov = ifelse(casestudy == "scRNAseq" & inform != "uninf", extC,  cov)) %>%
-  mutate(cov = ifelse(casestudy == "ChIPseq" & inform != "uninf", "region-width",  cov)) %>%
+  mutate(cov = ifelse(casestudy == "ChIPseq" & inform != "uninf", 
+                      ifelse(grepl("cov", filename), "coverage", "region-width"),  cov)) %>%
   mutate(cov = ifelse(casestudy == "GSEA" & inform != "uninf", "set-size",  cov)) %>%
@@
 
+colOrder <- c("GWAS", "RNAseq", "ChIPseq", "GSEA", "microbiome", "scRNAseq")
 rowOrder <- c("ihw", "bl", "adapt-glm", "lfdr", "fdrreg-e", "fdrreg-t")
 
+# calculate percentage diff as percentage of the # of rejections found by 
+# using no covariate. If no rejections are found using no covariate, and
+# rejections are found with the covariate, set to 100%. 
 dftab <- left_join(inftab, uninftab, by = "id") %>%
   dplyr::mutate(diff_inf = nrejects - uninf) %>%
-  dplyr::mutate(diff_inf_percent = diff_inf / nrejects) %>%
+  dplyr::mutate(diff_inf_percent = diff_inf / uninf) %>%
+  dplyr::mutate(diff_inf_percent = ifelse(uninf==0 & nrejects > 0, 1, 
+                                          diff_inf_percent)) %>%
   dplyr::mutate(method = factor(method, levels = rowOrder)) %>%
+  dplyr::mutate(casestudy = factor(casestudy, levels = colOrder)) %>%
   dplyr::group_by(casestudy, cov) %>%
@@
 
+# if no change, set diff_inf_percent to zero, even if no rejections in either case
+dftab <- dftab %>% 
+  mutate(diff_inf_percent = ifelse(nrejects == uninf, 0, diff_inf_percent))
+
 Fig_impactofinfcov <- ggplot(dftab, aes(x = dataset, y = method, fill = diff_inf_percent)) + 
         geom_tile() + 
-        scale_fill_gradientn("% Increase\nin rejections",
-                             labels = scales::percent, trans="log1p",
+        geom_text(aes(label = ifelse(is.na(top), "", #adaptive text lab color
+                                     ifelse(top > 0.7, scales::percent(top), ""))), 
+                  color = "white", size = 2) +
+        geom_text(aes(label = ifelse(is.na(top), "", 
+                                     ifelse(top <= 0.7, scales::percent(top), ""))), 
+                  color = "black", size = 2) +
+        scale_fill_gradientn("% Change in\nrejections",
+                             labels = scales::percent, 
                              colours = c("darkblue","white","darkred"), 
                              values = scales::rescale(c(-.05,0,1)),
-                             guide = "colorbar", limits=c(-.05,1)) +
+                             guide = "colorbar", oob=scales::squish,
+                             limits=c(-.05,1)) +
         theme_bw() +
@@
         theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
-              legend.position = c(0.985, 0.93), legend.justification = c(1, 1),
+              legend.position = c(0.99, 0.94), legend.justification = c(1, 1),
               legend.background = element_rect(fill = scales::alpha("white", 3/4), color = 'black'),
@@
               legend.title = element_text(size = 9),
-              legend.margin = margin(6, 10, 6, 10)) +
-        geom_text(aes(label = ifelse(is.na(top), "", scales::percent(top))),
-                  color = "black", size = 2)
+              legend.margin = margin(6, 10, 6, 10))
 
 Fig_impactofinfcov
-ggsave(file.path(outdir, "Figure_impactofinfcov.pdf"), width = 15, height = 4.25)
+ggsave(file.path(outdir, "Figure_impactofinfcov.pdf"), width = 15.5, height = 4.25)
+
 ```
@@
   group_by(casestudy, method) %>%
-  summarize(diff_inf_percent = mean(diff_inf_percent, na.rm = TRUE))
+  summarize(ndatasets = sum(!is.na(diff_inf_percent)),
+            diff_inf_percent = mean(diff_inf_percent, na.rm = TRUE))
 
@@
   group_by(casestudy) %>%
-  summarize(top = max(diff_inf_percent, na.rm=TRUE))
+  summarize(top = max(diff_inf_percent, na.rm=TRUE),
+            maxd = max(ndatasets))
 
 dftab <- left_join(dftab, maxVals, by = c("casestudy")) %>%
-  mutate(top = ifelse(diff_inf_percent == top, diff_inf_percent, NA))
+  mutate(top = ifelse(diff_inf_percent == top, diff_inf_percent, NA)) %>%
+  ungroup() %>%
+  mutate(casestudy = factor(casestudy, levels = colOrder)) %>%
+  mutate(casestudyN = paste0(casestudy, "(", maxd, ")")) %>%
+  mutate(casestudyN = factor(casestudyN, 
+                             levels = unique(casestudyN)[unique(as.numeric(casestudy))]))
 
@@
         geom_tile() + 
-        scale_fill_gradientn("% Increase\nin rejections",
-                             labels = scales::percent, trans="log1p",
+        scale_fill_gradientn("% Change in\nrejections",
+                             labels = scales::percent, 
                              colours = c("darkblue","white","darkred"), 
                              values = scales::rescale(c(-.05,0,1)),
-                             guide = "colorbar", limits=c(-.05,1)) +
+                             guide = "colorbar", limits=c(-.05,1),
+                             oob=scales::squish) +
         theme_bw() +
@@
         ylab("Method") +
-        ggtitle("Impact of informative covariate on number of rejections") +
+        ggtitle("Impact of informative covariate") +
         theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
-        geom_text(aes(label = ifelse(is.na(top), "", scales::percent(top))),
-                  color = "black", size = 2)
+        geom_text(aes(label = ifelse(is.na(top), "", #adaptive text lab color
+                                     ifelse(top > 0.7, scales::percent(top), ""))), 
+                  color = "white", size = 2) +
+        geom_text(aes(label = ifelse(is.na(top), "", 
+                                     ifelse(top <= 0.7, scales::percent(top), ""))), 
+                  color = "black", size = 2)  +
+        theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
+              legend.position = c(0.945, 0.94), legend.justification = c(1, 1),
+              legend.background = element_rect(fill = scales::alpha("white", 3/4), color = 'black'),
+              legend.direction = "horizontal",
+              legend.text = element_text(size = 7),
+              legend.title = element_text(size = 9),
+              legend.margin = margin(6, 10, 6, 10))
 
 Fig_impactofinfcov
-ggsave(file.path(outdir, "Figure_impactofinfcov_avg.pdf"), width = 5.75, height = 4)
+ggsave(file.path(outdir, "Figure_impactofinfcov_avg.pdf"), width = 5, height = 4)
+
+###################################
+# Add average simulation results 
+###################################
+org <- "yeast-"
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+resfile <- objects[grepl(org, objects) ] 
+resfile <- resfile[grepl("5", resfile) & !grepl("null", resfile)]
+
+# regexp to construct master data table separating case study datasets, methods
+# and covariates
+obj.names <- str_split(resfile, pattern = "/")
+casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
+filename <- gsub(".rds", "", unlist(lapply(obj.names, function(x) x[4] )))
+dataset <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[3]]))
+inform <-  ifelse(grepl("uninf", resfile), "uninf", "inf")
+type <-  ifelse(grepl("null", resfile), "null", "de")
+size <-  ifelse(grepl("10", resfile), "10", "5")
+aseq <- alpha
+
+tab <- data.frame()
+for (l in seq_along(resfile)){
+  x <- readRDS(resfile[l])
+
+  tmp2 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
+                  fill =  "FDR", annotate = NULL, 
+                  alpha = aseq) %>%
+    mutate(fdr = nrejects) %>%
+    select(method, casestudy, replicate, fdr, alpha)
+
+  tmp1 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
+                    fill =  "TPR", annotate = NULL, 
+                    alpha = aseq) %>%
+    mutate(tpr = nrejects) %>%
+    select(method, casestudy, replicate, tpr, alpha)
+
+  tmp <- left_join(tmp1, tmp2, by = c("method", "casestudy", "replicate", "alpha")) %>%
+      left_join(tmp3, by = c("method", "casestudy", "replicate", "alpha")) %>%
+      mutate(inform = inform[l],
+             type = type[l],
+             size = size[l])
+
+  tab <- bind_rows(tab, tmp)
+}
+
+myspread <- function(df, key, value) {
+    # quote key
+    keyq <- rlang::enquo(key)
+    # break value vector into quotes
+    valueq <- rlang::enquo(value)
+    s <- rlang::quos(!!valueq)
+    df %>% gather(variable, value, !!!s) %>%
+        unite(temp, !!keyq, variable) %>%
+        spread(temp, value)
+}
+
+dftab <- tab %>%
+  select(-casestudy, -tnr) %>%
+  myspread(inform, c(tpr, fdr))
+
+dftab <- dftab %>%
+  mutate(tpr_diff = inf_tpr - uninf_tpr,
+         fdr_diff = inf_fdr - uninf_fdr) %>%
+  group_by(type, size, method, alpha) %>%
+  summarize(tpr_se = sd(tpr_diff)/n(),
+            fdr_se = sd(fdr_diff)/n(),
+            tpr_diff = mean(tpr_diff),
+            fdr_diff = mean(fdr_diff)) %>%
+  ungroup() %>%
+  filter(method %in% rowOrder) %>%
+  mutate(method = factor(method, levels = rowOrder)) %>%
+  mutate(top_fdr = ifelse(fdr_diff == max(fdr_diff), fdr_diff, NA)) %>%
+  mutate(top_tpr = ifelse(tpr_diff == max(tpr_diff), tpr_diff, NA)) %>%
+  mutate(size ="sim(100)")
+
+pfdr <- ggplot(dftab, aes(x = size, y = method, fill = fdr_diff)) + 
+        geom_tile() + 
+        geom_text(aes(label = ifelse(is.na(top_fdr), "",  scales::percent(top_fdr))), 
+                  color = "black", size = 2) +
+        scale_fill_gradientn("% Change\nin FDR",
+                             labels = scales::percent, 
+                             colours = c("darkblue","white","darkred"), 
+                             values = scales::rescale(c(-.01,0,.01)),
+                             guide = "colorbar", oob=scales::squish,
+                             limits=c(-.01,0.01)) +
+        xlab("") + ylab("") +
+        theme_bw() +
+        theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
+              legend.justification = c(1, 0.5),
+              legend.text = element_text(size = 7),
+              legend.title = element_text(size = 8),
+              legend.margin = margin(1, 1, 1, 1))
+
+ptpr <- ggplot(dftab, aes(x = size, y = method, fill = tpr_diff)) + 
+        geom_tile() + 
+        geom_text(aes(label = ifelse(is.na(top_tpr), "",  scales::percent(top_tpr))), 
+                  color = "black", size = 2) +
+        scale_fill_gradientn("% Change\nin TPR",
+                             labels = scales::percent, 
+                             colours = c("darkblue","white","darkred"), 
+                             values = scales::rescale(c(-.01,0,.1)),
+                             guide = "colorbar", oob=scales::squish,
+                             limits=c(-.01,0.1)) +
+        xlab("") + ylab("") +
+        theme_bw() +
+        theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
+              legend.justification = c(1, 0.5),
+              legend.text = element_text(size = 7),
+              legend.title = element_text(size = 8),
+              legend.margin = margin(1, 1, 1, 1))
+
+ggdraw() +
+  draw_plot(pfdr, 0.53, 0.025, 0.24, 0.93) +
+  draw_plot(ptpr, 0.77, 0.025, 0.24, 0.93) +
+  draw_plot(Fig_impactofinfcov, 0, 0, 0.52, 1) +
+  draw_plot_label(c("A", "B"), c(0, 0.53), c(1, 1), size = 15) +
+  draw_plot_label("Simulation", 0.71, 0.06, size =11, fontface = "plain")
+ggsave(file.path(outdir, "Figure_impactofinfcov_avg_sim.pdf"), width=9.5, height=5)
+
 ```

### 2018-08-30 manuscript_figures/manuscript_figures.Rmd
Add figure to compare pi0 settings in yeast

@@
 
+## Figure contrasting FDR results in different yeast simulation settings
+
+```{r Figure2, fig.width = 10, fig.height = 8}
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+resfile <- objects[(grepl("yeast-", objects) | grepl("yeastH-", objects)) & 
+                   grepl("de5", objects) & 
+                  !grepl("uninf", objects)] 
+labs <- rep("2000 Differentially expressed genes", 2)
+labs[grepl("H", resfile)] <- "500 Differentially expressed genes"
+excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
+
+plist <- vector("list", length(resfile))
+for (l in seq_along(resfile)){
+
+  de5 <- readRDS(file=resfile[l])
+  res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
+
+  # Fdr by alpha
+  p1 <- plotsim_average(res5d, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) +
+                ggtitle("")
+
+  # tpr by alpha
+  p2 <- plotsim_average(res5d, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) + 
+                theme(legend.position = "none") + 
+                ggtitle("")
+
+  if (l == 1){
+    leg <- get_legend(p1)
+  }
+  p1 <- p1 + theme(legend.position = "none")
+
+  plist[[l]] <- plot_grid(p1, p2, ncol = 2, 
+                   labels = c("", labs[l]), label_x = -0.65 )
+}
+
+# combine rows into one figure
+Fig <- plot_grid(plotlist = plist, nrow = 2, 
+                 labels = LETTERS[1:2], label_size = 20)
+Fig <- plot_grid(Fig, leg, ncol = 2, rel_widths = c(1, 0.2))
+Fig
+ggsave(file.path(outdir, "Figure_yeast_varyingpi0.pdf"), width=10, height=8)
+
+```
+
 # Session information

### 2018-08-30 manuscript_figures/manuscript_figures_sims.Rmd
Change subtitles in Figure 3

@@
         expand_limits(y = 0) +
-        ggtitle(paste("Mean", met)) +
         theme_classic() +
@@
 
+## add title to 3A,B
+p3ab <- plot_grid(p3a + guides(color = FALSE, linetype = FALSE),
+                  p3b + guides(color = FALSE, linetype = FALSE),
+                  labels = LETTERS[1:2], ncol = 2)
+p3abTitle <- ggdraw() +
+    draw_label("Simulation performance across varying numbers of hypotheses",
+               fontface = 'bold')
+p3ab <- plot_grid(p3abTitle, p3ab, ncol = 1, rel_heights = c(0.1, 1))
+
 ## Figure 3C,D: FDR/TPR of methods across varying null proportion at alpha = 0.05
@@
 
+## add title to 3C,D
+p3cd <- plot_grid(p3c + guides(color = FALSE, linetype = FALSE),
+                  p3d + guides(color = FALSE, linetype = FALSE),
+                  labels = LETTERS[3:4], ncol = 2)
+p3cdTitle <- ggdraw() +
+    draw_label("Simulation performance across varying null proportion",
+               fontface = 'bold')
+p3cd <- plot_grid(p3cdTitle, p3cd, ncol = 1, rel_heights = c(0.1, 1))
+
 ## Figure 3E,F: FDR/TPR of methods across varying informativeness at alpha = 0.05
@@
 
-Fig3 <- plot_grid(p3a + guides(color = FALSE, linetype = FALSE),
-                p3b + guides(color = FALSE, linetype = FALSE),
-                p3c + guides(color = FALSE, linetype = FALSE),
-                p3d + guides(color = FALSE, linetype = FALSE),
-                p3e + guides(color = FALSE, linetype = FALSE),
+## add title to 3E,F
+p3ef <- plot_grid(p3e + guides(color = FALSE, linetype = FALSE),
                   p3f + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:6], ncol = 2)
+                  labels = LETTERS[5:6], ncol = 2)
+p3efTitle <- ggdraw() +
+    draw_label("Simulation performance across varying informativeness",
+               fontface = 'bold')
+p3ef <- plot_grid(p3efTitle, p3ef, ncol = 1, rel_heights = c(0.1, 1))
+
+## pull Figure 3 together
+Fig3 <- plot_grid(p3ab, p3cd, p3ef, ncol = 1)
 Fig3 <- plot_grid(Fig3, get_legend(p3a), rel_widths = c(1, .2))
@@
 res_noise <- dplyr::filter(resmet, setting == "informative-cubic")
+levels(res_noise$dist) <- gsub("Noise:", "Distribution:", levels(res_noise$dist))
 p3pa <- genplot(res_noise, cov = alpha, met = "FDR") +
@@
 
+## create title for 3plus
+p3pTitle <- ggdraw() +
+    draw_label("Simulation performance across test statistic distributions",
+               fontface = 'bold', size = 20)
+
+## pull Figure 3plus together
 Fig3p <- plot_grid(p3pa + guides(color = FALSE, linetype = FALSE),
@@
                    labels = LETTERS[1:2], ncol = 1)
+Fig3p <- plot_grid(p3pTitle, Fig3p, ncol = 1, rel_heights = c(0.1, 1))
 Fig3p <- plot_grid(Fig3p, get_legend(p3pa), rel_widths = c(7.5, 1))

### 2018-08-31 manuscript_figures/manuscript_figures_sims.Rmd
Modify Figure3 axis labels

@@
 
-```{r Figure2-load-data}
+```{r load-workspace}
 ## Load packages and source benchmark FDR
@@
 
-```{r}
+```{r load-sim-metrics}
 resmet <- readRDS(resmet_file)
@@
 
-```{r}
+```{r genplot-function}
 genplot <- function(tab, cov, xt = "", met = c("FDR", "TPR"), ebw = 0.0025) {
@@
 
-```{r, fig.width = 10, fig.height = 12}
+```{r Figure3, fig.width = 10, fig.height = 12}
 ## Figure 3A,B: FDR/TPR of methods across varying number of tests at alpha = 0.05
 res_ntests <- dplyr::filter(resmet, setting == "varyingntests", alpha == 0.05)
-p3a <- genplot(res_ntests, cov = as.factor(ntests), met = "FDR", ebw = 0.15) +
-    xlab("Number of Tests")
-p3b <- genplot(res_ntests, cov = as.factor(ntests), met = "TPR", ebw = 0.15) +
-    xlab("Number of Tests")
+p3a <- genplot(res_ntests, cov = ntests, met = "FDR", ebw = 0.05) +
+    scale_x_continuous("Number of Tests", trans = "log10",
+                       breaks = c(1e2, 5e2, 1e3, 5e3, 1e4, 5e4))
+p3b <- genplot(res_ntests, cov = ntests, met = "TPR", ebw = 0.05) +
+    scale_x_continuous("Number of Tests", trans = "log10",
+                       breaks = c(1e2, 5e2, 1e3, 5e3, 1e4, 5e4))
 
@@
 res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0", alpha == 0.05)
-p3c <- genplot(res_pi0, cov = pi0, met = "FDR", ebw = 2.5) +
-    scale_x_continuous("Proportion of Null Hypotheses",
-                       breaks = c(5, seq(10, 90, by = 10), 95, 99))
-p3d <- genplot(res_pi0, cov = pi0, met = "TPR", ebw = 2.5) +
-    scale_x_continuous("Proportion of Null Hypotheses",
-                       breaks = c(5, seq(10, 90, by = 10), 95, 99))
+p3c <- genplot(res_pi0, cov = 100-pi0, met = "FDR", ebw = 2.5) +
+    scale_x_continuous("Proportion of Non-Null Hypotheses",
+                       breaks = 100 - c(5, seq(10, 90, by = 10), 95, 99))
+p3d <- genplot(res_pi0, cov = 100-pi0, met = "TPR", ebw = 2.5) +
+    scale_x_continuous("Proportion of Non-Null Hypotheses",
+                       breaks = 100 - c(5, seq(10, 90, by = 10), 95, 99))
 
@@
 p3cdTitle <- ggdraw() +
-    draw_label("Simulation performance across varying null proportion",
+    draw_label("Simulation performance across varying non-null proportion",
                fontface = 'bold')
@@
 
-```{r, fig.width = 10, fig.height = 12}
+```{r Figure3plus, fig.width = 10, fig.height = 12}
 ## Figure 3plus A,B: FDR/TPR of methods across varying test statistics distribution

### 2018-08-31 datasets/simulations/simulations-varyinginformativeness-smooth.Rmd
Plot informativeness functions; try smooth informativeness

--- 2018-08-09 datasets/simulations/simulations-varyinginformativeness.Rmd
@@
 ---
-title: "Simulation Study: Varying Informativeness"
+title: "Simulation Study: Varying Informativeness (smooth)"
 author: "Patrick Kimes"
@@
 covariate is sampled uniformly from the interval [0, 1], and the conditional
-probability of a test being non-null is a step function of the covariate with
-two levels (over the intervals [0, 4/5], (4/5, 1]) where the size of the step
-is used to indicate the informativeness of the covariate.
+probability of a test being non-null is a smooth function of the covariate taking
+values between 0 and 1 based on the logistic function.
 
@@
 A tuning parameter is used to modify the informativeness of the covariate
-such that 0 corresponds to an uninformative covariate, and 1 corresponds to a
-fully informative covariate (i.e. all tests with covariate values in [0, 4/5]
-are null and all tests in (4/5, 1] are non-null).
+such that 0 corresponds to a completely uninformative covariate, and increasing
+"informativeness" corresponds to more non-null tests concentrating in a smaller
+range of the covariate. The functional relationship between the covariate and
+the null probability over varying informativeness is shown below.
 
@@
 
+First, we show the form of the relationship between the informative covariate and null
+proportion, pi0, across varying levels of "informativeness."
+
+```{r plot-informativeness, fig.width = 9, fig.height = 2.5}
+xseq <- seq(0, 1, by = .01)
+pi0trends <- lapply(seq(0, 1, by = .2), function(i) {
+    tibble(info = i, x = xseq,
+           pi0 = pi0_varyinfo80l(i)(xseq))
+})
+pi0trends <- bind_rows(pi0trends)
+
+ggplot(pi0trends, aes(x = x, y = pi0)) +
+    geom_line() +
+    geom_text(aes(label = round(pi0, 3), hjust = x, vjust = round(pi0)),
+              dplyr::filter(pi0trends, x == 0 | x == 1),
+              size = 3, color = 'blue') + 
+    theme_bw() +
+    scale_x_continuous("x (informative covariate)", breaks = 0:1) +
+    scale_y_continuous(labels = scales::percent) +
+    facet_grid(. ~ info, labeller = label_both) +
+    ggtitle("Covariate vs. pi0 for varying informativeness")
+```
+
 # Informativeness = 0 Setting
@@
 ```{r info00-parameters}
-pi0 <- pi0_varyinfo80(0.00)             # numeric: proportion of null hypotheses
+pi0 <- pi0_varyinfo80l(0.00)             # numeric: proportion of null hypotheses
 seed <- 1010
@@
 ```{r info20-parameters}
-pi0 <- pi0_varyinfo80(0.20)             # numeric: proportion of null hypotheses
+pi0 <- pi0_varyinfo80l(0.20)             # numeric: proportion of null hypotheses
 seed <- 1201
@@
 ```{r info40-parameters}
-pi0 <- pi0_varyinfo80(0.40)             # numeric: proportion of null hypotheses
+pi0 <- pi0_varyinfo80l(0.40)             # numeric: proportion of null hypotheses
 seed <- 51
@@
 ```{r info60-parameters}
-pi0 <- pi0_varyinfo80(0.60)             # numeric: proportion of null hypotheses
+pi0 <- pi0_varyinfo80l(0.60)             # numeric: proportion of null hypotheses
 seed <- 500
@@
 ```{r info80-parameters}
-pi0 <- pi0_varyinfo80(0.80)             # numeric: proportion of null hypotheses
+pi0 <- pi0_varyinfo80l(0.80)             # numeric: proportion of null hypotheses
 seed <- 608
@@
 ```{r info100-parameters}
-pi0 <- pi0_varyinfo80(1.00)             # numeric: proportion of null hypotheses
+pi0 <- pi0_varyinfo80l(1.00)             # numeric: proportion of null hypotheses
 seed <- 808

### 2018-08-31 datasets/R/helpers-simulations.R

@@ -184,6 +184,13 @@ pi0_varyinfo80 <- function(zz) {
     function(x) { ifelse(x > 4/5, 0.8 * (1 - zz), 0.2 * (4 + zz)) }
 }
 
+## logistic function to test different levels of "informativeness" with pi0 = 80%
+pi0_varyinfo80l <- function(zz) {
+    stopifnot(zz <= 1L, zz >= 0L)
+    function(x) { zz / (1+exp(5-25*x)) + (1 - zz) * 0.8 }
+}
+
+
 ## ##############################################################################
 ## CODE DERIVED FROM EXTERNAL SOURCE
 ## source: https://github.com/stephenslab/ash/blob/94a3347/code/dsc-shrink/datamakers/datamaker.R

### 2018-08-31 datasets/simulations/simulations-varyinginformativeness-smooth.Rmd
Fix varying info output file name

@@
 ## intermediary files we create below
-info00_file <- file.path(resdir, "varyinginfo-benchmark-level00.rds")
-info20_file <- file.path(resdir, "varyinginfo-benchmark-level20.rds")
-info40_file <- file.path(resdir, "varyinginfo-benchmark-level40.rds")
-info60_file <- file.path(resdir, "varyinginfo-benchmark-level60.rds")
-info80_file <- file.path(resdir, "varyinginfo-benchmark-level80.rds")
-info100_file <- file.path(resdir, "varyinginfo-benchmark-level100.rds")
+info00_file <- file.path(resdir, "varyinginfo-smooth-benchmark-level00.rds")
+info20_file <- file.path(resdir, "varyinginfo-smooth-benchmark-level20.rds")
+info40_file <- file.path(resdir, "varyinginfo-smooth-benchmark-level40.rds")
+info60_file <- file.path(resdir, "varyinginfo-smooth-benchmark-level60.rds")
+info80_file <- file.path(resdir, "varyinginfo-smooth-benchmark-level80.rds")
+info100_file <- file.path(resdir, "varyinginfo-smooth-benchmark-level100.rds")
 

### 2018-09-04 datasets/simulations/simulations-informative-cosine-nullAdaPT.Rmd
Add supplementary analyses of AdaPT with null option

--- 2018-08-06 datasets/simulations/simulations-informative-cosine.Rmd
@@
 ---
-title: "Simulation Study: Informative Covariate (cosine)"
+title: "Simulation Study: AdaPT with null model option (cosine covariate)" 
 author: "Patrick Kimes"
@@
 again with Gaussian, t-distributed, and Chi-Squared distributed noise.
+
+This set of simulations differs from `simulations-informative-cosine.Rmd` only
+in the implementation of the AdaPT method for multiple testing correction. This
+is a follow up analysis similar to that described and performed in
+`simulations-informative-step-nullAdaPT.Rmd`. For more details on the parameter settings
+of the two additional AdaPT methods included in the comparison, please check the
+other `-nullAdaPT.Rmd` analysis.
+
+The analysis in `simulations-informative-step-nullAdaPT.Rmd` was carried out
+to determine whether the additional `adapt-withnull` and `adapt-onlynull` methods
+could provide an improvement over the `adapt-glm` parameter setting in the case
+of low informativeness in the independent covariate. This analysis aims to examine
+how the `adapt-withnull` and `adalpt-onlynull` parameter settings perform in
+comparison to the default `adapt-glm` settings when the covariate is more
+informative.
 
@@
 ## intermediary files we create below
-gauss_file <- file.path(resdir, "informative-cosine-benchmark-gaussian.rds")
-tdist_file <- file.path(resdir, "informative-cosine-benchmark-t5.rds")
-tdist11_file <- file.path(resdir, "informative-cosine-benchmark-t11.rds")
-chisq_file <- file.path(resdir, "informative-cosine-benchmark-chisq4.rds")
+gauss_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-gaussian.rds")
+tdist_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-t5.rds")
+tdist11_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-t11.rds")
+chisq_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-chisq4.rds")
 
@@
                      control = list(lambda = 0.01))
+```
+
+We add a couple alternative calls to AdaPT to test whether the method performance improves in cases
+of low "informativeness" if given the option to not use the informative covariate. This option is
+coded by the null model, `"pi0 ~ 1"`, where only an intercept term is included in the model. Specifically,
+we add two alternative calls to the benchmark - one call with the null model added as one of the candidate
+models to `pi_formulas=` and `mu_formulas=`, and a second call with the null model given as the only
+candidate model to `pi_formulas=` and `mu_formulas=`.
+
+```{r}
+bdplus <- addBMethod(bdplus, "adapt-withnull",
+                     adaptMT::adapt_glm,
+                     function(x) { x$qvals },
+                     pvals = pval,
+                     x = data.frame(icov = ind_covariate),
+                     pi_formulas = c("~ 1", paste0("splines::ns(icov, df = ", seq(2, 10, 2), ")")),
+                     mu_formulas = c("~ 1", paste0("splines::ns(icov, df = ", seq(2, 10, 2), ")")),
+                     alphas = 0)
+
+bdplus <- addBMethod(bdplus, "adapt-onlynull",
+                     adaptMT::adapt_glm,
+                     function(x) { x$qvals },
+                     pvals = pval,
+                     x = data.frame(icov = ind_covariate),
+                     pi_formulas = "~ 1",
+                     mu_formulas = "~ 1",
+                     alphas = 0)
+
+candycols2 <- rbind(candycols,
+                    data.frame(Method = c("adapt-withnull", "adapt-onlynull"),
+                               col = c("darkgoldenrod3", "darkgoldenrod3"),
+                               lty = c("dashed", "dotted"),
+                               stringsAsFactors = FALSE))
 ```
@@
 ```{r gauss-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
+onerun <- simIteration(1, bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 plotsim_average(resdf, met="rejections", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE, 
+                palette = candycols2)
 
 plotsim_average(resdf, met="FDR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE, 
+                palette = candycols2)
 
 plotsim_average(resdf, met="TPR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE, 
+                palette = candycols2)
 
 plotsim_average(resdf, met="TNR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE) 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
+```
+
+Since we are primarily interested in comparing just the AdaPT methods, we plot these
+metrics excluding all other methods.
+
+```{r gauss-metrics-averages-adapt, results = "hide"}
+plotsim_average(resdf, met="rejections",
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdf, met="FDR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdf, met="TPR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdf, met="TNR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 ```
@@
 ```{r gauss-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size",
+                  palette = candycols2)
 
-covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate",
+                  palette = candycols2)
 ```
 
-Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
-the overlap of rejections between methods.
+We also look at the FDR as a function of the independent covariate.
 
-```{r gauss-metrics-upset, results = "hide"}
-if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
-    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
-} else {
-    message("Not enough methods found rejections at alpha ", ualpha, 
-            "; skipping upset plot")
-}
+```{r gauss-metrics-covlineplotFDR, results = "hide"}
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR",
+                  palette = candycols2)
 ```
@@
 plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
+```
+
+Again, we recreate the plots after subsetting on AdaPT methods.
+
+```{r gauss-metrics-differences-adapt, results = "hide"}
+plotsim_average(resdfiu, met="rejections", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdfiu, met="FDR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdfiu, met="TPR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdfiu, met="TNR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
 ```
@@
 ```{r t5-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
+onerun <- simIteration(1, bdplust, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 plotsim_average(resdf, met="rejections", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdf, met="FDR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdf, met="TPR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdf, met="TNR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE) 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
+```
+
+Since we are primarily interested in comparing just the AdaPT methods, we plot these
+metrics excluding all other methods.
+
+```{r t5-metrics-averages-adapt, results = "hide"}
+plotsim_average(resdf, met="rejections",
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdf, met="FDR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdf, met="TPR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdf, met="TNR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 ```
@@
 ```{r t5-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size",
+                  palette = candycols2)
 
-covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate",
+                  palette = candycols2)
 ```
 
-Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
-the overlap of rejections between methods.
+We also look at the FDR as a function of the independent covariate.
 
-```{r t5-metrics-upset, results = "hide"}
-if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
-    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
-} else {
-    message("Not enough methods found rejections at alpha ", ualpha, 
-            "; skipping upset plot")
-}
+```{r t5-metrics-covlineplotFDR, results = "hide"}
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR",
+                  palette = candycols2)
 ```
@@
 plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
+```
+
+Again, we recreate the plots after subsetting on AdaPT methods.
+
+```{r t5-metrics-differences-adapt, results = "hide"}
+plotsim_average(resdfiu, met="rejections", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdfiu, met="FDR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdfiu, met="TPR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdfiu, met="TNR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
 ```
@@
 
-Next, we consider a second setting with t-distributed noise.
+Next, we consider the setting with t-distributed noise.
 
@@
 ```{r t11-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
+onerun <- simIteration(1, bdplust, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 plotsim_average(resdf, met="rejections", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdf, met="FDR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdf, met="TPR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdf, met="TNR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE) 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
+```
+
+Since we are primarily interested in comparing just the AdaPT methods, we plot these
+metrics excluding all other methods.
+
+```{r t11-metrics-averages-adapt, results = "hide"}
+plotsim_average(resdf, met="rejections",
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdf, met="FDR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdf, met="TPR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdf, met="TNR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 ```
@@
 ```{r t11-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
+covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size",
+                  palette = candycols2)
 
-covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate",
+                  palette = candycols2)
 ```
 
-Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
-the overlap of rejections between methods.
+We also look at the FDR as a function of the independent covariate.
 
-```{r t11-metrics-upset, results = "hide"}
-if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
-    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
-} else {
-    message("Not enough methods found rejections at alpha ", ualpha, 
-            "; skipping upset plot")
-}
+```{r t11-metrics-covlineplotFDR, results = "hide"}
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR",
+                  palette = candycols2)
 ```
@@
 plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
+```
+
+Again, we recreate the plots after subsetting on AdaPT methods.
+
+```{r t11-metrics-differences-adapt, results = "hide"}
+plotsim_average(resdfiu, met="rejections", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdfiu, met="FDR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdfiu, met="TPR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdfiu, met="TNR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
 ```
@@
 ```{r}
-bdchi <- modifyBMethod(bd, "ashq", mode = "empirical")
+bdchi <- modifyBMethod(bdplus, "ashq", mode = "empirical")
+```
+
+We also drop the FDRreg methods from the chi-squared setting.
+
+```{r}
+bdchi <- dropBMethod(bdchi, "fdrreg-t")
+bdchi <- dropBMethod(bdchi, "fdrreg-e")
 ```
@@
 ```{r chisq4-one-simulation}
-onerun <- simIteration(bdchi, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
+onerun <- simIteration(1, bdchi, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                        icovariate = icovariate, null_dist = null_dist, execute = FALSE)
@@
 
-Since `ashq` is not appropriate when the test statistics are not approximately normal or t,
-we also plot the results without this method.
+Since we are primarily interested in comparing just the AdaPT methods, we plot these
+metrics excluding all other methods.
 
-```{r chisq4-metrics-averages-noash, results = "hide"}
-plotsim_average(resdf, met = "rejections", filter_set = c(excludeSet, "ashq"),
-                merge_ihw = TRUE, errorBars = TRUE)
+```{r chisq4-metrics-averages-adapt, results = "hide"}
+plotsim_average(resdf, met="rejections",
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 
-plotsim_average(resdf, met = "FDR", filter_set = c(excludeSet, "ashq"),
-                merge_ihw = TRUE, errorBars = TRUE)
+plotsim_average(resdf, met="FDR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 
-plotsim_average(resdf, met = "TPR", filter_set = c(excludeSet, "ashq"),
-                merge_ihw = TRUE, errorBars = TRUE)
+plotsim_average(resdf, met="TPR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 
-plotsim_average(resdf, met = "TNR", filter_set = c(excludeSet, "ashq"),
-                merge_ihw = TRUE, errorBars = TRUE) 
+plotsim_average(resdf, met="TNR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 ```
@@
 
-Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
-the overlap of rejections between methods. Again, without `ashq`.
+We also look at the FDR as a function of the independent covariate.
 
-```{r chisq4-metrics-upset, results = "hide"}
-if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = c(excludeSet, "ashq")) >= 3) {
-    aggupset(lapply(res_i, function(x) { x[, -which(colnames(x) == "ashq")] }),
-             alpha = ualpha, supplementary = FALSE, return_list = FALSE)
-} else {
-    message("Not enough methods found rejections at alpha ", ualpha, 
-            "; skipping upset plot")
-}
+```{r chisq4-metrics-covlineplotFDR, results = "hide"}
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR")
 ```
@@
 
+Again, we recreate the plots after subsetting on AdaPT methods.
+
+```{r chisq4-metrics-differences-adapt, results = "hide"}
+plotsim_average(resdfiu, met="rejections", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdfiu, met="FDR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdfiu, met="TPR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
+
+plotsim_average(resdfiu, met="TNR", 
+                filter_set = grep("adapt", unique(resdf$blabel),
+                                  invert = TRUE, value = TRUE), 
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
+```
+
 # Session Info

### 2018-09-04 datasets/simulations/simulations-informative-step-nullAdaPT.Rmd
Add supplementary analyses of AdaPT with null option

--- 2018-09-04 datasets/simulations/simulations-informative-cosine-nullAdaPT.Rmd
@@
 ---
-title: "Simulation Study: AdaPT with null model option (cosine covariate)" 
+title: "Simulation Study: AdaPT with null model option (step covariate)" 
 author: "Patrick Kimes"
@@
 the interval [0, 1], and the conditional probability of a test being non-null
-is a smooth (cosine) function of the covariate. We include simulation results
-again with Gaussian, t-distributed, and Chi-Squared distributed noise.
+is a non-smooth (monotone) step function of the covariate. We include simulation
+results again with Gaussian, t-distributed, and Chi-Squared distributed noise.
 
-This set of simulations differs from `simulations-informative-cosine.Rmd` only
-in the implementation of the AdaPT method for multiple testing correction. This
-is a follow up analysis similar to that described and performed in
-`simulations-informative-step-nullAdaPT.Rmd`. For more details on the parameter settings
-of the two additional AdaPT methods included in the comparison, please check the
-other `-nullAdaPT.Rmd` analysis.
+This set of simulations differs from `simulations-informative-step.Rmd` only
+in the implementation of the AdaPT method for multiple testing correction.
+In the primary set of simulation results (`simulations-informative-step.Rmd`),
+we observed that the AdaPT method can lose power when the informative covariate
+is only weakly informative. While the method still controls the FDR, we observe
+that the loss in power is substantial, with AdaPT achieving a lower mean TPR
+across 100 replications when using a weakly informative covariate rather than a
+completely uninformative (random) covariate.
 
-The analysis in `simulations-informative-step-nullAdaPT.Rmd` was carried out
-to determine whether the additional `adapt-withnull` and `adapt-onlynull` methods
-could provide an improvement over the `adapt-glm` parameter setting in the case
-of low informativeness in the independent covariate. This analysis aims to examine
-how the `adapt-withnull` and `adalpt-onlynull` parameter settings perform in
-comparison to the default `adapt-glm` settings when the covariate is more
-informative.
+In this set of simulations, we assess whether this loss of power with a weakly
+informative covariate can be mitigated by allowing the AdaPT model selection
+procedure to decide between the same collection of spline models used before, and
+an intercept-only (null) model, which makes no use of the covariate. This approach
+was suggested by the authors of the AdaPT method through personal communication.
+
+Two additional implementations of the AdaPT model are included in this analysis:
+`adapt-withnull`, the approach described above, and `adapt-onlynull`, where the method
+is forced to ignore the informative covariate. The latter (`onlynull`) implementation
+provides a baseline for comparing the two other AdaPT parameter settings.
 
@@
 ## intermediary files we create below
-gauss_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-gaussian.rds")
-tdist_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-t5.rds")
-tdist11_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-t11.rds")
-chisq_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-chisq4.rds")
+gauss_file <- file.path(resdir, "nullAdaPT-informative-step-benchmark-gaussian.rds")
+tdist_file <- file.path(resdir, "nullAdaPT-informative-step-benchmark-t5.rds")
+tdist11_file <- file.path(resdir, "nullAdaPT-informative-step-benchmark-t11.rds")
+chisq_file <- file.path(resdir, "nullAdaPT-informative-step-benchmark-chisq4.rds")
 
@@
 m <- 20000                        # integer: number of hypothesis tests
-pi0 <- pi0_cosine(0.90)           # numeric: proportion of null hypotheses
+pi0 <- pi0_step(0.90)             # numeric: proportion of null hypotheses
 icovariate <- runif               # functional: independent covariate

### 2018-09-04 datasets/simulations/simulations-informative-cubic-vash.Rmd
Add supplementary analyses of ASH with VASH shrinkage

--- 2018-08-09 datasets/simulations/simulations-informative-cubic.Rmd
@@
 ---
-title: "Simulation Study: Informative Covariate (cubic)"
+title: "Simulation Study: ASH with VASH shrinkage estimation"
 author: "Patrick Kimes"
@@
 is a smooth (cubic) function of the covariate. We include simulation results
-again with Gaussian, t-distributed, and Chi-Squared distributed noise.
+with only t-distributed noise.
+
+In primary analyses (`simulations-informative-cubic.Rmd` and others), we observed
+that the ASH method was anti-conservative when the test statistic was t-distributed,
+with the effect size following an approximate normal distribution, and standard
+errors sampled from a chi-squared distribution. This was the case even when specifying
+the `df=` argument of the `ash` method.
+
+In personal communications with the authors of the `ashr` package, it was suggested
+that this was possibly attributable to the approximation used when computing
+t-likelihoods, where the randomness of the observed standard errors are not properly
+accounted for. The authors suggested a solution to this problem based on passing
+shrunken standard error estimates and moderated degrees of freedom as input to ASH,
+rather than the observed standard errors and degrees of freedom. The `ashq` method
+in this analysis is run with shrunken standard errors and moderated degrees of
+freedom estimated using the `vashr` package available at the authors' GitHub site
+([`mengyin/vashr`](https://github.com/mengyin/vashr)). Since the focus of the analysis
+is on the ASH method which does not make use of an auxiliary informative covariate,
+the comparison of performance using informative and uninformative covariates is
+excluded from this write-up. Additionally, since the VASH approach is only necessary
+for t-distributed settings, we exclude the Gaussian and Chi-Squared settings as
+well.
 
@@
 ## intermediary files we create below
-gauss_file <- file.path(resdir, "informative-cubic-benchmark-gaussian.rds")
-tdist_file <- file.path(resdir, "informative-cubic-benchmark-t5.rds")
-tdist11_file <- file.path(resdir, "informative-cubic-benchmark-t11.rds")
-chisq_file <- file.path(resdir, "informative-cubic-benchmark-chisq4.rds")
+tdist_file <- file.path(resdir, "vashr-informative-cubic-benchmark-t5.rds")
+tdist11_file <- file.path(resdir, "vashr-informative-cubic-benchmark-t11.rds")
 
@@
 
-# Gaussian Setting
-
-First, we consider the setting with Gaussian noise.
-
-## Data Simulation
-
-```{r gauss-parameters}
-es_dist <- rnorm_generator(3)       # functional: dist of alternative test stats
-ts_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
-null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
-seed <- 608
-```
-
-We next run the simulations (including Scott's FDR Regression).
-
-```{r gauss-run-simulation}
-if (file.exists(gauss_file)) {
-    res <- readRDS(gauss_file)
-} else {
-    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
-                    pi0 = pi0, es_dist = es_dist, icovariate = icovariate,
-                    ts_dist = ts_dist, null_dist = null_dist,
-                    seed = seed, mc.cores = cores)
-    saveRDS(res, file = gauss_file)
-}
-res_i <- lapply(res, `[[`, "informative")
-res_u <- lapply(res, `[[`, "uninformative")
-```
-
-## Covariate Diagnostics
-
-Here, we show the relationship between the independent covariate and p-values for a
-single replication of the experiment.
-
-```{r gauss-one-simulation}
-onerun <- simIteration(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
-                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
-```
-
-```{r, gauss-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
-rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
-```
-
-```{r, gauss-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
-```
-
-## Benchmark Metrics
-
-We plot the averaged results across `r B` replications.
-
-```{r gauss-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(resdf, met="rejections", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
-
-plotsim_average(resdf, met="FDR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
-
-plotsim_average(resdf, met="TPR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
-
-plotsim_average(resdf, met="TNR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE) 
-```
-
-We also take a look at the distribution of rejects for each method as a function of
-the effect size and independent covariate.
-
-```{r gauss-metrics-covlineplot, results = "hide"}
-covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size")
-
-covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate")
-```
-
-We also look at the FDR as a function of the independent covariate.
-
-```{r gauss-metrics-covlineplotFDR, results = "hide"}
-covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR")
-```
-
-Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
-the overlap of rejections between methods.
-
-```{r gauss-metrics-upset, results = "hide"}
-if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
-    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
-} else {
-    message("Not enough methods found rejections at alpha ", ualpha, 
-            "; skipping upset plot")
-}
-```
-
-We also compare the simulation results with and without an informative covariate.
-
-```{r gauss-metrics-differences, results = "hide"}
-resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
-
-resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
-                                   performanceMetric, alpha, value),
-                            select(resdfu, rep, blabel, param.alpha, key,
-                                   performanceMetric, alpha, value),
-                            by = c("rep", "blabel", "param.alpha", "key",
-                                   "performanceMetric", "alpha"),
-                            suffix = c(".info", ".uninfo"))
-resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
-
-plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
-
-plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
-
-plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
-
-plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
-```
-
 # Student's t Setting (df = 5)
 
-Next, we consider the setting with t-distributed noise.
+We consider the setting with t-distributed noise.
 
@@
 ```{r}
-bdplust <- modifyBMethod(bdplus, "ashq", df = 5)
+library(vashr)
+
+bdplust <- modifyBMethod(bdplus, "ashq",
+                         df = 2 * vashr::vash(SE, df = 5, singlecomp = TRUE)$PosteriorShape[1],
+                         sebetahat = vashr::vash(SE, df = 5, singlecomp = TRUE)$sd.post)
 ```
@@
 res_i <- lapply(res, `[[`, "informative")
-res_u <- lapply(res, `[[`, "uninformative")
 ```
@@
 
-We also compare the simulation results with and without an informative covariate.
-
-```{r t5-metrics-differences, results = "hide"}
-resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
-
-resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
-                                   performanceMetric, alpha, value),
-                            select(resdfu, rep, blabel, param.alpha, key,
-                                   performanceMetric, alpha, value),
-                            by = c("rep", "blabel", "param.alpha", "key",
-                                   "performanceMetric", "alpha"),
-                            suffix = c(".info", ".uninfo"))
-resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
-
-plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
-
-plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
-
-plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
-
-plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
-```
-
 # Student's t Setting (df = 11)
@@
 ```{r}
-bdplust <- modifyBMethod(bdplus, "ashq", df = 11)
+bdplust <- modifyBMethod(bdplus, "ashq",
+                         df = 2 * vashr::vash(SE, df = 11, singlecomp = TRUE)$PosteriorShape[1],
+                         sebetahat = vashr::vash(SE, df = 11, singlecomp = TRUE)$sd.post)
 ```
@@
 
-We also compare the simulation results with and without an informative covariate.
-
-```{r t11-metrics-differences, results = "hide"}
-resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
-
-resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
-                                   performanceMetric, alpha, value),
-                            select(resdfu, rep, blabel, param.alpha, key,
-                                   performanceMetric, alpha, value),
-                            by = c("rep", "blabel", "param.alpha", "key",
-                                   "performanceMetric", "alpha"),
-                            suffix = c(".info", ".uninfo"))
-resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
-
-plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
-
-plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
-
-plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
-
-plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
-```
-
-# Chi-Squared Setting
-
-Finally, we consider the setting with chi-squared distributed noise.
-
-## Data Simulation
-
-```{r chisq4-parameters}
-es_dist <- rnorm_generator(15)       # functional: dist of alternative test stats
-ts_dist <- rchisq_perturber(4)  # functional: sampling dist/noise for test stats
-null_dist <- rchisq_pvaluer(4)     # functional: dist to calc p-values
-seed <- 1023
-```
-
-For the chi-squared distributed noise setting, we must change the "mode" setting for ASH
-from the default of 0 to `"estimate"` since the mode of null and alternative test statistics
-are no longer centered at 0. While both approximate Normality and unimodality of effects are
-violated in this simulation setting, by allowing the mode to be estimated, rather than forced
-to 0, should return more comparable results.
-
-```{r}
-bdchi <- modifyBMethod(bd, "ashq", mode = "empirical")
-```
-
-We next run the simulations. We do not include FDR Regression because the test statistics
-are not approximately normally distributed.
-
-```{r chisq4-run-simulation}
-if (file.exists(chisq_file)) {
-    res <- readRDS(chisq_file)
-} else {
-    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdchi, m = m,
-                    pi0 = pi0, es_dist = es_dist, icovariate = icovariate,
-                    ts_dist = ts_dist, null_dist = null_dist,
-                    seed = seed, mc.cores = cores)
-    saveRDS(res, file = chisq_file)
-}
-res_i <- lapply(res, `[[`, "informative")
-res_u <- lapply(res, `[[`, "uninformative")
-```
-
-## Covariate Diagnostics
-
-Here, we show the relationship between the independent covariate and p-values for a
-single replication of the experiment.
-
-```{r chisq4-one-simulation}
-onerun <- simIteration(bdchi, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
-                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
-```
-
-```{r, chisq4-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
-rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
-```
-
-```{r, chisq4-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
-strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
-```
-
-## Benchmark Metrics
-
-We plot the averaged results across `r B` replications.
-
-```{r chisq4-metrics-averages, results = "hide"}
-resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(resdf, met = "rejections", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
-
-plotsim_average(resdf, met = "FDR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
-
-plotsim_average(resdf, met = "TPR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
-
-plotsim_average(resdf, met = "TNR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE) 
-```
-
-Since `ashq` is not appropriate when the test statistics are not approximately normal or t,
-we also plot the results without this method.
-
-```{r chisq4-metrics-averages-noash, results = "hide"}
-plotsim_average(resdf, met = "rejections", filter_set = c(excludeSet, "ashq"),
-                merge_ihw = TRUE, errorBars = TRUE)
-
-plotsim_average(resdf, met = "FDR", filter_set = c(excludeSet, "ashq"),
-                merge_ihw = TRUE, errorBars = TRUE)
-
-plotsim_average(resdf, met = "TPR", filter_set = c(excludeSet, "ashq"),
-                merge_ihw = TRUE, errorBars = TRUE)
-
-plotsim_average(resdf, met = "TNR", filter_set = c(excludeSet, "ashq"),
-                merge_ihw = TRUE, errorBars = TRUE) 
-```
-
-We also take a look at the distribution of rejects for each method as a function of
-the effect size and independent covariate. Again, we plot these without `ashq`.
-
-```{r chisq4-metrics-covlineplot, results = "hide"}
-covariateLinePlot(lapply(res_i, function(x) { x[, -which(colnames(x) == "ashq")] }),
-                  alpha = ualpha, covname = "effect_size")
-
-covariateLinePlot(lapply(res_i, function(x) { x[, -which(colnames(x) == "ashq")] }),
-                  alpha = ualpha, covname = "ind_covariate")
-```
-
-We also look at the FDR as a function of the independent covariate.
-
-```{r chisq4-metrics-covlineplotFDR, results = "hide"}
-covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR")
-```
-
-Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
-the overlap of rejections between methods. Again, without `ashq`.
-
-```{r chisq4-metrics-upset, results = "hide"}
-if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = c(excludeSet, "ashq")) >= 3) {
-    aggupset(lapply(res_i, function(x) { x[, -which(colnames(x) == "ashq")] }),
-             alpha = ualpha, supplementary = FALSE, return_list = FALSE)
-} else {
-    message("Not enough methods found rejections at alpha ", ualpha, 
-            "; skipping upset plot")
-}
-```
-
-We also compare the simulation results with and without an informative covariate.
-
-```{r chisq4-metrics-differences, results = "hide"}
-resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))
-
-resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
-                                   performanceMetric, alpha, value),
-                            select(resdfu, rep, blabel, param.alpha, key,
-                                   performanceMetric, alpha, value),
-                            by = c("rep", "blabel", "param.alpha", "key",
-                                   "performanceMetric", "alpha"),
-                            suffix = c(".info", ".uninfo"))
-resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)
-
-plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
-
-plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
-
-plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
-
-plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
-```
-
 # Session Info

### 2018-09-04 datasets/simulations/simulations-informative-cosine-nullAdaPT.Rmd
Change wording from 'noise' to 'test stat dist' in writeups

@@
 is a smooth (cosine) function of the covariate. We include simulation results
-again with Gaussian, t-distributed, and Chi-Squared distributed noise.
+again with Gaussian, t-distributed, and Chi-Squared distributed test statistics.
 
@@
 As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
-for simulations with Gaussian or t-distributed noise. Again, we include both
+for simulations with Gaussian or t-distributed test statistics. Again, we include both
 `nulltype = "empirical"` and `nulltype = "theoretical"`.
@@
 
-First, we consider the setting with Gaussian noise.
+First, we consider the setting with Gaussian test statistics.
 
@@
 
-Next, we consider the setting with t-distributed noise.
+Next, we consider the setting with t-distributed test statistics.
 
@@
 
-For the t-distributed noise setting, we must specify the number of degrees of freedom
+For the t-distributed setting, we must specify the number of degrees of freedom
 for ASH. We add an additional parameter to the `ashq` method with the corresponding
-degrees of freedom of the noise distribution.
+degrees of freedom of the test statistic distribution.
 
@@
 
-Next, we consider the setting with t-distributed noise.
+Next, we consider a second setting with t-distributed test statistics.
 
@@
 
-Finally, we consider the setting with chi-squared distributed noise.
+Finally, we consider the setting with chi-squared distributed test statistics.
 
@@
 
-For the chi-squared distributed noise setting, we must change the "mode" setting for ASH
+For the chi-squared distributed setting, we must change the "mode" setting for ASH
 from the default of 0 to `"estimate"` since the mode of null and alternative test statistics

### 2018-09-04 datasets/simulations/simulations-informative-cosine.Rmd
Change wording from 'noise' to 'test stat dist' in writeups

@@
 
-Next, we consider a second setting with t-distributed noise.
+Next, we consider a second setting with t-distributed test statistics.
 

### 2018-09-04 datasets/simulations/simulations-informative-cubic.Rmd
Change wording from 'noise' to 'test stat dist' in writeups

@@
 is a smooth (cubic) function of the covariate. We include simulation results
-again with Gaussian, t-distributed, and Chi-Squared distributed noise.
+again with Gaussian, t-distributed, and Chi-Squared distributed test statistics.
 

### 2018-09-04 datasets/simulations/simulations-informative-sine.Rmd
Change wording from 'noise' to 'test stat dist' in writeups

@@
 is a smooth (sine) function of the covariate. We include simulation results
-again with Gaussian, t-distributed, and Chi-Squared distributed noise.
+again with Gaussian, t-distributed, and Chi-Squared distributed test statistics.
 

### 2018-09-04 datasets/simulations/simulations-informative-step-nullAdaPT.Rmd
Change wording from 'noise' to 'test stat dist' in writeups

@@
 is a non-smooth (monotone) step function of the covariate. We include simulation
-results again with Gaussian, t-distributed, and Chi-Squared distributed noise.
+results again with Gaussian, t-distributed, and Chi-Squared distributed test statistics.
 

### 2018-09-04 datasets/simulations/simulations-informative-step.Rmd
Change wording from 'noise' to 'test stat dist' in writeups

@@
 is a non-smooth (monotone) step function of the covariate. We include simulation results
-again with Gaussian, t-distributed, and Chi-Squared distributed noise.
+again with Gaussian, t-distributed, and Chi-Squared distributed test statistics.
 

### 2018-09-04 datasets/simulations/simulations-null.Rmd
Change wording from 'noise' to 'test stat dist' in writeups

@@
 
-First, we consider the null setting with gaussian noise.
+First, we consider the null setting with gaussian test statistics.
 
@@
 
-Next, we consider the null setting with t-distributed noise.
+Next, we consider the null setting with t-distributed test statistics.
 
@@
 
-Finally, we consider the null setting with chi-squared distributed noise.
+Finally, we consider the null setting with chi-squared distributed test statistics.
 

### 2018-09-04 datasets/simulations/simulations-varyingeffectsize-chisq.Rmd
Change wording from 'noise' to 'test stat dist' in writeups

@@
 ---
-title: "Simulation Study: Varying Effect Size"
+title: "Simulation Study: Varying Effect Size (chisq-4)"
 author: "Patrick Kimes"
@@
 presented in `simulations-varyingeffectsize.Rmd`, however, Chi-Squared
-distributed noise is used to perturb effects. Additionally, since the effect
+distributed test statistics are used. Additionally, since the effect
 must be non-negative (to be a valid value for the non-centrality parameter

### 2018-09-04 datasets/simulations/simulations-varyingeffectsize-t.Rmd
Change wording from 'noise' to 'test stat dist' in writeups

@@
 ---
-title: "Simulation Study: Varying Effect Size (t noise)"
+title: "Simulation Study: Varying Effect Size (t-11)"
 author: "Patrick Kimes"
@@
 `simulations-informative-cubic.Rmd`. This set of simulations is similar to those
-presented in `simulations-varyingeffectsize.Rmd`, however, t-distributed noise
-is used to perturb effects.
+presented in `simulations-varyingeffectsize.Rmd`, however, t-distributed test
+statistics are used.
 
@@
 As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
-for simulations with Gaussian or t-distributed noise. Again, we include both
+for simulations with Gaussian or t-distributed test statistics. Again, we include both
 `nulltype = "empirical"` and `nulltype = "theoretical"`. Since all settings in this
-series of simulations use test statistics simulated with Gaussian noise, we include
-Scott's FDR Regression in all of the comparisons.
+series of simulations use test statistics simulated with t-distributed test statistics,
+we include Scott's FDR Regression in all of the comparisons.
 
@@
 
-Since all simulation settings in this case study use t-distributed noise, we must specify
+Since all simulation settings in this case study use t-distributed test statistics, we must specify
 the number of degrees of freedom for ASH. We add an additional parameter to the `ashq`
-method with the corresponding degrees of freedom of the noise distribution.
+method with the corresponding degrees of freedom of the test statistic distribution.
 

### 2018-09-04 datasets/simulations/simulations-varyinginformativeness-smooth.Rmd
Change wording from 'noise' to 'test stat dist' in writeups

@@
 As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
-for simulations with Gaussian noise. Again, we include both
+for simulations with Gaussian test statistics. Again, we include both
 `nulltype = "empirical"` and `nulltype = "theoretical"`. Since all settings in this
-series of simulations use test statistics simulated with Gaussian noise, we include
+series of simulations use test statistics simulated with Gaussian test statistics, we include
 Scott's FDR Regression in all of the comparisons.

### 2018-09-04 datasets/simulations/simulations-varyinginformativeness.Rmd
Change wording from 'noise' to 'test stat dist' in writeups

@@
 ualpha <- 0.05
+```
+
+First, we show the form of the relationship between the informative covariate and null
+proportion, pi0, across varying levels of "informativeness."
+
+```{r plot-informativeness, fig.width = 9, fig.height = 2.5}
+xseq <- seq(0, 1, by = .01)
+pi0trends <- lapply(seq(0, 1, by = .2), function(i) {
+    tibble(info = i, x = xseq,
+           pi0 = pi0_varyinfo80(i)(xseq))
+})
+pi0trends <- bind_rows(pi0trends)
+
+ggplot(pi0trends, aes(x = x, y = pi0)) +
+    geom_line() +
+    geom_text(aes(label = round(pi0, 3), hjust = x, vjust = round(pi0)),
+              dplyr::filter(pi0trends, x == 0 | x == 1),
+              size = 3, color = 'blue') + 
+    theme_bw() +
+    scale_x_continuous("x (informative covariate)", breaks = 0:1) +
+    scale_y_continuous(labels = scales::percent) +
+    facet_grid(. ~ info, labeller = label_both) +
+    ggtitle("Covariate vs. pi0 for varying informativeness")
+```
+
+Alternatively, since the function is a step function, we can simply plot the levels
+of the two steps.
+
+```{r, plot-informativeness-v2, fig.width = 6, fig.height = 5}
+xseq <- seq(0, 1, by = .01)
+pi0trends <- lapply(xseq, function(i) {
+    tibble(info = i, x = 0:1,
+           pi0 = pi0_varyinfo80(i)(0:1))
+})
+pi0trends <- bind_rows(pi0trends)
+pi0trends <- dplyr::mutate(pi0trends,
+                           x = factor(x, levels = 0:1,
+                                      labels = c("[0, .8]", "(.8, 1]")))
+
+ggplot(pi0trends, aes(x = info, y = pi0, group = x)) +
+    geom_line() +
+    geom_text(aes(label = x, vjust = round(pi0)),
+              dplyr::filter(pi0trends, info == 1),
+              hjust = 1, size = 3, color = 'blue') + 
+    theme_bw() +
+    scale_x_continuous("Informativeness", labels = scales::percent) +
+    scale_y_continuous(labels = scales::percent) +
+    ggtitle("null proportions for varying informativeness")
 ```

### 2018-09-04 manuscript_figures/manuscript_figures.Rmd
Add several panels toFigure S1

--- 2018-08-30 manuscript_figures/manuscript_figures.Rmd
@@
 
+## Figure contrasting FDR results in different yeast simulation settings
+
+```{r Figure-yeast-pi0, fig.width = 10, fig.height = 8}
+
+# strongly informative covariate
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+resfile <- objects[(grepl("yeast-", objects) | grepl("yeastH-", objects) | 
+                    grepl("yeastII-", objects) | grepl("yeastIIH-", objects)) & 
+                   grepl("de5", objects) & 
+                  !grepl("uninf", objects)] 
+labs <- rep("30% Non-null", length(resfile))
+labs[grepl("H", resfile)] <- "7.5% Non-null"
+labs2 <- rep("Bimodal effect size", length(resfile))
+labs2[grepl("II", resfile)] <- "Unimodal effect size"
+excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
+
+plist <- vector("list", length(resfile))
+tab <- NULL
+for (l in seq_along(resfile)){
+
+  de5 <- readRDS(file=resfile[l])
+  res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
+  res5d$es_dist <- labs2[l]
+  res5d$pi0 <- labs[l]
+  tab <- rbind(tab, res5d)
+
+}
+
+FigStrong <- plotsim_average(tab, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE, grpVars = c("es_dist", "pi0")) +
+                ggtitle("") +
+  facet_grid( ~ es_dist + pi0) +
+  ylim(0,0.3)
+
+FigStrong
+
+# weakly informative covariate
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+resfile <- objects[(grepl("yeastU-", objects) | grepl("yeastHU-", objects) | 
+                    grepl("yeastIIU-", objects) | grepl("yeastIIHU-", objects)) & 
+                   grepl("de5", objects) & 
+                  !grepl("uninf", objects)] 
+labs <- rep("30% Non-null", length(resfile))
+labs[grepl("H", resfile)] <- "7.5% Non-null"
+labs2 <- rep("Bimodal effect size", length(resfile))
+labs2[grepl("II", resfile)] <- "Unimodal effect size"
+excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
+
+plist <- vector("list", length(resfile))
+tab <- NULL
+for (l in seq_along(resfile)){
+
+  de5 <- readRDS(file=resfile[l])
+  res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
+  res5d$es_dist <- labs2[l]
+  res5d$pi0 <- labs[l]
+  tab <- rbind(tab, res5d)
+
+}
+
+FigWeak <- plotsim_average(tab, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE, grpVars = c("es_dist", "pi0")) +
+                ggtitle("") +
+  ylim(0,0.3) +
+  facet_grid( ~ es_dist + pi0)
+
+FigWeak
+
+plot_grid(FigStrong, FigWeak, nrow = 2, 
+          labels = paste0(LETTERS[1:2], "    "
+                          c("Strongly informative covariate",
+                            "Weakly informative covariate")))
+ggsave(file.path(outdir, "Figure_yeast_varyingpi0.pdf"), width=10, height=8)
+
+```
+
 # Session information

### 2018-09-04 datasets/R/plotsim_average.R

@@ -28,6 +28,7 @@
 #'        those that use more than just the p-value (default = FALSE)
 #' @param diffplot logical indicating whether 'value' in tsb is a difference
 #'        between informative and uninformative covariates. (default = FALSE)
+#' @param grpVars vector of character names of additional columns of tsb to keep 
 #'
 #' @return
 #' a ggplot object.
@@ -36,14 +37,24 @@
 plotsim_average <- function(tsb, met, filter_set = NULL, merge_ihw = TRUE,
                             clean_names = FALSE, errorBars=FALSE,
                             palette = candycols, facetMethodType = FALSE,
-                            diffplot = FALSE) {
+                            diffplot = FALSE, grpVars = NULL){
 
     ## cacluate mean per replication
-    tsba <- tsb %>%
+    if(!is.null(grpVars)){
+      tsba <- tsb %>%
+        group_by(blabel, performanceMetric, alpha, param.alpha, key, 
+                 .dots = grpVars) %>%
+        summarize(n = sum(!is.na(value)),
+                  se = sd(value, na.rm = TRUE) / sqrt(n),
+                  value = mean(value, na.rm = TRUE))
+    }else{
+      tsba <- tsb %>%
         group_by(blabel, performanceMetric, alpha, param.alpha, key) %>%
         summarize(n = sum(!is.na(value)),
                   se = sd(value, na.rm = TRUE) / sqrt(n),
                   value = mean(value, na.rm = TRUE))
+   }
+   
    tsba$value[tsba$n == 0] <- NA
    tsba$se[tsba$n == 0] <- NA
     

### 2018-09-05 manuscript_figures/manuscript_figures.Rmd
Adjust labels and legend of Figure S1

--- 2018-08-30 manuscript_figures/manuscript_figures.Rmd
@@
 
+## Figure contrasting FDR results in different yeast simulation settings
+
+```{r Figure-yeast-pi0, fig.width = 10, fig.height = 8}
+# strongly informative covariate (top panel)
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+resfile <- objects[(grepl("yeast-", objects) | grepl("yeastH-", objects) | 
+                    grepl("yeastII-", objects) | grepl("yeastIIH-", objects)) & 
+                   grepl("de5", objects) & 
+                  !grepl("uninf", objects)] 
+labs <- rep("30% Non-null", length(resfile))
+labs[grepl("H", resfile)] <- "7.5% Non-null"
+labs2 <- rep("Unimodal effect size", length(resfile))
+labs2[grepl("II", resfile)] <- "Bimodal effect size"
+excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
+
+plist <- vector("list", length(resfile))
+tab <- NULL
+for (l in seq_along(resfile)){
+
+  de5 <- readRDS(file=resfile[l])
+  res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
+  res5d$es_dist <- labs2[l]
+  res5d$pi0 <- labs[l]
+  tab <- rbind(tab, res5d)
+}
+
+FigStrong <- plotsim_average(tab, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE, grpVars = c("es_dist", "pi0")) +
+                ggtitle("") +
+  facet_grid( ~ es_dist + pi0) +
+  ylim(0,0.3) 
+FigStrong
+
+# weakly informative covariate (bottom panel)
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+resfile <- objects[(grepl("yeastU-", objects) | grepl("yeastHU-", objects) | 
+                    grepl("yeastIIU-", objects) | grepl("yeastIIHU-", objects)) & 
+                   grepl("de5", objects) & 
+                  !grepl("uninf", objects)] 
+labs <- rep("30% Non-null", length(resfile))
+labs[grepl("H", resfile)] <- "7.5% Non-null"
+labs2 <- rep("Unimodal effect size", length(resfile))
+labs2[grepl("II", resfile)] <- "Bimodal effect size"
+excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
+
+plist <- vector("list", length(resfile))
+tab <- NULL
+for (l in seq_along(resfile)){
+
+  de5 <- readRDS(file=resfile[l])
+  res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
+  res5d$es_dist <- labs2[l]
+  res5d$pi0 <- labs[l]
+  tab <- rbind(tab, res5d)
+}
+
+FigWeak <- plotsim_average(tab, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE, grpVars = c("es_dist", "pi0")) +
+                ggtitle("") +
+  ylim(0,0.3) +
+  facet_grid( ~ es_dist + pi0)
+FigWeak
+
+leg <- get_legend(FigStrong)
+Fig <- plot_grid(FigStrong + theme(legend.position="none") +  
+                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)), 
+                 FigWeak + theme(legend.position="none") +  
+                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)),
+                 nrow = 2, 
+                 labels = paste0(LETTERS[1:2], "       ",
+                          c("Strongly informative covariate",
+                            "Weakly informative covariate")),
+                 hjust = 0)
+plot_grid(Fig, leg, ncol=2, rel_widths = c(1,0.1))
+ggsave(file.path(outdir, "Figure_yeast_varyingpi0.pdf"), width=10, height=6.5)
+
+```
+
 # Session information

### 2018-09-05 datasets/simulations/simulations-informative-cosine-nullAdaPT.Rmd
Add missing candycols2 specto nullAdaPT sim writeups

@@
 plotsim_average(resdf, met = "rejections", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE),
+                palette = candycols2)
 
 plotsim_average(resdf, met = "FDR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdf, met = "TPR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdf, met = "TNR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE) 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 ```
@@
 covariateLinePlot(lapply(res_i, function(x) { x[, -which(colnames(x) == "ashq")] }),
-                  alpha = ualpha, covname = "effect_size")
+                  alpha = ualpha, covname = "effect_size",
+                  palette = candycols2)
 
 covariateLinePlot(lapply(res_i, function(x) { x[, -which(colnames(x) == "ashq")] }),
-                  alpha = ualpha, covname = "ind_covariate")
+                  alpha = ualpha, covname = "ind_covariate",
+                  palette = candycols2)
 ```
@@
 ```{r chisq4-metrics-covlineplotFDR, results = "hide"}
-covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR")
+covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR",
+                  palette = candycols2)
 ```
@@
 plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
+                palette = candycols2)
 ```

### 2018-09-05 datasets/simulations/simulations-informative-step-nullAdaPT.Rmd
Add missing candycols2 specto nullAdaPT sim writeups

@@
 plotsim_average(resdf, met = "rejections", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdf, met = "FDR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdf, met = "TPR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE)
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 
 plotsim_average(resdf, met = "TNR", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE) 
+                merge_ihw = TRUE, errorBars = TRUE,
+                palette = candycols2)
 ```

### 2018-09-05 manuscript_figures/manuscript_figures.Rmd
Clean up labels, titles, and panel alignment in Figures 2 and S11

@@
                 merge_ihw = TRUE, errorBars=TRUE) +
-      theme(legend.position="none")
+      theme(legend.position="none") +  
+      ggtitle("Mean FDR (100 replications)") +
+      scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02))
 
@@
                 merge_ihw = TRUE, errorBars=TRUE) +
-      theme(legend.position="none")
+      theme(legend.position="none")  +  
+      ggtitle("Mean FDR (100 replications)") +
+      scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02))
 
@@
                 merge_ihw = TRUE, errorBars=TRUE) +
-      theme(legend.position="none")
+      theme(legend.position="none") +
+      ylab("Number of Rejections") +
+      ggtitle("Mean FDR (100 replications)") +
+      scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02))
 
@@
 leg <- get_legend(p4)
-p4 <- p4 + theme(legend.position="none") + 
-  ggtitle("Mean % significant by effect size %ile") + 
-  theme(axis.title = element_text(face="bold", size = 10),
-        plot.title = element_text(face="bold", size = 14)) +
-  xlab("Effect size (log2 fold change) percentile")
+p4 <- p4 +
+  ylab("Rejection rate") +
+  xlab("Effect size (log2 fold change) percentile") +
+  ggtitle("Rejection rate by effect size") + 
+  theme(legend.position="none")
 
 # Figure 2E - upsetr plot - save separately
-pdf(file.path(outdir, "Figure2E.pdf"), width = 10, height = 4, onefile=FALSE)
+pdf(file.path(outdir, "Figure2E.pdf"), width = 8, height = 4, onefile=FALSE)
 aggupset(de5, alpha=alpha, supplementary = FALSE, 
                        return_list = FALSE, nintersects = 12)
-grid.text("Mean overlap Over 100 Replications", x = 0.6, y = 0.97, 
-          gp = gpar(fontsize = 16, fontface = "bold"))
-grid.text("E", x = 0.02, y = 0.97, 
-          gp = gpar(fontsize = 24, fontface = "bold")) 
+grid.text("Mean overlap Over 100 Replications", x = 0.65, y = 0.97, 
+          gp = gpar(fontsize = 14, fontface = "bold"))
+grid.text("E", x = 0.01, y = 0.97, 
+          gp = gpar(fontsize = 22, fontface = "bold")) 
 dev.off()
@@
 Fig2 <- plot_grid(p1, p2, p3, p4, ncol = 2, 
-          labels = LETTERS[1:4], label_size = 20)
+          labels = LETTERS[1:4], label_size = 20, align = "hv")
 Fig2 <- plot_grid(Fig2, leg, ncol = 2, rel_widths = c(1, 0.2))
 Fig2
-ggsave(file.path(outdir, "Figure2A-D.pdf"), width=10, height=8)
+ggsave(file.path(outdir, "Figure2A-D.pdf"), width=8.5, height=6)
 
@@
                   size = 12) +
-  draw_plot_label("Overall mean expression pvalue", size = 11, 0.5, 0.14, fontface = "plain",
+  draw_plot_label("p-value by Overall mean expression group", size = 11, 0.5, 0.14, fontface = "plain",
                   hjust =0.5) +
-  draw_plot_label("Density", size = 11, 0, 0.36, fontface = "plain", angle=90)
+  draw_plot_label("Density", size = 11, 0, 0.3, fontface = "plain", angle=90)
 
@@
                   size = 12 ) +
-  draw_plot_label("Detection rate pvalue", size = 11, 0.5, 0.14, fontface = "plain",
+  draw_plot_label("p-value by Detection rate group", size = 11, 0.5, 0.14, fontface = "plain",
                   hjust =0.5) +
-  draw_plot_label("Density", size = 11, 0, 0.36, fontface = "plain", angle=90)
+  draw_plot_label("Density", size = 11, 0, 0.3, fontface = "plain", angle=90)
 
@@
   draw_plot_label("  GSEA case study: Mouse", 0.5, 1, hjust = 0.5, fontface = "bold" , size = 12) +
-  draw_plot_label("Gene Set size pvalue", size = 11, 0.5, 0.14, fontface = "plain",
+  draw_plot_label("p-value by Gene Set size group", size = 11, 0.5, 0.14, fontface = "plain",
                   hjust =0.5) +
-  draw_plot_label("Density", size = 11, 0, 0.36, fontface = "plain", angle=90)
+  draw_plot_label("Density", size = 11, 0, 0.3, fontface = "plain", angle=90)
 
@@
   draw_plot_label("GWAS case study", 0.5, 1, fontface = "bold", hjust = 0.5, size = 12) +
-  draw_plot_label("Minor Allele Frequency (MAF) pvalue", size = 11, 0.5, 0.14, fontface = "plain", 
+  draw_plot_label("p-value by Minor Allele Frequency (MAF) group", size = 11, 0.5, 0.14, fontface = "plain", 
                   hjust = 0.5) +
-  draw_plot_label("Density", size = 11, 0, 0.36, fontface = "plain", angle=90)
-
-p6 <- plot_grid(pa, pb, pc, pd, ncol=1, labels = LETTERS[1:4])
+  draw_plot_label("Density", size = 11, 0, 0.3, fontface = "plain", angle=90)
+
+p6 <- plot_grid(ggdraw(add_sub(pa, " ", size = 2)), 
+                ggdraw(add_sub(pb, " ", size = 2)), 
+                ggdraw(add_sub(pc, " ", size = 2)), 
+                ggdraw(add_sub(pd, " ", size = 2)), 
+                ncol=1, labels = LETTERS[1:4])
 p6
-ggsave(file.path(outdir, "Figure6.pdf"), width=7, height=7)
+ggsave(file.path(outdir, "Figure6.pdf"), width=7, height=8)
 

### 2018-09-05 datasets/R/effectSizePlot.R

@@ -173,7 +173,7 @@ covariateLinePlot <- function(sbl, alpha=0.05, nbins = 25,
   names(lty) <- as.character(df$Method)
 
   if (metric == "discoveries") {
-      ystr <- "% Significant"
+      ystr <- "Percent rejected"
   } else {
       ystr <- metric
   }
@@ -185,7 +185,10 @@ covariateLinePlot <- function(sbl, alpha=0.05, nbins = 25,
       scale_x_continuous(labels = scales::percent) +
       xlab(paste0(covname, " percentile")) +
       scale_color_manual(values = col) +
-      scale_linetype_manual(values = lty) 
+      scale_linetype_manual(values = lty) + 
+      theme_classic() + 
+      theme(axis.title = element_text(face="bold"),
+            plot.title = element_text(face="bold"))
     
     if (!is.list(sbl)){
       p <- p + ylab(ystr)
@@ -195,7 +198,7 @@ covariateLinePlot <- function(sbl, alpha=0.05, nbins = 25,
     }
     
     if(!is.null(trans)){
-      p <- p + scale_y_continuous(trans=trans)
+      p <- p + scale_y_continuous(trans = trans, labels = function(x) paste0(x, "%"))
     }
     
     if(!is.null(transx)){

### 2018-09-05 datasets/R/strat_hist.R

@@ -49,14 +49,15 @@ strat_hist <- function(dat, pvalue, covariate, binwidth=0.025, maxy=3, numQ=3,
                    colour="grey", fill="lightgrey") +
       aes(y=..density..)+
       theme_classic() +
-      theme(plot.title = element_text(size = 11),
+      theme(plot.subtitle = element_text(size = 11, margin = margin(b = -10), hjust = 0.5),
+            plot.title = element_text(size = 11),
             axis.title = element_text(size = 11)) +
       xlim(0,1) +
       scale_y_continuous(expand = c(0.02, 0)) +
       coord_cartesian(ylim=c(0, maxy)) + 
       xlab("") +
       ylab("") +
-      ggtitle(title) 
+      labs(title = "      ", subtitle = title) 
     
     if (axislabs)
       p <- p + xlab("p-value") + ylab("Density")

### 2018-09-05 manuscript_figures/manuscript_figures.Rmd
Redo color scale for Figure S4 and S10

@@
 
-Fig_impactofinfcov <- ggplot(dftab, aes(x = dataset, y = method, fill = diff_inf_percent)) + 
+trim_tails <- function(range = c(-Inf, Inf)) scales::trans_new("trim_tails", 
+                transform = function(x) {
+                  force(range)
+                  desired_breaks <- scales::extended_breaks(n = 7)(x[x >= range[1] & x <= range[2]])
+                  break_increment <- diff(desired_breaks)[1]
+                  x[x < range[1]] <- range[1] - break_increment
+                  x[x > range[2]] <- range[2] + break_increment
+                  x
+                },
+                inverse = function(x) x,
+
+                breaks = function(x) {
+                  force(range)
+                  scales::extended_breaks(n = 7)(x)
+                },
+                format = function(x) {
+                  force(range)
+                  x[1] <- paste("<", range[1])
+                  x[length(x)] <- paste(">", range[2])
+                  paste0(x,"%")
+                })
+
+Fig_impactofinfcov <- ggplot(dftab, aes(x = dataset, y = method, fill = diff_inf_percent*100)) + 
         geom_tile() + 
@@
                                      ifelse(top <= 0.7, scales::percent(top), ""))), 
-                  color = "black", size = 2) +
+                  color = "black", size = 2)+
         scale_fill_gradientn("% Change in\nrejections",
-                             labels = scales::percent, 
-                             colours = c("darkblue","white","darkred"), 
-                             values = scales::rescale(c(-.05,0,1)),
-                             guide = "colorbar", oob=scales::squish,
-                             limits=c(-.05,1)) +
+                             colors = c("darkblue","white","darkred"), 
+                             limits = c(-20, 100),
+                             values = scales::rescale(c(-20, 0, 100)), 
+                             trans = trim_tails(range = c(-20, 100)),
+                             oob = scales::squish) +
         theme_bw() +
@@
         theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
-              legend.position = c(0.99, 0.94), legend.justification = c(1, 1),
+              legend.position = c(0.99, 0.95), legend.justification = c(1, 1),
               legend.background = element_rect(fill = scales::alpha("white", 3/4), color = 'black'),
               legend.direction = "horizontal",
-              legend.text = element_text(size = 7),
-              legend.title = element_text(size = 9),
-              legend.margin = margin(6, 10, 6, 10))
+              legend.text = element_text(size = 7, angle=45, margin = margin(t=5)),
+              legend.title = element_text(size = 9, margin=margin(t=-10)),
+              legend.margin = margin(t=2, r=10, b=-8, l=10))
 
@@
 
-Fig_impactofinfcov <- ggplot(dftab, aes(x = casestudy, y = method, fill = diff_inf_percent)) + 
+Fig_impactofinfcov <- ggplot(dftab, aes(x = casestudy, y = method, fill = diff_inf_percent*100)) + 
         geom_tile() + 
         scale_fill_gradientn("% Change in\nrejections",
-                             labels = scales::percent, 
-                             colours = c("darkblue","white","darkred"), 
-                             values = scales::rescale(c(-.05,0,1)),
-                             guide = "colorbar", limits=c(-.05,1),
-                             oob=scales::squish) +
+                             colors = c("darkblue","white","darkred"), 
+                             limits = c(-20, 100),
+                             values = scales::rescale(c(-20, 0, 100)), 
+                             trans = trim_tails(range = c(-20, 100)),
+                             oob = scales::squish) +
         theme_bw() +
@@
               legend.direction = "horizontal",
-              legend.text = element_text(size = 7),
-              legend.title = element_text(size = 9),
-              legend.margin = margin(6, 10, 6, 10))
+              legend.text = element_text(size = 7, angle=45, margin = margin(t=5)),
+              legend.title = element_text(size = 9, margin=margin(t=-10)),
+              legend.margin = margin(t=2, r=10, b=-8, l=10))
 
@@
   tmp <- left_join(tmp1, tmp2, by = c("method", "casestudy", "replicate", "alpha")) %>%
-      left_join(tmp3, by = c("method", "casestudy", "replicate", "alpha")) %>%
       mutate(inform = inform[l],
@@
 dftab <- tab %>%
-  select(-casestudy, -tnr) %>%
+  select(-casestudy) %>%
   myspread(inform, c(tpr, fdr))

### 2018-09-06 datasets/simulations/simulations-uasettings-t.Rmd
Move simulation Rmds around, suppl into separate dir

--- 2018-09-04 datasets/simulations/simulations-varyingeffectsize-t.Rmd
@@
 ---
-title: "Simulation Study: Varying Effect Size (t-11)"
+title: "Simulation Study: Unimodal Settings (t test statistics)"
 author: "Patrick Kimes"
@@
 ## intermediary files we create below
-spiky_file <- file.path(resdir, "varyinges-t-benchmark-spiky.rds")
-flattop_file <- file.path(resdir, "varyinges-t-benchmark-flattop.rds")
-skew_file <- file.path(resdir, "varyinges-t-benchmark-skew.rds")
-bimodal_file <- file.path(resdir, "varyinges-t-benchmark-bimodal.rds")
+spiky_file <- file.path(resdir, "uasettings-t-benchmark-spiky.rds")
+flattop_file <- file.path(resdir, "uasettings-t-benchmark-flattop.rds")
+skew_file <- file.path(resdir, "uasettings-t-benchmark-skew.rds")
+bimodal_file <- file.path(resdir, "uasettings-t-benchmark-bimodal.rds")
 
@@
 `nulltype = "empirical"` and `nulltype = "theoretical"`. Since all settings in this
-series of simulations use test statistics simulated with t-distributed test statistics,
-we include Scott's FDR Regression in all of the comparisons.
+series of simulations use t-distributed test statistics, we include Scott's FDR Regression
+in all of the comparisons.
 
@@
 the number of degrees of freedom for ASH. We add an additional parameter to the `ashq`
-method with the corresponding degrees of freedom of the test statistic distribution.
+method with the corresponding degrees of freedom of the test statistics.
 
@@
 ```{r spiky-parameters}
-es_dist <- sampler_spiky       # functional: dist of alternative test stats
+es_dist <- function(n) { 2*sampler_spiky(n) }       # functional: dist of alternative test stats
 seed <- 778
@@
 ```{r flattop-parameters}
-es_dist <- sampler_flat_top       # functional: dist of alternative test stats
+es_dist <- function(n) { 2*sampler_flat_top(n) }       # functional: dist of alternative test stats
 seed <- 980
@@
 ```{r skew-parameters}
-es_dist <- sampler_skew       # functional: dist of alternative test stats
+es_dist <- function(n) { 2*sampler_skew(n) }       # functional: dist of alternative test stats
 seed <- 206
@@
 ```{r bimodal-parameters}
-es_dist <- sampler_bimodal       # functional: dist of alternative test stats
+es_dist <- function(n) { 2*sampler_bimodal(n) }       # functional: dist of alternative test stats
 seed <- 913

### 2018-09-06 datasets/simulations/simulations-uasettings.Rmd
Move simulation Rmds around, suppl into separate dir

--- 2018-09-06 datasets/simulations/simulations-uasettings-t.Rmd
@@
 ---
-title: "Simulation Study: Unimodal Settings (t test statistics)"
+title: "Simulation Study: Unimodal Settings"
 author: "Patrick Kimes"
@@
 setting. An informative covariate is included in the setting as described in
-`simulations-informative-cubic.Rmd`. This set of simulations is similar to those
-presented in `simulations-varyingeffectsize.Rmd`, however, t-distributed test
-statistics are used.
+`simulations-informative-cubic.Rmd`.
 
@@
 ## intermediary files we create below
-spiky_file <- file.path(resdir, "uasettings-t-benchmark-spiky.rds")
-flattop_file <- file.path(resdir, "uasettings-t-benchmark-flattop.rds")
-skew_file <- file.path(resdir, "uasettings-t-benchmark-skew.rds")
-bimodal_file <- file.path(resdir, "uasettings-t-benchmark-bimodal.rds")
+spiky_file <- file.path(resdir, "uasettings-benchmark-spiky.rds")
+flattop_file <- file.path(resdir, "varyinges-benchmark-flattop.rds")
+skew_file <- file.path(resdir, "uasettings-benchmark-skew.rds")
+bimodal_file <- file.path(resdir, "uasettings-benchmark-bimodal.rds")
 
@@
 `nulltype = "empirical"` and `nulltype = "theoretical"`. Since all settings in this
-series of simulations use t-distributed test statistics, we include Scott's FDR Regression
+series of simulations use Gaussian test statistics, we include Scott's FDR Regression
 in all of the comparisons.
@@
 
-Since all simulation settings in this case study use t-distributed test statistics, we must specify
-the number of degrees of freedom for ASH. We add an additional parameter to the `ashq`
-method with the corresponding degrees of freedom of the test statistics.
-
-```{r}
-bdplus <- modifyBMethod(bdplus, "ashq", df = 11)
-```
-
 All simulation settings will share the following parameters.
@@
 pi0 <- pi0_cubic(0.90)              # numeric: proportion of null hypotheses
-ts_dist <- rt_perturber(11)         # functional: sampling dist/noise for test stats
-null_dist <- rt_2pvaluer(11)        # functional: dist to calc p-values
+ts_dist <- rnorm_perturber(1)       # functional: sampling dist/noise for test stats
+null_dist <- rnorm_2pvaluer(1)      # functional: dist to calc p-values
 icovariate <- runif                 # functional: independent covariate

### 2018-09-06 datasets/simulations/simulations-varyinginformativeness-discrete.Rmd
Move simulation Rmds around, suppl into separate dir

--- 2018-09-04 datasets/simulations/simulations-varyinginformativeness.Rmd
@@
 ## intermediary files we create below
-info00_file <- file.path(resdir, "varyinginfo-benchmark-level00.rds")
-info20_file <- file.path(resdir, "varyinginfo-benchmark-level20.rds")
-info40_file <- file.path(resdir, "varyinginfo-benchmark-level40.rds")
-info60_file <- file.path(resdir, "varyinginfo-benchmark-level60.rds")
-info80_file <- file.path(resdir, "varyinginfo-benchmark-level80.rds")
-info100_file <- file.path(resdir, "varyinginfo-benchmark-level100.rds")
+info00_file <- file.path(resdir, "varyinginfo-discrete-benchmark-level00.rds")
+info20_file <- file.path(resdir, "varyinginfo-discrete-benchmark-level20.rds")
+info40_file <- file.path(resdir, "varyinginfo-discrete-benchmark-level40.rds")
+info60_file <- file.path(resdir, "varyinginfo-discrete-benchmark-level60.rds")
+info80_file <- file.path(resdir, "varyinginfo-discrete-benchmark-level80.rds")
+info100_file <- file.path(resdir, "varyinginfo-discrete-benchmark-level100.rds")
 

### 2018-09-06 datasets/simulations/supplementary/simulations-informative-cubic-vash.Rmd
Move simulation Rmds around, suppl into separate dir

--- 2018-09-04 datasets/simulations/simulations-informative-cubic-vash.Rmd
@@
 is a smooth (cubic) function of the covariate. We include simulation results
-with only t-distributed noise.
+with only t-distributed test statistics.
 
@@
 
-We consider the setting with t-distributed noise.
+We consider the setting with t-distributed test statistics.
 

### 2018-09-06 datasets/simulations/supplementary/simulations-uasettings-originalscale.Rmd
Move simulation Rmds around, suppl into separate dir

--- 2018-08-06 datasets/simulations/simulations-varyingeffectsize.Rmd
@@
 As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
-for simulations with Gaussian or t-distributed noise. Again, we include both
+for simulations with Gaussian or t-distributed test statistics. Again, we include both
 `nulltype = "empirical"` and `nulltype = "theoretical"`. Since all settings in this
-series of simulations use test statistics simulated with Gaussian noise, we include
+series of simulations use test statistics simulated with Gaussian test statistics, we include
 Scott's FDR Regression in all of the comparisons.

### 2018-09-06 datasets/simulations/simulations-informative-cosine.Rmd
Change t-5 effect sizes to match others for base cases

@@
 ```{r t5-parameters}
-es_dist <- rnorm_generator(6)    # functional: dist of alternative test stats
+es_dist <- rnorm_generator(3)    # functional: dist of alternative test stats
 ts_dist <- rt_perturber(5)  # functional: sampling dist/noise for test stats

### 2018-09-06 datasets/R/helpers-simulations.R
Fix allowed sim range

@@ -154,7 +154,7 @@ pi0_step <- function(pi0) {
 
 ## shifted/stretched cubic function
 pi0_cubic <- function(pi0) {
-    stopifnot(pi0 <= 1L, pi0 >= 2/3)
+    stopifnot(pi0 <= 1L, pi0 >= 3/4)
     function(x) { (1-x)^(1/3) * 4*(1-pi0) + 4 * pi0 - 3 }
 }
 

### 2018-09-06 datasets/simulations/simulations-uasettings.Rmd
Fix relative path of R directory for suppl Rmds; change ua output file names

@@
 spiky_file <- file.path(resdir, "uasettings-benchmark-spiky.rds")
-flattop_file <- file.path(resdir, "varyinges-benchmark-flattop.rds")
+flattop_file <- file.path(resdir, "uasettings-benchmark-flattop.rds")
 skew_file <- file.path(resdir, "uasettings-benchmark-skew.rds")

### 2018-09-06 datasets/simulations/supplementary/simulations-informative-cosine-nullAdaPT.Rmd
Fix relative path of R directory for suppl Rmds; change ua output file names

--- 2018-09-05 datasets/simulations/simulations-informative-cosine-nullAdaPT.Rmd
@@
 ## load helper functions
-for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
+for (f in list.files("../../R", "\\.(r|R)$", full.names = TRUE)) {
     source(f)
@@
 plotsim_average(resdf, met = "rejections", filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars = TRUE),
+                merge_ihw = TRUE, errorBars = TRUE,
                 palette = candycols2)

### 2018-09-06 datasets/simulations/supplementary/simulations-uasettings-originalscale.Rmd
Fix relative path of R directory for suppl Rmds; change ua output file names

@@
 ---
-title: "Simulation Study: Varying Effect Size"
+title: "Simulation Study: Unimodal Settings"
 author: "Patrick Kimes"
@@
 ## intermediary files we create below
-spiky_file <- file.path(resdir, "varyinges-benchmark-spiky.rds")
-flattop_file <- file.path(resdir, "varyinges-benchmark-flattop.rds")
-skew_file <- file.path(resdir, "varyinges-benchmark-skew.rds")
-bimodal_file <- file.path(resdir, "varyinges-benchmark-bimodal.rds")
+spiky_file <- file.path(resdir, "uasettings-benchmark-spiky.rds")
+flattop_file <- file.path(resdir, "uasettings-benchmark-flattop.rds")
+skew_file <- file.path(resdir, "uasettings-benchmark-skew.rds")
+bimodal_file <- file.path(resdir, "uasettings-benchmark-bimodal.rds")
 

### 2018-09-06 datasets/simulations/supplementary/simulations-uasettings-t-originalscale.Rmd
Fix relative path of R directory for suppl Rmds; change ua output file names

--- 2018-09-04 datasets/simulations/simulations-varyingeffectsize-t.Rmd
@@
 ---
-title: "Simulation Study: Varying Effect Size (t-11)"
+title: "Simulation Study: Unimodal Settings (t-11)"
 author: "Patrick Kimes"
@@
 `simulations-informative-cubic.Rmd`. This set of simulations is similar to those
-presented in `simulations-varyingeffectsize.Rmd`, however, t-distributed test
+presented in `simulations-uasettings-originalscale.Rmd`, however, t-distributed test
 statistics are used.

### 2018-09-06 manuscript_figures/manuscript_figures.Rmd
Fix typo in Fig 2 titles

@@
       theme(legend.position="none")  +  
-      ggtitle("Mean FDR (100 replications)") +
+      ggtitle("Mean TPR (100 replications)") +
       scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02))
@@
       ylab("Number of Rejections") +
-      ggtitle("Mean FDR (100 replications)") +
+      ggtitle("Mean Rejections (100 replications)") +
       scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02))
@@
 Fig2
-ggsave(file.path(outdir, "Figure2A-D.pdf"), width=8.5, height=6)
+ggsave(file.path(outdir, "Figure2A-D.pdf"), width=8.75, height=6)
 

### 2018-09-06 manuscript_figures/manuscript_figures_sims.Rmd
Update and clean simulation main and suppl plots

--- 2018-08-30 manuscript_figures/manuscript_figures_sims.Rmd
@@
 
-```{r}
-genplot <- function(tab, cov, xt = "", met = c("FDR", "TPR"), ebw = 0.0025) {
+```{r genplot-function}
+genplot <- function(tab, cov, type = c("info", "diff"), xt = "", met = c("FDR", "TPR"), ebw = 0.0025) {
     met <- match.arg(met)
     cov <- rlang::enquo(cov)
+    type <- match.arg(type)
+    if (type == "info") {
+        ymean <- rlang::quo(mean.info)
+        yse <- rlang::quo(se.info)
+    } else if (type == "diff") {
+        ymean <- rlang::quo(mean.diff)
+        yse <- rlang::quo(se.diff)
+    }        
     gp <- dplyr::filter(tab, key == met) %>%
-        ggplot(aes(x = !!cov, y = mean.info, color = Method, group = Method)) +
+        ggplot(aes(x = !!cov, y = !!ymean, color = Method, group = Method)) +
         geom_line(aes(linetype = Method), alpha = 0.85) +
-        geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = ebw, alpha=0.5) +
+        geom_errorbar(aes(ymin = !!ymean - !!yse, ymax = !!ymean + !!yse), width = ebw, alpha=0.5) +
         scale_linetype_manual(values = lty) + 
@@
               plot.title = element_text(face = "bold", size = 14))
-    if (met == "FDR") {
+    if (type == "diff") {
+        gp <- gp + geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) 
+    } else if (met == "FDR") {
         if (rlang::quo_name(cov) == "alpha") {
@@
 
-```{r, fig.width = 10, fig.height = 12}
-## Figure 3A,B: FDR/TPR of methods across varying number of tests at alpha = 0.05
-res_ntests <- dplyr::filter(resmet, setting == "varyingntests", alpha == 0.05)
-p3a <- genplot(res_ntests, cov = as.factor(ntests), met = "FDR", ebw = 0.15) +
-    xlab("Number of Tests")
-p3b <- genplot(res_ntests, cov = as.factor(ntests), met = "TPR", ebw = 0.15) +
-    xlab("Number of Tests")
-
-## Figure 3C,D: FDR/TPR of methods across varying null proportion at alpha = 0.05
-res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0", alpha == 0.05)
-p3c <- genplot(res_pi0, cov = pi0, met = "FDR", ebw = 2.5) +
-    scale_x_continuous("Proportion of Null Hypotheses",
-                       breaks = c(5, seq(10, 90, by = 10), 95, 99))
-p3d <- genplot(res_pi0, cov = pi0, met = "TPR", ebw = 2.5) +
-    scale_x_continuous("Proportion of Null Hypotheses",
-                       breaks = c(5, seq(10, 90, by = 10), 95, 99))
-
-## Figure 3E,F: FDR/TPR of methods across varying informativeness at alpha = 0.05
-res_info <- dplyr::filter(resmet, setting == "varyinginfo", alpha == 0.05)
-p3e <- genplot(res_info, cov = inform, met = "FDR", ebw = 2.5) +
+```{r Figure3, fig.width = 10, fig.height = 12}
+## Figure 3A,B: FDR/TPR of methods across varying informativeness at alpha = 0.05
+res_info <- dplyr::filter(resmet, setting == "varyinginfo-smooth", alpha == 0.05)
+p3a <- genplot(res_info, cov = inform, met = "FDR", ebw = 2.5) +
     scale_x_continuous("Informativeness", breaks = seq(0, 100, by = 20))
-p3f <- genplot(res_info, cov = inform, met = "TPR", ebw = 2.5) +
+p3b <- genplot(res_info, cov = inform, met = "TPR", ebw = 2.5) +
     scale_x_continuous("Informativeness", breaks = seq(0, 100, by = 20))
 
-Fig3 <- plot_grid(p3a + guides(color = FALSE, linetype = FALSE),
+## add title to 3A,B
+p3ab <- plot_grid(p3a + guides(color = FALSE, linetype = FALSE),
                   p3b + guides(color = FALSE, linetype = FALSE),
-                p3c + guides(color = FALSE, linetype = FALSE),
+                  labels = LETTERS[1:2], ncol = 2)
+p3abTitle <- ggdraw() +
+    draw_label("Simulation performance across varying informativeness",
+               fontface = 'bold')
+p3ab <- plot_grid(p3abTitle, p3ab, ncol = 1, rel_heights = c(0.1, 1))
+
+## Figure 3C,D: FDR/TPR of methods across varying number of tests at alpha = 0.05
+res_ntests <- dplyr::filter(resmet, setting == "varyingntests", alpha == 0.05)
+p3c <- genplot(res_ntests, cov = ntests, met = "FDR", ebw = 0.05) +
+    scale_x_continuous("Number of Tests", trans = "log10",
+                       breaks = c(1e2, 5e2, 1e3, 5e3, 1e4, 5e4))
+p3d <- genplot(res_ntests, cov = ntests, met = "TPR", ebw = 0.05) +
+    scale_x_continuous("Number of Tests", trans = "log10",
+                       breaks = c(1e2, 5e2, 1e3, 5e3, 1e4, 5e4))
+
+## add title to 3C,D
+p3cd <- plot_grid(p3c + guides(color = FALSE, linetype = FALSE),
                   p3d + guides(color = FALSE, linetype = FALSE),
-                p3e + guides(color = FALSE, linetype = FALSE),
+                  labels = LETTERS[3:4], ncol = 2)
+p3cdTitle <- ggdraw() +
+    draw_label("Simulation performance across varying numbers of hypotheses",
+               fontface = 'bold')
+p3cd <- plot_grid(p3cdTitle, p3cd, ncol = 1, rel_heights = c(0.1, 1))
+
+## Figure 3E,F: FDR/TPR of methods across varying null proportion at alpha = 0.05
+res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0", alpha == 0.05)
+p3e <- genplot(res_pi0, cov = 100-pi0, met = "FDR", ebw = 2.5) +
+    scale_x_continuous("Proportion of Non-Null Hypotheses",
+                       breaks = 100 - c(5, seq(10, 90, by = 10), 95, 99))
+p3f <- genplot(res_pi0, cov = 100-pi0, met = "TPR", ebw = 2.5) +
+    scale_x_continuous("Proportion of Non-Null Hypotheses",
+                       breaks = 100 - c(5, seq(10, 90, by = 10), 95, 99))
+
+## add title to 3E,F
+p3ef <- plot_grid(p3e + guides(color = FALSE, linetype = FALSE),
                   p3f + guides(color = FALSE, linetype = FALSE),
-                labels = LETTERS[1:6], ncol = 2)
+                  labels = LETTERS[5:6], ncol = 2)
+p3efTitle <- ggdraw() +
+    draw_label("Simulation performance across varying non-null proportion",
+               fontface = 'bold')
+p3ef <- plot_grid(p3efTitle, p3ef, ncol = 1, rel_heights = c(0.1, 1))
+
+## pull Figure 3 together
+Fig3 <- plot_grid(p3ab, p3cd, p3ef, ncol = 1)
 Fig3 <- plot_grid(Fig3, get_legend(p3a), rel_widths = c(1, .2))
@@
 
-# Figure 3+
+# Supplementary Figure (distributions)
 
@@
 Test statistics were simulated from Normal, t with 11 degrees of freedom, t with 5 degrees of freedom, and Chi-squared with 4 degrees of freedom.
+We plot TPR and FDR across 100 replicates over nominal FDR cutoffs of 0.01 to 0.10.
+Additionally, we plot the differences in TPR and FDR for methods when using an informative or an uninformative covariate.
 
-```{r, fig.width = 10, fig.height = 12}
-## Figure 3plus A,B: FDR/TPR of methods across varying test statistics distribution
+```{r SFigureSim1, fig.width = 14, fig.height = 16}
+## SFigure Sim1 A,B: FDR/TPR of methods across varying test statistics distribution
 res_noise <- dplyr::filter(resmet, setting == "informative-cubic")
-p3pa <- genplot(res_noise, cov = alpha, met = "FDR") +
+ps1a <- genplot(res_noise, cov = alpha, met = "FDR") +
     facet_grid(. ~ dist, labeller = label_parsed)
-p3pb <- genplot(res_noise, cov = alpha, met = "TPR") +
+ps1b <- genplot(res_noise, cov = alpha, met = "TPR") +
     facet_grid(. ~ dist, labeller = label_parsed)
 
-Fig3p <- plot_grid(p3pa + guides(color = FALSE, linetype = FALSE),
-                   p3pb + guides(color = FALSE, linetype = FALSE),
-                   labels = LETTERS[1:2], ncol = 1)
-Fig3p <- plot_grid(Fig3p, get_legend(p3pa), rel_widths = c(7.5, 1))
-Fig3p
-ggsave(file.path(outdir, "Figure3plus.pdf"), width=14, height=8)
+## SFigure Sim1 C,D: FDR/TPR difference of methods
+ps1c <- genplot(res_noise, cov = alpha, type = "diff", met = "FDR") +
+    facet_grid(. ~ dist, labeller = label_parsed)
+ps1d <- genplot(res_noise, cov = alpha, type = "diff", met = "TPR") +
+    facet_grid(. ~ dist, labeller = label_parsed)
+
+## create title for SFigure Sim1
+ps1Title <- ggdraw() +
+    draw_label("Simulation performance across test statistic distributions",
+               fontface = 'bold', size = 20)
+
+## pull SFigure Sim1 together
+SFigS1 <- plot_grid(ps1a + guides(color = FALSE, linetype = FALSE),
+                    ps1b + guides(color = FALSE, linetype = FALSE),
+                    ps1c + guides(color = FALSE, linetype = FALSE),
+                    ps1d + guides(color = FALSE, linetype = FALSE),
+                    labels = LETTERS[1:4], ncol = 1)
+SFigS1 <- plot_grid(ps1Title, SFigS1, ncol = 1, rel_heights = c(0.03, 1))
+SFigS1 <- plot_grid(SFigS1, get_legend(ps1a), rel_widths = c(7.5, 1))
+SFigS1
+ggsave(file.path(outdir, "SFigureSim1.pdf"), width=14, height=16)
+```
+
+# Supplementary Figure (unimodal assumption)
+
+This is a faceted figure with results from the pure simulation studies showing the behavior of methods across different sampling distributions of the effect size for non-null tests.
+The distributions of effect sizes used here were taken from the Stephens (Biostatistics, 2016) paper which originally introduced the ASH method.
+All effect size distributions are unimodal, with the exception of the appropriately named "bimodal" setting.
+The effect size distributions from the ASH manuscript have been doubled from their original implementations since TPRs were low (<10%) for all methods at their original magnitude.
+Again, the FDR and TPR as well as FDR and TPR differences are plotted for all methods.
+
+```{r SFigureSim2, fig.width = 10, fig.height = 14}
+res_ua <- dplyr::filter(resmet, setting == "uasettings", grepl("N\\(0, 1\\)", dist),
+                        signal != "bimodal")
+
+## SFigure Sim2 A,B: FDR/TPR of methods across unimodal effect size distribution
+ps2a <- genplot(res_ua, cov = alpha, met = "FDR") +
+    facet_grid(. ~ signal, labeller = label_parsed)
+ps2b <- genplot(res_ua, cov = alpha, met = "TPR") +
+    facet_grid(. ~ signal, labeller = label_parsed)
+
+## SFigure Sim2 C,D: FDR/TPR difference of methods
+ps2c <- genplot(res_ua, cov = alpha, type = "diff", met = "FDR") +
+    facet_grid(. ~ signal, labeller = label_parsed)
+ps2d <- genplot(res_ua, cov = alpha, type = "diff", met = "TPR") +
+    facet_grid(. ~ signal, labeller = label_parsed)
+
+## create title for SFigure Sim2
+ps2Title <- ggdraw() +
+    draw_label("Simulation performance across unimodal effect size distributions",
+               fontface = 'bold', size = 17)
+
+## pull SFigure Sim2 together
+SFigS2 <- plot_grid(ps2a + guides(color = FALSE, linetype = FALSE),
+                    ps2b + guides(color = FALSE, linetype = FALSE),
+                    ps2c + guides(color = FALSE, linetype = FALSE),
+                    ps2d + guides(color = FALSE, linetype = FALSE),
+                    labels = LETTERS[1:4], ncol = 1)
+SFigS2 <- plot_grid(ps2Title, SFigS2, ncol = 1, rel_heights = c(0.03, 1))
+SFigS2 <- plot_grid(SFigS2, get_legend(ps2a), rel_widths = c(6.5, 1))
+SFigS2
+ggsave(file.path(outdir, "SFigureSim2.pdf"), width=10, height=14)
+```
+
+# Supplementary Figure (covariate relationship)
+
+This is a faceted figure with results from the pure simulation studies showing the behavior of methods across different function relationships between the informative covariate and null probability (pi0) of a test.
+In addition to the more deliberately designed sequence of informative covariates studied in the "informativeness" set of simulations described above, we also considered four functional relationships between the covariate and pi0, named "step", "sine", "cosine", and "cubic".
+
+sampling distributions of the effect size for non-null tests.
+The distributions of effect sizes used here were taken from the Stephens (Biostatistics, 2016) paper which originally introduced the ASH method.
+All effect size distributions are unimodal, with the exception of the appropriately named "bimodal" setting.
+The effect size distributions from the ASH manuscript have been doubled from their original implementations since TPRs were low (<10%) for all methods at their original magnitude.
+Again, the FDR and TPR as well as FDR and TPR differences are plotted for all methods.
+
+```{r SFigureSim3, fig.width = 12, fig.height = 16}
+res_icov <- dplyr::filter(resmet, grepl("^informative", setting), grepl("N\\(0, 1\\)", dist))
+res_icov <- dplyr::mutate(res_icov, relationship = gsub("informative-", "", setting))
+
+## SFigure Sim3 A: Form of informative covariate and pi0 relationship
+pi0form <- tibble(x = seq(0, 1, .01))
+pi0form <- dplyr::mutate(pi0form,
+                         cosine = pi0_cosine(0.9)(x),
+                         cubic = pi0_cubic(0.9)(x),
+                         sine = pi0_sine(0.9)(x),
+                         step = pi0_step(0.9)(x))
+pi0form <- tidyr::gather(pi0form, relationship, pi0, -x)
+ps3a <- ggplot(pi0form, aes(x = x, y = pi0)) +
+    geom_line() +
+    expand_limits(y = c(0.5, 1)) + 
+    theme_classic() +
+    scale_x_continuous(breaks = seq(0, 1, .2)) + 
+    scale_y_continuous(labels = scales::percent) + 
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    facet_grid(. ~ relationship)
+
+## SFigure Sim3 B,C: FDR/TPR of methods across varying relationships
+ps3b <- genplot(res_icov, cov = alpha, met = "FDR") +
+    facet_grid(. ~ relationship, labeller = label_parsed)
+ps3c <- genplot(res_icov, cov = alpha, met = "TPR") +
+    facet_grid(. ~ relationship, labeller = label_parsed)
+
+## SFigure Sim3 D,E: FDR/TPR difference of methods
+ps3d <- genplot(res_icov, cov = alpha, type = "diff", met = "FDR") +
+    facet_grid(. ~ relationship, labeller = label_parsed)
+ps3e <- genplot(res_icov, cov = alpha, type = "diff", met = "TPR") +
+    facet_grid(. ~ relationship, labeller = label_parsed)
+
+## create title for SFigure Sim3
+ps3Title <- ggdraw() +
+    draw_label("Simulation performance across informative covariate relationship",
+               fontface = 'bold', size = 20)
+
+## pull SFigure Sim3 together
+SFigS3 <- plot_grid(ps3a + guides(color = FALSE, linetype = FALSE),
+                    ps3b + guides(color = FALSE, linetype = FALSE),
+                    ps3c + guides(color = FALSE, linetype = FALSE),
+                    ps3d + guides(color = FALSE, linetype = FALSE),
+                    ps3e + guides(color = FALSE, linetype = FALSE),
+                    labels = LETTERS[1:5], ncol = 1, rel_heights = c(.6, rep(1, 4)))
+SFigS3 <- plot_grid(ps3Title, SFigS3, ncol = 1, rel_heights = c(0.03, 1))
+SFigS3 <- plot_grid(SFigS3, get_legend(ps3b), rel_widths = c(6.5, 1))
+SFigS3
+ggsave(file.path(outdir, "SFigureSim3.pdf"), width=12, height=16)
 ```

### 2018-09-06 datasets/simulations/simulations-summary.Rmd
Update sim summary Rmd based on renamed/moved simulation settings

@@
 resmet <- dplyr::mutate(resmet, pi0 = as.numeric(gsub("nullprop", "", pi0)))
-resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("varyingpi0", setting), "gaussian", dist))
+resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyingpi0$", setting), "gaussian", dist))
+resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyingpi0-t$", setting), "t11", dist))
 
-resmet <- dplyr::mutate(resmet, signal = ifelse(grepl("varyinges", setting), dist, NA))
-resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyinges$", setting), "gaussian", dist))
-resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyinges-t$", setting), "t11", dist))
-resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyinges-chisq$", setting), "chisq4", dist))
-resmet <- dplyr::mutate(resmet, setting = ifelse(grepl("varyinges", setting), "varyinges", setting))
+resmet <- dplyr::mutate(resmet, signal = ifelse(grepl("uasettings", setting), dist, NA))
+resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^uasettings$", setting), "gaussian", dist))
+resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^uasettings-t$", setting), "t11", dist))
+resmet <- dplyr::mutate(resmet, setting = ifelse(grepl("uasettings", setting), "uasettings", setting))
 
@@
 
-We also replace the sampling noise labels with more plotting-friendly labels.
+We also replace the test statistic distribution labels with more plotting-friendly labels.
 
@@
 resmet <- dplyr::mutate(resmet, dist = factor(dist, levels = c("gaussian", "t11", "t5", "chisq4"),
-                                              labels = c(expression("Noise: " * N(0, 1)),
-                                                         expression("Noise: " * t[11]),
-                                                         expression("Noise: " * t[5]),
-                                                         expression("Noise: " * {chi^2}[4]))))
+                                              labels = c(expression("Distribution: " * N(0, 1)),
+                                                         expression("Distribution: " * t[11]),
+                                                         expression("Distribution: " * t[5]),
+                                                         expression("Distribution: " * {chi^2}[4]))))
 ```
@@
 covariate is sampled uniformly between 0 and 1 and the probability of the corresponding test
-being null is one value if the covariate is less than 0.8, and another value if the covariate
-is above 0.8. The "informativeness" is measured by the difference between these two probabilities,
-with informativeness 0 and 1 corresponding to no difference and maximal difference (i.e. 0 and 1),
-respectively. We plot the false discovery proportion and true positive rate at a nominal alpha
+being null is more or less dependent on the value of the covariate. The "informativeness" is
+greater when the value of the covariate is more informative of the probability of null being
+high or low. We plot the false discovery proportion and true positive rate at a nominal alpha
 cutoff of 0.05 for each method.
@@
 ```{r, fig.width = 10, fig.height = 4}
-res_info <- dplyr::filter(resmet, setting == "varyinginfo", alpha == 0.05)
+res_info <- dplyr::filter(resmet, setting == "varyinginfo-smooth", alpha == 0.05)
 
@@
 ```{r, fig.width = 18, fig.height = 8}
-res_ua <- dplyr::filter(resmet, setting == "varyinges", grepl("N\\(0, 1\\)", dist))
+res_ua <- dplyr::filter(resmet, setting == "uasettings", grepl("N\\(0, 1\\)", dist))
 
@@
 ```{r, fig.width = 18, fig.height = 8}
-res_ua <- dplyr::filter(resmet, setting == "varyinges", grepl("t\\[11\\]", dist))
+res_ua <- dplyr::filter(resmet, setting == "uasettings", grepl("t\\[11\\]", dist))
 

### 2018-09-07 manuscript_figures/manuscript_figures_sims.Rmd
Remove random text left in Rmd
Clean up notes in Rmd
Create AdaPT supplemental figure, add delta y-axis labels to plots

@@
 
-Running this
-
 ```{r load-workspace}
@@
 ## load helper functions
-for (f in list.files("../datasets/R", "\\.(r|R)$", full.names = TRUE)) {
+for (f in list.files(file.path("..", "datasets", "R"), "\\.(r|R)$", full.names = TRUE)) {
     source(f)
@@
 ```{r}
-col <- as.character(candycols$col)
-names(col) <- as.character(candycols$Method)
-lty <- as.character(candycols$lty)
-names(lty) <- as.character(candycols$Method)
+col <- c(as.character(candycols$col), "darkgoldenrod3")
+names(col) <- c(as.character(candycols$Method), "adapt-withnull")
+lty <- c(as.character(candycols$lty), "dashed")
+names(lty) <- c(as.character(candycols$Method), "adapt-withnull")
 ```
@@
         scale_color_manual(values = col) +
-        scale_y_continuous(met, labels = scales::percent) +
         expand_limits(y = 0) +
@@
         gp <- gp + geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2)
-    } else if (met == "FDR") {
+        gp <- gp + scale_y_continuous(bquote(atop(Mean ~ Delta ~ .(met), (informative-uninformative))),
+                                      labels = scales::percent)
+    } else {
+        gp <- gp + scale_y_continuous(met, labels = scales::percent)
+        if (met == "FDR") {
             if (rlang::quo_name(cov) == "alpha") {
@@
                 gp <- gp + geom_hline(yintercept = 0.05, lty = 2, color = "blue", alpha = 1/2) 
+            }
         }
@@
 ps1a <- genplot(res_noise, cov = alpha, met = "FDR") +
-    facet_grid(. ~ dist, labeller = label_parsed)
+    facet_grid(. ~ dist, labeller = label_parsed) +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0)
 ps1b <- genplot(res_noise, cov = alpha, met = "TPR") +
-    facet_grid(. ~ dist, labeller = label_parsed)
+    facet_grid(. ~ dist, labeller = label_parsed) +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0)
 
@@
 ps1c <- genplot(res_noise, cov = alpha, type = "diff", met = "FDR") +
-    facet_grid(. ~ dist, labeller = label_parsed)
+    facet_grid(. ~ dist, labeller = label_parsed) + 
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0)
 ps1d <- genplot(res_noise, cov = alpha, type = "diff", met = "TPR") +
-    facet_grid(. ~ dist, labeller = label_parsed)
+    facet_grid(. ~ dist, labeller = label_parsed) + 
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0) +
+    ylab(expression(paste("Mean ", Delta, "TPR (informative-uninformative)")))
 
@@
 ps2a <- genplot(res_ua, cov = alpha, met = "FDR") +
-    facet_grid(. ~ signal, labeller = label_parsed)
+    facet_grid(. ~ signal, labeller = label_parsed) +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0)
 ps2b <- genplot(res_ua, cov = alpha, met = "TPR") +
-    facet_grid(. ~ signal, labeller = label_parsed)
+    facet_grid(. ~ signal, labeller = label_parsed) +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0)
 
@@
 ps2c <- genplot(res_ua, cov = alpha, type = "diff", met = "FDR") +
-    facet_grid(. ~ signal, labeller = label_parsed)
+    facet_grid(. ~ signal, labeller = label_parsed) +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0) +
+    ylab(expression(paste("Mean ", Delta, "FDR (informative-uninformative)")))
 ps2d <- genplot(res_ua, cov = alpha, type = "diff", met = "TPR") +
-    facet_grid(. ~ signal, labeller = label_parsed)
+    facet_grid(. ~ signal, labeller = label_parsed) +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0) +
+    ylab(expression(paste("Mean ", Delta, "TPR (informative-uninformative)")))
 
@@
 In addition to the more deliberately designed sequence of informative covariates studied in the "informativeness" set of simulations described above, we also considered four functional relationships between the covariate and pi0, named "step", "sine", "cosine", and "cubic".
-
-sampling distributions of the effect size for non-null tests.
-The distributions of effect sizes used here were taken from the Stephens (Biostatistics, 2016) paper which originally introduced the ASH method.
-All effect size distributions are unimodal, with the exception of the appropriately named "bimodal" setting.
-The effect size distributions from the ASH manuscript have been doubled from their original implementations since TPRs were low (<10%) for all methods at their original magnitude.
 Again, the FDR and TPR as well as FDR and TPR differences are plotted for all methods.
+The relationship between the covariate (sampled from the interval between 0 and 1) and pi0 is plotted for each setting as well.
 
@@
 ps3b <- genplot(res_icov, cov = alpha, met = "FDR") +
-    facet_grid(. ~ relationship, labeller = label_parsed)
+    facet_grid(. ~ relationship, labeller = label_parsed) +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0) +
+    ylab(expression(paste("Mean ", Delta, "FDR (informative-uninformative)")))
 ps3c <- genplot(res_icov, cov = alpha, met = "TPR") +
-    facet_grid(. ~ relationship, labeller = label_parsed)
+    facet_grid(. ~ relationship, labeller = label_parsed) +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0) +
+    ylab(expression(paste("Mean ", Delta, "TPR (informative-uninformative)")))
 
@@
 ps3d <- genplot(res_icov, cov = alpha, type = "diff", met = "FDR") +
-    facet_grid(. ~ relationship, labeller = label_parsed)
+    facet_grid(. ~ relationship, labeller = label_parsed) +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0)
 ps3e <- genplot(res_icov, cov = alpha, type = "diff", met = "TPR") +
-    facet_grid(. ~ relationship, labeller = label_parsed)
+    facet_grid(. ~ relationship, labeller = label_parsed) +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0)
 
@@
 
+# Supplementary Figure (AdaPT with null model)
+
+This is a faceted figure with results from running the AdaPT method with non-default parameter settings.
+Based on communications with the authors of the methods, in addition to applying the method with parameter settings described in the accompanying package vignette, we attempted running the method with a non-default set of parameters.
+Namely, we run the method including a null relationship between the covariate and the null probability as one of the candidate "models", allowing the AdaPT method to ignore the specified covariate when it does not appear to be "informative enough".
+This model is denoted by `adapt-withnull`.
+The simulations were performed with the "step" informative covariate setting described above, under which the AdaPT method achieved lower TPR with the weakly informative covariate than with a completely uninformative covariate.
+
+Since these results are not included in the primary set of simulation results summarized in the `resmet` object, we recompute them from the output of the `supplementary/simulations-informative-step-nullAdaPT.Rmd` simulation case study.
+
+We only include a subset of methods in this plot because the primary focus is on the relative performance of AdaPT with default and modified parameters.
+
+```{r SFigureS4-parse-data}
+method_subset <- c("bonf", "bh", "qvalue", "adapt-glm", "adapt-withnull")
+
+nullAdapt <- readRDS(file.path("..", "datasets", "simulations", "supplementary", "results",
+                               "nullAdaPT-informative-step-benchmark-gaussian.rds"))
+
+nullAdapt_i <- lapply(nullAdapt, `[[`, "informative")
+nullAdapt_i <- plotsim_standardize(nullAdapt_i, alpha = seq(0.01, 0.10, 0.01))
+nullAdapt_i <- dplyr::select(nullAdapt_i, rep, blabel, param.alpha, key, performanceMetric, alpha, value)
+
+nullAdapt_u <- lapply(nullAdapt, `[[`, "uninformative")
+nullAdapt_u <- plotsim_standardize(nullAdapt_u, alpha = seq(0.01, 0.10, 0.01))
+nullAdapt_u <- dplyr::select(nullAdapt_u, rep, blabel, param.alpha, key, performanceMetric, alpha, value)
+
+nullAdapt <- dplyr::full_join(nullAdapt_i, nullAdapt_u,
+                              by = c("rep", "blabel", "param.alpha", "key", "performanceMetric", "alpha"),
+                              suffix = c(".info", ".uninfo"))
+
+nullAdapt <- dplyr::filter(nullAdapt, blabel %in% method_subset)
+
+nullAdapt <- dplyr::select(nullAdapt, -param.alpha, -key)
+nullAdapt <- dplyr::rename(nullAdapt, Method = blabel, key = performanceMetric)
+nullAdapt <- dplyr::mutate(nullAdapt, Method = factor(Method, levels = method_subset))
+
+adaptmet <- dplyr::mutate(nullAdapt,
+                          value.info = ifelse(is.na(value.info) & grepl("R$", key), 0, value.info),
+                          value.uninfo = ifelse(is.na(value.uninfo) & grepl("R$", key), 0, value.uninfo))
+
+adaptmet <- dplyr::group_by(adaptmet, Method, key, alpha)
+adaptmet <- dplyr::summarize(adaptmet,
+                             mean.info = mean(value.info, na.rm = TRUE),
+                             se.info = sd(value.info, na.rm = TRUE) / sqrt(sum(!is.na(value.info))),
+                             nNA.info = sum(is.na(value.info)),
+                             mean.diff = mean(value.info - value.uninfo, na.rm = TRUE),
+                             se.diff = sd(value.info - value.uninfo, na.rm = TRUE) /
+                                 sqrt(sum(!is.na(value.info - value.uninfo))))
+adaptmet <- dplyr::ungroup(adaptmet)
+```
+
+After computing the performance metrics, we now plot the results for AdaPT with default and modified parameter settings.
+Again, we plot the mean FDR and TPR at nominal FDR cutoffs of 0.01 to 0.10 across 100 replications, as well as the FDR and TPR differences with and without the informative covariate.
+
+```{r SFigureSim4, fig.width = 8, fig.height = 6}
+## SFigure Sim4 A,B: FDR/TPR of methods including adapt with null
+ps4a <- genplot(adaptmet, cov = alpha, met = "FDR") +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0)
+ps4b <- genplot(adaptmet, cov = alpha, met = "TPR") +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0)
+
+## SFigure Sim4 C,D: FDR/TPR difference of methods
+ps4c <- genplot(adaptmet, cov = alpha, type = "diff", met = "FDR") +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0)
+ps4d <- genplot(adaptmet, cov = alpha, type = "diff", met = "TPR") +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0)
+
+## create title for SFigure Sim4
+ps4Title <- ggdraw() +
+    draw_label("Simulation performance with modified AdaPT parameters",
+               fontface = 'bold', size = 10)
+
+## pull SFigure Sim4 together
+SFigS4 <- plot_grid(ps4a + guides(color = FALSE, linetype = FALSE),
+                    ps4b + guides(color = FALSE, linetype = FALSE),
+                    ps4c + guides(color = FALSE, linetype = FALSE),
+                    ps4d + guides(color = FALSE, linetype = FALSE),
+                    labels = LETTERS[1:4], ncol = 2)
+SFigS4 <- plot_grid(ps4Title, SFigS4, ncol = 1, rel_heights = c(0.06, 1))
+SFigS4 <- plot_grid(SFigS4, get_legend(ps4b), rel_widths = c(4.5, 1))
+SFigS4
+ggsave(file.path(outdir, "SFigureSim4.pdf"), width=8, height=6)
+```
+
 # Session information

### 2018-09-07 manuscript_figures/manuscript_figures_sims.Rmd
Change sim plot y-axis format slightly

@@
         gp <- gp + geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2)
-        gp <- gp + scale_y_continuous(bquote(atop(Mean ~ Delta ~ .(met), (informative-uninformative))),
+        gp <- gp + scale_y_continuous(bquote(Delta ~ .(met) ~ (informative-uninformative)),
                                       labels = scales::percent)
@@
     scale_x_continuous(breaks = seq(0, .1, .02)) +
-    expand_limits(x = 0)
+    expand_limits(x = 0) +
+    theme(axis.title.y = element_text(face = "bold", size = 9))
 ps4d <- genplot(adaptmet, cov = alpha, type = "diff", met = "TPR") +
     scale_x_continuous(breaks = seq(0, .1, .02)) +
-    expand_limits(x = 0)
+    expand_limits(x = 0) +
+    theme(axis.title.y = element_text(face = "bold", size = 9))
 

### 2018-09-07 manuscript_figures/manuscript_figures.Rmd
Add TPRs to Figure S1

@@
 for (l in seq_along(resfile)){
-
   de5 <- readRDS(file=resfile[l])
@@
 FigStrong
+
+FigStrongT <- plotsim_average(tab, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE, grpVars = c("es_dist", "pi0")) +
+                ggtitle("") +
+  ylim(0,0.88) +
+  facet_grid( ~ es_dist + pi0) 
+FigStrongT
 
@@
 
+FigWeakT <- plotsim_average(tab, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE, grpVars = c("es_dist", "pi0")) +
+                ggtitle("") +
+  ylim(0,0.88) +
+  facet_grid( ~ es_dist + pi0)
+FigWeakT
+
 leg <- get_legend(FigStrong)
-Fig <- plot_grid(FigStrong + theme(legend.position="none") +  
+
+FigA <- plot_grid(FigStrong + theme(legend.position="none") +  
                    scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)), 
-                 FigWeak + theme(legend.position="none") +  
+                  FigStrongT + theme(legend.position="none") +  
                    scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)),
+                 nrow = 2)
+
+FigB <- plot_grid(FigWeak + theme(legend.position="none") +  
+                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)), 
+                  FigWeakT + theme(legend.position="none") +  
+                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)),
+                 nrow = 2)
+
+Fig <- plot_grid(FigA, FigB,
                  nrow = 2, 
@@
 plot_grid(Fig, leg, ncol=2, rel_widths = c(1,0.1))
-ggsave(file.path(outdir, "Figure_yeast_varyingpi0.pdf"), width=10, height=6.5)
+ggsave(file.path(outdir, "Figure_yeast_varyingpi0.pdf"), width=10, height=12)
 

### 2018-09-10 manuscript_figures/manuscript_figures_sims.Rmd
Add unimodal distribution panel to suppl plot

@@
 p3cdTitle <- ggdraw() +
-    draw_label("Simulation performance across varying numbers of hypotheses",
+    draw_label("Simulation performance across varying number of hypotheses",
                fontface = 'bold')
@@
 ```{r SFigureSim1, fig.width = 14, fig.height = 16}
+res_noise <- dplyr::filter(resmet, setting == "informative-cubic")
+
 ## SFigure Sim1 A,B: FDR/TPR of methods across varying test statistics distribution
-res_noise <- dplyr::filter(resmet, setting == "informative-cubic")
 ps1a <- genplot(res_noise, cov = alpha, met = "FDR") +
@@
 
-```{r SFigureSim2, fig.width = 10, fig.height = 14}
-res_ua <- dplyr::filter(resmet, setting == "uasettings", grepl("N\\(0, 1\\)", dist),
-                        signal != "bimodal")
+```{r SFigureSim2, fig.width = 12, fig.height = 16}
+res_ua <- dplyr::filter(resmet, setting == "uasettings", grepl("N\\(0, 1\\)", dist))
 
-## SFigure Sim2 A,B: FDR/TPR of methods across unimodal effect size distribution
-ps2a <- genplot(res_ua, cov = alpha, met = "FDR") +
+## SFigure Sim2 A: Distirbution of unimodal effect sizes (based on random samples)
+set.seed(100)
+uadists <- tibble(x = seq(-10, 10, .01))
+uadists <- dplyr::mutate(uadists,
+                         bimodal = density(2*sampler_bimodal(1e7), from = -10, to = 10, n = length(x))$y,
+                         flattop  = density(2*sampler_flat_top(1e7), from = -10, to = 10, n = length(x))$y,
+                         skew = density(2*sampler_skew(1e7), from = -10, to = 10, n = length(x))$y,
+                         spiky = density(2*sampler_spiky(1e7), from = -10, to = 10, n = length(x))$y)
+uadists <- tidyr::gather(uadists, signal, density, -x)
+
+ps2a <- ggplot(uadists, aes(x = x, y = density)) +
+    geom_line() +
+    theme_classic() +
+    xlab("non-null effect size") + 
+    scale_y_continuous(labels = function(x) sprintf("%.2f", x)) + 
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    facet_grid(. ~ signal)
+
+## SFigure Sim2 B,C: FDR/TPR of methods across unimodal effect size distribution
+ps2b <- genplot(res_ua, cov = alpha, met = "FDR") +
     facet_grid(. ~ signal, labeller = label_parsed) +
@@
     expand_limits(x = 0)
-ps2b <- genplot(res_ua, cov = alpha, met = "TPR") +
+ps2c <- genplot(res_ua, cov = alpha, met = "TPR") +
     facet_grid(. ~ signal, labeller = label_parsed) +
@@
 
-## SFigure Sim2 C,D: FDR/TPR difference of methods
-ps2c <- genplot(res_ua, cov = alpha, type = "diff", met = "FDR") +
+## SFigure Sim2 D,E: FDR/TPR difference of methods
+ps2d <- genplot(res_ua, cov = alpha, type = "diff", met = "FDR") +
     facet_grid(. ~ signal, labeller = label_parsed) +
@@
     ylab(expression(paste("Mean ", Delta, "FDR (informative-uninformative)")))
-ps2d <- genplot(res_ua, cov = alpha, type = "diff", met = "TPR") +
+ps2e <- genplot(res_ua, cov = alpha, type = "diff", met = "TPR") +
     facet_grid(. ~ signal, labeller = label_parsed) +
@@
                     ps2d + guides(color = FALSE, linetype = FALSE),
-                    labels = LETTERS[1:4], ncol = 1)
+                    ps2e + guides(color = FALSE, linetype = FALSE),
+                    labels = LETTERS[1:5], ncol = 1, rel_heights = c(.6, rep(1, 4)))
 SFigS2 <- plot_grid(ps2Title, SFigS2, ncol = 1, rel_heights = c(0.03, 1))
-SFigS2 <- plot_grid(SFigS2, get_legend(ps2a), rel_widths = c(6.5, 1))
+SFigS2 <- plot_grid(SFigS2, get_legend(ps2b), rel_widths = c(6.5, 1))
 SFigS2
-ggsave(file.path(outdir, "SFigureSim2.pdf"), width=10, height=14)
+ggsave(file.path(outdir, "SFigureSim2.pdf"), width=12, height=16)
 ```

### 2018-09-10 manuscript_figures/manuscript_figures_sims.Rmd
Add panels of test statistic distributions

@@
 
-```{r SFigureSim1, fig.width = 14, fig.height = 16}
+```{r SFigureSim1, fig.width = 12, fig.height = 16}
 res_noise <- dplyr::filter(resmet, setting == "informative-cubic")
 
-## SFigure Sim1 A,B: FDR/TPR of methods across varying test statistics distribution
-ps1a <- genplot(res_noise, cov = alpha, met = "FDR") +
+## SFigure Sim2 A: Distirbution of unimodal effect sizes (based on random samples)
+set.seed(100)
+tsdists <- bind_rows(dplyr::mutate(simIteration(1, NULL, m = 2e4, pi0 = 0.9,
+                                                es_dist = rnorm_generator(3),
+                                                ts_dist = rnorm_perturber(1),
+                                                null_dist = rnorm_2pvaluer(1),
+                                                icovariate = runif, execute = FALSE),
+                                   dist = "\"Distribution: \" * N(0, 1)"),
+                     dplyr::mutate(simIteration(1, NULL, m = 2e4, pi0 = 0.9,
+                                                es_dist = rnorm_generator(3),
+                                                ts_dist = rt_perturber(5),
+                                                null_dist = rt_2pvaluer(5),
+                                                icovariate = runif, execute = FALSE),
+                                   dist = "\"Distribution: \" * t[5]"),
+                     dplyr::mutate(simIteration(1, NULL, m = 2e4, pi0 = 0.9,
+                                                es_dist = rnorm_generator(3),
+                                                ts_dist = rt_perturber(11),
+                                                null_dist = rt_2pvaluer(11),
+                                                icovariate = runif, execute = FALSE),
+                                   dist = "\"Distribution: \" * t[11]"),
+                     dplyr::mutate(simIteration(1, NULL, m = 2e4, pi0 = 0.9,
+                                                es_dist = function(x) { abs(rnorm_generator(15)(x)) },
+                                                ts_dist = rchisq_perturber(4),
+                                                null_dist = rchisq_pvaluer(4),
+                                                icovariate = runif, execute = FALSE),
+                                   dist = "\"Distribution: \" * {\n    chi^2\n}[4]"))
+tsdists <- dplyr::mutate(tsdists, dist = factor(dist, levels = levels(factor(dist))[c(2, 3, 4, 1)]))
+tsdists <- dplyr::rename(tsdists, truth = qvalue)
+tsdists <- dplyr::mutate(tsdists, truth = factor(truth, levels = 0:1, labels = c("null", "non-null")))
+
+## SFigure Sim1 A: Distirbution of test statistics (based on single replication)
+ps1a <- ggplot(tsdists, aes(x = test_statistic, y=..count..,
+                            group = truth, color = truth)) +
+    stat_density(geom = "line", position = "identity", adjust = 1/2) + 
+    theme_classic() +
+    scale_color_brewer("Truth", palette = "Set1", direction = -1) + 
+    scale_x_continuous("test statistic (single replication)") +
+    coord_cartesian(xlim = c(-5, 25)) + 
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    facet_grid(. ~ dist, labeller = label_parsed) +
+    theme(legend.position = c(0.99, 0.95), legend.justification = c(1, 1),
+          legend.background = element_rect(color = 'black', fill = 'grey90'))
+
+## SFigure Sim1 B,C: FDR/TPR of methods across varying test statistics distribution
+ps1b <- genplot(res_noise, cov = alpha, met = "FDR") +
     facet_grid(. ~ dist, labeller = label_parsed) +
@@
     expand_limits(x = 0)
-ps1b <- genplot(res_noise, cov = alpha, met = "TPR") +
+ps1c <- genplot(res_noise, cov = alpha, met = "TPR") +
     facet_grid(. ~ dist, labeller = label_parsed) +
@@
 
-## SFigure Sim1 C,D: FDR/TPR difference of methods
-ps1c <- genplot(res_noise, cov = alpha, type = "diff", met = "FDR") +
+## SFigure Sim1 D,E: FDR/TPR difference of methods
+ps1d <- genplot(res_noise, cov = alpha, type = "diff", met = "FDR") +
     facet_grid(. ~ dist, labeller = label_parsed) + 
@@
     expand_limits(x = 0)
-ps1d <- genplot(res_noise, cov = alpha, type = "diff", met = "TPR") +
+ps1e <- genplot(res_noise, cov = alpha, type = "diff", met = "TPR") +
     facet_grid(. ~ dist, labeller = label_parsed) + 
@@
 ## pull SFigure Sim1 together
-SFigS1 <- plot_grid(ps1a + guides(color = FALSE, linetype = FALSE),
+SFigS1 <- plot_grid(ps1a,
                     ps1b + guides(color = FALSE, linetype = FALSE),
@@
                     ps1d + guides(color = FALSE, linetype = FALSE),
-                    labels = LETTERS[1:4], ncol = 1)
+                    ps1e + guides(color = FALSE, linetype = FALSE),
+                    labels = LETTERS[1:5], ncol = 1, rel_heights = c(.6, rep(1, 4)))
 SFigS1 <- plot_grid(ps1Title, SFigS1, ncol = 1, rel_heights = c(0.03, 1))
-SFigS1 <- plot_grid(SFigS1, get_legend(ps1a), rel_widths = c(7.5, 1))
+SFigS1 <- plot_grid(SFigS1, get_legend(ps1b), rel_widths = c(6.5, 1))
 SFigS1
-ggsave(file.path(outdir, "SFigureSim1.pdf"), width=14, height=16)
+ggsave(file.path(outdir, "SFigureSim1.pdf"), width=12, height=16)
 ```
@@
     draw_label("Simulation performance across unimodal effect size distributions",
-               fontface = 'bold', size = 17)
+               fontface = 'bold', size = 20)
 

### 2018-09-11 manuscript_figures/manuscript_figures.Rmd
Add alternate color scales for Figures S4 and S9 - log-scale % and log2 Ratio

@@
   dplyr::mutate(diff_inf_percent = diff_inf / uninf) %>%
+  dplyr::mutate(diff_inf_ratio = log2(nrejects / uninf)) %>%
   dplyr::mutate(diff_inf_percent = ifelse(uninf==0 & nrejects > 0, 1, 
                                           diff_inf_percent)) %>%
+  dplyr::mutate(diff_inf_ratio = ifelse(uninf==0 & nrejects > 0, 1, 
+                                        diff_inf_ratio)) %>%
+  dplyr::mutate(diff_inf_ratio = ifelse(uninf==0 & nrejects==0, 0, 
+                                        diff_inf_ratio)) %>%
   dplyr::mutate(method = factor(method, levels = rowOrder)) %>%
@@
   group_by(dataset,cov) %>%
-  summarize(top = max(diff_inf_percent, na.rm=TRUE))
+  summarize(top = max(diff_inf_percent, na.rm=TRUE),
+            topR = max(diff_inf_ratio, na.rm =TRUE))
 
 dftab <- left_join(dftab, maxVals, by = c("dataset", "cov")) %>%
-  mutate(top = ifelse(diff_inf_percent == top, diff_inf_percent, NA))
+  dplyr::mutate(top = ifelse(diff_inf_percent == top, diff_inf_percent, NA),
+         topR = ifelse(diff_inf_ratio == topR, diff_inf_ratio, NA))
 
@@
 dftab <- dftab %>% 
-  mutate(diff_inf_percent = ifelse(nrejects == uninf, 0, diff_inf_percent))
-
-trim_tails <- function(range = c(-Inf, Inf)) scales::trans_new("trim_tails", 
+  dplyr::mutate(diff_inf_percent = ifelse(nrejects == uninf, 0, diff_inf_percent))
+
+# custom symmetric log scale for percent
+log_neg_percent <- function()
+              scales::trans_new("log_neg_percent", 
                 transform = function(x) {
-                  force(range)
-                  desired_breaks <- scales::extended_breaks(n = 7)(x[x >= range[1] & x <= range[2]])
-                  break_increment <- diff(desired_breaks)[1]
-                  x[x < range[1]] <- range[1] - break_increment
-                  x[x > range[2]] <- range[2] + break_increment
+                  x <- ifelse(x > 0, log(x + 1), -log(-x + 1))
                   x
                 },
-                inverse = function(x) x,
-
+                inverse = function(x) {
+                  x <- ifelse(x > 0, exp(x) - 1, -exp(-x) + 1)
+                  x
+                },
                 breaks = function(x) {
-                  force(range)
-                  scales::extended_breaks(n = 7)(x)
+                  if (max(x) > 200){
+                    c(-20, -5, 0, 5, 25, 100, 300)
+                  }else if (max(x) > 100){
+                    c(-20, -5, 0, 5, 25, 100, 200)
+                  }else{
+                    c(-20, -5, 0, 5, 25, 100)
+                  }
                 },
                 format = function(x) {
-                  force(range)
-                  x[1] <- paste("<", range[1])
-                  x[length(x)] <- paste(">", range[2])
                   paste0(x,"%")
                 })
+
+# log-scale color percent fill
+topRange <- max(dftab$diff_inf_percent, na.rm = TRUE)*100
+botRange <- min(dftab$diff_inf_percent, na.rm = TRUE)*100
 
@@
                              colors = c("darkblue","white","darkred"), 
-                             limits = c(-20, 100),
-                             values = scales::rescale(c(-20, 0, 100)), 
-                             trans = trim_tails(range = c(-20, 100)),
-                             oob = scales::squish) +
+                             limits = c(-20, topRange),
+                             values = scales::rescale(c(-transform(20), 0, transform(topRange))), 
+                             trans = log_neg_percent()) +
         theme_bw() +
@@
               legend.direction = "horizontal",
-              legend.text = element_text(size = 7, angle=45, margin = margin(t=5)),
+              legend.text = element_text(size = 7, angle=45, margin = margin(t=2)),
               legend.title = element_text(size = 9, margin=margin(t=-10)),
-              legend.margin = margin(t=2, r=10, b=-8, l=10))
+              legend.margin = margin(t=2, r=10, b=-6, l=10))
 
@@
 ggsave(file.path(outdir, "Figure_impactofinfcov.pdf"), width = 15.5, height = 4.25)
+
+# log-scale color log2 ratio fill
+topRange <- max(dftab$diff_inf_ratio, na.rm = TRUE)
+botRange <- min(dftab$diff_inf_ratio, na.rm = TRUE)
+
+Fig_impactofinfcov <- ggplot(dftab, aes(x = dataset, y = method, fill = diff_inf_ratio)) + 
+        geom_tile() + 
+        geom_text(aes(label = ifelse(is.na(topR), "", #adaptive text lab color
+                                     ifelse(topR > 1, signif(topR,2), ""))), 
+                  color = "white", size = 2) +
+        geom_text(aes(label = ifelse(is.na(topR), "", 
+                                     ifelse(topR <= 1, signif(topR,2), ""))), 
+                  color = "black", size = 2)+
+        scale_fill_gradientn("Fold Change\n(log2)",
+                             colors = c("darkblue","white","darkred"), 
+                             limits = c(-0.5, topRange), 
+                             values = scales::rescale(c(-0.5, 0, topRange))) +
+        theme_bw() +
+        xlab("Case Study") +
+        ylab("Method") +
+        facet_grid( ~ casestudy + cov, scales = "free", space = "free") +
+        ggtitle("Impact of informative covariate on number of rejections") +
+        theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
+              legend.position = c(0.99, 0.95), legend.justification = c(1, 1),
+              legend.background = element_rect(fill = scales::alpha("white", 3/4), color = 'black'),
+              legend.direction = "horizontal",
+              legend.text = element_text(size = 7, angle=45, margin = margin(t=0)),
+              legend.title = element_text(size = 9, margin=margin(t=-10)),
+              legend.margin = margin(t=2, r=10, b=-2, l=10))
+
+Fig_impactofinfcov
+ggsave(file.path(outdir, "Figure_impactofinfcov_fc.pdf"), width = 15.5, height = 4.25)
 
@@
   summarize(ndatasets = sum(!is.na(diff_inf_percent)),
-            diff_inf_percent = mean(diff_inf_percent, na.rm = TRUE))
+            diff_inf_percent = mean(diff_inf_percent, na.rm = TRUE), 
+            diff_inf_ratio = mean(diff_inf_ratio, na.rm = TRUE))
 
@@
   summarize(top = max(diff_inf_percent, na.rm=TRUE),
+            topR = max(diff_inf_ratio, na.rm=TRUE),
             maxd = max(ndatasets))
@@
   mutate(top = ifelse(diff_inf_percent == top, diff_inf_percent, NA)) %>%
+  mutate(topR = ifelse(diff_inf_ratio == topR, diff_inf_ratio, NA)) %>%
   ungroup() %>%
@@
 
+# log-scale color percent fill
+topRange <- max(dftab$diff_inf_percent, na.rm = TRUE)*100
+botRange <- min(dftab$diff_inf_percent, na.rm = TRUE)*100
+
 Fig_impactofinfcov <- ggplot(dftab, aes(x = casestudy, y = method, fill = diff_inf_percent*100)) + 
@@
 ggsave(file.path(outdir, "Figure_impactofinfcov_avg.pdf"), width = 5, height = 4)
+
+# log-scale color log2 ratio fill
+topRange <- max(dftab$diff_inf_ratio, na.rm = TRUE)
+botRange <- min(dftab$diff_inf_ratio, na.rm = TRUE)
+
+Fig_impactofinfcov <- ggplot(dftab, aes(x = casestudy, y = method, fill = diff_inf_ratio)) + 
+        geom_tile() + 
+        scale_fill_gradientn("Fold Change\n(log2)",
+                             colors = c("darkblue","white","darkred"), 
+                             limits = c(-0.5, topRange), 
+                             values = scales::rescale(c(-0.5, 0, topRange))) +
+        theme_bw() +
+        xlab("Case Study") +
+        ylab("Method") +
+        ggtitle("Impact of informative covariate") +
+        theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
+        geom_text(aes(label = ifelse(is.na(topR), "", #adaptive text lab color
+                                     ifelse(topR > 1, signif(topR, 2), ""))), 
+                  color = "white", size = 2) +
+        geom_text(aes(label = ifelse(is.na(topR), "", 
+                                     ifelse(topR <= 1, signif(topR, 2), ""))), 
+                  color = "black", size = 2)  +
+        theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
+              legend.position = c(0.945, 0.94), legend.justification = c(1, 1),
+              legend.background = element_rect(fill = scales::alpha("white", 3/4), color = 'black'),
+              legend.direction = "horizontal",
+              legend.text = element_text(size = 7, angle=45, margin = margin(t=2)),
+              legend.title = element_text(size = 9, margin=margin(t=-10)),
+              legend.margin = margin(t=2, r=10, b=-6, l=10))
+
+Fig_impactofinfcov
+ggsave(file.path(outdir, "Figure_impactofinfcov_avg_fc.pdf"), width = 5, height = 4)
 

### 2018-09-11 manuscript_figures/manuscript_figures.Rmd
Explicitly save alternate combined version of Fig S4; make legend labels more clear

@@
 
-Fig_impactofinfcov <- ggplot(dftab, aes(x = dataset, y = method, fill = diff_inf_ratio)) + 
+Fig_impactofinfcov_fc <- ggplot(dftab, aes(x = dataset, y = method, fill = diff_inf_ratio)) + 
         geom_tile() + 
@@
 
-Fig_impactofinfcov
+Fig_impactofinfcov_fc
 ggsave(file.path(outdir, "Figure_impactofinfcov_fc.pdf"), width = 15.5, height = 4.25)
@@
 dftab <- left_join(dftab, maxVals, by = c("casestudy")) %>%
-  mutate(top = ifelse(diff_inf_percent == top, diff_inf_percent, NA)) %>%
-  mutate(topR = ifelse(diff_inf_ratio == topR, diff_inf_ratio, NA)) %>%
+  dplyr::mutate(top = ifelse(diff_inf_percent == top, diff_inf_percent, NA)) %>%
+  dplyr::mutate(topR = ifelse(diff_inf_ratio == topR, diff_inf_ratio, NA)) %>%
   ungroup() %>%
-  mutate(casestudy = factor(casestudy, levels = colOrder)) %>%
-  mutate(casestudyN = paste0(casestudy, "(", maxd, ")")) %>%
-  mutate(casestudyN = factor(casestudyN, 
+  dplyr::mutate(casestudy = factor(casestudy, levels = colOrder)) %>%
+  dplyr::mutate(casestudyN = paste0(casestudy, "(", maxd, ")")) %>%
+  dplyr::mutate(casestudyN = factor(casestudyN, 
                              levels = unique(casestudyN)[unique(as.numeric(casestudy))]))
@@
 
-Fig_impactofinfcov <- ggplot(dftab, aes(x = casestudy, y = method, fill = diff_inf_ratio)) + 
+Fig_impactofinfcov_fc <- ggplot(dftab, aes(x = casestudy, y = method, fill = diff_inf_ratio)) + 
         geom_tile() + 
@@
 
-Fig_impactofinfcov
+Fig_impactofinfcov_fc
 ggsave(file.path(outdir, "Figure_impactofinfcov_avg_fc.pdf"), width = 5, height = 4)
@@
                   color = "black", size = 2) +
-        scale_fill_gradientn("% Change\nin FDR",
+        scale_fill_gradientn("Absolute\nFDR change",
                              labels = scales::percent, 
@@
                   color = "black", size = 2) +
-        scale_fill_gradientn("% Change\nin TPR",
+        scale_fill_gradientn("Absolute\nTPR change",
                              labels = scales::percent, 
@@
 
+ggdraw() +
+  draw_plot(pfdr, 0.53, 0.025, 0.24, 0.93) +
+  draw_plot(ptpr, 0.77, 0.025, 0.24, 0.93) +
+  draw_plot(Fig_impactofinfcov_fc, 0, 0, 0.52, 1) +
+  draw_plot_label(c("A", "B"), c(0, 0.53), c(1, 1), size = 15) +
+  draw_plot_label("Simulation", 0.71, 0.06, size =11, fontface = "plain")
+ggsave(file.path(outdir, "Figure_impactofinfcov_avg_sim_fc.pdf"), width=9.5, height=5)
+
 ```

### 2018-09-11 datasets/RNA-Seq/yeast-simulation-bimodalalternative-highpi0.Rmd
More stringent criteria for non-null null splits

@@
 library(R.utils)
+library(Hotelling)
 
@@
   # make sure null comparison is truly null: if PC1 or PC2 sig different,
+  # or if test of join means of PCs 1:4 is significant, 
   # reshuffle sample labels
   dds_test <- estimateSizeFactors(dds_test)
-  x <- t(counts(dds_test, normalize=TRUE))
-  pc <- prcomp(log(x + 0.5), scale.=TRUE)
-
-  a1 <- pc$x[colData(dds_test)$fake=="A",1]
-  b1 <- pc$x[colData(dds_test)$fake=="B",1] 
-  a2 <- pc$x[colData(dds_test)$fake=="A",2]
-  b2 <- pc$x[colData(dds_test)$fake=="B",2] 
-  p1 <- wilcox.test(a1, b1)$p.value
-  p2 <- wilcox.test(a2, b2)$p.value
-  tries <- 0
-
-  while(p1 < 0.10 || p2 < 0.10 && tries < 5){
-    colData(dds_test)$fake <- sample(colData(dds_test)$fake, ncol(dds_test))
   x <- t(counts(dds_test, normalize=TRUE))
@@
   p2 <- t.test(a2, b2)$p.value
-
+  p3 <- hotelling.test(pc$x[colData(dds_test)$fake == "A", 1:4],
+                       pc$x[colData(dds_test)$fake == "B", 1:4])$pval
+  tries <- 0
+
+  while(p1 < 0.20 || p2 < 0.20 || p3 < 0.20 && tries < 10){
+    colData(dds_test)$fake <- sample(colData(dds_test)$fake, ncol(dds_test))
+    x <- t(counts(dds_test, normalize=TRUE))
+    pc <- prcomp(log(x + 0.5), scale.=TRUE)
+
+    a1 <- pc$x[colData(dds_test)$fake=="A",1]
+    b1 <- pc$x[colData(dds_test)$fake=="B",1] 
+    a2 <- pc$x[colData(dds_test)$fake=="A",2]
+    b2 <- pc$x[colData(dds_test)$fake=="B",2] 
+    p1 <- t.test(a1, b1)$p.value
+    p2 <- t.test(a2, b2)$p.value
+    p3 <- hotelling.test(pc$x[colData(dds_test)$fake == "A", 1:4],
+                         pc$x[colData(dds_test)$fake == "B", 1:4])$pval
     tries <- tries + 1

### 2018-09-11 datasets/RNA-Seq/yeast-simulation-bimodalalternative.Rmd
More stringent criteria for non-null null splits

@@
   # make sure null comparison is truly null: if PC1 or PC2 sig different,
+  # or if test of join means of PCs 1:4 is significant,  
   # reshuffle sample labels
   dds_test <- estimateSizeFactors(dds_test)
-  x <- t(counts(dds_test, normalize=TRUE))
-  pc <- prcomp(log(x + 0.5), scale.=TRUE)
-
-  a1 <- pc$x[colData(dds_test)$fake=="A",1]
-  b1 <- pc$x[colData(dds_test)$fake=="B",1] 
-  a2 <- pc$x[colData(dds_test)$fake=="A",2]
-  b2 <- pc$x[colData(dds_test)$fake=="B",2] 
-  p1 <- wilcox.test(a1, b1)$p.value
-  p2 <- wilcox.test(a2, b2)$p.value
-  tries <- 0
-
-  while(p1 < 0.10 || p2 < 0.10 && tries < 5){
-    colData(dds_test)$fake <- sample(colData(dds_test)$fake, ncol(dds_test))
   x <- t(counts(dds_test, normalize=TRUE))
@@
   p2 <- t.test(a2, b2)$p.value
+  p3 <- hotelling.test(pc$x[colData(dds_test)$fake == "A", 1:4],
+                       pc$x[colData(dds_test)$fake == "B", 1:4])$pval
+  tries <- 0
+
+  while(p1 < 0.20 || p2 < 0.20 || p3 < 0.20 && tries < 10){
+    colData(dds_test)$fake <- sample(colData(dds_test)$fake, ncol(dds_test))
+    x <- t(counts(dds_test, normalize=TRUE))
+    pc <- prcomp(log(x + 0.5), scale.=TRUE)
+
+    a1 <- pc$x[colData(dds_test)$fake=="A",1]
+    b1 <- pc$x[colData(dds_test)$fake=="B",1] 
+    a2 <- pc$x[colData(dds_test)$fake=="A",2]
+    b2 <- pc$x[colData(dds_test)$fake=="B",2] 
+    p1 <- t.test(a1, b1)$p.value
+    p2 <- t.test(a2, b2)$p.value
+    p3 <- hotelling.test(pc$x[colData(dds_test)$fake == "A", 1:4],
+                        pc$x[colData(dds_test)$fake == "B", 1:4])$pval
 

### 2018-09-11 datasets/RNA-Seq/yeast-simulation-highpi0.Rmd
More stringent criteria for non-null null splits

@@
   # make sure null comparison is truly null: if PC1 or PC2 sig different, 
+  # or if test of join means of PCs 1:4 is significant, 
   # reshuffle sample labels
   dds_test <- estimateSizeFactors(dds_test)
-  x <- t(counts(dds_test, normalize=TRUE))
-  pc <- prcomp(log(x + 0.5), scale.=TRUE)
-
-  a1 <- pc$x[colData(dds_test)$fake=="A",1]
-  b1 <- pc$x[colData(dds_test)$fake=="B",1] 
-  a2 <- pc$x[colData(dds_test)$fake=="A",2]
-  b2 <- pc$x[colData(dds_test)$fake=="B",2] 
-  p1 <- wilcox.test(a1, b1)$p.value
-  p2 <- wilcox.test(a2, b2)$p.value
-  tries <- 0
-
-  while(p1 < 0.10 || p2 < 0.10 && tries < 5){
-    colData(dds_test)$fake <- sample(colData(dds_test)$fake, ncol(dds_test))
   x <- t(counts(dds_test, normalize=TRUE))
@@
   p2 <- t.test(a2, b2)$p.value
+  p3 <- hotelling.test(pc$x[colData(dds_test)$fake == "A", 1:4],
+                       pc$x[colData(dds_test)$fake == "B", 1:4])$pval
+  tries <- 0
+
+  while(p1 < 0.20 || p2 < 0.20 || p3 < 0.20 && tries < 10){
+    colData(dds_test)$fake <- sample(colData(dds_test)$fake, ncol(dds_test))
+    x <- t(counts(dds_test, normalize=TRUE))
+    pc <- prcomp(log(x + 0.5), scale.=TRUE)
+
+    a1 <- pc$x[colData(dds_test)$fake=="A",1]
+    b1 <- pc$x[colData(dds_test)$fake=="B",1] 
+    a2 <- pc$x[colData(dds_test)$fake=="A",2]
+    b2 <- pc$x[colData(dds_test)$fake=="B",2] 
+    p1 <- t.test(a1, b1)$p.value
+    p2 <- t.test(a2, b2)$p.value
+    p3 <- hotelling.test(pc$x[colData(dds_test)$fake == "A", 1:4],
+                         pc$x[colData(dds_test)$fake == "B", 1:4])$pval
 

### 2018-09-11 datasets/RNA-Seq/yeast-simulation.Rmd
More stringent criteria for non-null null splits

@@
   # make sure null comparison is truly null: if PC1 or PC2 sig different, 
+  # or if test of join means of PCs 1:4 is significant,  
   # reshuffle sample labels
   dds_test <- estimateSizeFactors(dds_test)
-  x <- t(counts(dds_test, normalize=TRUE))
-  pc <- prcomp(log(x + 0.5), scale.=TRUE)
-
-  a1 <- pc$x[colData(dds_test)$fake=="A",1]
-  b1 <- pc$x[colData(dds_test)$fake=="B",1] 
-  a2 <- pc$x[colData(dds_test)$fake=="A",2]
-  b2 <- pc$x[colData(dds_test)$fake=="B",2] 
-  p1 <- wilcox.test(a1, b1)$p.value
-  p2 <- wilcox.test(a2, b2)$p.value
-  tries <- 0
-
-  while(p1 < 0.10 || p2 < 0.10 && tries < 5){
-    colData(dds_test)$fake <- sample(colData(dds_test)$fake, ncol(dds_test))
   x <- t(counts(dds_test, normalize=TRUE))

### 2018-09-12 datasets/simulations/supplementary/simulations-uasettings-nonnull25.Rmd
Add/modify descriptions of UA setting Rmds; add UA sims w more non-null

--- 2018-09-06 datasets/simulations/simulations-uasettings.Rmd
@@
 ---
-title: "Simulation Study: Unimodal Settings"
+title: "Simulation Study: Unimodal Settings (25% non-null)"
 author: "Patrick Kimes"
@@
 setting. An informative covariate is included in the setting as described in
-`simulations-informative-cubic.Rmd`.
+`simulations-informative-cubic.Rmd`. The effect sizes for non-null tests are
+sampled from unimodal distributions composed of a mixture of normal distributions,
+as described in the original adaptive shrinkage (ASH) manuscript (Stephens, 2016).
+
+The simulation settings considered in this study are identical to those carried
+out in `simulations-uasettings.Rmd`, with the exception that the proportion of null
+hypotheses (pi0) is decreased from 90% to 75%. Here, we investigate the behavior of
+FDR-controlling methods under the unimodal setting when the proportion of non-null
+hypotheses is substantially larger (25%).
 
@@
 ## intermediary files we create below
-spiky_file <- file.path(resdir, "uasettings-benchmark-spiky.rds")
-flattop_file <- file.path(resdir, "varyinges-benchmark-flattop.rds")
-skew_file <- file.path(resdir, "uasettings-benchmark-skew.rds")
-bimodal_file <- file.path(resdir, "uasettings-benchmark-bimodal.rds")
+spiky_file <- file.path(resdir, "uasettings-25-benchmark-spiky.rds")
+flattop_file <- file.path(resdir, "uasettings-25-benchmark-flattop.rds")
+skew_file <- file.path(resdir, "uasettings-25-benchmark-skew.rds")
+bimodal_file <- file.path(resdir, "uasettings-25-benchmark-bimodal.rds")
 
@@
 m <- 20000                          # integer: number of hypothesis tests
-pi0 <- pi0_cubic(0.90)              # numeric: proportion of null hypotheses
+pi0 <- pi0_cubic(0.75)              # numeric: proportion of null hypotheses
 ts_dist <- rnorm_perturber(1)       # functional: sampling dist/noise for test stats

### 2018-09-12 datasets/RNA-Seq/yeast-simulation-bimodalalternative-highpi0.Rmd
Redefine informative covariate in yeast

@@
 
-We will investigate the overall mean expression as an informative and independent 
-(under the null hypothesis) covariate. 
+We will simulate a strongly informative and weakly informative independent covariate.
 We will also investigate the effects of an uninformative covariate by using a 
@@
 resfile_d10 <- file.path(resdir, "yeastIIH-results-de10.rds")
-resfile_d5_uw <- file.path(resdir, "yeastIIHU-results-de5.rds")
-resfile_d10_uw <- file.path(resdir, "yeastIIHU-results-de10.rds")
+resfile_d5_w <- file.path(resdir, "yeastIIHW-results-de5.rds")
+resfile_d10_w <- file.path(resdir, "yeastIIHW-results-de10.rds")
 
@@
 resfile_d10_uninfCov <- file.path(resdir, "yeastIIH-results-de10-uninfCov.rds")
-resfile_d5_uninfCov_uw <- file.path(resdir, "yeastIIHU-results-de5-uninfCov.rds")
-resfile_d10_uninfCov_uw <- file.path(resdir, "yeastIIHU-results-de10-uninfCov.rds")
+resfile_d5_uninfCov_w <- file.path(resdir, "yeastIIHW-results-de5-uninfCov.rds")
+resfile_d10_uninfCov_w <- file.path(resdir, "yeastIIHW-results-de10-uninfCov.rds")
 
@@
 #'  will specify only one worker per DEseq intance.
-#' @param uninformativeCovariate logical indicating whether to replace overall mean 
-#'  independent covariate with a randomly generated (uninformative) covariate. 
+#' @param uninformativeCovariate logical indicating whether to replace the
+#'  informative covariate with a randomly generated (uninformative) covariate. 
 #'  Default is FALSE.
@@
                              bd, uninformativeCovariate = FALSE, pvalHists = FALSE,
-                             weightByCovariate = TRUE){
-
-  uninf_covar <- rnorm(nrow(dds))
+                             strongCovariate = TRUE){
 
@@
   set.seed(as.numeric(X)*as.numeric(rseed))
+
+  covar <- abs(rnorm(nrow(dds)))
 
@@
   dds_test <- dds_test[pzero < 0.5,]
-  uninf_covar <- uninf_covar[pzero < 0.5]
+  covar <- covar[pzero < 0.5]
 
@@
   p2 <- t.test(a2, b2)$p.value
-  p3 <- hotelling.test(pc$x[colData(dds_test)$fake == "A", 1:4],
-                       pc$x[colData(dds_test)$fake == "B", 1:4])$pval
   tries <- 0
 
-  while(p1 < 0.20 || p2 < 0.20 || p3 < 0.20 && tries < 10){
+  while(p1 < 0.10 || p2 < 0.10 && tries < 10){
     colData(dds_test)$fake <- sample(colData(dds_test)$fake, ncol(dds_test))
@@
     p2 <- t.test(a2, b2)$p.value
-    p3 <- hotelling.test(pc$x[colData(dds_test)$fake == "A", 1:4],
-                         pc$x[colData(dds_test)$fake == "B", 1:4])$pval
     tries <- tries + 1
@@
     # pick random set of nDE genes to add signal to
-    if(weightByCovariate){
-      wts <- rowMeans(counts(dds_test, normalize = TRUE))
-    }else{
-      wts <- rep(1, nrow(dds_test))
+    if(strongCovariate){
+      wts <- covar
+    }else{
+      wts <- abs(covar + rnorm(length(covar)))
     }
@@
                                SE=resTEST$lfcSE,                 
-                               ind_covariate = resTEST$baseMean,
+                               ind_covariate = covar,
                                effect_size = resTEST$log2FoldChange, 
@@
   if (uninformativeCovariate)
-    geneExp <- mutate(geneExp, ind_covariate = uninf_covar)
+    geneExp <- mutate(geneExp, ind_covariate = abs(rnorm(length(covar))))
 
@@
 
-# Sim D5: Unweighted 5v5
-
-Here we'll repeat the above, but not explicitly weighting the DE gene selection by
-the informative covariate.
+# Sim D5: 5v5 Weak Covariate
+
+Here we'll repeat the above, but using a weaker informative covariate (that has noise
+added to it).
 
@@
 
-if (!file.exists(resfile_d5_uw)){
+if (!file.exists(resfile_d5_w)){
   de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores,
-                  weightByCovariate = FALSE)
-  saveRDS(de5, file = resfile_d5_uw)
-}else{
-  de5 <- readRDS(file = resfile_d5_uw)
+                  strongCovariate = FALSE)
+  saveRDS(de5, file = resfile_d5_w)
+}else{
+  de5 <- readRDS(file = resfile_d5_w)
 }
@@
 
-# Sim D10: Unweighted 10v10
-
-Here we'll repeat the above, but not explicitly weighting the DE gene selection by
-the informative covariate.
+# Sim D10: 10v10 Weak Covariate
+
+Here we'll repeat the above, but using a weaker informative covariate (that has noise
+added to it).
 
@@
 
-if(!file.exists(resfile_d10_uw)){
+if(!file.exists(resfile_d10_w)){
   de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores,
-                  weightByCovariate = FALSE)
-  saveRDS(de10, file = resfile_d10_uw)
-}else{
-  de10 <- readRDS(file = resfile_d10_uw)
+                  strongCovariate = FALSE)
+  saveRDS(de10, file = resfile_d10_w)
+}else{
+  de10 <- readRDS(file = resfile_d10_w)
 }
@@
 
-Here we repeat the previous four sections, using an uninformative covariate 
-instead of the overall mean expression.
+Here we repeat the previous four sections, using an uninformative covariate.
 
@@
 
-## Sim D5: Unweighted 5v5
-
-Here we'll repeat the above, but not explicitly weighting the DE gene selection by
-the informative covariate.
+## Sim D5: 5v5 Weak Covariate
+
+Here we'll repeat the above, but using a weaker informative covariate (that has noise
+added to it).
 This will be done for 100 random splits.
@@
 
-if (!file.exists(resfile_d5_uninfCov_uw)){
+if (!file.exists(resfile_d5_uninfCov_w)){
   de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
@@
                   uninformativeCovariate = TRUE, 
-                  weightByCovariate = FALSE,
+                  strongCovariate = FALSE,
                   mc.cores=nCores)
-  saveRDS(de5, file = resfile_d5_uninfCov_uw)
-}else{
-  de5 <- readRDS(file = resfile_d5_uninfCov_uw)
+  saveRDS(de5, file = resfile_d5_uninfCov_w)
+}else{
+  de5 <- readRDS(file = resfile_d5_uninfCov_w)
 }
@@
 
-## Sim D10: Unweighted 10v10
-
-Here we'll repeat the above, but not explicitly weighting the DE gene selection by
-the informative covariate.
+## Sim D10: 10v10 Weak Covariate
+
+Here we'll repeat the above, but using a weaker informative covariate (that has noise
+added to it).
 
@@
 
-if(!file.exists(resfile_d10_uninfCov_uw)){
+if(!file.exists(resfile_d10_uninfCov_w)){
   de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
@@
                    uninformativeCovariate = TRUE, 
-                   weightByCovariate = FALSE,
+                   strongCovariate = FALSE,
                    mc.cores=nCores)
-  saveRDS(de10, file = resfile_d10_uninfCov_uw)
-}else{
-  de10 <- readRDS(file = resfile_d10_uninfCov_uw)
+  saveRDS(de10, file = resfile_d10_uninfCov_w)
+}else{
+  de10 <- readRDS(file = resfile_d10_uninfCov_w)
 }
@@
 ```{r, fig.width=18, fig.height=12}
-plotMethodRanks(c(resfile_d5, resfile_d5_uw, resfile_d10, resfile_d10_uw,
-                  resfile_d5_uninfCov, resfile_d5_uninfCov_uw, resfile_d10_uninfCov, resfile_d10_uninfCov_uw), 
+plotMethodRanks(c(resfile_d5, resfile_d5_w, resfile_d10, resfile_d10_w,
+                  resfile_d5_uninfCov, resfile_d5_uninfCov_w, resfile_d10_uninfCov, resfile_d10_uninfCov_w), 
                 colLabels = c("5", "5 UW", "10", "10 UW",

### 2018-09-12 datasets/RNA-Seq/yeast-simulation-bimodalalternative.Rmd
Redefine informative covariate in yeast

--- 2018-08-15 datasets/RNA-Seq/yeast-simulation-bimodalalternative.Rmd
@@
 resfile_d10 <- file.path(resdir, "yeastII-results-de10.rds")
-resfile_d5_uw <- file.path(resdir, "yeastIIU-results-de5.rds")
-resfile_d10_uw <- file.path(resdir, "yeastIIU-results-de10.rds")
+resfile_d5_w <- file.path(resdir, "yeastIIW-results-de5.rds")
+resfile_d10_w <- file.path(resdir, "yeastIIW-results-de10.rds")
 
@@
 resfile_d10_uninfCov <- file.path(resdir, "yeastII-results-de10-uninfCov.rds")
-resfile_d5_uninfCov_uw <- file.path(resdir, "yeastIIU-results-de5-uninfCov.rds")
-resfile_d10_uninfCov_uw <- file.path(resdir, "yeastIIU-results-de10-uninfCov.rds")
+resfile_d5_uninfCov_w <- file.path(resdir, "yeastIIW-results-de5-uninfCov.rds")
+resfile_d10_uninfCov_w <- file.path(resdir, "yeastIIW-results-de10-uninfCov.rds")
 
@@
   set.seed(as.numeric(X)*as.numeric(rseed))
+
+  covar <- abs(rnorm(nrow(dds_full)))
 
@@
   p2 <- t.test(a2, b2)$p.value
+  tries <- 0
+
+  while(p1 < 0.10 || p2 < 0.10 && tries < 10){
+    colData(dds_test)$fake <- sample(colData(dds_test)$fake, ncol(dds_test))
+    x <- t(counts(dds_test, normalize=TRUE))
+    pc <- prcomp(log(x + 0.5), scale.=TRUE)
+
+    a1 <- pc$x[colData(dds_test)$fake=="A",1]
+    b1 <- pc$x[colData(dds_test)$fake=="B",1] 
+    a2 <- pc$x[colData(dds_test)$fake=="A",2]
+    b2 <- pc$x[colData(dds_test)$fake=="B",2] 
+    p1 <- t.test(a1, b1)$p.value
+    p2 <- t.test(a2, b2)$p.value
 
@@
 
-if (!file.exists(resfile_d5_uninfCov_uw)){
+if (!file.exists(resfile_d5_uninfCov_w)){
   de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd,
-                  uninformativeCovariate = TRUE, mc.cores=nCores)
-  saveRDS(de5, file = resfile_d5_uninfCov_uw)
-}else{
-  de5 <- readRDS(file = resfile_d5_uninfCov_uw)
+                  uninformativeCovariate = TRUE, 
+                  strongCovariate = FALSE,
+                  mc.cores=nCores)
+  saveRDS(de5, file = resfile_d5_uninfCov_w)
+}else{
+  de5 <- readRDS(file = resfile_d5_uninfCov_w)
 }
@@
 
-if(!file.exists(resfile_d10_uninfCov_uw)){
+if(!file.exists(resfile_d10_uninfCov_w)){
   de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                    nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   uninformativeCovariate = TRUE, mc.cores=nCores)
-  saveRDS(de10, file = resfile_d10_uninfCov_uw)
-}else{
-  de10 <- readRDS(file = resfile_d10_uninfCov_uw)
+                   uninformativeCovariate = TRUE, 
+                   strongCovariate = FALSE,
+                   mc.cores=nCores)
+  saveRDS(de10, file = resfile_d10_uninfCov_w)
+}else{
+  de10 <- readRDS(file = resfile_d10_uninfCov_w)
 }
@@
 
-Here we compare the method ranks for the different sample sizes, weighting, and
+Here we compare the method ranks for the different sample sizes, and informative
 covariate settings at alpha = 0.10.

### 2018-09-12 datasets/RNA-Seq/yeast-simulation-highpi0.Rmd
Redefine informative covariate in yeast

--- 2018-08-15 datasets/RNA-Seq/yeast-simulation-highpi0.Rmd
@@
 resfile_d10 <- file.path(resdir, "yeastH-results-de10.rds")
-resfile_d5_uw <- file.path(resdir, "yeastHU-results-de5.rds")
-resfile_d10_uw <- file.path(resdir, "yeastHU-results-de10.rds")
+resfile_d5_w <- file.path(resdir, "yeastHW-results-de5.rds")
+resfile_d10_w <- file.path(resdir, "yeastHW-results-de10.rds")
 
@@
 resfile_d10_uninfCov <- file.path(resdir, "yeastH-results-de10-uninfCov.rds")
-resfile_d5_uninfCov_uw <- file.path(resdir, "yeastHU-results-de5-uninfCov.rds")
-resfile_d10_uninfCov_uw <- file.path(resdir, "yeastHU-results-de10-uninfCov.rds")
+resfile_d5_uninfCov_w <- file.path(resdir, "yeastHW-results-de5-uninfCov.rds")
+resfile_d10_uninfCov_w <- file.path(resdir, "yeastHW-results-de10-uninfCov.rds")
 
@@
                              bd, uninformativeCovariate = FALSE, pvalHists = FALSE,
-                             weightByCovariate = TRUE){
-
-  uninf_covar <- rnorm(nrow(dds_full))
-
+                             strongCovariate = TRUE){
   # set random seed
   set.seed(as.numeric(X)*as.numeric(rseed))
+
+  covar <- abs(rnorm(nrow(dds_full)))
 
@@
     # pick random set of nDE genes to add signal to
-    if(weightByCovariate){
-      wts <- rowMeans(counts(dds_test, normalize = TRUE))
-    }else{
-      wts <- rep(1, nrow(dds_test))
+     if(strongCovariate){
+      wts <- covar
+    }else{
+      wts <- abs(covar + rnorm(length(covar)))
     }
@@
 
-if(!file.exists(resfile_d5_uninfCov_uw)){
+if(!file.exists(resfile_d5_uninfCov_w)){
   de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd,
-                  uninformativeCovariate = TRUE, mc.cores=nCores)
-  saveRDS(de5, file = resfile_d5_uninfCov_uw)
-}else{
-  de5 <- readRDS(file = resfile_d5_uninfCov_uw)
+                  uninformativeCovariate = TRUE, 
+                  strongCovariate=FALSE,
+                  mc.cores=nCores)
+  saveRDS(de5, file = resfile_d5_uninfCov_w)
+}else{
+  de5 <- readRDS(file = resfile_d5_uninfCov_w)
 }

### 2018-09-12 datasets/RNA-Seq/yeast-simulation.Rmd
Redefine informative covariate in yeast

--- 2018-08-15 datasets/RNA-Seq/yeast-simulation.Rmd
@@
 resfile_d10 <- file.path(resdir, "yeast-results-de10.rds")
-resfile_d5_uw <- file.path(resdir, "yeastU-results-de5.rds")
-resfile_d10_uw <- file.path(resdir, "yeastU-results-de10.rds")
+resfile_d5_w <- file.path(resdir, "yeastW-results-de5.rds")
+resfile_d10_w <- file.path(resdir, "yeastW-results-de10.rds")
 
@@
 resfile_d10_uninfCov <- file.path(resdir, "yeast-results-de10-uninfCov.rds")
-resfile_d5_uninfCov_uw <- file.path(resdir, "yeastU-results-de5-uninfCov.rds")
-resfile_d10_uninfCov_uw <- file.path(resdir, "yeastU-results-de10-uninfCov.rds")
+resfile_d5_uninfCov_w <- file.path(resdir, "yeastW-results-de5-uninfCov.rds")
+resfile_d10_uninfCov_w <- file.path(resdir, "yeastW-results-de10-uninfCov.rds")
 
@@
                              bd, uninformativeCovariate = FALSE, pvalHists = FALSE,
-                             weightByCovariate = TRUE){
-
-  uninf_covar <- rnorm(nrow(dds_full))
+                             strongCovariate = TRUE){
 
@@
   p2 <- t.test(a2, b2)$p.value
-
+  tries <- 0
+
+  while(p1 < 0.10 || p2 < 0.10  && tries < 10){
+    colData(dds_test)$fake <- sample(colData(dds_test)$fake, ncol(dds_test))
+    x <- t(counts(dds_test, normalize=TRUE))
+    pc <- prcomp(log(x + 0.5), scale.=TRUE)
+
+    a1 <- pc$x[colData(dds_test)$fake=="A",1]
+    b1 <- pc$x[colData(dds_test)$fake=="B",1] 
+    a2 <- pc$x[colData(dds_test)$fake=="A",2]
+    b2 <- pc$x[colData(dds_test)$fake=="B",2] 
+    p1 <- t.test(a1, b1)$p.value
+    p2 <- t.test(a2, b2)$p.value
     tries <- tries + 1
@@
 
-if(!file.exists(resfile_d10_uninfCov_uw)){
+if(!file.exists(resfile_d10_uninfCov_w)){
   de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
                    nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   uninformativeCovariate = TRUE, mc.cores=nCores)
-  saveRDS(de10, file = resfile_d10_uninfCov_uw)
-}else{
-  de10 <- readRDS(file = resfile_d10_uninfCov_uw)
+                   uninformativeCovariate = TRUE, 
+                   weakCovariate = FALSE,
+                   mc.cores=nCores)
+  saveRDS(de10, file = resfile_d10_uninfCov_w)
+}else{
+  de10 <- readRDS(file = resfile_d10_uninfCov_w)
 }
@@
 
-Here we compare the method ranks for the different sample sizes, weighting, and
-covariate settings at alpha = 0.10.
+Here we compare the method ranks for the different sample sizes and informativeness
+settings at alpha = 0.10.
 
 ```{r, fig.width=18, fig.height=12}
-plotMethodRanks(c(resfile_d5, resfile_d5_uw, resfile_d10, resfile_d10_uw,
-                  resfile_d5_uninfCov, resfile_d5_uninfCov_uw, resfile_d10_uninfCov, resfile_d10_uninfCov_uw), 
+plotMethodRanks(c(resfile_d5, resfile_d5_w, resfile_d10, resfile_d10_w,
+                  resfile_d5_uninfCov, resfile_d5_uninfCov_w, resfile_d10_uninfCov, resfile_d10_uninfCov_w), 
                 colLabels = c("5", "5 UW", "10", "10 UW",

### 2018-09-12 manuscript_figures/manuscript_figures.Rmd
Redefine informative covariate in yeast

@@
 objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
-resfile <- objects[(grepl("yeastU-", objects) | grepl("yeastHU-", objects) | 
-                    grepl("yeastIIU-", objects) | grepl("yeastIIHU-", objects)) & 
+resfile <- objects[(grepl("yeastW-", objects) | grepl("yeastHW-", objects) | 
+                    grepl("yeastIIW-", objects) | grepl("yeastIIHW-", objects)) & 
                    grepl("de5", objects) & 

### 2018-09-13 datasets/RNA-Seq/yeast-simulation-bimodalalternative-highpi0.Rmd
Refine informative covariate in yeast
Make strongly informative covariate stronger and weakly informative
covariate weaker.

@@
 # results files that will be generated
+# strong covariate
 resfile_d5 <- file.path(resdir, "yeastIIH-results-de5.rds")
 resfile_d10 <- file.path(resdir, "yeastIIH-results-de10.rds")
+
+# weak covariate
 resfile_d5_w <- file.path(resdir, "yeastIIHW-results-de5.rds")
@@
 
+# uninformative covariate
 resfile_d5_uninfCov <- file.path(resdir, "yeastIIH-results-de5-uninfCov.rds")
 resfile_d10_uninfCov <- file.path(resdir, "yeastIIH-results-de10-uninfCov.rds")
-resfile_d5_uninfCov_w <- file.path(resdir, "yeastIIHW-results-de5-uninfCov.rds")
-resfile_d10_uninfCov_w <- file.path(resdir, "yeastIIHW-results-de10-uninfCov.rds")
 
@@
 
-  covar <- abs(rnorm(nrow(dds)))
+  covar <- 1 / (1+exp(-runif(nrow(dds_full), 0, 10) + 5))
 
@@
     # pick random set of nDE genes to add signal to
-    if(strongCovariate){
-      wts <- covar
-    }else{
-      wts <- abs(covar + rnorm(length(covar)))
-    }
-    DE <- sample(1:nrow(dds_test), nDE, prob = wts)
+    DE <- sample(1:nrow(dds_test), nDE, prob = covar)
     truth[DE] <- TRUE
@@
 
-  if (uninformativeCovariate)
-    geneExp <- mutate(geneExp, ind_covariate = abs(rnorm(length(covar))))
+  if (uninformativeCovariate){
+    geneExp <- mutate(geneExp, ind_covariate = runif(length(covar)))
+  }else if(!strongCovariate){
+    geneExp <- mutate(geneExp, ind_covariate = pmin(1, abs(covar + rnorm(length(covar), 0, 0.25))))
+  }
 
@@
 ```
-
-# Sim D5: DE Comparison 5v5
+# Non-null Comparisons with Strongly Informative Covariate
+
+## D5S: DE 5v5 Strong
 
@@
 
-## Generate a list of SB results objects
+### Generate a list of SB results objects
 
@@
 
-## Plot average results over replications
+### Plot average results over replications
 
@@
 
-# Sim D10: DE Comparison 10v10
+# Non-null Comparisons with Strongly Informative Covariate
+
+## D10S: DE 10v10 Strong
 
@@
 ```
-
-# Sim D5: 5v5 Weak Covariate
+# Non-null Comparisons with Weakly Informative Covariate
+
+## D5W: DE 5v5 Weak
 
@@
 
-# Sim D10: 10v10 Weak Covariate
+## D10W: DE 10v10 Weak
 
@@
 
-# Uninformative covariate
-
-Here we repeat the previous four sections, using an uninformative covariate.
-
-## Sim D5: DE Comparison 5v5
+# Non-null Comparisons with Uninformative Covariate
+
+Here we repeat the previous sections, using an uninformative covariate.
+
+## D5U: DE 5v5 Uninformative
 
@@
 
-## Sim D10: DE Comparison 10v10
+## D10U: DE 10v10 Uninformative
 
@@
 
-## Sim D5: 5v5 Weak Covariate
-
-Here we'll repeat the above, but using a weaker informative covariate (that has noise
-added to it).
-This will be done for 100 random splits.
-
-### Generate a list of SB results objects
-
-```{r d5-uc-uw, results='hide', message=FALSE}
-rseed <- 445
-sampleSize <- 5
-nDE <- 500
-
-if (!file.exists(resfile_d5_uninfCov_w)){
-  de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
-                  nDE=nDE, sampleSize=sampleSize, bd=bd,
-                  uninformativeCovariate = TRUE, 
-                  strongCovariate = FALSE,
-                  mc.cores=nCores)
-  saveRDS(de5, file = resfile_d5_uninfCov_w)
-}else{
-  de5 <- readRDS(file = resfile_d5_uninfCov_w)
-}
-```
-
-### Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(de5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(res5d, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res5d, met="FDR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res5d, met="TPR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(de5, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(de5, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE)
-
-}else{
-  message("Not enough methods found rejections at alpha ", ualpha, 
-          "; skipping upset plot")
-}
-
-```
-
-## Sim D10: 10v10 Weak Covariate
-
-Here we'll repeat the above, but using a weaker informative covariate (that has noise
-added to it).
-
-### Generate a list of SB results objects
-
-```{r d10-uc-uw, results='hide', message=FALSE}
-rseed <- 868
-sampleSize <- 10
-nDE <- 500
-
-if(!file.exists(resfile_d10_uninfCov_w)){
-  de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
-                   nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   uninformativeCovariate = TRUE, 
-                   strongCovariate = FALSE,
-                   mc.cores=nCores)
-  saveRDS(de10, file = resfile_d10_uninfCov_w)
-}else{
-  de10 <- readRDS(file = resfile_d10_uninfCov_w)
-}
-```
-
-### Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(de10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(res10d, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res10d, met="FDR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res10d, met="TPR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(de10, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(de10, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res10d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(de10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
-}else{
-  message("Not enough methods found rejections at alpha ", ualpha, 
-          "; skipping upset plot")
-}
-
-```
-
 # Simulation setting comparison
@@
 plotMethodRanks(c(resfile_d5, resfile_d5_w, resfile_d10, resfile_d10_w,
-                  resfile_d5_uninfCov, resfile_d5_uninfCov_w, resfile_d10_uninfCov, resfile_d10_uninfCov_w), 
-                colLabels = c("5", "5 UW", "10", "10 UW",
-                              "5 UI", "5 UI UW", "10 UI", "10 UI UW"), 
+                  resfile_d5_uninfCov, resfile_d10_uninfCov), 
+                colLabels = c("5 S", "5 W", "10 S", "10 W",
+                              "5 UI", "10 UI"), 
                 alpha = 0.10, xlab = "Comparison", 

### 2018-09-13 datasets/RNA-Seq/yeast-simulation-bimodalalternative.Rmd
Refine informative covariate in yeast
Make strongly informative covariate stronger and weakly informative
covariate weaker.

@@
 # results files that will be generated
+# strong covariate
 resfile_d5 <- file.path(resdir, "yeastII-results-de5.rds")
 resfile_d10 <- file.path(resdir, "yeastII-results-de10.rds")
+
+# weak covariate
 resfile_d5_w <- file.path(resdir, "yeastIIW-results-de5.rds")
@@
 
+# uninformative covariate
 resfile_d5_uninfCov <- file.path(resdir, "yeastII-results-de5-uninfCov.rds")
 resfile_d10_uninfCov <- file.path(resdir, "yeastII-results-de10-uninfCov.rds")
-resfile_d5_uninfCov_w <- file.path(resdir, "yeastIIW-results-de5-uninfCov.rds")
-resfile_d10_uninfCov_w <- file.path(resdir, "yeastIIW-results-de10-uninfCov.rds")
 
@@
 
-  covar <- abs(rnorm(nrow(dds_full)))
+  covar <- 1 / (1+exp(-runif(nrow(dds_full), 0, 10) + 5))
 
@@
 
-# Sim D5: DE Comparison 5v5
+# Non-null Comparisons with Strongly Informative Covariate
+
+## D5S: DE 5v5 Strong
 
 Here we'll carry out a DE comparison 5 versus 5 Snf2 samples,
-where the groups are selected randomly and 2000 DE genes are added. 
+where the groups are selected randomly and 2000 DE genes are added, and a
+strongly informative covariate is used.
 This will be done for 100 random splits. Note that we do not need to run null
@@
 
-# Sim D10: DE Comparison 10v10
+## D10S: DE 10v10 Strong
 
@@
 
-# Sim D5: 5v5 Weak Covariate
+# Non-null Comparisons with Strongly Informative Covariate
+
+## D5S: DE 5v5 Weak
 
@@
 
-# Uninformative covariate
+# Non-null Comparisons with Uninformative Covariate
 
@@
 
-## Sim D5: DE Comparison 5v5
+## D5U: DE 5v5 Uninformative
 
@@
 
-## Sim D5: 5v5 Weak Covariate
-
-Here we'll repeat the above, but using a weaker informative covariate (that has noise
-added to it).
-This will be done for 100 random splits.
-
-### Generate a list of SB results objects
-
-```{r d5-uc-uw, results='hide', message=FALSE}
-rseed <- 445
-sampleSize <- 5
-nDE <- 2000
-
-if (!file.exists(resfile_d5_uninfCov_w)){
-  de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
-                  nDE=nDE, sampleSize=sampleSize, bd=bd,
-                  uninformativeCovariate = TRUE, 
-                  strongCovariate = FALSE,
-                  mc.cores=nCores)
-  saveRDS(de5, file = resfile_d5_uninfCov_w)
-}else{
-  de5 <- readRDS(file = resfile_d5_uninfCov_w)
-}
-```
-
-### Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(de5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(res5d, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res5d, met="FDR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res5d, met="TPR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(de5, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(de5, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE)
-
-}else{
-  message("Not enough methods found rejections at alpha ", ualpha, 
-          "; skipping upset plot")
-}
-
-```
-
-## Sim D10: 10v10 Weak Covariate
-
-Here we'll repeat the above, but using a weaker informative covariate (that has noise
-added to it).
-
-### Generate a list of SB results objects
-
-```{r d10-uc-uw, results='hide', message=FALSE}
-rseed <- 868
-sampleSize <- 10
-nDE <- 2000
-
-if(!file.exists(resfile_d10_uninfCov_w)){
-  de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
-                   nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   uninformativeCovariate = TRUE, 
-                   strongCovariate = FALSE,
-                   mc.cores=nCores)
-  saveRDS(de10, file = resfile_d10_uninfCov_w)
-}else{
-  de10 <- readRDS(file = resfile_d10_uninfCov_w)
-}
-```
-
-### Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(de10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(res10d, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res10d, met="FDR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res10d, met="TPR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(de10, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(de10, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res10d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(de10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
-}else{
-  message("Not enough methods found rejections at alpha ", ualpha, 
-          "; skipping upset plot")
-}
-
-```
-
 # Simulation setting comparison

### 2018-09-13 datasets/RNA-Seq/yeast-simulation-highpi0.Rmd
Refine informative covariate in yeast
Make strongly informative covariate stronger and weakly informative
covariate weaker.

@@
 # results files that will be generated
+# strong covariate
 resfile_d5 <- file.path(resdir, "yeastH-results-de5.rds")
 resfile_d10 <- file.path(resdir, "yeastH-results-de10.rds")
+
+# weak covariate
 resfile_d5_w <- file.path(resdir, "yeastHW-results-de5.rds")
@@
 
+# uninformative covariate
 resfile_d5_uninfCov <- file.path(resdir, "yeastH-results-de5-uninfCov.rds")
 resfile_d10_uninfCov <- file.path(resdir, "yeastH-results-de10-uninfCov.rds")
-resfile_d5_uninfCov_w <- file.path(resdir, "yeastHW-results-de5-uninfCov.rds")
-resfile_d10_uninfCov_w <- file.path(resdir, "yeastHW-results-de10-uninfCov.rds")
 
@@
     # pick random set of nDE genes to add signal to
-     if(strongCovariate){
-      wts <- covar
-    }else{
-      wts <- abs(covar + rnorm(length(covar)))
-    }
-    DE <- sample(1:nrow(dds_test), nDE, prob = wts)
+    DE <- sample(1:nrow(dds_test), nDE, prob = covar)
     truth[DE] <- TRUE
@@
 
-# Sim D5: DE Comparison 5v5
+# Non-null Comparisons with Strongly Informative Covariate
+
+## D5S: DE 5v5 Strong
 
@@
 
-# Sim D5: 5v5 Weak Covariate
+# Non-null Comparisons with Weakly Informative Covariate
+
+## D5W: DE 5v5 Weak
 
@@
 
-## Sim D5: 5v5 Weak Covariate
-
-Here we'll repeat the above, but using a weaker informative covariate (that has noise
-added to it).
-This will be done for 100 random splits.
-
-### Generate a list of SB results objects
-
-```{r d5-uc-uw, results='hide', message=FALSE}
-rseed <- 445
-sampleSize <- 5
-nDE <- 500
-
-if(!file.exists(resfile_d5_uninfCov_w)){
-  de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
-                  nDE=nDE, sampleSize=sampleSize, bd=bd,
-                  uninformativeCovariate = TRUE, 
-                  strongCovariate=FALSE,
-                  mc.cores=nCores)
-  saveRDS(de5, file = resfile_d5_uninfCov_w)
-}else{
-  de5 <- readRDS(file = resfile_d5_uninfCov_w)
-}
-```
-
-### Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(de5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(res5d, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-plotsim_average(res5d, met="FDR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res5d, met="TPR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(de5, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(de5, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if(numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE)
-
-}else{
-  message("Not enough methods found rejections at alpha ", ualpha, 
-          "; skipping upset plot")
-}
-
-```
-
-## Sim D10: 10v10 Weak Covariate
-
-Here we'll repeat the above, but using a weaker informative covariate (that has noise
-added to it).
-
-### Generate a list of SB results objects
-
-```{r d10-uc-uw, results='hide', message=FALSE}
-rseed <- 868
-sampleSize <- 10
-nDE <- 500
-
-if(!file.exists(resfile_d10_uninfCov_w)){
-  de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
-                   nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   uninformativeCovariate = TRUE, 
-                   strongCovariate = FALSE,
-                   mc.cores=nCores)
-  saveRDS(de10, file = resfile_d10_uninfCov_w)
-}else{
-  de10 <- readRDS(file = resfile_d10_uninfCov_w)
-}
-```
-
-### Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(de10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(res10d, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res10d, met="FDR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res10d, met="TPR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(de10, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(de10, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res10d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(de10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
-}else{
-  message("Not enough methods found rejections at alpha ", ualpha, 
-          "; skipping upset plot")
-}
-
-```
-
 # Simulation setting comparison

### 2018-09-13 datasets/RNA-Seq/yeast-simulation.Rmd
Refine informative covariate in yeast
Make strongly informative covariate stronger and weakly informative
covariate weaker.

@@
 # results files that will be generated
+# null
 resfile_n5 <- file.path(resdir, "yeast-results-null5.rds")
 resfile_n10 <- file.path(resdir, "yeast-results-null10.rds")
+
+# strong covariate
 resfile_d5 <- file.path(resdir, "yeast-results-de5.rds")
 resfile_d10 <- file.path(resdir, "yeast-results-de10.rds")
+
+# weak covariate
 resfile_d5_w <- file.path(resdir, "yeastW-results-de5.rds")
@@
 
-resfile_n5_uninfCov <- file.path(resdir, "yeast-results-null5-uninfCov.rds")
-resfile_n10_uninfCov <- file.path(resdir, "yeast-results-null10-uninfCov.rds")
+# uninformative covariate
 resfile_d5_uninfCov <- file.path(resdir, "yeast-results-de5-uninfCov.rds")
 resfile_d10_uninfCov <- file.path(resdir, "yeast-results-de10-uninfCov.rds")
-resfile_d5_uninfCov_w <- file.path(resdir, "yeastW-results-de5-uninfCov.rds")
-resfile_d10_uninfCov_w <- file.path(resdir, "yeastW-results-de10-uninfCov.rds")
 
@@
 
-# Sim N5: Null Comparison 5v5
+# Null Comparisons
+
+## N5: Null 5v5
 
@@
 
-# Sim N10: Null Comparison 10v10
+## N10: Null 10v10
 
@@
 where the groups are selected randomly. This will be done for 100 random splits.
-
-## Generate a list of SB results objects
+Note that here the covariate-aware methods use a covariate that by definition is
+uninformative (since there are no non-null observations).
+
+### Generate a list of SB results objects
 
@@
 where the groups are selected randomly and 2000 DE genes are added. 
+We'll use a strongly informative covariate.
 This will be done for 100 random splits.
 
-## Generate a list of SB results objects
+### Generate a list of SB results objects
 
@@
 where the groups are selected randomly and 2000 DE genes are added.
+We'll use a strongly informative covariate.
 This will be done for 100 random splits.
 
-## Generate a list of SB results objects
+### Generate a list of SB results objects
 
@@
 
-# Sim D5: 5v5 Weak Covariate
-
-Here we'll repeat the above, but using a weaker informative covariate (that has noise
-added to it).
-
-## Generate a list of SB results objects
+# Non-null Comparisons with Weakly Informative Covariate
+
+## D5W: DE 5v5 Weak
+
+Here we'll repeat the previous section, but using a weaker 
+informative covariate (that has noise added to it).
+
+### Generate a list of SB results objects
 
@@
 
-# Uninformative covariate
-
-Here we repeat the previous four sections, using an uninformative covariate.
-
-## Sim N5: Null Comparison 5v5
-
-Here we'll repeat the above, but for a null comparison 5 versus 5 Snf2 samples,
-where the groups are selected randomly. This will be done for 100 random splits.
-
-### Generate a list of SB results objects
-
-```{r n5-uc, results='hide', message=FALSE}
-rseed <- 225
-sampleSize <- 5
-nDE <- 0
-
-if (!file.exists(resfile_n5_uninfCov)){
-  null5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed,
-                   nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   uninformativeCovariate = TRUE, mc.cores = nCores)
-  saveRDS(null5, file = resfile_n5_uninfCov)
-}else{
-  null5 <- readRDS(file = resfile_n5_uninfCov)
-}
-```
-
-### Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(null5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res5 <- plotsim_standardize(null5, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(res5, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res5, met="TNR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(null5, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(null5, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res5, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(null5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
-}else{
-  message("Not enough methods found rejections at alpha ", ualpha, 
-          "; skipping upset plot")
-}
-
-```
-
-## Sim N10: Null Comparison 10v10
-
-Here we'll repeat the above, but for a null comparison 10 versus 10 Snf2 samples,
-where the groups are selected randomly. This will be done for 100 random splits.
-
-### Generate a list of SB results objects
-
-```{r n10-uc, results='hide', message=FALSE}
-rseed <- 837
-sampleSize <- 10
-nDE <- 0
-
-if (!file.exists(resfile_n10_uninfCov)){
-  null10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
-                   nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   uninformativeCovariate = TRUE, mc.cores=nCores)
-  saveRDS(null10, file = resfile_n10_uninfCov)
-}else{
-  null10 <- readRDS(file = resfile_n10_uninfCov)
-}
-```
-
-### Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(null10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res10 <- plotsim_standardize(null10, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(res10, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res10, met="TNR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(null10, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(null10, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res10, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(null10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
-}else{
-  message("Not enough methods found rejections at alpha ", ualpha, 
-          "; skipping upset plot")
-}
-
-```
-
-## Sim D5: DE Comparison 5v5
+# Non-null Comparisons with Uninformative Covariate
+
+Here we repeat the previous section, using an uninformative covariate.
+
+## D5U: DE 5v5 Uninformative
 
@@
 
-## Sim D5: 5v5 Weak Covariate
-
-Here we'll repeat the above, but using a weaker informative covariate (that has noise
-added to it).
-This will be done for 100 random splits.
-
-### Generate a list of SB results objects
-
-```{r d5-uc-uw, results='hide', message=FALSE}
-rseed <- 445
-sampleSize <- 5
-nDE <- 2000
-
-if (!file.exists(resfile_d5_uninfCov_w)){
-  de5 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
-                  nDE=nDE, sampleSize=sampleSize, bd=bd,
-                  uninformativeCovariate = TRUE, 
-                  strongCovariate = FALSE,
-                  mc.cores=nCores)
-  saveRDS(de5, file = resfile_d5_uninfCov_w)
-}else{
-  de5 <- readRDS(file = resfile_d5_uninfCov_w)
-}
-```
-
-### Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(de5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(res5d, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res5d, met="FDR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res5d, met="TPR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(de5, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(de5, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE)
-
-}else{
-  message("Not enough methods found rejections at alpha ", ualpha, 
-          "; skipping upset plot")
-}
-
-```
-
-## Sim D10: 10v10 Weak Covariate
-
-Here we'll repeat the above, but using a weaker informative covariate (that has noise
-added to it).
-
-### Generate a list of SB results objects
-
-```{r d10-uc-uw, results='hide', message=FALSE}
-rseed <- 868
-sampleSize <- 10
-nDE <- 2000
-
-if(!file.exists(resfile_d10_uninfCov_w)){
-  de10 <- mclapply(X=1:B, FUN=simulateOneSplit, rseed=rseed, 
-                   nDE=nDE, sampleSize=sampleSize, bd=bd,
-                   uninformativeCovariate = TRUE, 
-                   weakCovariate = FALSE,
-                   mc.cores=nCores)
-  saveRDS(de10, file = resfile_d10_uninfCov_w)
-}else{
-  de10 <- readRDS(file = resfile_d10_uninfCov_w)
-}
-```
-
-### Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(de10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(res10d, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res10d, met="FDR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res10d, met="TPR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(de10, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(de10, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res10d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(de10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
-}else{
-  message("Not enough methods found rejections at alpha ", ualpha, 
-          "; skipping upset plot")
-}
-
-```
-
 # Simulation setting comparison

### 2018-09-13 manuscript_figures/manuscript_figures_sims.Rmd
Add sim plot of unimodal effects with 25percent nonnull

@@
 
-methodset <- c("bonf", "bh", "ihw", "qvalue", "bl-df03", "lfdr",
+methodset <- c("bonf", "bh", "ihw", "qvalue", "bl", "lfdr",
                "fdrreg-e", "fdrreg-t", "ashq", "adapt-glm")
@@
 
+# Supplementary Figure (unimodal assumption, 25% non-null)
+
+This is a faceted figure with results from the pure simulation studies investigating the behavior of methods acorss different sampling distributions of the effect size for non-null tests, similar to Supplementary Figure S2 above.
+However, in contrast to the above simulation settings where only 10% of the hypotheses were non-null, these simulations were performed with a higher percentage (25%) of non-null tests.
+Again, the distributions of effect sizes used were taken from the Stephens (Biostatistics, 2016) paper, with all effect sizes doubled.
+
+As with the AdaPT simulations above, since these results are not included in the primary set of simulation results summarized in the `resmet` object, we recompute them from the output of the `supplementary/simulations-uasettings-nonnull25.Rmd` simulation case study.
+
+```{r SFigureS5-parse-data}
+ua25files <- list.files(file.path("..", "datasets", "simulations", "supplementary", "results"),
+                        "uasettings-25-benchmark", full.names = TRUE)
+
+ua25 <- lapply(ua25files, readRDS)
+
+ua25_i <- lapply(ua25, lapply, `[[`, "informative")
+ua25_i <- lapply(ua25_i, plotsim_standardize, alpha = seq(0.01, 0.10, 0.01))
+names(ua25_i) <- gsub("uasettings-25-benchmark-(.*?)\\.rds", "\\1", basename(ua25files))
+ua25_i <- bind_rows(ua25_i, .id = "signal")
+ua25_i <- dplyr::select(ua25_i, signal, rep, blabel, param.alpha, key, performanceMetric, alpha, value)
+
+ua25_u <- lapply(ua25, lapply, `[[`, "uninformative")
+ua25_u <- lapply(ua25_u, plotsim_standardize, alpha = seq(0.01, 0.10, 0.01))
+names(ua25_u) <- gsub("uasettings-25-benchmark-(.*?)\\.rds", "\\1", basename(ua25files))
+ua25_u <- bind_rows(ua25_u, .id = "signal")
+ua25_u <- dplyr::select(ua25_u, signal, rep, blabel, param.alpha, key, performanceMetric, alpha, value)
+
+ua25 <- dplyr::full_join(ua25_i, ua25_u,
+                              by = c("signal", "rep", "blabel", "param.alpha", "performanceMetric", "alpha"),
+                              suffix = c(".info", ".uninfo"))
+
+## clean up IHW methods
+ua25 <- dplyr::filter(ua25, !grepl("ihw-", blabel) | param.alpha == as.character(alpha))
+ua25 <- dplyr::mutate(ua25, blabel = ifelse(grepl("ihw-", blabel), "ihw", blabel))
+ua25 <- dplyr::mutate(ua25, blabel = ifelse(blabel == "bl-df03", "bl", blabel))
+ua25 <- dplyr::rename(ua25, Method = blabel, key = performanceMetric)
+ua25 <- dplyr::select(ua25, -param.alpha)
+ua25 <- dplyr::mutate(ua25, Method = factor(Method, levels = methodset))
+ua25 <- dplyr::filter(ua25, !is.na(Method))
+
+ua25met <- dplyr::mutate(ua25,
+                         value.info = ifelse(is.na(value.info) & grepl("R$", key), 0, value.info),
+                         value.uninfo = ifelse(is.na(value.uninfo) & grepl("R$", key), 0, value.uninfo))
+
+ua25met <- dplyr::group_by(ua25met, Method, signal, key, alpha)
+ua25met <- dplyr::summarize(ua25met,
+                            mean.info = mean(value.info, na.rm = TRUE),
+                            se.info = sd(value.info, na.rm = TRUE) / sqrt(sum(!is.na(value.info))),
+                            nNA.info = sum(is.na(value.info)),
+                            mean.diff = mean(value.info - value.uninfo, na.rm = TRUE),
+                            se.diff = sd(value.info - value.uninfo, na.rm = TRUE) /
+                                sqrt(sum(!is.na(value.info - value.uninfo))))
+ua25met <- dplyr::ungroup(ua25met)
+```
+
+After computing the performance metrics, we now plot the results for all methods.
+Again, we plot the mean FDR and TPR at nominal FDR cutoffs of 0.01 to 0.10 across 100 replications, as well as the FDR and TPR differences with and without the informative covariate.
+
+```{r SFigureSim5, fig.width = 8, fig.height = 6}
+## SFigure Sim5 B,C: FDR/TPR of methods across unimodal effect size distribution
+ps5b <- genplot(ua25met, cov = alpha, met = "FDR") +
+    facet_grid(. ~ signal, labeller = label_parsed) +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0)
+ps5c <- genplot(ua25met, cov = alpha, met = "TPR") +
+    facet_grid(. ~ signal, labeller = label_parsed) +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0)
+
+## SFigure Sim5 D,E: FDR/TPR difference of methods
+ps5d <- genplot(ua25met, cov = alpha, type = "diff", met = "FDR") +
+    facet_grid(. ~ signal, labeller = label_parsed) +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0) +
+    ylab(expression(paste("Mean ", Delta, "FDR (informative-uninformative)")))
+ps5e <- genplot(ua25met, cov = alpha, type = "diff", met = "TPR") +
+    facet_grid(. ~ signal, labeller = label_parsed) +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0) +
+    ylab(expression(paste("Mean ", Delta, "TPR (informative-uninformative)")))
+
+## create title for SFigure Sim5
+ps5Title <- ggdraw() +
+    draw_label("Simulation performance across unimodal effect size distributions w/ 25% non-null",
+               fontface = 'bold', size = 15)
+
+## pull SFigure Sim5 together
+SFigS5 <- plot_grid(ps2a + guides(color = FALSE, linetype = FALSE),
+                    ps5b + guides(color = FALSE, linetype = FALSE),
+                    ps5c + guides(color = FALSE, linetype = FALSE),
+                    ps5d + guides(color = FALSE, linetype = FALSE),
+                    ps5e + guides(color = FALSE, linetype = FALSE),
+                    labels = LETTERS[1:5], ncol = 1, rel_heights = c(.6, rep(1, 4)))
+SFigS5 <- plot_grid(ps5Title, SFigS5, ncol = 1, rel_heights = c(0.03, 1))
+SFigS5 <- plot_grid(SFigS5, get_legend(ps5b), rel_widths = c(6.5, 1))
+SFigS5
+ggsave(file.path(outdir, "SFigureSim5.pdf"), width=12, height=16)
+```
+
 # Session information

### 2018-09-14 datasets/RNA-Seq/yeast-simulation-bimodalalternative-highpi0.Rmd
Edit section headers
Standardize random seeds ffor different covariate settings

@@
 
-# Non-null Comparisons with Strongly Informative Covariate
-
 ## D10S: DE 10v10 Strong
@@
 ```
+
 # Non-null Comparisons with Weakly Informative Covariate
@@
 ```{r d5-uw, results='hide', message=FALSE}
-rseed <- 445
+rseed <- 198
 sampleSize <- 5
@@
 ```{r d10-uw, results='hide', message=FALSE}
-rseed <- 868
+rseed <- 961
 sampleSize <- 10

### 2018-09-14 datasets/RNA-Seq/yeast-simulation-bimodalalternative.Rmd
Edit section headers
Standardize random seeds ffor different covariate settings

@@
 
-# Non-null Comparisons with Strongly Informative Covariate
+# Non-null Comparisons with Weakly Informative Covariate
 

### 2018-09-14 manuscript_figures/manuscript_figures.Rmd
Add panels for different yeast settings in Fig S3
Also remove fc version of Figs S4 and S9

@@
 # Supplementary Figures
-
-## Uninformative covariate impact on simulations
-
-Here we'll contrast the results of informative and uninformative covariates 
-on the yeast and polyester simulation (DE sample size 5). We'll plot
-differences in TPR and FDR due to informative covariate by alpha level.
-
-```{r, fig.width = 10, fig.height = 4}
-for (org in c("yeast-", "polyester")){
-
-  objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
-  resfile <- objects[grepl(org, objects) ] 
-  resfile <- resfile[grepl("5", resfile) & !grepl("null", resfile)]
-
-  # regexp to construct master data table separating case study datasets, methods
-  # and covariates
-  obj.names <- str_split(resfile, pattern = "/")
-  casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
-  filename <- gsub(".rds", "", unlist(lapply(obj.names, function(x) x[4] )))
-  dataset <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[3]]))
-  inform <-  ifelse(grepl("uninf", resfile), "uninf", "inf")
-  type <-  ifelse(grepl("null", resfile), "null", "de")
-  size <-  ifelse(grepl("10", resfile), "10", "5")
-  aseq <- seq(0.01, 0.10, by = 0.01)
-
-  tab <- data.frame()
-  for (l in seq_along(resfile)){
-    x <- readRDS(resfile[l])
-
-    tmp2 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
-                    fill =  "FDR", annotate = NULL, 
-                    alpha = aseq) %>%
-      mutate(fdr = nrejects) %>%
-      select(method, casestudy, replicate, fdr, alpha)
-    tmp3 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
-                    fill =  "TNR", annotate = NULL, 
-                    alpha = aseq) %>%
-      mutate(tnr = nrejects) %>%
-      select(method, casestudy, replicate, tnr, alpha)
-
-    if (type[l] != "null"){
-      tmp1 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
-                      fill =  "TPR", annotate = NULL, 
-                      alpha = aseq) %>%
-        mutate(tpr = nrejects) %>%
-        select(method, casestudy, replicate, tpr, alpha)
-
-      tmp <- left_join(tmp1, tmp2, by = c("method", "casestudy", "replicate", "alpha")) %>%
-        left_join(tmp3, by = c("method", "casestudy", "replicate", "alpha")) %>%
-        mutate(inform = inform[l],
-               type = type[l],
-               size = size[l])
-    }else{
-      tmp <- left_join(tmp2, tmp3, by = c("method", "casestudy", "replicate", "alpha")) %>%
-        mutate(inform = inform[l],
-               type = type[l],
-               size = size[l])
-    }
-
-    tab <- bind_rows(tab, tmp)
-  }
-
-myspread <- function(df, key, value) {
-    # quote key
-    keyq <- rlang::enquo(key)
-    # break value vector into quotes
-    valueq <- rlang::enquo(value)
-    s <- rlang::quos(!!valueq)
-    df %>% gather(variable, value, !!!s) %>%
-        unite(temp, !!keyq, variable) %>%
-        spread(temp, value)
-}
-
-dftab <- tab %>%
-  select(-casestudy) %>%
-  myspread(inform, c(tpr, fdr, tnr))
-
-dftab <- dftab %>%
-  mutate(tpr_diff = inf_tpr - uninf_tpr,
-         fdr_diff = inf_fdr - uninf_fdr,
-         tnr_diff = inf_tnr - uninf_tnr) %>%
-  group_by(type, size, method, alpha) %>%
-  summarize(tpr_se = sd(tpr_diff)/n(),
-            fdr_se = sd(fdr_diff)/n(),
-            tnr_se = sd(tnr_diff)/n(),
-            tpr_diff = mean(tpr_diff),
-            fdr_diff = mean(fdr_diff),
-            tnr_diff = mean(tnr_diff))
-
-col <- as.character(candycols$col)
-names(col) <- as.character(candycols$Method)
-lty <- as.character(candycols$lty)
-names(lty) <- as.character(candycols$Method)
-
-lineplot_fdr <- ggplot(dftab,
-                       aes(x = alpha, y = fdr_diff, color = method, linetype = method)) +
-    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
-    geom_line(alpha = 0.85) +
-    geom_errorbar(width=.003, alpha = 0.5, aes(ymin = fdr_diff - fdr_se, 
-                                ymax = fdr_diff + fdr_se)) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
-    scale_y_continuous(expression(paste("Mean ", Delta, "FDR (informative-uninformative)")), 
-                       breaks = seq(-1, 1, .005), labels = scales::percent) + 
-    ggtitle("FDR difference with informative covariate") +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 11, hjust = 0.8))
-
-lineplot_tpr <- ggplot(dftab, aes(x = alpha, y = tpr_diff, color = method, linetype = method)) +
-    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
-    geom_errorbar(width=.003, alpha = 0.5, aes(ymin = tpr_diff - tpr_se, 
-                                ymax = tpr_diff + tpr_se)) +
-    geom_line(alpha = 0.85) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
-    scale_y_continuous(expression(paste("Mean ", Delta, "TPR (informative-uninformative)")), 
-                       labels = scales::percent) + 
-    ggtitle("TPR difference with informative covariate") +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 11))
-
-  gp <- plot_grid(lineplot_fdr + guides(color = FALSE, linetype = FALSE),
-                  lineplot_tpr + guides(color = FALSE, linetype = FALSE),
-                  labels = LETTERS[1:2], nrow = 1)
-  gp <- plot_grid(gp, get_legend(lineplot_fdr), rel_widths = c(1, 1/8))
-  gp
-  ggsave(file.path(outdir, paste0("Figure_infCov_", gsub("-", "", org), ".pdf")), 
-                   height = 3.5, width = 8)
-}
-```
 
@@
 # custom symmetric log scale for percent
+transform <- function(x) { 
+  x <- ifelse(x > 0, log(x + 1), -log(-x + 1))
+  x
+}
+
 log_neg_percent <- function()
@@
         facet_grid( ~ casestudy + cov, scales = "free", space = "free") +
-        ggtitle("Impact of informative covariate on number of rejections") +
         theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
@@
 ggsave(file.path(outdir, "Figure_impactofinfcov.pdf"), width = 15.5, height = 4.25)
-
-# log-scale color log2 ratio fill
-topRange <- max(dftab$diff_inf_ratio, na.rm = TRUE)
-botRange <- min(dftab$diff_inf_ratio, na.rm = TRUE)
-
-Fig_impactofinfcov_fc <- ggplot(dftab, aes(x = dataset, y = method, fill = diff_inf_ratio)) + 
-        geom_tile() + 
-        geom_text(aes(label = ifelse(is.na(topR), "", #adaptive text lab color
-                                     ifelse(topR > 1, signif(topR,2), ""))), 
-                  color = "white", size = 2) +
-        geom_text(aes(label = ifelse(is.na(topR), "", 
-                                     ifelse(topR <= 1, signif(topR,2), ""))), 
-                  color = "black", size = 2)+
-        scale_fill_gradientn("Fold Change\n(log2)",
-                             colors = c("darkblue","white","darkred"), 
-                             limits = c(-0.5, topRange), 
-                             values = scales::rescale(c(-0.5, 0, topRange))) +
-        theme_bw() +
-        xlab("Case Study") +
-        ylab("Method") +
-        facet_grid( ~ casestudy + cov, scales = "free", space = "free") +
-        ggtitle("Impact of informative covariate on number of rejections") +
-        theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
-              legend.position = c(0.99, 0.95), legend.justification = c(1, 1),
-              legend.background = element_rect(fill = scales::alpha("white", 3/4), color = 'black'),
-              legend.direction = "horizontal",
-              legend.text = element_text(size = 7, angle=45, margin = margin(t=0)),
-              legend.title = element_text(size = 9, margin=margin(t=-10)),
-              legend.margin = margin(t=2, r=10, b=-2, l=10))
-
-Fig_impactofinfcov_fc
-ggsave(file.path(outdir, "Figure_impactofinfcov_fc.pdf"), width = 15.5, height = 4.25)
 
@@
         ylab("Method") +
-        ggtitle("Impact of informative covariate") +
         theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
@@
 ggsave(file.path(outdir, "Figure_impactofinfcov_avg.pdf"), width = 5, height = 4)
-
-# log-scale color log2 ratio fill
-topRange <- max(dftab$diff_inf_ratio, na.rm = TRUE)
-botRange <- min(dftab$diff_inf_ratio, na.rm = TRUE)
-
-Fig_impactofinfcov_fc <- ggplot(dftab, aes(x = casestudy, y = method, fill = diff_inf_ratio)) + 
-        geom_tile() + 
-        scale_fill_gradientn("Fold Change\n(log2)",
-                             colors = c("darkblue","white","darkred"), 
-                             limits = c(-0.5, topRange), 
-                             values = scales::rescale(c(-0.5, 0, topRange))) +
-        theme_bw() +
-        xlab("Case Study") +
-        ylab("Method") +
-        ggtitle("Impact of informative covariate") +
-        theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
-        geom_text(aes(label = ifelse(is.na(topR), "", #adaptive text lab color
-                                     ifelse(topR > 1, signif(topR, 2), ""))), 
-                  color = "white", size = 2) +
-        geom_text(aes(label = ifelse(is.na(topR), "", 
-                                     ifelse(topR <= 1, signif(topR, 2), ""))), 
-                  color = "black", size = 2)  +
-        theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
-              legend.position = c(0.945, 0.94), legend.justification = c(1, 1),
-              legend.background = element_rect(fill = scales::alpha("white", 3/4), color = 'black'),
-              legend.direction = "horizontal",
-              legend.text = element_text(size = 7, angle=45, margin = margin(t=2)),
-              legend.title = element_text(size = 9, margin=margin(t=-10)),
-              legend.margin = margin(t=2, r=10, b=-6, l=10))
-
-Fig_impactofinfcov_fc
-ggsave(file.path(outdir, "Figure_impactofinfcov_avg_fc.pdf"), width = 5, height = 4)
 
@@
 
-ggdraw() +
-  draw_plot(pfdr, 0.53, 0.025, 0.24, 0.93) +
-  draw_plot(ptpr, 0.77, 0.025, 0.24, 0.93) +
-  draw_plot(Fig_impactofinfcov_fc, 0, 0, 0.52, 1) +
-  draw_plot_label(c("A", "B"), c(0, 0.53), c(1, 1), size = 15) +
-  draw_plot_label("Simulation", 0.71, 0.06, size =11, fontface = "plain")
-ggsave(file.path(outdir, "Figure_impactofinfcov_avg_sim_fc.pdf"), width=9.5, height=5)
-
 ```
@@
   facet_grid( ~ es_dist + pi0) +
-  ylim(0,0.3) 
+  ylim(0,0.17) 
 FigStrong
@@
                 ggtitle("") +
-  ylim(0,0.88) +
+  ylim(0,0.72) +
   facet_grid( ~ es_dist + pi0) 
@@
                 ggtitle("") +
-  ylim(0,0.3) +
+  ylim(0,0.17) +
   facet_grid( ~ es_dist + pi0)
@@
                 ggtitle("") +
-  ylim(0,0.88) +
+  ylim(0,0.72) +
   facet_grid( ~ es_dist + pi0)
@@
 
+## Uninformative covariate impact on simulations
+
+Here we'll contrast the results of informative and uninformative covariates 
+on the yeast and polyester simulation (DE sample size 5). We'll plot
+differences in TPR and FDR due to informative covariate by alpha level.
+
+```{r, fig.width = 10, fig.height = 4}
+org <- "yeast"
+
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+resfile <- objects[grepl(org, objects) ] 
+resfile <- resfile[grepl("5", resfile) & !grepl("null", resfile)]
+
+# regexp to construct master data table separating case study datasets, methods
+# and covariates
+obj.names <- str_split(resfile, pattern = "/")
+casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
+filename <- gsub(".rds", "", unlist(lapply(obj.names, function(x) x[4] )))
+dataset <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[3]]))
+inform <-  ifelse(grepl("uninf", resfile), "uninf", "inf")
+type <-  ifelse(grepl("null", resfile), "null", "de")
+size <-  ifelse(grepl("10", resfile), "10", "5")
+mode <-  ifelse(grepl("II", resfile), "Bimodal effect Size", "Unimodal Effect Size")
+pi0 <-  ifelse(grepl("H", resfile), "7.5% Non-null", "30% Non-null")
+strength <- ifelse(grepl("W", resfile), "Weakly informative", "Strongly informative")
+strength <- ifelse(grepl("uninf", resfile), "Uninformative", strength)
+aseq <- seq(0.01, 0.10, by = 0.01)
+
+tab <- data.frame()
+for (l in seq_along(resfile)){
+  x <- readRDS(resfile[l])
+
+  tmp2 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
+                  fill =  "FDR", annotate = NULL, 
+                  alpha = aseq) %>%
+    mutate(fdr = nrejects) %>%
+    select(method, casestudy, replicate, fdr, alpha)
+
+  if (type[l] != "null"){
+    tmp1 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
+                    fill =  "TPR", annotate = NULL, 
+                    alpha = aseq) %>%
+      mutate(tpr = nrejects) %>%
+      select(method, casestudy, replicate, tpr, alpha)
+
+    tmp <- left_join(tmp1, tmp2, by = c("method", "casestudy", "replicate", "alpha")) %>%
+      mutate(inform = inform[l],
+             type = type[l],
+             size = size[l],
+             mode = mode[l],
+             pi0 = pi0[l],
+             strength = strength[l])
+  }else{
+    tmp <- tmp2  %>%
+      mutate(inform = inform[l],
+             type = type[l],
+             size = size[l],
+             mode = mode[l],
+             pi0 = pi0[l],
+             strength = strength[l])
+  }
+
+  tab <- bind_rows(tab, tmp)
+}
+
+tab <- tab %>% select(-c(type, size))
+dftab <- tab %>% filter(inform == "inf") %>%
+  select(-inform) %>%
+  rename(fdr = "inf_fdr",
+         tpr = "inf_tpr")
+uninftab <- tab %>% filter(inform == "uninf") %>%
+  select(-c(strength, inform, casestudy)) %>%
+  rename(fdr = "uninf_fdr",
+         tpr = "uninf_tpr")
+
+dftab <- left_join(dftab, uninftab, 
+                   by = c("method", "replicate", 
+                          "mode", "pi0", "alpha"))
+
+dftab <- dftab %>%
+  mutate(tpr_diff = inf_tpr - uninf_tpr,
+         fdr_diff = inf_fdr - uninf_fdr) %>%
+  group_by(method, alpha, mode, pi0, strength) %>%
+  summarize(tpr_se = sd(tpr_diff)/n(),
+            fdr_se = sd(fdr_diff)/n(),
+            tpr_diff = mean(tpr_diff),
+            fdr_diff = mean(fdr_diff))
+
+col <- as.character(candycols$col)
+names(col) <- as.character(candycols$Method)
+lty <- as.character(candycols$lty)
+names(lty) <- as.character(candycols$Method)
+
+FigWeak <- dftab %>% filter(strength == "Weakly informative") %>%
+  ggplot(aes(x = alpha, y = fdr_diff, color = method, linetype = method)) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    geom_line(alpha = 0.85) +
+    geom_errorbar(width=.003, alpha = 0.5, aes(ymin = fdr_diff - fdr_se, 
+                                ymax = fdr_diff + fdr_se)) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
+    scale_y_continuous(expression(paste(Delta, "FDR (informative-uninformative)")), 
+                       breaks = seq(-0.03, 1, .003), labels = scales::percent,
+                       limits = c(-0.008, 0.004)) + 
+    ggtitle("") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 11, hjust = 0.8)) +
+  facet_grid( ~ mode + pi0)
+FigWeak
+
+FigWeakT <- dftab %>% filter(strength == "Weakly informative") %>%
+  ggplot(aes(x = alpha, y = tpr_diff, color = method, linetype = method)) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    geom_errorbar(width=.003, alpha = 0.5, aes(ymin = tpr_diff - tpr_se, 
+                                ymax = tpr_diff + tpr_se)) +
+    geom_line(alpha = 0.85) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
+    scale_y_continuous(expression(paste(Delta, "TPR (informative-uninformative)")), 
+                       labels = scales::percent, limits=c(-0.02, 0.07),
+                       breaks = seq(-0.02, 0.07, by = 0.02)) + 
+    ggtitle("") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 11)) +
+  facet_grid( ~ mode + pi0)
+FigWeakT
+
+FigStrong <- dftab %>% filter(strength == "Strongly informative") %>%
+  ggplot(aes(x = alpha, y = fdr_diff, color = method, linetype = method)) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    geom_line(alpha = 0.85) +
+    geom_errorbar(width=.003, alpha = 0.5, aes(ymin = fdr_diff - fdr_se, 
+                                ymax = fdr_diff + fdr_se)) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
+    scale_y_continuous(expression(paste(Delta, "FDR (informative-uninformative)")), 
+                       breaks = seq(-0.03, 1, .003), labels = scales::percent,
+                       limits = c(-0.008, 0.004)) + 
+    ggtitle("") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 11, hjust = 0.8)) +
+  facet_grid( ~ mode + pi0)
+FigStrong
+
+FigStrongT <- dftab %>% filter(strength == "Strongly informative") %>%
+  ggplot(aes(x = alpha, y = tpr_diff, color = method, linetype = method)) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    geom_errorbar(width=.003, alpha = 0.5, aes(ymin = tpr_diff - tpr_se, 
+                                ymax = tpr_diff + tpr_se)) +
+    geom_line(alpha = 0.85) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
+    scale_y_continuous(expression(paste(Delta, "TPR (informative-uninformative)")), 
+                       labels = scales::percent, limits=c(-0.02, 0.07),
+                       breaks = seq(-0.02, 0.07, by = 0.02)) + 
+    ggtitle("") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 11)) +
+  facet_grid( ~ mode + pi0)
+FigStrongT
+
+leg <- get_legend(FigStrong)
+
+FigA <- plot_grid(FigStrong + theme(legend.position="none") +  
+                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)), 
+                  FigStrongT + theme(legend.position="none") +  
+                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)),
+                 nrow = 2, axis = "rlbt")
+
+FigB <- plot_grid(FigWeak + theme(legend.position="none") +  
+                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)), 
+                  FigWeakT + theme(legend.position="none") +  
+                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)),
+                 nrow = 2, axis = "rlbt")
+
+Fig <- plot_grid(FigA, FigB,
+                 nrow = 2, 
+                 labels = paste0(LETTERS[1:2], "       ",
+                          c("Strongly informative covariate",
+                            "Weakly informative covariate")),
+                 hjust = 0, axis = "rlbt")
+plot_grid(Fig, leg, ncol=2, rel_widths = c(1,0.1))
+ggsave(file.path(outdir, "Figure_infCov_yeast.pdf"), width=10, height=12)
+
+```
+
 # Session information

### 2018-09-17 manuscript_figures/manuscript_figures.Rmd
Adjust positions of Fig S4 after removing title

@@
 ggdraw() +
-  draw_plot(pfdr, 0.53, 0.025, 0.24, 0.93) +
-  draw_plot(ptpr, 0.77, 0.025, 0.24, 0.93) +
+  draw_plot(pfdr, 0.53, 0.03, 0.24, 0.97) +
+  draw_plot(ptpr, 0.77, 0.03, 0.24, 0.97) +
   draw_plot(Fig_impactofinfcov, 0, 0, 0.52, 1) +

### 2018-09-17 manuscript_figures/manuscript_figures.Rmd
Edit subtitles Fig1 & Alignment of Fig S1 and S3

@@
       theme(legend.position="none") +  
-      ggtitle("Mean FDR (100 replications)") +
+      ggtitle("FDR by alpha level") +
       scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02))
@@
       theme(legend.position="none")  +  
-      ggtitle("Mean TPR (100 replications)") +
+      ggtitle("TPR by alpha level") +
       scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02))
@@
       ylab("Number of Rejections") +
-      ggtitle("Mean Rejections (100 replications)") +
+      ggtitle("Rejections by alpha level") +
       scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02))
@@
                        return_list = FALSE, nintersects = 12)
-grid.text("Mean overlap Over 100 Replications", x = 0.65, y = 0.97, 
+grid.text("Mean overlap of rejection sets", x = 0.65, y = 0.97, 
           gp = gpar(fontsize = 14, fontface = "bold"))
@@
                    scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)),
-                 nrow = 2)
+                 nrow = 2, align = "hv")
 
@@
                    scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)),
-                 nrow = 2, axis = "rlbt")
+                 nrow = 2, align = "hv")
 

### 2018-09-17 manuscript_figures/manuscript_figures_sims.Rmd
Remove plot titles from supplementary simulation plots

@@
 
-```{r SFigureSim1, fig.width = 12, fig.height = 16}
+```{r SFigureSim1, fig.width = 12, fig.height = 15}
 res_noise <- dplyr::filter(resmet, setting == "informative-cubic")
@@
 
-## create title for SFigure Sim1
-ps1Title <- ggdraw() +
-    draw_label("Simulation performance across test statistic distributions",
-               fontface = 'bold', size = 20)
+## ## create title for SFigure Sim1
+## ps1Title <- ggdraw() +
+##     draw_label("Simulation performance across test statistic distributions",
+##                fontface = 'bold', size = 20)
 
@@
                     labels = LETTERS[1:5], ncol = 1, rel_heights = c(.6, rep(1, 4)))
-SFigS1 <- plot_grid(ps1Title, SFigS1, ncol = 1, rel_heights = c(0.03, 1))
+## SFigS1 <- plot_grid(ps1Title, SFigS1, ncol = 1, rel_heights = c(0.03, 1))
 SFigS1 <- plot_grid(SFigS1, get_legend(ps1b), rel_widths = c(6.5, 1))
 SFigS1
-ggsave(file.path(outdir, "SFigureSim1.pdf"), width=12, height=16)
+ggsave(file.path(outdir, "SFigureSim1.pdf"), width = 12, height = 15)
 ```
@@
 
-```{r SFigureSim2, fig.width = 12, fig.height = 16}
+```{r SFigureSim2, fig.width = 12, fig.height = 15}
 res_ua <- dplyr::filter(resmet, setting == "uasettings", grepl("N\\(0, 1\\)", dist))
@@
 
-## create title for SFigure Sim2
-ps2Title <- ggdraw() +
-    draw_label("Simulation performance across unimodal effect size distributions",
-               fontface = 'bold', size = 20)
+## ## create title for SFigure Sim2
+## ps2Title <- ggdraw() +
+##     draw_label("Simulation performance across unimodal effect size distributions",
+##                fontface = 'bold', size = 20)
 
@@
                     labels = LETTERS[1:5], ncol = 1, rel_heights = c(.6, rep(1, 4)))
-SFigS2 <- plot_grid(ps2Title, SFigS2, ncol = 1, rel_heights = c(0.03, 1))
+## SFigS2 <- plot_grid(ps2Title, SFigS2, ncol = 1, rel_heights = c(0.03, 1))
 SFigS2 <- plot_grid(SFigS2, get_legend(ps2b), rel_widths = c(6.5, 1))
 SFigS2
-ggsave(file.path(outdir, "SFigureSim2.pdf"), width=12, height=16)
+ggsave(file.path(outdir, "SFigureSim2.pdf"), width = 12, height = 15)
 ```
@@
 
-```{r SFigureSim3, fig.width = 12, fig.height = 16}
+```{r SFigureSim3, fig.width = 12, fig.height = 15}
 res_icov <- dplyr::filter(resmet, grepl("^informative", setting), grepl("N\\(0, 1\\)", dist))
@@
 
-## create title for SFigure Sim3
-ps3Title <- ggdraw() +
-    draw_label("Simulation performance across informative covariate relationship",
-               fontface = 'bold', size = 20)
+## ## create title for SFigure Sim3
+## ps3Title <- ggdraw() +
+##     draw_label("Simulation performance across informative covariate relationship",
+##                fontface = 'bold', size = 20)
 
@@
                     labels = LETTERS[1:5], ncol = 1, rel_heights = c(.6, rep(1, 4)))
-SFigS3 <- plot_grid(ps3Title, SFigS3, ncol = 1, rel_heights = c(0.03, 1))
+## SFigS3 <- plot_grid(ps3Title, SFigS3, ncol = 1, rel_heights = c(0.03, 1))
 SFigS3 <- plot_grid(SFigS3, get_legend(ps3b), rel_widths = c(6.5, 1))
 SFigS3
-ggsave(file.path(outdir, "SFigureSim3.pdf"), width=12, height=16)
+ggsave(file.path(outdir, "SFigureSim3.pdf"), width = 12, height = 15)
 ```
@@
 
-## create title for SFigure Sim4
-ps4Title <- ggdraw() +
-    draw_label("Simulation performance with modified AdaPT parameters",
-               fontface = 'bold', size = 10)
+## ## create title for SFigure Sim4
+## ps4Title <- ggdraw() +
+##     draw_label("Simulation performance with modified AdaPT parameters",
+##                fontface = 'bold', size = 10)
 
@@
                     labels = LETTERS[1:4], ncol = 2)
-SFigS4 <- plot_grid(ps4Title, SFigS4, ncol = 1, rel_heights = c(0.06, 1))
+## SFigS4 <- plot_grid(ps4Title, SFigS4, ncol = 1, rel_heights = c(0.06, 1))
 SFigS4 <- plot_grid(SFigS4, get_legend(ps4b), rel_widths = c(4.5, 1))
 SFigS4
-ggsave(file.path(outdir, "SFigureSim4.pdf"), width=8, height=6)
+ggsave(file.path(outdir, "SFigureSim4.pdf"), width = 8, height = 6)
 ```
@@
 
-## create title for SFigure Sim5
-ps5Title <- ggdraw() +
-    draw_label("Simulation performance across unimodal effect size distributions w/ 25% non-null",
-               fontface = 'bold', size = 15)
+## ## create title for SFigure Sim5
+## ps5Title <- ggdraw() +
+##     draw_label("Simulation performance across unimodal effect size distributions w/ 25% non-null",
+##                fontface = 'bold', size = 15)
 
@@
                     labels = LETTERS[1:5], ncol = 1, rel_heights = c(.6, rep(1, 4)))
-SFigS5 <- plot_grid(ps5Title, SFigS5, ncol = 1, rel_heights = c(0.03, 1))
+## SFigS5 <- plot_grid(ps5Title, SFigS5, ncol = 1, rel_heights = c(0.03, 1))
 SFigS5 <- plot_grid(SFigS5, get_legend(ps5b), rel_widths = c(6.5, 1))
 SFigS5
-ggsave(file.path(outdir, "SFigureSim5.pdf"), width=12, height=16)
+ggsave(file.path(outdir, "SFigureSim5.pdf"), width = 12, height = 15)
 ```

### 2018-09-18 datasets/RNA-Seq/yeast-simulation-bimodalalternative-highpi0.Rmd
Add plot of logistic informative covariate

@@
 
-First, here's a function that will generate sampling weights for test statistics
+First we'll illustrate the logistic curve used to simulate the informative covariate. 
+In this study we'll sample from a logistic curve and use these values as probability 
+weights when  selecting DE genes. A weakly informative covariate will be generated by 
+adding noise to the strongly informative covariate.
+
+```{r}
+x <- seq(0,10, by = 0.1)
+infcov <- 1 / (1+exp(-x + 5))
+plot(x, infcov, type = "l", ylab = "informative covariate")
+
+plot(infcov,  pmin(1, abs(infcov + rnorm(length(x), 0, 0.25))), 
+     ylab = "weak covariate", xlab = "true covariate")
+```
+
+Next, here's a function that will generate sampling weights for test statistics
 of DE genes that downweights observations near zero, as well as extremely large values.

### 2018-09-18 datasets/RNA-Seq/yeast-simulation-highpi0.Rmd
Add plot of logistic informative covariate

@@
 # Simulation set up
+
+First we'll illustrate the logistic curve used to simulate the informative covariate. 
+In this study we'll sample from a logistic curve and use these values as probability 
+weights when  selecting DE genes. A weakly informative covariate will be generated by 
+adding noise to the strongly informative covariate.
+
+```{r}
+x <- seq(0,10, by = 0.1)
+infcov <- 1 / (1+exp(-x + 5))
+plot(x, infcov, type = "l", ylab = "informative covariate")
+
+plot(infcov,  pmin(1, abs(infcov + rnorm(length(x), 0, 0.25))), 
+     ylab = "weak covariate", xlab = "true covariate")
+```
 

### 2018-09-18 datasets/RNA-Seq/polyester-simulation.Rmd
Standardize alpha scale in polyester
Remove mean covariate; add weak covariate to Polyester

@@
 # results files that will be generated below
+# null
 resfile_n5 <- file.path(resdir, "polyester-results-null5.rds")
 resfile_n10 <- file.path(resdir, "polyester-results-null10.rds")
+
+# strong covariate
 resfile_d5 <- file.path(resdir, "polyester-results-de5.rds")
@@
 
-resfile_n5_uninfCov <- file.path(resdir, "polyester-results-null5-uninfCov.rds")
-resfile_n10_uninfCov <- file.path(resdir, "polyester-results-null10-uninfCov.rds")
+# weak covariate 
+resfile_d5_w <- file.path(resdir, "polyesterW-results-de5.rds")
+resfile_d10_w <- file.path(resdir, "polyesterW-results-de10.rds")
+
+# uninformative covariate
 resfile_d5_uninfCov <- file.path(resdir, "polyester-results-de5-uninfCov.rds")
@@
 
-We use the [polyester](https://bioconductor.org/packages/release/bioc/html/polyester.html) Bioconductor package to simulate RNA-Seq reads.
-The package requires a count matrix as input which estimates the mean variance relationship.
-Here we start with the yeast data with 48 biological replicates in each of
+We use the [polyester](https://bioconductor.org/packages/release/bioc/html/polyester.html) 
+Bioconductor package to simulate RNA-Seq reads. The package requires a count matrix
+as input which estimates the mean variance relationship. Here we start with the 
+yeast data with 48 biological replicates in each of 
 two conditions (analyzed in this [publication](https://www.ncbi.nlm.nih.gov/pubmed/26206307/)). 
@@
 #' @param bd is the bench design object
-#' @param uninformativeCovariate logical indicating whether to replace overall mean 
-#'  independent covariate with a randomly generated (uninformative) covariate. 
-#'  Default is FALSE.
+#' @param uninformativeCovariate logical indicating whether to use an uninformative
+#'  covariate (default is FALSE)
+#' @param strongCovariate logical indicating whether to use a strongly informative
+#'  covariate (default is TRUE)
 #' @param rseed random seed
@@
                              nDE, sampleSize, bd, uninformativeCovariate = FALSE,
-                             ngenes = nrow(counts)){
-  uninf_covar <- rnorm(ngenes)
+                             ngenes = nrow(counts),
+                             strongCovariate = TRUE){
 
@@
   set.seed(as.numeric(X)*as.numeric(rseed))
+
+  covar <- 1 / (1+exp(-runif(nrow(counts), 0, 10) + 5))
 
@@
   p0 <- p0[pzero < 0.5]
-  uninf_covar <- uninf_covar[pzero < 0.5]
+  covar <- covar[pzero < 0.5]
 
@@
   if(nDE > 0){
-    wts <- rowMeans(dat_alt)
-    DE <- sample(1:nrow(dat_alt), nDE, prob = wts)
+    DE <- sample(1:nrow(dat_alt), nDE, prob = covar)
     truth[DE] <- TRUE
@@
   geneExp <- tbl_df(data.frame(pval=res_alt$pvalue, SE=res_alt$lfcSE, 
-                               ind_covariate = res_alt$baseMean, 
+                               ind_covariate = covar, 
                                effect_size=res_alt$log2FoldChange, 
@@
 
-  if (uninformativeCovariate)
-    geneExp <- mutate(geneExp, ind_covariate = uninf_covar)
+  if (uninformativeCovariate){
+    geneExp <- mutate(geneExp, ind_covariate = runif(length(covar)))
+  }else if(!strongCovariate){
+    geneExp <- mutate(geneExp, ind_covariate = pmin(1, abs(covar + rnorm(length(covar), 0, 0.25))))
+  }
 
@@
   null5 <- readRDS(file=resfile_n5)
-}
-```
-
-## Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(null5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res5 <- plotsim_standardize(null5, alpha = seq(0.01, 0.10, 0.01))
-
-# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
-plotsim_average(res5, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res5, met="TNR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(null5, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(null5, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res5, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(null5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
-} else {
-  message("Not enough methods found rejections at alpha ", ualpha, 
-          "; skipping upset plot")
-}
-
-```
-
-# Sim N10: Null Comparison 10v10
-
-Here we'll repeat the above, but for a null comparison 10 versus 10 samples. 
-This will be done for 100 simulations.
-
-## Generate a list of SB results objects
-
-```{r, results='hide', message=FALSE}
-sampleSize <- 10
-nDE <- 0
-rseed <- 837
-
-if (!file.exists(resfile_n10)){
-  null10 <- mclapply(X=1:B, rseed=837, FUN=simulateOneSplit,
-                     nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
-  saveRDS(null10, file=resfile_n10)
-} else {
-  null10 <- readRDS(file=resfile_n10)
-}
-```
-
-## Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(null10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res10 <- plotsim_standardize(null10, alpha = seq(0.01, 0.10, 0.01))
-
-plotsim_average(res10, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res10, met="TNR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(null10, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(null10, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res10, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(null10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
-} else {
-  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
-}
-
-```
-
-# Sim D5: DE Comparison 5v5
-
-Here we'll repeat the above, but for a DE comparison 5 versus 5 samples
-and 2000 DE genes are added. This will be done for 100 simulations.
-
-## Generate a list of SB results objects
-
-```{r, results='hide', message=FALSE}
-sampleSize <- 5
-nDE <- 2000
-rseed <- 198
-
-if (!file.exists(resfile_d5)){
-  de5 <- mclapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, 
-                  nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
-  saveRDS(de5, file=resfile_d5)
-} else {
-  de5 <- readRDS(file=resfile_d5)
-}
-```
-
-## Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(de5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.20, 0.01))
-
-plotsim_average(res5d, met="rejections",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res5d, met="FDR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res5d, met="TPR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(de5, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(de5, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
-} else {
-  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
-}
-
-```
-
-# Sim D10: DE Comparison 10v10
-
-Here we'll repeat the above, but for a DE comparison 10 versus 10 samples
-and 2000 DE genes are added. This will be done for 100 simulations.
-
-## Generate a list of SB results objects
-
-```{r, results='hide', message=FALSE}
-sampleSize <- 10
-nDE <- 2000
-rseed <- 961
-
-if (!file.exists(resfile_d10)){
-  de10 <- mclapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, 
-                   nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores = nCores)
-  saveRDS(de10, file=resfile_d10)
-} else {
-  de10 <- readRDS(file=resfile_d10)
-}
-
-```
-
-## Plot average results over replications
-
-Plot results.
-
-```{r}
-# Check for missing results (if any methods threw an error for relevant metrics).
-rowSums(sapply(de10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
-
-res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.20, 0.01))
-
-plotsim_average(res10d, met="rejections",filter_set = excludeSet, 
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res10d, met="FDR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) 
-plotsim_average(res10d, met="TPR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE)
-
-covariateLinePlot(de10, alpha=0.05, covname="log2FC", nbins=25, 
-                 trans="log1p")
-covariateLinePlot(de10, alpha=0.05, covname="ind_covariate", nbins=25, 
-                 trans="log1p")
-
-if (numberMethodsReject(res10d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
-  aggupset(de10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
-} else {
-  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
-}
-```
-
-# Sample size comparison
-
-Here we compare the method ranks for the different sample sizes at alpha = 0.10.
-
-```{r}
-plotMethodRanks(c(resfile_d5, resfile_d10), 
-                colLabels = c("DE 5", "DE 10"), 
-                alpha = 0.10, xlab = "Comparison",
-                excludeMethods = NULL)
-```
-
-# Uninformative Covariate
-
-## Sim N5: Null Comparison 5v5
-
-Here we run a null comparison 5 versus 5 RNA-Seq samples. 
-This will be done for 100 simulations.
-
-### Generate a list of SB results objects
-
-```{r, results='hide', message=FALSE}
-sampleSize <- 5
-nDE <- 0
-rseed <- 225
-
-if (!file.exists(resfile_n5_uninfCov)){
-  null5 <- mclapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
-                    nDE=nDE, sampleSize=sampleSize, 
-                   bd=bd, uninformativeCovariate = TRUE, mc.cores=nCores)
-  saveRDS(null5, file=resfile_n5_uninfCov)
-} else {
-  null5 <- readRDS(file=resfile_n5_uninfCov)
 }
@@
 
-## Sim N10: Null Comparison 10v10
+## N10: Null 10v10
 
@@
 
-if (!file.exists(resfile_n10_uninfCov)){
-  null10 <- mclapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
-                     nDE=nDE, sampleSize=sampleSize, 
-                   bd=bd, uninformativeCovariate = TRUE, mc.cores=nCores)
-  saveRDS(null10, file=resfile_n10_uninfCov)
+if (!file.exists(resfile_n10)){
+  null10 <- mclapply(X=1:B, rseed=837, FUN=simulateOneSplit,
+                     nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
+  saveRDS(null10, file=resfile_n10)
 } else {
-  null10 <- readRDS(file=resfile_n10_uninfCov)
+  null10 <- readRDS(file=resfile_n10)
 }
@@
 
-## Sim D5: DE Comparison 5v5
+# Non-null Comparisons with Strongly Informative Covariate
+
+## D5S: DE 5v5 Strong
 
@@
 and 2000 DE genes are added. This will be done for 100 simulations.
+
+### Generate a list of SB results objects
+
+```{r, results='hide', message=FALSE}
+sampleSize <- 5
+nDE <- 2000
+rseed <- 198
+
+if (!file.exists(resfile_d5)){
+  de5 <- mclapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, 
+                  nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores=nCores)
+  saveRDS(de5, file=resfile_d5)
+} else {
+  de5 <- readRDS(file=resfile_d5)
+}
+```
+
+### Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(res5d, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5d, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5d, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(de5, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(de5, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
+  aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
+}
+
+```
+
+## D10S: DE 10v10 Strong
+
+Here we'll repeat the above, but for a DE comparison 10 versus 10 samples
+and 2000 DE genes are added. This will be done for 100 simulations.
+
+### Generate a list of SB results objects
+
+```{r, results='hide', message=FALSE}
+sampleSize <- 10
+nDE <- 2000
+rseed <- 961
+
+if (!file.exists(resfile_d10)){
+  de10 <- mclapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, 
+                   nDE=nDE, sampleSize=sampleSize, bd=bd, mc.cores = nCores)
+  saveRDS(de10, file=resfile_d10)
+} else {
+  de10 <- readRDS(file=resfile_d10)
+}
+
+```
+
+### Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(res10d, met="rejections",filter_set = excludeSet, 
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10d, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10d, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(de10, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(de10, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res10d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
+  aggupset(de10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
+}
+```
+
+# Non-null Comparisons with Weakly Informative Covariate
+
+## D5W: DE 5v5 Weak
+
+Here we'll repeat the above, but using a weakly informative covariate.
+
+### Generate a list of SB results objects
+
+```{r, results='hide', message=FALSE}
+sampleSize <- 5
+nDE <- 2000
+rseed <- 198
+
+if (!file.exists(resfile_d5_w)){
+  de5 <- mclapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, 
+                  nDE=nDE, sampleSize=sampleSize, bd=bd, 
+                  strongCovariate = FALSE, mc.cores=nCores)
+  saveRDS(de5, file=resfile_d5_w)
+} else {
+  de5 <- readRDS(file=resfile_d5_w)
+}
+```
+
+### Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(res5d, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5d, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5d, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(de5, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(de5, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
+  aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
+}
+
+```
+
+## D10W: DE 10v10 Weak
+
+Here we'll repeat the above, but for a DE comparison 10 versus 10 samples.
+
+### Generate a list of SB results objects
+
+```{r, results='hide', message=FALSE}
+sampleSize <- 10
+nDE <- 2000
+rseed <- 961
+
+if (!file.exists(resfile_d10_w)){
+  de10 <- mclapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, 
+                   nDE=nDE, sampleSize=sampleSize, bd=bd, 
+                   strongCovariate = FALSE,
+                   mc.cores = nCores)
+  saveRDS(de10, file=resfile_d10_w)
+} else {
+  de10 <- readRDS(file=resfile_d10_w)
+}
+
+```
+
+### Plot average results over replications
+
+Plot results.
+
+```{r}
+# Check for missing results (if any methods threw an error for relevant metrics).
+rowSums(sapply(de10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))
+
+res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.10, 0.01))
+
+plotsim_average(res10d, met="rejections",filter_set = excludeSet, 
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10d, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10d, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE)
+
+covariateLinePlot(de10, alpha=0.05, covname="log2FC", nbins=25, 
+                 trans="log1p")
+covariateLinePlot(de10, alpha=0.05, covname="ind_covariate", nbins=25, 
+                 trans="log1p")
+
+if (numberMethodsReject(res10d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
+  aggupset(de10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
+} else {
+  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
+}
+```
+
+# Non-null Comparisons with Uninformative Covariate
+
+## D5U: DE 5v5 Uninformative
+
+Here we'll repeat the above, but using a completely uninformative covariate.
 
@@
 
-res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.20, 0.01))
+res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
 
@@
 
-## Sim D10: DE Comparison 10v10
+## D10U: DE 10v10 Uninformative
 
-Here we'll repeat the above, but for a DE comparison 10 versus 10 samples
-and 2000 DE genes are added. This will be done for 100 simulations.
+Here we'll repeat the above, but for a DE comparison 10 versus 10 samples.
 
@@
 
-res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.20, 0.01))
+res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.10, 0.01))
 
@@
 
-## Sample size comparison
+# Sample size comparison
 
-Here we compare the method ranks for the different sample sizes at alpha = 0.10.
+Here we compare the method ranks for the different sample sizes and informative
+covariate settings at alpha = 0.10.
 
 ```{r}
-# uninformative covariate
-plotMethodRanks(c(resfile_d5_uninfCov, resfile_d10_uninfCov), 
-                colLabels = c("DE 5", "DE 10"), 
+plotMethodRanks(c(resfile_d5, resfile_d5_w, resfile_d5_uninfCov,
+                  resfile_d10, resfile_d10_w, resfile_d10_uninfCov), 
+                colLabels = c("DE5 S", "DE5 W", "DE5 U", "DE10 S", "DE10 W", "DE10 U"), 
                 alpha = 0.10, xlab = "Comparison",

### 2018-09-18 manuscript_figures/manuscript_figures_sims.Rmd

@@
 
+# Supplementary Figure (informativeness covariates)
+
+This is a faceted figure illustrating the functional relationship between the informative covariate and the the null proportion, pi0, when using either the continuous or discrete "informativeness" functions.
+The relationship is shown over a range of informativeness values for the two settings.
+These functions were used to investigate the impact of increasing/descreasing informativeness of the covariates while fixing other simulation conditions.
+
+```{r SFigureSim6, fig.width = 12, fig.height = 4}
+## SFigure Sim6 A: Informativeness covariate relationships (continuous)
+cinfovar <- tibble(informativeness = seq(0, 1, by = 0.2))
+cinfovar <- dplyr::mutate(cinfovar, x = list(seq(0, 1, .01)))
+cinfovar <- dplyr::mutate(cinfovar, pi0 = lapply(informativeness, function(zz) pi0_varyinfo80l(zz)(x[[1]])))
+cinfovar <- dplyr::mutate(cinfovar, informativeness = informativeness * 100)
+cinfovar <- tidyr::unnest(cinfovar)
+
+ps6a <- ggplot(cinfovar, aes(x = x, y = pi0)) +
+    geom_line() +
+    expand_limits(y = c(0.5, 1)) + 
+    theme_classic() +
+    scale_x_continuous(breaks = seq(0, 1, .2)) + 
+    scale_y_continuous(labels = scales::percent) + 
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    facet_grid(. ~ informativeness, labeller = label_both)
+
+## SFigure Sim6 B: Informativeness covariate relationships (discrete)
+dinfovar <- tibble(informativeness = seq(0, 1, by = 0.2))
+dinfovar <- dplyr::mutate(dinfovar, x = list(seq(0, 1, .01)))
+dinfovar <- dplyr::mutate(dinfovar, pi0 = lapply(informativeness, function(zz) pi0_varyinfo80(zz)(x[[1]])))
+dinfovar <- dplyr::mutate(dinfovar, informativeness = informativeness * 100)
+dinfovar <- tidyr::unnest(dinfovar)
+
+ps6b <- ggplot(dinfovar, aes(x = x, y = pi0)) +
+    geom_line() +
+    expand_limits(y = c(0.5, 1)) + 
+    theme_classic() +
+    scale_x_continuous(breaks = seq(0, 1, .2)) + 
+    scale_y_continuous(labels = scales::percent) + 
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    facet_grid(. ~ informativeness, labeller = label_both)
+
+## pull SFigure Sim6 together
+SFigS6 <- plot_grid(ps6a, ps6b, labels = LETTERS[1:2], ncol = 1)
+SFigS6
+ggsave(file.path(outdir, "SFigureSim6.pdf"), width = 12, height = 4)
+```
+
 # Session information

### 2018-10-09 datasets/R/plotsim_standardize.R
Add FPR to performance metrics for simulations

@@ -29,6 +29,16 @@ plotsim_standardize <- function(res, alpha = seq(0.01, 0.10, 0.01)) {
                       function( query, truth, alpha = 0.1) {
                           mean(query < alpha, na.rm = TRUE)
                       })
+    
+    # add FPR metric for ROC curve
+    sbl <- lapply(sbl, addPerformanceMetric,
+                  assay="qvalue", evalMetric="FPR",
+                  evalFunction = 
+                    function( query, truth, alpha=0.1 ){
+                    sum( query <= alpha & truth == 0, na.rm = TRUE ) / 
+                         sum( truth == 0, na.rm = TRUE )
+                    })
+    
     # override default FDR perf. metric st fdr is zero (instead of NA) when
     # there are zero rejections
     sbl <- lapply(sbl, addPerformanceMetric,

### 2018-10-09 datasets/R/plotsim_average.R
Add ROC plot to simulation plotter

@@ -6,8 +6,10 @@
 #' 
 #' @param tsb standardized metric data.frame generated using
 #'        standardize_results.
-#' @param met name of metric to plot - must be one of the performance
-#'        metrics added by default with "addDefaultMetrics" or FWER.
+#' @param met name(s) of metric to plot - must be one of the performance
+#'        metrics added by default with "addDefaultMetrics" or FWER or FPR.
+#'        May be a vector of names of length 2 - in this case the first will
+#'        be used for the x-axis and the second for the y-axis. 
 #' @param filter_set character vector of "blabel" IDs of methods which
 #'        shouldn't be included in the plot. Alternatively, a subset of
 #'        methods can be chosen by filtering the data.table before passing
@@ -38,7 +40,9 @@ plotsim_average <- function(tsb, met, filter_set = NULL, merge_ihw = TRUE,
                             clean_names = FALSE, errorBars=FALSE,
                             palette = candycols, facetMethodType = FALSE,
                             diffplot = FALSE, grpVars = NULL){
-
+   if (length(met)>2)
+     stop("Can only plot 2 metrics at a time")
+  
     ## cacluate mean per replication
     if(!is.null(grpVars)){
       tsba <- tsb %>%
@@ -69,10 +73,21 @@ plotsim_average <- function(tsb, met, filter_set = NULL, merge_ihw = TRUE,
         tsba$blabel[grepl("^ihw-", tsba$blabel)] <- "ihw"
     }
     
-    # filter by performance metric 
-    tsba <- tsba %>% 
-      filter(performanceMetric == met) %>%
-      dplyr::mutate(Method = gsub("-df03", "", blabel)) 
+    # filter by performance metric
+    tsba_m <- NULL 
+    for (m in seq_along(met)){
+      tmp <- tsba %>% 
+        filter(performanceMetric == met[m]) %>%
+        dplyr::mutate(Method = gsub("-df03", "", blabel)) 
+      if (m > 1){
+        tsba_m <- left_join(tsba_m, tmp, 
+                            by = c("Method", "alpha", "n", 
+                                   "param.alpha", "blabel"))
+      }else{
+        tsba_m <- tmp
+      }
+    }
+    tsba <- tsba_m
            
     if (clean_names) {
         ulabs <- unique(tsba$blabel)
[…]

### 2018-10-09 manuscript_figures/manuscript_figures.Rmd
Consistent subtitle labels

@@
 size <-  ifelse(grepl("10", resfile), "10", "5")
-mode <-  ifelse(grepl("II", resfile), "Bimodal effect Size", "Unimodal Effect Size")
+mode <-  ifelse(grepl("II", resfile), "Bimodal effect size", "Unimodal effect size")
 pi0 <-  ifelse(grepl("H", resfile), "7.5% Non-null", "30% Non-null")

### 2018-10-09 datasets/RNA-Seq/polyester-simulation.Rmd
Update authorship of polyester

@@
 title: "Polyester RNA-seq simulation study"
-author: "Stephanie Hicks"
+author: "Stephanie Hicks & Keegan Korthauer"
 output: 

### 2018-10-09 datasets/RNA-Seq/polyester-simulation.Rmd
Add ROC curves to yeast & polyester

@@
                 merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5d, met=c("FPR", "TPR"), filter_set = excludeSet,
+                merge_ihw = TRUE)
 
@@
                 merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10d, met=c("FPR", "TPR"), filter_set = excludeSet,
+                merge_ihw = TRUE)
 

### 2018-10-09 datasets/RNA-Seq/yeast-simulation-bimodalalternative.Rmd
Add ROC curves to yeast & polyester

@@
                 merge_ihw = TRUE, errorBars=TRUE)
+plotsim_average(res5d, met=c("FPR", "TPR"), filter_set = excludeSet,
+                merge_ihw = TRUE)
 

### 2018-10-11 manuscript_figures/manuscript_figures.Rmd
Swap panel of Fig 1 to include ROC

@@
 de5 <- readRDS(file=resfile)
-excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
+excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05", "bonf", "fdrreg-e")
 res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
@@
 
-# Figure 2C: Total number of discoveries varied across method
-p3 <- plotsim_average(res5d, met="rejections",filter_set = excludeSet,
+# Figure 2C: ROC curve
+p3 <- plotsim_average(res5d, met=c("FPR", "TPR"),filter_set = excludeSet,
                 merge_ihw = TRUE, errorBars=TRUE) +
       theme(legend.position="none") +
-      ylab("Number of Rejections") +
-      ggtitle("Rejections by alpha level") +
+      #ylab("Number of Rejections") +
+      ggtitle("ROC (FPR vs TPR)") +
       scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02))

### 2018-10-12 manuscript_figures/manuscript_figures.Rmd
Fix axis label in Fig1

@@
       theme(legend.position="none") +
-      #ylab("Number of Rejections") +
       ggtitle("ROC (FPR vs TPR)") +
-      scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02))
+      scale_x_continuous("FPR", breaks=seq(0, 1, by=0.01))
 

### 2018-10-16 manuscript_figures/manuscript_figures_main.Rmd
Add separate Rmd for reworked main figures

---
title: "Manuscript Figures for Genome Biology"
author: "Rafalab"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This Rmd generates the main manuscript figures. Additional Rmds
generate the supplementary case study and in silico experiment figures 
(`manuscript_figures.Rmd`) and supplementary case study figures 
(`manuscript_figures_sims.Rmd`).

# Set up workspace

```{r Figure2-load-data}
# Load packages and source benchmark FDR
library(tidyr)
library(dplyr)
library(ggplot2)
library(magrittr)
library(cowplot)
library(tibble)
library(ggthemes)
library(grid)
library(SummarizedBenchmark)

## load helper functions
for (f in list.files("../datasets/R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

# Assumes sb objects for the case studies and in silico experiments are in 
# the following location, which contains subfolders
# for each casestudy (if this isn't true, then parsing the case study and 
# dataset names later on will be incorrect)
path <- "../results"

# Also assumes that simulation results summary file is in the following location
# The `result-metrics.rds` was generated in the 
# `datasets/simulations/simulations-summary.Rmd` file.
resmet_file <- file.path("..", "datasets", "simulations", "results-summary", "result-metrics.rds")

# set up results directory
outdir <- "./figures"
dir.create(outdir, showWarnings = FALSE)

# set alpha cutoff for plots that are fixed at a certain value of alpha
alpha.thresh <- 0.05

# methods to include in all figures ( exclude bonferroni and fdrreg-e)
methodset <- c("bh", "ihw", paste0("ihw-a0", 1:9), "ihw-a10", 
               "qvalue", "bl", "bl-df03", "lfdr",
               "fdrreg-t", "ashq", "adapt-glm")
```

We use the standardize candy color scheme and line types for the plots.

```{r}
col <- as.character(candycols$col)
names(col) <- as.character(candycols$Method)
lty <- as.character(candycols$lty)
names(lty) <- as.character(candycols$Method)
```

To generate the figures in this document, the simulation results must first be aggregated by running the code at `datasets/simulations/simulations-summary.Rmd`.

```{r load-sim-metrics}
resmet <- readRDS(resmet_file)
```

Unfortunately, because the plots in this section illustrate FDR and TPR acorss varying simulation settings rather than varying nominal FDR cutoff, we cannot simply use the `plotsim_average()` function to generate plots. Instead, we define the following helper function to consistently generate plots similar to those output by `plotsim_average()` with arbitrary covariate, e.g. number of tests or proportion of null hypotheses, as the x-axis. The function assumes that the input table only includes FDR or TPR values for each method at a single nominal alpha cutoff. In these figures, we only plot FDR and TPR values at the nominal alpha cutoff of 0.05.

```{r genplot-function}
genplot <- function(tab, cov, type = c("info", "diff"), xt = "", met = c("FDR", "TPR"), ebw = 0.0025) {
    met <- match.arg(met)
    cov <- rlang::enquo(cov)
    type <- match.arg(type)
    if (type == "info") {
        ymean <- rlang::quo(mean.info)
        yse <- rlang::quo(se.info)
    } else if (type == "diff") {
        ymean <- rlang::quo(mean.diff)
        yse <- rlang::quo(se.diff)
    }        
    gp <- dplyr::filter(tab, key == met) %>%
        ggplot(aes(x = !!cov, y = !!ymean, color = Method, group = Method)) +
        geom_line(aes(linetype = Method), alpha = 0.85) +
        geom_errorbar(aes(ymin = !!ymean - !!yse, ymax = !!ymean + !!yse), width = ebw, alpha=0.5) +
        scale_linetype_manual(values = lty) + 
        scale_color_manual(values = col) +
        expand_limits(y = 0) +
        theme_classic() +
        theme(axis.title = element_text(face = "bold", size = 10),
              plot.title = element_text(face = "bold", size = 12))
    if (type == "diff") {
        gp <- gp + geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2)
        gp <- gp + scale_y_continuous(bquote(Delta ~ .(met) ~ (informative-uninformative)),
                                      labels = scales::percent)
    } else {
        gp <- gp + scale_y_continuous(met, labels = scales::percent)
        if (met == "FDR") {
            if (rlang::quo_name(cov) == "alpha") {
                gp <- gp + geom_abline(lty = 2, color = "blue", alpha = 1/2) 
            } else {
                gp <- gp + geom_hline(yintercept = 0.05, lty = 2, color = "blue", alpha = 1/2) 
            }
        }
    }
    gp
}
```

# Figure 1 - FDR

This figure contains the main FDR results for the simulations and *in silico* 
experiments. Panel (A) contains the FDR results for a representative setting of
the yeast *in silico* experiments, showing that all methods control FDR. 
Panels (B) and (C) contain the FDR results for simulations that vary the number
of hypotheses and the proportion of null hypotheses, respectively. These results
show that some methods fail to control FDR under certain settings. We restrict
this plot to have the same y-axes.

Instead of three separate panels showing disjoint experiments/settings, this result
may be better shown by making a heat map with one column per experiment/setting, 
and one row per method, where the color shading indicates whether FDR is controlled
at the target level (for a fixed alpha level, say 0.05). The color scale could be 
divergent, such that points that lose FDR control are highlighted, and points that
are overly conservative are also identifiable.

```{r Figure3, fig.width = 10, fig.height = 12}
## Figure 1A: FDR of methods in representative yeast setting at alpha.thresh
objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
resfile <- objects[grepl("yeast-", objects) & 
                   grepl("de5", objects) & 
                  !grepl("uninf", objects)] 
de5 <- readRDS(file=resfile)
res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01)) %>% 
  filter(blabel %in% methodset)
p1a <- plotsim_average(res5d, met="FDR",
                merge_ihw = TRUE, errorBars=TRUE) +
      #ggtitle(expression(paste(bold("Yeast "), bolditalic("in silico"), bold(" experiment")))) +
      ggtitle("FDR control in yeast") +
      scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)) +
      expand_limits(y=0.12) +
      theme(plot.title = element_text(face = "bold", size = 12))

## Figure 1B: FDR of methods across varying number of tests at alpha.thresh
res_ntests <- dplyr::filter(resmet, setting == "varyingntests", 
                            alpha == alpha.thresh, Method %in% methodset)
p1b <- genplot(res_ntests, cov = ntests, met = "FDR", ebw = 0.05) +
    scale_x_continuous("Number of Tests", trans = "log10",
                       breaks = c(1e2, 5e2, 1e3, 5e3, 1e4, 5e4)) +
    guides(color = FALSE, linetype = FALSE) +
    ggtitle("Varying number of tests") +
    coord_cartesian(ylim = c(0,0.12))

## Figure 1C: FDR of methods across varying null proportion at alpha.thresh
res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0", 
                         alpha == alpha.thresh, Method %in% methodset)
p1c <- genplot(res_pi0, cov = 100-pi0, met = "FDR", ebw = 2.5) +
    scale_x_continuous("Proportion of Non-Null Hypotheses",
                       breaks = 100 - c(5, seq(10, 90, by = 10), 95, 99)) + 
  guides(color = FALSE, linetype = FALSE) + 
  ggtitle("Varying non-null proportion") +
  expand_limits(y=0.12)

## pull Figure 1 together
Fig1 <- plot_grid(p1a + theme(legend.position="none"), p1b, p1c, nrow = 1)
Fig1 <- plot_grid(Fig1, get_legend(p1a), rel_widths = c(1, .2))
Fig1
ggsave(file.path(outdir, "Fig1.pdf"), width=10, height=3)
```

# Figure 2 - Power

# Figure 3 - Robustness

# Figure 4 - Applicability

# Figure 5 - Summary

The summary of recommendations is created manually.

# Session information

```{r}
sessionInfo()
```

### 2018-10-17 manuscript_figures/manuscript_figures_main.Rmd
Add figure 2 (TPR)

@@
         theme_classic() +
-        theme(axis.title = element_text(face = "bold", size = 10),
+        theme(axis.title = element_text(face = "bold", size = 11),
               plot.title = element_text(face = "bold", size = 12))
@@
 experiments. Panel (A) contains the FDR results for a representative setting of
-the yeast *in silico* experiments, showing that all methods control FDR. 
-Panels (B) and (C) contain the FDR results for simulations that vary the number
-of hypotheses and the proportion of null hypotheses, respectively. These results
-show that some methods fail to control FDR under certain settings. We restrict
-this plot to have the same y-axes.
+the yeast *in silico* experiments, showing that all methods control FDR. Panel 
+(B) contains the FDR results for the polyester read count simulation, showing that
+most methods control the FDR, with the exception of lfdr and ash, which have slightly
+inflated FDR. Panels (C) and (D) contain the FDR results for simulations that 
+vary the number of hypotheses and the proportion of null hypotheses, respectively. 
+These results show that ash and lfdr fail to control FDR under certain settings.
 
-Instead of three separate panels showing disjoint experiments/settings, this result
-may be better shown by making a heat map with one column per experiment/setting, 
-and one row per method, where the color shading indicates whether FDR is controlled
-at the target level (for a fixed alpha level, say 0.05). The color scale could be 
-divergent, such that points that lose FDR control are highlighted, and points that
-are overly conservative are also identifiable.
-
-```{r Figure3, fig.width = 10, fig.height = 12}
-## Figure 1A: FDR of methods in representative yeast setting at alpha.thresh
+```{r Figure1, fig.width = 7.5, fig.height = 6.5}
+# FDR in in silico experiments across alpha values 
+## Figure 1A: FDR of methods in representative yeast setting at varying alpha levels
 objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
@@
                 merge_ihw = TRUE, errorBars=TRUE) +
-      #ggtitle(expression(paste(bold("Yeast "), bolditalic("in silico"), bold(" experiment")))) +
-      ggtitle("FDR control in yeast") +
-      scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)) +
-      expand_limits(y=0.12) +
+      ggtitle("Resampling with spike-ins") +
+      scale_x_continuous(expression(paste(bold(alpha), bold(" level"))),
+                         breaks=seq(0, 1, by=0.02)) +
+      scale_y_continuous("FDR", labels=scales::percent, breaks=seq(0, 1, by=0.02)) +
+      expand_limits(y=0.11) +
       theme(plot.title = element_text(face = "bold", size = 12))
 
-## Figure 1B: FDR of methods across varying number of tests at alpha.thresh
+## Figure 1B: FDR of methods in representative polyester setting at varying alpha levels
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+resfile <- objects[grepl("polyester-", objects) & 
+                   grepl("de5", objects) & 
+                  !grepl("uninf", objects)] 
+de5 <- readRDS(file=resfile)
+res5dp <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01)) %>% 
+  filter(blabel %in% methodset)
+p1b <- plotsim_average(res5dp, met="FDR",
+                merge_ihw = TRUE, errorBars=TRUE) +
+      ggtitle("Polyester count simulation") +
+      scale_x_continuous(expression(paste(bold(alpha), bold(" level"))),
+                         breaks=seq(0, 1, by=0.02)) +
+      scale_y_continuous("", labels=scales::percent, breaks=seq(0, 1, by=0.02)) +
+      expand_limits(y=0.11) +
+      theme(plot.title = element_text(face = "bold", size = 12)) + 
+  theme(legend.position="none")
+
+p1abTitle <- ggdraw() +
+    draw_label(expression(paste(bold("FDR control in RNA-seq "), 
+                                bolditalic("in silico"), bold(" experiments"))))
+
+# Panels C and D -> FDR in sims across settings at a single specified alpha
+## Figure 1C: FDR of methods across varying number of tests at alpha.thresh
 res_ntests <- dplyr::filter(resmet, setting == "varyingntests", 
                             alpha == alpha.thresh, Method %in% methodset)
-p1b <- genplot(res_ntests, cov = ntests, met = "FDR", ebw = 0.05) +
-    scale_x_continuous("Number of Tests", trans = "log10",
-                       breaks = c(1e2, 5e2, 1e3, 5e3, 1e4, 5e4)) +
+p1c <- genplot(res_ntests, cov = ntests, met = "FDR", ebw = 0.1) +
+    scale_x_continuous("Number of tests", trans = "log10",
+                       breaks = c(1e2, 5e2, 1e3, 5e3, 1e4, 5e4),
+                       labels = c("100", "500", "1000", "5000", "1e4", "5e4")) +
     guides(color = FALSE, linetype = FALSE) +
+    scale_y_continuous("FDR", labels = scales::percent, breaks=seq(0,1,by=0.02), 
+                       limits=c(0,.12)) +
     ggtitle("Varying number of tests") +
@@
 
-## Figure 1C: FDR of methods across varying null proportion at alpha.thresh
+## Figure 1D: FDR of methods across varying null proportion at alpha.thresh
 res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0", 
-                         alpha == alpha.thresh, Method %in% methodset)
-p1c <- genplot(res_pi0, cov = 100-pi0, met = "FDR", ebw = 2.5) +
-    scale_x_continuous("Proportion of Non-Null Hypotheses",
-                       breaks = 100 - c(5, seq(10, 90, by = 10), 95, 99)) + 
+                         alpha == alpha.thresh, Method %in% methodset,
+                         pi0 >= 10) # restrict to pi0 at least 10%
+p1d <- genplot(res_pi0, cov = 100-pi0, met = "FDR", ebw = 2.5) +
+    scale_x_continuous("Proportion non-null",
+                       breaks = 100 - c(seq(10, 90, by = 10), 95, 99)) + 
   guides(color = FALSE, linetype = FALSE) + 
   ggtitle("Varying non-null proportion") +
-  expand_limits(y=0.12)
+  scale_y_continuous("", labels = scales::percent, breaks=seq(0,1,by=0.02), 
+                       limits=c(0,.12))
+
+p1cdTitle <- ggdraw() +
+    draw_label(expression(paste(bold("FDR across simulation settings ("), 
+                                bold(alpha), bold(" = 0.05)"))))
 
 ## pull Figure 1 together
-Fig1 <- plot_grid(p1a + theme(legend.position="none"), p1b, p1c, nrow = 1)
+top <- plot_grid(p1a + theme(legend.position="none"), p1b, nrow = 1)
+bot <- plot_grid(p1c, p1d, nrow = 1)
+Fig1 <- plot_grid(p1abTitle, top, p1cdTitle, bot, nrow=4, 
+                  rel_heights = c(0.1, 1, 0.1, 1))
 Fig1 <- plot_grid(Fig1, get_legend(p1a), rel_widths = c(1, .2))
 Fig1
-ggsave(file.path(outdir, "Fig1.pdf"), width=10, height=3)
+ggsave(file.path(outdir, "Fig1.pdf"), width=7.5, height=6.5)
 ```
@@
 # Figure 2 - Power
+
+This figure contains the main TPR results for the simulations and *in silico* 
+experiments, and is analagous to Figure 1. 
+Panels (A) and (B) contain the TPR results for a representative setting of
+the yeast *in silico* experiments and polyester simulations, showing that in 
+general, modern FDR-controlling methods have modestly higher power than classic methods. 
+In addition, lfdr, fdrreg-t, and adapt tend to have highest power compared to other
+methods.
+Panels (C) and (D) contain the TPR results for simulations that 
+vary the number of hypotheses and the proportion of null hypotheses, respectively. 
+These results show that adapt and lfdr are very conservative under some settings 
+of the number of tests, and that the relative ranking of methods is somewhat 
+robust under varying proportions of null hypotheses.
+
+```{r Figure2, fig.width = 7.5, fig.height = 6.5}
+# TPR in in silico experiments across alpha values 
+## Figure 2A: TPR of methods in representative yeast setting at varying alpha levels
+p2a <- plotsim_average(res5d, met="TPR",
+                merge_ihw = TRUE, errorBars=TRUE) +
+      ggtitle("Resampling with spike-ins") +
+      scale_x_continuous(expression(paste(bold(alpha), bold(" level"))),
+                         breaks=seq(0, 1, by=0.02)) +
+      scale_y_continuous("TPR", labels=scales::percent, breaks=seq(0, 1, by=0.05)) +
+      theme(plot.title = element_text(face = "bold", size = 12))
+
+## Figure 2B: TPR of methods in representative polyester setting at varying alpha levels
+p2b <- plotsim_average(res5dp, met="TPR",
+                merge_ihw = TRUE, errorBars=TRUE) +
+      ggtitle("Polyester count simulation") +
+      scale_x_continuous(expression(paste(bold(alpha), bold(" level"))),
+                         breaks=seq(0, 1, by=0.02)) +
+      scale_y_continuous("", labels=scales::percent, breaks=seq(0, 1, by=0.05)) +
+      theme(plot.title = element_text(face = "bold", size = 12)) + 
+  theme(legend.position="none")
+
+p2abTitle <- ggdraw() +
+    draw_label(expression(paste(bold("TPR in RNA-seq "), 
+                                bolditalic("in silico"), bold(" experiments"))))
+
+# Panels C and D -> TPR in sims across settings at a single specified alpha
+## Figure 2C: TPR of methods across varying number of tests at alpha.thresh
+res_ntests <- dplyr::filter(resmet, setting == "varyingntests", 
+                            alpha == alpha.thresh, Method %in% methodset)
+p2c <- genplot(res_ntests, cov = ntests, met = "TPR", ebw = 0.1) +
+    scale_x_continuous("Number of tests", trans = "log10",
+                       breaks = c(1e2, 5e2, 1e3, 5e3, 1e4, 5e4),
+                       labels = c("100", "500", "1000", "5000", "1e4", "5e4")) +
+    guides(color = FALSE, linetype = FALSE) +
+    scale_y_continuous("TPR", labels = scales::percent, breaks=seq(0,1,by=0.1)) +
+    ggtitle("Varying number of tests")
+
+## Figure 1D: FDR of methods across varying null proportion at alpha.thresh
+res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0", 
+                         alpha == alpha.thresh, Method %in% methodset,
+                         pi0 >= 10) # restrict to pi0 at least 10%
+p2d <- genplot(res_pi0, cov = 100-pi0, met = "TPR", ebw = 2.5) +
+    scale_x_continuous("Proportion non-null",
+                       breaks = 100 - c(seq(10, 90, by = 10), 95, 99)) + 
+  guides(color = FALSE, linetype = FALSE) + 
+  ggtitle("Varying non-null proportion") +
+  scale_y_continuous("", labels = scales::percent, breaks=seq(0,1,by=0.20))
+
+p2cdTitle <- ggdraw() +
+    draw_label(expression(paste(bold("TPR across simulation settings ("), 
+                                bold(alpha), bold(" = 0.05)"))))
+
+## pull Figure 1 together
+top <- plot_grid(p2a + theme(legend.position="none"), p2b, nrow = 1)
+bot <- plot_grid(p2c, p2d, nrow = 1)
+Fig2 <- plot_grid(p2abTitle, top, p2cdTitle, bot, nrow=4, 
+                  rel_heights = c(0.1, 1, 0.1, 1))
+Fig2 <- plot_grid(Fig2, get_legend(p2a), rel_widths = c(1, .2))
+Fig2
+ggsave(file.path(outdir, "Fig2.pdf"), width=7.5, height=6.5)
+```
 
@@
 
-The summary of recommendations is created manually.
+The summary of recommendations figure is drawn manually.
 

### 2018-10-17 manuscript_figures/manuscript_figures_main.Rmd
Change lfdr to dotted line below 4000 tests

@@
 
-We use the standardize candy color scheme and line types for the plots.
+We use the standardize candy color scheme and line types for the plots. We'll
+add the "lfdr*" method, which indicates lfdr was applied with fewer than 200 
+tests per bin (out of 20 bins).
 
@@
 This figure contains the main FDR results for the simulations and *in silico* 
-experiments. Panel (A) contains the FDR results for a representative setting of
+experiments. Panel (A, left) contains the FDR results for a representative setting of
 the yeast *in silico* experiments, showing that all methods control FDR. Panel 
-(B) contains the FDR results for the polyester read count simulation, showing that
+(A, right) contains the FDR results for the polyester read count simulation, showing that
 most methods control the FDR, with the exception of lfdr and ash, which have slightly
-inflated FDR. Panels (C) and (D) contain the FDR results for simulations that 
+inflated FDR. Panels (B) contains the FDR results for simulations that 
 vary the number of hypotheses and the proportion of null hypotheses, respectively. 
@@
                             alpha == alpha.thresh, Method %in% methodset) 
-p1c <- genplot(res_ntests, cov = ntests, met = "FDR", ebw = 0.1) +
+p1c <- genplot(res_ntests %>% filter(ntests >= 4000 | Method != "lfdr"), 
+               cov = ntests, met = "FDR", ebw = 0.1) +
     scale_x_continuous("Number of tests", trans = "log10",
                        breaks = c(1e2, 5e2, 1e3, 5e3, 1e4, 5e4),
-                       labels = c("100", "500", "1000", "5000", "1e4", "5e4")) +
+                       labels = c("100", "500", "1e3", "5e3", "1e4", "5e4")) +
     guides(color = FALSE, linetype = FALSE) +
-    scale_y_continuous("FDR", labels = scales::percent, breaks=seq(0,1,by=0.02), 
-                       limits=c(0,.12)) +
+    scale_y_continuous("FDR", labels = scales::percent, breaks=seq(0,1,by=0.02)) +
     ggtitle("Varying number of tests") +
-    coord_cartesian(ylim = c(0,0.12))
+    coord_cartesian(ylim = c(0,0.12)) + 
+    geom_line(data = res_ntests %>% 
+                filter(ntests <= 5000 & Method == "lfdr", performanceMetric == "FDR"),
+              aes(x = ntests, y = mean.info), 
+              linetype = "dashed", color = "dodgerblue3", alpha = 0.88) +
+    geom_errorbar(data = res_ntests %>% 
+                 filter(ntests <= 5000 & Method == "lfdr", performanceMetric == "FDR"),
+                 aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.1, alpha=0.5)
 
@@
 Fig1 <- plot_grid(p1abTitle, top, p1cdTitle, bot, nrow=4, 
-                  rel_heights = c(0.1, 1, 0.1, 1))
+                  rel_heights = c(0.1, 1, 0.1, 1),
+                  labels = c("A", "", "B", ""))
 Fig1 <- plot_grid(Fig1, get_legend(p1a), rel_widths = c(1, .2))
@@
 experiments, and is analagous to Figure 1. 
-Panels (A) and (B) contain the TPR results for a representative setting of
+Panel (A) contains the TPR results for a representative setting of
 the yeast *in silico* experiments and polyester simulations, showing that in 
@@
 methods.
-Panels (C) and (D) contain the TPR results for simulations that 
+Panel (B) contains the TPR results for simulations that 
 vary the number of hypotheses and the proportion of null hypotheses, respectively. 
@@
                             alpha == alpha.thresh, Method %in% methodset)
-p2c <- genplot(res_ntests, cov = ntests, met = "TPR", ebw = 0.1) +
+p2c <- genplot(res_ntests %>% filter(ntests >= 4000 | Method != "lfdr"),
+               cov = ntests, met = "TPR", ebw = 0.1) +
     scale_x_continuous("Number of tests", trans = "log10",
                        breaks = c(1e2, 5e2, 1e3, 5e3, 1e4, 5e4),
-                       labels = c("100", "500", "1000", "5000", "1e4", "5e4")) +
+                       labels = c("100", "500", "1e3", "5e3", "1e4", "5e4")) +
     guides(color = FALSE, linetype = FALSE) +
     scale_y_continuous("TPR", labels = scales::percent, breaks=seq(0,1,by=0.1)) +
-    ggtitle("Varying number of tests")
+    ggtitle("Varying number of tests")  + 
+    geom_line(data = res_ntests %>% 
+                filter(ntests <= 5000 & Method == "lfdr", performanceMetric == "TPR"),
+              aes(x = ntests, y = mean.info), 
+              linetype = "dashed", color = "dodgerblue3", alpha = 0.88) +
+    geom_errorbar(data = res_ntests %>% 
+                 filter(ntests <= 5000 & Method == "lfdr", performanceMetric == "TPR"),
+                 aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.1, alpha=0.5)
 
@@
 Fig2 <- plot_grid(p2abTitle, top, p2cdTitle, bot, nrow=4, 
-                  rel_heights = c(0.1, 1, 0.1, 1))
+                  rel_heights = c(0.1, 1, 0.1, 1),
+                  labels = c("A", "", "B", ""))
 Fig2 <- plot_grid(Fig2, get_legend(p2a), rel_widths = c(1, .2))
@@
 # Figure 4 - Applicability
+
+This figure shows the applicability of each of the methods to a variety of data types
+within the case studies, as well as to different types of statistical tests both 
+within the case studies and simulation settings. Panel (A)
 

### 2018-10-17 manuscript_figures/manuscript_figures_main.Rmd
Add Figure 4

@@
 within the case studies, as well as to different types of statistical tests both 
-within the case studies and simulation settings. Panel (A)
+within the case studies and simulation settings. Panel (A) displays the performance 
+across different test statistic distributions, and panel (B) displays the performance
+across different case studies.
+
+```{r Figure4, fig.width = 11, fig.height = 11}
+res_noise <- dplyr::filter(resmet, setting == "informative-cubic") %>%
+  filter(Method %in% methodset)
+
+## Distribution of unimodal effect sizes (based on random samples)
+set.seed(100)
+tsdists <- bind_rows(dplyr::mutate(simIteration(1, NULL, m = 2e4, pi0 = 0.9,
+                                                es_dist = rnorm_generator(3),
+                                                ts_dist = rnorm_perturber(1),
+                                                null_dist = rnorm_2pvaluer(1),
+                                                icovariate = runif, execute = FALSE),
+                                   dist = "\"Distribution: \" * N(0, 1)"),
+                     dplyr::mutate(simIteration(1, NULL, m = 2e4, pi0 = 0.9,
+                                                es_dist = rnorm_generator(3),
+                                                ts_dist = rt_perturber(5),
+                                                null_dist = rt_2pvaluer(5),
+                                                icovariate = runif, execute = FALSE),
+                                   dist = "\"Distribution: \" * t[5]"),
+                     dplyr::mutate(simIteration(1, NULL, m = 2e4, pi0 = 0.9,
+                                                es_dist = rnorm_generator(3),
+                                                ts_dist = rt_perturber(11),
+                                                null_dist = rt_2pvaluer(11),
+                                                icovariate = runif, execute = FALSE),
+                                   dist = "\"Distribution: \" * t[11]"),
+                     dplyr::mutate(simIteration(1, NULL, m = 2e4, pi0 = 0.9,
+                                                es_dist = function(x) { abs(rnorm_generator(15)(x)) },
+                                                ts_dist = rchisq_perturber(4),
+                                                null_dist = rchisq_pvaluer(4),
+                                                icovariate = runif, execute = FALSE),
+                                   dist = "\"Distribution: \" * {\n    chi^2\n}[4]"))
+tsdists <- dplyr::mutate(tsdists, dist = factor(dist, levels = levels(factor(dist))[c(2, 3, 4, 1)]))
+tsdists <- dplyr::rename(tsdists, truth = qvalue)
+tsdists <- dplyr::mutate(tsdists, truth = factor(truth, levels = 0:1, labels = c("null", "non-null")))
+
+## Figure 4A: Performance across distribution of test statistics (based on single replicate)
+p4a <- ggplot(tsdists, aes(x = test_statistic, y=..count..,
+                            group = truth, color = truth)) +
+    stat_density(geom = "line", position = "identity", adjust = 1/2) + 
+    theme_classic() +
+    scale_color_brewer("Truth", palette = "Set1", direction = -1) + 
+    scale_x_continuous("test statistic (single replicate)") +
+    coord_cartesian(xlim = c(-5, 25)) + 
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    facet_grid(. ~ dist, labeller = label_parsed) 
+p4a <- plot_grid(p4a, genplot(res_noise, cov = alpha, met = "FDR") +
+                   facet_grid(. ~ dist, labeller = label_parsed) +
+                   scale_x_continuous(expression(paste(bold(alpha), bold(" level"))),
+                         breaks=seq(0, 1, by=0.02)) +
+                   expand_limits(x = 0) , axis = "trbl", align="hv",
+                 nrow=2, rel_heights = c(0.75,1))
+
+Titlep4a <- ggdraw() +
+    draw_label(expression(paste(bold("FDR across test statistic distributions in simulation"))))
+p4a <- plot_grid(Titlep4a, p4a, nrow = 2, rel_heights = c(0.1,1), labels = c("A", ""))
+
+## Figure 4B: Case study heatmap demonstrating variability in applicability
+
+## read list of case study results
+objects <- list.files(path, recursive=TRUE, pattern="rds", full.names=TRUE)
+
+## exclude yeast and polyester simulations, as well as uninformative covariate
+## analyses
+objects <- objects[!grepl("yeast", objects) & !grepl("polyester", objects) &
+                   !grepl("uninf", objects)]
+
+## create shorter names for case studies
+object_labs <- gsub("\\.rds", "", basename(objects))
+object_labs <- gsub("-benchmark", "", object_labs)
+
+## read in case study results (proportion of rejects vs. case study max)
+objdat <- tidy_df(objects=objects, colLabels=object_labs, 
+                  fill="propMaxRejections", annotate=NULL,
+                  alpha =0.10)
+objdat <- as_tibble(objdat)
+objdat <- tidyr::complete(objdat, method, casestudy)
+
+## exclude 'enigma-so4-otus' results with no significant calls
+objdat <- dplyr::filter(objdat, casestudy != "enigma-so4-otus")
+
+## create table of case study-to-assay mapping
+studies <- tibble(
+    casestudy = c("cbp-csaw", "h3k4me3-csaw", "human", "mouse", "bmi-maf", "bmi-samplesize",
+                  "enigma-al-otus", "enigma-ph-otus", "enigma-so4-otus", "goodrich-genus",
+                  "papa-genus", "schubert-otus", "brain", "mir200c",
+                  "human-mast-det", "human-mast-mean", "human-scdd-det",
+                  "human-scdd-mean", "human-wilcox-det", "human-wilcox-mean",
+                  "mouse-mast-det", "mouse-mast-mean", "mouse-scdd-det",
+                  "mouse-scdd-mean", "mouse-wilcox-det", "mouse-wilcox-mean"),
+    assay = c(rep("ChIP-seq", 2), rep("GSEA", 2), rep("GWAS", 2), rep("Microbiome", 6),
+              rep("RNA-seq", 2), rep("scRNA-seq", 12)))
+
+## add assay labels to results table
+objdat <- left_join(objdat, studies, by = "casestudy")
+objdat <- dplyr::filter(objdat, !is.na(assay))
+
+## create table of rejected proportion for top methods per case study
+objdat_props <- dplyr::group_by(objdat, casestudy)
+objdat_props <- dplyr::filter(objdat_props, nrejects == max(nrejects, na.rm = TRUE))
+objdat_props <- dplyr::ungroup(objdat_props)
+objdat_props <- dplyr::mutate(objdat_props, lab = paste0(100*round(propPossible, 4), "%"))
+objdat_props$plotrow <- "real"
+
+## create table of total tests for each case study
+objdat_tab <- dplyr::select(objdat_props, casestudy, nrejects, propPossible, assay)
+objdat_tab <- dplyr::distinct(objdat_tab)
+objdat_tab <- dplyr::mutate(objdat_tab, `Total Tests` = scales::comma(round(nrejects / propPossible)))
+objdat_tab <- dplyr::select(objdat_tab, casestudy, assay, `Total Tests`) 
+objdat_tab <- tidyr::gather(objdat_tab, method, cnt, -casestudy, -assay)
+
+## hack to get total counts and heatmap in same plot
+objdat_joint <- bind_rows(list(real = objdat, fake = objdat_tab), .id = "plotrow")
+
+# filter for methods in methodset
+objdat_joint <- filter(objdat_joint, method %in% c(methodset, "Total Tests"))
+
+## specify method order
+method_order <- c("bh", "ihw", "qvalue", "bl", "adapt-glm", "lfdr", "fdrreg-t", "ashq")
+objdat_joint <- dplyr::mutate(objdat_joint, method = factor(method, levels = c(method_order, "Total Tests")))
+
+# add notation for missing because "shouldn't"
+objdat_joint <- objdat_joint %>% 
+  mutate(should = ifelse(method == "lfdr" & is.na(nrejects), "*", ""))
+
+p4b <- ggplot(objdat_joint, aes(x = casestudy, y = method, fill = propMaxRej)) +
+    scale_x_discrete("Case Study", expand = c(0, 0)) +
+    scale_y_discrete("Method", expand = c(0, 0)) +
+    geom_tile() +
+    geom_text(aes(label = lab), data = filter(objdat_props, method %in% methodset),
+              size = 1.8, color = "white") +
+    geom_text(aes(label = should), size = 5, color = "black") +
+    geom_tile(data = filter(objdat_joint, plotrow == "fake"), fill = "white") + 
+    geom_text(aes(label = cnt), data = dplyr::filter(objdat_joint, !is.na(cnt)), size = 2) + 
+    scale_fill_distiller("% Rejected\n(relative to max)",
+                         palette = "Blues", direction = 1, limits = c(0, 1),
+                         labels = scales::percent) +
+    facet_grid(plotrow ~ assay, scales = "free", space = "free") +
+    theme_bw() + 
+    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
+          strip.text.y = element_blank(),
+          legend.position = c(0,-0.32), legend.justification = c(0, 1),
+          legend.direction = "horizontal",
+          legend.text = element_text(size = 7),
+          legend.title = element_text(size = 9),
+          legend.margin = margin(6, 10, 6, 10))
+Titlep4b <- ggdraw() +
+    draw_label(expression(paste(bold("Number of rejections in case studies"))))
+p4b <- plot_grid(Titlep4b, p4b, nrow = 2, rel_heights = c(0.075,1), 
+                 labels = c("B", ""))
+
+## pull Figure 4 together
+Fig4 <- plot_grid(p4a, p4b, ncol = 1, rel_heights = c(1,1))
+Fig4
+ggsave(file.path(outdir, "Fig4.pdf"), width = 11, height = 10.5)
+```
 

### 2018-10-18 datasets/microbiome/microbiome-baxter.Rmd
Add results plots for mean abunandance covariate
Add uninformative covariate comparison to CRC genus level

--- 2018-05-13 datasets/microbiome/microbiome-baxter.Rmd
@@
 gns_bench_file_abun <- file.path(sbdir, "baxter-genus-mean-abun-benchmark.rds")
+gns_bench_file_uninf <- file.path(sbdir, "baxter-genus-mean-uninf-benchmark.rds")
 ```
@@
 
+Add random (uninformative) covariate to test at the genus level.
+
+```{r}
+set.seed(72664)
+res$rand_covar <- rnorm(nrow(res))
+```
+
 ## Covariate Diagnostics
@@
 rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
+```
+
+### Random
+
+```{r, baxter-uninf-hist, fig.width=10, fig.height=3.2}
+strat_hist(res, pvalue="pval", covariate="rand_covar", maxy=8, binwidth=0.05)
+```
+
+```{r, baxter-uninf-scatter, fig.width=4.5, fig.height=3.5}
+rank_scatter(res, pvalue="pval", covariate="rand_covar")
 ```
@@
 
+## Multiple-Testing Correction - Random
+
+Let's use `rand_covar` as our `ind_covariate`.
+
+```{r, baxter-genus-labelcovariate-uninf}
+res <- dplyr::mutate(res, ind_covariate = rand_covar)
+```
+
+First, we'll create an object of `BenchDesign` class to hold the data and 
+add the benchmark methods to the `BenchDesign` object. We remove ASH from
+the comparison.
+
+Then, we'll construct the `SummarizedBenchmark` object, which will run
+the functions specified in each method (these are actually sourced in from the
+helper scripts).
+
+```{r, baxter-genus-sb-uninf, results="hide", message=FALSE, eval=TRUE}
+if (!file.exists(gns_bench_file_uninf)) {
+    bd <- initializeBenchDesign()
+    bd <- dropBMethod(bd, "ashq")
+    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
+    metadata(sb)$data_download_link <-
+                   "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
+    saveRDS(sb, file = gns_bench_file_uninf)
+} else {
+    sb <- readRDS(gns_bench_file_uninf)
+}
+```
+## Benchmark Metrics - Random
+
+Next, we'll add the default performance metric for q-value assays. First, we have
+to rename the assay to 'qvalue'.
+
+```{r, baxter-genus-uninf-metrics, eval=TRUE}
+assayNames(sb) <- "qvalue"
+sb <- addDefaultMetrics(sb)
+```
+
+Now, we'll plot the results.
+
+```{r, baxter-genus-uninf-plot, width=15, height=15, eval=TRUE}
+rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
+rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
+```
+
+Note: Benjamini-Hochberg (bh) overlaps exactly with the IHW results.
+
+```{r, baxter-genus-uninf-overlap}
+plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
+```
+```{r}
+covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
+```
+
 # OTU/Genus comparison
@@
                   gns_bench_file, gns_bench_file_log_ubiquity,
-                  gns_bench_file_abun), 
+                  gns_bench_file_abun, gns_bench_file_uninf), 
                 colLabels = c("OTU", "OTU - log", "Genus", "Genus - log",
-                              "Genus - abun"), 
+                              "Genus - abun", "Genus - uninf"), 
                 alpha = 0.10, xlab = "Comparison")

### 2018-10-19 manuscript_figures/manuscript_figures.Rmd
Add microbiome CRC dataset to summary figures

@@
 # Also remove'log' ubiquity since redundant
-null <- which(grepl(c("goodrich|papa"), objects) & 
+null <- which(grepl(c("goodrich|papa|baxter"), objects) & 
               grepl(c("otu"), objects) | 
-              grepl("baxter", objects) |
               grepl("log", objects) | 
@@
 
-## Case study rankings
-
-This is a figure similar to Figure 4, except that we 
-include all case studies in a single plot.
-
-```{r}
-## read list of case study results
-objects <- list.files(path, recursive=TRUE, pattern="rds", full.names=TRUE)
-
-## exclude yeast and polyester simulations, as well as uninformative covariate
-## analyses
-objects <- objects[!grepl("yeast", objects) & !grepl("polyester", objects) &
-                   !grepl("uninf", objects)]
-
-## create shorter names for case studies
-object_labs <- gsub("\\.rds", "", basename(objects))
-object_labs <- gsub("-benchmark", "", object_labs)
-
-## read in case study results (proportion of rejects vs. case study max)
-objdat <- tidy_df(objects=objects, colLabels=object_labs, 
-                  fill="propMaxRejections", annotate=NULL,
-                  alpha =0.10)
-objdat <- as_tibble(objdat)
-objdat <- tidyr::complete(objdat, method, casestudy)
-
-## exclude 'enigma-so4-otus' results with no significant calls
-objdat <- dplyr::filter(objdat, casestudy != "enigma-so4-otus")
-
-## create table of case study-to-assay mapping
-studies <- tibble(
-    casestudy = c("cbp-csaw", "h3k4me3-csaw", "human", "mouse", "bmi-maf", "bmi-samplesize",
-                  "enigma-al-otus", "enigma-ph-otus", "enigma-so4-otus", "goodrich-genus",
-                  "papa-genus", "schubert-otus", "brain", "mir200c",
-                  "human-mast-det", "human-mast-mean", "human-scdd-det",
-                  "human-scdd-mean", "human-wilcox-det", "human-wilcox-mean",
-                  "mouse-mast-det", "mouse-mast-mean", "mouse-scdd-det",
-                  "mouse-scdd-mean", "mouse-wilcox-det", "mouse-wilcox-mean"),
-    assay = c(rep("ChIP-seq", 2), rep("GSEA", 2), rep("GWAS", 2), rep("Microbiome", 6),
-              rep("RNA-seq", 2), rep("scRNA-seq", 12)))
-
-## add assay labels to results table
-objdat <- left_join(objdat, studies, by = "casestudy")
-objdat <- dplyr::filter(objdat, !is.na(assay))
-
-## create table of rejected proportion for top methods per case study
-objdat_props <- dplyr::group_by(objdat, casestudy)
-objdat_props <- dplyr::filter(objdat_props, nrejects == max(nrejects, na.rm = TRUE))
-objdat_props <- dplyr::ungroup(objdat_props)
-objdat_props <- dplyr::mutate(objdat_props, lab = paste0(100*round(propPossible, 4), "%"))
-objdat_props$plotrow <- "real"
-
-## create table of total tests for each case study
-objdat_tab <- dplyr::select(objdat_props, casestudy, nrejects, propPossible, assay)
-objdat_tab <- dplyr::distinct(objdat_tab)
-objdat_tab <- dplyr::mutate(objdat_tab, `Total Tests` = scales::comma(round(nrejects / propPossible)))
-objdat_tab <- dplyr::select(objdat_tab, casestudy, assay, `Total Tests`) 
-objdat_tab <- tidyr::gather(objdat_tab, method, cnt, -casestudy, -assay)
-
-## hack to get total counts and heatmap in same plot
-objdat_joint <- bind_rows(list(real = objdat, fake = objdat_tab), .id = "plotrow")
-
-## specify method order
-method_order <- c("bonf", "bh", "ihw", "qvalue", "bl", "adapt-glm", "lfdr",
-                  "fdrreg-e", "fdrreg-t", "ashq")
-objdat_joint <- dplyr::mutate(objdat_joint, method = factor(method, levels = c(method_order, "Total Tests")))
-```
-
-```{r, fig.width = 12, fig.height = 5}
-SFigAlt <- ggplot(objdat_joint, aes(x = casestudy, y = method, fill = propMaxRej)) +
-    scale_x_discrete("Case Study", expand = c(0, 0)) +
-    scale_y_discrete("Method", expand = c(0, 0)) +
-    geom_tile() +
-    geom_text(aes(label = lab), data = objdat_props, size = 1.8, color = "white") +
-    geom_tile(data = filter(objdat_joint, plotrow == "fake"), fill = "white") + 
-    geom_text(aes(label = cnt), data = dplyr::filter(objdat_joint, !is.na(cnt)), size = 2) + 
-    scale_fill_distiller("% Rejected\n(relative to max)",
-                         palette = "Blues", direction = 1, limits = c(0, 1),
-                         labels = scales::percent) +
-    facet_grid(plotrow ~ assay, scales = "free", space = "free") +
-    theme_bw() + 
-    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
-          strip.text.y = element_blank(),
-          legend.position = c(.99, .85), legend.justification = c(1, 1),
-          legend.background = element_rect(fill = scales::alpha("white", 3/4), color = 'black'),
-          legend.direction = "horizontal",
-          legend.text = element_text(size = 7),
-          legend.title = element_text(size = 9),
-          legend.margin = margin(6, 10, 6, 10))
-SFigAlt
-
-ggsave(file.path(outdir, "FigureSAlt.pdf"), SFigAlt, width = 12, height = 5)
-```
-
 ## Informativeness of covariates in case studies

### 2018-10-19 manuscript_figures/manuscript_figures_main.Rmd
Add microbiome CRC dataset to summary figures

@@
     casestudy = c("cbp-csaw", "h3k4me3-csaw", "human", "mouse", "bmi-maf", "bmi-samplesize",
-                  "enigma-al-otus", "enigma-ph-otus", "enigma-so4-otus", "goodrich-genus",
-                  "papa-genus", "schubert-otus", "brain", "mir200c",
+                  "enigma-al-otus", "enigma-ph-otus", "enigma-so4-otus", "baxter-genus", 
+                  "goodrich-genus", "papa-genus", "schubert-otus", "brain", "mir200c",
                   "human-mast-det", "human-mast-mean", "human-scdd-det",
@@
                   "mouse-scdd-mean", "mouse-wilcox-det", "mouse-wilcox-mean"),
-    assay = c(rep("ChIP-seq", 2), rep("GSEA", 2), rep("GWAS", 2), rep("Microbiome", 6),
+    assay = c(rep("ChIP-seq", 2), rep("GSEA", 2), rep("GWAS", 2), rep("Microbiome", 7),
               rep("RNA-seq", 2), rep("scRNA-seq", 12)))
@@
 Fig4
-ggsave(file.path(outdir, "Fig4.pdf"), width = 11, height = 10.5)
+ggsave(file.path(outdir, "Fig4.pdf"), width = 11.2, height = 10.5)
 ```

### 2018-10-19 manuscript_figures/manuscript_figures_supp.Rmd
Reorder supp table 4 rows; adjust legends of of casestudy info. covariate impact heatmap

--- 2018-10-19 manuscript_figures/manuscript_figures.Rmd
@@
 tabS3 <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
-                        tableOnly = TRUE)
+                        tableOnly = TRUE, excludeMethods = NULL)
 tabS3 <- tabS3 %>% 
@@
   select(casestudy, method, mean_prop) %>%
-  spread(casestudy, mean_prop)
+  spread(casestudy, mean_prop) %>% 
+  mutate(method =  factor(method, levels = rev(Fig4All.rowOrder))) %>%
+  arrange(method)
 
@@
                   if (max(x) > 200){
-                    c(-20, -5, 0, 5, 25, 100, 300)
+                    c(-100, -25, -5, 0, 5, 25, 100, 300)
                   }else if (max(x) > 100){
-                    c(-20, -5, 0, 5, 25, 100, 200)
+                    c(-100, -25, -5, 0, 5, 25, 100, 200)
                   }else{
-                    c(-20, -5, 0, 5, 25, 100)
+                    c(-100, -25, -5, 0, 5, 25, 100)
                   }
@@
                              colors = c("darkblue","white","darkred"), 
-                             limits = c(-20, topRange),
-                             values = scales::rescale(c(-transform(20), 0, transform(topRange))), 
+                             limits = c(min(c(-20, botRange)), topRange),
+                             values = scales::rescale(c(-transform(max(20,abs(botRange))), 
+                                                        0, transform(topRange))), 
                              trans = log_neg_percent()) +
@@
         theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
-              legend.position = c(0.99, 0.95), legend.justification = c(1, 1),
+              legend.position = c(0.995, 0.95), legend.justification = c(1, 1),
               legend.background = element_rect(fill = scales::alpha("white", 3/4), color = 'black'),
@@
                              colors = c("darkblue","white","darkred"), 
-                             limits = c(-20, topRange),
+                             limits = c(-25, topRange),
                              values = scales::rescale(c(-transform(20), 0, transform(topRange))), 

### 2018-10-19 datasets/RNA-Seq/yeast-simulation.Rmd
Add plot of reject proportion and FWER to yeast null sims

@@
                 merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5, met="rejectprop",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res5, met="FWER",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
 plotsim_average(res5, met="TNR",filter_set = excludeSet,
@@
 plotsim_average(res10, met="rejections",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10, met="rejectprop",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) 
+plotsim_average(res10, met="FWER",filter_set = excludeSet,
                 merge_ihw = TRUE, errorBars=TRUE) 

### 2018-10-21 manuscript_figures/manuscript_ratings.Rmd
Add first pass at computing final summary rankings

---
title: "Final Ratings for Genome Biology"
author: "Rafalab"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This Rmd is used to determine the ratings for each method shown in Figure 4.

# Set up workspace

```{r Figure2-load-data}
## Load packages and source benchmark FDR
library(tidyverse)
library(cowplot)
library(ggthemes)
library(grid)
##library(SummarizedBenchmark)
devtools::load_all("../../SummarizedBenchmark")

## load helper functions
for (f in list.files("../datasets/R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

# Assumes sb objects for the case studies and in silico experiments are in 
# the following location, which contains subfolders
# for each casestudy (if this isn't true, then parsing the case study and 
# dataset names later on will be incorrect)
path <- "../results"

# Also assumes that simulation results summary file is in the following location
# The `result-metrics.rds` was generated in the 
# `datasets/simulations/simulations-summary.Rmd` file.
resmet_file <- file.path("..", "datasets", "simulations", "results-summary", "result-metrics.rds")

# set up results directory
outdir <- "./figures"
dir.create(outdir, showWarnings = FALSE)

# set alpha cutoff for plots that are fixed at a certain value of alpha
alpha.thresh <- 0.05

# methods to include in all figures (exclude bonferroni and fdrreg-e)
methodset <- c("bh", "ihw", "qvalue", "bl", "lfdr",
               "fdrreg-t", "ashq", "adapt-glm")
```

We use the standardize candy color scheme and line types for the plots. We'll
add the "lfdr*" method, which indicates lfdr was applied with fewer than 200 
tests per bin (out of 20 bins).

```{r}
col <- as.character(candycols$col)
names(col) <- as.character(candycols$Method)
lty <- as.character(candycols$lty)
names(lty) <- as.character(candycols$Method)
```

To generate the figures in this document, the simulation results must first be aggregated by running the code at `datasets/simulations/simulations-summary.Rmd`.

```{r load-sim-metrics}
resmet <- readRDS(resmet_file)
```

We similarly load all other results from the following files.

```{r}
objects <- list.files(path, recursive = TRUE, pattern = "rds", full.names = TRUE)
```

# Summarize Data Sets

## Yeast

```{r}
objects_yeast <- grep("yeast", objects, value = TRUE)

## read just yeast samples
if (file.exists("yeastres.rds")) {
   yeastres <- readRDS("yeastres.rds")
} else {
    yeastres <- lapply(objects_yeast, readRDS)
    yeastres <- lapply(yeastres, plotsim_standardize, alpha = alpha.thresh)

    names(yeastres) <- basename(objects_yeast)
    yeastres <- bind_rows(yeastres, .id = "series")
    saveRDS(yeastres, "yeastres.rds")
}

## clean up the results
yeast <- dplyr::select(yeastres, series, rep, blabel, param.alpha, key, alpha, value)
yeast <- dplyr::filter(yeast, as.numeric(param.alpha) == alpha | is.na(param.alpha))
yeast <- dplyr::mutate(yeast, blabel = ifelse(grepl("ihw", blabel), "ihw", blabel))
yeast <- dplyr::mutate(yeast, blabel = ifelse(blabel == "bl-df03", "bl", blabel))
yeast <- dplyr::filter(yeast, blabel %in% methodset)
yeast <- dplyr::select(yeast, -param.alpha, -alpha)
yeast <- dplyr::rename(yeast, Method = blabel)

## use 0s for "rates" that are reported as NAs
yeast <- dplyr::mutate(yeast, value = ifelse(is.na(value) & grepl("R$", key), 0, value))

## compute mean and standard errors for each method in each sim for each metric
yeast <- dplyr::group_by(yeast, series, Method, key)
yeast <- dplyr::summarize(yeast, mean = mean(value, na.rm = TRUE),
                          se = sd(value, na.rm = TRUE) / sqrt(sum(!is.na(value))))
yeast <- dplyr::ungroup(yeast)

## clean up the simulation names
yeast <- dplyr::mutate(yeast, series = gsub("\\.rds", "", series))
yeast <- tidyr::separate(yeast, series, c("setting_base", "setting_res", "setting_de", "setting_cov"),
                         fill = "right")
yeast <- dplyr::select(yeast, -setting_res)
yeast <- dplyr::mutate(yeast, setting_mode = ifelse(grepl("II", setting_base), "bimodal", "unimodal"))
yeast <- dplyr::mutate(yeast, setting_pi0 = ifelse(grepl("H", setting_base), "7.5%", "30%"))

yeast <- dplyr::mutate(yeast, setting_cov = ifelse(is.na(setting_cov), "strongInfoCov", setting_cov))
yeast <- dplyr::mutate(yeast, setting_cov = ifelse(grepl("W", setting_base), "weakInfoCov", setting_cov))
yeast <- dplyr::select(yeast, -setting_base) 
```

We also rank each method in each setting according to FDR and TPR.

```{r}
yeast <- dplyr::group_by(yeast, setting_de, setting_cov, setting_mode, setting_pi0, key)
yeast <- dplyr::mutate(yeast, rank = rank(-mean))
yeast <- dplyr::ungroup(yeast)
```

## Case Studies

```{r}
objects_dat <- objects[!grepl("polyester", objects) & !grepl("yeast", objects)]

## remove microbiome objects from 'null' comparisons
## removing baxter, goodrich, papa OTU-level, and So4-enigma (since they mostly found
## nothing for all methods and the genus-level analysis is also present)
## Also remove'log' ubiquity since redundant
null <- grepl(c("goodrich|papa"), objects_dat) &
    grepl(c("otu"), objects_dat) |
    grepl("baxter", objects_dat) |
    grepl("log", objects_dat) |
    grepl("so4", objects_dat)
objects_dat <- objects_dat[!null]

## remove chip-seq promoter analysis (since it is the only one that uses
## ash and we also include the csaw analysis on the same dataset).
promot <- grepl("promot", objects_dat)
objects_dat <- objects_dat[!promot]

# remove random covariate analyses
uninf <- grepl("uninf", objects_dat)
objects_dat <- objects_dat[!uninf]

## read just real data samples
if (file.exists("datres.rds")) {
    datres <- readRDS("datres.rds")
} else {
    datres <- lapply(objects_dat, readRDS)
    saveRDS(datres, "datres.rds")
}

## gather all results
datres <- lapply(datres, assay, "bench")
datres <- lapply(datres, as_tibble)
names(datres) <- gsub("^\\.\\./results/(.*?)-benchmark(.*)\\.rds", "\\1\\2", objects_dat)
datres <- bind_rows(datres, .id = "cs")
datres <- tidyr::separate(datres, cs, c("casestudy", "dataset"), sep = "/")
datres <- tidyr::gather(datres, Method, adjp, -casestudy, -dataset)

## clean up names
datres <- dplyr::group_by(datres, casestudy, dataset, Method)
datres <- dplyr::summarize(datres, ntotal = n(), nNA = sum(is.na(adjp)),
                           nrej = sum(adjp < alpha.thresh, na.rm = TRUE))
datres <- dplyr::ungroup(datres)

## NAs only appear in cases where the method wasn't applied so we can go ahead
## and replace the nrej values with NA for these methods
datres <- dplyr::mutate(datres, nrej = ifelse(nNA/ntotal > .99, NA, nrej))

## clean up method names
iihw <- paste0("ihw-a", sprintf("%02d", 100*alpha.thresh))
datres <- dplyr::mutate(datres, Method = ifelse(grepl(iihw, Method), "ihw", Method))
datres <- dplyr::mutate(datres, Method = ifelse(Method == "bl-df03", "bl", Method))
datres <- dplyr::filter(datres, Method %in% methodset)

## add method rank in each category
datres <- dplyr::group_by(datres, casestudy, dataset)
datres <- dplyr::mutate(datres, rank = rank(-nrej)) 
datres <- dplyr::mutate(datres, rank = ifelse(is.na(nrej), NA, rank))
datres <- dplyr::ungroup(datres)

##saveRDS(datres, "datres-table.rds")
```

## Simulations

We can use the results stored in `resmet` for simulation settings.
For easier plotting, we replace create a shorter distribution string column.

```{r}
resmet <- dplyr::mutate(resmet, sdist = factor(dist, sort(levels(dist)), c("chisq", "normal", "t11", "t5")))
```

We only look at results at the desired alpha cutoff, and resmet `fdrreg-e` from this comparison.

```{r}
resmet <- dplyr::filter(resmet, Method != "fdrreg-e", alpha == alpha.thresh)
```

We also rank each method in each setting according to FDR and TPR.

```{r}
resmet <- dplyr::group_by(resmet, setting, dist, inform, pi0, signal, ntests, sdist, key)
resmet <- dplyr::mutate(resmet, rank = rank(-mean.info))
resmet <- dplyr::ungroup(resmet)
```

For consistency across plotting of settings, we relabel some settings.

```{r}
resmet <- dplyr::mutate(resmet, setting = ifelse(setting == "varyingpi0-t", "varyingpi0", setting))
```

# Exploratory Figures

After summarizing all yeast experiments, simulations, and case studies, we take a look at the
aggregated results.

## Yeast

```{r, fig.width = 8, fig.height = 12}
dplyr::filter(yeast, key == "FDR") %>%
    ggplot(aes(x = interaction(setting_mode, setting_cov, setting_pi0, setting_de),
               y = mean, ymin = mean - se, ymax = mean+se, color = Method)) +
    geom_point(size = 1/2, position = position_dodge(width = 1/2)) +
    geom_errorbar(position = position_dodge(width = 1/2)) +
    facet_grid(Method ~ setting_de, scales = "free_x", space = "free_x") +
    expand_limits(y = 0) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    ylab("FDR") +
    xlab(NULL) +
    theme_bw() +
    geom_hline(yintercept = 0.05, linetype = 2, color = "blue", alpha = 1/2) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1, size = 6)) +
    ggtitle(paste0("Yeast Experiments: FDR control across settings, alpha = ", alpha.thresh))

dplyr::filter(yeast, key == "TPR", !grepl("null", setting_de)) %>%
    ggplot(aes(x = interaction(setting_mode, setting_pi0),
               y = mean, ymin = mean - se, ymax = mean+se, color = Method)) +
    geom_point(size = 1/2, position = position_dodge(width = 1/2)) +
    geom_errorbar(position = position_dodge(width = 1/2)) +
    facet_grid(setting_de ~ setting_cov) +
    expand_limits(y = 0) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    ylab("TPR") +
    xlab(NULL) +
    theme_bw() +
    geom_hline(yintercept = 0.05, linetype = 2, color = "blue", alpha = 1/2) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1, size = 6)) +
    ggtitle(paste0("Yeast Experiments: TPR across settings, alpha = ", alpha.thresh))
```

Now we also look at the ranks of methods.

```{r, fig.width = 8, fig.height = 12}
dplyr::filter(yeast, key == "FDR") %>%
    ggplot(aes(x = interaction(setting_mode, setting_cov, setting_pi0, setting_de),
               y = rank, color = Method)) +
    geom_point(size = 4, position = position_dodge(width = 1/2)) +
    facet_grid(Method ~ setting_de, scales = "free_x", space = "free_x") +
    expand_limits(y = 0) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    ylab("FDR rank") +
    xlab(NULL) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1, size = 6)) +
    ggtitle(paste0("Yeast Experiments: method rank for TPR, alpha = ", alpha.thresh))

dplyr::filter(yeast, key == "TPR", !grepl("null", setting_de)) %>%
    ggplot(aes(x = interaction(setting_mode, setting_cov, setting_pi0, setting_de),
               y = rank, color = Method)) +
    geom_point(size = 4, position = position_dodge(width = 1/2)) +
    facet_grid(Method ~ setting_de, scales = "free_x", space = "free_x") +
    expand_limits(y = 0) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    ylab("TPR rank") +
    xlab(NULL) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1, size = 6)) +
    ggtitle(paste0("Yeast Experiments: method rank for TPR, alpha = ", alpha.thresh))
```

## Case Studies

```{r, fig.width = 10, fig.height = 5}
ggplot(datres, aes(x = dataset, y = nrej / ntotal, color = Method)) +
    geom_point(size = 3/4, position = position_dodge(width = 1/2)) +
    facet_grid(. ~ casestudy, scales = "free_x", space = "free_x") +
    expand_limits(y = c(0, 1)) +
    scale_color_manual(values = col) +
    scale_y_continuous("Percent Rejections", labels = scales::percent) +
    xlab(NULL) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
    ggtitle(paste0("Case Study: method rejection proportion, alpha = ", alpha.thresh))
```

```{r, fig.width = 10, fig.height = 14}
ggplot(datres, aes(x = dataset, y = rank, color = Method, group = Method)) +
    geom_line(size = 1/2, alpha = 1/4, position = position_dodge(width = 1/4)) +
    geom_point(size = 1, position = position_dodge(width = 1/4)) +
    facet_grid(Method ~ casestudy, scales = "free_x", space = "free_x") +
    scale_color_manual(values = col) +
    scale_y_continuous("Rank", breaks = 1:8) +
    xlab(NULL) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
    ggtitle(paste0("Case Study: method rank for total rejections, alpha = ", alpha.thresh))
```

```{r, fig.width = 10, fig.height = 6}
ggplot(datres, aes(x = dataset, y = rank, color = Method, group = Method)) +
    geom_point(size = 4, position = position_dodge(width = 1/2)) +
    facet_grid(. ~ casestudy, scales = "free_x", space = "free_x") +
    scale_color_manual(values = col) +
    scale_y_continuous("Rank", breaks = 1:8) +
    xlab(NULL) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
    ggtitle(paste0("Case Study: method rank for total rejections, alpha = ", alpha.thresh))
```

## Simulations

```{r}
dplyr::filter(resmet, key == "FDR") %>%
    tidyr::unite(subsetting, sdist, inform, pi0, signal, ntests) %>%
    ggplot(aes(x = subsetting,
               y = mean.info, ymin = mean.info - se.info, ymax = mean.info + se.info, color = Method)) +
    geom_point(size = 1/2, position = position_dodge(width = 1/2)) +
    geom_errorbar(position = position_dodge(width = 1/2)) +
    facet_grid(Method ~ setting, space = "free_x", scales = "free_x") +
    expand_limits(y = 0) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    ylab("FDR") +
    xlab(NULL) +
    theme_bw() +
    geom_hline(yintercept = 0.05, linetype = 2, color = "blue", alpha = 1/2) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1, size = 6)) +
    ggtitle(paste0("Simulations: FDR control across settings, alpha = ", alpha.thresh))

dplyr::filter(yeast, key == "TPR", !grepl("null", setting_de)) %>%
    ggplot(aes(x = interaction(setting_mode, setting_pi0),
               y = mean, ymin = mean - se, ymax = mean+se, color = Method)) +
    geom_point(size = 1/2, position = position_dodge(width = 1/2)) +
    geom_errorbar(position = position_dodge(width = 1/2)) +
    facet_grid(setting_de ~ setting_cov) +
    expand_limits(y = 0) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    ylab("TPR") +
    xlab(NULL) +
    theme_bw() +
    geom_hline(yintercept = 0.05, linetype = 2, color = "blue", alpha = 1/2) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1, size = 6)) +
    ggtitle(paste0("Yeast Experiments: TPR across settings, alpha = ", alpha.thresh))
```

# Ratings

We determine cutoffs using the summarized results.

## FDR Control

FDR control ratings are determined using only the yeast experiments and simulation
results where ground truth is known.

```{r}
## FDR control across simulation settings
fdrcontrol_sims <- resmet %>%
    dplyr::filter(key == "FDR", !grepl("null", setting), sdist != "chisq") %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(nExceeds = sum((mean.info - se.info) > alpha.thresh),
                     pExceeds = round(nExceeds / n() * 100, 2)) %>%
    dplyr::ungroup()
dplyr::arrange(fdrcontrol_sims, pExceeds)

## FDR control across yeast settings
fdrcontrol_yeast <- yeast %>%
    dplyr::filter(key == "FDR", !grepl("null", setting_de)) %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(nExceeds = sum((mean - se) > alpha.thresh),
                     pExceeds = round(nExceeds / n() * 100, 2)) %>%
    dplyr::ungroup()
dplyr::arrange(fdrcontrol_yeast, pExceeds)

## combine
fdrcontrol <- dplyr::left_join(fdrcontrol_sims, fdrcontrol_yeast, by = "Method",
                               suffix = c(".sims", ".yeast"))
dplyr::arrange(fdrcontrol, pExceeds.sims + pExceeds.yeast)
```

## Power

Similarly, power ratings are only determined using only the yeast experiments and
simulation results where ground truth is known.

```{r}
## power across simulation settings
power_sims <- resmet %>%
    dplyr::filter(key == "TPR", alpha == alpha.thresh,
                  !grepl("null", setting), sdist != "chisq") %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(meanRank = mean(rank),
                     medianRank = median(rank)) %>%
    dplyr::ungroup()
dplyr::arrange(power_sims, meanRank)

## power across yeast settings
power_yeast <- yeast %>%
    dplyr::filter(key == "TPR", !grepl("null", setting_de)) %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(meanRank = mean(rank),
                     medianRank = median(rank)) %>%
    dplyr::ungroup()
dplyr::arrange(power_yeast, meanRank)

## combine
power <- dplyr::left_join(power_sims, power_yeast, by = "Method",
                          suffix = c(".sims", ".yeast"))
dplyr::arrange(power, meanRank.sims + meanRank.yeast)
```

## Robustness

Next, we define metrics for the robustness metric, which quantifies the amount of consistency,
or robustness, of each method across simulations, yeast experiments, and case studies. For
this metric, we only look at modern methods, i.e. exclude Bonferroni correction, BH, and
Storey's q-value. We do this, as we quantify "robustness" as the relative gain of each method
over the average of BH and q-value in FDR and TPR (and proportion of rejections for case studies).

```{r}
## robustness as variance of FDR/TPR as % increase vs BH/qvalue
robustness_sims <- resmet %>%
    dplyr::filter(key == "TPR" | key == "FDR", alpha == alpha.thresh,
                  !grepl("null", setting), sdist != "chisq",
                  Method != "bonf") %>%
    dplyr::select(Method, key, mean.info, setting, sdist, inform, pi0, signal, ntests) %>%
    tidyr::spread(Method, mean.info) %>%
    tidyr::gather(Method, mean.info, -key, -setting, -sdist, -inform,
                  -pi0, -signal, -ntests, -qvalue, -bh) %>%
    dplyr::group_by(Method, key) %>%
    dplyr::summarize(meanGain = mean(mean.info / (bh + qvalue) * 2),
                     sdGainBH = sd(mean.info / bh),
                     sdGainQ = sd(mean.info / qvalue),
                     sdGain = sd(mean.info / (bh + qvalue) * 2),
                     cvGain = sdGain / meanGain) %>%
    dplyr::ungroup()
dplyr::arrange(robustness_sims, key, -cvGain)

## robustness as variance of FDR/TPR as % increase vs BH/qvalue
robustness_yeast <- yeast %>%
    dplyr::filter(key == "TPR" | key == "FDR", !grepl("null", setting_de)) %>%
    dplyr::select(Method, mean, key, setting_de, setting_cov, setting_mode, setting_pi0) %>%
    tidyr::spread(Method, mean) %>%
    tidyr::gather(Method, mean, -key, -setting_de, -setting_cov, -setting_pi0,
                  -setting_mode, -qvalue, -bh) %>%
    dplyr::group_by(Method, key) %>%
    dplyr::summarize(meanGain = mean(mean / (bh + qvalue) * 2),
                     sdGainBH = sd(mean / bh),
                     sdGainQ = sd(mean / qvalue),
                     sdGain = sd(mean / (bh + qvalue) * 2),
                     cvGain = sdGain / meanGain) %>%
    dplyr::ungroup()
dplyr::arrange(robustness_yeast, key, -cvGain)

robustness_data <- datres %>%
    dplyr::mutate(prej = nrej / ntotal) %>%
    dplyr::select(-rank, -nrej, -ntotal, -nNA) %>%
    tidyr::spread(Method, prej) %>%
    tidyr::gather(Method, prej, -casestudy, -dataset, -qvalue, -bh) %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(meanGain = mean(prej / (bh + qvalue) * 2, na.rm = TRUE),
                     sdGainBH = sd(prej / bh, na.rm = TRUE),
                     sdGainQ = sd(prej / qvalue, na.rm = TRUE),
                     sdGain = sd(prej / (bh + qvalue) * 2, na.rm = TRUE),
                     cvGain = sdGain / meanGain) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(key = "PropRej")
dplyr::arrange(robustness_data, -cvGain)

## combine
robustness <- dplyr::full_join(dplyr::select(robustness_sims, Method, key, cvGain),
                               dplyr::select(robustness_yeast, Method, key, cvGain),
                               by = c("key", "Method"),
                               suffix = c(".sims", ""))
robustness <- dplyr::full_join(robustness,
                               dplyr::select(robustness_data, Method, key, cvGain),
                               by = c("key", "Method"),
                               suffix = c(".yeast", ".data"))
dplyr::arrange(robustness, key, cvGain.sims + cvGain.yeast, cvGain.data)
```

## Applicability

Finally, for "applicability" we take a look at the number of times each method was not applied
in the case studies.

```{r}
applicability_data <- datres %>%
    dplyr::group_by(casestudy, Method) %>%
    dplyr::summarize(meanNA = mean(is.na(nrej))) %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(meanNA = mean(meanNA)) %>%
    dplyr::ungroup()
dplyr::arrange(applicability_data, meanNA)
```

### 2018-10-22 manuscript_figures/manuscript_ratings.Rmd
Add figures with tables for manuscript-ratings

@@
 
-```{r Figure2-load-data}
+```{r load-workspace, message = FALSE}
 ## Load packages and source benchmark FDR
@@
 # `datasets/simulations/simulations-summary.Rmd` file.
-resmet_file <- file.path("..", "datasets", "simulations", "results-summary", "result-metrics.rds")
+simres_file <- file.path("..", "datasets", "simulations", "results-summary", "result-metrics.rds")
 
@@
 ```{r load-sim-metrics}
-resmet <- readRDS(resmet_file)
+simres <- readRDS(simres_file)
 ```
@@
     datres <- lapply(objects_dat, readRDS)
-    saveRDS(datres, "datres.rds")
-}
 
@@
 
-## clean up names
+    ## summarize results
     datres <- dplyr::group_by(datres, casestudy, dataset, Method)
@@
     datres <- dplyr::ungroup(datres)
+
+    saveRDS(datres, "datres.rds")
+}
 
@@
 datres <- dplyr::ungroup(datres)
-
-##saveRDS(datres, "datres-table.rds")
 ```
@@
 
-We can use the results stored in `resmet` for simulation settings.
+We can use the results stored in `simres` for simulation settings.
 For easier plotting, we replace create a shorter distribution string column.
@@
 ```{r}
-resmet <- dplyr::mutate(resmet, sdist = factor(dist, sort(levels(dist)), c("chisq", "normal", "t11", "t5")))
+simres <- dplyr::mutate(simres, sdist = factor(dist, sort(levels(dist)), c("chisq", "normal", "t11", "t5")))
 ```
 
-We only look at results at the desired alpha cutoff, and resmet `fdrreg-e` from this comparison.
+We only look at results at the desired alpha cutoff, and remove `fdrreg-e` from this comparison.
 
 ```{r}
-resmet <- dplyr::filter(resmet, Method != "fdrreg-e", alpha == alpha.thresh)
+simres <- dplyr::filter(simres, Method != "fdrreg-e", Method != "bonf", alpha == alpha.thresh)
 ```
@@
 ```{r}
-resmet <- dplyr::group_by(resmet, setting, dist, inform, pi0, signal, ntests, sdist, key)
-resmet <- dplyr::mutate(resmet, rank = rank(-mean.info))
-resmet <- dplyr::ungroup(resmet)
+simres <- dplyr::group_by(simres, setting, dist, inform, pi0, signal, ntests, sdist, key)
+simres <- dplyr::mutate(simres, rank = rank(-mean.info))
+simres <- dplyr::ungroup(simres)
 ```
@@
 ```{r}
-resmet <- dplyr::mutate(resmet, setting = ifelse(setting == "varyingpi0-t", "varyingpi0", setting))
+simres <- dplyr::mutate(simres, setting = ifelse(setting == "varyingpi0-t", "varyingpi0", setting))
 ```
@@
 ```{r}
-dplyr::filter(resmet, key == "FDR") %>%
+dplyr::filter(simres, key == "FDR") %>%
     tidyr::unite(subsetting, sdist, inform, pi0, signal, ntests) %>%
@@
 ## FDR control across simulation settings
-fdrcontrol_sims <- resmet %>%
+fdrcontrol_sims <- simres %>%
     dplyr::filter(key == "FDR", !grepl("null", setting), sdist != "chisq") %>%
@@
     dplyr::summarize(nExceeds = sum((mean.info - se.info) > alpha.thresh),
-                     pExceeds = round(nExceeds / n() * 100, 2)) %>%
+                     pExceeds = round(nExceeds / n(), 4)) %>%
     dplyr::ungroup()
-dplyr::arrange(fdrcontrol_sims, pExceeds)
 
@@
     dplyr::summarize(nExceeds = sum((mean - se) > alpha.thresh),
-                     pExceeds = round(nExceeds / n() * 100, 2)) %>%
+                     pExceeds = round(nExceeds / n(), 4)) %>%
     dplyr::ungroup()
-dplyr::arrange(fdrcontrol_yeast, pExceeds)
 
@@
 dplyr::arrange(fdrcontrol, pExceeds.sims + pExceeds.yeast)
+```
+
+We plot the results as well.
+
+```{r, fig.width = 6, fig.height = 4}
+bind_rows(yeast = fdrcontrol_yeast, sims = fdrcontrol_sims, .id = "analysis") %>%
+    ggplot(aes(x = analysis, y = pExceeds, color = Method, group = Method)) +
+    geom_line(aes(linetype = Method), alpha = 1/2, position = position_dodge(width = 1/4)) +
+    geom_point(position = position_dodge(width = 1/4)) + 
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_discrete(NULL) + 
+    scale_y_continuous("Percent Settings Exceeding nominal FDR", labels = scales::percent) + 
+    expand_limits(y = 0) +
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    ggtitle("FDR control across simulations and yeast experiments") 
 ```
@@
 ## power across simulation settings
-power_sims <- resmet %>%
+power_sims <- simres %>%
     dplyr::filter(key == "TPR", alpha == alpha.thresh,
@@
     dplyr::ungroup()
-dplyr::arrange(power_sims, meanRank)
 
@@
     dplyr::ungroup()
-dplyr::arrange(power_yeast, meanRank)
 
@@
 dplyr::arrange(power, meanRank.sims + meanRank.yeast)
+```
+
+We plot the results as well.
+
+```{r, fig.width = 6, fig.height = 4}
+bind_rows(yeast = power_yeast, sims = power_sims, .id = "analysis") %>%
+    dplyr::mutate(Method = reorder(Method, meanRank)) %>%
+    ggplot(aes(x = Method, y = meanRank, color = Method, group = analysis)) +
+    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
+    geom_point() + 
+    scale_color_manual(values = col, guide = FALSE) +
+    scale_x_discrete(NULL) + 
+    scale_y_continuous("Mean TPR rank", breaks = 0:10) + 
+    expand_limits(y = 1) +
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    ggtitle("TPR rank across simulations and yeast experiments") 
 ```
@@
 ```{r}
-## robustness as variance of FDR/TPR as % increase vs BH/qvalue
-robustness_sims <- resmet %>%
+## robustness across simulation settings
+robustness_sims <- simres %>%
     dplyr::filter(key == "TPR" | key == "FDR", alpha == alpha.thresh,
@@
     dplyr::ungroup()
-dplyr::arrange(robustness_sims, key, -cvGain)
 
-## robustness as variance of FDR/TPR as % increase vs BH/qvalue
+## robustness across yeast experiments
 robustness_yeast <- yeast %>%
@@
     dplyr::ungroup()
-dplyr::arrange(robustness_yeast, key, -cvGain)
 
+## robustness across case studies
 robustness_data <- datres %>%
@@
     dplyr::mutate(key = "PropRej")
-dplyr::arrange(robustness_data, -cvGain)
 
@@
 dplyr::arrange(robustness, key, cvGain.sims + cvGain.yeast, cvGain.data)
+```
+
+We plot the results as well.
+
+```{r, fig.width = 6, fig.height = 8}
+bind_rows(yeast = robustness_yeast, sims = robustness_sims,
+          casestudies = robustness_data, .id = "analysis") %>%
+    dplyr::mutate(Method = reorder(Method, sdGain)) %>%
+    ggplot(aes(x = Method, y = sdGain, color = Method, group = analysis)) +
+    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
+    geom_point() + 
+    scale_color_manual(values = col, guide = FALSE) +
+    scale_x_discrete(NULL) + 
+    scale_y_continuous("SD of percent relative to BH/qvalue") + 
+    expand_limits(y = 1) +
+    facet_grid(key ~ ., scales = "free") + 
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    ggtitle("Variation across simulations, yeast experiments,\ncase studies")
+
+bind_rows(yeast = robustness_yeast, sims = robustness_sims,
+          casestudies = robustness_data, .id = "analysis") %>%
+    dplyr::mutate(Method = reorder(Method, cvGain)) %>%
+    ggplot(aes(x = Method, y = cvGain, color = Method, group = analysis)) +
+    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
+    geom_point() + 
+    scale_color_manual(values = col, guide = FALSE) +
+    scale_x_discrete(NULL) + 
+    scale_y_continuous("CV of percent relative to BH/qvalue") + 
+    expand_limits(y = 0) +
+    facet_grid(key ~ ., scales = "free") + 
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    ggtitle("Variation across simulations, yeast experiments,\ncase studies")
+
+bind_rows(yeast = robustness_yeast, sims = robustness_sims,
+          casestudies = robustness_data, .id = "analysis") %>%
+    dplyr::mutate(Method = reorder(Method, cvGain)) %>%
+    ggplot(aes(x = Method, y = cvGain, color = Method, group = key)) +
+    geom_line(aes(linetype = key), color = 'black', alpha = 1/4) +
+    geom_point() + 
+    scale_color_manual(values = col, guide = FALSE) +
+    scale_x_discrete(NULL) + 
+    scale_y_continuous("CV of percent relative to BH/qvalue") +
+    scale_linetype("Metric") + 
+    expand_limits(y = 0) +
+    facet_grid(analysis ~ ., scales = "free_y") + 
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    ggtitle("Variation across simulations, yeast experiments,\ncase studies")
+
+bind_rows(yeast = robustness_yeast, sims = robustness_sims,
+          casestudies = robustness_data, .id = "analysis") %>%
+    dplyr::mutate(Method = reorder(Method, sdGain)) %>%
+    ggplot(aes(x = Method, y = sdGain, color = Method, group = key)) +
+    geom_line(aes(linetype = key), color = 'black', alpha = 1/4) +
+    geom_point() + 
+    scale_color_manual(values = col, guide = FALSE) +
+    scale_x_discrete(NULL) + 
+    scale_y_continuous("SD of percent relative to BH/qvalue") +
+    scale_linetype("Metric") + 
+    expand_limits(y = 0) +
+    facet_grid(analysis ~ ., scales = "free_y") + 
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    ggtitle("Variation across simulations, yeast experiments,\ncase studies") 
 ```

### 2018-10-25 datasets/microbiome/microbiome-enigma.Rmd

@@
 
-Here we analyze a subset of the ENIGMA data ([Smith, M. and Rocha, A. et al 2015](http://dx.doi.org/10.1128/mBio.00326-15)).
+Here we analyze a subset of the ENIGMA data ([Smith, M. and Rocha, A. et al 2015](https://doi.org/10.1128/mBio.00326-15)).
 The dataset includes water samples from 67 wells with well-characterized biochemical metadata (i.e. pH, heavy metal concentrations, etc).
-We acquired the data from a collaborator and have put it in the `benchmark-fdr/datasets/microbiome/DATA/enigma` folder.
+We'll download the processed OTU table and associated metadata from Zenodo and put it in the `data/enigma_results` folder.
 
@@
 if (!file.exists(file.path(enigma_dir, "enigma.metadata.txt"))) {
-    download.file("https://www.dropbox.com/s/0r1zjp26u10l0p4/enigma.metadata.txt?dl=1",
+    download.file("https://zenodo.org/record/1455793/files/enigma.metadata.txt",
                   destfile = file.path(enigma_dir, "enigma.metadata.txt"))
@@
 if (!file.exists(file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"))) {
-    download.file("https://www.dropbox.com/s/u9qh86cd7c0vbn9/enigma.otu_table_resampled_updated_r.txt.gz?dl=1",
+    download.file("https://zenodo.org/record/1455793/files/enigma.otu_table_resampled_updated_r.txt.gz",
                   destfile = file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"))
@@
 ## load OTU table and metadata
-print(file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"))
 otu <- read.table(file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"), 

### 2018-10-25 datasets/microbiome/microbiome-goodrich.Rmd

@@
 Here we download and analyze the Goodrich obesity (OB) dataset
-[(Goodrich et al., 2014)](http://dx.doi.org/10.1016/j.cell.2014.09.053).
+[(Goodrich et al., 2014)](https://doi.org/10.1016/j.cell.2014.09.053).
 The dataset includes fecal samples from 428 individuals with lean body mass index (BMI < 25),
 and 185 obese individuals (BMI > 30). We'll download the processed OTU tables from
-Zenodo and unzip them in the `benchmark-fdr/datasets/microbiome/data` folder.
+Zenodo and unzip them in the `data/ob_goodrich_results` folder.
 
-We expect few or no truly differentially abundant OTUs ([Finucane et al., 2014](http://dx.doi.org/10.1371/journal.pone.0084689) and [Walters et al., 2014](https://dx.doi.org/10.1016%2Fj.febslet.2014.09.039)), so this study provides a real-life "null" example.
+We expect few or no truly differentially abundant OTUs ([Finucane et al., 2014](https://doi.org/10.1371/journal.pone.0084689) and [Walters et al., 2014](https://doi.org/10.1016%2Fj.febslet.2014.09.039)), so this study provides a real-life "null" example.
 

### 2018-10-25 datasets/microbiome/microbiome-schubert.Rmd

@@
 
-Here we download and analyze the Schubert diarrhea dataset [(Schubert et al., 2014)](http://dx.doi.org/10.1128/mBio.01021-14).
+Here we download and analyze the Schubert diarrhea dataset [(Schubert et al., 2014)](https://doi.org/10.1128/mBio.01021-14).
 The dataset includes 154 healthy nondiarrheal, 93 Clostridium difficile infection (CDI) associated diarrheal,
 and 89 non-CDI associated diarrheal stool samples. We'll download the processed OTU tables from
-Zenodo and unzip them in the `benchmark-fdr/datasets/microbiome/data` folder.
+Zenodo and unzip them in the `data/cdi_schubert_results` folder.
 
@@
 cdi_dir <- file.path(datdir, "cdi_schubert_results")
+dir.create(cdi_dir, showWarnings = FALSE)
 

### 2018-10-25 manuscript_figures/manuscript_ratings.Rmd

@@
 bind_rows(yeast = fdrcontrol_yeast, sims = fdrcontrol_sims, .id = "analysis") %>%
-    ggplot(aes(x = analysis, y = pExceeds, color = Method, group = Method)) +
-    geom_line(aes(linetype = Method), alpha = 1/2, position = position_dodge(width = 1/4)) +
-    geom_point(position = position_dodge(width = 1/4)) + 
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
+    dplyr::mutate(Method = reorder(Method, pExceeds)) %>%
+    ggplot(aes(x = Method, y = pExceeds, color = Method, group = analysis)) +
+    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
+    geom_point() + 
+    scale_color_manual(values = col, guide = FALSE) +
     scale_x_discrete(NULL) + 
-    scale_y_continuous("Percent Settings Exceeding nominal FDR", labels = scales::percent) + 
-    expand_limits(y = 0) +
+    scale_y_continuous("Percent of settings exceeding nominal FDR", labels = scales::percent) + 
     theme_classic() +
@@
     scale_y_continuous("Mean TPR rank", breaks = 0:10) + 
+    expand_limits(y = 1) +
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    ggtitle("TPR rank across simulations and yeast experiments")
+
+bind_rows(yeast = power_yeast, sims = power_sims, .id = "analysis") %>%
+    dplyr::mutate(Method = reorder(Method, medianRank)) %>%
+    ggplot(aes(x = Method, y = medianRank, color = Method, group = analysis)) +
+    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
+    geom_point() + 
+    scale_color_manual(values = col, guide = FALSE) +
+    scale_x_discrete(NULL) + 
+    scale_y_continuous("Median TPR rank", breaks = 0:10) + 
     expand_limits(y = 1) +
@@
 
-```{r, fig.width = 6, fig.height = 8}
+```{r, fig.width = 6, fig.height = 6}
 bind_rows(yeast = robustness_yeast, sims = robustness_sims,
@@
 ```
+
+We also include data sets where 0 rejections were made by `adapt-glm` as cases where
+the method could not be applied.
+
+```{r}
+applicability_data <- datres %>%
+    dplyr::mutate(nrej = ifelse(nrej == 0 & Method == "adapt-glm", NA, nrej)) %>%
+    dplyr::group_by(casestudy, Method) %>%
+    dplyr::summarize(meanNA = mean(is.na(nrej))) %>%
+    dplyr::group_by(Method) %>%
+    dplyr::summarize(meanNA = mean(meanNA)) %>%
+    dplyr::ungroup()
+dplyr::arrange(applicability_data, meanNA)
+```

### 2018-10-25 manuscript_figures/manuscript_figures_main.Rmd
Make alpha threshold consistent for Fcase study results figure

@@
                   fill="propMaxRejections", annotate=NULL,
-                  alpha =0.10)
+                  alpha = alpha.thresh)
 objdat <- as_tibble(objdat)

### 2018-10-25 manuscript_figures/manuscript_figures_supp.Rmd
Make alpha threshold consistent in Supplementary figures

@@
 
-# Figure 1
-
-This is a schema of Inputs, Assumptions and Outputs for the various FDR methods. Claire is leading this charge.
+This file contains Supplementary Figures to complement the main results figures 
+in `manuscript_figures_main.Rmd`.
 
@@
 # set alpha cutoff for plots that are fixed at a certain value of alpha
-alpha <- 0.05
+alpha.thresh <- 0.05
 ```
@@
 # Figure 2D: Some methods are tailored to detecting smaller effect sizes
-p4 <- covariateLinePlot(de5, alpha=alpha, covname="log2FC", trans="log1p",
+p4 <- covariateLinePlot(de5, alpha=alpha.thresh, covname="log2FC", trans="log1p",
                         nbins=25)
@@
 pdf(file.path(outdir, "Figure2E.pdf"), width = 8, height = 4, onefile=FALSE)
-aggupset(de5, alpha=alpha, supplementary = FALSE, 
+aggupset(de5, alpha=alpha.thresh, supplementary = FALSE, 
                        return_list = FALSE, nintersects = 12)
@@
 This is a heatmap of methods by case studies with color representing the ranking
-by of number of rejections @ FDR cutoff `r alpha`.
+by of number of rejections @ FDR cutoff `r alpha.thresh`.
 
@@
                      "fdrreg-e", "fdrreg-t", "ashq")
-Fig4All <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
+Fig4All <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha.thresh,
                         excludeMethods = NULL, rowOrder = Fig4All.rowOrder) +
@@
 library(xtable)
-tabS3 <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
+tabS3 <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha.thresh,
                         tableOnly = TRUE, excludeMethods = NULL)
@@
 # generate detailed case study table for creating sim vs cs ranks boxplots
-casestudy.tab <- plotMethodRanks(objects, colLabels = dataset, alpha = alpha,
+casestudy.tab <- plotMethodRanks(objects, colLabels = dataset, alpha = alpha.thresh,
                                  tableOnly = TRUE)
@@
                 colLabels = c("sim", "sim"), 
-                alpha = alpha, xlab = "Comparison",
+                alpha = alpha.thresh, xlab = "Comparison",
                 fill = "FDR",
@@
                          low = "#2171B5", 
-                         midpoint = alpha) +
+                         midpoint = alpha.thresh) +
           labs(fill = "Mean FDR") +
@@
                 colLabels = c("sim", "sim"), 
-                alpha = alpha, xlab = "Comparison",
+                alpha = alpha.thresh, xlab = "Comparison",
                 fill = "TPR",
@@
 # generate detailed yeast sim table for creating sim vs cs ranks boxplots
-yeastsim.tab <- plotMethodRanks(objs, colLabels = labs, alpha = alpha,
+yeastsim.tab <- plotMethodRanks(objs, colLabels = labs, alpha = alpha.thresh,
                                  tableOnly = TRUE)
@@
 res <- addDefaultMetrics(res)
-pA <- covariateLinePlot(res, alpha=alpha, covname="ind_covariate", trans="log1p",
+pA <- covariateLinePlot(res, alpha=alpha.thresh, covname="ind_covariate", trans="log1p",
                         nbins=25) +
@@
 res <- addDefaultMetrics(res)
-pB <- covariateLinePlot(res, alpha=alpha, covname="detection", trans="log1p",
+pB <- covariateLinePlot(res, alpha=alpha.thresh, covname="detection", trans="log1p",
                         nbins=25) +
@@
 res <- addDefaultMetrics(res)
-pD <- covariateLinePlot(res, alpha=alpha, covname="ind_covar_AF", trans="log1p",
+pD <- covariateLinePlot(res, alpha=alpha.thresh, covname="ind_covar_AF", trans="log1p",
                         nbins=25) +
@@
 tab <- tidy_df(objects=objects, colLabels=filename, 
-              fill="meanRank", annotate=NULL, alpha =0.10) %>%
+              fill="meanRank", annotate=NULL, alpha = alpha.thresh) %>%
   rename(casestudy = "filename") %>%
@@
 size <-  ifelse(grepl("10", resfile), "10", "5")
-aseq <- alpha
 
@@
                   fill =  "FDR", annotate = NULL, 
-                  alpha = aseq) %>%
+                  alpha = alpha.thresh) %>%
     mutate(fdr = nrejects) %>%
@@
                     fill =  "TPR", annotate = NULL, 
-                    alpha = aseq) %>%
+                    alpha = alpha.thresh) %>%
     mutate(tpr = nrejects) %>%
@@
                     fill =  "TPR", annotate = NULL, 
-                    alpha = aseq) %>%
+                    alpha = alpha.thresh) %>%
       mutate(tpr = nrejects) %>%

### 2018-10-26 manuscript_figures/manuscript_ratings.Rmd
Create supplementary plot for ratings

@@
 library(grid)
-##library(SummarizedBenchmark)
-devtools::load_all("../../SummarizedBenchmark")
+library(SummarizedBenchmark)
 
@@
 
-## Robustness
+## Consistency
 
-Next, we define metrics for the robustness metric, which quantifies the amount of consistency,
-or robustness, of each method across simulations, yeast experiments, and case studies. For
-this metric, we only look at modern methods, i.e. exclude Bonferroni correction, BH, and
-Storey's q-value. We do this, as we quantify "robustness" as the relative gain of each method
+Next, we define metrics for the consistency metric, which quantifies the amount of consistency,
+of each method across simulations and case studies. While the yeast experiments
+could also be included as part of the consistency calculation, it is excluded, primarily because
+all settings are variations on differential expression in RNA-seq, and therefore contains far
+less variable settings as compared to the simulations and case studies.
+
+For this metric, we only look at modern methods, i.e. exclude Bonferroni correction, BH, and
+Storey's q-value. We do this, as we quantify "consistency" as the relative gain of each method
 over the average of BH and q-value in FDR and TPR (and proportion of rejections for case studies).
 
+We exclude both `ashq` and `fdrreg-t` from the analysis of the case studies here because they
+were applied in substantially fewer case studies (only 4/26). The gain over BH/q-value can vary
+substantially based on the informativeness of the covariate and therefore, if a method was applied
+to far fewer case studies, it may have an arbitrarily smaller or larger amount of variability across
+case studies simply due to the composition of the subset relative to the composition of the
+entire collection of case studies.
+
 ```{r}
-## robustness across simulation settings
-robustness_sims <- simres %>%
+## consistency across simulation settings
+consistency_sims <- simres %>%
     dplyr::filter(key == "TPR" | key == "FDR", alpha == alpha.thresh,
@@
     dplyr::summarize(meanGain = mean(mean.info / (bh + qvalue) * 2),
+                     meanRGain = mean(log2(mean.info / (bh + qvalue) * 2)),
                      sdGainBH = sd(mean.info / bh),
@@
                      sdGain = sd(mean.info / (bh + qvalue) * 2),
+                     sdRGain = sd(log2(.001 + mean.info / (bh + qvalue) * 2)),
                      cvGain = sdGain / meanGain) %>%
@@
 
-## robustness across yeast experiments
-robustness_yeast <- yeast %>%
-    dplyr::filter(key == "TPR" | key == "FDR", !grepl("null", setting_de)) %>%
-    dplyr::select(Method, mean, key, setting_de, setting_cov, setting_mode, setting_pi0) %>%
-    tidyr::spread(Method, mean) %>%
-    tidyr::gather(Method, mean, -key, -setting_de, -setting_cov, -setting_pi0,
-                  -setting_mode, -qvalue, -bh) %>%
-    dplyr::group_by(Method, key) %>%
-    dplyr::summarize(meanGain = mean(mean / (bh + qvalue) * 2),
-                     sdGainBH = sd(mean / bh),
-                     sdGainQ = sd(mean / qvalue),
-                     sdGain = sd(mean / (bh + qvalue) * 2),
-                     cvGain = sdGain / meanGain) %>%
-    dplyr::ungroup()
-
-## robustness across case studies
-robustness_data <- datres %>%
-    dplyr::mutate(prej = nrej / ntotal) %>%
+## consistency across case studies
+consistency_data <- datres %>%
+    dplyr::filter(Method != "ashq", Method != "fdrreg-t", !is.na(nrej)) %>%
+    dplyr::mutate(prej = (nrej+1) / ntotal) %>%
     dplyr::select(-rank, -nrej, -ntotal, -nNA) %>%
@@
     dplyr::summarize(meanGain = mean(prej / (bh + qvalue) * 2, na.rm = TRUE),
+                     meanRGain = mean(log2(prej / (bh + qvalue) * 2), na.rm = TRUE),
                      sdGainBH = sd(prej / bh, na.rm = TRUE),
@@
                      sdGain = sd(prej / (bh + qvalue) * 2, na.rm = TRUE),
+                     sdRGain = sd(log2(prej / (bh + qvalue) * 2), na.rm = TRUE),
                      cvGain = sdGain / meanGain) %>%
     dplyr::ungroup() %>%
-    dplyr::mutate(key = "PropRej")
-
-## combine
-robustness <- dplyr::full_join(dplyr::select(robustness_sims, Method, key, cvGain),
-                               dplyr::select(robustness_yeast, Method, key, cvGain),
-                               by = c("key", "Method"),
-                               suffix = c(".sims", ""))
-robustness <- dplyr::full_join(robustness,
-                               dplyr::select(robustness_data, Method, key, cvGain),
-                               by = c("key", "Method"),
-                               suffix = c(".yeast", ".data"))
-dplyr::arrange(robustness, key, cvGain.sims + cvGain.yeast, cvGain.data)
+    dplyr::mutate(key = "%reject")
 ```
@@
 ```{r, fig.width = 6, fig.height = 6}
-bind_rows(yeast = robustness_yeast, sims = robustness_sims,
-          casestudies = robustness_data, .id = "analysis") %>%
-    dplyr::mutate(Method = reorder(Method, sdGain)) %>%
-    ggplot(aes(x = Method, y = sdGain, color = Method, group = analysis)) +
-    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
+bind_rows(Sims = consistency_sims, CS = consistency_data, .id = "analysis") %>%
+    dplyr::mutate(Method = reorder(Method, sdGain),
+                  key = factor(paste0(key, " (", analysis, ")")),
+                  key = factor(key, levels = levels(key)[c(2, 3, 1)])) %>%
+    ggplot(aes(x = Method, y = sdGain, color = Method, group = key)) +
+    geom_line(aes(linetype = key), color = 'black', alpha = 1/2) +
     geom_point() + 
@@
     scale_x_discrete(NULL) + 
-    scale_y_continuous("SD of percent relative to BH/qvalue") + 
+    scale_y_continuous("SD of % relative to BH/qvalue") + 
     expand_limits(y = 1) +
-    facet_grid(key ~ ., scales = "free") + 
     theme_classic() +
@@
           plot.title = element_text(face = "bold", size = 14)) +
-    ggtitle("Variation across simulations, yeast experiments,\ncase studies")
+    ggtitle("Variation across simulations and case studies")
 
-bind_rows(yeast = robustness_yeast, sims = robustness_sims,
-          casestudies = robustness_data, .id = "analysis") %>%
-    dplyr::mutate(Method = reorder(Method, cvGain)) %>%
-    ggplot(aes(x = Method, y = cvGain, color = Method, group = analysis)) +
-    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
+bind_rows(Sims = consistency_sims, CS = consistency_data, .id = "analysis") %>%
+    dplyr::mutate(Method = reorder(Method, cvGain),
+                  key = factor(paste0(key, " (", analysis, ")")),
+                  Metric = factor(key, levels = levels(key)[c(2, 3, 1)])) %>%
+    ggplot(aes(x = Method, y = cvGain, color = Method, group = Metric)) +
+    geom_line(aes(linetype = Metric), color = 'black', alpha = 1/2) +
     geom_point() + 
@@
     scale_x_discrete(NULL) + 
-    scale_y_continuous("CV of percent relative to BH/qvalue") + 
-    expand_limits(y = 0) +
-    facet_grid(key ~ ., scales = "free") + 
+    scale_y_continuous("CV of % relative to BH/qvalue") + 
+    expand_limits(y = 1) +
     theme_classic() +
@@
           plot.title = element_text(face = "bold", size = 14)) +
-    ggtitle("Variation across simulations, yeast experiments,\ncase studies")
+    ggtitle("Variation across simulations and case studies")
 
-bind_rows(yeast = robustness_yeast, sims = robustness_sims,
-          casestudies = robustness_data, .id = "analysis") %>%
-    dplyr::mutate(Method = reorder(Method, cvGain)) %>%
-    ggplot(aes(x = Method, y = cvGain, color = Method, group = key)) +
-    geom_line(aes(linetype = key), color = 'black', alpha = 1/4) +
+bind_rows(Sims = consistency_sims, CS = consistency_data, .id = "analysis") %>%
+    dplyr::mutate(Method = reorder(Method, sdRGain),
+                  key = factor(paste0(key, " (", analysis, ")")),
+                  key = factor(key, levels = levels(key)[c(2, 3, 1)])) %>% 
+    ggplot(aes(x = Method, y = sdRGain, color = Method, group = key)) +
+    geom_line(aes(linetype = key), color = 'black', alpha = 1/2) +
     geom_point() + 
@@
     scale_x_discrete(NULL) + 
-    scale_y_continuous("CV of percent relative to BH/qvalue") +
-    scale_linetype("Metric") + 
-    expand_limits(y = 0) +
-    facet_grid(analysis ~ ., scales = "free_y") + 
+    scale_y_continuous("SD of log-ratio to BH/qvalue") + 
+    expand_limits(y = 1) +
     theme_classic() +
@@
           plot.title = element_text(face = "bold", size = 14)) +
-    ggtitle("Variation across simulations, yeast experiments,\ncase studies")
+    ggtitle("Variation across simulations and case studies")
+```
 
-bind_rows(yeast = robustness_yeast, sims = robustness_sims,
-          casestudies = robustness_data, .id = "analysis") %>%
-    dplyr::mutate(Method = reorder(Method, sdGain)) %>%
-    ggplot(aes(x = Method, y = sdGain, color = Method, group = key)) +
-    geom_line(aes(linetype = key), color = 'black', alpha = 1/4) +
+Additionally, we look at the proportion of times a method performed better than BH and Storey's
+q-value.
+
+```{r}
+## consistency (definition 2) across simulation settings
+consistency2_sims <- simres %>%
+    dplyr::filter(key == "TPR", alpha == alpha.thresh,
+                  !grepl("null", setting), sdist != "chisq",
+                  Method != "bonf") %>%
+    dplyr::select(Method, key, mean.info, setting, sdist, inform, pi0, signal, ntests) %>%
+    tidyr::spread(Method, mean.info) %>%
+    tidyr::gather(Method, mean.info, -key, -setting, -sdist, -inform,
+                  -pi0, -signal, -ntests, -qvalue, -bh) %>%
+    dplyr::group_by(Method, key) %>%
+    dplyr::summarize(propBoth = mean(mean.info < pmax(bh, qvalue)),
+                     propBoth95 = mean(mean.info < 0.95 * pmax(bh, qvalue)),
+                     propEither = mean(mean.info < pmin(bh, qvalue)),
+                     propEither95 = mean(mean.info < 0.95 * pmin(bh, qvalue))) %>%
+    dplyr::ungroup()
+
+## consistency (definition 2) across case studies
+consistency2_data <- datres %>%
+    dplyr::filter(Method != "ashq", Method != "fdrreg-t", !is.na(nrej)) %>%
+    dplyr::select(-rank, -ntotal, -nNA) %>%
+    tidyr::spread(Method, nrej) %>%
+    tidyr::gather(Method, nrej, -casestudy, -dataset, -qvalue, -bh) %>%
+    dplyr::group_by(Method) %>%
+    dplyr::summarize(propBoth = mean(nrej < pmax(bh, qvalue), na.rm = TRUE),
+                     propBoth95 = mean(nrej < 0.95 * pmax(bh, qvalue), na.rm = TRUE),
+                     propEither = mean(nrej < pmin(bh, qvalue), na.rm = TRUE),
+                     propEither95 = mean(nrej < 0.95 * pmin(bh, qvalue), na.rm = TRUE)) %>%
+    dplyr::ungroup() %>%
+    dplyr::mutate(key = "%reject")
+```
+
+We also plot these results for the lowerbound case of out-performing either classical approach (BH or Storey's q-value).
+
+```{r, fig.width = 6, fig.height = 6}
+bind_rows(Sims = consistency2_sims, CS = consistency2_data, .id = "analysis") %>%
+    dplyr::mutate(Method = reorder(Method, propEither95)) %>%
+    ggplot(aes(x = Method, y = propEither95, color = Method, group = analysis)) +
+    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/2) +
     geom_point() + 
@@
     scale_x_discrete(NULL) + 
-    scale_y_continuous("SD of percent relative to BH/qvalue") +
-    scale_linetype("Metric") + 
-    expand_limits(y = 0) +
-    facet_grid(analysis ~ ., scales = "free_y") + 
+    scale_y_continuous("Percent of settings w/ TPR or #rejections\n <95% of BH/q-value", labels = scales::percent) + 
     theme_classic() +
@@
           plot.title = element_text(face = "bold", size = 14)) +
-    ggtitle("Variation across simulations, yeast experiments,\ncase studies") 
+    ggtitle("Consistency of gains over BH or q-value\nacross simulations and case studies")
 ```
@@
 ```
+
+# Figures
+
+Figures for the manuscript are also generated.
+
+```{r, fig.width = 10, fig.height = 8}
+p1a <- bind_rows(Yeast = fdrcontrol_yeast, Sims = fdrcontrol_sims, .id = "Analysis") %>%
+    dplyr::mutate(Method = reorder(Method, pExceeds),
+                  Analysis = factor(Analysis, levels = c("Yeast", "Sims"))) %>%
+    ggplot(aes(x = Method, y = pExceeds, color = Method, group = Analysis)) +
+    geom_line(aes(linetype = Analysis), color = 'black', alpha = 1/2) +
+    geom_point(size = 3) + 
+    scale_color_manual(values = col, guide = FALSE) +
+    scale_x_discrete(NULL) + 
+    scale_y_continuous("% settings exceeding nominal FDR", labels = scales::percent) + 
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    ggtitle("FDR Control")
+
+p1b <- bind_rows(Yeast = power_yeast, Sims = power_sims, .id = "Analysis") %>%
+    dplyr::mutate(Method = reorder(Method, meanRank),
+                  Analysis = factor(Analysis, levels = c("Yeast", "Sims"))) %>%
+    ggplot(aes(x = Method, y = meanRank, color = Method, group = Analysis)) +
+    geom_line(aes(linetype = Analysis), color = 'black', alpha = 1/2) +
+    geom_point(size = 3) + 
+    scale_color_manual(values = col, guide = FALSE) +
+    scale_x_discrete(NULL) + 
+    scale_y_continuous("Mean TPR rank", breaks = 0:10) + 
+    expand_limits(y = 1) +
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    ggtitle("Power")
+
+p1c <- bind_rows(Sims = consistency2_sims, CS = consistency2_data, .id = "Analysis") %>%
+    dplyr::mutate(Method = reorder(Method, propEither95),
+                  Analysis = factor(paste0(Analysis, " (", key, ")")),
+                  Aanlysis = factor(Analysis, levels = levels(Analysis)[c(2, 3, 1)])) %>%
+    ggplot(aes(x = Method, y = propEither95, color = Method, group = Analysis)) +
+    geom_line(aes(linetype = Analysis), color = 'black', alpha = 1/2) +
+    geom_point(size = 3) + 
+    scale_color_manual(values = col, guide = FALSE) +
+    scale_x_discrete(NULL) + 
+    scale_y_continuous("% settings w/ metric\n < 95% of BH and q-value", labels = scales::percent) + 
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    ggtitle("Consistency (I)")
+
+p1d <- bind_rows(Sims = consistency_sims, CS = consistency_data, .id = "Analysis") %>%
+    dplyr::mutate(Method = reorder(Method, sdRGain),
+                  Analysis = factor(paste0(Analysis, " (", key, ")"))) %>%
+    ggplot(aes(x = Method, y = sdRGain, color = Method, group = Analysis)) +
+    geom_line(aes(linetype = Analysis), color = 'black', alpha = 1/2) +
+    geom_point(size = 3) + 
+    scale_color_manual(values = col, guide = FALSE) +
+    scale_x_discrete(NULL) + 
+    scale_y_continuous("SD of log-ratio relative to\nmean of BH and q-value") + 
+    expand_limits(y = 1) +
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    ggtitle("Consistency (II)")
+
+p1title <- ggdraw() +
+    draw_label("Summary Recommendation Metrics",
+               fontface = 'bold')
+
+p1ab <- plot_grid(p1a + guides(linetype = FALSE),
+                  p1b + guides(linetype = FALSE),
+                  labels = LETTERS[1:2], nrow = 1)
+p1ab <- plot_grid(p1ab, get_legend(p1a), nrow = 1, rel_widths = c(1, .2))
+
+p1cd <- plot_grid(p1c + guides(linetype = FALSE),
+                  p1d + guides(linetype = FALSE),
+                  labels = LETTERS[3:4], nrow = 1)
+p1cd <- plot_grid(p1cd, get_legend(p1d), nrow = 1, rel_widths = c(1, .2))
+
+FigS <- plot_grid(p1title, p1ab, p1cd, ncol = 1, rel_heights = c(.15, 1, 1))
+FigS
+
+ggsave(file.path(outdir, "SFigureMetrics.pdf"), width=10, height=8)
+```

### 2018-10-26 manuscript_figures/manuscript_ratings.Rmd
Add plot with common x-axis; make fixes to plots

@@
 bind_rows(Sims = consistency2_sims, CS = consistency2_data, .id = "analysis") %>%
-    dplyr::mutate(Method = reorder(Method, propEither95)) %>%
-    ggplot(aes(x = Method, y = propEither95, color = Method, group = analysis)) +
+    dplyr::mutate(Method = reorder(Method, propEither)) %>%
+    ggplot(aes(x = Method, y = propEither, color = Method, group = analysis)) +
     geom_line(aes(linetype = analysis), color = 'black', alpha = 1/2) +
@@
     scale_x_discrete(NULL) + 
-    scale_y_continuous("Percent of settings w/ TPR or #rejections\n <95% of BH/q-value", labels = scales::percent) + 
+    scale_y_continuous("Percent of settings w/ TPR or #rejections\n of BH/q-value", labels = scales::percent) + 
     theme_classic() +
@@
 p1c <- bind_rows(Sims = consistency2_sims, CS = consistency2_data, .id = "Analysis") %>%
-    dplyr::mutate(Method = reorder(Method, propEither95),
-                  Analysis = factor(paste0(Analysis, " (", key, ")")),
-                  Aanlysis = factor(Analysis, levels = levels(Analysis)[c(2, 3, 1)])) %>%
-    ggplot(aes(x = Method, y = propEither95, color = Method, group = Analysis)) +
+    dplyr::mutate(Method = reorder(Method, propEither),
+                  Analysis = factor(paste0(Analysis, " (", key, ")"))) %>%
+    ggplot(aes(x = Method, y = propEither, color = Method, group = Analysis)) +
     geom_line(aes(linetype = Analysis), color = 'black', alpha = 1/2) +
@@
     scale_x_discrete(NULL) + 
-    scale_y_continuous("% settings w/ metric\n < 95% of BH and q-value", labels = scales::percent) + 
+    scale_y_continuous("% settings w/ metric < classic methods", labels = scales::percent) + 
     theme_classic() +
@@
           plot.title = element_text(face = "bold", size = 14)) +
-    ggtitle("Consistency (I)")
+    ggtitle("Consistency of Gain")
 
@@
     dplyr::mutate(Method = reorder(Method, sdRGain),
-                  Analysis = factor(paste0(Analysis, " (", key, ")"))) %>%
+                  Analysis = factor(paste0(Analysis, " (", key, ")")),
+                  Analysis = factor(Analysis, levels = levels(Analysis)[c(1, 3, 2)])) %>%
     ggplot(aes(x = Method, y = sdRGain, color = Method, group = Analysis)) +
@@
     scale_x_discrete(NULL) + 
-    scale_y_continuous("SD of log-ratio relative to\nmean of BH and q-value") + 
+    scale_y_continuous("SD of log-ratio relative to classic methods") + 
     expand_limits(y = 1) +
@@
           plot.title = element_text(face = "bold", size = 14)) +
-    ggtitle("Consistency (II)")
+    ggtitle("Relative Consistency")
 
@@
 
-FigS <- plot_grid(p1title, p1ab, p1cd, ncol = 1, rel_heights = c(.15, 1, 1))
+FigS <- plot_grid(p1title, p1ab, p1cd, ncol = 1, rel_heights = c(.2, 1, 1))
 FigS
@@
 ```
+
+Additionally, we consider creating the above plot with a common x-axis.
+
+```{r, fig.width = 10, fig.height = 8}
+consistency_order <- c("bh", "qvalue", "bl", "ihw", "lfdr", "fdrreg-t", "ashq", "adapt-glm")
+
+p2a <- bind_rows(Yeast = fdrcontrol_yeast, Sims = fdrcontrol_sims, .id = "Analysis") %>%
+    dplyr::mutate(Method = reorder(Method, pExceeds),
+                  Analysis = factor(Analysis, levels = c("Yeast", "Sims"))) %>%
+    ggplot(aes(x = Method, y = pExceeds, color = Method, group = Analysis)) +
+    geom_point(aes(color = Analysis), size = 3, position = position_dodge(width = 1/4)) + 
+    scale_color_brewer(palette = "Set2") + 
+    scale_x_discrete(NULL, limits = consistency_order) + 
+    scale_y_continuous("% settings exceeding nominal FDR", labels = scales::percent) + 
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    ggtitle("FDR Control")
+
+p2b <- bind_rows(Yeast = power_yeast, Sims = power_sims, .id = "Analysis") %>%
+    dplyr::mutate(Method = reorder(Method, meanRank),
+                  Analysis = factor(Analysis, levels = c("Yeast", "Sims"))) %>%
+    ggplot(aes(x = Method, y = meanRank, color = Method, group = Analysis)) +
+    geom_point(aes(color = Analysis), size = 3, position = position_dodge(width = 1/4)) + 
+    scale_color_brewer(palette = "Set2") + 
+    scale_x_discrete(NULL, limits = consistency_order) + 
+    scale_y_continuous("Mean TPR rank", breaks = 0:10) + 
+    expand_limits(y = 1) +
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    ggtitle("Power")
+
+p2c <- bind_rows(Sims = consistency2_sims, CS = consistency2_data, .id = "Analysis") %>%
+    dplyr::mutate(Method = reorder(Method, propEither),
+                  Analysis = factor(paste0(Analysis, " (", key, ")"))) %>%
+    ggplot(aes(x = Method, y = propEither, color = Method, group = Analysis)) +
+    geom_point(aes(color = Analysis), size = 3, position = position_dodge(width = 1/4)) +
+    scale_color_brewer(palette = "Set1") + 
+    scale_x_discrete(NULL, limits = consistency_order) + 
+    scale_y_continuous("% settings w/ metric < classic methods", labels = scales::percent) + 
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    ggtitle("Consistency of Gain")
+
+p2d <- bind_rows(Sims = consistency_sims, CS = consistency_data, .id = "Analysis") %>%
+    dplyr::mutate(Method = reorder(Method, sdRGain),
+                  Analysis = factor(paste0(Analysis, " (", key, ")")),
+                  Analysis = factor(Analysis, levels = levels(Analysis)[c(1, 3, 2)])) %>%
+    ggplot(aes(x = Method, y = sdRGain, color = Method, group = Analysis)) +
+    geom_point(aes(color = Analysis), size = 3, position = position_dodge(width = 1/4)) + 
+    scale_color_brewer(palette = "Set1") + 
+    scale_x_discrete(NULL, limits = consistency_order) + 
+    scale_y_continuous("SD of log-ratio relative to classic methods") + 
+    expand_limits(y = 1) +
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    ggtitle("Relative Consistency")
+
+p2title <- ggdraw() +
+    draw_label("Summary Recommendation Metrics",
+               fontface = 'bold')
+
+p2ab <- plot_grid(p2a + guides(color = FALSE),
+                  p2b + guides(color = FALSE),
+                  labels = LETTERS[1:2], nrow = 1)
+p2ab <- plot_grid(p2ab, get_legend(p2a), nrow = 1, rel_widths = c(1, .2))
+
+p2cd <- plot_grid(p2c + guides(color = FALSE),
+                  p2d + guides(color = FALSE),
+                  labels = LETTERS[3:4], nrow = 1)
+p2cd <- plot_grid(p2cd, get_legend(p2d), nrow = 1, rel_widths = c(1, .2))
+
+FigS2 <- plot_grid(p2title, p2ab, p2cd, ncol = 1, rel_heights = c(.2, 1, 1))
+FigS2
+
+ggsave(file.path(outdir, "SFigureMetrics-fixedorder.pdf"), width=10, height=8)
+```

### 2018-10-26 manuscript_figures/manuscript_figures_supp.Rmd
Fix error introduced in previous commit

@@
 res <- addDefaultMetrics(res)
-pC <- covariateLinePlot(res, alpha=alpha, covname="ind_covariate", trans="log1p",
+pC <- covariateLinePlot(res, alpha=alpha.thresh, covname="ind_covariate", trans="log1p",
                         nbins=25) +

### 2018-10-26 manuscript_figures/manuscript_ratings.Rmd
Change legend title in ratings figure

--- 2018-10-26 manuscript_figures/manuscript_ratings.Rmd
@@
 ```{r, fig.width = 10, fig.height = 8}
-p1a <- bind_rows(Yeast = fdrcontrol_yeast, Sims = fdrcontrol_sims, .id = "Analysis") %>%
+p1a <- bind_rows(Yeast = fdrcontrol_yeast, Sims = fdrcontrol_sims, .id = "Study") %>%
     dplyr::mutate(Method = reorder(Method, pExceeds),
-                  Analysis = factor(Analysis, levels = c("Yeast", "Sims"))) %>%
-    ggplot(aes(x = Method, y = pExceeds, color = Method, group = Analysis)) +
-    geom_line(aes(linetype = Analysis), color = 'black', alpha = 1/2) +
+                  Study = factor(Study, levels = c("Yeast", "Sims"))) %>%
+    ggplot(aes(x = Method, y = pExceeds, color = Method, group = Study)) +
+    geom_line(aes(linetype = Study), color = 'black', alpha = 1/2) +
     geom_point(size = 3) + 
@@
 
-p1b <- bind_rows(Yeast = power_yeast, Sims = power_sims, .id = "Analysis") %>%
+p1b <- bind_rows(Yeast = power_yeast, Sims = power_sims, .id = "Study") %>%
     dplyr::mutate(Method = reorder(Method, meanRank),
-                  Analysis = factor(Analysis, levels = c("Yeast", "Sims"))) %>%
-    ggplot(aes(x = Method, y = meanRank, color = Method, group = Analysis)) +
-    geom_line(aes(linetype = Analysis), color = 'black', alpha = 1/2) +
+                  Study = factor(Study, levels = c("Yeast", "Sims"))) %>%
+    ggplot(aes(x = Method, y = meanRank, color = Method, group = Study)) +
+    geom_line(aes(linetype = Study), color = 'black', alpha = 1/2) +
     geom_point(size = 3) + 
@@
 
-p1c <- bind_rows(Sims = consistency2_sims, CS = consistency2_data, .id = "Analysis") %>%
-    dplyr::mutate(Method = reorder(Method, propEither95),
-                  Analysis = factor(paste0(Analysis, " (", key, ")")),
-                  Aanlysis = factor(Analysis, levels = levels(Analysis)[c(2, 3, 1)])) %>%
-    ggplot(aes(x = Method, y = propEither95, color = Method, group = Analysis)) +
-    geom_line(aes(linetype = Analysis), color = 'black', alpha = 1/2) +
+p1c <- bind_rows(Sims = consistency2_sims, CS = consistency2_data, .id = "Metric") %>%
+    dplyr::mutate(Method = reorder(Method, propEither),
+                  Metric = factor(paste0(Metric, " (", key, ")"))) %>%
+    ggplot(aes(x = Method, y = propEither, color = Method, group = Metric)) +
+    geom_line(aes(linetype = Metric), color = 'black', alpha = 1/2) +
     geom_point(size = 3) + 
@@
           plot.title = element_text(face = "bold", size = 14)) +
-    ggtitle("Consistency (I)")
+    ggtitle("Consistency of Gain")
 
-p1d <- bind_rows(Sims = consistency_sims, CS = consistency_data, .id = "Analysis") %>%
+p1d <- bind_rows(Sims = consistency_sims, CS = consistency_data, .id = "Metric") %>%
     dplyr::mutate(Method = reorder(Method, sdRGain),
-                  Analysis = factor(paste0(Analysis, " (", key, ")"))) %>%
-    ggplot(aes(x = Method, y = sdRGain, color = Method, group = Analysis)) +
-    geom_line(aes(linetype = Analysis), color = 'black', alpha = 1/2) +
+                  Metric = factor(paste0(Metric, " (", key, ")")),
+                  Metric = factor(Metric, levels = levels(Metric)[c(1, 3, 2)])) %>%
+    ggplot(aes(x = Method, y = sdRGain, color = Method, group = Metric)) +
+    geom_line(aes(linetype = Metric), color = 'black', alpha = 1/2) +
     geom_point(size = 3) + 
@@
 ```
+
+Additionally, we consider creating the above plot with a common x-axis.
+
+```{r, fig.width = 10, fig.height = 8}
+consistency_order <- c("bh", "qvalue", "bl", "ihw", "lfdr", "fdrreg-t", "ashq", "adapt-glm")
+
+p2a <- bind_rows(Yeast = fdrcontrol_yeast, Sims = fdrcontrol_sims, .id = "Study") %>%
+    dplyr::mutate(Method = reorder(Method, pExceeds),
+                  Study = factor(Study, levels = c("Yeast", "Sims"))) %>%
+    ggplot(aes(x = Method, y = pExceeds, color = Study, group = Study)) +
+    geom_point(size = 3, position = position_dodge(width = 1/4)) + 
+    scale_color_brewer(palette = "Set2") + 
+    scale_x_discrete(NULL, limits = consistency_order) + 
+    scale_y_continuous("% settings exceeding nominal FDR", labels = scales::percent) + 
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    ggtitle("FDR Control")
+
+p2b <- bind_rows(Yeast = power_yeast, Sims = power_sims, .id = "Study") %>%
+    dplyr::mutate(Method = reorder(Method, meanRank),
+                  Study = factor(Study, levels = c("Yeast", "Sims"))) %>%
+    ggplot(aes(x = Method, y = meanRank, color = Study, group = Study)) +
+    geom_point(size = 3, position = position_dodge(width = 1/4)) + 
+    scale_color_brewer(palette = "Set2") + 
+    scale_x_discrete(NULL, limits = consistency_order) + 
+    scale_y_continuous("Mean TPR rank", breaks = 0:10) + 
+    expand_limits(y = 1) +
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    ggtitle("Power")
+
+p2c <- bind_rows(Sims = consistency2_sims, CS = consistency2_data, .id = "Metric") %>%
+    dplyr::mutate(Method = reorder(Method, propEither),
+                  Metric = factor(paste0(Metric, " (", key, ")"))) %>%
+    ggplot(aes(x = Method, y = propEither, color = Metric, group = Metric)) +
+    geom_point(size = 3, position = position_dodge(width = 1/4)) +
+    scale_color_brewer(palette = "Set1") + 
+    scale_x_discrete(NULL, limits = consistency_order) + 
+    scale_y_continuous("% settings w/ metric < classic methods", labels = scales::percent) + 
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    ggtitle("Consistency of Gain")
+
+p2d <- bind_rows(Sims = consistency_sims, CS = consistency_data, .id = "Metric") %>%
+    dplyr::mutate(Method = reorder(Method, sdRGain),
+                  Metric = factor(paste0(Metric, " (", key, ")")),
+                  Metric = factor(Metric, levels = levels(Metric)[c(1, 3, 2)])) %>%
+    ggplot(aes(x = Method, y = sdRGain, color = Metric, group = Metric)) +
+    geom_point(size = 3, position = position_dodge(width = 1/4)) + 
+    scale_color_brewer(palette = "Set1") + 
+    scale_x_discrete(NULL, limits = consistency_order) + 
+    scale_y_continuous("SD of log-ratio relative to classic methods") + 
+    expand_limits(y = 1) +
+    theme_classic() +
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    ggtitle("Relative Consistency")
+
+p2title <- ggdraw() +
+    draw_label("Summary Recommendation Metrics",
+               fontface = 'bold')
+
+p2ab <- plot_grid(p2a + guides(color = FALSE),
+                  p2b + guides(color = FALSE),
+                  labels = LETTERS[1:2], nrow = 1)
+p2ab <- plot_grid(p2ab, get_legend(p2a), nrow = 1, rel_widths = c(1, .2))
+
+p2cd <- plot_grid(p2c + guides(color = FALSE),
+                  p2d + guides(color = FALSE),
+                  labels = LETTERS[3:4], nrow = 1)
+p2cd <- plot_grid(p2cd, get_legend(p2d), nrow = 1, rel_widths = c(1, .2))
+
+FigS2 <- plot_grid(p2title, p2ab, p2cd, ncol = 1, rel_heights = c(.2, 1, 1))
+FigS2
+
+ggsave(file.path(outdir, "SFigureMetrics-fixedorder.pdf"), width=10, height=8)
+```

### 2018-10-27 manuscript_figures/manuscript_figures_main.Rmd
Add labels for lines difficult to distinguish in Figs 1 and 2

@@
 
-```{r Figure2-load-data}
+```{r setup}
 # Load packages and source benchmark FDR
@@
 library(SummarizedBenchmark)
+library(ggrepel)
 
@@
       scale_y_continuous("FDR", labels=scales::percent, breaks=seq(0, 1, by=0.02)) +
-      expand_limits(y=0.11) +
+      expand_limits(y=0.11, x = 0.105) +
       theme(plot.title = element_text(face = "bold", size = 12))
@@
   filter(blabel %in% methodset)
+datsub <- res5dp %>% filter(performanceMetric == "FDR",
+                                        abs(alpha-0.10) < 1e-6,
+                                        blabel %in% c("lfdr", "ashq")) %>%
+                                 group_by(blabel, alpha) %>%
+                                 summarize(value = mean(value)) %>%
+                                 mutate(Method = blabel) 
 p1b <- plotsim_average(res5dp, met="FDR",
@@
       scale_y_continuous("", labels=scales::percent, breaks=seq(0, 1, by=0.02)) +
-      expand_limits(y=0.11) +
+      expand_limits(y=0.11, x = 0.105) +
       theme(plot.title = element_text(face = "bold", size = 12)) + 
-  theme(legend.position="none")
+      theme(legend.position="none") +
+      geom_label_repel(data=datsub,
+                 aes(alpha, value, label = Method),
+                 min.segment.length = unit(0, 'lines'),
+                 nudge_x = c(-0.01, 10), nudge_y = c(0,0.01),
+                 label.padding = 0.1, 
+                 box.padding = 0.1, size = 2.25)
 
@@
 ## Figure 2A: TPR of methods in representative yeast setting at varying alpha levels
+datsub <- res5d %>% filter(performanceMetric == "TPR",
+                                        abs(alpha-0.10) < 1e-6,
+                                        blabel %in% c("adapt-glm", "lfdr", "fdrreg-t")) %>%
+                                 group_by(blabel, alpha) %>%
+                                 summarize(value = mean(value)) %>%
+                                 mutate(Method = blabel) %>%
+                                 arrange(desc(Method))
 p2a <- plotsim_average(res5d, met="TPR",
@@
       scale_x_continuous(expression(paste(bold(alpha), bold(" level"))),
-                         breaks=seq(0, 1, by=0.02)) +
+                         breaks=seq(0, 1, by=0.02), 
+                         limits=c(0,0.114)) +
       scale_y_continuous("TPR", labels=scales::percent, breaks=seq(0, 1, by=0.05)) +
-      theme(plot.title = element_text(face = "bold", size = 12))
+      theme(plot.title = element_text(face = "bold", size = 12)) +
+      geom_label_repel(data=datsub,
+                 aes(alpha, value, label = Method),
+                 min.segment.length = unit(0, 'lines'),
+                 nudge_x = c(10,10,-0.01), size = 2.25, nudge_y=c(0.008,0.008, 0.008), #l,f,a
+                 fill = "white", alpha = 1,
+                 box.padding = 0.1, label.padding = 0.1)
 
 ## Figure 2B: TPR of methods in representative polyester setting at varying alpha levels
+datsub <- res5dp %>% filter(performanceMetric == "TPR",
+                                        abs(alpha-0.10) < 1e-6,
+                                        blabel %in% c("lfdr", "fdrreg-t")) %>%
+                                 group_by(blabel, alpha) %>%
+                                 summarize(value = mean(value)) %>%
+                                 mutate(Method = blabel) 
 p2b <- plotsim_average(res5dp, met="TPR",
@@
       scale_x_continuous(expression(paste(bold(alpha), bold(" level"))),
-                         breaks=seq(0, 1, by=0.02)) +
+                         breaks=seq(0, 1, by=0.02), 
+                         limits = c(0,0.114)) +
       scale_y_continuous("", labels=scales::percent, breaks=seq(0, 1, by=0.05)) +
       theme(plot.title = element_text(face = "bold", size = 12)) + 
-  theme(legend.position="none")
+      theme(legend.position="none") +
+      geom_label_repel(data=datsub,
+                 aes(alpha, value, label = Method),
+                 min.segment.length = unit(0, 'lines'),
+                 nudge_x = 10, size = 2.25, 
+                 box.padding = 0.1, label.padding = 0.1,
+                 fill = "white", alpha = 1)
 
@@
                   labels = c("A", "", "B", ""))
-Fig2 <- plot_grid(Fig2, get_legend(p2a), rel_widths = c(1, .2))
+Fig2 <- plot_grid(Fig2, get_legend(p1a), rel_widths = c(1, .2))
 Fig2

### 2018-10-30 manuscript_figures/manuscript_figures_main.Rmd
Reorder and rename figures

--- 2018-10-25 manuscript_figures/manuscript_figures_main.Rmd
@@
 
-```{r Figure2-load-data}
+```{r workspacesetup}
 # Load packages and source benchmark FDR
@@
 
-# Figure 1 - FDR
+# Figure 1 - Methods summary table
+
+This figure is created manually using keynote.
+
+# Figure 2 - FDR
 
@@
 
-```{r Figure1, fig.width = 7.5, fig.height = 6.5}
+```{r Figure2, fig.width = 7.5, fig.height = 6.5}
 # FDR in in silico experiments across alpha values 
-## Figure 1A: FDR of methods in representative yeast setting at varying alpha levels
+## Figure 2A: FDR of methods in representative yeast setting at varying alpha levels
 objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
@@
   filter(blabel %in% methodset)
-p1a <- plotsim_average(res5d, met="FDR",
+p2a <- plotsim_average(res5d, met="FDR",
                 merge_ihw = TRUE, errorBars=TRUE) +
@@
       scale_y_continuous("FDR", labels=scales::percent, breaks=seq(0, 1, by=0.02)) +
-      expand_limits(y=0.11) +
+      expand_limits(y=0.11, x = 0.105) +
       theme(plot.title = element_text(face = "bold", size = 12))
 
-## Figure 1B: FDR of methods in representative polyester setting at varying alpha levels
+## Figure 2B: FDR of methods in representative polyester setting at varying alpha levels
 objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
@@
   filter(blabel %in% methodset)
-p1b <- plotsim_average(res5dp, met="FDR",
+datsub <- res5dp %>% filter(performanceMetric == "FDR",
+                                        abs(alpha-0.10) < 1e-6,
+                                        blabel %in% c("lfdr", "ashq")) %>%
+                                 group_by(blabel, alpha) %>%
+                                 summarize(value = mean(value)) %>%
+                                 mutate(Method = blabel) 
+p2b <- plotsim_average(res5dp, met="FDR",
                 merge_ihw = TRUE, errorBars=TRUE) +
@@
       scale_y_continuous("", labels=scales::percent, breaks=seq(0, 1, by=0.02)) +
-      expand_limits(y=0.11) +
+      expand_limits(y=0.11, x = 0.105) +
       theme(plot.title = element_text(face = "bold", size = 12)) + 
-  theme(legend.position="none")
+      theme(legend.position="none") +
+      geom_label_repel(data=datsub,
+                 aes(alpha, value, label = Method),
+                 min.segment.length = unit(0, 'lines'),
+                 nudge_x = c(-0.01, 10), nudge_y = c(0,0.01),
+                 label.padding = 0.1, 
+                 box.padding = 0.1, size = 2.25)
 
-p1abTitle <- ggdraw() +
+p2abTitle <- ggdraw() +
     draw_label(expression(paste(bold("FDR control in RNA-seq "), 
@@
 # Panels C and D -> FDR in sims across settings at a single specified alpha
-## Figure 1C: FDR of methods across varying number of tests at alpha.thresh
+## Figure 2C: FDR of methods across varying number of tests at alpha.thresh
 res_ntests <- dplyr::filter(resmet, setting == "varyingntests", 
                             alpha == alpha.thresh, Method %in% methodset) 
-p1c <- genplot(res_ntests %>% filter(ntests >= 4000 | Method != "lfdr"), 
+p2c <- genplot(res_ntests %>% filter(ntests >= 4000 | Method != "lfdr"), 
                cov = ntests, met = "FDR", ebw = 0.1) +
@@
 
-## Figure 1D: FDR of methods across varying null proportion at alpha.thresh
+## Figure 2D: FDR of methods across varying null proportion at alpha.thresh
 res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0", 
@@
                          pi0 >= 10) # restrict to pi0 at least 10%
-p1d <- genplot(res_pi0, cov = 100-pi0, met = "FDR", ebw = 2.5) +
+p2d <- genplot(res_pi0, cov = 100-pi0, met = "FDR", ebw = 2.5) +
     scale_x_continuous("Proportion non-null",
@@
 
-p1cdTitle <- ggdraw() +
+p2cdTitle <- ggdraw() +
     draw_label(expression(paste(bold("FDR across simulation settings ("), 
@@
 
-## pull Figure 1 together
-top <- plot_grid(p1a + theme(legend.position="none"), p1b, nrow = 1)
-bot <- plot_grid(p1c, p1d, nrow = 1)
-Fig1 <- plot_grid(p1abTitle, top, p1cdTitle, bot, nrow=4, 
+## pull Figure 2 together
+top <- plot_grid(p2a + theme(legend.position="none"), p2b, nrow = 1)
+bot <- plot_grid(p2c, p2d, nrow = 1)
+Fig2 <- plot_grid(p2abTitle, top, p2cdTitle, bot, nrow=4, 
                   rel_heights = c(0.1, 1, 0.1, 1),
                   labels = c("A", "", "B", ""))
-Fig1 <- plot_grid(Fig1, get_legend(p1a), rel_widths = c(1, .2))
-Fig1
-ggsave(file.path(outdir, "Fig1.pdf"), width=7.5, height=6.5)
+Fig2 <- plot_grid(Fig2, get_legend(p2a), rel_widths = c(1, .2))
+Fig2
+ggsave(file.path(outdir, "Figure2.pdf"), width=7.5, height=6.5)
 ```
 
-# Figure 2 - Power
+# Figure 3 - Power
 
 This figure contains the main TPR results for the simulations and *in silico* 
-experiments, and is analagous to Figure 1. 
+experiments, and is analagous to Figure 2. 
 Panel (A) contains the TPR results for a representative setting of
@@
 
-```{r Figure2, fig.width = 7.5, fig.height = 6.5}
+```{r Figure3, fig.width = 7.5, fig.height = 6.5}
 # TPR in in silico experiments across alpha values 
-## Figure 2A: TPR of methods in representative yeast setting at varying alpha levels
-p2a <- plotsim_average(res5d, met="TPR",
+## Figure 3A: TPR of methods in representative yeast setting at varying alpha levels
+datsub <- res5d %>% filter(performanceMetric == "TPR",
+                                        abs(alpha-0.10) < 1e-6,
+                                        blabel %in% c("adapt-glm", "lfdr", "fdrreg-t")) %>%
+                                 group_by(blabel, alpha) %>%
+                                 summarize(value = mean(value)) %>%
+                                 mutate(Method = blabel) %>%
+                                 arrange(desc(Method))
+p3a <- plotsim_average(res5d, met="TPR",
                 merge_ihw = TRUE, errorBars=TRUE) +
@@
       scale_x_continuous(expression(paste(bold(alpha), bold(" level"))),
-                         breaks=seq(0, 1, by=0.02)) +
+                         breaks=seq(0, 1, by=0.02), 
+                         limits=c(0,0.114)) +
       scale_y_continuous("TPR", labels=scales::percent, breaks=seq(0, 1, by=0.05)) +
-      theme(plot.title = element_text(face = "bold", size = 12))
+      theme(plot.title = element_text(face = "bold", size = 12)) +
+      geom_label_repel(data=datsub,
+                 aes(alpha, value, label = Method),
+                 min.segment.length = unit(0, 'lines'),
+                 nudge_x = c(10,10,-0.01), size = 2.25, nudge_y=c(0.008,0.008, 0.008), #l,f,a
+                 fill = "white", alpha = 1,
+                 box.padding = 0.1, label.padding = 0.1)
 
-## Figure 2B: TPR of methods in representative polyester setting at varying alpha levels
-p2b <- plotsim_average(res5dp, met="TPR",
+## Figure 3B: TPR of methods in representative polyester setting at varying alpha levels
+datsub <- res5dp %>% filter(performanceMetric == "TPR",
+                                        abs(alpha-0.10) < 1e-6,
+                                        blabel %in% c("lfdr", "fdrreg-t")) %>%
+                                 group_by(blabel, alpha) %>%
+                                 summarize(value = mean(value)) %>%
+                                 mutate(Method = blabel) 
+p3b <- plotsim_average(res5dp, met="TPR",
                 merge_ihw = TRUE, errorBars=TRUE) +
@@
       scale_x_continuous(expression(paste(bold(alpha), bold(" level"))),
-                         breaks=seq(0, 1, by=0.02)) +
+                         breaks=seq(0, 1, by=0.02), 
+                         limits = c(0,0.114)) +
       scale_y_continuous("", labels=scales::percent, breaks=seq(0, 1, by=0.05)) +
       theme(plot.title = element_text(face = "bold", size = 12)) + 
-  theme(legend.position="none")
+      theme(legend.position="none") +
+      geom_label_repel(data=datsub,
+                 aes(alpha, value, label = Method),
+                 min.segment.length = unit(0, 'lines'),
+                 nudge_x = 10, size = 2.25, 
+                 box.padding = 0.1, label.padding = 0.1,
+                 fill = "white", alpha = 1)
 
-p2abTitle <- ggdraw() +
+p3abTitle <- ggdraw() +
     draw_label(expression(paste(bold("TPR in RNA-seq "), 
@@
 # Panels C and D -> TPR in sims across settings at a single specified alpha
-## Figure 2C: TPR of methods across varying number of tests at alpha.thresh
+## Figure 3C: TPR of methods across varying number of tests at alpha.thresh
 res_ntests <- dplyr::filter(resmet, setting == "varyingntests", 
                             alpha == alpha.thresh, Method %in% methodset)
-p2c <- genplot(res_ntests %>% filter(ntests >= 4000 | Method != "lfdr"),
+p3c <- genplot(res_ntests %>% filter(ntests >= 4000 | Method != "lfdr"),
                cov = ntests, met = "TPR", ebw = 0.1) +
@@
 
-## Figure 1D: FDR of methods across varying null proportion at alpha.thresh
+## Figure 3D: FDR of methods across varying null proportion at alpha.thresh
 res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0", 
@@
                          pi0 >= 10) # restrict to pi0 at least 10%
-p2d <- genplot(res_pi0, cov = 100-pi0, met = "TPR", ebw = 2.5) +
+p3d <- genplot(res_pi0, cov = 100-pi0, met = "TPR", ebw = 2.5) +
     scale_x_continuous("Proportion non-null",
@@
 
-p2cdTitle <- ggdraw() +
+p3cdTitle <- ggdraw() +
     draw_label(expression(paste(bold("TPR across simulation settings ("), 
@@
 
-## pull Figure 1 together
-top <- plot_grid(p2a + theme(legend.position="none"), p2b, nrow = 1)
-bot <- plot_grid(p2c, p2d, nrow = 1)
-Fig2 <- plot_grid(p2abTitle, top, p2cdTitle, bot, nrow=4, 
+## pull Figure 3 together
+top <- plot_grid(p3a + theme(legend.position="none"), p3b, nrow = 1)
+bot <- plot_grid(p3c, p3d, nrow = 1)
+Fig3 <- plot_grid(p3abTitle, top, p3cdTitle, bot, nrow=4, 
                   rel_heights = c(0.1, 1, 0.1, 1),
                   labels = c("A", "", "B", ""))
-Fig2 <- plot_grid(Fig2, get_legend(p2a), rel_widths = c(1, .2))
-Fig2
-ggsave(file.path(outdir, "Fig2.pdf"), width=7.5, height=6.5)
+Fig3 <- plot_grid(Fig3, get_legend(p2a), rel_widths = c(1, .2))
+Fig3
+ggsave(file.path(outdir, "Figure3.pdf"), width=7.5, height=6.5)
 ```
-
-# Figure 3 - Robustness
 
@@
 
-## Figure 4A: Performance across distribution of test statistics (based on single replicate)
+## Figure 3A: Performance across distribution of test statistics (based on single replicate)
 p4a <- ggplot(tsdists, aes(x = test_statistic, y=..count..,
@@
 
-## Figure 4B: Case study heatmap demonstrating variability in applicability
+## Figure 3B: Case study heatmap demonstrating variability in applicability
 
@@
 Fig4
-ggsave(file.path(outdir, "Fig4.pdf"), width = 11.2, height = 10.5)
+ggsave(file.path(outdir, "Figure4.pdf"), width = 11.2, height = 10.5)
 ```
 
-# Figure 5 - Summary
+# Figure 5 - Consistency
+
+This figure is generated in `manuscript_ratings.Rmd`.
+
+# Figure 6 - Summary
 

### 2018-10-30 manuscript_figures/manuscript_figures_supp.Rmd
Reorder and rename figures

@@
 
-# Figure 2
-
-This is a faceted figure with results using the yeast data with 48 biological replicates analyzed in this [publication](https://www.ncbi.nlm.nih.gov/pubmed/26206307/). A complete analysis 
-can be found in the `/datasets/RNA-Seq/yeast-simulation.Rmd`. This is the 5v5 comparison.
-
-```{r Figure2-load-data}
+# Workspace setup
+
+```{r wkspace setup}
 # Load packages and source benchmark FDR
@@
 
-```{r Figure2, fig.width = 10, fig.height = 8}
+# Figure S1
+
+This is a faceted figure with results using the yeast data with 48 biological replicates analyzed in this [publication](https://www.ncbi.nlm.nih.gov/pubmed/26206307/). A complete analysis 
+can be found in the `/datasets/RNA-Seq/yeast-simulation.Rmd`. This is the 5v5 comparison.
+
+```{r FigureS1, fig.width = 10, fig.height = 8}
 objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
@@
 
-# Figure 2A: Not all methods control FDR at nominal alpha level
+# Figure S1A: Not all methods control FDR at nominal alpha level
 p1 <- plotsim_average(res5d, met="FDR",filter_set = excludeSet,
@@
 
-# Figure 2B: TPR of methods across nominal alpha level
+# Figure S1B: TPR of methods across nominal alpha level
 p2 <- plotsim_average(res5d, met="TPR",filter_set = excludeSet,
@@
 
-# Figure 2C: ROC curve
+# Figure S1C: ROC curve
 p3 <- plotsim_average(res5d, met=c("FPR", "TPR"),filter_set = excludeSet,
@@
 
-# Figure 2D: Some methods are tailored to detecting smaller effect sizes
+# Figure S1D: Some methods are tailored to detecting smaller effect sizes
 p4 <- covariateLinePlot(de5, alpha=alpha.thresh, covname="log2FC", trans="log1p",
@@
 
-# Figure 2E - upsetr plot - save separately
-pdf(file.path(outdir, "Figure2E.pdf"), width = 8, height = 4, onefile=FALSE)
+# Figure S1E - upsetr plot - save separately
+pdf(file.path(outdir, "FigureS1E.pdf"), width = 8, height = 4, onefile=FALSE)
 aggupset(de5, alpha=alpha.thresh, supplementary = FALSE, 
@@
 # Figure 2 (ALL)
-Fig2 <- plot_grid(p1, p2, p3, p4, ncol = 2, 
+FigS1 <- plot_grid(p1, p2, p3, p4, ncol = 2, 
           labels = LETTERS[1:4], label_size = 20, align = "hv")
-Fig2 <- plot_grid(Fig2, leg, ncol = 2, rel_widths = c(1, 0.2))
-Fig2
-ggsave(file.path(outdir, "Figure2A-D.pdf"), width=8.75, height=6)
-
-```
-
-# Figure 3
-
-Replaced with a pure simulation figure - see `manuscript_figures_sims.Rmd`.
-
-# Figure 4
-
-This is a heatmap of methods by case studies with color representing the ranking
-by of number of rejections @ FDR cutoff `r alpha.thresh`.
-
-```{r, Figure 4, fig.width = 9, fig.height = 7}
-library(SummarizedBenchmark)
-library(stringr)
-
+FigS1 <- plot_grid(FigS1, leg, ncol = 2, rel_widths = c(1, 0.2))
+FigS1
+ggsave(file.path(outdir, "FigureS1A-D.pdf"), width=8.75, height=6)
+
+```
+
+# Figure S2
+
+This Figure contrasts FDR results in different yeast simulation settings.
+
+```{r Figure-yeast-pi0, fig.width = 10, fig.height = 8}
+# strongly informative covariate (top panel)
 objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
-
-# remove simulation objects
-sim <- unique(c(which(grepl("simulations", objects)),
-               which(grepl("yeast", objects)),
-               which(grepl("polyester", objects))))
-if (length(sim) > 0){
-  objects <- objects[-sim]
-}
-
-# remove microbiome objects from 'null' comparisons
-# removing baxter, goodrich, papa OTU-level, and So4-enigma (since they mostly found 
-# nothing for all methods and the genus-level analysis is also present)
-# Also remove'log' ubiquity since redundant
-null <- which(grepl(c("goodrich|papa|baxter"), objects) & 
-              grepl(c("otu"), objects) | 
-              grepl("log", objects) | 
-              grepl("so4", objects))
-
-if (length(null) > 0){
-  objects <- objects[-null]
-}
-
-# remove chip-seq promoter analysis (since it is the only one that uses 
-# ash and we also include the csaw analysis on the same dataset).
-promot <- which(grepl("promot", objects))
-if (length(promot) > 0){
-  objects <- objects[-promot]
-}
-
-# remove random covariate analyses
-uninf <- which(grepl("uninf", objects))
-if (length(uninf) > 0){
-  objects <- objects[-uninf]
-}
-
-# grab casestudy names
-obj.names <- str_split(objects, pattern = "/")
+resfile <- objects[(grepl("yeast-", objects) | grepl("yeastH-", objects) | 
+                    grepl("yeastII-", objects) | grepl("yeastIIH-", objects)) & 
+                   grepl("de5", objects) & 
+                  !grepl("uninf", objects)] 
+labs <- rep("30% Non-null", length(resfile))
+labs[grepl("H", resfile)] <- "7.5% Non-null"
+labs2 <- rep("Unimodal effect size", length(resfile))
+labs2[grepl("II", resfile)] <- "Bimodal effect size"
+excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
+
+plist <- vector("list", length(resfile))
+
+if (file.exists("yeast_summary_tab_all.rds")){
+  tab <- readRDS("yeast_summary_tab_all.rds")
+}else{
+  tab <- NULL
+  for (l in seq_along(resfile)){
+    de5 <- readRDS(file=resfile[l])
+    res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
+    res5d$es_dist <- labs2[l]
+    res5d$pi0 <- labs[l]
+    tab <- rbind(tab, res5d)
+  }
+  saveRDS(tab, file = "yeast_summary_tab_all.rds")
+}
+
+FigStrong <- plotsim_average(tab, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE, grpVars = c("es_dist", "pi0")) +
+                ggtitle("") +
+  facet_grid( ~ es_dist + pi0) +
+  ylim(0,0.17) 
+FigStrong
+
+FigStrongT <- plotsim_average(tab, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE, grpVars = c("es_dist", "pi0")) +
+                ggtitle("") +
+  ylim(0,0.72) +
+  facet_grid( ~ es_dist + pi0) 
+FigStrongT
+
+# weakly informative covariate (bottom panel)
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+resfile <- objects[(grepl("yeastW-", objects) | grepl("yeastHW-", objects) | 
+                    grepl("yeastIIW-", objects) | grepl("yeastIIHW-", objects)) & 
+                   grepl("de5", objects) & 
+                  !grepl("uninf", objects)] 
+labs <- rep("30% Non-null", length(resfile))
+labs[grepl("H", resfile)] <- "7.5% Non-null"
+labs2 <- rep("Unimodal effect size", length(resfile))
+labs2[grepl("II", resfile)] <- "Bimodal effect size"
+excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
+
+plist <- vector("list", length(resfile))
+tab <- NULL
+for (l in seq_along(resfile)){
+  de5 <- readRDS(file=resfile[l])
+  res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
+  res5d$es_dist <- labs2[l]
+  res5d$pi0 <- labs[l]
+  tab <- rbind(tab, res5d)
+}
+
+FigWeak <- plotsim_average(tab, met="FDR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE, grpVars = c("es_dist", "pi0")) +
+                ggtitle("") +
+  ylim(0,0.17) +
+  facet_grid( ~ es_dist + pi0)
+FigWeak
+
+FigWeakT <- plotsim_average(tab, met="TPR",filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE, grpVars = c("es_dist", "pi0")) +
+                ggtitle("") +
+  ylim(0,0.72) +
+  facet_grid( ~ es_dist + pi0)
+FigWeakT
+
+leg <- get_legend(FigStrong)
+
+FigA <- plot_grid(FigStrong + theme(legend.position="none") +  
+                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)), 
+                  FigStrongT + theme(legend.position="none") +  
+                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)),
+                 nrow = 2, align = "hv")
+
+FigB <- plot_grid(FigWeak + theme(legend.position="none") +  
+                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)), 
+                  FigWeakT + theme(legend.position="none") +  
+                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)),
+                 nrow = 2, align = "hv")
+
+Fig <- plot_grid(FigA, FigB,
+                 nrow = 2, 
+                 labels = paste0(LETTERS[1:2], "       ",
+                          c("Strongly informative covariate",
+                            "Weakly informative covariate")),
+                 hjust = 0)
+plot_grid(Fig, leg, ncol=2, rel_widths = c(1,0.1))
+ggsave(file.path(outdir, "FigureS2.pdf"), width=10, height=12)
+
+```
+
+# Figures S3-S8
+
+Figures S3-S8 are created in `manuscript_figures_sims.Rmd`.
+
+# Figures S9
+
+Here we'll contrast the results of informative and uninformative covariates 
+on the yeast simulation (DE sample size 5). We'll plot
+differences in TPR and FDR due to informative covariate by alpha level.
+
+```{r, fig.width = 10, fig.height = 4}
+org <- "yeast"
+
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+resfile <- objects[grepl(org, objects) ] 
+resfile <- resfile[grepl("5", resfile) & !grepl("null", resfile)]
+
+# regexp to construct master data table separating case study datasets, methods
+# and covariates
+obj.names <- str_split(resfile, pattern = "/")
 casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
-dataset <- unlist(lapply(obj.names, function(x) x[4] ))
-
-# not excluding any methods
-Fig4All.rowOrder <- c("bonf", "bh", "ihw", "qvalue", "bl", "adapt-glm", "lfdr",
-                     "fdrreg-e", "fdrreg-t", "ashq")
-Fig4All <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha.thresh,
-                        excludeMethods = NULL, rowOrder = Fig4All.rowOrder) +
-        theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
-              legend.position = c(0.97, 0.92), legend.justification = c(1, 1),
-              legend.background = element_rect(fill = scales::alpha("white", 3/4), color = 'black'),
-              legend.direction = "horizontal",
-              legend.text = element_text(size = 7),
-              legend.title = element_text(size = 9),
-              legend.margin = margin(6, 10, 6, 10))
-
-Fig4All
-```
-
-Extract a table of the values in Figure 4.
-
-```{r}
-# Supplementary table S3 
-library(xtable)
-tabS3 <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha.thresh,
-                        tableOnly = TRUE, excludeMethods = NULL)
-tabS3 <- tabS3 %>% 
-  mutate(mean_prop = ifelse(is.na(mean_prop), "-",
-                            paste0(round(mean_prop,2), " (", round(min_prop,2), 
-                            "-", round(max_prop, 2), ")"))) %>%
-  select(casestudy, method, mean_prop) %>%
-  spread(casestudy, mean_prop) %>% 
-  mutate(method =  factor(method, levels = rev(Fig4All.rowOrder))) %>%
-  arrange(method)
-
-print(xtable(tabS3), include.rownames=FALSE)
-
-# generate detailed case study table for creating sim vs cs ranks boxplots
-casestudy.tab <- plotMethodRanks(objects, colLabels = dataset, alpha = alpha.thresh,
-                                 tableOnly = TRUE)
-```
-
-Create more panels with heatmaps of yeast simulation results.
-
-```{r, fig.width = 14, fig.height = 5}
-objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
-
-# grab casestudy names
-obj.names <- str_split(objects, pattern = "/")
-casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
-dataset <- unlist(lapply(obj.names, function(x) x[4] ))
-
-# non-null
-wcs <- which(grepl("RNAseq", casestudy) & 
-            grepl(c("yeast-"), dataset) &
-            grepl("de", dataset) &
-            !grepl("scRNAseq", casestudy) &
-            !grepl("uninf", dataset))
-objs <- objects[wcs]
-labs <- gsub(".rds", "", 
-             unlist(lapply(str_split(dataset[wcs], pattern = "-results-"), 
-                      function(x) x[2])))
-
-pfdr <- plotMethodRanks(objs, 
-                colLabels = c("sim", "sim"), 
-                alpha = alpha.thresh, xlab = "Comparison",
-                fill = "FDR",
-                rowOrder = Fig4All.rowOrder, 
-                annotate = "FDR", excludeMethods = NULL,
-                maxColor = "black") +
-          scale_fill_gradient2(high = "darkred", 
-                         mid = "white", 
-                         low = "#2171B5", 
-                         midpoint = alpha.thresh) +
-          labs(fill = "Mean FDR") +
-          theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
-              legend.justification = c(1, 0.5),
-              legend.text = element_text(size = 7),
-              legend.title = element_text(size = 8),
-              legend.margin = margin(1, 1, 1, 1)) +
-          ylab("") + xlab("") +  
-          theme(plot.margin=unit(c(5.5, 5.5, 30, 5.5), "points"))
-
-ptpr <- plotMethodRanks(objs, 
-                colLabels = c("sim", "sim"), 
-                alpha = alpha.thresh, xlab = "Comparison",
-                fill = "TPR",
-                rowOrder = Fig4All.rowOrder,
-                annotate = "TPR", excludeMethods = NULL) +
-          theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
-              legend.justification = c(1, 0.5),
-              legend.text = element_text(size = 7),
-              legend.title = element_text(size = 8),
-              legend.margin = margin(1, 1, 1, 1)) +
-          ylab("") + xlab("") +  
-          theme(plot.margin=unit(c(5.5, 5.5, 30, 5.5), "points"))
-
-ggdraw() +
-  draw_plot(pfdr, 0.53, 0, 0.24, 1) +
-  draw_plot(ptpr, 0.77, 0, 0.24, 1) +
-  draw_plot(Fig4All, 0, 0, 0.52, 1) +
-  draw_plot_label(c("A", "B"), c(0, 0.53), c(1, 1), size = 15) +
-  draw_plot_label("Simulation", 0.69, 0.06, size =11, fontface = "plain")
-ggsave(file.path(outdir, "Figure4.pdf"), width=9.5, height=5)
-
-# generate detailed yeast sim table for creating sim vs cs ranks boxplots
-yeastsim.tab <- plotMethodRanks(objs, colLabels = labs, alpha = alpha.thresh,
-                                 tableOnly = TRUE)
-```
-
-Create boxplot of ranking in sims versus rankings in case studies using the
-`casestudy.tab` and `yeastsim.tab` objects created above.
-
-```{r, fig.width = 6, fig.height = 4}
-casestudy.tab <- mutate(casestudy.tab, prop_casestudy = mean_prop) %>%
-  select(casestudy, method, prop_casestudy)
-yeastsim.tab <- yeastsim.tab %>% 
-  mutate(prop_yeastsim = mean_prop) %>% 
-  filter(casestudy == "de5") %>%
-  select(casestudy, method, prop_yeastsim)
-
-rank.tab <- left_join(casestudy.tab, yeastsim.tab, by = "method") %>%
-  left_join(candycols %>% dplyr::mutate(method=Method), by = "method")
-
-col <- as.character(rank.tab$col)
-names(col) <- as.character(rank.tab$method)
-
-rank.tab$method <- factor(rank.tab$method,
-                          levels = yeastsim.tab$method[order(yeastsim.tab$prop_yeastsim)])
-x <- match(rank.tab$casestudy.x, dataset)
-rank.tab$casestudy <- casestudy[x]
-
-ggplot(rank.tab %>% filter(method != "bonf"), 
-       aes(y = prop_casestudy, x = prop_yeastsim, group = prop_yeastsim, color=Method)) +
-  geom_boxplot(aes(x = prop_yeastsim, y = prop_casestudy),
-                   alpha = 0.3, position = position_dodge(width = 0),
-                   width = 0.02) +
-  scale_color_manual(values=col) +
-  theme_bw() +
-  xlab("Proportion of maximum rejections in yeast simulation (n=5)") +  
-  ylab("Proportion of maximum rejections in casestudies") +
-  labs(shape = "Case Study") 
-ggsave(file.path(outdir, "Figure_simVcasestudyRanks_scaled.pdf"), width=6, height=4)
-
-```
-
-# Figure 5
-
-This is a multi-panel Figure displaying examples of covariate line plots in the
-case studies. We'll create a 4-panel figure with the following case studies:
-RNAseq, scRNAseq, GSEA, GWAS.
-
-```{r, Figure 5, fig.width = 9, fig.height = 7}
-
-# RNAseq
-objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
-resfile <- objects[grepl("RNAseq", objects) & grepl("brain", objects) & !grepl("uninf", objects)] 
-res <- readRDS(file=resfile)
-assayNames(res) <- "qvalue"
-res <- addDefaultMetrics(res)
-pA <- covariateLinePlot(res, alpha=alpha.thresh, covname="ind_covariate", trans="log1p",
-                        nbins=25) +
-  ggtitle("RNA-seq case study: Brain") +
-  xlab("Overall mean expression percentile")
-
-leg <-  get_legend(pA)
-pA <- pA + theme(legend.position="none")
-
-# scRNAseq
-resfile <- objects[grepl("RNAseq", objects) & grepl("human", objects) &
-                   grepl("mast", objects) & grepl("det", objects) & !grepl("uninf", objects)] 
-res <- readRDS(file=resfile)
-assayNames(res) <- "qvalue"
-res <- addDefaultMetrics(res)
-pB <- covariateLinePlot(res, alpha=alpha.thresh, covname="detection", trans="log1p",
-                        nbins=25) +
-  ggtitle("scRNA-seq case study: Human, MAST") +
-  xlab("Detection rate percentile") + 
-  theme(legend.position="none")
-
-# GSEA
-resfile <- objects[grepl("GSEA", objects) & grepl("mouse", objects) & !grepl("uninf", objects)] 
-res <- readRDS(file=resfile)
-assayNames(res) <- "qvalue"
-res <- addDefaultMetrics(res)
-pC <- covariateLinePlot(res, alpha=alpha.thresh, covname="ind_covariate", trans="log1p",
-                        nbins=25) +
-  ggtitle("GSEA case study: Mouse") +
-  xlab("Gene Set size percentile") + 
-  theme(legend.position="none")
-
-# GWAS
-resfile <- objects[grepl("GWAS", objects) & grepl("maf", objects)] 
-res <- readRDS(file=resfile)
-assayNames(res) <- "qvalue"
-res <- addDefaultMetrics(res)
-pD <- covariateLinePlot(res, alpha=alpha.thresh, covname="ind_covar_AF", trans="log1p",
-                        nbins=25) +
-  ggtitle("GWAS case study") +
-  xlab("Minor Allele Frequency (MAF) percentile") +
-  theme(legend.position="none")
-
-p5 <- plot_grid(pA, pB, pC, pD, ncol=2, labels = LETTERS[1:4])
-p5 <- plot_grid(p5, leg, ncol=2, rel_widths = c(1,0.2))
-p5
-ggsave(file.path(outdir, "Figure5.pdf"), width=9, height=7)
-
-```
-
-# Figure 6
-
-This is a multi-panel Figure displaying examples of covariate vs p-value scatterplots
-and/or histograms in the case studies. 
-It will be used to discuss how we evaluated the use of the 
-covariates as independent and informative in the case studies.
-
-We'll do so for the same 4 case studies as in figure 5.
-
-```{r, Figure 6}
-# RNAseq 
-objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
-resfile <- objects[grepl("RNAseq", objects) & grepl("brain", objects) & !grepl("uninf", objects)] 
-sb <- readRDS(file=resfile)
-res <- data.frame(rowData(sb)) %>%
-  dplyr::mutate(pval = assays(sb)$bench[,"unadjusted"])
-pa <- strat_hist(res, pvalue = "pval",
-           covariate = "ind_covariate", maxy = 8.75, axislabs = FALSE) +
-  draw_plot_label("  RNA-seq case study: Brain", 0.5, 1, hjust = 0.5,fontface = "bold", 
-                  size = 12) +
-  draw_plot_label("p-value by Overall mean expression group", size = 11, 0.5, 0.14, fontface = "plain",
-                  hjust =0.5) +
-  draw_plot_label("Density", size = 11, 0, 0.3, fontface = "plain", angle=90)
-
-# scRNAseq
-resfile <- objects[grepl("RNAseq", objects) & grepl("human", objects) &
-                   grepl("mast", objects) & grepl("det", objects) & !grepl("uninf", objects)] 
-sb <- readRDS(file=resfile)
-
-res <- data.frame(rowData(sb)) %>%
-  dplyr::mutate(pval = assays(sb)$bench[,"unadjusted"]) 
-pb <- strat_hist(res, pvalue = "pval",
-           covariate = "detection", maxy = 12, axislabs = FALSE)+
-  draw_plot_label("  scRNA-seq case study: Human, MAST",0.5, 1, hjust = 0.5, fontface = "bold",
-                  size = 12 ) +
-  draw_plot_label("p-value by Detection rate group", size = 11, 0.5, 0.14, fontface = "plain",
-                  hjust =0.5) +
-  draw_plot_label("Density", size = 11, 0, 0.3, fontface = "plain", angle=90)
-
-# GSEA
-resfile <- objects[grepl("GSEA", objects) & grepl("mouse", objects) & !grepl("uninf", objects)] 
-sb <- readRDS(file=resfile)
-res <- data.frame(rowData(sb)) %>%
-  dplyr::mutate(pval = assays(sb)$bench[,"unadjusted"])
-
-pc <- strat_hist(res, pvalue = "pval",
-           covariate = "ind_covariate", maxy = 8.5, axislabs = FALSE) +
-  draw_plot_label("  GSEA case study: Mouse", 0.5, 1, hjust = 0.5, fontface = "bold" , size = 12) +
-  draw_plot_label("p-value by Gene Set size group", size = 11, 0.5, 0.14, fontface = "plain",
-                  hjust =0.5) +
-  draw_plot_label("Density", size = 11, 0, 0.3, fontface = "plain", angle=90)
-
-# GWAS
-resfile <- objects[grepl("GWAS", objects) & grepl("maf", objects)] 
-sb <- readRDS(file=resfile)
-res <- data.frame(rowData(sb)) %>%
-  dplyr::mutate(pval = assays(sb)$bench[,"unadjusted"])
-
-pd <- strat_hist(res, pvalue = "pval",
-           covariate = "ind_covar_AF", maxy = 3, axislabs = FALSE) +
-  draw_plot_label("GWAS case study", 0.5, 1, fontface = "bold", hjust = 0.5, size = 12) +
-  draw_plot_label("p-value by Minor Allele Frequency (MAF) group", size = 11, 0.5, 0.14, fontface = "plain", 
-                  hjust = 0.5) +
-  draw_plot_label("Density", size = 11, 0, 0.3, fontface = "plain", angle=90)
-
-p6 <- plot_grid(ggdraw(add_sub(pa, " ", size = 2)), 
-                ggdraw(add_sub(pb, " ", size = 2)), 
-                ggdraw(add_sub(pc, " ", size = 2)), 
-                ggdraw(add_sub(pd, " ", size = 2)), 
-                ncol=1, labels = LETTERS[1:4])
-p6
-ggsave(file.path(outdir, "Figure6.pdf"), width=7, height=8)
-
-```
-
-# Supplementary Figures
-
-## Informativeness of covariates in case studies
-
+filename <- gsub(".rds", "", unlist(lapply(obj.names, function(x) x[4] )))
+dataset <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[3]]))
+inform <-  ifelse(grepl("uninf", resfile), "uninf", "inf")
+type <-  ifelse(grepl("null", resfile), "null", "de")
+size <-  ifelse(grepl("10", resfile), "10", "5")
+mode <-  ifelse(grepl("II", resfile), "Bimodal effect size", "Unimodal effect size")
+pi0 <-  ifelse(grepl("H", resfile), "7.5% Non-null", "30% Non-null")
+strength <- ifelse(grepl("W", resfile), "Weakly informative", "Strongly informative")
+strength <- ifelse(grepl("uninf", resfile), "Uninformative", strength)
+aseq <- seq(0.01, 0.10, by = 0.01)
+
+if (file.exists("yeast_summary_tab.rds")){
+  tab <- readRDS("yeast_summary_tab.rds")
+}else{
+  tab <- data.frame()
+  for (l in seq_along(resfile)){
+    x <- readRDS(resfile[l])
+
+    tmp2 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
+                    fill =  "FDR", annotate = NULL, 
+                    alpha = aseq) %>%
+      mutate(fdr = nrejects) %>%
+      select(method, casestudy, replicate, fdr, alpha)
+
+    if (type[l] != "null"){
+      tmp1 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
+                      fill =  "TPR", annotate = NULL, 
+                      alpha = aseq) %>%
+        mutate(tpr = nrejects) %>%
+        select(method, casestudy, replicate, tpr, alpha)
+
+      tmp <- left_join(tmp1, tmp2, by = c("method", "casestudy", "replicate", "alpha")) %>%
+        mutate(inform = inform[l],
+               type = type[l],
+               size = size[l],
+               mode = mode[l],
+               pi0 = pi0[l],
+               strength = strength[l])
+    }else{
+      tmp <- tmp2  %>%
+        mutate(inform = inform[l],
+               type = type[l],
+               size = size[l],
+               mode = mode[l],
+               pi0 = pi0[l],
+               strength = strength[l])
+    }
+
+    tab <- bind_rows(tab, tmp)
+  }
+  saveRDS(tab, file = "yeast_summary_tab.rds")
+}
+
+tab <- tab %>% select(-c(type, size))
+dftab <- tab %>% filter(inform == "inf") %>%
+  select(-inform) %>%
+  rename(fdr = "inf_fdr",
+         tpr = "inf_tpr")
+uninftab <- tab %>% filter(inform == "uninf") %>%
+  select(-c(strength, inform, casestudy)) %>%
+  rename(fdr = "uninf_fdr",
+         tpr = "uninf_tpr")
+
+dftab <- left_join(dftab, uninftab, 
+                   by = c("method", "replicate", 
+                          "mode", "pi0", "alpha"))
+
+dftab <- dftab %>%
+  mutate(tpr_diff = inf_tpr - uninf_tpr,
+         fdr_diff = inf_fdr - uninf_fdr) %>%
+  group_by(method, alpha, mode, pi0, strength) %>%
+  summarize(tpr_se = sd(tpr_diff)/n(),
+            fdr_se = sd(fdr_diff)/n(),
+            tpr_diff = mean(tpr_diff),
+            fdr_diff = mean(fdr_diff))
+
+col <- as.character(candycols$col)
+names(col) <- as.character(candycols$Method)
+lty <- as.character(candycols$lty)
+names(lty) <- as.character(candycols$Method)
+
+FigWeak <- dftab %>% filter(strength == "Weakly informative") %>%
+  ggplot(aes(x = alpha, y = fdr_diff, color = method, linetype = method)) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    geom_line(alpha = 0.85) +
+    geom_errorbar(width=.003, alpha = 0.5, aes(ymin = fdr_diff - fdr_se, 
+                                ymax = fdr_diff + fdr_se)) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
+    scale_y_continuous(expression(paste(Delta, "FDR (informative-uninformative)")), 
+                       breaks = seq(-0.03, 1, .003), labels = scales::percent,
+                       limits = c(-0.008, 0.004)) + 
+    ggtitle("") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 11, hjust = 0.8)) +
+  facet_grid( ~ mode + pi0)
+FigWeak
+
+FigWeakT <- dftab %>% filter(strength == "Weakly informative") %>%
+  ggplot(aes(x = alpha, y = tpr_diff, color = method, linetype = method)) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    geom_errorbar(width=.003, alpha = 0.5, aes(ymin = tpr_diff - tpr_se, 
+                                ymax = tpr_diff + tpr_se)) +
+    geom_line(alpha = 0.85) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
+    scale_y_continuous(expression(paste(Delta, "TPR (informative-uninformative)")), 
+                       labels = scales::percent, limits=c(-0.02, 0.07),
+                       breaks = seq(-0.02, 0.07, by = 0.02)) + 
+    ggtitle("") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 11)) +
+  facet_grid( ~ mode + pi0)
+FigWeakT
+
+FigStrong <- dftab %>% filter(strength == "Strongly informative") %>%
+  ggplot(aes(x = alpha, y = fdr_diff, color = method, linetype = method)) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    geom_line(alpha = 0.85) +
+    geom_errorbar(width=.003, alpha = 0.5, aes(ymin = fdr_diff - fdr_se, 
+                                ymax = fdr_diff + fdr_se)) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
+    scale_y_continuous(expression(paste(Delta, "FDR (informative-uninformative)")), 
+                       breaks = seq(-0.03, 1, .003), labels = scales::percent,
+                       limits = c(-0.008, 0.004)) + 
+    ggtitle("") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 11, hjust = 0.8)) +
+  facet_grid( ~ mode + pi0)
+FigStrong
+
+FigStrongT <- dftab %>% filter(strength == "Strongly informative") %>%
+  ggplot(aes(x = alpha, y = tpr_diff, color = method, linetype = method)) +
+    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
+    geom_errorbar(width=.003, alpha = 0.5, aes(ymin = tpr_diff - tpr_se, 
+                                ymax = tpr_diff + tpr_se)) +
+    geom_line(alpha = 0.85) +
+    scale_linetype_manual(values = lty) + 
+    scale_color_manual(values = col) +
+    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
+    scale_y_continuous(expression(paste(Delta, "TPR (informative-uninformative)")), 
+                       labels = scales::percent, limits=c(-0.02, 0.07),
+                       breaks = seq(-0.02, 0.07, by = 0.02)) + 
+    ggtitle("") +
+    theme_classic() +
+    theme(axis.title = element_text(face="bold", size = 10),
+          plot.title = element_text(face="bold", size = 11)) +
+  facet_grid( ~ mode + pi0)
+FigStrongT
+
+leg <- get_legend(FigStrong)
+
+FigA <- plot_grid(FigStrong + theme(legend.position="none") +  
+                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)), 
+                  FigStrongT + theme(legend.position="none") +  
+                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)),
+                 nrow = 2, align = "hv")
+
+FigB <- plot_grid(FigWeak + theme(legend.position="none") +  
+                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)), 
+                  FigWeakT + theme(legend.position="none") +  
+                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)),
+                 nrow = 2, align = "hv")
+
+Fig <- plot_grid(FigA, FigB,
+                 nrow = 2, 
+                 labels = paste0(LETTERS[1:2], "       ",
+                          c("Strongly informative covariate",
+                            "Weakly informative covariate")),
+                 hjust = 0, axis = "rlbt")
+plot_grid(Fig, leg, ncol=2, rel_widths = c(1,0.1))
+ggsave(file.path(outdir, "FigureS9.pdf"), width=10, height=12)
+```
+
+# Figures S10 & S16
+
+These figures explore the informativeness of covariates in case studies. 
 Here we create a heatmap similar to figure 4, but instead of containing the mean
@@
 Fig_impactofinfcov
-ggsave(file.path(outdir, "Figure_impactofinfcov.pdf"), width = 15.5, height = 4.25)
+ggsave(file.path(outdir, "FigureS16.pdf"), width = 15.5, height = 4.25)
 
@@
 Fig_impactofinfcov
-ggsave(file.path(outdir, "Figure_impactofinfcov_avg.pdf"), width = 5, height = 4)
 
@@
 
+if(file.exists("yeast_tab_covimpact.rds")){
+  tab <- readRDS("yeast_tab_covimpact.rds")
+}else{
   tab <- data.frame()
@@
     tab <- bind_rows(tab, tmp)
+  }
+  save(tab, file="yeast_tab_covimpact.rds")
 }
@@
   draw_plot_label("Simulation", 0.71, 0.06, size =11, fontface = "plain")
-ggsave(file.path(outdir, "Figure_impactofinfcov_avg_sim.pdf"), width=9.5, height=5)
-
-```
-
-## Figure contrasting FDR results in different yeast simulation settings
-
-```{r Figure-yeast-pi0, fig.width = 10, fig.height = 8}
-# strongly informative covariate (top panel)
+ggsave(file.path(outdir, "FigureS10.pdf"), width=9.5, height=5)
+
+```
+
+# Figure S11
+
+Figure S11 is created in `manuscript_figures_sims.Rmd`.
+
+# Figure S12
+
+This is a multi-panel Figure displaying examples of covariate line plots in the
+case studies. We'll create a 4-panel figure with the following case studies:
+RNAseq, scRNAseq, GSEA, GWAS.
+
+```{r, FigureS12, fig.width = 9, fig.height = 7}
+
+# RNAseq
 objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
-resfile <- objects[(grepl("yeast-", objects) | grepl("yeastH-", objects) | 
-                    grepl("yeastII-", objects) | grepl("yeastIIH-", objects)) & 
-                   grepl("de5", objects) & 
-                  !grepl("uninf", objects)] 
-labs <- rep("30% Non-null", length(resfile))
-labs[grepl("H", resfile)] <- "7.5% Non-null"
-labs2 <- rep("Unimodal effect size", length(resfile))
-labs2[grepl("II", resfile)] <- "Bimodal effect size"
-excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
-
-plist <- vector("list", length(resfile))
-tab <- NULL
-for (l in seq_along(resfile)){
-  de5 <- readRDS(file=resfile[l])
-  res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
-  res5d$es_dist <- labs2[l]
-  res5d$pi0 <- labs[l]
-  tab <- rbind(tab, res5d)
-}
-
-FigStrong <- plotsim_average(tab, met="FDR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE, grpVars = c("es_dist", "pi0")) +
-                ggtitle("") +
-  facet_grid( ~ es_dist + pi0) +
-  ylim(0,0.17) 
-FigStrong
-
-FigStrongT <- plotsim_average(tab, met="TPR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE, grpVars = c("es_dist", "pi0")) +
-                ggtitle("") +
-  ylim(0,0.72) +
-  facet_grid( ~ es_dist + pi0) 
-FigStrongT
-
-# weakly informative covariate (bottom panel)
+resfile <- objects[grepl("RNAseq", objects) & grepl("brain", objects) & !grepl("uninf", objects)] 
+res <- readRDS(file=resfile)
+assayNames(res) <- "qvalue"
+res <- addDefaultMetrics(res)
+pA <- covariateLinePlot(res, alpha=alpha.thresh, covname="ind_covariate", trans="log1p",
+                        nbins=25) +
+  ggtitle("RNA-seq case study: Brain") +
+  xlab("Overall mean expression percentile")
+
+leg <-  get_legend(pA)
+pA <- pA + theme(legend.position="none")
+
+# scRNAseq
+resfile <- objects[grepl("RNAseq", objects) & grepl("human", objects) &
+                   grepl("mast", objects) & grepl("det", objects) & !grepl("uninf", objects)] 
+res <- readRDS(file=resfile)
+assayNames(res) <- "qvalue"
+res <- addDefaultMetrics(res)
+pB <- covariateLinePlot(res, alpha=alpha.thresh, covname="detection", trans="log1p",
+                        nbins=25) +
+  ggtitle("scRNA-seq case study: Human, MAST") +
+  xlab("Detection rate percentile") + 
+  theme(legend.position="none")
+
+# GSEA
+resfile <- objects[grepl("GSEA", objects) & grepl("mouse", objects) & !grepl("uninf", objects)] 
+res <- readRDS(file=resfile)
+assayNames(res) <- "qvalue"
+res <- addDefaultMetrics(res)
+pC <- covariateLinePlot(res, alpha=alpha.thresh, covname="ind_covariate", trans="log1p",
+                        nbins=25) +
+  ggtitle("GSEA case study: Mouse") +
+  xlab("Gene Set size percentile") + 
+  theme(legend.position="none")
+
+# GWAS
+resfile <- objects[grepl("GWAS", objects) & grepl("maf", objects)] 
+res <- readRDS(file=resfile)
+assayNames(res) <- "qvalue"
+res <- addDefaultMetrics(res)
+pD <- covariateLinePlot(res, alpha=alpha.thresh, covname="ind_covar_AF", trans="log1p",
+                        nbins=25) +
+  ggtitle("GWAS case study") +
+  xlab("Minor Allele Frequency (MAF) percentile") +
+  theme(legend.position="none")
+
+ps12 <- plot_grid(pA, pB, pC, pD, ncol=2, labels = LETTERS[1:4])
+ps12 <- plot_grid(ps12, leg, ncol=2, rel_widths = c(1,0.2))
+ps12
+ggsave(file.path(outdir, "FigureS12.pdf"), width=9, height=7)
+```
+
+# Figure S13
+
+This is a multi-panel Figure displaying examples of covariate vs p-value scatterplots
+and/or histograms in the case studies. 
+It will be used to discuss how we evaluated the use of the 
+covariates as independent and informative in the case studies.
+
+We'll do so for the same 4 case studies as in figure S12
+
+```{r, FigureS13}
+# RNAseq 
 objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
-resfile <- objects[(grepl("yeastW-", objects) | grepl("yeastHW-", objects) | 
-                    grepl("yeastIIW-", objects) | grepl("yeastIIHW-", objects)) & 
-                   grepl("de5", objects) & 
-                  !grepl("uninf", objects)] 
-labs <- rep("30% Non-null", length(resfile))
-labs[grepl("H", resfile)] <- "7.5% Non-null"
-labs2 <- rep("Unimodal effect size", length(resfile))
-labs2[grepl("II", resfile)] <- "Bimodal effect size"
-excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
-
-plist <- vector("list", length(resfile))
-tab <- NULL
-for (l in seq_along(resfile)){
-  de5 <- readRDS(file=resfile[l])
-  res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))
-  res5d$es_dist <- labs2[l]
-  res5d$pi0 <- labs[l]
-  tab <- rbind(tab, res5d)
-}
-
-FigWeak <- plotsim_average(tab, met="FDR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE, grpVars = c("es_dist", "pi0")) +
-                ggtitle("") +
-  ylim(0,0.17) +
-  facet_grid( ~ es_dist + pi0)
-FigWeak
-
-FigWeakT <- plotsim_average(tab, met="TPR",filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE, grpVars = c("es_dist", "pi0")) +
-                ggtitle("") +
-  ylim(0,0.72) +
-  facet_grid( ~ es_dist + pi0)
-FigWeakT
-
-leg <- get_legend(FigStrong)
-
-FigA <- plot_grid(FigStrong + theme(legend.position="none") +  
-                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)), 
-                  FigStrongT + theme(legend.position="none") +  
-                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)),
-                 nrow = 2, align = "hv")
-
-FigB <- plot_grid(FigWeak + theme(legend.position="none") +  
-                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)), 
-                  FigWeakT + theme(legend.position="none") +  
-                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)),
-                 nrow = 2, align = "hv")
-
-Fig <- plot_grid(FigA, FigB,
-                 nrow = 2, 
-                 labels = paste0(LETTERS[1:2], "       ",
-                          c("Strongly informative covariate",
-                            "Weakly informative covariate")),
-                 hjust = 0)
-plot_grid(Fig, leg, ncol=2, rel_widths = c(1,0.1))
-ggsave(file.path(outdir, "Figure_yeast_varyingpi0.pdf"), width=10, height=12)
-
-```
-
-## Uninformative covariate impact on simulations
-
-Here we'll contrast the results of informative and uninformative covariates 
-on the yeast and polyester simulation (DE sample size 5). We'll plot
-differences in TPR and FDR due to informative covariate by alpha level.
-
-```{r, fig.width = 10, fig.height = 4}
-org <- "yeast"
+resfile <- objects[grepl("RNAseq", objects) & grepl("brain", objects) & !grepl("uninf", objects)] 
+sb <- readRDS(file=resfile)
+res <- data.frame(rowData(sb)) %>%
+  dplyr::mutate(pval = assays(sb)$bench[,"unadjusted"])
+pa <- strat_hist(res, pvalue = "pval",
+           covariate = "ind_covariate", maxy = 8.75, axislabs = FALSE) +
+  draw_plot_label("  RNA-seq case study: Brain", 0.5, 1, hjust = 0.5,fontface = "bold", 
+                  size = 12) +
+  draw_plot_label("p-value by Overall mean expression group", size = 11, 0.5, 0.14, fontface = "plain",
+                  hjust =0.5) +
+  draw_plot_label("Density", size = 11, 0, 0.3, fontface = "plain", angle=90)
+
+# scRNAseq
+resfile <- objects[grepl("RNAseq", objects) & grepl("human", objects) &
+                   grepl("mast", objects) & grepl("det", objects) & !grepl("uninf", objects)] 
+sb <- readRDS(file=resfile)
+
+res <- data.frame(rowData(sb)) %>%
+  dplyr::mutate(pval = assays(sb)$bench[,"unadjusted"]) 
+pb <- strat_hist(res, pvalue = "pval",
+           covariate = "detection", maxy = 12, axislabs = FALSE)+
+  draw_plot_label("  scRNA-seq case study: Human, MAST",0.5, 1, hjust = 0.5, fontface = "bold",
+                  size = 12 ) +
+  draw_plot_label("p-value by Detection rate group", size = 11, 0.5, 0.14, fontface = "plain",
+                  hjust =0.5) +
+  draw_plot_label("Density", size = 11, 0, 0.3, fontface = "plain", angle=90)
+
+# GSEA
+resfile <- objects[grepl("GSEA", objects) & grepl("mouse", objects) & !grepl("uninf", objects)] 
+sb <- readRDS(file=resfile)
+res <- data.frame(rowData(sb)) %>%
+  dplyr::mutate(pval = assays(sb)$bench[,"unadjusted"])
+
+pc <- strat_hist(res, pvalue = "pval",
+           covariate = "ind_covariate", maxy = 8.5, axislabs = FALSE) +
+  draw_plot_label("  GSEA case study: Mouse", 0.5, 1, hjust = 0.5, fontface = "bold" , size = 12) +
+  draw_plot_label("p-value by Gene Set size group", size = 11, 0.5, 0.14, fontface = "plain",
+                  hjust =0.5) +
+  draw_plot_label("Density", size = 11, 0, 0.3, fontface = "plain", angle=90)
+
+# GWAS
+resfile <- objects[grepl("GWAS", objects) & grepl("maf", objects)] 
+sb <- readRDS(file=resfile)
+res <- data.frame(rowData(sb)) %>%
+  dplyr::mutate(pval = assays(sb)$bench[,"unadjusted"])
+
+pd <- strat_hist(res, pvalue = "pval",
+           covariate = "ind_covar_AF", maxy = 3, axislabs = FALSE) +
+  draw_plot_label("GWAS case study", 0.5, 1, fontface = "bold", hjust = 0.5, size = 12) +
+  draw_plot_label("p-value by Minor Allele Frequency (MAF) group", size = 11, 0.5, 0.14, fontface = "plain", 
+                  hjust = 0.5) +
+  draw_plot_label("Density", size = 11, 0, 0.3, fontface = "plain", angle=90)
+
+ps13 <- plot_grid(ggdraw(add_sub(pa, " ", size = 2)), 
+                ggdraw(add_sub(pb, " ", size = 2)), 
+                ggdraw(add_sub(pc, " ", size = 2)), 
+                ggdraw(add_sub(pd, " ", size = 2)), 
+                ncol=1, labels = LETTERS[1:4])
+ps13
+ggsave(file.path(outdir, "FigureS13.pdf"), width=7, height=8)
+
+```
+
+# Figure S14
+
+This is a heatmap of methods by case studies with color representing the ranking
+by of number of rejections @ FDR cutoff `r alpha.thresh`.
+
+```{r, FigureS14, fig.width = 9, fig.height = 7}
+library(SummarizedBenchmark)
+library(stringr)
 
 objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
-resfile <- objects[grepl(org, objects) ] 
-resfile <- resfile[grepl("5", resfile) & !grepl("null", resfile)]
-
-# regexp to construct master data table separating case study datasets, methods
-# and covariates
-obj.names <- str_split(resfile, pattern = "/")
+
+# remove simulation objects
+sim <- unique(c(which(grepl("simulations", objects)),
+               which(grepl("yeast", objects)),
+               which(grepl("polyester", objects))))
+if (length(sim) > 0){
+  objects <- objects[-sim]
+}
+
+# remove microbiome objects from 'null' comparisons
+# removing baxter, goodrich, papa OTU-level, and So4-enigma (since they mostly found 
+# nothing for all methods and the genus-level analysis is also present)
+# Also remove'log' ubiquity since redundant
+null <- which(grepl(c("goodrich|papa|baxter"), objects) & 
+              grepl(c("otu"), objects) | 
+              grepl("log", objects) | 
+              grepl("so4", objects))
+
+if (length(null) > 0){
+  objects <- objects[-null]
+}
+
+# remove chip-seq promoter analysis (since it is the only one that uses 
+# ash and we also include the csaw analysis on the same dataset).
+promot <- which(grepl("promot", objects))
+if (length(promot) > 0){
+  objects <- objects[-promot]
+}
+
+# remove random covariate analyses
+uninf <- which(grepl("uninf", objects))
+if (length(uninf) > 0){
+  objects <- objects[-uninf]
+}
+
+# grab casestudy names
+obj.names <- str_split(objects, pattern = "/")
 casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
-filename <- gsub(".rds", "", unlist(lapply(obj.names, function(x) x[4] )))
-dataset <- unlist(lapply(str_split(filename, pattern = "-"), function(x) x[[3]]))
-inform <-  ifelse(grepl("uninf", resfile), "uninf", "inf")
-type <-  ifelse(grepl("null", resfile), "null", "de")
-size <-  ifelse(grepl("10", resfile), "10", "5")
-mode <-  ifelse(grepl("II", resfile), "Bimodal effect size", "Unimodal effect size")
-pi0 <-  ifelse(grepl("H", resfile), "7.5% Non-null", "30% Non-null")
-strength <- ifelse(grepl("W", resfile), "Weakly informative", "Strongly informative")
-strength <- ifelse(grepl("uninf", resfile), "Uninformative", strength)
-aseq <- seq(0.01, 0.10, by = 0.01)
-
-tab <- data.frame()
-for (l in seq_along(resfile)){
-  x <- readRDS(resfile[l])
-
-  tmp2 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
-                  fill =  "FDR", annotate = NULL, 
-                  alpha = alpha.thresh) %>%
-    mutate(fdr = nrejects) %>%
-    select(method, casestudy, replicate, fdr, alpha)
-
-  if (type[l] != "null"){
-    tmp1 <- tidy_df(x, colLabels = rep(filename[l], length(x)), 
-                    fill =  "TPR", annotate = NULL, 
-                    alpha = alpha.thresh) %>%
-      mutate(tpr = nrejects) %>%
-      select(method, casestudy, replicate, tpr, alpha)
-
-    tmp <- left_join(tmp1, tmp2, by = c("method", "casestudy", "replicate", "alpha")) %>%
-      mutate(inform = inform[l],
-             type = type[l],
-             size = size[l],
-             mode = mode[l],
-             pi0 = pi0[l],
-             strength = strength[l])
-  }else{
-    tmp <- tmp2  %>%
-      mutate(inform = inform[l],
-             type = type[l],
-             size = size[l],
-             mode = mode[l],
-             pi0 = pi0[l],
-             strength = strength[l])
-  }
-
-  tab <- bind_rows(tab, tmp)
-}
-
-tab <- tab %>% select(-c(type, size))
-dftab <- tab %>% filter(inform == "inf") %>%
-  select(-inform) %>%
-  rename(fdr = "inf_fdr",
-         tpr = "inf_tpr")
-uninftab <- tab %>% filter(inform == "uninf") %>%
-  select(-c(strength, inform, casestudy)) %>%
-  rename(fdr = "uninf_fdr",
-         tpr = "uninf_tpr")
-
-dftab <- left_join(dftab, uninftab, 
-                   by = c("method", "replicate", 
-                          "mode", "pi0", "alpha"))
-
-dftab <- dftab %>%
-  mutate(tpr_diff = inf_tpr - uninf_tpr,
-         fdr_diff = inf_fdr - uninf_fdr) %>%
-  group_by(method, alpha, mode, pi0, strength) %>%
-  summarize(tpr_se = sd(tpr_diff)/n(),
-            fdr_se = sd(fdr_diff)/n(),
-            tpr_diff = mean(tpr_diff),
-            fdr_diff = mean(fdr_diff))
-
-col <- as.character(candycols$col)
-names(col) <- as.character(candycols$Method)
-lty <- as.character(candycols$lty)
-names(lty) <- as.character(candycols$Method)
-
-FigWeak <- dftab %>% filter(strength == "Weakly informative") %>%
-  ggplot(aes(x = alpha, y = fdr_diff, color = method, linetype = method)) +
-    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
-    geom_line(alpha = 0.85) +
-    geom_errorbar(width=.003, alpha = 0.5, aes(ymin = fdr_diff - fdr_se, 
-                                ymax = fdr_diff + fdr_se)) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
-    scale_y_continuous(expression(paste(Delta, "FDR (informative-uninformative)")), 
-                       breaks = seq(-0.03, 1, .003), labels = scales::percent,
-                       limits = c(-0.008, 0.004)) + 
-    ggtitle("") +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 11, hjust = 0.8)) +
-  facet_grid( ~ mode + pi0)
-FigWeak
-
-FigWeakT <- dftab %>% filter(strength == "Weakly informative") %>%
-  ggplot(aes(x = alpha, y = tpr_diff, color = method, linetype = method)) +
-    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
-    geom_errorbar(width=.003, alpha = 0.5, aes(ymin = tpr_diff - tpr_se, 
-                                ymax = tpr_diff + tpr_se)) +
-    geom_line(alpha = 0.85) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
-    scale_y_continuous(expression(paste(Delta, "TPR (informative-uninformative)")), 
-                       labels = scales::percent, limits=c(-0.02, 0.07),
-                       breaks = seq(-0.02, 0.07, by = 0.02)) + 
-    ggtitle("") +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 11)) +
-  facet_grid( ~ mode + pi0)
-FigWeakT
-
-FigStrong <- dftab %>% filter(strength == "Strongly informative") %>%
-  ggplot(aes(x = alpha, y = fdr_diff, color = method, linetype = method)) +
-    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
-    geom_line(alpha = 0.85) +
-    geom_errorbar(width=.003, alpha = 0.5, aes(ymin = fdr_diff - fdr_se, 
-                                ymax = fdr_diff + fdr_se)) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
-    scale_y_continuous(expression(paste(Delta, "FDR (informative-uninformative)")), 
-                       breaks = seq(-0.03, 1, .003), labels = scales::percent,
-                       limits = c(-0.008, 0.004)) + 
-    ggtitle("") +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 11, hjust = 0.8)) +
-  facet_grid( ~ mode + pi0)
-FigStrong
-
-FigStrongT <- dftab %>% filter(strength == "Strongly informative") %>%
-  ggplot(aes(x = alpha, y = tpr_diff, color = method, linetype = method)) +
-    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
-    geom_errorbar(width=.003, alpha = 0.5, aes(ymin = tpr_diff - tpr_se, 
-                                ymax = tpr_diff + tpr_se)) +
-    geom_line(alpha = 0.85) +
-    scale_linetype_manual(values = lty) + 
-    scale_color_manual(values = col) +
-    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) + 
-    scale_y_continuous(expression(paste(Delta, "TPR (informative-uninformative)")), 
-                       labels = scales::percent, limits=c(-0.02, 0.07),
-                       breaks = seq(-0.02, 0.07, by = 0.02)) + 
-    ggtitle("") +
-    theme_classic() +
-    theme(axis.title = element_text(face="bold", size = 10),
-          plot.title = element_text(face="bold", size = 11)) +
-  facet_grid( ~ mode + pi0)
-FigStrongT
-
-leg <- get_legend(FigStrong)
-
-FigA <- plot_grid(FigStrong + theme(legend.position="none") +  
-                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)), 
-                  FigStrongT + theme(legend.position="none") +  
-                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)),
-                 nrow = 2, align = "hv")
-
-FigB <- plot_grid(FigWeak + theme(legend.position="none") +  
-                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)), 
-                  FigWeakT + theme(legend.position="none") +  
-                   scale_x_continuous("alpha cutoff", breaks=seq(0, 1, by=0.02)),
-                 nrow = 2, align = "hv")
-
-Fig <- plot_grid(FigA, FigB,
-                 nrow = 2, 
-                 labels = paste0(LETTERS[1:2], "       ",
-                          c("Strongly informative covariate",
-                            "Weakly informative covariate")),
-                 hjust = 0, axis = "rlbt")
-plot_grid(Fig, leg, ncol=2, rel_widths = c(1,0.1))
-ggsave(file.path(outdir, "Figure_infCov_yeast.pdf"), width=10, height=12)
+dataset <- unlist(lapply(obj.names, function(x) x[4] ))
+
+# not excluding any methods
+FigS14All.rowOrder <- c("bonf", "bh", "ihw", "qvalue", "bl", "adapt-glm", "lfdr",
+                     "fdrreg-e", "fdrreg-t", "ashq")
+FigS14All <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha.thresh,
+                        excludeMethods = NULL, rowOrder = FigS14All.rowOrder) +
+        theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
+              legend.position = c(0.97, 0.92), legend.justification = c(1, 1),
+              legend.background = element_rect(fill = scales::alpha("white", 3/4), color = 'black'),
+              legend.direction = "horizontal",
+              legend.text = element_text(size = 7),
+              legend.title = element_text(size = 9),
+              legend.margin = margin(6, 10, 6, 10))
+
+FigS14All
+```
+
+Extract a table of the values in Figure S14.
+
+```{r}
+# Supplementary table S3 
+library(xtable)
+tabS3 <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha.thresh,
+                        tableOnly = TRUE, excludeMethods = NULL)
+tabS3 <- tabS3 %>% 
+  mutate(mean_prop = ifelse(is.na(mean_prop), "-",
+                            paste0(round(mean_prop,2), " (", round(min_prop,2), 
+                            "-", round(max_prop, 2), ")"))) %>%
+  select(casestudy, method, mean_prop) %>%
+  spread(casestudy, mean_prop) %>% 
+  mutate(method =  factor(method, levels = rev(FigS14All.rowOrder))) %>%
+  arrange(method)
+
+print(xtable(tabS3), include.rownames=FALSE)
+
+# generate detailed case study table for creating sim vs cs ranks boxplots
+casestudy.tab <- plotMethodRanks(objects, colLabels = dataset, alpha = alpha.thresh,
+                                 tableOnly = TRUE)
+```
+
+Create more panels with heatmaps of yeast simulation results.
+
+```{r, fig.width = 14, fig.height = 5}
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+
+# grab casestudy names
+obj.names <- str_split(objects, pattern = "/")
+casestudy <- unlist(lapply(obj.names, function(x) x[3] ))
+dataset <- unlist(lapply(obj.names, function(x) x[4] ))
+
+# non-null
+wcs <- which(grepl("RNAseq", casestudy) & 
+            grepl(c("yeast-"), dataset) &
+            grepl("de", dataset) &
+            !grepl("scRNAseq", casestudy) &
+            !grepl("uninf", dataset))
+objs <- objects[wcs]
+labs <- gsub(".rds", "", 
+             unlist(lapply(str_split(dataset[wcs], pattern = "-results-"), 
+                      function(x) x[2])))
+
+pfdr <- plotMethodRanks(objs, 
+                colLabels = c("sim", "sim"), 
+                alpha = alpha.thresh, xlab = "Comparison",
+                fill = "FDR",
+                rowOrder = FigS14All.rowOrder, 
+                annotate = "FDR", excludeMethods = NULL,
+                maxColor = "black") +
+          scale_fill_gradient2(high = "darkred", 
+                         mid = "white", 
+                         low = "#2171B5", 
+                         midpoint = alpha.thresh) +
+          labs(fill = "Mean FDR") +
+          theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
+              legend.justification = c(1, 0.5),
+              legend.text = element_text(size = 7),
+              legend.title = element_text(size = 8),
+              legend.margin = margin(1, 1, 1, 1)) +
+          ylab("") + xlab("") +  
+          theme(plot.margin=unit(c(5.5, 5.5, 30, 5.5), "points"))
+
+ptpr <- plotMethodRanks(objs, 
+                colLabels = c("sim", "sim"), 
+                alpha = alpha.thresh, xlab = "Comparison",
+                fill = "TPR",
+                rowOrder = FigS14All.rowOrder,
+                annotate = "TPR", excludeMethods = NULL) +
+          theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1),
+              legend.justification = c(1, 0.5),
+              legend.text = element_text(size = 7),
+              legend.title = element_text(size = 8),
+              legend.margin = margin(1, 1, 1, 1)) +
+          ylab("") + xlab("") +  
+          theme(plot.margin=unit(c(5.5, 5.5, 30, 5.5), "points"))
+
+ggdraw() +
+  draw_plot(pfdr, 0.53, 0, 0.24, 1) +
+  draw_plot(ptpr, 0.77, 0, 0.24, 1) +
+  draw_plot(FigS14All, 0, 0, 0.52, 1) +
+  draw_plot_label(c("A", "B"), c(0, 0.53), c(1, 1), size = 15) +
+  draw_plot_label("Simulation", 0.69, 0.06, size =11, fontface = "plain")
+ggsave(file.path(outdir, "FigureS14.pdf"), width=9.5, height=5)
+
+# generate detailed yeast sim table for creating sim vs cs ranks boxplots
+yeastsim.tab <- plotMethodRanks(objs, colLabels = labs, alpha = alpha.thresh,
+                                 tableOnly = TRUE)
+```
+
+# Figure S15
+
+Create boxplot of ranking in sims versus rankings in case studies using the
+`casestudy.tab` and `yeastsim.tab` objects created above.
+
+```{r, fig.width = 6, fig.height = 4}
+casestudy.tab <- mutate(casestudy.tab, prop_casestudy = mean_prop) %>%
+  select(casestudy, method, prop_casestudy)
+yeastsim.tab <- yeastsim.tab %>% 
+  mutate(prop_yeastsim = mean_prop) %>% 
+  filter(casestudy == "de5") %>%
+  select(casestudy, method, prop_yeastsim)
+
+rank.tab <- left_join(casestudy.tab, yeastsim.tab, by = "method") %>%
+  left_join(candycols %>% dplyr::mutate(method=Method), by = "method")
+
+col <- as.character(rank.tab$col)
+names(col) <- as.character(rank.tab$method)
+
+rank.tab$method <- factor(rank.tab$method,
+                          levels = yeastsim.tab$method[order(yeastsim.tab$prop_yeastsim)])
+x <- match(rank.tab$casestudy.x, dataset)
+rank.tab$casestudy <- casestudy[x]
+
+ggplot(rank.tab %>% filter(method != "bonf"), 
+       aes(y = prop_casestudy, x = prop_yeastsim, group = prop_yeastsim, color=Method)) +
+  geom_boxplot(aes(x = prop_yeastsim, y = prop_casestudy),
+                   alpha = 0.3, position = position_dodge(width = 0),
+                   width = 0.02) +
+  scale_color_manual(values=col) +
+  theme_bw() +
+  xlab("Proportion of maximum rejections in yeast simulation (n=5)") +  
+  ylab("Proportion of maximum rejections in casestudies") +
+  labs(shape = "Case Study") 
+ggsave(file.path(outdir, "FigureS15.pdf"), width=6, height=4)
 

### 2018-10-30 datasets/GSEA/fgsea-human.Rmd
Added data description and replaced dropbox links for zenodo links.

@@
 The objective of this document is to compare different methods for multiple testing correction in the context of Gene Set Enrichment Analysis (GSEA).
-In order to get a list of genes, we will use RNA-seq data consisting of cortex and cerebellum samples of a subset of the GTEx data.
+We use RNA-seq data consisting of cortex and cerebellum samples of a subset of the GTEx individuals.
+Shortly, samples were downloaded using the *Short Read Archive Toolkit* and mapped to the human reference genome version GRCh38 using *STAR v2.4.2a*.
+*htseq-count* was used to tabulate the number of uniquely mapping reads for each gene.
+We used the `DESeq2` package to format the data into a `DESeqDataSet` object.
 
-Here we use the `fgsea` Bioconductor package to implement the GSEA method. This is 
-a Functional Class Scoring approach, which does not require setting an arbitrary threshold
-for Differential Expression, but instead relies on the gene's rank (here we rank by
-DESeq2 test statistic).
+We use the `fgsea` Bioconductor package to implement the GSEA method. This is 
+a Functional Class Scoring approach, which does not require setting an arbitrary threshold for Differential Expression, but instead relies on the gene's rank (here we rank by DESeq2 test statistic).
 
@@
 
-The dataset consists of 10 samples (6 cerebellum and 4 cortex samples), all from male individuals.
-We download a *DESeqDataSet* if not present locally and modify the design to test for differences in gene expression between cerebellums and cortex.
+We download the *DESeqDataSet* from zenodo, if not present locally, and modify the design to test for differences in gene expression between cerebellums and cortex.
 
@@
 if (!file.exists(count_file)) {
-    download.file("https://www.dropbox.com/s/itrs2a6cnmurnck/gsea-human-counts.rds?dl=1",
+    download.file("https://zenodo.org/record/1475409/files/gsea-human-counts.rds?download=1",
                   destfile = count_file)

### 2018-10-30 datasets/GSEA/fgsea-mouse.Rmd
Added data description and replaced dropbox links for zenodo links.

@@
 The objective of this vignette is to test different multiple testing methods in the context of Gene Set Enrichment Analysis (GSEA).
-To do this, we will use mouse data from the paper by [Cabezas-Wallscheid et al.
+To do this, we use data from the paper by [Cabezas-Wallscheid et al.
 (Cell stem Cell, 2014)](https://www.sciencedirect.com/science/article/pii/S1934590914003014?via%3Dihub).
-The dataset consists of RNA-seq data from mouse hematopoietic stem cells and 4 multipotent progenitor lineages.
-We will test for difference in gene expression between two different cell fractions and perform GSEA in the resulting list of genes.
+The data consist of RNA-seq data from mouse hematopoietic stem cells and multipotent progenitor lineages.
+The raw fastq data is available through the ArrayExpress database (http://www.ebi.ac.uk/arrayexpress) under accession number E-MTAB-2262.
+These data were mapped to the mouse reference genome GRCm38 (ENSEMBL release 69) using the Genomic Short-Read Nucleotide Alignment program (version 2012-07-20).
+We used htseq-count to count the number of reads overlapping with each gene and used the DESeq2 package to format the data as a DESeqDataSet R object.
 
 Here we use the `fgsea` Bioconductor package to implement the GSEA method. This is 
-a Functional Class Scoring approach, which does not require setting an arbitrary threshold
-for Differential Expression, but instead relies on the gene's rank (here we rank by
-DESeq2 test statistic).
+a Functional Class Scoring approach, which does not require setting an arbitrary threshold for Differential Expression, but instead relies on the gene's rank (here we rank by DESeq2 test statistic).
 
@@
 if (!file.exists(count_file)) {
-    download.file("https://www.dropbox.com/s/jq6giqeclxcdypq/gsea-mouse-counts.rds?dl=1",
+    download.file("https://zenodo.org/record/1475409/files/gsea-mouse-counts.rds?download=1",
                   destfile = count_file)
@@
 if (!file.exists(goset_file)) {
+    library(biomaRt)
     mart <- useMart("ensembl", "mmusculus_gene_ensembl")

### 2018-10-30 datasets/RNA-Seq/RNAseq-gtex.Rmd
Added data description and replaced dropbox links for zenodo links.

@@
 
-The first dataset consists of 10 samples each from two regions of the human basal ganglia, the nucleus accumbens and the putamen, obtained from the GTEx project.
-Shortly, samples were downloaded using the \emph{Short Read Archive Toolkit} and mapped to the human reference genome version GRCh38 using \emph{STAR v2.4.2a}.
-\emph{htseq-count} was used to tabulate the number of uniquely mapping reads for each gene.
+The data consists of 20 samples from two regions of the human basal ganglia, the nucleus accumbens and the putamen, from the GTEx project.
+Shortly, samples were downloaded using the *Short Read Archive Toolkit* and mapped to the human reference genome version GRCh38 using *STAR v2.4.2a*.
+*htseq-count* was used to tabulate the number of uniquely mapping reads for each gene.
+We used *DESeq2* to format the data into a *DESeqDataSet* object.
 
@@
 
-We download a DESeqDataSet object containing gene level counts for GTEx samples from the nucleus accumbens and the putamen.
+We download the DESeqDataSet object from zenodo that contains the gene level counts for GTEx samples.
 
@@
 if (!file.exists(count_file)) {
-    download.file("https://www.dropbox.com/s/9g06cx9vyonhdao/rnaseq-brain-counts.rds?dl=1",
-                  destfile = count_file)
+    download.file("https://zenodo.org/record/1475409/files/rnaseq-brain-counts.rds?download=1", destfile = count_file)
 }

### 2018-10-30 manuscript_figures/manuscript_figures_sims.Rmd

--- 2018-09-13 manuscript_figures/manuscript_figures_sims.Rmd
@@
 
-# Figure 3
+# Figure S3
 
@@
 
-```{r Figure3, fig.width = 10, fig.height = 12}
-## Figure 3A,B: FDR/TPR of methods across varying informativeness at alpha = 0.05
+```{r FigureS3, fig.width = 10, fig.height = 12}
+## Figure S3A,B: FDR/TPR of methods across varying informativeness at alpha = 0.05
 res_info <- dplyr::filter(resmet, setting == "varyinginfo-smooth", alpha == 0.05)
@@
 
-## Figure 3C,D: FDR/TPR of methods across varying number of tests at alpha = 0.05
+## Figure S3C,D: FDR/TPR of methods across varying number of tests at alpha = 0.05
 res_ntests <- dplyr::filter(resmet, setting == "varyingntests", alpha == 0.05)
@@
 
-## Figure 3E,F: FDR/TPR of methods across varying null proportion at alpha = 0.05
+## Figure S3E,F: FDR/TPR of methods across varying null proportion at alpha = 0.05
 res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0", alpha == 0.05)
@@
 
-## pull Figure 3 together
-Fig3 <- plot_grid(p3ab, p3cd, p3ef, ncol = 1)
-Fig3 <- plot_grid(Fig3, get_legend(p3a), rel_widths = c(1, .2))
-Fig3
-ggsave(file.path(outdir, "Figure3.pdf"), width=10, height=12)
+## pull Figure S3 together
+FigS3 <- plot_grid(p3ab, p3cd, p3ef, ncol = 1)
+FigS3 <- plot_grid(FigS3, get_legend(p3a), rel_widths = c(1, .2))
+FigS3
+ggsave(file.path(outdir, "FigureS3.pdf"), width=10, height=12)
 ```
 
-# Supplementary Figure (distributions)
+# Figure S4 (covariate relationship)
+
+This is a faceted figure with results from the pure simulation studies showing the behavior of methods across different function relationships between the informative covariate and null probability (pi0) of a test.
+In addition to the more deliberately designed sequence of informative covariates studied in the "informativeness" set of simulations described above, we also considered four functional relationships between the covariate and pi0, named "step", "sine", "cosine", and "cubic".
+Again, the FDR and TPR as well as FDR and TPR differences are plotted for all methods.
+The relationship between the covariate (sampled from the interval between 0 and 1) and pi0 is plotted for each setting as well.
+
+```{r FigureS4, fig.width = 12, fig.height = 15}
+res_icov <- dplyr::filter(resmet, grepl("^informative", setting), grepl("N\\(0, 1\\)", dist))
+res_icov <- dplyr::mutate(res_icov, relationship = gsub("informative-", "", setting))
+
+## Figure S3 A: Form of informative covariate and pi0 relationship
+pi0form <- tibble(x = seq(0, 1, .01))
+pi0form <- dplyr::mutate(pi0form,
+                         cosine = pi0_cosine(0.9)(x),
+                         cubic = pi0_cubic(0.9)(x),
+                         sine = pi0_sine(0.9)(x),
+                         step = pi0_step(0.9)(x))
+pi0form <- tidyr::gather(pi0form, relationship, pi0, -x)
+ps4a <- ggplot(pi0form, aes(x = x, y = pi0)) +
+    geom_line() +
+    expand_limits(y = c(0.5, 1)) + 
+    theme_classic() +
+    scale_x_continuous(breaks = seq(0, 1, .2)) + 
+    scale_y_continuous(labels = scales::percent) + 
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    facet_grid(. ~ relationship)
+
+## Figure S4 B,C: FDR/TPR of methods across varying relationships
+ps4b <- genplot(res_icov, cov = alpha, met = "FDR") +
+    facet_grid(. ~ relationship, labeller = label_parsed) +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0) +
+    ylab(expression(paste("Mean ", Delta, "FDR (informative-uninformative)")))
+ps4c <- genplot(res_icov, cov = alpha, met = "TPR") +
+    facet_grid(. ~ relationship, labeller = label_parsed) +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0) +
+    ylab(expression(paste("Mean ", Delta, "TPR (informative-uninformative)")))
+
+## figure S4 D,E: FDR/TPR difference of methods
+ps4d <- genplot(res_icov, cov = alpha, type = "diff", met = "FDR") +
+    facet_grid(. ~ relationship, labeller = label_parsed) +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0)
+ps4e <- genplot(res_icov, cov = alpha, type = "diff", met = "TPR") +
+    facet_grid(. ~ relationship, labeller = label_parsed) +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0)
+
+## ## create title for SFigure Sim4
+## ps4Title <- ggdraw() +
+##     draw_label("Simulation performance across informative covariate relationship",
+##                fontface = 'bold', size = 20)
+
+## pull FigureS4 together
+SFigS4 <- plot_grid(ps4a + guides(color = FALSE, linetype = FALSE),
+                    ps4b + guides(color = FALSE, linetype = FALSE),
+                    ps4c + guides(color = FALSE, linetype = FALSE),
+                    ps4d + guides(color = FALSE, linetype = FALSE),
+                    ps4e + guides(color = FALSE, linetype = FALSE),
+                    labels = LETTERS[1:5], ncol = 1, rel_heights = c(.6, rep(1, 4)))
+## SFigS4 <- plot_grid(ps4Title, SFigS4, ncol = 1, rel_heights = c(0.03, 1))
+SFigS4 <- plot_grid(SFigS4, get_legend(ps4b), rel_widths = c(6.5, 1))
+SFigS4
+ggsave(file.path(outdir, "FigureS4.pdf"), width = 12, height = 15)
+```
+
+# Figure S5 (distributions)
 
@@
 
-```{r SFigureSim1, fig.width = 12, fig.height = 16}
+```{r FigureS5, fig.width = 12, fig.height = 15}
 res_noise <- dplyr::filter(resmet, setting == "informative-cubic")
 
-## SFigure Sim2 A: Distirbution of unimodal effect sizes (based on random samples)
+## Figure S5 A: Distirbution of unimodal effect sizes (based on random samples)
 set.seed(100)
@@
 
-## SFigure Sim1 A: Distirbution of test statistics (based on single replication)
-ps1a <- ggplot(tsdists, aes(x = test_statistic, y=..count..,
+## Figure S5 A: Distirbution of test statistics (based on single replication)
+ps5a <- ggplot(tsdists, aes(x = test_statistic, y=..count..,
                             group = truth, color = truth)) +
@@
 
-## SFigure Sim1 B,C: FDR/TPR of methods across varying test statistics distribution
-ps1b <- genplot(res_noise, cov = alpha, met = "FDR") +
+## Figure S5 B,C: FDR/TPR of methods across varying test statistics distribution
+ps5b <- genplot(res_noise, cov = alpha, met = "FDR") +
     facet_grid(. ~ dist, labeller = label_parsed) +
@@
     expand_limits(x = 0)
-ps1c <- genplot(res_noise, cov = alpha, met = "TPR") +
+ps5c <- genplot(res_noise, cov = alpha, met = "TPR") +
     facet_grid(. ~ dist, labeller = label_parsed) +
@@
 
-## SFigure Sim1 D,E: FDR/TPR difference of methods
-ps1d <- genplot(res_noise, cov = alpha, type = "diff", met = "FDR") +
+## Figure S5 D,E: FDR/TPR difference of methods
+ps5d <- genplot(res_noise, cov = alpha, type = "diff", met = "FDR") +
     facet_grid(. ~ dist, labeller = label_parsed) + 
@@
     expand_limits(x = 0)
-ps1e <- genplot(res_noise, cov = alpha, type = "diff", met = "TPR") +
+ps5e <- genplot(res_noise, cov = alpha, type = "diff", met = "TPR") +
     facet_grid(. ~ dist, labeller = label_parsed) + 
@@
 
-## create title for SFigure Sim1
-ps1Title <- ggdraw() +
-    draw_label("Simulation performance across test statistic distributions",
-               fontface = 'bold', size = 20)
+## ## create title for SFigure Sim1
+## ps5Title <- ggdraw() +
+##     draw_label("Simulation performance across test statistic distributions",
+##                fontface = 'bold', size = 20)
 
-## pull SFigure Sim1 together
-SFigS1 <- plot_grid(ps1a,
-                    ps1b + guides(color = FALSE, linetype = FALSE),
-                    ps1c + guides(color = FALSE, linetype = FALSE),
-                    ps1d + guides(color = FALSE, linetype = FALSE),
-                    ps1e + guides(color = FALSE, linetype = FALSE),
+## pull Figure S5 together
+SFigS5 <- plot_grid(ps5a,
+                    ps5b + guides(color = FALSE, linetype = FALSE),
+                    ps5c + guides(color = FALSE, linetype = FALSE),
+                    ps5d + guides(color = FALSE, linetype = FALSE),
+                    ps5e + guides(color = FALSE, linetype = FALSE),
                     labels = LETTERS[1:5], ncol = 1, rel_heights = c(.6, rep(1, 4)))
-SFigS1 <- plot_grid(ps1Title, SFigS1, ncol = 1, rel_heights = c(0.03, 1))
-SFigS1 <- plot_grid(SFigS1, get_legend(ps1b), rel_widths = c(6.5, 1))
-SFigS1
-ggsave(file.path(outdir, "SFigureSim1.pdf"), width=12, height=16)
+## FigS5 <- plot_grid(ps5Title, SFigS5, ncol = 1, rel_heights = c(0.03, 1))
+SFigS5 <- plot_grid(SFigS5, get_legend(ps5b), rel_widths = c(6.5, 1))
+SFigS5
+ggsave(file.path(outdir, "FigureS5.pdf"), width = 12, height = 15)
 ```
 
-# Supplementary Figure (unimodal assumption)
+# Figure S6 (unimodal assumption)
 
@@
 
-```{r SFigureSim2, fig.width = 12, fig.height = 16}
+```{r FigureS6, fig.width = 12, fig.height = 15}
 res_ua <- dplyr::filter(resmet, setting == "uasettings", grepl("N\\(0, 1\\)", dist))
 
-## SFigure Sim2 A: Distirbution of unimodal effect sizes (based on random samples)
+## FigureS6 A: Distirbution of unimodal effect sizes (based on random samples)
 set.seed(100)
@@
 
-ps2a <- ggplot(uadists, aes(x = x, y = density)) +
+ps6a <- ggplot(uadists, aes(x = x, y = density)) +
     geom_line() +
@@
 
-## SFigure Sim2 B,C: FDR/TPR of methods across unimodal effect size distribution
-ps2b <- genplot(res_ua, cov = alpha, met = "FDR") +
+## Figure S6 B,C: FDR/TPR of methods across unimodal effect size distribution
+ps6b <- genplot(res_ua, cov = alpha, met = "FDR") +
     facet_grid(. ~ signal, labeller = label_parsed) +
@@
     expand_limits(x = 0)
-ps2c <- genplot(res_ua, cov = alpha, met = "TPR") +
+ps6c <- genplot(res_ua, cov = alpha, met = "TPR") +
     facet_grid(. ~ signal, labeller = label_parsed) +
@@
 
-## SFigure Sim2 D,E: FDR/TPR difference of methods
-ps2d <- genplot(res_ua, cov = alpha, type = "diff", met = "FDR") +
+## Figure S6 D,E: FDR/TPR difference of methods
+ps6d <- genplot(res_ua, cov = alpha, type = "diff", met = "FDR") +
     facet_grid(. ~ signal, labeller = label_parsed) +
@@
     ylab(expression(paste("Mean ", Delta, "FDR (informative-uninformative)")))
-ps2e <- genplot(res_ua, cov = alpha, type = "diff", met = "TPR") +
+ps6e <- genplot(res_ua, cov = alpha, type = "diff", met = "TPR") +
     facet_grid(. ~ signal, labeller = label_parsed) +
@@
 
-## create title for SFigure Sim2
-ps2Title <- ggdraw() +
-    draw_label("Simulation performance across unimodal effect size distributions",
-               fontface = 'bold', size = 20)
+## ## create title for Figure S6
+## ps6Title <- ggdraw() +
+##     draw_label("Simulation performance across unimodal effect size distributions",
+##                fontface = 'bold', size = 20)
 
-## pull SFigure Sim2 together
-SFigS2 <- plot_grid(ps2a + guides(color = FALSE, linetype = FALSE),
-                    ps2b + guides(color = FALSE, linetype = FALSE),
-                    ps2c + guides(color = FALSE, linetype = FALSE),
-                    ps2d + guides(color = FALSE, linetype = FALSE),
-                    ps2e + guides(color = FALSE, linetype = FALSE),
+## pull Figure S6 together
+SFigS6 <- plot_grid(ps6a + guides(color = FALSE, linetype = FALSE),
+                    ps6b + guides(color = FALSE, linetype = FALSE),
+                    ps6c + guides(color = FALSE, linetype = FALSE),
+                    ps6d + guides(color = FALSE, linetype = FALSE),
+                    ps6e + guides(color = FALSE, linetype = FALSE),
                     labels = LETTERS[1:5], ncol = 1, rel_heights = c(.6, rep(1, 4)))
-SFigS2 <- plot_grid(ps2Title, SFigS2, ncol = 1, rel_heights = c(0.03, 1))
-SFigS2 <- plot_grid(SFigS2, get_legend(ps2b), rel_widths = c(6.5, 1))
-SFigS2
-ggsave(file.path(outdir, "SFigureSim2.pdf"), width=12, height=16)
+## SFigS2 <- plot_grid(ps2Title, SFigS2, ncol = 1, rel_heights = c(0.03, 1))
+SFigS6 <- plot_grid(SFigS6, get_legend(ps6b), rel_widths = c(6.5, 1))
+SFigS6
+ggsave(file.path(outdir, "FigureS6.pdf"), width = 12, height = 15)
 ```
 
-# Supplementary Figure (covariate relationship)
+# Figure S7 (unimodal assumption, 25% non-null)
 
-This is a faceted figure with results from the pure simulation studies showing the behavior of methods across different function relationships between the informative covariate and null probability (pi0) of a test.
-In addition to the more deliberately designed sequence of informative covariates studied in the "informativeness" set of simulations described above, we also considered four functional relationships between the covariate and pi0, named "step", "sine", "cosine", and "cubic".
-Again, the FDR and TPR as well as FDR and TPR differences are plotted for all methods.
-The relationship between the covariate (sampled from the interval between 0 and 1) and pi0 is plotted for each setting as well.
-
-```{r SFigureSim3, fig.width = 12, fig.height = 16}
-res_icov <- dplyr::filter(resmet, grepl("^informative", setting), grepl("N\\(0, 1\\)", dist))
-res_icov <- dplyr::mutate(res_icov, relationship = gsub("informative-", "", setting))
-
-## SFigure Sim3 A: Form of informative covariate and pi0 relationship
-pi0form <- tibble(x = seq(0, 1, .01))
-pi0form <- dplyr::mutate(pi0form,
-                         cosine = pi0_cosine(0.9)(x),
-                         cubic = pi0_cubic(0.9)(x),
-                         sine = pi0_sine(0.9)(x),
-                         step = pi0_step(0.9)(x))
-pi0form <- tidyr::gather(pi0form, relationship, pi0, -x)
-ps3a <- ggplot(pi0form, aes(x = x, y = pi0)) +
-    geom_line() +
-    expand_limits(y = c(0.5, 1)) + 
-    theme_classic() +
-    scale_x_continuous(breaks = seq(0, 1, .2)) + 
-    scale_y_continuous(labels = scales::percent) + 
-    theme(axis.title = element_text(face = "bold", size = 10),
-          plot.title = element_text(face = "bold", size = 14)) +
-    facet_grid(. ~ relationship)
-
-## SFigure Sim3 B,C: FDR/TPR of methods across varying relationships
-ps3b <- genplot(res_icov, cov = alpha, met = "FDR") +
-    facet_grid(. ~ relationship, labeller = label_parsed) +
-    scale_x_continuous(breaks = seq(0, .1, .02)) +
-    expand_limits(x = 0) +
-    ylab(expression(paste("Mean ", Delta, "FDR (informative-uninformative)")))
-ps3c <- genplot(res_icov, cov = alpha, met = "TPR") +
-    facet_grid(. ~ relationship, labeller = label_parsed) +
-    scale_x_continuous(breaks = seq(0, .1, .02)) +
-    expand_limits(x = 0) +
-    ylab(expression(paste("Mean ", Delta, "TPR (informative-uninformative)")))
-
-## SFigure Sim3 D,E: FDR/TPR difference of methods
-ps3d <- genplot(res_icov, cov = alpha, type = "diff", met = "FDR") +
-    facet_grid(. ~ relationship, labeller = label_parsed) +
-    scale_x_continuous(breaks = seq(0, .1, .02)) +
-    expand_limits(x = 0)
-ps3e <- genplot(res_icov, cov = alpha, type = "diff", met = "TPR") +
-    facet_grid(. ~ relationship, labeller = label_parsed) +
-    scale_x_continuous(breaks = seq(0, .1, .02)) +
-    expand_limits(x = 0)
-
-## create title for SFigure Sim3
-ps3Title <- ggdraw() +
-    draw_label("Simulation performance across informative covariate relationship",
-               fontface = 'bold', size = 20)
-
-## pull SFigure Sim3 together
-SFigS3 <- plot_grid(ps3a + guides(color = FALSE, linetype = FALSE),
-                    ps3b + guides(color = FALSE, linetype = FALSE),
-                    ps3c + guides(color = FALSE, linetype = FALSE),
-                    ps3d + guides(color = FALSE, linetype = FALSE),
-                    ps3e + guides(color = FALSE, linetype = FALSE),
-                    labels = LETTERS[1:5], ncol = 1, rel_heights = c(.6, rep(1, 4)))
-SFigS3 <- plot_grid(ps3Title, SFigS3, ncol = 1, rel_heights = c(0.03, 1))
-SFigS3 <- plot_grid(SFigS3, get_legend(ps3b), rel_widths = c(6.5, 1))
-SFigS3
-ggsave(file.path(outdir, "SFigureSim3.pdf"), width=12, height=16)
-```
-
-# Supplementary Figure (AdaPT with null model)
-
-This is a faceted figure with results from running the AdaPT method with non-default parameter settings.
-Based on communications with the authors of the methods, in addition to applying the method with parameter settings described in the accompanying package vignette, we attempted running the method with a non-default set of parameters.
-Namely, we run the method including a null relationship between the covariate and the null probability as one of the candidate "models", allowing the AdaPT method to ignore the specified covariate when it does not appear to be "informative enough".
-This model is denoted by `adapt-withnull`.
-The simulations were performed with the "step" informative covariate setting described above, under which the AdaPT method achieved lower TPR with the weakly informative covariate than with a completely uninformative covariate.
-
-Since these results are not included in the primary set of simulation results summarized in the `resmet` object, we recompute them from the output of the `supplementary/simulations-informative-step-nullAdaPT.Rmd` simulation case study.
-
-We only include a subset of methods in this plot because the primary focus is on the relative performance of AdaPT with default and modified parameters.
-
-```{r SFigureS4-parse-data}
-method_subset <- c("bonf", "bh", "qvalue", "adapt-glm", "adapt-withnull")
-
-nullAdapt <- readRDS(file.path("..", "datasets", "simulations", "supplementary", "results",
-                               "nullAdaPT-informative-step-benchmark-gaussian.rds"))
-
-nullAdapt_i <- lapply(nullAdapt, `[[`, "informative")
-nullAdapt_i <- plotsim_standardize(nullAdapt_i, alpha = seq(0.01, 0.10, 0.01))
-nullAdapt_i <- dplyr::select(nullAdapt_i, rep, blabel, param.alpha, key, performanceMetric, alpha, value)
-
-nullAdapt_u <- lapply(nullAdapt, `[[`, "uninformative")
-nullAdapt_u <- plotsim_standardize(nullAdapt_u, alpha = seq(0.01, 0.10, 0.01))
-nullAdapt_u <- dplyr::select(nullAdapt_u, rep, blabel, param.alpha, key, performanceMetric, alpha, value)
-
-nullAdapt <- dplyr::full_join(nullAdapt_i, nullAdapt_u,
-                              by = c("rep", "blabel", "param.alpha", "key", "performanceMetric", "alpha"),
-                              suffix = c(".info", ".uninfo"))
-
-nullAdapt <- dplyr::filter(nullAdapt, blabel %in% method_subset)
-
-nullAdapt <- dplyr::select(nullAdapt, -param.alpha, -key)
-nullAdapt <- dplyr::rename(nullAdapt, Method = blabel, key = performanceMetric)
-nullAdapt <- dplyr::mutate(nullAdapt, Method = factor(Method, levels = method_subset))
-
-adaptmet <- dplyr::mutate(nullAdapt,
-                          value.info = ifelse(is.na(value.info) & grepl("R$", key), 0, value.info),
-                          value.uninfo = ifelse(is.na(value.uninfo) & grepl("R$", key), 0, value.uninfo))
-
-adaptmet <- dplyr::group_by(adaptmet, Method, key, alpha)
-adaptmet <- dplyr::summarize(adaptmet,
-                             mean.info = mean(value.info, na.rm = TRUE),
-                             se.info = sd(value.info, na.rm = TRUE) / sqrt(sum(!is.na(value.info))),
-                             nNA.info = sum(is.na(value.info)),
-                             mean.diff = mean(value.info - value.uninfo, na.rm = TRUE),
-                             se.diff = sd(value.info - value.uninfo, na.rm = TRUE) /
-                                 sqrt(sum(!is.na(value.info - value.uninfo))))
-adaptmet <- dplyr::ungroup(adaptmet)
-```
-
-After computing the performance metrics, we now plot the results for AdaPT with default and modified parameter settings.
-Again, we plot the mean FDR and TPR at nominal FDR cutoffs of 0.01 to 0.10 across 100 replications, as well as the FDR and TPR differences with and without the informative covariate.
-
-```{r SFigureSim4, fig.width = 8, fig.height = 6}
-## SFigure Sim4 A,B: FDR/TPR of methods including adapt with null
-ps4a <- genplot(adaptmet, cov = alpha, met = "FDR") +
-    scale_x_continuous(breaks = seq(0, .1, .02)) +
-    expand_limits(x = 0)
-ps4b <- genplot(adaptmet, cov = alpha, met = "TPR") +
-    scale_x_continuous(breaks = seq(0, .1, .02)) +
-    expand_limits(x = 0)
-
-## SFigure Sim4 C,D: FDR/TPR difference of methods
-ps4c <- genplot(adaptmet, cov = alpha, type = "diff", met = "FDR") +
-    scale_x_continuous(breaks = seq(0, .1, .02)) +
-    expand_limits(x = 0) +
-    theme(axis.title.y = element_text(face = "bold", size = 9))
-ps4d <- genplot(adaptmet, cov = alpha, type = "diff", met = "TPR") +
-    scale_x_continuous(breaks = seq(0, .1, .02)) +
-    expand_limits(x = 0) +
-    theme(axis.title.y = element_text(face = "bold", size = 9))
-
-## create title for SFigure Sim4
-ps4Title <- ggdraw() +
-    draw_label("Simulation performance with modified AdaPT parameters",
-               fontface = 'bold', size = 10)
-
-## pull SFigure Sim4 together
-SFigS4 <- plot_grid(ps4a + guides(color = FALSE, linetype = FALSE),
-                    ps4b + guides(color = FALSE, linetype = FALSE),
-                    ps4c + guides(color = FALSE, linetype = FALSE),
-                    ps4d + guides(color = FALSE, linetype = FALSE),
-                    labels = LETTERS[1:4], ncol = 2)
-SFigS4 <- plot_grid(ps4Title, SFigS4, ncol = 1, rel_heights = c(0.06, 1))
-SFigS4 <- plot_grid(SFigS4, get_legend(ps4b), rel_widths = c(4.5, 1))
-SFigS4
-ggsave(file.path(outdir, "SFigureSim4.pdf"), width=8, height=6)
-```
-
-# Supplementary Figure (unimodal assumption, 25% non-null)
-
-This is a faceted figure with results from the pure simulation studies investigating the behavior of methods acorss different sampling distributions of the effect size for non-null tests, similar to Supplementary Figure S2 above.
+This is a faceted figure with results from the pure simulation studies investigating the behavior of methods acorss different sampling distributions of the effect size for non-null tests, similar to Supplementary Figure S6 above.
 However, in contrast to the above simulation settings where only 10% of the hypotheses were non-null, these simulations were performed with a higher percentage (25%) of non-null tests.
@@
 
-```{r SFigureS5-parse-data}
+```{r FigureS7-parse-data}
 ua25files <- list.files(file.path("..", "datasets", "simulations", "supplementary", "results"),
@@
 
-```{r SFigureSim5, fig.width = 8, fig.height = 6}
-## SFigure Sim5 B,C: FDR/TPR of methods across unimodal effect size distribution
-ps5b <- genplot(ua25met, cov = alpha, met = "FDR") +
+```{r FigureS7, fig.width = 8, fig.height = 6}
+## Figure S7 B,C: FDR/TPR of methods across unimodal effect size distribution
+ps7b <- genplot(ua25met, cov = alpha, met = "FDR") +
     facet_grid(. ~ signal, labeller = label_parsed) +
@@
     expand_limits(x = 0)
-ps5c <- genplot(ua25met, cov = alpha, met = "TPR") +
+ps7c <- genplot(ua25met, cov = alpha, met = "TPR") +
     facet_grid(. ~ signal, labeller = label_parsed) +
@@
 
-## SFigure Sim5 D,E: FDR/TPR difference of methods
-ps5d <- genplot(ua25met, cov = alpha, type = "diff", met = "FDR") +
+## Figure S7 D,E: FDR/TPR difference of methods
+ps7d <- genplot(ua25met, cov = alpha, type = "diff", met = "FDR") +
     facet_grid(. ~ signal, labeller = label_parsed) +
@@
     ylab(expression(paste("Mean ", Delta, "FDR (informative-uninformative)")))
-ps5e <- genplot(ua25met, cov = alpha, type = "diff", met = "TPR") +
+ps7e <- genplot(ua25met, cov = alpha, type = "diff", met = "TPR") +
     facet_grid(. ~ signal, labeller = label_parsed) +
@@
 
-## create title for SFigure Sim5
-ps5Title <- ggdraw() +
-    draw_label("Simulation performance across unimodal effect size distributions w/ 25% non-null",
-               fontface = 'bold', size = 15)
+## ## create title for SFigure Sim5
+## ps7Title <- ggdraw() +
+##     draw_label("Simulation performance across unimodal effect size distributions w/ 25% non-null",
+##                fontface = 'bold', size = 15)
 
-## pull SFigure Sim5 together
-SFigS5 <- plot_grid(ps2a + guides(color = FALSE, linetype = FALSE),
-                    ps5b + guides(color = FALSE, linetype = FALSE),
-                    ps5c + guides(color = FALSE, linetype = FALSE),
-                    ps5d + guides(color = FALSE, linetype = FALSE),
-                    ps5e + guides(color = FALSE, linetype = FALSE),
+## pull Figure S7 together
+SFigS7 <- plot_grid(ps6a + guides(color = FALSE, linetype = FALSE),
+                    ps7b + guides(color = FALSE, linetype = FALSE),
+                    ps7c + guides(color = FALSE, linetype = FALSE),
+                    ps7d + guides(color = FALSE, linetype = FALSE),
+                    ps7e + guides(color = FALSE, linetype = FALSE),
                     labels = LETTERS[1:5], ncol = 1, rel_heights = c(.6, rep(1, 4)))
-SFigS5 <- plot_grid(ps5Title, SFigS5, ncol = 1, rel_heights = c(0.03, 1))
-SFigS5 <- plot_grid(SFigS5, get_legend(ps5b), rel_widths = c(6.5, 1))
-SFigS5
-ggsave(file.path(outdir, "SFigureSim5.pdf"), width=12, height=16)
+## SFigS5 <- plot_grid(ps7Title, SFigS7, ncol = 1, rel_heights = c(0.03, 1))
+SFigS7 <- plot_grid(SFigS7, get_legend(ps7b), rel_widths = c(6.5, 1))
+SFigS7
+ggsave(file.path(outdir, "FigureS7.pdf"), width = 12, height = 15)
+```
+
+# Figure S8 (informativeness covariates)
+
+This is a faceted figure illustrating the functional relationship between the informative covariate and the the null proportion, pi0, when using either the continuous or discrete "informativeness" functions.
+The relationship is shown over a range of informativeness values for the two settings.
+These functions were used to investigate the impact of increasing/descreasing informativeness of the covariates while fixing other simulation conditions.
+
+```{r FigureS8, fig.width = 12, fig.height = 4}
+## Figure S8 A: Informativeness covariate relationships (continuous)
+cinfovar <- tibble(informativeness = seq(0, 1, by = 0.2))
+cinfovar <- dplyr::mutate(cinfovar, x = list(seq(0, 1, .01)))
+cinfovar <- dplyr::mutate(cinfovar, pi0 = lapply(informativeness, function(zz) pi0_varyinfo80l(zz)(x[[1]])))
+cinfovar <- dplyr::mutate(cinfovar, informativeness = informativeness * 100)
+cinfovar <- tidyr::unnest(cinfovar)
+
+ps8a <- ggplot(cinfovar, aes(x = x, y = pi0)) +
+    geom_line() +
+    expand_limits(y = c(0.5, 1)) + 
+    theme_classic() +
+    scale_x_continuous(breaks = seq(0, 1, .2)) + 
+    scale_y_continuous(labels = scales::percent) + 
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    facet_grid(. ~ informativeness, labeller = label_both)
+
+## Figure S6 B: Informativeness covariate relationships (discrete)
+dinfovar <- tibble(informativeness = seq(0, 1, by = 0.2))
+dinfovar <- dplyr::mutate(dinfovar, x = list(seq(0, 1, .01)))
+dinfovar <- dplyr::mutate(dinfovar, pi0 = lapply(informativeness, function(zz) pi0_varyinfo80(zz)(x[[1]])))
+dinfovar <- dplyr::mutate(dinfovar, informativeness = informativeness * 100)
+dinfovar <- tidyr::unnest(dinfovar)
+
+ps8b <- ggplot(dinfovar, aes(x = x, y = pi0)) +
+    geom_line() +
+    expand_limits(y = c(0.5, 1)) + 
+    theme_classic() +
+    scale_x_continuous(breaks = seq(0, 1, .2)) + 
+    scale_y_continuous(labels = scales::percent) + 
+    theme(axis.title = element_text(face = "bold", size = 10),
+          plot.title = element_text(face = "bold", size = 14)) +
+    facet_grid(. ~ informativeness, labeller = label_both)
+
+## pull Figure S6 together
+SFigS8 <- plot_grid(ps8a, ps8b, labels = LETTERS[1:2], ncol = 1)
+SFigS8
+ggsave(file.path(outdir, "FigureS8.pdf"), width = 12, height = 4)
+```
+
+# Figure S11 (AdaPT with null model)
+
+This is a faceted figure with results from running the AdaPT method with non-default parameter settings.
+Based on communications with the authors of the methods, in addition to applying the method with parameter settings described in the accompanying package vignette, we attempted running the method with a non-default set of parameters.
+Namely, we run the method including a null relationship between the covariate and the null probability as one of the candidate "models", allowing the AdaPT method to ignore the specified covariate when it does not appear to be "informative enough".
+This model is denoted by `adapt-withnull`.
+The simulations were performed with the "step" informative covariate setting described above, under which the AdaPT method achieved lower TPR with the weakly informative covariate than with a completely uninformative covariate.
+
+Since these results are not included in the primary set of simulation results summarized in the `resmet` object, we recompute them from the output of the `supplementary/simulations-informative-step-nullAdaPT.Rmd` simulation case study.
+
+We only include a subset of methods in this plot because the primary focus is on the relative performance of AdaPT with default and modified parameters.
+
+```{r FigureS11-parse-data}
+method_subset <- c("bonf", "bh", "qvalue", "adapt-glm", "adapt-withnull")
+
+nullAdapt <- readRDS(file.path("..", "datasets", "simulations", "supplementary", "results",
+                               "nullAdaPT-informative-step-benchmark-gaussian.rds"))
+
+nullAdapt_i <- lapply(nullAdapt, `[[`, "informative")
+nullAdapt_i <- plotsim_standardize(nullAdapt_i, alpha = seq(0.01, 0.10, 0.01))
+nullAdapt_i <- dplyr::select(nullAdapt_i, rep, blabel, param.alpha, key, performanceMetric, alpha, value)
+
+nullAdapt_u <- lapply(nullAdapt, `[[`, "uninformative")
+nullAdapt_u <- plotsim_standardize(nullAdapt_u, alpha = seq(0.01, 0.10, 0.01))
+nullAdapt_u <- dplyr::select(nullAdapt_u, rep, blabel, param.alpha, key, performanceMetric, alpha, value)
+
+nullAdapt <- dplyr::full_join(nullAdapt_i, nullAdapt_u,
+                              by = c("rep", "blabel", "param.alpha", "key", "performanceMetric", "alpha"),
+                              suffix = c(".info", ".uninfo"))
+
+nullAdapt <- dplyr::filter(nullAdapt, blabel %in% method_subset)
+
+nullAdapt <- dplyr::select(nullAdapt, -param.alpha, -key)
+nullAdapt <- dplyr::rename(nullAdapt, Method = blabel, key = performanceMetric)
+nullAdapt <- dplyr::mutate(nullAdapt, Method = factor(Method, levels = method_subset))
+
+adaptmet <- dplyr::mutate(nullAdapt,
+                          value.info = ifelse(is.na(value.info) & grepl("R$", key), 0, value.info),
+                          value.uninfo = ifelse(is.na(value.uninfo) & grepl("R$", key), 0, value.uninfo))
+
+adaptmet <- dplyr::group_by(adaptmet, Method, key, alpha)
+adaptmet <- dplyr::summarize(adaptmet,
+                             mean.info = mean(value.info, na.rm = TRUE),
+                             se.info = sd(value.info, na.rm = TRUE) / sqrt(sum(!is.na(value.info))),
+                             nNA.info = sum(is.na(value.info)),
+                             mean.diff = mean(value.info - value.uninfo, na.rm = TRUE),
+                             se.diff = sd(value.info - value.uninfo, na.rm = TRUE) /
+                                 sqrt(sum(!is.na(value.info - value.uninfo))))
+adaptmet <- dplyr::ungroup(adaptmet)
+```
+
+After computing the performance metrics, we now plot the results for AdaPT with default and modified parameter settings.
+Again, we plot the mean FDR and TPR at nominal FDR cutoffs of 0.01 to 0.10 across 100 replications, as well as the FDR and TPR differences with and without the informative covariate.
+
+```{r FigureSim11, fig.width = 8, fig.height = 6}
+## Figure S11 A,B: FDR/TPR of methods including adapt with null
+ps11a <- genplot(adaptmet, cov = alpha, met = "FDR") +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0)
+ps11b <- genplot(adaptmet, cov = alpha, met = "TPR") +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0)
+
+## Figure S11 C,D: FDR/TPR difference of methods
+ps11c <- genplot(adaptmet, cov = alpha, type = "diff", met = "FDR") +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0) +
+    theme(axis.title.y = element_text(face = "bold", size = 9))
+ps11d <- genplot(adaptmet, cov = alpha, type = "diff", met = "TPR") +
+    scale_x_continuous(breaks = seq(0, .1, .02)) +
+    expand_limits(x = 0) +
+    theme(axis.title.y = element_text(face = "bold", size = 9))
+
+## ## create title for SFigure Sim4
+## ps11Title <- ggdraw() +
+##     draw_label("Simulation performance with modified AdaPT parameters",
+##                fontface = 'bold', size = 10)
+
+## pull SFigure Sim4 together
+SFigS11 <- plot_grid(ps11a + guides(color = FALSE, linetype = FALSE),
+                    ps11b + guides(color = FALSE, linetype = FALSE),
+                    ps11c + guides(color = FALSE, linetype = FALSE),
+                    ps11d + guides(color = FALSE, linetype = FALSE),
+                    labels = LETTERS[1:4], ncol = 2)
+## SFigS11 <- plot_grid(ps11Title, SFigS11, ncol = 1, rel_heights = c(0.06, 1))
+SFigS11 <- plot_grid(SFigS11, get_legend(ps11b), rel_widths = c(4.5, 1))
+SFigS11
+ggsave(file.path(outdir, "FigureS11.pdf"), width = 8, height = 6)
 ```

### 2018-10-31 datasets/simulations/simulations-informative-cubic.Rmd
Update text at top of some simulation Rmds

@@
 In this set of simulations, we consider settings with both null and non-null
-tests with an informative covariate. The covariate is sampled uniformly from
+tests with an informative and uninformative covariate. The informative covariate is sampled uniformly from
 the interval [0, 1], and the conditional probability of a test being non-null
-is a smooth (cubic) function of the covariate. We include simulation results
+is a smooth (cubic) function of the covariate. The uninformative covariate is simply
+uninformly sampled independently from the interval [0, 1]. The uninformative covariate is included
+as a baseline to compare the informative covariate against. We include simulation results
 again with Gaussian, t-distributed, and Chi-Squared distributed test statistics.

### 2018-10-31 datasets/simulations/simulations-informative-sine.Rmd
Update text at top of some simulation Rmds

@@
 In this set of simulations, we consider settings with both null and non-null
-tests with an informative covariate. The covariate is sampled uniformly from
+tests with an informative and uninformative covariate. The informative covariate is sampled uniformly from
 the interval [0, 1], and the conditional probability of a test being non-null
-is a smooth (sine) function of the covariate. We include simulation results
+is a smooth (sine) function of the covariate. The uninformative covariate is simply
+uninformly sampled independently from the interval [0, 1]. The uninformative covariate is included
+as a baseline to compare the informative covariate against. We include simulation results
 again with Gaussian, t-distributed, and Chi-Squared distributed test statistics.

### 2018-10-31 datasets/simulations/simulations-informative-step.Rmd
Update text at top of some simulation Rmds

@@
 In this set of simulations, we consider settings with both null and non-null
-tests with an informative covariate. The covariate is sampled uniformly from
+tests with an informative and uninformative covariate. The informative covariate is sampled uniformly from
 the interval [0, 1], and the conditional probability of a test being non-null
-is a non-smooth (monotone) step function of the covariate. We include simulation results
+is a non-smooth monotone step function of the covariate. The uninformative covariate is simply
+uninformly sampled independently from the interval [0, 1]. The uninformative covariate is included
+as a baseline to compare the informative covariate against. We include simulation results
 again with Gaussian, t-distributed, and Chi-Squared distributed test statistics.

### 2018-10-31 datasets/simulations/simulations-uasettings.Rmd
Add/modify descriptions of UA setting Rmds; add UA sims w more non-null
Update text at top of some simulation Rmds

@@
 tests with varying distribution of effect sizes under the non-null (alternative)
-setting. An informative covariate is included in the setting as described in
-`simulations-informative-cubic.Rmd`.
+setting. Both informative and uninformative covariates are included in the setting as described in
+`simulations-informative-cubic.Rmd`. The effect sizes for non-null tests are
+sampled from unimodal distributions composed of a mixture of normal distributions,
+as described in the original adaptive shrinkage (ASH) manuscript (Stephens, 2016).
 

### 2018-10-31 datasets/simulations/supplementary/simulations-informative-step-nullAdaPT.Rmd
Fix relative path of R directory for suppl Rmds; change ua output file names
Update text at top of some simulation Rmds

--- 2018-09-05 datasets/simulations/simulations-informative-step-nullAdaPT.Rmd
@@
 In this set of simulations, we consider settings with both null and non-null
-tests with an informative covariate. The covariate is sampled uniformly from
-the interval [0, 1], and the conditional probability of a test being non-null
-is a non-smooth (monotone) step function of the covariate. We include simulation
-results again with Gaussian, t-distributed, and Chi-Squared distributed test statistics.
+tests with informative and non-informative covariates as described in
+`simulations-informative-step.Rmd`.
 

### 2018-10-31 datasets/GSEA/goseq-human.Rmd
Replace dropbox links with zenodo

@@
 The objective of this document is to compare different methods for multiple testing correction in the context of Gene Set Enrichment Analysis (GSEA).
-In order to get a list of genes, we will use RNA-seq data consisting of cortex and cerebellum samples of a subset of the GTEx data.
+We use RNA-seq data consisting of cortex and cerebellum samples of a subset of the GTEx individuals.
+Shortly, samples were downloaded using the *Short Read Archive Toolkit* and mapped to the human reference genome version GRCh38 using *STAR v2.4.2a*.
+*htseq-count* was used to tabulate the number of uniquely mapping reads for each gene.
+We used the `DESeq2` package to format the data into a `DESeqDataSet` object.
 

### 2018-10-31 datasets/GSEA/goseq-mouse.Rmd
Replace dropbox links with zenodo

@@
 (Cell stem Cell, 2014)](https://www.sciencedirect.com/science/article/pii/S1934590914003014?via%3Dihub).
-The dataset consists of RNA-seq data from mouse hematopoietic stem cells and 4 multipotent progenitor lineages.
-We will test for difference in gene expression between two different cell fractions and perform GSEA in the resulting list of genes.
+The data consist of RNA-seq data from mouse hematopoietic stem cells and multipotent progenitor lineages.
+The raw fastq data is available through the ArrayExpress database (http://www.ebi.ac.uk/arrayexpress) under accession number E-MTAB-2262.
+These data were mapped to the mouse reference genome GRCm38 (ENSEMBL release 69) using the Genomic Short-Read Nucleotide Alignment program (version 2012-07-20).
+We used htseq-count to count the number of reads overlapping with each gene and used the DESeq2 package to format the data as a DESeqDataSet R object.
 
@@
 ```
-
-**Why no filtering on gene_biotype == "protein_coding" for this data?
-AR: I think I did this in the other vignette to keep running times short.
-This shouldn't change the result since anyway protein coding genes are those with GO categories and genes without GO assignments are ignored.**
 

### 2018-10-31 datasets/simulations/simulations-informative-cosine.Rmd

@@
 In this set of simulations, we consider settings with both null and non-null
-tests with an informative covariate. The covariate is sampled uniformly from
+tests with an informative and uninformative covariate. The informative covariate is sampled uniformly from
 the interval [0, 1], and the conditional probability of a test being non-null
-is a smooth (cosine) function of the covariate. We include simulation results
+is a smooth (cosine) function of the covariate. The uninformative covariate is simply
+uninformly sampled independently from the interval [0, 1]. The uninformative covariate is included
+as a baseline to compare the informative covariate against. We include simulation results
 again with Gaussian, t-distributed, and Chi-Squared distributed test statistics.

### 2018-10-31 datasets/simulations/simulations-uasettings-t.Rmd

@@
 ---
-title: "Simulation Study: Unimodal Settings (t test statistics)"
+title: "Simulation Study: Unimodal Settings (t-11 test statistics)"
 author: "Patrick Kimes"
@@
 tests with varying distribution of effect sizes under the non-null (alternative)
-setting. An informative covariate is included in the setting as described in
-`simulations-informative-cubic.Rmd`. This set of simulations is similar to those
-presented in `simulations-varyingeffectsize.Rmd`, however, t-distributed test
-statistics are used.
+setting. Both informative and uninformative covariates are included in the setting as described in
+`simulations-informative-cubic.Rmd`. The effect sizes for non-null tests are
+sampled from unimodal distributions composed of a mixture of normal distributions,
+as described in the original adaptive shrinkage (ASH) manuscript (Stephens, 2016).
+
+This set of simulations is similar to those presented in `simulations-uasettings.Rmd`.
+However, t-distributed test statistics with 11 degrees of freedom are considered
+rather than normally distributed test statistics.
 

### 2018-10-31 datasets/simulations/simulations-varyingntests.Rmd

@@
 tests. Since many newer FDR controlling approaches require fitting some model to
-the independent covariates, they may be sensitive to lower number of tests. An
-informative covariate is included in the setting as described in
+the independent covariates, they may be sensitive to lower numbers of tests. Both
+informative and uninformative covariates are included in these settings as described in
 `simulations-informative-sine.Rmd`.

### 2019-03-28 manuscript_figures/manuscript_figures_main.Rmd
Response to reviewer 1, comment 1
Add iCOBRA-style plots to combine Figures 2A-3A, and Figures 2B-3B.

@@
 ## exclude 'enigma-so4-otus' results with no significant calls
-objdat <- dplyr::filter(objdat, casestudy != "enigma-so4-otus")
+objdat <- dplyr::filter(objdat, !grepl("so4", casestudy))
 
@@
                   "enigma-al-otus", "enigma-ph-otus", "enigma-so4-otus", "baxter-genus", 
-                  "goodrich-genus", "papa-genus", "schubert-otus", "brain", "mir200c",
+                  "goodrich-genus", "papa-genus", "schubert-otus",
+                  "enigma-al-otus-abun", "enigma-ph-otus-abun", "enigma-so4-otus-abun",
+                  "baxter-genus-mean-abun", "goodrich-genus-abun", "papa-genus-abun",
+                  "schubert-otus-abun",
+                  "brain", "mir200c",
                   "human-mast-det", "human-mast-mean", "human-scdd-det",
@@
                   "mouse-scdd-mean", "mouse-wilcox-det", "mouse-wilcox-mean"),
-    assay = c(rep("ChIP-seq", 2), rep("GSEA", 2), rep("GWAS", 2), rep("Microbiome", 7),
+    assay = c(rep("ChIP-seq", 2), rep("GSEA", 2), rep("GWAS", 2), rep("Microbiome", 14),
               rep("RNA-seq", 2), rep("scRNA-seq", 12)))
@@
 objdat <- left_join(objdat, studies, by = "casestudy")
-objdat <- dplyr::filter(objdat, !is.na(assay))
+objdat <- dplyr::filter(objdat, !is.na(assay == "Microbiome"))
+
+# add covariate label to microbiome ubiquity 
+objdat <- objdat %>%
+          mutate(casestudy = ifelse(assay == "Microbiome", 
+                                    ifelse(!grepl("abun", casestudy),
+                                           paste0(casestudy, "-ubiq"),
+                                           casestudy),
+                                    casestudy)) %>%
+          mutate(casestudy = gsub("mean-abun", "abun", casestudy))
 
@@
           strip.text.y = element_blank(),
-          legend.position = c(0,-0.32), legend.justification = c(0, 1),
+          legend.position = c(0,-0.38), legend.justification = c(0, 1),
           legend.direction = "horizontal",
@@
 
+# Revisions - Reviewer 1, comment 1:
+
+This is a figure that combines Figure 2A and 3A into one plot (mean FDR vs mean TPR), with open or
+closed circles representing whether or not the nominal error rate is controlled (on average) at three 
+points (0.01, 0.05, 0.10).
+
+```{r}
+excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05", "bonf", "fdrreg-e")
+r1.1a <- plotsim_average(res5d, met=c("FDR", "TPR"), filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) +
+      theme(legend.position="none") +
+      ggtitle("FDR vs TPR") +
+      scale_x_continuous("FDR", breaks=seq(0, 1, by=0.01))
+
+res5d.sum <- res5d %>% 
+  group_by(blabel, performanceMetric, alpha, param.alpha, key) %>%
+  summarize(n = sum(!is.na(value)),
+            se = sd(value, na.rm = TRUE) / sqrt(n),
+            value = mean(value, na.rm = TRUE)) %>%
+  filter(round(alpha,2) %in% c(0.01, 0.05, 0.1),
+         performanceMetric %in% c("FDR", "TPR"),
+         !blabel %in% excludeSet,
+         is.na(param.alpha) | (param.alpha == alpha))
+
+tsba_m <- NULL 
+met = c("FDR", "TPR")
+for (m in seq_along(met)){
+     tmp <- res5d.sum %>% 
+     filter(performanceMetric == met[m]) %>%
+     dplyr::mutate(Method = gsub("-df03", "", blabel)) 
+     if (m > 1){
+      tsba_m <- left_join(tsba_m, tmp, 
+                          by = c("Method", "alpha", "n", 
+                                 "param.alpha", "blabel"))
+  }else{
+      tsba_m <- tmp
+  }
+}
+res5d.sum <- tsba_m
+res5d.sum <- res5d.sum %>%
+  mutate(control = value.x < alpha) %>%
+  mutate(col2 = ifelse(control, col, "white"))
+
+# add color palette
+res5d.sum$Method[grepl("^ihw-", res5d.sum$Method)] <- "ihw"
+res5d.sum <- dplyr::left_join(res5d.sum, candycols, by="Method")
+
+r1.1a <- r1.1a +
+  geom_point(data=res5d.sum, aes(x = value.x, y = value.y, 
+             shape = control), size = 3.5) +
+  geom_point(data=res5d.sum %>% filter(!control), aes(x = value.x, y = value.y), 
+             shape = 19, size = 2.5, color = "white") +
+  scale_shape_manual(values = c(1, 19)) + 
+  ggtitle("In silico experiments") +
+  scale_x_continuous(expression(paste(bold(alpha), bold(" level"))),
+                         breaks=seq(0, 1, by=0.02))
+```
+
+Next we'll do the same for 2B and 3B (polyester).
+
+```{r}
+excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05", "bonf", "fdrreg-e")
+r1.1b <- plotsim_average(res5dp, met=c("FDR", "TPR"), filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) +
+      theme(legend.position="none") +
+      ggtitle("FDR vs TPR") +
+      scale_x_continuous("FDR", breaks=seq(0, 1, by=0.01))
+
+res5d.sum <- res5dp %>% 
+  group_by(blabel, performanceMetric, alpha, param.alpha, key) %>%
+  summarize(n = sum(!is.na(value)),
+            se = sd(value, na.rm = TRUE) / sqrt(n),
+            value = mean(value, na.rm = TRUE)) %>%
+  filter(round(alpha,2) %in% c(0.01, 0.05, 0.1),
+         performanceMetric %in% c("FDR", "TPR"),
+         !blabel %in% excludeSet,
+         is.na(param.alpha) | (param.alpha == alpha))
+
+tsba_m <- NULL 
+met = c("FDR", "TPR")
+for (m in seq_along(met)){
+     tmp <- res5d.sum %>% 
+     filter(performanceMetric == met[m]) %>%
+     dplyr::mutate(Method = gsub("-df03", "", blabel)) 
+     if (m > 1){
+      tsba_m <- left_join(tsba_m, tmp, 
+                          by = c("Method", "alpha", "n", 
+                                 "param.alpha", "blabel"))
+  }else{
+      tsba_m <- tmp
+  }
+}
+res5d.sum <- tsba_m
+res5d.sum <- res5d.sum %>%
+  mutate(control = value.x < alpha) %>%
+  mutate(col2 = ifelse(control, col, "white"))
+
+# add color palette
+res5d.sum$Method[grepl("^ihw-", res5d.sum$Method)] <- "ihw"
+res5d.sum <- dplyr::left_join(res5d.sum, candycols, by="Method")
+
+r1.1b <- r1.1b +
+  geom_point(data=res5d.sum, aes(x = value.x, y = value.y, 
+             shape = control), size = 3.5) +
+  geom_point(data=res5d.sum %>% filter(!control), aes(x = value.x, y = value.y), 
+             shape = 19, size = 2.5, color = "white") +
+  geom_point(data=res5d.sum %>% filter(!control), aes(x = value.x, y = value.y), 
+             shape = 1, size = 3.5) +
+  scale_shape_manual(values = c(1, 19)) +
+  ggtitle("Polyester") +
+  scale_x_continuous(expression(paste(bold(alpha), bold(" level"))),
+                         breaks=seq(0, 1, by=0.02))
+```
+
+Put them together and save.
+
+```{r}
+r1.1 <- plot_grid(r1.1a, r1.1b, rel_widths = c(1,1.05))
+r1.1 <- plot_grid(r1.1, get_legend(p2a), rel_widths = c(1, .2)) 
+r1.1
+ggsave(file.path(outdir, "Figure_R1.1.pdf"), width = 8, height = 3)
+```
+
 # Session information

### 2019-03-28 manuscript_figures/manuscript_figures_main.Rmd
Add max prop rejects for mir200c case study in Fig4b

--- 2018-10-30 manuscript_figures/manuscript_figures_main.Rmd
@@
 
-## Figure 3B: Case study heatmap demonstrating variability in applicability
+## Figure 4B: Case study heatmap demonstrating variability in applicability
 
@@
 objdat <- left_join(objdat, studies, by = "casestudy")
-objdat <- dplyr::filter(objdat, !is.na(assay))
+objdat <- dplyr::filter(objdat, !is.na(assay == "Microbiome"))
+
+# add covariate label to microbiome ubiquity 
+objdat <- objdat %>%
+          mutate(casestudy = ifelse(assay == "Microbiome", 
+                                    ifelse(!grepl("abun", casestudy),
+                                           paste0(casestudy, "-ubiq"),
+                                           casestudy),
+                                    casestudy)) %>%
+          mutate(casestudy = gsub("mean-abun", "abun", casestudy))
 
 ## create table of rejected proportion for top methods per case study
-objdat_props <- dplyr::group_by(objdat, casestudy)
+objdat_props <- dplyr::group_by(objdat, casestudy) %>% filter(method %in% methodset)
 objdat_props <- dplyr::filter(objdat_props, nrejects == max(nrejects, na.rm = TRUE))
@@
     geom_tile() +
-    geom_text(aes(label = lab), data = filter(objdat_props, method %in% methodset),
+    geom_text(aes(label = lab), data = objdat_props,
               size = 1.8, color = "white") +

### 2019-03-28 manuscript_figures/manuscript_figures_main.Rmd
fix x-axis label

--- 2019-03-28 manuscript_figures/manuscript_figures_main.Rmd
@@
   ggtitle("In silico experiments") +
-  scale_x_continuous(expression(paste(bold(alpha), bold(" level"))),
-                         breaks=seq(0, 1, by=0.02))
+  scale_x_continuous(breaks=seq(0, 1, by=0.02))
 ```
@@
   ggtitle("Polyester") +
-  scale_x_continuous(expression(paste(bold(alpha), bold(" level"))),
-                         breaks=seq(0, 1, by=0.02))
+  scale_x_continuous(breaks=seq(0, 1, by=0.02))
 ```

### 2019-04-02 manuscript_figures/manuscript_figures_main.Rmd
move revised figure 1.1 to supplementary

@@
 
-# Revisions - Reviewer 1, comment 1:
-
-This is a figure that combines Figure 2A and 3A into one plot (mean FDR vs mean TPR), with open or
-closed circles representing whether or not the nominal error rate is controlled (on average) at three 
-points (0.01, 0.05, 0.10).
-
-```{r}
-excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05", "bonf", "fdrreg-e")
-r1.1a <- plotsim_average(res5d, met=c("FDR", "TPR"), filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) +
-      theme(legend.position="none") +
-      ggtitle("FDR vs TPR") +
-      scale_x_continuous("FDR", breaks=seq(0, 1, by=0.01))
-
-res5d.sum <- res5d %>% 
-  group_by(blabel, performanceMetric, alpha, param.alpha, key) %>%
-  summarize(n = sum(!is.na(value)),
-            se = sd(value, na.rm = TRUE) / sqrt(n),
-            value = mean(value, na.rm = TRUE)) %>%
-  filter(round(alpha,2) %in% c(0.01, 0.05, 0.1),
-         performanceMetric %in% c("FDR", "TPR"),
-         !blabel %in% excludeSet,
-         is.na(param.alpha) | (param.alpha == alpha))
-
-tsba_m <- NULL 
-met = c("FDR", "TPR")
-for (m in seq_along(met)){
-     tmp <- res5d.sum %>% 
-     filter(performanceMetric == met[m]) %>%
-     dplyr::mutate(Method = gsub("-df03", "", blabel)) 
-     if (m > 1){
-      tsba_m <- left_join(tsba_m, tmp, 
-                          by = c("Method", "alpha", "n", 
-                                 "param.alpha", "blabel"))
-  }else{
-      tsba_m <- tmp
-  }
-}
-res5d.sum <- tsba_m
-res5d.sum <- res5d.sum %>%
-  mutate(control = value.x < alpha) %>%
-  mutate(col2 = ifelse(control, col, "white"))
-
-# add color palette
-res5d.sum$Method[grepl("^ihw-", res5d.sum$Method)] <- "ihw"
-res5d.sum <- dplyr::left_join(res5d.sum, candycols, by="Method")
-
-r1.1a <- r1.1a +
-  geom_point(data=res5d.sum, aes(x = value.x, y = value.y, 
-             shape = control), size = 3.5) +
-  geom_point(data=res5d.sum %>% filter(!control), aes(x = value.x, y = value.y), 
-             shape = 19, size = 2.5, color = "white") +
-  scale_shape_manual(values = c(1, 19)) + 
-  ggtitle("In silico experiments") +
-  scale_x_continuous(breaks=seq(0, 1, by=0.02))
-```
-
-Next we'll do the same for 2B and 3B (polyester).
-
-```{r}
-excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05", "bonf", "fdrreg-e")
-r1.1b <- plotsim_average(res5dp, met=c("FDR", "TPR"), filter_set = excludeSet,
-                merge_ihw = TRUE, errorBars=TRUE) +
-      theme(legend.position="none") +
-      ggtitle("FDR vs TPR") +
-      scale_x_continuous("FDR", breaks=seq(0, 1, by=0.01))
-
-res5d.sum <- res5dp %>% 
-  group_by(blabel, performanceMetric, alpha, param.alpha, key) %>%
-  summarize(n = sum(!is.na(value)),
-            se = sd(value, na.rm = TRUE) / sqrt(n),
-            value = mean(value, na.rm = TRUE)) %>%
-  filter(round(alpha,2) %in% c(0.01, 0.05, 0.1),
-         performanceMetric %in% c("FDR", "TPR"),
-         !blabel %in% excludeSet,
-         is.na(param.alpha) | (param.alpha == alpha))
-
-tsba_m <- NULL 
-met = c("FDR", "TPR")
-for (m in seq_along(met)){
-     tmp <- res5d.sum %>% 
-     filter(performanceMetric == met[m]) %>%
-     dplyr::mutate(Method = gsub("-df03", "", blabel)) 
-     if (m > 1){
-      tsba_m <- left_join(tsba_m, tmp, 
-                          by = c("Method", "alpha", "n", 
-                                 "param.alpha", "blabel"))
-  }else{
-      tsba_m <- tmp
-  }
-}
-res5d.sum <- tsba_m
-res5d.sum <- res5d.sum %>%
-  mutate(control = value.x < alpha) %>%
-  mutate(col2 = ifelse(control, col, "white"))
-
-# add color palette
-res5d.sum$Method[grepl("^ihw-", res5d.sum$Method)] <- "ihw"
-res5d.sum <- dplyr::left_join(res5d.sum, candycols, by="Method")
-
-r1.1b <- r1.1b +
-  geom_point(data=res5d.sum, aes(x = value.x, y = value.y, 
-             shape = control), size = 3.5) +
-  geom_point(data=res5d.sum %>% filter(!control), aes(x = value.x, y = value.y), 
-             shape = 19, size = 2.5, color = "white") +
-  geom_point(data=res5d.sum %>% filter(!control), aes(x = value.x, y = value.y), 
-             shape = 1, size = 3.5) +
-  scale_shape_manual(values = c(1, 19)) +
-  ggtitle("Polyester") +
-  scale_x_continuous(breaks=seq(0, 1, by=0.02))
-```
-
-Put them together and save.
-
-```{r}
-r1.1 <- plot_grid(r1.1a, r1.1b, rel_widths = c(1,1.05))
-r1.1 <- plot_grid(r1.1, get_legend(p2a), rel_widths = c(1, .2)) 
-r1.1
-ggsave(file.path(outdir, "Figure_R1.1.pdf"), width = 8, height = 3)
-```
-
 # Session information

### 2019-04-02 manuscript_figures/manuscript_figures_supp.Rmd

@@
         number_sections: true
+editor_options: 
+  chunk_output_type: console
 ---
@@
 
+```
+
+# Figure SR1 (added in response to reviewer comments)
+
+This is a figure that combines Figure 2A and 3A into one plot (mean FDR vs mean TPR), with open or
+closed circles representing whether or not the nominal error rate is controlled (on average) at three 
+points (0.01, 0.05, 0.10).
+
+```{r}
+# methods to include in all figures ( exclude bonferroni and fdrreg-e)
+alpha <- alpha.thresh
+methodset <- c("bh", "ihw", paste0("ihw-a0", 1:9), "ihw-a10", 
+               "qvalue", "bl", "bl-df03", "lfdr",
+               "fdrreg-t", "ashq", "adapt-glm")
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+resfile <- objects[grepl("yeast-", objects) & 
+                   grepl("de5", objects) & 
+                  !grepl("uninf", objects)] 
+de5 <- readRDS(file=resfile)
+res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01)) %>% 
+  filter(blabel %in% methodset)
+excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05", "bonf", "fdrreg-e")
+r1.1a <- plotsim_average(res5d, met=c("FDR", "TPR"), filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) +
+      theme(legend.position="none") +
+      ggtitle("FDR vs TPR") +
+      scale_x_continuous("FDR", breaks=seq(0, 1, by=0.01))
+
+res5d.sum <- res5d %>% 
+  group_by(blabel, performanceMetric, alpha, param.alpha, key) %>%
+  summarize(n = sum(!is.na(value)),
+            se = sd(value, na.rm = TRUE) / sqrt(n),
+            value = mean(value, na.rm = TRUE)) %>%
+  filter(round(alpha,2) %in% c(0.01, 0.05, 0.1),
+         performanceMetric %in% c("FDR", "TPR"),
+         !blabel %in% excludeSet,
+         is.na(param.alpha) | (param.alpha == alpha))
+
+tsba_m <- NULL 
+met = c("FDR", "TPR")
+for (m in seq_along(met)){
+     tmp <- res5d.sum %>% 
+     filter(performanceMetric == met[m]) %>%
+     dplyr::mutate(Method = gsub("-df03", "", blabel)) 
+     if (m > 1){
+      tsba_m <- left_join(tsba_m, tmp, 
+                          by = c("Method", "alpha", "n", 
+                                 "param.alpha", "blabel"))
+  }else{
+      tsba_m <- tmp
+  }
+}
+res5d.sum <- tsba_m
+res5d.sum <- res5d.sum %>%
+  mutate(control = value.x < alpha)
+
+# add color palette
+res5d.sum$Method[grepl("^ihw-", res5d.sum$Method)] <- "ihw"
+res5d.sum <- dplyr::left_join(res5d.sum, candycols, by="Method") %>%
+  mutate(col2 = ifelse(control, col, "white"))
+
+r1.1a <- r1.1a +
+  geom_point(data=res5d.sum, aes(x = value.x, y = value.y, 
+             shape = control), size = 3.5) +
+  geom_point(data=res5d.sum %>% filter(!control), aes(x = value.x, y = value.y), 
+             shape = 19, size = 2.5, color = "white") +
+  scale_shape_manual(values = c(1, 19)) + 
+  ggtitle("In silico experiments") +
+  scale_x_continuous(breaks=seq(0, 1, by=0.02))
+```
+
+Next we'll do the same for 2B and 3B (polyester).
+
+```{r}
+objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
+resfile <- objects[grepl("polyester-", objects) & 
+                   grepl("de5", objects) & 
+                  !grepl("uninf", objects)] 
+de5 <- readRDS(file=resfile)
+res5dp <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01)) %>% 
+  filter(blabel %in% methodset)
+excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05", "bonf", "fdrreg-e")
+r1.1b <- plotsim_average(res5dp, met=c("FDR", "TPR"), filter_set = excludeSet,
+                merge_ihw = TRUE, errorBars=TRUE) +
+      theme(legend.position="none") +
+      ggtitle("FDR vs TPR") +
+      scale_x_continuous("FDR", breaks=seq(0, 1, by=0.01))
+
+res5d.sum <- res5dp %>% 
+  group_by(blabel, performanceMetric, alpha, param.alpha, key) %>%
+  summarize(n = sum(!is.na(value)),
+            se = sd(value, na.rm = TRUE) / sqrt(n),
+            value = mean(value, na.rm = TRUE)) %>%
+  filter(round(alpha,2) %in% c(0.01, 0.05, 0.1),
+         performanceMetric %in% c("FDR", "TPR"),
+         !blabel %in% excludeSet,
+         is.na(param.alpha) | (param.alpha == alpha))
+
+tsba_m <- NULL 
+met = c("FDR", "TPR")
+for (m in seq_along(met)){
+     tmp <- res5d.sum %>% 
+     filter(performanceMetric == met[m]) %>%
+     dplyr::mutate(Method = gsub("-df03", "", blabel)) 
+     if (m > 1){
+      tsba_m <- left_join(tsba_m, tmp, 
+                          by = c("Method", "alpha", "n", 
+                                 "param.alpha", "blabel"))
+  }else{
+      tsba_m <- tmp
+  }
+}
+res5d.sum <- tsba_m
+res5d.sum <- res5d.sum %>%
+  mutate(control = value.x < alpha)
+
+# add color palette
+res5d.sum$Method[grepl("^ihw-", res5d.sum$Method)] <- "ihw"
+res5d.sum <- dplyr::left_join(res5d.sum, candycols, by="Method") %>%
+  mutate(col2 = ifelse(control, col, "white"))
+
+r1.1b <- r1.1b +
+  geom_point(data=res5d.sum, aes(x = value.x, y = value.y, 
+             shape = control), size = 3.5) +
+  geom_point(data=res5d.sum %>% filter(!control), aes(x = value.x, y = value.y), 
+             shape = 19, size = 2.5, color = "white") +
+  geom_point(data=res5d.sum %>% filter(!control), aes(x = value.x, y = value.y), 
+             shape = 1, size = 3.5) +
+  scale_shape_manual(values = c(1, 19)) +
+  ggtitle("Polyester") +
+  scale_x_continuous(breaks=seq(0, 1, by=0.02))
+```
+
+Put them together and save.
+
+```{r}
+r1.1 <- plot_grid(r1.1a, r1.1b, rel_widths = c(1,1.05))
+r1.1 <- plot_grid(r1.1, leg, rel_widths = c(1, .2)) 
+r1.1
+ggsave(file.path(outdir, "Figure_SR1.pdf"), width = 8, height = 3)
 ```

### 2019-04-05 datasets/simulations/nulldependent/simulations-nulldependent-cubic.Rmd
Add simulation looking at dependence under null

---
title: "Simulation Study: Null Dependent Covariate (cubic)"
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

# Summary

In this set of simulations, we consider settings with both null and non-null
tests with an informative and uninformative covariate. The informative covariate is sampled uniformly from the
interval [0, 1], and the conditional probability of a test being non-null
is a smooth (cubic) function of the covariate. The uninformative covariate is simply
uninformly sampled independently from the interval [0, 1]. The uninformative covariate is included
as a baseline to compare the informative covariate against. We include simulation results
again with Gaussian distributed test statistics.

Additionally, with the informative covariate, a dependence is introduced between the coviarate 
the p-value under the null distribution.

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(parallel)
library(cowplot)

## load helper functions
for (f in list.files("../../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## new helper function
source("helpers-simulations-nulldependent.R")

## project data/results folders
resdir <- "results"
dir.create(resdir, showWarnings = FALSE, recursive = TRUE)

## intermediary files we create below
weak_file <- file.path(resdir, "nulldependent-cubic-benchmark-weaker.rds")
strong_file <- file.path(resdir, "nulldependent-cubic-benchmark-stronger.rds")

## number of cores for parallelization
cores <- 20
B <- 100

## define bechmarking design
bd <- initializeBenchDesign()
```

As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
since all settings here use Gaussian test statistics. Again, we include both
`nulltype = "empirical"` and `nulltype = "theoretical"`.

```{r}
bdplus <- bd
bdplus <- addBMethod(bdplus, "fdrreg-t",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'theoretical',
                     control = list(lambda = 0.01))
bdplus <- addBMethod(bdplus, "fdrreg-e",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'empirical',
                     control = list(lambda = 0.01))
```

All simulation settings will share the following parameters.

```{r parameters-shared}
m <- 20000                        # integer: number of hypothesis tests
pi0 <- pi0_cubic(0.90)            # numeric: proportion of null hypotheses
icovariate <- runif               # functional: independent covariate

es_dist <- rnorm_generator(3)     # functional: dist of alternative test stats
ts_dist <- rnorm_perturber(1)     # functional: sampling dist/noise for test stats
null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
null_inv <- function(p) { (2*rbinom(length(p), 1, .5) - 1) * qnorm(p / 2, 0, 1) }
seed <- 608
```

Simulation results will be presented excluding a subset of methods, and
for certain plots (upset plots), a single alpha cutoff will be used.

```{r}
excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
ualpha <- 0.05
```

# Null Dependence

Here, we show null dependence on the variate by just showing the distribution of p-values when
all tests are null (pi0 = 1). We show the distribution of p-values for two settings:
weaker and stronger dependence between the covariate and p-values under the null.

## Weaker Dependence

```{r weak-null-one-simulation}
onerun_w <- simNullDependent(bdplus, m = 20000, pi0 = 1, es_dist = es_dist, ts_dist = ts_dist,
                             icovariate = icovariate, null_dist = null_dist, null_inv = null_inv,
                             null_dependence = 0.3, execute = FALSE)
```

```{r, weak-null-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun_w, pvalue = "pval", covariate = "ind_covariate") +
    ggtitle("Dist. of null p-values w/ weaker dependence")
```

```{r, weak-null-diag-hist, results = "hide", fig.width=10, fig.height=4}
strat_hist(onerun_w, pvalue = "pval", covariate = "ind_covariate", maxy = 7, numQ = 5) +
    ggtitle("Dist. of null p-values w/ weaker dependence")
```

## Stronger Dependence

```{r strong-null-one-simulation}
onerun_s <- simNullDependent(bdplus, m = 20000, pi0 = 1, es_dist = es_dist, ts_dist = ts_dist,
                             icovariate = icovariate, null_dist = null_dist, null_inv = null_inv,
                             null_dependence = 0.7, execute = FALSE)
```

```{r, strong-null-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun_s, pvalue = "pval", covariate = "ind_covariate") +
    ggtitle("Dist. of null p-values w/ stronger dependence")
```

```{r, strong-null-diag-hist, results = "hide", fig.width=10, fig.height=4}
strat_hist(onerun_s, pvalue = "pval", covariate = "ind_covariate", maxy = 7, numQ = 5) +
    ggtitle("Dist. of null p-values w/ stronger dependence")
```

## Multipanel Plot

We also create the above plots as a multipanel plot.

```{r weak-4-panels, fig.width = 9, fig.height = 3.5}
gp_oW <- rank_scatter(onerun_w, pvalue = "pval", covariate = "ind_covariate") +
    ggtitle("Dist. of null p-values w/ weaker dependence")
gp_oS <- rank_scatter(onerun_s, pvalue = "pval", covariate = "ind_covariate") +
    ggtitle("Dist. of null p-values w/ stronger dependence")

gp_o <- plot_grid(gp_oW + expand_limits(y = c(0, 5)),
                  gp_oS + expand_limits(y = c(0, 5)),
                  labels = LETTERS[1:2], nrow = 1)
gp_o
```

We next run the simulations.

# Weaker Dependence

First, we take a look at the results with weaker dependence.

```{r weak-run-simulation}
if (file.exists(weak_file)) {
    res <- readRDS(weak_file)
} else {
    res <- mclapply(X = 1:B, FUN = simNullDependent, bench = bdplus, m = m,
                    pi0 = pi0, es_dist = es_dist, icovariate = icovariate,
                    ts_dist = ts_dist, null_dist = null_dist, null_inv = null_inv,
                    null_dependence = 0.3,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = weak_file)
}

res_dep <- lapply(res, `[[`, "null_dependent")
res_indep <- lapply(res, `[[`, "null_independent")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment. Unlike the plots above, these include both
null and non-null tests.

```{r weak-one-simulation}
onerun <- simNullDependent(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                           icovariate = icovariate, null_dist = null_dist, null_inv = null_inv,
                           null_dependence = 0.3, execute = FALSE)
```

```{r, weak-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, weak-diag-hist, results = "hide", fig.width=10, fig.height=4}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 7, numQ = 5)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r weak-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_dep, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r weak-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_dep, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res_dep, alpha = ualpha, covname = "ind_covariate")
```

We also look at the FDR as a function of the independent covariate.

```{r weak-metrics-covlineplotFDR, results = "hide"}
covariateLinePlot(res_dep, alpha = ualpha, covname = "ind_covariate", metric = "FDR")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r weak-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res_dep, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

We also compare the simulation results with and without dependence between the covariate
under the nul.

```{r weak-metrics-differences, results = "hide"}
resdfu <- plotsim_standardize(res_indep, alpha = seq(0.01, 0.10, 0.01))

resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            select(resdfu, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            by = c("rep", "blabel", "param.alpha", "key",
                                   "performanceMetric", "alpha"),
                            suffix = c(".dep", ".indep"))
resdfiu <- dplyr::mutate(resdfiu, value = value.dep - value.indep)

plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    ylab("rejections (dep - indep)")

plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("FDR (dep - indep)", labels = scales::percent)

plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("TPR (dep - indep)", labels = scales::percent)

plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("TNR (dep - indep)", labels = scales::percent)
```

## Multipanel Plot

Here, we create a multipanel plot to consolidate the results to a single figure.

```{r weak-4-panels, fig.width = 11, fig.height = 7.5}
gp_wA <- plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                         merge_ihw = TRUE, errorBars = TRUE) 
gp_wB <- plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                         merge_ihw = TRUE, errorBars = TRUE) 
gp_wC <- plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                         merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("FDR (dep - indep)", labels = scales::percent)
gp_wD <- plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                         merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("TPR (dep - indep)", labels = scales::percent)

gp_w <- plot_grid(gp_wA + theme(legend.position = "none"),
                  gp_wB + theme(legend.position = "none"),
                  gp_wC + theme(legend.position = "none"),
                  gp_wD + theme(legend.position = "none"),
                  labels = LETTERS[1:4], nrow = 2, ncol = 2)
gp_w <- plot_grid(gp_w, get_legend(gp_wA),
                  rel_widths = c(1, .2))
gp_w
```

# Stronger Dependence

Next, we take a look at the results with stronger dependence.

```{r strong-run-simulation}
if (file.exists(strong_file)) {
    res <- readRDS(strong_file)
} else {
    res <- mclapply(X = 1:B, FUN = simNullDependent, bench = bdplus, m = m,
                    pi0 = pi0, es_dist = es_dist, icovariate = icovariate,
                    ts_dist = ts_dist, null_dist = null_dist, null_inv = null_inv,
                    null_dependence = 0.7,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = strong_file)
}

res_dep <- lapply(res, `[[`, "null_dependent")
res_indep <- lapply(res, `[[`, "null_independent")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment. Unlike the plots above, these include both
null and non-null tests.

```{r strong-one-simulation}
onerun <- simNullDependent(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                           icovariate = icovariate, null_dist = null_dist, null_inv = null_inv,
                           null_dependence = 0.3, execute = FALSE)
```

```{r, strong-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, strong-diag-hist, results = "hide", fig.width=10, fig.height=4}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 7, numQ = 5)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r strong-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_dep, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r strong-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_dep, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res_dep, alpha = ualpha, covname = "ind_covariate")
```

We also look at the FDR as a function of the independent covariate.

```{r strong-metrics-covlineplotFDR, results = "hide"}
covariateLinePlot(res_dep, alpha = ualpha, covname = "ind_covariate", metric = "FDR")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r strong-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res_dep, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

We also compare the simulation results with and without dependence between the covariate
under the nul.

```{r strong-metrics-differences, results = "hide"}
resdfu <- plotsim_standardize(res_indep, alpha = seq(0.01, 0.10, 0.01))

resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            select(resdfu, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            by = c("rep", "blabel", "param.alpha", "key",
                                   "performanceMetric", "alpha"),
                            suffix = c(".dep", ".indep"))
resdfiu <- dplyr::mutate(resdfiu, value = value.dep - value.indep)

plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    ylab("rejections (dep - indep)")

plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("FDR (dep - indep)", labels = scales::percent)

plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("TPR (dep - indep)", labels = scales::percent)

plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("TNR (dep - indep)", labels = scales::percent)
```

## Multipanel Plot

Here, we create a multipanel plot to consolidate the results to a single figure.

```{r strong-4-panels, fig.width = 11, fig.height = 7.5}
gp_sA <- plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                         merge_ihw = TRUE, errorBars = TRUE) 
gp_sB <- plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                         merge_ihw = TRUE, errorBars = TRUE) 
gp_sC <- plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                         merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("FDR (dep - indep)", labels = scales::percent)
gp_sD <- plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                         merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("TPR (dep - indep)", labels = scales::percent)

gp_s <- plot_grid(gp_sA + theme(legend.position = "none"),
                  gp_sB + theme(legend.position = "none"),
                  gp_sC + theme(legend.position = "none"),
                  gp_sD + theme(legend.position = "none"),
                  labels = LETTERS[1:4], nrow = 2, ncol = 2)
gp_s <- plot_grid(gp_s, get_legend(gp_sA),
                  rel_widths = c(1, .2))
gp_s
```

# Session Info

```{r}
sessionInfo()
```

### 2019-04-05 datasets/simulations/nulldependent/helpers-simulations-nulldependent.R

@@ -1,0 +1,152 @@
+#' Simulate Collection of Test Statistics
+#'
+#' Function for simulating a large set of (independent) test results.
+#' For Normal and t-distributed test statistics, effect sizes are directly simulated
+#' from a specified true effect size distribution (`es_dist`) and perturbed according to a
+#' specified noise distribution with correspond standard errors to obtain the observed set of
+#' test statistics (`ts_dist`). For Chi-squared distributed test statistics, the notions of
+#' effect size and standard error are not well defined, and we therefore just set the standard
+#' error to 1 and effect size equal to the test statistic. This does not matter for our simulations,
+#' where the effect size and standard error are only used by a single method (ASH) which is
+#' not included in the comparisons under the Chi-squared distribution. This simulator resembles
+#' the simulations carried out in the manuscripts of the Boca-Leek, ASH, and IHW methods.
+#' (For IHW, in particular, the simulations illustrating the "size investing strategy" fall under
+#' this class of simulations.)
+#'
+#' @param X simulation index
+#' @param bench BenchDesign object of methods to run on simulated data set.
+#' @param m number of test statistics.
+#' @param pi0 proportion of null hypotheses. Can either be a single numeric value
+#'        between 0 and 1 or a function which takes a numeric vector (i.e. 'informative
+#'        covariate' values) and returns a vector of the same length containing null
+#'        probabilities between 0 and 1. 
+#' @param icovariate specification for the independent covariate, must be a function
+#'        which takes an integer value and returns a vector of the specified length to
+#'        be used as the independent covariate.
+#' @param es_dist expected effect size under the alternative, can either be a single
+#'        numeric value or a function which takes an integer as input and returns a
+#'        numeric vector of effect sizes of the specified length.
+#' @param ts_dist sampling distribution of effect size, must be a function taking two
+#'        inputs, a vector of true effect sizes and a logical value whether to return
+#'        the perturbed effect sizes or the corresponding standard error. The perturbed
+#'        effect sizes and standard errors are used to compute the test statistics.
[…]

### 2019-04-16 datasets/simulations/nulldependent/simulations-nulldependent-cubic.Rmd
Fix typo in null dep Rmd

@@
                            icovariate = icovariate, null_dist = null_dist, null_inv = null_inv,
-                           null_dependence = 0.3, execute = FALSE)
+                           null_dependence = 0.7, execute = FALSE)
 ```
